### 2018-11-30 biblio.Rmd
Moving to new directory

---
title: "Annotated Bibliography"
author: |
  | Philina English
bibliography: pbs-rangeshifts.bib
csl: apa-cv.csl
output: html_document
#biblio-style: chicago-annote
---
# csl: smithsonian-institution-scholarly-press-note.csl
# csl: apa-cv
# csl: chicago-annotated-bibliography.csl
# csl: style-manual-australian-government-note.csl
# csl: chicago-note-bibliography-16th-edition.csl

# bookdown::pdf_document2:
#  toc: no

# bibliography: [packages.bib, papers.bib]
# biblio-style: apalike

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
# I usually load my libraries up front to keep things organized
library(tidyverse)
library(bookdown)
library(knitr)
library(kableExtra)
library(ggthemes)
library(ggExtra)
library(stringr)
```

```{r}
# library(citr) # went to addin, insert citations, but too slow/didn't seem to work
```

```{r}
library(rbbt)
bbt_cite()
bbt_bib_zotero()
```

[@thorson_model-based_2016]

Key

(1) Citation

(2) Introduction

(3) Aims & Research methods

(4) Scope

(5) Usefulness (to your research/ to a particular topic)

(6) Limitations

(7) Conclusions

(8) Reflection (explain how this work illuminates your topic or how it will fit in with your research)

### 2018-11-30 code_notes.Rmd
Moving to new directory

---
title: "Code Notes"
author: "Philina English"
date: "October 24, 2018"
output: html_document
---

# installing key packages

```{r}
install.packages("tidyverse")
install.packages(c("devtools", "roxygen2", "testthat", "knitr"))
install.packages("devtools")
library(devtools)
devtools::install_github("nwfsc-assess/geostatistical_delta-GLMM")
install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
```

# ggplot2

```{r}
library(ggplot2)
 + coord_flip() # switches the x and y axes eg. if you want horizontal boxplots, long labels that are hard to get them to fit without overlapping on x-axis
 + coord_quickmap()

```

# Bibliography

```{r}
library(rbbt)
bbt_cite()
bbt_bib_zotero()
```

```{r}
library(RefManageR)
biblio <- ReadZotero(user = 863479, .params = list(collection = 'items'))
```

### 2018-12-03 range-shift-lit.Rmd
modify csl for "inline" annotations and make range-shift-lit into .Rmd

---
title: "Range shift literature review"
author: Philina English
date: "December 3, 2018"
bibliography: pbs-rangeshifts.bib
csl: apa-annotated-bibliograph-annotation-inline.csl 
output: pdf_document
---

```{r}

```

[@parmesan_globally_2003]

[@thorson_model-based_2016]

[@pinsky2013]

# References

### 2018-12-03 code_notes.Rmd
update code_notes

@@
 ```{r}
-install.packages("tidyverse")
-install.packages(c("devtools", "roxygen2", "testthat", "knitr"))
-install.packages("devtools")
-library(devtools)
-devtools::install_github("nwfsc-assess/geostatistical_delta-GLMM")
-install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
+# install.packages("tidyverse")
+# install.packages(c("devtools", "roxygen2", "testthat", "knitr"))
+# install.packages("devtools")
+# library(devtools)
+# devtools::install_github("nwfsc-assess/geostatistical_delta-GLMM")
+# install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
 ```
@@
 ```
+
+```{r eval=False}
+library(tidyverse)
+tidyverse_update()
+```

### 2018-12-04 range-shift-lit.Rmd
add a couple ref from Sunday's paper

@@
 csl: apa-annotated-bibliograph-annotation-inline.csl 
-output: pdf_document
+output: html_document
 ---
 
-```{r}
+[@parry_climate_2007]
 
-```
+[@burrows2011]
 
 [@parmesan_globally_2003]
+
+[@poloczanska2013]
 

### 2018-12-05 range-shift-lit.Rmd
interim commit on lit review notes

@@
 
+##Changing climate
 [@parry_climate_2007]
 
+### Changes in the marine environment
 [@burrows2011]
+
+##--> Many organisms must "adapt, move, or die"
+--when the environment changes, if they are to persist, organisms must be able to cope with change, change themselves, or move ([@maggini_are_2011])
+
+[@root_fingerprints_2003]
+
+###When species move: shifting distributions
+####General patterns 
+--many terrestrial species from a range of taxanomic groups have moved to higher latitudes and elevations ([@hickling_distributions_2006])
+
+--despite similar climate velocities, within regions, some species have rapidly extended their geographical distributions, while others have moved less, or have even moved in the opposite direction
 
@@
 
-[@thorson_model-based_2016]
+[@sunday_species_2015]
+
+- but climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases ([@thorson_model-based_2016])
+
+###Stages of distribution change and the importance of species traits
+[@bates_defining_2014]
+
+[@poyry_species_2009]
+###Modeling responses to environmental gradients
+[@maggini_are_2011]
 

### 2018-12-05 range-shift-lit.Rmd
additional range shift ref

@@
 
-##--> Many organisms must "adapt, move, or die"
+##Many organisms must "adapt, move, or die"
 --when the environment changes, if they are to persist, organisms must be able to cope with change, change themselves, or move ([@maggini_are_2011])
@@
 
-- but climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases ([@thorson_model-based_2016])
+--but climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases ([@thorson_model-based_2016])
+
+--even statistically and biologically significant models based on species traits can have limited exlanatory power and therefore conservation value ([@angert_species_2011])
 

### 2018-12-05 range-shift-lit.Rmd
more citations added

--- 2018-12-04 range-shift-lit.Rmd
@@
 
+##Changing climate
 [@parry_climate_2007]
 
+### Changes in the marine environment
 [@burrows2011]
+
+##In response organisms "adapt, move, or die"
+--when the environment changes, if they are to persist, organisms must be able to cope with change, change themselves, or move ([@maggini_are_2011])
+
+[@root_fingerprints_2003]
+
+###When species move: shifting distributions
+####General patterns 
+--many terrestrial species from a range of taxanomic groups have moved to higher latitudes and elevations ([@hickling_distributions_2006,@chen_rapid_2011])
+
+--despite similar climate velocities, within regions, some species have rapidly extended their geographical distributions, while others have moved less, or have even moved in the opposite direction
 
@@
 
-[@thorson_model-based_2016]
+[@sunday_species_2015]
 
-[@pinsky2013]
+--even statistically and biologically significant models based on species traits can have limited exlanatory power and therefore conservation value ([@angert_species_2011])
+
+###Stages of distribution change and the importance of species traits
+[@bates_defining_2014]
+
+[@poyry_species_2009]
+
+###Modeling responses to environmental change
+
+--distances moved were correlated with warming and generally sufficient to track temperatures latitudinally, but not elevationally [@chen_rapid_2011]
+
+--climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases ([@thorson_model-based_2016])
+
+--sea surface temperature was not correlated with range shifts [@przeslawski_using_2012]
+
+[@maggini_are_2011]
 

### 2018-12-06 range-shift-lit.Rmd
draft report for contract

@@
 ---
-title: "Range shift literature review"
+title: "Groundfish range shifts in response to climate: literature outline"
 author: Philina English
-date: "December 3, 2018"
+date: "December 5, 2018"
 bibliography: pbs-rangeshifts.bib
 csl: apa-annotated-bibliograph-annotation-inline.csl 
-output: html_document
+output: pdf_document
 ---
@@
 ### Changes in the marine environment
+
+#### Temperature
 [@burrows2011]
+
+#### Dissolved O2 
+[@keeling_ocean_2010]
+[@pierce_declining_2012]
+[@long_finding_2016]
+
+But [@schmidtko_decline_2017]
+
+[@crawford_declining_2013]
 
@@
 ####General patterns 
---many terrestrial species from a range of taxanomic groups have moved to higher latitudes and elevations ([@hickling_distributions_2006,@chen_rapid_2011])
+--many terrestrial species from a range of taxonomic groups have moved to higher latitudes and elevations ([@hickling_distributions_2006,@chen_rapid_2011])
 
@@
 
---even statistically and biologically significant models based on species traits can have limited exlanatory power and therefore conservation value ([@angert_species_2011])
+--even statistically and biologically significant models based on species traits can have limited explanatory power and therefore conservation value ([@angert_species_2011])
 
@@
 
-###Modeling responses to environmental change
+###Use abundance-occupancy patterns to assess locality restricted responses to climate?
 
---distances moved were correlated with warming and generally sufficient to track temperatures latitudinally, but not elevationally [@chen_rapid_2011]
+--centre of gravity (COG) as a measure of range shift? 
+[@thorson_model-based_2016]
 
---climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases ([@thorson_model-based_2016])
+--incorporating body size a key variable in understanding both importance of dispersal and trophic interactions in occupancy changes in fishes
+
+[@webb_birds_2011]
+
+--both more detailed assessments of distribution change than COG, including accounting for age structure are important for butterfish in the Atlantic 
+[@adams_age-specific_2017]
+
+##Modeling occupancy-density responses to environmental change
+
+--occupancy shift distances were correlated with warming and generally sufficient to track temperatures latitudinally, but not elevationally [@chen_rapid_2011]
+
+--climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases (**Thorson et al.
+2016**)
 
@@
 
+###Comparison with analagous elevation change in terrestrial environments?
+
+--through consideration of the appropriate ecological ‘units’ (species or size class), comparisons between different systems greatly benefits the search for general understanding of the processes structuring communities (**Steele 1991; Webb et al.
+2011**)
+
+--marine systems are inherently more variable than terrestrial systems at decadal timescales due to the tight coupling of physical and biological processes in the sea ([@steele_can_1991-1])
+
+--rapid, spatially-extensive human impacts on terrestrial ecosystems make them behave more like oceanic systems (in that ecological and physical processes occur at similar scales), and yet temporal variability in the distribution and abundance of fish species is still greater (**Webb et al.
+2011**)
+
+##Groundfish in Pacific Canada
+
+--large number of taxa
+
+--includes economically relevant stocks --> also human food sources
+
+--live at depth--most stable and slow to change environment, yet showing signs of responding to temperature change 
+[@dulvy_climate_2008]
+
+--region with low-average rate of warming on a global scale (**Burrows et al. 2011**)
+
+--region with higher than average rate of deoxygenation (**Keeling et al. 2010;Pierce et al. 2012; Long et al. 2016**) and a guild that appears to respond to very strongly 
+[@koslow_fish_2013]
+
+[@keller_occurrence_2015]
+
+[@keller_species-specific_2017]
+
 # References

### 2018-12-06 range-shift-lit-review-notes.Rmd
renaming

--- 2018-12-06 range-shift-lit.Rmd
@@
 bibliography: pbs-rangeshifts.bib
-csl: apa-annotated-bibliograph-annotation-inline.csl 
-output: pdf_document
+csl: apa-annotated-bibliograph-inline.csl 
+output: html_document
 ---

### 2019-01-21 literature/range-shift-lit-review-notes.Rmd
cleanup

--- 2018-12-06 range-shift-lit-review-notes.Rmd
@@
 date: "December 5, 2018"
-bibliography: pbs-rangeshifts.bib
+bibliography: pbs-rangeshifts-edit.bib
 csl: apa-annotated-bibliograph-inline.csl 

### 2019-01-21 analysis/resources/SensorDataExample.r

@@ -1,0 +1,45 @@
+# Simple example to demonstrate extraction of sensor data from GFBioSQL
+
+library(gfplot)
+require("RODBC")
+
+# SQL to extract salinity data from set #1 of the 2018 West Coast Haida Gwaii
+# synoptic trawl survey. TRIP_ID 84230 = 2018 WCHG synoptic survey;
+# FE_MAJOR_LEVEL_ID 1 = Set #1; SENSOR_DATA_ATTRIBUTE 4 = Dissolved O2;
+# SENSOR_DATA_UNIT_CODE 5 = Millilitres per litre
+sql1 <- paste("SELECT SD.TIME_STAMP,",
+              "SD.SENSOR_DATA_VALUE",
+              "FROM FISHING_EVENT FE",
+              "INNER JOIN FISHING_EVENT_SENSOR_DATAFILE FESD ON",
+              "FE.FISHING_EVENT_ID = FESD.FISHING_EVENT_ID",
+              "INNER JOIN SENSOR_DATA SD ON",
+              "FESD.SENSOR_DATAFILE_NAME = SD.SENSOR_DATAFILE_NAME AND",
+              "TIME_STAMP BETWEEN FE_BEGIN_DEPLOYMENT_TIME AND",
+              "FE_END_RETRIEVAL_TIME",
+              "WHERE FE.TRIP_ID = 84230 AND",
+              "FE.FE_MAJOR_LEVEL_ID = 1 AND",
+              "SD.SENSOR_DATA_ATTRIBUTE_CODE = 4 AND",
+              "SD.SENSOR_DATA_UNIT_CODE = 5",
+              "ORDER BY FE.FE_MAJOR_LEVEL_ID,",
+              "SD.TIME_STAMP")
+
+# SQL to extract the beginning and end of bottom contact time
+sql2 <- paste("SELECT FE_MAJOR_LEVEL_ID,",
+              "FE_BEGIN_BOTTOM_CONTACT_TIME,",
+              "FE_END_BOTTOM_CONTACT_TIME",
+              "FROM FISHING_EVENT FE",
+              "WHERE TRIP_ID = 84230 AND",
+              "FE_MAJOR_LEVEL_ID = 1")
+
+# Fetch the data frames
+# cnn <- gfplot:::db_connection()
+
[…]

### 2019-01-22 manual-geostat-TMB.Rmd
geostat model code test

---
title: "manual-geostat-TMB"
author: "Philina English"
date: "1/21/2019"
output: html_document
---

```{r}
library(dplyr)
library(gfplot)
library(ggplot2)
```

```{r}
survey <- "SYN QCS"

d <- gfplot::get_survey_sets("pacific cod", ssid = 1)
 # d <- dplyr::filter(d, !(year == 2014 & survey_abbrev == "SYN WCHG")) # something went wrong in 2014

col <- if (grepl("SYN", survey)) "density_kgpm2" else "density_ppkm2" # grep says if SYN is in survey returns logical (l), use kgpm2

dat <- gfplot:::tidy_survey_sets(d, survey, years = seq(1, 1e6), # say which surveys and years to include
    density_column = col) # make utm = x,y rename density, remove

#glimpse(dat)
  if (mean(dat$present) < 0.05) stop("Not enough data.") # stops if too little data

#  .scale <- if (grepl("SYN", survey)) 1000 else 1 # for computational stability # may be needed for INLA
#  dat <- dplyr::mutate(dat, density = density * .scale) # may be needed for INLA

  if (any(is.na(dat$depth)))
    dat <- gfplot:::interp_survey_bathymetry(dat)$data
    dat <- gfplot:::scale_survey_predictors(dat) # choose interp data if needed, and scale
```

```{r}
#  if (grepl("SYN", survey)) {
    grid_locs <- gfplot:::make_prediction_grid(
      dplyr::filter(dat, year == max(dat$year)), survey = survey,
      cell_width = 2)$grid # cell width of 2 km
    grid_locs <- dplyr::rename(grid_locs, depth = akima_depth)
#  } else {
#    grid_locs <- if (surv == "HBLL OUT N") gfplot::hbll_n_grid$grid else gfplot::hbll_s_grid$grid
 # }
  grid_locs$year <- NULL

#  formula <- if (include_depth) {
#    stats::as.formula(density ~ 0 + as.factor(year))
#  } else {
#    stats::as.formula(density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2)
# }
```

```{r}
  spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 100)

  m <- sdmTMB::sdmTMB(
            formula = density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
            data = dat, time = "year", 
            spde = spde, 
            family = sdmTMB::tweedie(link = "log"),
            anisotropy = FALSE, 
            silent = FALSE)

  names(m) 
```

```{r}
  predictions <- stats::predict(m, newdata = grid_locs)$data
# predictions$resids <- residuals(m) # randomized quantile residuals but not working
  glimpse(predictions) # est = predicted values

# A short function for plotting our predictions:
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

```{r}
#ggplot(predictions, aes(X, Y, col = resids)) + scale_colour_gradient2() +
#  geom_point() + facet_wrap(~year)

#hist(predictions$resids)
#qqnorm(predictions$resids);abline(a = 0, b = 1)
```

```{r}
plot_map(predictions, "exp(est)") +
  scale_fill_viridis_c(trans = "sqrt") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}

plot_map(predictions, "exp(est_fe)") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

```{r}
plot_map(predictions, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()
```

```{r}
plot_map(predictions, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()

```

index-standardization

### 2019-01-22 manual-geostat-TMB.Rmd
annotate manual test

@@
 
+#### Load packages
+
 ```{r}
@@
 
+#### Retrieve data
 ```{r}
@@
  # d <- dplyr::filter(d, !(year == 2014 & survey_abbrev == "SYN WCHG")) # something went wrong in 2014
+
+```
+
+#### Tidy data and fill in missing depth values
+```{r}
 
@@
 
+#### Make prediction grid for plotting
 ```{r}
@@
 
+#### Spatial model
 ```{r}
@@
 
+#### Save preditions from model
 ```{r}
@@
 
+#### Check residuals
 ```{r}
@@
 
+#### Plot overall model
 ```{r}
@@
 
+#### Plot fixed effects
 ```{r}
@@
 
+#### Plot spatial random effects
 ```{r}
@@
 
+#### Plot spatiotemporal random effects
 ```{r}
@@
 ```
-
-index-standardization

### 2019-01-22 manual-geostat-TMB.Rmd
Fix residuals (they need to be calculated on the original data... of course)

@@
 ```{r}
-  predictions <- stats::predict(m, newdata = grid_locs)$data
-# predictions$resids <- residuals(m) # randomized quantile residuals but not working
+predictions <- predict(m, newdata = grid_locs)$data
 glimpse(predictions) # est = predicted values
+
+# Make predictions on original data:
+predictions_original <- predict(m)$data
+predictions_original$resids <- residuals(m) # randomized quantile residuals
+glimpse(predictions_original) # est = predicted values
 
@@
 ```{r}
-#ggplot(predictions, aes(X, Y, col = resids)) + scale_colour_gradient2() +
-#  geom_point() + facet_wrap(~year)
+ggplot(predictions_original, aes(X, Y, col = resids)) + scale_colour_gradient2() +
+ geom_point() + facet_wrap(~year)
 
-#hist(predictions$resids)
-#qqnorm(predictions$resids);abline(a = 0, b = 1)
+hist(predictions_original$resids)
+qqnorm(predictions_original$resids);abline(a = 0, b = 1)
 ```

### 2019-01-22 analysis/manual-geostat-TMB.Rmd
Move 'manual-geostas-TMB.Rmd'

--- 2019-01-22 manual-geostat-TMB.Rmd
@@
 # d <- dplyr::filter(d, !(year == 2014 & survey_abbrev == "SYN WCHG")) # something went wrong in 2014
-
 ```
@@
 ```{r}
-
 col <- if (grepl("SYN", survey)) "density_kgpm2" else "density_ppkm2" # grep says if SYN is in survey returns logical (l), use kgpm2
@@
 ```{r}
-#  if (grepl("SYN", survey)) {
 grid_locs <- gfplot:::make_prediction_grid(
@@
 grid_locs <- dplyr::rename(grid_locs, depth = akima_depth)
-#  } else {
-#    grid_locs <- if (surv == "HBLL OUT N") gfplot::hbll_n_grid$grid else gfplot::hbll_s_grid$grid
- # }
 grid_locs$year <- NULL
-
-#  formula <- if (include_depth) {
-#    stats::as.formula(density ~ 0 + as.factor(year))
-#  } else {
-#    stats::as.formula(density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2)
-# }
 ```
@@
 ```{r}
-  spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 100)
+spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 50)
 
@@
 
-#### Save preditions from model
+#### Save predictions from model
 ```{r}

### 2019-01-23 analysis/manual-geostat-TMB.Rmd
Increase knots
knots=100

@@
 ```{r}
-spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 50)
+spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 100)
 
@@
   scale_fill_gradient2()
-
 ```

### 2019-02-01 analysis/binom-sdmTMB.Rmd
Bolker and TMB notes

---
title: "Binomial TMB"
output: html_notebook
---

```{r}
d <- subset(pcod, year >= 2011) # subset for example speed
pcod_spde <- make_spde(d$X, d$Y, n_knots = 100) # only 100 knots for example speed
plot_spde(pcod_spde)

# Tweedie:
m <- sdmTMB(
  d, density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
  time = "year", spde = pcod_spde, family = tweedie(link = "log"),
  silent = FALSE)

# Contents of the output object:
names(m)
m$model
TMB::sdreport(m$tmb_obj)
r <- m$tmb_obj$report()
names(r)

# Binomial:
pcod_binom <- d
pcod_binom$present <- ifelse(pcod_binom$density > 0, 1L, 0L)
m_bin <- sdmTMB(pcod_binom,
  present ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
  time = "year", spde = pcod_spde, family = binomial(link = "logit"))

# Gaussian:
pcod_gaus <- subset(d, density > 0 & year >= 2013)
pcod_spde_gaus <- make_spde(pcod_gaus$X, pcod_gaus$Y, n_knots = 50)
m_pos <- sdmTMB(pcod_gaus,
  log(density) ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
  time = "year", spde = pcod_spde_gaus)

## Not run: 
# Stan sampling (warning: slow going and priors are flat).

# Must load tmbstan first and then TMB and/or sdmTMB
# or you will get the error `"is_Null_NS" not resolved from current
# namespace (rstan)`
# Restart R session, then:
library(tmbstan)

# Then:
library(sdmTMB)

# Then:
set.seed(42)
pcod_pos <- subset(pcod, year > 2013 & density > 0)
pcod_pos_spde <- make_spde(pcod_pos$X/10, pcod_pos$Y/10, n_knots = 200) # scale UTMs for #' Stan
m <- sdmTMB(pcod_pos,
 log(density) ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
 time = "year", spde = pcod_pos_spde)
m_stan <- tmbstan(m$tmb_obj, chains = 1, iter = 200, cores=1,
  init = "last.par.best", control = list(adapt_delta = 0.80, max_treedepth = 20),
  seed = 123, laplace = T)

pars <- c('b_j', 'ln_tau_O', 'ln_tau_E', 'ln_kappa', 'ln_phi')
m_stan2 <- tmbstan(m$tmb_obj, chains = 1, iter = 200, cores=1,
  init = "last.par.best", control = list(adapt_delta = 0.80, max_treedepth = 20),
  seed = 123, laplace = F, pars = pars)

m_stan

```

### 2019-02-01 analysis/resources/Bolker-notes.Rmd
Bolker and TMB notes

---
title: "Bolker code notes"
output: html_notebook
---

```{r setup}

```

```{r}

curve(2 * x/(1 + x), from = 0, to = 8, ylim = c(0, 10))

micmen <- function(x, a = 2, b = 1) { 
  a * x/(b + x)
}

curve(micmen(x))
```

```{r}
curve(micmen(x), from = 0, to = 8, ylim = c(0, 10))
curve(micmen(x, b = 3), add = TRUE, col = 2)
curve(micmen(x, a = 8), add = TRUE, col = 3)
abline(h = 8)
```

```{r}
xvec <- seq(0, 10, by = 0.1)

yvec=micmen(xvec)

#OR WITH SAPPLY

sapply (yvec=sapply(xvec,micmen))

#OR WITH FOR LOOP

yvec <- NA
for (i in 1:length(xvec)) { 
    yvec[i]=micmen(xvec[i])
  }

plot(xvec, yvec)
lines(xvec, yvec, add = TRUE)
```

# Chapter 5: Simulations

```{r}
x = 1:20 # x varies systematically

a=2 
b=1

# Calculate the deterministic part of the model
y_det = a + b * x

# Pick 20 random normal deviates with the mean equal to the deter- ministic equation and σ = 2
y = rnorm(20, mean = y_det, sd = 2)

#OR

y = y_det+rnorm(20,sd=2) # BUT only works for normal distributions

```

```{r}
x = runif(50, min = 0, max = 5) # simulating samples from a gradient in the predictor variable

a = 20 
b=1 
k=5

y_det = a * b/(b + x)

y = rnbinom(50, mu = y_det, size = k)

plot(x, y)
curve(a*b/(b+x), add =TRUE)

```

```{r}
g = factor(rep(1:2, each = 25))

a = c(20, 10)
b = c(1, 2) 
k=5

y_det = a[g]/(b[g] + x)

y = rnbinom(50, mu = y_det, size = k)

plot(x, y, col = g)
curve(a[1]/(b[1] + x), add=TRUE)
curve(a[2]/(b[2] + x), col='red', add=TRUE)
```

```{r}
N = 603
a = 0.696
b = 9.79
mu = 25.32
zprob = 0.123
k = 0.932
# Define a function for the recruitment probability:
recrprob <- function(S) { 
      a/(1 + (a/b) * S) 
  }
#Now simulate the number of settlers and the number of recruits, using rzinbinom from the emdbook package:
library(emdbook)
?rzinbinom
S = rzinbinom(N, mu = mu, size = k, zprob = zprob)
recr = rbinom(N, prob = recrprob(S), size = S)

plot(S, recr, xlim = c(0, 200), ylim = c(0, 20))
#curve(recrprob(S), add = TRUE) 
```

```{r}

set.seed(1001)
L = 30
dispdist = 2

nparents = 50
offspr_per_parent = 10
noffspr = nparents * offspr_per_parent

# If you wanted to allow different numbers of offspring for each parent — for example, drawn from a Poisson distribution — you could use offspr_per_parent=rpois(nparents,lambda) and then rep(..., times=offspr_per_parent). Instead of specifying that each parent’s coordinates should be repeated the same number of times, you would be telling R to repeat each parent’s coordinates according to its number of offspring.
# offspr_per_parent = rpois(nparents,lambda)
# rep(..., times=offspr_per_parent)

# Pick locations for the parents:
parent_x = runif(nparents, min = 0, max = L)
parent_y = runif(nparents, min = 0, max = L)

# Pick angles and distances for dispersal:
angle = runif(noffspr, min = 0, max = 2 * pi)
dist = rexp(noffspr, 1/dispdist)

# Add the offspring displacements to the parent coordinates (using rep(...,each=offspr_per_parent)):
offspr_x = rep(parent_x, each = offspr_per_parent) +
   cos(angle) * dist
offspr_y = rep(parent_y, each = offspr_per_parent) +
   sin(angle) * dist
```

Next we calculate the neighborhood density, or the number of individuals within 2 m of each plant (not counting itself). Figure 5.4(b) shows this distribution, along with a fitted negative binomial distribution. This calculation reduces the spatial pattern to a simpler non-spatial distribution of crowding.
```{r}
pos <- cbind(offspr_x, offspr_y)
ndist <- as.matrix(dist(pos, upper = TRUE, diag = TRUE))
nbrcrowd = apply(ndist < 2, 1, sum) - 1

hist(nbrcrowd, breaks=30)
```

Next we use a relationship that Pacala and Silander found between end-of-year mass (M) and competition index (C).

They fitted this relationship based on a competition index estimated as a function of the neighborhood density of conspecific (pigweed) and heterospecific (velvetleaf) competitors, C = 1 + cppnp + cvpnv. For this example, I simply made up a proportionality constant to match the observed range of competition indices. Pacala and Silander found that biomass M ∼ Gamma(shape = m/(1 + C), scale = α), with m = 2.3 and α = 0.49.

```{r}
ci = nbrcrowd * 3
M = 2.3
alpha = 0.49
mass_det = M/(1 + ci)
mass = rgamma(length(mass_det), scale = mass_det, shape = alpha)

plot(ci, log(mass))

```

Finally, we simulate seed set as a function of biomass, again using a relationship estimated by Pacala and Silander. 
Seed set is proportional to mass, with negative binomial errors: S ∼ NegBin(μ = bM,k), with b = 271.6, k = 0.569.

```{r}
b = 271.6
k = 0.569
seed_det = b * mass
seed = rnbinom(length(seed_det), mu = seed_det, size = k)

plot(log(mass),log(1+seed))
```

## Power analysis

```{r}
N = 20
x = runif(20, min = 0, max = 5)
a = 2 
b = 1
sd = 2

y_det = a + b * x

y = rnorm(N, mean = y_det, sd = sd)

m = lm(y ~ x)

coef(summary(m))["x", "Pr(>|t|)"]

nsim = 400
pval = numeric(nsim)

for (i in 1:nsim) {
    y_det = a + b * x
    y = rnorm(N, mean = y_det, sd = sd)
m = lm(y ~ x)
pval[i] = coef(summary(m))["x", "Pr(>|t|)"] 
}

sum(pval < 0.05)/nsim
```

```{r}
bvec = seq(-2, 2, by = 0.1)
power.b = numeric(length(bvec))
for (j in 1:length(bvec)) {
    b= bvec[j]
    for (i in 1:nsim) {
            y_det = a + b * x
            y = rnorm(N, mean = y_det, sd = sd)
            m = lm(y ~ x)
            pval[i] = coef(summary(m))["x", "Pr(>|t|)"]
    }
    power.b[j] = sum(pval < 0.05)/nsim
}

plot(bvec, power.b)
```

I modified code to simulate change with sample size
```{r}
a = 2 
b = 1
sd = 2

Nvec = seq(10, 50, by = 5)
power.N = numeric(length(Nvec))
for (j in 1:length(Nvec)) {
    N= Nvec[j]
    for (i in 1:nsim) {
            x = runif(N, min = 0, max = 5)      
            y_det = a + b * x
            y = rnorm(N, mean = y_det, sd = sd)
            m = lm(y ~ x)
            pval[i] = coef(summary(m))["x", "Pr(>|t|)"]
    }
    power.N[j] = sum(pval < 0.05)/nsim
}

plot(Nvec, power.N)
```

# Chapter 6: Likelihood and all that
## 6.2 PARAMETER ESTIMATION: SINGLE DISTRIBUTIONS
### 6.2.1 Maximum likelihood
#### R code for a binomial negative log-likelihood function:

```{r}
binomNLL1 = function(p, k, N) {
    -sum(dbinom(k, prob = p, size = N, log = TRUE)) 
}
```

The dbinom function calculates the binomial likelihood for a specified data set (vector of number of successes) k, probability p, and number of trials N; the log=TRUE option gives the log-probability instead of the probability (more accurately than taking the log of the product of the probabilities); -sum adds the log-likelihoods and changes the sign to get an overall negative log-likelihood for the data set.
Load the data and extract the subset we plan to work with:

```{r}
data(ReedfrogPred)
x = subset(ReedfrogPred, pred == "pred" & density ==
    10 & size == "small")
k = x$surv
x
```

We can use the optim function to numerically optimize (by default, minimizing rather than maximizing) this function. You need to give optim the objective function — the function you want to minimize (binomNLL1 in this case) — and a vector of starting parameters. You can also give it other information, such as a data set, to be passed on to the objective function. The starting parameters don’t have to be very accurate (if we had accurate estimates already we wouldn’t need optim), but they do have to be reasonable. That’s why we spent so much time in Chapters 3 and 4 on eyeballing curves and the method of moments.

```{r}
O1 = optim(fn = binomNLL1, par = c(p = 0.5), N = 10,
    k = k, method = "BFGS")

O1$par
exp(-O1$value)
```

optim insists that you specify all of the parameters packed into a single numeric vector in your negative log-likelihood function. mle prefers the parameters as a list. mle2 will accept either

mle2 assumes that the objective function is a negative log-likelihood function. The names of the arguments are easier to understand: minuslogl instead of fn for the negative log-likelihood function, start instead of par for the starting parameters, and data for additional parameters and data.

```{r}
library(bbmle)
m1 = mle2(minuslogl = binomNLL1, start = list(p = 0.5),
    data = list(N = 10, k = k))
m1

m2 = mle2(k ~ dbinom(prob = p, size = 10), start = list(p = 0.5), data = list(k = x$surv))
m2
```

## Gamma
#### 6.2.1.2 Myxomatosis data: Gamma likelihood

```{r}
data(MyxoTiter_sum)
myxdat = subset(MyxoTiter_sum, grade == 1)

gammaNLL1 = function(shape, scale) {
    -sum(dgamma(myxdat$titer, shape = shape, scale = scale, log = TRUE))
}

```

find starting parameters for the Gamma distribution: 
use the method of moments (Chapter 4) to determine reasonable starTing values for the scale (=variance/mean=coefficient of variation [CV]) and shape(=variance/mean2=mean/CV) parameters
```{r}
gm = mean(myxdat$titer)
cv = var(myxdat$titer)/mean(myxdat$titer)
```

fit the data:
```{r}
m3 = mle2(gammaNLL1, start = list(shape = gm/cv,
   scale = cv))
m3
```

```{r}
m3 = mle2(myxdat$titer ~ dgamma(shape, scale = scale),
          start = list(shape = gm/cv, scale = cv), 
          data = list(myxdat$titer))
m3
```

```{r}
f1 = MASS::fitdistr(myxdat$titer, "gamma")
f1
```

#### 6.2.2.2 Gamma distribution: multiparameter distributions and non-conjugate priors

```{r}
prior.as = function(a, s) {
  dgamma(a, shape = 0.01, scale = 100) * dgamma(s, shape = 0.1, scale = 10)
}

unscaled.posterior = function(a, s) {
  prior.as(a, s) * exp(-gammaNLL1(shape = a, scale = s)) 
}
```

## 6.3 MORE COMPLEX FUNCTIONS
### 6.3.1 Maximum likelihood
#### 6.3.1.1 Tadpole predation
Since the distribution and density functions in R (such as dbinom) operate on vectors just as do the random-deviate functions (such as rbinom) used in Chapter 5, I can translate this model definition directly into R, using a numeric vector p={a, s} for the parameters:
```{r}
binomNLL2 = function(p, N, k) {
    a = p[1]
    h = p[2]
    predprob = a/(1 + a * h * N)
    -sum(dbinom(k, prob = predprob, size = N, log = TRUE))
}
```

eyeballing the initial slope and asymptote gives us crude starting estimates of a (initial slope) at around 0.5 and h (1/asymptote) at around 1/80 = 0.0125

```{r}
data(ReedfrogFuncresp)
attach(ReedfrogFuncresp)
O2 = optim(fn = binomNLL2, par = c(a = 0.5, h = 0.0125),
    N = Initial, k = Killed)
```

mle2 for this purpose, you would normally have to rewrite the negative log-likelihood function with the parameters a and h as separate arguments (i.e. function(a,h,p,N,k)). However, mle2 will let you pass the parameters inside a vector as long as you use parnames to attach the names of the parameters to the function.
```{r}
parnames(binomNLL2) = c("a", "h")
m2 = mle2(binomNLL2, start = c(a = 0.5, h = 0.0125),
   data = list(N = Initial, k = Killed))
m2
```

#### 6.3.1.2 Myxomatosis virus
```{r}
gammaNLL2 = function(a, b, shape) {
  meantiter = a * myxdat$day * exp(-b * myxdat$day)
  -sum(dgamma(myxdat$titer, shape = shape, scale = meantiter/shape, log = TRUE))
}
```

```{r}
m4 = mle2(gammaNLL2, start = list(a = 1, b = 0.2,
     shape = 50), method = "Nelder-Mead")
m4
```

```{r}
mle2(titer ~ dgamma(shape, scale = a * day * exp(-b *
     day)/shape), start = list(a = 1, b = 0.2, shape = 50),
     data = myxdat, method = "Nelder-Mead")

```

### 6.4.1.1 The Likelihood Ratio Test

R can compute profiles and profile confidence limits automatically.

Given an mle2 fit m, profile(m) will compute a likelihood profile and confint(m) will compute profile confidence limits. plot(profile(m2)) will plot the profile, square-root transformed so that a quadratic profile will appear V-shaped (or linear if you specify absVal=FALSE). This transforma- tion makes it easier to see whether the profile is quadratic, since it’s easier to see whether a line is straight than it is to see whether it’s quadratic. Computing the profile can be slow, so if you want to plot the profile and find confidence limits, or find several different confidence limits, you can save the profile and then use confint on the profile:

```{r}
profile(m2)
confint(m2)
plot(profile(m2), absVal=FALSE)
p2 = profile(m2)
confint(p2)
```

```{r}
binomNLL2.a = function(p, N, k, a) {
  h = p[1]
 p = a/(1 + a * h * N)
 -sum(dbinom(k, prob = p, size = N, log = TRUE))
}

```

```{r}
avec = seq(0.3, 0.8, length = 100)
aprof = numeric(100)
for (i in 1:100) {
  aprof[i] = optim(binomNLL2.a, par = 0.02, k = ReedfrogFuncresp$Killed,
    N = ReedfrogFuncresp$Initial, a = avec[i],
    method = "BFGS")$value
}

plot(avec,aprof)
```

```{r}
prof.lower = aprof[1:which.min(aprof)]
prof.avec = avec[1:which.min(aprof)]

approx(prof.lower, prof.avec, xout = -logLik(m2) +
     qchisq(0.95, 1)/2)
```

### 2019-02-11 analysis/src/simpletest.cpp
simpletest -- first attempt to write TMB code

@@ -1,0 +1,14 @@
+// Simple linear regression.
+#include <TMB.hpp>
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  DATA_VECTOR(y);
+  DATA_VECTOR(x);
+  PARAMETER(a);
+  PARAMETER(b);
+  PARAMETER(logSigma);
+  // ADREPORT(exp(2*logSigma));
+  Type nll = -sum(dnorm(y, a+b*x, exp(logSigma), true));
+  return nll;
+}

### 2019-02-12 analysis/INLA/INLA-sole-example.Rmd
INLA sole example - change in effect of a through t

---
title: "Sole Example from Wang et al. 2018"
output: html_notebook
---
```{r install, eval=FALSE}
library(devtools)
install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
install_github("julianfaraway/brinla")
install.packages('gamair')
```

# Sole eggs in the Bristol Channel
## Selected code from Wang et al. 2018 example

sole eggs per square meter of sea surface in each of 4 identifiable egg developmental stages, at each of a number of sampling stations in the Bristol Channel on the west coast of England. The samples were taken from 5 research cruises over the spawning season.
Let’s load the dataset and look at its structure:

```{r setup}
library(INLA)
library(brinla)
library(gamair)
data(sole, package = 'gamair')
```

```{r}
str(sole)
# sole$stage
```

la and lo are the latitude and longitude of sampling station,
t is the time of midpoint of the cruise on which this sample was taken, 
eggs is the egg density per square meter of sea surface, 
stage is one of 4 stages the sample relates, 
a.0 is the age of the youngest possible egg in this sample, and a.1 is of the oldest.
Following Wood (2006), we calculate the width (off) and average age (a) of the corresponding egg class:

```{r}
solr <- sole
solr$off <- log(sole$a.1 - sole$a.0)
solr$a <- (sole$a.1 + sole$a.0)/2

boxplot(a ~ stage, data = solr)
points(a.0 ~ stage, col = "red", data = solr)
points(a.1 ~ stage, col = "red", data = solr)
```

```{r}
solr$t <- solr$t - mean(sole$t)
solr$t <- solr$t/var(sole$t)^0.5
solr$la <- solr$la - mean(sole$la)
solr$lo <- solr$lo - mean(sole$lo)

hist(solr$eggs, breaks = 30)
boxplot(eggs ~ t, data = solr)
boxplot(eggs ~ stage, data = solr)
```

at each station counts for the four different egg stages are all taken from the same net sample

```{r}
solr$station <- as.numeric(factor(with(solr, paste(-la, -lo, -t, sep=""))))
```

```{r}
solr$eggs <- solr$eggs*1000
solr$off <- solr$off + log(1000)
```

```{r}
length(which(solr$eggs == 0))/length(solr$eggs)
```

 zero-inflated negative binomial (ZINB) likelihood

eggsi ∼ ZINB(ρ,n,pi), μi =n(1−pi)/pi

log(μi) = log(offi) + β0 + ai + lai∗ti + lai∗t2i + loi∗ti + loi∗t2i + ai*f1(ti)+ f2(loi,lai) + stationj(i)

(9.8)

where ρ is the zero-probability parameter, n is the overdispersion parameter, pi is the
probability of “success” in ith trial, and μi is the mean of ith egg density.

```{r}
loc <- cbind(solr$lo, solr$la)
mesh <- inla.mesh.2d(loc, max.edge = c(0.1, 0.2))
node <- mesh$idx$loc
tps <- bri.tps.prior(mesh, constr = TRUE)
```

```{r}
formula1 <- eggs ~ a + I(la*t) + I(la*t^2) + I(lo*t) + I(lo*t^2) + f(t, a, model = 'rw2', scale.model = TRUE) + f(node, model = tps, diagonal = 1e-6) + f(station, model = 'iid')
```

```{r}
result1 <- inla(formula1, 
                family = 'zeroinflatednbinomial0', 
                offset = off, 
                data = solr, 
                control.predictor = list(compute = TRUE), 
                control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE))

```

```{r}
result1 <- result1
round(result1$summary.fixed, 4)
```

```{r}
bri.band.ggplot(result1, name = 't', type = 'random')
```

```{r}
bri.fixed.plot(result1)
```

```{r}
plot(result1, single = FALSE)
```

```{r, warning=FALSE}
proj <- inla.mesh.projector(mesh)
spa.mean <- inla.mesh.project(proj, result1$summary.random$node$mean)
library(fields)
image.plot(proj$x, proj$y, spa.mean, xlim=c(-1.5, 2), ylim=c(-1, 1))
points(loc, pch = 19, cex = 0.2)
```

The posterior distributions of the hyperparameters are summarized using:

```{r}
tmp <- bri.hyperpar.summary(result1)
row.names(tmp) <- c("Overdispersion", "Zero-probability", 'SD for t',
'Theta1 for node', 'SD for station') 
round(tmp, 4)
```

```{r}
length(which(result1$cpo$failure > 0))
```

# My modifications:

## What about getting effect of a to change through time?

```{r}
solr$t.age <- solr$t # duplicates t for the a slope change through time

formula.t.a <- eggs ~ I(a) + I(a^2) + I(la*t) + I(la*t^2) + I(lo*t) + I(lo*t^2) + f(t, model = 'rw2', scale.model = TRUE) + f(t.age, a, model = 'rw2', scale.model = TRUE) + f(node, model = tps, diagonal = 1e-6) + f(station, model = 'iid')

# without quadratic fixed effect of age
formula.t.a.nq <- eggs ~ I(a) + I(la*t) + I(la*t^2) + I(lo*t) + I(lo*t^2) + f(t, model = 'rw2', scale.model = TRUE) + f(t.age, a, model = 'rw2', scale.model = TRUE) + f(node, model = tps, diagonal = 1e-6) + f(station, model = 'iid')

# without age varying with time
formula.t <- eggs ~ I(a) + I(a^2) + I(la*t) + I(la*t^2) + I(lo*t) + I(lo*t^2) + f(t, model = 'rw2', scale.model = TRUE) + f(node, model = tps, diagonal = 1e-6) + f(station, model = 'iid')
```

```{r}
result.t.a <- inla(formula.t.a, 
                family = 'zeroinflatednbinomial0', 
                offset = off, 
                data = solr, 
                control.predictor = list(compute = TRUE), 
                control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config=TRUE))
```

```{r}
result.t.a.nq <- inla(formula.t.a.nq, 
                family = 'zeroinflatednbinomial0', 
                offset = off, 
                data = solr, 
                control.predictor = list(compute = TRUE), 
                control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config=TRUE))
```

```{r}
result.t <- inla(formula.t, 
                family = 'zeroinflatednbinomial0', 
                offset = off, 
                data = solr, 
                control.predictor = list(compute = TRUE), 
                control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config=TRUE))
```

```{r}
c(result.t.a$dic$dic, result.t.a.nq$dic$dic, result.t$dic$dic, result1$dic$dic)
```

```{r}
result <- result.t.a.nq
round(result$summary.fixed, 4)
#round(result$summary.fixed[, c("mean", "0.025quant", "0.975quant")], 4)
```

```{r}
summary(result)$hyperpar[, c("mean", "0.025quant", "0.975quant")]
```

```{r}
bri.band.ggplot(result, name = 't', type = 'random')
```

```{r}
?bri.band.ggplot
bri.band.ggplot(result, name = 't.age', type = 'random')
```

```{r}
bri.fixed.plot(result)
```

```{r, warning=FALSE}
proj <- inla.mesh.projector(mesh)
spa.mean <- inla.mesh.project(proj, result$summary.random$node$mean)
library(fields)
image.plot(proj$x, proj$y, spa.mean, xlim=c(-1.5, 2), ylim=c(-1, 1))
points(loc, pch = 19, cex = 0.2)
```

```{r}
result$cpo$cpo  
result$cpo$failure
```

### 2019-02-19 analysis/src/simple.cpp
practice lm in tmb

@@ -1,42 +1,17 @@
 // Normal linear mixed model specified through sparse design matrices.
 #include <TMB.hpp>
-
 template<class Type>
 Type objective_function<Type>::operator() ()
 {
-  DATA_VECTOR(x);         // Observations
-  DATA_SPARSE_MATRIX(B);  // Random effect design matrix
-  DATA_SPARSE_MATRIX(A);  // Fixed effect design matrix
-  PARAMETER_VECTOR(u);    // Random effects vector
-  PARAMETER_VECTOR(beta); // Fixed effects vector
-  PARAMETER(logsdu);      // Random effect standard deviations
-  PARAMETER(logsd0);      // Measurement standard deviation
-
-  // Distribution of random effect (u):
-  Type ans = 0;
-  ans -= dnorm(u, Type(0), exp(logsdu), true).sum();
-
-  // Optionally: How to simulate the random effects
-  SIMULATE {
-    u = rnorm(u.size(), Type(0), exp(logsdu));
-    REPORT(u);
-  }
-
-  // Distribution of obs given random effects (x|u):
-  vector<Type> y = A * beta + B * u;
-  ans -= dnorm(x, y, exp(logsd0), true).sum();
-
-  // Optionally: How to simulate the data
-  SIMULATE {
-    x = rnorm(y, exp(logsd0));
-    REPORT(x);
-  }
-
-  // Apply delta method on sd0:
-  ADREPORT( exp(logsd0) );
-
-  // Report posterior mode and mean of sum(exp(u))
-  ADREPORT( sum(exp(u)) );
-
-  return ans;
+  // data:
+  DATA_VECTOR(y);
+  DATA_VECTOR(x);
+  // parameters:
+  PARAMETER(a); // intercept
+  PARAMETER(b); // slope
+  PARAMETER(logSigma);
+  // procedures: (transformed parameters)
+  // Type sigma = exp(logSigma);
+  Type nll = -sum(dnorm(y, a+b*x, exp(logSigma), true));
+  return nll;
 }

### 2019-02-19 analysis/tweedie-TMB.Rmd
attemp tweedie in TMB - not compiling yet

---
title: "Tweedie in TMB"
output: html_notebook
---
Started using tmb.template

```{r set-up}
## Load TMB TRUE
library(TMB)
library(tweedie)

## Make C++ file
# TMB::template("tweedie_tmb.cpp")
## Compile and load the model
compile("tweedie_tmb.cpp")
dyn.load(dynlib("tweedie_tmb"))
```

Simulate data

```{r}
N <- 100
x <- runif(N, -2, 2)
a <- 3
b <- 1

power <- 1.5 # must be between 1 and 2 for compound poisson-gamma
phi <- 1 # dispersion must be positive
eta <- a + b * x 
mu <- exp(eta) 
y <- rtweedie(N, power = power, mu = mu, phi = phi)

# plot simulated data 
plot(x, y)
lines(sort(x), mu[order(x)])

hist(y, breaks=30)
```

```{r}
## Data and parameters
data <- list(x = x, y = y)
parameters <- list(a=0, b=0, power = 1.5, log_phi = log(1) )

## Make a function object
obj <- MakeADFun(data, parameters, DLL="tweedie_tmb")

## Call function minimizer
opt <- nlminb(obj$par, obj$fn, obj$gr)

## Get parameter uncertainties and convergence diagnostics
sdr <- sdreport(obj)
sdr

```

# check against values are returned with glm
```{r}
tweedie_glm <- glm(y ~ x, family = tweedie(var.power = 1.5, link.power = 0))
summary(tweedie_glm)
coef(tweedie_glm)
```

### 2019-02-20 analysis/tweedie-TMB.Rmd
tweedie in TMB that runs!

@@
 ## Data and parameters
-data <- list(x = x, y = y)
-parameters <- list(a=0, b=0, power = 1.5, log_phi = log(1) )
+data <- list(y = y, x = x)
+parameters <- list(a=0, b=0, log_phi = log(1), p = 0 )
 
@@
 sdr
-
 ```
 
-# check against values are returned with glm
+# find maximum likelihood using nlminb
+
+```{r}
+# function to generate negative log-likelihood for a parameter value
+nll_tweedie <- function(par) {
+  eta <- par[1] + par[2] * x
+  mu <- exp(eta)
+  p <- plogis(par[3]) + 1
+  phi <- exp(par[4])
+  # sum of negative log likelihoods:
+  loglik <- log(dtweedie(y, power = p, mu = mu, phi = phi))
+  nll <- -sum(loglik) #[is.finite(loglik)]
+  print(nll)
+  nll
+}
+
+tweedie_ml <- nlminb(start = c(a = 0, b = 0, ipx = 0, log_phi = log(1)), nll_tweedie)
+tweedie_ml$par 
+```
+
+# check both methods against values are returned with glm
 ```{r}
@@
 coef(tweedie_glm)
+sdr
+tweedie_ml$par 
 ```

### 2019-02-20 analysis/tweedie_tmb.cpp

@@ -11,18 +11,19 @@ Type objective_function<Type>::operator() ()
   PARAMETER(log_phi);
   PARAMETER(p);
   
-  // Objective function is sum of negative log likelihood components
-  Type power = invlogit(p) + Type(1.0);
-   
-  // ------------------ Linear predictor----------------------------------------
-  
-  vector<Type> eta = a + b * x;
-  vector<Type> mu = exp(eta);
-  
   // ------------------ negative log likelihood --------------------------
-  Type nll = 0;  // likelihood of data
-  Type nll -= sum(dtweedie(y, mu, exp(log_phi), power, true));
-
+  Type power = invlogit(p) + Type(1.0);
+  Type phi = exp(log_phi);
+  
+  int n = y.size(); // get number of data points to loop over
+  
+  Type nll = 0.0; // initialize negative log likelihood
+  
+  for(int i = 0; i < n; i++){ // C++ starts loops at 0!
+    // get negative log likelihood (last argument is log = TRUE)
+    nll -= dtweedie(y[i], exp(a + b * x[i]), phi, power, true);
+  }
+  
   return nll;
 
   /* Quick Reference

### 2019-02-20 analysis/tweedie-TMB.Rmd
Add profile examples

@@
 sdr
+
+## Get likelihood profiles for individual parameters:
+a_profile <- tmbprofile(obj, "a")
+b_profile <- tmbprofile(obj, "b")
+
+plot(a_profile)
+plot(b_profile)
+confint(a_profile)
+confint(b_profile)
 ```

### 2019-02-20 analysis/tweedie-TMB.Rmd
tidy tweedie TMB

--- 2019-02-20 analysis/tweedie-TMB.Rmd
@@
 ## Compile and load the model
-compile("tweedie_tmb.cpp")
-dyn.load(dynlib("tweedie_tmb"))
+compile("src/tweedie_tmb.cpp")
+dyn.load(dynlib("src/tweedie_tmb"))
 ```
 
-Simulate data
+# Simulate data
 
@@
 
+# Send data to TMB function to get likelihood profiles for parameters
+
 ```{r}
@@
 sdr
+
+## Get likelihood profiles for individual parameters:
+a_profile <- tmbprofile(obj, "a")
+b_profile <- tmbprofile(obj, "b")
+confint(a_profile)
+confint(b_profile)
+
+plot(a_profile)
+plot(b_profile)
+
 ```
 
-# find maximum likelihood using nlminb
+# For comparision, find maximum likelihood parameter estimates using nlminb
 
@@
 
-# check both methods against values are returned with glm
+# Check both methods against values are returned with a glm
+
 ```{r}
@@
 coef(tweedie_glm)
-sdr
+sdr$par.fixed
 tweedie_ml$par 

### 2019-02-20 analysis/m_tweedie_TMB.Rmd
mixed effect tweedie in TMB

---
title: "Mixed effect tweedie model in TMB"
output: html_notebook
---

```{r}
# use tmb.template snippet
```

```{r set-up}
## Load TMB 
library(TMB)
library(tweedie)

#library(statmod)

# Make C++ file
# TMB::template("m_tweedie_tmb.cpp")

## Compile and load the model
compile("src/m_tweedie_tmb.cpp")
dyn.load(dynlib("src/m_tweedie_tmb"))
```

# Simulate data with a tweedie distribution and mixed-effects...

```{r}
# Simulate predictors

n_groups = 10
n_within = 10
n_total = n_groups*n_within

group_sd = 1

group <- rep( 1:n_groups, each = n_within)
group_beta <- rnorm(length(unique(group)), mean = 0, sd = group_sd)

x <- runif(n_total, -2, 2)
a <- 3
b <- 1

power <- 1.5 # must be between 1 and 2 for compound poisson-gamma
phi <- 1 # dispersion must be positive
eta <- a + group_beta[group] + b * x 
mu <- exp(eta)

# Simulate response
y = rtweedie(n_total, power = power, mu = mu, phi = phi)

d = data.frame(y, x, group)

# plot simulated data 
plot(x, y)
hist(y, breaks=30)
```

# plot data by group to see if it looks right
```{r}
library(ggplot2)
library(viridis)

ggplot(d, aes(x, y, group = as.factor(group), color = as.factor(group))) + geom_point() + geom_smooth(method = 'loess', formula = 'y ~ x', span = 3, se = FALSE) + scale_color_viridis_d()
```

```{r}
## Data and parameters
data <- list("g_i" = group-1, "y_i" = y, "x" = x)
parameters <- list("a" = 0, "b" = 0, "log_phi" = log(1), "p" = 0, "log_SDZ"= 2, "z_g" = rep(0, length(unique(group))))
random = c("z_g")
#if(REML==TRUE ) Random = union( Random, "a")

## Make a function object
obj <- MakeADFun(data, parameters, random = random, DLL="m_tweedie_tmb")

## Call function minimizer
opt <- nlminb(obj$par, obj$fn, obj$gr)

## Get parameter uncertainties and convergence diagnostics
sdr <- sdreport(obj)
sdr
```

```{r}
## Get likelihood profiles for individual parameters:
a_profile <- tmbprofile(obj, "a")
b_profile <- tmbprofile(obj, "b")
confint(a_profile)
confint(b_profile)

plot(a_profile)
plot(b_profile)
```

### 2019-02-20 analysis/src/m_tweedie_tmb.cpp

@@ -1,0 +1,53 @@
+#include <TMB.hpp>
+
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  // ------------------ Data ---------------------------------------------
+  DATA_IVECTOR(g_i);
+  DATA_VECTOR(y_i);
+  DATA_VECTOR(x);
+  
+  
+  // ------------------ Parameters ---------------------------------------------
+  PARAMETER( a ); // intercept
+  PARAMETER( b ); // slope
+  PARAMETER( p ); // tweedie power
+  PARAMETER( log_phi ); // dispersion / sigma
+  
+  PARAMETER( log_SDZ ); // omega / random 
+  PARAMETER_VECTOR( z_g );
+  
+
+  
+  // ------------------ negative log likelihood --------------------------
+   
+  Type power = invlogit(p) + Type(1.0);
+  Type phi = exp(log_phi);
+  Type SDZ = exp(log_SDZ);
+  
+  int n = y_i.size(); // get number of data points to loop over
+  
+  Type nll = 0.0;  // initialize negative log likelihood
+  
+  // Probability of data conditional on fixed and random effect values
+  for( int i = 0; i < n; i++){
+    nll -= dtweedie(y_i[i], exp(a + z_g(g_i[i]) + b * x[i]), phi, power, true );
+  }
+
+  // Probability of random coefficients
+  for( int g=0; g < z_g.size(); g++){
+    nll -= dnorm( z_g[g], Type(0.0), SDZ, true );
+  }
+  
+  // Reporting
+  REPORT( SDZ );
+  REPORT(power);
+  REPORT(phi);
+  
+  ADREPORT( SDZ );
+  ADREPORT( power );
+  ADREPORT( phi );
+  
+  return nll;
+}

### 2019-03-08 analysis/INLA/test-INLA-sensor.Rmd
start working with trawl sensor data

---
title: "R Notebook"
output: html_notebook
---

```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 
library(gfplot)
library(lattice)
library(INLA)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(mapdata)
library(lubridate)
```

```{r}
ssids <- get_ssids()
ssids
```

```{r}
raw <- readRDS("data-cache/pacific-cod.rds")
#dat$survey_abbrev
glimpse(raw)
```

```{r}
testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2015, 2016, 2017, 2018))
#testdat <- gfplot:::interp_survey_bathymetry(testdat)
glimpse(testdat)
View(testdat[[1]])
```

```{r}
# get_ssids()
sensor_syn <- get_sensor_data_trawl(ssid = c(1, 3, 4, 16), spread_attributes = TRUE)
# View(sensor_syn)
saveRDS(sensor_syn, "data-cache/allsynoptic_sensor_07mar19_time2.rds")
# write.csv(sensor, "synoptic_3_sensor_data_07mar19.csv")
# sensor <- read_csv("data-cache/synoptic_3_sensor_data_07mar19.csv", col_types = cols(do_mlpL_avg = col_double(), 
#                     do_mlpL_max = col_double(), do_mlpL_min = col_double(), salinity_PSU_avg = col_double(), salinity_PSU_max = col_double(), 
#                     salinity_PSU_min = col_double()), na = "NA")
sensor <- readRDS("data-cache/allsynoptic_sensor_07mar19_time2.rds")
View(sensor)
```

```{r}
dat <- left_join(testdat[[1]], sensor, by="fishing_event_id", suffix = c("", ".sensor"))
dat$year_f <- factor(dat$year)

View(dat)

dat <- dat %>% mutate(do_range = do_mlpL_max - do_mlpL_min, salinity_range = salinity_PSU_max -
  salinity_PSU_min, temp_range = temperature_C_max - temperature_C_min, depth_range = depth_m_max - depth_m_min)
glimpse(dat)
```

```{r}
plot(depth ~ depth_m_avg, data = dat) # both depth variables seem to agree

plot(temp_range ~ depth_range, data = dat) 
plot(do_range ~ depth_range, data = dat) 
plot(salinity_range ~ depth_range, data = dat)

plot(temperature_C_avg ~ temp_range, data = dat) #variance in temp estimate is highest for middle values
plot(do_mlpL_avg ~ do_range, data = dat) # one outlier at range = 8, others < 4
plot(do_range ~ depth, data = dat) # but outlier is at middle depth
plot(salinity_PSU_avg ~ salinity_range, data = dat) # 5 extreme values range > 10 and abnormally low mean salinity 
plot(salinity_PSU_avg ~ depth, cex = salinity_range/3, ylim=c(25,35), data = dat)
plot(salinity_PSU_max ~ depth, cex = salinity_range/3, ylim=c(25,35), data = dat)
plot(do_mlpL_avg ~ depth, cex = do_range, data = dat) 
plot(temperature_C_avg ~ depth, cex = temp_range, data = dat)

plot(do_mlpL_avg ~ temperature_C_avg, cex = depth/100, data = dat)
plot(do_mlpL_avg ~ salinity_PSU_avg, cex = depth/100, data = dat)

plot(-log(depth) ~ do_mlpL_avg,  cex = depth_range/35, data = dat) 
plot(-log(depth) ~ temperature_C_avg,  cex = depth_range/35, data = dat)
plot(-log(depth) ~ salinity_PSU_max,  cex = depth_range/35, data = dat)

```

```{r}
dat2 <- dat %>% filter(temperature_C_avg!="NA")

ggplot(dat2, aes(x=depth, y=temperature_C_avg)) + 
  geom_point(alpha=0.25) + 
  geom_point(data=subset(dat2, temp_range>1), alpha=0.5) + 
  geom_errorbar(data=subset(dat2, temp_range>1),aes(ymin=temperature_C_min, ymax=temperature_C_max),alpha= 0.75, colour="blue") + 
  geom_errorbarh(data=subset(dat2, temp_range>1),aes(xmin=depth_m_min, xmax=depth_m_max),alpha= 0.75, colour="red") + 
  ylim(2,12) + 
  facet_wrap(~survey_desc, scales = "free")
```

```{r}
dat2 <- dat %>% filter(do_mlpL_avg!="NA")

ggplot(dat2, aes(x=depth, y=do_mlpL_avg)) + 
  geom_point(alpha=0.25) + 
  geom_point(data=subset(dat2, do_range>1), alpha=0.5) + 
  geom_errorbar(data=subset(dat2, do_range>1), aes(ymin=do_mlpL_min, ymax=do_mlpL_max), alpha= 0.75, colour="blue") + 
  geom_errorbarh(data=subset(dat2, do_range>1), aes(xmin=depth_m_min, xmax=depth_m_max), alpha= 0.75, colour="red") + 
  ylim(0,10) + 
  facet_wrap(~survey_desc, scales = "free")
```

```{r}
dat2 <- dat %>% filter(salinity_PSU_avg!="NA")

ggplot(dat2, aes(x=depth, y=salinity_PSU_avg)) + 
  geom_point(alpha=0.25) + 
  geom_point(data=subset(dat2, salinity_range>1), alpha=0.5) + 
  geom_errorbar(data=subset(dat2, salinity_range>1), aes(ymin=salinity_PSU_min, ymax=salinity_PSU_max), alpha= 0.75, colour="blue") + 
  geom_errorbarh(data=subset(dat2, salinity_range>1), aes(xmin=depth_m_min, xmax=depth_m_max), alpha= 0.75, colour="red") + 
  #ylim(0,10) + 
  facet_wrap(~survey_desc, scales = "free")
```

```{r}
out <- dat %>% filter(depth_range>80)
out

pair <- dat %>% filter(ssid=="1") %>% select(., depth_range, temp_range, do_range, salinity_range, depth_m_avg,do_mlpL_avg,temperature_C_avg,salinity_PSU_max)
pairs(pair)  
```

Scale predictors by modifying code from gfplot to include environmental variables...
Changes should be added to gfplot?
```{r}
scale_survey_predictors <- function(dat) {
  if (sum(is.na(dat$depth)) > 0) {
    dat$depth[is.na(dat$depth)] <- dat$akima_depth[is.na(dat$depth)]
  }
  mutate(dat,
    # depth
    depth_mean = mean(log(depth), na.rm = TRUE),
    depth_c = (log(depth) - depth_mean[1]),
    depth_sd = sd(log(depth), na.rm = TRUE),
    depth_scaled = (log(depth) - depth_mean[1]) / depth_sd[1],
    depth_scaled2 = depth_scaled^2,
    # temperature
    temp_mean = mean(log(temperature_C_avg), na.rm = TRUE),
    temp_c = (log(temperature_C_avg) - temp_mean[1]),
    temp_sd = sd(log(temperature_C_avg), na.rm = TRUE),
    temp_scaled = (log(temperature_C_avg) - temp_mean[1]) / temp_sd[1],
    temp_scaled2 = temp_scaled^2,
    # dissolved O2
    do_mean = mean(log(do_mlpL_avg), na.rm = TRUE),
    do_c = (log(do_mlpL_avg) - do_mean[1]),
    do_sd = sd(log(do_mlpL_avg), na.rm = TRUE),
    do_scaled = (log(do_mlpL_avg) - do_mean[1]) / do_sd[1],
    do_scaled2 = do_scaled^2,
    # salinity
    salinity_mean = mean(log(salinity_PSU_avg), na.rm = TRUE),
    salinity_c = (log(salinity_PSU_avg) - salinity_mean[1]),
    salinity_sd = sd(log(salinity_PSU_avg), na.rm = TRUE),
    salinity_scaled = (log(salinity_PSU_avg) - salinity_mean[1]) / salinity_sd[1],
    salinity_scaled2 = salinity_scaled^2,
    # to put spatial decay parameter on right scale
    X10 = X / 10, Y10 = Y / 10 
  )
}
```

```{r}
scaled <- scale_survey_predictors(dat)
# glimpse(scaled)
synhs <- scaled %>% filter(do_mlpL_avg!="NA") %>% filter (density!="NA") 
glimpse(synhs)
```

```{r}
ggplot(synhs, aes(depth)) + geom_histogram() 
ggplot(synhs, aes(density*1000)) + geom_histogram() 
ggplot(synhs, aes(depth, (density*1000))) + geom_point() 
#ggplot(synhs, aes(depth_m, log(density*1000+1))) + geom_point() 
```

######################################
# Start Bayesian analysis
# One more time the model:

# Movement_ij      ~ Gamma(mu_ik, r)
# E(Movement_ij)   = mu_ij
# var(Movement_ij) = mu_ij^2 / r 
# mu_ij = exp(Intercept + fRepro + Annual Trend + Seasonal Trend + a_i)

# Below we define a model with a factor fRepro, a long term
# trend for Year, and seasonal trend for DayInYear, and
# a random intercept for bear.

# Each trend is modelled as a random walk trend.
# So each of them has a sigma.

```{r}

####
names(d) <- tolower(names(d))
d$species_common_name <- tolower(d$species_common_name)
d$species_science_name <- tolower(d$species_science_name)
d$year <- lubridate::year(d$trip_start_date)

sp <- filter(d, species_common_name %in% "pacific ocean perch") %>% 
  filter(!is.na(catch_weight)) %>% 
  filter(year %in% seq(2004, 2012, 2))

dat <- sp
dat <- filter(dat, start_lon > -128, start_lon < 125, start_lat > 48, start_lat < 50.3)
nrow(dat)

dat <- filter(dat, fe_bottom_water_temperature > 1, fe_bottom_water_temperature < 12,
  !is.na(fe_bottom_water_temp_depth), !is.na(fe_bottom_water_temperature))
nrow(dat)

dat <- select(dat, year, start_lon, start_lat, catch_weight, fe_bottom_water_temp_depth) %>% 
  rename(X = start_lon, Y = start_lat)

attr(dat, "projection") <- "LL"

dat <- PBSmapping::convUL(dat)
dat <- as.data.frame(na.omit(dat))

subcoords = cbind(dat$X, dat$Y)

bnd = inla.nonconvex.hull(subcoords, convex = 30)
mesh1 = inla.mesh.2d(
  boundary = bnd,
  max.edge = c(15, 40),
  cutoff = 5,
  offset = 1
)
plot(mesh1)
points(dat$X, dat$Y, col = "red")
summary(mesh1)

##########
A.data <- inla.spde.make.A(mesh1, loc = cbind(dat$X, dat$Y))

# Make SPDE based on mesh
spde = inla.spde2.matern(mesh1, alpha = 3 / 2)

n = nrow(dat)
YEARS <- unique(dat$year)
k = length(unique(dat$year))

# Make a design matrix where the first year is the intercept
# ...
dm <- matrix(nrow = nrow(dat), ncol = 2)
dm[,1] <- 1
dm[,2] <- dat$fe_bottom_water_temp_depth

iset <- inla.spde.make.index("i2D", n.spde = mesh1$n, n.group = k)

X.1 = dm
Covar.names <- c("intercept", "depth")
XX.list <- as.list(X.1)
effect.list <- list()
#   effect.list[[1]] <- c(iset, list(Intercept=1))
effect.list[[1]] <- c(iset)
for (Z in 1:ncol(X.1))
  effect.list[[Z + 1]] <- XX.list[[Z]]
names(effect.list) <- c("1", Covar.names)

### Make data stack.
A <-
  inla.spde.make.A(
    mesh = mesh1,
    loc = cbind(dat$X, dat$Y),
    group = dat$year
  )
A.list = list()
A.list[[1]] = A
for (Z in 1:ncol(X.1))
  A.list[[Z + 1]] <- 1

### Make projection points stack.
Ntrials <- rep(1, length(dat$Y))

sdat <-
  inla.stack(
    tag = 'stdata',
    data = list(
      y = dat$Y,
      link = 1,
      Ntrials = Ntrials
    ),
    A = A.list,
    effects = effect.list
  )

formula = as.formula(
  paste0(
    "y ~ -1 +",
    paste(Covar.names, collapse = "+"),
    "+ f(i2D, model=spde, group = i2D.group, control.group = list(model='ar1'))"
  )
)		# field evolves with AR1 by year

inlaModel <-
  inla(
    formula,
    family = "gamma",
    data = inla.stack.data(sdat),
    control.predictor = list(compute = TRUE, A = inla.stack.A(sdat)),
    verbose = TRUE,
    debug = TRUE,
    keep = FALSE,
    control.compute = list(dic = TRUE, cpo = TRUE, config = TRUE),
    control.fixed = list(correlation.matrix = TRUE),
    control.inla = list(lincomb.derived.correlation.matrix = TRUE)
  )
```

```{r}

synhs$year_f2 <- synhs$year_f
synhs$year_f3 <- synhs$year_f

f1 <- temperature_C2 ~ depth_scaled + depth_scaled2 + 
                 f(year_f, model = "rw2") +
                 f(year_f2, depth_m, model = "rw2") +
                 f(year_f3, depth2, model = "rw2")

I1 <- inla(f1, 
           control.predictor = list(compute = TRUE),
           family = "gamma",
           data = synhs
           )
summary(I1)

# We can access the fitted values:
Fit1 <- I1$summary.fitted.values[,"mean"]

# And using these fitted values we can
# calculate Pearson residuals. These can
# be used for model validation purposes.

# Let us plot the 2 trends.
# Here is how we access them:
Yearsm <- I1$summary.random$Year
Daysm  <- I1$summary.random$DayInYear

# The rest is simple plotting
#par(mfrow = c(1,2), mar = c(5,5,2,2), cex.lab = 1.5)
plot(Yearsm[,1:2], type='l',
     xlab = 'Year', 
     ylab = 'Smoother',
     ylim = c(-0.2, 0.2) )
abline(h=0, lty=3)
lines(Yearsm[, c(1, 4)], lty=2)
lines(Yearsm[, c(1, 6)], lty=2)
text(1988, 0.2, "A", cex = 1.5)

plot(Daysm[,1:2], type='l',
     xlab = 'DayInYear', 
     ylab = 'Smoother',
     ylim = c(-0.5, 0.6) )
abline(h=0, lty=3)
lines(Daysm[, c(1, 4)], lty=2)
lines(Daysm[, c(1, 6)], lty=2)
text(0, 0.6, "B", cex = 1.5)

####################################

# Here is a technical problem.
# Hyper-parameters are on different scales, 
# which makes it difficult to find appropriate 
# shape and scale values for each hyper-parameter. 
# Sørbye (2013) recommends applying a scaling 
# so that priors for the precision parameters 
# for certain models (e.g. rw1, rw2, besag, bym) 
# become comparable. This is achieved by either 
# adding scale.model = TRUE to the rw1 and rw2 
# models, or by typing:

inla.setOption(scale.model.default = TRUE)

# We can now specify one set of scale and shape 
# values (instead of different ones for the rw1 and 
# rw2 models), and that should in principle work better.

# Let's use the PC prior instead of the gamma prior.
# After pottering around with some values
# for U we tried:

U <- 1.5
hyper.prec = list(theta = list(
                   prior = "pc.prec", 
                   param = c(U, 0.01)))

f2 <- Movement ~ fRepro + 
                 f(Year, 
                   model = "rw1",
                   hyper = hyper.prec) +
                 f(DayInYear, 
                   model = "rw2",
                   hyper = hyper.prec
                   ) +
                 f(BearN, model = "iid")

I2 <- inla(f2, 
           control.predictor = list(compute = TRUE),
           family = "gamma",
           data = PB
           )
summary(I2)
Fit2 <- I2$summary.fitted.values[,"mean"]

# Access the trends
Yearsm <- I2$summary.random$Year
Daysm  <- I2$summary.random$DayInYear

# And plot the trends
par(mfrow = c(1,2), mar = c(5,5,2,2), cex.lab = 1.5)
Time <- Yearsm[,1] #* sd(PB$Year) + mean(PB$Year)
plot(x = Time,
     y = Yearsm[,2], type='l',
     xlab = 'Year', 
     ylab = 'Trend',
     ylim = c(-0.5, 0.5) )
abline(h=0, lty=3)
lines(x = Time, y = Yearsm[, c(4)], lty=2)
lines(x = Time, y = Yearsm[, c(6)], lty=2)
text(1988, 0.5, "A", cex = 1.5)

Time <- Daysm[,1] #* sd(PB$DayInYear) + mean(PB$DayInYear)
plot(x = Time,
     y = Daysm[,2], type='l',
     xlab = 'DayInYear', 
     ylab = 'Trend',
     ylim = c(-0.5, 0.5) )
abline(h=0, lty=3)
lines(Time, Daysm[, c(4)], lty=2)
lines(Time, Daysm[, c(6)], lty=2)
text(0, 0.5, "B", cex = 1.5)

# Nice and smooth.

# For your information: The rw1 and rw2 are
# also your smoother options (in case you want
# to do GAM). Time allowing we will discuss
# better options.

###################
# TRY WITH interaction
inla.list.models()
# te = two-dimensional smoother

library(dplyr)

PB$Year2 <- PB$Year
PB$DayInYear2 <- PB$DayInYear

PB2 <- PB %>% filter(Year > 1991)

U <- 1.5

hyper.prec = list(theta = list(
  prior = "pc.prec", 
  param = c(U, 0.05)))

f3 <- Movement ~ fRepro + 
  f(Year, 
    model = "rw2",
    hyper = hyper.prec) +
  f(DayInYear, Year2,
    model = "rw2",
    hyper = hyper.prec) +  
  f(BearN, model = "iid")

I3 <- inla(f3, 
           control.predictor = list(compute = TRUE),
           family = "gamma",
           data = PB2
)
summary(I3)
Fit3 <- I3$summary.fitted.values[,"mean"]

# Access the trends
Yearsm <- I3$summary.random$Year
Daysm  <- I3$summary.random$DayInYear

# And plot the trends
#par(mfrow = c(1,2), mar = c(5,5,2,2), cex.lab = 1.5)
Time <- Yearsm[,1] #* sd(PB$Year) + mean(PB$Year)
plot(x = Time,
     y = Yearsm[,2], type='l',
     xlab = 'Year', 
     ylab = 'Trend',
     ylim = c(-0.5, 0.5) )
abline(h=0, lty=3)
lines(x = Time, y = Yearsm[, c(4)], lty=2)
lines(x = Time, y = Yearsm[, c(6)], lty=2)
text(1988, 0.5, "A", cex = 1.5)

Time <- Daysm[,1] #* sd(PB$DayInYear) + mean(PB$DayInYear)
plot(x = Time,
     y = Daysm[,2], type='l',
     xlab = 'DayInYear', 
     ylab = 'Trend',
     ylim = c(-0.5, 0.5) )
abline(h=0, lty=3)
lines(Time, Daysm[, c(4)], lty=2)
lines(Time, Daysm[, c(6)], lty=2)
text(0, 0.5, "B", cex = 1.5)

```

### 2019-03-11 analysis/INLA/test-INLA-sensor.Rmd
messy trial code with INLA... but some of it works now!

@@
 testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2015, 2016, 2017, 2018))
-#testdat <- gfplot:::interp_survey_bathymetry(testdat)
+testdat <- gfplot:::interp_survey_bathymetry(testdat)
 glimpse(testdat)
-View(testdat[[1]])
+#View(testdat[[1]])
 ```
@@
 # get_ssids()
-sensor_syn <- get_sensor_data_trawl(ssid = c(1, 3, 4, 16), spread_attributes = TRUE)
+# sensor_syn <- get_sensor_data_trawl(ssid = c(1, 3, 4, 16), spread_attributes = TRUE)
 # View(sensor_syn)
-saveRDS(sensor_syn, "data-cache/allsynoptic_sensor_07mar19_time2.rds")
+# saveRDS(sensor_syn, "data-cache/allsynoptic_sensor_07mar19_time2.rds")
 # write.csv(sensor, "synoptic_3_sensor_data_07mar19.csv")
@@
 sensor <- readRDS("data-cache/allsynoptic_sensor_07mar19_time2.rds")
-View(sensor)
+# View(sensor)
 ```
@@
 dat <- left_join(testdat[[1]], sensor, by="fishing_event_id", suffix = c("", ".sensor"))
+# dat <- left_join(testdat, sensor, by="fishing_event_id", suffix = c("", ".sensor"))
 dat$year_f <- factor(dat$year)
-
-View(dat)
-
+# View(dat)
 dat <- dat %>% mutate(do_range = do_mlpL_max - do_mlpL_min, salinity_range = salinity_PSU_max -
   salinity_PSU_min, temp_range = temperature_C_max - temperature_C_min, depth_range = depth_m_max - depth_m_min)
-glimpse(dat)
+# glimpse(dat)
 ```
@@
 plot(-log(depth) ~ salinity_PSU_max,  cex = depth_range/35, data = dat)
-
 ```
@@
   ylim(2,12) + 
-  facet_wrap(~survey_desc, scales = "free")
+  facet_grid(rows = vars(year), cols = vars(ssid), scales = "free")
 ```
@@
 # glimpse(scaled)
-synhs <- scaled %>% filter(do_mlpL_avg!="NA") %>% filter (density!="NA") 
-glimpse(synhs)
+temp_dat <- scaled %>% filter (density!="NA") %>% filter ( ssid =="1") %>% filter(temperature_C_avg!="NA") # %>% filter(year==2017)
+# glimpse(temp_dat)
+# do_dat <- scaled %>% filter (density!="NA") %>% filter ( ssid =="1") %>%filter(do_mlpL_avg!="NA")
+#pg <- readRDS("prediction-grids/QCSpred-grid-interp-cell-width-2.rds")
+#pg <- as.array(pg)
+
 ```
@@
 ```{r}
-ggplot(synhs, aes(depth)) + geom_histogram() 
-ggplot(synhs, aes(density*1000)) + geom_histogram() 
-ggplot(synhs, aes(depth, (density*1000))) + geom_point() 
+ggplot(sens_dat, aes(depth)) + geom_histogram() 
+ggplot(sens_dat, aes(density*1000)) + geom_histogram() 
+ggplot(sens_dat, aes(depth, (density*1000))) + geom_point() 
 #ggplot(synhs, aes(depth_m, log(density*1000+1))) + geom_point() 
@@
 ```{r}
+fit_inla_sensor <- function(dat, response = "temperature_C_avg",
+                     n_knots = 50,
+                     family = "binomial",
+                     max_edge = c(20, 100),
+                     kmeans = FALSE,
+                     plot = FALSE,
+                     fit_model = TRUE,
+                     extend = list(n = 8, offset = -0.1),
+                     offset = c(5, 25),
+                     cutoff = 10,
+                     include_depth = TRUE,
+                     verbose = FALSE,
+                     debug = FALSE,
+                     trials = 10) {
+  d <- dat
+  coords <- as.matrix(unique(d[, c("X", "Y")]))
 
-####
-names(d) <- tolower(names(d))
-d$species_common_name <- tolower(d$species_common_name)
-d$species_science_name <- tolower(d$species_science_name)
-d$year <- lubridate::year(d$trip_start_date)
+  # use kmeans() to calculate centers:
+  if (kmeans) {
+    km <- stats::kmeans(x = coords, centers = n_knots)
+    mesh <- INLA::inla.mesh.create(km$centers,
+      extend = extend
+    )
+  } else {
+    bnd <- INLA::inla.nonconvex.hull(coords)
+    mesh <- INLA::inla.mesh.2d(
+      offset = offset,
+      boundary = bnd,
+      max.edge = max_edge,
+      cutoff = cutoff
+    )
+  }
 
-sp <- filter(d, species_common_name %in% "pacific ocean perch") %>% 
-  filter(!is.na(catch_weight)) %>% 
-  filter(year %in% seq(2004, 2012, 2))
+  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], ")")
+  if (plot) {
+    graphics::plot(mesh, asp = 1)
+    graphics::points(coords, col = "red")
+  }
 
-dat <- sp
-dat <- filter(dat, start_lon > -128, start_lon < 125, start_lat > 48, start_lat < 50.3)
-nrow(dat)
+  spde <- INLA::inla.spde2.matern(mesh, alpha = 3 / 2)
+  dat <- data.frame(
+    y = d[, response, drop = TRUE],
+    xcoo = d$X,
+    ycoo = d$Y
+  )
 
-dat <- filter(dat, fe_bottom_water_temperature > 1, fe_bottom_water_temperature < 12,
-  !is.na(fe_bottom_water_temp_depth), !is.na(fe_bottom_water_temperature))
-nrow(dat)
+  dat$intercept <- 1
 
-dat <- select(dat, year, start_lon, start_lat, catch_weight, fe_bottom_water_temp_depth) %>% 
-  rename(X = start_lon, Y = start_lat)
+  if (include_depth) {
+    dat$depth <- d$depth_scaled
+    dat$depth2 <- d$depth_scaled2
+  }
 
-attr(dat, "projection") <- "LL"
+  # construct index for ar1 model:
+  iset <- INLA::inla.spde.make.index(
+    "i2D",
+    n.spde = mesh$n
+  )
 
-dat <- PBSmapping::convUL(dat)
-dat <- as.data.frame(na.omit(dat))
-
-subcoords = cbind(dat$X, dat$Y)
-
-bnd = inla.nonconvex.hull(subcoords, convex = 30)
-mesh1 = inla.mesh.2d(
-  boundary = bnd,
-  max.edge = c(15, 40),
-  cutoff = 5,
-  offset = 1
-)
-plot(mesh1)
-points(dat$X, dat$Y, col = "red")
-summary(mesh1)
-
-##########
-A.data <- inla.spde.make.A(mesh1, loc = cbind(dat$X, dat$Y))
-
-# Make SPDE based on mesh
-spde = inla.spde2.matern(mesh1, alpha = 3 / 2)
-
-n = nrow(dat)
-YEARS <- unique(dat$year)
-k = length(unique(dat$year))
-
-# Make a design matrix where the first year is the intercept
-# ...
-dm <- matrix(nrow = nrow(dat), ncol = 2)
-dm[,1] <- 1
-dm[,2] <- dat$fe_bottom_water_temp_depth
-
-iset <- inla.spde.make.index("i2D", n.spde = mesh1$n, n.group = k)
-
-X.1 = dm
-Covar.names <- c("intercept", "depth")
+  # make the covariates:
+  X.1 <- dat[, -c(1:3), drop = FALSE]
+  covar_names <- colnames(X.1)
   XX.list <- as.list(X.1)
   effect.list <- list()
-#   effect.list[[1]] <- c(iset, list(Intercept=1))
   effect.list[[1]] <- c(iset)
-for (Z in 1:ncol(X.1))
+  for (Z in seq_len(ncol(X.1)))
     effect.list[[Z + 1]] <- XX.list[[Z]]
-names(effect.list) <- c("1", Covar.names)
+  names(effect.list) <- c("1", covar_names)
 
-### Make data stack.
-A <-
-  inla.spde.make.A(
-    mesh = mesh1,
-    loc = cbind(dat$X, dat$Y),
-    group = dat$year
+  A <- INLA::inla.spde.make.A(
+    mesh = mesh,
+    loc = cbind(dat$xcoo, dat$ycoo)
   )
-A.list = list()
-A.list[[1]] = A
-for (Z in 1:ncol(X.1))
+  A.list <- list()
+  A.list[[1]] <- A
+  for (Z in seq_len(ncol(X.1)))
     A.list[[Z + 1]] <- 1
 
-### Make projection points stack.
-Ntrials <- rep(1, length(dat$Y))
-
-sdat <-
-  inla.stack(
-    tag = 'stdata',
-    data = list(
-      y = dat$Y,
-      link = 1,
-      Ntrials = Ntrials
-    ),
+  # make projection points stack:
+  sdat <- INLA::inla.stack(
+    tag = "stdata",
+    data = list(y = dat$y),
     A = A.list,
@@
 
-formula = as.formula(
-  paste0(
+  formula <- as.formula(paste(
     "y ~ -1 +",
-    paste(Covar.names, collapse = "+"),
-    "+ f(i2D, model=spde, group = i2D.group, control.group = list(model='ar1'))"
+    paste(covar_names, collapse = "+"),
+    "+ f(i2D, model=spde)"
+  ))
+
+  if (fit_model) {
+    try_fitting <- TRUE
+    trial_i <- 1
+    while (try_fitting && trial_i <= trials) {
+      model <- try(INLA::inla(formula,
+        family = family,
+        data = INLA::inla.stack.data(sdat),
+        control.predictor = list(
+          compute = TRUE,
+          A = INLA::inla.stack.A(sdat)
+        ),
+        control.fixed = list(
+          mean = 0, prec = 1 / (5^2),
+          mean.intercept = 0, prec.intercept = 1 / (20^2)
+        ),
+        control.compute = list(config = TRUE),
+        verbose = verbose,
+        debug = debug,
+        keep = FALSE
+      ))
+      trial_i <- trial_i + 1
+      if (class(model) != "try-error") {
+        try_fitting <- FALSE
+      } else {
+        # if (trial_i > 5) browser()
+        message("INLA error. Trying again.")
+      }
+    }
+    if (class(model) == "try-error") {
+      model <- NA
+    }
+  } else {
+    model <- NA
+  }
+  list(
+    model = model, mesh = mesh, spde = spde, data = dat,
+    formula = formula, iset = iset
   )
-)		# field evolves with AR1 by year
+}
+```
 
-inlaModel <-
-  inla(
-    formula,
-    family = "gamma",
-    data = inla.stack.data(sdat),
-    control.predictor = list(compute = TRUE, A = inla.stack.A(sdat)),
-    verbose = TRUE,
-    debug = TRUE,
-    keep = FALSE,
-    control.compute = list(dic = TRUE, cpo = TRUE, config = TRUE),
-    control.fixed = list(correlation.matrix = TRUE),
-    control.inla = list(lincomb.derived.correlation.matrix = TRUE)
-  )
+```{r}
+obj <- fit_inla_sensor(temp_dat, response = "temperature_C_avg",
+                       n_knots = 50,
+                       family = "gaussian",
+                       max_edge = c(20, 100),
+                       kmeans = FALSE,
+                       plot = FALSE,
+                       fit_model = TRUE,
+                       extend = list(n = 8, offset = -0.1),
+                       offset = c(5, 25),
+                       cutoff = 10,
+                       include_depth = TRUE,
+                       verbose = FALSE,
+                       debug = FALSE,
+                       trials = 2)
+
+pg <- gfplot:::make_prediction_grid(filter(temp_dat, year %in% c(2016, 2017)), survey ="SYN QCS")$grid
+
+cell_area <- gfplot:::make_prediction_grid(filter(temp_dat, year %in% c(2016, 2017)), survey = "SYN QCS")$cell_area
+
+pred_temp <- predict_inla_sens(obj=obj, pred_grid = pg , samples = 100L,
+                         include_depth = TRUE)
+```
+
+Project onto grid:
+
+```{r posterior-draws}
+pg_one_year <- filter(pg, year == min(pg$year))
+predicted <- predict_inla_sens(obj = obj, pred_grid = pg_one_year, samples = 100, include_depth = TRUE)
 ```
@@
 
-synhs$year_f2 <- synhs$year_f
-synhs$year_f3 <- synhs$year_f
+# alternative to function, hard code
 
-f1 <- temperature_C2 ~ depth_scaled + depth_scaled2 + 
+samples = 10L
+pred_grid <- pg
+
+predict_inla_sens <- function(obj, pred_grid, samples = 100L,
+                         include_depth = TRUE) {
+
+  mesh <- obj$mesh
+  model <- obj$model
+  iset <- obj$iset
+
+  inla.mcmc <- INLA::inla.posterior.sample(n = samples, model)
+
+  # get indices of various effects:
+  latent_names <- rownames(inla.mcmc[[1]]$latent)
+  re_indx <- grep("^i2D", latent_names)
+  if (include_depth) {
+    depth_fe_indx1 <- grep("^depth$", latent_names)
+    depth_fe_indx2 <- grep("^depth2$", latent_names)
+  }
+  intercept_indx <- grep("^intercept$", latent_names)
+
+  # read in locations and knots, from projection matrix
+  grid_locs <- pred_grid[, c("X", "Y")]
+
+  # projections will be stored as an array:
+  projected_latent_grid <-
+    array(0, dim = c(nrow(grid_locs), samples))
+  proj_matrix <- INLA::inla.spde.make.A(mesh, loc = as.matrix(grid_locs))
+
+  # do MCMC projections:
+  indx <- which(iset$i2D.group == 1)
+  for (i in seq_len(samples)) {
+
+    if (include_depth) {
+      depth_effects <-
+        pred_grid[, "depth_scaled", drop = TRUE] * inla.mcmc[[i]]$latent[depth_fe_indx1] +
+        pred_grid[, "depth_scaled2", drop = TRUE] * inla.mcmc[[i]]$latent[depth_fe_indx2]
+    } else {
+      depth_effects <- 0
+    }
+
+    projected_latent_grid[, i] <-
+      as.numeric(proj_matrix %*% inla.mcmc[[i]]$latent[re_indx][indx]) +
+      inla.mcmc[[i]]$latent[intercept_indx] +
+      depth_effects
+  }
+
+  projected_latent_grid
+}
+
+```
+
+Calculate probability * Gamma median values on the grid:
+
+```{r prediction-grid-calcs}
+pg_one_year <- filter(pg, year == min(pg$year))
+attributes(predicted)
+# out <- stats::plogis(p_bin) * exp(p_pos)
+out_med <- apply(predicted, c(1), median)
+# out_bin <- stats::plogis(p_bin)
+# out_bin <- apply(out_bin, c(1, 3), median)
+dim(predicted)
+
+pg_all <- bind_rows(replicate(1, pg_one_year, simplify = FALSE))
+
+out_long <- reshape2::melt(out_med)
+pg_all$pred <- out_long$value
+```
+
+Spatiotemporal plot:
+
+```{r spatio-temporal-plot, fig.width=9, fig.height=9}
+coast <- gfplot:::load_coastline(range(d$lon), range(d$lat), utm_zone = 9)
+isobath <- gfplot:::load_isobath(range(d$lon), range(d$lat),
+  bath = c(100, 200, 500), utm_zone = 9
+)
+map_padding <- c(-5, 5)
+
+make_plot <- function(dat, fill = "sqrt(pred)") {
+  g <- ggplot(dat, aes(X * 10, Y * 10)) +
+    geom_raster(aes_string(fill = fill)) +
+    scale_fill_viridis_c(option = "C") +
+    #facet_wrap(~ year) +
+    geom_polygon(
+      data = coast, aes_string(x = "X", y = "Y", group = "PID"),
+      fill = "grey80"
+    ) +
+    geom_point(
+      data = filter(d, present == 1),
+      aes(x = X * 10, y = Y * 10, size = density), inherit.aes = FALSE,
+      pch = 21, col = "white", alpha = 0.3
+    ) +
+    geom_point(
+      data = filter(d, present == 0),
+      aes(x = X * 10, y = Y * 10), inherit.aes = FALSE,
+      pch = 4, col = "white", alpha = 0.2
+    )
+
+  g <- g + geom_path(
+    data = isobath, aes_string(
+      x = "X", y = "Y",
+      group = "paste(PID, SID)"
+    ),
+    inherit.aes = FALSE, lwd = 0.4, col = "grey70", alpha = 0.4
+  )
+
+  g <- g + xlab("UTM 9N Easting (km)") + ylab("UTM 9N Northing (km)") +
+    scale_size(range = c(0, 9)) +
+    guides(size = FALSE, fill = FALSE) +
+    theme_pbs() +
+    coord_equal(
+      expand = FALSE, xlim = range(pg_all$X * 10) + map_padding,
+      ylim = range(pg_all$Y * 10) + map_padding
+    )
+
+  g
+}
+# make_plot(pg_all)
+median(pg_all$pred)
+
+make_plot(pg_all, "pred") + 
+  scale_fill_gradient2(
+    midpoint = median(pg_all$pred),
+    low = scales::muted("blue"), 
+    mid = "white",
+    high = scales::muted("red"))
+
+make_plot(pg_all, "depth_scaled") + 
+  scale_fill_gradient2(
+    midpoint = 0,
+    high = scales::muted("blue"), 
+    mid = "white",
+    low = scales::muted("red"))
+```
+
+Run using Zuur's polar bear code as template for letting relationship between temp and depth change through time...
+
+```{r}
+temp_dat$year_f2 <- temp_dat$year_f
+temp_dat$year_f3 <- temp_dat$year_f
+
+f1 <- temperature_C_avg ~ depth_scaled + depth_scaled2 + 
                  f(year_f, model = "rw2") +
-                 f(year_f2, depth_m, model = "rw2") +
-                 f(year_f3, depth2, model = "rw2")
+                 f(year_f2, depth_scaled, model = "rw2") +
+                 f(year_f3, depth_scaled2, model = "rw2")
 
@@
            control.predictor = list(compute = TRUE),
-           family = "gamma",
-           data = synhs
+           family = "gaussian",
+           data = temp_dat
            )
@@
 # Here is how we access them:
-Yearsm <- I1$summary.random$Year
-Daysm  <- I1$summary.random$DayInYear
+Yearsm <- I1$summary.random$year_f
+Depthre  <- I1$summary.random$year_f2
+Depth2re  <- I1$summary.random$year_f3
+```
 
+```{r}
 # The rest is simple plotting
@@
 
-plot(Daysm[,1:2], type='l',
-     xlab = 'DayInYear', 
+plot(Depthre[,1:2], type='l',
+     xlab = 'Depth', 
      ylab = 'Smoother',
@@
 abline(h=0, lty=3)
-lines(Daysm[, c(1, 4)], lty=2)
-lines(Daysm[, c(1, 6)], lty=2)
+lines(Depthre[, c(1, 4)], lty=2)
+lines(Depthre[, c(1, 6)], lty=2)
 text(0, 0.6, "B", cex = 1.5)
+```
 
+```{r}
 ####################################
@@
 
-###################
-# TRY WITH interaction
-inla.list.models()
-# te = two-dimensional smoother
+```
 
-library(dplyr)
+Sean's code that I dont' remember where i found...
 
-PB$Year2 <- PB$Year
-PB$DayInYear2 <- PB$DayInYear
+```{r}
 
-PB2 <- PB %>% filter(Year > 1991)
+####
+# names(d) <- tolower(names(d))
+# d$species_common_name <- tolower(d$species_common_name)
+# d$species_science_name <- tolower(d$species_science_name)
+# d$year <- lubridate::year(d$trip_start_date)
+# 
+# sp <- filter(d, species_common_name %in% "pacific ocean perch") %>% 
+#   filter(!is.na(catch_weight)) %>% 
+#   filter(year %in% seq(2004, 2012, 2))
+# 
+# dat <- sp
+# dat <- filter(dat, start_lon > -128, start_lon < 125, start_lat > 48, start_lat < 50.3)
+# nrow(dat)
+# 
+# dat <- filter(dat, fe_bottom_water_temperature > 1, fe_bottom_water_temperature < 12,
+#   !is.na(fe_bottom_water_temp_depth), !is.na(fe_bottom_water_temperature))
+# nrow(dat)
+# 
+# dat <- select(dat, year, start_lon, start_lat, catch_weight, fe_bottom_water_temp_depth) %>% 
+#   rename(X = start_lon, Y = start_lat)
 
-U <- 1.5
+dat <- temp_dat
 
-hyper.prec = list(theta = list(
-  prior = "pc.prec", 
-  param = c(U, 0.05)))
+attr(dat, "projection") <- "LL"
 
-f3 <- Movement ~ fRepro + 
-  f(Year, 
-    model = "rw2",
-    hyper = hyper.prec) +
-  f(DayInYear, Year2,
-    model = "rw2",
-    hyper = hyper.prec) +  
-  f(BearN, model = "iid")
+dat <- PBSmapping::convUL(dat)
+dat <- as.data.frame(na.omit(dat))
 
-I3 <- inla(f3, 
-           control.predictor = list(compute = TRUE),
+subcoords = cbind(dat$X, dat$Y)
+
+bnd = inla.nonconvex.hull(subcoords, convex = 30)
+mesh1 = inla.mesh.2d(
+  boundary = bnd,
+  max.edge = c(15, 40),
+  cutoff = 5,
+  offset = 1
+)
+plot(mesh1)
+points(dat$X, dat$Y, col = "red")
+summary(mesh1)
+
+##########
+A.data <- inla.spde.make.A(mesh1, loc = cbind(dat$X, dat$Y))
+
+# Make SPDE based on mesh
+spde = inla.spde2.matern(mesh1, alpha = 3 / 2)
+
+n = nrow(dat)
+YEARS <- unique(dat$year)
+k = length(unique(dat$year))
+
+# Make a design matrix where the first year is the intercept
+# ...
+dm <- matrix(nrow = nrow(dat), ncol = 2)
+dm[,1] <- 1
+
+#########
+
+dm[,2] <- dat$fe_bottom_water_temp_depth
+
+########
+
+iset <- inla.spde.make.index("i2D", n.spde = mesh1$n, n.group = k)
+
+X.1 = dm
+Covar.names <- c("intercept", "depth")
+XX.list <- as.list(X.1)
+effect.list <- list()
+#   effect.list[[1]] <- c(iset, list(Intercept=1))
+effect.list[[1]] <- c(iset)
+for (Z in 1:ncol(X.1))
+  effect.list[[Z + 1]] <- XX.list[[Z]]
+names(effect.list) <- c("1", Covar.names)
+
+### Make data stack.
+A <-
+  inla.spde.make.A(
+    mesh = mesh1,
+    loc = cbind(dat$X, dat$Y),
+    group = dat$year
+  )
+A.list = list()
+A.list[[1]] = A
+for (Z in 1:ncol(X.1))
+  A.list[[Z + 1]] <- 1
+
+### Make projection points stack.
+Ntrials <- rep(1, length(dat$Y))
+
+sdat <-
+  inla.stack(
+    tag = 'stdata',
+    data = list(
+      y = dat$Y,
+      link = 1,
+      Ntrials = Ntrials
+    ),
+    A = A.list,
+    effects = effect.list
+  )
+
+formula = as.formula(
+  paste0(
+    "y ~ -1 +",
+    paste(Covar.names, collapse = "+"),
+    "+ f(i2D, model=spde, group = i2D.group, control.group = list(model='ar1'))"
+  )
+)		# field evolves with AR1 by year
+
+inlaModel <-
+  inla(
+    formula,
     family = "gamma",
-           data = PB2
+    data = inla.stack.data(sdat),
+    control.predictor = list(compute = TRUE, A = inla.stack.A(sdat)),
+    verbose = TRUE,
+    debug = TRUE,
+    keep = FALSE,
+    control.compute = list(dic = TRUE, cpo = TRUE, config = TRUE),
+    control.fixed = list(correlation.matrix = TRUE),
+    control.inla = list(lincomb.derived.correlation.matrix = TRUE)
   )
-summary(I3)
-Fit3 <- I3$summary.fitted.values[,"mean"]
-
-# Access the trends
-Yearsm <- I3$summary.random$Year
-Daysm  <- I3$summary.random$DayInYear
-
-# And plot the trends
-#par(mfrow = c(1,2), mar = c(5,5,2,2), cex.lab = 1.5)
-Time <- Yearsm[,1] #* sd(PB$Year) + mean(PB$Year)
-plot(x = Time,
-     y = Yearsm[,2], type='l',
-     xlab = 'Year', 
-     ylab = 'Trend',
-     ylim = c(-0.5, 0.5) )
-abline(h=0, lty=3)
-lines(x = Time, y = Yearsm[, c(4)], lty=2)
-lines(x = Time, y = Yearsm[, c(6)], lty=2)
-text(1988, 0.5, "A", cex = 1.5)
-
-Time <- Daysm[,1] #* sd(PB$DayInYear) + mean(PB$DayInYear)
-plot(x = Time,
-     y = Daysm[,2], type='l',
-     xlab = 'DayInYear', 
-     ylab = 'Trend',
-     ylim = c(-0.5, 0.5) )
-abline(h=0, lty=3)
-lines(Time, Daysm[, c(4)], lty=2)
-lines(Time, Daysm[, c(6)], lty=2)
-text(0, 0.5, "B", cex = 1.5)
-
 ```

### 2019-03-13 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
Switch R files to Rmd to save the plots for a presentation

# Fit spatial temporal models to bottom temperature

```{r}
library(tidyverse)
library(sdmTMB)

d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))

set.seed(39201114)
withheld_fe <- sample(d_trawl$fishing_event_id, size = round(nrow(d_trawl) * .2))
d_withheld <- filter(d_trawl, fishing_event_id %in% withheld_fe)
d_fit <- filter(d_trawl, !fishing_event_id %in% withheld_fe)
nrow(d_fit)
nrow(d_withheld)
unique(d_withheld$year)
unique(d_fit$year)

# FIXME: need to work out a bug in production on this new data set
# It's crashing TMB.
d_fit <- d_trawl
d_withheld <- d_trawl
spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 200)
plot_spde(spde)

m_temp1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
  # time_varying =  ~ depth_scaled + depth_scaled2,
  time = "year", spde = spde, family = gaussian(link = "identity"), ar1_fields = FALSE,
  anisotropy = FALSE, include_spatial = TRUE, silent = TRUE)

m_temp1$model$par

# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
exp(m_temp1$model$par[["ln_phi"]])

predictions <- predict(m_temp1)
predictions$data$residuals <- residuals(m_temp1)

plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", colour = column)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed()
}

plot_map(predictions$data, "est") +
  scale_colour_viridis_c() +
  ggtitle("Prediction (fixed effects + all random effects)")

plot_map(predictions$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_colour_viridis_c()

plot_map(predictions$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_colour_gradient2()

plot_map(predictions$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_colour_gradient2()

ggplot(predictions$data, aes(est, temperature_c)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  geom_abline()

ggplot(predictions$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions5$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# Looks a bit curved, let's try with a quadratic

m_temp2 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 2),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2 <- predict(m_temp2)
predictions2$data$residuals <- residuals(m_temp2)

ggplot(predictions2$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions2$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions2$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# Try with a third degree polynomial:
m_temp3 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions3 <- predict(m_temp3)
predictions3$data$residuals <- residuals(m_temp3)

ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# That's getting pretty straight although it looks a bit heteroscedastic

# Perhaps this would work better with log temperature as the response
# since it basically isn't going to go lower than 0 degree celsius

d_fit$log_temperature_c <- log(d_fit$temperature_c)
m_temp4 <- sdmTMB(d_fit, log_temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions4 <- predict(m_temp4)
predictions4$data$residuals <- residuals(m_temp4)

ggplot(predictions4$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions4$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

ggplot(predictions4$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# That might've helped a little but not a lot. Let's try with depth and not log depth.

m_temp5 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_m, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions5 <- predict(m_temp5)
predictions5$data$residuals <- residuals(m_temp5)

ggplot(predictions5$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions5$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()

# I don't think that helps with the heteroscedasticity but it made the curvature of the residuals a little bit worse.

ggplot(predictions5$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(predictions5$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth()

aic <- function(m) {
  k <- length(m$model$par)
  nll <- m$model$objective
  2 * k - 2 * (-nll)
}
aic(m_temp1)
aic(m_temp2)
aic(m_temp3)
aic(m_temp4)
aic(m_temp5)

# Conclusion: log depth, 3rd order polynomial with depth

ggplot(predictions3$data, aes(est, temperature_c)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  geom_abline()

# R2
cor(predictions3$data$est, predictions3$data$temperature_c)^2

# Extrapolate:

predictions3_nd <- predict(m_temp3, newdata = sdmTMB::qcs_grid)

ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_color_gradient2()

# A short function for plotting our predictions:
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

plot_map(predictions3_nd$data, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")

plot_map(predictions3_nd$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")

plot_map(predictions3_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()

plot_map(predictions3_nd$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()

temp_cog <- get_cog(predictions3_nd)

temp_cog %>% reshape2::dcast(year ~ coord, value.var = "est") %>%
  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
  scale_color_viridis_c()

temp_cog %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  facet_wrap(~coord, scales = "free_y")
```

### 2019-03-13 analysis/tmb-sensor-explore/3-fit-do-salinity.Rmd
Switch R files to Rmd to save the plots for a presentation

# Fit spatiotemporal models to dissolved oxygen and salinity

```{r}
library(sdmTMB)
library(tidyverse)
d_fit <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
d_fit$log_do_mlpl <- log(d_fit$do_mlpl)

d <- filter(d_fit, !is.na(do_mlpl), !is.na(salinity_psu))
spde <- make_spde(d$X, d$Y, n_knots = 150)
unique(d$year)
plot_spde(spde)

hist(d$do_mlpl)
hist(log(d$do_mlpl))
min(d$do_mlpl)

m_do1 <- sdmTMB(d, log_do_mlpl ~ 1 + depth_scaled,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)

m_do1$model$par
exp(m_do1$model$par[["ln_phi"]])

get_diag <- function(m, response = "log_do_mlpl") {
  predictions1 <- predict(m)
  predictions1$data$residuals <- residuals(m)

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point() +
      coord_fixed()
  }

  g <- plot_map(predictions1$data, "est") +
    scale_colour_viridis_c() +
    ggtitle("Prediction (fixed effects + all random effects)")
  print(g)

  g <- plot_map(predictions1$data, "est_fe") +
    ggtitle("Prediction (fixed effects only)") +
    scale_colour_viridis_c()
  print(g)

  g <- plot_map(predictions1$data, "est_re_s") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- ggplot(predictions1$data, aes_string("est", response)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()
  print(g)

  g <- ggplot(predictions1$data, aes(est, residuals)) +
    geom_point() +
    geom_smooth()
  print(g)

  g <- ggplot(predictions1$data, aes(depth_scaled, residuals)) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions1$data, aes(X, Y, colour = residuals)) +
    geom_point() +
    coord_fixed() +
    scale_color_gradient2()
  print(g)

  aic <- function(m) {
    k <- length(m$model$par)
    nll <- m$model$objective
    2 * k - 2 * (-nll)
  }

  r2 <- cor(predictions1$data$est, predictions1$data$log_do_mlpl)^2
  print(r2)

  print(aic(m))
}

get_diag(m_do1)

m_do2 <- sdmTMB(d, log_do_mlpl ~ 1 + poly(depth_scaled, 2),
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_do2)

m_do3 <- sdmTMB(d, log_do_mlpl ~ 1 + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_do3)

.grid <- sdmTMB::qcs_grid %>% filter(year == 2017)
predictions3_nd <- predict(m_do3, newdata = .grid)

plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

plot_map(predictions3_nd$data, "exp(est)") +
  scale_fill_viridis_c(option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)") +
  geom_point(data = d, aes(X, Y, size = do_mlpl), inherit.aes = FALSE, pch = 21) +
  scale_size_area(max_size = 10)

plot_map(predictions3_nd$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")

plot_map(predictions3_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()

# Salinity:

hist(d$salinity_psu, breaks = 40)
hist(log(d$salinity_psu))
hist(exp(d$salinity_psu), breaks = 40)

m_salinity3 <- sdmTMB(d, salinity_psu ~ 1 + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity3, response = "salinity_psu")

m_salinity4 <- sdmTMB(d,
  salinity_psu ~ 1 + depth_scaled * log_do_mlpl,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity4, response = "salinity_psu")

m_salinity5 <- sdmTMB(d,
  salinity_psu ~ 1 + poly(depth_scaled, 3) * log_do_mlpl,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity5, response = "salinity_psu")

m_salinity5 <- sdmTMB(d,
  salinity_psu ~ 1 + poly(depth_scaled, 3) * temperature_c,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity5, response = "salinity_psu")

m_salinity5 <- sdmTMB(d,
  salinity_psu ~ 1 + poly(depth_scaled, 3) * log_do_mlpl + temperature_c,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity5, response = "salinity_psu")

# much harder to fit
```

### 2019-03-13 analysis/tmb-sensor-explore/4-mountain-rw-examples.Rmd
Switch R files to Rmd to save the plots for a presentation

# Fit time-varying quadratic models with depth and temperature

```{r}
library(dplyr)
library(ggplot2)
survey <- "SYN QCS"

fit_mountains <- function(survey_set_data, type = c("temperature", "depth"), n_knots = 100) {

  type <- match.arg(type)

  d <- survey_set_data
  # sd <- gfplot::get_sensor_data_trawl(ssid = 1, spread_attributes = FALSE)
  sd <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl.rds"))
  sd <- sd %>%
    select(-count, -start_time, -end_time, -min, -max) %>%
    mutate(attribute = tolower(attribute)) %>%
    distinct() %>%
    reshape2::dcast(fishing_event_id + year + ssid + survey_desc ~ attribute, value.var = "avg")

  col <- if (grepl("SYN", survey)) "density_kgpm2" else "density_ppkm2"
  dat <- gfplot:::tidy_survey_sets(d, survey, years = seq(1, 1e6),
    density_column = col)

  dat <- left_join(dat, select(sd, temperature_c, year, fishing_event_id), 
    by = c("year", "fishing_event_id")) %>%
    rename(temperature = temperature_c)

  nrow(dat)
  dat <- filter(dat, !is.na(temperature), !is.na(depth))
  nrow(dat)

  spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = n_knots)

  dat <- mutate(dat,
    temperature_mean = mean(temperature),
    temperature_sd = sd(temperature),
    depth_mean = mean(depth),
    depth_sd = sd(depth),
    depth_scaled = (depth - mean(depth)) / sd(depth),
    temperature_scaled = (temperature - mean(temperature)) / sd(temperature),
    temperature_scaled2 = temperature_scaled^2,
    depth_scaled2 = depth_scaled^2,
  )

  if (type == "depth") {
    formula <-  ~ 0 + depth_scaled + depth_scaled2
  }
  if (type == "temperature"){
    formula <-  ~ 0 + temperature_scaled + temperature_scaled2
  }
  m <- sdmTMB::sdmTMB(
    formula = density ~ 0 + as.factor(year),
    time_varying = formula,
    data = dat, time = "year",
    ar1_fields = FALSE,
    spde = spde, include_spatial = TRUE,
    family = sdmTMB::tweedie(link = "log"),
    anisotropy = FALSE,
    silent = TRUE)

  get_y_hat <- function(b0, b1, b2, year, predictor, mean_column, sd_column) {
    x_pred <- seq(min(dat[[predictor]]), max(dat[[predictor]]), length.out = 300)
    data.frame(
      x = x_pred * dat[[sd_column]][[1]] + dat[[mean_column]][[1]],
      y_hat = exp(b0 + b1 * x_pred + b2 * x_pred^2),
      year = year)
  }

  r <- m$tmb_obj$report()
  r$b_rw_t
  n_t <- nrow(r$b_rw_t)
  yrs <- sort(unique(dat$year))

  pred_temperature <- purrr::map_df(seq_len(n_t), function(.t) {
    get_y_hat(r$b_j[.t], r$b_rw_t[.t,1], r$b_rw_t[.t,2], yrs[.t], 
      sd_column = if (type == "depth") "depth_sd" else "temperature_sd" ,
      mean_column = if (type == "depth") "depth_mean" else "temperature_mean", 
      predictor = if (type == "depth") "depth_scaled" else "temperature_scaled")
  })

  pred_temperature
}

plot_mountains <- function(dat, type = c("depth", "temperature")) {
  type <- match.arg(type)
  mountain_scaler <- max(dat$y_hat)
  mountain_scaler <- mountain_scaler * 10000
  ggplot(dat, aes_string(x = "x", 
    ymax = "100000*y_hat + year*mountain_scaler", ymin = "0 + year*mountain_scaler",
    group = "year", colour = "year", fill = "year")) +
    geom_ribbon(lwd = 0.5, alpha = 0.2) +
    xlab(if (type == "depth") "Depth" else "Temperature") +
    scale_color_viridis_c(option = "C") +
    scale_fill_viridis_c(option = "C") +
    ylab("Predicted density") +
    gfplot::theme_pbs() +
    coord_cartesian(expand = FALSE) + 
    theme(axis.text.y = element_blank())
}
```

```{r, fig.asp=1.3, fig.width=5}
# d <- gfplot::get_survey_sets("pacific cod", ssid = 1)
d <- readRDS(here::here("../gfsynopsis/report/data-cache/petrale-sole.rds"))
f <- fit_mountains(d$survey_sets, type = "temperature")
plot_mountains(f, type = "temperature")
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")

d <- readRDS(here::here("../gfsynopsis/report/data-cache/pacific-cod.rds"))
f <- fit_mountains(d$survey_sets, type = "temperature")
plot_mountains(f, type = "temperature")
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")

d <- readRDS(here::here("../gfsynopsis/report/data-cache/pacific-ocean-perch.rds"))
f <- fit_mountains(d$survey_sets, type = "temperature")
plot_mountains(f, type = "temperature")
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")

d <- readRDS(here::here("../gfsynopsis/report/data-cache/rougheye-blackspotted-rockfish-complex.rds"))
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")
```

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
running INLA with multiple years

@@
 ######################################
-# Start Bayesian analysis
-# One more time the model:
-
-# Movement_ij      ~ Gamma(mu_ik, r)
-# E(Movement_ij)   = mu_ij
-# var(Movement_ij) = mu_ij^2 / r 
-# mu_ij = exp(Intercept + fRepro + Annual Trend + Seasonal Trend + a_i)
-
-# Below we define a model with a factor fRepro, a long term
-# trend for Year, and seasonal trend for DayInYear, and
-# a random intercept for bear.
-
-# Each trend is modelled as a random walk trend.
-# So each of them has a sigma.
+ Start Bayesian analysis
+ One more time the model:
+ Movement_ij      ~ Gamma(mu_ik, r)
+ E(Movement_ij)   = mu_ij
+ var(Movement_ij) = mu_ij^2 / r
+ mu_ij = exp(Intercept + fRepro + Annual Trend + Seasonal Trend + a_i)
+ Below we define a model with a factor fRepro, a long term
+ trend for Year, and seasonal trend for DayInYear, and
+ a random intercept for bear.
+ Each trend is modelled as a random walk trend.
+ So each of them has a sigma.
 
 ```{r}
-fit_inla_sensor <- function(dat, response = "temperature_C_avg",
+
+d <- temp_dat
+response = "temperature_C_avg"
+n_knots = 50
+family = "gaussian"
+plot = TRUE
+fit_model = TRUE
+extend = list(n = 8, offset = -0.1)
+RangeGuess = 100 
+offset = c(RangeGuess/5, RangeGuess)
+max_edge = c(RangeGuess/5,RangeGuess)
+cutoff = (RangeGuess/5)/5
+prior_range = c(RangeGuess, 0.01) # P(range < 0.05) = 0.01
+prior_sigma = c(0.5, 0.5)
+kmeans = FALSE    
+verbose = FALSE
+debug = FALSE
+trials = 2
+covariates = c("depth_scaled", "depth_scaled2")
+time = "year"
+
+fit_inla_sensor <- function(d, response = "temperature_C_avg",
                      n_knots = 50,
                      family = "binomial",
-                     max_edge = c(20, 100),
-                     kmeans = FALSE,
                      plot = FALSE,
@@
                      extend = list(n = 8, offset = -0.1),
-                     offset = c(5, 25),
-                     cutoff = 10,
-                     include_depth = TRUE,
+                     kmeans = FALSE,                   
+                     RangeGuess = 100, 
+                     offset = c(RangeGuess/5, RangeGuess),                 
+                     max_edge = c(RangeGuess/5,RangeGuess),
+                     cutoff = (RangeGuess/5)/5,
+                     prior_range = c(RangeGuess, 0.01), # P(range < 0.05) = 0.01
+                     prior_sigma = c(0.5, 0.5),
+                     time = "year",
+                     covariates = NULL,
                      verbose = FALSE,
@@
                      trials = 10) {
-  d <- dat
+
   coords <- as.matrix(unique(d[, c("X", "Y")]))
@@
     bnd <- INLA::inla.nonconvex.hull(coords)
+
     mesh <- INLA::inla.mesh.2d(
@@
   }
+  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh$n)
 
-  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], ")")
   if (plot) {
@@
 
-  spde <- INLA::inla.spde2.matern(mesh, alpha = 3 / 2)
+  # create data.frame including response ("y"), coordinates("xcoo", "ycoo"), covariates, and "intercept"
   dat <- data.frame(
@@
     xcoo = d$X,
-    ycoo = d$Y
+    ycoo = d$Y, 
+    raw_time = d[, time, drop = TRUE],
+    select(d, covariates)
   )
 
-  dat$intercept <- 1
+  # make a design matrix where the first year is the intercept:
+  YO  <- order(dat$raw_time)
+  dat <- dat[YO,]
 
-  if (include_depth) {
-    dat$depth <- d$depth_scaled
-    dat$depth2 <- d$depth_scaled2
+  time_k <- as.numeric(as.factor(dat$raw_time))
+
+  #time_k <- dat$raw_time - min(dat$raw_time) + 1
+  k <- max(time_k)
+  time_lab <- paste0("Y", seq(1, k))
+  dat[time_lab] <- 0
+  dat[, time_lab[1]] <- 1
+  for (j in seq_along(time_lab)) {
+    dat[time_k == j, time_lab[j]] <- 1
   }
+  #names(dat)[names(dat)=="Y1"] <- "intercept"
 
-  # construct index for ar1 model:
+  # create an object for a Matérn model
+  spde <- INLA::inla.spde2.matern(mesh = mesh, alpha = 3 / 2) #, # WHAT DOES ALPHS DO?????
+  #                                prior.range = prior_range, 
+  #                                prior.sigma = prior_sigma)
+
+  # construct index for ar1 model ---- DON'T QUITE UNDERSTAND THIS
   iset <- INLA::inla.spde.make.index(
-    "i2D",
-    n.spde = mesh$n
+    name = "spatial.field",
+    n.spde = mesh$n,
+    n.group = k
   )
 
-  # make the covariates:
-  X.1 <- dat[, -c(1:3), drop = FALSE]
-  covar_names <- colnames(X.1)
-  XX.list <- as.list(X.1)
+  # isolate list of covariates:
+  X.1              <- dat[, -c(1:4), drop = FALSE] # remove "y", "xcoo", "ycoo", "real_time"
+  covar_names      <- colnames(X.1) # list covariates
+  XX.list          <- as.list(X.1) # turn data.frame to a list
   effect.list      <- list()
   effect.list[[1]] <- c(iset)
+
  for (Z in seq_len(ncol(X.1)))
     effect.list[[Z + 1]] <- XX.list[[Z]]
+
   names(effect.list) <- c("1", covar_names)
 
+ # projection matrix to map to observed points
  A <- INLA::inla.spde.make.A(
     mesh = mesh,
-    loc = cbind(dat$xcoo, dat$ycoo)
+    loc = cbind(dat$xcoo, dat$ycoo),
+    group = time_k
   )
@@
   A.list[[1]] <- A
+
   for (Z in seq_len(ncol(X.1)))
@@
   # make projection points stack:
-  sdat <- INLA::inla.stack(
-    tag = "stdata",
+  sdat <- INLA::inla.stack(tag = "stdata",
                            data = list(y = dat$y),
@@
     paste(covar_names, collapse = "+"),
-    "+ f(i2D, model=spde)"
+    "+ f(spatial.field, model=spde)"
   ))
@@
                    data = INLA::inla.stack.data(sdat),
-        control.predictor = list(
-          compute = TRUE,
-          A = INLA::inla.stack.A(sdat)
-        ),
-        control.fixed = list(
-          mean = 0, prec = 1 / (5^2),
-          mean.intercept = 0, prec.intercept = 1 / (20^2)
-        ),
+                   control.predictor = list(compute = TRUE,
+                                            A = INLA::inla.stack.A(sdat)),
+                   control.fixed     = list(mean = 0, 
+                                            prec = 1 / (5^2),
+                                            mean.intercept = 0, 
+                                            prec.intercept = 1 / (20^2)),
                    control.compute = list(config = TRUE),
@@
                    keep = FALSE
-      ))
+                 )
+      )
       trial_i <- trial_i + 1
@@
                        family = "gaussian",
-                       max_edge = c(20, 100),
-                       kmeans = FALSE,
-                       plot = FALSE,
+                       plot = TRUE,
                        fit_model = TRUE,
                        extend = list(n = 8, offset = -0.1),
-                       offset = c(5, 25),
-                       cutoff = 10,
-                       include_depth = TRUE,
+                       RangeGuess = 100, 
+                       covariates = c("depth_scaled", "depth_scaled2"),
                        verbose = FALSE,
@@
 
-pg <- gfplot:::make_prediction_grid(filter(temp_dat, year %in% c(2016, 2017)), survey ="SYN QCS")$grid
-
+pg <- gfplot:::make_prediction_grid(temp_dat, survey ="SYN QCS")$grid #filter(temp_dat, year %in% c(2016, 2017))
 cell_area <- gfplot:::make_prediction_grid(filter(temp_dat, year %in% c(2016, 2017)), survey = "SYN QCS")$cell_area
-
-pred_temp <- predict_inla_sens(obj=obj, pred_grid = pg , samples = 100L,
-                         include_depth = TRUE)
 ```
@@
 Project onto grid:
-
-```{r posterior-draws}
-pg_one_year <- filter(pg, year == min(pg$year))
-predicted <- predict_inla_sens(obj = obj, pred_grid = pg_one_year, samples = 100, include_depth = TRUE)
-```
 
@@
 
-samples = 10L
-pred_grid <- pg
+# samples = 10L
+# pred_grid <- pg_one_year
 
-predict_inla_sens <- function(obj, pred_grid, samples = 100L,
-                         include_depth = TRUE) {
-
+predict_inla_sens <- function(obj, pred_grid, 
+                                   samples = 100L) {
   mesh <- obj$mesh
@@
   latent_names   <- rownames(inla.mcmc[[1]]$latent)
-  re_indx <- grep("^i2D", latent_names)
-  if (include_depth) {
-    depth_fe_indx1 <- grep("^depth$", latent_names)
-    depth_fe_indx2 <- grep("^depth2$", latent_names)
-  }
-  intercept_indx <- grep("^intercept$", latent_names)
+  yr_fe_indx     <- grep("^Y[0-9]+$", latent_names)
+  re_indx        <- grep("^spatial.field", latent_names)
+  depth_fe_indx1 <- grep("^depth_scaled$", latent_names)
+  depth_fe_indx2 <- grep("^depth_scaled2$", latent_names)
+  #intercept_indx <- grep("^intercept$", latent_names)
 
@@
 
-  # projections will be stored as an array:
-  projected_latent_grid <-
-    array(0, dim = c(nrow(grid_locs), samples))
+  # define empty arrays to store projections:
+  latent_grid   <- array(0, dim = c(nrow(grid_locs), samples, length(yr_fe_indx)))
+  spatial_field <- array(0, dim = c(nrow(grid_locs), samples, length(yr_fe_indx)))
   proj_matrix   <- INLA::inla.spde.make.A(mesh, loc = as.matrix(grid_locs))
 
-  # do MCMC projections:
-  indx <- which(iset$i2D.group == 1)
+  # loop over all years; do MCMC projections for that year
+  for (yr in seq_along(yr_fe_indx)) {
+   indx <- which(iset$spatial.field.group == 1) 
       for (i in seq_len(samples)) {
+      if (yr > 1) {
+        year_diff <- inla.mcmc[[i]]$latent[yr_fe_indx[[yr]]]
+      } else {
+        year_diff <- 0
+      }
 
-    if (include_depth) {
-      depth_effects <-
+      # constant covariate effects across years (currently hard coded...)
+      fixed_covariate_effects <- 
         pred_grid[, "depth_scaled", drop = TRUE] * inla.mcmc[[i]]$latent[depth_fe_indx1] +
         pred_grid[, "depth_scaled2", drop = TRUE] * inla.mcmc[[i]]$latent[depth_fe_indx2]
-    } else {
-      depth_effects <- 0
+
+      # spatial temporal 
+      spatial_field[, i, yr] <-
+         as.numeric(proj_matrix %*% inla.mcmc[[i]]$latent[re_indx][indx]) +
+         #inla.mcmc[[i]]$latent[intercept_indx] + 
+         inla.mcmc[[i]]$latent[yr_fe_indx[[1]]] + # base year
+         year_diff # 0 for the base year, diff otherwise
+
+      # model predictions (current covariates are hard coded)   
+      latent_grid[, i, yr] <- spatial_field[, i, yr] + fixed_covariate_effects
    }
-
-    projected_latent_grid[, i] <-
-      as.numeric(proj_matrix %*% inla.mcmc[[i]]$latent[re_indx][indx]) +
-      inla.mcmc[[i]]$latent[intercept_indx] +
-      depth_effects
   }
-
-  projected_latent_grid
+  list(latent_grid = latent_grid, depth_effects = depth_effects, spatial_field = spatial_field)
 }
@@
 
-Calculate probability * Gamma median values on the grid:
+```{r posterior-draws}
+#pg_one_year <- filter(pg, year == min(pg$year))
+projected <- predict_inla_sens(obj = obj, pred_grid = pg, samples = 10)
+```
+
+Plot values on the grid:
 
 ```{r prediction-grid-calcs}
+attributes(projected)
 pg_one_year <- filter(pg, year == min(pg$year))
-attributes(predicted)
+attributes(projected$latent_grid)
 # out <- stats::plogis(p_bin) * exp(p_pos)
-out_med <- apply(predicted, c(1), median)
+out_med <- apply(projected$latent_grid, c(1,3), median)
 # out_bin <- stats::plogis(p_bin)
 # out_bin <- apply(out_bin, c(1, 3), median)
-dim(predicted)
+pg_all <- bind_rows(replicate(dim(projected$latent_grid)[3], pg, simplify = FALSE))
+out_long <- reshape2::melt(out_med)
+pg_all$time <- out_long$Var2
+pg_all$pred <- out_long$value
 
-pg_all <- bind_rows(replicate(1, pg_one_year, simplify = FALSE))
+# pg$year_k <- pg$year - min(pg$year) + 1
+# count_k <- unique(year_k)
 
-out_long <- reshape2::melt(out_med)
-pg_all$pred <- out_long$value
+```
+
+```{r prediction-grid-calcs}
+attributes(projected$spatial_field)
+# out <- stats::plogis(p_bin) * exp(p_pos)
+out_med_sp <- apply(projected$spatial_field, c(1,3), median)
+out_long_sp <- reshape2::melt(out_med_sp)
+pg_all$sp_field <- out_long_sp$value
 ```
@@
     scale_fill_viridis_c(option = "C") +
-    #facet_wrap(~ year) +
     geom_polygon(
@@
 
-make_plot(pg_all, "pred") + 
+make_plot(pg_all, "pred") + facet_wrap(~ time) +
   scale_fill_gradient2(
@@
     high = scales::muted("red"))
+
+median(pg_all$sp_field)
+make_plot(pg_all, "sp_field") + facet_wrap(~ time) +
+  scale_fill_gradient2(
+    midpoint = median(pg_all$sp_field),
+    high = scales::muted("blue"), 
+    mid = "white",
+    low = scales::muted("red"))
 

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
AR1 spatial field running in INLA to predict temperature for all 6 years

@@
 ```{r}
-testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2015, 2016, 2017, 2018))
-testdat <- gfplot:::interp_survey_bathymetry(testdat)
+testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))
+#testdat <- gfplot:::interp_survey_bathymetry(testdat)
 glimpse(testdat)
@@
 ```{r}
-dat2 <- dat %>% filter(temperature_C_avg!="NA")
+dat2 <- dat %>% filter(temperature_C_avg!="NA") %>% filter(ssid==1)
 
@@
 
-  coords <- as.matrix(unique(d[, c("X", "Y")]))
-
-  # use kmeans() to calculate centers:
-  if (kmeans) {
-    km <- stats::kmeans(x = coords, centers = n_knots)
-    mesh <- INLA::inla.mesh.create(km$centers,
-      extend = extend
-    )
-  } else {
-    bnd <- INLA::inla.nonconvex.hull(coords)
-
-    mesh <- INLA::inla.mesh.2d(
-      offset = offset,
-      boundary = bnd,
-      max.edge = max_edge,
-      cutoff = cutoff
-    )
-  }
-  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh$n)
-
-  if (plot) {
-    graphics::plot(mesh, asp = 1)
-    graphics::points(coords, col = "red")
-  }
-
    # create data.frame including response ("y"), coordinates("xcoo", "ycoo"), covariates, and "intercept"
@@
   k <- max(time_k)
+
   time_lab <- paste0("Y", seq(1, k))
@@
 
+  ## Select time knots and generate time mesh
+  t_knots = seq(1, k, length = k)
+  mesh1 = inla.mesh.1d(loc = t_knots, degree = 1, boundary = "free")
+
+  ## Generate spatial mesh
+  coords <- as.matrix(unique(d[, c("X", "Y")]))
+
+  # use kmeans() to calculate centers:
+  if (kmeans) {
+    km <- stats::kmeans(x = coords, centers = n_knots)
+    mesh <- INLA::inla.mesh.create(km$centers,
+      extend = extend
+    )
+  } else {
+    bnd <- INLA::inla.nonconvex.hull(coords)
+
+    mesh2d <- INLA::inla.mesh.2d(
+      offset = offset,
+      boundary = bnd,
+      max.edge = max_edge,
+      cutoff = cutoff
+    )
+  }
+  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh$n)
+
+  if (plot) {
+    graphics::plot(mesh, asp = 1)
+    graphics::points(coords, col = "red")
+  }
+
   # create an object for a Matérn model
-  spde <- INLA::inla.spde2.matern(mesh = mesh, alpha = 3 / 2) #, # WHAT DOES ALPHS DO?????
+  spde <- INLA::inla.spde2.matern(mesh = mesh2d, alpha = 3 / 2) #, # WHAT DOES ALPHA DO?????
   #                                prior.range = prior_range, 
@@
 
+  ## build the space time indices
+  # index = inla.spde.make.index("space", n.spde = spde$n.spde, n.group = mesh1$m)
   # construct index for ar1 model ---- DON'T QUITE UNDERSTAND THIS
@@
     name = "spatial.field",
-    n.spde = mesh$n,
-    n.group = k
+    n.spde = spde$n.spde, 
+    n.group = mesh1$m
   )
@@
  A <- INLA::inla.spde.make.A(
-    mesh = mesh,
+    mesh = mesh2d,
     loc = cbind(dat$xcoo, dat$ycoo),
-    group = time_k
+    group = time_k, group.mesh = mesh1
   )
@@
   )
+# 
+#   formula <- as.formula(paste(
+#     "y ~ -1 + ",
+#     paste(covar_names, collapse = "+"), 
+#     "+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))"))
+#
 
-  formula <- as.formula(paste(
-    "y ~ -1 +",
-    paste(covar_names, collapse = "+"),
-    "+ f(spatial.field, model=spde)"
-  ))
+  #covariates <- paste(covar_names, collapse = "+")
+
+  formula <- y ~ -1 + depth_scaled+depth_scaled2+Y1+Y2+Y3+Y4+Y5+Y6 +
+   f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))
+
+ # formula <- as.formula(paste(
+ #    "y ~ -1 + ",
+ #    paste(covar_names, collapse = "+"), "+ 
+ #      f(year_f, model = "rw2") +
+ #      f(year_f2, depth_scaled, model = "rw2") +
+ #      f(year_f3, depth_scaled2, model = "rw2") +
+ #      f(spatial.field, model=spde)"
+ #  ))
+ #
 
@@
 ```{r}
-
 # alternative to function, hard code
-
 # samples = 10L
@@
   spatial_field <- array(0, dim = c(nrow(grid_locs), samples, length(yr_fe_indx)))
-  proj_matrix   <- INLA::inla.spde.make.A(mesh, loc = as.matrix(grid_locs))
+  proj_matrix   <- INLA::inla.spde.make.A(mesh2d, loc = as.matrix(grid_locs))
 

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
tweak of markdown

@@
 
-```{r}
+```{r , eval=TRUE, include=FLASE}
 # alternative to function, hard code
@@
 attributes(projected$latent_grid)
-# out <- stats::plogis(p_bin) * exp(p_pos)
 out_med <- apply(projected$latent_grid, c(1,3), median)
-# out_bin <- stats::plogis(p_bin)
-# out_bin <- apply(out_bin, c(1, 3), median)
 pg_all <- bind_rows(replicate(dim(projected$latent_grid)[3], pg, simplify = FALSE))
@@
 pg_all$pred <- out_long$value
-
-# pg$year_k <- pg$year - min(pg$year) + 1
-# count_k <- unique(year_k)
-
 ```
 
-```{r prediction-grid-calcs}
+```{r }
 attributes(projected$spatial_field)
-# out <- stats::plogis(p_bin) * exp(p_pos)
 out_med_sp <- apply(projected$spatial_field, c(1,3), median)
@@
 }
-# make_plot(pg_all)
+
+```
+
+```{r}
 median(pg_all$pred)
-
 make_plot(pg_all, "pred") + facet_wrap(~ time) +
@@
     low = scales::muted("red"))
+```
 ```

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
fix formula for inla function

@@
 testdat <- tidy_survey_sets(raw, c("SYN HS","SYN QCS","SYN WCVI","SYN WCHG"), years = c(2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))
-#testdat <- gfplot:::interp_survey_bathymetry(testdat)
-glimpse(testdat)
-#View(testdat[[1]])
 ```
 
+Get sensor data
 ```{r}
@@
 sensor <- readRDS("data-cache/allsynoptic_sensor_07mar19_time2.rds")
+glimpse(sensor)
 # View(sensor)
@@
 
+Join sensor data with trawl data from tidy_survey_sets function
 ```{r}
-dat <- left_join(testdat[[1]], sensor, by="fishing_event_id", suffix = c("", ".sensor"))
-# dat <- left_join(testdat, sensor, by="fishing_event_id", suffix = c("", ".sensor"))
+dat <- left_join(testdat, sensor, by="fishing_event_id", suffix = c("", ".sensor"))
 dat$year_f <- factor(dat$year)
@@
 # glimpse(dat)
+```
+
+Optional: Interpolate missing depth data
+```{r, eval=FALSE}
+testdat2 <- gfplot:::interp_survey_bathymetry(testdat)
+glimpse(testdat2[[1]])
+# if bathymetry ran...
+dat <- left_join(testdat2[[1]], sensor, by="fishing_event_id", suffix = c("", ".sensor"))
+dat$year_f <- factor(dat$year)
+# View(dat)
+dat <- dat %>% mutate(do_range = do_mlpL_max - do_mlpL_min, salinity_range = salinity_PSU_max -
+  salinity_PSU_min, temp_range = temperature_C_max - temperature_C_min, depth_range = depth_m_max - depth_m_min)
 ```
@@
 ```{r}
-scaled <- scale_survey_predictors(dat)
+scaled <- dat %>% filter(depth != "NA") %>% scale_survey_predictors()
 # glimpse(scaled)
@@
 ```{r}
-ggplot(sens_dat, aes(depth)) + geom_histogram() 
-ggplot(sens_dat, aes(density*1000)) + geom_histogram() 
-ggplot(sens_dat, aes(depth, (density*1000))) + geom_point() 
+ggplot(temp_dat, aes(depth)) + geom_histogram() 
+ggplot(temp_dat, aes(density*1000)) + geom_histogram() 
+ggplot(temp_dat, aes(depth, (density*1000))) + geom_point() 
 #ggplot(synhs, aes(depth_m, log(density*1000+1))) + geom_point() 
@@
 time = "year"
+re_model_type = "rw1"
+formula = full_model
 
@@
                      fit_model = TRUE,
+                     re_model_type = "rw1",
                      extend = list(n = 8, offset = -0.1),
@@
                      debug = FALSE,
-                     trials = 10) {
+                     trials = 10,
+                     formula = full_model) {
 
-   # create data.frame including response ("y"), coordinates("xcoo", "ycoo"), covariates, and "intercept"
+  # create data.frame including response ("y"), coordinates("xcoo", "ycoo"), covariates, but not "intercept" as that will be input as time 1
   dat <- data.frame(
@@
 
-  # make a design matrix where the first year is the intercept:
+  # make sure dat is ordered temporally
   YO  <- order(dat$raw_time)
@@
 
+  # number years sequentially, removing placeholders for missing years
   time_k <- as.numeric(as.factor(dat$raw_time))
-
-  #time_k <- dat$raw_time - min(dat$raw_time) + 1
   k <- max(time_k)
 
-  time_lab <- paste0("Y", seq(1, k))
-  dat[time_lab] <- 0
-  dat[, time_lab[1]] <- 1
+  # make design matrix with first time as intercept
+  time_lab <- paste0("Y", seq(1, k)) # make vector of each year labeled "Y1", "Y2", ..."
+  dat[time_lab] <- 0 # fill empty matrix with 0
+  dat[, time_lab[1]] <- 1 # first time point assigned 1 for intercept
   for (j in seq_along(time_lab)) {
@@
   }
-  #names(dat)[names(dat)=="Y1"] <- "intercept"
 
-  ## Select time knots and generate time mesh
+  ## Generate time mesh
   t_knots = seq(1, k, length = k)
@@
   }
-  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh$n)
+  message("INLA max_edge = ", "c(", max_edge[[1]], ", ", max_edge[[2]], "), n = ", mesh2d$n)
 
   if (plot) {
-    graphics::plot(mesh, asp = 1)
+    graphics::plot(mesh2d, asp = 1)
     graphics::points(coords, col = "red")
@@
   )
-# 
-#   formula <- as.formula(paste(
-#     "y ~ -1 + ",
-#     paste(covar_names, collapse = "+"), 
-#     "+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))"))
-#
 
-  #covariates <- paste(covar_names, collapse = "+")
-
-  formula <- y ~ -1 + depth_scaled+depth_scaled2+Y1+Y2+Y3+Y4+Y5+Y6 +
-   f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model = "ar1"))
-
- # formula <- as.formula(paste(
- #    "y ~ -1 + ",
- #    paste(covar_names, collapse = "+"), "+ 
- #      f(year_f, model = "rw2") +
- #      f(year_f2, depth_scaled, model = "rw2") +
- #      f(year_f3, depth_scaled2, model = "rw2") +
- #      f(spatial.field, model=spde)"
- #  ))
- #
+  full_model <- as.formula(paste(
+    'y ~ -1 + ',
+    paste(covar_names, collapse = '+'),
+    '+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model ="', re_model_type,'"))'))
 
@@
                        fit_model = TRUE,
+                       re_model_type = "rw1",
                        extend = list(n = 8, offset = -0.1),
@@
                        debug = FALSE,
-                       trials = 2)
+                       trials = 2
+                       )
 
@@
 
-```{r , eval=TRUE, include=FLASE}
+```{r , eval=TRUE, include=FALSE}
 # alternative to function, hard code

### 2019-03-13 analysis/INLA/test-INLA-sensor.Rmd
experiment with depth mesh

@@
 time = "year"
-re_model_type = "rw1"
-formula = full_model
+sp_model_type = "ar1"
+formula = "full_model"
 
@@
                      fit_model = TRUE,
-                     re_model_type = "rw1",
+                     sp_model_type = "ar1",
                      extend = list(n = 8, offset = -0.1),
@@
 
-  ## Generate time mesh
-  t_knots = seq(1, k, length = k)
-  mesh1 = inla.mesh.1d(loc = t_knots, degree = 1, boundary = "free")
-
   ## Generate spatial mesh
@@
     km <- stats::kmeans(x = coords, centers = n_knots)
-    mesh <- INLA::inla.mesh.create(km$centers,
+    mesh2d <- INLA::inla.mesh.create(km$centers,
       extend = extend
@@
 
-  if (plot) {
-    graphics::plot(mesh2d, asp = 1)
-    graphics::points(coords, col = "red")
-  }
-
   # create an object for a Matérn model
@@
   # index = inla.spde.make.index("space", n.spde = spde$n.spde, n.group = mesh1$m)
-  # construct index for ar1 model ---- DON'T QUITE UNDERSTAND THIS
+  # construct index for ar1 model 
   iset <- INLA::inla.spde.make.index(
@@
     n.spde = spde$n.spde, 
-    n.group = mesh1$m
+    n.group = k
   )
+
+  ## Generate depth mesh
+  d_knots = seq(min(d$depth_scaled), max(d$depth_scaled), length.out = 30)
+  #mesh0 = inla.mesh.1d(loc = d_knots, degree = 0, boundary = c('neumann', 'free'))
+  mesh1 = inla.mesh.1d(loc = d_knots, degree = 1, boundary = c('neumann', 'free'))
+  #mesh2 = inla.mesh.1d(loc = d_knots, degree = 2, boundary = c('neumann', 'free'))
+    # graphics::plot(mesh0$loc,mesh0$mid)
+    # graphics::points(mesh1$loc,mesh1$mid)
+    # graphics::points(mesh2$loc,mesh2$mid, col = "red")
+
+  idepth <- INLA::inla.spde.make.index("depth.field", n.spde = mesh1$m, n.group = k)
+
+  if (plot) {
+    graphics::plot(mesh2d, asp = 1)
+    graphics::points(coords, col = "red")
+  }
 
@@
   effect.list[[1]] <- c(iset)
+  effect.list[[2]] <- c(idepth)
 
  for (Z in seq_len(ncol(X.1)))
-    effect.list[[Z + 1]] <- XX.list[[Z]]
+    effect.list[[Z + 2]] <- XX.list[[Z]]
 
-  names(effect.list) <- c("1", covar_names)
+  names(effect.list) <- c("1", "d_knots", covar_names)
 
@@
     loc = cbind(dat$xcoo, dat$ycoo),
-    group = time_k, group.mesh = mesh1
+    group = time_k
+  )
+ A2 <- INLA::inla.spde.make.A(
+    mesh = mesh1,
+    loc = dat$depth_scaled,
+    group = time_k
   )
@@
   A.list[[1]] <- A
+  A.list[[2]] <- A2
 
   for (Z in seq_len(ncol(X.1)))
-    A.list[[Z + 1]] <- 1
+    A.list[[Z + 2]] <- 1
 
@@
     paste(covar_names, collapse = '+'),
-    '+ f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model ="', re_model_type,'"))'))
+    '+f(depth_scaled, time_k, model = "rw1", scale.model = TRUE) + f(spatial.field, model = spde, group = spatial.field.group, control.group = list(model ="', sp_model_type,'"))'))
 
@@
   }
-  list(
-    model = model, mesh = mesh, spde = spde, data = dat,
+#  obj <- 
+    list(model = model, mesh = mesh2d, spde = spde, data = dat,
     formula = formula, iset = iset
@@
     midpoint = median(pg_all$sp_field),
-    high = scales::muted("blue"), 
+    high = scales::muted("red"), 
     mid = "white",
-    low = scales::muted("red"))
+    low = scales::muted("blue"))
 

### 2019-03-15 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
Tweaks to temperature fitting script
Tweak fit temp

@@
 d_withheld <- d_trawl
-spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 200)
+spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 250)
 plot_spde(spde)
@@
 
-ggplot(predictions5$data, aes(est, residuals)) +
-  geom_point() +
-  geom_smooth()
-
 ggplot(predictions$data, aes(depth_scaled, residuals)) +
@@
   time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
+  silent = FALSE)
 
@@
 predictions3_nd <- predict(m_temp3, newdata = sdmTMB::qcs_grid)
+saveRDS(predictions3_nd$data, file = "analysis/tmb-sensor-explore/qcs-temp.rds")
 

### 2019-03-16 analysis/tmb-sensor-explore/2-fit-temperature.Rmd

@@
 # Try with a third degree polynomial:
-m_temp3 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
+m_temp3 <- sdmTMB(d_fit, 
+  temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = FALSE)
+
+m_temp3_rw <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + poly(depth_scaled, 3),
+  include_spatial = FALSE, ar1_fields = TRUE,
   time = "year", spde = spde, family = gaussian(link = "identity"),
@@
 
-ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+predictions3_rw <- predict(m_temp3_rw)
+predictions3_rw$data$residuals <- residuals(m_temp3_rw)
+
+ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
   geom_point() +

### 2019-04-05 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
Identify stage in code where predict error is occuring

@@
 
-```{r}
+```{r setup}
 library(tidyverse)
@@
 set.seed(39201114)
-withheld_fe <- sample(d_trawl$fishing_event_id, size = round(nrow(d_trawl) * .2))
+```
+
+```{r}
+withheld_fe <- sample(d_trawl$fishing_event_id, size = round(nrow(d_trawl)*.2))
+
 d_withheld <- filter(d_trawl, fishing_event_id %in% withheld_fe)
+
 d_fit <- filter(d_trawl, !fishing_event_id %in% withheld_fe)
-nrow(d_fit)
-nrow(d_withheld)
+
+nrow(d_fit) # 3774
+nrow(d_withheld) # 943
 unique(d_withheld$year) 
 unique(d_fit$year)
+length(unique(d_fit$year))
 
@@
 # It's crashing TMB.
-d_fit <- d_trawl
-d_withheld <- d_trawl
+
+#d_fit <- d_trawl
+#d_withheld <- d_trawl
+```
+
+```{r}
 spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 250)
 plot_spde(spde)
+```
 
+```{r}
 m_temp1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
   # time_varying =  ~ depth_scaled + depth_scaled2,
-  time = "year", spde = spde, family = gaussian(link = "identity"), ar1_fields = FALSE,
-  anisotropy = FALSE, include_spatial = TRUE, silent = TRUE)
+  time = "year", spde = spde, family = gaussian(link = "identity"), 
+  ar1_fields = FALSE,
+  anisotropy = FALSE, 
+  include_spatial = TRUE, 
+  silent = FALSE)
 
+```
+
+```{r}
 m_temp1$model$par
-
 # 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
 exp(m_temp1$model$par[["ln_phi"]])
+```
 
+```{r}
+# FIXME: Error in `$<-.data.frame`(`*tmp*`, "est", value = c(7.12313541783276, :replacement has 4716 rows, data has 4717
 predictions <- predict(m_temp1)
+
+attributes(m_temp1)
+
+# find source of error in predict function
+object <- m_temp1
+
+ nd <- object$data # this has 3,774 rows of data
+ lp <- object$tmb_obj$env$last.par # this has 6583 length(lp)
+  r <- object$tmb_obj$report(lp)
+    length(r$omega_s) # 386
+    length(r$epsilon_st) # this is 6176
+    length(r$b_rw_t) # this has 16 time steps
+    length(r$eta_i) # this is a vector of 3773
+
+    nd$est <- r$eta_i # ERROR OCCURS HERE
+    nd$est_fe <- r$eta_fixed_i # ERROR OCCURS HERE
+
+attributes(r)
+
+```
+
+```{r}
 predictions$data$residuals <- residuals(m_temp1)
+```
 
+```{r}
 plot_map <- function(dat, column = "est") {
@@
 }
+```
 
+```{r}
 plot_map(predictions$data, "est") +
@@
   ggtitle("Prediction (fixed effects + all random effects)")
+```
 
+```{r}
 plot_map(predictions$data, "est_fe") +

### 2019-04-08 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
prelim comparison withheld data with model estimates... found error in how predict function deals with newdata

@@
 ```{r}
-# FIXME: Error in `$<-.data.frame`(`*tmp*`, "est", value = c(7.12313541783276, :replacement has 4716 rows, data has 4717
 predictions <- predict(m_temp1)
-
-attributes(m_temp1)
-
-# find source of error in predict function
-object <- m_temp1
-
- nd <- object$data # this has 3,774 rows of data
- lp <- object$tmb_obj$env$last.par # this has 6583 length(lp)
-  r <- object$tmb_obj$report(lp)
-    length(r$omega_s) # 386
-    length(r$epsilon_st) # this is 6176
-    length(r$b_rw_t) # this has 16 time steps
-    length(r$eta_i) # this is a vector of 3773
-
-    nd$est <- r$eta_i # ERROR OCCURS HERE
-    nd$est_fe <- r$eta_fixed_i # ERROR OCCURS HERE
-
-attributes(r)
-
 ```
@@
   scale_colour_viridis_c()
+```
 
+```{r}
 plot_map(predictions$data, "est_re_s") +
@@
   scale_colour_gradient2()
+```
 
+```{r}
 plot_map(predictions$data, "est_re_st") +
@@
   scale_colour_gradient2()
+```
 
+```{r}
 ggplot(predictions$data, aes(est, temperature_c)) +
@@
   geom_abline()
+```
 
+```{r}
 ggplot(predictions$data, aes(est, residuals)) +
@@
   facet_wrap(~year)
+```
 
+```{r}
 ggplot(predictions$data, aes(depth_scaled, residuals)) +
@@
   facet_wrap(~year)
+```
 
+```{r}
+ggplot(predictions$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
 ggplot(predictions$data, aes(depth_scaled, residuals)) +
@@
   geom_smooth()
+```
 
+```{r}
 ggplot(predictions$data, aes(X, Y, colour = residuals)) +
@@
   scale_color_gradient2()
+```
 
-# Looks a bit curved, let's try with a quadratic
+ Looks a bit curved, let's try with a quadratic
 
+```{r}
 m_temp2 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 2),
@@
 predictions2$data$residuals <- residuals(m_temp2)
+```
 
+```{r}
 ggplot(predictions2$data, aes(depth_scaled, residuals)) +
@@
   facet_wrap(~year)
+```
 
+```{r}
 ggplot(predictions2$data, aes(depth_scaled, residuals)) +
@@
   geom_smooth()
+```
 
+```{r}
 ggplot(predictions2$data, aes(X, Y, colour = residuals)) +
@@
   scale_color_gradient2()
+```
 
-# Try with a third degree polynomial:
+Try with a third degree polynomial:
+
+```{r}
 m_temp3 <- sdmTMB(d_fit, 
@@
 
+predictions3 <- predict(m_temp3)
+predictions3$data$residuals <- residuals(m_temp3)
+```
+
+```{r}
 m_temp3_rw <- sdmTMB(d_fit,
@@
 
-predictions3 <- predict(m_temp3)
-predictions3$data$residuals <- residuals(m_temp3)
-
 predictions3_rw <- predict(m_temp3_rw)
 predictions3_rw$data$residuals <- residuals(m_temp3_rw)
+```
 
+```{r}
 ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
@@
   facet_wrap(~year)
+```
 
+```{r}
 ggplot(predictions3$data, aes(depth_scaled, residuals)) +
@@
   geom_smooth()
+```
 
+```{r}
 ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
@@
   scale_color_gradient2()
+```
 
@@
 
+```{r}
 d_fit$log_temperature_c <- log(d_fit$temperature_c)
@@
 predictions4$data$residuals <- residuals(m_temp4)
+```
 
+```{r}
 ggplot(predictions4$data, aes(depth_scaled, residuals)) +
@@
   facet_wrap(~year)
+```
 
+```{r}
 ggplot(predictions4$data, aes(depth_scaled, residuals)) +
@@
   geom_smooth()
+```
 
+```{r}
 ggplot(predictions4$data, aes(X, Y, colour = residuals)) +
@@
 
+```{r}
 m_temp5 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_m, 3),
@@
 predictions5$data$residuals <- residuals(m_temp5)
+```
 
+```{r}
 ggplot(predictions5$data, aes(depth_scaled, residuals)) +
@@
   facet_wrap(~year)
+```
 
+```{r}
 ggplot(predictions5$data, aes(depth_scaled, residuals)) +
@@
   geom_smooth()
+```
 
@@
 
+```{r}
 ggplot(predictions5$data, aes(est, residuals)) +
@@
   facet_wrap(~year)
+```
 
+```{r}
 ggplot(predictions5$data, aes(est, residuals)) +
@@
   geom_smooth()
+```
 
+```{r}
 aic <- function(m) {
@@
 aic(m_temp5)
+```
 
@@
 
+```{r}
 ggplot(predictions3$data, aes(est, temperature_c)) +
@@
   geom_abline()
+```
 
 # R2
+
+```{r}
 cor(predictions3$data$est, predictions3$data$temperature_c)^2
+```
 
@@
 
+```{r}
 predictions3_nd <- predict(m_temp3, newdata = sdmTMB::qcs_grid)
-saveRDS(predictions3_nd$data, file = "analysis/tmb-sensor-explore/qcs-temp.rds")
+saveRDS(predictions3_nd$data, file = "qcs-temp.rds")
+```
 
+```{r}
 ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
@@
   scale_color_gradient2()
+```
 
 # A short function for plotting our predictions:
+
+```{r}
 plot_map <- function(dat, column = "est") {
@@
 }
+```
 
+```{r}
 plot_map(predictions3_nd$data, "est") +
@@
   ggtitle("Prediction (fixed effects + all random effects)")
+```
 
+```{r}
 plot_map(predictions3_nd$data, "est_fe") +
@@
   scale_fill_viridis_c(trans = "sqrt")
+```
 
+```{r}
 plot_map(predictions3_nd$data, "est_re_s") +
@@
   scale_fill_gradient2()
+```
 
+```{r}
 plot_map(predictions3_nd$data, "est_re_st") +
@@
   scale_fill_gradient2()
+```
 
+```{r}
+m_temp3$model$par
+# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
+exp(m_temp3$model$par[["ln_phi"]])
+
+p_withheld <- predict(m_temp3, newdata = d_withheld)
+
+#FIXME: p_withheld$data$temperature_c is all zeros?!?!?! 
+# temp work around (first make sure both df in same order)
+arrange(p_withheld$data, fishing_event_id) 
+arrange(d_withheld, fishing_event_id) 
+p_withheld$data$temperature_c <- d_withheld$temperature_c
+```
+
+```{r}
+ggplot(p_withheld$data, aes(est, temperature_c)) +
+  geom_point() +
+  facet_wrap(~year) +
+  coord_fixed() +
+  xlim(3,12) +
+  ylim(3,12) +
+  geom_abline()
+```
+
+```{r}
 temp_cog <- get_cog(predictions3_nd)
@@
   scale_color_viridis_c()
+```
 
+```{r}
 temp_cog %>%

### 2019-04-08 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
change seed and add model configurations

@@
 
-set.seed(39201114)
+#set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
+set.seed(392011149)
 ```
@@
 ```{r}
+m_temp3_rw_sp <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + poly(depth_scaled, 3),
+  include_spatial = TRUE, ar1_fields = TRUE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = FALSE)
+
+predictions3_rw_sp <- predict(m_temp3_rw)
+predictions3_rw_sp$data$residuals <- residuals(m_temp3_rw)
+```
+
+```{r}
+m_temp3_st <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + poly(depth_scaled, 3),
+  include_spatial = TRUE, ar1_fields = TRUE, spatial_trend = TRUE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = FALSE)
+
+predictions3_st <- predict(m_temp3_st)
+predictions3_st$data$residuals <- residuals(m_temp3_st)
+```
+
+```{r}
+ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
 ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(predictions3_st$data, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 d_fit$log_temperature_c <- log(d_fit$temperature_c)
-m_temp4 <- sdmTMB(d_fit, log_temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
+m_temp4 <- sdmTMB(d_fit, 
+  log_temperature_c ~ 0 + as.factor(year),
+  time_varying = ~ 0 + poly(depth_scaled, 3),
+  #time_varying =  ~ 0 + depth_scaled,
+  include_spatial = TRUE, ar1_fields = TRUE, spatial_trend = TRUE,
   time = "year", spde = spde, family = gaussian(link = "identity"),
@@
 aic(m_temp3)
+aic(m_temp3_rw)
+aic(m_temp3_st)
 aic(m_temp4)
@@
 predictions3_nd <- predict(m_temp3, newdata = sdmTMB::qcs_grid)
-saveRDS(predictions3_nd$data, file = "qcs-temp.rds")
+#saveRDS(predictions3_nd$data, file = "qcs-temp.rds")
 ```
@@
 p_withheld$data$temperature_c <- d_withheld$temperature_c
+
+cor(p_withheld$data$est, p_withheld$data$temperature_c)^2 # 0.88 in first 
 ```
@@
 ```{r}
-ggplot(p_withheld$data, aes(est, temperature_c)) +
+ggplot(p_withheld$data, aes(temperature_c,est)) +
   geom_point() +
@@
   coord_fixed() +
-  xlim(3,12) +
-  ylim(3,12) +
+  xlim(3,15) +
+  ylim(3,15) +
+  geom_abline()
+```
+
+```{r}
+m_temp3_rw$model$par
+# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
+exp(m_temp3_rw$model$par[["ln_phi"]])
+
+p_withheld_rw <- predict(m_temp3_rw, newdata = d_withheld)
+
+#FIXME: p_withheld$data$temperature_c is all zeros?!?!?! 
+# temp work around (first make sure both df in same order)
+arrange(p_withheld_rw$data, fishing_event_id) 
+arrange(d_withheld, fishing_event_id) 
+p_withheld_rw$data$temperature_c <- d_withheld$temperature_c
+
+cor(p_withheld_rw$data$est, p_withheld_rw$data$temperature_c)^2 # 0.86?
+```
+
+```{r}
+ggplot(p_withheld_rw$data, aes(temperature_c,est)) +
+  geom_point() +
+  facet_wrap(~year) +
+  coord_fixed() +
+  xlim(3,15) +
+  ylim(3,15) +
+  geom_abline()
+```
+
+```{r}
+m_temp3_st$model$par
+# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
+exp(m_temp3_st$model$par[["ln_phi"]])
+
+p_withheld_st <- predict(m_temp3_st, newdata = d_withheld)
+
+#FIXME: p_withheld$data$temperature_c is all zeros?!?!?! 
+# temp work around (first make sure both df in same order)
+arrange(p_withheld_st$data, fishing_event_id) 
+arrange(d_withheld, fishing_event_id) 
+p_withheld_st$data$temperature_c <- d_withheld$temperature_c
+
+cor(p_withheld_st$data$est, p_withheld_st$data$temperature_c)^2
+```
+
+```{r}
+ggplot(p_withheld_st$data, aes(temperature_c,est)) +
+  geom_point() +
+  facet_wrap(~year) +
+  coord_fixed() +
+  xlim(3,15) +
+  ylim(3,15) +
+  geom_abline()
+```
+
+```{r}
+m_temp4$model$par
+# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
+exp(m_temp4$model$par[["ln_phi"]])
+
+p_withheld4 <- predict(m_temp4, newdata = d_withheld)
+
+#FIXME: p_withheld$data$temperature_c is all zeros?!?!?! 
+# temp work around (first make sure both df in same order)
+arrange(p_withheld4$data, fishing_event_id) 
+arrange(d_withheld, fishing_event_id) 
+p_withheld4$data$temperature_c <- d_withheld$temperature_c
+
+cor(p_withheld4$data$est, p_withheld4$data$temperature_c)^2 # 0.86
+```
+
+```{r}
+ggplot(p_withheld4$data, aes(log(temperature_c),est)) +
+  geom_point() +
+  facet_wrap(~year) +
+  coord_fixed() +
+  xlim(1, 3) +
+  ylim(1, 3) +
   geom_abline()

### 2019-04-09 analysis/tmb-sensor-explore/2-fit-temperature-aic.Rmd
aic selection of temp model

# Fit spatial temporal models to bottom temperature

```{r setup}
library(tidyverse)
library(sdmTMB)

d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))

set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
#set.seed(392011149)
d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
```

```{r}
withheld_fe <- sample(d_trawl$fishing_event_id, size = round(nrow(d_trawl)*.2))

d_withheld <- filter(d_trawl, fishing_event_id %in% withheld_fe)

d_fit <- filter(d_trawl, !fishing_event_id %in% withheld_fe)

nrow(d_fit) # 3774
nrow(d_withheld) # 943
unique(d_withheld$year) 
unique(d_fit$year)
length(unique(d_fit$year))

# FIXME: need to work out a bug in production on this new data set
# It's crashing TMB.

#d_fit <- d_trawl
#d_withheld <- d_trawl
```

```{r}
spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 250)
plot_spde(spde)
```

Functions for model exploration:
```{r}
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", colour = column)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed()
}

aic <- function(m) {
  k <- length(m$model$par)
  nll <- m$model$objective
  2 * k - 2 * (-nll)
}
```

Various candidate models:

```{r}
m_temp1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
  time = "year", spde = spde, family = gaussian(link = "identity"), 
  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
  anisotropy = FALSE, 
  silent = TRUE)

predictions <- predict(m_temp1)
predictions$data$residuals <- residuals(m_temp1)
cor(predictions$data$est, predictions$data$temperature_c)^2
aic(m_temp1)
```

AR1 fields slightly improves linear model AIC, but not R^2
```{r}
m_temp1_ar1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
  time = "year", spde = spde, family = gaussian(link = "identity"), 
  ar1_fields = TRUE, include_spatial = TRUE, 
  anisotropy = FALSE, 
  silent = TRUE)

predictions_ar1 <- predict(m_temp1_ar1)
predictions_ar1$data$residuals <- residuals(m_temp1_ar1)
cor(predictions_ar1$data$est, predictions_ar1$data$temperature_c)^2
aic(m_temp1_ar1)
```

Time-varying slope for depth effect makes huge improvement in AIC, but none for R^2
```{r}
m_temp1_rw <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year),   
  time_varying = ~ 0 + depth_scaled,
  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions1_rw <- predict(m_temp1_rw)
predictions1_rw$data$residuals <- residuals(m_temp1_rw)
cor(predictions1_rw$data$est, predictions1_rw$data$temperature_c)^2
aic(m_temp1_rw)
```

AR1 field in addition to time-varying slope does NOT improve AIC
```{r}
m_temp1_rw_ar1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year),   
  time_varying = ~ 0 + depth_scaled,
  ar1_fields = TRUE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions1_rw_ar1 <- predict(m_temp1_rw_ar1)
predictions1_rw_ar1$data$residuals <- residuals(m_temp1_rw_ar1)
cor(predictions1_rw_ar1$data$est, predictions1_rw_ar1$data$temperature_c)^2
aic(m_temp1_rw_ar1)
```

Time-varying has much better AIC than spatial trend varying
```{r}
m_temp1_st <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year) + depth_scaled,
  #time_varying =  ~ 0 + depth_scaled,
  include_spatial = TRUE, ar1_fields = FALSE, spatial_trend = TRUE, 
  # seems to be only config that converges
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions1_st <- predict(m_temp1_st)
predictions1_st$data$residuals <- residuals(m_temp1_st)
cor(predictions1_st$data$est, predictions1_st$data$temperature_c)^2 # 
aic(m_temp1_st)
```

Both time and space varying... doesn't improve any indices
```{r}
m_temp1_rw_st <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year), 
  time_varying =  ~ 0 + depth_scaled,
  include_spatial = TRUE, ar1_fields = FALSE, spatial_trend = TRUE, 
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions1_rw_st <- predict(m_temp1_rw_st)
predictions1_rw_st$data$residuals <- residuals(m_temp1_rw_st)
cor(predictions1_rw_st$data$est, predictions1_rw_st$data$temperature_c)^2 # 
aic(m_temp1_rw_st)
```

What about polynomial response to depth?

Ignore simplest configurations because both time-varying, and a fixed spatial field, seem critically important.

```{r, eval=FALSE}
m_temp2 <- sdmTMB(d_fit, 
  temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 2),
  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2 <- predict(m_temp2)
predictions2$data$residuals <- residuals(m_temp2)
cor(predictions2$data$est, predictions2$data$temperature_c)^2
aic(m_temp2)
```

```{r, eval=FALSE}
m_temp2_nosp <- sdmTMB(d_fit, 
  temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 2),
  ar1_fields = FALSE, include_spatial = FALSE, 
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2_nosp <- predict(m_temp2_nosp)
predictions2_nosp$data$residuals <- residuals(m_temp2_nosp)
cor(predictions2_nosp$data$est, predictions2_nosp$data$temperature_c)^2
aic(m_temp2_nosp)
```

```{r}
m_temp2_rw <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year),   
  time_varying = ~ 0 + poly(depth_scaled, 2),
  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2_rw <- predict(m_temp2_rw)
predictions2_rw$data$residuals <- residuals(m_temp2_rw)
cor(predictions2_rw$data$est, predictions2_rw$data$temperature_c)^2
aic(m_temp2_rw)
```

```{r, eval=FALSE}
m_temp2_rw_nosp <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled, 2),
  include_spatial = FALSE, ar1_fields = FALSE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2_rw_nosp <- predict(m_temp2_rw_nosp)
predictions2_rw_nosp$data$residuals <- residuals(m_temp2_rw_nosp)
cor(predictions2_rw_nosp$data$est, predictions2_rw_nosp$data$temperature_c)^2
aic(m_temp2_rw_nosp)
```

```{r, eval=FALSE}
m_temp2_st <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled, 2),
  include_spatial = TRUE, ar1_fields = FALSE, spatial_trend = TRUE, 
  # seems to be only config that converges
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions2_st <- predict(m_temp2_st)
predictions2_st$data$residuals <- residuals(m_temp2_st)
cor(predictions2_st$data$est, predictions2_st$data$temperature_c)^2 # 
aic(m_temp2_st)
```

What about 3 order polynomial?
```{r, eval=FALSE}
m_temp3 <- sdmTMB(d_fit, 
  temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions3 <- predict(m_temp3)
predictions3$data$residuals <- residuals(m_temp3)

cor(predictions3$data$est, predictions3$data$temperature_c)^2 # 
```

```{r}
m_temp3_rw <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled, 3),
  ar1_fields = FALSE, include_spatial = TRUE, 
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions3_rw <- predict(m_temp3_rw)
predictions3_rw$data$residuals <- residuals(m_temp3_rw)
cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 # 
aic(m_temp3_rw)
```

```{r, eval = FALSE}
m_temp3_rw_nosp <- sdmTMB(d_fit,
  temperature_c ~ 0 + as.factor(year),
  time_varying =  ~ 0 + poly(depth_scaled, 3),
  include_spatial = FALSE, ar1_fields = FALSE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE)

predictions3_rw_nosp <- predict(m_temp3_rw_nosp)
predictions3_rw_nosp$data$residuals <- residuals(m_temp3_rw_nosp)
cor(predictions3_rw_nosp$data$est, predictions3_rw_nosp$data$temperature_c)^2 # 
```

# Conclusion: log depth, both 2nd and 3rd order polynomials, that is time-varying

```{r}
ggplot(predictions2_rw$data, aes(temperature_c, est)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  geom_abline()
```

Residual plots:
```{r}
ggplot(predictions2_rw$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3_rw$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions2_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions2_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

# R2

```{r}
cor(predictions2_rw$data$est, predictions2_rw$data$temperature_c)^2
```

# Extrapolate:

#FIXME: this is crashing R
```{r}
# predictions_nd <- predict(m_temp3_rw, newdata = sdmTMB::qcs_grid)
#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
```

# A short function for plotting our predictions:

```{r}
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

```{r}
plot_map(predictions_nd$data, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
plot_map(predictions_nd$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

```{r}
plot_map(predictions_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()
```

```{r}
plot_map(predictions_nd$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()
```

```{r}
# m_temp2_rw$model$par
# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
# exp(m_temp2_rw$model$par[["ln_phi"]])

p_withheld2_rw <- predict(m_temp2_rw, newdata = d_withheld)
cor(p_withheld2_rw$data$est, p_withheld2_rw$data$temperature_c)^2 # 0.87 in first 
```

```{r}
ggplot(p_withheld2_rw$data, aes(temperature_c,est)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  xlim(3,15) +
  ylim(3,15) +
  geom_abline()
```

```{r}
# m_temp3_rw$model$par
# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
# exp(m_temp3_rw$model$par[["ln_phi"]])

p_withheld3_rw <- predict(m_temp3_rw, newdata = d_withheld)

#FIXME: p_withheld$data$temperature_c is all zeros?!?!?! 
# temp work around (first make sure both df in same order)
# arrange(p_withheld3_rw$data, fishing_event_id) 
# arrange(d_withheld, fishing_event_id) 
# p_withheld3_rw$data$temperature_c <- d_withheld$temperature_c

cor(p_withheld3_rw$data$est, p_withheld3_rw$data$temperature_c)^2 # 0.87
```

```{r}
ggplot(p_withheld2_rw$data, aes(temperature_c,est)) +
  geom_point() +
  facet_wrap(~year) +
  coord_fixed() +
  xlim(3,15) +
  ylim(3,15) +
  geom_abline()
```

```{r}
plot_map(predictions2_rw$data, "est") +
  scale_colour_viridis_c() +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
plot_map(predictions$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_colour_gradient2()
```

```{r}
plot_map(predictions$data, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_colour_gradient2()
```

```{r}
temp_cog <- get_cog(predictions_nd)

temp_cog %>% reshape2::dcast(year ~ coord, value.var = "est") %>%
  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
  scale_color_viridis_c()
```

```{r}
temp_cog %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  facet_wrap(~coord, scales = "free_y")
```

### 2019-04-11 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
candidate temperature models without poly(depth_scaled, 2) errors

--- 2019-04-08 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 
-set.seed(39201114)
+#set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
+set.seed(392011149)
+d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
+d_trawl$depth2 <- (d_trawl$depth_scaled)^2
+d_trawl$depth3 <- (d_trawl$depth_scaled)^3
 ```
@@
 withheld_fe <- sample(d_trawl$fishing_event_id, size = round(nrow(d_trawl)*.2))
-
 d_withheld <- filter(d_trawl, fishing_event_id %in% withheld_fe)
-
 d_fit <- filter(d_trawl, !fishing_event_id %in% withheld_fe)
@@
 ```{r}
-spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 250)
+spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 100)
 plot_spde(spde)
@@
 
-```{r}
-m_temp1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
-  # time_varying =  ~ depth_scaled + depth_scaled2,
-  time = "year", spde = spde, family = gaussian(link = "identity"), 
-  ar1_fields = FALSE,
-  anisotropy = FALSE, 
-  include_spatial = TRUE, 
-  silent = FALSE)
-
-```
-
-```{r}
-m_temp1$model$par
-# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
-exp(m_temp1$model$par[["ln_phi"]])
-```
-
-```{r}
-predictions <- predict(m_temp1)
-```
-
-```{r}
-predictions$data$residuals <- residuals(m_temp1)
-```
-
+Functions for model exploration:
 ```{r}
@@
 }
+
+aic <- function(m) {
+  k <- length(m$model$par)
+  nll <- m$model$objective
+  2 * k - 2 * (-nll)
+}
+```
+
+Various candidate models:
+
+```{r}
+m_temp1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
+  time = "year", spde = spde, family = gaussian(link = "identity"), 
+  ar1_fields = FALSE, include_spatial = FALSE, 
+  silent = TRUE)
+
+predictions <- predict(m_temp1)
+predictions$data$residuals <- residuals(m_temp1)
+cor(predictions$data$est, predictions$data$temperature_c)^2
+aic(m_temp1)
+
+m_temp1_nosp  <- m_temp1
+```
+
+AR1 fields no improvement in R^2
+```{r, eval=FALSE}
+m_temp1_ar1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
+  time = "year", spde = spde, family = gaussian(link = "identity"), 
+  ar1_fields = TRUE, include_spatial = FALSE, 
+  anisotropy = FALSE, 
+  silent = TRUE)
+
+predictions_ar1 <- predict(m_temp1_ar1)
+predictions_ar1$data$residuals <- residuals(m_temp1_ar1)
+cor(predictions_ar1$data$est, predictions_ar1$data$temperature_c)^2
+aic(m_temp1_ar1)
+```
+
+Time-varying slope for depth, little imporvement for R^2
+```{r}
+m_temp1_rw <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year),   
+  time_varying = ~ 0 + depth_scaled,
+  ar1_fields = FALSE, include_spatial = FALSE, #these are defaults
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions1_rw <- predict(m_temp1_rw)
+predictions1_rw$data$residuals <- residuals(m_temp1_rw)
+cor(predictions1_rw$data$est, predictions1_rw$data$temperature_c)^2
+aic(m_temp1_rw)
+```
+
+NO improvement
+```{r, eval=FALSE}
+m_temp1_rw_ar1 <- sdmTMB(d_fit, 
+  temperature_c ~ 0 + as.factor(year),   
+  time_varying = ~ 0 + depth_scaled,
+  ar1_fields = TRUE, include_spatial = TRUE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions1_rw_ar1 <- predict(m_temp1_rw_ar1)
+predictions1_rw_ar1$data$residuals <- residuals(m_temp1_rw_ar1)
+cor(predictions1_rw_ar1$data$est, predictions1_rw_ar1$data$temperature_c)^2
+aic(m_temp1_rw_ar1)
+```
+
+NO improvement
+```{r, eval=FALSE}
+m_temp1_st <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year) + depth_scaled,
+  #time_varying =  ~ 0 + depth_scaled,
+  include_spatial = TRUE, ar1_fields = FALSE, spatial_trend = TRUE, 
+  # seems to be only config that converges
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions1_st <- predict(m_temp1_st)
+predictions1_st$data$residuals <- residuals(m_temp1_st)
+cor(predictions1_st$data$est, predictions1_st$data$temperature_c)^2 # 
+aic(m_temp1_st)
+```
+
+Both time and space varying... doesn't improve R^2
+```{r, eval=FALSE}
+m_temp1_rw_st <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year), 
+  time_varying =  ~ 0 + depth_scaled,
+  include_spatial = TRUE, ar1_fields = FALSE, spatial_trend = TRUE, 
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions1_rw_st <- predict(m_temp1_rw_st)
+predictions1_rw_st$data$residuals <- residuals(m_temp1_rw_st)
+cor(predictions1_rw_st$data$est, predictions1_rw_st$data$temperature_c)^2 # 
+aic(m_temp1_rw_st)
+```
+
+What about polynomial response to depth?
+
+```{r, eval=FALSE}
+m_temp2 <- sdmTMB(d_fit, 
+  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2,
+  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions2 <- predict(m_temp2)
+predictions2$data$residuals <- residuals(m_temp2)
+cor(predictions2$data$est, predictions2$data$temperature_c)^2
+aic(m_temp2)
 ```
@@
 ```{r}
-plot_map(predictions$data, "est") +
-  scale_colour_viridis_c() +
-  ggtitle("Prediction (fixed effects + all random effects)")
+m_temp2_nosp <- sdmTMB(d_fit, 
+  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2,
+  ar1_fields = FALSE, include_spatial = FALSE, 
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions2_nosp <- predict(m_temp2_nosp)
+predictions2_nosp$data$residuals <- residuals(m_temp2_nosp)
+cor(predictions2_nosp$data$est, predictions2_nosp$data$temperature_c)^2
+aic(m_temp2_nosp)
 ```
@@
 ```{r}
-plot_map(predictions$data, "est_fe") +
-  ggtitle("Prediction (fixed effects only)") +
-  scale_colour_viridis_c()
+m_temp2_rw_nosp <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + depth_scaled + depth2,
+  include_spatial = FALSE, ar1_fields = FALSE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions2_rw_nosp <- predict(m_temp2_rw_nosp)
+predictions2_rw_nosp$data$residuals <- residuals(m_temp2_rw_nosp)
+cor(predictions2_rw_nosp$data$est, predictions2_rw_nosp$data$temperature_c)^2
+aic(m_temp2_rw_nosp)
+```
+
+NO improvement 
+```{r, eval=FALSE}
+m_temp2_st <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year), #+ poly(depth_scaled, 2),
+  time_varying =  ~ 0 + poly(depth_scaled, 2),
+  include_spatial = FALSE, ar1_fields = FALSE, spatial_trend = TRUE, 
+  # seems to be only config that converges
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions2_st <- predict(m_temp2_st)
+predictions2_st$data$residuals <- residuals(m_temp2_st)
+cor(predictions2_st$data$est, predictions2_st$data$temperature_c)^2 # 
+aic(m_temp2_st)
+```
+
+What about 3 order polynomial?
+```{r, eval=FALSE}
+m_temp3 <- sdmTMB(d_fit, 
+  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2 + depth3,
+  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions3 <- predict(m_temp3)
+predictions3$data$residuals <- residuals(m_temp3)
+
+cor(predictions3$data$est, predictions3$data$temperature_c)^2 # 
+```
+
+Slightly better without a fixed spatial field!
+```{r}
+m_temp3_nosp <- sdmTMB(d_fit, 
+  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2 + depth3,
+  ar1_fields = FALSE, include_spatial = FALSE, 
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions3_nosp <- predict(m_temp3_nosp)
+predictions3_nosp$data$residuals <- residuals(m_temp3_nosp)
+
+cor(predictions3_nosp$data$est, predictions3_nosp$data$temperature_c)^2 # 
+```
+
+Better without AR1
+```{r, eval=FALSE}
+m_temp3_ar1 <- sdmTMB(d_fit, 
+  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2 + depth3,
+  ar1_fields = TRUE, include_spatial = FALSE, 
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions3_ar1 <- predict(m_temp3_ar1)
+predictions3_ar1$data$residuals <- residuals(m_temp3_ar1)
+
+cor(predictions3_ar1$data$est, predictions3_ar1$data$temperature_c)^2 # 
+```
+
+```{r, eval=FALSE}
+m_temp3_rw <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + depth_scaled + depth2 + depth3,
+  ar1_fields = FALSE, include_spatial = TRUE, 
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions3_rw <- predict(m_temp3_rw)
+predictions3_rw$data$residuals <- residuals(m_temp3_rw)
+cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 # 
+aic(m_temp3_rw)
+```
+
+Also better without fixed spatial field but not quite as good as 2nd order
+```{r}
+m_temp3_rw_nosp <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + depth_scaled + depth2 + depth3,
+  include_spatial = FALSE, ar1_fields = FALSE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions3_rw_nosp <- predict(m_temp3_rw_nosp)
+predictions3_rw_nosp$data$residuals <- residuals(m_temp3_rw_nosp)
+cor(predictions3_rw_nosp$data$est, predictions3_rw_nosp$data$temperature_c)^2 # 
+```
+
+Not converging 
+```{r, eval=FALSE}
+m_temp3_st <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year), #+ poly(depth_scaled, 2),
+  time_varying =  ~ 0 + depth_scaled + depth2 + depth3,
+  include_spatial = FALSE, ar1_fields = FALSE, spatial_trend = TRUE, 
+  # seems to be only config that converges
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions3_st <- predict(m_temp3_st)
+predictions3_st$data$residuals <- residuals(m_temp3_st)
+cor(predictions3_st$data$est, predictions3_st$data$temperature_c)^2 # 
+```
+
+# Conclusion:
+
+```{r}
+# m_temp2_rw$model$par
+# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
+# exp(m_temp2_rw$model$par[["ln_phi"]])
+
+p_withheld1 <- predict(m_temp1_nosp, newdata = d_withheld)
+cor(p_withheld1$data$est, p_withheld1$data$temperature_c)^2 # 0.87 in first 
+```
+```{r}
+p_withheld1_rw <- predict(m_temp1_rw, newdata = d_withheld)
+cor(p_withheld1_rw$data$est, p_withheld1_rw$data$temperature_c)^2 # 0.87 in first 
+```
+
+# 4
+```{r}
+p_withheld2 <- predict(m_temp2_nosp, newdata = d_withheld)
+cor(p_withheld2$data$est, p_withheld2$data$temperature_c)^2 # 
+```
+
+# 2
+```{r}
+p_withheld2_rw <- predict(m_temp2_rw_nosp, newdata = d_withheld)
+cor(p_withheld2_rw$data$est, p_withheld2_rw$data$temperature_c)^2 # 
+```
+
+# 3
+```{r}
+p_withheld3 <- predict(m_temp3_nosp, newdata = d_withheld)
+cor(p_withheld3$data$est, p_withheld3$data$temperature_c)^2 # 0.87
+```
+
+# 1
+```{r}
+p_withheld3_rw <- predict(m_temp3_rw_nosp, newdata = d_withheld)
+cor(p_withheld3_rw$data$est, p_withheld3_rw$data$temperature_c)^2 # 0.87
 ```
@@
 ```{r}
-plot_map(predictions$data, "est_re_s") +
-  ggtitle("Spatial random effects only") +
-  scale_colour_gradient2()
-```
-
-```{r}
-plot_map(predictions$data, "est_re_st") +
-  ggtitle("Spatiotemporal random effects only") +
-  scale_colour_gradient2()
-```
-
-```{r}
-ggplot(predictions$data, aes(est, temperature_c)) +
+ggplot(p_withheld3_rw$data, aes(temperature_c,est)) +
   geom_point() +
@@
   coord_fixed() +
+  xlim(3,15) +
+  ylim(3,15) +
   geom_abline()
@@
 
+Residual plots:
 ```{r}
-ggplot(predictions$data, aes(est, residuals)) +
+ggplot(predictions3_rw_nosp$data, aes(est, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions$data, aes(depth_scaled, residuals)) +
+ggplot(predictions2_rw_nosp$data, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions$data, aes(depth_scaled, residuals)) +
+ggplot(predictions2_rw_nosp$data, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions$data, aes(X, Y, colour = residuals)) +
-  geom_point() +
-  facet_wrap(~year) +
-  coord_fixed() +
-  scale_color_gradient2()
-```
-
- Looks a bit curved, let's try with a quadratic
-
-```{r}
-m_temp2 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 2),
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions2 <- predict(m_temp2)
-predictions2$data$residuals <- residuals(m_temp2)
-```
-
-```{r}
-ggplot(predictions2$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions2$data, aes(depth_scaled, residuals)) +
+ggplot(predictions3_rw_nosp$data, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 
-```{r}
-ggplot(predictions2$data, aes(X, Y, colour = residuals)) +
-  geom_point() +
-  facet_wrap(~year) +
-  coord_fixed() +
-  scale_color_gradient2()
-```
+# Extrapolate:
 
-Try with a third degree polynomial:
+#FIXME: this is crashing R
 
 ```{r}
-m_temp3 <- sdmTMB(d_fit, 
-  temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
+m_temp2_rw_poly <- sdmTMB(d_fit,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + poly(depth_scaled,2),
+  include_spatial = FALSE, ar1_fields = FALSE,
   time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = FALSE)
-
-predictions3 <- predict(m_temp3)
-predictions3$data$residuals <- residuals(m_temp3)
+  silent = TRUE)
 ```
@@
 ```{r}
-m_temp3_rw <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year),
-  time_varying =  ~ 0 + poly(depth_scaled, 3),
-  include_spatial = FALSE, ar1_fields = TRUE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = FALSE)
-
-predictions3_rw <- predict(m_temp3_rw)
-predictions3_rw$data$residuals <- residuals(m_temp3_rw)
-```
-
-```{r}
-ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions3$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth()
-```
-
-```{r}
-ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
-  geom_point() +
-  facet_wrap(~year) +
-  coord_fixed() +
-  scale_color_gradient2()
-```
-
-# That's getting pretty straight although it looks a bit heteroscedastic
-
-# Perhaps this would work better with log temperature as the response
-# since it basically isn't going to go lower than 0 degree celsius
-
-```{r}
-d_fit$log_temperature_c <- log(d_fit$temperature_c)
-m_temp4 <- sdmTMB(d_fit, log_temperature_c ~ 0 + as.factor(year) + poly(depth_scaled, 3),
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions4 <- predict(m_temp4)
-predictions4$data$residuals <- residuals(m_temp4)
-```
-
-```{r}
-ggplot(predictions4$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions4$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth()
-```
-
-```{r}
-ggplot(predictions4$data, aes(X, Y, colour = residuals)) +
-  geom_point() +
-  facet_wrap(~year) +
-  coord_fixed() +
-  scale_color_gradient2()
-```
-
-# That might've helped a little but not a lot. Let's try with depth and not log depth.
-
-```{r}
-m_temp5 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + poly(depth_m, 3),
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions5 <- predict(m_temp5)
-predictions5$data$residuals <- residuals(m_temp5)
-```
-
-```{r}
-ggplot(predictions5$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions5$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth()
-```
-
-# I don't think that helps with the heteroscedasticity but it made the curvature of the residuals a little bit worse.
-
-```{r}
-ggplot(predictions5$data, aes(est, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions5$data, aes(est, residuals)) +
-  geom_point() +
-  geom_smooth()
-```
-
-```{r}
-aic <- function(m) {
-  k <- length(m$model$par)
-  nll <- m$model$objective
-  2 * k - 2 * (-nll)
-}
-aic(m_temp1)
-aic(m_temp2)
-aic(m_temp3)
-aic(m_temp4)
-aic(m_temp5)
-```
-
-# Conclusion: log depth, 3rd order polynomial with depth
-
-```{r}
-ggplot(predictions3$data, aes(est, temperature_c)) +
-  geom_point() +
-  facet_wrap(~year) +
-  coord_fixed() +
-  geom_abline()
-```
-
-# R2
-
-```{r}
-cor(predictions3$data$est, predictions3$data$temperature_c)^2
-```
-
-# Extrapolate:
-
-```{r}
-predictions3_nd <- predict(m_temp3, newdata = sdmTMB::qcs_grid)
-saveRDS(predictions3_nd$data, file = "qcs-temp.rds")
-```
-
-```{r}
-ggplot(predictions3$data, aes(X, Y, colour = residuals)) +
-  geom_point() +
-  facet_wrap(~year) +
-  coord_fixed() +
-  scale_color_gradient2()
+predictions_nd <- predict(m_temp2_rw_poly, newdata = qcs_grid)
+#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
 ```
@@
 ```{r}
-plot_map(predictions3_nd$data, "est") +
+plot_map(predictions_nd$data, "est") +
   scale_fill_viridis_c(trans = "sqrt", option = "C") +
@@
 ```{r}
-plot_map(predictions3_nd$data, "est_fe") +
+plot_map(predictions_nd$data, "est_fe") +
   ggtitle("Prediction (fixed effects only)") +
@@
 
-```{r}
-plot_map(predictions3_nd$data, "est_re_s") +
+```{r, eval=FALSE}
+plot_map(predictions_nd$data, "est_re_s") +
   ggtitle("Spatial random effects only") +
@@
 ```{r}
-plot_map(predictions3_nd$data, "est_re_st") +
+plot_map(predictions_nd$data, "est_re_st") +
   ggtitle("Spatiotemporal random effects only") +
@@
 ```{r}
-m_temp3$model$par
-# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
-exp(m_temp3$model$par[["ln_phi"]])
-
-p_withheld <- predict(m_temp3, newdata = d_withheld)
-
-#FIXME: p_withheld$data$temperature_c is all zeros?!?!?! 
-# temp work around (first make sure both df in same order)
-arrange(p_withheld$data, fishing_event_id) 
-arrange(d_withheld, fishing_event_id) 
-p_withheld$data$temperature_c <- d_withheld$temperature_c
-```
-
-```{r}
-ggplot(p_withheld$data, aes(est, temperature_c)) +
-  geom_point() +
-  facet_wrap(~year) +
-  coord_fixed() +
-  xlim(3,12) +
-  ylim(3,12) +
-  geom_abline()
-```
-
-```{r}
-temp_cog <- get_cog(predictions3_nd)
+temp_cog <- get_cog(predictions_nd)
 

### 2019-04-12 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
add loglik_cv for model comparision

@@
 d_trawl$depth3 <- (d_trawl$depth_scaled)^3
+d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
 ```
@@
 plot_spde(spde)
+```
+
+10-fold cross-validation of candidate models
+```{r}
+temp1 <- loglik_cv(d_trawl, n_knots = 100, 
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp1$nll)
+
+temp2 <- loglik_cv(d_trawl, n_knots = 100, 
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp2$nll)
+
+temp3 <- loglik_cv(d_trawl, n_knots = 100, 
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp3$nll)
+
+# allow slope with depth to vary through between years with a random walk
+temp1_rw <- loglik_cv(d_trawl, n_knots = 100,
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp1_rw$nll)
+
+temp2_rw <- loglik_cv(d_trawl, n_knots = 100, 
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp2_rw$nll)
+
+temp3_rw <- loglik_cv(d_trawl, n_knots = 100, 
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp3_rw$nll)
 ```
@@
 
-#FIXME: this is crashing R
-
-```{r}
+# there is probably a problem with doing this... but need to figure out how to add poly args to predict function
+```{r eval=FALSE}
 m_temp2_rw_poly <- sdmTMB(d_fit,
@@
 ```{r}
-predictions_nd <- predict(m_temp2_rw_poly, newdata = qcs_grid)
+qcs_grid$depth2 <- (qcs_grid$depth_scaled)^2
+qcs_grid$depth3 <- (qcs_grid$depth_scaled)^3
+```
+
+```{r}
+predictions_nd <- predict(m_temp2_rw_nosp, newdata = qcs_grid)
 #saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")

### 2019-04-12 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
update calls to cross validation code

--- 2019-04-11 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 plot_spde(spde)
+```
+
+10-fold cross-validation of candidate models
+```{r}
+temp1 <- loglik_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp1$sum_loglik)
+
+temp2 <- loglik_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp2$sum_loglik)
+
+temp3 <- loglik_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp3$sum_loglik)
+
+# allow slope with depth to vary through between years with a random walk
+temp1_rw <- loglik_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp1_rw$sum_loglik)
+
+temp2_rw <- loglik_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp2_rw$sum_loglik)
+
+temp3_rw <- loglik_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp3_rw$sum_loglik)
 ```

### 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
batch cross-validation of temp models

--- 2019-04-11 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 library(sdmTMB)
+library(future)
 
 d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
-
 #set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
 set.seed(392011149)
-d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
-d_trawl$depth2 <- (d_trawl$depth_scaled)^2
-d_trawl$depth3 <- (d_trawl$depth_scaled)^3
+#d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
+d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
+```
+
+### Functions for model exploration:
+```{r}
+generate_cv_args <- function(
+                formula = formula,
+                time_varying = time_varying,
+                ar1_fields = c(FALSE,TRUE), 
+                include_spatial = c(FALSE,TRUE),
+                n_knots = c(30,80) ){
+  arguments <- expand.grid(
+    formula = formula,
+    time_varying = time_varying,
+    include_spatial = include_spatial,
+    ar1_fields = ar1_fields,
+    n_knots = n_knots
+  )
+  args <- as_tibble(arguments)
+  args
+}
+
+plot_cv_meshes <- function(cv_out) {
+  k <- max(cv_out$data$cv_fold)
+  op <- graphics::par(
+      mfrow = c((sqrt(k)), ceiling(sqrt(k))),
+      mar = c(1, 1, 1, 1)
+    )
+  for (i in seq_len(k)){
+    plot_spde(cv_out$models[[i]]$spde)
+  }
+graphics::par(op)
+}
+
+aic <- function(m) {
+  k <- length(m$model$par)
+  nll <- m$model$objective
+  2 * k - 2 * (-nll)
+}
+```
+
+# 10-fold cross-validation
+
+## All fixed slope models
+```{r}
+cv_args <- generate_cv_args (formula = list(
+    temperature_c ~ 0 + as.factor(year) + depth_scaled,
+    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
+    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3),
+    time_varying = list(NULL),
+    ar1_fields = c(FALSE, TRUE), 
+    include_spatial = c(FALSE,TRUE),
+    n_knots = c(30,80,150)
+  )
+
+plan(multiprocess)
+sdmTMB_cv_batch1 <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
+cv_args1 <- cv_args
+cv_loglik <- c()
+for (i in seq_len(nrow(cv_args1))) {
+  cv_loglik[[i]] <- round(sdmTMB_cv_batch1[[i]]$sum_loglik, 1)
+}
+cv_args1$formula <- as.character(cv_args1$formula)
+cv_args1$cv_loglik <- cv_loglik
+View(cv_args1)
+cv_args_notv <- as.data.frame(cv_args1) %>% select(., -time_varying)
+glimpse(cv_args_notv)
+write.csv(cv_args_notv, file="cv-all-non-time-varying.csv")
+saveRDS(sdmTMB_cv_batch1, file = "cv-all-non-time-varying.rds")
+# readRDS(file = "sdmTMB_cv_nontv.rds")
+```
+
+## All time-varying slope models
+```{r}
+cv_args <- generate_cv_args (formula = list(
+    temperature_c ~ 0 + as.factor(year)),
+    time_varying = list(
+    ~ 0 + depth_scaled,
+    ~ 0 + depth_scaled + depth_scaled2,
+    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
+    ar1_fields = c(FALSE, TRUE), 
+    include_spatial = c(FALSE,TRUE),
+    n_knots = c(30,80,150)
+    )
+
+plan(multiprocess)
+sdmTMB_cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
+
+cv_loglik <- c()
+for (i in seq_len(nrow(cv_args))) {
+  cv_loglik[[i]] <- c(sdmTMB_cv_batch[[i]]$sum_loglik)
+}
+cv_args2 <- cv_args
+cv_args2$formula <- as.character(cv_args$formula)
+cv_args2$time_varying <- as.character(cv_args$time_varying)
+cv_args2$cv_loglik <- cv_loglik
+glimpse(cv_args2)
+cv_args_tv <- as.data.frame(cv_args2) 
+write.csv(cv_args_tv, file="cv-all-time-varying.csv")
+saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")
+# readRDS(file = "sdmTMB_cv_nontv.rds")
+```
+
+# A functions for plotting our predictions:
+
+```{r}
+plot_map <- function(dat, column = "est") {
+  ggplot(dat, aes_string("X", "Y", colour = column)) +
+    geom_point() +
+    facet_wrap(~year) +
+    coord_fixed()
+}
+
+plot_map_raster <- function(dat, column = "est") {
+  ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    facet_wrap(~year) +
+    coord_fixed()
+}
+```
+
+## Plotting predictions
+
+```{r}
+predictions_nd <- predict(m_temp2_rw_nosp, newdata = qcs_grid)
+#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
+```
+
+```{r}
+plot_map(predictions_nd$data, "est") +
+  scale_fill_viridis_c(trans = "sqrt", option = "C") +
+  ggtitle("Prediction (fixed effects + all random effects)")
+```
+
+```{r}
+plot_map(predictions_nd$data, "est_fe") +
+  ggtitle("Prediction (fixed effects only)") +
+  scale_fill_viridis_c(trans = "sqrt")
+```
+
+```{r, eval=FALSE}
+plot_map(predictions_nd$data, "est_re_s") +
+  ggtitle("Spatial random effects only") +
+  scale_fill_gradient2()
+```
+
+```{r}
+plot_map(predictions_nd$data, "est_re_st") +
+  ggtitle("Spatiotemporal random effects only") +
+  scale_fill_gradient2()
+```
+
+########################
+########################
+
+Various candidate models
+
+########################
+########################
+
+```{r}
+spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 100)
+p <- plot_spde(spde)
+p
 ```
@@
 length(unique(d_fit$year))
-
-# FIXME: need to work out a bug in production on this new data set
-# It's crashing TMB.
-
-#d_fit <- d_trawl
-#d_withheld <- d_trawl
 ```
 
+Misc models
 ```{r}
-spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 100)
-plot_spde(spde)
+temp2 <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+plot_cv_meshes(temp2)
+sum(temp2$sum_loglik)
+
+temp3 <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp3$sum_loglik)
+
+# allow slope with depth to vary through between years with a random walk
+temp1_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp1_rw$sum_loglik)
+
+temp2_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp2_rw$sum_loglik)
+
+temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = FALSE, include_spatial = FALSE)
+sum(temp3_rw$sum_loglik)
 ```
-
-Functions for model exploration:
-```{r}
-plot_map <- function(dat, column = "est") {
-  ggplot(dat, aes_string("X", "Y", colour = column)) +
-    geom_point() +
-    facet_wrap(~year) +
-    coord_fixed()
-}
-
-aic <- function(m) {
-  k <- length(m$model$par)
-  nll <- m$model$objective
-  2 * k - 2 * (-nll)
-}
-```
-
-Various candidate models:
 
@@
 ```{r, eval=FALSE}
+#d_trawl$depth2 <- (d_trawl$depth_scaled)^2
+#d_trawl$depth3 <- (d_trawl$depth_scaled)^3
+
 m_temp2 <- sdmTMB(d_fit, 
@@
 
-# Extrapolate:
-
-#FIXME: this is crashing R
-
-```{r}
+# there is probably a problem with doing this... but need to figure out how to add poly args to predict function
+```{r eval=FALSE}
 m_temp2_rw_poly <- sdmTMB(d_fit,
@@
   silent = TRUE)
-```
-
-```{r}
-predictions_nd <- predict(m_temp2_rw_poly, newdata = qcs_grid)
-#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
-```
-
-# A short function for plotting our predictions:
-
-```{r}
-plot_map <- function(dat, column = "est") {
-  ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    facet_wrap(~year) +
-    coord_fixed()
-}
-```
-
-```{r}
-plot_map(predictions_nd$data, "est") +
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
-  ggtitle("Prediction (fixed effects + all random effects)")
-```
-
-```{r}
-plot_map(predictions_nd$data, "est_fe") +
-  ggtitle("Prediction (fixed effects only)") +
-  scale_fill_viridis_c(trans = "sqrt")
-```
-
-```{r, eval=FALSE}
-plot_map(predictions_nd$data, "est_re_s") +
-  ggtitle("Spatial random effects only") +
-  scale_fill_gradient2()
-```
-
-```{r}
-plot_map(predictions_nd$data, "est_re_st") +
-  ggtitle("Spatiotemporal random effects only") +
-  scale_fill_gradient2()
 ```

### 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
batch cross-validation of temp models

@@
 plan(multiprocess)
-sdmTMB_cv_batch1 <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
+sdmTMB_cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
 cv_args1 <- cv_args
@@
 write.csv(cv_args_notv, file="cv-all-non-time-varying.csv")
-saveRDS(sdmTMB_cv_batch1, file = "cv-all-non-time-varying.rds")
-# readRDS(file = "sdmTMB_cv_nontv.rds")
 ```
@@
 write.csv(cv_args_tv, file="cv-all-time-varying.csv")
-saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")
-# readRDS(file = "sdmTMB_cv_nontv.rds")
 ```

### 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd

--- 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 write.csv(cv_args_tv, file="cv-all-time-varying.csv")
+
 saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")

### 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd

--- 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 # 10-fold cross-validation 
-
+#FIXME: need to extract information on whether or not models converge
 ## All fixed slope models
@@
 plan(multiprocess)
-sdmTMB_cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
+sdmTMB_cv_batch1 <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
 cv_args1 <- cv_args
@@
 cv_args1$cv_loglik <- cv_loglik
-View(cv_args1)
 cv_args_notv <- as.data.frame(cv_args1) %>% select(., -time_varying)
+# View(cv_args_notv)
+# write.csv(cv_args_notv, file = "cv-all-non-time-varying.csv")
+# saveRDS(sdmTMB_cv_batch1, file = "cv-all-non-time-varying.rds")
 glimpse(cv_args_notv)
-write.csv(cv_args_notv, file="cv-all-non-time-varying.csv")
 ```
@@
 sdmTMB_cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
-
-cv_loglik <- c()
+cv_args2 <- cv_args
+cv_args2$cv_loglik <- c()
 for (i in seq_len(nrow(cv_args))) {
-  cv_loglik[[i]] <- c(sdmTMB_cv_batch[[i]]$sum_loglik)
+  cv_args2$cv_loglik[[i]] <- c(sdmTMB_cv_batch[[i]]$sum_loglik)
 }
-cv_args2 <- cv_args
 cv_args2$formula <- as.character(cv_args$formula)
 cv_args2$time_varying <- as.character(cv_args$time_varying)
-cv_args2$cv_loglik <- cv_loglik
-glimpse(cv_args2)
 cv_args_tv <- as.data.frame(cv_args2) 
-write.csv(cv_args_tv, file="cv-all-time-varying.csv")
+# View(cv_args_tv)
+# write.csv(cv_args_tv, file = "cv-all-time-varying.csv")
+# saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")
+glimpse(cv_args_tv)
+```
+
+## vary n_knots for top model
+```{r}
+cv_args_knots <- generate_cv_args (formula = list(
+    temperature_c ~ 0 + as.factor(year)),
+    time_varying = list(
+    ~ 0 + depth_scaled + depth_scaled2,
+    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
+    ar1_fields = c(FALSE,TRUE), 
+    include_spatial = c(FALSE,TRUE),
+    n_knots = c(60,90,110,130,160)
+    )
+
+plan(multiprocess)
+sdmTMB_cv_knots2 <- purrr::pmap(cv_args_knots, sdmTMB_cv, data = d_trawl)
+cv_args3 <- cv_args_knots
+cv_args3$cv_loglik <- c()
+for (i in seq_len(nrow(cv_args3))) {
+  cv_args3$cv_loglik[[i]] <- c(sdmTMB_cv_knots2[[i]]$sum_loglik)
+}
+cv_args3$formula <- as.character(cv_args3$formula)
+cv_args3$time_varying <- as.character(cv_args3$time_varying)
+cv_args_knots2 <- as.data.frame(cv_args3) 
+# View(cv_args_knots2)
+# write.csv(cv_args_knots2, file = "cv-varying_knots2.csv")
+# saveRDS(sdmTMB_cv_knots2, file = "cv-varying_knots2.rds")
+glimpse(cv_args_knots2)
+```
+
+Load saved batches:
+```{r}
+non_time_varying_batch <- readRDS(file = "cv-all-non-time-varying.rds")
+time_varying_batch <- readRDS(file = "cv-all-time-varying.rds")
+knot_varying_batch <- readRDS(file = "cv-varying_knots.rds")
+knot_varying_batch2 <- readRDS(file = "cv-varying_knots2.rds")
+```
+
+Current top model
+```{r}
+plan(multiprocess)
+temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 130, x = "X", y ="Y", k_folds = 10, 
+                   formula = temperature_c ~ 0 + as.factor(year),
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time = "year", family = gaussian(link = "identity"), 
+                   ar1_fields = TRUE, include_spatial = TRUE)
+temp3_rw$models
+sum(temp3_rw$sum_loglik)
+```
+
+```{r}
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 130)
+plot_spde(spde)
+```
+
+```{r}
+m_temp3_rw <- sdmTMB(d_trawl,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying =  ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+  ar1_fields = TRUE, include_spatial = TRUE, 
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE)
+
+predictions3_rw <- predict(m_temp3_rw)
+predictions3_rw$data$residuals <- residuals(m_temp3_rw)
+cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 # 
+aic(m_temp3_rw)
+```
+
+Residual plots:
+```{r}
+ggplot(predictions3_rw$data, aes(est, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth()
 ```
@@
 ## Plotting predictions
-
 ```{r}
-predictions_nd <- predict(m_temp2_rw_nosp, newdata = qcs_grid)
+qcs_grid$depth_scaled3 <- (qcs_grid$depth_scaled)^3
+predictions_nd <- predict(m_temp3_rw, newdata = qcs_grid)
 #saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
@@
 ```{r}
-plot_map(predictions_nd$data, "est") +
+plot_map_raster(predictions_nd$data, "est") +
   scale_fill_viridis_c(trans = "sqrt", option = "C") +
@@
 ```{r}
-plot_map(predictions_nd$data, "est_fe") +
+plot_map_raster(predictions_nd$data, "est_fe") +
   ggtitle("Prediction (fixed effects only)") +
@@
 ```{r, eval=FALSE}
-plot_map(predictions_nd$data, "est_re_s") +
+plot_map_raster(predictions_nd$data, "est_re_s") +
   ggtitle("Spatial random effects only") +
@@
 ```{r}
-plot_map(predictions_nd$data, "est_re_st") +
+plot_map_raster(predictions_nd$data, "est_re_st") +
   ggtitle("Spatiotemporal random effects only") +
@@
 
-########################
-########################
-
-Various candidate models
-
-########################
-########################
-
 ```{r}
-spde <- make_spde(d_fit$X, d_fit$Y, n_knots = 100)
-p <- plot_spde(spde)
-p
-```
-
-```{r}
-withheld_fe <- sample(d_trawl$fishing_event_id, size = round(nrow(d_trawl)*.2))
-d_withheld <- filter(d_trawl, fishing_event_id %in% withheld_fe)
-d_fit <- filter(d_trawl, !fishing_event_id %in% withheld_fe)
-
-nrow(d_fit) # 3774
-nrow(d_withheld) # 943
-unique(d_withheld$year) 
-unique(d_fit$year)
-length(unique(d_fit$year))
-```
-
-Misc models
-```{r}
-temp2 <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
-                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
-                   time = "year", family = gaussian(link = "identity"), 
-                   ar1_fields = FALSE, include_spatial = FALSE)
-plot_cv_meshes(temp2)
-sum(temp2$sum_loglik)
-
-temp3 <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
-                   formula = temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
-                   time = "year", family = gaussian(link = "identity"), 
-                   ar1_fields = FALSE, include_spatial = FALSE)
-sum(temp3$sum_loglik)
-
-# allow slope with depth to vary through between years with a random walk
-temp1_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
-                   formula = temperature_c ~ 0 + as.factor(year),
-                   time_varying = ~ 0 + depth_scaled,
-                   time = "year", family = gaussian(link = "identity"), 
-                   ar1_fields = FALSE, include_spatial = FALSE)
-sum(temp1_rw$sum_loglik)
-
-temp2_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
-                   formula = temperature_c ~ 0 + as.factor(year),
-                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
-                   time = "year", family = gaussian(link = "identity"), 
-                   ar1_fields = FALSE, include_spatial = FALSE)
-sum(temp2_rw$sum_loglik)
-
-temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 30, x = "X", y ="Y", k_folds = 5, plot_spde = TRUE,
-                   formula = temperature_c ~ 0 + as.factor(year),
-                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
-                   time = "year", family = gaussian(link = "identity"), 
-                   ar1_fields = FALSE, include_spatial = FALSE)
-sum(temp3_rw$sum_loglik)
-```
-
-```{r}
-m_temp1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
-  time = "year", spde = spde, family = gaussian(link = "identity"), 
-  ar1_fields = FALSE, include_spatial = FALSE, 
-  silent = TRUE)
-
-predictions <- predict(m_temp1)
-predictions$data$residuals <- residuals(m_temp1)
-cor(predictions$data$est, predictions$data$temperature_c)^2
-aic(m_temp1)
-
-m_temp1_nosp  <- m_temp1
-```
-
-AR1 fields no improvement in R^2
-```{r, eval=FALSE}
-m_temp1_ar1 <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year) + depth_scaled,
-  time = "year", spde = spde, family = gaussian(link = "identity"), 
-  ar1_fields = TRUE, include_spatial = FALSE, 
-  anisotropy = FALSE, 
-  silent = TRUE)
-
-predictions_ar1 <- predict(m_temp1_ar1)
-predictions_ar1$data$residuals <- residuals(m_temp1_ar1)
-cor(predictions_ar1$data$est, predictions_ar1$data$temperature_c)^2
-aic(m_temp1_ar1)
-```
-
-Time-varying slope for depth, little imporvement for R^2
-```{r}
-m_temp1_rw <- sdmTMB(d_fit, temperature_c ~ 0 + as.factor(year),   
-  time_varying = ~ 0 + depth_scaled,
-  ar1_fields = FALSE, include_spatial = FALSE, #these are defaults
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions1_rw <- predict(m_temp1_rw)
-predictions1_rw$data$residuals <- residuals(m_temp1_rw)
-cor(predictions1_rw$data$est, predictions1_rw$data$temperature_c)^2
-aic(m_temp1_rw)
-```
-
-NO improvement
-```{r, eval=FALSE}
-m_temp1_rw_ar1 <- sdmTMB(d_fit, 
-  temperature_c ~ 0 + as.factor(year),   
-  time_varying = ~ 0 + depth_scaled,
-  ar1_fields = TRUE, include_spatial = TRUE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions1_rw_ar1 <- predict(m_temp1_rw_ar1)
-predictions1_rw_ar1$data$residuals <- residuals(m_temp1_rw_ar1)
-cor(predictions1_rw_ar1$data$est, predictions1_rw_ar1$data$temperature_c)^2
-aic(m_temp1_rw_ar1)
-```
-
-NO improvement
-```{r, eval=FALSE}
-m_temp1_st <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year) + depth_scaled,
-  #time_varying =  ~ 0 + depth_scaled,
-  include_spatial = TRUE, ar1_fields = FALSE, spatial_trend = TRUE, 
-  # seems to be only config that converges
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions1_st <- predict(m_temp1_st)
-predictions1_st$data$residuals <- residuals(m_temp1_st)
-cor(predictions1_st$data$est, predictions1_st$data$temperature_c)^2 # 
-aic(m_temp1_st)
-```
-
-Both time and space varying... doesn't improve R^2
-```{r, eval=FALSE}
-m_temp1_rw_st <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year), 
-  time_varying =  ~ 0 + depth_scaled,
-  include_spatial = TRUE, ar1_fields = FALSE, spatial_trend = TRUE, 
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions1_rw_st <- predict(m_temp1_rw_st)
-predictions1_rw_st$data$residuals <- residuals(m_temp1_rw_st)
-cor(predictions1_rw_st$data$est, predictions1_rw_st$data$temperature_c)^2 # 
-aic(m_temp1_rw_st)
-```
-
-What about polynomial response to depth?
-
-```{r, eval=FALSE}
-#d_trawl$depth2 <- (d_trawl$depth_scaled)^2
-#d_trawl$depth3 <- (d_trawl$depth_scaled)^3
-
-m_temp2 <- sdmTMB(d_fit, 
-  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2,
-  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions2 <- predict(m_temp2)
-predictions2$data$residuals <- residuals(m_temp2)
-cor(predictions2$data$est, predictions2$data$temperature_c)^2
-aic(m_temp2)
-```
-
-```{r}
-m_temp2_nosp <- sdmTMB(d_fit, 
-  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2,
-  ar1_fields = FALSE, include_spatial = FALSE, 
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions2_nosp <- predict(m_temp2_nosp)
-predictions2_nosp$data$residuals <- residuals(m_temp2_nosp)
-cor(predictions2_nosp$data$est, predictions2_nosp$data$temperature_c)^2
-aic(m_temp2_nosp)
-```
-
-```{r}
-m_temp2_rw_nosp <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year),
-  time_varying =  ~ 0 + depth_scaled + depth2,
-  include_spatial = FALSE, ar1_fields = FALSE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions2_rw_nosp <- predict(m_temp2_rw_nosp)
-predictions2_rw_nosp$data$residuals <- residuals(m_temp2_rw_nosp)
-cor(predictions2_rw_nosp$data$est, predictions2_rw_nosp$data$temperature_c)^2
-aic(m_temp2_rw_nosp)
-```
-
-NO improvement 
-```{r, eval=FALSE}
-m_temp2_st <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year), #+ poly(depth_scaled, 2),
-  time_varying =  ~ 0 + poly(depth_scaled, 2),
-  include_spatial = FALSE, ar1_fields = FALSE, spatial_trend = TRUE, 
-  # seems to be only config that converges
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions2_st <- predict(m_temp2_st)
-predictions2_st$data$residuals <- residuals(m_temp2_st)
-cor(predictions2_st$data$est, predictions2_st$data$temperature_c)^2 # 
-aic(m_temp2_st)
-```
-
-What about 3 order polynomial?
-```{r, eval=FALSE}
-m_temp3 <- sdmTMB(d_fit, 
-  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2 + depth3,
-  ar1_fields = FALSE, include_spatial = TRUE, #these are defaults
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions3 <- predict(m_temp3)
-predictions3$data$residuals <- residuals(m_temp3)
-
-cor(predictions3$data$est, predictions3$data$temperature_c)^2 # 
-```
-
-Slightly better without a fixed spatial field!
-```{r}
-m_temp3_nosp <- sdmTMB(d_fit, 
-  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2 + depth3,
-  ar1_fields = FALSE, include_spatial = FALSE, 
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions3_nosp <- predict(m_temp3_nosp)
-predictions3_nosp$data$residuals <- residuals(m_temp3_nosp)
-
-cor(predictions3_nosp$data$est, predictions3_nosp$data$temperature_c)^2 # 
-```
-
-Better without AR1
-```{r, eval=FALSE}
-m_temp3_ar1 <- sdmTMB(d_fit, 
-  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth2 + depth3,
-  ar1_fields = TRUE, include_spatial = FALSE, 
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions3_ar1 <- predict(m_temp3_ar1)
-predictions3_ar1$data$residuals <- residuals(m_temp3_ar1)
-
-cor(predictions3_ar1$data$est, predictions3_ar1$data$temperature_c)^2 # 
-```
-
-```{r, eval=FALSE}
-m_temp3_rw <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year),
-  time_varying =  ~ 0 + depth_scaled + depth2 + depth3,
-  ar1_fields = FALSE, include_spatial = TRUE, 
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions3_rw <- predict(m_temp3_rw)
-predictions3_rw$data$residuals <- residuals(m_temp3_rw)
-cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 # 
-aic(m_temp3_rw)
-```
-
-Also better without fixed spatial field but not quite as good as 2nd order
-```{r}
-m_temp3_rw_nosp <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year),
-  time_varying =  ~ 0 + depth_scaled + depth2 + depth3,
-  include_spatial = FALSE, ar1_fields = FALSE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions3_rw_nosp <- predict(m_temp3_rw_nosp)
-predictions3_rw_nosp$data$residuals <- residuals(m_temp3_rw_nosp)
-cor(predictions3_rw_nosp$data$est, predictions3_rw_nosp$data$temperature_c)^2 # 
-```
-
-Not converging 
-```{r, eval=FALSE}
-m_temp3_st <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year), #+ poly(depth_scaled, 2),
-  time_varying =  ~ 0 + depth_scaled + depth2 + depth3,
-  include_spatial = FALSE, ar1_fields = FALSE, spatial_trend = TRUE, 
-  # seems to be only config that converges
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
-
-predictions3_st <- predict(m_temp3_st)
-predictions3_st$data$residuals <- residuals(m_temp3_st)
-cor(predictions3_st$data$est, predictions3_st$data$temperature_c)^2 # 
-```
-
-# Conclusion:
-
-```{r}
-# m_temp2_rw$model$par
-# 2 * plogis(m_temp1$model$par[["ar1_phi"]]) - 1
-# exp(m_temp2_rw$model$par[["ln_phi"]])
-
-p_withheld1 <- predict(m_temp1_nosp, newdata = d_withheld)
-cor(p_withheld1$data$est, p_withheld1$data$temperature_c)^2 # 0.87 in first 
-```
-```{r}
-p_withheld1_rw <- predict(m_temp1_rw, newdata = d_withheld)
-cor(p_withheld1_rw$data$est, p_withheld1_rw$data$temperature_c)^2 # 0.87 in first 
-```
-
-# 4
-```{r}
-p_withheld2 <- predict(m_temp2_nosp, newdata = d_withheld)
-cor(p_withheld2$data$est, p_withheld2$data$temperature_c)^2 # 
-```
-
-# 2
-```{r}
-p_withheld2_rw <- predict(m_temp2_rw_nosp, newdata = d_withheld)
-cor(p_withheld2_rw$data$est, p_withheld2_rw$data$temperature_c)^2 # 
-```
-
-# 3
-```{r}
-p_withheld3 <- predict(m_temp3_nosp, newdata = d_withheld)
-cor(p_withheld3$data$est, p_withheld3$data$temperature_c)^2 # 0.87
-```
-
-# 1
-```{r}
-p_withheld3_rw <- predict(m_temp3_rw_nosp, newdata = d_withheld)
-cor(p_withheld3_rw$data$est, p_withheld3_rw$data$temperature_c)^2 # 0.87
-```
-
-```{r}
-ggplot(p_withheld3_rw$data, aes(temperature_c,est)) +
+ggplot(predictions3_rw$data, aes(temperature_c, est)) +
   geom_point() +
@@
   geom_abline()
-```
-
-Residual plots:
-```{r}
-ggplot(predictions3_rw_nosp$data, aes(est, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions2_rw_nosp$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(predictions2_rw_nosp$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth()
-```
-
-```{r}
-ggplot(predictions3_rw_nosp$data, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth()
-```
-
-# there is probably a problem with doing this... but need to figure out how to add poly args to predict function
-```{r eval=FALSE}
-m_temp2_rw_poly <- sdmTMB(d_fit,
-  temperature_c ~ 0 + as.factor(year),
-  time_varying =  ~ 0 + poly(depth_scaled,2),
-  include_spatial = FALSE, ar1_fields = FALSE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
 ```

### 2019-04-15 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
add seed and convergence reporting to batch processes

@@
                 include_spatial = c(FALSE,TRUE),
-                n_knots = c(30,80) ){
+                n_knots = c(30,80),
+                seed = c(999)
+  ){
   arguments <- expand.grid(
@@
     ar1_fields = ar1_fields,
-    n_knots = n_knots
+    n_knots = n_knots,
+    seed = seed
   )
@@
     time_varying = list(NULL),
-    ar1_fields = c(FALSE, TRUE), 
+    ar1_fields = c(FALSE), 
     include_spatial = c(FALSE,TRUE),
-    n_knots = c(30,80,150)
+    n_knots = c(150),
+    seed = c(111,222,333)
   )
@@
 cv_args1 <- cv_args
-cv_loglik <- c()
+cv_args1$cv_loglik <- c()
+cv_args1$all_converged <- c()
 for (i in seq_len(nrow(cv_args1))) {
-  cv_loglik[[i]] <- round(sdmTMB_cv_batch1[[i]]$sum_loglik, 1)
+  cv_args1$cv_loglik[[i]] <- round(sdmTMB_cv_batch[[i]]$sum_loglik, 1)
+  cv_args1$all_converged[[i]] <- all(sdmTMB_cv_batch[[i]]$converg)
 }
 cv_args1$formula <- as.character(cv_args1$formula)
-cv_args1$cv_loglik <- cv_loglik
-cv_args_notv <- as.data.frame(cv_args1) %>% select(., -time_varying)
+cv_args1$time_varying <- "NA"
+
+cv_args_notv <- as.data.frame(cv_args1) 
 # View(cv_args_notv)
-# write.csv(cv_args_notv, file = "cv-all-non-time-varying.csv")
-# saveRDS(sdmTMB_cv_batch1, file = "cv-all-non-time-varying.rds")
+# write.csv(cv_args_notv, file = "cv-non-time-varying.csv")
+# saveRDS(sdmTMB_cv_batch1, file = "cv-non-time-varying.rds")
 glimpse(cv_args_notv)
@@
     include_spatial = c(FALSE,TRUE),
-    n_knots = c(30,80,150)
+    n_knots = c(150),
+    seed = c(111,222,333)
     )
@@
 cv_args2$cv_loglik <- c()
+cv_args2$all_converged <- c()
 for (i in seq_len(nrow(cv_args))) {
   cv_args2$cv_loglik[[i]] <- c(sdmTMB_cv_batch[[i]]$sum_loglik)
+  cv_args2$all_converged[[i]] <- all(sdmTMB_cv_batch[[i]]$converg)
 }
@@
 # View(cv_args_tv)
-# write.csv(cv_args_tv, file = "cv-all-time-varying.csv")
-# saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")
+# write.csv(cv_args_tv, file = "cv-time-varying.csv")
+# saveRDS(sdmTMB_cv_batch, file = "cv-time-varying.rds")
 glimpse(cv_args_tv)
@@
 ```{r}
-cv_args_knots <- generate_cv_args (formula = list(
+cv_args_knots <- generate_cv_args(formula = list(
     temperature_c ~ 0 + as.factor(year)),
     time_varying = list(
-    ~ 0 + depth_scaled + depth_scaled2,
-    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
-    ar1_fields = c(FALSE,TRUE), 
-    include_spatial = c(FALSE,TRUE),
-    n_knots = c(60,90,110,130,160)
+    ~ 0 + depth_scaled + depth_scaled2),
+    #~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
+    ar1_fields = c(FALSE), 
+    include_spatial = c(TRUE),
+    n_knots = c(80,120,180),
+    seed = c(111,222,333)
     )
@@
 plan(multiprocess)
-sdmTMB_cv_knots2 <- purrr::pmap(cv_args_knots, sdmTMB_cv, data = d_trawl)
+sdmTMB_cv_knots <- purrr::pmap(cv_args_knots, sdmTMB_cv, data = d_trawl)
+cv_args3 <- cv_args_knots
+cv_args3$cv_loglik <- c()
+cv_args3$all_converged <- c()
+for (i in seq_len(nrow(cv_args3))) {
+  cv_args3$cv_loglik[[i]] <- c(sdmTMB_cv_knots[[i]]$sum_loglik)
+  cv_args3$all_converged[[i]] <- all(sdmTMB_cv_knots[[i]]$converg)
+}
+cv_args3$formula <- as.character(cv_args3$formula)
+cv_args3$time_varying <- as.character(cv_args3$time_varying)
+cv_knots <- as.data.frame(cv_args3) 
+# View(cv_knots)
+# write.csv(cv_knots, file = "cv-varying-knots.csv")
+# saveRDS(sdmTMB_cv_knots, file = "cv-varying-knots.rds")
+glimpse(cv_knots)
+```
+
+Change seeds and refine knots
+```{r}
+cv_args_knots <- generate_cv_args(formula = list(
+    temperature_c ~ 0 + as.factor(year)),
+    time_varying = list(
+    ~ 0 + depth_scaled + depth_scaled2),
+    #~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
+    ar1_fields = c(FALSE,TRUE), 
+    include_spatial = c(TRUE),
+    n_knots = c(135,145,155),
+    seed = c(444,555)
+    )
+
+plan(multiprocess)
+sdmTMB_cv_knots <- purrr::pmap(cv_args_knots, sdmTMB_cv, data = d_trawl)
 cv_args3 <- cv_args_knots
@@
 cv_args3$time_varying <- as.character(cv_args3$time_varying)
-cv_args_knots2 <- as.data.frame(cv_args3) 
-# View(cv_args_knots2)
-# write.csv(cv_args_knots2, file = "cv-varying_knots2.csv")
-# saveRDS(sdmTMB_cv_knots2, file = "cv-varying_knots2.rds")
-glimpse(cv_args_knots2)
+cv_args_knots <- as.data.frame(cv_args3) 
+# View(cv_args_knots)
+# write.csv(cv_args_knots, file = "cv-refine-knots.csv")
+# saveRDS(sdmTMB_cv_knots, file = "cv-refine-knots.rds")
+glimpse(cv_args_knots)
 ```
@@
 ```{r}
-non_time_varying_batch <- readRDS(file = "cv-all-non-time-varying.rds")
-time_varying_batch <- readRDS(file = "cv-all-time-varying.rds")
-knot_varying_batch <- readRDS(file = "cv-varying_knots.rds")
-knot_varying_batch2 <- readRDS(file = "cv-varying_knots2.rds")
+non_time_varying_batch <- readRDS(file = "cv-non-time-varying.rds")
+time_varying_batch <- readRDS(file = "cv-time-varying.rds")
+knot_varying_batch <- readRDS(file = "cv-varying-knots.rds")
+knot_refining_batch <- readRDS(file = "cv-refine-knots.rds")
+
+all_cv_loglik <- rbind(cv_args_notv, cv_args_tv, cv_args_knots, deparse.level = 1, make.row.names=TRUE)
+View(all_cv_loglik)
+
+by_seed <- all_cv_loglik %>% group_by(seed)
+for (i in unique(all_cv_loglik$seed)){
+  for (j in seq_len(nrow(diff_ll))) {
+  by_seed[[i]]$delta_ll <- c()
+  by_seed[[i]]$delta_ll[[j]] <- by_seed[[i]]$cv_loglik[[j]]-min(by_seed[[i]]$cv_loglik)
+  }
+}
+all_cv_loglik <- by_seed %>% ungroup()
 ```
@@
                    formula = temperature_c ~ 0 + as.factor(year),
-                   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+                   time_varying = ~ 0 + depth_scaled + depth_scaled2,
                    time = "year", family = gaussian(link = "identity"), 
-                   ar1_fields = TRUE, include_spatial = TRUE)
+                   ar1_fields = FALS, include_spatial = TRUE)
 temp3_rw$models

### 2019-04-15 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
testing batch processes for sdmTMB_cv

--- 2019-04-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
+---
+output: html_document
+editor_options: 
+  chunk_output_type: inline
+---
 # Fit spatial temporal models to bottom temperature
@@
 library(future)
-
+plan(multiprocess)
 d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
-#set.seed(39201114) # original seed, predictions max out at .88 cor with true temp
-set.seed(392011149)
 #d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
 d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
+unique(d_trawl$year)
 ```
@@
 ```{r}
+aic <- function(m) {
+  k <- length(m$model$par)
+  nll <- m$model$objective
+  2 * k - 2 * (-nll)
+}
+
 generate_cv_args <- function(
@@
                 time_varying = time_varying,
-                ar1_fields = c(FALSE,TRUE), 
-                include_spatial = c(FALSE,TRUE),
-                n_knots = c(30,80) ){
+                ar1_fields = c(FALSE), # simiple, fast defaults
+                include_spatial = c(FALSE), # simiple, fast defaults
+                n_knots = c(100), # simiple, fast defaults
+                seed = c(999)
+  ){
   arguments <- expand.grid(
@@
     ar1_fields = ar1_fields,
-    n_knots = n_knots
+    n_knots = n_knots,
+    seed = seed
   )
-  args <- as_tibble(arguments)
-  args
+  #browser()
+  as.data.frame(arguments)
 }
 
-plot_cv_meshes <- function(cv_out) {
-  k <- max(cv_out$data$cv_fold)
+sdmTMB_cv_batch <- function (cv_args, data, ...) {
+  cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl, ...)
+  cv_out <- cv_args
+  cv_out$cv_loglik <- c()
+  cv_out$all_converged <- c()
+  cv_out$max_grad <- c()
+  for (i in seq_len(nrow(cv_args))) {
+    cv_out$cv_loglik[[i]] <- round(cv_batch[[i]]$sum_loglik, 1)
+    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converg)
+    cv_out$max_grad[[i]] <- max(cv_batch[[i]]$max_grad)
+    if (is.null(cv_out$time_varying[[i]])) 
+      cv_out$time_varying[[i]] <- as.character(NA)
+  }
+  cv_out$formula <- as.character(cv_out$formula)
+  cv_out$time_varying <- as.character(cv_out$time_varying)
+  list(summary = as.data.frame(cv_out), cv_model_set = cv_batch)
+}
+
+plot_cv_meshes <- function(sdmTMB_cv_output) {
+  cv_output <- sdmTMB_cv_output
+  k <- max(cv_output[["data"]]$cv_fold)
+  if (k == 2 | k == 5 | k == 10) {
     op <- graphics::par(
@@
       )
+  } else { 
+    op <- graphics::par(
+      mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
+      mar = c(1, 1, 1, 1)
+    )
+  }
   for (i in seq_len(k)){
-    plot_spde(cv_out$models[[i]]$spde)
+    plot_spde(cv_output$models[[i]]$spde)
   }
@@
 
-aic <- function(m) {
-  k <- length(m$model$par)
-  nll <- m$model$objective
-  2 * k - 2 * (-nll)
-}
+```
+
+Short example: 
+```{r, eval=FALSE}
+cv_args <- generate_cv_args (formula = list(
+    temperature_c ~ 0 + as.factor(year) + depth_scaled),
+    time_varying = list(NULL),
+    ar1_fields = c(FALSE), 
+    include_spatial = c(FALSE),
+    n_knots = c(30), 
+    seed = c(111,222)
+  )
+
+cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
+glimpse(cv$summary)
+plot_cv_meshes(cv$cv_model_set[[1]])
+plot_cv_meshes(cv$cv_model_set[[2]])
 ```
@@
 # 10-fold cross-validation
-#FIXME: need to extract information on whether or not models converge
-## All fixed slope models
+
+## Just compare fixed slope models with simplest random effect structure
 ```{r}
@@
     time_varying = list(NULL),
-    ar1_fields = c(FALSE, TRUE), 
-    include_spatial = c(FALSE,TRUE),
-    n_knots = c(30,80,150)
+    ar1_fields = c(FALSE), 
+    include_spatial = c(FALSE),
+    n_knots = c(150), #start with close to projected optimum n_knots 
+    seed = c(1111,2221,3331)
   )
 
-plan(multiprocess)
-sdmTMB_cv_batch1 <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
-cv_args1 <- cv_args
-cv_loglik <- c()
-for (i in seq_len(nrow(cv_args1))) {
-  cv_loglik[[i]] <- round(sdmTMB_cv_batch1[[i]]$sum_loglik, 1)
-}
-cv_args1$formula <- as.character(cv_args1$formula)
-cv_args1$cv_loglik <- cv_loglik
-cv_args_notv <- as.data.frame(cv_args1) %>% select(., -time_varying)
-# View(cv_args_notv)
-# write.csv(cv_args_notv, file = "cv-all-non-time-varying.csv")
-# saveRDS(sdmTMB_cv_batch1, file = "cv-all-non-time-varying.rds")
-glimpse(cv_args_notv)
+non_time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
+View(non_time_varying$summary)
+write.csv(non_time_varying$summary, file = "cv-non-time-varying.csv")
+saveRDS(non_time_varying, file = "cv-non-time-varying.rds")
+glimpse(non_time_varying$summary)
 ```
 
-## All time-varying slope models
+## Compare time-varying slope models still with simplest random effect structure
 ```{r}
@@
     ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
-    ar1_fields = c(FALSE, TRUE), 
-    include_spatial = c(FALSE,TRUE),
-    n_knots = c(30,80,150)
+    ar1_fields = c(FALSE), 
+    include_spatial = c(FALSE),
+    n_knots = c(150), # keep same n_knots
+    seed = c(1111,2221,3331) # keep same seeds
     )
 
-plan(multiprocess)
-sdmTMB_cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl)
-cv_args2 <- cv_args
-cv_args2$cv_loglik <- c()
-for (i in seq_len(nrow(cv_args))) {
-  cv_args2$cv_loglik[[i]] <- c(sdmTMB_cv_batch[[i]]$sum_loglik)
-}
-cv_args2$formula <- as.character(cv_args$formula)
-cv_args2$time_varying <- as.character(cv_args$time_varying)
-cv_args_tv <- as.data.frame(cv_args2) 
-# View(cv_args_tv)
-# write.csv(cv_args_tv, file = "cv-all-time-varying.csv")
-# saveRDS(sdmTMB_cv_batch, file = "cv-all-time-varying.rds")
-glimpse(cv_args_tv)
+time_varying  <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
+
+View(time_varying$summary)
+write.csv(time_varying$summary, file = "cv-time-varying.csv")
+saveRDS(time_varying, file = "cv-time-varying.rds")
+glimpse(time_varying$summary)
 ```
 
-## vary n_knots for top model
+## Now vary n_knots and random effect configuation for top model(s) 
 ```{r}
-cv_args_knots <- generate_cv_args (formula = list(
+cv_args <- generate_cv_args(formula = list(
     temperature_c ~ 0 + as.factor(year)),
@@
     ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
-    ar1_fields = c(FALSE,TRUE), 
-    include_spatial = c(FALSE,TRUE),
-    n_knots = c(60,90,110,130,160)
+    ar1_fields = c(FALSE, TRUE), 
+    include_spatial = c(FALSE, TRUE),
+    n_knots = c(85,110,135),
+    seed = c(1111,2221,3331)
     )
 
-plan(multiprocess)
-sdmTMB_cv_knots2 <- purrr::pmap(cv_args_knots, sdmTMB_cv, data = d_trawl)
-cv_args3 <- cv_args_knots
-cv_args3$cv_loglik <- c()
-for (i in seq_len(nrow(cv_args3))) {
-  cv_args3$cv_loglik[[i]] <- c(sdmTMB_cv_knots2[[i]]$sum_loglik)
-}
-cv_args3$formula <- as.character(cv_args3$formula)
-cv_args3$time_varying <- as.character(cv_args3$time_varying)
-cv_args_knots2 <- as.data.frame(cv_args3) 
-# View(cv_args_knots2)
-# write.csv(cv_args_knots2, file = "cv-varying_knots2.csv")
-# saveRDS(sdmTMB_cv_knots2, file = "cv-varying_knots2.rds")
-glimpse(cv_args_knots2)
+vary_rf_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
+
+View(vary_rf_knots$summary)
+write.csv(vary_rf_knots$summary, file = "cv-varying-rf.csv")
+saveRDS(vary_rf_knots, file = "cv-varying-rf.rds")
+glimpse(vary_rf_knots$summary)
+```
+
+## Change seeds 
+```{r}
+cv_args <- generate_cv_args(formula = list(
+    temperature_c ~ 0 + as.factor(year)),
+    time_varying = list(
+    ~ 0 + depth_scaled + depth_scaled2,
+    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
+    ar1_fields = c(FALSE), 
+    include_spatial = c(TRUE),
+    n_knots = c(120),
+    seed = c(4441,5551,6661,7771,8881,9991)
+    )
+
+more_seeds <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
+
+View(more_seeds$summary)
+write.csv(more_seeds$summary, file = "cv-more-seeds.csv")
+saveRDS(more_seeds, file = "cv-more-seeds.rds")
+glimpse(more_seeds$summary)
 ```
@@
 ```{r}
-non_time_varying_batch <- readRDS(file = "cv-all-non-time-varying.rds")
-time_varying_batch <- readRDS(file = "cv-all-time-varying.rds")
-knot_varying_batch <- readRDS(file = "cv-varying_knots.rds")
-knot_varying_batch2 <- readRDS(file = "cv-varying_knots2.rds")
+# non_time_varying <- readRDS(file = "cv-non-time-varying.rds")
+# time_varying <- readRDS(file = "cv-time-varying.rds")
+# vary_rf <- readRDS(file = "cv-varying-rf.rds")
+# knot_refining <- readRDS(file = "cv-refine-knots.rds")
+
+# non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
+# time_varying <- read.csv(file = "cv-time-varying.csv")
+# vary_rf <- read.csv(file = "cv-varying-rf.csv")
+# knot_refining <- read.csv(file = "cv-refine-knots.csv")
+
+all_cv_loglik <- rbind(non_time_varying$summary, time_varying$summary, non_tv$summary, tvar$summary, vary_rf$summary, vary_rf_knots$summary, knot_refining$summary, more_seeds$summary, deparse.level = 1, make.row.names=TRUE)
+View(all_cv_loglik)
+
+by_seed <- all_cv_loglik %>% group_by(seed) %>% 
+  mutate(., max_ll_per_seed = max(cv_loglik)) %>% 
+  mutate(., delta_ll = cv_loglik - max_ll_per_seed)
+
+View(by_seed)
+all_cv_loglik <- by_seed %>% ungroup()
+
+write.csv(all_cv_loglik, file = "all_cv_loglik.csv")
 ```

### 2019-04-18 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
add 'day of event' SST to temperature models; not appearing very useful... what about trying a mean from the month prior to the survey?

@@
 library(future)
+library(dplyr)
+
 plan(multiprocess)
-d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
-#d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
-d_trawl$depth_scaled3 <- (d_trawl$depth_scaled)^3
+# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
+# nrow(d_trawl1)
+d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
+glimpse(d_trawl)
+nrow(d_trawl)
+head(d_trawl)
+# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
+d_trawl <- mutate(d_trawl,
+  depth_scaled = arm::rescale(log(depth_m)),
+  depth_scaled2 = depth_scaled^2,
+  depth_scaled3 = depth_scaled^3
+)
 unique(d_trawl$year)
+d_trawl[is.na(d_trawl$year), ]$year <- 2003
 ```
@@
 ```{r}
-aic <- function(m) {
-  k <- length(m$model$par)
-  nll <- m$model$objective
-  2 * k - 2 * (-nll)
-}
+# aic <- function(m) {
+#   k <- length(m$model$par)
+#   nll <- m$model$objective
+#   2 * k - 2 * (-nll)
+# }
 
@@
                              n_knots = c(100), # simiple, fast defaults
-                seed = c(999)
-  ){
+                             seed = c(999)) {
   arguments <- expand.grid(
@@
   )
-  #browser()
   as.data.frame(arguments)
@@
 
-sdmTMB_cv_batch <- function (cv_args, data, ...) {
+sdmTMB_cv_batch <- function(cv_args, data, ...) {
   cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl, ...)
@@
     cv_out$cv_loglik[[i]] <- round(cv_batch[[i]]$sum_loglik, 1)
-    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converg)
+    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converged)
     cv_out$max_grad[[i]] <- max(cv_batch[[i]]$max_grad)
-    if (is.null(cv_out$time_varying[[i]])) 
+    if (is.null(cv_out$time_varying[[i]])) {
       cv_out$time_varying[[i]] <- as.character(NA)
+    }
   }
@@
   cv_out$time_varying <- as.character(cv_out$time_varying)
+
+  cv_out <- cv_out %>%
+    group_by(seed) %>%
+    mutate(., max_ll_per_seed = max(cv_loglik)) %>%
+    mutate(., delta_ll = cv_loglik - max_ll_per_seed) %>%
+    ungroup()
+  row.names(cv_out) <- NULL
   list(summary = as.data.frame(cv_out), cv_model_set = cv_batch)
@@
 
-plot_cv_meshes <- function(sdmTMB_cv_output) {
-  cv_output <- sdmTMB_cv_output
-  k <- max(cv_output[["data"]]$cv_fold)
-  if (k == 2 | k == 5 | k == 10) {
-    op <- graphics::par(
-      mfrow = c((sqrt(k)), ceiling(sqrt(k))),
-      mar = c(1, 1, 1, 1)
-      )
-  } else { 
-    op <- graphics::par(
-      mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
-      mar = c(1, 1, 1, 1)
-    )
-  }
-  for (i in seq_len(k)){
-    plot_spde(cv_output$models[[i]]$spde)
-  }
-  graphics::par(op)
-}
-
+# # this code keeps breaking... working, fails, worked again, then failed again...
+# plot_cv_meshes <- function(sdmTMB_cv_output) {
+#   cv_output <- sdmTMB_cv_output
+#   k <- max(cv_output$data$cv_fold) # should probably rename cv_fold to fold_id
+#   if (k == 2 | k == 5 | k == 10) {
+#     op <- graphics::par(
+#       mfrow = c((sqrt(k)), ceiling(sqrt(k))),
+#       mar = c(1, 1, 1, 1)
+#       )
+#   } else {
+#     op <- graphics::par(
+#       mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
+#       mar = c(1, 1, 1, 1)
+#     )
+#   }
+#   for (i in seq_len(k)){
+#     plot_spde(cv_output$models[[i]]$spde)
+# #FIXME: Error in xy.coords(x, y, xlabel, ylabel, log) :
+# #  'x' is a list, but does not have components 'x' and 'y'
+#   }
+#   graphics::par(op)
+# }
 ```
@@
 ```{r, eval=FALSE}
-cv_args <- generate_cv_args (formula = list(
-    temperature_c ~ 0 + as.factor(year) + depth_scaled),
+cv_args <- generate_cv_args(
+  formula = list(
+    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3
+  ),
   time_varying = list(NULL),
@@
   include_spatial = c(FALSE),
-    n_knots = c(30), 
-    seed = c(111,222)
+  n_knots = c(135, 140, 145, 150),
+  seed = c(1)
 )
@@
 glimpse(cv$summary)
-plot_cv_meshes(cv$cv_model_set[[1]])
-plot_cv_meshes(cv$cv_model_set[[2]])
+write.csv(cv$summary, file = "cv-test.csv")
+cv_test <- read_csv(file = "cv-test.csv")
+glimpse(cv)
+View(cv$summary)
+# plot_cv_meshes(cv$cv_model_set[[1]])
+# plot_cv_meshes(cv$cv_model_set[[2]])
 ```
@@
 ```{r}
-cv_args <- generate_cv_args (formula = list(
+cv_args <- generate_cv_args(
+  formula = list(
+    temperature_c ~ 1 + SST + depth_scaled,
+    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2,
+    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2 + depth_scaled3,
     temperature_c ~ 0 + as.factor(year) + depth_scaled,
     temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
-    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3),
+    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
+    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled,
+    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2,
+    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3,
+    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled,
+    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2,
+    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2 + SST * depth_scaled3
+  ),
   time_varying = list(NULL),
@@
   include_spatial = c(FALSE),
-    n_knots = c(150), #start with close to projected optimum n_knots 
-    seed = c(1111,2221,3331)
+  n_knots = c(100),
+  seed = c(1111, 2222, 3333, 4444)
 )
@@
 ## Compare time-varying slope models still with simplest random effect structure
+All of top models for any seed contained year and non showed a linear effect of depth
 ```{r}
-cv_args <- generate_cv_args (formula = list(
-    temperature_c ~ 0 + as.factor(year)),
+cv_args <- generate_cv_args(
+  formula = list(
+    temperature_c ~ 0 + as.factor(year) + SST,
+    temperature_c ~ 0 + as.factor(year)
+  ),
   time_varying = list(
-    ~ 0 + depth_scaled,
     ~ 0 + depth_scaled + depth_scaled2,
-    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
+    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
+  ),
   ar1_fields = c(FALSE),
   include_spatial = c(FALSE),
-    n_knots = c(150), # keep same n_knots
-    seed = c(1111,2221,3331) # keep same seeds
+  n_knots = c(100), # keep same n_knots
+  seed = c(1111, 2222, 3333, 4444) # keep same seeds
 )
@@
 
-## Now vary n_knots and random effect configuation for top model(s) 
+## Now vary random effect configuation for top model(s) 
 ```{r}
-cv_args <- generate_cv_args(formula = list(
-    temperature_c ~ 0 + as.factor(year)),
+cv_args <- generate_cv_args(
+  formula = list(
+    temperature_c ~ 0 + as.factor(year) + SST,
+    temperature_c ~ 0 + as.factor(year)
+  ),
   time_varying = list(
-    ~ 0 + depth_scaled + depth_scaled2,
-    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
+    ~ 0 + depth_scaled + depth_scaled2
+  ),
   ar1_fields = c(FALSE, TRUE),
   include_spatial = c(FALSE, TRUE),
-    n_knots = c(85,110,135),
-    seed = c(1111,2221,3331)
+  n_knots = c(100),
+  seed = c(1111, 2222, 3333, 4444)
 )
@@
 ```{r}
-cv_args <- generate_cv_args(formula = list(
-    temperature_c ~ 0 + as.factor(year)),
+cv_args <- generate_cv_args(
+    formula = list(
+    temperature_c ~ 0 + as.factor(year) + SST,
+    temperature_c ~ 0 + as.factor(year)
+  ),
   time_varying = list(
-    ~ 0 + depth_scaled + depth_scaled2,
-    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3),
-    ar1_fields = c(FALSE), 
+    ~ 0 + depth_scaled + depth_scaled2
+  ),
+  ar1_fields = c(FALSE, TRUE),
   include_spatial = c(TRUE),
   n_knots = c(120),
-    seed = c(4441,5551,6661,7771,8881,9991)
+  seed = c(5555, 6666, 7777, 8888, 9999)
 )
@@
 Load saved batches:
-```{r}
+```{r, eval=FALSE}
 # non_time_varying <- readRDS(file = "cv-non-time-varying.rds")
@@
 
-# non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
-# time_varying <- read.csv(file = "cv-time-varying.csv")
-# vary_rf <- read.csv(file = "cv-varying-rf.csv")
-# knot_refining <- read.csv(file = "cv-refine-knots.csv")
+non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
+time_varying <- read.csv(file = "cv-time-varying.csv")
+vary_rf <- read.csv(file = "cv-varying-rf.csv")
+non_time_varying1 <- read.csv(file = "cv-non-time-varying1.csv")
+time_varying1 <- read.csv(file = "cv-time-varying1.csv")
+vary_rf1 <- read.csv(file = "cv-varying-rf1.csv")
+refine_knots1 <- read.csv(file = "cv-refine-knots1.csv")
+more_seeds <- read.csv(file = "cv-more-seeds.csv")
+more_seeds1 <- read.csv(file = "cv-more-seeds1.csv")
 
-all_cv_loglik <- rbind(non_time_varying$summary, time_varying$summary, non_tv$summary, tvar$summary, vary_rf$summary, vary_rf_knots$summary, knot_refining$summary, more_seeds$summary, deparse.level = 1, make.row.names=TRUE)
-View(all_cv_loglik)
+all_cv_loglik <- rbind(
+  non_time_varying,
+  time_varying,
+  vary_rf,
+  non_time_varying1,
+  time_varying1,
+  vary_rf1,
+  refine_knots1,
+  more_seeds,
+  more_seeds1,
+  deparse.level = 1, make.row.names = TRUE
+)
 
-by_seed <- all_cv_loglik %>% group_by(seed) %>% 
+by_seed <- all_cv_loglik %>%
+  group_by(seed) %>%
   mutate(., max_ll_per_seed = max(cv_loglik)) %>%
@@
 plan(multiprocess)
-temp3_rw <- sdmTMB_cv(data = d_trawl, n_knots = 130, x = "X", y ="Y", k_folds = 10, 
-                   formula = temperature_c ~ 0 + as.factor(year),
+temp3_rw <- sdmTMB_cv(
+  data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
+  formula = temperature_c ~ 0 + as.factor(year) + SST,
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
   time = "year", family = gaussian(link = "identity"),
-                   ar1_fields = FALS, include_spatial = TRUE)
-temp3_rw$models
+  ar1_fields = FALSE, include_spatial = TRUE, seed = 999
+)
+# temp3_rw$models
 sum(temp3_rw$sum_loglik)
@@
 m_temp3_rw <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year),
+  temperature_c ~ 0 + as.factor(year) + SST,
   time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
-  ar1_fields = TRUE, include_spatial = TRUE, 
+  ar1_fields = FALSE, include_spatial = TRUE,
   time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE)
+  silent = TRUE
+)
 
@@
 cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 #
-aic(m_temp3_rw)
+# aic(m_temp3_rw)
 ```
@@
 ## Plotting predictions
+#FIXME: DON'T HAVE SST TEMP FOR PREDICITON GRID YET...
+
 ```{r}
@@
 predictions_nd <- predict(m_temp3_rw, newdata = qcs_grid)
-#saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
+# saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
 ```
@@
   ggtitle("Spatial random effects only") +
-  scale_fill_gradient2()
+  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
 ```
@@
   ggtitle("Spatiotemporal random effects only") +
-  scale_fill_gradient2()
+  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
 ```
@@
   coord_fixed() +
-  xlim(3,15) +
-  ylim(3,15) +
+  xlim(3, 15) +
+  ylim(3, 15) +
   geom_abline()
@@
 
-temp_cog %>% reshape2::dcast(year ~ coord, value.var = "est") %>%
+temp_cog %>%
+  reshape2::dcast(year ~ coord, value.var = "est") %>%
   ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +

### 2019-04-19 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
exploring effect of daily SST on bottom temperature

@@
 plan(multiprocess)
+```
+
+Load data
+```{r}
 # d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
@@
 d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
-glimpse(d_trawl)
 nrow(d_trawl)
-head(d_trawl)
 # d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
+unique(d_trawl$year)
+d_trawl[is.na(d_trawl$year), ]$year <- 2003
+d_trawl <- d_trawl[!is.na(d_trawl$depth_m), ]
+d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
+d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
 d_trawl <- mutate(d_trawl,
   depth_scaled = arm::rescale(log(depth_m)),
+  SST_scaled = arm::rescale(SST),
   depth_scaled2 = depth_scaled^2,
-  depth_scaled3 = depth_scaled^3
+  depth_scaled3 = depth_scaled^3,
+  shallow = ifelse(depth_m>50,0,1)
 )
-unique(d_trawl$year)
-d_trawl[is.na(d_trawl$year), ]$year <- 2003
+glimpse(d_trawl)
+```
+
+Explore SST 
+```{r}
+ggplot(d_trawl, aes_string("X", "Y", colour = "SST")) +
+    geom_point() +
+    facet_wrap(~year) +
+    coord_fixed() +
+    scale_color_viridis_c()
+```
+
+```{r}
+ggplot(dplyr::filter(d_trawl, depth_m <50), aes(SST,temperature_c, colour = depth_m)) + geom_point()
+```
+
+```{r}
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 130)
+plot_spde(spde)
+```
+
+```{r}
+m_temp3 <- sdmTMB(d_trawl,
+  temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
+  ar1_fields = FALSE, include_spatial = TRUE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE
+)
+
+m_temp3_nosst <- sdmTMB(d_trawl,
+  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
+  ar1_fields = FALSE, include_spatial = TRUE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE
+)
+
+m_temp3
+m_temp3_nosst
+```
+
+```{r}
+
+predictions3 <- predict(m_temp3)
+predictions3b <- predict(m_temp3_nosst)
+
+predictions3$data$residuals <- residuals(m_temp3)
+predictions3b$data$residuals <- residuals(m_temp3_nosst)
+
+nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth_m)
+
+ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
+  geom_point() + geom_abline(intercept = 0, slope = 1)
 ```
@@
 glimpse(cv$summary)
-write.csv(cv$summary, file = "cv-test.csv")
-cv_test <- read_csv(file = "cv-test.csv")
-glimpse(cv)
+write.csv(cv$summary, file = "cv-output/cv-test.csv")
+cv_test <- read_csv(file = "cv-output/cv-test.csv")
 View(cv$summary)
@@
 View(non_time_varying$summary)
-write.csv(non_time_varying$summary, file = "cv-non-time-varying.csv")
-saveRDS(non_time_varying, file = "cv-non-time-varying.rds")
+write.csv(non_time_varying$summary, file = "cv-output/cv-non-time-varying.csv")
+saveRDS(non_time_varying, file = "cv-output/cv-non-time-varying.rds")
 glimpse(non_time_varying$summary)
@@
 ## Compare time-varying slope models still with simplest random effect structure
-All of top models for any seed contained year and non showed a linear effect of depth
+All of top models for any seed contained year and none showed a linear effect of depth
+All are better than fixed slope models
 ```{r}
@@
 View(time_varying$summary)
-write.csv(time_varying$summary, file = "cv-time-varying.csv")
-saveRDS(time_varying, file = "cv-time-varying.rds")
+write.csv(time_varying$summary, file = "cv-output/cv-time-varying.csv")
+saveRDS(time_varying, file = "cv-output/cv-time-varying.rds")
 glimpse(time_varying$summary)
@@
 ## Now vary random effect configuation for top model(s) 
+a separate spatial field is in all models with highest likelihoods for each seed
+AR1 doesn't make much difference (differences between years appear independent)
 ```{r}
@@
 
-vary_rf_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
+vary_rf <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
 
-View(vary_rf_knots$summary)
-write.csv(vary_rf_knots$summary, file = "cv-varying-rf.csv")
-saveRDS(vary_rf_knots, file = "cv-varying-rf.rds")
-glimpse(vary_rf_knots$summary)
+View(vary_rf$summary)
+write.csv(vary_rf$summary, file = "cv-output/cv-varying-rf.csv")
+saveRDS(vary_rf, file = "cv-output/cv-varying-rf.rds")
+glimpse(vary_rf$summary)
 ```
 
-## Change seeds 
+## Change knot number
+
 ```{r}
@@
   ),
-  ar1_fields = c(FALSE, TRUE),
+  ar1_fields = c(FALSE),
   include_spatial = c(TRUE),
-  n_knots = c(120),
+  n_knots = c(80,120,140),
+  seed = c(1111, 2222, 3333, 4444)
+)
+
+more_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
+
+View(more_knots$summary)
+write.csv(more_knots$summary, file = "cv-output/cv-more-seeds.csv")
+saveRDS(more_knots, file = "cv-output/cv-more-seeds.rds")
+glimpse(more_knots$summary)
+```
+
+## Change seeds 
+Also increased knot number too 
+Model without SST is always very slightly higher ll
+
+```{r}
+cv_args <- generate_cv_args(
+    formula = list(
+    temperature_c ~ 0 + as.factor(year) + SST,
+    temperature_c ~ 0 + as.factor(year)
+  ),
+  time_varying = list(
+    ~ 0 + depth_scaled + depth_scaled2
+  ),
+  ar1_fields = c(FALSE),
+  include_spatial = c(TRUE),
+  n_knots = c(130),
   seed = c(5555, 6666, 7777, 8888, 9999)
@@
 View(more_seeds$summary)
-write.csv(more_seeds$summary, file = "cv-more-seeds.csv")
-saveRDS(more_seeds, file = "cv-more-seeds.rds")
+write.csv(more_seeds$summary, file = "cv-output/cv-more-seeds.csv")
+saveRDS(more_seeds, file = "cv-output/cv-more-seeds.rds")
 glimpse(more_seeds$summary)
@@
 ```{r, eval=FALSE}
-# non_time_varying <- readRDS(file = "cv-non-time-varying.rds")
-# time_varying <- readRDS(file = "cv-time-varying.rds")
-# vary_rf <- readRDS(file = "cv-varying-rf.rds")
-# knot_refining <- readRDS(file = "cv-refine-knots.rds")
+# non_time_varying <- readRDS(file = "cv-output/cv-non-time-varying.rds")
+# time_varying <- readRDS(file = "cv-output/cv-time-varying.rds")
+# vary_rf <- readRDS(file = "cv-output/cv-varying-rf.rds")
+# knot_refining <- readRDS(file = "cv-output/cv-refine-knots.rds")
 
-non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
-time_varying <- read.csv(file = "cv-time-varying.csv")
-vary_rf <- read.csv(file = "cv-varying-rf.csv")
-non_time_varying1 <- read.csv(file = "cv-non-time-varying1.csv")
-time_varying1 <- read.csv(file = "cv-time-varying1.csv")
-vary_rf1 <- read.csv(file = "cv-varying-rf1.csv")
-refine_knots1 <- read.csv(file = "cv-refine-knots1.csv")
-more_seeds <- read.csv(file = "cv-more-seeds.csv")
-more_seeds1 <- read.csv(file = "cv-more-seeds1.csv")
+non_time_varying <- read.csv(file = "cv-output/cv-non-time-varying.csv")
+time_varying <- read.csv(file = "cv-output/cv-time-varying.csv")
+vary_rf <- read.csv(file = "cv-output/cv-varying-rf.csv")
+non_time_varying1 <- read.csv(file = "cv-output/cv-non-time-varying1.csv")
+time_varying1 <- read.csv(file = "cv-output/cv-time-varying1.csv")
+vary_rf1 <- read.csv(file = "cv-output/cv-varying-rf1.csv")
+refine_knots1 <- read.csv(file = "cv-output/cv-refine-knots1.csv")
+more_seeds <- read.csv(file = "cv-output/cv-more-seeds.csv")
+more_seeds1 <- read.csv(file = "cv-output/cv-more-seeds1.csv")
 
@@
 plan(multiprocess)
-temp3_rw <- sdmTMB_cv(
+temp_rw <- sdmTMB_cv(
   data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
@@
 )
-# temp3_rw$models
-sum(temp3_rw$sum_loglik)
+sum(temp_rw$sum_loglik)
 ```
@@
 ```{r}
-m_temp3_rw <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year) + SST,
-  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+m_temp_rw <- sdmTMB(d_trawl,
+  temperature_c ~ 0 + as.factor(year),
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
   ar1_fields = FALSE, include_spatial = TRUE,
@@
 
-predictions3_rw <- predict(m_temp3_rw)
-predictions3_rw$data$residuals <- residuals(m_temp3_rw)
-cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 #
+predictions_rw <- predict(m_temp_rw)
+predictions_rw$data$residuals <- residuals(m_temp_rw)
+cor(predictions_rw$data$est, predictions_rw$data$temperature_c)^2 #
 # aic(m_temp3_rw)
@@
 ```{r}
-ggplot(predictions3_rw$data, aes(est, residuals)) +
+ggplot(predictions_rw$data, aes(est, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
+ggplot(predictions_rw$data, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
+ggplot(predictions_rw$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth()
+```
+
+Does a fixed interaction with SST do any better?
+
+```{r}
+m_temp3 <- sdmTMB(d_trawl,
+  temperature_c ~ 0 + as.factor(year) + SST*depth_scaled + SST*depth_scaled2 + SST*depth_scaled3,
+  ar1_fields = FALSE, include_spatial = TRUE,
+  time = "year", spde = spde, family = gaussian(link = "identity"),
+  silent = TRUE
+)
+
+predictions3 <- predict(m_temp3)
+predictions3$data$residuals <- residuals(m_temp3)
+cor(predictions3$data$est, predictions3$data$temperature_c)^2 #
+```
+
+```{r}
+ggplot(predictions3$data, aes(est, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(predictions3$data, aes(depth_scaled, residuals)) +
   geom_point() +

### 2019-04-19 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
move rds files to data folder

@@
 ```{r}
-# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
+# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/data/sensor-data-processed.rds"))
 # nrow(d_trawl1)
-d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
+d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-SST.rds"))
 nrow(d_trawl)

### 2019-04-19 analysis/tmb-sensor-explore/1-get-SST-data.R

@@ -13,14 +13,14 @@
 #   ssid = c(1),
 #   attribute = c("temperature"),
 #   spread_attributes = FALSE)
-## saveRDS(trawl_depth, file = "analysis/tmb-sensor-explore/dat-sensor-trawl-depth.rds")
-## saveRDS(trawl_temp, file = "analysis/tmb-sensor-explore/dat-sensor-trawl-temp.rds")
+## saveRDS(trawl_depth, file = "analysis/tmb-sensor-explore/data/dat-sensor-trawl-depth.rds")
+## saveRDS(trawl_temp, file = "analysis/tmb-sensor-explore/data/dat-sensor-trawl-temp.rds")
 #
 # TEMPORARY CODE TO WRANGLE SENSOR DATA FROM MULTIPLE SENSORS
 #TODO: need to refine gfdata SQL query output before finalizing any of this...
 
-# trawl_depth <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-depth.rds"))
-# trawl_temp <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-temp.rds"))
+# trawl_depth <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-depth.rds"))
+# trawl_temp <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-temp.rds"))
 # library(dplyr)
 # trawl_date <- trawl_temp %>% select (fishing_event_id, survey_desc, start_time, end_time) %>% distinct() 
 # trawl_date$date <- format(as.Date(trawl_date$start_time), "%Y-%m-%d")
@@ -44,9 +44,9 @@
 # sd_trawl <- right_join(trawl_date, sd_trawl1, by ="fishing_event_id")
 # 
 # glimpse(sd_trawl)
-## saveRDS(sd_trawl, file = "analysis/tmb-sensor-explore/dat-sensor-trawl-processed.rds")
+## saveRDS(sd_trawl, file = "analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")
 
-sd_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-processed.rds"))
+sd_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds"))
[…]

### 2019-04-23 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
test if meanSST does any better than SST--yes, but not enough to really be useful

--- 2019-04-19 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 ```{r}
-# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
+# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/data/sensor-data-processed.rds"))
 # nrow(d_trawl1)
-d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
+d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST.rds"))
 nrow(d_trawl)
@@
 unique(d_trawl$year)
-d_trawl[is.na(d_trawl$year), ]$year <- 2003
 d_trawl <- d_trawl[!is.na(d_trawl$depth_m), ]
@@
 d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
+d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
 d_trawl <- mutate(d_trawl,
   depth_scaled = arm::rescale(log(depth_m)),
-  SST_scaled = arm::rescale(SST),
+  SST_scaled = arm::rescale(meanSST),
   depth_scaled2 = depth_scaled^2,
   depth_scaled3 = depth_scaled^3,
-  shallow = ifelse(depth_m>50,0,1)
+  shallow = ifelse(depth_m>40,0,1)
 )
+d_trawl[is.na(d_trawl$shallow), ]
+d_trawl[is.na(d_trawl$SST_scaled), ]
+
 glimpse(d_trawl)
@@
 ```{r}
-ggplot(dplyr::filter(d_trawl, depth_m <50), aes(SST,temperature_c, colour = depth_m)) + geom_point()
+ggplot(dplyr::filter(d_trawl, depth_m <60), aes(SST,temperature_c, colour = -depth_m)) + geom_point()
+```
+
+Explore mean  SST 
+```{r}
+ggplot(d_trawl, aes_string("X", "Y", colour = "meanSST")) +
+    geom_point() +
+    facet_wrap(~year) +
+    coord_fixed() +
+    scale_color_viridis_c()
+```
+
+```{r}
+ggplot(dplyr::filter(d_trawl, depth_m <50), aes(meanSST,temperature_c, colour = -depth_m)) + geom_point()
 ```
@@
 m_temp3 <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
+  temperature_c ~ 0 + SST_scaled*shallow, 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
   ar1_fields = FALSE, include_spatial = TRUE,
@@
 
-m_temp3_nosst <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
+m_temp3b <- sdmTMB(d_trawl,
+  temperature_c ~ 0 + as.factor(year), 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
   ar1_fields = FALSE, include_spatial = TRUE,
@@
 predictions3 <- predict(m_temp3)
-predictions3b <- predict(m_temp3_nosst)
+predictions3b <- predict(m_temp3b)
+#predictions3b <- predict(m_temp3_nosst)
 
 predictions3$data$residuals <- residuals(m_temp3)
-predictions3b$data$residuals <- residuals(m_temp3_nosst)
+predictions3b$data$residuals <- residuals(m_temp3b)
 
@@
   geom_point() + geom_abline(intercept = 0, slope = 1)
+```
+Residual plots:
+```{r}
+ggplot(predictions3$data, aes(est, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+ggplot(predictions3b$data, aes(est, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  ylim(-7,8) +
+  geom_smooth()
+ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  ylim(-7,8) +
+  geom_smooth()
 ```
@@
   formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3
+    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3
   ),
@@
   formula = list(
-    temperature_c ~ 1 + SST + depth_scaled,
-    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2,
-    temperature_c ~ 1 + SST + depth_scaled + depth_scaled2 + depth_scaled3,
+    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled,
+    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2,
+    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
+    temperature_c ~ 1 + SST_scaled + depth_scaled,
+    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2,
+    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
     temperature_c ~ 0 + as.factor(year) + depth_scaled,
@@
     temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
-    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled,
-    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2,
-    temperature_c ~ 0 + as.factor(year) + SST + depth_scaled + depth_scaled2 + depth_scaled3,
-    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled,
-    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2,
-    temperature_c ~ 0 + as.factor(year) + SST * depth_scaled + SST * depth_scaled2 + SST * depth_scaled3
+    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2 + SST_scaled * depth_scaled3
   ),
@@
   formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST,
+    #temperature_c ~ 0 + SST_scaled*shallow,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
+    #temperature_c ~ 1 + SST_scaled,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled,
     temperature_c ~ 0 + as.factor(year)
@@
   formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled,
     temperature_c ~ 0 + as.factor(year)
@@
     formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST,
+    # temperature_c ~ 0 + as.factor(year) + SST_scaled,
     temperature_c ~ 0 + as.factor(year)
@@
   ),
-  ar1_fields = c(FALSE),
+  ar1_fields = c(FALSE,TRUE),
   include_spatial = c(TRUE),
-  n_knots = c(80,120,140),
-  seed = c(1111, 2222, 3333, 4444)
+  n_knots = c(140,180,220,250),
+  #seed = c(1111, 2222, 3333, 4444) 
+  seed = c(5555, 6666, 7777, 8888, 9999)
 )
@@
 View(more_knots$summary)
-write.csv(more_knots$summary, file = "cv-output/cv-more-seeds.csv")
-saveRDS(more_knots, file = "cv-output/cv-more-seeds.rds")
+write.csv(more_knots$summary, file = "cv-output/cv-more-knots2.csv")
+saveRDS(more_knots, file = "cv-output/cv-more-knots2.rds")
 glimpse(more_knots$summary)
@@
 Also increased knot number too 
-Model without SST is always very slightly higher ll
+Models without SST_scaled are usually better
 
@@
     formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
+    temperature_c ~ 0 + as.factor(year) + SST_scaled,
     temperature_c ~ 0 + as.factor(year)
@@
   ),
-  ar1_fields = c(FALSE),
-  include_spatial = c(TRUE),
+  ar1_fields = c(FALSE, TRUE),
+  include_spatial = c(TRUE, FALSE),
   n_knots = c(130),
@@
   data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
-  formula = temperature_c ~ 0 + as.factor(year) + SST,
+  formula = temperature_c ~ 0 + as.factor(year) + SST_scaled,
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
 ```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 130)
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 220)
 plot_spde(spde)
@@
 
-Does a fixed interaction with SST do any better?
+Does a fixed interaction with SST_scaled do any better?
 
@@
 m_temp3 <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year) + SST*depth_scaled + SST*depth_scaled2 + SST*depth_scaled3,
+  temperature_c ~ 0 + as.factor(year) + SST_scaled*depth_scaled + SST_scaled*depth_scaled2 + SST_scaled*depth_scaled3,
   ar1_fields = FALSE, include_spatial = TRUE,

### 2019-04-25 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
tweaks to temp models

@@
 ```{r}
-# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/data/sensor-data-processed.rds"))
-# nrow(d_trawl1)
-d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST.rds"))
-nrow(d_trawl)
-# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
-unique(d_trawl$year)
-d_trawl <- d_trawl[!is.na(d_trawl$depth_m), ]
+all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
+# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
+# glimpse(all_sensor)
+# 
+# replace obviously faulty depth data with NA
+all_sensor$depth_m[all_sensor$depth_m < 10] <- NA
+
+all_depth <- all_sensor %>%
+  # filter missing location data and trial year
+  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%
+
+  # convert lat and lon to UTMs
+  dplyr::mutate(X = longitude, Y = latitude) %>%
+  gfplot:::ll2utm(., utm_zone = 9) %>%
+
+  # interpolate missing depth values
+  dplyr::rename(depth = depth_m) %>%
+  gfplot:::interp_survey_bathymetry()
+
+unique(all_depth$data$year)
+
+d_trawl <- all_depth$data %>% 
+  # filter(ssid != 1) %>%
+  # filter(ssid != 3) %>%
+  # filter(ssid != 4) %>%
+  # filter(ssid != 16) %>%
+  gfplot:::scale_survey_predictors()
+
+hist(d_trawl$temperature_c)
+
+# remove obvious sensor fail, d_trawl %>% filter(ssid==4, temperature_c >12), fishing event id = 481861
+d_trawl <- d_trawl[d_trawl$fishing_event_id != 481861, ]
+d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
 d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
@@
 d_trawl <- mutate(d_trawl,
-  depth_scaled = arm::rescale(log(depth_m)),
+  DOY = lubridate::yday(date),
+  DOY_scaled = arm::rescale(lubridate::yday(date)),
+  #depth_scaled = arm::rescale(log(depth_m)),
   SST_scaled = arm::rescale(meanSST),
-  depth_scaled2 = depth_scaled^2,
+  #depth_scaled2 = depth_scaled^2,
   depth_scaled3 = depth_scaled^3,
-  shallow = ifelse(depth_m>40,0,1)
+  shallow = ifelse(depth>35,0,1)
 )
-d_trawl[is.na(d_trawl$shallow), ]
-d_trawl[is.na(d_trawl$SST_scaled), ]
+# d_trawl[is.na(d_trawl$shallow), ]
+# d_trawl[is.na(d_trawl$SST_scaled), ]
 
 glimpse(d_trawl)
+
+d_trawl <- d_trawl %>% dplyr::filter(year>2003)
 ```
@@
 ```
+```{r}
+ggplot(d_trawl, aes(DOY, SST, colour = -depth)) + geom_point() + facet_wrap(~year) 
+```
 
 ```{r}
-ggplot(dplyr::filter(d_trawl, depth_m <60), aes(SST,temperature_c, colour = -depth_m)) + geom_point()
+ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = -depth)) + geom_point()
+ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = DOY)) + geom_point()
+
 ```
@@
 ```{r}
-ggplot(dplyr::filter(d_trawl, depth_m <50), aes(meanSST,temperature_c, colour = -depth_m)) + geom_point()
+ggplot(dplyr::filter(d_trawl, depth <35), aes(meanSST,temperature_c, colour = -depth)) + geom_point()
+```
+
+EXLORE TEMPERATURE at DEPTH
+
+```{r}
+ggplot(d_trawl, aes_string("X", "Y", colour = "temperature_c")) +
+    geom_point() +
+    facet_wrap(~year) +
+    coord_fixed() +
+    scale_color_viridis_c()
 ```
@@
 ```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 130)
+ggplot(d_trawl, aes(DOY, temperature_c, colour = -depth, alpha=0.5)) + geom_point() + facet_wrap(~year) 
+```
+
+```{r}
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 650)
 plot_spde(spde)
@@
 m_temp3 <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + SST_scaled*shallow, 
-  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+  temperature_c ~ 0 + as.factor(year) + shallow*poly(DOY_scaled,2), 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
   ar1_fields = FALSE, include_spatial = TRUE,
@@
   temperature_c ~ 0 + as.factor(year), 
-  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
   ar1_fields = FALSE, include_spatial = TRUE,
@@
 m_temp3
-m_temp3_nosst
+m_temp3b
 ```
@@
 
-nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth_m)
+nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth)
 
@@
   geom_smooth() +
+  ylim(-8,8) +
+  xlim(2,15) +
   facet_wrap(~year)
@@
   geom_smooth() +
+  ylim(-8,8) +
+   xlim(2,15) +
   facet_wrap(~year)
@@
   geom_smooth() +
+  ylim(-7,8) +
+  facet_wrap(~year)
+ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
+  geom_point() +
+  geom_smooth() +
+  ylim(-7,8) +
   facet_wrap(~year)
@@
   geom_point() +
-  ylim(-7,8) +
+  ylim(-8,8) +
   geom_smooth()

### 2019-05-14 analysis/resources/simple-sdmTMB.Rmd
move exploratory code to analysis/resources

--- 2019-02-01 analysis/binom-sdmTMB.Rmd
@@
 ---
-title: "Binomial TMB"
+title: "Practice TMB"
 output: html_notebook
 ---
+
+```{r setup}
+    knitr::opts_knit$set(root.dir = normalizePath("/Users/capitan/github/dfo/gfranges/")) 
+library(sdmTMB)
+
+```
+
+Load data from cache
+
+```{r}
+d <- readRDS("data-cache/pacific-cod.rds")
+d$density <- d$density_kgpm2
+```
+
+```{r}
+# Gaussian:
+pcod_gaus <- subset(d, density > 0 & year >= 2013)
+
+pcod_spde_gaus <- make_spde(pcod_gaus$X, pcod_gaus$Y, n_knots = 50) #need to calculate X and Y (utms?)
+m_pos <- sdmTMB(pcod_gaus,
+  log(density) ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
+  time = "year", spde = pcod_spde_gaus)
+
+```
 

### 2019-05-14 analysis/resources/tweedie-TMB.Rmd
move exploratory code to analysis/resources

--- 2019-02-20 analysis/tweedie-TMB.Rmd
@@
 ```{r set-up}
-## Load TMB TRUE
+## Load TMB 
 library(TMB)
@@
 data <- list(y = y, x = x)
-parameters <- list(a=0, b=0, log_phi = log(1), p = 0 )
+parameters <- list(a=0, b=0, p = 0, log_phi = log(1) )
 
@@
 plot(b_profile)
-
 ```
@@
   mu <- exp(eta)
-  p <- plogis(par[3]) + 1
   phi <- exp(par[4])
+  power <- plogis(par[3]) + 1
   # sum of negative log likelihoods:
-  loglik <- log(dtweedie(y, power = p, mu = mu, phi = phi))
+  loglik <- log(dtweedie(y, power = power, mu = mu, phi = phi))
   nll <- -sum(loglik) #[is.finite(loglik)]
@@
 
-tweedie_ml <- nlminb(start = c(a = 0, b = 0, ipx = 0, log_phi = log(1)), nll_tweedie)
+tweedie_ml <- nlminb(start = c(a = 0, b = 0,  logis_p = 0, log_phi = log(1)), nll_tweedie)
 tweedie_ml$par 
@@
 ```{r}
+library(statmod)
 tweedie_glm <- glm(y ~ x, family = tweedie(var.power = 1.5, link.power = 0))

### 2019-05-15 analysis/VOCC/dist-vocc.Rmd
RMD of analysis including draft function for producing vocc data from sdmTMB output

---
title: "Distanced-based VOCC analysis"
author: "Philina English"
date: '2019-05-14'
output: html_document
---

```{r setup, include=FALSE}
# install.packages("yaImpute") # install package for k-nearest neighbour (kNN) search
# install.packages("raster") # required for calcslope + all raster functions
# install.package(gfplot) # required for some plotting and data functions

library(dplyr)
library(ggplot2)
library(sdmTMB)
```

Retrieve raw data
```{r}
all_depth <- readRDS("../tmb-sensor-explore/data/dat-sensor-trawl-all-depth.rds")

# add and scale predictors after filtering for survey(s) of interest
data <- all_depth$data %>%
  filter(depth_max>10) %>% # removes data if sensor stayed too near surface
# dplyr::filter(year > 2003) %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

data <- data[data$fishing_event_id != 481861, ]
# add any new predictors
# data <- data %>%
#   dplyr::mutate(
#   DOY = lubridate::yday(date),
#   shallow = ifelse(depth > 35, 0, 1)
# )
```

Create prediction grids for each survey area so that only years with data are included
```{r}
# choose base year(s) to create grid from
dummy_year <- c(2004, 2005)

# create grids for each ssid separately 
ssid <- 1
survey_abbrev <- "SYN QCS"
nd_1 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_1$year)

ssid <- 3
survey_abbrev <- "SYN HS"
nd_3 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_3$year)

ssid <- 4
survey_abbrev <- "SYN WCVI"
nd_4 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_4$year)

ssid <- 16
survey_abbrev <- "SYN WCHG"
nd_16 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year = 2006)
unique(nd_16$year)

nd <- rbind(nd_1, nd_3, nd_4, nd_16)
# nd <- nd_4 # for faster test option
nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))

```

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
# # choose the spatial range to model
dat <- data # %>% filter(ssid == 4)
nd <- nd # %>% filter(ssid == 4)

spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 500)
sdmTMB::plot_spde(spde)

m_temp <- sdmTMB::sdmTMB(dat,
  temperature_c ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = gaussian(link = "identity"),
  ar1_fields = TRUE, # maybe TRUE is better for all areas combined?
  include_spatial = TRUE,
  silent = FALSE
)

# stopifnot(m_temp$model$convergence == 0L)
# m_temp
# # Warning messages:
# #   1: In doTryCatch(return(expr), name, parentenv, handler) :
# #   restarting interrupted promise evaluation
# 
saveRDS(m_temp, file = "data/m_temp_allyears.rds")
```

Calculate predicted values
```{r}
m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
# m_temp <- readRDS("../VOCC/data/m_temp_allpost2003.rds")

predictions <- predict(m_temp, newdata = nd)

plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

p <- plot_map(predictions , "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```
```{r}
# choose the spatial range to build raster on
predicted1 <- predictions %>% filter(ssid == 1) 
predicted3 <- predictions %>% filter(ssid == 3) 
predicted4 <- predictions %>% filter(ssid == 4) 
predicted16 <- predictions %>% filter(ssid == 16)

p <- plot_map(predicted1 , "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```

```{r}

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted3, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-hs.rds")

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted1, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-qcs.rds")

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted4, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-wcvi.rds")

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted16, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-wchg.rds")
```

```{r}

vector_data <- function(climate_data, 
  ssid = NULL,
  input_cell_size = 2, 
  scale_fac = 2,
  indices = c(1, 1, 1, 2, 2, 3, 3, 3),
  thresholds = c(0.75),
  delta_t = 10,
  time_
  ) {

climate_data <- climate_data %>% filter(ssid == ssid)

if (!is.list(climate_data)) {

rbrick <- make_raster_brick(climate_data, scale_fac = scale_fac)
mnraster_brick <- raster::stackApply(rbrick, indices = indices, fun = mean)
start_raster <- mnraster_brick[[1]]
end_raster <- mnraster_brick[[3]]

# make sparate named lists containing climate rasters or dataframes
# data with just one climate variable
start_data <- list(var_1 = start_raster)
end_data <- list(var_1 = end_raster)

} else {
rbrick <- list()
for (i in seq_len(climate_data)) {

rbrick[[i]] <- make_raster_brick(climate_data[[i]], scale_fac = scale_fac)
mnraster_brick <- raster::stackApply(rbrick[[i]], indices = indices, fun = mean)

start_raster <- mnraster_brick[[1]]
end_raster <- mnraster_brick[[3]]
var <- list(start_raster = start_raster, end_raster = end_raster)
}

start_data <- list(var[[i]] = var$start_raster[[i]])
end_data <- list(var[[i]] = var$end_raster[[i]])
}

browser()
start_data #
end_data  #

out <- dist_based_vocc(
  start_data = start_data,
  end_data = end_data,
  x = "x",
  y = "y",
  variable_names = c("index_1"), # what the layer within each element is called
  thresholds = thresholds,
  cell_size =  input_cell_size*scal_fac,
  delta_t = delta_t,
  raster = TRUE
  )

slopedat <-  calcslope(rbrick, time_slices = 10) # vocc::calcslope for comparison
out <- left_join(out, slopedat, by = c("x", "y")) %>% select(-icell)
out$C_per_decade <- out$slope * 10 / out$time_scale
out$km_per_decade <- out$distance * 10 / delta_t

}

vector_data (predictions, 
  ssid = 1,
  input_cell_size = 2, 
  scale_fac = 2,
  indices = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
  thresholds = c(0.75),
  delta_t = 10,
  )
```

### 2019-05-17 analysis/VOCC/dist-vocc.Rmd
fix calcslope to always have time axis in single units

@@
 library(sdmTMB)
+library(gfranges)
 ```
@@
   filter(depth_max>10) %>% # removes data if sensor stayed too near surface
-# dplyr::filter(year > 2003) %>% 
   # filter(ssid != 1) %>%
@@
 # # choose the spatial range to model
-dat <- data # %>% filter(ssid == 4)
+dat <- data %>% dplyr::filter(year > 2003) #%>% filter(ssid == 4)
 nd <- nd # %>% filter(ssid == 4)
@@
 m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
+
+# can also be filtered to include only year with multiple surveys 
 # m_temp <- readRDS("../VOCC/data/m_temp_allpost2003.rds")
+# nd <- nd %>% dplyr::filter(year > 2003)
 
@@
 predicted16 <- predictions %>% filter(ssid == 16) 
+predicted1n3 <- predictions %>% filter(ssid != 16) %>% filter(ssid != 4) %>% filter(year > 2004)
 
-p <- plot_map(predicted1 , "est") +
+p <- plot_map(predicted1n3 , "est") +
   scale_fill_viridis_c(trans = "sqrt", option = "C") +
@@
 saveRDS(rbrick, file = "analysis/data/rbrick-temp-wchg.rds")
+
+rbrick <- make_raster_brick(predicted1n3, scale_fac = 2)
+saveRDS(rbrick, file = "analysis/data/rbrick-temp-hs-qcs.rds")
+
 ```
@@
 
-vector_data <- function(climate_data, 
+make_vector_data <- function(climate_data, 
   ssid = NULL,
+  start_year = NULL,
   input_cell_size = 2, 
   scale_fac = 2,
-  indices = c(1, 1, 1, 2, 2, 3, 3, 3),
-  thresholds = c(0.75),
-  delta_t = 10,
-  time_
+  delta_t_total = 10,
+  delta_t_step = 2,
+  time_slices = 7, 
+  indices = c(1, 1, 2, 2, 2, 3, 3),
+  thresholds = c(0.75)
   ) {
+#browser()
+if (!is.null(ssid)) climate_data <- climate_data[climate_data$ssid %in% ssid, ]
+if (!is.null(start_year)) climate_data <- climate_data %>% dplyr::filter(year >= start_year)
 
-climate_data <- climate_data %>% filter(ssid == ssid)
+#if (!identical(time_slices, length(indices)))
+  length_indices <- length(indices)
+  if(!isTRUE(time_slices==length_indices))
+      stop("Must have an indice assigned to each time slice,",
+      "therefore length('indices') must equal 'time_slices'.",
+      call. = FALSE)
 
-if (!is.list(climate_data)) {
+#FIXME: need way of determining how many variables(in separate dataframes) are in dataset  
+#if (!is.list(climate_data)) {
+# if (length(thresholds) > 1)
+#       stop("If multiple climate variables (and corresponding thresholds),", 
+#         "data must be in list form.", call. = FALSE)
 
@@
 
-} else {
-rbrick <- list()
-for (i in seq_len(climate_data)) {
+# } else {
+#     if (!identical(length(climate_data), length(thresholds)))
+#       stop("Must have a 'thresholds' value for each climate variable", call. = FALSE)
+# rbrick <- list()
+# for (i in seq_len(climate_data)) {
+#   
+# rbrick[[i]] <- make_raster_brick(climate_data[[i]], scale_fac = scale_fac)
+# mnraster_brick <- raster::stackApply(rbrick[[i]], indices = indices, fun = mean)
+# 
+# start_raster <- mnraster_brick[[1]]
+# end_raster <- mnraster_brick[[3]]
+# var <- list(start_raster = start_raster, end_raster = end_raster)
+# }
 
-rbrick[[i]] <- make_raster_brick(climate_data[[i]], scale_fac = scale_fac)
-mnraster_brick <- raster::stackApply(rbrick[[i]], indices = indices, fun = mean)
+# start_data <- list(var = var$start_raster[[i]])
+# end_data <- list(var = var$end_raster[[i]])
+#}
 
-start_raster <- mnraster_brick[[1]]
-end_raster <- mnraster_brick[[3]]
-var <- list(start_raster = start_raster, end_raster = end_raster)
-}
-
-start_data <- list(var[[i]] = var$start_raster[[i]])
-end_data <- list(var[[i]] = var$end_raster[[i]])
-}
-
-browser()
 start_data #
@@
   cell_size =  input_cell_size*scal_fac,
-  delta_t = delta_t,
+  delta_t = delta_t_total,
   raster = TRUE
@@
 
-slopedat <-  calcslope(rbrick, time_slices = 10) # vocc::calcslope for comparison
+# FIXME: need to change function to deal with different time steps within a brick
+slopedat <-  calcslope(rbrick, delta_t = delta_t_step) # vocc::calcslope for comparison
 out <- left_join(out, slopedat, by = c("x", "y")) %>% select(-icell)
-out$C_per_decade <- out$slope * 10 / out$time_scale
-out$km_per_decade <- out$distance * 10 / delta_t
-
+out$C_per_decade <- out$slope * 10
+out$km_per_decade <- (out$distance/delta_t)*10 
+out
 }
 
-vector_data (predictions, 
-  ssid = 1,
+make_vector_data (predictions, 
+  ssid = c(1, 3),
+  start_year = 2005,
   input_cell_size = 2, 
   scale_fac = 2,
-  indices = c(1, 1, 1, 2, 2, 2, 3, 3, 3),
-  thresholds = c(0.75),
-  delta_t = 10,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  time_slices = 7,
+  indices = c(1, 1, 2, 2, 2, 3, 3),
+  thresholds = c(0.75)
   )

### 2019-05-17 R/calcslope.R

@@ -1,11 +1,11 @@
 #' Calculate trend in varible for pixels in a raster brick
 #'
 #' @param rx Raster brick containing layers for multiple years. 
-#' @param time_units Number of years to calculate slope across. 
+#' @param delta_t Number of time steps between layers, if not 1 allows slope to be calculated for a single unit of time. 
 #' @param na.rm Logical for if NAs are removed.
 #'
 #' @export
-calcslope <- function(rx, time_units = 10, na.rm = TRUE) {
+calcslope <- function(rx, delta_t = 2, na.rm = TRUE) {
   
   # this function is a modified version of Chris Brown's (https://github.com/cbrown5/vocc/blob/master/R/calcslope.R)
   # gives same result in intial tests, but may be slower
@@ -13,15 +13,14 @@ calcslope <- function(rx, time_units = 10, na.rm = TRUE) {
   
   icell <- seq(1, raster::ncell(rx))
   coord <- raster::xyFromCell(rx, icell)
-  #  browser()
   y <- t(raster::getValues(rx))
   t <- row(y) # matrix of times for all cells ... assumes equal time between slices
   #FIXME: Can we make this robust to uneven timesteps?
-  
+  # browser()
   x1 <- y
   x1[!is.na(x1)] <- 1 # matrix of 1s for cells with y data
   N <- apply(x1, 2, sum, na.rm = na.rm) # total time slices for each cell ("2" means by column)
-  x <- t * x1 # matrix with NaN for squares and times without data
+  x <- (t - 1) * delta_t * x1 # t * x1 if sampled every year = matrix with NaN for squares and times without data
   #sumx <- apply(x, 2, sum, na.rm = na.rm)
   meanx <- apply(x, 2, sum, na.rm = na.rm)/N # mean time for each cell
   # subtract meanx values (from vector) from each value in the corresponding column
@@ -34,6 +33,6 @@ calcslope <- function(rx, time_units = 10, na.rm = TRUE) {
   numerator <- apply(xy, 2, sum, na.rm = na.rm) 
   deltax2 <- deltax^2 # matrix of values for denominator
   denom <- apply(deltax2, 2, sum, na.rm = na.rm)
-  slope <- (numerator/denom)*time_units
-  data.frame(slope = slope, N = N, time_scale = time_units, coord, icell)
+  slope <- numerator/denom
+  data.frame(slope = slope, N = N, delta_t = delta_t, coord, icell)
 }

### 2019-05-17 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
misc changes to tmb-sensor-explore

@@
 # replace obviously faulty depth data with NA
-all_sensor$depth_m[all_sensor$depth_m < 10] <- NA
+#all_sensor$depth_m[all_sensor$depth_m < 10] <- NA # 5,139
 
 all_depth <- all_sensor %>%
+  #filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
   # filter missing location data and trial year
@@
 d_trawl <- all_depth$data %>% 
+  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
+  filter(depth_max>10) %>% 
   # filter(ssid != 1) %>%
@@
 
-# remove obvious sensor fail, d_trawl %>% filter(ssid==4, temperature_c >12), fishing event id = 481861
-d_trawl <- d_trawl[d_trawl$fishing_event_id != 481861, ]
 d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
@@
   DOY_scaled = arm::rescale(lubridate::yday(date)),
+  DOY_scaled2 = arm::rescale(lubridate::yday(date))^2,
   #depth_scaled = arm::rescale(log(depth_m)),
@@
   depth_scaled3 = depth_scaled^3,
-  shallow = ifelse(depth>35,0,1)
+  shallow = ifelse(depth>35,0,1),
+    deep = ifelse(depth<35,0,1)
 )
@@
 ```{r}
-ggplot(dplyr::filter(d_trawl, depth <35), aes(meanSST,temperature_c, colour = -depth)) + geom_point()
+ggplot(dplyr::filter(d_trawl1, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
 ```
@@
 ```{r}
+ggplot(d_trawl, aes(log(depth),temperature_c, colour = depth_max)) + geom_point() + facet_wrap(~ssid)
+```
+
+```{r}
 ggplot(d_trawl, aes(DOY, temperature_c, colour = -depth, alpha=0.5)) + geom_point() + facet_wrap(~year) 
@@
 ```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 650)
+d_trawl1 <- d_trawl %>% filter(ssid == 3)
+spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 150)
 plot_spde(spde)
+sum(d_trawl1$shallow)
 ```
@@
 ```{r}
-m_temp3 <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year) + shallow*poly(DOY_scaled,2), 
-  time_varying = ~ 0 + depth_scaled + depth_scaled2,
-  ar1_fields = FALSE, include_spatial = TRUE,
+d_trawl1$year_scaled <- arm::rescale(d_trawl1$year)
+d_trawl1$x_scaled <- arm::rescale(d_trawl1$X)
+d_trawl1$y_scaled <- arm::rescale(d_trawl1$Y)
+
+m_temp_SST <- sdmTMB(d_trawl1,
+  temperature_c ~ year_scaled + poly(SST_scaled,3)*depth_scaled , #+ SST_scaled*x_scaled + SST_scaled*y_scaled
+  #time_varying = ~ 0 + depth_scaled2,
+  ar1_fields = FALSE, include_spatial = FALSE,
   time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE
+  silent = FALSE
 )
 
-m_temp3b <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year), 
-  time_varying = ~ 0 + depth_scaled + depth_scaled2,
-  ar1_fields = FALSE, include_spatial = TRUE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE
-)
+p_SST <- predict(m_temp_SST)
 
-m_temp3
-m_temp3b
+p_SST$residuals <- residuals(m_temp_SST)
+
+ggplot(p_SST, aes(depth_scaled, residuals)) +
+  geom_point(alpha=0.4) +
+  ylim(-8,8) +
+  geom_smooth()
+
+```
+
+```{r}
+library(gbm)
+
+m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl1, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
+
+m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl1, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
+
+plot(m,i.var=2)
+plot(m,i.var=3)
+plot(m,i.var=4)
+plot(m,i.var=5)
+
+plot(m2,i.var=1)
+plot(m2,i.var=2)
+plot(m,i.var=3)
+plot(m,i.var=4)
+
+plot(m2,i.var=1:2)
+plot(m2,i.var=c(1,3))
+plot(m2,i.var=c(1,4))
+d_trawl$r <- predict(m, n.trees = 2000) - d_trawl1$temperature_c
+
+ggplot(d_trawl, aes(depth_scaled, r)) +
+  geom_point(alpha=0.4) +
+  ylim(-8,8) +
+  geom_smooth()
 ```
@@
 
-predictions3 <- predict(m_temp3)
-predictions3b <- predict(m_temp3b)
+p_SST <- predict(m_temp_SST)
+p <- predict(m_temp)
 #predictions3b <- predict(m_temp3_nosst)
 
-predictions3$data$residuals <- residuals(m_temp3)
-predictions3b$data$residuals <- residuals(m_temp3b)
+p_SST$residuals <- residuals(m_temp_SST)
+p$residuals <- residuals(m_temp)
 
-nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth)
-
-ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
-  geom_point() + geom_abline(intercept = 0, slope = 1)
+# nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth)
+# 
+# ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
+#   geom_point() + geom_abline(intercept = 0, slope = 1)
 ```
@@
 ```{r}
-ggplot(predictions3$data, aes(est, residuals)) +
+ggplot(p_SST, aes(est, residuals)) +
   geom_point() +
@@
   facet_wrap(~year)
-ggplot(predictions3b$data, aes(est, residuals)) +
+ggplot(p, aes(est, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+ggplot(p_SST, aes(depth_scaled, residuals)) +
   geom_point() +
@@
   facet_wrap(~year)
-ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
+ggplot(p, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+p_SST$depth_mean <- exp(p_SST$depth_scaled*0.66+5)
+glimpse(d_trawl)
+ggplot(d_trawl, aes(depth_scaled, r)) +
+  geom_point(alpha=0.4) +
+  ylim(-8,8) +
+  geom_smooth()
+```
+
+```{r}
+p1 <- ggplot(p_SST, aes(depth_scaled, residuals)) +
   geom_point() +
@@
   geom_smooth()
-ggplot(predictions3b$data, aes(depth_scaled, residuals)) +
+p2 <- ggplot(p, aes(depth_scaled, residuals)) +
   geom_point() +
@@
   geom_smooth()
+gridExtra::grid.arrange(p1, p2, nrow = 2)
 ```

### 2019-05-17 analysis/tmb-sensor-explore/1-get-sensor-dat.R

@@ -1,9 +1,9 @@
 # gfplot::get_sensor_attributes()
 
-
+# library(gfdata)
 # get and save sensor data from different surveys
 
-d_trawl <- gfplot::get_sensor_data_trawl(ssid = c(1, 3, 4, 16), spread_attributes = FALSE)
+d_trawl <- gfdata::get_sensor_data_trawl(ssid = c(1, 3, 4, 16), spread_attributes = FALSE)
 saveRDS(d_trawl, file = "analysis/tmb-sensor-explore/dat-sensor-trawl.rds")
 
 # d_ll <- gfplot::get_sensor_data_ll_ctd(c(22, 36), sensor_min_max = TRUE)
@@ -12,6 +12,7 @@ saveRDS(d_trawl, file = "analysis/tmb-sensor-explore/dat-sensor-trawl.rds")
 
 
 # get and save data for different species
+# pcod <- gfdata::get_survey_sets(join_sample_ids = TRUE, species = "pacific cod", ssid = c(1))
 
 pcod <- gfplot::get_survey_sets(join_sample_ids = TRUE, species = "pacific cod", ssid = c(1, 3, 4, 16))
 saveRDS(pcod, file = "analysis/tmb-sensor-explore/pacific-cod.rds")

### 2019-05-17 analysis/VOCC/dist-vocc.Rmd
test alternate data option

@@
 data <- all_depth$data %>%
-  filter(depth_max>10) %>% # removes data if sensor stayed too near surface
+  filter(depth_m>10) %>% # removes data if sensor stayed too near surface
   # filter(ssid != 1) %>%
@@
 # )
+```
+
+Or use old data 
+```{r eval=FALSE}
+data <- readRDS("../tmb-sensor-explore/data/sensor-data-processed.rds")
+data <- data %>% mutate(depth = depth_m) %>% 
+  filter(depth_m >10) %>% # removes data if sensor stayed too near surface
+  gfplot:::scale_survey_predictors()
+
+data <- data[data$fishing_event_id != 481861, ]
 ```
@@
 nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
-
 ```
@@
 # # choose the spatial range to model
-dat <- data %>% dplyr::filter(year > 2003) #%>% filter(ssid == 4)
-nd <- nd # %>% filter(ssid == 4)
+dat <- data %>% dplyr::filter(year > 2003) %>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
+nd <- nd #%>% filter(ssid == 4)
 
@@
 # 
+# saveRDS(m_temp, file = "data/m_temp_post2003.rds")
 saveRDS(m_temp, file = "data/m_temp_allyears.rds")
@@
 ```{r}
-m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
+m_temp <- readRDS("../VOCC/data/m_temp_post2003.rds")
+# m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
 
@@
 # nd <- nd %>% dplyr::filter(year > 2003)
+nd <- nd # %>% filter(ssid == 4)
 

### 2019-05-17 analysis/VOCC/dist-vocc.Rmd
refining plot, calcslope, and analysis function

@@
 data <- all_depth$data %>%
-  filter(depth_m>10) %>% # removes data if sensor stayed too near surface
+  filter(depth_max>10) %>% # removes data if sensor stayed too near surface
   # filter(ssid != 1) %>%
@@
 # # choose the spatial range to model
-dat <- data %>% dplyr::filter(year > 2003) %>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
+dat <- data %>% dplyr::filter(year > 2003) #%>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
 nd <- nd #%>% filter(ssid == 4)
@@
 ```{r}
-m_temp <- readRDS("../VOCC/data/m_temp_post2003.rds")
-# m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
+# m_temp <- readRDS("../VOCC/data/m_temp_post2003.rds")
+ m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
 
@@
 ```
+
 ```{r}
@@
 ```{r}
-
 make_vector_data <- function(climate_data, 
@@
   ) {
-#browser()
+
 if (!is.null(ssid)) climate_data <- climate_data[climate_data$ssid %in% ssid, ]
@@
 
-start_data #
-end_data  #
-
 out <- dist_based_vocc(
@@
   thresholds = thresholds,
-  cell_size =  input_cell_size*scal_fac,
+  cell_size =  input_cell_size*scale_fac,
   delta_t = delta_t_total,
@@
 # FIXME: need to change function to deal with different time steps within a brick
-slopedat <-  calcslope(rbrick, delta_t = delta_t_step) # vocc::calcslope for comparison
+slopedat <-  calcslope(rbrick, delta_t_step = delta_t_step) # vocc::calcslope for comparison
 out <- left_join(out, slopedat, by = c("x", "y")) %>% select(-icell)
 out$C_per_decade <- out$slope * 10
-out$km_per_decade <- (out$distance/delta_t)*10 
+out$km_per_decade <- (out$distance/delta_t_total)*10 
 out
@@
 
-make_vector_data (predictions, 
+```
+
+```{r}
+out <- make_vector_data (predictions, 
   ssid = c(1, 3),
@@
   )
+glimpse(out)
 ```
+
+```{r}
+gvocc <- plot_vocc(out,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey97", 
+  vec_aes = "distance",
+  vec_lwd_range = c(0.5,0.6),
+  max_vec_plotted = 100,
+  fill_col = "var_1_s",
+  fill_label = "Mean\ntemperature\n2005 & 2007",
+  raster_alpha = 1,
+  vec_alpha = 0.8
+)
+gvocc
+```
+
+```{r}
+gtrend <- plot_vocc(out,
+  fill_col = "C_per_decade",
+  fill_label = "Temperature\ntrend (°C/decade)\nfor 2005-2017",
+  raster_alpha = 1,
+  vec_aes = NULL
+)
+gtrend
+```
+
+```{r}
+# png(file = "accasp-wg-fig.png",   # The directory you want to save the file in
+# res = 600,
+# units = 'in',
+# width = 10, # The width of the plot in inches
+# height = 7) # The height of the plot in inches
+
+gridExtra::grid.arrange(gtrend, gvocc, nrow = 1)
+#dev.off()
+```

### 2019-05-17 R/calcslope.R

@@ -5,7 +5,7 @@
 #' @param na.rm Logical for if NAs are removed.
 #'
 #' @export
-calcslope <- function(rx, delta_t = 2, na.rm = TRUE) {
+calcslope <- function(rx, delta_t_step = 2, na.rm = TRUE) {
   
   # this function is a modified version of Chris Brown's (https://github.com/cbrown5/vocc/blob/master/R/calcslope.R)
   # gives same result in intial tests, but may be slower
@@ -20,7 +20,7 @@ calcslope <- function(rx, delta_t = 2, na.rm = TRUE) {
   x1 <- y
   x1[!is.na(x1)] <- 1 # matrix of 1s for cells with y data
   N <- apply(x1, 2, sum, na.rm = na.rm) # total time slices for each cell ("2" means by column)
-  x <- (t - 1) * delta_t * x1 # t * x1 if sampled every year = matrix with NaN for squares and times without data
+  x <- (t - 1) * delta_t_step * x1 # t * x1 if sampled every year = matrix with NaN for squares and times without data
   #sumx <- apply(x, 2, sum, na.rm = na.rm)
   meanx <- apply(x, 2, sum, na.rm = na.rm)/N # mean time for each cell
   # subtract meanx values (from vector) from each value in the corresponding column
@@ -34,5 +34,5 @@ calcslope <- function(rx, delta_t = 2, na.rm = TRUE) {
   deltax2 <- deltax^2 # matrix of values for denominator
   denom <- apply(deltax2, 2, sum, na.rm = na.rm)
   slope <- numerator/denom
-  data.frame(slope = slope, N = N, delta_t = delta_t, coord, icell)
+  data.frame(slope = slope, N = N, time_step = delta_t_step, coord, icell)
 }

### 2019-05-17 R/plot.R

@@ -30,7 +30,7 @@ plot_vocc <- function(df,
                       arrowhead_size = 0.015,
                       vec_lwd_range = c(0.7, 0.8),
                       vec_alpha = 1,
-                      max_vec_plotted = 100,
+                      max_vec_plotted = max(df$distance),
                       low_fill = "Steel Blue 4",
                       mid_fill = "white",
                       high_fill = "Red 3",
@@ -40,23 +40,20 @@ plot_vocc <- function(df,
                       coast = NULL,
                       contours = NULL) {
   df <- df[order(-df$distance), ] # order so smaller vectors are on top?
-  # Add arrows indicating target cells
-  if (!is.null(max_vec_plotted)) {
-    if (max(df$distance) > max_vec_plotted) {
+  
+  if (max(df$distance) > max_vec_plotted) {
       df[df$distance > max_vec_plotted, ]$target_X <- NA
       df[df$distance > max_vec_plotted, ]$target_Y <- NA
-    }
-    # browser()
+  }
 
-
-    gvocc <- ggplot2::ggplot(df, aes(x, y)) +
+  gvocc <- ggplot2::ggplot(df, aes(x, y)) +
       xlab("UTM") + ylab("UTM") +
       coord_fixed(xlim = range(df$x) + c(-3, 3), ylim = range(df$y) + c(-3, 3)) +
       gfplot::theme_pbs()
 
-
-    if (!is.null(fill_col)) {
-      fill <- df[[fill_col]]
+  #### Add fill #### 
+  if (!is.null(fill_col)) {
+   fill <- df[[fill_col]]
 
       if (min(fill, na.rm = TRUE) > 0) {
         gvocc <- gvocc +
@@ -73,56 +70,21 @@ plot_vocc <- function(df,
       }
     }
 
-    # Add bathymetry
+    #### Add bathymetry ####
     if (!is.null(contours)) {
-      # add premade bathymetry layer (must be in utms)
-      isobath <- contours
-
+      # add premade contour layers (will all be same colour and must be in utms)
       gvocc <- gvocc +
         geom_path(
-          data = isobath[isobath$PID == 100, ],
-          aes_string(
-            x = "X", y = "Y",
-            group = "paste(PID, SID)"
-          ),
-          inherit.aes = FALSE, lwd = 0.4, alpha = 0.4, colour = "grey85"
-        ) +
-        geom_path(
-          data = isobath[isobath$PID == 200, ],
-          aes_string(
-            x = "X", y = "Y",
[…]

### 2019-05-17 analysis/VOCC/dist-vocc.Rmd
example analysis code using make_vector_data and plot_vocc functions

@@
 nd <- rbind(nd_1, nd_3, nd_4, nd_16)
-# nd <- nd_4 # for faster test option
 nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
@@
 ```{r}
-# m_temp <- readRDS("../VOCC/data/m_temp_post2003.rds")
 m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
 
-# can also be filtered to include only year with multiple surveys 
+# can also be filtered to exclude certain years
 # m_temp <- readRDS("../VOCC/data/m_temp_allpost2003.rds")
 # nd <- nd %>% dplyr::filter(year > 2003)
-nd <- nd # %>% filter(ssid == 4)
 
@@
 
-```{r}
-# choose the spatial range to build raster on
+Explore subsets ot the spatial range 
+```{r eval=FALSE}
 predicted1 <- predictions %>% filter(ssid == 1) 
@@
 ```{r}
-
-# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
-rbrick <- make_raster_brick(predicted3, scale_fac = 2)
-saveRDS(rbrick, file = "analysis/data/rbrick-temp-hs.rds")
-
-# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
-rbrick <- make_raster_brick(predicted1, scale_fac = 2)
-saveRDS(rbrick, file = "analysis/data/rbrick-temp-qcs.rds")
-
-# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
-rbrick <- make_raster_brick(predicted4, scale_fac = 2)
-saveRDS(rbrick, file = "analysis/data/rbrick-temp-wcvi.rds")
-
-# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
-rbrick <- make_raster_brick(predicted16, scale_fac = 2)
-saveRDS(rbrick, file = "analysis/data/rbrick-temp-wchg.rds")
-
-rbrick <- make_raster_brick(predicted1n3, scale_fac = 2)
-saveRDS(rbrick, file = "analysis/data/rbrick-temp-hs-qcs.rds")
-
-```
-
-```{r}
-make_vector_data <- function(climate_data, 
-  ssid = NULL,
-  start_year = NULL,
-  input_cell_size = 2, 
-  scale_fac = 2,
-  delta_t_total = 10,
-  delta_t_step = 2,
-  time_slices = 7, 
-  indices = c(1, 1, 2, 2, 2, 3, 3),
-  thresholds = c(0.75)
-  ) {
-
-if (!is.null(ssid)) climate_data <- climate_data[climate_data$ssid %in% ssid, ]
-if (!is.null(start_year)) climate_data <- climate_data %>% dplyr::filter(year >= start_year)
-
-#if (!identical(time_slices, length(indices)))
-length_indices <- length(indices)
-if(!isTRUE(time_slices==length_indices))
-      stop("Must have an indice assigned to each time slice,",
-      "therefore length('indices') must equal 'time_slices'.",
-      call. = FALSE)
-
-#FIXME: need way of determining how many variables(in separate dataframes) are in dataset  
-#if (!is.list(climate_data)) {
-# if (length(thresholds) > 1)
-#       stop("If multiple climate variables (and corresponding thresholds),", 
-#         "data must be in list form.", call. = FALSE)
-
-rbrick <- make_raster_brick(climate_data, scale_fac = scale_fac)
-mnraster_brick <- raster::stackApply(rbrick, indices = indices, fun = mean)
-start_raster <- mnraster_brick[[1]]
-end_raster <- mnraster_brick[[3]]
-
-# make sparate named lists containing climate rasters or dataframes
-# data with just one climate variable
-start_data <- list(var_1 = start_raster)
-end_data <- list(var_1 = end_raster)
-
-# } else {
-#     if (!identical(length(climate_data), length(thresholds)))
-#       stop("Must have a 'thresholds' value for each climate variable", call. = FALSE)
-# rbrick <- list()
-# for (i in seq_len(climate_data)) {
-#   
-# rbrick[[i]] <- make_raster_brick(climate_data[[i]], scale_fac = scale_fac)
-# mnraster_brick <- raster::stackApply(rbrick[[i]], indices = indices, fun = mean)
-# 
-# start_raster <- mnraster_brick[[1]]
-# end_raster <- mnraster_brick[[3]]
-# var <- list(start_raster = start_raster, end_raster = end_raster)
-# }
-
-# start_data <- list(var = var$start_raster[[i]])
-# end_data <- list(var = var$end_raster[[i]])
-#}
-
-out <- dist_based_vocc(
-  start_data = start_data,
-  end_data = end_data,
-  x = "x",
-  y = "y",
-  variable_names = c("index_1"), # what the layer within each element is called
-  thresholds = thresholds,
-  cell_size =  input_cell_size*scale_fac,
-  delta_t = delta_t_total,
-  raster = TRUE
-  )
-
-# FIXME: need to change function to deal with different time steps within a brick
-slopedat <-  calcslope(rbrick, delta_t_step = delta_t_step) # vocc::calcslope for comparison
-out <- left_join(out, slopedat, by = c("x", "y")) %>% select(-icell)
-out$C_per_decade <- out$slope * 10
-out$km_per_decade <- (out$distance/delta_t_total)*10 
-out
-}
-
-```
-
-```{r}
 out <- make_vector_data (predictions, 
-  ssid = c(1, 3),
+  ssid = c(1),
   start_year = 2005,
@@
   delta_t_step = 2,
-  time_slices = 7,
-  indices = c(1, 1, 2, 2, 2, 3, 3),
-  thresholds = c(0.75)
+  time_steps = 7,
+  indices = c(1, 1, 1, 2, 3, 3, 3),
+  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
   )

### 2019-05-17 analysis/VOCC/dist-vocc.Rmd
style files

@@
 data <- all_depth$data %>%
-  filter(depth_max>10) %>% # removes data if sensor stayed too near surface
+  filter(depth_max > 10) %>% # removes data if sensor stayed too near surface
   # filter(ssid != 1) %>%
@@
 data <- readRDS("../tmb-sensor-explore/data/sensor-data-processed.rds")
-data <- data %>% mutate(depth = depth_m) %>% 
-  filter(depth_m >10) %>% # removes data if sensor stayed too near surface
+data <- data %>%
+  mutate(depth = depth_m) %>%
+  filter(depth_m > 10) %>% # removes data if sensor stayed too near surface
   gfplot:::scale_survey_predictors()
@@
 # # choose the spatial range to model
-dat <- data %>% dplyr::filter(year > 2003) #%>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
-nd <- nd #%>% filter(ssid == 4)
+dat <- data %>% dplyr::filter(year > 2003) # %>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
+nd <- nd # %>% filter(ssid == 4)
 
@@
 
-p <- plot_map(predictions , "est") +
+p <- plot_map(predictions, "est") +
   scale_fill_viridis_c(trans = "sqrt", option = "C") +
@@
 
-p <- plot_map(predicted1n3 , "est") +
+p <- plot_map(predicted1n3, "est") +
   scale_fill_viridis_c(trans = "sqrt", option = "C") +
@@
 ```{r}
-out <- make_vector_data (predictions, 
+out <- make_vector_data(predictions,
   ssid = c(1),
@@
   vec_aes = "distance",
-  vec_lwd_range = c(0.5,0.6),
+  vec_lwd_range = c(0.5, 0.6),
   max_vec_plotted = 100,
@@
 gridExtra::grid.arrange(gtrend, gvocc, nrow = 1)
-#dev.off()
+# dev.off()
 ```

### 2019-05-17 R/calcslope.R

@@ -1,38 +1,38 @@
 #' Calculate trend in varible for pixels in a raster brick
 #'
-#' @param rx Raster brick containing layers for multiple years. 
-#' @param delta_t Number of time steps between layers, if not 1 allows slope to be calculated for a single unit of time. 
+#' @param rx Raster brick containing layers for multiple years.
+#' @param delta_t Number of time steps between layers, if not 1 allows slope to be calculated for a single unit of time.
 #' @param na.rm Logical for if NAs are removed.
 #'
 #' @export
 calcslope <- function(rx, delta_t_step = 2, na.rm = TRUE) {
-  
+
   # this function is a modified version of Chris Brown's (https://github.com/cbrown5/vocc/blob/master/R/calcslope.R)
   # gives same result in intial tests, but may be slower
   # use vocc::calcslope from devtools::install_github("seananderson/vocc") to apply his method instead
-  
+
   icell <- seq(1, raster::ncell(rx))
   coord <- raster::xyFromCell(rx, icell)
   y <- t(raster::getValues(rx))
   t <- row(y) # matrix of times for all cells ... assumes equal time between slices
-  #FIXME: Can we make this robust to uneven timesteps?
+  # FIXME: Can we make this robust to uneven timesteps?
   # browser()
   x1 <- y
   x1[!is.na(x1)] <- 1 # matrix of 1s for cells with y data
   N <- apply(x1, 2, sum, na.rm = na.rm) # total time slices for each cell ("2" means by column)
   x <- (t - 1) * delta_t_step * x1 # t * x1 if sampled every year = matrix with NaN for squares and times without data
-  #sumx <- apply(x, 2, sum, na.rm = na.rm)
-  meanx <- apply(x, 2, sum, na.rm = na.rm)/N # mean time for each cell
+  # sumx <- apply(x, 2, sum, na.rm = na.rm)
+  meanx <- apply(x, 2, sum, na.rm = na.rm) / N # mean time for each cell
   # subtract meanx values (from vector) from each value in the corresponding column
-  deltax <- sweep(x, 2, meanx, `-`) 
-  #sumy <- apply(y, 2, sum, na.rm = na.rm)
-  meany <- apply(y, 2, sum, na.rm = na.rm)/N # mean y for each cell
+  deltax <- sweep(x, 2, meanx, `-`)
+  # sumy <- apply(y, 2, sum, na.rm = na.rm)
[…]

### 2019-05-17 R/make-vector-data.R

@@ -1,75 +1,75 @@
 #' Make VOCC vector data from sdmTMB output predict funciton
 #'
-#' @param climate_data List of dataframes each containing output from predict. 
+#' @param climate_data List of dataframes each containing output from predict.
 #' @param ssid Vector of ssid to be included.
 #' @param start_year Starting year for VOCC calculation.
 #' @param input_cell_size Cell size of predict grid.
-#' @param scale_fac Factor by which cells are aggregated for VOCC. 
-#' @param delta_t_total Mean time between start and end climate values. 
-#' @param delta_t_step Time between each time step in same units as above. 
-#' @param time_steps Number of time steps. 
+#' @param scale_fac Factor by which cells are aggregated for VOCC.
+#' @param delta_t_total Mean time between start and end climate values.
+#' @param delta_t_step Time between each time step in same units as above.
+#' @param time_steps Number of time steps.
 #' @param indices Vector of length equal to number of time steps,
-#'    where 1 = starting time step(s), 2 = middle time step(s) not included in VOCC calc, 3 = end time step(s). 
-#' @param thresholds 
+#'    where 1 = starting time step(s), 2 = middle time step(s) not included in VOCC calc, 3 = end time step(s).
+#' @param thresholds
 #'
 #' @export
-make_vector_data <- function(climate_data, 
-  ssid = NULL,
-  start_year = NULL,
-  input_cell_size = 2, 
-  scale_fac = 2,
-  time_var = "year",
-  delta_t_total = 10,
-  delta_t_step = 2,
-  time_steps = 7, 
-  indices = c(1, 1, 2, 2, 2, 3, 3),
-  thresholds = c(0.75)
-) {
-  
+make_vector_data <- function(climate_data,
+                             ssid = NULL,
+                             start_year = NULL,
+                             input_cell_size = 2,
+                             scale_fac = 2,
+                             time_var = "year",
+                             delta_t_total = 10,
+                             delta_t_step = 2,
+                             time_steps = 7,
+                             indices = c(1, 1, 2, 2, 2, 3, 3),
+                             thresholds = c(0.75)) {
   if (!is.null(ssid)) climate_data <- climate_data[climate_data$ssid %in% ssid, ]
   if (!is.null(start_year)) climate_data <- climate_data %>% dplyr::filter(year >= start_year)
-  
[…]

### 2019-05-17 R/plot.R

@@ -40,154 +40,154 @@ plot_vocc <- function(df,
                       coast = NULL,
                       contours = NULL) {
   df <- df[order(-df$distance), ] # order so smaller vectors are on top?
-  
+
   if (max(df$distance) > max_vec_plotted) {
-      df[df$distance > max_vec_plotted, ]$target_X <- NA
-      df[df$distance > max_vec_plotted, ]$target_Y <- NA
+    df[df$distance > max_vec_plotted, ]$target_X <- NA
+    df[df$distance > max_vec_plotted, ]$target_Y <- NA
   }
 
   gvocc <- ggplot2::ggplot(df, aes(x, y)) +
-      xlab("UTM") + ylab("UTM") +
-      coord_fixed(xlim = range(df$x) + c(-3, 3), ylim = range(df$y) + c(-3, 3)) +
-      gfplot::theme_pbs()
+    xlab("UTM") + ylab("UTM") +
+    coord_fixed(xlim = range(df$x) + c(-3, 3), ylim = range(df$y) + c(-3, 3)) +
+    gfplot::theme_pbs()
 
-  #### Add fill #### 
+  #### Add fill ####
   if (!is.null(fill_col)) {
-   fill <- df[[fill_col]]
+    fill <- df[[fill_col]]
 
-      if (min(fill, na.rm = TRUE) > 0) {
-        gvocc <- gvocc +
-          geom_raster(aes(fill = fill), alpha = raster_alpha) +
-          scale_fill_viridis_c(trans = "sqrt") +
-          labs(fill = fill_label) +
-          theme(legend.position = c(0.17, 0.17))
-      } else {
-        gvocc <- gvocc +
-          geom_raster(aes(fill = fill), alpha = raster_alpha) +
-          scale_fill_gradient2(low = low_fill, mid = mid_fill, high = high_fill) +
-          labs(fill = fill_label) +
-          theme(legend.position = c(0.17, 0.17))
-      }
+    if (min(fill, na.rm = TRUE) > 0) {
+      gvocc <- gvocc +
+        geom_raster(aes(fill = fill), alpha = raster_alpha) +
+        scale_fill_viridis_c(trans = "sqrt") +
+        labs(fill = fill_label) +
+        theme(legend.position = c(0.17, 0.17))
+    } else {
+      gvocc <- gvocc +
+        geom_raster(aes(fill = fill), alpha = raster_alpha) +
[…]

### 2019-06-06 analysis/VOCC/dist-vocc.Rmd
exploratory biomass change plotting

@@
 library(gfranges)
+
+# Basic map function
+plot_map <- function(dat, column = "est") {
+  ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    facet_wrap(~year) +
+    coord_fixed() + theme(axis.title=element_blank())
+}
 ```
@@
 dat <- data %>% dplyr::filter(year > 2003) # %>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
-nd <- nd # %>% filter(ssid == 4)
+#nd <- nd_1 # %>% filter(ssid == 4)
 
@@
 
-predictions <- predict(m_temp, newdata = nd)
+temp <- predict(m_temp, newdata = nd)
+temperature <- temp %>% 
+  # filter(ssid!=3) %>% 
+  filter(ssid!=4) %>% 
+  filter(ssid!=16) %>% 
+  filter(year>2004)
 
-plot_map <- function(dat, column = "est") {
-  ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    facet_wrap(~year) +
-    coord_fixed()
-}
+glimpse(temperature)
 
-p <- plot_map(predictions, "est") +
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
-  ggtitle("Prediction (fixed effects + all random effects)")
-print(p)
+temp <- plot_map(temperature, "est") +
+  scale_fill_viridis_c(trans = "sqrt", option = "D") +
+  ggtitle("Bottom temperature") + theme(legend.position = "none") #c(0.9, 0), legend.justification = c(0.9, 0), legend.title = element_blank() ) 
+print(temp)
 ```
@@
 ```{r}
-out <- make_vector_data(predictions,
-  ssid = c(1),
-  start_year = 2005,
+out <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2005,
+  skip_time = 2011,
   input_cell_size = 2,
-  scale_fac = 2,
+  scale_fac = 1,
   delta_t_total = 10,
   delta_t_step = 2,
-  time_steps = 7,
-  indices = c(1, 1, 1, 2, 3, 3, 3),
+  indices = c(1, 1, 1, 2, 2, 2),
   thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
 )
-glimpse(out)
+# glimpse(out)
 ```
@@
   mid_col = "white",
-  high_col = "grey97",
+  high_col = "grey87",
   vec_aes = "distance",
-  vec_lwd_range = c(0.5, 0.6),
+  vec_lwd_range = c(0.2, 0.5),
   max_vec_plotted = 100,
-  fill_col = "var_1_s",
-  fill_label = "Mean\ntemperature\n2005 & 2007",
+  fill_col = "var_1_e",
+  fill_label = "Mean\ntemperature\n2013-2017",
   raster_alpha = 1,
-  vec_alpha = 0.8
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D"
 )
+gvocc <- gvocc + ggtitle("VOCC vectors for <0.5°C change")
 gvocc
@@
 gtrend <- plot_vocc(out,
-  fill_col = "C_per_decade",
-  fill_label = "Temperature\ntrend (°C/decade)\nfor 2005-2017",
+  fill_col = "units_per_decade",
+  fill_label = "°C per decade",
   raster_alpha = 1,
@@
 )
+gtrend <- gtrend + ggtitle("Bottom temperature trend 2005-2017")
 gtrend
@@
 ```
+
+Step through pairs of years?
+
+```{r}
+
+step1 <- make_vector_data(temperature,
+  ssid = c(1, 3),
+  start_time = 2007,
+  end_time = 2017,
+  skip_time = c(2011,2013),
+  input_cell_size = 2,
+  scale_fac = 2,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1,1,2,2), 
+  thresholds = c(.75) # vector of plus/minus threshold(s) to define climate match.
+)
+
+gvocc <- plot_vocc(step1,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey97",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.5, 0.6),
+  max_vec_plotted = 100,
+  fill_col = "var_1_e",
+  fill_label = "Mean temperature\n2015 & 2017",
+  raster_alpha = 1,
+  vec_alpha = 0.8
+)
+gvocc
+```
+
+# Biomass data
+
+```{r}
+biomass <- readRDS("data/maturity-biomass-data-pop-1n3.rds")
+#doy <- data %>% select(fishing_event_id, DOY)
+#biomass <- left_join(biomass, doy)
+```
+
+```{r}
+p <- ggplot(biomass, aes(X, Y, size = density, colour = density)) + 
+  geom_point(alpha=0.5) + scale_colour_viridis_c(trans = "sqrt", option = "C", guide = "legend") + facet_wrap(~year)
+print(p)
+
+p <- ggplot(biomass, aes(X, Y, size = adult_density, colour = adult_density)) + 
+  geom_point(alpha=0.5) + scale_colour_viridis_c(trans = "sqrt", option = "C", guide = "legend") + facet_wrap(~year)
+print(p)
+
+p <- ggplot(biomass, aes(X, Y, size = imm_density, colour = imm_density))  +
+  geom_point(alpha=0.5) + scale_colour_viridis_c(trans = "sqrt", option = "C", guide = "legend") + facet_wrap(~year)
+print(p)
+```
+
+Run sdmTMB model (if not done previously).
+```{r eval=FALSE}
+# # choose the spatial range to model
+dat <- biomass %>% gfplot:::interp_survey_bathymetry() 
+dat <- dat$data %>% gfplot:::scale_survey_predictors()
+#nd <- nd %>% filter(ssid != 16) %>% filter(ssid != 4)
+
+spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 500)
+sdmTMB::plot_spde(spde)
+
+adult_biomass <- sdmTMB::sdmTMB(dat,
+  adult_density ~ 0 + as.factor(year),
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  time = "year", spde = spde,
+  family = tweedie(link = "log"),
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  silent = FALSE
+)
+
+imm_biomass <- sdmTMB::sdmTMB(dat,
+  imm_density ~ 0 + as.factor(year),
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  time = "year", spde = spde,
+  family = tweedie(link = "log"),
+  ar1_fields = FALSE, 
+  include_spatial = TRUE,
+  silent = FALSE
+)
+
+saveRDS(adult_biomass, file = "data/m_adult_biomass.rds")
+saveRDS(imm_biomass, file = "data/m_imm_biomass.rds")
+```
+
+Calculate predicted values
+```{r}
+adult_biomass <- readRDS("data/m_adult_biomass.rds")
+imm_biomass <- readRDS("data/m_imm_biomass.rds")
+
+nd <- nd %>% mutate(year = as.integer(year)) %>% 
+  # filter(ssid!=3) %>% 
+  filter(ssid!=4) %>% 
+  filter(ssid!=16) %>%
+  dplyr::filter(year > 2003)
+
+adult_predictions <- predict(adult_biomass, newdata = nd)
+imm_predictions <- predict(imm_biomass, newdata = nd)
+```
+
+```{r}
+predictions <- adult_predictions
+predictions$est_exp <- exp(predictions$est)
+# sqrt_trans <- scales::trans_new(
+#   name = "logp",
+#   trans = function(x) sqrt(x),
+#   inverse = function(x) ((x))^2,
+#   breaks = scales::log_breaks()
+# )
+p_adult <- plot_map(predictions, "est_exp") + 
+  scale_fill_viridis_c(trans = "sqrt", option = "C") +
+  ggtitle("Pacific Ocean Perch mature biomass") + theme(legend.position = c(0.9, 0), legend.justification = c(0.9, 0), legend.title = element_blank() ) 
+print(p_adult)
+```
+
+```{r}
+adult_predictions$est_exp <- exp(adult_predictions$est)
+adult <- make_vector_data(adult_predictions,
+  ssid = c(1,3),
+  start_time = 2005,
+  skip_time = 2011,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1, 1, 1, 2, 2, 2),
+  variable_names = "est_exp",
+  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
+)
+#glimpse(adult )
+```
+
+```{r eval=FALSE}
+gvobiomass <- plot_vocc(adult,
+  # low_col = "yellow",
+  # mid_col = "grey",
+  # high_col = "black",
+  # vec_aes = "units_per_decade",
+  # vec_lwd_range = c(0.7,0.005 ),
+  # vec_alpha = 0.7,
+  # max_vec_plotted = 30,
+  fill_col = "var_1_e",
+  fill_label = "Mean mature \nbiomass (kg/m2)\n in 2013-2017",
+  raster_alpha = 1,
+  viridis_option = "C",
+  vec_aes = NULL
+)
+gvobiomass <- gvobiomass +
+  ggtitle("Pacific Ocean Perch")
+
+gvobiomass 
+```
+
+```{r}
+biotrend_ad <- plot_vocc(adult,
+  fill_col = "units_per_decade",
+  fill_label = "Mature \nbiomass trend \n(kg/m2/decade)\nfor 2005-2017",
+  raster_alpha = 1,
+  vec_aes = NULL
+)
+biotrend_ad
+```
+
+```{r}
+predictions <- imm_predictions
+predictions$est_exp <- exp(predictions$est)
+
+p <- plot_map(predictions, "est_exp") + 
+  scale_fill_viridis_c(trans = "sqrt", option = "C") +
+  ggtitle("Prediction (fixed effects + all random effects)")
+print(p)
+```
+
+```{r}
+imm_predictions$est_exp <- exp(imm_predictions$est)
+
+out_im <- make_vector_data(imm_predictions,
+  ssid = c(1,3),
+  start_time = 2005,
+  skip_time = 2011,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1, 1, 1, 2, 2, 2),
+  variable_names = "est_exp",
+  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
+)
+# glimpse(out_im)
+```
+
+```{r}
+biotrend_im <- plot_vocc(out_im,
+  fill_col = "units_per_decade",
+  fill_label = "Immature \nbiomass trend \n(kg/m2/decade)\nfor 2005-2017",
+  raster_alpha = 1,
+  vec_aes = NULL
+)
+biotrend_im
+```
+
+```{r}
+png(file = "exploratory-pop-figs.png",   # The directory you want to save the file in
+res = 600,
+units = 'in',
+width = 14, # The width of the plot in inches
+height = 12) # The height of the plot in inches
+
+gridExtra::grid.arrange(temp, gvocc, gtrend, 
+  # p_adult,
+    gvobiomass, 
+  biotrend_ad, biotrend_im, nrow = 2)
+dev.off()
+```

### 2019-06-08 analysis/VOCC/dist-vocc.Rmd
calculate age at maturity with random effect of year

@@
   start_time = 2005,
-  skip_time = 2011,
+  #skip_time = 2011,
   input_cell_size = 2,
@@
   delta_t_step = 2,
-  indices = c(1, 1, 1, 2, 2, 2),
+  indices = c(1, 1, 1, 2, 2, 2, 2),
   thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
@@
   axis_lables = FALSE,
-  viridis_option = "D"
+  viridis_option = "D",
+  transform_col = no_trans
 )
@@
   raster_alpha = 1,
-  vec_aes = NULL
+  vec_aes = NULL,
+  transform_col = no_trans
 )
@@
 ```{r}
-biomass <- readRDS("data/maturity-biomass-data-pop-1n3.rds")
+library(dplyr)
+library(ggplot2)
+library(gfplot)
+library(gfdata)
+
+events <- readRDS("analysis/VOCC/data/event-data-pop-1n3.rds")
+fish <- readRDS("analysis/VOCC/data/bio-data-pop-1n3.rds")
+
+biomass <- split_catch_maturity (events, fish, 
+  survey = c("SYN HS","SYN QCS"), 
+  years = NULL,
+  cutoff_quantile = c(.9995),
+  plot = TRUE
+)
+
+saveRDS(biomass, file = "analysis/VOCC/data/maturity-biomass-data-pop-1n3-re.rds")
+
+# disagreement between catch_weight and bio sample for single fishing event with 17 large POP. 
+# events[events$fishing_event_id==1506954,]$catch_weight
+# fish[fish$fishing_event_id==1506954,]$length
+# tidy_sets[tidy_sets$fishing_event_id==1506954,]$catch_weight
+
+```
+
+```{r}
+biomass <- readRDS("data/maturity-biomass-data-pop-1n3-re.rds")
 #doy <- data %>% select(fishing_event_id, DOY)
@@
 nd <- nd %>% mutate(year = as.integer(year)) %>%
-  # filter(ssid!=3) %>% 
+  # dplyr::filter(year > 2003) %>% 
+  filter(ssid!=3) %>% 
   filter(ssid!=4) %>% 
-  filter(ssid!=16) %>%
-  dplyr::filter(year > 2003)
+  filter(ssid!=16) 
+#unique(biomass$year)
 
@@
 predictions <- adult_predictions
-predictions$est_exp <- exp(predictions$est)
-# sqrt_trans <- scales::trans_new(
-#   name = "logp",
-#   trans = function(x) sqrt(x),
-#   inverse = function(x) ((x))^2,
-#   breaks = scales::log_breaks()
-# )
+predictions$est_exp <- exp(predictions$est)*10000
+
+breaks <- scales::trans_breaks(fourth_root_power[["transform"]], fourth_root_power[["inverse"]], n=5) 
+labels <- function(x) { format(x, digits=1) }
+
 p_adult <- plot_map(predictions, "est_exp") + 
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
+  scale_fill_viridis_c(trans = fourth_root_power, breaks=breaks, labels = labels, option = "C") +
   ggtitle("Pacific Ocean Perch mature biomass") + theme(legend.position = c(0.9, 0), legend.justification = c(0.9, 0), legend.title = element_blank() ) 
@@
 ```{r}
-adult_predictions$est_exp <- exp(adult_predictions$est)
+adult_predictions$est_exp <- exp(adult_predictions$est)*10000
 adult <- make_vector_data(adult_predictions,
@@
   delta_t_step = 2,
-  indices = c(1, 1, 1, 2, 2, 2),
+  indices = c(1, 1, 1, 2, 2, 2, 2),
   variable_names = "est_exp",
@@
   fill_col = "var_1_e",
-  fill_label = "Mean mature \nbiomass (kg/m2)\n in 2013-2017",
+  fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
   raster_alpha = 1,
   viridis_option = "C",
-  vec_aes = NULL
+  vec_aes = NULL,
+  transform_col = fourth_root_power
 )
@@
   fill_col = "units_per_decade",
-  fill_label = "Mature \nbiomass trend \n(kg/m2/decade)\nfor 2005-2017",
+  fill_label = "Mature \nbiomass trend \n(kg/ha/decade)\nfor 2005-2017",
   raster_alpha = 1,
-  vec_aes = NULL
+  vec_aes = NULL,
+  transform_col = fourth_root_power
 )
@@
 predictions <- imm_predictions
-predictions$est_exp <- exp(predictions$est)
+predictions$est_exp <- exp(predictions$est)*10000
 
 p <- plot_map(predictions, "est_exp") + 
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
+  scale_fill_viridis_c(trans = fourth_root_power, breaks=breaks, labels = labels, option = "C") +
   ggtitle("Prediction (fixed effects + all random effects)")
@@
 ```{r}
-imm_predictions$est_exp <- exp(imm_predictions$est)
+imm_predictions$est_exp <- exp(imm_predictions$est)*10000
 
@@
   fill_col = "units_per_decade",
-  fill_label = "Immature \nbiomass trend \n(kg/m2/decade)\nfor 2005-2017",
+  fill_label = "Immature \nbiomass trend \n(kg/ha/decade)\nfor 2005-2017",
   raster_alpha = 1,
-  vec_aes = NULL
+  vec_aes = NULL,
+  transform_col = fourth_root_power
 )
@@
 ```
+```{r}
+png(file = "POP-maturity-year-re.png",   # The directory you want to save the file in
+res = 600,
+units = 'in',
+width = 8, # The width of the plot in inches
+height = 10)
+
+plot_mat_ogive(m)
+dev.off()
+```
 
 ```{r}
-png(file = "exploratory-pop-figs.png",   # The directory you want to save the file in
+png(file = "exploratory-pop-figs-trans.png",   # The directory you want to save the file in
 res = 600,

### 2019-06-12 analysis/VOCC/biomass-trends.Rmd
biomass plots rmd

---
title: "Biomass trend maps"
author: "Philina English"
date: '2019-06-07'
output: html_document
---

```{r}
species <- "Quillback Rockfish"
species <- "Yelloweye Rockfish"
species <- "North Pacific Spiny Dogfish" 
species <- "Silvergray Rockfish"
species <- "Arrowtooth Flounder"
```

```{r}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
model_ssid <- c(1,3)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r, message=FALSE, warning=FALSE}
survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
fish <- readRDS(paste0("raw/bio-data-", spp, ""))
bath <- readRDS("data/bathymetry-data")
# for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
# fish <- fish[fish$fishing_event_id!=1506954,] 
```

Create prediction grids for each survey area so that only years with data are included
```{r}
# choose base year(s) to create grid from
dummy_year <- c(2005,2006)
dat <- bath$data %>% scale_survey_predictors()
# create grids for each ssid separately
ssid <- 1
survey_abbrev <- "SYN QCS"
nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_1$year)

ssid <- 3
survey_abbrev <- "SYN HS"
nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_3$year)

ssid <- 4
survey_abbrev <- "SYN WCVI"
nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_4$year)

ssid <- 16
survey_abbrev <- "SYN WCHG"
nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_16$year)

nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>%
  mutate(year = as.numeric(year)) 
#nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
```

Calculate maturity weighted biomass
```{r, eval=FALSE}
maturity <- split_catch_maturity (survey_sets, fish, bath,
  survey = c("SYN HS","SYN QCS"), 
  years = NULL,
  cutoff_quantile = c(.9995),
  plot = TRUE
)

saveRDS(maturity$data, file = paste0("data/data-by-maturity-", spp, ".rds"))
```

```{r, eval=FALSE}
png(file = paste0("figs/", spp, "-mass-model.png"),   # The directory you want to save the file in
res = 600,
units = 'in',
width = 8, # The width of the plot in inches
height = 10)

maturity$mass_model

dev.off()
print(maturity$mass_model)
```

```{r, eval=FALSE}
png(file = paste0("figs/", spp, "-maturity-year-re.png"),   # The directory you want to save the file in
res = 600,
units = 'in',
width = 8, # The width of the plot in inches
height = 10)

maturity$maturity

dev.off()
print(maturity$maturity)
```

Choose spatial zone
```{r}
biomass <- readRDS(paste0("data/data-by-maturity-", spp, ".rds"))
data <- biomass %>% gfranges:::scale_survey_predictors()

nd <- nd_all %>% filter(ssid %in% model_ssid) 
data <- data %>% filter(ssid %in% model_ssid)
```

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
sdmTMB::plot_spde(spde)

if (any(names(data)=="adult_density")) {
adult_biomass <- sdmTMB::sdmTMB(data,
  adult_density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  enable_priors = TRUE,
  # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
  silent = FALSE
)

imm_biomass <- sdmTMB::sdmTMB(data,
  imm_density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = FALSE, 
  include_spatial = TRUE,
  enable_priors = TRUE,
  silent = FALSE
)

saveRDS(adult_biomass, file = paste0("data/model-mature-biomass-", spp, "-1n3.rds"))

saveRDS(imm_biomass, file = paste0("data/model-imm-biomass-", spp, "-1n3.rds"))

} else {

total_biomass <- sdmTMB::sdmTMB(dat,
  density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  silent = FALSE
)

saveRDS(total_biomass, file = paste0("data/model-total-biomass-", spp, "-1n3.rds"))

}
```

Calculate predicted values for available models
#FIXME: tryCatch not working here, try this?
```{r, error=TRUE}
if (length(unique(fish$maturity_code))>2) {
    adult_biomass <- readRDS(paste0("data/model-mature-biomass-", spp, "-1n3.rds"))
    nd$year<- as.integer(nd$year)
    adult_predictions <- predict(adult_biomass, newdata = nd)
    adult_predictions$est_exp <- exp(adult_predictions$est)*10000
    max_raster <- quantile(adult_predictions$est_exp, .9999)
    max_adult <-  round(max(adult_predictions$est_exp))
  } else {
    rm(adult_predictions)
    total_biomass <- readRDS(paste0("data/model-total-biomass-", spp, "-1n3.rds"))
    nd$year<- as.integer(nd$year)
    predictions <- predict(total_biomass, newdata = nd)
    predictions$total_bio <- exp(predictions$est)*10000
    max_raster <- quantile(predictions$total_bio, .9999)
    max_bio <-  round(max(predictions$total_bio))
  }
```

```{r, error=TRUE}
if (exists("adult_predictions")) {
p_adult <- plot_facet_map(adult_predictions, "est_exp", 
  transform_col = fourth_root_power, 
  raster_limits = c(0, max_raster)) + 
  labs(fill="kg/ha") +
  ggtitle(paste0("", species, " mature biomass (max =",max_adult,"kg/ha)"))
print(p_adult) 
} else { 
  p_adult <- plot_facet_map(predictions, "total_bio", 
    #raster_limits = c(0, max_raster),
    transform_col = fourth_root_power) + 
  labs(fill="kg/ha") +
  ggtitle(paste0("", species, " total biomass (max = ",max_bio,"kg/ha)"))
print(p_adult) 
  }
```

```{r, message=FALSE, warning=FALSE}
if (exists("adult_predictions")) {

out_mature1 <- make_vector_data(adult_predictions,
  ssid = c(1),
  start_time = 2003,
  #start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
  #indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
)

out_mature3 <- make_vector_data(adult_predictions,
  ssid = c(3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
)

out_mature <- rbind(out_mature1, out_mature3)
} else { rm(out_mature)}
```

```{r, message=FALSE, warning=FALSE}
out_total <- make_vector_data(predictions,
  ssid = c(1,3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "total_bio",
  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
)
```

```{r eval=FALSE}
current_biomass <- plot_vocc(out_total,
  fill_col = "var_1_e",
  fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
  raster_alpha = 1,
  viridis_option = "C",
  vec_aes = NULL,
  transform_col = fourth_root_power
)
current_biomass <- current_biomass +
  ggtitle(paste0("", species, " total biomass"))

current_biomass
```

```{r, error=TRUE}
if (exists("out_mature")) {
data <- out_mature1
data <- out_mature
biotrend_ad <- plot_vocc(data,
  fill_col = "units_per_decade",
  fill_label = "Mature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_ad
} else {
  biotrend_ad <- plot_vocc(out_total,
  fill_col = "units_per_decade",
  fill_label = "Biomass trend \n(kg/ha/decade)\nfor 2003-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_ad
}
```

```{r, error=TRUE}
if (length(unique(fish$maturity_code))>2) {
    imm_biomass <- readRDS(paste0("data/model-imm-biomass-", spp, "-1n3.rds"))

    nd$year<- as.integer(nd$year)
    imm_predictions <- predict(imm_biomass, newdata = nd)
    imm_predictions$est_exp <- exp(imm_predictions$est)*10000

    predictions <- adult_predictions
    predictions$prop_imm <- imm_predictions$est_exp/(adult_predictions$est_exp+imm_predictions$est_exp)
    predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp

  }else{ 
    rm(imm_predictions) 
}
```

```{r}
if (exists("imm_predictions")) {
p_imm <- plot_facet_map(imm_predictions, "est_exp", transform_col = fourth_root_power, 
  raster_limits = c(0, max_raster)) + 
  labs(fill="kg/ha") +
  ggtitle(paste0("", species, " immature biomass"))
p_imm
} else {
  p_imm <- grid::grid.rect(gp=grid::gpar(col="white"))
}
```

```{r, message=FALSE, warning=FALSE}
if (exists("imm_predictions")) {

out_im1 <- make_vector_data(imm_predictions,
  ssid = c(1),
  start_time = 2003,
  #start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
  #indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
)

out_im3 <- make_vector_data(imm_predictions,
  ssid = c(3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
)
# glimpse(out_im)

out_imm <- rbind(out_im1, out_im3)
} else { rm(out_imm)}
```

```{r}
if (exists("out_imm")) {
data <- out_im1
data <- out_imm
biotrend_im <- plot_vocc(data,
  fill_col = "units_per_decade",
  fill_label = "Immature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_im
} else { biotrend_im <- grid::grid.rect(gp=grid::gpar(col="white"))}
```

#FIXME: failing
```{r}
if (exists("imm_predictions")) {

p_prop_imm <- plot_facet_map(predictions, "prop_imm") + 
  labs(fill="% by mass") +
  ggtitle(paste0("", species, " proportion immature"))
print(p_prop_imm)

} else { p_prop_imm <- grid::grid.rect(gp=grid::gpar(col="white"))}
```

Calculate predicted values for temperature
```{r, eval=FALSE}
m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
temperature <- predict(m_temp, newdata = nd_all)
temperature <- filter(temperature, ssid %in% model_ssid)
```

```{r, eval=FALSE}
temp <-plot_facet_map(temperature, "est", transform_col = no_trans, viridis_option = "D") + 
  labs(fill="°C") +
  ggtitle("Bottom temperature")
print(temp)
```

```{r, eval=FALSE}
png(file = paste0("figs/exploratory-", spp, "-figs.png"),   # The directory you want to save the file in
res = 600,
units = 'in',
width = 14, # The width of the plot in inches
height = 12) # The height of the plot in inches

gridExtra::grid.arrange(
    p_adult, p_imm, 
    p_prop_imm, #current_biomass, 
    biotrend_ad, biotrend_im, 
    temp,
    nrow = 2)
dev.off()
```

### 2019-06-12 analysis/VOCC/dist-vocc.Rmd
update temperature plots with all biannual-step plots

--- 2019-05-17 analysis/VOCC/dist-vocc.Rmd
@@
 library(gfranges)
+
 ```
@@
 
-nd <- rbind(nd_1, nd_3, nd_4, nd_16)
-nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
+nd_all <- rbind(nd_1, nd_3, nd_4, nd_16)
+nd <- nd_all %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
 ```
@@
 
-# can also be filtered to exclude certain years
+# # must be filtered to match years in model
 # m_temp <- readRDS("../VOCC/data/m_temp_allpost2003.rds")
@@
 
-predictions <- predict(m_temp, newdata = nd)
-
-plot_map <- function(dat, column = "est") {
-  ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    facet_wrap(~year) +
-    coord_fixed()
-}
-
-p <- plot_map(predictions, "est") +
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
-  ggtitle("Prediction (fixed effects + all random effects)")
-print(p)
+temp <- predict(m_temp, newdata = nd)
 ```
 
-Explore subsets ot the spatial range 
+Define subset of the spatial range 
 ```{r eval=FALSE}
-predicted1 <- predictions %>% filter(ssid == 1)
-predicted3 <- predictions %>% filter(ssid == 3)
-predicted4 <- predictions %>% filter(ssid == 4)
-predicted16 <- predictions %>% filter(ssid == 16)
-predicted1n3 <- predictions %>% filter(ssid != 16) %>% filter(ssid != 4) %>% filter(year > 2004)
+temperature <- temp %>% 
+  # filter(ssid!=3) %>% 
+  filter(ssid!=4) %>% 
+  filter(ssid!=16) %>% 
+  filter(year>2004)
 
-p <- plot_map(predicted1n3, "est") +
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
-  ggtitle("Prediction (fixed effects + all random effects)")
-print(p)
+#glimpse(temperature)
+
+temp <-plot_facet_map(temperature, "est", transform_col = no_trans) + 
+  labs(fill="kg/ha") +
+  ggtitle("Bottom temperature")
+print(temp)
+```
+
+Biannual changes starting 2005 and ending 2017
+```{r}
+out1 <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2005,
+  end_time = 2007,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 2,
+  delta_t_step = 2,
+  indices = c(1, 2),
+  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
+)
+gvocc1 <- plot_vocc(out1,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey87",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.2, 0.5),
+  max_vec_plotted = 50,
+  NA_label = ".",
+  fill_col = "var_1_e",
+  fill_label = "2007",
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans,
+  raster_limits = c(5, 13)
+)
+#gvocc1 <- gvocc1 + ggtitle("VOCC vectors for <1°C change in bottom temperature")
+gvocc1
 ```
@@
 ```{r}
-out <- make_vector_data(predictions,
-  ssid = c(1),
-  start_year = 2005,
+out2 <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2007,
+  end_time = 2009,
   input_cell_size = 2,
-  scale_fac = 2,
-  delta_t_total = 10,
+  scale_fac = 1,
+  delta_t_total = 2,
   delta_t_step = 2,
-  time_steps = 7,
-  indices = c(1, 1, 1, 2, 3, 3, 3),
+  indices = c(1, 2),
   thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
 )
-glimpse(out)
+gvocc2 <- plot_vocc(out2,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey87",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.2, 0.5),
+  max_vec_plotted = 50,
+  NA_label = ".",
+  fill_col = "var_1_e",
+  fill_label = "2009",
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans,
+  raster_limits = c(5, 13)
+)
+
+#gvocc2 <- gvocc2 + ggtitle(" ")
+gvocc2
 ```
@@
 ```{r}
-gvocc <- plot_vocc(out,
+out3 <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2009,
+  end_time = 2011,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 2,
+  delta_t_step = 2,
+  indices = c(1, 2),
+  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
+)
+gvocc3 <- plot_vocc(out3,
   low_col = "white",
   mid_col = "white",
-  high_col = "grey97",
+  high_col = "grey87",
   vec_aes = "distance",
-  vec_lwd_range = c(0.5, 0.6),
+  vec_lwd_range = c(0.2, 0.5),
+  max_vec_plotted = 50,
+  NA_label = ".",
+  fill_col = "var_1_e",
+  fill_label = "2011",
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans,
+  raster_limits = c(5, 13)
+)
+#gvocc3 <- gvocc3 + ggtitle(" ")
+gvocc3
+```
+
+```{r}
+out4 <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2011,
+  end_time = 2013,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 2,
+  delta_t_step = 2,
+  indices = c(1, 2),
+  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
+)
+gvocc4 <- plot_vocc(out4,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey87",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.2, 0.5),
+  max_vec_plotted = 50,
+  NA_label = ".",
+  fill_col = "var_1_e",
+  fill_label = "2013",
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans,
+  raster_limits = c(5, 13)
+)
+#gvocc4 <- gvocc4 + ggtitle(" ")
+gvocc4
+```
+
+```{r}
+out5 <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2013,
+  end_time = 2015,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 2,
+  delta_t_step = 2,
+  indices = c(1, 2),
+  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
+)
+gvocc5 <- plot_vocc(out5,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey87",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.2, 0.5),
+  max_vec_plotted = 50,
+  NA_label = ".",
+  fill_col = "var_1_e",
+  fill_label = "2015",
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans,
+  raster_limits = c(5, 13)
+)
+#gvocc5 <- gvocc5 + ggtitle(" ")
+gvocc5
+```
+
+```{r}
+out6 <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2015,
+  end_time = 2017,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 2,
+  delta_t_step = 2,
+  indices = c(1, 2),
+  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
+)
+gvocc6 <- plot_vocc(out6,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey87",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.2, 0.5),
+  max_vec_plotted = 50,
+  NA_label = ".",
+  fill_col = "var_1_e",
+  fill_label = "2017",
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans,
+  raster_limits = c(5, 13)
+)
+#gvocc6 <- gvocc6 + ggtitle(" ")
+gvocc6
+```
+
+```{r}
+png(file = "biannual_temp_change.png",   # The directory you want to save the file in
+res = 600,
+units = 'in',
+width = 10, # The width of the plot in inches
+height = 8) # The height of the plot in inches
+
+gridExtra::grid.arrange(gvocc1, gvocc2, gvocc3, gvocc4, gvocc5, gvocc6, nrow = 2, top = grid::textGrob("VOCC vectors for <0.5°C biannual changes in bottom temperature"))#,gp=gpar(fontsize=20,font=3)) 
+dev.off()
+```
+
+Cumulative change between last two decades 
+VOCC for 0.5 degree change and 1 degree change
+```{r}
+out0.5d <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2005,
+  #skip_time = 2011,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1, 1, 1, 2, 2, 2, 2),
+  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
+)
+```
+
+```{r}
+gvocc <- plot_vocc(out0.5,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey87",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.2, 0.5),
   max_vec_plotted = 100,
-  fill_col = "var_1_s",
-  fill_label = "Mean\ntemperature\n2005 & 2007",
+  fill_col = "var_1_e",
+  fill_label = "2013-2017",
   raster_alpha = 1,
-  vec_alpha = 0.8
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans
 )
+gvocc <- gvocc + ggtitle("VOCC vectors for <0.5°C change")
 gvocc
@@
 ```{r}
-gtrend <- plot_vocc(out,
-  fill_col = "C_per_decade",
-  fill_label = "Temperature\ntrend (°C/decade)\nfor 2005-2017",
+out1d <- make_vector_data(temperature,
+  ssid = c(1,3),
+  start_time = 2005,
+  #skip_time = 2011,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1, 1, 1, 2, 2, 2, 2),
+  thresholds = c(1) # vector of plus/minus threshold(s) to define climate match.
+)
+```
+
+```{r}
+gvocc1d <- plot_vocc(out1d,
+  low_col = "white",
+  mid_col = "white",
+  high_col = "grey87",
+  vec_aes = "distance",
+  vec_lwd_range = c(0.2, 0.5),
+  max_vec_plotted = 100,
+  fill_col = "var_1_e",
+  fill_label = "2013-2017",
   raster_alpha = 1,
-  vec_aes = NULL
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  viridis_option = "D",
+  transform_col = no_trans
 )
+gvocc1d <- gvocc1d + ggtitle("VOCC vectors for <1°C change")
+gvocc1d
+```
+
+```{r}
+gtrend <- plot_vocc(out1d,
+  fill_col = "units_per_decade",
+  fill_label = "°C per decade",
+  raster_alpha = 1,
+  vec_aes = NULL,
+  transform_col = no_trans
+)
+gtrend <- gtrend + ggtitle("Bottom temperature trend 2005-2017")
 gtrend
@@
 ```{r}
-# png(file = "accasp-wg-fig.png",   # The directory you want to save the file in
-# res = 600,
-# units = 'in',
-# width = 10, # The width of the plot in inches
-# height = 7) # The height of the plot in inches
+png(file = "temp-2005-2017-1n3.png",   # The directory you want to save the file in
+res = 600,
+units = 'in',
+width = 10, # The width of the plot in inches
+height = 7) # The height of the plot in inches
 
-gridExtra::grid.arrange(gtrend, gvocc, nrow = 1)
-# dev.off()
+gridExtra::grid.arrange(gtrend, gvocc, gvocc1d, nrow = 1)
+dev.off()
 ```

### 2019-06-13 analysis/VOCC/biomass-trends.Rmd
tweak for code usability across survey areas and retrieval of previously run models for speed, change orientation of exploratory fig to vertical

@@
 
+Choose one value for each parameter:
+
+Species
 ```{r}
@@
 species <- "Yelloweye Rockfish"
-species <- "North Pacific Spiny Dogfish" 
+species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
 species <- "Silvergray Rockfish"
@@
 
+Region of interest
 ```{r}
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-model_ssid <- c(1,3)
+region <- "Both odd year surveys"
+region <- "West Coast Vancouver Island"
+region <- "West Coast Haida Gwaii"
+region <- "Queen Charlotte Sound"
+region <- "Hecate Strait"
+region <- "Both West Coast surveys"
+
+## region <- "All synoptic surveys" # Probably not possible
 ```
+
+Is a new spatial needed?
+```{r}
+update_spatial_model <- TRUE
+update_spatial_model <- FALSE
+```
+
+Run all subsequent code...
 
@@
 
-```{r, message=FALSE, warning=FALSE}
-survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
-bath <- readRDS("data/bathymetry-data")
-# for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
-# fish <- fish[fish$fishing_event_id!=1506954,] 
+```{r}
+
+if(region == "Both odd year surveys") {
+survey <- c("SYN QCS","SYN HS")
+model_ssid <- c(1,3)
+years <- NULL
+}
+
+if(region == "West Coast Vancouver Island") {
+survey <- c("SYN WCVI")
+model_ssid <- c(4)
+years <- NULL
+}
+
+if(region == "West Coast Haida Gwaii") {
+survey <- c("SYN WCHG")
+model_ssid <- c(16)
+years <- NULL
+}
+
+if(region == "Queen Charlotte Sound") {
+survey <- c("SYN QCS")
+model_ssid <- c(1)
+years <- NULL
+}
+
+if(region == "Hecate Strait") {
+survey <- c("SYN HS")
+model_ssid <- c(3)
+years <- NULL
+}
+
+# if(region == "Both West Coast surveys") {
+# survey <- c("SYN WCVI","SYN WCHG")
+# model_ssid <- c(4,16)
+# years <- NULL
+# }
+
+# if(region == "All synoptic surveys") {
+# survey <- c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG")
+# model_ssid <- c(1,3,4,16)
+# years <- NULL
+# }
+
+ssid_string <- paste0(model_ssid, collapse = "n")
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
 ```
 
-Create prediction grids for each survey area so that only years with data are included
-```{r}
+Make prediction grids for all surveys and years, if not done before.
+```{r, message=FALSE, warning=FALSE}
+try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
+
+if (!exists("nd_all")) {
+bath <- readRDS("data/bathymetry-data")
 # choose base year(s) to create grid from
@@
 #nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
+
+saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
+}
 ```
 
-Calculate maturity weighted biomass
-```{r, eval=FALSE}
+Calculate maturity weighted biomass if not done before
+```{r}
+try(biomass <- readRDS(paste0("data/data-by-maturity-", spp,"-", ssid_string, ".rds")))
+
+if (!exists("biomass")) {
+
+survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
+fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+bath <- readRDS("data/bathymetry-data")
+# for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
+# fish <- fish[fish$fishing_event_id!=1506954,]
+
 maturity <- split_catch_maturity (survey_sets, fish, bath,
-  survey = c("SYN HS","SYN QCS"), 
-  years = NULL,
+  #survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"), 
+  survey = survey,
+  years = years,
   cutoff_quantile = c(.9995),
@@
 
-saveRDS(maturity$data, file = paste0("data/data-by-maturity-", spp, ".rds"))
+saveRDS(maturity$data, file = paste0("data/data-by-maturity-", spp, "-", ssid_string, ".rds"))
+}
+
 ```
 
-```{r, eval=FALSE}
-png(file = paste0("figs/", spp, "-mass-model.png"),   # The directory you want to save the file in
+```{r}
+if(exists("maturity")) {
+png(file = paste0("figs/", spp, "-mass-model-", ssid_string, ".png"),   # The directory you want to save the file in
 res = 600,
@@
 print(maturity$mass_model)
+}
 ```
 
-```{r, eval=FALSE}
-png(file = paste0("figs/", spp, "-maturity-year-re.png"),   # The directory you want to save the file in
+```{r}
+if(exists("maturity")) {
+png(file = paste0("figs/", spp, "-maturity-", ssid_string, ".png"),   # The directory you want to save the file in
 res = 600,
@@
 print(maturity$maturity)
+}
 ```
 
-Choose spatial zone
-```{r}
-biomass <- readRDS(paste0("data/data-by-maturity-", spp, ".rds"))
+Run sdmTMB model (if not done previously).
+```{r eval=FALSE}
+if (update_spatial_model) {
+
+biomass <- readRDS(paste0("data/data-by-maturity-", spp, "-", ssid_string, ".rds"))
 data <- biomass %>% gfranges:::scale_survey_predictors()
@@
 data <- data %>% filter(ssid %in% model_ssid)
-```
 
-Run sdmTMB model (if not done previously).
-```{r eval=FALSE}
 spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
@@
 if (any(names(data)=="adult_density")) {
+
 adult_biomass <- sdmTMB::sdmTMB(data,
@@
 
-saveRDS(adult_biomass, file = paste0("data/model-mature-biomass-", spp, "-1n3.rds"))
+saveRDS(adult_biomass, file = paste0("data/model-mature-biomass-", spp, "-", ssid_string, ".rds"))
 
-saveRDS(imm_biomass, file = paste0("data/model-imm-biomass-", spp, "-1n3.rds"))
+saveRDS(imm_biomass, file = paste0("data/model-imm-biomass-", spp, "-", ssid_string, ".rds"))
 
@@
 
-saveRDS(total_biomass, file = paste0("data/model-total-biomass-", spp, "-1n3.rds"))
+saveRDS(total_biomass, file = paste0("data/model-total-biomass-", spp, "-", ssid_string, ".rds"))
 
+}
 }
@@
 
-Calculate predicted values for available models
-#FIXME: tryCatch not working here, try this?
+Calculate predicted values for available models, if needed
 ```{r, error=TRUE}
+if (isFALSE(update_spatial_model)) {
+  try(adult_predictions <- readRDS(paste0("data/predictions-", spp, "-", ssid_string, "-mature-biomass.rds")))
+  try(imm_predictions <- readRDS(paste0("data/predictions-", spp, "-", ssid_string, "-imm-biomass.rds")))
+  try(predictions <- readRDS(paste0("data/predictions-", spp, "-", ssid_string, "-total-biomass.rds")))
+}
+
+if (!exists("adult_predictions")) {  
 if (length(unique(fish$maturity_code))>2) {
-    adult_biomass <- readRDS(paste0("data/model-mature-biomass-", spp, "-1n3.rds"))
     nd$year<- as.integer(nd$year)
+
+    adult_biomass <- readRDS(paste0("data/model-mature-biomass-", spp, "-", ssid_string, ".rds"))
     adult_predictions <- predict(adult_biomass, newdata = nd)
+    saveRDS(adult_predictions, file = paste0("data/predictions-", spp, "-", ssid_string, "-mature-biomass.rds"))
+
+    imm_biomass <- readRDS(paste0("data/model-imm-biomass-", spp, "-", ssid_string, ".rds"))
+    imm_predictions <- predict(imm_biomass, newdata = nd)
+    saveRDS(imm_predictions, file = paste0("data/predictions-", spp, "-", ssid_string, "-imm-biomass.rds"))
+
+  } else {
+    rm(adult_predictions)
+    rm(imm_predictions)
+
+    if (!exists(predictions)) {  
+    total_biomass <- readRDS(paste0("data/model-total-biomass-", spp, "-", ssid_string, ".rds"))
+    nd$year<- as.integer(nd$year)
+    predictions <- predict(total_biomass, newdata = nd)
+    saveRDS(predictions, file = paste0("data/predictions-", spp, "-", ssid_string, "-total-biomass.rds"))
+    }
+    }
+}
+
+if (exists("adult_predictions")) {
     adult_predictions$est_exp <- exp(adult_predictions$est)*10000
+    imm_predictions$est_exp <- exp(imm_predictions$est)*10000
+
+    adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
+    imm_predictions$prop_imm <- imm_predictions$est_exp/(adult_predictions$est_exp+imm_predictions$est_exp)
+
     max_raster <- quantile(adult_predictions$est_exp, .9999)
     max_adult <-  round(max(adult_predictions$est_exp))
+
 } else {  
-    rm(adult_predictions)
-    total_biomass <- readRDS(paste0("data/model-total-biomass-", spp, "-1n3.rds"))
-    nd$year<- as.integer(nd$year)
-    predictions <- predict(total_biomass, newdata = nd)
     predictions$total_bio <- exp(predictions$est)*10000
@@
   transform_col = fourth_root_power, 
-  raster_limits = c(0, max_raster)) + 
+  raster_limits = c(0, max_raster),
+  legend_position = "none"
+  ) + 
   labs(fill="kg/ha") +
-  ggtitle(paste0("", species, " mature biomass (max =",max_adult,"kg/ha)"))
+  ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
 print(p_adult) 
@@
   labs(fill="kg/ha") +
-  ggtitle(paste0("", species, " total biomass (max = ",max_bio,"kg/ha)"))
+  ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
 print(p_adult) 
@@
 if (exists("adult_predictions")) {
+  predictions <- adult_predictions 
+}
 
-out_mature1 <- make_vector_data(adult_predictions,
+if ( 1 %in% model_ssid) {
+out_mature1 <- make_vector_data(predictions,
   ssid = c(1),
@@
   #start_time = 2005,
-  #skip_time = 2011,
+  skip_time = 2004,
   input_cell_size = 2,
@@
   delta_t_step = 2,
-  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
+  indices = c(1, 1, 1, 1, 2, 2, 2, 2),
   #indices = c(1, 1, 1, 2, 2, 2, 2),
   variable_names = "est_exp",
-  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
+  thresholds = c(0.001) 
 )
+} else { out_mature1 <- NULL }
 
-out_mature3 <- make_vector_data(adult_predictions,
+if ( 3 %in% model_ssid) {  
+out_mature3 <- make_vector_data(predictions,
   ssid = c(3),
   start_time = 2005,
-  #skip_time = 2011,
   input_cell_size = 2,
@@
   variable_names = "est_exp",
-  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
+  thresholds = c(0.001) 
 )
+} else { out_mature3 <- NULL }
 
-out_mature <- rbind(out_mature1, out_mature3)
-} else { rm(out_mature)}
-```
-
-```{r, message=FALSE, warning=FALSE}
-out_total <- make_vector_data(predictions,
-  ssid = c(1,3),
-  start_time = 2005,
-  #skip_time = 2011,
+if ( 4 %in% model_ssid) {  
+out_mature4 <- make_vector_data(predictions,
+  ssid = c(4),
+  start_time = 2004,
   input_cell_size = 2,
@@
   delta_t_step = 2,
-  indices = c(1, 1, 1, 2, 2, 2, 2),
-  variable_names = "total_bio",
-  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
+  indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+  variable_names = "est_exp",
+  thresholds = c(0.001) 
 )
+} else { out_mature4 <- NULL }
+
+if ( 16 %in% model_ssid) {  
+out_mature16 <- make_vector_data(predictions,
+  ssid = c(16),
+  start_time = 2004,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+  variable_names = "est_exp",
+  thresholds = c(0.001) 
+)
+} else { out_mature16 <- NULL }
+out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)
+
+if (length(unique(fish$maturity_code))<2) {
+  out_total <- out_mature
+} 
 ```
 
+Optional plot of current total biomass 
 ```{r eval=FALSE}
+if (exists("out_total")) {
 current_biomass <- plot_vocc(out_total,
@@
 current_biomass
+}
 ```
@@
 if (exists("out_mature")) {
-data <- out_mature1
 data <- out_mature
@@
   fill_col = "units_per_decade",
-  fill_label = "Mature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
+  fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
   raster_alpha = 1,
   vec_aes = NULL,
-  transform_col = fourth_root_power
-)
+  # axis_lables = TRUE,
+  transform_col = fourth_root_power,
+  legend_position = c(0.2, 0.25)
+)+ ggtitle(paste0("", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))))
 biotrend_ad
@@
   fill_col = "units_per_decade",
-  fill_label = "Biomass trend \n(kg/ha/decade)\nfor 2003-2017",
+  fill_label = "Biomass \ntrend \n(kg/ha/decade)",
   raster_alpha = 1,
   vec_aes = NULL,
-  transform_col = fourth_root_power
-)
+  # axis_lables = TRUE,
+  transform_col = fourth_root_power,
+  legend_position = c(0.2, 0.25)
+)+ ggtitle(paste0("", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))))
 biotrend_ad
-}
-```
-
-```{r, error=TRUE}
-if (length(unique(fish$maturity_code))>2) {
-    imm_biomass <- readRDS(paste0("data/model-imm-biomass-", spp, "-1n3.rds"))
-
-    nd$year<- as.integer(nd$year)
-    imm_predictions <- predict(imm_biomass, newdata = nd)
-    imm_predictions$est_exp <- exp(imm_predictions$est)*10000
-
-    predictions <- adult_predictions
-    predictions$prop_imm <- imm_predictions$est_exp/(adult_predictions$est_exp+imm_predictions$est_exp)
-    predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
-
-  }else{ 
-    rm(imm_predictions) 
 }
@@
   labs(fill="kg/ha") +
-  ggtitle(paste0("", species, " immature biomass"))
+  ggtitle(paste0("", species, " immature biomass \n "))
 p_imm
@@
 if (exists("imm_predictions")) {
+  predictions <- imm_predictions
 
-out_im1 <- make_vector_data(imm_predictions,
+if ( 1 %in% model_ssid) {
+out_im1 <- make_vector_data(predictions,
   ssid = c(1),
@@
   delta_t_step = 2,
-  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
+  indices = c(1, 1, 1, 1, 2, 2, 2, 2),
   #indices = c(1, 1, 1, 2, 2, 2, 2),
   variable_names = "est_exp",
-  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
+  thresholds = c(0.001) 
 )
+} else { out_im1 <- NULL }
 
-out_im3 <- make_vector_data(imm_predictions,
+if ( 3 %in% model_ssid) {  
+out_im3 <- make_vector_data(predictions,
   ssid = c(3),
   start_time = 2005,
-  #skip_time = 2011,
   input_cell_size = 2,
@@
   variable_names = "est_exp",
-  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
+  thresholds = c(0.001) 
 )
-# glimpse(out_im)
+} else { out_im3 <- NULL }
 
-out_imm <- rbind(out_im1, out_im3)
-} else { rm(out_imm)}
+if ( 4 %in% model_ssid) {  
+out_im4 <- make_vector_data(predictions,
+  ssid = c(4),
+  start_time = 2004,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+  variable_names = "est_exp",
+  thresholds = c(0.001) 
+)
+} else { out_im4 <- NULL }
+
+if ( 16 %in% model_ssid) {  
+out_im16 <- make_vector_data(predictions,
+  ssid = c(16),
+  start_time = 2004,
+  input_cell_size = 2,
+  scale_fac = 1,
+  delta_t_total = 10,
+  delta_t_step = 2,
+  indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+  variable_names = "est_exp",
+  thresholds = c(0.001) 
+)
+} else { out_im16 <- NULL }
+out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
+
+} else { rm(out_imm) }
+
 ```
@@
   fill_col = "units_per_decade",
-  fill_label = "Immature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
+  fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
   raster_alpha = 1,
   vec_aes = NULL,
-  transform_col = fourth_root_power
-)
+  transform_col = fourth_root_power,
+  #axis_lables = TRUE,
+  legend_position = c(0.2, 0.25)
+) + ggtitle(paste0("", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))))
 biotrend_im
@@
 
-#FIXME: failing
 ```{r}
@@
 
-p_prop_imm <- plot_facet_map(predictions, "prop_imm") + 
-  labs(fill="% by mass") +
+p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") + 
+  labs(fill="% \nby mass") +
   ggtitle(paste0("", species, " proportion immature"))
@@
 
-Calculate predicted values for temperature
+Retrieve (or calculate) predicted values for temperature
 ```{r, eval=FALSE}
-m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
-temperature <- predict(m_temp, newdata = nd_all)
+# m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
+# temperature <- predict(m_temp, newdata = nd_all)
+# saveRDS(temperature, file = "data/predicted_temp_allyears.rds")
+
+temperature <- readRDS("data/predicted_temp_allyears.rds")
 temperature <- filter(temperature, ssid %in% model_ssid)
@@
 ```{r, eval=FALSE}
-png(file = paste0("figs/exploratory-", spp, "-figs.png"),   # The directory you want to save the file in
+png(file = paste0("figs/exploratory-", spp, "-", ssid_string, ".png"),   # The directory you want to save the file in
 res = 600,
 units = 'in',
-width = 14, # The width of the plot in inches
-height = 12) # The height of the plot in inches
+width = 9.5, # The width of the plot in inches
+height = 12.5) # The height of the plot in inches
 
@@
     p_adult, p_imm, 
-    p_prop_imm, #current_biomass, 
     biotrend_ad, biotrend_im, 
-    temp,
-    nrow = 2)
+    #current_biomass, 
+    p_prop_imm, temp,
+    nrow = 3)
 dev.off()
+
 ```

### 2019-06-20 analysis/VOCC/biomass-trends.Rmd
make species subfolders for data and figures, stitch together of separate region models and save separate plots for each ssid, fix problem saving maturity plots

--- 2019-06-12 analysis/VOCC/biomass-trends.Rmd
@@
 species <- "Yelloweye Rockfish"
-species <- "North Pacific Spiny Dogfish" 
+species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
 species <- "Silvergray Rockfish"
 species <- "Arrowtooth Flounder"
+species <- "Lingcod"
+
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+
+# folder to hold figs for this species
+dir.create(file.path("figs", spp))
+dir.create(file.path("data", spp))
 ```
 
+Region of interest
 ```{r}
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-model_ssid <- c(1,3)
+region <- "Both odd year surveys"
+region <- "West Coast Vancouver Island"
+region <- "West Coast Haida Gwaii"
+# region <- "Queen Charlotte Sound"
+# region <- "Hecate Strait"
+# region <- "Both West Coast surveys"
+region <- "All synoptic surveys" # Probably not possible
 ```
+
+Is a new spatial needed?
+```{r}
+update_spatial_model <- TRUE
+# update_spatial_model <- FALSE
+priors <- FALSE
+# priors <- TRUE
+```
+
+Run all subsequent code...
 
@@
 
-```{r, message=FALSE, warning=FALSE}
-survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
-bath <- readRDS("data/bathymetry-data")
-# for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
-# fish <- fish[fish$fishing_event_id!=1506954,] 
+```{r}
+
+if (region == "Both odd year surveys") {
+  survey <- c("SYN QCS", "SYN HS")
+  model_ssid <- c(1, 3)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  years <- NULL
+}
+
+if (region == "West Coast Vancouver Island") {
+  survey <- c("SYN WCVI")
+  model_ssid <- c(4)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  years <- NULL
+}
+
+if (region == "West Coast Haida Gwaii") {
+  survey <- c("SYN WCHG")
+  model_ssid <- c(16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  years <- NULL
+}
+
+# if(region == "Queen Charlotte Sound") {
+# survey <- c("SYN QCS")
+# model_ssid <- c(1)
+# ssid_string <- paste0(model_ssid, collapse = "n")
+# years <- NULL
+# }
+#
+# if(region == "Hecate Strait") {
+# survey <- c("SYN HS")
+# model_ssid <- c(3)
+# ssid_string <- paste0(model_ssid, collapse = "n")
+# years <- NULL
+# }
+
+# if(region == "Both West Coast surveys") {
+# survey <- c("SYN WCVI","SYN WCHG")
+# model_ssid <- c(4,16)
+# years <- NULL
+# }
+
+if (region == "All synoptic surveys") {
+  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  years <- NULL
+}
 ```
 
-Create prediction grids for each survey area so that only years with data are included
-```{r}
+Make prediction grids for all surveys and years, if not done before.
+```{r, message=FALSE, warning=FALSE}
+try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
+
+if (!exists("nd_all")) {
+  bath <- readRDS("data/bathymetry-data")
+
   # choose base year(s) to create grid from
-dummy_year <- c(2005,2006)
+  dummy_year <- c(2005, 2006)
   dat <- bath$data %>% scale_survey_predictors()
+
   # create grids for each ssid separately
@@
     mutate(year = as.numeric(year))
-#nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
+  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
+
+  saveRDS(nd_all, file = paste0("data/", spp, "/nd_all_synoptic.rds"))
+}
 ```
 
-Calculate maturity weighted biomass
-```{r, eval=FALSE}
-maturity <- split_catch_maturity (survey_sets, fish, bath,
-  survey = c("SYN HS","SYN QCS"), 
-  years = NULL,
+Calculate maturity weighted biomass if not done before
+```{r}
+rm(biomass)
+try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
+
+if (!exists("biomass")) {
+  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
+  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+  bath <- readRDS("data/bathymetry-data")
+  # for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
+  # fish <- fish[fish$fishing_event_id!=1506954,]
+
+  maturity <- split_catch_maturity(survey_sets, fish, bath,
+    # survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
+    survey = c("SYN HS", "SYN WCHG"),
+    #  survey = survey,
+    years = years,
     cutoff_quantile = c(.9995),
@@
 
-saveRDS(maturity$data, file = paste0("data/data-by-maturity-", spp, ".rds"))
+  saveRDS(maturity$data, file = paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
+}
 ```
 
-```{r, eval=FALSE}
-png(file = paste0("figs/", spp, "-mass-model.png"),   # The directory you want to save the file in
-res = 600,
-units = 'in',
+```{r}
+if (exists("maturity")) {
+  print(maturity$mass_model)
+  ggsave(
+    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
+    dpi = 600,
     width = 8, # The width of the plot in inches
-height = 10)
+    height = 10
+  )
+}
 
-maturity$mass_model
-
-dev.off()
-print(maturity$mass_model)
+# try(print(maturity$mass_model))
 ```
 
-```{r, eval=FALSE}
-png(file = paste0("figs/", spp, "-maturity-year-re.png"),   # The directory you want to save the file in
+```{r}
+if (exists("maturity")) {
+  png(
+    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), # The directory you want to save the file in
     res = 600,
-units = 'in',
+    units = "in",
     width = 8, # The width of the plot in inches
-height = 10)
+    height = 10
+  )
+  print(maturity$maturity)
+  dev.off()
+}
 
-maturity$maturity
-
-dev.off()
-print(maturity$maturity)
-```
-
-Choose spatial zone
-```{r}
-biomass <- readRDS(paste0("data/data-by-maturity-", spp, ".rds"))
-data <- biomass %>% gfranges:::scale_survey_predictors()
-
-nd <- nd_all %>% filter(ssid %in% model_ssid) 
-data <- data %>% filter(ssid %in% model_ssid)
+try(print(maturity$maturity))
 ```
@@
 ```{r eval=FALSE}
+if (update_spatial_model) {
+  biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
+  data <- biomass %>% gfranges:::scale_survey_predictors()
+
+  data <- data %>% filter(ssid %in% model_ssid)
+
+  if (region == "All synoptic surveys") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 1200)
+  }
+
+  if (region == "Both odd year surveys") {
     spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
+  }
+
+  if (region == "West Coast Vancouver Island") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
+  }
+
+  if (region == "West Coast Haida Gwaii") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 300)
+  }
+
   sdmTMB::plot_spde(spde)
 
-if (any(names(data)=="adult_density")) {
+  if (any(names(data) == "adult_density")) {
+    starttime1 <- Sys.time()
     adult_biomass <- sdmTMB::sdmTMB(data,
@@
       include_spatial = TRUE,
-  enable_priors = TRUE,
-  # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      enable_priors = priors,
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
     )
+    endtime1 <- Sys.time()
+    time1 <- round(starttime1 - endtime1)
 
+    saveRDS(adult_biomass, file = paste0("data/", spp, "/model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, time1, ".rds"))
+
+    starttime2 <- Sys.time()
     imm_biomass <- sdmTMB::sdmTMB(data,
@@
       include_spatial = TRUE,
-  enable_priors = TRUE,
+      enable_priors = priors,
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
     )
+    endtime2 <- Sys.time()
+    time2 <- round(starttime2 - endtime2)
 
-saveRDS(adult_biomass, file = paste0("data/model-mature-biomass-", spp, "-1n3.rds"))
-
-saveRDS(imm_biomass, file = paste0("data/model-imm-biomass-", spp, "-1n3.rds"))
-
+    saveRDS(imm_biomass, file = paste0("data/", spp, "/model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, time2, ".rds"))
   } else {
-
+    starttime3 <- Sys.time()
     total_biomass <- sdmTMB::sdmTMB(dat,
@@
       ar1_fields = TRUE,
-  include_spatial = TRUE,
+      include_spatial = TRUE, ,
+      enable_priors = priors,
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
     )
+    endtime3 <- Sys.time()
+    time3 <- round(starttime3 - endtime3)
 
-saveRDS(total_biomass, file = paste0("data/model-total-biomass-", spp, "-1n3.rds"))
-
+    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, ".rds"))
+  }
 }
@@
 
-Calculate predicted values for available models
-#FIXME: tryCatch not working here, try this?
+Calculate predicted values for available models, if needed
 ```{r, error=TRUE}
-if (length(unique(fish$maturity_code))>2) {
-    adult_biomass <- readRDS(paste0("data/model-mature-biomass-", spp, "-1n3.rds"))
-    nd$year<- as.integer(nd$year)
+rm(adult_predictions)
+rm(imm_predictions)
+rm(predictions)
+
+if (isFALSE(update_spatial_model)) {
+  if (region != "All synoptic surveys") {
+    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
+    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-imm-biomass-priors-", priors, ".rds")))
+    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
+  } else {
+    # Bind together predictions from different models
+    try ({
+      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-priors-", priors, ".rds"))
+      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-priors-", priors, ".rds"))
+      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-priors-", priors, ".rds"))
+      adult_predictions <- rbind(pred1n3, pred4, pred16) 
+      # model_ssid <- unique(adult_predictions$ssid)
+      # ssid_string <- paste0(model_ssid, collapse = "n")
+    })
+
+    try ({
+      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-imm-biomass-priors-", priors, ".rds"))
+      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-imm-biomass-priors-", priors, ".rds"))
+      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-imm-biomass-priors-", priors, ".rds"))
+      imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm) 
+      # model_ssid <- unique(imm_predictions$ssid)
+      # ssid_string <- paste0(model_ssid, collapse = "n")
+    })
+
+    try ({
+      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-priors-", priors, ".rds"))
+      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-priors-", priors, ".rds"))
+      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-priors-", priors, ".rds"))
+      predictions <- rbind(pred1n3_all, pred4_all, pred16_all) # ,
+      #  model_ssid <- unique(predictions$ssid),
+      # ssid_string <- paste0(model_ssid, collapse = "n")
+    })
+
+    if (!exists("adult_predictions")) paste("Maturity split biomass predictions are not available.")
+    if (!exists("predictions")) paste("Total biomass predictions are not available.")
+  }
+} else {
+  if (region == "All synoptic surveys") {
+    paste("To update spatial models, you need to build and save each region's model separately.")
+  }
+}
+
+if (!exists("adult_predictions")) {
+  if (length(unique(fish$maturity_code)) > 2) {
+    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
+    nd <- nd_all %>% filter(ssid %in% model_ssid)
+    nd$year <- as.integer(nd$year)
+
+    adult_biomass <- readRDS(list.files(
+      path = paste0("data/", spp, "/"),
+      pattern = paste0("model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, "-*"),
+      all.files = F, full.names = T
+    ))
+
     adult_predictions <- predict(adult_biomass, newdata = nd)
-    adult_predictions$est_exp <- exp(adult_predictions$est)*10000
+    saveRDS(adult_predictions,
+      file = paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
+    )
+
+    imm_biomass <- readRDS(list.files(
+      path = paste0("data/", spp, "/"),
+      pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, "-*"),
+      all.files = F, full.names = T
+    ))
+    imm_predictions <- predict(imm_biomass, newdata = nd)
+    saveRDS(imm_predictions,
+      file = paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-imm-biomass-priors-", priors, ".rds")
+    )
+  } else {
+    rm(adult_predictions)
+    rm(imm_predictions)
+
+    if (!exists(predictions)) {
+      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
+      nd <- nd_all %>% filter(ssid %in% model_ssid)
+      nd$year <- as.integer(nd$year)
+
+      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, ".rds"))
+      nd$year <- as.integer(nd$year)
+      predictions <- predict(total_biomass, newdata = nd)
+      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"))
+    }
+  }
+}
+```
+
+Bind together predictions from different models
+```{r, eval=FALSE}
+if (region == "All synoptic surveys") {
+  if (isFALSE(update_spatial_model)) {
+    try(
+      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-priors-", priors, ".rds")),
+      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-priors-", priors, ".rds")),
+      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-priors-", priors, ".rds")),
+
+      adult_predictions <- rbind(pred1n3, pred4, pred16),
+      model_ssid <- unique(adult_predictions$ssid),
+      ssid_string <- paste0(model_ssid, collapse = "n")
+    )
+
+    try(
+      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-imm-biomass-priors-", priors, ".rds")),
+      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-imm-biomass-priors-", priors, ".rds")),
+      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-imm-biomass-priors-", priors, ".rds")),
+
+      imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm),
+      model_ssid <- unique(imm_predictions$ssid),
+      ssid_string <- paste0(model_ssid, collapse = "n")
+    )
+
+    try(
+      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-priors-", priors, ".rds")),
+      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-priors-", priors, ".rds")),
+      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-priors-", priors, ".rds")),
+
+      predictions <- rbind(pred1n3_all, pred4_all, pred16_all),
+      model_ssid <- unique(predictions$ssid),
+      ssid_string <- paste0(model_ssid, collapse = "n")
+    )
+  } else {
+    paste("To update spatial models, you need to build and save each region's model separately.")
+  }
+}
+
+if (!exists(adult_predictions)) paste("Maturity split biomass predictions were not available.")
+if (!exists(predictions)) paste("Total biomass predictions were not available.")
+```
+
+Caluclate variables of interest from predictions
+```{r}
+if (exists("adult_predictions")) {
+  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
+  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000
+
+  adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
+  imm_predictions$prop_imm <- imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
+
   max_raster <- quantile(adult_predictions$est_exp, .9999)
@@
 } else {
-    rm(adult_predictions)
-    total_biomass <- readRDS(paste0("data/model-total-biomass-", spp, "-1n3.rds"))
-    nd$year<- as.integer(nd$year)
-    predictions <- predict(total_biomass, newdata = nd)
-    predictions$total_bio <- exp(predictions$est)*10000
+  predictions$total_bio <- exp(predictions$est) * 10000
   max_raster <- quantile(predictions$total_bio, .9999)
@@
 if (exists("adult_predictions")) {
-p_adult <- plot_facet_map(adult_predictions, "est_exp", 
+  p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
     transform_col = fourth_root_power,
-  raster_limits = c(0, max_raster)) + 
-  labs(fill="kg/ha") +
-  ggtitle(paste0("", species, " mature biomass (max =",max_adult,"kg/ha)"))
-print(p_adult) 
+    raster_limits = c(0, max_raster),
+    legend_position = "none"
+  ) +
+    labs(fill = "kg/ha") +
+    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
+  print(p_adult_all)
 } else {
-  p_adult <- plot_facet_map(predictions, "total_bio", 
-    #raster_limits = c(0, max_raster),
-    transform_col = fourth_root_power) + 
-  labs(fill="kg/ha") +
-  ggtitle(paste0("", species, " total biomass (max = ",max_bio,"kg/ha)"))
+  p_adult_all <- plot_facet_map(predictions, "total_bio",
+    # raster_limits = c(0, max_raster),
+    transform_col = fourth_root_power
+  ) +
+    labs(fill = "kg/ha") +
+    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
+  print(p_adult_all)
+}
+```
+
+```{r}
+if (exists("adult_predictions")) {
+  # if (length(model_ssid)>1){
+  if (1 %in% model_ssid) {
+    adult_predictions1 <- adult_predictions %>% filter(ssid == 1)
+  } else {
+    adult_predictions1 <- NULL
+  }
+
+  if (3 %in% model_ssid) {
+    adult_predictions3 <- adult_predictions %>% filter(ssid == 3)
+  } else {
+    adult_predictions3 <- NULL
+  }
+
+  if (4 %in% model_ssid) {
+    adult_predictions4 <- adult_predictions %>% filter(ssid == 4)
+  } else {
+    adult_predictions4 <- NULL
+  }
+
+  if (16 %in% model_ssid) {
+    adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
+  } else {
+    adult_predictions16 <- NULL
+  }
+
+  ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)
+
+  ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]
+
+  p_adult <- list()
+  for (i in seq_len(length(model_ssid))) {
+    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
+      raster_limits = c(0, max_raster),
+      # legend_position = "none"
+      transform_col = fourth_root_power
+    ) +
+      labs(fill = "kg/ha") +
+      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
+    p_adult[[i]]
+  }
   print(p_adult)
 }
+# }
 ```
@@
 if (exists("adult_predictions")) {
+  predictions <- adult_predictions
+  # model_ssid <- unique(predictions$ssid)
+  # ssid_string <- paste0(model_ssid, collapse = "n")
+}
 
-out_mature1 <- make_vector_data(adult_predictions,
+if (1 %in% model_ssid) {
+  out_mature1 <- make_vector_data(predictions,
     ssid = c(1),
     start_time = 2003,
-  #start_time = 2005,
-  #skip_time = 2011,
+    # start_time = 2005,
+    skip_time = 2004,
     input_cell_size = 2,
@@
     delta_t_step = 2,
-  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
-  #indices = c(1, 1, 1, 2, 2, 2, 2),
+    indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+    # indices = c(1, 1, 1, 2, 2, 2, 2),
     variable_names = "est_exp",
-  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
+    thresholds = c(0.001)
   )
+} else {
+  out_mature1 <- NULL
+}
 
-out_mature3 <- make_vector_data(adult_predictions,
+if (3 %in% model_ssid) {
+  out_mature3 <- make_vector_data(predictions,
     ssid = c(3),
     start_time = 2005,
-  #skip_time = 2011,
     input_cell_size = 2,
@@
     variable_names = "est_exp",
-  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
+    thresholds = c(0.001)
   )
+} else {
+  out_mature3 <- NULL
+}
 
-out_mature <- rbind(out_mature1, out_mature3)
-} else { rm(out_mature)}
-```
+if (4 %in% model_ssid) {
+  out_mature4 <- make_vector_data(predictions,
+    ssid = c(4),
+    start_time = 2004,
+    input_cell_size = 2,
+    scale_fac = 1,
+    delta_t_total = 10,
+    delta_t_step = 2,
+    indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+    variable_names = "est_exp",
+    thresholds = c(0.001)
+  )
+} else {
+  out_mature4 <- NULL
+}
 
-```{r, message=FALSE, warning=FALSE}
-out_total <- make_vector_data(predictions,
-  ssid = c(1,3),
-  start_time = 2005,
-  #skip_time = 2011,
+if (16 %in% model_ssid) {
+  out_mature16 <- make_vector_data(predictions,
+    ssid = c(16),
+    start_time = 2004,
+    skip_time = 2007,
     input_cell_size = 2,
@@
     indices = c(1, 1, 1, 2, 2, 2, 2),
-  variable_names = "total_bio",
-  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
+    variable_names = "est_exp",
+    thresholds = c(0.001)
   )
+} else {
+  out_mature16 <- NULL
+}
+out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)
+
+saveRDS(out_mature,
+  file = paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
+)
+
+if (!exists("adult_predictions")) {
+  out_total <- out_mature
+  saveRDS(out_total,
+    file = paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
+  )
+}
 ```
 
+Optional plot of current total biomass 
 ```{r eval=FALSE}
+if (exists("out_total")) {
   current_biomass <- plot_vocc(out_total,
@@
   current_biomass
+}
 ```
@@
 ```{r, error=TRUE}
+try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
+
 if (exists("out_mature")) {
-data <- out_mature1
   data <- out_mature
@@
     fill_col = "units_per_decade",
-  fill_label = "Mature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
+    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
     raster_alpha = 1,
     vec_aes = NULL,
-  transform_col = fourth_root_power
-)
+    # axis_lables = TRUE,
+    white_zero = TRUE,
+    transform_col = fourth_root_power,
+    legend_position = c(0.2, 0.25)
+  ) + ggtitle(paste0("", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))))
   biotrend_ad
 } else {
+  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
+
   biotrend_ad <- plot_vocc(out_total,
     fill_col = "units_per_decade",
-  fill_label = "Biomass trend \n(kg/ha/decade)\nfor 2003-2017",
+    fill_label = "Biomass \ntrend \n(kg/ha/decade)",
     raster_alpha = 1,
     vec_aes = NULL,
-  transform_col = fourth_root_power
-)
+    # axis_lables = TRUE,
+    white_zero = TRUE,
+    transform_col = fourth_root_power,
+    legend_position = c(0.2, 0.25)
+  ) + ggtitle(paste0("", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))))
   biotrend_ad
@@
 ```{r, error=TRUE}
-if (length(unique(fish$maturity_code))>2) {
-    imm_biomass <- readRDS(paste0("data/model-imm-biomass-", spp, "-1n3.rds"))
-
-    nd$year<- as.integer(nd$year)
-    imm_predictions <- predict(imm_biomass, newdata = nd)
-    imm_predictions$est_exp <- exp(imm_predictions$est)*10000
-
-    predictions <- adult_predictions
-    predictions$prop_imm <- imm_predictions$est_exp/(adult_predictions$est_exp+imm_predictions$est_exp)
-    predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
-
-  }else{ 
-    rm(imm_predictions) 
+if (exists("imm_predictions")) {
+  p_imm_all <- plot_facet_map(imm_predictions, "est_exp",
+    transform_col = fourth_root_power,
+    raster_limits = c(0, max_raster)
+  ) +
+    labs(fill = "kg/ha") +
+    ggtitle(paste0("", species, " immature biomass \n "))
+  p_imm_all
+} else {
+  p_imm_all <- grid::grid.rect(gp = grid::gpar(col = "white"))
 }
@@
 ```{r}
-if (exists("imm_predictions")) {
-p_imm <- plot_facet_map(imm_predictions, "est_exp", transform_col = fourth_root_power, 
-  raster_limits = c(0, max_raster)) + 
-  labs(fill="kg/ha") +
-  ggtitle(paste0("", species, " immature biomass"))
-p_imm
+if (exists("adult_predictions")) {
+  # if (length(model_ssid)>1){
+  if (1 %in% model_ssid) {
+    imm_predictions1 <- imm_predictions %>% filter(ssid == 1)
   } else {
-  p_imm <- grid::grid.rect(gp=grid::gpar(col="white"))
+    imm_predictions1 <- NULL
   }
+
+  if (3 %in% model_ssid) {
+    imm_predictions3 <- imm_predictions %>% filter(ssid == 3)
+  } else {
+    imm_predictions3 <- NULL
+  }
+
+  if (4 %in% model_ssid) {
+    imm_predictions4 <- imm_predictions %>% filter(ssid == 4)
+  } else {
+    imm_predictions4 <- NULL
+  }
+
+  if (16 %in% model_ssid) {
+    imm_predictions16 <- imm_predictions %>% filter(ssid == 16)
+  } else {
+    imm_predictions16 <- NULL
+  }
+
+  im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)
+
+  im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]
+
+  p_imm <- list()
+  for (i in seq_len(length(model_ssid))) {
+    p_imm[[i]] <- plot_facet_map(im_prediction_list[[i]], "est_exp",
+      raster_limits = c(0, max_raster),
+      # legend_position = "none",
+      transform_col = fourth_root_power
+    ) +
+      labs(fill = "kg/ha") +
+      ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))
+    p_imm[[i]]
+  }
+  print(p_imm)
+}
+# }
 ```
@@
 if (exists("imm_predictions")) {
+  predictions <- imm_predictions
 
-out_im1 <- make_vector_data(imm_predictions,
+  if (1 %in% model_ssid) {
+    out_im1 <- make_vector_data(predictions,
       ssid = c(1),
       start_time = 2003,
-  #start_time = 2005,
-  #skip_time = 2011,
+      # start_time = 2005,
+      skip_time = 2004,
       input_cell_size = 2,
@@
       delta_t_step = 2,
-  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
-  #indices = c(1, 1, 1, 2, 2, 2, 2),
+      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+      # indices = c(1, 1, 1, 2, 2, 2, 2),
       variable_names = "est_exp",
-  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
+      thresholds = c(0.001)
     )
+  } else {
+    out_im1 <- NULL
+  }
 
-out_im3 <- make_vector_data(imm_predictions,
+  if (3 %in% model_ssid) {
+    out_im3 <- make_vector_data(predictions,
       ssid = c(3),
       start_time = 2005,
-  #skip_time = 2011,
       input_cell_size = 2,
@@
       variable_names = "est_exp",
-  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
+      thresholds = c(0.001)
     )
-# glimpse(out_im)
+  } else {
+    out_im3 <- NULL
+  }
 
-out_imm <- rbind(out_im1, out_im3)
-} else { rm(out_imm)}
+  if (4 %in% model_ssid) {
+    out_im4 <- make_vector_data(predictions,
+      ssid = c(4),
+      start_time = 2004,
+      input_cell_size = 2,
+      scale_fac = 1,
+      delta_t_total = 10,
+      delta_t_step = 2,
+      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+      variable_names = "est_exp",
+      thresholds = c(0.001)
+    )
+  } else {
+    out_im4 <- NULL
+  }
+
+  if (16 %in% model_ssid) {
+    out_im16 <- make_vector_data(predictions,
+      ssid = c(16),
+      start_time = 2004,
+      input_cell_size = 2,
+      scale_fac = 1,
+      delta_t_total = 10,
+      delta_t_step = 2,
+      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+      variable_names = "est_exp",
+      thresholds = c(0.001)
+    )
+  } else {
+    out_im16 <- NULL
+  }
+
+  out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
+  saveRDS(out_imm,
+    file = paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
+  )
+} else {
+  rm(out_imm)
+}
 ```
@@
 ```{r}
+try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
+
 if (exists("out_imm")) {
-data <- out_im1
   data <- out_imm
@@
     fill_col = "units_per_decade",
-  fill_label = "Immature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
+    fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
     raster_alpha = 1,
     vec_aes = NULL,
-  transform_col = fourth_root_power
-)
+    white_zero = TRUE,
+    transform_col = fourth_root_power,
+    # axis_lables = TRUE,
+    legend_position = c(0.2, 0.25)
+  ) + ggtitle(paste0("", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))))
   biotrend_im
-} else { biotrend_im <- grid::grid.rect(gp=grid::gpar(col="white"))}
+} else {
+  biotrend_im <- grid::grid.rect(gp = grid::gpar(col = "white"))
+}
 ```
 
-#FIXME: failing
-```{r}
+```{r, error=TRUE}
 if (exists("imm_predictions")) {
-
-p_prop_imm <- plot_facet_map(predictions, "prop_imm") + 
-  labs(fill="% by mass") +
+  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
+    labs(fill = "% \nby mass") +
     ggtitle(paste0("", species, " proportion immature"))
   print(p_prop_imm)
-
-} else { p_prop_imm <- grid::grid.rect(gp=grid::gpar(col="white"))}
+} else {
+  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
+}
 ```
 
-Calculate predicted values for temperature
+Retrieve (or calculate) predicted values for temperature
 ```{r, eval=FALSE}
-m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
-temperature <- predict(m_temp, newdata = nd_all)
+# m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
+# temperature <- predict(m_temp, newdata = nd_all)
+# saveRDS(temperature, file = "data/predicted_temp_allyears.rds")
+
+temperature <- readRDS("data/predicted_temp_allyears.rds")
 temperature <- filter(temperature, ssid %in% model_ssid)
@@
 ```{r, eval=FALSE}
-temp <-plot_facet_map(temperature, "est", transform_col = no_trans, viridis_option = "D") + 
-  labs(fill="°C") +
+temp <- plot_facet_map(temperature, "est", transform_col = no_trans, viridis_option = "D") +
+  labs(fill = "°C") +
   ggtitle("Bottom temperature")
@@
 
-```{r, eval=FALSE}
-png(file = paste0("figs/exploratory-", spp, "-figs.png"),   # The directory you want to save the file in
+```{r}
+png(
+  file = paste0("figs/", spp, "/trends-", spp, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
   res = 600,
-units = 'in',
-width = 14, # The width of the plot in inches
-height = 12) # The height of the plot in inches
+  units = "in",
+  width = 11, # The width of the plot in inches
+  height = 8
+)
+gridExtra::grid.arrange(
+  # p_adult, p_imm,
+  biotrend_ad, biotrend_im,
+  nrow = 1
+)
+dev.off()
+```
+
+```{r}
+biomass_plots <- c(p_adult, p_imm)
+
+png(
+  file = paste0("figs/", spp, "/biomass-", spp, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  res = 600,
+  units = "in",
+  width = 16, # The width of the plot in inches
+  height = 8.5
+)
+gridExtra::grid.arrange(
+  grobs = biomass_plots,
+  nrow = 2
+)
+dev.off()
+```
+
+```{r}
+png(
+  file = paste0("figs/", spp, "/exploratory-", spp, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  res = 600,
+  units = "in",
+  width = 9.5, # The width of the plot in inches
+  height = 12.5
+) # The height of the plot in inches
 
 gridExtra::grid.arrange(
-    p_adult, p_imm, 
-    p_prop_imm, #current_biomass, 
+  p_adult_all, p_imm_all,
   biotrend_ad, biotrend_im,
-    temp,
-    nrow = 2)
+  # current_biomass,
+  p_prop_imm, temp,
+  nrow = 3
+)
 dev.off()

### 2019-06-21 analysis/VOCC/biomass-trends.Rmd
new covariates ready for the prediction grid; false start on integrating them with fishing_event data...

@@
 if (!exists("nd_all")) {
-  bath <- readRDS("data/bathymetry-data")
 
@@
   dummy_year <- c(2005, 2006)
-  dat <- bath$data %>% scale_survey_predictors()
+
+  # bath <- readRDS("data/bathymetry-data")
+  # dat <- bath$data %>% scale_survey_predictors()
+
+  newcov <- readRDS("data/new_covariates.rds")
+  dat <- newcov %>% gfranges::scale_survey_predictors()
 
@@
   biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
-  data <- biomass %>% gfranges:::scale_survey_predictors()
+  covars <- readRDS("event-covariates.rds")
+  data <- left_join(biomass, covars)
+
+  data <- data %>% gfranges:::scale_survey_predictors()
 

### 2019-06-21 analysis/VOCC/get-footprint.R

@@ -33,13 +33,6 @@ shp2raster <- function(shp, mask.raster, label, value, transform = FALSE, proj.f
 }
 
 
-
-# put prediction data in raster form
-nd_all <- readRDS("analysis/VOCC/data/nd_all_synoptic.rds")
-nd <- nd_all %>% filter(year %in% c(2005,2006)) %>%
-  mutate(X=X*1000, Y=Y*1000) # change utms to meters from Kms
-nd_raster <- raster::rasterFromXYZ( nd %>% dplyr::select(X, Y, depth), crs = proj)
-
 # retrieve trawl shapefile
 trawl_footprint <-readOGR(dsn="analysis/VOCC/data/trawl-footprint",layer="Trawl_footprint")
 plot(trawl_footprint)
@@ -48,15 +41,47 @@ plot(trawl_footprint)
 proj.from <- proj4string(trawl_footprint)
 proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
 
+# put prediction data in raster form
+nd_all <- readRDS("analysis/VOCC/data/nd_all_synoptic.rds")
+nd <- nd_all %>% filter(year %in% c(2005,2006)) %>%
+  mutate(X=X*1000, Y=Y*1000) # change utms to meters from Kms
+nd_raster <- raster::rasterFromXYZ( nd %>% dplyr::select(X, Y, depth), crs = proj.to)
 
 # extract all polygons and convert to a raster where cells
 # associated with raster have a value of 1 and everything else has a
 # value of 0.
 footprint <- shp2raster(shp = trawl_footprint,
-  mask.raster = nd_raster, label = "trawled", transform = TRUE, proj.from = proj.from, 
+  mask.raster = nd_raster, label = "trawled", 
+  transform = TRUE, proj.from = proj.from,
   proj.to = proj.to, value = 1)
 
+coords <- as.data.frame(raster::rasterToPoints(footprint))[, c("x", "y")]
+trawled <- as.data.frame(raster::rasterToPoints(footprint))[, 3]
+glimpse(trawled)
+trawl <- cbind(coords, trawled)
+trawl$X <- trawl$x/1000
+trawl$Y <- trawl$y/1000
 
+saveRDS(trawl, file = "analysis/VOCC/data/trawl-footprint.rds")
+
+
+
+# put event data in raster form
[…]

### 2019-06-24 analysis/VOCC/biomass-trends.Rmd
sdmTMB and predict working with added covariates

@@
 
-Region of interest
+Is a new spatial needed?
 ```{r}
+update_spatial_model <- TRUE
+priors <- FALSE
+# priors <- TRUE
+
+# Choose region to update model for:
+
 region <- "Both odd year surveys"
@@
 region <- "West Coast Haida Gwaii"
+
 # region <- "Queen Charlotte Sound"
@@
 # region <- "Both West Coast surveys"
-region <- "All synoptic surveys" # Probably not possible
 ```
 
-Is a new spatial needed?
+If not...
 ```{r}
-update_spatial_model <- TRUE
-# update_spatial_model <- FALSE
+update_spatial_model <- FALSE
 priors <- FALSE
 # priors <- TRUE
+region <- "All synoptic surveys" 
 ```
@@
 
-  # bath <- readRDS("data/bathymetry-data")
-  # dat <- bath$data %>% scale_survey_predictors()
-
-  newcov <- readRDS("data/new_covariates.rds")
-  dat <- newcov %>% gfranges::scale_survey_predictors()
+  bath <- readRDS("data/bathymetry-data")
+  dat <- bath$data %>% scale_survey_predictors()
 
@@
     mutate(year = as.numeric(year))
+
+  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky)
+  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))
+
   # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
 
-  saveRDS(nd_all, file = paste0("data/", spp, "/nd_all_synoptic.rds"))
+  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
 }
@@
   biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
-  covars <- readRDS("event-covariates.rds")
-  data <- left_join(biomass, covars)
+  covars <- readRDS("data/event-covariates.rds")
+  data <- dplyr::left_join(biomass,covars)
 
@@
     adult_biomass <- sdmTMB::sdmTMB(data,
-      adult_density ~ 0 + as.factor(year),
+      adult_density ~ 0 + as.factor(year) + trawled + muddy + rocky,
       time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
 
-    saveRDS(adult_biomass, file = paste0("data/", spp, "/model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, time1, ".rds"))
+    saveRDS(adult_biomass, file = paste0("data/", spp, "/model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, time1, "-newcovs.rds"))
 
@@
     imm_biomass <- sdmTMB::sdmTMB(data,
-      imm_density ~ 0 + as.factor(year),
+      imm_density ~ 0 + as.factor(year) + trawled + muddy + rocky,
       time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
 
-    saveRDS(imm_biomass, file = paste0("data/", spp, "/model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, time2, ".rds"))
+    saveRDS(imm_biomass, file = paste0("data/", spp, "/model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, time2, "-newcovs.rds"))
   } else {
@@
     total_biomass <- sdmTMB::sdmTMB(dat,
-      density ~ 0 + as.factor(year),
+      density ~ 0 + as.factor(year)+ trawled + muddy + rocky,
       time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
 
-    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, ".rds"))
+    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, "-newcovs.rds"))
   }
@@
     nd <- nd_all %>% filter(ssid %in% model_ssid) 
+    nd <- na.omit(nd)
     nd$year <- as.integer(nd$year)
@@
       nd <- nd_all %>% filter(ssid %in% model_ssid)
+      nd <- na.omit(nd)
       nd$year <- as.integer(nd$year)

### 2019-06-24 analysis/VOCC/combine-covariates.R

@@ -1,16 +1,17 @@
 library(dplyr)
 
 events <- readRDS("analysis/VOCC/data/events-trawled.rds")
-
+events2 <- events %>% dplyr::select(-catch_weight, -present, -density,-X, -Y, -depth, -akima_depth)
+glimpse(events)
 substrate <- readRDS("analysis/VOCC/data/events-w-substrate.rds")
 
 substrate2 <- as_tibble(substrate) 
 substrate2 <- substrate2 %>% dplyr::select(fishing_event_id, mixed, muddy, sandy, rocky)
 
-events <- as_tibble(events)
-events_w_covs <- left_join(events, substrate2)
+events2 <- as_tibble(events2)
+events_w_covs <- left_join(events2, substrate2)
 
-saveRDS(events_w_covs, file = "analysis/VOCC/data/events_covariates.rds")
+saveRDS(events_w_covs, file = "analysis/VOCC/data/event-covariates.rds")
 
 
 

### 2019-06-25 analysis/VOCC/biomass-trends.Rmd
add ability to select different covariates and include these selections in file names

@@
 species <- "Quillback Rockfish"
-species <- "Yelloweye Rockfish"
+# species <- "Yelloweye Rockfish"
+# priors <- TRUE
+covariates <- "+trawled+mixed+rocky"
+
 species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
@@
 species <- "Lingcod"
+```
 
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-# folder to hold figs for this species
-dir.create(file.path("figs", spp))
-dir.create(file.path("data", spp))
+# Choose covariates  
+```{r}
+covariates <- "+trawled+muddy+rocky+mixed"
+covariates <- "+trawled+mixed+rocky"
+covariates <- "+muddy+mixed+rocky"
+covariates <- "+mixed+rocky"
+covariates <- "+trawled+mixed"
 ```
@@
 priors <- FALSE
-# priors <- TRUE
+#priors <- TRUE
 
 # Choose region to update model for:
-
 region <- "Both odd year surveys"
@@
 
-# region <- "Queen Charlotte Sound"
-# region <- "Hecate Strait"
-# region <- "Both West Coast surveys"
 ```
@@
 region <- "All synoptic surveys"
+
+if (region == "All synoptic surveys") {
+  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  years <- NULL
+}
 ```
@@
 library(gfranges)
-```
 
-```{r}
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+
+# folder to hold figs for this species
+dir.create(file.path("figs", spp))
+dir.create(file.path("data", spp))
+fish <- readRDS(paste0("raw/bio-data-", spp, ""))
 
@@
 
+```
+
+```{r}
+
 # if(region == "Queen Charlotte Sound") {
@@
 # }
-
-# if(region == "Both West Coast surveys") {
-# survey <- c("SYN WCVI","SYN WCHG")
-# model_ssid <- c(4,16)
-# years <- NULL
-# }
-
-if (region == "All synoptic surveys") {
-  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
-  model_ssid <- c(1, 3, 4, 16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  years <- NULL
-}
 ```
@@
 
-  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky)
+  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed)
   nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))
@@
     # survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
-    survey = c("SYN HS", "SYN WCHG"),
-    #  survey = survey,
+    # survey = c("SYN HS", "SYN WCHG"),
+    survey = survey,
     years = years,
@@
 Run sdmTMB model (if not done previously).
-```{r eval=FALSE}
+```{r}
 if (update_spatial_model) {
@@
   if (any(names(data) == "adult_density")) {
+
+    adult_formula <- as.formula(paste("adult_density ~ 0 + as.factor(year)", covariates, ""))
+
     starttime1 <- Sys.time()
     adult_biomass <- sdmTMB::sdmTMB(data,
-      adult_density ~ 0 + as.factor(year) + trawled + muddy + rocky,
+      adult_formula,
       time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
 
-    saveRDS(adult_biomass, file = paste0("data/", spp, "/model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, time1, "-newcovs.rds"))
+    saveRDS(adult_biomass, file = paste0("data/", spp, "/mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time1, ".rds"))
+
+    imm_formula <- as.formula(paste("imm_density ~ 0 + as.factor(year)", covariates, ""))
 
@@
     imm_biomass <- sdmTMB::sdmTMB(data, 
-      imm_density ~ 0 + as.factor(year) + trawled + muddy + rocky,
+      imm_formula,
       time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
 
-    saveRDS(imm_biomass, file = paste0("data/", spp, "/model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, time2, "-newcovs.rds"))
+    saveRDS(imm_biomass, file = paste0("data/", spp, "/mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time2, ".rds"))
   } else {
     starttime3 <- Sys.time()
+    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
+
     total_biomass <- sdmTMB::sdmTMB(dat,
-      density ~ 0 + as.factor(year)+ trawled + muddy + rocky,
+      dens_formula,
       time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
       ar1_fields = TRUE,
-      include_spatial = TRUE, ,
+      include_spatial = TRUE,
       enable_priors = priors,
@@
 
-    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, "-newcovs.rds"))
+    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
   }
@@
   if (region != "All synoptic surveys") {
-    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
-    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-imm-biomass-priors-", priors, ".rds")))
-    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
+    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
+    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
+    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
   } else {
@@
     try ({
-      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-priors-", priors, ".rds"))
-      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-priors-", priors, ".rds"))
-      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-priors-", priors, ".rds"))
+      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-prior-", priors, ".rds"))
+      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-prior-", priors, ".rds"))
+      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-prior-", priors, ".rds"))
       adult_predictions <- rbind(pred1n3, pred4, pred16) 
@@
     try ({
-      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-imm-biomass-priors-", priors, ".rds"))
-      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-imm-biomass-priors-", priors, ".rds"))
-      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-imm-biomass-priors-", priors, ".rds"))
+      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-imm-biomass-prior-", priors, ".rds"))
+      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-imm-biomass-prior-", priors, ".rds"))
+      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-imm-biomass-prior-", priors, ".rds"))
       imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm) 
@@
     try ({
-      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-priors-", priors, ".rds"))
-      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-priors-", priors, ".rds"))
-      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-priors-", priors, ".rds"))
+      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-prior-", priors, ".rds"))
+      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-prior-", priors, ".rds"))
+      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-prior-", priors, ".rds"))
       predictions <- rbind(pred1n3_all, pred4_all, pred16_all) # ,
@@
 if (!exists("adult_predictions")) {
+    fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+
   if (length(unique(fish$maturity_code)) > 2) {
+
+    adult_biomass <- readRDS(list.files(
+      path = paste0("data/", spp, "/"),
+      #pattern = paste0("mod-mat-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
+      pattern = paste0("mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
+      all.files = F, full.names = T
+    ))
+
     nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-    nd <- nd_all %>% filter(ssid %in% model_ssid) 
+    nd <- nd_all %>% filter(ssid %in% model_ssid) %>% filter(year %in% unique(adult_biomass$data$year))
     nd <- na.omit(nd)
@@
 
-    adult_biomass <- readRDS(list.files(
-      path = paste0("data/", spp, "/"),
-      pattern = paste0("model-mature-biomass-", spp, "-", ssid_string, "-priors-", priors, "-*"),
-      all.files = F, full.names = T
-    ))
-
     adult_predictions <- predict(adult_biomass, newdata = nd)
     saveRDS(adult_predictions,
-      file = paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
+      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-mature-biomass-prior-", priors, ".rds")
     )
@@
       path = paste0("data/", spp, "/"),
-      pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-priors-", priors, "-*"),
+      #pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
+      pattern = paste0("mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
       all.files = F, full.names = T
@@
     saveRDS(imm_predictions,
-      file = paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-imm-biomass-priors-", priors, ".rds")
+      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-imm-biomass-prior-", priors, ".rds")
     )
@@
 
-      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, "-", ssid_string, "-priors-", priors, ".rds"))
+      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
       nd$year <- as.integer(nd$year)
       predictions <- predict(total_biomass, newdata = nd)
-      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"))
+      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
     }
@@
     try(
-      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-priors-", priors, ".rds")),
-      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-priors-", priors, ".rds")),
-      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-priors-", priors, ".rds")),
+      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-mature-biomass-prior-", priors, ".rds")),
+      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-mature-biomass-prior-", priors, ".rds")),
+      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-mature-biomass-prior-", priors, ".rds")),
 
@@
     try(
-      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-imm-biomass-priors-", priors, ".rds")),
-      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-imm-biomass-priors-", priors, ".rds")),
-      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-imm-biomass-priors-", priors, ".rds")),
+      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-imm-biomass-prior-", priors, ".rds")),
+      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-imm-biomass-prior-", priors, ".rds")),
+      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-imm-biomass-prior-", priors, ".rds")),
 
@@
     try(
-      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-priors-", priors, ".rds")),
-      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-priors-", priors, ".rds")),
-      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-priors-", priors, ".rds")),
+      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-prior-", priors, ".rds")),
+      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-prior-", priors, ".rds")),
+      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-prior-", priors, ".rds")),
 
@@
   max_raster <- quantile(adult_predictions$est_exp, .9999)
-  max_adult <- round(max(adult_predictions$est_exp))
+  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)
 } else {
@@
   p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
-    transform_col = fourth_root_power,
     raster_limits = c(0, max_raster),
-    legend_position = "none"
+    legend_position = "none",
+    transform_col = fourth_root_power
   ) +
@@
 saveRDS(out_mature,
-  file = paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
+  file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
 )
@@
   saveRDS(out_total,
-    file = paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
+    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
   )
@@
 ```{r, error=TRUE}
-try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
+try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
 
@@
 } else {
-  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
+  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
 
@@
   saveRDS(out_imm,
-    file = paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
+    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
   )
@@
 ```{r}
-try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
+try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
 
@@
 png(
-  file = paste0("figs/", spp, "/trends-", spp, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/trends-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/biomass-", spp, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/biomass-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/exploratory-", spp, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/exploratory-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
   res = 600,

### 2019-06-26 analysis/VOCC/biomass-trends.Rmd
make "trawled" time-varying

@@
 # priors <- TRUE
-covariates <- "+trawled+mixed+rocky"
 
@@
 ```{r}
+covariates <- "+muddy+mixed+rocky"
+covariates <- "+mixed+rocky"
+
 covariates <- "+trawled+muddy+rocky+mixed"
 covariates <- "+trawled+mixed+rocky"
-covariates <- "+muddy+mixed+rocky"
-covariates <- "+mixed+rocky"
 covariates <- "+trawled+mixed"
@@
 
+covs <- gsub("\\+", "-", covariates)
 spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
@@
 
+  # scale predictors before filtering to ensure mean and SD are global
   data <- data %>% gfranges:::scale_survey_predictors()
@@
 
-  if (region == "All synoptic surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 1200)
-  }
+  # if (region == "All synoptic surveys") {
+  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 1200)
+  # }
 
   if (region == "Both odd year surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
   }
@@
   if (region == "West Coast Vancouver Island") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
   }
@@
   if (region == "West Coast Haida Gwaii") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 300)
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
   }
@@
       adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
       time = "year", spde = spde,
@@
 
-    saveRDS(adult_biomass, file = paste0("data/", spp, "/mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time1, ".rds"))
+    #saveRDS(adult_biomass, 
+    # file = paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, time1, ".rds"))
+    saveRDS(adult_biomass, 
+      file = paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
 
@@
       imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
       time = "year", spde = spde,
@@
 
-    saveRDS(imm_biomass, file = paste0("data/", spp, "/mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time2, ".rds"))
+    #saveRDS(imm_biomass, 
+    # file = paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, time2, ".rds"))
+    saveRDS(imm_biomass, 
+     file = paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
   } else {
@@
 
-    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
+    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
   }
 }
+```
+
+```{r eval=FALSE}
+k <- length(adult_biomass$model$par)
+nll <- adult_biomass$model$objective
+AIC <- 2*k+2*nll
+AIC #(-5500.438)
 ```
@@
   if (region != "All synoptic surveys") {
-    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
-    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
-    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
+    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
+    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
+    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
   } else {
@@
 
-    adult_biomass <- readRDS(list.files(
-      path = paste0("data/", spp, "/"),
-      #pattern = paste0("mod-mat-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
-      pattern = paste0("mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
-      all.files = F, full.names = T
-    ))
+    # adult_biomass <- readRDS(list.files(
+    #   path = paste0("data/", spp, "/"),
+    #   #pattern = paste0("mod-mat-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
+    #   pattern = paste0("mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, "-*"),
+    #   all.files = F, full.names = T
+    # ))
+
+     adult_biomass <- readRDS(paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
 
@@
     saveRDS(adult_predictions,
-      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-mature-biomass-prior-", priors, ".rds")
+      file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-mature-biomass-prior-", priors, ".rds")
     )
 
-    imm_biomass <- readRDS(list.files(
-      path = paste0("data/", spp, "/"),
-      #pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
-      pattern = paste0("mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
-      all.files = F, full.names = T
-    ))
+    # imm_biomass <- readRDS(list.files(
+    #   path = paste0("data/", spp, "/"),
+    #   #pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
+    #   pattern = paste0("mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, "-*"),
+    #   all.files = F, full.names = T
+    # ))
+
+    imm_biomass <- readRDS(paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
     imm_predictions <- predict(imm_biomass, newdata = nd)
     saveRDS(imm_predictions,
-      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-imm-biomass-prior-", priors, ".rds")
+      file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-imm-biomass-prior-", priors, ".rds")
     )
@@
 
-      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
+      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
       nd$year <- as.integer(nd$year)
       predictions <- predict(total_biomass, newdata = nd)
-      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
+      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
     }
@@
     try(
-      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-mature-biomass-prior-", priors, ".rds")),
-      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-mature-biomass-prior-", priors, ".rds")),
-      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-mature-biomass-prior-", priors, ".rds")),
+      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds")),
+      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds")),
+      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds")),
 
@@
     try(
-      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-imm-biomass-prior-", priors, ".rds")),
-      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-imm-biomass-prior-", priors, ".rds")),
-      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-imm-biomass-prior-", priors, ".rds")),
+      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds")),
+      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds")),
+      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds")),
 
@@
 saveRDS(out_mature,
-  file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
+  file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
 )
@@
   saveRDS(out_total,
-    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
+    file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
   )
@@
 
-Optional plot of current total biomass 
-```{r eval=FALSE}
-if (exists("out_total")) {
-  current_biomass <- plot_vocc(out_total,
-    fill_col = "var_1_e",
-    fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
-    raster_alpha = 1,
-    viridis_option = "C",
-    vec_aes = NULL,
-    transform_col = fourth_root_power
-  )
-  current_biomass <- current_biomass +
-    ggtitle(paste0("", species, " total biomass"))
-
-  current_biomass
-}
+Attempting to plot of current mature biomass with trawl footprint...
+```{r }
+# if (exists("out_mature")) {
+#   
+#   trawl_footprint <- sf::st_read(dsn="data/trawl-footprint",layer="Trawl_footprint")
+#   proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
+# 
+#   trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)
+# 
+#   
+#   current_biomass <- plot_vocc(out_mature,
+#     fill_col = "var_1_e",
+#     fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
+#     raster_alpha = 1,
+#     viridis_option = "C",
+#     vec_aes = NULL,
+#     transform_col = fourth_root_power
+#   )
+#   current_biomass <- current_biomass +
+#   #ggplot(trawl_footprint) + #, aes(x=x/1000, y=y/1000)
+#     geom_sf(data= trawl_footprint, aes(geometry=geometry/1000), inherit.aes = FALSE) #+
+#     #coord_sf(datum = NULL) 
+#   
+#   +
+#     ggtitle(paste0("", species, " total biomass"))  
+# 
+#   current_biomass
+# }
 ```
@@
 ```{r, error=TRUE}
-try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
+try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
 
@@
 } else {
-  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
+  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
 
@@
   saveRDS(out_imm,
-    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
+    file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
   )
@@
 ```{r}
-try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
+try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
 
@@
 png(
-  file = paste0("figs/", spp, "/trends-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/biomass-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/exploratory-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
   res = 600,

### 2019-07-11 analysis/tmb-sensor-explore/3-fit-do.Rmd
draft models of DO

---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Fit spatial temporal models to bottom temperature

```{r setup}
library(tidyverse)
library(gfplot)
library(sdmTMB)
library(gfranges)

#library(future)
#library(dplyr)

# plan(multiprocess)
```

Load data
```{r}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/all-sensor-data-processed.rds"))
# glimpse(all_sensor)

d_trawl <- all_sensor %>%
  # filter missing do data 
  dplyr::filter(!is.na(do_mlpl)) %>%
  # remove 2007 data because clearly faulty or scaled differently
  dplyr::filter(year>2007) %>%
  # add day of year variable
  mutate(DOY = lubridate::yday(fe_event_start), temp = temperature_c, 
    raw_depth = depth, depth = log(depth), log_do = log(do_mlpl)) %>%
  # scale potential covariates across all surveys
  scale_predictors(predictors = c(quo(depth), quo(DOY), quo(temp)))

d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]

d_trawl <- d_trawl %>% mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
  filter(exclude != 1)

glimpse(d_trawl)
```

Model diagnostics functions
```{r}
get_diag <- function(m, response = "log_do", variable = "depth_scaled") {
  predictions <- predict(m)
  predictions$residuals <- residuals(m)

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point() +
      coord_fixed()
  }

  g <- plot_map(predictions, "est") +
    scale_colour_viridis_c() +
    ggtitle("Prediction (fixed effects + all random effects)")
  print(g)

  g <- plot_map(predictions, "est_non_rf") +
    ggtitle("Prediction (fixed effects only)") +
    scale_colour_viridis_c()
  print(g)

  g <- plot_map(predictions, "est_rf") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- ggplot(predictions, aes_string("est", response)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()
  print(g)

  g <- ggplot(predictions, aes(est, residuals)) +
    geom_point() +
    geom_smooth()
  print(g)

  g <- ggplot(predictions, aes(predictions[[variable]], residuals)) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point() +
    coord_fixed() +
    scale_color_gradient2()
  print(g)

  aic <- function(m) {
    k <- length(m$model$par)
    nll <- m$model$objective
    2 * k - 2 * (-nll)
  }

  print("R^2:")
  r2 <- cor(predictions$est, predictions[[response]])^2
  print(r2)

  print("")
  print("AIC:")
  print(aic(m))
}
```

```{r}
ggplot(d_trawl, aes(X,Y, colour = do_mlpl)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(X,Y, colour = temp)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(depth, do_mlpl, colour = DOY, size = sqrt(do_change))) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()

```

```{r}
ggplot(d_trawl, aes(raw_depth, temp, colour = DOY, size = sqrt(temp_range))) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(DOY, -depth , colour = do_mlpl)) + geom_point() + facet_wrap(~year) +
  scale_color_viridis_c() + ylab("-log(depth)")
```

```{r}
hist(d_trawl$do_mlpl)
# log do_mlpl to allow gaussian 
hist(d_trawl$log_do)
```

```{r}
# # filter surveys if desired
# d_trawl <- d_trawl %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
  # filter(depth_m_max>10) %>%

```

```{r}

d_trawl1 <- d_trawl #%>% filter(ssid != 16) %>% filter(ssid != 4)    
spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 800)
plot_spde(spde)
```

```{r}
    starttime1 <- Sys.time()

    temp_model_all <- sdmTMB::sdmTMB(d_trawl1,
      temp_scaled ~ 0 + as.factor(year),
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = gaussian(link = "identity"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = FALSE,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

#saveRDS(do_model, file = (here::here("analysis/tmb-sensor-explore/models/model-do-1n3.rds")))
time1
temp_model_all
```

```{r}
get_diag(temp_model_all, response = "temp_scaled", variable = "depth_scaled")
```

Add temperature predictions to newdata grid for DO predictions
```{r}

temperature <- readRDS(here::here("analysis/VOCC/data/predicted_temp_allyears.rds"))
years <- unique(d_trawl1$year) 
temperature <- filter(temperature, year %in% years)
unique(temperature$year)

temperature$temp <- temperature$est
#temperature$temp_scaled2 <- temperature$est^2
temperature$raw_depth <- temperature$depth
temperature$depth <- log(temperature$depth)

newdata <- temperature %>% 
  select(-est, - est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st, -temperature_c) %>%
  scale_predictors(predictors = c(quo(depth),quo(temp))) %>% 
  mutate(depth = raw_depth)

glimpse(newdata)

```

```{r}
    starttime1 <- Sys.time()

    do_model_depth2 <- sdmTMB::sdmTMB(d_trawl1,
      log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ depth_scaled3,
      time = "year", spde = spde,
      family = gaussian(link = "identity"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = FALSE,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

saveRDS(do_model_depth2, file = (here::here("analysis/tmb-sensor-explore/models/model-do-depth2.rds")))
time1
do_model_depth2
```

```{r}
do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY.rds"))

get_diag(do_model_DOY,response = "log_do", variable = "log_depth_scaled")

get_diag(do_model_DOY,response = "log_do", variable = "temp_scaled")
```

```{r}
do_model_depth3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth3.rds"))

get_diag(do_model_depth3,response = "log_do")

p_nd <- predict(do_model_depth3, newdata = newdata)

glimpse(p_nd)
```

```{r}
plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

p_nd$do_est <- exp(p_nd$est)
hist(p_nd$do_est)
max(p_nd$do_est)

plot_map_raster(p_nd, "do_est") +
  scale_fill_viridis_c(trans = "log",) +
#  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
do_model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-1.rds"))
get_diag(do_model1)
```

```{r}
plot_map_raster(p_nd, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
plot_map_raster(p_nd, "est_non_rf") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

```{r, eval=FALSE}
#glimpse(p_nd)
plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
```

```{r}
plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
```

```{r}
#do_cog <- get_cog(p_nd)

#temp_cog %>%
#  reshape2::dcast(year ~ coord, value.var = "est") %>%
#  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
#  scale_color_viridis_c()
```

```{r}
#temp_cog %>%
#  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
#  geom_line() +
#  geom_ribbon(alpha = 0.2) +
#  facet_wrap(~coord, scales = "free_y")
```

### 2019-07-12 analysis/VOCC/dist-vocc-temp-do-4.Rmd
VOCC using temp and DO together for WCVI

---
title: "Distanced-based VOCC analysis for WCVI"
author: "Philina English"
date: '2019-07-11'
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfranges)
```

Load predicted values
```{r}
pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
```

Define subset of the spatial range 
```{r eval=FALSE}
predicted <- pred_temp_do %>% filter(ssid==4)

predicted <- list(predicted, predicted)

do <-plot_facet_map(predicted, "do_est", transform_col = no_trans) + 
  labs(fill="ml/L") +
  ggtitle("Bottom DO")
print(do)
```

Biannual changes starting 2005 and ending 2017
```{r}
out1 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"), 
  ssid = c(4),
  start_time = 2008,
  end_time = 2010,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
  thresholds = c(1,1) # vector of plus/minus threshold(s) to define climate match.
)
out1$do_est.units_per_decade <- out1$do_est.units_per_decade/5
out1$temp.units_per_decade <- out1$temp.units_per_decade/5

gvocc1a <- plot_vocc(out1,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
 # viridis_option = "C",
  #viridis_dir = -1#,
  #transform_col = fourth_root_power,
 #raster_limits = c(0, 6)
)

gvocc1b <- plot_vocc(out1,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
 # transform_col = fourth_root_power,
  #raster_limits = c(5, 13)
)

gvocc1 <- gridExtra::grid.arrange(gvocc1b, gvocc1a, nrow = 1, top= "VOCC vectors to maintain <1°C or 1 ml/L change between 2008 and 2010") 
```

```{r}
out2 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"), 
  ssid = c(4),
  start_time = 2010,
  end_time = 2012,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
  thresholds = c(1,1) # vector of plus/minus threshold(s) to define climate match.
)
out2$do_est.units_per_decade <- out2$do_est.units_per_decade/5
out2$temp.units_per_decade <- out2$temp.units_per_decade/5

gvocc2a <- plot_vocc(out2,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc2b <- plot_vocc(out2,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc2 <- gridExtra::grid.arrange(gvocc2b, gvocc2a, nrow = 1, top= "VOCC vectors to maintain <1°C or 1 ml/L change between 2010 and 2012") 
```

`
```{r}
out3 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"), 
  ssid = c(4),
  start_time = 2012,
  end_time = 2014,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
  thresholds = c(1,1) # vector of plus/minus threshold(s) to define climate match.
)
out3$do_est.units_per_decade <- out3$do_est.units_per_decade/5
out3$temp.units_per_decade <- out3$temp.units_per_decade/5

gvocc3a <- plot_vocc(out3,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc3b <- plot_vocc(out3,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc3 <- gridExtra::grid.arrange(gvocc3b, gvocc3a, nrow = 1, top= "VOCC vectors to maintain <1°C or 1 ml/L change between 2012 and 2014") 
```

```{r}
out4 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"), 
  ssid = c(4),
  start_time = 2014,
  end_time = 2016,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
  thresholds = c(1,1) # vector of plus/minus threshold(s) to define climate match.
)
out4$do_est.units_per_decade <- out4$do_est.units_per_decade/5
out4$temp.units_per_decade <- out4$temp.units_per_decade/5

gvocc4a <- plot_vocc(out4,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc4b <- plot_vocc(out4,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc4 <- gridExtra::grid.arrange(gvocc4b, gvocc4a, nrow = 1, top= "VOCC vectors to maintain <1°C or 1 ml/L change between 2014 and 2016") 
```

```{r}
out5 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"), 
  ssid = c(4),
  start_time = 2016,
  end_time = 2018,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
  thresholds = c(1,1) # vector of plus/minus threshold(s) to define climate match.
)
out5$do_est.units_per_decade <- out5$do_est.units_per_decade/5
out5$temp.units_per_decade <- out5$temp.units_per_decade/5

gvocc5a <- plot_vocc(out5,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc5b <- plot_vocc(out5,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE#,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc5 <- gridExtra::grid.arrange(gvocc5b, gvocc5a, nrow = 1, top= "VOCC vectors to maintain <1°C or 1 ml/L change between 2016 and 2018") 
```

```{r}
png(file = "biannual_temp-do_change-4.png",   # The directory you want to save the file in
res = 600,
units = 'in',
width = 10, # The width of the plot in inches
height = 8) # The height of the plot in inches

gridExtra::grid.arrange(gvocc1, gvocc2, gvocc3, gvocc4, gvocc5, nrow = 2, top = grid::textGrob("VOCC vectors for <1 ml/L biannual changes in bottom DO"))#,gp=gpar(fontsize=20,font=3)) 
dev.off()
```

Cumulative change between last two decades 
VOCC for 1 ml/L change and 1 degree change
```{r}
out0.5d <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"), 
  ssid = c(4),
  start_time = 2008,
  skip_time = 2016,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 6,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2),
  thresholds = c(0.5,0.5) # vector of plus/minus threshold(s) to define climate match.
)
```

```{r}
gvocc <- plot_vocc(out0.5d,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 100,
  fill_col = "1_e",
  fill_label = "mean DO",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  viridis_option = "C",
  viridis_dir = -1,
  transform_col = no_trans#,
  )
gvocc <- gvocc + ggtitle("VOCC vectors for <0.5 ml/L change (excluding 2016)")
gvocc
```

```{r}
gtrend1 <- plot_vocc(out0.5d,
  fill_col = "temp.units_per_decade",
  fill_label = "°C per decade",
  #high_fill = "Steel Blue 4",
  #low_fill = "Red 3",
  raster_alpha = 1,
  vec_aes = NULL,  
  viridis_option = "D",
  transform_col = no_trans
)
gtrend1 <- gtrend1 + ggtitle("Temperature trend 2008-2018 (excluding 2016)")
gtrend1
```

```{r}
gtrend <- plot_vocc(out0.5d,
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L per decade",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = no_trans
)
gtrend <- gtrend + ggtitle("Bottom DO trend 2008-2018 (excluding 2016)")
gtrend
```

```{r}
png(file = "temp-do-2008-2018-4.png",   # The directory you want to save the file in
res = 600,
units = 'in',
width = 10, # The width of the plot in inches
height = 7) # The height of the plot in inches

gridExtra::grid.arrange( gvocc, gtrend1, gtrend, nrow = 1)
dev.off()
```

### 2019-07-13 analysis/VOCC/biomass-trends.Rmd
add any_rock substrate variable that combines rocky and mixed

@@
 covariates <- "+mixed+rocky"
+covariates <- "+muddy+any_rock"
 
-covariates <- "+trawled+muddy+rocky+mixed"
-covariates <- "+trawled+mixed+rocky"
-covariates <- "+trawled+mixed"
+# covariates <- "+trawled+muddy+rocky+mixed"
+# covariates <- "+trawled+mixed+rocky"
+# covariates <- "+trawled+mixed"
+
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
-try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
+#try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
 
@@
 
-  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed)
+  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
   nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))

### 2019-07-13 analysis/VOCC/combine-covariates.R

@@ -10,8 +10,10 @@ substrate2 <- substrate2 %>% dplyr::select(fishing_event_id, mixed, muddy, sandy
 
 events2 <- as_tibble(events2)
 events_w_covs <- left_join(events2, substrate2)
+events_w_covs$any_rock <- events_w_covs$rocky + events_w_covs$mixed
 
 saveRDS(events_w_covs, file = "analysis/VOCC/data/event-covariates.rds")
+events_w_covs <- readRDS("analysis/VOCC/data/event-covariates.rds")
 
 
 
@@ -57,12 +59,12 @@ glimpse(nd_test)
 
 #View(nd_test)
 
-
 nd_combind <- nd_test %>% select(-x, -y, -substrate_100m_1step)
+nd_combind$any_rock <- nd_combind$rocky + nd_combind$mixed
 
-
+glimpse(nd_combind)
 saveRDS(nd_combind, file = "analysis/VOCC/data/new_covariates.rds")
-
+#nd_combind <- readRDS("analysis/VOCC/data/new_covariates.rds")
 
 
 # THIS SHOULD probably be done by make grids so that only years with samples are created...

### 2019-07-13 analysis/tmb-sensor-explore/3-fit-do.Rmd
more plots and model configs for DO

@@
   g <- plot_map(predictions, "est_rf") +
+    ggtitle("All random effects only") +
+    scale_colour_gradient2()
+  print(g)
+
+  g <- plot_map(predictions, "omega_s") +
     ggtitle("Spatial random effects only") +
+    scale_colour_gradient2()
+ print(g)
+
+   g <- plot_map(predictions, "epsilon_st") +
+    ggtitle("Spatiotemporal random effects only") +
+    facet_wrap(~year) +
     scale_colour_gradient2()
@@
 
-  g <- ggplot(predictions, aes(predictions[[variable]], residuals)) +
+  g <- ggplot(predictions, aes_string(variable, "residuals")) +
     geom_point() +
@@
     coord_fixed() +
-    scale_color_gradient2()
+    scale_color_gradient2()+
+    facet_wrap(~year)
   print(g)
@@
 ```{r}
+ggplot(d_trawl, aes(temp, log_do, colour = DOY, size = sqrt(temp_range))) +
+  geom_point() +
+  facet_wrap(~year) +
+  scale_color_viridis_c()
+```
+
+```{r}
 ggplot(d_trawl, aes(DOY, -depth , colour = do_mlpl)) + geom_point() + facet_wrap(~year) +
@@
 
-```{r}
+Test temp model on this subset of data
+DOY doesn't improve model
+```{r eval=FALSE}
     starttime1 <- Sys.time()
 
-    temp_model_all <- sdmTMB::sdmTMB(d_trawl1,
-      temp_scaled ~ 0 + as.factor(year),
+    temp_model_DOY <- sdmTMB::sdmTMB(d_trawl1,
+      temp_scaled ~ 0 + as.factor(year), # + DOY_scaled,
       time_varying = ~ 0 + depth_scaled + depth_scaled2,
@@
 
-#saveRDS(do_model, file = (here::here("analysis/tmb-sensor-explore/models/model-do-1n3.rds")))
-time1
+saveRDS(temp_model_all, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-2008-on.rds")))
+
+# saveRDS(temp_model_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-2008-on-DOY.rds")))
+# time1
+# temp_model_DOY
 temp_model_all
@@
 
-```{r}
+```{r eval=FALSE}
 get_diag(temp_model_all, response = "temp_scaled", variable = "depth_scaled")
@@
 ```{r}
-
 temperature <- readRDS(here::here("analysis/VOCC/data/predicted_temp_allyears.rds"))
@@
 temperature$depth <- log(temperature$depth)
+temperature$DOY_scaled <- 0
 
@@
 glimpse(newdata)
-
 ```
 
-```{r}
+Current best model of DO
+```{r eval=FALSE}
     starttime1 <- Sys.time()
 
-    do_model_depth2 <- sdmTMB::sdmTMB(d_trawl1,
-      log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ depth_scaled3,
+    do_model_re_DOY <- sdmTMB::sdmTMB(d_trawl1,
+      log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 ,
+      time_varying = ~ 0 + DOY_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
       time = "year", spde = spde,
@@
 
-saveRDS(do_model_depth2, file = (here::here("analysis/tmb-sensor-explore/models/model-do-depth2.rds")))
+saveRDS(do_model_re_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds")))
 time1
-do_model_depth2
+do_model_re_DOY
+```
+
+```{r}
+do_model_re_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds"))
+do_model_re_DOY
+```
+
+```{r}
+get_diag(do_model_all_fe, response = "log_do", variable = "depth_scaled")
+
 ```
@@
 do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY.rds"))
+do_model_DOY
 
-get_diag(do_model_DOY,response = "log_do", variable = "log_depth_scaled")
+get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")
 
-get_diag(do_model_DOY,response = "log_do", variable = "temp_scaled")
+# get_diag(do_model_DOY, response = "log_do", variable = "temp_scaled")
 ```
 
+Other models tried:
+
+```{r eval=FALSE}
+do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-DOY.rds"))
+do_model_DOY
+#p_nd <- predict(do_model_DOY, newdata = newdata)
+get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")
+
+```
+
+```{r eval=FALSE}
+do_model_depth2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth2.rds"))
+do_model_depth2
+get_diag(do_model_depth2,response = "log_do", variable = "depth_scaled")
+#p_nd <- predict(do_model_depth2, newdata = newdata)
+```
+
+```{r eval = FALSE}
+do_model_depth3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth3.rds"))
+get_diag(do_model_depth3,response = "log_do")
+#p_nd <- predict(do_model_depth3, newdata = newdata)
+```
+
+```{r eval=FALSE}
+do_model_all <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all.rds"))
+#p_nd <- predict(do_model_all, newdata = newdata)
+do_model_all
+get_diag(do_model_all, response = "log_do", variable = "depth_scaled")
+```
+
+Predict DO for whole grid using quadratic of predicted temperature values and mean DOY
 ```{r}
-do_model_depth3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth3.rds"))
+p_nd <- predict(do_model_DOY, newdata = newdata)
+p_nd$do_est <- exp(p_nd$est)
 
-get_diag(do_model_depth3,response = "log_do")
+p_nd <- p_nd %>% filter(depth>15)
+hist(p_nd$do_est)
 
-p_nd <- predict(do_model_depth3, newdata = newdata)
+max(p_nd$do_est)
+max(d_trawl1$do_mlpl)
+glimpse(p_nd)
+min(p_nd$raw_depth)
+min(d_trawl1$raw_depth)
+p_nd[p_nd$do_est>8, ]
 
-glimpse(p_nd)
+p_nd <- p_nd %>% filter(depth>15)
+p_nd_filtered[p_nd_filtered$do_est>8, ]
+
+saveRDS(p_nd, file = (here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
 ```
@@
     geom_raster() +
-    facet_wrap(~year) +
     coord_fixed()
@@
 
-p_nd$do_est <- exp(p_nd$est)
-hist(p_nd$do_est)
-max(p_nd$do_est)
+plot_map_raster(p_nd, "do_est") +
+  scale_fill_viridis_c(trans = "log") +
+  facet_wrap(~year) +
+  ggtitle("Prediction (fixed effects + all random effects)")
 
-plot_map_raster(p_nd, "do_est") +
-  scale_fill_viridis_c(trans = "log",) +
-#  scale_fill_viridis_c(trans = "sqrt", option = "C") +
-  ggtitle("Prediction (fixed effects + all random effects)")
-```
-
-```{r}
-do_model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-1.rds"))
-get_diag(do_model1)
-```
-
-```{r}
-plot_map_raster(p_nd, "est") +
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
-  ggtitle("Prediction (fixed effects + all random effects)")
 ```
@@
   ggtitle("Prediction (fixed effects only)") +
-  scale_fill_viridis_c(trans = "sqrt")
+  facet_wrap(~year) +
+  scale_fill_viridis_c()
+```
+
+```{r}
+plot_map_raster(p_nd, "est_rf") +
+  ggtitle("Prediction (all random effects only)") +
+  facet_wrap(~year) +
+  scale_fill_viridis_c()
 ```
@@
   ggtitle("Spatial random effects only") +
-  facet_wrap(~year) +
-  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
+   scale_fill_gradient2( high = "Steel Blue 4",
+  low = "Red 3")+
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
 ```
@@
   facet_wrap(~year) +
-  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
+  scale_fill_gradient2( high = "Steel Blue 4",
+  low = "Red 3")
+```
+
+```{r}
+p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
+plot_map_raster(p_nd1, "epsilon_st") +
+  ggtitle("Spatiotemporal random effects only") +
+  facet_wrap(~year) +
+  scale_fill_gradient2( high = "Steel Blue 4",
+  low = "Red 3", limits = c(-0.4,0.5)) +
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+```
+```{r}
+p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
+plot_map_raster(p_nd4, "epsilon_st") +
+  ggtitle("Spatiotemporal random effects only") +
+  facet_wrap(~year) +
+  scale_fill_gradient2( high = "Steel Blue 4",
+  low = "Red 3", limits = c(-0.4,0.5)) +
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+```
+
+```{r}
+p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
+plot_map_raster(p_nd16, "epsilon_st") +
+  ggtitle("Spatiotemporal random effects only") +
+  facet_wrap(~year) +
+  scale_fill_gradient2( high = "Steel Blue 4",
+  low = "Red 3", limits = c(-0.4,0.5)) +
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
 ```

### 2019-07-17 analysis/tmb-sensor-explore/3-fit-do.Rmd
minor organizational edits

@@
 library(gfranges)
-
-#library(future)
-#library(dplyr)
-
-# plan(multiprocess)
 ```
@@
   # filter(depth_m_max>10) %>% 
-
 ```
@@
 
-    do_model_re_DOY <- sdmTMB::sdmTMB(d_trawl1,
-      log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 ,
-      time_varying = ~ 0 + DOY_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
+    do_model_all_DOY <- sdmTMB::sdmTMB(d_trawl1,
+      log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + DOY_scaled,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
       time = "year", spde = spde,
@@
 
-saveRDS(do_model_re_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds")))
+saveRDS(do_model_all_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY2.rds")))
 time1
-do_model_re_DOY
-```
-
-```{r}
-do_model_re_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds"))
-do_model_re_DOY
-```
-
-```{r}
-get_diag(do_model_all_fe, response = "log_do", variable = "depth_scaled")
-
+do_model_all_DOY
 ```
@@
 Other models tried:
+
+```{r eval=FALSE}
+#with both DOY and depth as time-varying random effects
+do_model_re_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds"))
+do_model_re_DOY
+get_diag(do_model_re_DOY, response = "log_do", variable = "depth_scaled")
+```
 
@@
   ggtitle("Prediction (fixed effects + all random effects)")
-
 ```

### 2019-07-17 analysis/VOCC/vocc-test.R

@@ -1,11 +1,11 @@
 start_data <- list(
-  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4), do_est = c(1.50, 1.01, 1.01, 1.01)),
-  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4),   temp = c(1.01, 1.01, 1.01, 1.01))
+  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4),   do = c(1.50, 1.01, 1.01, 1.01)),
+  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4), temp = c(1.01, 1.01, 1.01, 1.01))
 )
 
 end_data <- list(
-  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4), do_est = c(0.55, 1.00, 1.00, 10.0)),
-  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4),   temp = c(1.00, 1.45, 1.00, 1.00))
+  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4),   do = c(0.55, 1.00, 1.00, 10.0)),
+  data.frame(x = c(1, 2, 3, 4), y = c(1, 2, 3, 4), temp = c(1.00, 1.45, 1.00, 1.00))
 )
 
 
@@ -15,8 +15,8 @@ dist_based_vocc(
   end_data = end_data,
   x = "x",
   y = "y",
-  variable_names = c("do_est", "temp"), # what the layer within each element is called
-  plus_minus = c(0.5, 0.5), # vector of plus/minus threshold(s) to define climate match 
+  variable_names = c("do", "temp"), # what the layer within each element is called
+  plus_minus = c(0.5, 0.5), # vector of plus/minus threshold(s) to define climate match
   match_logic = c("==", "=="), # will impliment rounding and max_thresholds will be ignored
   cell_size = 1,
   delta_t = 1,
@@ -30,7 +30,7 @@ dist_based_vocc(
   end_data = end_data,
   x = "x",
   y = "y",
-  variable_names = c("do_est", "temp"), 
+  variable_names = c("do", "temp"),
   round_fact = 10,
   min_thresholds = c(1, 1), # vectors of actual lower threshold values plus_minus*2
   max_thresholds = c(1, 1), # vectors of actual higher threshold values plus_minus*2
[…]

### 2019-07-17 analysis/tmb-sensor-explore/1-get-all-sensor-dat.R

@@ -12,23 +12,23 @@ names(d) <- tolower(names(d))
 head(d)
 
 
-.d <- d %>% #left_join(d, fe, by = "fishing_event_id") %>%
-  group_by(fishing_event_id) %>% 
+.d <- d %>% # left_join(d, fe, by = "fishing_event_id") %>%
+  group_by(fishing_event_id) %>%
   mutate(
-    #lat = latitude,
-    #lon = longitude,
-    do_mlpl.y = mean(do, na.rm = TRUE), 
-    do_mlpl_min.y = min(do, na.rm = TRUE), 
-    do_mlpl_max.y = max(do, na.rm = TRUE), 
-    salinity_psu.y = mean(salinity, na.rm = TRUE), 
-    salinity_psu_min.y = min(salinity, na.rm = TRUE), 
+    # lat = latitude,
+    # lon = longitude,
+    do_mlpl.y = mean(do, na.rm = TRUE),
+    do_mlpl_min.y = min(do, na.rm = TRUE),
+    do_mlpl_max.y = max(do, na.rm = TRUE),
+    salinity_psu.y = mean(salinity, na.rm = TRUE),
+    salinity_psu_min.y = min(salinity, na.rm = TRUE),
     salinity_psu_max.y = max(salinity, na.rm = TRUE),
     do_event_start = min(fe_event_time),
     do_event_end = max(fe_event_time)
-  ) %>% 
-  #rename(X = longitude, Y = latitude) %>%
-  add_tally(do, name= "do_mlpl_N.y") %>% 
-  add_tally(salinity, name = "salinity_psu_N.y") %>% 
+  ) %>%
+  # rename(X = longitude, Y = latitude) %>%
+  add_tally(do, name = "do_mlpl_N.y") %>%
+  add_tally(salinity, name = "salinity_psu_N.y") %>%
   mutate(do_mlpl_N.y = as.integer(do_mlpl_N.y), salinity_psu_N.y = as.integer(salinity_psu_N.y)) %>%
   select(-do, -salinity, -fe_event_time)
 
@@ -40,14 +40,21 @@ head(.d)
 glimpse(.d)
 # merge with newer sensor data
 new_sensor_data <- readRDS("analysis/tmb-sensor-explore/data/dat-sensor-trawl.rds")
-#library(tidyverse)
+# library(tidyverse)
 head(new_sensor_data)
 View(new_sensor_data)
[…]

### 2019-07-18 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
update rmd code for producing figures of environmental variables

@@
 predictions_rw <- predict(m_temp_rw)
-predictions_rw$data$residuals <- residuals(m_temp_rw)
-cor(predictions_rw$data$est, predictions_rw$data$temperature_c)^2 #
+predictions_rw$residuals <- residuals(m_temp_rw)
+cor(predictions_rw$est, predictions_rw$temperature_c)^2 #
 # aic(m_temp3_rw)
@@
 ```{r}
-ggplot(predictions_rw$data, aes(est, residuals)) +
+ggplot(predictions_rw, aes(est, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions_rw$data, aes(depth_scaled, residuals)) +
+ggplot(predictions_rw, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 predictions3 <- predict(m_temp3)
-predictions3$data$residuals <- residuals(m_temp3)
-cor(predictions3$data$est, predictions3$data$temperature_c)^2 #
+predictions3$residuals <- residuals(m_temp3)
+cor(predictions3$est, predictions3$temperature_c)^2 #
 ```
@@
 ```{r}
-ggplot(predictions3$data, aes(est, residuals)) +
+ggplot(predictions3, aes(est, residuals)) +
   geom_point() +
@@
 ```{r}
-ggplot(predictions3$data, aes(depth_scaled, residuals)) +
+ggplot(predictions3, aes(depth_scaled, residuals)) +
   geom_point() +
@@
 
-# A functions for plotting our predictions:
+#FIXME: DON'T HAVE SST TEMP FOR PREDICITON GRID YET...
+
+## Plotting predictions from saved temperature model
 
 ```{r}
-plot_map <- function(dat, column = "est") {
-  ggplot(dat, aes_string("X", "Y", colour = column)) +
-    geom_point() +
-    facet_wrap(~year) +
-    coord_fixed()
-}
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears.rds")
+# View(p_nd)
+```
 
+```{r}
 plot_map_raster <- function(dat, column = "est") {
@@
     geom_raster() +
+    coord_fixed()+
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+}
+
+plot_map_raster(p_nd, "est") +
+  scale_fill_viridis_c(trans = "log") +
   facet_wrap(~year) +
-    coord_fixed()
-}
-```
-
-## Plotting predictions
-#FIXME: DON'T HAVE SST TEMP FOR PREDICITON GRID YET...
-
-```{r}
-qcs_grid$depth_scaled3 <- (qcs_grid$depth_scaled)^3
-predictions_nd <- predict(m_temp3_rw, newdata = qcs_grid)
-# saveRDS(predictions3_rw_nd$data, file = "qcs-temp.rds")
-```
-
-```{r}
-plot_map_raster(predictions_nd$data, "est") +
-  scale_fill_viridis_c(trans = "sqrt", option = "C") +
   ggtitle("Prediction (fixed effects + all random effects)")
@@
 ```{r}
-plot_map_raster(predictions_nd$data, "est_fe") +
-  ggtitle("Prediction (fixed effects only)") +
-  scale_fill_viridis_c(trans = "sqrt")
-```
+p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
 
-```{r, eval=FALSE}
-plot_map_raster(predictions_nd$data, "est_re_s") +
-  ggtitle("Spatial random effects only") +
-  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
+breaks <- scales::trans_breaks(no_trans[["transform"]], no_trans[["inverse"]],n = 6)
+labels <- function(x) format(x, digits = 2, scientific = FALSE)
+
+pred_1 <- plot_map_raster(p_nd1, "est") +
+  scale_fill_viridis_c(labels = labels, 
+    #trans = "sqrt",
+    option = "C",
+    #breaks = c(0,2,4,6,8), 
+    limits = c(5, 11), 
+    na.value = "red") +
+  facet_wrap(~year) +
+  labs(fill = "°C") + 
+  theme(legend.position = "none") +
+  ggtitle("Prediction (fixed effects + all random effects)") 
+pred_1
 ```
@@
 ```{r}
-plot_map_raster(predictions_nd$data, "est_re_st") +
-  ggtitle("Spatiotemporal random effects only") +
-  scale_fill_gradient2(low = "navy", mid = "white", high = "darkred")
+p_nd4 <- p_nd %>% filter(ssid==4)
+pred_4 <- plot_map_raster(p_nd4, "est") +
+  scale_fill_viridis_c(labels = labels, 
+     #trans = "sqrt",
+    option = "C",
+    #breaks = c(0,2,4,6,8), 
+    limits = c(5, 11), 
+    na.value = "black")+
+  facet_wrap(~year) + 
+  theme(legend.position = c(.85,.12)) +
+  labs(fill = "°C")
+  #ggtitle(" ")
+pred_4
 ```
@@
 ```{r}
-ggplot(predictions3_rw$data, aes(temperature_c, est)) +
-  geom_point() +
+plot_map_raster(p_nd, "est_non_rf") +
+  ggtitle("Prediction (fixed effects only)") +
   facet_wrap(~year) +
-  coord_fixed() +
-  xlim(3, 15) +
-  ylim(3, 15) +
-  geom_abline()
+  scale_fill_viridis_c()
 ```
@@
 ```{r}
-temp_cog <- get_cog(predictions_nd)
+plot_map_raster(p_nd, "est_rf") +
+  ggtitle("Prediction (all random effects only)") +
+  facet_wrap(~year) +
+  scale_fill_viridis_c()
+```
+
+```{r}
+#glimpse(p_nd)
+
+omega_s_all <- plot_map_raster(p_nd, "omega_s") +
+  ggtitle("Spatial random effects only") +
+   scale_fill_gradient2( low = "Steel Blue 4",
+ high = "Red 3", limits = c(-1.5,2.5))+
+  labs(fill = "") + 
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))
+
+omega_s_all
+```
+
+```{r}
+ plot_map_raster(p_nd, "epsilon_st") +
+  ggtitle("Spatiotemporal random effects only") +
+  facet_wrap(~year) +
+  scale_fill_gradient2( low = "Steel Blue 4",
+ high = "Red 3")
+```
+
+```{r}
+p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
+st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
+  ggtitle("Spatiotemporal random effects") +
+  facet_wrap(~year) +
+  scale_fill_gradient2(
+    low = "Steel Blue 4",
+    high = "Red 3", 
+    limits = c(-1,1), 
+    na.value = "red") +
+  labs(fill = "epsilon") + 
+  gfplot::theme_pbs() + 
+  theme(
+  legend.position = "none", 
+  axis.title.x = element_blank(), axis.title.y = element_blank())
+st_1n3
+```
+
+```{r}
+p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
+st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
+  #ggtitle(" ") +
+  facet_wrap(~year) +
+  scale_fill_gradient2( low = "Steel Blue 4",
+ high = "Red 3", limits = c(-1,1)) +
+          labs(fill = " ") + 
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.15))
+st_4
+```
+
+```{r}
+p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
+st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
+  #ggtitle(" ") +
+  facet_wrap(~year) +
+  scale_fill_gradient2( low = "Steel Blue 4",
+ high = "Red 3", limits = c(-1,1)) +
+  xlim(65,345) +
+    ylim(5800,6100) +
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
+st_16
+```
+
+```{r}
+png(
+  file = "figs/temp-predictions.png", 
+  res = 600,
+  units = "in",
+  width = 11, 
+  height = 10
+)
+
+gridExtra::grid.arrange(
+  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16 ),
+  widths = c(2, 2, 1.75),
+  heights = c(2, 1.35),
+  layout_matrix = rbind(c(1, 2, 3),
+                        c(4, 5, 6)),
+  top = grid::textGrob("Bottom temperature estimates from sdmTMB model")
+)
+
+# gridExtra::grid.arrange( nrow = 2, heights = c(2,1.5), 
+#   top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")) 
+dev.off()
+```
+
+```{r}
+temp_cog <- get_cog(p_nd)
 

### 2019-07-18 analysis/tmb-sensor-explore/3-fit-do.Rmd
update rmd code for producing figures of environmental variables

@@
   scale_color_viridis_c()
+```
+
+```{r}
+ggplot(d_trawl, aes(temp, do_mlpl, colour = as.factor(ssid))) +
+  geom_point() +
+  facet_wrap(~year) +
+  scale_color_viridis_d()
+```
+
+```{r}
+ggplot(filter(d_trawl, year==2016), aes(temp, do_mlpl, colour = as.factor(ssid))) +
+  geom_point() +
+  #facet_wrap(~year) +
+  scale_color_viridis_d(begin=0.65)
+```
+
+```{r}
+ggplot(filter(d_trawl, year==2017), aes(temp, do_mlpl, colour = as.factor(ssid))) +
+  geom_point() +
+  #facet_wrap(~year) +
+  scale_color_viridis_d(end = 0.35)
 ```
@@
 
+Plot predictions from saved DO predictions file
+
 ```{r}
+p_nd <- readRDS((here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
+
 plot_map_raster <- function(dat, column = "est") {
@@
     geom_raster() +
-    coord_fixed()
+    coord_fixed()+
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
 }
@@
   ggtitle("Prediction (fixed effects + all random effects)")
+```
+
+```{r}
+p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
+
+breaks <- scales::trans_breaks(no_trans[["transform"]], no_trans[["inverse"]],n = 6)
+labels <- function(x) format(x, digits = 2, scientific = FALSE)
+
+pred_1 <- plot_map_raster(p_nd1, "do_est") +
+  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 9), na.value = "red") +
+  facet_wrap(~year) +
+  labs(fill = "ml/L DO") +
+  ggtitle("Prediction (fixed effects + all random effects)") + 
+  theme(legend.position = c(.85,.25))
+pred_1
+```
+
+```{r}
+p_nd4 <- p_nd %>% filter(ssid==4)
+pred_4 <- plot_map_raster(p_nd4, "do_est") +
+  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 9), na.value = "red")+
+  facet_wrap(~year) +
+  labs(fill = "ml/L DO") +
+  #ggtitle(" ")+ 
+  theme(legend.position = "none")
+pred_4
 ```
@@
 #glimpse(p_nd)
-plot_map_raster(p_nd, "omega_s") +
+
+omega_s_all <- plot_map_raster(p_nd, "omega_s") +
   ggtitle("Spatial random effects only") +
    scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3")+
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+  low = "Red 3", limits = c(-0.65,1))+
+  labs(fill = " ") + 
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))
+
+omega_s_all
 ```
@@
 p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
-plot_map_raster(p_nd1, "epsilon_st") +
-  ggtitle("Spatiotemporal random effects only") +
+st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
+  ggtitle("Spatiotemporal random effects") +
   facet_wrap(~year) +
   scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3", limits = c(-0.4,0.5)) +
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+  low = "Red 3", limits = c(-0.5,0.5)) +
+        labs(fill = " ") + 
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.25))
+st_1n3
 ```
+
 ```{r}
 p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
-plot_map_raster(p_nd4, "epsilon_st") +
-  ggtitle("Spatiotemporal random effects only") +
+st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
+  #ggtitle(" ") +
   facet_wrap(~year) +
   scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3", limits = c(-0.4,0.5)) +
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+  low = "Red 3", limits = c(-0.5,0.5)) +
+          labs(fill = "epsilon") + 
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
+st_4
 ```
@@
 p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
-plot_map_raster(p_nd16, "epsilon_st") +
-  ggtitle("Spatiotemporal random effects only") +
+st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
+  #ggtitle(" ") +
   facet_wrap(~year) +
   scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3", limits = c(-0.4,0.5)) +
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+  low = "Red 3", limits = c(-0.5,0.5)) +
+  xlim(65,345) +
+    ylim(5800,6100) +
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
+st_16
+```
+
+```{r}
+png(
+  file = "figs/do-predictions.png", 
+  res = 600,
+  units = "in",
+  width = 11, 
+  height = 6.5
+)
+
+gridExtra::grid.arrange(
+  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16 ),
+  widths = c(2, 2, 1.75),
+  heights = c(2, 1.5),
+  layout_matrix = rbind(c(1, 2, 3),
+                        c(4, 5, 6)),
+  top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")
+)
+
+# gridExtra::grid.arrange( nrow = 2, heights = c(2,1.5), 
+#   top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")) 
+dev.off()
 ```

### 2019-07-24 analysis/VOCC/1-biomass-do-models.Rmd
separate .Rmd files for biomass models based on each of depth, temp and DO

---
title: "Spatiotemporal models of species biomass split by maturity class"
author: "Philina English"
date: '2019-07-18'
output: html_document
params: 
    species: "Arrowtooth Flounder"

---

Choose one value for each parameter:

Species
```{r}
species <- params$species

# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

Region 
```{r}
region <- "Both odd year surveys"
# region <- "West Coast Vancouver Island"
#  region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
#priors <- FALSE
priors <- TRUE

covariates <- "+muddy+any_rock"
# covariates <- ""

# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

#update_spatial_model <- TRUE
```

Combine biomass and sensor data
```{r}
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars) %>% mutate(depth_akima = depth) %>% select(-depth)

  CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
  data <- dplyr::left_join(data, CTD) %>% 
  mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% filter(exclude != 1) %>% # added on july 24 after first round of models run...
    filter(year > 2007) # 2007 DO data is flawed

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth), temp = temperature_c) 
  do_data <- gfranges::scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- do_data %>% mutate(depth = raw_depth) %>% 
    filter(ssid %in% model_ssid) %>% filter(!is.na(do_mlpl))# %>% filter(year != 2016)
```

```{r}
  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  sdmTMB::plot_spde(spde)
```

Run DO sdmTMB model
```{r}
  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime1 <- Sys.time()
    adult_biomass_d <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime2 <- Sys.time()
    imm_biomass_d <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

  } else {

    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

    total_biomass_d <- sdmTMB::sdmTMB(data,
      dens_formula,
      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
```

### 2019-07-24 analysis/VOCC/1-biomass-models.Rmd
separate .Rmd files for biomass models based on each of depth, temp and DO

---
title: "Spatiotemporal models of species biomass split by maturity class"
author: "Philina English"
date: '2019-07-18'
output: html_document
params: 
    species: "Arrowtooth Flounder"

---

Choose one value for each parameter:

Species
```{r}
species <- params$species

# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
# species <- "Quillback Rockfish" # no maturity classes and WCHG not work
# species <- "Pacific Ocean Perch" 
species <- "Yelloweye Rockfish"
```

Region 
```{r}
# region <- "Both odd year surveys"
  region <- "West Coast Vancouver Island"
 region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
priors <- FALSE
# priors <- TRUE

# covariates <- ""

  covariates <- "+muddy+any_rock"

# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

#update_spatial_model <- TRUE
```

Make prediction grids for all surveys and years, if not done before.
```{r, message=FALSE, warning=FALSE}
try(nd_all <- readRDS("data/nd_all_synoptic.rds"))

if (!exists("nd_all")) {

  # choose base year(s) to create grid from
  dummy_year <- c(2005, 2006)

  bath <- readRDS("data/bathymetry-data")
  #dat <- bath$data %>% scale_survey_predictors()
  dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
  dat <- gfranges::scale_predictors(dat, predictors = c(quo(depth)))

  # create grids for each ssid separately
  ssid <- 1
  survey_abbrev <- "SYN QCS"
  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_1$year)

  ssid <- 3
  survey_abbrev <- "SYN HS"
  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_3$year)

  ssid <- 4
  survey_abbrev <- "SYN WCVI"
  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_4$year)

  ssid <- 16
  survey_abbrev <- "SYN WCHG"
  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_16$year)

  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
    mutate(year = as.numeric(year))

  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))

  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))

  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
}
```

Calculate maturity weighted biomass if not done before
```{r}
rm(biomass)
try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))

if (!exists("biomass")) {
  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
  bath <- readRDS("data/bathymetry-data")
  # for POP: exclude sample with unusually large fish 
  # sizes are impossible given recorded catch weight
  # fish <- fish[fish$fishing_event_id!=1506954,]
 rm(maturity)
 try( maturity <- split_catch_maturity(survey_sets, fish, bath,
    # survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
    # survey = c("SYN HS", "SYN WCHG"),
    survey = survey,
    years = years,
    cutoff_quantile = c(.9995),
    plot = TRUE
  ))

  if(!exists("maturity")) {
  maturity <- split_catch_maturity(survey_sets, fish, bath,
    survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
    years = years,
    cutoff_quantile = c(.9995),
    plot = TRUE
  )}

  saveRDS(maturity$data, file = paste0("data/", spp, 
    "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
}
```

```{r}
if (exists("maturity")) {
  print(maturity$mass_model)
  ggsave(
    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
    dpi = 600,
    width = 8, 
    height = 10
  )
}
```

```{r}
if (exists("maturity")) {
  png(
    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), 
    res = 600,
    units = "in",
    width = 8, 
    height = 10
  )
  print(maturity$maturity)
  dev.off()
}

try(print(maturity$maturity))
```

Run climate independent sdmTMB model
```{r}
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars)

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
  data <- gfranges::scale_predictors(data, predictors = c(quo(depth)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)

  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  sdmTMB::plot_spde(spde)

  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime1 <- Sys.time()
    adult_biomass <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime2 <- Sys.time()
    imm_biomass <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass, 
     file = paste0("data/", spp, 
       "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
       ))

  } else {

    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

    total_biomass <- sdmTMB::sdmTMB(data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
```

Save predictions for full spatial grid

```{r}

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

if (length(unique(fish$maturity_code)) > 2) {

    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
    nd <- nd_all %>% 
      filter(ssid %in% model_ssid) %>% 
      filter(year %in% unique(adult_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    adult_biomass <- readRDS(paste0("data/", spp, 
      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
    adult_predictions <- predict(adult_biomass, newdata = nd)
    saveRDS(adult_predictions, file = paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
    ))

    imm_biomass <- readRDS(paste0("data/", spp, 
      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions, file = paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
    ))
  } else {
    rm(adult_predictions)
    rm(imm_predictions)

    if (!exists(predictions)) {
      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
      nd <- nd_all %>% filter(ssid %in% model_ssid)
      nd <- na.omit(nd)
      nd$year <- as.integer(nd$year)

      total_biomass <- readRDS(paste0("data/", spp, 
        "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
      predictions <- predict(total_biomass, newdata = nd)
      saveRDS(predictions, file = paste0("data/", spp, 
        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
      ))
    }
  }
```

### 2019-07-24 analysis/VOCC/3-variable-response-plots.Rmd
.Rmd to make all mountain plots for a species

---
title: "Plot responses to model covariates to inform threshold choices for VOCC"
author: "Philina English"
date: '2019-07-18'
output: html_document
---

Species
```{r}

# # Species run so far...
# species <- "Pacific Cod"
species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"

# species <- "Yelloweye Rockfish"
# species <- "Arrowtooth Flounder"

```

Choose model details
```{r}
# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
covariates <- "+muddy+any_rock"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  fig.path = paste0("figs/", spp, "/"),
  echo = FALSE, warning = FALSE, message = FALSE
)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

Dissolved O2 plots
```{r}
rm(
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16
)
priors <- TRUE
# covariates <- ""
try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL

model_list <- list(
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
)

model_list <- model_list[!sapply(model_list, is.null)]
```

```{r}
d <- list()
do_plots <- list()
for (i in seq_len(length(model_list))) {
  d[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
  do_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Dissolved O2", xlimits = c(0,8)) + ggtitle(names(model_list[i]))
}
print(do_plots)
```

```{r}
png(
  file = paste0("figs/", spp, "/do-", spp, covs, "-priors-TRUE.png"), 
  res = 600,
  units = "in",
  width = 8.5, 
  height = 11
)
gridExtra::grid.arrange(
  grobs = do_plots,
  ncol = 2,
  top = grid::textGrob(paste(species))
)
dev.off()
```

Temperature plots
```{r}
rm(
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16
)
priors <- TRUE
try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL

temp_model_list <- list(
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
)

temp_model_list <- temp_model_list[!sapply(temp_model_list, is.null)]
```

```{r}
d <- list()
temp_plots <- list()
for (i in seq_len(length(temp_model_list))) {
  d[[i]] <- time_varying_density(temp_model_list[[i]], predictor = "temp")
  #d[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
  temp_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Temperature", xlimits = c(2,10)) + ggtitle(names(temp_model_list[i]))
}
print(temp_plots)
```

```{r}
png(
  file = paste0("figs/", spp, "/temp-", spp, covs, "-priors-TRUE.png"), 
  res = 600,
  units = "in",
  width = 8.5, 
  height = 11
)
gridExtra::grid.arrange(
  grobs = temp_plots,
  ncol = 2,
  top = grid::textGrob(paste(species))
)
dev.off()
```

Depth plots
```{r}
rm(
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16
)
priors <- FALSE
try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL

depth_model_list <- list(
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
)

depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
```

```{r}
d <- list()
depth_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
  d[[i]]$x <- exp(d[[i]]$x)
  depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0,1000)) + ggtitle(names(depth_model_list[i]))
}
print(depth_plots)
```

```{r}
png(
  file = paste0("figs/", spp, 
    "/depth-", spp, covs, "-priors-", priors, ".png"), 
  res = 600,
  units = "in",
  width = 8.5, 
  height = 11
)
gridExtra::grid.arrange(
  grobs = depth_plots,
  ncol = 2,
  top = grid::textGrob(paste(species))
)
dev.off()
```

Plot time-varying effect of trawling at mean depth
```{r}
trawl_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  r <- depth_model_list[[i]]$tmb_obj$report()
  b3 <- r$b_rw_t[, 3]
  yrs <- sort(unique(depth_model_list[[i]]$data$year))
  b_j <- depth_model_list[[i]]$model$par[1:length(yrs)]

  SE <- 0.35 #Place-holder

  trawled <- data.frame(year = yrs, intercept = b_j, trawling = b3)

  # trawl_plots[[i]] <- ggplot(trawled, aes(year, y=exp(trawling) ) ) + 
  #   geom_point(colour ="red") + 
  #   #geom_pointrange(aes(ymin=exp(intercept-0.35), ymax=exp(intercept+0.35))) +
  #   ylab("Difference in kg/ha within trawled zone") + ggtitle(names(depth_model_list[i]))

  trawl_plots[[i]] <- ggplot(trawled, aes(year, exp(intercept))) + geom_point() +
    geom_point(aes(year, y = exp(intercept + trawling)), colour = "red") +
    geom_pointrange(aes(ymin = exp(intercept - (1.96 * SE)), ymax = exp(intercept + (1.96 * SE)))) + # using mean se on year effects
    ylim(min(exp(trawled$intercept - trawled$trawling - (1.96 * SE))), max(exp(trawled$intercept + trawled$trawling + (1.96 * SE)))) +
    ylab("Predicted kg/ha at mean depth") + ggtitle(names(depth_model_list[i]))
}
print(trawl_plots)
```

### Code for calculation SE on time-varying effects...

```{r eval=FALSE}
# FIXME: need to add to sdmTMB!
r <- m$tmb_obj$report()
summary(m)
yrs <- sort(unique(m$data$year))
.sd <- summary(m$sd_report)
.sd <- as.data.frame(.sd)
.sd$par <- row.names(.sd)
.sd <- .sd[grepl("^b_rw_t", .sd$par), ]
varnames <- colnames(model.matrix(m$time_varying, m$data))

rep(varnames, each = length(yrs))
.sd$parameter <- rep(varnames, each = length(yrs))
.sd$year <- rep(seq_along(yrs), each = length(varnames))
row.names(.sd) <- NULL
.sd$par <- NULL

names(.sd$Estimate) <- "estimate"
names(.sd[2]) <- "se"
.sd$ci <- .sd$se * 1.96
m$time_varying

nrow(.sd)

b3 <- r$b_rw_t[, 3]
yrs <- sort(unique(m$data$year))
b_j <- m$tmb_params$b_j[1:8]

trawled <- data.frame(year = yrs, intercept = b_j, trawling = b3)

```

### 2019-07-26 R/plot-maps.R
refine plot-map functions: remove utm axis lables from faceted maps and make bathymetry lines finer

@@ -148,7 +148,7 @@ plot_vocc <- function(df,
           x = "X", y = "Y",
           group = "paste(PID, SID)"
         ),
-        inherit.aes = FALSE, lwd = 0.4, alpha = 0.4, colour = "grey55"
+        inherit.aes = FALSE, lwd = 0.1, alpha = 0.4, colour = "grey55"
       )
   } else {
     # add bathymetry layer from gfplot
@@ -176,7 +176,7 @@ plot_vocc <- function(df,
             x = "X", y = "Y",
             group = "paste(PID, SID)"
           ),
-          inherit.aes = FALSE, lwd = 0.4, alpha = 0.4, colour = "grey85"
+          inherit.aes = FALSE, lwd = 0.2, alpha = 0.4, colour = "grey85"
         ) +
         geom_path(
           data = isobath[isobath$PID == 200, ],
@@ -184,7 +184,7 @@ plot_vocc <- function(df,
             x = "X", y = "Y",
             group = "paste(PID, SID)"
           ),
-          inherit.aes = FALSE, lwd = 0.4, alpha = 0.4, colour = "grey70"
+          inherit.aes = FALSE, lwd = 0.2, alpha = 0.4, colour = "grey70"
         ) +
         geom_path(
           data = isobath[isobath$PID == 300, ],
@@ -192,7 +192,7 @@ plot_vocc <- function(df,
             x = "X", y = "Y",
             group = "paste(PID, SID)"
           ),
-          inherit.aes = FALSE, lwd = 0.4, alpha = 0.4, colour = "grey55"
+          inherit.aes = FALSE, lwd = 0.2, alpha = 0.4, colour = "grey55"
         ) +
         geom_path(
           data = isobath[isobath$PID == 400, ],
@@ -200,7 +200,7 @@ plot_vocc <- function(df,
             x = "X", y = "Y",
             group = "paste(PID, SID)"
           ),
-          inherit.aes = FALSE, lwd = 0.4, alpha = 0.4, colour = "grey40"
+          inherit.aes = FALSE, lwd = 0.2, alpha = 0.4, colour = "grey40"
         ) +
         geom_path(
           data = isobath[isobath$PID == 500, ],
@@ -208,7 +208,7 @@ plot_vocc <- function(df,
             x = "X", y = "Y",
             group = "paste(PID, SID)"
           ),
[…]

### 2019-07-27 analysis/VOCC/4-vocc-by-spp.Rmd
initial draft merge of climate vectors with biomass estimates

---
title: "Biomass predictions and VOCC"
author: "Philina English"
date: '2019-07-18'
output: html_document
---

This file calls models built by "1-biomass-models.Rmd" and stiches the predictions together.

Species
```{r}
# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

Region 
```{r}
# region <- "All synoptic surveys"

# region <- "Both odd year surveys"
region <- "West Coast Vancouver Island"
# region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
update_spatial_model <- FALSE
priors <- FALSE
# priors <- TRUE

# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
covariates <- "+muddy+any_rock"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

#fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

```

```{r}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

adult_predictions <- readRDS(paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
adult_predictions <- filter(adult_predictions, ssid %in% model_ssid)

imm_predictions <- readRDS(paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))

imm_predictions <- filter(imm_predictions, ssid %in% model_ssid)

```

```{r}
glimpse(adult_predictions)
adult_change <- make_trend_data(adult_predictions,
        ssid = c(4),
        start_time = 2008,
        skip_time = 2016,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 6,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2),
        variable_names = "est_exp"
      )
```

Import vectors for single climate variable at a time
```{r}
variable_names <- c("DO")

declining_var <- "do_est"
min_threshold <- 0.2
vocc <- readRDS(paste0("data/", 
      "vocc-", ssid_string, declining_var, min_threshold, "decline.rds"))

# increasing_var <- "temp" 
# max_threshold <- 0.5
# vocc <- readRDS(paste0("data/", 
#       "vocc-", ssid_string, increasing_var, max_threshold, "increase.rds"))
```

```{r}
glimpse(adult_change)
#unique(adult_predictions$ssid)
glimpse(vocc)

start <- adult_change %>% rename(start_density = var_1_s, end_density = var_1_e) %>% select(icell, x, y, start_density, end_density)
end <- adult_change %>% rename(start_density = var_1_s, end_density = var_1_e, target_X = x, target_Y = y) %>% select(icell,  start_density, end_density, target_X, target_Y)

if (length(variable_names) == 1) {
vectors <- vocc  %>% mutate(start_cell = icell, vector_id = paste0(id, "_to_", tid), 
  initial_value = var_1_s, source_end_value = var_1_e, target_value = as.numeric(target_values)) %>% 
  filter(distance > 1) %>%
  select(vector_id, start_cell, x, y, target_X, target_Y, distance, initial_value, source_end_value, target_value, km_per_decade, N)
} else {
vectors <- vocc  %>% mutate(start_cell = icell, vector_id = paste0(id, "_to_", tid), 
  initial_value = s, source_end_value = e, target_value = target_values) %>% 
  filter(distance > 1) %>%
  select(vector_id, start_cell, x, y, target_X, target_Y, distance, initial_value, source_end_value, target_value, km_per_decade, N) # need to replace N ...
}
vector_source <- left_join(vectors, start, by=c("x", "y")) %>% mutate(cell_type = "source")
vector_target <- left_join(vectors, end, by=c("target_X", "target_Y")) %>% mutate(cell_type = "target")

glimpse(vector_source)
glimpse(vector_target)

vectors <- rbind(vector_source, vector_target)
glimpse(vectors)
```

```{r}
do_vect_plot <- ggplot(filter(vectors, distance < 100), aes(distance, y = end_density-start_density, group = cell_type, colour= cell_type)) + 
  geom_jitter(width = 0.25, alpha = 0.45, size = 2) + 
  ylab(paste("Change in biomass over", vectors$N[1], "years")) + 
  xlab(paste("Distance traveled to maintain DO decline of <", min_threshold, "")) + 
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + gfplot::theme_pbs() + theme(legend.position = c(0.8,0.8)) + ggtitle(paste(survey, species))
do_vect_plot 
```

```{r}
test <- lmerTest::lmer(end_density ~ cell_type * distance + start_density  + (1|vector_id), data = filter(vectors, distance < 30))
summary(test)
```

```{r}
png(
  file = paste0("figs/", spp, "/vocc-biomass-change", spp, ssid_string, ".png"),
  res = 600,
  units = "in",
  width = 5,
  height = 4
)

do_vect_plot 
dev.off()
```

### 2019-08-02 analysis/VOCC/4-vocc-by-spp.Rmd
code linking vocc and change in biomass estimates; explores a variety of attempts to model of the relationship despite ugly residuals

@@
 
-This file calls models built by "1-biomass-models.Rmd" and stiches the predictions together.
+This file ...
 
 Species
+
 ```{r}
 # # Species run so far...
-# species <- "Arrowtooth Flounder"
-# species <- "Pacific Cod"
-# species <- "Sablefish"
+ species <- "Arrowtooth Flounder"
+ species <- "Pacific Cod"
+ species <- "Sablefish"
 # species <- "Silvergray Rockfish"
-species <- "Lingcod"
+#  species <- "Lingcod"
 #  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
@@
 # species <- "Yelloweye Rockfish"
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
 ```
@@
 # region <- "All synoptic surveys"
-
-# region <- "Both odd year surveys"
+ region <- "Both odd year surveys"
  region <- "West Coast Vancouver Island"
@@
 
-Choose model details
+Make climate vectors for biannual changes between surveys
 ```{r}
-update_spatial_model <- FALSE
-priors <- FALSE
-# priors <- TRUE
+trend <- FALSE
+start_time <- 2015
+end_time <- 2018
 
-# covariates <- "+muddy+mixed+rocky"
-# covariates <- "+mixed+rocky"
-covariates <- "+muddy+any_rock"
-# covariates <- "+trawled+muddy+rocky+mixed"
-# covariates <- "+trawled+mixed+rocky"
-# covariates <- "+trawled+mixed"
+climate <- "temperature"
+temp_threshold <- c(1)
+optimal_temp <- 7 # yellow-eye 4
+
+climate <- "do"
+do_threshold <- c(0.5)
+optimal_do <- 2.5 # yellow-eye 4
+
+# climate <- "DO and temperature"
+# max_thresholds <- c(Inf, 1)
+# min_thresholds <- c(0.5, Inf)
+
+```
+
+Make climate vectors for average changes and lower threshold values
+```{r}
+# trend <- TRUE
+# 
+# climate <- "temperature"
+# temp_trend_threshold <- c(0.25)
+
+# climate <- "DO"
+# do_trend_threshold <- c(0.2)
+
+# climate <- "both DO and temperature"
+# min_trend_thresholds <- c(0.2, Inf)
+# max_trend_thresholds <- c(Inf, 0.5)
 ```
@@
 ```{r global_options, include=FALSE}
-covs <- gsub("\\+", "-", covariates)
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-# folder to hold figs for this species
-dir.create(file.path("figs", spp))
-dir.create(file.path("data", spp))
-
-knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
-                      #fig.path=paste0("figs/", spp, "/"),
-                      echo=FALSE, warning=FALSE, message=FALSE)
-```
-
-```{r, message=FALSE, warning=FALSE}
 library(dplyr)
@@
 library(sdmTMB)
-library(gfranges)
+#library(gfranges)
 
-#fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+knitr::opts_chunk$set(
+  fig.width = 11, fig.height = 8.5,
+  echo = FALSE, warning = FALSE, message = FALSE
+)
+```
 
-if (region == "All synoptic surveys") {
-  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
-  model_ssid <- c(1, 3, 4, 16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  years <- NULL
-}
-
+```{r, message=FALSE, warning=FALSE}
 if (region == "Both odd year surveys") {
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+  trend_indices_do <- c(1, 1, 1, 2, 2)
+  trend_indices_temp <- c(1, 1, 1, 2, 2)
   years <- NULL
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+  trend_indices_do <- c(1, 1, 1, 2, 2)
+  trend_indices_temp <- c(1, 1, 1, 2, 2, 2)
   years <- NULL
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+  trend_indices_do <- c(1, 1, 1, 2, 2)
+  trend_indices_temp <- c(1, 1, 1, 2, 2)
   years <- NULL
+}
+```
+
+```{r}
+# Set default cell sizes and scale
+input_cell_size <- 2
+scale_fac <- 1
+delta_t_step <- 2
+skip_time <- NULL
+
+if (climate == "temperature") {
+  variable_names <- c("temp")
+  min_thresholds <- c(Inf)
+  delta_threshold <- temp_threshold
+  optimal_value <- optimal_temp
+
+  if (trend) {
+    max_thresholds <- temp_trend_threshold
+    start_time <- 2008 # change if can add back in prior 2008 data
+    end_time <- NULL
+    delta_t_total <- 6 # change if can add back in prior 2008 data
+    indices <- trend_indices_temp
+  } else {
+    max_thresholds <- temp_threshold
+    indices <- c(1, 2)
+    delta_t_total <- 2
+  }
+}
+
+if (climate == "do") {
+  variable_names <- c("do_est")
+  max_thresholds <- c(Inf)
+  delta_threshold <- -1*(do_threshold)
+  optimal_value <- optimal_do
+
+  if (trend) {
+    min_thresholds <- do_trend_threshold
+    start_time <- 2008
+    skip_time <- 2016
+    end_time <- NULL
+    delta_t_total <- 6
+    indices <- trend_indices_do
+  } else {
+    min_thresholds <- do_threshold
+    indices <- c(1, 2)
+    delta_t_total <- 2
+  }
+}
+
+if (climate == "do and temperature") {
+  variable_names <- c("do_est", "temp")
+
+  if (trend) {
+    min_thresholds <- min_trend_thresholds
+    max_thresholds <- max_trend_thresholds
+    start_time <- 2008
+    skip_time <- 2016
+    end_time <- NULL
+    delta_t_total <- 6
+    indices <- trend_indices_do
+  } else {
+    min_thresholds <- min_thresholds
+    max_thresholds <- max_thresholds
+    indices <- c(1, 2)
+    delta_t_total <- 2
+  }
+}
+
+min_string <- paste0(min_thresholds, collapse = "-")
+max_string <- paste0(max_thresholds, collapse = "-")
+climate <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
+```
+
+Calculate climate vectors
+```{r}
+rm(vocc)
+try({
+  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate, "-",  
+    min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+})
+
+if (!exists("vocc")) {
+pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
+
+if (length(variable_names) > 1) {
+  climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
+} else {
+  climate_predictions <- pred_temp_do
+}
+
+starttime <- Sys.time()
+vocc <- make_vector_data(climate_predictions,
+  variable_names = variable_names,
+  ssid = model_ssid,
+  start_time = start_time,
+  end_time = end_time,
+  skip_time = skip_time,
+  input_cell_size = input_cell_size,
+  scale_fac = scale_fac,
+  delta_t_total = delta_t_total,
+  delta_t_step = delta_t_step,
+  indices = indices,
+  min_thresholds = min_thresholds,
+  max_thresholds = max_thresholds,
+  round_fact = 10
+)
+endtime <- Sys.time()
+time_vocc <- round(starttime - endtime)
+
+vocc$var_1_min <- min_thresholds[1]
+vocc$var_2_min <- min_thresholds[2]
+vocc$var_1_max<- max_thresholds[1]
+vocc$var_2_max<- max_thresholds[2]
+
+saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate, "-",  
+  min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
 }
@@
 
+Get matching biomass data for the above time periods
+
+Get all biomass predictions 
 ```{r}
+priors <- FALSE
+# priors <- TRUE
+
+covariates <- "+muddy+any_rock"
+
+covs <- gsub("\\+", "-", covariates)
+
 rm(adult_predictions)
@@
 
+try({
   adult_predictions <- readRDS(paste0("data/", spp,
   "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
-adult_predictions <- filter(adult_predictions, ssid %in% model_ssid)
+})
 
+try({
   imm_predictions <- readRDS(paste0("data/", spp,
   "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))
+})
 
-imm_predictions <- filter(imm_predictions, ssid %in% model_ssid)
+try({
+  predictions <- readRDS(paste0("data/", spp,
+  "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
+})
+
+if (!exists("adult_predictions")) adult_predictions <- NULL
+if (!exists("imm_predictions")) imm_predictions <- NULL
+if (!exists("predictions")) predictions <- NULL
+
+biomass_list <- list(
+  adult = adult_predictions, 
+  imm = imm_predictions, 
+  total = predictions
+)
+
+biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
+```
+
+Extract biomass values to match spatial and temporal range of climate vectors
+```{r}
+vocc_by_sp <- list()
+for (i in seq_along(biomass_list)){
+
+biomass_change <- make_trend_data(biomass_list[[i]],
+  ssid = model_ssid,
+  start_time = start_time,
+  skip_time = skip_time,
+  end_time = end_time,
+  input_cell_size = input_cell_size,
+  scale_fac = scale_fac,
+  delta_t_total = delta_t_total,
+  delta_t_step = delta_t_step,
+  indices = indices,
+  variable_names = "est_exp"
+)
+
+start <- biomass_change %>% 
+  rename(start_density = var_1_s, end_density = var_1_e) %>% 
+  select(icell, x, y, start_density, end_density)
+
+end <- biomass_change %>% 
+  rename(start_density = var_1_s, end_density = var_1_e, target_X = x, target_Y = y) %>% 
+  select(icell, start_density, end_density, target_X, target_Y)
+
+source_density <- biomass_change %>% rename(source_density = var_1_s) %>% select(source_density, x, y)
+
+if (length(variable_names) == 1) {
+  vectors <- vocc %>%
+    mutate(
+      start_cell = icell, vector_id = paste0(id, "_to_", tid),
+      initial_value = var_1_s, source_end_value = var_1_e, target_value = as.numeric(target_values)
+    ) %>%
+    filter(distance > 1)  %>%
+    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
+      initial_value, source_end_value, target_value, 
+      km_per_decade, N) 
+  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
+  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))
+
+} else {
+  vectors <- vocc %>%
+    mutate(
+      start_cell = icell, vector_id = paste0(id, "_to_", tid),
+      initial_value = s, source_end_value = e, target_value = target_values
+    ) %>%
+    filter(distance > 1) %>%
+    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
+      initial_value, source_end_value, target_value, 
+      km_per_decade, N) 
+  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
+  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))
+
+}
+
+vector_source <- left_join(vectors, start, by = c("x", "y")) %>% 
+  mutate(cell_type = "source", end_climate = source_end_value, source_density = start_density)
+
+vector_target <- left_join(vectors, end, by = c("target_X", "target_Y")) %>% 
+  mutate(cell_type = "target", end_climate = target_value)
+vector_target <- left_join(vector_target, source_density, by = c("x", "y")) #%>%  
+  # filter(end_climate > 2)
+
+vocc_by_sp[[i]] <- rbind(vector_source, vector_target) %>% 
+  filter(!is.na(icell)) %>% filter(source_density > 0.000001) %>% 
+  mutate(perc_change = (end_density - start_density)/start_density, 
+    cell_type = as.factor(cell_type)) %>% 
+  rename(X = x, Y = y)
+
+icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
+mean.change <- mean(icells$perc_change, na.rm = TRUE)
+sd.change <- sd(icells$perc_change, na.rm = TRUE)
+
+mean.density <- mean(icells$source_density, na.rm = TRUE)
+sd.density <- sd(icells$source_density, na.rm = TRUE)
+
+mean.climate <- mean(icells$end_climate, na.rm = TRUE)
+sd.climate <- sd(icells$end_climate, na.rm = TRUE)
+
+# glimpse(icells)
+# range(icells$perc_change)
+# range(vocc_by_sp[[i]]$perc_change)
+
+Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
+Xmin <- quantile(vocc_by_sp[[i]]$X, 0.0005)
+
+vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% mutate(std_change = (perc_change)/sd.change, 
+  std_source_density = (source_density - mean.density)/(sd.density*2), 
+  std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
+  mean_change = mean.change, sd_change = sd.change,
+  mean_source_density = mean.density, sd_density = sd.density,
+  mean_climate = mean.climate, sd_climate = sd.climate,
+  std_end_climate2 = std_end_climate^2, 
+  distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, # set intercept at scale of cells
+  distance2 = distance*distance
+  ) %>% filter ( X >= Xmin & X <= Xmax )
+
+# range(vocc_by_sp[[i]]$X)
+
+#glimpse(vocc_by_sp[[i]])
+}
+names(vocc_by_sp) <- names(biomass_list)
+```
+
+Check final climate-biomass relationship
+```{r}
+data <- vocc_by_sp[["adult"]]
+#glimpse(data)
+
+mature_plot <- ggplot(data, aes(x = end_climate, y = (end_density # - start_density
+  ), colour = cell_type, #size = start_density, 
+  group = cell_type)) +
+  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
+  #   , aes(x = end_climate, y = (end_density #- start_density
+  #  )), inherit.aes = FALSE) +
+  scale_size_continuous(guide = "none") +
+  scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("Final biomass")) +
+  xlab(paste("Final value of", climate,"for all cells")) +
+  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
+  ggtitle(paste(survey, "mature", species))
+
+mature_plot
+```
+
+Check distance threshold of 50 looks reasonable
+```{r}
+data <- vocc_by_sp[["adult"]]
+#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)
+
+mature_plot2 <- ggplot(data, aes(distance, y = std_change, 
+  group = cell_type, size = start_density, colour = end_climate)) +
+  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+  geom_smooth(method="lm", formula = y~x+I(x^2) )+
+  facet_wrap(~cell_type) +#, scales = "free") +
+  scale_size_continuous(guide = "none") +
+  # scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
+  ggtitle(paste(survey, "mature", species))
+
+mature_plot2
+```
+
+```{r warnings=FALSE}
+data <- vocc_by_sp[["adult"]]
+#data <- filter(vocc_by_sp[["adult"]], distance < 50)
+#data <- filter(vocc_by_sp[["adult"]], distance > 4 & distance < 50)
+
+mature_plot3 <- ggplot(data, aes(x = source_density, y = std_change, 
+    colour = end_climate, 
+    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
+    group = cell_type)) +
+  geom_jitter(width = 0.01) +
+  geom_smooth(method="lm") +
+  facet_wrap(~cell_type) + #, scales = "free") +
+  scale_alpha_continuous(guide = "none") +
+  scale_size_continuous(guide = "none") +
+  scale_x_continuous(trans = sqrt) +
+  # scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Source density (dark points are closer to source)")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+
+mature_plot3
+```
+
+```{r}
+data <- vocc_by_sp[["adult"]]
+data <- filter(vocc_by_sp[["adult"]], distance < 50)
+
+mature_plot4 <- ggplot(filter(data, distance < 10), aes(source_density, y = std_change , 
+    colour = end_climate, 
+    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
+    group = cell_type)) +
+  geom_jitter(width = 0.01) +
+  geom_smooth(method="lm") +
+  facet_wrap(~cell_type) +#, scales = "free") +
+  scale_alpha_continuous(guide = "none") +
+  scale_size_continuous(guide = "none") +
+  scale_x_continuous(trans = sqrt) +
+  #scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Source density (dark points are closer to source)")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+
+mature_plot4
+```
+
+```{r}
+data <- vocc_by_sp[["adult"]]
+data <- filter(vocc_by_sp[["adult"]], distance < 50)
+
+mature_plot5 <- ggplot(data, aes(x = end_climate - initial_value , y = #std_change
+    (end_density - start_density)
+  , colour = cell_type, size = start_density)) +
+  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+  scale_size_continuous(guide = "none") +
+  #scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("kg/ha change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Change in", climate,"")) +
+  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "mature", species))
+mature_plot5
+```
+
+```{r}
+# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
+#             std_end_climate + std_end_climate2 + 
+#             (1 | vector_id), 
+#  #   offset = log(start_density), 
+#  #   family = Gamma(link = "inverse")
+#      data = data)
+# hist(data$end_density)
+# hist(data$std_change)
+data <- vocc_by_sp[["adult"]]
+
+test2 <- lmerTest::lmer(std_change ~ cell_type * std_source_density + 
+    cell_type * distance + cell_type * distance2 +
+    std_end_climate + std_end_climate2 + (1 | vector_id), REML = F, data = data)
+
+summary(test2)
+plot(test2)
+```
+
+Robust lmm
+```{r eval=FALSE}
+library(lme4)
+data <- vocc_by_sp[["adult"]]
+
+test3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
+    cell_type * distance + cell_type * distance2 + 
+    std_end_climate + std_end_climate2 + (1 | vector_id), data = data)
+
+summary(test3)
+plot(test3, ask = FALSE)
+```
+
+Boosted regression
+```{r}
+data <- vocc_by_sp[["adult"]]
+
+m <- gbm::gbm(
+  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
+  data = data, 
+  distribution = "gaussian",
+  n.trees = 2000, 
+  interaction.depth = 3, 
+  shrinkage = 0.02
+  )
+
+# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
+#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
+```
+
+```{r}
+summary(m)
+```
+
+```{r}
+plot(m,1)
+plot(m,2)
+plot(m,3)
+plot(m,4)
+plot(m,c(1,2))
+plot(m,c(1,3))
+plot(m,c(1,4))
+plot(m,c(2,3))
+plot(m,c(2,4))
+plot(m,c(1,2,4))
+plot(m,c(2,3,4))
+plot(m,c(5,6))
+```
+
+```{r}
+data <- vocc_by_sp[["adult"]]
+
+  if (region == "Both odd year surveys") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
+  }
+
+  if (region == "West Coast Vancouver Island") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 100)
+  }
+
+  if (region == "West Coast Haida Gwaii") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
+  }
+
+  sdmTMB::plot_spde(spde)
+```
+
+```{r}
+#data$end_climate2 <- data$end_climate^2
+#data <- vocc_by_sp[["adult"]]
+data$std_source_density2 <- data$std_source_density^2
+test <- sdmTMB(data = data,
+      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 +
+    cell_type * distance + cell_type * distance2 + 
+      std_end_climate + std_end_climate2, 
+      # std_change ~ cell_type * source_density + cell_type * distance  + end_climate + end_climate2,
+      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      silent = TRUE)
+test
+```
+
+```{r}
+p <- predict(test)
+p$residuals <- residuals(test)
+#idata$resids <- p$residuals
+#glimpse(p)
+ggplot(p, aes(est, residuals)) +
+  geom_point(alpha=0.4) +
+  ylim(-8,8) +
+  geom_smooth()
+
+plot_map_raster <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    coord_fixed()+
+    gfplot::theme_pbs() + 
+    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+}
+
+plot_map_raster(p, "est")
+plot_map_raster(p, "est_rf")
+plot_map_raster(p, "omega_s")
+plot_map_raster(p, "residuals")
+
+```
+
+Check final climate-biomass relationship
+```{r}
+idata <- vocc_by_sp[["imm"]]
+#glimpse(data)
+
+immature_plot1 <- ggplot(idata, aes(x = end_climate, y = (end_density # - start_density
+  ), colour = cell_type, #size = start_density, 
+  group = cell_type)) +
+  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
+  #geom_smooth() +
+  scale_size_continuous(guide = "none") +
+  scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("Final biomass")) +
+  xlab(paste("Final value of", climate,"for all cells")) +
+  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
+  ggtitle(paste(survey, "immature", species))
+
+immature_plot1
+```
+
+Check distance threshold of 50 looks reasonable
+```{r warnings=FALSE}
+idata <- vocc_by_sp[["imm"]]
+#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)
+
+immature_plot2 <- ggplot(idata, aes(distance, y = std_change, 
+  group = cell_type, size = start_density, colour = end_climate)) +
+  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+  geom_smooth(method="lm", formula = y~x+I(x^2) )+
+  facet_wrap(~cell_type) + #, scales = "free") +
+  scale_size_continuous(guide = "none") +
+  scale_y_continuous(trans = sqrt) +
+  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "immature", species))
+
+immature_plot2
+```
+
+```{r warnings=FALSE}
+idata <- vocc_by_sp[["imm"]]
+#idata <- filter(vocc_by_sp[["imm"]], distance < 50)
+#idata <- filter(vocc_by_sp[[""imm"]], distance > 4 & distance < 50)
+
+immature_plot3 <- ggplot(idata, aes(x = source_density, y = std_change, 
+    colour = end_climate, 
+    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
+    group = cell_type)) +
+  geom_jitter(width = 0.01) +
+  geom_smooth(method="lm") +
+  facet_wrap(~cell_type) + #, scales = "free") +
+  scale_alpha_continuous(guide = "none") +
+  scale_size_continuous(guide = "none") +
+  scale_x_continuous(trans = sqrt) +
+  scale_y_continuous(trans = sqrt) +
+  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Source density (dark points are closer to source)")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+
+immature_plot3
+```
+
+```{r warnings=FALSE}
+idata <- vocc_by_sp[["imm"]]
+idata <- filter(vocc_by_sp[["imm"]], distance < 50)
+
+immature_plot4 <- ggplot(filter(idata, distance < 10), aes(source_density, y = std_change , 
+    colour = end_climate, 
+    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
+    group = cell_type)) +
+  geom_jitter(width = 0.01) +
+  geom_smooth(method="lm") +
+  facet_wrap(~cell_type) +#, scales = "free") +
+  scale_alpha_continuous(guide = "none") +
+  scale_size_continuous(guide = "none") +
+  scale_x_continuous(trans = sqrt) +
+  scale_y_continuous(trans = sqrt) +
+  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Source density (dark points are closer to source)")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+
+immature_plot4
+```
+
+```{r}
+idata <- vocc_by_sp[["imm"]]
+idata <- filter(vocc_by_sp[["imm"]], distance < 50)
+
+immature_plot5 <- ggplot(idata, aes(x = end_climate - initial_value , y = (end_density - start_density
+  ), colour = cell_type, size = start_density)) +
+  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+  scale_size_continuous(guide = "none") +
+  #scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("kg/ha change in biomass over", vectors$N[1], "years")) +
+  xlab(paste("Change in", climate,"")) +
+  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "immature", species))
+immature_plot5
+```
+
+```{r}
+idata <- vocc_by_sp[["imm"]]
+
+itest <- lmerTest::lmer(std_change ~ 
+      cell_type * std_source_density + 
+      cell_type * distance + cell_type * distance2 +
+      std_end_climate + std_end_climate2 + (1 | vector_id), 
+      REML = F, data = idata)
+summary(itest)
+plot(itest)
+```
+```{r eval=FALSE}
+library(lme4)
+itest3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
+    cell_type * distance + cell_type * distance2 + 
+    std_end_climate + std_end_climate2 + (1 | vector_id), data = idata)
+summary(itest3)
+plot(itest3, ask = FALSE)
+```
+
+```{r}
+mi <- gbm::gbm(
+  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
+  data = idata, 
+  distribution = "gaussian",
+  n.trees = 2000, 
+  interaction.depth = 3, 
+  shrinkage = 0.02
+  )
+
+# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
+#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
+```
+
+```{r}
+summary(mi)
+```
+
+```{r}
+plot(mi,1)
+plot(mi,2)
+plot(mi,3)
+plot(mi,4)
+plot(mi,c(1,2))
+plot(mi,c(1,3))
+plot(mi,c(1,4))
+plot(mi,c(2,3))
+plot(mi,c(2,4))
+plot(mi,c(1,2,4))
+plot(mi,c(2,3,4))
+plot(mi,c(5,6))
+```
+
+```{r}
+  if (region == "Both odd year surveys") {
+    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 200)
+  }
+
+  if (region == "West Coast Vancouver Island") {
+    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 100)
+  }
+
+  if (region == "West Coast Haida Gwaii") {
+    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 60)
+  }
+
+  sdmTMB::plot_spde(ispde)
+```
+
+```{r }
+#glimpse(idata)
+
+idata$std_source_density2 <- idata$std_source_density^2
+itest <- sdmTMB(data = idata, 
+      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 + 
+      cell_type * distance + cell_type * distance2 + 
+      std_end_climate + std_end_climate2, 
+      spde = ispde, silent = TRUE)
+itest
 
@@
 ```{r}
-glimpse(adult_predictions)
-adult_change <- make_trend_data(adult_predictions,
-        ssid = c(4),
-        start_time = 2008,
-        skip_time = 2016,
-        input_cell_size = 2,
-        scale_fac = 1,
-        delta_t_total = 6,
-        delta_t_step = 2,
-        indices = c(1, 1, 1, 2, 2),
-        variable_names = "est_exp"
-      )
-```
+ip <- predict(itest)
+ip$residuals <- residuals(itest)
+#idata$resids <- p$residuals
+#glimpse(p)
+ggplot(ip, aes(est, residuals)) +
+  geom_point(alpha=0.4) +
+  ylim(-8,8) +
+  geom_smooth()
 
-Import vectors for single climate variable at a time
-```{r}
-variable_names <- c("DO")
+plot_map_raster <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    coord_fixed()+
+    gfplot::theme_pbs() + 
+    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+}
 
-declining_var <- "do_est"
-min_threshold <- 0.2
-vocc <- readRDS(paste0("data/", 
-      "vocc-", ssid_string, declining_var, min_threshold, "decline.rds"))
-
-# increasing_var <- "temp" 
-# max_threshold <- 0.5
-# vocc <- readRDS(paste0("data/", 
-#       "vocc-", ssid_string, increasing_var, max_threshold, "increase.rds"))
+plot_map_raster(ip, "est")
+plot_map_raster(ip, "est_rf")
+plot_map_raster(ip, "omega_s")
+plot_map_raster(ip, "residuals")
 ```
@@
 ```{r}
-glimpse(adult_change)
-#unique(adult_predictions$ssid)
-glimpse(vocc)
-
-start <- adult_change %>% rename(start_density = var_1_s, end_density = var_1_e) %>% select(icell, x, y, start_density, end_density)
-end <- adult_change %>% rename(start_density = var_1_s, end_density = var_1_e, target_X = x, target_Y = y) %>% select(icell,  start_density, end_density, target_X, target_Y)
-
-if (length(variable_names) == 1) {
-vectors <- vocc  %>% mutate(start_cell = icell, vector_id = paste0(id, "_to_", tid), 
-  initial_value = var_1_s, source_end_value = var_1_e, target_value = as.numeric(target_values)) %>% 
-  filter(distance > 1) %>%
-  select(vector_id, start_cell, x, y, target_X, target_Y, distance, initial_value, source_end_value, target_value, km_per_decade, N)
-} else {
-vectors <- vocc  %>% mutate(start_cell = icell, vector_id = paste0(id, "_to_", tid), 
-  initial_value = s, source_end_value = e, target_value = target_values) %>% 
-  filter(distance > 1) %>%
-  select(vector_id, start_cell, x, y, target_X, target_Y, distance, initial_value, source_end_value, target_value, km_per_decade, N) # need to replace N ...
-}
-vector_source <- left_join(vectors, start, by=c("x", "y")) %>% mutate(cell_type = "source")
-vector_target <- left_join(vectors, end, by=c("target_X", "target_Y")) %>% mutate(cell_type = "target")
-
-glimpse(vector_source)
-glimpse(vector_target)
-
-vectors <- rbind(vector_source, vector_target)
-glimpse(vectors)
+# png(
+#   file = paste0("figs/", spp, "/vocc-biomass-change", spp, ssid_string, ".png"),
+#   res = 600,
+#   units = "in",
+#   width = 5,
+#   height = 4
+# )
+# 
+# do_vect_plot
+# dev.off()
 ```
-
-```{r}
-do_vect_plot <- ggplot(filter(vectors, distance < 100), aes(distance, y = end_density-start_density, group = cell_type, colour= cell_type)) + 
-  geom_jitter(width = 0.25, alpha = 0.45, size = 2) + 
-  ylab(paste("Change in biomass over", vectors$N[1], "years")) + 
-  xlab(paste("Distance traveled to maintain DO decline of <", min_threshold, "")) + 
-  scale_colour_viridis_d(begin = 0.1, end = 0.5) + gfplot::theme_pbs() + theme(legend.position = c(0.8,0.8)) + ggtitle(paste(survey, species))
-do_vect_plot 
-```
-
-```{r}
-test <- lmerTest::lmer(end_density ~ cell_type * distance + start_density  + (1|vector_id), data = filter(vectors, distance < 30))
-summary(test)
-```
-
-```{r}
-png(
-  file = paste0("figs/", spp, "/vocc-biomass-change", spp, ssid_string, ".png"),
-  res = 600,
-  units = "in",
-  width = 5,
-  height = 4
-)
-
-do_vect_plot 
-dev.off()
-```

### 2019-08-02 analysis/VOCC/1-biomass-models.Rmd
minor tweaks to 1-biomass-models.rmd

@@
 # species <- "Silvergray Rockfish"
-# species <- "Lingcod"
+ species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
@@
 # species <- "Pacific Ocean Perch" 
-species <- "Yelloweye Rockfish"
+# species <- "Yelloweye Rockfish"
 ```
@@
 ```{r}
-# region <- "Both odd year surveys"
+ region <- "Both odd year surveys"
 region <- "West Coast Vancouver Island"
- region <- "West Coast Haida Gwaii"
+#  region <- "West Coast Haida Gwaii"
 ```
@@
 
-# covariates <- ""
-
+covariates <- ""
+covs <- gsub("\\+", "-", "+muddy+any_rock")
 covariates <- "+muddy+any_rock"
@@
 ```{r global_options, include=FALSE}
-covs <- gsub("\\+", "-", covariates)
+# covs <- gsub("\\+", "-", covariates)
 spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
@@
 
-Run climate independent sdmTMB model
+Load and filter data 
 ```{r}
@@
   data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)
+```
+
+Run climate independent sdmTMB model
+```{r}
 
@@
       dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
       time = "year", spde = spde,
@@
 ```{r}
-
+#covs <- gsub("\\+", "-", covariates)
 fish <- readRDS(paste0("raw/bio-data-", spp, ""))
@@
     nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
+    adult_biomass <- readRDS(paste0("data/", spp, 
+      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    ))
+
     nd <- nd_all %>% 
@@
 
-    adult_biomass <- readRDS(paste0("data/", spp, 
-      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
     adult_predictions <- predict(adult_biomass, newdata = nd)

### 2019-08-02 analysis/VOCC/3-variable-response-plots.Rmd
minor tweaks to 3-variable-response-plots.rmd

@@
 # species <- "Pacific Cod"
-species <- "Sablefish"
+# species <- "Sablefish"
 # species <- "Silvergray Rockfish"
-# species <- "Lingcod"
+ species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
@@
 # covariates <- ""
+
+try({
+  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+
+  adult_survey_all <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+
+  imm_survey_all <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+if (!exists("adult_survey_all")) adult_survey_all <- NULL
+if (!exists("imm_survey_all")) imm_survey_all <- NULL
+
 try({
@@
 model_list <- list(
+  adult_survey_all=adult_survey_all, imm_survey_all=imm_survey_all,
   adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
@@
   d[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
-  do_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Dissolved O2", xlimits = c(0,8)) + ggtitle(names(model_list[i]))
+  do_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Dissolved O2", xlimits = c(0,8)) + ggtitle(paste(species, names(model_list[i])))
 }
@@
   #d[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
-  temp_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Temperature", xlimits = c(2,10)) + ggtitle(names(temp_model_list[i]))
+  temp_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Temperature", xlimits = c(2,10)) +
+    ggtitle(paste(species,names(temp_model_list[i])))
 }
@@
   d[[i]]$x <- exp(d[[i]]$x)
-  depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0,1000)) + ggtitle(names(depth_model_list[i]))
+  depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0,1000)) + 
+    ggtitle(paste(species, names(depth_model_list[i])))
 }

### 2019-08-02 analysis/VOCC/2-biomass-trends.Rmd
new rmd making biomass trend plots using faster function that by-passes vector creation

---
title: "Biomass trends and figures"
author: "Philina English"
date: '2019-07-18'
output: html_document
---

This file calls models built by "1-biomass-models.Rmd" and stiches the predictions saved therein together.

# Choose params

Species
```{r}
# # Species run so far...
species <- "Arrowtooth Flounder"
 species <- "Pacific Cod"
species <- "Sablefish"
 species <- "Silvergray Rockfish"
species <- "Lingcod"
 species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
 species <- "Pacific Ocean Perch"
 species <- "Yelloweye Rockfish"
```

Region 
```{r}
region <- "All synoptic surveys"

# region <- "Both odd year surveys"
# region <- "West Coast Vancouver Island"
# region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
update_spatial_model <- FALSE
priors <- FALSE
# priors <- TRUE

# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
covariates <- "+muddy+any_rock"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

Bind together predictions from different models
```{r}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

      rm(pred1n3)
  rm(pred4)
rm(pred16)

      rm(pred1n3_imm)
  rm(pred4_imm)
rm(pred16_imm)

    rm(pred1n3_all)
  rm(pred4_all)
rm(pred16_all)
#if (isFALSE(update_spatial_model)) {
if (region != "All synoptic surveys") {

    try(adult_predictions <- readRDS(paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
    try(imm_predictions <- readRDS(paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))

    if (!exists("adult_predictions")) {
      paste("Maturity split biomass predictions are not available for this survey.")

      try(predictions <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))

      if (!exists("predictions")) {
        paste("Total biomass predictions for this survey are also not available.")
      }
    }

} else {
    # Bind together predictions from different models
    try({
      pred1n3 <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"))
      pred4 <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"))
      pred16 <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"))
    })
    if (!exists("pred16")) pred16 <- NULL
    if (!exists("pred4")) pred4 <- NULL
    adult_predictions <- rbind(pred1n3, pred4, pred16)

    try({
      pred1n3_imm <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"))
      pred4_imm <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))
      pred16_imm <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"))
    })

    if (!exists("pred16_imm")) pred16_imm <- NULL
    if (!exists("pred4_imm")) pred4_imm <- NULL
    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)

    if (!exists("adult_predictions")) {
      paste("Maturity split biomass predictions are not available.")

      try({
      pred1n3_all <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))
      pred4_all <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))
      pred16_all <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))
      predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
      })

      if (!exists("predictions")) paste("Total biomass predictions are not available.")
    }
}
```

Caluclate variables of interest from predictions
```{r}
if (exists("adult_predictions")) {

  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

  adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
  imm_predictions$prop_imm <- imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)

  max_raster <- quantile(adult_predictions$est_exp, .999)
  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)

  saveRDS(adult_predictions, file = paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
  saveRDS(imm_predictions, file = paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))

} else {

  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .999)
  max_bio <- round(max(predictions$total_bio))

  saveRDS(predictions, file = paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
}
```

# Plots and trends

```{r, error=TRUE}
if (exists("adult_predictions")) {

  p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
    raster_limits = c(0, max_raster),
    legend_position = "none",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
  print(p_adult_all)

} else {

  p_adult_all <- plot_facet_map(predictions, "total_bio",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))

  print(p_adult_all)
}
```

```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    adult_predictions1 <- adult_predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- adult_predictions %>% filter(ssid != 16) %>% filter(ssid != 4) 
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- adult_predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
  } else {
    adult_predictions16 <- NULL
  }

  ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)
  ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]

  p_adult <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  print(p_adult)
}
```

```{r, message=FALSE, warning=FALSE}
#try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
rm(out_mature)
if (!exists("out_mature")) {
  if (exists("adult_predictions")) {
    predictions <- adult_predictions
  } else {
    try(out_total <- readRDS(paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
    )))
  }

  if (!exists("out_total")) {
    if (1 %in% model_ssid) {
      out_mature1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = 2003,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature1 <- NULL
    }

    if (3 %in% model_ssid) {
      out_mature3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = 2005,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_mature3 <- NULL
    }

    if (4 %in% model_ssid) {
      out_mature4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature4 <- NULL
    }

    if (16 %in% model_ssid) {
      out_mature16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = 2004,
        skip_time = 2007,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_mature16 <- NULL
    }

    out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)

    saveRDS(out_mature, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"
    ))

    if (!exists("adult_predictions")) {
      out_total <- out_mature
      saveRDS(out_total, file = paste0(
        "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
      ))
    }
  }
}
```

```{r }
#FIXME: Attempting to plot of current mature biomass with trawl footprint...
# if (exists("out_mature")) {

trawl_footprint <- sf::st_read(dsn="data/trawl-footprint",layer="Trawl_footprint")
proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)

current_biomass <- plot_vocc(out_mature,
    fill_col = "var_1_e",
    fill_label = "Mean mature \nbiomass (kg/ha)\npost-2010",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
current_biomass <- current_biomass +
   geom_sf(data= trawl_footprint, aes(geometry=geometry/1000), colour = "gray", fill = "gray", alpha= 0.2, lwd=0.25, inherit.aes = FALSE) +
   coord_sf(xlim = c(180,820), ylim = c(5350,6050), datum = NULL) +
   xlab("") + 
   ylab("") +
   theme(panel.grid.major = element_line(colour = "transparent")) +
   ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
current_biomass
# }
```

```{r, error=TRUE}
if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,    
    high_fill = "deepskyblue3",#"lightseagreen", # limegreen",
    low_fill = "red", # "deeppink3",# "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))
  ))
  biotrend_ad
} else {
  biotrend_ad <- plot_vocc(out_total,
    fill_col = "units_per_decade",
    fill_label = "Biomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,    
    high_fill = "deepskyblue3", # "greenyellow", # limegreen", 
    low_fill = "red", #"steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))
  ))
  biotrend_ad
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_imm_all <- plot_facet_map(imm_predictions, "est_exp",
    transform_col = fourth_root_power,
    raster_limits = c(0, max_raster)
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n "))

  p_imm_all
} else {
  p_imm_all <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    imm_predictions1 <- imm_predictions %>% filter(ssid == 1)
  } else {
    imm_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    imm_predictions3 <- imm_predictions %>% filter(ssid != 16) %>% filter(ssid != 4) 
  } else {
    imm_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    imm_predictions4 <- imm_predictions %>% filter(ssid == 4)
  } else {
    imm_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    imm_predictions16 <- imm_predictions %>% filter(ssid == 16)
  } else {
    imm_predictions16 <- NULL
  }

  im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)
  im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]

  p_imm <- list()
  for (i in seq_len(length(model_ssid))) {
    p_imm[[i]] <- plot_facet_map(im_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))

    p_imm[[i]]
  }
  print(p_imm)
}
```

```{r, message=FALSE, warning=FALSE}
#try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
rm(out_imm)
if (!exists("out_imm")) {
  if (exists("imm_predictions")) {
    predictions <- imm_predictions

    if (1 %in% model_ssid) {
      out_im1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = 2003,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im1 <- NULL
    }

    if (3 %in% model_ssid) {
      out_im3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = 2005,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im3 <- NULL
    }

    if (4 %in% model_ssid) {
      out_im4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im4 <- NULL
    }

    if (16 %in% model_ssid) {
      out_im16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im16 <- NULL
    }

    out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
    saveRDS(out_imm, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds"
    ))
  } else {
    rm(out_imm)
  }
}
```

```{r}
if (exists("out_imm")) {
  data <- out_imm
  biotrend_im <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE, 
    high_fill = "deepskyblue3",# "greenyellow", # limegreen",
    low_fill = "red", #"steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))
  ))
  biotrend_im
} else {
  biotrend_im <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

Trend plots
```{r}
png(
  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 11,
  height = 8.5
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

Spatiotemporal biomass plots
```{r}
biomass_plots <- c(p_adult, p_imm)

png(
  file = paste0("figs/", spp, "/biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

Plot medley
```{r}
png(
  file = paste0("figs/", spp, "/exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 9.5,
  height = 12.5
)

gridExtra::grid.arrange(
  current_biomass,
  p_prop_imm,
  p_adult_all, p_imm_all,
  biotrend_ad, biotrend_im,
  nrow = 3
)
dev.off()
```

### 2019-08-02 analysis/VOCC/4-vocc-by-spp.Rmd
change N to timespan in vocc table; start drafting code to save model betas from biomass-vocc comparisons

@@
       initial_value, source_end_value, target_value, 
-      km_per_decade, N) 
+      km_per_decade, timespan) 
   if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
@@
 # range(icells$perc_change)
-# range(vocc_by_sp[[i]]$perc_change)
+# range(vocc_by_sp[[i]]$std_change)
 
@@
 Xmin <- quantile(vocc_by_sp[[i]]$X, 0.0005)
+std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
 
@@
   # scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
@@
   scale_x_continuous(trans = sqrt) +
-  # scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  scale_y_continuous(trans = sqrt) +
+  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Source density (dark points are closer to source)")) +
@@
   #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Source density (dark points are closer to source)")) +
@@
   #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("kg/ha change in biomass over", vectors$N[1], "years")) +
+  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Change in", climate,"")) +
@@
 ```{r}
+
 p <- predict(test)
@@
   scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
@@
   scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$N[1], "years")) +
+  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Source density (dark points are closer to source)")) +
@@
 ```{r}
+m <- test
+
+varnames <- colnames(m$tmb_data$X_ij)
+#b_j <- m$model$par
+#source_intercept <- .sd[1,]
+#target_intercept <- b_j[2]
+#source_slope <- b_j[3]
+
+.sd <- summary(m$sd_report)
+.sd <- as.data.frame(.sd)
+names(.sd) <- c("beta","se")
+.sd$ci <- .sd$se*1.96
+var <- varnames
+var[1] <- "intercept"
+.sd <- .sd[1:length(var), ]
+row.names(.sd) <- NULL
+.sd <-cbind(var, .sd)
+
+.sd$species <- spp 
+.sd$climate <- climate 
+.sd$min_thresholds <- min_thresholds
+.sd$max_thresholds <- max_thresholds
+.sd$optimum <- optimal_value
+.sd$start_time <- start_time
+.sd$timespan <- delta_t_total
+
+# intercept = mean std percent change in biomass for cell with avarage source biomass
+data$mean_source_density[1]
+range(data$source_density)
+```
+
+```{r}
 # png(

### 2019-08-02 R/make-vector-data.R

@@ -147,7 +147,6 @@ make_vector_data <- function(data,
     slopedat$x <- x
     slopedat$y <- y
     slopedat$icell <- icell ####
-    slopedat$N <- delta_t_step
   }
   out <- dist_based_vocc(
     start_data = start_data,
@@ -167,5 +166,6 @@ make_vector_data <- function(data,
 
   out <- left_join(out, slopedat, by = c("x", "y"))
   out$km_per_decade <- (out$distance / delta_t_total) * 10
+  out$timespan <- delta_t_total
   out
 }

### 2019-08-06 analysis/VOCC/4-vocc-by-spp.Rmd
linking vocc and species density code, still in development

@@
  species <- "Arrowtooth Flounder"
- species <- "Pacific Cod"
- species <- "Sablefish"
+# species <- "Pacific Cod"
+# species <- "Sablefish"
 # species <- "Silvergray Rockfish"
@@
 # region <- "All synoptic surveys"
- region <- "Both odd year surveys"
  region <- "West Coast Vancouver Island"
+# region <- "Both odd year surveys"
 # region <- "West Coast Haida Gwaii"
@@
 trend <- FALSE
-start_time <- 2015
-end_time <- 2018
-
-climate <- "temperature"
-temp_threshold <- c(1)
-optimal_temp <- 7 # yellow-eye 4
-
-climate <- "do"
-do_threshold <- c(0.5)
-optimal_do <- 2.5 # yellow-eye 4
-
-# climate <- "DO and temperature"
-# max_thresholds <- c(Inf, 1)
-# min_thresholds <- c(0.5, Inf)
-
+start_time <- 2008
+end_time <- 2011
+
+#climate <- "temperature"
+temp_threshold <- c(0.75)
+optimal_temp <- 6.5 # yellow-eye 4
+
+#climate <- "do"
+do_threshold <- c(0.25)
+optimal_do <- 2 # yellow-eye 4
+
+climate <- "do and temperature"
+max_thresholds <- c(Inf, 0.75)
+min_thresholds <- c(0.25, Inf)
+optimal_value <- c(6.5, 2)
 ```
@@
 library(sdmTMB)
-#library(gfranges)
+library(gfranges)
 
@@
 
-if (region == "West Coast Haida Gwaii") {
-  survey <- c("SYN WCHG")
-  model_ssid <- c(16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  trend_indices_do <- c(1, 1, 1, 2, 2)
-  trend_indices_temp <- c(1, 1, 1, 2, 2)
-  years <- NULL
-}
+# if (region == "West Coast Haida Gwaii") {
+#   survey <- c("SYN WCHG")
+#   model_ssid <- c(16)
+#   ssid_string <- paste0(model_ssid, collapse = "n")
+#   trend_indices_do <- c(1, 1, 1, 2, 2)
+#   trend_indices_temp <- c(1, 1, 1, 2, 2)
+#   years <- NULL
+# }
 ```
@@
 max_string <- paste0(max_thresholds, collapse = "-")
-climate <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
-```
-
-Calculate climate vectors
+
+optimal_string <- paste0(optimal_value, collapse = "-")
+climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate_string)))
+```
+
+# Calculate climate vectors
+
 ```{r}
@@
 try({
-  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate, "-",  
-    min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
+    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
 })
@@
 
-saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate, "-",  
-  min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
+    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+}
+# glimpse(vocc)
+```
+
+### Plot climate vectors
+```{r}
+
+if(climate== "temperature") {
+vocc$diff <- vocc$units_per_decade/10*delta_t_total
+
+vocc_plot <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+)
+vocc_plot + ggtitle(climate)
+}
+
+if (climate == "do") {
+vocc$diff <- vocc$units_per_decade/10*delta_t_total
+
+vocc_plot <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+)
+vocc_plot + ggtitle(climate)
+}
+
+if (climate == "do and temperature") {
+
+vocc$temp_diff <- vocc$temp.units_per_decade/10*delta_t_total
+
+vocc_plot_temp <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "temp_diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+)+ ggtitle("temperature")
+
+vocc$do_diff <- vocc$do_est.units_per_decade/10*delta_t_total
+
+vocc_plot_do <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "do_diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+) + ggtitle("do")
+
+plot <- gridExtra::grid.arrange(
+  vocc_plot_temp, vocc_plot_do,
+  ncol = 2,
+  top = grid::textGrob(paste(climate, "-", optimal_value, "-",
+    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+)
+
 }
@@
 biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
-```
-
-Extract biomass values to match spatial and temporal range of climate vectors
+
+# glimpse(biomass_list)
+```
+
+# Extract biomass values to match spatial and temporal range of climate vectors
 ```{r}
@@
 end <- biomass_change %>% 
-  rename(start_density = var_1_s, end_density = var_1_e, target_X = x, target_Y = y) %>% 
+  rename(start_density = var_1_s, 
+         end_density = var_1_e, 
+         target_X = x, target_Y = y) %>% 
   select(icell, start_density, end_density, target_X, target_Y)
@@
     mutate(
-      start_cell = icell, vector_id = paste0(id, "_to_", tid),
-      initial_value = var_1_s, source_end_value = var_1_e, target_value = as.numeric(target_values)
+      start_cell = icell, 
+      vector_id = paste0(id, "_to_", tid),
+      initial_value = var_1_s, 
+      source_end_value = var_1_e, 
+      target_value = as.numeric(target_values)
     ) %>%
@@
       km_per_decade, timespan)
+
   if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
@@
 } else {
+
   vectors <- vocc %>%
@@
       km_per_decade, timespan)
-  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
-  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))
-
+
+  source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), function(x) data.frame(t(as.numeric(x))))
+  vectors <- cbind(vectors, source_end)
+
+  for (j in seq_along(variable_names)){
+  vectors <- vectors %>% filter( paste0("X",j) < (optimal_value[j] - min_thresholds[j]) )
+  }
+
+}
+
+if (length(vectors[ ,2]) == 0) {
+  try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
+
+  coefs$var <- "NA"
+  coefs$beta <- "NA"
+  coefs$se <- "NA"
+  coefs$species <- spp 
+  coefs$lifestage <- "NA"
+  coefs$climate <- climate 
+  coefs$min_thresholds <- min_thresholds
+  coefs$max_thresholds <- max_thresholds
+  coefs$optimum <- optimal_value
+  coefs$start_time <- start_time
+  coefs$timespan <- delta_t_total
+  coefs$region <- region
+  coefs$ssid_string <- ssid_string
+
+  rbind(other_coefs, coefs)
+  saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")
+
+  stop("No vectors.")
 }
@@
 
+# combine source and target cell data
 vocc_by_sp[[i]] <- rbind(vector_source, vector_target) %>% 
-  filter(!is.na(icell)) %>% filter(source_density > 0.000001) %>% 
-  mutate(perc_change = (end_density - start_density)/start_density, 
-    cell_type = as.factor(cell_type)) %>% 
-  rename(X = x, Y = y)
-
+  filter(!is.na(icell)) %>% 
+# remove potential source cells with no fish  
+  #FIXME: should their be a flexible min density of fish for a source cell?
+  filter(source_density > 0.000001)
+
+# get mean and sd for densities without cell duplication
 icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
@@
 
+# calculate standardized density varibles
+vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
+  mutate(
+    perc_change = (end_density - start_density)/start_density, 
+    cell_type = as.factor(cell_type),
+    std_change = (perc_change)/sd.change, 
+    std_source_density = (source_density - mean.density)/(sd.density*2), 
+    mean_change = mean.change, sd_change = sd.change,
+    mean_source_density = mean.density, sd_density = sd.density,
+# set distance intercept at cell size
+    distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, 
+    distance2 = distance*distance) %>% 
+  rename(X = x, Y = y)
+
+if (length(variable_names) == 1) {
 mean.climate <- mean(icells$end_climate, na.rm = TRUE)
 sd.climate <- sd(icells$end_climate, na.rm = TRUE)
+
+vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
+  mutate(
+    mean_climate = mean.climate, sd_climate = sd.climate,
+    std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
+    std_end_climate2 = std_end_climate^2
+    )
+
+}else{
+
+if (length(variable_names) == 2) {
+
+vectors <- vectors %>% rename(
+  source_end_climateA = X1, source_end_climateB = X2
+  )
+
+mean.climateA <- mean(icells$source_end_climateA, na.rm = TRUE)
+sd.climateA <- sd(icells$source_end_climateA, na.rm = TRUE)
+mean.climateB <- mean(icells$source_end_climateB, na.rm = TRUE)
+sd.climateB <- sd(icells$source_end_climateB, na.rm = TRUE)
+
+vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
+  mutate(
+    mean_climateA = mean.climateA, sd_climate = sd.climateA,
+    std_end_climateA = (source_end_climateA - mean.climateA)/(sd.climateA*2),
+    std_end_climateA2 = std_end_climateA^2,
+    mean_climateB = mean.climateB, sd_climate = sd.climateB,
+    std_end_climateB = (source_end_climateB - mean.climateB)/(sd.climateB*2),
+    std_end_climateB2 = std_end_climateB^2
+    )
+} else {
+ stop("Doesn't currently support testing vectors bases on more than two climate variables at once.")
+}
+}
+
+}
+names(vocc_by_sp) <- names(biomass_list)
 
@@
 # range(vocc_by_sp[[i]]$std_change)
-
-Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
-Xmin <- quantile(vocc_by_sp[[i]]$X, 0.0005)
-std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
-
-vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% mutate(std_change = (perc_change)/sd.change, 
-  std_source_density = (source_density - mean.density)/(sd.density*2), 
-  std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
-  mean_change = mean.change, sd_change = sd.change,
-  mean_source_density = mean.density, sd_density = sd.density,
-  mean_climate = mean.climate, sd_climate = sd.climate,
-  std_end_climate2 = std_end_climate^2, 
-  distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, # set intercept at scale of cells
-  distance2 = distance*distance
-  ) %>% filter ( X >= Xmin & X <= Xmax )
-
 # range(vocc_by_sp[[i]]$X)
 
-#glimpse(vocc_by_sp[[i]])
-}
-names(vocc_by_sp) <- names(biomass_list)
-```
+# glimpse(vocc_by_sp[[i]])
+# trim spatial outliers 
+# Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
+# Xmin <- quantile(vocc_by_sp[[i]]$X, 0.005)
+# std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
+# 
+# vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax )
+
+```
+
+# Adult
+
+## Adult plots
 
@@
 
-mature_plot2 <- ggplot(data, aes(distance, y = std_change, 
+mature_plot2 <- ggplot(data, aes(x = distance, y = std_change, 
   group = cell_type, size = start_density, colour = end_climate)) +
@@
   geom_jitter(width = 0.01) +
-  geom_smooth(method="lm") +
+  geom_smooth(method="lm", aes(group = cell_type)) +
   facet_wrap(~cell_type) +#, scales = "free") +
@@
 
+## Adult models
+
+### Regular mixed-model
+
 ```{r}
@@
 
-Boosted regression
+### Boosted regression
 ```{r}
@@
 
+### sdmTMB
+
 ```{r}
 data <- vocc_by_sp[["adult"]]
+# Xmax <- quantile(data$X, 0.9995)
+# Xmin <- quantile(data$X, 0.005)
+# std_max <- quantile(data$std_change, 0.9995)
+# 
+# data <- data %>% filter ( X >= Xmin & X <= Xmax )
+
+xy <- as.data.frame(cbind(data$X, data$Y))
 
   if (region == "Both odd year surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
+    n_knots <- 200
   }
@@
   if (region == "West Coast Vancouver Island") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 100)
-  }
-
-  if (region == "West Coast Haida Gwaii") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
-  }
-
+    n_knots <- 100
+  }
+
+  # if (region == "West Coast Haida Gwaii") {
+  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
+  # }
+
+if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(xy))}
+
+  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = n_knots)
   sdmTMB::plot_spde(spde)
+
 ```
@@
 ```
+
+# Immature
+
+### Immature plots
 
@@
 
+## Immature models
+
+### Regular mixed-model
 ```{r}
@@
 
+# Xmax <- quantile(idata$X, 0.9995)
+# Xmin <- quantile(idata$X, 0.005)
+# std_max <- quantile(idata$std_change, 0.9995)
+# 
+# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )
+#  
 itest <- lmerTest::lmer(std_change ~ 
@@
 
-```{r}
+### Boosted regression
+
+```{r}
+idata <- vocc_by_sp[["imm"]]
+
+# Xmax <- quantile(idata$X, 0.9995)
+# Xmin <- quantile(idata$X, 0.005)
+# std_max <- quantile(idata$std_change, 0.9995)
+# 
+# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )
+
 mi <- gbm::gbm(
@@
 
-```{r}
+### sdmTMB
+
+```{r}
+idata <- vocc_by_sp[["imm"]]
+
+# Xmax <- quantile(data$X, 0.9995)
+# Xmin <- quantile(data$X, 0.005)
+# std_max <- quantile(data$std_change, 0.9995)
+# 
+# data <- data %>% filter ( X >= Xmin & X <= Xmax )
+
+ixy <- as.data.frame(cbind(idata$X, idata$Y))
+
   if (region == "Both odd year surveys") {
-    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 200)
+    n_knots <- 200
   }
@@
   if (region == "West Coast Vancouver Island") {
-    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 100)
-  }
-
-  if (region == "West Coast Haida Gwaii") {
-    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 60)
-  }
-
+    n_knots <- 100
+  }
+
+  # if (region == "West Coast Haida Gwaii") {
+  #    n_knots <- 60
+  # }
+
+if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(ixy))}
+
+  ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = n_knots)
   sdmTMB::plot_spde(ispde)
+
 ```
@@
 varnames <- colnames(m$tmb_data$X_ij)
-#b_j <- m$model$par
-#source_intercept <- .sd[1,]
-#target_intercept <- b_j[2]
-#source_slope <- b_j[3]
-
-.sd <- summary(m$sd_report)
-.sd <- as.data.frame(.sd)
-names(.sd) <- c("beta","se")
-.sd$ci <- .sd$se*1.96
+
+coefs <- summary(m$sd_report)
+coefs <- as.data.frame(coefs)
+names(coefs) <- c("beta","se")
+coefs$ci <- coefs$se*1.96
 var <- varnames
 var[1] <- "intercept"
-.sd <- .sd[1:length(var), ]
-row.names(.sd) <- NULL
-.sd <-cbind(var, .sd)
-
-.sd$species <- spp 
-.sd$climate <- climate 
-.sd$min_thresholds <- min_thresholds
-.sd$max_thresholds <- max_thresholds
-.sd$optimum <- optimal_value
-.sd$start_time <- start_time
-.sd$timespan <- delta_t_total
-
-# intercept = mean std percent change in biomass for cell with avarage source biomass
+coefs <- coefs[1:length(var), ]
+row.names(coefs) <- NULL
+
+coefs <-cbind(var, coefs)
+#coefs$var
+#coefs$beta
+#coefs$se
+coefs$species <- spp 
+coefs$lifestage <- "adult"
+coefs$climate <- climate 
+coefs$min_thresholds <- min_thresholds
+coefs$max_thresholds <- max_thresholds
+coefs$optimum <- optimal_value
+coefs$start_time <- start_time
+coefs$timespan <- delta_t_total
+coefs$region <- region
+coefs$ssid_string <- ssid_string
+
+glimpse(coefs)
+rm(other_coefs)
+try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
+
+if (exists("other_coefs")) {
+coefs <- rbind(other_coefs, coefs)
+}
+saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")
+
 data$mean_source_density[1]
 range(data$source_density)
+```
+
+```{r}
+m <- itest
+
+varnames <- colnames(m$tmb_data$X_ij)
+
+coefs <- summary(m$sd_report)
+coefs <- as.data.frame(coefs)
+names(coefs) <- c("beta","se")
+coefs$ci <- coefs$se*1.96
+var <- varnames
+var[1] <- "intercept"
+coefs <- coefs[1:length(var), ]
+row.names(coefs) <- NULL
+coefs <-cbind(var, coefs)
+
+coefs$species <- spp 
+coefs$lifestage <- "immature"
+coefs$climate <- climate 
+coefs$min_thresholds <- min_thresholds
+coefs$max_thresholds <- max_thresholds
+coefs$optimum <- optimal_value
+coefs$start_time <- start_time
+coefs$timespan <- delta_t_total
+coefs$region <- region
+coefs$ssid_string <- ssid_string
+
+glimpse(coefs)
+try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
+if (exists("other_coefs")) {
+coefs <- rbind(other_coefs, coefs)
+}
+saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")
+
+# intercept = mean std percent change in biomass for cell with avarage source biomass
+idata$mean_source_density[1]
+range(idata$source_density)
 ```

### 2019-08-06 analysis/VOCC/4-vocc-by-spp.Rmd
get two varible vocc plots working

@@
 min_thresholds <- c(0.25, Inf)
-optimal_value <- c(6.5, 2)
+optimal_value <- c(2, 6.5)
 ```
@@
 optimal_string <- paste0(optimal_value, collapse = "-")
-climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate_string)))
+climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
 ```
@@
 
-plot <- gridExtra::grid.arrange(
+vocc_plot <- gridExtra::grid.arrange(
   vocc_plot_temp, vocc_plot_do,
   ncol = 2,
-  top = grid::textGrob(paste(climate, "-", optimal_value, "-",
-    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+  top = grid::textGrob(paste("do of ", optimal_value[1], "less", min_thresholds[1],"or temperature of ", optimal_value[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
 )
@@
 }
-
+vocc_plot
 ```
@@
   for (j in seq_along(variable_names)){
-  vectors <- vectors %>% filter( paste0("X",j) < (optimal_value[j] - min_thresholds[j]) )
+  browser()
+    vectors <-  filter( vectors, paste0("X",j) < (optimal_value[j] - min_thresholds[j]) )
   }
@@
   try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
-
+  coefs <- other_coefs[1,]
   coefs$var <- "NA"
@@
   coefs$climate <- climate 
-  coefs$min_thresholds <- min_thresholds
-  coefs$max_thresholds <- max_thresholds
-  coefs$optimum <- optimal_value
+  coefs$min_thresholds <- min_string
+  coefs$max_thresholds <- max_string
+  coefs$optimum <- optimal_string
   coefs$start_time <- start_time
@@
 ## Immature models
-
 ### Regular mixed-model
+
 ```{r}

### 2019-08-06 analysis/VOCC/2-biomass-trends.Rmd
minor edits to model building rmds

@@
  species <- "Arrowtooth Flounder"
- species <- "Pacific Cod"
-species <- "Sablefish"
- species <- "Silvergray Rockfish"
-species <- "Lingcod"
- species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
+#  species <- "Pacific Cod"
+# species <- "Sablefish"
+#  species <- "Silvergray Rockfish"
+# species <- "Lingcod"
+#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
 # species <- "Quillback Rockfish"
- species <- "Pacific Ocean Perch"
- species <- "Yelloweye Rockfish"
+# species <- "Pacific Ocean Perch"
+# species <- "Yelloweye Rockfish"
 ```
@@
 
-  max_raster <- quantile(adult_predictions$est_exp, .999)
+  max_raster <- quantile(adult_predictions$est_exp, 0.999)
   max_adult <- signif(max(adult_predictions$est_exp), digits = 2)

### 2019-08-06 analysis/VOCC/3-variable-response-plots.Rmd
minor edits to model building rmds

@@
 ```{r}
-
+#species <- params$species
 # # Species run so far...
+species <- "Arrowtooth Flounder"
 # species <- "Pacific Cod"
@@
 # species <- "Silvergray Rockfish"
- species <- "Lingcod"
+# species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
+
 # species <- "Quillback Rockfish"
 # species <- "Pacific Ocean Perch"
-
 # species <- "Yelloweye Rockfish"
-# species <- "Arrowtooth Flounder"
-
 ```

### 2019-08-06 analysis/VOCC/vocc-example.Rmd
example vocc rmd

--- 2019-08-02 analysis/VOCC/4-vocc-by-spp.Rmd
@@
 ---
-title: "Biomass predictions and VOCC"
+title: "vocc example"
 author: "Philina English"
-date: '2019-07-18'
+date: "06/08/2019"
 output: html_document
@@
 
-This file ...
-
-Species
-
+Load climate predictions
 ```{r}
-# # Species run so far...
- species <- "Arrowtooth Flounder"
- species <- "Pacific Cod"
- species <- "Sablefish"
-# species <- "Silvergray Rockfish"
-#  species <- "Lingcod"
-#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-#  species <- "Quillback Rockfish"
-#  species <- "Pacific Ocean Perch"
-# species <- "Yelloweye Rockfish"
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+pred_temp_do <- readRDS(here::here("predicted-DO.rds"))
 ```
 
-Region 
+Choose settings spatial and temporal settings
 ```{r}
-# region <- "All synoptic surveys"
- region <- "Both odd year surveys"
 region <- "West Coast Vancouver Island"
-# region <- "West Coast Haida Gwaii"
+trend <- FALSE
+start_time <- 2008
+end_time <- 2011
 ```
 
-Make climate vectors for biannual changes between surveys
+Choose what variables to base vectors on 
 ```{r}
-trend <- FALSE
-start_time <- 2015
-end_time <- 2018
+# JUST TEMPERATURE
+# climate <- "temperature"
+# temp_threshold <- c(0.75)
+# optimal_temp <- 6.5
 
-climate <- "temperature"
-temp_threshold <- c(1)
-optimal_temp <- 7 # yellow-eye 4
+# JUST DO
+# climate <- "do"
+# do_threshold <- c(0.25)
+# optimal_do <- 2
 
-climate <- "do"
-do_threshold <- c(0.5)
-optimal_do <- 2.5 # yellow-eye 4
-
-# climate <- "DO and temperature"
-# max_thresholds <- c(Inf, 1)
-# min_thresholds <- c(0.5, Inf)
-
-```
-
-Make climate vectors for average changes and lower threshold values
-```{r}
-# trend <- TRUE
-# 
-# climate <- "temperature"
-# temp_trend_threshold <- c(0.25)
-
-# climate <- "DO"
-# do_trend_threshold <- c(0.2)
-
-# climate <- "both DO and temperature"
-# min_trend_thresholds <- c(0.2, Inf)
-# max_trend_thresholds <- c(Inf, 0.5)
+# BOTH
+climate <- "do and temperature"
+max_thresholds <- c(Inf, 0.75)
+min_thresholds <- c(0.25, Inf)
+optimal_value <- c(2, 6.5)
 ```
@@
 library(gfplot)
-library(gfdata)
-library(sdmTMB)
-#library(gfranges)
-
-knitr::opts_chunk$set(
-  fig.width = 11, fig.height = 8.5,
-  echo = FALSE, warning = FALSE, message = FALSE
-)
+library(gfranges)
 ```
@@
 max_string <- paste0(max_thresholds, collapse = "-")
-climate <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
+optimal_string <- paste0(optimal_value, collapse = "-")
+climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
 ```
 
-Calculate climate vectors
+# Calculate climate vectors
+
 ```{r}
-rm(vocc)
-try({
-  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate, "-",  
-    min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
-})
-
-if (!exists("vocc")) {
-pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
 
@@
 
-saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate, "-",  
-  min_string,"-", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+glimpse(vocc)
+```
+
+### Plot climate vectors
+```{r}
+
+if(climate== "temperature") {
+vocc$diff <- vocc$units_per_decade/10*delta_t_total
+
+vocc_plot <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans
+)
+vocc_plot + ggtitle(climate)
 }
 
-```
+if (climate == "do") {
+vocc$diff <- vocc$units_per_decade/10*delta_t_total
 
-Get matching biomass data for the above time periods
-
-Get all biomass predictions 
-```{r}
-priors <- FALSE
-# priors <- TRUE
-
-covariates <- "+muddy+any_rock"
-
-covs <- gsub("\\+", "-", covariates)
-
-rm(adult_predictions)
-rm(imm_predictions)
-rm(predictions)
-
-try({
-  adult_predictions <- readRDS(paste0("data/", spp,
-  "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
-})
-
-try({
-  imm_predictions <- readRDS(paste0("data/", spp,
-  "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))
-})
-
-try({
-  predictions <- readRDS(paste0("data/", spp,
-  "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
-})
-
-if (!exists("adult_predictions")) adult_predictions <- NULL
-if (!exists("imm_predictions")) imm_predictions <- NULL
-if (!exists("predictions")) predictions <- NULL
-
-biomass_list <- list(
-  adult = adult_predictions, 
-  imm = imm_predictions, 
-  total = predictions
+vocc_plot <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans
 )
-
-biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
-```
-
-Extract biomass values to match spatial and temporal range of climate vectors
-```{r}
-vocc_by_sp <- list()
-for (i in seq_along(biomass_list)){
-
-biomass_change <- make_trend_data(biomass_list[[i]],
-  ssid = model_ssid,
-  start_time = start_time,
-  skip_time = skip_time,
-  end_time = end_time,
-  input_cell_size = input_cell_size,
-  scale_fac = scale_fac,
-  delta_t_total = delta_t_total,
-  delta_t_step = delta_t_step,
-  indices = indices,
-  variable_names = "est_exp"
-)
-
-start <- biomass_change %>% 
-  rename(start_density = var_1_s, end_density = var_1_e) %>% 
-  select(icell, x, y, start_density, end_density)
-
-end <- biomass_change %>% 
-  rename(start_density = var_1_s, end_density = var_1_e, target_X = x, target_Y = y) %>% 
-  select(icell, start_density, end_density, target_X, target_Y)
-
-source_density <- biomass_change %>% rename(source_density = var_1_s) %>% select(source_density, x, y)
-
-if (length(variable_names) == 1) {
-  vectors <- vocc %>%
-    mutate(
-      start_cell = icell, vector_id = paste0(id, "_to_", tid),
-      initial_value = var_1_s, source_end_value = var_1_e, target_value = as.numeric(target_values)
-    ) %>%
-    filter(distance > 1)  %>%
-    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
-      initial_value, source_end_value, target_value, 
-      km_per_decade, timespan) 
-  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
-  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))
-
-} else {
-  vectors <- vocc %>%
-    mutate(
-      start_cell = icell, vector_id = paste0(id, "_to_", tid),
-      initial_value = s, source_end_value = e, target_value = target_values
-    ) %>%
-    filter(distance > 1) %>%
-    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
-      initial_value, source_end_value, target_value, 
-      km_per_decade, timespan) 
-  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
-  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))
-
+vocc_plot + ggtitle(climate)
 }
 
-vector_source <- left_join(vectors, start, by = c("x", "y")) %>% 
-  mutate(cell_type = "source", end_climate = source_end_value, source_density = start_density)
+if (climate == "do and temperature") {
 
-vector_target <- left_join(vectors, end, by = c("target_X", "target_Y")) %>% 
-  mutate(cell_type = "target", end_climate = target_value)
-vector_target <- left_join(vector_target, source_density, by = c("x", "y")) #%>%  
-  # filter(end_climate > 2)
+vocc$temp_diff <- vocc$temp.units_per_decade/10*delta_t_total
 
-vocc_by_sp[[i]] <- rbind(vector_source, vector_target) %>% 
-  filter(!is.na(icell)) %>% filter(source_density > 0.000001) %>% 
-  mutate(perc_change = (end_density - start_density)/start_density, 
-    cell_type = as.factor(cell_type)) %>% 
-  rename(X = x, Y = y)
+vocc_plot_temp <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "temp_diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans
+)+ ggtitle(paste("Temperature within", max_thresholds[2], "°C"))
 
-icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
-mean.change <- mean(icells$perc_change, na.rm = TRUE)
-sd.change <- sd(icells$perc_change, na.rm = TRUE)
+vocc$do_diff <- vocc$do_est.units_per_decade/10*delta_t_total
 
-mean.density <- mean(icells$source_density, na.rm = TRUE)
-sd.density <- sd(icells$source_density, na.rm = TRUE)
+vocc_plot_do <- plot_vocc(vocc,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "do_diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans
+) + ggtitle(paste("DO within", min_thresholds[1], "ml/L") )
 
-mean.climate <- mean(icells$end_climate, na.rm = TRUE)
-sd.climate <- sd(icells$end_climate, na.rm = TRUE)
-
-# glimpse(icells)
-# range(icells$perc_change)
-# range(vocc_by_sp[[i]]$std_change)
-
-Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
-Xmin <- quantile(vocc_by_sp[[i]]$X, 0.0005)
-std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
-
-vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% mutate(std_change = (perc_change)/sd.change, 
-  std_source_density = (source_density - mean.density)/(sd.density*2), 
-  std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
-  mean_change = mean.change, sd_change = sd.change,
-  mean_source_density = mean.density, sd_density = sd.density,
-  mean_climate = mean.climate, sd_climate = sd.climate,
-  std_end_climate2 = std_end_climate^2, 
-  distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, # set intercept at scale of cells
-  distance2 = distance*distance
-  ) %>% filter ( X >= Xmin & X <= Xmax )
-
-# range(vocc_by_sp[[i]]$X)
-
-#glimpse(vocc_by_sp[[i]])
+vocc_plot <- gridExtra::grid.arrange(
+  vocc_plot_temp, vocc_plot_do,
+  ncol = 2,
+  top = grid::textGrob("VOCC: vectors based on two climate variables at once")
+)
 }
-names(vocc_by_sp) <- names(biomass_list)
+vocc_plot
 ```
-
-Check final climate-biomass relationship
-```{r}
-data <- vocc_by_sp[["adult"]]
-#glimpse(data)
-
-mature_plot <- ggplot(data, aes(x = end_climate, y = (end_density # - start_density
-  ), colour = cell_type, #size = start_density, 
-  group = cell_type)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
-  #   , aes(x = end_climate, y = (end_density #- start_density
-  #  )), inherit.aes = FALSE) +
-  scale_size_continuous(guide = "none") +
-  scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Final biomass")) +
-  xlab(paste("Final value of", climate,"for all cells")) +
-  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
-  ggtitle(paste(survey, "mature", species))
-
-mature_plot
-```
-
-Check distance threshold of 50 looks reasonable
-```{r}
-data <- vocc_by_sp[["adult"]]
-#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)
-
-mature_plot2 <- ggplot(data, aes(distance, y = std_change, 
-  group = cell_type, size = start_density, colour = end_climate)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y~x+I(x^2) )+
-  facet_wrap(~cell_type) +#, scales = "free") +
-  scale_size_continuous(guide = "none") +
-  # scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
-  ggtitle(paste(survey, "mature", species))
-
-mature_plot2
-```
-
-```{r warnings=FALSE}
-data <- vocc_by_sp[["adult"]]
-#data <- filter(vocc_by_sp[["adult"]], distance < 50)
-#data <- filter(vocc_by_sp[["adult"]], distance > 4 & distance < 50)
-
-mature_plot3 <- ggplot(data, aes(x = source_density, y = std_change, 
-    colour = end_climate, 
-    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
-    group = cell_type)) +
-  geom_jitter(width = 0.01) +
-  geom_smooth(method="lm") +
-  facet_wrap(~cell_type) + #, scales = "free") +
-  scale_alpha_continuous(guide = "none") +
-  scale_size_continuous(guide = "none") +
-  scale_x_continuous(trans = sqrt) +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-mature_plot3
-```
-
-```{r}
-data <- vocc_by_sp[["adult"]]
-data <- filter(vocc_by_sp[["adult"]], distance < 50)
-
-mature_plot4 <- ggplot(filter(data, distance < 10), aes(source_density, y = std_change , 
-    colour = end_climate, 
-    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
-    group = cell_type)) +
-  geom_jitter(width = 0.01) +
-  geom_smooth(method="lm") +
-  facet_wrap(~cell_type) +#, scales = "free") +
-  scale_alpha_continuous(guide = "none") +
-  scale_size_continuous(guide = "none") +
-  scale_x_continuous(trans = sqrt) +
-  #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-mature_plot4
-```
-
-```{r}
-data <- vocc_by_sp[["adult"]]
-data <- filter(vocc_by_sp[["adult"]], distance < 50)
-
-mature_plot5 <- ggplot(data, aes(x = end_climate - initial_value , y = #std_change
-    (end_density - start_density)
-  , colour = cell_type, size = start_density)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  scale_size_continuous(guide = "none") +
-  #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Change in", climate,"")) +
-  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "mature", species))
-mature_plot5
-```
-
-```{r}
-# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
-#             std_end_climate + std_end_climate2 + 
-#             (1 | vector_id), 
-#  #   offset = log(start_density), 
-#  #   family = Gamma(link = "inverse")
-#      data = data)
-# hist(data$end_density)
-# hist(data$std_change)
-data <- vocc_by_sp[["adult"]]
-
-test2 <- lmerTest::lmer(std_change ~ cell_type * std_source_density + 
-    cell_type * distance + cell_type * distance2 +
-    std_end_climate + std_end_climate2 + (1 | vector_id), REML = F, data = data)
-
-summary(test2)
-plot(test2)
-```
-
-Robust lmm
-```{r eval=FALSE}
-library(lme4)
-data <- vocc_by_sp[["adult"]]
-
-test3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
-    cell_type * distance + cell_type * distance2 + 
-    std_end_climate + std_end_climate2 + (1 | vector_id), data = data)
-
-summary(test3)
-plot(test3, ask = FALSE)
-```
-
-Boosted regression
-```{r}
-data <- vocc_by_sp[["adult"]]
-
-m <- gbm::gbm(
-  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
-  data = data, 
-  distribution = "gaussian",
-  n.trees = 2000, 
-  interaction.depth = 3, 
-  shrinkage = 0.02
-  )
-
-# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
-#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
-```
-
-```{r}
-summary(m)
-```
-
-```{r}
-plot(m,1)
-plot(m,2)
-plot(m,3)
-plot(m,4)
-plot(m,c(1,2))
-plot(m,c(1,3))
-plot(m,c(1,4))
-plot(m,c(2,3))
-plot(m,c(2,4))
-plot(m,c(1,2,4))
-plot(m,c(2,3,4))
-plot(m,c(5,6))
-```
-
-```{r}
-data <- vocc_by_sp[["adult"]]
-
-  if (region == "Both odd year surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-  }
-
-  if (region == "West Coast Vancouver Island") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 100)
-  }
-
-  if (region == "West Coast Haida Gwaii") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
-  }
-
-  sdmTMB::plot_spde(spde)
-```
-
-```{r}
-#data$end_climate2 <- data$end_climate^2
-#data <- vocc_by_sp[["adult"]]
-data$std_source_density2 <- data$std_source_density^2
-test <- sdmTMB(data = data,
-      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 +
-    cell_type * distance + cell_type * distance2 + 
-      std_end_climate + std_end_climate2, 
-      # std_change ~ cell_type * source_density + cell_type * distance  + end_climate + end_climate2,
-      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = TRUE)
-test
-```
-
-```{r}
-
-p <- predict(test)
-p$residuals <- residuals(test)
-#idata$resids <- p$residuals
-#glimpse(p)
-ggplot(p, aes(est, residuals)) +
-  geom_point(alpha=0.4) +
-  ylim(-8,8) +
-  geom_smooth()
-
-plot_map_raster <- function(dat, column = "est") {
-    ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    coord_fixed()+
-    gfplot::theme_pbs() + 
-    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
-}
-
-plot_map_raster(p, "est")
-plot_map_raster(p, "est_rf")
-plot_map_raster(p, "omega_s")
-plot_map_raster(p, "residuals")
-
-```
-
-Check final climate-biomass relationship
-```{r}
-idata <- vocc_by_sp[["imm"]]
-#glimpse(data)
-
-immature_plot1 <- ggplot(idata, aes(x = end_climate, y = (end_density # - start_density
-  ), colour = cell_type, #size = start_density, 
-  group = cell_type)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
-  #geom_smooth() +
-  scale_size_continuous(guide = "none") +
-  scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Final biomass")) +
-  xlab(paste("Final value of", climate,"for all cells")) +
-  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
-  ggtitle(paste(survey, "immature", species))
-
-immature_plot1
-```
-
-Check distance threshold of 50 looks reasonable
-```{r warnings=FALSE}
-idata <- vocc_by_sp[["imm"]]
-#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)
-
-immature_plot2 <- ggplot(idata, aes(distance, y = std_change, 
-  group = cell_type, size = start_density, colour = end_climate)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y~x+I(x^2) )+
-  facet_wrap(~cell_type) + #, scales = "free") +
-  scale_size_continuous(guide = "none") +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species))
-
-immature_plot2
-```
-
-```{r warnings=FALSE}
-idata <- vocc_by_sp[["imm"]]
-#idata <- filter(vocc_by_sp[["imm"]], distance < 50)
-#idata <- filter(vocc_by_sp[[""imm"]], distance > 4 & distance < 50)
-
-immature_plot3 <- ggplot(idata, aes(x = source_density, y = std_change, 
-    colour = end_climate, 
-    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
-    group = cell_type)) +
-  geom_jitter(width = 0.01) +
-  geom_smooth(method="lm") +
-  facet_wrap(~cell_type) + #, scales = "free") +
-  scale_alpha_continuous(guide = "none") +
-  scale_size_continuous(guide = "none") +
-  scale_x_continuous(trans = sqrt) +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-immature_plot3
-```
-
-```{r warnings=FALSE}
-idata <- vocc_by_sp[["imm"]]
-idata <- filter(vocc_by_sp[["imm"]], distance < 50)
-
-immature_plot4 <- ggplot(filter(idata, distance < 10), aes(source_density, y = std_change , 
-    colour = end_climate, 
-    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
-    group = cell_type)) +
-  geom_jitter(width = 0.01) +
-  geom_smooth(method="lm") +
-  facet_wrap(~cell_type) +#, scales = "free") +
-  scale_alpha_continuous(guide = "none") +
-  scale_size_continuous(guide = "none") +
-  scale_x_continuous(trans = sqrt) +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-immature_plot4
-```
-
-```{r}
-idata <- vocc_by_sp[["imm"]]
-idata <- filter(vocc_by_sp[["imm"]], distance < 50)
-
-immature_plot5 <- ggplot(idata, aes(x = end_climate - initial_value , y = (end_density - start_density
-  ), colour = cell_type, size = start_density)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  scale_size_continuous(guide = "none") +
-  #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Change in", climate,"")) +
-  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species))
-immature_plot5
-```
-
-```{r}
-idata <- vocc_by_sp[["imm"]]
-
-itest <- lmerTest::lmer(std_change ~ 
-      cell_type * std_source_density + 
-      cell_type * distance + cell_type * distance2 +
-      std_end_climate + std_end_climate2 + (1 | vector_id), 
-      REML = F, data = idata)
-summary(itest)
-plot(itest)
-```
-```{r eval=FALSE}
-library(lme4)
-itest3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
-    cell_type * distance + cell_type * distance2 + 
-    std_end_climate + std_end_climate2 + (1 | vector_id), data = idata)
-summary(itest3)
-plot(itest3, ask = FALSE)
-```
-
-```{r}
-mi <- gbm::gbm(
-  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
-  data = idata, 
-  distribution = "gaussian",
-  n.trees = 2000, 
-  interaction.depth = 3, 
-  shrinkage = 0.02
-  )
-
-# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
-#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
-```
-
-```{r}
-summary(mi)
-```
-
-```{r}
-plot(mi,1)
-plot(mi,2)
-plot(mi,3)
-plot(mi,4)
-plot(mi,c(1,2))
-plot(mi,c(1,3))
-plot(mi,c(1,4))
-plot(mi,c(2,3))
-plot(mi,c(2,4))
-plot(mi,c(1,2,4))
-plot(mi,c(2,3,4))
-plot(mi,c(5,6))
-```
-
-```{r}
-  if (region == "Both odd year surveys") {
-    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 200)
-  }
-
-  if (region == "West Coast Vancouver Island") {
-    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 100)
-  }
-
-  if (region == "West Coast Haida Gwaii") {
-    ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = 60)
-  }
-
-  sdmTMB::plot_spde(ispde)
-```
-
-```{r }
-#glimpse(idata)
-
-idata$std_source_density2 <- idata$std_source_density^2
-itest <- sdmTMB(data = idata, 
-      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 + 
-      cell_type * distance + cell_type * distance2 + 
-      std_end_climate + std_end_climate2, 
-      spde = ispde, silent = TRUE)
-itest
-
-```
-
-```{r}
-ip <- predict(itest)
-ip$residuals <- residuals(itest)
-#idata$resids <- p$residuals
-#glimpse(p)
-ggplot(ip, aes(est, residuals)) +
-  geom_point(alpha=0.4) +
-  ylim(-8,8) +
-  geom_smooth()
-
-plot_map_raster <- function(dat, column = "est") {
-    ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    coord_fixed()+
-    gfplot::theme_pbs() + 
-    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
-}
-
-plot_map_raster(ip, "est")
-plot_map_raster(ip, "est_rf")
-plot_map_raster(ip, "omega_s")
-plot_map_raster(ip, "residuals")
-```
-
-```{r}
-m <- test
-
-varnames <- colnames(m$tmb_data$X_ij)
-#b_j <- m$model$par
-#source_intercept <- .sd[1,]
-#target_intercept <- b_j[2]
-#source_slope <- b_j[3]
-
-.sd <- summary(m$sd_report)
-.sd <- as.data.frame(.sd)
-names(.sd) <- c("beta","se")
-.sd$ci <- .sd$se*1.96
-var <- varnames
-var[1] <- "intercept"
-.sd <- .sd[1:length(var), ]
-row.names(.sd) <- NULL
-.sd <-cbind(var, .sd)
-
-.sd$species <- spp 
-.sd$climate <- climate 
-.sd$min_thresholds <- min_thresholds
-.sd$max_thresholds <- max_thresholds
-.sd$optimum <- optimal_value
-.sd$start_time <- start_time
-.sd$timespan <- delta_t_total
-
-# intercept = mean std percent change in biomass for cell with avarage source biomass
-data$mean_source_density[1]
-range(data$source_density)
-```
-
-```{r}
-# png(
-#   file = paste0("figs/", spp, "/vocc-biomass-change", spp, ssid_string, ".png"),
-#   res = 600,
-#   units = "in",
-#   width = 5,
-#   height = 4
-# )
-# 
-# do_vect_plot
-# dev.off()
-```

### 2019-08-07 analysis/VOCC/vocc-example.Rmd
Make average (trend-based) vocc work as well and revise plot titles

@@
 region <- "West Coast Vancouver Island"
-trend <- FALSE
+trend <- TRUE # IF TREND will automatically apply different thresholds for average change
+
+# set years if not using an average trend
+if (!trend) { 
 start_time <- 2008
 end_time <- 2011
+} 
 ```
@@
 # JUST TEMPERATURE
-# climate <- "temperature"
-# temp_threshold <- c(0.75)
-# optimal_temp <- 6.5
+climate <- "temperature"
+temp_threshold <- c(0.75)
+optimal_temp <- 6.5
+temp_trend_threshold <- c(0.25)
 
-# JUST DO
+# # JUST DO
 # climate <- "do"
@@
 # optimal_do <- 2 
+# do_trend_threshold <- c(0.1)
 
-# BOTH
-climate <- "do and temperature"
-max_thresholds <- c(Inf, 0.75)
-min_thresholds <- c(0.25, Inf)
-optimal_value <- c(2, 6.5)
+# # BOTH
+# climate <- "do and temperature"
+# max_thresholds <- c(Inf, 0.75)
+# min_thresholds <- c(0.25, Inf)
+# optimal_value <- c(2, 6.5) 
+# # should be lower than used for raw changes
+# min_trend_thresholds <- c(0.1, Inf) 
+# max_trend_thresholds <- c(Inf, 0.25)
 ```
@@
   transform_col = no_trans
-)
-vocc_plot + ggtitle(climate)
+) + ggtitle(paste("Temperature within", max_thresholds, "°C"))
 }
@@
   transform_col = no_trans
-)
-vocc_plot + ggtitle(climate)
+) + ggtitle(paste("DO within", min_thresholds, "ml/L") )
 }

### 2019-08-09 analysis/VOCC/vocc-example.Rmd
add white_zero = TRUE to example plots

@@
 
-glimpse(vocc)
+#glimpse(vocc)
 ```
@@
   raster_alpha = 1,
+  white_zero = TRUE,
   vec_alpha = 0.25,
@@
   raster_alpha = 1,
+  white_zero = TRUE,
   high_fill = "Steel Blue 4",
@@
 }
+
 vocc_plot

### 2019-08-09 analysis/VOCC/2-biomass-trends.Rmd
update variable name

@@
 current_biomass <- plot_vocc(out_mature,
-    fill_col = "var_1_e",
+    fill_col = "est_exp_e",
     fill_label = "Mean mature \nbiomass (kg/ha)\npost-2010",

### 2019-08-09 analysis/VOCC/4-vocc-by-spp.Rmd
beginning to simplify vocc-biomass merge code

@@
 Species
+```{r global_options, include=FALSE}
+# library(dplyr)
+# library(ggplot2)
+# library(gfplot)
+# library(gfdata)
+# library(sdmTMB)
+# library(gfranges)
+# 
+# knitr::opts_chunk$set(
+#   fig.width = 11, fig.height = 8.5,
+#   echo = FALSE, warning = FALSE, message = FALSE
+# )
+```
 
@@
 
-#climate <- "temperature"
+climate <- "temperature"
 temp_threshold <- c(0.75)
@@
 
-climate <- "do and temperature"
+#climate <- "do and temperature"
 max_thresholds <- c(Inf, 0.75)
 min_thresholds <- c(0.25, Inf)
+critical_thresholds <- c(-0.25, 0.75)
 optimal_value <- c(2, 6.5)
@@
 Run all subsequent code...
-```{r global_options, include=FALSE}
-library(dplyr)
-library(ggplot2)
-library(gfplot)
-library(gfdata)
-library(sdmTMB)
-library(gfranges)
-
-knitr::opts_chunk$set(
-  fig.width = 11, fig.height = 8.5,
-  echo = FALSE, warning = FALSE, message = FALSE
-)
-```
 
@@
 rm(vocc)
-try({
-  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
-    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
-})
+# try({
+#   vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
+#     min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+# })
 
@@
 start <- biomass_change %>% 
-  rename(start_density = var_1_s, end_density = var_1_e) %>% 
+  rename(start_density = est_exp_s, end_density = est_exp_e) %>% 
   select(icell, x, y, start_density, end_density)
@@
 end <- biomass_change %>% 
-  rename(start_density = var_1_s, 
-         end_density = var_1_e, 
+  rename(start_density = est_exp_s, 
+         end_density = est_exp_e, 
          target_X = x, target_Y = y) %>% 
@@
 
-source_density <- biomass_change %>% rename(source_density = var_1_s) %>% select(source_density, x, y)
-
-if (length(variable_names) == 1) {
+source_density <- biomass_change %>% 
+  rename(source_density = est_exp_s) %>% 
+  select(source_density, x, y)
+
+# if (length(variable_names) == 1) {
+#   
+#   vectors <- vocc %>%
+#     mutate(
+#       start_cell = icell, 
+#       vector_id = paste0(id, "_to_", tid),
+#       initial_value = UQ(rlang::sym(paste0(variable_names, "_s"))), #var_1_s, 
+#       source_end_value = UQ(rlang::sym(paste0(variable_names, "_e"))), #var_1_e, 
+#       target_value = as.numeric(target_values)
+#     ) %>%
+#     filter(distance > 1)  %>%
+#     select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
+#       initial_value, source_end_value, target_value, 
+#       km_per_decade, timespan) 
+#   
+#   if (climate == "do") vectors <- vectors %>% 
+#       filter(source_end_value < (optimal_do - do_threshold))
+#   if (climate == "temperature") vectors <- vectors %>% 
+#       filter(source_end_value > (optimal_temp + temp_threshold))
+# 
+# } else { # number of variables > 1
+  # 
+  # vectors <- vocc %>%
+  #   mutate(
+  #     start_cell = icell, vector_id = paste0(id, "_to_", tid),
+  #     initial_value = s, source_end_value = e, target_value = target_values
+  #   ) %>%
+  #   filter(distance > 1) %>%
+  #   select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
+  #     initial_value, source_end_value, target_value, 
+  #     km_per_decade, timespan) 
+  # 
+  # # split source_end_value into multiple vectors
+  # source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), 
+  #                  function(x) data.frame(t(as.numeric(x))))
+  # 
+  # vectors <- cbind(vectors, source_end) 
+  # 
+  # for (j in seq_along(variable_names)){
+  #   
+  #   source_end_climatej <- paste0("X",j)
+  #   vectors <- vectors %>% filter( 
+  #     (!!sym(source_end_climatej)) < (optimal_value[j] + critical_thresholds[j])
+  #     )
+  # }
+#  
+#   #climate <- "do and temperature"
+# # max_thresholds <- c(Inf, 0.75)
+# # min_thresholds <- c(0.25, Inf)
+# # critical_thresholds <- c(-0.25, 0.75)
+# # optimal_value <- c(2, 6.5)
+# #
+
+  for (j in seq_along(variable_names)){
+
   vectors <- vocc %>%
@@
       vector_id = paste0(id, "_to_", tid),
-      initial_value = var_1_s, 
-      source_end_value = var_1_e, 
-      target_value = as.numeric(target_values)
-    ) %>%
-    filter(distance > 1)  %>%
-    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
-      initial_value, source_end_value, target_value, 
-      km_per_decade, timespan)
-
-  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
-  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))
-
-} else {
-
-  vectors <- vocc %>%
-    mutate(
-      start_cell = icell, vector_id = paste0(id, "_to_", tid),
-      initial_value = s, source_end_value = e, target_value = target_values
-    ) %>%
-    filter(distance > 1) %>%
-    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
-      initial_value, source_end_value, target_value, 
-      km_per_decade, timespan)
-
-  source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), function(x) data.frame(t(as.numeric(x))))
-  vectors <- cbind(vectors, source_end)
-
-  for (j in seq_along(variable_names)){
-  browser()
-    vectors <-  filter( vectors, paste0("X",j) < (optimal_value[j] - min_thresholds[j]) )
-  }
+      source_end_value = e, # may need to turn into numeric vector later...
+      target_value = target_values
+    ) %>% filter(distance > 1)
+
+      if (min_thresholds[j] != Inf) { 
+        vectors <- filter(vectors, UQ(rlang::sym(paste0(variable_names[j], "_s"))) < optimal_value[j] - min_thresholds[j]) 
+        }
+      if (max_thresholds[j] != Inf) {
+        vectors <- filter(vectors, UQ(rlang::sym(paste0(variable_names[j], "_s"))) > optimal_value[j] + max_thresholds[j])
+        }
+
+  #FIXME: decide which variables might cause problems in future merges and remove them...   
+  #   vectors <- vectors %>% select(-N) 
+  #
 
@@
 icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
-mean.change <- mean(icells$perc_change, na.rm = TRUE)
-sd.change <- sd(icells$perc_change, na.rm = TRUE)
 
@@
     cell_type = as.factor(cell_type),
-    std_change = (perc_change)/sd.change, 
+    #std_change = (perc_change)/sd.change, 
     std_source_density = (source_density - mean.density)/(sd.density*2), 
@@
 
+icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
+
+mean.change <- mean(icells$perc_change, na.rm = TRUE)
+sd.change <- sd(icells$perc_change, na.rm = TRUE)
+
+vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
+  mutate(std_change = (perc_change)/sd.change)
+
 if (length(variable_names) == 1) {
@@
 
-vectors <- vectors %>% rename(
+vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% rename(
   source_end_climateA = X1, source_end_climateB = X2
@@
 
+#browser()
+
+icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
 mean.climateA <- mean(icells$source_end_climateA, na.rm = TRUE)
@@
     )
+
+#browser()
+  end_climate <- purrr::map_df(strsplit(vocc_by_sp[[i]]$end_climate, " "), 
+                   function(x) data.frame(t(as.numeric(x))))
+
+  vocc_by_sp[[i]] <- cbind(vocc_by_sp[[i]], end_climate) %>%
+    rename(end_climate_chr = end_climate, end_climate = X1, end_climateB = X2 )
+
 } else {
- stop("Doesn't currently support testing vectors bases on more than two climate variables at once.")
+
+ stop("Doesn't currently support testing vectors based on more than two climate variables at once.")
 }
@@
 # vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax ) 
-
 ```
@@
 data <- vocc_by_sp[["adult"]]
-#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)
+#data <- filter(vocc_by_sp[["adult"]],  distance < 90) # distance > 1 &
+#data <- filter(vocc_by_sp[["adult"]], distance > 1 ) #  &
 

### 2019-08-09 analysis/VOCC/4-vocc-by-spp.Rmd
streamline merge of vocc and density using lapply (required subsequent code modifications still incomplete)

@@
 
-climate <- "temperature"
+#climate <- "temperature"
 temp_threshold <- c(0.75)
@@
 
-#climate <- "do and temperature"
+climate <- "do and temperature"
 max_thresholds <- c(Inf, 0.75)
@@
 critical_thresholds <- c(-0.25, 0.75)
-optimal_value <- c(2, 6.5)
+optimal_values <- c(2, 6.5)
 ```
@@
   delta_threshold <- temp_threshold
-  optimal_value <- optimal_temp
+  optimal_values <- optimal_temp
 
@@
   delta_threshold <- -1*(do_threshold)
-  optimal_value <- optimal_do
+  optimal_values <- optimal_do
 
@@
 
-optimal_string <- paste0(optimal_value, collapse = "-")
+optimal_string <- paste0(optimal_values, collapse = "-")
 climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
@@
   ncol = 2,
-  top = grid::textGrob(paste("do of ", optimal_value[1], "less", min_thresholds[1],"or temperature of ", optimal_value[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
+  top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1],"or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
 )
@@
 ```{r}
-vocc_by_sp <- list()
-for (i in seq_along(biomass_list)){
-
-biomass_change <- make_trend_data(biomass_list[[i]],
-  ssid = model_ssid,
-  start_time = start_time,
-  skip_time = skip_time,
-  end_time = end_time,
-  input_cell_size = input_cell_size,
-  scale_fac = scale_fac,
-  delta_t_total = delta_t_total,
-  delta_t_step = delta_t_step,
-  indices = indices,
-  variable_names = "est_exp"
-)
-
-start <- biomass_change %>% 
-  rename(start_density = est_exp_s, end_density = est_exp_e) %>% 
-  select(icell, x, y, start_density, end_density)
-
-end <- biomass_change %>% 
-  rename(start_density = est_exp_s, 
-         end_density = est_exp_e, 
-         target_X = x, target_Y = y) %>% 
-  select(icell, start_density, end_density, target_X, target_Y)
-
-source_density <- biomass_change %>% 
-  rename(source_density = est_exp_s) %>% 
-  select(source_density, x, y)
-
-# if (length(variable_names) == 1) {
-#   
-#   vectors <- vocc %>%
-#     mutate(
-#       start_cell = icell, 
-#       vector_id = paste0(id, "_to_", tid),
-#       initial_value = UQ(rlang::sym(paste0(variable_names, "_s"))), #var_1_s, 
-#       source_end_value = UQ(rlang::sym(paste0(variable_names, "_e"))), #var_1_e, 
-#       target_value = as.numeric(target_values)
-#     ) %>%
-#     filter(distance > 1)  %>%
-#     select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
-#       initial_value, source_end_value, target_value, 
-#       km_per_decade, timespan) 
-#   
-#   if (climate == "do") vectors <- vectors %>% 
-#       filter(source_end_value < (optimal_do - do_threshold))
-#   if (climate == "temperature") vectors <- vectors %>% 
-#       filter(source_end_value > (optimal_temp + temp_threshold))
-# 
-# } else { # number of variables > 1
-  # 
-  # vectors <- vocc %>%
-  #   mutate(
-  #     start_cell = icell, vector_id = paste0(id, "_to_", tid),
-  #     initial_value = s, source_end_value = e, target_value = target_values
-  #   ) %>%
-  #   filter(distance > 1) %>%
-  #   select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
-  #     initial_value, source_end_value, target_value, 
-  #     km_per_decade, timespan) 
-  # 
-  # # split source_end_value into multiple vectors
-  # source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), 
-  #                  function(x) data.frame(t(as.numeric(x))))
-  # 
-  # vectors <- cbind(vectors, source_end) 
-  # 
-  # for (j in seq_along(variable_names)){
-  #   
-  #   source_end_climatej <- paste0("X",j)
-  #   vectors <- vectors %>% filter( 
-  #     (!!sym(source_end_climatej)) < (optimal_value[j] + critical_thresholds[j])
-  #     )
-  # }
-#  
-#   #climate <- "do and temperature"
-# # max_thresholds <- c(Inf, 0.75)
-# # min_thresholds <- c(0.25, Inf)
+
+# #   #climate <- "do and temperature"
+# max_thresholds <- c(Inf, 0.5)
+# min_thresholds <- c(0.1, Inf)
 # # critical_thresholds <- c(-0.25, 0.75)
-# # optimal_value <- c(2, 6.5)
-# #
-
-  for (j in seq_along(variable_names)){
-
-  vectors <- vocc %>%
-    mutate(
-      start_cell = icell,
-      vector_id = paste0(id, "_to_", tid),
-      source_end_value = e, # may need to turn into numeric vector later...
-      target_value = target_values
-    ) %>% filter(distance > 1)
-
-      if (min_thresholds[j] != Inf) { 
-        vectors <- filter(vectors, UQ(rlang::sym(paste0(variable_names[j], "_s"))) < optimal_value[j] - min_thresholds[j]) 
-        }
-      if (max_thresholds[j] != Inf) {
-        vectors <- filter(vectors, UQ(rlang::sym(paste0(variable_names[j], "_s"))) > optimal_value[j] + max_thresholds[j])
-        }
-
-  #FIXME: decide which variables might cause problems in future merges and remove them...   
-  #   vectors <- vectors %>% select(-N) 
-  #
-
-}
+# optimal_values <- c(2, 6.5)
+
+vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds)
 
 if (length(vectors[ ,2]) == 0) {
-  try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
+  other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")
   coefs <- other_coefs[1,]
@@
 
-  stop("No vectors.")
-}
-
-vector_source <- left_join(vectors, start, by = c("x", "y")) %>% 
-  mutate(cell_type = "source", end_climate = source_end_value, source_density = start_density)
+  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
+}
+
+#glimpse(vectors)
+```
+
+# Combine vector and biomass data
+
+```{r}
+vocc_by_sp <- lapply(biomass_list, function(i) {
+
+biomass_change <- make_trend_data(i,
+  ssid = model_ssid,
+  start_time = start_time,
+  skip_time = skip_time,
+  end_time = end_time,
+  input_cell_size = input_cell_size,
+  scale_fac = scale_fac,
+  delta_t_total = delta_t_total,
+  delta_t_step = delta_t_step,
+  indices = indices,
+  variable_names = "est_exp"
+)
+
+vectors <- vectors %>%
+      mutate(
+        start_cell = icell,
+        vector_id = paste0(id, "_to_", tid),
+        distance2 = distance*distance,
+        source_end_value = e, # may need to turn into numeric vector later...
+        target_value = target_values
+      )
+
+start <- biomass_change %>% 
+  rename(start_density = est_exp_s, end_density = est_exp_e) %>% 
+  select(icell, x, y, start_density, end_density)
+
+vector_source <- left_join(vectors, start, by = c( "icell", "x", "y")) %>% 
+  mutate(cell_type = "source", end_climate = source_end_value, 
+    source_density = start_density)
+
+end <- biomass_change %>% 
+  rename(start_density = est_exp_s, 
+         end_density = est_exp_e, 
+         target_X = x, target_Y = y) %>% 
+  select(start_density, end_density, target_X, target_Y)
 
@@
   mutate(cell_type = "target", end_climate = target_value)
-vector_target <- left_join(vector_target, source_density, by = c("x", "y")) #%>%  
-  # filter(end_climate > 2)
+
+source_density <- biomass_change %>% 
+  rename(source_density = est_exp_s) %>% 
+  select(source_density, x, y)
+
+vector_target <- left_join(vector_target, source_density, by = c("x", "y"))
 
 # combine source and target cell data
-vocc_by_sp[[i]] <- rbind(vector_source, vector_target) %>% 
+vocc_by_sp <- rbind(vector_source, vector_target) %>% 
+  mutate(cell_type = as.factor(cell_type)) %>%
   filter(!is.na(icell)) %>%
-# remove potential source cells with no fish  
-  #FIXME: should their be a flexible min density of fish for a source cell?
-  filter(source_density > 0.000001)
+
+  # remove potential source cells with lowest 1% quantile of fish biomass  
+  filter(source_density > quantile(source_density, 0.01) ) %>%
+
+  # calculate percent change in biomass
+  mutate(perc_change = (end_density - start_density)/start_density)
+
+end_climate <- purrr::map_df(strsplit(vocc_by_sp$end_climate, " "), 
+                   function(x) data.frame(t(as.numeric(x))))
+
+names(end_climate) <- paste0("end_", variable_names)
+
+vocc_by_sp <- cbind(vocc_by_sp, end_climate) %>%
+    rename(end_climate_chr = end_climate)
 
 # get mean and sd for densities without cell duplication
-icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
+icells <- vocc_by_sp %>% distinct(icell, .keep_all = TRUE)
 
@@
 sd.density <- sd(icells$source_density, na.rm = TRUE)
-
-# calculate standardized density varibles
-vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
-  mutate(
-    perc_change = (end_density - start_density)/start_density, 
-    cell_type = as.factor(cell_type),
-    #std_change = (perc_change)/sd.change, 
-    std_source_density = (source_density - mean.density)/(sd.density*2), 
-    mean_change = mean.change, sd_change = sd.change,
-    mean_source_density = mean.density, sd_density = sd.density,
-# set distance intercept at cell size
-    distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, 
-    distance2 = distance*distance) %>% 
-  rename(X = x, Y = y)
-
-icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
-
 mean.change <- mean(icells$perc_change, na.rm = TRUE)
@@
 
-vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
-  mutate(std_change = (perc_change)/sd.change)
-
-if (length(variable_names) == 1) {
-mean.climate <- mean(icells$end_climate, na.rm = TRUE)
-sd.climate <- sd(icells$end_climate, na.rm = TRUE)
-
-vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
+for (j in seq_along(variable_names)){
+
+mean.var <- paste0("mean.", variable_names[j])
+sd.var <- paste0("sd.", variable_names[j])
+
+icells <- mutate(icells,
+  (!! mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j], "_e"))), na.rm = TRUE),
+  (!! sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j], "_e"))), na.rm = TRUE)
+)
+}
+
+# calculate standardized density variables
+vocc_by_sp <- vocc_by_sp %>% 
   mutate(
-    mean_climate = mean.climate, sd_climate = sd.climate,
-    std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
-    std_end_climate2 = std_end_climate^2
+    std_source_density = (source_density - mean.density)/(sd.density*2), 
+    mean_source_density = mean.density, sd_density = sd.density,
+    mean_change = mean.change, sd_change = sd.change,
+    std_change = (perc_change)/sd.change)
+
+# calculate standardized climate variables
+for (j in seq_along(variable_names)){
+
+mean.var <- paste0("mean.", variable_names[j])
+sd.var <- paste0("sd.", variable_names[j])
+std.var <- paste0("std.", variable_names[j])
+std.var2 <- paste0("std.", variable_names[j],"2")
+
+vocc_by_sp <- mutate(vocc_by_sp,
+  (!! mean.var) := icells[1, mean.var],
+  (!! sd.var) := icells[1, sd.var],
+  (!! std.var) := (UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ icells[1, sd.var],
+  (!! std.var2) := ((UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ icells[1, sd.var])^2
   )
-
-}else{
-
-if (length(variable_names) == 2) {
-
-vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% rename(
-  source_end_climateA = X1, source_end_climateB = X2
-  )
-
-#browser()
-
-icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
-mean.climateA <- mean(icells$source_end_climateA, na.rm = TRUE)
-sd.climateA <- sd(icells$source_end_climateA, na.rm = TRUE)
-mean.climateB <- mean(icells$source_end_climateB, na.rm = TRUE)
-sd.climateB <- sd(icells$source_end_climateB, na.rm = TRUE)
-
-vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
+}
+
+vocc_by_sp <- vocc_by_sp %>% 
   mutate(
-    mean_climateA = mean.climateA, sd_climate = sd.climateA,
-    std_end_climateA = (source_end_climateA - mean.climateA)/(sd.climateA*2),
-    std_end_climateA2 = std_end_climateA^2,
-    mean_climateB = mean.climateB, sd_climate = sd.climateB,
-    std_end_climateB = (source_end_climateB - mean.climateB)/(sd.climateB*2),
-    std_end_climateB2 = std_end_climateB^2
-    )
-
-#browser()
-  end_climate <- purrr::map_df(strsplit(vocc_by_sp[[i]]$end_climate, " "), 
-                   function(x) data.frame(t(as.numeric(x))))
-
-  vocc_by_sp[[i]] <- cbind(vocc_by_sp[[i]], end_climate) %>%
-    rename(end_climate_chr = end_climate, end_climate = X1, end_climateB = X2 )
-
-} else {
-
- stop("Doesn't currently support testing vectors based on more than two climate variables at once.")
-}
-}
-
-}
-names(vocc_by_sp) <- names(biomass_list)
-
-# glimpse(icells)
-# range(icells$perc_change)
-# range(vocc_by_sp[[i]]$std_change)
-# range(vocc_by_sp[[i]]$X)
-
-# glimpse(vocc_by_sp[[i]])
+    std_source_density = (source_density - mean.density)/(sd.density*2), 
+    mean_source_density = mean.density, sd_density = sd.density,
+    mean_change = mean.change, sd_change = sd.change,
+    std_change = (perc_change)/sd.change)
+
+# rename spatial coordinates for inclusion in models and plots without naming conflicts
+vocc_by_sp <- vocc_by_sp %>% 
+  rename(X = x, Y = y)
+
+vocc_by_sp
+})
+
+# glimpse(vocc_by_sp)
 # trim spatial outliers 
@@
 
-mature_plot <- ggplot(data, aes(x = end_climate, y = (end_density # - start_density
+#for (j in seq_along(variable_names)) {
+
+mature_plot <- lapply(variable_names, function(j) { 
+ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density # - start_density
   ), colour = cell_type, #size = start_density, 
   group = cell_type)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+    geom_point() +
+  #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
   geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
@@
   ylab(paste("Final biomass")) +
-  xlab(paste("Final value of", climate,"for all cells")) +
+  xlab(paste("Final value of", variable_names,"for all cells")) +
   scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
@@
   ggtitle(paste(survey, "mature", species))
-
-mature_plot
+})
+print(mature_plot)
 ```
@@
 coefs$max_thresholds <- max_thresholds
-coefs$optimum <- optimal_value
+coefs$optimum <- optimal_values
 coefs$start_time <- start_time

### 2019-08-13 analysis/VOCC/4-vocc-by-spp.Rmd
use log_change as response; save lmer coefs

@@
 ```{r global_options, include=FALSE}
-# library(dplyr)
-# library(ggplot2)
+library(dplyr)
+library(ggplot2)
 # library(gfplot)
 # library(gfdata)
-# library(sdmTMB)
-# library(gfranges)
+library(sdmTMB)
+library(gfranges)
 # 
@@
 climate <- "do and temperature"
-max_thresholds <- c(Inf, 0.75)
-min_thresholds <- c(0.25, Inf)
-critical_thresholds <- c(-0.25, 0.75)
+max_thresholds <- c(Inf, 0.5)
+min_thresholds <- c(0.1, Inf)
 optimal_values <- c(2, 6.5)
+
+# critical_thresholds <- c(-0.25, 0.75)
+
 ```
@@
 # min_thresholds <- c(0.1, Inf)
-# # critical_thresholds <- c(-0.25, 0.75)
 # optimal_values <- c(2, 6.5)
@@
 if (length(vectors[ ,2]) == 0) {
-  other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")
+  other_coefs <- readRDS(paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
   coefs <- other_coefs[1,]
@@
   coefs$se <- "NA"
+  coefs$model <- "NA"
   coefs$species <- spp 
@@
 
-  rbind(other_coefs, coefs)
-  saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")
+  coefs <- rbind(other_coefs, coefs) %>% distinct()
+  saveRDS(coefs, file = paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
 
@@
 }
-
+#glimpse(vocc)
 #glimpse(vectors)
@@
 )
-
+# browser()
 vectors <- vectors %>%
@@
   rename(start_density = est_exp_s, end_density = est_exp_e) %>% 
-  select(icell, x, y, start_density, end_density)
+  dplyr::select(icell, x, y, start_density, end_density)
 
@@
          target_X = x, target_Y = y) %>% 
-  select(start_density, end_density, target_X, target_Y)
+  dplyr::select(start_density, end_density, target_X, target_Y)
 
@@
   rename(source_density = est_exp_s) %>% 
-  select(source_density, x, y)
+  dplyr::select(source_density, x, y)
 
@@
   # calculate percent change in biomass
-  mutate(perc_change = (end_density - start_density)/start_density)
+  mutate(
+    log_change = log(end_density/start_density),
+    perc_change = end_density/start_density
+    )
 
@@
     std_source_density = (source_density - mean.density)/(sd.density*2),
+    std_source_density2 = (source_density - mean.density)/(sd.density*2)^2,
     mean_source_density = mean.density, sd_density = sd.density,
-    mean_change = mean.change, sd_change = sd.change,
-    std_change = (perc_change)/sd.change)
+    mean_change = mean.change, sd_change = sd.change
+    )
 
@@
   (!! sd.var) := icells[1, sd.var],
-  (!! std.var) := (UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ icells[1, sd.var],
+  (!! std.var) := (UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ (icells[1, sd.var]*2),
   (!! std.var2) := ((UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var])/ icells[1, sd.var])^2
@@
 }
-
-vocc_by_sp <- vocc_by_sp %>% 
-  mutate(
-    std_source_density = (source_density - mean.density)/(sd.density*2), 
-    mean_source_density = mean.density, sd_density = sd.density,
-    mean_change = mean.change, sd_change = sd.change,
-    std_change = (perc_change)/sd.change)
 
@@
 mature_plot <- lapply(variable_names, function(j) { 
-ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density # - start_density
-  ), colour = cell_type, #size = start_density, 
+ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density, colour = cell_type, size = start_density, 
   group = cell_type)) +
-    geom_point() +
-  #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
-  #   , aes(x = end_climate, y = (end_density #- start_density
-  #  )), inherit.aes = FALSE) +
-  scale_size_continuous(guide = "none") +
+  geom_point(alpha = 0.5) +
+  # geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = log_change
+  # ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
+  # scale_size_continuous(guide = "none") +
   scale_y_continuous(trans = fourth_root_power) +
+  xlab(paste("Final value of", j,"for cells")) +
   ylab(paste("Final biomass")) +
-  xlab(paste("Final value of", variable_names,"for all cells")) +
   scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
@@
 
-mature_plot2 <- ggplot(data, aes(x = distance, y = std_change, 
-  group = cell_type, size = start_density, colour = end_climate)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y~x+I(x^2) )+
+mature_plot2 <- lapply(variable_names, function(j) { 
+  ggplot(data, aes(x = distance, y = log_change, 
+      group = cell_type, size = start_density, 
+      colour = UQ(rlang::sym(paste0("end_", j))))) +
+    geom_point(alpha = 0.5) +
+    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+    geom_smooth(method="lm", formula = y~x+I(x^2), se = FALSE )+
+    scale_size_continuous(guide = "none") +
+    scale_y_continuous(trans = fourth_root_power) +
     facet_wrap(~cell_type) +#, scales = "free") +
-  scale_size_continuous(guide = "none") +
-  # scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
+    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
+    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
     # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
+    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
     ggtitle(paste(survey, "mature", species))
-
+})
 mature_plot2
@@
 
-mature_plot3 <- ggplot(data, aes(x = source_density, y = std_change, 
-    colour = end_climate, 
+mature_plot3 <- lapply(variable_names, function(j) { 
+  ggplot(data, aes(x = source_density, y = log_change, 
+          colour = UQ(rlang::sym(paste0("end_", j))),
+    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
+    group = cell_type)) +
+  geom_jitter(width = 0.01) +
+  geom_smooth(method="lm", se = FALSE) +
+  facet_wrap(~cell_type) + #, scales = "free") +
+  scale_alpha_continuous(guide = "none") +
+  scale_size_continuous(guide = "none") +
+  scale_x_continuous(trans = sqrt) +
+  scale_y_continuous(trans = sqrt) +
+  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
+  xlab(paste("Source density (dark points are closer to source)")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+})
+mature_plot3
+```
+
+```{r}
+data <- vocc_by_sp[["adult"]]
+data <- filter(vocc_by_sp[["adult"]], distance < 50)
+
+mature_plot4 <- lapply(variable_names, function(j) {  
+ggplot(filter(data, distance < 10), aes(source_density, y = log_change , 
+    colour = UQ(rlang::sym(paste0("end_", j))),
+    alpha = 1/(distance + input_cell_size), #size = 1/distance,
+    group = cell_type)) +
+  geom_jitter(width = 0.01) +
+  geom_smooth(method="lm", aes(group = cell_type)) +
+  facet_wrap(~cell_type) +#, scales = "free") +
+  scale_alpha_continuous(guide = "none") +
+  scale_size_continuous(guide = "none") +
+  scale_x_continuous(trans = sqrt) +
+  #scale_y_continuous(trans = fourth_root_power) +
+  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
+  xlab(paste("Source density (dark points are closer to source)")) +
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+})
+mature_plot4
+```
+
+```{r}
+data <- vocc_by_sp[["adult"]]
+data <- filter(vocc_by_sp[["adult"]], distance < 50)
+
+mature_plot5 <- lapply(variable_names, function(j) {  
+  ggplot(data, aes(
+      x = UQ(rlang::sym(paste0(j, ".units_per_decade"))) , 
+      y = (end_density - start_density), 
+      colour = cell_type, 
+      size = start_density)) +
+    geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+    #scale_size_continuous(guide = "none") +
+    #scale_y_continuous(trans = fourth_root_power) +
+    ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
+    xlab(paste("Change in", j," in units per decade")) +
+    scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
+    ggtitle(paste(survey, "mature", species))
+})
+mature_plot5
+```
+
+## Adult models
+
+### Regular mixed-model
+
+```{r}
+# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
+#             std_end_climate + std_end_climate2 + 
+#             (1 | vector_id), 
+#  #   offset = log(start_density), 
+#  #   family = Gamma(link = "inverse")
+#      data = data)
+# hist(data$end_density)
+# hist(data$log_change)
+data <- vocc_by_sp[["adult"]]
+
+model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * std_source_density2 + cell_type * distance ") # + cell_type * distance2
+
+for (j in (variable_names) ){
+  model_formula <- paste0("", model_formula, " + std.", j, "+ std.", j, "2"
+  )
+}
+
+model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
+
+test <- lmerTest::lmer(as.formula(model_formula), REML = F, data = data)
+
+summary(test)
+# plot(test2)
+
+library(sjPlot)
+#library(sjlabelled)
+#library(sjmisc)
+sjPlot::plot_model(test, title = "Log biomass change")
+```
+
+```{r}
+beta <- lme4::fixef(test) 
+se <- arm::se.fixef(test) 
+var<- names(beta)
+
+coefs <- as.data.frame(cbind(beta, se))
+coefs$ci <- coefs$se*1.96
+var[1] <- "intercept"
+row.names(coefs) <- NULL
+coefs <-cbind(var, coefs)
+coefs$model <- model_formula
+coefs$species <- spp 
+coefs$lifestage <- "adult"
+coefs$climate <- climate 
+coefs$min_thresholds <- min_thresholds
+coefs$max_thresholds <- max_thresholds
+coefs$optimum <- optimal_string
+coefs$start_time <- start_time
+coefs$timespan <- delta_t_total
+coefs$region <- region
+coefs$ssid_string <- ssid_string
+
+glimpse(coefs)
+rm(other_coefs)
+try({other_coefs <- readRDS(paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))})
+
+if (exists("other_coefs")) {
+coefs <- rbind(other_coefs, coefs)
+}
+saveRDS(coefs, file = paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))
+```
+
+Robust lmm
+```{r eval=FALSE}
+library(lme4)
+data <- vocc_by_sp[["adult"]]
+
+test3 <- robustlmm::rlmer(as.formula(model_formula), data = data)
+
+summary(test3)
+plot(test3, ask = FALSE)
+```
+
+Maybe try GAM?
+```{r}
+# library()
+# 
+# model_formula <- paste0("log_change ~ cell_type + s(std_source_density) + s(distance) ") # + cell_type * distance2
+# 
+# for (j in (variable_names) ){
+#   model_formula <- paste0("", model_formula, " + s(std.", j, ") + "
+#   )
+# }
+# 
+# model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
+# 
+# 
+# test2 <- gamm4::....(as.formula(model_formula), REML = F, data = data)
+#
+
+```
+
+### Boosted regression
+```{r}
+data <- vocc_by_sp[["adult"]]
+
+gbm_formula <- paste0("log_change ~ std_source_density + distance + cell_type + X + Y") # + cell_type * distance2
+
+for (j in (variable_names) ){
+  gbm_formula <- paste0(gbm_formula, " + std.", j)
+}
+
+m <- gbm::gbm(
+  as.formula(gbm_formula), 
+  data = data, 
+  distribution = "gaussian",
+  n.trees = 2000, 
+  interaction.depth = 3, 
+  shrinkage = 0.02
+  )
+
+# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
+#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
+```
+
+```{r}
+summary(m)
+```
+
+```{r}
+plot(m,1)
+plot(m,2)
+plot(m,3)
+plot(m,6)
+if (length(variable_names) > 1 ) plot(m,7)
+if (length(variable_names) > 2 ) plot(m,8)
+
+plot(m,c(1,2))
+plot(m,c(1,3))
+plot(m,c(1,6))
+
+if (length(variable_names) > 1 ) plot(m,c(1,7))
+
+plot(m,c(2,3))
+plot(m,c(2,6))
+if (length(variable_names) > 1 ) plot(m,c(2,7))
+
+plot(m,c(1,2,6))
+plot(m,c(2,6,3))
+plot(m,c(4,5))
+```
+
+### sdmTMB
+
+```{r}
+data <- vocc_by_sp[["adult"]]
+# Xmax <- quantile(data$X, 0.9995)
+# Xmin <- quantile(data$X, 0.005)
+# std_max <- quantile(data$log_change, 0.9995)
+# 
+# data <- data %>% filter ( X >= Xmin & X <= Xmax )
+
+xy <- as.data.frame(cbind(data$X, data$Y))
+
+  if (region == "Both odd year surveys") {
+    n_knots <- 200
+  }
+
+  if (region == "West Coast Vancouver Island") {
+    n_knots <- 100
+  }
+
+  # if (region == "West Coast Haida Gwaii") {
+  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
+  # }
+
+if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(xy))}
+
+  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = n_knots)
+  sdmTMB::plot_spde(spde)
+
+```
+
+```{r}
+data <- vocc_by_sp[["adult"]]
+
+tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + distance2") # + cell_type * std_source_density2
+
+for (j in (variable_names) ){
+  tmb_formula <- paste0("", tmb_formula, " + std.", j, "+ std.", j, "2"
+  )
+}
+
+test <- sdmTMB(data = data,
+      as.formula(tmb_formula), 
+      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      silent = TRUE)
+test
+```
+
+```{r}
+
+p <- predict(test)
+p$residuals <- residuals(test)
+#idata$resids <- p$residuals
+#glimpse(p)
+ggplot(p, aes(est, residuals)) +
+  geom_point(alpha=0.4) +
+  ylim(-8,8) +
+  geom_smooth()
+
+plot_map_raster <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    coord_fixed()+
+    gfplot::theme_pbs() + 
+    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+}
+
+plot_map_raster(p, "est")
+plot_map_raster(p, "est_rf")
+plot_map_raster(p, "omega_s")
+plot_map_raster(p, "residuals")
+
+```
+
+# Immature
+
+### Immature plots
+
+Check final climate-biomass relationship
+```{r}
+data <- vocc_by_sp[["imm"]]
+#glimpse(data)
+
+#for (j in seq_along(variable_names)) {
+
+immature_plot <- lapply(variable_names, function(j) { 
+ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density # - start_density
+  ), colour = cell_type, size = start_density, 
+  group = cell_type)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density
+  ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
+  scale_size_continuous(guide = "none") +
+  scale_y_continuous(trans = fourth_root_power) +
+  xlab(paste("Final value of", j,"for cells")) +
+  ylab(paste("Final biomass")) +
+  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
+  ggtitle(paste(survey, "immature", species))
+})
+immature_plot
+```
+
+Check distance threshold of 50 looks reasonable
+```{r}
+data <- vocc_by_sp[["imm"]]
+#data <- filter(vocc_by_sp[["imm"]],  distance < 90) # distance > 1 &
+#data <- filter(vocc_by_sp[["imm"]], distance > 1 ) #  &
+
+immature_plot2 <- lapply(variable_names, function(j) { 
+  ggplot(data, aes(x = distance, y = log_change, 
+      group = cell_type, size = start_density, 
+      colour = UQ(rlang::sym(paste0("end_", j))))) +
+    geom_point(alpha = 0.5) +
+    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
+    geom_smooth(method="lm", formula = y~x+I(x^2) )+
+    scale_size_continuous(guide = "none") +
+    scale_y_continuous(trans = fourth_root_power) +
+    facet_wrap(~cell_type) +#, scales = "free") +
+    ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
+    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
+    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
+    # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
+    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
+    ggtitle(paste(survey, "immature", species))
+})
+immature_plot2
+```
+
+```{r warnings=FALSE}
+data <- vocc_by_sp[["imm"]]
+#data <- filter(vocc_by_sp[["imm"]], distance < 50)
+#data <- filter(vocc_by_sp[["imm"]], distance > 4 & distance < 50)
+
+immature_plot3 <- lapply(variable_names, function(j) { 
+  ggplot(data, aes(x = source_density, y = log_change, 
+          colour = UQ(rlang::sym(paste0("end_", j))),
     alpha = 1/(distance+ input_cell_size), #size = 1/distance,
@@
   xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
   gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-mature_plot3
-```
-
-```{r}
-data <- vocc_by_sp[["adult"]]
-data <- filter(vocc_by_sp[["adult"]], distance < 50)
-
-mature_plot4 <- ggplot(filter(data, distance < 10), aes(source_density, y = std_change , 
-    colour = end_climate, 
-    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
+  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+})
+immature_plot3
+```
+
+```{r}
+data <- vocc_by_sp[["imm"]]
+data <- filter(vocc_by_sp[["imm"]], distance < 50)
+
+immature_plot4 <- lapply(variable_names, function(j) {  
+ggplot(filter(data, distance < 10), aes(source_density, y = log_change , 
+    colour = UQ(rlang::sym(paste0("end_", j))),
+    alpha = 1/(distance + input_cell_size), #size = 1/distance,
     group = cell_type)) +
@@
   xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
+  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
   gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-mature_plot4
-```
-
-```{r}
-data <- vocc_by_sp[["adult"]]
-data <- filter(vocc_by_sp[["adult"]], distance < 50)
-
-mature_plot5 <- ggplot(data, aes(x = end_climate - initial_value , y = #std_change
-    (end_density - start_density)
-  , colour = cell_type, size = start_density)) +
+  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+})
+immature_plot4
+```
+
+```{r}
+data <- vocc_by_sp[["imm"]]
+data <- filter(vocc_by_sp[["imm"]], distance < 50)
+
+immature_plot5 <- lapply(variable_names, function(j) {  
+  ggplot(data, aes(
+      x = UQ(rlang::sym(paste0(j, ".units_per_decade"))) , 
+      y = (end_density - start_density), 
+      colour = cell_type, 
+      size = start_density)) +
     geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  scale_size_continuous(guide = "none") +
+    #scale_size_continuous(guide = "none") +
     #scale_y_continuous(trans = fourth_root_power) +
     ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Change in", climate,"")) +
+    xlab(paste("Change in", j," in units per decade")) +
     scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "mature", species))
-mature_plot5
-```
-
-## Adult models
+    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.7)) + 
+    ggtitle(paste(survey, "immature", species))
+})
+immature_plot5
+```
+
+## Immature models
 
@@
 ```{r}
-# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
-#             std_end_climate + std_end_climate2 + 
-#             (1 | vector_id), 
-#  #   offset = log(start_density), 
-#  #   family = Gamma(link = "inverse")
-#      data = data)
-# hist(data$end_density)
-# hist(data$std_change)
-data <- vocc_by_sp[["adult"]]
-
-test2 <- lmerTest::lmer(std_change ~ cell_type * std_source_density + 
-    cell_type * distance + cell_type * distance2 +
-    std_end_climate + std_end_climate2 + (1 | vector_id), REML = F, data = data)
-
-summary(test2)
-plot(test2)
+idata <- vocc_by_sp[["imm"]]
+
+model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * std_source_density2 + cell_type * distance ") # + cell_type * distance2
+
+for (j in (variable_names) ){
+  model_formula <- paste0("", model_formula, " + std.", j, "+ std.", j, "2"
+  )
+}
+
+model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
+
+itest <- lmerTest::lmer(as.formula(model_formula), REML = F, data = idata)
+
+summary(itest)
+
+#library(sjlabelled)
+#library(sjmisc)
+sjPlot::plot_model(itest, title = "Log biomass change")
+```
+
+```{r}
+beta <- lme4::fixef(itest) 
+se <- arm::se.fixef(itest) 
+var<- names(beta)
+
+coefs <- as.data.frame(cbind(beta, se))
+coefs$ci <- coefs$se*1.96
+var[1] <- "intercept"
+row.names(coefs) <- NULL
+coefs <-cbind(var, coefs)
+coefs$model <- model_formula
+coefs$species <- spp 
+coefs$lifestage <- "immature"
+coefs$climate <- climate 
+coefs$min_thresholds <- min_thresholds
+coefs$max_thresholds <- max_thresholds
+coefs$optimum <- optimal_string
+coefs$start_time <- start_time
+coefs$timespan <- delta_t_total
+coefs$region <- region
+coefs$ssid_string <- ssid_string
+
+glimpse(coefs)
+rm(other_coefs)
+try({other_coefs <- readRDS(paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))})
+
+if (exists("other_coefs")) {
+coefs <- rbind(other_coefs, coefs)
+}
+saveRDS(coefs, file = paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))
 ```
@@
 library(lme4)
-data <- vocc_by_sp[["adult"]]
-
-test3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
-    cell_type * distance + cell_type * distance2 + 
-    std_end_climate + std_end_climate2 + (1 | vector_id), data = data)
+
+test3 <- robustlmm::rlmer(as.formula(model_formula), data = idata)
 
@@
 ```{r}
-data <- vocc_by_sp[["adult"]]
-
-m <- gbm::gbm(
-  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
-  data = data, 
-  distribution = "gaussian",
-  n.trees = 2000, 
-  interaction.depth = 3, 
-  shrinkage = 0.02
-  )
-
-# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
-#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
-```
-
-```{r}
-summary(m)
-```
-
-```{r}
-plot(m,1)
-plot(m,2)
-plot(m,3)
-plot(m,4)
-plot(m,c(1,2))
-plot(m,c(1,3))
-plot(m,c(1,4))
-plot(m,c(2,3))
-plot(m,c(2,4))
-plot(m,c(1,2,4))
-plot(m,c(2,3,4))
-plot(m,c(5,6))
-```
-
-### sdmTMB
-
-```{r}
-data <- vocc_by_sp[["adult"]]
-# Xmax <- quantile(data$X, 0.9995)
-# Xmin <- quantile(data$X, 0.005)
-# std_max <- quantile(data$std_change, 0.9995)
+idata <- vocc_by_sp[["imm"]]
+
+# gbm_formula <- paste0("log_change ~ std_source_density + distance + cell_type + X + Y") # + cell_type * distance2
 # 
-# data <- data %>% filter ( X >= Xmin & X <= Xmax )
-
-xy <- as.data.frame(cbind(data$X, data$Y))
-
-  if (region == "Both odd year surveys") {
-    n_knots <- 200
-  }
-
-  if (region == "West Coast Vancouver Island") {
-    n_knots <- 100
-  }
-
-  # if (region == "West Coast Haida Gwaii") {
-  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
+# for (j in (variable_names) ){
+#   gbm_formula <- paste0(gbm_formula, " + std.", j)
 # }
 
-if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(xy))}
-
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = n_knots)
-  sdmTMB::plot_spde(spde)
-
-```
-
-```{r}
-#data$end_climate2 <- data$end_climate^2
-#data <- vocc_by_sp[["adult"]]
-data$std_source_density2 <- data$std_source_density^2
-test <- sdmTMB(data = data,
-      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 +
-    cell_type * distance + cell_type * distance2 + 
-      std_end_climate + std_end_climate2, 
-      # std_change ~ cell_type * source_density + cell_type * distance  + end_climate + end_climate2,
-      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = TRUE)
-test
-```
-
-```{r}
-
-p <- predict(test)
-p$residuals <- residuals(test)
-#idata$resids <- p$residuals
-#glimpse(p)
-ggplot(p, aes(est, residuals)) +
-  geom_point(alpha=0.4) +
-  ylim(-8,8) +
-  geom_smooth()
-
-plot_map_raster <- function(dat, column = "est") {
-    ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    coord_fixed()+
-    gfplot::theme_pbs() + 
-    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
-}
-
-plot_map_raster(p, "est")
-plot_map_raster(p, "est_rf")
-plot_map_raster(p, "omega_s")
-plot_map_raster(p, "residuals")
-
-```
-
-# Immature
-
-### Immature plots
-
-Check final climate-biomass relationship
-```{r}
-idata <- vocc_by_sp[["imm"]]
-#glimpse(data)
-
-immature_plot1 <- ggplot(idata, aes(x = end_climate, y = (end_density # - start_density
-  ), colour = cell_type, #size = start_density, 
-  group = cell_type)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
-  #geom_smooth() +
-  scale_size_continuous(guide = "none") +
-  scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Final biomass")) +
-  xlab(paste("Final value of", climate,"for all cells")) +
-  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
-  ggtitle(paste(survey, "immature", species))
-
-immature_plot1
-```
-
-Check distance threshold of 50 looks reasonable
-```{r warnings=FALSE}
-idata <- vocc_by_sp[["imm"]]
-#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)
-
-immature_plot2 <- ggplot(idata, aes(distance, y = std_change, 
-  group = cell_type, size = start_density, colour = end_climate)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  geom_smooth(method="lm", formula = y~x+I(x^2) )+
-  facet_wrap(~cell_type) + #, scales = "free") +
-  scale_size_continuous(guide = "none") +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species))
-
-immature_plot2
-```
-
-```{r warnings=FALSE}
-idata <- vocc_by_sp[["imm"]]
-#idata <- filter(vocc_by_sp[["imm"]], distance < 50)
-#idata <- filter(vocc_by_sp[[""imm"]], distance > 4 & distance < 50)
-
-immature_plot3 <- ggplot(idata, aes(x = source_density, y = std_change, 
-    colour = end_climate, 
-    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
-    group = cell_type)) +
-  geom_jitter(width = 0.01) +
-  geom_smooth(method="lm") +
-  facet_wrap(~cell_type) + #, scales = "free") +
-  scale_alpha_continuous(guide = "none") +
-  scale_size_continuous(guide = "none") +
-  scale_x_continuous(trans = sqrt) +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-immature_plot3
-```
-
-```{r warnings=FALSE}
-idata <- vocc_by_sp[["imm"]]
-idata <- filter(vocc_by_sp[["imm"]], distance < 50)
-
-immature_plot4 <- ggplot(filter(idata, distance < 10), aes(source_density, y = std_change , 
-    colour = end_climate, 
-    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
-    group = cell_type)) +
-  geom_jitter(width = 0.01) +
-  geom_smooth(method="lm") +
-  facet_wrap(~cell_type) +#, scales = "free") +
-  scale_alpha_continuous(guide = "none") +
-  scale_size_continuous(guide = "none") +
-  scale_x_continuous(trans = sqrt) +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-
-immature_plot4
-```
-
-```{r}
-idata <- vocc_by_sp[["imm"]]
-idata <- filter(vocc_by_sp[["imm"]], distance < 50)
-
-immature_plot5 <- ggplot(idata, aes(x = end_climate - initial_value , y = (end_density - start_density
-  ), colour = cell_type, size = start_density)) +
-  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-  scale_size_continuous(guide = "none") +
-  #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Change in", climate,"")) +
-  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species))
-immature_plot5
-```
-
-## Immature models
-### Regular mixed-model
-
-```{r}
-idata <- vocc_by_sp[["imm"]]
-
-# Xmax <- quantile(idata$X, 0.9995)
-# Xmin <- quantile(idata$X, 0.005)
-# std_max <- quantile(idata$std_change, 0.9995)
-# 
-# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )
-#  
-itest <- lmerTest::lmer(std_change ~ 
-      cell_type * std_source_density + 
-      cell_type * distance + cell_type * distance2 +
-      std_end_climate + std_end_climate2 + (1 | vector_id), 
-      REML = F, data = idata)
-summary(itest)
-plot(itest)
-```
-```{r eval=FALSE}
-library(lme4)
-itest3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
-    cell_type * distance + cell_type * distance2 + 
-    std_end_climate + std_end_climate2 + (1 | vector_id), data = idata)
-summary(itest3)
-plot(itest3, ask = FALSE)
-```
-
-### Boosted regression
-
-```{r}
-idata <- vocc_by_sp[["imm"]]
-
-# Xmax <- quantile(idata$X, 0.9995)
-# Xmin <- quantile(idata$X, 0.005)
-# std_max <- quantile(idata$std_change, 0.9995)
-# 
-# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )
-
-mi <- gbm::gbm(
-  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
+im <- gbm::gbm(
+  as.formula(gbm_formula), 
   data = idata, 
@@
 ```{r}
-summary(mi)
-```
-
-```{r}
-plot(mi,1)
-plot(mi,2)
-plot(mi,3)
-plot(mi,4)
-plot(mi,c(1,2))
-plot(mi,c(1,3))
-plot(mi,c(1,4))
-plot(mi,c(2,3))
-plot(mi,c(2,4))
-plot(mi,c(1,2,4))
-plot(mi,c(2,3,4))
-plot(mi,c(5,6))
+summary(im)
+```
+
+```{r}
+plot(im,1)
+plot(im,2)
+plot(im,3)
+plot(im,6)
+if (length(variable_names) > 1 ) plot(im,7)
+if (length(variable_names) > 2 ) plot(im,8)
+
+plot(im,c(1,2))
+plot(im,c(1,3))
+plot(im,c(1,6))
+
+if (length(variable_names) > 1 ) plot(im,c(1,7))
+
+plot(im,c(2,3))
+plot(im,c(2,6))
+if (length(variable_names) > 1 ) plot(im,c(2,7))
+
+plot(im,c(1,2,6))
+plot(im,c(2,3,6))
+plot(im,c(4,5))
 ```
@@
 idata <- vocc_by_sp[["imm"]]
-
 # Xmax <- quantile(data$X, 0.9995)
 # Xmin <- quantile(data$X, 0.005)
-# std_max <- quantile(data$std_change, 0.9995)
+# std_max <- quantile(data$log_change, 0.9995)
 # 
@@
 
-ixy <- as.data.frame(cbind(idata$X, idata$Y))
+xy <- as.data.frame(cbind(idata$X, idata$Y))
 
@@
   # if (region == "West Coast Haida Gwaii") {
-  #    n_knots <- 60
+  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
   # }
 
-if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(ixy))}
+if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(xy))}
 
@@
 
-```{r }
-#glimpse(idata)
-
-idata$std_source_density2 <- idata$std_source_density^2
+```{r}
+idata <- vocc_by_sp[["imm"]]
+
+# tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + distance2") # + cell_type * std_source_density2
+# 
+# for (j in (variable_names) ){
+#   tmb_formula <- paste0("", tmb_formula, " + std.", j, "+ std.", j, "2"
+#   )
+# }
+
 itest <- sdmTMB(data = idata,
-      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 + 
-      cell_type * distance + cell_type * distance2 + 
-      std_end_climate + std_end_climate2, 
-      spde = ispde, silent = TRUE)
+      as.formula(tmb_formula), 
+      spde = ispde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      silent = TRUE)
 itest
-
 ```
@@
 #coefs$se
+coefs$model <- model_formula
 coefs$species <- spp 
@@
 coefs$max_thresholds <- max_thresholds
-coefs$optimum <- optimal_values
+coefs$optimum <- optimal_string
 coefs$start_time <- start_time
@@
 rm(other_coefs)
-try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
+try({other_coefs <- readRDS(paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))})
 
@@
 }
-saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")
+saveRDS(coefs, file = paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
 
@@
 coefs <-cbind(var, coefs)
-
+coefs$model <- model_formula
 coefs$species <- spp 
@@
 glimpse(coefs)
-try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
+try({other_coefs <- readRDS(paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))})
 if (exists("other_coefs")) {
@@
 }
-saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")
+saveRDS(coefs, file = "paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
 

### 2019-08-14 analysis/VOCC/3-variable-response-plots.Rmd
calculate optimal values for enviro variables

@@
 ```{r}
-#species <- params$species
+# species <- params$species
 # # Species run so far...
-species <- "Arrowtooth Flounder"
+# species <- "Arrowtooth Flounder"
 # species <- "Pacific Cod"
@@
 # species <- "Silvergray Rockfish"
-# species <- "Lingcod"
+ species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
@@
 
-Dissolved O2 plots
+Load Dissolved O2 models
 ```{r}
@@
 model_list <- list(
-  adult_survey_all=adult_survey_all, imm_survey_all=imm_survey_all,
+  adult_survey_all = adult_survey_all, imm_survey_all = imm_survey_all,
   adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
@@
 
-```{r}
-d <- list()
-do_plots <- list()
-for (i in seq_len(length(model_list))) {
-  d[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
-  do_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Dissolved O2", xlimits = c(0,8)) + ggtitle(paste(species, names(model_list[i])))
-}
-print(do_plots)
-```
-
-```{r}
-png(
-  file = paste0("figs/", spp, "/do-", spp, covs, "-priors-TRUE.png"), 
-  res = 600,
-  units = "in",
-  width = 8.5, 
-  height = 11
-)
-gridExtra::grid.arrange(
-  grobs = do_plots,
-  ncol = 2,
-  top = grid::textGrob(paste(species))
-)
-dev.off()
-```
-
-Temperature plots
+Load temperature models
 ```{r}
@@
 
+Save optimal values
+
+```{r}
+td <- list()
+
+for (i in seq_len(length(temp_model_list))) {
+  td[[i]] <- time_varying_density(temp_model_list[[i]], predictor = "temp")
+  #browser()
+  if (length(td[[i]])>0) td[[i]]$model <- names(temp_model_list[i])
+}
+
+tdat <- do.call("rbind", td)
+
+tdat <- tdat %>%
+  group_by(model) %>%
+  mutate(maxy = max(y_hat)) %>%
+  ungroup() %>%
+  mutate(meanymax = mean(maxy), sdymax = sd(maxy)) %>%
+  filter(maxy < (meanymax + 2 * sdymax))
+
+hist(tdat$maxy)
+
+optimal_temp <- get_optimal_value(tdat, xlimits = c(3, 8))
+
+do <- list()
+
+for (i in seq_len(length(model_list))) {
+ do[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
+ if (length(do[[i]])>0) do[[i]]$model <- names(model_list[i])
+}
+odat <- do.call("rbind", do)
+
+odat <- odat %>%
+  group_by(model) %>%
+  mutate(maxy = max(y_hat)) %>%
+  ungroup() %>%
+  mutate(meanymax = mean(maxy), sdymax = sd(maxy)) %>%
+  filter(maxy < (meanymax + 2 * sdymax))
+
+hist(odat$maxy)
+
+optimal_do <- get_optimal_value(odat, xlimits = c(0.5, 6))
+
+date <- Sys.time()
+
+params <- data.frame(species, optimal_do, optimal_temp, date)
+
+other_params <- readRDS(paste0("data/optimal-values-by-species.rds"))
+params <- rbind(other_params, params)
+saveRDS(params, file = "data/optimal-values-by-species.rds")
+
+params
+```
+
+Build DO plots
+
+```{r}
+d <- list()
+do_plots <- list()
+for (i in seq_len(length(model_list))) {
+  d[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
+  if (length(d[[i]]) == 0) {
+    do_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } else {
+    do_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Dissolved O2", xlimits = c(0, 8)) + ggtitle(paste(species, names(model_list[i])))
+  }
+}
+print(do_plots)
+```
+
+```{r}
+png(
+  file = paste0("figs/", spp, "/do-", spp, covs, "-priors-TRUE.png"),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 11
+)
+gridExtra::grid.arrange(
+  grobs = do_plots,
+  ncol = 2,
+  top = grid::textGrob(paste(species))
+)
+dev.off()
+```
+
+Build Temperature plots
+
 ```{r}
@@
   d[[i]] <- time_varying_density(temp_model_list[[i]], predictor = "temp")
-  #d[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
-  temp_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Temperature", xlimits = c(2,10)) +
-    ggtitle(paste(species,names(temp_model_list[i])))
+
+  if (length(d[[i]]) == 0) {
+    temp_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } else {
+    temp_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Temperature", xlimits = c(2, 10)) +
+      ggtitle(paste(species, names(temp_model_list[i])))
+  }
 }
@@
 
-Depth plots
+Load Depth models
 ```{r}
@@
 
+Build depth plots
 ```{r}
@@
   d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
+
+  if (length(d[[i]]) == 0) {
+    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } else {
     d[[i]]$x <- exp(d[[i]]$x)
-  depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0,1000)) + 
+    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, 1000)) +
       ggtitle(paste(species, names(depth_model_list[i])))
+  }
 }
@@
 png(
-  file = paste0("figs/", spp, 
-    "/depth-", spp, covs, "-priors-", priors, ".png"), 
+  file = paste0(
+    "figs/", spp,
+    "/depth-", spp, covs, "-priors-", priors, ".png"
+  ),
   res = 600,
@@
 Plot time-varying effect of trawling at mean depth
+
 ```{r}
@@
 
-  SE <- 0.35 #Place-holder
+  SE <- 0.35 # Place-holder
 
@@
 trawled <- data.frame(year = yrs, intercept = b_j, trawling = b3)
-
 ```

### 2019-08-14 R/plot-mountains.R

@@ -66,21 +66,45 @@ plot_mountains <- function(dat, time_varying = TRUE, variable_label = "Depth", x
 #' @param predictor Prefix for scaled parameter in model
 #'
 #' @export
-#'
 time_varying_density <- function(m, predictor = "depth") {
+  
   get_y_hat <- function(b0, b1, b2, year, 
     predictor, mean_column, sd_column
   ) {
-    
-   x_pred <- seq(min(m$data[[predictor]], na.rm = TRUE), max(m$data[[predictor]], na.rm = TRUE), length.out = 300)
+ 
+   x_pred <- seq(min(m$data[[predictor]], na.rm = TRUE), 
+     max(m$data[[predictor]], na.rm = TRUE), length.out = 300)
+   
+   # exclude curves that are increasing at extremes and therefore not capturing maxima
+   maxvalue <- exp(b0 + b1 * max(x_pred) + b2 * (max(x_pred)^2))
+   nearmax <- exp(
+     b0 + b1 * (max(x_pred) - m$data[[sd_column]][[1]]/10) + 
+       b2 * ((max(x_pred) - m$data[[sd_column]][[1]]/10)^2 )
+   )
+   
+   minvalue <- exp(b0 + b1 * min(x_pred) + b2 * (min(x_pred)^2))
+   nearmin <- exp(
+     b0 + b1 * (min(x_pred) + m$data[[sd_column]][[1]]/10) + 
+       b2 * ((min(x_pred) + m$data[[sd_column]][[1]]/10)^2 )
+   )
 
+   #browser()
+   
+   if (nearmax > maxvalue & nearmin > minvalue) {
+       
    data.frame(
-      x = (x_pred * m$data[[sd_column]][[1]] + m$data[[mean_column]][[1]]), # if depth, actually is log_depth so must exp(x) for raw depth
-      y_hat = exp(b0 + b1 * x_pred + b2 * x_pred^2), # leave these values as predicted for un-trawled zone
+      # if depth, actually is log_depth so must exp(x) for raw depth
+      x = (x_pred * m$data[[sd_column]][[1]] + m$data[[mean_column]][[1]]), 
+      # leave these values as predicted for un-trawled zone
+      y_hat = exp(b0 + b1 * x_pred + b2 * x_pred^2), 
       year = year
     )
-
+       
+  } else {
+     NULL
   }
+  }
+   
   r <- m$tmb_obj$report()
   r$b_rw_t
   b_j <- m$model$par
[…]

### 2019-08-14 analysis/VOCC/4-vocc-by-spp.Rmd
automate optimum value selection and save coefs

@@
 # )
+options(scipen = 999)
+options(digits = 2) 
 ```
@@
 # # Species run so far...
- species <- "Arrowtooth Flounder"
+# species <- "Arrowtooth Flounder"
 # species <- "Pacific Cod"
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish"
+ species <- "Silvergray Rockfish"
 # species <- "Lingcod"
@@
 # region <- "All synoptic surveys"
- region <- "West Coast Vancouver Island"
-# region <- "Both odd year surveys"
+# region <- "West Coast Vancouver Island"
+ region <- "Both odd year surveys"
 # region <- "West Coast Haida Gwaii"
@@
 #climate <- "temperature"
-temp_threshold <- c(0.75)
-optimal_temp <- 6.5 # yellow-eye 4
+temp_threshold <- c(0.5)
 
 #climate <- "do"
-do_threshold <- c(0.25)
-optimal_do <- 2 # yellow-eye 4
+do_threshold <- c(0.1)
 
 climate <- "do and temperature"
-max_thresholds <- c(Inf, 0.5)
-min_thresholds <- c(0.1, Inf)
-optimal_values <- c(2, 6.5)
-
-# critical_thresholds <- c(-0.25, 0.75)
-
+max_thresholds <- c(Inf, temp_threshold)
+min_thresholds <- c(do_threshold, Inf)
+```
+
+Get most up to date optimal values from mountain plots for most abundant year
+
+```{r}
+params <- readRDS(paste0("data/optimal-values-by-species.rds"))
+params <- params[rev(order(as.Date(params$date))), ]
+params <- params[!duplicated(params$species),]
+
+optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
+optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)
+
+optimal_values <- c(optimal_do, optimal_temp)
 ```
@@
 Run all subsequent code...
-
 ```{r, message=FALSE, warning=FALSE}
@@
 rm(vocc)
-# try({
-#   vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
-#     min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
-# })
+try({
+  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-",
+    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
+})
 
@@
 
-saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
+saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate_string, "-",
     min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
@@
 
-### Plot climate vectors
-```{r}
+### Plot all climate vectors
+```{r eval = FALSE}
 
@@
 
-Get matching biomass data for the above time periods
-
-Get all biomass predictions 
-```{r}
-priors <- FALSE
-# priors <- TRUE
-
-covariates <- "+muddy+any_rock"
-
-covs <- gsub("\\+", "-", covariates)
-
-rm(adult_predictions)
-rm(imm_predictions)
-rm(predictions)
-
-try({
-  adult_predictions <- readRDS(paste0("data/", spp,
-  "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
-})
-
-try({
-  imm_predictions <- readRDS(paste0("data/", spp,
-  "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))
-})
-
-try({
-  predictions <- readRDS(paste0("data/", spp,
-  "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
-})
-
-if (!exists("adult_predictions")) adult_predictions <- NULL
-if (!exists("imm_predictions")) imm_predictions <- NULL
-if (!exists("predictions")) predictions <- NULL
-
-biomass_list <- list(
-  adult = adult_predictions, 
-  imm = imm_predictions, 
-  total = predictions
-)
-
-biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
-
-# glimpse(biomass_list)
-```
-
-# Extract biomass values to match spatial and temporal range of climate vectors
-```{r}
-
-# #   #climate <- "do and temperature"
-# max_thresholds <- c(Inf, 0.5)
-# min_thresholds <- c(0.1, Inf)
-# optimal_values <- c(2, 6.5)
-
+Filter vectors to match optimal range for chosen species
+```{r}
 vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds)
@@
 if (length(vectors[ ,2]) == 0) {
-  other_coefs <- readRDS(paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
+  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
   coefs <- other_coefs[1,]
@@
   coefs <- rbind(other_coefs, coefs) %>% distinct()
-  saveRDS(coefs, file = paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
+  saveRDS(coefs, file = paste0("data/", climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
 
@@
 ```
+
+### Plot remaining climate vectors for regions where change exceeds thresholds
+```{r}
+
+if(climate== "temperature") {
+vocc_plot <- plot_vocc(vectors,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  white_zero = TRUE,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+)
+vocc_plot + ggtitle(climate)
+}
+
+if (climate == "do") {
+vocc_plot <- plot_vocc(vectors,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  white_zero = TRUE,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+)
+vocc_plot + ggtitle(climate)
+}
+
+if (climate == "do and temperature") {
+
+vocc_plot_temp <- plot_vocc(vectors,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "temp_diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  white_zero = TRUE,
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+)+ ggtitle("temperature")
+
+vocc$do_diff <- vocc$do_est.units_per_decade/10*delta_t_total
+
+vocc_plot_do <- plot_vocc(vectors,
+  vec_aes = "distance",
+  #max_vec_plotted = 100,
+  min_vec_plotted = input_cell_size,
+  NA_label = ".",
+  fill_col = "do_diff",
+  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+  raster_alpha = 1,
+  white_zero = TRUE,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  vec_alpha = 0.25,
+  axis_lables = FALSE,
+  transform_col = no_trans#,
+  #raster_limits = c(5, 13)
+) + ggtitle("do")
+
+vocc_plot <- gridExtra::grid.arrange(
+  vocc_plot_temp, vocc_plot_do,
+  ncol = 2,
+  top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1],"or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
+)
+
+}
+vocc_plot
+```
+
+Get matching biomass data for the above time periods
+
+Get all biomass predictions 
+```{r}
+priors <- FALSE
+# priors <- TRUE
+
+covariates <- "+muddy+any_rock"
+
+covs <- gsub("\\+", "-", covariates)
+
+rm(adult_predictions)
+rm(imm_predictions)
+rm(predictions)
+
+try({
+  adult_predictions <- readRDS(paste0("data/", spp,
+  "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
+})
+
+try({
+  imm_predictions <- readRDS(paste0("data/", spp,
+  "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))
+})
+
+try({
+  predictions <- readRDS(paste0("data/", spp,
+  "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
+})
+
+if (!exists("adult_predictions")) adult_predictions <- NULL
+if (!exists("imm_predictions")) imm_predictions <- NULL
+if (!exists("predictions")) predictions <- NULL
+
+biomass_list <- list(
+  adult = adult_predictions, 
+  imm = imm_predictions, 
+  total = predictions
+)
+
+biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
+
+# glimpse(biomass_list)
+```
+
+# Extract biomass values to match spatial and temporal range of climate vectors
 
@@
 )
-# browser()
+#browser()
 vectors <- vectors %>%
@@
         vector_id = paste0(id, "_to_", tid),
+        distance = ifelse(distance>120, 120, distance),
         distance2 = distance*distance,
@@
   filter(!is.na(icell)) %>% 
-
-  # remove potential source cells with lowest 1% quantile of fish biomass  
-  filter(source_density > quantile(source_density, 0.01) ) %>%
+  filter(!is.na(source_density)) %>% filter(!is.na(end_density)) %>%
+
+  # remove potential source cells with lowest 10% quantile of fish biomass  
+  filter(source_density > quantile(source_density, 0.1)) %>%
 
@@
     )
-
+#browser()
 end_climate <- purrr::map_df(strsplit(vocc_by_sp$end_climate, " "), 
@@
 })
-
-# glimpse(vocc_by_sp)
-# trim spatial outliers 
-# Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
-# Xmin <- quantile(vocc_by_sp[[i]]$X, 0.005)
-# std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
-# 
-# vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax ) 
 ```
@@
 data <- vocc_by_sp[["adult"]]
-#glimpse(data)
-
-#for (j in seq_along(variable_names)) {
 
@@
   geom_point(alpha = 0.5) +
-  # geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = log_change
-  # ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
+  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
+    formula = y ~ x + I(x^2), inherit.aes = FALSE) + # + I(x^3)
   # scale_size_continuous(guide = "none") +
-  scale_y_continuous(trans = fourth_root_power) +
+  scale_y_continuous(trans = log10) +
   xlab(paste("Final value of", j,"for cells")) +
@@
   scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
+  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.7)) + 
   ggtitle(paste(survey, "mature", species))
@@
 #data <- filter(vocc_by_sp[["adult"]],  distance < 90) # distance > 1 &
-#data <- filter(vocc_by_sp[["adult"]], distance > 1 ) #  &
 
@@
     #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-    geom_smooth(method="lm", formula = y~x+I(x^2), se = FALSE )+
+    geom_smooth(method="lm",  se = FALSE )+ # formula = y~x+I(x^2),
     scale_size_continuous(guide = "none") +
-    scale_y_continuous(trans = fourth_root_power) +
     facet_wrap(~cell_type) +#, scales = "free") +
@@
   scale_x_continuous(trans = sqrt) +
-  scale_y_continuous(trans = sqrt) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
+  #scale_y_continuous(trans = sqrt) +
+  ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Source density (dark points are closer to source)")) +
@@
   gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+  ggtitle(paste(survey, "mature", species))
 })
@@
   #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
+  ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
   xlab(paste("Source density (dark points are closer to source)")) +
@@
 
-```{r}
-data <- vocc_by_sp[["adult"]]
-data <- filter(vocc_by_sp[["adult"]], distance < 50)
-
-mature_plot5 <- lapply(variable_names, function(j) {  
-  ggplot(data, aes(
-      x = UQ(rlang::sym(paste0(j, ".units_per_decade"))) , 
-      y = (end_density - start_density), 
-      colour = cell_type, 
-      size = start_density)) +
-    geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-    #scale_size_continuous(guide = "none") +
-    #scale_y_continuous(trans = fourth_root_power) +
-    ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
-    xlab(paste("Change in", j," in units per decade")) +
-    scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-    ggtitle(paste(survey, "mature", species))
-})
-mature_plot5
-```
-
 ## Adult models
@@
 ```{r}
-# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
-#             std_end_climate + std_end_climate2 + 
-#             (1 | vector_id), 
-#  #   offset = log(start_density), 
-#  #   family = Gamma(link = "inverse")
-#      data = data)
-# hist(data$end_density)
-# hist(data$log_change)
-data <- vocc_by_sp[["adult"]]
-
-model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * std_source_density2 + cell_type * distance ") # + cell_type * distance2
+data <- vocc_by_sp[["adult"]] #%>% filter(distance < 50)
+
+model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance") # + cell_type * std_source_density2 + cell_type * distance2") #
 
@@
 
-library(sjPlot)
+#library(sjPlot)
 #library(sjlabelled)
 #library(sjmisc)
-sjPlot::plot_model(test, title = "Log biomass change")
+sjPlot::plot_model(test, title = paste("Log adult", species, "biomass change"))
 ```
@@
 
-glimpse(coefs)
 rm(other_coefs)
@@
 if (exists("other_coefs")) {
-coefs <- rbind(other_coefs, coefs)
-}
+coefs <- rbind(other_coefs, coefs) %>% distinct() %>% 
+ mutate_if(is.numeric, funs(signif(.,digits = 2)))
+}
+
+coefs <- coefs %>% distinct()
+
 saveRDS(coefs, file = paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))
@@
 # 
-
 ```
@@
 
-m <- gbm::gbm(
+mboost <- gbm::gbm(
   as.formula(gbm_formula), 
@@
   )
-
-# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
-#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
-```
-
-```{r}
-summary(m)
-```
-
-```{r}
-plot(m,1)
-plot(m,2)
-plot(m,3)
-plot(m,6)
-if (length(variable_names) > 1 ) plot(m,7)
-if (length(variable_names) > 2 ) plot(m,8)
-
-plot(m,c(1,2))
-plot(m,c(1,3))
-plot(m,c(1,6))
-
-if (length(variable_names) > 1 ) plot(m,c(1,7))
-
-plot(m,c(2,3))
-plot(m,c(2,6))
-if (length(variable_names) > 1 ) plot(m,c(2,7))
-
-plot(m,c(1,2,6))
-plot(m,c(2,6,3))
-plot(m,c(4,5))
+```
+
+```{r}
+summary(mboost)
+```
+
+```{r}
+plot(mboost,3)
+
+plot(mboost,c(1,6,3))
+if (length(variable_names) > 1 ) plot(mboost,c(1,7,3))
+if (length(variable_names) > 2 ) plot(mboost,c(1,8,3))
+
+plot(mboost,c(2,6,3))
+if (length(variable_names) > 1 ) plot(mboost,c(2,7,3))
+if (length(variable_names) > 2 ) plot(mboost,c(2,8,3))
+```
+
+Other gbm plots
+```{r eval=FALSE}
+plot(mboost, 1)
+plot(mboost,2)
+plot(mboost,6)
+if (length(variable_names) > 1 ) plot(mboost,7)
+if (length(variable_names) > 2 ) plot(mboost,8)
+
+plot(mboost,c(1,2))
+plot(mboost,c(1,3))
+plot(mboost,c(1,6))
+
+if (length(variable_names) > 1 ) plot(mboost,c(1,7))
+if (length(variable_names) > 2 ) plot(mboost,c(1,8))
+
+plot(mboost,c(2,3))
+plot(mboost,c(2,6))
+if (length(variable_names) > 1 ) plot(mboost,c(2,7))
+if (length(variable_names) > 2 ) plot(mboost,c(2,8))
+
+plot(mboost,c(4,5))
 ```
@@
 
-tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + distance2") # + cell_type * std_source_density2
+tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance") # + cell_type * distance2 cell_type * std_source_density2 +
 
@@
 
-test <- sdmTMB(data = data,
+mtmb <- sdmTMB(data = data,
       as.formula(tmb_formula), 
@@
       silent = TRUE)
-test
-```
-
-```{r}
-
-p <- predict(test)
-p$residuals <- residuals(test)
+mtmb 
+```
+
+```{r}
+p <- predict(mtmb)
+p$residuals <- residuals(mtmb)
 #idata$resids <- p$residuals
@@
 
-#for (j in seq_along(variable_names)) {
-
 immature_plot <- lapply(variable_names, function(j) { 
-ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density # - start_density
-  ), colour = cell_type, size = start_density, 
+ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density), 
+    colour = cell_type, size = start_density, 
     group = cell_type)) +
   geom_point(alpha = 0.5) +
-  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density
-  ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
+  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
+    formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
   scale_size_continuous(guide = "none") +
-  scale_y_continuous(trans = fourth_root_power) +
+  scale_y_continuous(trans = log10) +
   xlab(paste("Final value of", j,"for cells")) +
@@
     #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-    geom_smooth(method="lm", formula = y~x+I(x^2) )+
+    geom_smooth(method="lm") + #, formula = y~x+I(x^2) 
     scale_size_continuous(guide = "none") +
-    scale_y_continuous(trans = fourth_root_power) +
     facet_wrap(~cell_type) +#, scales = "free") +
-    ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
+    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
     xlab(paste("Distance traveled to stay within critical range of", j,"")) +
@@
     # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
+    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.7)) + 
     ggtitle(paste(survey, "immature", species))
@@
   gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
+  ggtitle(paste(survey, "immature", species))
 })
@@
 
-```{r}
-data <- vocc_by_sp[["imm"]]
-data <- filter(vocc_by_sp[["imm"]], distance < 50)
-
-immature_plot4 <- lapply(variable_names, function(j) {  
-ggplot(filter(data, distance < 10), aes(source_density, y = log_change , 
-    colour = UQ(rlang::sym(paste0("end_", j))),
-    alpha = 1/(distance + input_cell_size), #size = 1/distance,
-    group = cell_type)) +
-  geom_jitter(width = 0.01) +
-  geom_smooth(method="lm", aes(group = cell_type)) +
-  facet_wrap(~cell_type) +#, scales = "free") +
-  scale_alpha_continuous(guide = "none") +
-  scale_size_continuous(guide = "none") +
-  scale_x_continuous(trans = sqrt) +
-  #scale_y_continuous(trans = fourth_root_power) +
-  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
-  xlab(paste("Source density (dark points are closer to source)")) +
-  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
-  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
-  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
-})
-immature_plot4
-```
-
-```{r}
-data <- vocc_by_sp[["imm"]]
-data <- filter(vocc_by_sp[["imm"]], distance < 50)
-
-immature_plot5 <- lapply(variable_names, function(j) {  
-  ggplot(data, aes(
-      x = UQ(rlang::sym(paste0(j, ".units_per_decade"))) , 
-      y = (end_density - start_density), 
-      colour = cell_type, 
-      size = start_density)) +
-    geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
-    #scale_size_continuous(guide = "none") +
-    #scale_y_continuous(trans = fourth_root_power) +
-    ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
-    xlab(paste("Change in", j," in units per decade")) +
-    scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
-    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.7)) + 
-    ggtitle(paste(survey, "immature", species))
-})
-immature_plot5
-```
-
 ## Immature models
@@
 
-model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * std_source_density2 + cell_type * distance ") # + cell_type * distance2
-
-for (j in (variable_names) ){
-  model_formula <- paste0("", model_formula, " + std.", j, "+ std.", j, "2"
-  )
-}
-
-model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
+# for (j in (variable_names) ){
+#   model_formula <- paste0("", model_formula, " + std.", j, "+ std.", j, "2"
+#   )
+# }
+# 
+# model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
 
@@
 #library(sjmisc)
-sjPlot::plot_model(itest, title = "Log biomass change")
+sjPlot::plot_model(itest, title = paste("Log immature", species, "biomass change"))
 ```
@@
 
-glimpse(coefs)
 rm(other_coefs)
-try({other_coefs <- readRDS(paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))})
+try({other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))})
 
 if (exists("other_coefs")) {
-coefs <- rbind(other_coefs, coefs)
-}
-saveRDS(coefs, file = paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))
+coefs <- rbind(other_coefs, coefs) %>% distinct() %>% 
+ mutate_if(is.numeric, funs(signif(.,digits = 2)))
+}
+
+coeds <- coefs %>% distinct()
+
+saveRDS(coefs, file = paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
 ```
@@
   )
-
-# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
-#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
 ```
@@
 ```{r}
+plot(im,3)
+
+plot(im,c(1,6,3))
+if (length(variable_names) > 1 ) plot(im,c(1,7,3))
+if (length(variable_names) > 2 ) plot(im,c(1,8,3))
+
+plot(im,c(2,6,3))
+if (length(variable_names) > 1 ) plot(im,c(2,7,3))
+if (length(variable_names) > 2 ) plot(im,c(2,8,3))
+```
+
+Other gbm plots
+```{r eval=FALSE}
 plot(im,1)
 plot(im,2)
-plot(im,3)
 plot(im,6)
@@
 if (length(variable_names) > 1 ) plot(im,c(1,7))
+if (length(variable_names) > 2 ) plot(im,c(1,8))
 
@@
 if (length(variable_names) > 1 ) plot(im,c(2,7))
-
-plot(im,c(1,2,6))
-plot(im,c(2,3,6))
+if (length(variable_names) > 2 ) plot(im,c(2,8))
+
 plot(im,c(4,5))
@@
 idata <- vocc_by_sp[["imm"]]
-# Xmax <- quantile(data$X, 0.9995)
-# Xmin <- quantile(data$X, 0.005)
-# std_max <- quantile(data$log_change, 0.9995)
-# 
-# data <- data %>% filter ( X >= Xmin & X <= Xmax )
 
@@
 
-itest <- sdmTMB(data = idata,
+itmb  <- sdmTMB(data = idata,
       as.formula(tmb_formula), 
@@
       silent = TRUE)
-itest
-```
-
-```{r}
-ip <- predict(itest)
-ip$residuals <- residuals(itest)
+itmb
+```
+
+```{r}
+ip <- predict(itmb)
+ip$residuals <- residuals(itmb)
 #idata$resids <- p$residuals
@@
 ```{r}
-m <- test
+m <- mtmb
 
@@
 if (exists("other_coefs")) {
-coefs <- rbind(other_coefs, coefs)
+coefs <- rbind(other_coefs, coefs) %>% distinct() %>% 
+ mutate_if(is.numeric, funs(signif(.,digits = 2)))
 }
 saveRDS(coefs, file = paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
-
-data$mean_source_density[1]
-range(data$source_density)
-```
-
-```{r}
-m <- itest
+```
+
+```{r}
+m <- itmb
 
@@
 names(coefs) <- c("beta","se")
-coefs$ci <- coefs$se*1.96
+#coefs$beta <- signif(coefs$beta, digits = 3)
+#coefs$se <- signif(coefs$se, digits = 3)
+coefs$ci <- signif(coefs$se*1.96, digits = 3)
 var <- varnames
@@
 coefs <-cbind(var, coefs)
-coefs$model <- model_formula
 coefs$species <- spp 
@@
 coefs$ssid_string <- ssid_string
-
-glimpse(coefs)
-try({other_coefs <- readRDS(paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))})
+coefs$model <- model_formula
+
+try({other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))})
+
 if (exists("other_coefs")) {
-coefs <- rbind(other_coefs, coefs)
-}
-saveRDS(coefs, file = "paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
+coefs <- rbind(other_coefs, coefs) %>% distinct() %>% 
+ mutate_if(is.numeric, funs(signif(.,digits = 2)))
+}
+
+saveRDS(coefs, file =paste0("data/vocc-", climate_string, "-by-spp-sdmTMB-coefs.RDS"))
 
@@
 
-```{r}
-# png(
-#   file = paste0("figs/", spp, "/vocc-biomass-change", spp, ssid_string, ".png"),
-#   res = 600,
-#   units = "in",
-#   width = 5,
-#   height = 4
-# )
-# 
-# do_vect_plot
-# dev.off()
-```
+```{r eval=FALSE}
+lmer_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
+
+lmer_coefs$beta_upper <- lmer_coefs$beta + lmer_coefs$ci 
+lmer_coefs$beta_lower <- lmer_coefs$beta - lmer_coefs$ci
+
+View(lmer_coefs)
+
+beta_target <- lmer_coefs %>% filter(ssid_string == "1n3") %>% filter(var == "cell_typetarget")
+
+View(beta_target)
+
+```

### 2019-08-16 analysis/VOCC/4-vocc-BACI.Rmd
first pass on BACI of biomass change for VOCC source cells

---
title: "BACI of biomass change for VOCC source cells"
author: "Philina English"
date: '2019-08-15'
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

#### Set species
```{r echo=TRUE}
# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, echo=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

params <- readRDS(paste0("data/optimal-values-by-species.rds"))
params <- params[rev(order(as.Date(params$date))), ]
params <- params[!duplicated(params$species), ]

optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)

optimal_values <- c(optimal_do, optimal_temp)
```

#### Set vector type
```{r echo=TRUE}
 climate <- "temperature"
# climate <- "do"
# climate <- "do and temperature"
```

#### Make climate vectors for biannual changes between surveys
```{r echo=TRUE}
trend <- FALSE
start_time <- 2012
end_time <- 2015
temp_threshold <- c(0.1)
do_threshold <- c(0.1)
```

# OR

#### Make climate vectors for average changes and lower threshold values
```{r echo=TRUE}
# trend <- TRUE
# temp_trend_threshold <- c(0.25)
# do_trend_threshold <- c(0.2)
```

#### Set Region 
```{r echo=TRUE}
# region <- "West Coast Vancouver Island"
region <- "both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

Run all subsequent code...
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
# library(gfplot)
# library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2, 2)
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}
```

```{r, include=FALSE}
# Set default cell sizes and scale
input_cell_size <- 2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  delta_threshold <- temp_threshold
  optimal_values <- optimal_temp

  if (trend) {
    max_thresholds <- temp_trend_threshold
    start_time <- 2008 # change if can add back in prior 2008 data
    end_time <- NULL
    delta_t_total <- 6 # change if can add back in prior 2008 data
    indices <- trend_indices_temp
  } else {
    max_thresholds <- temp_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  delta_threshold <- -1 * (do_threshold)
  optimal_values <- optimal_do

  if (trend) {
    min_thresholds <- do_trend_threshold
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- do_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do and temperature") {
  variable_names <- c("do_est", "temp")

  if (trend) {
    min_thresholds <- c(do_trend_threshold, Inf)
    max_thresholds <- c(Inf, temp_trend_threshold)
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- c(do_threshold, Inf)
    max_thresholds <- c(Inf, temp_threshold)
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")

optimal_string <- paste0(optimal_values, collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
```

# Calculate climate vectors

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
})

if (!exists("vocc")) {
  pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }

  starttime <- Sys.time()
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = min_thresholds,
    max_thresholds = max_thresholds,
    round_fact = 10
  )
  endtime <- Sys.time()
  time_vocc <- round(starttime - endtime)

  vocc$var_1_min <- min_thresholds[1]
  vocc$var_2_min <- min_thresholds[2]
  vocc$var_1_max <- max_thresholds[1]
  vocc$var_2_max <- max_thresholds[2]

  saveRDS(vocc, file = paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
}
# glimpse(vocc)
```

Plot all possible vectors
```{r echo=FALSE}
if (climate == "temperature") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```

Filter vectors to match optimal range for chosen species
```{r}
vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds,
  cell_size = input_cell_size, dist_intercept = input_cell_size/2,
  max_dist = 120
)

if (length(vectors[, 2]) == 0) {
  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
  coefs <- other_coefs[1, ]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$model <- "NA"
  coefs$species <- spp
  coefs$lifestage <- "NA"
  coefs$climate <- climate
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  coefs <- rbind(other_coefs, coefs) %>% distinct()
  saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))

  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
}
# glimpse(vocc)
# glimpse(vectors)
```

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r echo=FALSE, warning=FALSE}

if (climate == "temperature") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vectors$temp_diff <- vectors$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vectors$do_diff <- vectors$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("DO of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```
# Add species biomass data

#### Get all biomass predictions 
```{r, echo=FALSE, warning=FALSE}
priors <- FALSE
# priors <- TRUE

covariates <- "+muddy+any_rock"

covs <- gsub("\\+", "-", covariates)

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

try({
  adult_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
  ))
})

try({
  imm_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
  ))
})

try({
  predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
  ))
})

if (!exists("adult_predictions")) adult_predictions <- NULL
if (!exists("imm_predictions")) imm_predictions <- NULL
if (!exists("predictions")) predictions <- NULL

biomass_list <- list(
  adult = adult_predictions,
  imm = imm_predictions,
  total = predictions
)

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]

# glimpse(biomass_list)
```

#### Extract biomass values to match spatial and temporal range of climate vectors

```{r}
vocc_by_sp <- lapply(biomass_list, function(i) {
  biomass_change <- make_trend_data(i,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )

  biomass_covs <- i %>% select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>% distinct()

  biomass_change <- biomass_change %>% 
    rename(start_density = est_exp_s, end_density = est_exp_e) %>%
    dplyr::select(icell, x, y, start_density, end_density)
    # dplyr::select(-N, -time_step, - units_per_decade, -slope)

  biomass  <- left_join(biomass_change,  biomass_covs, by = c("x" = "X", "y"="Y"))

  # keep just one row for each cell that exceeded the end climate threshold 
  source_cells <- vectors %>% distinct(icell, .keep_all = TRUE) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  source_biomass <- left_join(source_cells, biomass, by = c("icell", "x", "y")) %>%
    mutate(cell_type = "source") %>% filter(!is.na(start_density))

  # identify all possible control cells, assume distance = 0 given 
  control_cells <- anti_join(vocc, vectors, by = c("icell", "x", "y")) %>% distinct(icell, .keep_all = TRUE) %>% 
    mutate(distance = 0) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  control_biomass <- left_join(control_cells, biomass, by = c("icell", "x", "y")) %>%
    mutate(cell_type = "control") %>% filter(!is.na(start_density))

#browser()
  alldata <- rbind(source_biomass, control_biomass) 
  #alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0))
  #fit <- McSpatial::matchprop(cell_type ~ start_density + depth, data= alldata, ytreat = "source")
  fit_formula <- paste0("cell_type ~ start_density + depth") 
 # fit_formula <- paste0("cell_type ~ start_density + depth + trawled + muddy + any_rock + omega_s")

  for (j in (variable_names)) {
  fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  }

  matched <- McSpatial::matchprop(as.formula(fit_formula), data = alldata, ytreat = "source")

  matched_time0 <- matched %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
  matched_time1 <- matched %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)

  for (j in seq_along(variable_names)) {
    climate <- paste0(variable_names[j])
    matched_time0 <- matched_time0 %>% 
      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    matched_time1 <- matched_time1 %>% 
      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
   }

  vocc_by_sp <- rbind(matched_time0, matched_time1)

  # calculate standardized climate variables
  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])
    std.var <- paste0("std_", variable_names[j])
    std.var2 <- paste0("std_", variable_names[j], "2")

    vocc_by_sp <- mutate(
      vocc_by_sp,
      (!!mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!std.var) := ((UQ(rlang::sym(paste0(variable_names[j]))) - 
          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) / 
          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE))*2),
      (!!std.var2) := (( UQ(rlang::sym(paste0(variable_names[j]))) - mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) ) / (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)*2))^2
    )
  }

  # rename spatial coordinates for inclusion in models and plots without naming conflicts
  vocc_by_sp <- vocc_by_sp %>%
    rename(X = x, Y = y)

  vocc_by_sp
})
```

## Adults

```{r}
data <- vocc_by_sp[["adult"]]
#View(data)
data$log_density <-  log(data$density)

ggplot(data, aes(y=log_density, x=temp, colour=as.factor(after))) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C")

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)
```

```{r}
m <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + (1|matchobs), data=data)
summary(m)
```

```{r}
sjPlot::plot_model(m, title = paste(species, "adult"))
```

## Immatures

```{r}
idata <- vocc_by_sp[["imm"]]
#View(data)
idata$log_density <-  log(idata$density)

ggplot(idata, aes(y=log_density, x=temp, colour=as.factor(after))) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C")

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(idata, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)
```

```{r}
mi <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + (1|matchobs), data=idata)
summary(mi)
```

```{r}
sjPlot::plot_model(mi, title = paste(species, "immature"))
```

### 2019-08-20 analysis/VOCC/4-vocc-BACI.Rmd
add gamm to BACI analysis

@@
 end_time <- 2015
-temp_threshold <- c(0.1)
+temp_threshold <- c(0.5)
 do_threshold <- c(0.1)
@@
 library(ggplot2)
-# library(gfplot)
+library(gfplot)
 # library(gfdata)
 library(sdmTMB)
+#library(itsadug)
 library(gfranges)
@@
   min_thresholds <- c(Inf)
-  delta_threshold <- temp_threshold
+  #delta_threshold <- temp_threshold
   optimal_values <- optimal_temp
@@
   max_thresholds <- c(Inf)
-  delta_threshold <- -1 * (do_threshold)
+  #delta_threshold <- -1 * (do_threshold)
   optimal_values <- optimal_do
@@
 
-if (length(vectors[, 2]) == 0) {
-  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
-  coefs <- other_coefs[1, ]
-  coefs$var <- "NA"
-  coefs$beta <- "NA"
-  coefs$se <- "NA"
-  coefs$model <- "NA"
-  coefs$species <- spp
-  coefs$lifestage <- "NA"
-  coefs$climate <- climate
-  coefs$min_thresholds <- min_string
-  coefs$max_thresholds <- max_string
-  coefs$optimum <- optimal_string
-  coefs$start_time <- start_time
-  coefs$timespan <- delta_t_total
-  coefs$region <- region
-  coefs$ssid_string <- ssid_string
-
-  coefs <- rbind(other_coefs, coefs) %>% distinct()
-  saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
-
-  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
-}
+# if (length(vectors[, 2]) == 0) {
+#   other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
+#   coefs <- other_coefs[1, ]
+#   coefs$var <- "NA"
+#   coefs$beta <- "NA"
+#   coefs$se <- "NA"
+#   coefs$model <- "NA"
+#   coefs$species <- spp
+#   coefs$lifestage <- "NA"
+#   coefs$climate <- climate
+#   coefs$min_thresholds <- min_string
+#   coefs$max_thresholds <- max_string
+#   coefs$optimum <- optimal_string
+#   coefs$start_time <- start_time
+#   coefs$timespan <- delta_t_total
+#   coefs$region <- region
+#   coefs$ssid_string <- ssid_string
+# 
+#   coefs <- rbind(other_coefs, coefs) %>% distinct()
+#   saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
+# 
+#   stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
+# }
 # glimpse(vocc)
@@
 
+Prep data
 ```{r}
@@
 data$log_density <-  log(data$density)
+data$distance2 <- data$distance^2
 
-ggplot(data, aes(y=log_density, x=temp, colour=as.factor(after))) +
+data <- mutate(data, time = ifelse(after==1, "end", "begin"))
+data$matchobs <- as.factor(data$matchobs)
+data$origobs  <- as.factor(data$origobs)   
+```
+
+Plot data
+```{r}
+ggplot(data, aes(y=log_density, x=temp, colour=time)) +
   geom_point(alpha=0.35) +
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~after)
+  facet_wrap(~time)
 
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~after)
+  facet_wrap(~time)
 ```
 
+Basic BACI
 ```{r}
-m <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + (1|matchobs), data=data)
+
+m <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + (1|icell) + (1|matchobs), data=data)
 summary(m)
@@
 ```{r}
-sjPlot::plot_model(m, title = paste(species, "adult"))
+sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,7,5,6,3,4), title = paste(species, "adult"))
+```
+
+GAMM
+```{r}
+gam4 <- gamm4::gamm4(log_density~ time * cell_type + s(std_temp) + s(distance) + s(X,Y), 
+  random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+
+plot(gam4$gam,pages=1)
+
+summary(gam4$gam) ## summary of gam
+summary(gam4$mer) ## underlying mixed model
+anova(gam4$gam)
+
+# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
+#   parametricOnly = TRUE, main="") + scale_y_reverse()
+
+plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
+   parametricOnly = TRUE, main="", xlab ="log_density") 
 ```
@@
 idata$log_density <-  log(idata$density)
+idata$distance2 <- idata$distance^2
 
-ggplot(idata, aes(y=log_density, x=temp, colour=as.factor(after))) +
+idata <- mutate(idata, time = ifelse(after==1, "end", "begin"))
+idata$matchobs <- as.factor(idata$matchobs)
+idata$origobs  <- as.factor(idata$origobs)   
+```
+
+Plot data
+```{r}
+ggplot(idata, aes(y=log_density, x=temp, colour=time)) +
   geom_point(alpha=0.35) +
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~after)
+  facet_wrap(~time)
 ```
 
+Basic BACI
 ```{r}
-mi <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + (1|matchobs), data=idata)
+mi <- lmerTest::lmer(log_density ~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + (1|icell) + (1|matchobs), data=idata)
 summary(mi)
@@
 ```{r}
-sjPlot::plot_model(mi, title = paste(species, "immature"))
+sjPlot::plot_model(mi, type = "std2", title = paste(species, "immature"))
 ```
+
+GAMM
+```{r}
+igam4 <- gamm4::gamm4(log_density~ time * cell_type + s(std_temp) + s(distance) + s(X,Y), 
+  random = ~(1|matchobs) + (1|origobs), data=idata) # + s(X,Y, bs="re")
+
+plot(igam4$gam,pages=1)
+
+summary(igam4$gam) ## summary of gam
+summary(igam4$mer) ## underlying mixed model
+anova(igam4$gam)
+
+# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
+#   parametricOnly = TRUE, main="") + scale_y_reverse()
+
+plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
+   parametricOnly = TRUE, main="", xlab ="log_density") 
+```
+
+Misc exploring gamm 
+```{r eval=FALSE}
+# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
+# summary(mi)
+# 
+# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(distance) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
+# summary(mi)
+# mi
+# 
+# 
+# idata$after <- as.factor(idata$after)
+# idata$matchobs <- as.factor(idata$matchobs)
+# 
+# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(distance) + #s(X) + s(Y) + 
+#     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")
+# 
+# 
+# plot(m4b$gam,pages=1)
+# 
+# summary(m4b$gam) ## summary of gam
+# summary(m4b$mer) ## underlying mixed model
+# anova(m4b$gam) 
+# 
+# gamtabs(m1)
+# 
+# #library(itsadug)
+# plot_parametric(m4b$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")))
+# 
+# library(mgcv)
+```

### 2019-08-20 analysis/VOCC/4-vocc-BACI.Rmd
automate changing variables

@@
 ```{r echo=TRUE}
- climate <- "temperature"
-# climate <- "do"
+# climate <- "temperature"
+ climate <- "do"
 # climate <- "do and temperature"
@@
 #View(data)
+
 data$log_density <-  log(data$density)
@@
 ```{r}
-ggplot(data, aes(y=log_density, x=temp, colour=time)) +
+lapply(variable_names, function(j) {
+ggplot(data, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
   geom_point(alpha=0.35) +
   scale_colour_viridis_d(option="C") 
+})
 
@@
 ```{r}
+for (j in (variable_names)) {
+  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + distance + distance2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs)")
+}
 
-m <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + (1|icell) + (1|matchobs), data=data)
+m <- lmerTest::lmer(as.formula(baci_formula), data=data)
 summary(m)
@@
 ```{r}
-gam4 <- gamm4::gamm4(log_density~ time * cell_type + s(std_temp) + s(distance) + s(X,Y), 
-  random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+
+for (j in (variable_names)) {
+  gamm_formula <- paste0("log_density ~ time * cell_type + s(distance) + s(std_", j, ") + s(X,Y)")
+}
+
+gam4 <- gamm4::gamm4(as.formula(gamm_formula),random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
 
@@
 ```{r}
-ggplot(idata, aes(y=log_density, x=temp, colour=time)) +
+lapply(variable_names, function(j) {
+ggplot(idata, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
   geom_point(alpha=0.35) +
   scale_colour_viridis_d(option="C") 
-
+})
 ggplot(idata, aes(log_density, fill = cell_type)) +
@@
 ```{r}
-mi <- lmerTest::lmer(log_density ~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + (1|icell) + (1|matchobs), data=idata)
+mi <- lmerTest::lmer(as.formula(baci_formula), data=idata)
 summary(mi)
@@
 ```{r}
-igam4 <- gamm4::gamm4(log_density~ time * cell_type + s(std_temp) + s(distance) + s(X,Y), 
-  random = ~(1|matchobs) + (1|origobs), data=idata) # + s(X,Y, bs="re")
+# for (j in (variable_names)) {
+#   gamm_formula <- paste0("log_density ~ time * cell_type + s(distance) + s(std_", j, ") + s(X,Y)")
+# }
+
+igam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=idata) # + s(X,Y, bs="re")
 

### 2019-08-21 analysis/VOCC/4-vocc-BACI.Rmd
exploring match function

@@
 ```{r echo=TRUE}
-# climate <- "temperature"
- climate <- "do"
+ climate <- "temperature"
+# climate <- "do"
 # climate <- "do and temperature"
@@
   alldata <- rbind(source_biomass, control_biomass) 
-  #alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0))
+  alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
+    select(-var_2_min, -var_2_max) %>% 
+    rename(vect_dist = distance)
+
+  # control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.05)
+  # alldata <- filter(alldata, start_density >= control_min)
+
+  low_biomass <- quantile(alldata$start_density, 0.1)
+  alldata <- filter(alldata, start_density >= low_biomass )
+
+  # if (order == "decreasing") 
+  #alldata <-   alldata[rev(order(alldata$start_density)), ]
+  # 
+  # if (order == "increasing") 
+  alldata <- alldata[(order(alldata$start_density)), ]
+
+  #browser()
+
+  alldata$depth2 <- alldata$depth^2
+
+  #fit_formula <- paste0("treat ~ start_density") 
+  #fit_formula <- paste0("treat ~ start_density + depth") 
+  fit_formula <- paste0("treat ~ start_density + depth + trawled + muddy + any_rock + omega_s + x + y")
+
+  # need to be standardized? seems not
+  # alldata$start_density.st <- (arm::rescale(alldata$start_density))*10 
+  # alldata$depth.st <- arm::rescale(alldata$depth) 
+  # alldata$muddy.st <- arm::rescale(alldata$muddy) 
+  # alldata$any_rock.st <- arm::rescale(alldata$any_rock) 
+  # alldata$omega_s.st <- arm::rescale(alldata$omega_s) 
+  # alldata$trawled.st <- arm::rescale(alldata$trawled) 
+  # fit_formula <- paste0("treat ~ start_density.st + depth.st + trawled.st + muddy.st + any_rock.st") #+ + omega_s
+
   #fit <- McSpatial::matchprop(cell_type ~ start_density + depth, data= alldata, ytreat = "source")
-  fit_formula <- paste0("cell_type ~ start_density + depth") 
- # fit_formula <- paste0("cell_type ~ start_density + depth + trawled + muddy + any_rock + omega_s")
 
@@
 
-  matched <- McSpatial::matchprop(as.formula(fit_formula), data = alldata, ytreat = "source")
+  # matched <- MatchIt::matchit(as.formula(fit_formula), data = alldata, method="nearest", discard="both")
+  # mdata <- MatchIt::match.data(matched)
 
-  matched_time0 <- matched %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
-  matched_time1 <- matched %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)
+  # Uses logit matching for which most matches are higher that treatment values
+  # mdata <- McSpatial::matchprop(as.formula(fit_formula), data = alldata, ytreat = 1, discard = "both", m.order = "decreasing")
+
+  # Uses n mahalanobis distances
+  mdata <- McSpatial::matchprop(as.formula(fit_formula), data = alldata, ytreat = 1, distance="probit", discard = "both", reestimate=TRUE) #m.order = "random", 
+  #mdata <- McSpatial::matchmahal(as.formula(fit_formula), data = alldata, ytreat = 1, distance="mahal", discard = "both") #, reestimate=TRUE)
+  #mdata <- McSpatial::matchmahal(as.formula(fit_formula), data = alldata, ytreat = 1, distance="probit", discard = "both") #, reestimate=TRUE) , m.order = "random"
+
+  mdata <- mdata %>% rename(distance = vect_dist) #%>% select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)
+
+  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
+  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)
 
@@
 data <- vocc_by_sp[["adult"]]
-#View(data)
+# View(data)
 
@@
 ```{r}
+
+ggplot(filter(data, as.numeric(matchobs) <= 150), aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
+  geom_point(alpha=0.4, size =4) +
+  scale_shape_manual(values= c(4, 1)) +
+  scale_colour_viridis_d(option="C", guide = FALSE)
+
 lapply(variable_names, function(j) {
@@
 ggplot(data, aes(log_density, fill = cell_type)) +
-  geom_histogram(alpha=0.5) +
+  geom_histogram(alpha=0.45) +
   scale_fill_viridis_d(option="C") +
@@
 
-gam4 <- gamm4::gamm4(as.formula(gamm_formula),random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+
+# for (j in (variable_names)) {
+#   gamm_formula <- paste0("log_density ~ time * cell_type + s(distance) + s(std_", j, ") + te(X,Y)")
+# }
+# gam <- mgcv::gamm(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
 
@@
 
+#my_cols <- c("#E69F00", "#56B4E9")
 plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
-   parametricOnly = TRUE, main="", xlab ="log_density") 
+   parametricOnly = TRUE, main="", xlab ="log_density"
+  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
+  ) 
 ```

### 2019-08-22 analysis/VOCC/4-vocc-BACI.Rmd
exact matching option

--- 2019-08-20 analysis/VOCC/4-vocc-BACI.Rmd
@@
 ```{r}
+#i <- biomass_list[[1]]
+
 vocc_by_sp <- lapply(biomass_list, function(i) {
@@
   alldata <- rbind(source_biomass, control_biomass) 
-  #alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0))
-  #fit <- McSpatial::matchprop(cell_type ~ start_density + depth, data= alldata, ytreat = "source")
-  fit_formula <- paste0("cell_type ~ start_density + depth") 
- # fit_formula <- paste0("cell_type ~ start_density + depth + trawled + muddy + any_rock + omega_s")
+  alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
+    select(-var_2_min, -var_2_max) %>% 
+    rename(vect_dist = distance)
+  alldata$treat <- as.factor(alldata$treat)
+  ggplot(alldata, aes(x=cell_type, y=log(start_density))) + geom_boxplot()
 
+  ## still including this trimming to reduce size of dataset 
+  control_min <- min(alldata$start_density[alldata$treat=="0"])
+  source_min <- min(alldata$start_density[alldata$treat=="1"])
+  # control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.001)
+  # source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.001)
+  group_min <- max(control_min, source_min)
+
+  control_max <- max(alldata$start_density[alldata$treat=="0"])
+  source_max <- max(alldata$start_density[alldata$treat=="1"])
+  # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 0.95)
+  # source_max <- quantile(alldata$start_density[alldata$treat=="1"], 0.95)
+  group_max <- min(control_max, source_max)
+  # 
+   alldata <- filter(alldata, start_density >= group_min) #%>% filter(start_density <= group_max)
+  # 
+  # num_source <- nrow(alldata[alldata$treat=="1", ])
+  #
+
+  # alldata <- filter(alldata, start_density >= control_min)
+
+  # low_biomass <- quantile(alldata$start_density, 0.1)
+  # alldata <- filter(alldata, start_density >= low_biomass )
+
+  # if (order == "decreasing") 
+  #alldata <-   alldata[rev(order(alldata$start_density)), ]
+  # 
+  # if (order == "increasing") 
+ # alldata <- alldata[rev(order(alldata$start_density)), ]
+
+  # need to be standardized? seems not
+  alldata$depth2 <- alldata$depth^2
+  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
+  alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
+  alldata$depth.st <- arm::rescale(log(alldata$depth)) 
+  alldata$muddy.st <- arm::rescale(alldata$muddy) 
+  alldata$any_rock.st <- arm::rescale(alldata$any_rock) 
+  #alldata$omega_s.st <- arm::rescale(alldata$omega_s)
+
+ ## Using mysterious cem optimization ... doesn't work   
+  # fit_formula <- paste0("treat ~ start_density_st") 
+  # fit_formula <- paste0("treat ~ start_density_st + depth.st") 
+  # fit_formula <- paste0("treat ~ start_density_st + depth + trawled + muddy + any_rock + omega_s + x + y") 
+  # fit_formula <- paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") #+ + omega_s
+  # 
+  # # for (j in (variable_names)) {
+  # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
+  # #   }
+  # 
+  # # # options("optmatch_max_problem_size" = Inf)
+  # # matched <- MatchIt::matchit(as.formula(fit_formula), data = alldata, method = "optimal", ratio = 1)
+  # #fit_formula <- paste0("treat ~ depth.st")
+  # matched <- MatchIt::matchit(as.formula(fit_formula),
+  #   method = "cem",
+  #   distance = "probit",
+  #   discard="control",
+  #   data = alldata)
+
+  # Using exact match and rounding
+  fit_formula <- paste0("treat ~ depth.st +  muddy.st + any_rock.st") #+ trawled + omega_s
   for (j in (variable_names)) {
@@
 
-  matched <- McSpatial::matchprop(as.formula(fit_formula), data = alldata, ytreat = "source")
+  matched <- MatchIt::matchit(as.formula(fit_formula), exact = c("start_density.st"),
+    method = "nearest", m.order = "random",
+    #method = "full",
+    distance = "probit",
+    discard="control",
+    data = alldata)
 
-  matched_time0 <- matched %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
-  matched_time1 <- matched %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)
+  mdata <- MatchIt::match.data(matched, distance = "pscore")
+
+  summary(matched)
+  plot(matched)
+
+  sources <- mdata[row.names(matched$match.matrix), ]
+  sources$origobs <- row.names(sources)   
+  sources$matchobs <- matched$match.matrix[,1]
+
+  controls <- mdata[matched$match.matrix, ]
+  controls$origobs <- row.names(controls)   
+  controls$matchobs <- matched$match.matrix[,1]
+
+  # controls2 <- mdata[matched$match.matrix, ]
+  # controls2$origobs <- row.names(controls)   
+  # controls2$matchobs <- matched$match.matrix[,2]
+
+  mdata <- rbind(sources, controls) %>% tidyr::drop_na() %>% filter(matchobs!="-1")
+
+  #mdata <- mdata %>% rename(distance = vect_dist) #%>% select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)
+
+  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
+  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)
 
@@
 ```{r}
-data <- vocc_by_sp[["adult"]]
-#View(data)
+data <- vocc_by_sp [["adult"]]
+# View(data)
 
 data$log_density <-  log(data$density)
-data$distance2 <- data$distance^2
+data$vect_dist2 <- data$vect_dist^2
 
@@
 ```{r}
+ggplot(data, 
+#ggplot(filter(data, as.numeric(matchobs) <= 550), 
+  aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
+  geom_point(alpha=0.4, size =4) +
+  scale_shape_manual(values= c(4, 1)) +
+  scale_colour_viridis_d(option="C", guide = FALSE) +
+  theme_bw()
+
 lapply(variable_names, function(j) {
@@
   geom_point(alpha=0.35) +
-  scale_colour_viridis_d(option="C") 
+  scale_colour_viridis_d(option="C") +
+  theme_bw()
 })
+
+ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
+  geom_point(alpha=0.35) +
+  scale_colour_viridis_d(option="C") +
+  facet_wrap(~time) +
+  theme_bw()
 
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)
+  facet_wrap(~time)+
+  theme_bw()
 
 ggplot(data, aes(log_density, fill = cell_type)) +
-  geom_histogram(alpha=0.5) +
+  geom_histogram(alpha=0.45) +
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)
+  facet_wrap(~time)+
+  theme_bw()
 
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)
+  facet_wrap(~time)+
+  theme_bw()
 ```
@@
 for (j in (variable_names)) {
-  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + distance + distance2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs)")
+  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs)")
 }
@@
 ```{r}
-sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,7,5,6,3,4), title = paste(species, "adult"))
+sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "adult"))
 ```
@@
 for (j in (variable_names)) {
-  gamm_formula <- paste0("log_density ~ time * cell_type + s(distance) + s(std_", j, ") + s(X,Y)")
+  gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
 }
 
-gam4 <- gamm4::gamm4(as.formula(gamm_formula),random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+
+# for (j in (variable_names)) {
+#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + te(X,Y)")
+# }
+# gam <- mgcv::gamm(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
 
@@
 
-summary(gam4$gam) ## summary of gam
 summary(gam4$mer) ## underlying mixed model
@@
    parametricOnly = TRUE, main="", xlab ="log_density")
+
+summary(gam4$gam) ## summary of gam
 ```
@@
 idata$log_density <-  log(idata$density)
-idata$distance2 <- idata$distance^2
+idata$vect_dist2 <- idata$vect_dist^2
 
@@
 ```{r}
+ggplot(filter(data, as.numeric(matchobs) <= 150), aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
+  geom_point(alpha=0.4, size =4) +
+  scale_shape_manual(values= c(4, 1)) +
+  scale_colour_viridis_d(option="C", guide = FALSE) +
+  theme_bw()
+
 lapply(variable_names, function(j) {
@@
 # for (j in (variable_names)) {
-#   gamm_formula <- paste0("log_density ~ time * cell_type + s(distance) + s(std_", j, ") + s(X,Y)")
+#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
 # }
@@
 ```{r eval=FALSE}
-# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
+# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + vect_dist + vect_dist2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
 # summary(mi)
 # 
-# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(distance) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
+# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(vect_dist) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
 # summary(mi)
@@
 # 
-# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(distance) + #s(X) + s(Y) + 
+# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(vect_dist) + #s(X) + s(Y) + 
 #     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")

### 2019-08-23 analysis/VOCC/4-vocc-BACI.Rmd
refine matchit: best seems to be combined exact matching of density and nearestneighbor for other variables

@@
 author: "Philina English"
-date: '2019-08-15'
+date: '2019-08-22'
 output: html_document
@@
 # # Species run so far...
-# species <- "Arrowtooth Flounder"
+ species <- "Arrowtooth Flounder"
 # species <- "Pacific Cod"
 # species <- "Sablefish"
-species <- "Silvergray Rockfish"
+# species <- "Silvergray Rockfish"
 # species <- "Lingcod"
@@
 end_time <- 2015
-temp_threshold <- c(0.5)
+temp_threshold <- c(1)
 do_threshold <- c(0.1)
@@
 
-Plot all possible vectors
+### Plot all possible vectors
 ```{r echo=FALSE}
@@
 
-Filter vectors to match optimal range for chosen species
+### Filter vectors to match optimal range for chosen species
 ```{r}
@@
 ```{r echo=FALSE, warning=FALSE}
-
 if (climate == "temperature") {
@@
 
-#### Get all biomass predictions 
-```{r, echo=FALSE, warning=FALSE}
+Get all biomass predictions 
+```{r, echo=FALSE, warning=FALSE, error=FALSE, message=F}
 priors <- FALSE
@@
 
-#### Extract biomass values to match spatial and temporal range of climate vectors
+Extract biomass values to match spatial and temporal range of climate vectors
+```{r}
+alldata <- lapply(biomass_list, function(i) {
 
-```{r}
-#i <- biomass_list[[1]]
-
-vocc_by_sp <- lapply(biomass_list, function(i) {
   biomass_change <- make_trend_data(i,
@@
   # identify all possible control cells, assume distance = 0 given 
-  control_cells <- anti_join(vocc, vectors, by = c("icell", "x", "y")) %>% distinct(icell, .keep_all = TRUE) %>% 
+  control_cells <- anti_join(vocc, vectors, by = c("icell", "x", "y")) %>% 
+    distinct(icell, .keep_all = TRUE) %>% 
     mutate(distance = 0) %>% 
@@
 
-#browser()
   alldata <- rbind(source_biomass, control_biomass) 
@@
   alldata$treat <- as.factor(alldata$treat)
-  ggplot(alldata, aes(x=cell_type, y=log(start_density))) + geom_boxplot()
 
-  ## still including this trimming to reduce size of dataset 
-  control_min <- min(alldata$start_density[alldata$treat=="0"])
-  source_min <- min(alldata$start_density[alldata$treat=="1"])
-  # control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.001)
-  # source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.001)
+  alldata
+
+})
+```
+
+```{r, include=FALSE, echo=FALSE}
+lapply(alldata, function(i){ggplot(i, aes(x=cell_type, y=log(start_density))) + geom_boxplot()})
+```
+
+# Select matched control cells
+```{r}
+matched <- lapply(alldata, function(alldata) {  
+  ## trim non-overlapping values to reduce size of dataset 
+  control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.001)
+  source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.001)
+  # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 1)
+  # source_max <- quantile(alldata$start_density[alldata$treat=="1"], 1)
+
   group_min <- max(control_min, source_min)
-
-  control_max <- max(alldata$start_density[alldata$treat=="0"])
-  source_max <- max(alldata$start_density[alldata$treat=="1"])
-  # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 0.95)
-  # source_max <- quantile(alldata$start_density[alldata$treat=="1"], 0.95)
-  group_max <- min(control_max, source_max)
-  # 
+  # group_max <- min(control_max, source_max)
   alldata <- filter(alldata, start_density >= group_min) #%>% filter(start_density <= group_max)
-  # 
-  # num_source <- nrow(alldata[alldata$treat=="1", ])
-  #
-
-  # alldata <- filter(alldata, start_density >= control_min)
-
-  # low_biomass <- quantile(alldata$start_density, 0.1)
-  # alldata <- filter(alldata, start_density >= low_biomass )
-
-  # if (order == "decreasing") 
-  #alldata <-   alldata[rev(order(alldata$start_density)), ]
-  # 
-  # if (order == "increasing") 
- # alldata <- alldata[rev(order(alldata$start_density)), ]
 
@@
 
- ## Using mysterious cem optimization ... doesn't work   
-  # fit_formula <- paste0("treat ~ start_density_st") 
-  # fit_formula <- paste0("treat ~ start_density_st + depth.st") 
-  # fit_formula <- paste0("treat ~ start_density_st + depth + trawled + muddy + any_rock + omega_s + x + y") 
-  # fit_formula <- paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") #+ + omega_s
-  # 
-  # # for (j in (variable_names)) {
-  # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
-  # #   }
-  # 
-  # # # options("optmatch_max_problem_size" = Inf)
-  # # matched <- MatchIt::matchit(as.formula(fit_formula), data = alldata, method = "optimal", ratio = 1)
-  # #fit_formula <- paste0("treat ~ depth.st")
-  # matched <- MatchIt::matchit(as.formula(fit_formula),
-  #   method = "cem",
-  #   distance = "probit",
-  #   discard="control",
-  #   data = alldata)
+ ## Using optimization ... doesn't work   
+  # if(
+  #   class(
+  #   try(
+  #     fit_formula <- paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") 
+  #   # fit_formula <- paste0("treat ~ start_density_st") 
+  #   # fit_formula <- paste0("treat ~ start_density_st + depth.st") 
+  #   # # for (j in (variable_names)) {
+  #   # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
+  #   # #   }
+  #     # options("optmatch_max_problem_size" = Inf)
+  #     matched <- MatchIt::matchit(as.formula(fit_formula), method = "optimal", data = alldata)
+  #     , silent = TRUE)
+  #     ) == "try-error"){
 
   # Using exact match and rounding
-  fit_formula <- paste0("treat ~ depth.st +  muddy.st + any_rock.st") #+ trawled + omega_s
-  for (j in (variable_names)) {
-    fit_formula <- paste0("", fit_formula, " + ", j, "_s")
-    }
+  fit_formula <- paste0("treat ~ depth.st + muddy.st + any_rock.st + trawled + omega_s + x + y") #
+  # for (j in (variable_names)) {
+  #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
+  #   }
+
+  # Choose data order
+  # alldata <- alldata[(order(alldata$y)), ]
+  # if (order == "increasing") alldata <- alldata[rev(order(alldata$start_density)), ]
+  # if (order == "decreasing") alldata <- alldata[rev(order(alldata$start_density)), ]
 
@@
     method = "nearest", m.order = "random",
-    #method = "full",
     distance = "probit",
@@
 
+  #}
+
+  matched
+
+})
+```
+
+Print diagnostics for matching
+```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
+lapply(matched, function(i){  
+  summary(i)
+  plot(i)
+  })
+```
+
+Prep matched data for analysis
+```{r}
+vocc_by_sp <- lapply(matched, function(matched) {    
   mdata <- MatchIt::match.data(matched, distance = "pscore")
-
-  summary(matched)
-  plot(matched)
 
@@
 
+  # if selecting 2 matches
   # controls2 <- mdata[matched$match.matrix, ]
@@
 
-  #mdata <- mdata %>% rename(distance = vect_dist) #%>% select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)
-
-  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
-  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)
+  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) 
+  #%>% dplyr::select(-start_density, -end_density)
+  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) 
+  #%>% dplyr::select(-start_density, -end_density)
 
@@
           (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE))*2),
-      (!!std.var2) := (( UQ(rlang::sym(paste0(variable_names[j]))) - mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) ) / (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)*2))^2
+      (!!std.var2) := (( UQ(rlang::sym(paste0(variable_names[j]))) - 
+          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) ) / 
+          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)*2))^2
     )
@@
   # rename spatial coordinates for inclusion in models and plots without naming conflicts
-  vocc_by_sp <- vocc_by_sp %>%
+  data <- vocc_by_sp %>%
     rename(X = x, Y = y)
+      #%>% rename(distance = vect_dist) %>% select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)
+  # other variable prep
+  data$log_density <-  log(data$density)
+  data$vect_dist2 <- data$vect_dist^2
+  data <- mutate(data, time = ifelse(after==1, "end", "begin"))
+  data$matchobs <- as.factor(data$matchobs)
+  data$origobs  <- as.factor(data$origobs)
 
-  vocc_by_sp
+  data
+
 })
@@
 ## Adults
-
-Prep data
 ```{r}
@@
 # View(data)
-
-data$log_density <-  log(data$density)
-data$vect_dist2 <- data$vect_dist^2
-
-data <- mutate(data, time = ifelse(after==1, "end", "begin"))
-data$matchobs <- as.factor(data$matchobs)
-data$origobs  <- as.factor(data$origobs)   
 ```
 
-Plot data
+### Plot matches
 ```{r}
@@
 #ggplot(filter(data, as.numeric(matchobs) <= 550), 
-  aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
-  geom_point(alpha=0.4, size =4) +
-  scale_shape_manual(values= c(4, 1)) +
+  aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
+  geom_point(alpha=0.4, size = 2) +
+  scale_shape_manual(values= c(0, 15)) +
   scale_colour_viridis_d(option="C", guide = FALSE) +
@@
 
-Basic BACI
+### Basic BACI
 ```{r}
@@
 
-GAMM
+### GAMM
 ```{r}
-
 for (j in (variable_names)) {
@@
 
-# for (j in (variable_names)) {
-#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + te(X,Y)")
-# }
-# gam <- mgcv::gamm(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+summary(gam4$gam) ## summary of gam
 
 plot(gam4$gam,pages=1)
+```
 
-summary(gam4$mer) ## underlying mixed model
-anova(gam4$gam)
-
-# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
-#   parametricOnly = TRUE, main="") + scale_y_reverse()
-
-plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
-   parametricOnly = TRUE, main="", xlab ="log_density")
-
-summary(gam4$gam) ## summary of gam
+```{r,  results="hide", include=FALSE}
+plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), parametricOnly = TRUE, main="", xlab ="log_density"
+  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
+  ) 
 ```
@@
 ## Immatures
-
-Prep data
 ```{r}
@@
 #View(data)
-idata$log_density <-  log(idata$density)
-idata$vect_dist2 <- idata$vect_dist^2
-
-idata <- mutate(idata, time = ifelse(after==1, "end", "begin"))
-idata$matchobs <- as.factor(idata$matchobs)
-idata$origobs  <- as.factor(idata$origobs)   
 ```
 
-Plot data
+### Plot matches
 ```{r}
-ggplot(filter(data, as.numeric(matchobs) <= 150), aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
-  geom_point(alpha=0.4, size =4) +
-  scale_shape_manual(values= c(4, 1)) +
+ggplot(data, 
+#ggplot(filter(data, as.numeric(matchobs) <= 550), 
+  aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
+  geom_point(alpha=0.4, size = 2) +
+  scale_shape_manual(values= c(0, 15)) +
   scale_colour_viridis_d(option="C", guide = FALSE) +
@@
 ```{r}
-sjPlot::plot_model(mi, type = "std2", title = paste(species, "immature"))
+sjPlot::plot_model(mi, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "immature"))
 ```
 
-GAMM
+### GAMM
 ```{r}
@@
 
+summary(igam4$gam) ## summary of gam
+
 plot(igam4$gam,pages=1)
+```
 
-summary(igam4$gam) ## summary of gam
-summary(igam4$mer) ## underlying mixed model
-anova(igam4$gam)
-
-# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
-#   parametricOnly = TRUE, main="") + scale_y_reverse()
-
-#my_cols <- c("#E69F00", "#56B4E9")
-plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
-   parametricOnly = TRUE, main="", xlab ="log_density"
+```{r, results="hide", include=FALSE}
+plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), parametricOnly = TRUE, main="", xlab ="log_density"
   #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
@@
 ```{r eval=FALSE}
+# summary(igam4$mer) ## underlying mixed model
+# anova(igam4$gam)
+
+# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), parametricOnly = TRUE, main="")
+
 # mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + vect_dist + vect_dist2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)

### 2019-08-24 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
loop vocc functions across multiple time periods... need to check details of exact matching for multiple variables (may need to split by year before matching)

--- 2019-08-23 analysis/VOCC/4-vocc-BACI.Rmd
@@
 # # Species run so far...
- species <- "Arrowtooth Flounder"
-# species <- "Pacific Cod"
+# species <- "Arrowtooth Flounder"
+ species <- "Pacific Cod"
 # species <- "Sablefish"
@@
 ```{r echo=TRUE}
-trend <- FALSE
-start_time <- 2012
-end_time <- 2015
+#trend <- FALSE
 temp_threshold <- c(1)
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  trend_indices_do <- c(1, 1, 1, 2, 2)
-  trend_indices_temp <- c(1, 1, 1, 2, 2)
+  start_years <- list("2009"=2009, "2011"=2011, "2013"=2013, "2015"=2015) 
+  # maybe remove 2011 & 2015 once we can add "2017"=2017?
   years <- NULL
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  trend_indices_do <- c(1, 1, 1, 2, 2)
-  trend_indices_temp <- c(1, 1, 1, 2, 2, 2)
+  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) # prob need to remove 2010 & 2014
   years <- NULL
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  trend_indices_do <- c(1, 1, 1, 2, 2)
-  trend_indices_temp <- c(1, 1, 1, 2, 2)
+  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) # prob need to remove 2010 & 2014
   years <- NULL
@@
 input_cell_size <- 2
+dist_intercept <- input_cell_size/2
 scale_fac <- 1
@@
   optimal_values <- optimal_temp
-
-  if (trend) {
-    max_thresholds <- temp_trend_threshold
-    start_time <- 2008 # change if can add back in prior 2008 data
-    end_time <- NULL
-    delta_t_total <- 6 # change if can add back in prior 2008 data
-    indices <- trend_indices_temp
-  } else {
   max_thresholds <- temp_threshold
@@
   delta_t_total <- 2
-  }
   }
@@
   optimal_values <- optimal_do
-
-  if (trend) {
-    min_thresholds <- do_trend_threshold
-    start_time <- 2008
-    skip_time <- 2016
-    end_time <- NULL
-    delta_t_total <- 6
-    indices <- trend_indices_do
-  } else {
   min_thresholds <- do_threshold
@@
   }
-}
 
@@
   variable_names <- c("do_est", "temp")
-
-  if (trend) {
-    min_thresholds <- c(do_trend_threshold, Inf)
-    max_thresholds <- c(Inf, temp_trend_threshold)
-    start_time <- 2008
-    skip_time <- 2016
-    end_time <- NULL
-    delta_t_total <- 6
-    indices <- trend_indices_do
-  } else {
   min_thresholds <- c(do_threshold, Inf)
@@
 rm(vocc)
+#start_time <- NULL
+vocc_by_year <- lapply(start_years, function(start_time) {
+
 try({
@@
   }
-
-  starttime <- Sys.time()
+  end_time <- start_time + 2
   vocc <- make_vector_data(climate_predictions,
@@
   )
-  endtime <- Sys.time()
-  time_vocc <- round(starttime - endtime)
 
@@
 }
+ vocc
+})  
 # glimpse(vocc)
@@
 ```{r echo=FALSE}
+all_vector_plots <- lapply(vocc_by_year, function(vocc) {    
+start_time <- vocc$start_time[1]
+
 if (climate == "temperature") {
@@
 vocc_plot
+})
+print(all_vector_plots)
 ```
@@
 ```{r}
+
+trimmed_vectors <- lapply(vocc_by_year, function(vocc) {    
+#start_time <- vocc$start_time[1]
+
 vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds,
-  cell_size = input_cell_size, dist_intercept = input_cell_size/2,
+  cell_size = input_cell_size, dist_intercept = dist_intercept,
   max_dist = 120
 )
+})
+
+all_vectors <- do.call("rbind", trimmed_vectors)
 
@@
 if (climate == "temperature") {
+all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
+raster_min <- min(all_vectors$diff)
+raster_max <- max(all_vectors$diff)
+}
+
+if (climate == "do") {
+all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
+raster_min <- min(all_vectors$diff)
+raster_max <- max(all_vectors$diff)
+}
+
+if (climate == "do and temperature") {
+  all_vectors$temp_diff <- all_vectors$temp.units_per_decade / 10 * delta_t_total
+  all_vectors$do_diff <- all_vectors$do_est.units_per_decade / 10 * delta_t_total
+  raster_min_temp <- min(all_vectors$temp_diff)
+  raster_max_temp <- max(all_vectors$temp_diff)
+  raster_min_do <- min(all_vectors$do_diff)
+  raster_max_do <- max(all_vectors$do_diff)
+}
+
+trimmed_vector_plots <- lapply(trimmed_vectors, function(vectors) {
+
+try({
+
+  start_time <- vectors$start_time[1]
+
+  if (climate == "temperature") {
   vectors$diff <- vectors$units_per_decade / 10 * delta_t_total
@@
     raster_alpha = 1,
+    raster_limits = c(raster_min, raster_max),
     white_zero = TRUE,
@@
     raster_alpha = 1,
+    raster_limits = c(raster_min_temp, raster_max_temp),
     white_zero = TRUE,
@@
     raster_alpha = 1,
+    raster_limits = c(raster_min_do, raster_max_do),
     white_zero = TRUE,
@@
 }
+
 vocc_plot
+})
+
+})
+
+print(trimmed_vector_plots)
 ```
@@
 ```{r}
+
 alldata <- lapply(biomass_list, function(i) {
 
-  biomass_change <- make_trend_data(i,
+  biomass_by_year <- list()
+
+  for (j in seq_along(start_years)) {
+   start_time <- start_years[[j]]
+   end_time <- start_time + 2
+
+   biomass_by_year[[j]] <- make_trend_data(i,
     ssid = model_ssid,
@@
   )
+  }
+
+  biomass_change <- do.call("rbind", biomass_by_year)
 
@@
     rename(start_density = est_exp_s, end_density = est_exp_e) %>%
-    dplyr::select(icell, x, y, start_density, end_density)
+    dplyr::select(icell, x, y, start_time, start_density, end_density)
     # dplyr::select(-N, -time_step, - units_per_decade, -slope)
@@
   # keep just one row for each cell that exceeded the end climate threshold 
-  source_cells <- vectors %>% distinct(icell, .keep_all = TRUE) %>% 
+  source_cells <- all_vectors %>% filter (distance > (- dist_intercept)) %>% 
+    distinct(icell, start_time, .keep_all = TRUE) %>% 
     select(-target_X, -target_Y, -target_values, -tid)
 
-  source_biomass <- left_join(source_cells, biomass, by = c("icell", "x", "y")) %>%
+  source_biomass <- left_join(source_cells, biomass, by = c("icell", "x", "y", "start_time")) %>%
     mutate(cell_type = "source") %>% filter(!is.na(start_density))
@@
   # identify all possible control cells, assume distance = 0 given
-  control_cells <- anti_join(vocc, vectors, by = c("icell", "x", "y")) %>% 
-    distinct(icell, .keep_all = TRUE) %>% 
-    mutate(distance = 0) %>% 
+
+  control_cells <- all_vectors %>% filter (distance == (- dist_intercept)) %>% 
+    distinct(icell, start_time, .keep_all = TRUE) %>% 
+    mutate(distance = (- dist_intercept)) %>% 
     select(-target_X, -target_Y, -target_values, -tid)
 
-  control_biomass <- left_join(control_cells, biomass, by = c("icell", "x", "y")) %>%
+  control_biomass <- left_join(control_cells, biomass, by = c("icell", "x", "y", "start_time")) %>%
     mutate(cell_type = "control") %>% filter(!is.na(start_density))
@@
   alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
-    select(-var_2_min, -var_2_max) %>% 
+    select(-var_2_min, -var_2_max, -mean_target_X, -mean_target_Y) %>% 
     rename(vect_dist = distance)
@@
   ## trim non-overlapping values to reduce size of dataset 
-  control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.001)
-  source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.001)
+  control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.01)
+  source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.01)
   # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 1)
@@
 
-  matched <- MatchIt::matchit(as.formula(fit_formula), exact = c("start_density.st"),
+  matched <- MatchIt::matchit(as.formula(fit_formula), exact = c("start_density.st", "start_time"),
     method = "nearest", m.order = "random",
@@
 
-  matched
+  #matched
 
@@
   data$origobs  <- as.factor(data$origobs)   
+  data$start_time <- as.factor(data$start_time)
 
@@
   aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
-  geom_point(alpha=0.4, size = 2) +
+  geom_point(alpha=0.8, size = 1.5) +
   scale_shape_manual(values= c(0, 15)) +
   scale_colour_viridis_d(option="C", guide = FALSE) +
+  facet_wrap(~start_time) +
   theme_bw()
@@
 
+dat_summary <- data %>% 
+  group_by(start_time, cell_type) %>% 
+  tally() %>% filter(cell_type=="source")
+
 ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)+
+  geom_text(data = dat_summary,
+            aes(cell_type, max(data$log_density), label = n), hjust = -0.3) +
+  facet_grid(time~start_time)+
   theme_bw()
@@
 ```{r}
+#FIXME: need to nest random effects
+
 for (j in (variable_names)) {
-  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs)")
+  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs) + (1|start_time)")
 }
@@
 
-gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")
+#FIXME: need to nest random effects
+gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs) + (1|start_time), data=data) # + s(X,Y, bs="re")
 
@@
 ```{r}
-ggplot(data, 
+ggplot(idata, 
 #ggplot(filter(data, as.numeric(matchobs) <= 550), 
@@
   scale_colour_viridis_d(option="C", guide = FALSE) +
+  facet_wrap(~start_time) +
   theme_bw()
@@
 
+idat_summary <- idata %>% 
+  group_by(start_time, cell_type) %>% 
+  tally() %>% filter(cell_type=="source")
+
 ggplot(idata, aes(y=log_density, x=cell_type, fill=cell_type)) +
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)
+  geom_text(data = idat_summary,
+            aes(cell_type, max(data$log_density), label = n), hjust = -0.3) +
+  facet_grid(time~start_time)+
+  theme_bw()
 ```
@@
 ```{r}
+
 mi <- lmerTest::lmer(as.formula(baci_formula), data=idata)
@@
 
-igam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=idata) # + s(X,Y, bs="re")
+#FIXME: need to nest random effects
+
+igam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs) + (1|start_time), data=idata) # + s(X,Y, bs="re")
 

### 2019-08-25 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
working multi-year BACI

@@
 ---
-title: "BACI of biomass change for VOCC source cells"
+title: "BACI of biomass change for VOCC source cells for all years"
 author: "Philina English"
-date: '2019-08-22'
+date: '2019-08-24'
 output: html_document
@@
  climate <- "temperature"
+ temp_threshold <- c(0.5)
+
 # climate <- "do"
+# do_threshold <- c(0.1)
+
 # climate <- "do and temperature"
-```
-
-#### Make climate vectors for biannual changes between surveys
-```{r echo=TRUE}
-#trend <- FALSE
-temp_threshold <- c(1)
-do_threshold <- c(0.1)
-```
-
-# OR
-
-#### Make climate vectors for average changes and lower threshold values
-```{r echo=TRUE}
-# trend <- TRUE
-# temp_trend_threshold <- c(0.25)
-# do_trend_threshold <- c(0.2)
 ```
@@
     raster_alpha = 1,
+    raster_limits = c(-4, 3),
     white_zero = TRUE,
@@
 biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
-
 # glimpse(biomass_list)
@@
 ```{r}
-
-alldata <- lapply(biomass_list, function(i) {
+alldata <- lapply(biomass_list, function(split_bio_data) {
 
   biomass_by_year <- list()
-
   for (j in seq_along(start_years)) {
@@
 
-   biomass_by_year[[j]] <- make_trend_data(i,
+   biomass_by_year[[j]] <- make_trend_data(split_bio_data,
     ssid = model_ssid,
@@
 
-  biomass_covs <- i %>% select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>% distinct()
+  biomass_covs <- split_bio_data %>% 
+    select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>% 
+    distinct()
 
@@
 
-  # identify all possible control cells, assume distance = 0 given
+  # identify all possible control cells have been given a distance == - dist_intercept
 
@@
   alldata
-
 })
@@
 ```{r, include=FALSE, echo=FALSE}
-lapply(alldata, function(i){ggplot(i, aes(x=cell_type, y=log(start_density))) + geom_boxplot()})
+lapply(alldata, function(i){ggplot(i, aes(x=cell_type, y=log(start_density))) + geom_boxplot()+facet_wrap(~start_time)})
 ```
@@
   # Using exact match and rounding
-  fit_formula <- paste0("treat ~ depth.st + muddy.st + any_rock.st + trawled + omega_s + x + y") #
+  fit_formula <- paste0("treat ~ depth.st + muddy.st + any_rock.st + 
+    trawled + omega_s + x + y") #
   # for (j in (variable_names)) {
@@
 
-  matched <- MatchIt::matchit(as.formula(fit_formula), exact = c("start_density.st", "start_time"),
+  matched <- MatchIt::matchit(as.formula(fit_formula), 
+    exact = c("start_density.st", "start_time"),
     method = "nearest", m.order = "random",
@@
             aes(cell_type, max(data$log_density), label = n), hjust = -0.3) +
-  facet_grid(time~start_time)+
+  facet_grid(start_time~time)+
   theme_bw()
@@
 
-#FIXME: need to nest random effects
 gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs) + (1|start_time), data=data) # + s(X,Y, bs="re")

### 2019-08-27 analysis/VOCC/1-biomass-do-models-fixed.Rmd
make optimal values of temp and do based on fixed effect rather than time-varying models

---
title: "Spatiotemporal models of species biomass with time-varying DO"
author: "Philina English"
date: '2019-07-18'
output: html_document
params: 
    species: "Arrowtooth Flounder"

---

Choose one value for each parameter:

Species
```{r}
species <- params$species

# # Species run so far...
# species <- "Arrowtooth Flounder"
 species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

Region 
```{r}
 region <- "All synoptic surveys"
# region <- "Both odd year surveys"
# region <- "West Coast Vancouver Island"
# region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
 priors <- FALSE
# priors <- TRUE
# covariates <- "+muddy+any_rock"
 covariates <- ""
# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
#update_spatial_model <- TRUE
```

Combine biomass and sensor data
```{r}
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars) %>% mutate(depth_akima = depth) %>% select(-depth)

  CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
  data <- dplyr::left_join(data, CTD) %>% 
    mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
    #mutate(exclude = if_else(do_mlpl>6, 1, 0)) %>% #try excluding high values?
    filter(exclude != 1) %>% # added on july 24 after first round of models run...
    filter(year > 2007) # 2007 DO data is flawed

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth), temp = temperature_c) 
  do_data <- gfranges::scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- do_data %>% mutate(depth = raw_depth) %>% #filter(year != 2016) %>%
    filter(ssid %in% model_ssid) %>% filter(!is.na(do_mlpl)) 
```

```{r}
  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "All synoptic surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
  }

  sdmTMB::plot_spde(spde)
```

Run DO sdmTMB model
```{r}
  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))

    starttime1 <- Sys.time()
    adult_biomass_d <- sdmTMB::sdmTMB(data,
      adult_formula,
      #time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))

    starttime2 <- Sys.time()
    imm_biomass_d <- sdmTMB::sdmTMB(data, 
      imm_formula,
      #time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

  } else {

    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + depth_scaled + depth_scaled2", covariates, ""))

    total_biomass_d <- sdmTMB::sdmTMB(data,
      dens_formula,
      #time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2 + do_mlpl_scaled3, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
```

### 2019-08-27 analysis/VOCC/1-biomass-temp-models-fixed.Rmd
make optimal values of temp and do based on fixed effect rather than time-varying models

---
title: "Spatiotemporal models of species biomass split by maturity class"
author: "Philina English"
date: '2019-07-18'
output: html_document
params: 
    species: "Arrowtooth Flounder"

---

Choose one value for each parameter:

Species
```{r}
species <- params$species
# # Species run so far...
species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds

# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

Region 
```{r}
 region <- "All synoptic surveys"
# region <- "Both odd year surveys"
# region <- "West Coast Vancouver Island"
# region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
priors <- FALSE
# priors <- TRUE
# covariates <- "+muddy+any_rock"
covariates <- ""
# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
#update_spatial_model <- TRUE
```

Combine biomass and sensor data
```{r}
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars) %>% mutate(depth_akima = depth) %>% select(-depth)

  CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
  data <- dplyr::left_join(data, CTD)

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth), temp = temperature_c) 
  do_data <- gfranges::scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- do_data %>% mutate(depth = raw_depth) %>% 
    filter(ssid %in% model_ssid) %>% filter(!is.na(temp))# %>% filter(year != 2016)
```

```{r}
  if (region == "All synoptic surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 800)
  }

  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  sdmTMB::plot_spde(spde)
```

Run temperature sdmTMB model
```{r}
  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))

    starttime1 <- Sys.time()
    adult_biomass_t <- sdmTMB::sdmTMB(data,
      adult_formula,
      #time_varying = ~ 0 + temp_scaled + temp_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass_t, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))

    starttime2 <- Sys.time()
    imm_biomass_t <- sdmTMB::sdmTMB(data, 
      imm_formula,
      #time_varying = ~ 0 + temp_scaled + temp_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass_t, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

  } else {

    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""))

    total_biomass_t <- sdmTMB::sdmTMB(data,
      dens_formula,
     # time_varying = ~ 0 + temp_scaled + temp_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass_t, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
```

### 2019-08-27 analysis/VOCC/3-fixed-var-response-plots.Rmd
make optimal values of temp and do based on fixed effect rather than time-varying models

--- 2019-08-14 analysis/VOCC/3-variable-response-plots.Rmd
@@
 # # Species run so far...
-# species <- "Arrowtooth Flounder"
+ species <- "Arrowtooth Flounder"
 # species <- "Pacific Cod"
@@
 # species <- "Silvergray Rockfish"
- species <- "Lingcod"
+# species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
@@
 ```{r}
+covariates <- ""
 # covariates <- "+muddy+mixed+rocky"
 # covariates <- "+mixed+rocky"
-covariates <- "+muddy+any_rock"
+#covariates <- "+muddy+any_rock"
 # covariates <- "+trawled+muddy+rocky+mixed"
@@
 rm(
+  adult, imm,
   adult_survey1, imm_survey1,
@@
 )
-priors <- TRUE
+priors <- FALSE
 # covariates <- ""
@@
 
-  adult_survey_all <- readRDS(paste0(
+  adult <- readRDS(paste0(
     "data/", spp,
-    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
 
-  imm_survey_all <- readRDS(paste0(
+  imm <- readRDS(paste0(
     "data/", spp,
-    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
 })
-if (!exists("adult_survey_all")) adult_survey_all <- NULL
-if (!exists("imm_survey_all")) imm_survey_all <- NULL
+if (!exists("adult")) adult <- NULL
+if (!exists("imm")) imm <- NULL
 
@@
     "data/", spp,
-    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
     "data/", spp,
-    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
 model_list <- list(
-  adult_survey_all = adult_survey_all, imm_survey_all = imm_survey_all,
+  adult = adult, imm = imm,
   adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
@@
 ```{r}
-rm(
+rm(adult, imm,
   adult_survey1, imm_survey1,
@@
 )
-priors <- TRUE
+priors <- FALSE
+covariates <- "+muddy+any_rock"
+covs <- gsub("\\+", "-", covariates)
+
+try({
+  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+
+  adult <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+
+  imm <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+if (!exists("adult")) adult <- NULL
+if (!exists("imm")) imm <- NULL
+
 try({
@@
     "data/", spp,
-    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
     "data/", spp,
-    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
 temp_model_list <- list(
+  adult = adult, imm = imm, 
   adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
@@
 for (i in seq_len(length(temp_model_list))) {
-  td[[i]] <- time_varying_density(temp_model_list[[i]], predictor = "temp")
+  td[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
   #browser()
   if (length(td[[i]])>0) td[[i]]$model <- names(temp_model_list[i])
+  if (length(td[[i]])>0) td[[i]]$optimal_temp <- get_optimal_value(td[[i]], xlimits = c(0, 12))
 }
 
-tdat <- do.call("rbind", td)
-
-tdat <- tdat %>%
+tdat <- do.call("rbind", td) %>%
   group_by(model) %>%
-  mutate(maxy = max(y_hat)) %>%
-  ungroup() %>%
-  mutate(meanymax = mean(maxy), sdymax = sd(maxy)) %>%
-  filter(maxy < (meanymax + 2 * sdymax))
-
-hist(tdat$maxy)
-
-optimal_temp <- get_optimal_value(tdat, xlimits = c(3, 8))
+  mutate(age = ifelse(grepl("imm", model), "imm", "adult"),
+    t_maxy = max(y_hat)) %>% 
+  select(model, age, optimal_temp, t_maxy) %>%
+  distinct()
 
@@
 for (i in seq_len(length(model_list))) {
- do[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
+ do[[i]] <- fixed_density(model_list[[i]], predictor = "do_mlpl")
  if (length(do[[i]])>0) do[[i]]$model <- names(model_list[i])
+ if (length(do[[i]])>0) do[[i]]$optimal_do <- get_optimal_value(do[[i]], xlimits = c(0.5, 6))
 }
-odat <- do.call("rbind", do)
 
-odat <- odat %>%
+odat <- do.call("rbind", do) %>%
   group_by(model) %>%
-  mutate(maxy = max(y_hat)) %>%
-  ungroup() %>%
-  mutate(meanymax = mean(maxy), sdymax = sd(maxy)) %>%
-  filter(maxy < (meanymax + 2 * sdymax))
+  mutate(age = ifelse(grepl("imm", model), "imm", "adult"),
+    do_maxy = max(y_hat)) %>% 
+  select(model, age, optimal_do, do_maxy) %>%
+  distinct()
 
-hist(odat$maxy)
+params <- full_join(tdat, odat, by=c("model", "age"))
+#params <- tdat
+params$species <- species
+#params$date <- Sys.Date()
 
-optimal_do <- get_optimal_value(odat, xlimits = c(0.5, 6))
-
-date <- Sys.time()
-
-params <- data.frame(species, optimal_do, optimal_temp, date)
-
-other_params <- readRDS(paste0("data/optimal-values-by-species.rds"))
+other_params <- readRDS(paste0("data/optimal-values-by-species-by-age.rds"))
 params <- rbind(other_params, params)
-saveRDS(params, file = "data/optimal-values-by-species.rds")
+saveRDS(params, file = "data/optimal-values-by-species-by-age.rds")
 
@@
 for (i in seq_len(length(model_list))) {
-  d[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
+  d[[i]] <- fixed_density(model_list[[i]], predictor = "do_mlpl")
   if (length(d[[i]]) == 0) {
@@
 for (i in seq_len(length(temp_model_list))) {
-  d[[i]] <- time_varying_density(temp_model_list[[i]], predictor = "temp")
+  d[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
 
@@
     d[[i]]$x <- exp(d[[i]]$x)
-    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, 1000)) +
+    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, 800)) +
       ggtitle(paste(species, names(depth_model_list[i])))

### 2019-08-27 R/plot-mountains.R

@@ -127,87 +127,6 @@ time_varying_density <- function(m, predictor = "depth") {
   pred_density
 }
 
-time_varying_density3 <- function(m, predictor = "depth") {
-  get_y_hat <- function(b0, b1, b2, b3, year, 
-    predictor, mean_column, sd_column
-  ) {
-    
-    x_pred <- seq(min(m$data[[predictor]], na.rm = TRUE), max(m$data[[predictor]], na.rm = TRUE), length.out = 300)
-    
-    data.frame(
-      x = (x_pred * m$data[[sd_column]][[1]] + m$data[[mean_column]][[1]]), # if depth, actually is log_depth so must exp(x) for raw depth
-      y_hat = exp(b0 + b1 * x_pred + b2 * x_pred^2 + b3 * x_pred^3), # leave these values as predicted for un-trawled zone
-      year = year
-    )
-    
-  }
-  r <- m$tmb_obj$report()
-  r$b_rw_t
-  b_j <- m$model$par
-  n_t <- nrow(r$b_rw_t)
-  yrs <- sort(unique(m$data$year))
-  ssid <- m$data$ssid
-  
-  pred_density <- purrr::map_df(seq_len(n_t), function(.t) {
-    get_y_hat(
-      b0 = b_j[.t],
-      b1 = r$b_rw_t[.t, 1],
-      b2 = r$b_rw_t[.t, 2],
-      b3 = r$b_rw_t[.t, 3],
-      year = yrs[.t],
-      sd_column = paste0(predictor, "_sd"),
-      mean_column = paste0(predictor, "_mean"),
-      predictor = paste0(predictor, "_scaled")
-    )
-  })
-  pred_density
-}
-
-
-#' Calculate density curve for scaled, quadratic parameters with non-AR1 year interaction
-#'
-#' @param m Output of sdmTMB model
-#' @param predictor Prefix for scaled parameter in model
-#'
-#' @export
-#'
-fixed_interaction_density <- function(m, predictor = "temp") {
-  get_y_hat <- function(b0, b1, b2, year, 
-    predictor, mean_column, sd_column
-  ) {
-    x_pred <- seq(min(m$data[[predictor]], na.rm = TRUE), max(m$data[[predictor]], na.rm = TRUE), length.out = 300)
[…]

### 2019-08-27 analysis/VOCC/1-biomass-models.Rmd
add all surveys to biomass-model options

@@
 ```{r}
- region <- "Both odd year surveys"
-region <- "West Coast Vancouver Island"
+ region <- "All synoptic surveys"
+# region <- "Both odd year surveys"
+# region <- "West Coast Vancouver Island"
 # region <- "West Coast Haida Gwaii"
@@
 
-#update_spatial_model <- TRUE
+if (region == "All synoptic surveys") {
+  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  years <- NULL
+}
 ```

### 2019-08-27 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
split BACI code to run separately by maturity class

@@
 ---
+
+Should optimal temperature vary for age classes? Seems likely. 
+Therefore need separate VOCC trimming and matching for age classes -> change function structure to choose age upfront
+
+Should matches be selected based on exact depth categories in addition to density (all within time periods)?
+What does it mean for the analysis that the depth range at which VOCC occur within a species temperature range varies dramatically between time periods?
 
@@
 # # Species run so far...
-# species <- "Arrowtooth Flounder"
- species <- "Pacific Cod"
+ species <- "Arrowtooth Flounder"
+# species <- "Pacific Cod"
 # species <- "Sablefish"
@@
 
-#### Get species-specific optimal values from mountain plots for most abundant year
-```{r, echo=FALSE}
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-params <- readRDS(paste0("data/optimal-values-by-species.rds"))
-params <- params[rev(order(as.Date(params$date))), ]
-params <- params[!duplicated(params$species), ]
-
-optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
-optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)
-
-optimal_values <- c(optimal_do, optimal_temp)
+#### Set age class
+```{r echo=TRUE}
+age <- "mature" # or mostly so
+# age <- "imm"
 ```
@@
 library(sdmTMB)
-#library(itsadug)
+# library(itsadug)
 library(gfranges)
+```
+
+#### Get species-specific optimal values from mountain plots for most abundant year
+```{r, echo=FALSE}
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+
+params <- readRDS(paste0("data/optimal-values-by-species-by-age.rds"))
+
+if (age == "imm") {
+  params <- filter(params, age == "imm")
+} else {
+  params <- filter(params, age != "imm")
+}
+params <- params[rev(order(params$maxy)), ]
+params <- params[!duplicated(params$species), ]
+
+#optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
+optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)
+
+#optimal_values <- c(optimal_do, optimal_temp)
 ```
@@
   min_thresholds <- c(Inf)
-  #delta_threshold <- temp_threshold
   optimal_values <- optimal_temp
@@
   max_thresholds <- c(Inf)
-  #delta_threshold <- -1 * (do_threshold)
   optimal_values <- optimal_do
@@
 rm(vocc)
-#start_time <- NULL
+
 vocc_by_year <- lapply(start_years, function(start_time) {
@@
 all_vectors <- do.call("rbind", trimmed_vectors)
-
-# if (length(vectors[, 2]) == 0) {
-#   other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
-#   coefs <- other_coefs[1, ]
-#   coefs$var <- "NA"
-#   coefs$beta <- "NA"
-#   coefs$se <- "NA"
-#   coefs$model <- "NA"
-#   coefs$species <- spp
-#   coefs$lifestage <- "NA"
-#   coefs$climate <- climate
-#   coefs$min_thresholds <- min_string
-#   coefs$max_thresholds <- max_string
-#   coefs$optimum <- optimal_string
-#   coefs$start_time <- start_time
-#   coefs$timespan <- delta_t_total
-#   coefs$region <- region
-#   coefs$ssid_string <- ssid_string
-# 
-#   coefs <- rbind(other_coefs, coefs) %>% distinct()
-#   saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
-# 
-#   stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
-# }
-# glimpse(vocc)
-# glimpse(vectors)
 ```
@@
     fill_col = "diff",
-    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+    fill_label = paste("Change from\n", 
+      start_time, "to", start_time + delta_t_total, ""),
     raster_alpha = 1,
@@
     fill_col = "temp_diff",
-    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+    fill_label = paste("Change from\n", 
+      start_time, "to", start_time + delta_t_total, ""),
     raster_alpha = 1,
@@
     fill_col = "do_diff",
-    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
+    fill_label = paste("Change from\n", 
+      start_time, "to", start_time + delta_t_total, ""),
     raster_alpha = 1,
@@
     ncol = 2,
-    top = grid::textGrob(paste("DO of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
+    top = grid::textGrob(paste(
+      "DO of ", optimal_values[1], "less", min_thresholds[1], 
+      "or temperature of ", optimal_values[2], "plus", max_thresholds[2], 
+      " over ", start_time, "-", start_time + delta_t_total, ""))
   )
 }
-
 vocc_plot
 })
-
 })
@@
 priors <- FALSE
-# priors <- TRUE
-
 covariates <- "+muddy+any_rock"
-
 covs <- gsub("\\+", "-", covariates)
 
-rm(adult_predictions)
-rm(imm_predictions)
-rm(predictions)
+rm(biomass_predictions)
 
+if (age == "imm") {
   try({
-  adult_predictions <- readRDS(paste0(
+  biomass_predictions <- readRDS(paste0(
+    "data/", spp,
+    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
+  ))
+})
+} else {
+try({
+  biomass_predictions <- readRDS(paste0(
     "data/", spp,
@@
 })
-
+if (!exists("biomass_predictions")) {
 try({
-  imm_predictions <- readRDS(paste0(
-    "data/", spp,
-    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
-  ))
-})
-
-try({
-  predictions <- readRDS(paste0(
+  biomass_predictions <- readRDS(paste0(
     "data/", spp,
@@
 })
-
-if (!exists("adult_predictions")) adult_predictions <- NULL
-if (!exists("imm_predictions")) imm_predictions <- NULL
-if (!exists("predictions")) predictions <- NULL
-
-biomass_list <- list(
-  adult = adult_predictions,
-  imm = imm_predictions,
-  total = predictions
-)
-
-biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
-# glimpse(biomass_list)
+}
+}
 ```
@@
 ```{r}
-alldata <- lapply(biomass_list, function(split_bio_data) {
+#alldata <- lapply(biomass_list, function(biomass_predictions) {
 
@@
 
-   biomass_by_year[[j]] <- make_trend_data(split_bio_data,
+   biomass_by_year[[j]] <- make_trend_data(biomass_predictions,
     ssid = model_ssid,
@@
 
-  biomass_covs <- split_bio_data %>% 
+  biomass_covs <- biomass_predictions %>% 
     select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>% 
@@
   # keep just one row for each cell that exceeded the end climate threshold 
-  source_cells <- all_vectors %>% filter (distance > (- dist_intercept)) %>% 
+  source_cells <- all_vectors %>% filter(distance > (-dist_intercept)) %>% 
     distinct(icell, start_time, .keep_all = TRUE) %>% 
@@
 
-  source_biomass <- left_join(source_cells, biomass, by = c("icell", "x", "y", "start_time")) %>%
+  source_biomass <- left_join(source_cells, biomass, 
+    by = c("icell", "x", "y", "start_time")) %>%
     mutate(cell_type = "source") %>% filter(!is.na(start_density))
@@
   # identify all possible control cells have been given a distance == - dist_intercept
-
-  control_cells <- all_vectors %>% filter (distance == (- dist_intercept)) %>% 
+  control_cells <- all_vectors %>% filter(distance == (-dist_intercept)) %>% 
     distinct(icell, start_time, .keep_all = TRUE) %>% 
@@
 
-  control_biomass <- left_join(control_cells, biomass, by = c("icell", "x", "y", "start_time")) %>%
+  control_biomass <- left_join(control_cells, biomass, 
+    by = c("icell", "x", "y", "start_time")) %>%
     mutate(cell_type = "control") %>% filter(!is.na(start_density))
@@
   alldata
-})
+#})
 ```
 
-```{r, include=FALSE, echo=FALSE}
-lapply(alldata, function(i){ggplot(i, aes(x=cell_type, y=log(start_density))) + geom_boxplot()+facet_wrap(~start_time)})
+```{r eval=FALSE, include=FALSE, echo=FALSE}
+# lapply(alldata, function(i){ggplot(i, aes(x=cell_type, y=log(start_density))) + geom_boxplot()+facet_wrap(~start_time)})
 ```
@@
 ```{r}
-matched <- lapply(alldata, function(alldata) {  
+rm(vocc_by_sp)
+try({
+  vocc_by_sp <- readRDS(paste0(
+    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
+    min_string, "to", max_string, "-multiyear-all.rds"
+  ))
+})
+
+match_formula <- paste0("treat ~ 
+  muddy.st + any_rock.st + x + y") #depth.st + trawled + omega_s + 
+  # for (j in (variable_names)) {
+  #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
+  #   }
+
+if (!exists("vocc_by_sp")){
+# matched <- lapply(alldata, function(alldata) {
+
   ## trim non-overlapping values to reduce size of dataset 
@@
   # group_max <- min(control_max, source_max)
-  alldata <- filter(alldata, start_density >= group_min) #%>% filter(start_density <= group_max)
+  alldata <- filter(alldata, start_density >= group_min) 
+  #%>% filter(start_density <= group_max)
 
@@
   alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
-  alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
-  alldata$depth.st <- arm::rescale(log(alldata$depth)) 
+  alldata$depth.st <- round(arm::rescale(log(alldata$depth))*5)/5
   alldata$muddy.st <- arm::rescale(alldata$muddy) 
@@
  ## Using optimization ... doesn't work   
-  # if(
-  #   class(
-  #   try(
-  #     fit_formula <- paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") 
-  #   # fit_formula <- paste0("treat ~ start_density_st") 
-  #   # fit_formula <- paste0("treat ~ start_density_st + depth.st") 
-  #   # # for (j in (variable_names)) {
-  #   # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
-  #   # #   }
+  # alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
+  # fit_formula <- 
+  #   paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") 
+  # # for (j in (variable_names)) {
+  # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
+  # #   }
   # # options("optmatch_max_problem_size" = Inf)
-  #     matched <- MatchIt::matchit(as.formula(fit_formula), method = "optimal", data = alldata)
-  #     , silent = TRUE)
-  #     ) == "try-error"){
+  # matched <- MatchIt::matchit(as.formula(fit_formula), 
+  #     method = "optimal", data = alldata)
 
-  # Using exact match and rounding
-  fit_formula <- paste0("treat ~ depth.st + muddy.st + any_rock.st + 
-    trawled + omega_s + x + y") #
-  # for (j in (variable_names)) {
-  #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
-  #   }
-
-  # Choose data order
+  # # Using exact match and rounding
+  # # Choose data order if not using random
   # alldata <- alldata[(order(alldata$y)), ]
@@
 
-  matched <- MatchIt::matchit(as.formula(fit_formula), 
-    exact = c("start_density.st", "start_time"),
+  matched <- MatchIt::matchit(as.formula(match_formula), 
+    exact = c("start_density.st", "start_time", "depth.st"),
     method = "nearest", m.order = "random",
     distance = "probit",
-    discard="control",
+    discard="both",
     data = alldata)
-
-  #}
-
-  #matched
-
-})
+  # matched
+# })
+}
 ```
@@
 ```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
-lapply(matched, function(i){  
+if (!exists("vocc_by_sp")){
+match_plots <- lapply(matched, function(i){  
   summary(i)
@@
   })
+match_plots
+}
 ```
@@
 ```{r}
-vocc_by_sp <- lapply(matched, function(matched) {    
+if (!exists("vocc_by_sp")){
+#vocc_by_sp <- lapply(matched, function(matched) {    
   mdata <- MatchIt::match.data(matched, distance = "pscore")
@@
   controls$matchobs <- matched$match.matrix[,1]
-
-  # if selecting 2 matches
-  # controls2 <- mdata[matched$match.matrix, ]
-  # controls2$origobs <- row.names(controls)   
-  # controls2$matchobs <- matched$match.matrix[,2]
 
@@
 
-  # rename spatial coordinates for inclusion in models and plots without naming conflicts
+  # rename spatial coordinates to avoid naming conflict with y~x notation
   data <- vocc_by_sp %>%
     rename(X = x, Y = y)
-      #%>% rename(distance = vect_dist) %>% select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)
+    # %>% rename(distance = vect_dist) %>% 
+    # select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)
+
   # other variable prep
   data$log_density <-  log(data$density)
+  data$log_depth <-  log(data$depth)
+  data$std_log_depth <- arm::rescale(log(data$depth))
   data$vect_dist2 <- data$vect_dist^2
@@
 
-  data
-
-})
+#  data
+#})
+vocc_by_sp <- list()
+vocc_by_sp[[1]] <- data
+vocc_by_sp[[3]] <- match_formula  
+vocc_by_sp[[4]] <- match_plots
+saveRDS(vocc_by_sp, file = paste0(
+    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
+    min_string, "to", max_string, "-multiyear-all.rds"
+  ))
+}
 ```
@@
 ```{r}
+try({
+  vocc_by_sp <- readRDS(paste0(
+    "data/ BACI-", species, "-", ssid_string, "-", climate_string, "-",
+    min_string, "to", max_string, "-multiyear-all.rds"
+  ))
+})
 data <- vocc_by_sp [["adult"]]
@@
 ```{r}
+dat_summary <- data %>% 
+  group_by(start_time, cell_type) %>%
+  tally() %>% filter(cell_type=="source")
+data_depth <- data %>% 
+  group_by(start_time) %>% 
+  mutate(mean_depth = mean(depth))
+
 ggplot(data, 
@@
   aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
-  geom_point(alpha=0.8, size = 1.5) +
+  geom_point(alpha=0.8, size = 1) +
   scale_shape_manual(values= c(0, 15)) +
@@
   scale_colour_viridis_d(option="C") +
+  facet_grid(start_time~cell_type) +
   theme_bw()
@@
 
-ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
-  geom_point(alpha=0.35) +
-  scale_colour_viridis_d(option="C") +
-  facet_wrap(~time) +
+ggplot(data, aes(std_log_depth, fill = cell_type)) +
+  geom_density(alpha=0.5) +
+  scale_fill_viridis_d(option="C") +
+  facet_wrap(~start_time) +
   theme_bw()
+#  
+# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
+#   geom_point(alpha=0.35) +
+#   scale_colour_viridis_d(option="C") +
+#     facet_grid(start_time~time) +
+#   theme_bw()
 
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)+
+  facet_grid(start_time~time, scales = 'fixed') +
   theme_bw()
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)+
+  facet_grid(time~start_time) +
   theme_bw()
@@
   geom_text(data = dat_summary,
-            aes(cell_type, max(data$log_density), label = n), hjust = -0.3) +
-  facet_grid(start_time~time)+
+            aes(cell_type, max(data$log_density)-1, label = n), hjust = -0.3) +
+  facet_grid(time~start_time) +
   theme_bw()
@@
 ```{r}
-#FIXME: need to nest random effects
-
 for (j in (variable_names)) {
@@
 ```{r}
+ gamm_formula <- paste0("log_density ~ ", 
+  # "time * start_time + ",
+  #"start_time * cell_type + ",
+  "time * cell_type + s(vect_dist) + s(std_log_depth) + s(X,Y)")
+
 for (j in (variable_names)) {
-  gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
+ gamm_formula <- paste0(gamm_formula, "+ s(std_", j, ")")
  }
 
-gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs) + (1|start_time), data=data) # + s(X,Y, bs="re")
+gam4 <- gamm4::gamm4(as.formula(gamm_formula), 
+  random = ~(1|matchobs) + (1|icell) 
+   + (1|start_time)
+  # + (time|start_time) 
+  # + (cell_type|start_time)
+  , data=data)
 
-summary(gam4$gam) ## summary of gam
+summary(gam4$gam)
 
@@
 ```{r,  results="hide", include=FALSE}
-plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), parametricOnly = TRUE, main="", xlab ="log_density"
+plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source"), start_time=c("2009","2011","2013","2015")), parametricOnly = TRUE, main="", xlab ="log_density"
   #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
@@
 ```{r}
+idat_summary <- idata %>% 
+  group_by(start_time, cell_type) %>%
+  tally() %>% filter(cell_type=="source")
+idata_depth <- idata %>% 
+  group_by(start_time) %>% 
+  mutate(mean_depth = mean(depth))
+
 ggplot(idata, 
@@
   aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
-  geom_point(alpha=0.4, size = 2) +
+  geom_point(alpha=0.4, size = 1) +
   scale_shape_manual(values= c(0, 15)) +
@@
   geom_point(alpha=0.35) +
-  scale_colour_viridis_d(option="C") 
+  scale_colour_viridis_d(option="C")+
+  facet_grid(start_time~cell_type) +
+  theme_bw() 
 })
+
+# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
+#   geom_point(alpha=0.35) +
+#   scale_colour_viridis_d(option="C") +
+#     facet_grid(start_time~time) +
+#   geom_text(data = idata_depth,
+#     aes(150, 0.5, label = round(mean_depth)), hjust = -0.3, inherit.aes = FALSE) +
+#   theme_bw()
+ggplot(idata, aes(depth, fill = cell_type)) +
+  geom_density(alpha=0.5) +
+  scale_fill_viridis_d(option="C") +
+  facet_wrap(~start_time) +
+  theme_bw()
+
 ggplot(idata, aes(log_density, fill = cell_type)) +
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)
+  facet_grid(start_time~time) +
+  geom_text(data = idat_summary,
+            aes(1, 0.75, label = n), hjust = -0.3) +
+  theme_bw()
 
@@
   scale_fill_viridis_d(option="C") +
-  facet_wrap(~time)
-
-idat_summary <- idata %>% 
-  group_by(start_time, cell_type) %>% 
-  tally() %>% filter(cell_type=="source")
+  facet_grid(time~start_time) +
+  theme_bw()
 
@@
             aes(cell_type, max(data$log_density), label = n), hjust = -0.3) +
-  facet_grid(start_time~time)+
+  facet_grid(time~start_time) +
   theme_bw()
@@
 ```{r}
-
 mi <- lmerTest::lmer(as.formula(baci_formula), data=idata)
@@
 # for (j in (variable_names)) {
-#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
+#  gamm_formula <- paste0("log_density ~ ", 
+#   # "time * start_time + ",
+#   "start_time * cell_type + ",
+#   "time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
 #  }
+igam4 <- gamm4::gamm4(as.formula(gamm_formula), 
+  random = ~(1|matchobs) + (1|icell) 
+   + (1|start_time)
+  # + (time|start_time) 
+  # + (cell_type|start_time)
+  , data=idata)
 
-#FIXME: need to nest random effects
-
-igam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs) + (1|start_time), data=idata) # + s(X,Y, bs="re")
-
-summary(igam4$gam) ## summary of gam
+summary(igam4$gam)
 

### 2019-08-28 analysis/VOCC/1-biomass-models.Rmd
fix switch for amalgamating catch data across surveys when calculating length at maturity

@@
 # species <- "Arrowtooth Flounder"
-# species <- "Pacific Cod"
+#species <- "Pacific Cod"
 # species <- "Sablefish"
 # species <- "Silvergray Rockfish"
- species <- "Lingcod"
+# species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
@@
 # species <- "Pacific Ocean Perch" 
-# species <- "Yelloweye Rockfish"
+ species <- "Yelloweye Rockfish"
 ```
@@
 # priors <- TRUE
-
+# covariates <- "+muddy+any_rock"
 covariates <- ""
 covs <- gsub("\\+", "-", "+muddy+any_rock")
-covariates <- "+muddy+any_rock"
-
+#covs <- gsub("\\+", "-", covariates)
 # covariates <- "+muddy+mixed+rocky"
@@
 
-Calculate maturity weighted biomass if not done before
+# Split biomass by maturity
+
+#### Calculate maturity weighted biomass if not done before
 ```{r}
  rm(biomass)
-try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
+ rm(maturity)
+# try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
 
@@
   # fish <- fish[fish$fishing_event_id!=1506954,]
- rm(maturity)
- try( maturity <- split_catch_maturity(survey_sets, fish, bath,
-    # survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
-    # survey = c("SYN HS", "SYN WCHG"),
+
+ try({ 
+   maturity <- split_catch_maturity(survey_sets, fish, bath,
     survey = survey,
@@
     cutoff_quantile = c(.9995),
-    plot = TRUE
-  ))
+    plot = TRUE)
+   })
 
-  if(!exists("maturity")) {
+  if(is.null(maturity$maturity)) {
   maturity <- split_catch_maturity(survey_sets, fish, bath,
@@
     plot = TRUE
-  )}
+  )
+  maturity$data <- filter(maturity$data, ssid == model_ssid) 
+  }
 
@@
 
+#### Save length~weight plots
 ```{r}
@@
 
+#### Save length at maturity plot 
 ```{r}
@@
 
+# Build spatiotemporal density model
 Load and filter data 

### 2019-08-28 R/mature-biomass-est.R

@@ -72,7 +72,7 @@ years_w_maturity <- fish %>% group_by(year) %>%
 levels_per_year <- unique(years_w_maturity$maturity_levels)
 
 if (max(levels_per_year)<3) {
-  return(list(data = tidy_sets, maturity = NA, mass_model = NA))
+  return(list(data = tidy_sets, maturity = NULL, mass_model = NULL))
 } 
 
 if (min(levels_per_year)<3) {

### 2019-08-28 analysis/VOCC/1-biomass-do-models-fixed.Rmd
calculate fixed climate response curves with time-varying depth

@@
 # species <- "Arrowtooth Flounder"
- species <- "Pacific Cod"
+# species <- "Pacific Cod"
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish"
+ species <- "Silvergray Rockfish"
 # species <- "Lingcod"
@@
 ```{r}
- priors <- FALSE
-# priors <- TRUE
+# priors <- FALSE
+ priors <- TRUE
 # covariates <- "+muddy+any_rock"
@@
 # covariates <- "+trawled+mixed"
+covs <- gsub("\\+", "-", covariates)
+covs <- "-tv-depth"
 ```
@@
 ```{r global_options, include=FALSE}
-covs <- gsub("\\+", "-", covariates)
+
 spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
@@
 library(gfranges)
-
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
 
@@
 }
-#update_spatial_model <- TRUE
 ```
@@
   # scale predictors before filtering to ensure mean and SD are global
-  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth), temp = temperature_c) 
-  do_data <- gfranges::scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
+  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), temp = temperature_c) 
+  do_data <- scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
        # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
   data <- do_data %>% mutate(depth = raw_depth) %>% #filter(year != 2016) %>%
-    filter(ssid %in% model_ssid) %>% filter(!is.na(do_mlpl)) 
+    filter(ssid %in% model_ssid) %>% 
+    filter(!is.na(do_mlpl)) %>% filter(!is.na(depth_scaled)) %>% 
+    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, do_mlpl_scaled, do_mlpl_scaled2, depth_mean, depth_sd, do_mlpl_mean, do_mlpl_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, do_mlpl) %>% filter(!is.na(depth_scaled))
 ```
@@
     adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + depth_scaled + depth_scaled2", covariates, ""
+      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""
       ))
@@
       adult_formula,
-      #time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
       time = "year", spde = spde,
@@
     imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + depth_scaled + depth_scaled2", covariates, ""
+      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""
       ))
@@
       imm_formula,
-      #time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
       time = "year", spde = spde,
@@
     starttime3 <- Sys.time()
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + depth_scaled + depth_scaled2", covariates, ""))
+    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""))
 
@@
       dens_formula,
-      #time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2 + do_mlpl_scaled3, # + trawled
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
       time = "year", spde = spde,

### 2019-08-28 analysis/VOCC/1-biomass-temp-models-fixed.Rmd
calculate fixed climate response curves with time-varying depth

@@
 # # Species run so far...
-species <- "Arrowtooth Flounder"
+# species <- "Arrowtooth Flounder"
 # species <- "Pacific Cod"
@@
 ```{r}
-priors <- FALSE
-# priors <- TRUE
+# priors <- FALSE
+priors <- TRUE
+
 # covariates <- "+muddy+any_rock"
@@
 ```{r global_options, include=FALSE}
-covs <- gsub("\\+", "-", covariates)
 spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
@@
   # scale predictors before filtering to ensure mean and SD are global
-  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth), temp = temperature_c) 
-  do_data <- gfranges::scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
-       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
-  data <- do_data %>% mutate(depth = raw_depth) %>% 
-    filter(ssid %in% model_ssid) %>% filter(!is.na(temp))# %>% filter(year != 2016)
+  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), temp = temperature_c) 
+  do_data <- scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
+
+  data <- do_data %>% mutate(depth = raw_depth) %>% # filter(year != 2016) %>% 
+    filter(ssid %in% model_ssid) %>% 
+    filter(!is.na(raw_depth)) %>% filter(!is.na(temp)) %>%
+    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, temp_scaled, temp_scaled2, depth_mean, depth_sd, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, temp)
 ```
@@
     adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
+      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", 
+      covariates, ""
       ))
@@
       adult_formula,
-      #time_varying = ~ 0 + temp_scaled + temp_scaled2,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2,
       time = "year", spde = spde,
@@
     imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
+      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", 
+      #+ depth_scaled + depth_scaled2", 
+      covariates, ""
       ))
@@
       imm_formula,
-      #time_varying = ~ 0 + temp_scaled + temp_scaled2,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2,
       time = "year", spde = spde,
@@
     starttime3 <- Sys.time()
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""))
+    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", 
+      #+ depth_scaled + depth_scaled2", 
+      covariates, ""))
 
@@
       dens_formula,
-     # time_varying = ~ 0 + temp_scaled + temp_scaled2,
+      time_varying = ~ 0  + depth_scaled + depth_scaled2,
       time = "year", spde = spde,

### 2019-08-28 analysis/VOCC/3-fixed-var-response-plots.Rmd
calculate and save optimal values and max predicted biomass from fixed response to climate variables

--- 2019-08-06 analysis/VOCC/3-variable-response-plots.Rmd
@@
 ```{r}
-#species <- params$species
+# species <- params$species
 # # Species run so far...
@@
 # species <- "Silvergray Rockfish"
-# species <- "Lingcod"
+ species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-
 # species <- "Quillback Rockfish"
-# species <- "Pacific Ocean Perch"
+ species <- "Pacific Ocean Perch"
 # species <- "Yelloweye Rockfish"
@@
 # covariates <- "+trawled+mixed"
+
+# priors <- FALSE
+# covariates <- "+muddy+any_rock"
+# covs <- gsub("\\+", "-", covariates)
+priors <- TRUE
+covs <- "-tv-depth"
 ```
@@
 
-Dissolved O2 plots
+Load Dissolved O2 models
 ```{r}
-rm(
+rm(adult, imm,
   adult_survey1, imm_survey1,
   adult_survey4, imm_survey4,
-  adult_survey16, imm_survey16
-)
-priors <- TRUE
-# covariates <- ""
+  adult_survey16, imm_survey16)
 
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey_all <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  adult <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
-
-  imm_survey_all <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  imm <- readRDS(paste0("data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
 })
-if (!exists("adult_survey_all")) adult_survey_all <- NULL
-if (!exists("imm_survey_all")) imm_survey_all <- NULL
+if (!exists("adult")) adult <- NULL
+if (!exists("imm")) imm <- NULL
 
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey1 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  adult_survey1 <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
-
-  imm_survey1 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  imm_survey1 <- readRDS(paste0("data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey4 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  adult_survey4 <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
-
-  imm_survey4 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  imm_survey4 <- readRDS(paste0("data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey16 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  adult_survey16 <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
-
-  imm_survey16 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  imm_survey16 <- readRDS(paste0("data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
 })
-
 if (!exists("adult_survey16")) adult_survey16 <- NULL
@@
 model_list <- list(
-  adult_survey_all=adult_survey_all, imm_survey_all=imm_survey_all,
+  adult = adult, imm = imm,
+  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
+  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
+  adult_survey16 = adult_survey16#, imm_survey16 = imm_survey16
+)
+model_list <- model_list[!sapply(model_list, is.null)]
+```
+
+```{r}
+# test load small sdm model to avoid crashing R
+try({imm})
+# or maybe it's crashing when NaNs produced error goes into density function?
+```
+
+Load temperature models
+```{r}
+rm(adult, imm,
+  adult_survey1, imm_survey1,
+  adult_survey4, imm_survey4,
+  adult_survey16, imm_survey16)
+
+try({
+  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  adult <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  imm <- readRDS(paste0("data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+if (!exists("adult")) adult <- NULL
+if (!exists("imm")) imm <- NULL
+
+try({
+  survey <- c("SYN QCS", "SYN HS")
+  model_ssid <- c(1, 3)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  adult_survey1 <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  imm_survey1 <- readRDS(paste0("data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+if (!exists("adult_survey1")) adult_survey1 <- NULL
+if (!exists("imm_survey1")) imm_survey1 <- NULL
+
+try({
+  survey <- c("SYN WCVI")
+  model_ssid <- c(4)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  adult_survey4 <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  imm_survey4 <- readRDS(paste0("data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+if (!exists("adult_survey4")) adult_survey4 <- NULL
+if (!exists("imm_survey4")) imm_survey4 <- NULL
+
+try({
+  survey <- c("SYN WCHG")
+  model_ssid <- c(16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  adult_survey16 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  imm_survey16 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+if (!exists("adult_survey16")) adult_survey16 <- NULL
+if (!exists("imm_survey16")) imm_survey16 <- NULL
+
+temp_model_list <- list(
+  adult = adult, imm = imm, 
   adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
@@
 )
-
-model_list <- model_list[!sapply(model_list, is.null)]
+temp_model_list <- temp_model_list[!sapply(temp_model_list, is.null)]
 ```
@@
 ```{r}
-d <- list()
+do <- list()
 do_plots <- list()
+
 for (i in seq_len(length(model_list))) {
-  d[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
-  do_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Dissolved O2", xlimits = c(0,8)) + ggtitle(paste(species, names(model_list[i])))
+ do[[i]] <- fixed_density(model_list[[i]], predictor = "do_mlpl")
+ if (length(do[[i]])>0) {
+   do[[i]]$species <- species
+   do[[i]]$model <- names(model_list[i])
+   do[[i]]$covs <- covs 
+   do[[i]]$optimal_do <- get_optimal_value(do[[i]], xlimits = c(0.5, 6))
+   do_plots[[i]] <- 
+     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0, 8)) +
+        ggtitle(paste(species, names(model_list[i])))
+   } else {
+    do_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  }
+}
+
+odat <- do.call("rbind", do) 
+if (length(odat)>0) {
+  odat <- odat %>% group_by(model) %>%
+  mutate(age = ifelse(grepl("imm", model), "imm", "adult"),
+    do_maxy = max(y_hat)) %>% 
+  select(species, age, model, optimal_do, do_maxy, covs) %>% 
+  distinct() 
+} else {
+  odat <- NULL
 }
@@
 ```{r}
+td <- list()
+temp_plots <- list()
+for (i in seq_len(length(temp_model_list))) {
+  td[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
+  if (length(td[[i]])>0) {
+    td[[i]]$species <- species
+    td[[i]]$model <- names(temp_model_list[i])
+    td[[i]]$covs <- covs
+    td[[i]]$optimal_temp <- get_optimal_value(td[[i]], xlimits = c(0, 12))
+    temp_plots[[i]] <- plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 10)) +
+      ggtitle(paste(species, names(temp_model_list[i])))
+  } else {
+    temp_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  }
+}
+
+tdat <- do.call("rbind", td) 
+if (length(tdat)>0) {
+  tdat <- tdat %>% group_by(model) %>%
+  mutate(age = ifelse(grepl("imm", model), "imm", "adult"),
+    t_maxy = max(y_hat)) %>% 
+  select(species, age, model, optimal_temp, t_maxy, covs) %>%
+  distinct() 
+} else {
+  tdat <- NULL
+}
+print(temp_plots)
+```
+
+Save optimal values
+```{r}
+params <- full_join(tdat, odat, by=c("species", "age", "model", "covs"))
+
+other_params <- readRDS(paste0("data/optimal-values-by-species-by-age-tv-depth.rds"))
+params <- rbind(other_params, params) 
+#%>% select(species, age, model, optimal_temp, t_maxy, optimal_do, do_maxy, covs)
+saveRDS(params, file = "data/optimal-values-by-species-by-age-tv-depth.rds")
+
+params
+```
+
+Save DO plots
+```{r}
 png(
-  file = paste0("figs/", spp, "/do-", spp, covs, "-priors-TRUE.png"), 
+  file = paste0("figs/", spp, "/do-fixed-", spp, covs, "-priors-", priors, ".png"),
   res = 600,
@@
 
-Temperature plots
-```{r}
-rm(
-  adult_survey1, imm_survey1,
-  adult_survey4, imm_survey4,
-  adult_survey16, imm_survey16
-)
-priors <- TRUE
-try({
-  survey <- c("SYN QCS", "SYN HS")
-  model_ssid <- c(1, 3)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey1 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-
-  imm_survey1 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult_survey1")) adult_survey1 <- NULL
-if (!exists("imm_survey1")) imm_survey1 <- NULL
-
-try({
-  survey <- c("SYN WCVI")
-  model_ssid <- c(4)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey4 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-
-  imm_survey4 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult_survey4")) adult_survey4 <- NULL
-if (!exists("imm_survey4")) imm_survey4 <- NULL
-
-try({
-  survey <- c("SYN WCHG")
-  model_ssid <- c(16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey16 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-
-  imm_survey16 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-
-if (!exists("adult_survey16")) adult_survey16 <- NULL
-if (!exists("imm_survey16")) imm_survey16 <- NULL
-
-temp_model_list <- list(
-  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
-  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
-  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
-)
-
-temp_model_list <- temp_model_list[!sapply(temp_model_list, is.null)]
-```
-
-```{r}
-d <- list()
-temp_plots <- list()
-for (i in seq_len(length(temp_model_list))) {
-  d[[i]] <- time_varying_density(temp_model_list[[i]], predictor = "temp")
-  #d[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
-  temp_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Temperature", xlimits = c(2,10)) +
-    ggtitle(paste(species,names(temp_model_list[i])))
-}
-print(temp_plots)
-```
-
+Save Temperature plots
 ```{r}
 png(
-  file = paste0("figs/", spp, "/temp-", spp, covs, "-priors-TRUE.png"), 
+  file = paste0("figs/", spp, "/temp-fixed-", spp, covs, "-priors-", priors, ".png"),
   res = 600,
@@
 priors <- FALSE
+covariates <- "+muddy+any_rock"
+covs <- gsub("\\+", "-", covariates)
 try({
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey1 <- readRDS(paste0(
-    "data/", spp,
+  adult_survey1 <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
-
-  imm_survey1 <- readRDS(paste0(
-    "data/", spp,
+  imm_survey1 <- readRDS(paste0("data/", spp,
     "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey4 <- readRDS(paste0(
-    "data/", spp,
+  adult_survey4 <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
-
-  imm_survey4 <- readRDS(paste0(
-    "data/", spp,
+  imm_survey4 <- readRDS(paste0("data/", spp,
     "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-
-  adult_survey16 <- readRDS(paste0(
-    "data/", spp,
+  adult_survey16 <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
-
-  imm_survey16 <- readRDS(paste0(
-    "data/", spp,
+  imm_survey16 <- readRDS(paste0("data/", spp,
     "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
 )
-
 depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
@@
   d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
+
+  if (length(d[[i]]) == 0) {
+    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } else {
     d[[i]]$x <- exp(d[[i]]$x)
-  depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0,1000)) + 
+    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, 800)) +
       ggtitle(paste(species, names(depth_model_list[i])))
+  }
 }
@@
 ### Code for calculation SE on time-varying effects... 
-
 ```{r eval=FALSE}

### 2019-08-29 analysis/VOCC/1-biomass-do-models-fixed-both.Rmd
model do and temp while accounting for the effect of the other (these models are identical except for order of covariates)

--- 2019-08-28 analysis/VOCC/1-biomass-do-models-fixed.Rmd
@@
 # species <- "Sablefish"
- species <- "Silvergray Rockfish"
+# species <- "Silvergray Rockfish"
 # species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Quillback Rockfish"
+ species <- "Quillback Rockfish"
 # species <- "Pacific Ocean Perch"
@@
 covs <- gsub("\\+", "-", covariates)
-covs <- "-tv-depth"
+covs <- "-both-tv-depth"
 ```
@@
     filter(ssid %in% model_ssid) %>% 
-    filter(!is.na(do_mlpl)) %>% filter(!is.na(depth_scaled)) %>% 
-    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, do_mlpl_scaled, do_mlpl_scaled2, depth_mean, depth_sd, do_mlpl_mean, do_mlpl_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, do_mlpl) %>% filter(!is.na(depth_scaled))
+    filter(!is.na(do_mlpl)) %>% filter(!is.na(depth_scaled)) %>% filter(!is.na(temp))%>% 
+    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, depth_mean, depth_sd, do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, temp_scaled, temp_scaled2, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, do_mlpl) %>% filter(!is.na(depth_scaled))
 ```
@@
     adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""
+      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""
       ))
@@
       adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
       time = "year", spde = spde,
@@
     imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""
+      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""
       ))
@@
       imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
       time = "year", spde = spde,
@@
     starttime3 <- Sys.time()
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""))
+    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""))
 
@@
       dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
       time = "year", spde = spde,

### 2019-08-29 analysis/VOCC/1-biomass-temp-models-fixed-both.Rmd
model do and temp while accounting for the effect of the other (these models are identical except for order of covariates)

--- 2019-08-28 analysis/VOCC/1-biomass-temp-models-fixed.Rmd
@@
 # species <- "Silvergray Rockfish"
- species <- "Lingcod"
+# species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-
-# species <- "Quillback Rockfish"
+species <- "Quillback Rockfish"
 # species <- "Pacific Ocean Perch"
@@
   CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
-  data <- dplyr::left_join(data, CTD)
+  data <- dplyr::left_join(data, CTD) %>% 
+    mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
+    #mutate(exclude = if_else(do_mlpl>6, 1, 0)) %>% #try excluding high values?
+    filter(exclude != 1) %>% # added on july 24 after first round of models run...
+    filter(year > 2007) # 2007 DO data is flawed
 
@@
     filter(ssid %in% model_ssid) %>% 
-    filter(!is.na(raw_depth)) %>% filter(!is.na(temp)) %>%
-    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, temp_scaled, temp_scaled2, depth_mean, depth_sd, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, temp)
+    filter(!is.na(raw_depth)) %>% filter(!is.na(temp)) %>% select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2,  depth_mean, depth_sd, do_mlpl_scaled, do_mlpl_scaled2, do_mlpl_mean, do_mlpl_sd, temp_scaled, temp_scaled2, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, temp)
 ```
@@
     adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", 
+      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + do_mlpl_scaled + do_mlpl_scaled2", 
       covariates, ""
@@
     imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", 
+      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + do_mlpl_scaled + do_mlpl_scaled2", 
       #+ depth_scaled + depth_scaled2", 
@@
     starttime3 <- Sys.time()
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", 
-      #+ depth_scaled + depth_scaled2", 
+    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + do_mlpl_scaled + do_mlpl_scaled2", 
       covariates, ""))

### 2019-08-29 analysis/VOCC/3-fixed-var-response-plots.Rmd
add density plots for raw biomass vs climate measurements

@@
 # # Species run so far...
- species <- "Arrowtooth Flounder"
+# species <- "Arrowtooth Flounder" # still throws error for some models
 # species <- "Pacific Cod"
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish"
- species <- "Lingcod"
+# species <- "Silvergray Rockfish" # this one looks good!
+# species <- "Lingcod"
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Quillback Rockfish"
- species <- "Pacific Ocean Perch"
+ species <- "Quillback Rockfish"
+# species <- "Pacific Ocean Perch"
 # species <- "Yelloweye Rockfish"
@@
 covs <- "-tv-depth"
+#covs <- "-both-tv-depth"
 ```
@@
 # test load small sdm model to avoid crashing R
-try({imm})
+# try({imm})
 # or maybe it's crashing when NaNs produced error goes into density function?
@@
 
+## Density curves from model coefficients
+
+#### Calculate DO curves
 ```{r}
@@
 
+#### Calculate temperature curves
 ```{r}
@@
 
-Save optimal values
+## Check optimals against raw capture denstiies
+
+#### Combine biomass and sensor data
 ```{r}
-params <- full_join(tdat, odat, by=c("species", "age", "model", "covs"))
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  biomass <- readRDS(paste0(
+    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
+    ))
+  covars <- readRDS("data/event-covariates.rds")
+  data <- dplyr::left_join(biomass, covars) %>% mutate(depth_akima = depth) %>% select(-depth)
 
-other_params <- readRDS(paste0("data/optimal-values-by-species-by-age-tv-depth.rds"))
-params <- rbind(other_params, params) 
-#%>% select(species, age, model, optimal_temp, t_maxy, optimal_do, do_maxy, covs)
-saveRDS(params, file = "data/optimal-values-by-species-by-age-tv-depth.rds")
+  CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
+  data <- dplyr::left_join(data, CTD) %>% 
+    mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
+    #mutate(exclude = if_else(do_mlpl>6, 1, 0)) %>% #try excluding high values?
+    filter(exclude != 1) %>% # added on july 24 after first round of models run...
+    filter(year > 2007) # 2007 DO data is flawed
 
-params
+  # scale predictors before filtering to ensure mean and SD are global
+  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), temp = temperature_c) 
+```
+
+#### Plot raw DO
+```{r}
+odata <- data %>% filter(!is.na(do_mlpl)) %>% 
+  mutate(ad_density_class = ifelse(adult_density==0, 0, (round(log10((adult_density*10000)+1)*2)+1)),
+    im_density_class = ifelse(imm_density==0, 0, (round(log10((imm_density*10000)+1)*2)+1))) %>% 
+  group_by(ad_density_class) %>% mutate(ad_mean_density = mean(adult_density, na.rm = TRUE)) %>% ungroup() %>%
+  group_by(im_density_class) %>% mutate(im_mean_density = mean(imm_density, na.rm = TRUE)) %>% ungroup()
+
+ad_density_class_labs <- unique(signif(odata$ad_mean_density*10000),2) # convert to kg/ha
+ad_density_class_labs <- ad_density_class_labs[order(ad_density_class_labs)]
+ad_density_class_labs <- formatC(ad_density_class_labs, digits = 1, format = "fg")
+ad_density_class_labs
+
+im_density_class_labs <- unique(signif(odata$im_mean_density*10000),2) # convert to kg/ha
+im_density_class_labs <- im_density_class_labs[order(im_density_class_labs)]
+im_density_class_labs <- formatC(im_density_class_labs, digits = 1, format = "fg")
+im_density_class_labs
+
+do_count_ad <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(ad_density_class))) +
+  geom_density(position = "stack") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  geom_vline(xintercept=odat[odat$model =="adult", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
+  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Count") +
+  ggtitle(paste("Adult"))
+
+do_prop_ad <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(ad_density_class))) +
+  geom_density(position = "fill") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  geom_vline(xintercept=odat[odat$model =="adult", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
+  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Proportion") +
+  ggtitle(paste("Adult"))
+
+do_count_im <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(im_density_class))) +
+  geom_density(position = "stack") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
+  geom_vline( xintercept = odat[odat$model =="imm", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
+  geom_vline( xintercept = odat[odat$age =="imm", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Count") +
+  ggtitle(paste("Immature"))
+
+do_prop_im <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(im_density_class))) +
+  geom_density(position = "fill") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
+  geom_vline( xintercept = odat[odat$model =="imm", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
+  geom_vline( xintercept = odat[odat$age =="imm", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Proportion") +
+  ggtitle(paste("Immature"))
+
+do_count_ad
+do_prop_ad
+do_count_im
+do_prop_im
+```
+
+#### Plot raw temperature
+```{r}
+tdata <- data %>% filter(!is.na(temp)) %>% 
+  mutate(ad_density_class = ifelse(adult_density==0, 0, (round(log10((adult_density*10000)+1)*2)+1)),
+    im_density_class = ifelse(imm_density==0, 0, (round(log10((imm_density*10000)+1))+1)*2)) %>% 
+  group_by(ad_density_class) %>% mutate(ad_mean_density = mean(adult_density, na.rm = TRUE)) %>% ungroup() %>%
+  group_by(im_density_class) %>% mutate(im_mean_density = mean(imm_density, na.rm = TRUE)) %>% ungroup()
+
+ad_density_class_labs <- unique(signif(tdata$ad_mean_density*10000),2) # convert to kg/ha
+ad_density_class_labs <- ad_density_class_labs[order(ad_density_class_labs)]
+ad_density_class_labs <- formatC(ad_density_class_labs, digits = 1, format = "fg")
+ad_density_class_labs
+
+im_density_class_labs <- unique(signif(tdata$im_mean_density*10000),2) # convert to kg/ha
+im_density_class_labs <- im_density_class_labs[order(im_density_class_labs)]
+im_density_class_labs <- formatC(im_density_class_labs, digits = 1, format = "fg")
+im_density_class_labs
+
+#FIXME: breaks when no optimals are available
+t_count_ad <- ggplot(tdata, aes(x = temp, stat(count),fill = as.factor(ad_density_class))) + 
+  geom_density(position = "stack") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
+  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Temperature from CTD") +
+  ylab("Count") +
+  ggtitle(paste("Adult"))
+
+t_prop_ad <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(ad_density_class))) +
+  geom_density(position = "fill") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
+  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Temperature from CTD") +
+  ylab("Proportion") +
+  ggtitle(paste("Adult"))
+
+t_count_im <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(im_density_class))) + 
+  geom_density(position = "stack") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
+  geom_vline( xintercept = tdat[tdat$model =="imm", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
+  geom_vline( xintercept = tdat[tdat$age =="imm", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Temperature from CTD") +
+  ylab("Count") +
+  ggtitle(paste("Immature"))
+
+t_prop_im <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(im_density_class))) +
+  geom_density(position = "fill") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
+  geom_vline( xintercept = tdat[tdat$model =="imm", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
+  geom_vline( xintercept = tdat[tdat$age =="imm", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Temperature from CTD") +
+  ylab("Proportion") +
+  ggtitle(paste("Immature"))
+
+t_count_ad
+t_prop_ad 
+t_count_im 
+t_prop_im 
 ```
@@
   top = grid::textGrob(paste(species))
+)
+dev.off()
+```
+
+#### Save plots of raw densities vs model optimals 
+```{r}
+png(
+  file = paste0("figs/", spp, "/check-optimals-", spp, covs, "-priors-", priors, ".png"),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 11
+)
+gridExtra::grid.arrange( do_count_ad, do_count_im, do_prop_ad, do_prop_im, t_count_ad, t_count_im, t_prop_ad, t_prop_im, 
+  ncol = 2,
+  top = grid::textGrob(paste("Raw densities of", species, "with modeled optimals in vertical lines (if solid includes all surveys)"))
+)
+dev.off()
+```
+
+## Save optimal values
+```{r eval=FALSE}
+other_params <- readRDS(paste0("data/optimal-values-by-species-by-age-tv-depth.rds"))
+
+if(!is.null(odat)) {
+  if(!is.null(tdat)) {
+    params <- full_join(tdat, odat, by=c("species", "age", "model", "covs"))
+    all_params <- full_join(other_params, params)
+  } else {
+  try({all_params <- full_join(other_params, odat)})
+  }
+} else {
+  try({all_params <- full_join(other_params, tdat)})
+}
+
+all_params <- all_params %>% distinct()
+saveRDS(all_params, file = "data/optimal-values-by-species-by-age-tv-depth.rds")
+all_params
+```
+
+Alternative presense-absense code
+```{r eval=FALSE}
+odata <- data %>% filter(!is.na(do_mlpl)) #%>% filter(do_mlpl < 1.5)
+
+odata1 <- filter(odata, adult_density > 0)
+# ymax <- which.max(density(odata1$do_mlpl)$y)
+# do_optimal_raw <- density(odata1$do_mlpl)$x[ymax]
+
+oplot <- ggplot(odata1) + 
+  geom_density(aes(x = do_mlpl), colour = "blue") +
+  # geom_vline(aes(xintercept=do_optimal_raw),
+  #           color="blue", linetype="dashed", size=1) +
+  geom_density(data=filter(odata, adult_density == 0), aes(x = do_mlpl), colour = "red", inherit.aes = FALSE) + theme_bw() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Frequency") +
+  ggtitle(paste("Adult"))
+oplot
+
+oploti <- ggplot(filter(odata, imm_density > 0)) + geom_density(aes(x = do_mlpl), colour = "blue") + 
+  geom_density(data=filter(odata, imm_density == 0), aes(x = do_mlpl), colour = "red", inherit.aes = FALSE) + theme_bw() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Frequency") +
+  ggtitle(paste("Immature"))
+oploti
+
+tdata <- data %>% filter(!is.na(temp)) #%>% filter(do_mlpl < 1.5)
+tplot <-ggplot(filter(tdata, adult_density > 0)) + geom_density(aes(x = temp), colour = "blue") + 
+  geom_density(data=filter(tdata, adult_density == 0), aes(x = temp), colour = "red", inherit.aes = FALSE) + theme_bw() +
+  xlab("Temperature from CTD") +
+  ylab("Frequency") +
+  ggtitle(paste("Adult"))
+tplot
+tploti <-ggplot(filter(tdata, imm_density > 0)) + geom_density(aes(x = temp), colour = "blue") + 
+  geom_density(data=filter(tdata, imm_density == 0), aes(x = temp), colour = "red", inherit.aes = FALSE) + theme_bw() +
+  xlab("Temperature from CTD") +
+  ylab("Frequency") +
+  ggtitle(paste("Immature"))
+tploti
+```
+
+Save raw presense-absence density plots
+```{r eval=FALSE}
+png(
+  file = paste0("figs/", spp, "/raw-presense-by-climate-", spp, "-density-plots.png"),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 11
+)
+gridExtra::grid.arrange(
+  oplot,  tplot, oploti, tploti,
+  ncol = 2,
+  top = grid::textGrob(paste("Proportion of tows with", species, "(blue) vs without (red)"), gp=grid::gpar(fontsize=16))
 )
@@
 ```
+
+Get all biomass predictions 
+#### Get all biomass predictions 
+```{r, eval= FALSE, echo=FALSE, warning=FALSE}
+priors <- FALSE
+covariates <- "+muddy+any_rock"
+covs <- gsub("\\+", "-", covariates)
+
+rm(adult_predictions)
+rm(imm_predictions)
+rm(predictions)
+pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
+
+try({adult_predictions <- readRDS(paste0("data/", spp,
+    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
+  ))%>% filter(year > 2007)
+adult_predictions <- left_join(adult_predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
+})
+
+try({imm_predictions <- readRDS(paste0("data/", spp,
+    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
+  )) %>% filter(year > 2007)
+imm_predictions <- left_join(imm_predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
+})
+
+try({predictions <- readRDS(paste0("data/", spp,
+    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
+  ))%>% filter(year > 2007)
+predictions <- left_join(predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
+})
+
+if (!exists("adult_predictions")) adult_predictions <- NULL
+if (!exists("imm_predictions")) imm_predictions <- NULL
+if (!exists("predictions")) predictions <- NULL
+
+biomass_list <- list(
+  adult = adult_predictions,
+  imm = imm_predictions,
+  total = predictions
+)
+
+biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
+
+# 
+# ggplot(biomass_list[[1]], aes(x = do_est, y = log(est_exp))) + 
+#   geom_point(alpha = 0.3) #+ #geom_smooth()
+#   geom_smooth(method = "lm", formula = y ~ x + I(x^2) , size = 1) #+ I(x^3)
+```

### 2019-08-29 analysis/VOCC/1-biomass-temp-models-fixed.Rmd
various minor housekeeping

@@
 ```{r}
- region <- "All synoptic surveys"
-# region <- "Both odd year surveys"
-# region <- "West Coast Vancouver Island"
-# region <- "West Coast Haida Gwaii"
+# region <- "All synoptic surveys"
+ region <- "Both odd year surveys"
+ region <- "West Coast Vancouver Island"
+ region <- "West Coast Haida Gwaii"
 ```
@@
     filter(ssid %in% model_ssid) %>% 
-    filter(!is.na(raw_depth)) %>% filter(!is.na(temp)) %>%
-    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, temp_scaled, temp_scaled2, depth_mean, depth_sd, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, temp)
+    filter(!is.na(raw_depth)) %>% filter(!is.na(temp)) #%>% select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, temp_scaled, temp_scaled2, depth_mean, depth_sd, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, temp)
 ```

### 2019-08-29 analysis/VOCC/2-biomass-trends.Rmd
various minor housekeeping

@@
 
+  # to convert from kg/m2 to kg/hectare multiply by 10000
   adult_predictions$est_exp <- exp(adult_predictions$est) * 10000

### 2019-08-29 analysis/VOCC/3-variable-response-plots.Rmd
various minor housekeeping

@@
 
-date <- Sys.time()
+date <- Sys.Date()
 

### 2019-08-29 R/survey-data-processing.R

@@ -108,17 +108,15 @@ add_missing_depths <- function(dat, survey, years, bath) {
 #' Scale survey predictors
 #'
 #' @param dat
-#' @param predictors List of variable names must by wrapped in 'quo()'.
+#' @param predictors List of variable names must be wrapped in 'quo()'.
 #'
 #' @export
 #'
-scale_predictors <- function(dat,
-                             predictors = c(quo(log_depth))) {
+scale_predictors <- function(dat, predictors = c(quo(log_depth))) {
   # to put spatial decay parameter on right scale
 
   for (i in seq_len(length(predictors))) {
     var <- predictors[[i]]
-
     var_mean <- paste0(quo_name(var), "_mean")
     var_sd <- paste0(quo_name(var), "_sd")
     var_scaled <- paste0(quo_name(var), "_scaled")

### 2019-08-29 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
optimal values integration with BACI code... incomplete

@@
 
-Should optimal temperature vary for age classes? Seems likely. 
-Therefore need separate VOCC trimming and matching for age classes -> change function structure to choose age upfront
+Modified to allow optimal temperature to vary by maturity class. 
+Therefore need to run this file separately for each species-class combination.
 
 Should matches be selected based on exact depth categories in addition to density (all within time periods)?
+
 What does it mean for the analysis that the depth range at which VOCC occur within a species temperature range varies dramatically between time periods?
@@
 # climate <- "do"
-# do_threshold <- c(0.1)
+# do_threshold <- c(0.5)
 
@@
 # region <- "West Coast Haida Gwaii"
+```
+
+#### Set default cell sizes and scale
+```{r, echo=FALSE}
+input_cell_size <- 2
+dist_intercept <- input_cell_size/2
+scale_fac <- 1
+delta_t_step <- 2
+skip_time <- NULL
 ```
@@
 #### Get species-specific optimal values from mountain plots for most abundant year
-```{r, echo=FALSE}
+```{r, include=FALSE}
 spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
@@
 }
-params <- params[rev(order(params$maxy)), ]
-params <- params[!duplicated(params$species), ]
 
-#optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
-optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)
+params <- params[params$species == species, ]
 
-#optimal_values <- c(optimal_do, optimal_temp)
+if (climate == "temperature") {
+  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
+  optimal_temps <- na.omit(optimal_temps)
+  optimal_temp <- round(optimal_temps[1], digits = 2)
+  variable_names <- c("temp")
+  min_thresholds <- c(Inf)
+  optimal_values <- optimal_temp
+  max_thresholds <- temp_threshold
+  indices <- c(1, 2)
+  delta_t_total <- 2
+  }
+
+if (climate == "do") {
+  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
+  optimal_dos <- na.omit(optimal_dos)
+  optimal_do <- round(optimal_dos[1], digits = 2)
+  variable_names <- c("do_est")
+  max_thresholds <- c(Inf)
+  optimal_values <- optimal_do
+  min_thresholds <- do_threshold
+  indices <- c(1, 2)
+  delta_t_total <- 2
+  }
+
+if (climate == "do and temperature") {
+  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
+  optimal_temps <- na.omit(optimal_temps)
+  optimal_temp <- round(optimal_temps[1], digits = 2)
+  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
+  optimal_dos <- na.omit(optimal_dos)
+  optimal_do <- round(optimal_dos[1], digits = 2)
+  optimal_values <- c(optimal_do, optimal_temp)
+  variable_names <- c("do_est", "temp")
+  min_thresholds <- c(do_threshold, Inf)
+  max_thresholds <- c(Inf, temp_threshold)
+  indices <- c(1, 2)
+  delta_t_total <- 2
+  }
+
+min_string <- paste0(min_thresholds, collapse = "-")
+max_string <- paste0(max_thresholds, collapse = "-")
+
+optimal_string <- paste0(optimal_values, collapse = "-")
+climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
 ```
@@
 
-```{r, include=FALSE}
-# Set default cell sizes and scale
-input_cell_size <- 2
-dist_intercept <- input_cell_size/2
-scale_fac <- 1
-delta_t_step <- 2
-skip_time <- NULL
-
-if (climate == "temperature") {
-  variable_names <- c("temp")
-  min_thresholds <- c(Inf)
-  optimal_values <- optimal_temp
-  max_thresholds <- temp_threshold
-  indices <- c(1, 2)
-  delta_t_total <- 2
-  }
-
-if (climate == "do") {
-  variable_names <- c("do_est")
-  max_thresholds <- c(Inf)
-  optimal_values <- optimal_do
-  min_thresholds <- do_threshold
-  indices <- c(1, 2)
-  delta_t_total <- 2
-  }
-
-if (climate == "do and temperature") {
-  variable_names <- c("do_est", "temp")
-  min_thresholds <- c(do_threshold, Inf)
-  max_thresholds <- c(Inf, temp_threshold)
-  indices <- c(1, 2)
-  delta_t_total <- 2
-  }
-
-min_string <- paste0(min_thresholds, collapse = "-")
-max_string <- paste0(max_thresholds, collapse = "-")
-
-optimal_string <- paste0(optimal_values, collapse = "-")
-climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
-```
-
 # Calculate climate vectors
-
 ```{r}
@@
 if (!exists("vocc_by_sp")){
-match_plots <- lapply(matched, function(i){  
-  summary(i)
-  plot(i)
+match_plots <- apply(matched, function(matched){  
+  summary(matched)
+  plot(matched)
  })
@@
 
-## Adults
 ```{r}
@@
   vocc_by_sp <- readRDS(paste0(
-    "data/ BACI-", species, "-", ssid_string, "-", climate_string, "-",
+    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
     min_string, "to", max_string, "-multiyear-all.rds"
@@
 })
-data <- vocc_by_sp [["adult"]]
+data <- vocc_by_sp [[1]]
 # View(data)
@@
 
-```{r,  results="hide", include=FALSE}
+```{r, eval=FALSE, results="hide", include=FALSE}
 plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source"), start_time=c("2009","2011","2013","2015")), parametricOnly = TRUE, main="", xlab ="log_density"
@@
 ```
-
-## Immatures
-```{r}
-idata <- vocc_by_sp[["imm"]]
-#View(data)
-```
-
-### Plot matches
-```{r}
-idat_summary <- idata %>% 
-  group_by(start_time, cell_type) %>%
-  tally() %>% filter(cell_type=="source")
-idata_depth <- idata %>% 
-  group_by(start_time) %>% 
-  mutate(mean_depth = mean(depth))
-
-ggplot(idata, 
-#ggplot(filter(data, as.numeric(matchobs) <= 550), 
-  aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
-  geom_point(alpha=0.4, size = 1) +
-  scale_shape_manual(values= c(0, 15)) +
-  scale_colour_viridis_d(option="C", guide = FALSE) +
-  facet_wrap(~start_time) +
-  theme_bw()
-
-lapply(variable_names, function(j) {
-ggplot(idata, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
-  geom_point(alpha=0.35) +
-  scale_colour_viridis_d(option="C")+
-  facet_grid(start_time~cell_type) +
-  theme_bw() 
-})
-
-# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
-#   geom_point(alpha=0.35) +
-#   scale_colour_viridis_d(option="C") +
-#     facet_grid(start_time~time) +
-#   geom_text(data = idata_depth,
-#     aes(150, 0.5, label = round(mean_depth)), hjust = -0.3, inherit.aes = FALSE) +
-#   theme_bw()
-ggplot(idata, aes(depth, fill = cell_type)) +
-  geom_density(alpha=0.5) +
-  scale_fill_viridis_d(option="C") +
-  facet_wrap(~start_time) +
-  theme_bw()
-
-ggplot(idata, aes(log_density, fill = cell_type)) +
-  geom_density(alpha=0.5) +
-  scale_fill_viridis_d(option="C") +
-  facet_grid(start_time~time) +
-  geom_text(data = idat_summary,
-            aes(1, 0.75, label = n), hjust = -0.3) +
-  theme_bw()
-
-ggplot(idata, aes(log_density, fill = cell_type)) +
-  geom_histogram(alpha=0.5) +
-  scale_fill_viridis_d(option="C") +
-  facet_grid(time~start_time) +
-  theme_bw()
-
-ggplot(idata, aes(y=log_density, x=cell_type, fill=cell_type)) +
-  geom_boxplot(alpha=0.5) +
-  scale_fill_viridis_d(option="C") +
-  geom_text(data = idat_summary,
-            aes(cell_type, max(data$log_density), label = n), hjust = -0.3) +
-  facet_grid(time~start_time) +
-  theme_bw()
-```
-
-### Basic BACI
-```{r}
-mi <- lmerTest::lmer(as.formula(baci_formula), data=idata)
-summary(mi)
-```
-
-```{r}
-sjPlot::plot_model(mi, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "immature"))
-```
-
-### GAMM
-```{r}
-# for (j in (variable_names)) {
-#  gamm_formula <- paste0("log_density ~ ", 
-#   # "time * start_time + ",
-#   "start_time * cell_type + ",
-#   "time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
-#  }
-igam4 <- gamm4::gamm4(as.formula(gamm_formula), 
-  random = ~(1|matchobs) + (1|icell) 
-   + (1|start_time)
-  # + (time|start_time) 
-  # + (cell_type|start_time)
-  , data=idata)
-
-summary(igam4$gam)
-
-plot(igam4$gam,pages=1)
-```
-
-```{r, results="hide", include=FALSE}
-plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), parametricOnly = TRUE, main="", xlab ="log_density"
-  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
-  ) 
-```
-
-Misc exploring gamm 
-```{r eval=FALSE}
-# summary(igam4$mer) ## underlying mixed model
-# anova(igam4$gam)
-
-# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), parametricOnly = TRUE, main="")
-
-# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + vect_dist + vect_dist2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
-# summary(mi)
-# 
-# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(vect_dist) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
-# summary(mi)
-# mi
-# 
-# 
-# idata$after <- as.factor(idata$after)
-# idata$matchobs <- as.factor(idata$matchobs)
-# 
-# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(vect_dist) + #s(X) + s(Y) + 
-#     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")
-# 
-# 
-# plot(m4b$gam,pages=1)
-# 
-# summary(m4b$gam) ## summary of gam
-# summary(m4b$mer) ## underlying mixed model
-# anova(m4b$gam) 
-# 
-# gamtabs(m1)
-# 
-# #library(itsadug)
-# plot_parametric(m4b$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")))
-# 
-# library(mgcv)
-```

### 2019-09-06 analysis/VOCC/1-biomass-do-models-fixed-both.Rmd
add species; allow to run for those not split by maturity

@@
 # species <- "Arrowtooth Flounder"
+# species <- "Petrale Sole"
+# species <- "English Sole"
+# species <- "Dover Sole" # maturity data error for WCHG
+# species <- "Southern Rock Sole"
+# species <- "Flathead Sole"
+
+# species <- "Walleye Pollock"
 # species <- "Pacific Cod"
+# species <- "Pacific Ocean Perch"
+
+# species <- "Silvergray Rockfish"
+# species <- "Quillback Rockfish"
+# species <- "Yelloweye Rockfish"
+# species <- "Bocaccio"
+# species <- "Sharpchin Rockfish"
+# species <- "Splitnose Rockfish"
+# species <- "Canary Rockfish"
+# species <- "Rougheye/Blackspotted Rockfish Complex"
+# species <- "Redbanded Rockfish"
+# species <- "Greenstriped Rockfish"
+
+# species <- "Longspine Thornyhead"
+# species <- "Shortspine Thornyhead",
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish"
 # species <- "Lingcod"
+
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
- species <- "Quillback Rockfish"
-# species <- "Pacific Ocean Perch"
-# species <- "Yelloweye Rockfish"
+# species <- "Longnose Skate"
+species <- "Spotted Ratfish"
+
 ```
@@
     filter(ssid %in% model_ssid) %>% 
-    filter(!is.na(do_mlpl)) %>% filter(!is.na(depth_scaled)) %>% filter(!is.na(temp))%>% 
-    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, depth_mean, depth_sd, do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, temp_scaled, temp_scaled2, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, do_mlpl) %>% filter(!is.na(depth_scaled))
+    filter(!is.na(depth_scaled)) %>% filter(!is.na(do_mlpl)) %>% filter(!is.na(temp))
+
+if (any(names(data) == "adult_density")) { 
+  data <- data %>% select(X, Y, X10, Y10,
+      adult_density, imm_density,
+      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
+      do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, 
+      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
+      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
+      depth, do_mlpl) 
+ } else {
+    data <- data %>% select(X, Y, X10, Y10,
+      density,
+      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
+      do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, 
+      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
+      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
+      depth, do_mlpl) 
+}
 ```

### 2019-09-06 analysis/VOCC/3-fixed-var-response-plots.Rmd
save thresholds from quadratic roots

@@
 ```{r}
-# species <- params$species
+
 # # Species run so far...
-# species <- "Arrowtooth Flounder" # still throws error for some models
+ species <- "Arrowtooth Flounder" # running more models
+
+# species <- "Petrale Sole"
+# species <- "English Sole"
+# species <- "Dover Sole"
+# species <- "Southern Rock Sole"
+
+ species <- "Walleye Pollock" # running more models
 # species <- "Pacific Cod"
+# species <- "Pacific Ocean Perch" # beautiful curves, no depth change
+
+# species <- "Silvergray Rockfish" # beautiful curves, no depth change
+ species <- "Quillback Rockfish" # overall a bit flat... running survey specific models
+# species <- "Yelloweye Rockfish" # beautiful curves, with depth change
+ species <- "Bocaccio" # good curves, no depth change ... should run survey specific models
+ species <- "Sharpchin Rockfish"
+# species <- "Splitnose Rockfish"
+# species <- "Canary Rockfish"
+# species <- "Rougheye/Blackspotted Rockfish Complex"
+# species <- "Redbanded Rockfish"
+# species <- "Greenstriped Rockfish"
+
+# species <- "Longspine Thornyhead"
+# species <- "Shortspine Thornyhead",
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish" # this one looks good!
 # species <- "Lingcod"
+
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
- species <- "Quillback Rockfish"
-# species <- "Pacific Ocean Perch"
-# species <- "Yelloweye Rockfish"
+# species <- "Longnose Skate"
+#  species <- "Spotted Ratfish"
 ```
@@
 priors <- TRUE
-covs <- "-tv-depth"
-#covs <- "-both-tv-depth"
+#covs <- "-tv-depth"
+covs <- "-both-tv-depth"
+num_params <- 2
 ```
@@
   adult_survey4, imm_survey4,
-  adult_survey16, imm_survey16)
+  adult_survey16, imm_survey16,
+  total, total1, total4, total16
+  )
 
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  adult <- readRDS(paste0("data/", spp,
+  try({adult <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   })
+  try({total <- readRDS(paste0("data/", spp,
+    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  })
+})
 if (!exists("adult")) adult <- NULL
 if (!exists("imm")) imm <- NULL
+if (!exists("total")) total <- NULL
 
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey1 <- readRDS(paste0("data/", spp,
+  try({adult_survey1 <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   })
+  try({total1 <- readRDS(paste0("data/", spp,
+    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  })
+  })
 if (!exists("adult_survey1")) adult_survey1 <- NULL
 if (!exists("imm_survey1")) imm_survey1 <- NULL
+if (!exists("total1")) total1 <- NULL
 
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey4 <- readRDS(paste0("data/", spp,
+  try({adult_survey4 <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   })
+  try({total4 <- readRDS(paste0("data/", spp,
+    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  })
+})
 if (!exists("adult_survey4")) adult_survey4 <- NULL
 if (!exists("imm_survey4")) imm_survey4 <- NULL
+if (!exists("total4")) total4 <- NULL
 
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey16 <- readRDS(paste0("data/", spp,
+  try({adult_survey16 <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   })
+  try({total16 <- readRDS(paste0("data/", spp,
+    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  })
+})
 if (!exists("adult_survey16")) adult_survey16 <- NULL
 if (!exists("imm_survey16")) imm_survey16 <- NULL
+if (!exists("total16")) total16 <- NULL
 
-model_list <- list(
-  adult = adult, imm = imm,
+model_list <- list(total1 = total1, total4 = total4, total16 = total16, 
   adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
   adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
-  adult_survey16 = adult_survey16#, imm_survey16 = imm_survey16
+  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16,
+  total = total, adult = adult, imm = imm
 )
@@
 
-Load temperature models
-```{r}
-rm(adult, imm,
-  adult_survey1, imm_survey1,
-  adult_survey4, imm_survey4,
-  adult_survey16, imm_survey16)
-
-try({
-  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
-  model_ssid <- c(1, 3, 4, 16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  adult <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  imm <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult")) adult <- NULL
-if (!exists("imm")) imm <- NULL
-
-try({
-  survey <- c("SYN QCS", "SYN HS")
-  model_ssid <- c(1, 3)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey1 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  imm_survey1 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult_survey1")) adult_survey1 <- NULL
-if (!exists("imm_survey1")) imm_survey1 <- NULL
-
-try({
-  survey <- c("SYN WCVI")
-  model_ssid <- c(4)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey4 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  imm_survey4 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult_survey4")) adult_survey4 <- NULL
-if (!exists("imm_survey4")) imm_survey4 <- NULL
-
-try({
-  survey <- c("SYN WCHG")
-  model_ssid <- c(16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey16 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  imm_survey16 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult_survey16")) adult_survey16 <- NULL
-if (!exists("imm_survey16")) imm_survey16 <- NULL
-
-temp_model_list <- list(
-  adult = adult, imm = imm, 
-  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
-  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
-  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
-)
-temp_model_list <- temp_model_list[!sapply(temp_model_list, is.null)]
-```
-
 ## Density curves from model coefficients
@@
 for (i in seq_len(length(model_list))) {
+
+ # check if models converged
+ if(model_list[[i]]$model$convergence==0){
   do[[i]] <- fixed_density(model_list[[i]], predictor = "do_mlpl")
+
   if (length(do[[i]])>0) {
    do[[i]]$species <- species
+   do[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
    do[[i]]$model <- names(model_list[i])
    do[[i]]$covs <- covs 
-   do[[i]]$optimal_do <- get_optimal_value(do[[i]], xlimits = c(0.5, 6))
+   roots <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.5)
+   do[[i]]$optimal_do <- roots[["optimal"]]
+   do[[i]]$lower_do_50 <- roots[["lower_threshold"]]
+   do[[i]]$upper_do_50 <- roots[["upper_threshold"]]
+   roots75 <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.75)
+   do[[i]]$lower_do_75 <- roots75[["lower_threshold"]]
+   do[[i]]$upper_do_75 <- roots75[["upper_threshold"]]
+   roots25 <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.25)
+   do[[i]]$lower_do_25 <- roots25[["lower_threshold"]]
+   do[[i]]$upper_do_25 <- roots25[["upper_threshold"]]
    do_plots[[i]] <- 
      plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0, 8)) +
-        ggtitle(paste(species, names(model_list[i])))
+        ggtitle(paste(species, names(model_list[i]))) +
+        geom_vline(xintercept=do[[i]]$optimal_do, size= 1.2, colour = "darkgrey") + 
+        geom_vline(xintercept=do[[i]]$lower_do_50, linetype="longdash", alpha = 0.75) + 
+        geom_vline(xintercept=do[[i]]$upper_do_50, linetype="longdash", alpha = 0.75) +
+        geom_vline(xintercept=do[[i]]$lower_do_75, linetype="dotted", alpha = 0.5) + 
+        geom_vline(xintercept=do[[i]]$upper_do_75, linetype="dotted", alpha = 0.5) +
+        geom_vline(xintercept=do[[i]]$lower_do_25, linetype="dotted", alpha = 0.5) + 
+        geom_vline(xintercept=do[[i]]$upper_do_25, linetype="dotted", alpha = 0.5) 
+   } else {
+    do_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+   }
  } else {
@@
   odat <- odat %>% group_by(model) %>%
-  mutate(age = ifelse(grepl("imm", model), "imm", "adult"),
-    do_maxy = max(y_hat)) %>% 
-  select(species, age, model, optimal_do, do_maxy, covs) %>% 
+  mutate(do_maxy = max(y_hat)) %>% 
+  select(-x, -y_hat, -year) %>% 
   distinct() 
@@
 temp_plots <- list()
-for (i in seq_len(length(temp_model_list))) {
-  td[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
+for (i in seq_len(length(model_list))) {
+ if(model_list[[i]]$model$convergence==0){
+  td[[i]] <- fixed_density(model_list[[i]], predictor = "temp", fixed_params = num_params)
   if (length(td[[i]])>0) {
     td[[i]]$species <- species
-    td[[i]]$model <- names(temp_model_list[i])
+    td[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
+    td[[i]]$model <- names(model_list[i])
     td[[i]]$covs <- covs
-    td[[i]]$optimal_temp <- get_optimal_value(td[[i]], xlimits = c(0, 12))
-    temp_plots[[i]] <- plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 10)) + 
-      ggtitle(paste(species, names(temp_model_list[i])))
+   roots <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.5)
+   td[[i]]$optimal_temp <- roots[["optimal"]]
+   td[[i]]$lower_t_50 <- roots[["lower_threshold"]]
+   td[[i]]$upper_t_50 <- roots[["upper_threshold"]]
+   roots75 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.75)
+   td[[i]]$lower_t_75 <- roots75[["lower_threshold"]]
+   td[[i]]$upper_t_75 <- roots75[["upper_threshold"]]
+   roots25 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.25)
+   td[[i]]$lower_t_25 <- roots25[["lower_threshold"]]
+   td[[i]]$upper_t_25 <- roots25[["upper_threshold"]]
+   temp_plots[[i]] <- 
+     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 10)) +
+        ggtitle(paste(species, names(model_list[i]))) +
+        geom_vline(xintercept=td[[i]]$optimal_t, size= 1.2, colour = "darkgrey") + 
+        geom_vline(xintercept=td[[i]]$lower_t_50, linetype="longdash", alpha = 0.75) + 
+        geom_vline(xintercept=td[[i]]$upper_t_50, linetype="longdash", alpha = 0.75) +
+        geom_vline(xintercept=td[[i]]$lower_t_75, linetype="dotted", alpha = 0.5) + 
+        geom_vline(xintercept=td[[i]]$upper_t_75, linetype="dotted", alpha = 0.5) +
+        geom_vline(xintercept=td[[i]]$lower_t_25, linetype="dotted", alpha = 0.5) + 
+        geom_vline(xintercept=td[[i]]$upper_t_25, linetype="dotted", alpha = 0.5)
+
+  } else {
+    temp_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } 
  } else {
@@
   tdat <- tdat %>% group_by(model) %>%
-  mutate(age = ifelse(grepl("imm", model), "imm", "adult"),
-    t_maxy = max(y_hat)) %>% 
-  select(species, age, model, optimal_temp, t_maxy, covs) %>%
+  mutate(t_maxy = max(y_hat)) %>% 
+  select(-x, -y_hat, -year) %>% 
   distinct() 
@@
 
+Build time-varying depth plots
+```{r }
+max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
+d <- list()
+depth_plots <- list()
+for (i in seq_len(length(model_list))) {
+ if(model_list[[i]]$model$convergence==0){
+  d[[i]] <- time_varying_density(model_list[[i]], predictor = "depth")
+
+  if (length(d[[i]]) == 0) {
+    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } else {
+    d[[i]]$x <- exp(d[[i]]$x)
+    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, max_depth_found)) +
+      ggtitle(paste(species, names(model_list[i])))
+  }
+ } else {
+    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } 
+}
+print(depth_plots)
+```
+
 #### Plot raw DO
 ```{r}
+if(!is.null(data$adult_density)){
 odata <- data %>% filter(!is.na(do_mlpl)) %>% 
@@
   ylab("Count") +
-  ggtitle(paste("Adult"))
+  ggtitle(paste("Adult", species))
 
@@
   ylab("Proportion") +
-  ggtitle(paste("Adult"))
+  ggtitle(paste("Adult", species))
 
@@
   ylab("Count") +
-  ggtitle(paste("Immature"))
+  ggtitle(paste("Immature", species))
 
@@
   ylab("Proportion") +
-  ggtitle(paste("Immature"))
+  ggtitle(paste("Immature", species))
 
-do_count_ad
-do_prop_ad
-do_count_im
-do_prop_im
+raw_do_plots <- list(do_count_ad, do_prop_ad, do_count_im, do_prop_im )
+print(raw_do_plots)
+
+} else {
+
+odata <- data %>% filter(!is.na(do_mlpl)) %>% 
+  mutate(density_class = ifelse(density==0, 0, (round(log10((density*10000)+1)*2)+1))) %>% 
+  group_by(density_class) %>% mutate(mean_density = mean(density, na.rm = TRUE)) %>% ungroup()
+
+density_class_labs <- unique(signif(odata$mean_density*10000),2) # convert to kg/ha
+density_class_labs <- density_class_labs[order(density_class_labs)]
+density_class_labs <- formatC(density_class_labs, digits = 1, format = "fg")
+density_class_labs
+
+do_count <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(density_class))) +
+  geom_density(position = "stack") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
+  geom_vline(xintercept=odat[odat$model =="total", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
+  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Count") +
+  ggtitle(paste("Total", species))
+
+do_prop <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(density_class))) +
+  geom_density(position = "fill") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
+  geom_vline(xintercept=odat[odat$model =="total", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
+  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Dissolved O2 from CTD (ml/L)") +
+  ylab("Proportion") +
+  ggtitle(paste("Total", species))
+
+raw_do_plots <- list(do_count,do_prop)
+print(raw_do_plots)
+}
 ```
@@
 ```{r}
+if(!is.null(data$adult_density)){
 tdata <- data %>% filter(!is.na(temp)) %>% 
@@
 
-#FIXME: breaks when no optimals are available
 t_count_ad <- ggplot(tdata, aes(x = temp, stat(count),fill = as.factor(ad_density_class))) + 
@@
   ylab("Proportion") +
-  ggtitle(paste("Immature"))
+  ggtitle(paste("Immature", species))
 
-t_count_ad
-t_prop_ad 
-t_count_im 
-t_prop_im 
+raw_t_plots <- list(t_count_ad, t_prop_ad, t_count_im, t_prop_im )
+print(raw_t_plots)
+
+} else {
+
+tdata <- data %>% filter(!is.na(temp)) %>% 
+  mutate(density_class = ifelse(density==0, 0, (round(log10((density*10000)+1)*2)+1))) %>% 
+  group_by(density_class) %>% mutate(mean_density = mean(density, na.rm = TRUE)) %>% ungroup()
+
+density_class_labs <- unique(signif(tdata$mean_density*10000),2) # convert to kg/ha
+density_class_labs <- density_class_labs[order(density_class_labs)]
+density_class_labs <- formatC(density_class_labs, digits = 1, format = "fg")
+density_class_labs
+
+t_count <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(density_class))) +
+  geom_density(position = "stack") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
+  geom_vline(xintercept=tdat[tdat$model =="total", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
+  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Temperature from CTD") +
+  ylab("Count") +
+  ggtitle(paste("Total", species))
+
+t_prop <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(density_class))) +
+  geom_density(position = "fill") +
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
+  geom_vline(xintercept=tdat[tdat$model =="total", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
+  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
+  theme_pbs() +
+  xlab("Temperature from CTD") +
+  ylab("Proportion") +
+  ggtitle(paste("Total", species))
+
+raw_t_plots <- list(t_count,t_prop)
+print(raw_t_plots)
+}
 ```
@@
 )
-gridExtra::grid.arrange( do_count_ad, do_count_im, do_prop_ad, do_prop_im, t_count_ad, t_count_im, t_prop_ad, t_prop_im, 
+gridExtra::grid.arrange(grobs = c(raw_do_plots, raw_t_plots),
   ncol = 2,
@@
 
+#### Save plots of time-varying depth from climate model
+```{r}
+png(
+  file = paste0(
+    "figs/", spp,
+    "/depth-", spp, covs, "-priors-", priors, ".png"
+  ),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 11
+)
+gridExtra::grid.arrange(
+  grobs = depth_plots,
+  ncol = 2,
+  top = grid::textGrob(paste(species))
+)
+dev.off()
+```
+
 ## Save optimal values
-```{r eval=FALSE}
-other_params <- readRDS(paste0("data/optimal-values-by-species-by-age-tv-depth.rds"))
+```{r }
+other_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
 
@@
 all_params <- all_params %>% distinct()
-saveRDS(all_params, file = "data/optimal-values-by-species-by-age-tv-depth.rds")
+# # If data far a particular species needs redoing
+# all_params <- subset(other_params, species!="Longnose Skate")
+saveRDS(all_params, file = "data/root-values-by-species-by-age.rds")
 all_params
@@
 
-Load Depth models
-```{r}
+Load Depth models (if using old depth only models)
+```{r eval=FALSE}
 rm(
@@
 Build depth plots
-```{r}
+```{r eval=FALSE}
+
+max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
 d <- list()
@@
     d[[i]]$x <- exp(d[[i]]$x)
-    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, 800)) +
+    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, max_depth_found)) +
       ggtitle(paste(species, names(depth_model_list[i])))
@@
 
-```{r}
+```{r eval=FALSE}
 png(
@@
     "figs/", spp,
-    "/depth-", spp, covs, "-priors-", priors, ".png"
+    "/depth-only-", spp, covs, "-priors-", priors, ".png"
   ),
@@
 
+# TRAWLING
 Plot time-varying effect of trawling at mean depth
-```{r}
+```{r eval=FALSE}
 trawl_plots <- list()
@@
 #### Get all biomass predictions 
-```{r, eval= FALSE, echo=FALSE, warning=FALSE}
+```{r, eval=FALSE, echo=FALSE, warning=FALSE}
 priors <- FALSE

### 2019-09-06 analysis/VOCC/1-biomass-models.Rmd
housekeeping

@@
 # species <- "Arrowtooth Flounder"
-#species <- "Pacific Cod"
+# species <- "Petrale Sole"
+# species <- "English Sole"
+# species <- "Dover Sole" # maturity data error for WCHG
+# species <- "Southern Rock Sole"
+# species <- "Flathead Sole"
+
+# species <- "Walleye Pollock"
+# species <- "Pacific Cod"
+# species <- "Pacific Ocean Perch"
+
+# species <- "Silvergray Rockfish"
+# species <- "Quillback Rockfish"
+# species <- "Yelloweye Rockfish"
+# species <- "Bocaccio"
+# species <- "Sharpchin Rockfish"
+# species <- "Splitnose Rockfish"
+# species <- "Canary Rockfish"
+# species <- "Rougheye/Blackspotted Rockfish Complex"
+ species <- "Redbanded Rockfish"
+# species <- "Greenstriped Rockfish"
+
+# species <- "Longspine Thornyhead"
+# species <- "Shortspine Thornyhead",
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish"
 # species <- "Lingcod"
-# species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
-# species <- "Quillback Rockfish" # no maturity classes and WCHG not work
-# species <- "Pacific Ocean Perch" 
- species <- "Yelloweye Rockfish"
+
+# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
+# species <- "Longnose Skate"
+ species <- "Spotted Ratfish"
+
 ```
@@
   years <- NULL
-}
-```
-
-Make prediction grids for all surveys and years, if not done before.
-```{r, message=FALSE, warning=FALSE}
-try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
-
-if (!exists("nd_all")) {
-
-  # choose base year(s) to create grid from
-  dummy_year <- c(2005, 2006)
-
-  bath <- readRDS("data/bathymetry-data")
-  #dat <- bath$data %>% scale_survey_predictors()
-  dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
-  dat <- gfranges::scale_predictors(dat, predictors = c(quo(depth)))
-
-  # create grids for each ssid separately
-  ssid <- 1
-  survey_abbrev <- "SYN QCS"
-  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_1$year)
-
-  ssid <- 3
-  survey_abbrev <- "SYN HS"
-  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_3$year)
-
-  ssid <- 4
-  survey_abbrev <- "SYN WCVI"
-  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_4$year)
-
-  ssid <- 16
-  survey_abbrev <- "SYN WCHG"
-  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_16$year)
-
-  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
-    mutate(year = as.numeric(year))
-
-  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
-  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))
-
-  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
-
-  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
 }
@@
 
+  if (region == "All synoptic surveys") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
+  }
+
   sdmTMB::plot_spde(spde)
@@
 
+Make prediction grids for all surveys and years, if not done before.
+```{r, message=FALSE, warning=FALSE}
+try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
+
+if (!exists("nd_all")) {
+
+  # choose base year(s) to create grid from
+  dummy_year <- c(2005, 2006)
+
+  bath <- readRDS("data/bathymetry-data")
+  #dat <- bath$data %>% scale_survey_predictors()
+  dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
+  dat <- gfranges::scale_predictors(dat, predictors = c(quo(depth)))
+
+  # create grids for each ssid separately
+  ssid <- 1
+  survey_abbrev <- "SYN QCS"
+  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
+  unique(nd_1$year)
+
+  ssid <- 3
+  survey_abbrev <- "SYN HS"
+  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
+  unique(nd_3$year)
+
+  ssid <- 4
+  survey_abbrev <- "SYN WCVI"
+  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
+  unique(nd_4$year)
+
+  ssid <- 16
+  survey_abbrev <- "SYN WCHG"
+  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
+  unique(nd_16$year)
+
+  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
+    mutate(year = as.numeric(year))
+
+  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
+  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))
+
+  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
+
+  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
+}
+```
+
 Save predictions for full spatial grid

### 2019-09-10 analysis/VOCC/1-split-biomass-by maturity.Rmd
give splitting my maturity it's own rmd

--- 2019-08-28 analysis/VOCC/1-biomass-models.Rmd
@@
 # species <- "Arrowtooth Flounder"
-#species <- "Pacific Cod"
+# species <- "Petrale Sole"
+# species <- "English Sole"
+# species <- "Dover Sole" # maturity data error for WCHG
+# species <- "Southern Rock Sole"
+# species <- "Flathead Sole"
+
+# species <- "Silvergray Rockfish"
+# species <- "Quillback Rockfish" # summer-birthing
+species <- "Yelloweye Rockfish" # summer-birthing, overfished, 
+# species <- "Bocaccio" # winter-birthing, overfished
+# species <- "Sharpchin Rockfish"
+# species <- "Splitnose Rockfish"
+
+# species <- "Rougheye/Blackspotted Rockfish Complex"
+# species <- "Redbanded Rockfish"
+# species <- "Greenstriped Rockfish"
+# # species <- "Copper Rockfish" # small sample
+# species <- "Darkblotched Rockfish"
+# # species <- "Shortbelly Rockfish" # small sample
+
+# species <- "Pacific Ocean Perch" # schooling
+# species <- "Widow Rockfish" # schooling
+# species <- "Yellowtail Rockfish" # schooling
+# species <- "Canary Rockfish" # schooling, winter-birthing
+
+# species <- "Longspine Thornyhead"
+# species <- "Shortspine Thornyhead"
+
+# species <- "Walleye Pollock"
+# species <- "Pacific Cod"
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish"
 # species <- "Lingcod"
-# species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
-# species <- "Quillback Rockfish" # no maturity classes and WCHG not work
-# species <- "Pacific Ocean Perch" 
- species <- "Yelloweye Rockfish"
+# species <- "Pacific Hake"
+
+# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
+# species <- "Longnose Skate"
+# species <- "Big Skate"
+# species <- "Spotted Ratfish"
+# # species <- "Sandpaper Skate" # small sample
+# # species <- "Brown Cat Shark" # small sample
 ```
@@
 ```{r}
- region <- "All synoptic surveys"
-# region <- "Both odd year surveys"
+# region <- "All synoptic surveys"
+ region <- "Both odd year surveys"
  # region <- "West Coast Vancouver Island"
  # region <- "West Coast Haida Gwaii"
-```
-
-Choose model details
-```{r}
-priors <- FALSE
-# priors <- TRUE
-# covariates <- "+muddy+any_rock"
-covariates <- ""
-covs <- gsub("\\+", "-", "+muddy+any_rock")
-#covs <- gsub("\\+", "-", covariates)
-# covariates <- "+muddy+mixed+rocky"
-# covariates <- "+mixed+rocky"
-# covariates <- "+trawled+muddy+rocky+mixed"
-# covariates <- "+trawled+mixed+rocky"
-# covariates <- "+trawled+mixed"
 ```
@@
 
-Make prediction grids for all surveys and years, if not done before.
-```{r, message=FALSE, warning=FALSE}
-try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
-
-if (!exists("nd_all")) {
-
-  # choose base year(s) to create grid from
-  dummy_year <- c(2005, 2006)
-
-  bath <- readRDS("data/bathymetry-data")
-  #dat <- bath$data %>% scale_survey_predictors()
-  dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
-  dat <- gfranges::scale_predictors(dat, predictors = c(quo(depth)))
-
-  # create grids for each ssid separately
-  ssid <- 1
-  survey_abbrev <- "SYN QCS"
-  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_1$year)
-
-  ssid <- 3
-  survey_abbrev <- "SYN HS"
-  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_3$year)
-
-  ssid <- 4
-  survey_abbrev <- "SYN WCVI"
-  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_4$year)
-
-  ssid <- 16
-  survey_abbrev <- "SYN WCHG"
-  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_16$year)
-
-  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
-    mutate(year = as.numeric(year))
-
-  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
-  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))
-
-  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
-
-  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
-}
-```
-
 # Split biomass by maturity
@@
   )
+  }
   maturity$data <- filter(maturity$data, ssid == model_ssid) 
-  }
+  # %>% select(-adult_density, -imm_density) # if maturity model is deemed unuseful
 
@@
 ```
-
-# Build spatiotemporal density model
-Load and filter data 
-```{r}
-biomass <- readRDS(paste0(
-    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
-    ))
-  covars <- readRDS("data/event-covariates.rds")
-  data <- dplyr::left_join(biomass, covars)
-
-  # scale predictors before filtering to ensure mean and SD are global
-  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
-  data <- gfranges::scale_predictors(data, predictors = c(quo(depth)))
-       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
-  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)
-```
-
-Run climate independent sdmTMB model
-```{r}
-
-  if (region == "Both odd year surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
-  }
-
-  if (region == "West Coast Vancouver Island") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-  }
-
-  if (region == "West Coast Haida Gwaii") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-  }
-
-  sdmTMB::plot_spde(spde)
-
-  if (any(names(data) == "adult_density")) {
-
-    adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year)", covariates, ""
-      ))
-
-    starttime1 <- Sys.time()
-    adult_biomass <- sdmTMB::sdmTMB(data,
-      adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime1 <- Sys.time()
-    time1 <- round(starttime1 - endtime1)
-
-    saveRDS(adult_biomass, 
-      file = paste0("data/", spp, 
-        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-        ))
-
-    imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year)", covariates, ""
-      ))
-
-    starttime2 <- Sys.time()
-    imm_biomass <- sdmTMB::sdmTMB(data, 
-      imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = FALSE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime2 <- Sys.time()
-    time2 <- round(starttime2 - endtime2)
-
-    saveRDS(imm_biomass, 
-     file = paste0("data/", spp, 
-       "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-       ))
-
-  } else {
-
-    starttime3 <- Sys.time()
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
-
-    total_biomass <- sdmTMB::sdmTMB(data,
-      dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime3 <- Sys.time()
-    time3 <- round(starttime3 - endtime3)
-
-    saveRDS(total_biomass, file = paste0("data/", spp, 
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-      ))
-  }
-```
-
-Save predictions for full spatial grid
-
-```{r}
-#covs <- gsub("\\+", "-", covariates)
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
-
-rm(adult_predictions)
-rm(imm_predictions)
-rm(predictions)
-
-if (length(unique(fish$maturity_code)) > 2) {
-
-    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-    adult_biomass <- readRDS(paste0("data/", spp, 
-      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
-
-    nd <- nd_all %>% 
-      filter(ssid %in% model_ssid) %>% 
-      filter(year %in% unique(adult_biomass$data$year))
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    adult_predictions <- predict(adult_biomass, newdata = nd)
-    saveRDS(adult_predictions, file = paste0("data/", spp, 
-      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
-    ))
-
-    imm_biomass <- readRDS(paste0("data/", spp, 
-      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
-    imm_predictions <- predict(imm_biomass, newdata = nd)
-    saveRDS(imm_predictions, file = paste0("data/", spp, 
-      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
-    ))
-  } else {
-    rm(adult_predictions)
-    rm(imm_predictions)
-
-    if (!exists(predictions)) {
-      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-      nd <- nd_all %>% filter(ssid %in% model_ssid)
-      nd <- na.omit(nd)
-      nd$year <- as.integer(nd$year)
-
-      total_biomass <- readRDS(paste0("data/", spp, 
-        "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-      ))
-      predictions <- predict(total_biomass, newdata = nd)
-      saveRDS(predictions, file = paste0("data/", spp, 
-        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
-      ))
-    }
-  }
-```

### 2019-09-10 analysis/VOCC/2-biomass-depth-only-model.Rmd
climate independent biomass model that converges most relaibly is global with trawled and ssid as fixed effects

--- 2019-08-28 analysis/VOCC/1-biomass-models.Rmd
@@
 ---
-title: "Spatiotemporal models of species biomass split by maturity class"
+title: "Spatiotemporal models of species biomass with time-varying depth and trawl footprint, and optional substrate variables"
 author: "Philina English"
@@
 # # Species run so far...
-# species <- "Arrowtooth Flounder"
-#species <- "Pacific Cod"
-# species <- "Sablefish"
-# species <- "Silvergray Rockfish"
-# species <- "Lingcod"
-# species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
-# species <- "Quillback Rockfish" # no maturity classes and WCHG not work
-# species <- "Pacific Ocean Perch" 
- species <- "Yelloweye Rockfish"
+# species <- "Arrowtooth Flounder" # temp good, not DO, imm moving deeper?
+
+# species <- "Petrale Sole"
+# species <- "English Sole"
+# species <- "Dover Sole"
+# species <- "Southern Rock Sole"
+
+# species <- "Walleye Pollock" # survey 4 or 16 might work
+# species <- "Pacific Cod"
+
+## Sebastes
+
+## TRY BOTH - not schooling
+ # species <- "Yelloweye Rockfish" # beautiful curves, with depth change ***
+ #  species <- "Bocaccio" # good curves, no depth change ... most survey-specific models fail to find minimum ***
+#  species <- "Sharpchin Rockfish" # beautiful curves, imm depth change ***
+#  species <- "Greenstriped Rockfish" # beautiful curves for adults, no depth change ***
+
+# ## TRY BOTH - schooling
+#  species <- "Pacific Ocean Perch" # beautiful curves, no depth change, schooling ***
+#  species <- "Silvergray Rockfish" # beautiful curves, no depth change, schooling ***
+
+## TRY JUST TEMPERATURE 
+# species <- "Rougheye/Blackspotted Rockfish Complex" # good temp, not DO
+# species <- "Redbanded Rockfish" # good temp, not DO, maybe moving shallower
+ species <- "Splitnose Rockfish" # good temp, survey 4 DO ok
+
+# species <- "Yellowtail Rockfish" #  good temp for survey 4 only, no DO, schooling
+# species <- "Quillback Rockfish" # like warm = overall a bit flat... territorial! 
+# species <- "Canary Rockfish" # schooling, winter-birthing, and depth change!, but curves at edges... adult1 looks ok?
+#  species <- "Widow Rockfish" # schooling, and depth change, but curves at edges
+
+# species <- "Longspine Thornyhead" # too deep to produce curves
+# species <- "Shortspine Thornyhead"
+# species <- "Sablefish" # temp might work, no DO, no consistant depth change
+# species <- "Lingcod" # curves ok, overall a bit flat
+# # species <- "Pacific Hake" # not useful
+
+# species <- "North Pacific Spiny Dogfish" # too flat, but with depth change! note: pooled maturity 
+# species <- "Longnose Skate" # do too flat, temp from 4 might be useful, seems to go deeper?
+# species <- "Big Skate"
+# species <- "Sandpaper Skate"
+# species <- "Brown Cat Shark"
+# species <- "Spotted Ratfish" # curves not useful because to abundant in shallow
 ```
@@
 # priors <- TRUE
+covariates <- "+trawled+as.factor(ssid)"
+ covs <- "-trawled-ssid"
+
 # covariates <- "+muddy+any_rock"
-covariates <- ""
-covs <- gsub("\\+", "-", "+muddy+any_rock")
-#covs <- gsub("\\+", "-", covariates)
 # covariates <- "+muddy+mixed+rocky"
 # covariates <- "+mixed+rocky"
-# covariates <- "+trawled+muddy+rocky+mixed"
-# covariates <- "+trawled+mixed+rocky"
-# covariates <- "+trawled+mixed"
+# covariates <- "+rocky"
+# covariates <- "+mixed_scaled+rocky_scaled+muddy_scaled"
+
+# covs <- gsub("\\+", "-", covariates)
+# covs <- gsub("\\+", "-", "+muddy+any_rock")
+# covs <- gsub("\\+", "-", "-variable-substrate")
+# covs <- gsub("\\+", "-", "-all-substrates")
 ```
@@
 
+# Build spatiotemporal density model
+Load and filter data 
+```{r}
+biomass <- readRDS(paste0(
+    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
+    ))
+  covars <- readRDS("data/event-covariates.rds")
+  data <- dplyr::left_join(biomass, covars)
+
+  # scale predictors before filtering to ensure mean and SD are global
+  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
+  data <- gfranges::scale_predictors(data, #predictors = c(quo(depth)))
+        predictors = c(quo(depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
+  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)
+```
+
+Make mesh
+```{r}
+  if (region == "Both odd year surveys") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
+  }
+
+  if (region == "West Coast Vancouver Island") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
+  }
+
+  if (region == "West Coast Haida Gwaii") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
+  }
+
+  if (region == "All synoptic surveys") {
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
+  }
+
+  sdmTMB::plot_spde(spde)
+```
+
+Run climate independent sdmTMB model
+```{r}
+
+  if (any(names(data) == "adult_density")) {
+
+    adult_formula <- as.formula(paste(
+      "adult_density ~ 0 + as.factor(year)", covariates, ""
+      ))
+
+    starttime1 <- Sys.time()
+    adult_biomass <- sdmTMB::sdmTMB(data,
+      adult_formula,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2 , #+ trawled
+      time = "year", spde = spde,
+      family = tweedie(link = "log"),
+      ar1_fields = TRUE,
+      include_spatial = TRUE,
+      enable_priors = priors,
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      silent = FALSE
+    )
+    endtime1 <- Sys.time()
+    time1 <- round(starttime1 - endtime1)
+
+    saveRDS(adult_biomass, 
+      file = paste0("data/", spp, 
+        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+        ))
+
+    imm_formula <- as.formula(paste(
+      "imm_density ~ 0 + as.factor(year)", covariates, ""
+      ))
+
+    starttime2 <- Sys.time()
+    imm_biomass <- sdmTMB::sdmTMB(data, 
+      imm_formula,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
+      time = "year", spde = spde,
+      family = tweedie(link = "log"),
+      ar1_fields = FALSE,
+      include_spatial = TRUE,
+      enable_priors = priors,
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      silent = FALSE
+    )
+    endtime2 <- Sys.time()
+    time2 <- round(starttime2 - endtime2)
+
+    saveRDS(imm_biomass, 
+     file = paste0("data/", spp, 
+       "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+       ))
+
+  } else {
+
+    starttime3 <- Sys.time()
+    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
+
+    total_biomass <- sdmTMB::sdmTMB(data,
+      dens_formula,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2,# + trawled
+      time = "year", spde = spde,
+      family = tweedie(link = "log"),
+      ar1_fields = TRUE,
+      include_spatial = TRUE,
+      enable_priors = priors,
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      silent = FALSE
+    )
+    endtime3 <- Sys.time()
+    time3 <- round(starttime3 - endtime3)
+
+    saveRDS(total_biomass, file = paste0("data/", spp, 
+      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+      ))
+  }
+```
+
 Make prediction grids for all surveys and years, if not done before.
@@
   saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
-}
-```
-
-# Split biomass by maturity
-
-#### Calculate maturity weighted biomass if not done before
-```{r}
- rm(biomass)
- rm(maturity)
-# try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
-
-if (!exists("biomass")) {
-  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
-  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
-  bath <- readRDS("data/bathymetry-data")
-  # for POP: exclude sample with unusually large fish 
-  # sizes are impossible given recorded catch weight
-  # fish <- fish[fish$fishing_event_id!=1506954,]
-
- try({ 
-   maturity <- split_catch_maturity(survey_sets, fish, bath,
-    survey = survey,
-    years = years,
-    cutoff_quantile = c(.9995),
-    plot = TRUE)
-   })
-
-  if(is.null(maturity$maturity)) {
-  maturity <- split_catch_maturity(survey_sets, fish, bath,
-    survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
-    years = years,
-    cutoff_quantile = c(.9995),
-    plot = TRUE
-  )
-  maturity$data <- filter(maturity$data, ssid == model_ssid) 
-  }
-
-  saveRDS(maturity$data, file = paste0("data/", spp, 
-    "/data-by-maturity-", spp, "-", ssid_string, ".rds"
-    ))
-}
-```
-
-#### Save length~weight plots
-```{r}
-if (exists("maturity")) {
-  print(maturity$mass_model)
-  ggsave(
-    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
-    dpi = 600,
-    width = 8, 
-    height = 10
-  )
-}
-```
-
-#### Save length at maturity plot 
-```{r}
-if (exists("maturity")) {
-  png(
-    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), 
-    res = 600,
-    units = "in",
-    width = 8, 
-    height = 10
-  )
-  print(maturity$maturity)
-  dev.off()
-}
-
-try(print(maturity$maturity))
-```
-
-# Build spatiotemporal density model
-Load and filter data 
-```{r}
-biomass <- readRDS(paste0(
-    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
-    ))
-  covars <- readRDS("data/event-covariates.rds")
-  data <- dplyr::left_join(biomass, covars)
-
-  # scale predictors before filtering to ensure mean and SD are global
-  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
-  data <- gfranges::scale_predictors(data, predictors = c(quo(depth)))
-       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
-  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)
-```
-
-Run climate independent sdmTMB model
-```{r}
-
-  if (region == "Both odd year surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
-  }
-
-  if (region == "West Coast Vancouver Island") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-  }
-
-  if (region == "West Coast Haida Gwaii") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-  }
-
-  sdmTMB::plot_spde(spde)
-
-  if (any(names(data) == "adult_density")) {
-
-    adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year)", covariates, ""
-      ))
-
-    starttime1 <- Sys.time()
-    adult_biomass <- sdmTMB::sdmTMB(data,
-      adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime1 <- Sys.time()
-    time1 <- round(starttime1 - endtime1)
-
-    saveRDS(adult_biomass, 
-      file = paste0("data/", spp, 
-        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-        ))
-
-    imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year)", covariates, ""
-      ))
-
-    starttime2 <- Sys.time()
-    imm_biomass <- sdmTMB::sdmTMB(data, 
-      imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = FALSE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime2 <- Sys.time()
-    time2 <- round(starttime2 - endtime2)
-
-    saveRDS(imm_biomass, 
-     file = paste0("data/", spp, 
-       "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-       ))
-
-  } else {
-
-    starttime3 <- Sys.time()
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
-
-    total_biomass <- sdmTMB::sdmTMB(data,
-      dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime3 <- Sys.time()
-    time3 <- round(starttime3 - endtime3)
-
-    saveRDS(total_biomass, file = paste0("data/", spp, 
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-      ))
 }

### 2019-09-10 analysis/VOCC/2-biomass-trends.Rmd
climate independent biomass model that converges most relaibly is global with trawled and ssid as fixed effects

@@
 ```{r}
+
 # # Species run so far...
- species <- "Arrowtooth Flounder"
+# species <- "Arrowtooth Flounder" # temp good, not DO, imm moving deeper?
+
+# species <- "Petrale Sole"
+# species <- "English Sole"
+# species <- "Dover Sole"
+# species <- "Southern Rock Sole"
+
+# species <- "Walleye Pollock" # survey 4 or 16 might work
 # species <- "Pacific Cod"
-# species <- "Sablefish"
-#  species <- "Silvergray Rockfish"
-# species <- "Lingcod"
-#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Quillback Rockfish"
-# species <- "Pacific Ocean Perch"
-# species <- "Yelloweye Rockfish"
+
+## Sebastes
+
+## TRY BOTH - not schooling
+ # species <- "Yelloweye Rockfish" # beautiful curves, with depth change ***
+ # species <- "Bocaccio" # good curves, no depth change ... most survey-specific models fail to find minimum ***
+#  species <- "Sharpchin Rockfish" # beautiful curves, imm depth change ***
+#  species <- "Greenstriped Rockfish" # beautiful curves for adults, no depth change ***
+# 
+# ## TRY BOTH - schooling
+#  species <- "Pacific Ocean Perch" # beautiful curves, no depth change, schooling ***
+#  species <- "Silvergray Rockfish" # beautiful curves, no depth change, schooling ***
+
+## TRY JUST TEMPERATURE 
+# species <- "Rougheye/Blackspotted Rockfish Complex" # good temp, not DO
+# species <- "Redbanded Rockfish" # good temp, not DO, maybe moving shallower
+# species <- "Splitnose Rockfish" # good temp, not DO
+
+# species <- "Yellowtail Rockfish" #  good temp for survey 4 only, no DO, schooling
+# species <- "Quillback Rockfish" # like warm = overall a bit flat... territorial! 
+# species <- "Canary Rockfish" # schooling, winter-birthing, and depth change!, but curves at edges... adult1 looks ok?
+#  species <- "Widow Rockfish" # schooling, and depth change, but curves at edges
+
+# species <- "Longspine Thornyhead" # too deep to produce curves
+# species <- "Shortspine Thornyhead"
+# species <- "Sablefish" # temp might work, no DO, no consistant depth change
+# species <- "Lingcod" # curves ok, overall a bit flat
+# # species <- "Pacific Hake" # not useful
+
+# species <- "North Pacific Spiny Dogfish" # too flat, but with depth change! note: pooled maturity 
+# species <- "Longnose Skate" # do too flat, temp from 4 might be useful, seems to go deeper?
+# species <- "Big Skate"
+# species <- "Sandpaper Skate"
+# species <- "Brown Cat Shark"
+# species <- "Spotted Ratfish" # curves not useful because to abundant in shallow
 ```
@@
 # priors <- TRUE
-
+covariates <- ""
+covariates <- "+trawled"
+covs <- gsub("\\+", "-", covariates)
+covariates <- "+trawled+as.factor(ssid)"
+ covs <- gsub("\\+", "-", "+trawled+ssid")
 # covariates <- "+muddy+mixed+rocky"
 # covariates <- "+mixed+rocky"
-covariates <- "+muddy+any_rock"
+# covariates <- "+muddy+any_rock"
 # covariates <- "+trawled+muddy+rocky+mixed"
@@
 } else {
-    # Bind together predictions from different models
-    try({
-      pred1n3 <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"))
-      pred4 <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"))
-      pred16 <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"))
-    })
+
+   try({adult_predictions <- readRDS(paste0("data/", spp, 
+      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
+    ))})
+   try({imm_predictions <- readRDS(paste0("data/", spp, 
+      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
+    ))})
+
+  if (!exists("adult_predictions")) {
+    # OR bind together predictions from different models
+   try({pred1n3 <- readRDS(paste0("data/", spp, 
+        "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"))})
+   try({pred4 <- readRDS(paste0("data/", spp, 
+        "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"))}) 
+   try({pred16 <- readRDS(paste0("data/", spp, 
+        "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"))})
+
+    if (!exists("pred1n3")) pred1n3 <- NULL
     if (!exists("pred16")) pred16 <- NULL
@@
       pred1n3_imm <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"))
+        "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"))})
+   try({
       pred4_imm <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))
+        "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))})
+   try({
       pred16_imm <- readRDS(paste0("data/", spp, 
@@
 
+    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
     if (!exists("pred16_imm")) pred16_imm <- NULL
@@
 
+      try({predictions <- readRDS(paste0("data/", spp, 
+        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
+      ))})
+
+      if (!exists("predictions")) {
+
+      try({pred1n3_all <- readRDS(paste0("data/", spp, 
+        "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))})
+      try({pred4_all <- readRDS(paste0("data/", spp, 
+        "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))})
+      try({pred16_all <- readRDS(paste0("data/", spp, 
+        "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))})
       try({
-      pred1n3_all <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))
-      pred4_all <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))
-      pred16_all <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))
       predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
@@
       if (!exists("predictions")) paste("Total biomass predictions are not available.")
+      }
+    }
   }
@@
 
+  model_ssid <- unique(adult_predictions$ssid)
+
 } else {
@@
       "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
+
+  model_ssid <- unique(predictions$ssid)
 }
@@
   if (16 %in% model_ssid) {
+    if (any(adult_predictions$ssid == 16)) {
     adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
+    } else {
+    adult_predictions16 <- NULL
+    }
   } else {
@@
 ```{r }
-#FIXME: Attempting to plot of current mature biomass with trawl footprint...
-# if (exists("out_mature")) {
+if (exists("out_mature")) {
 
@@
 current_biomass
-# }
+}
 ```

### 2019-09-10 analysis/VOCC/3-fixed-var-response-plots.Rmd
add more species

@@
 # # Species run so far...
- species <- "Arrowtooth Flounder" # running more models
+# species <- "Arrowtooth Flounder" # temp good, not DO, imm moving deeper?
 
@@
 
- species <- "Walleye Pollock" # running more models
+# species <- "Walleye Pollock" # survey 4 or 16 might work
 # species <- "Pacific Cod"
-# species <- "Pacific Ocean Perch" # beautiful curves, no depth change
 
-# species <- "Silvergray Rockfish" # beautiful curves, no depth change
- species <- "Quillback Rockfish" # overall a bit flat... running survey specific models
-# species <- "Yelloweye Rockfish" # beautiful curves, with depth change
- species <- "Bocaccio" # good curves, no depth change ... should run survey specific models
- species <- "Sharpchin Rockfish"
-# species <- "Splitnose Rockfish"
-# species <- "Canary Rockfish"
-# species <- "Rougheye/Blackspotted Rockfish Complex"
-# species <- "Redbanded Rockfish"
-# species <- "Greenstriped Rockfish"
+## Sebastes
+# 
+# ## TRY BOTH - not schooling
+#  species <- "Yelloweye Rockfish" # beautiful curves, with depth change ***
+#  species <- "Bocaccio" # good curves, no depth change ... most survey-specific models fail to find minimum ***
+#  species <- "Sharpchin Rockfish" # beautiful curves, imm depth change ***
+#  species <- "Greenstriped Rockfish" # beautiful curves for adults, no depth change ***
 
-# species <- "Longspine Thornyhead"
-# species <- "Shortspine Thornyhead",
-# species <- "Sablefish"
-# species <- "Lingcod"
+# ## TRY BOTH - schooling
+#  species <- "Pacific Ocean Perch" # beautiful curves, no depth change, schooling ***
+#  species <- "Silvergray Rockfish" # beautiful curves, no depth change, schooling ***
 
-# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Longnose Skate"
-#  species <- "Spotted Ratfish"
+# ## TRY JUST TEMPERATURE 
+# # species <- "Rougheye/Blackspotted Rockfish Complex" # good temp, not DO
+# # species <- "Redbanded Rockfish" # good temp, not DO, maybe moving shallower
+#  species <- "Splitnose Rockfish" # good temp, survey 4 DO ok
+
+# # species <- "Yellowtail Rockfish" #  good temp for survey 4 only, no DO, schooling
+# # species <- "Quillback Rockfish" # like warm = overall a bit flat... territorial! 
+# # species <- "Canary Rockfish" # schooling, winter-birthing, and depth change!, but curves at edges... adult1 looks ok?
+#  species <- "Widow Rockfish" # schooling, and depth change, but curves at edges
+
+# species <- "Longspine Thornyhead" # too deep to produce curves
+# species <- "Shortspine Thornyhead"
+# species <- "Sablefish" # temp might work, no DO, no consistant depth change
+# species <- "Lingcod" # curves ok, overall a bit flat
+# # species <- "Pacific Hake" # not useful
+
+# species <- "North Pacific Spiny Dogfish" # too flat, but with depth change! note: pooled maturity 
+# species <- "Longnose Skate" # do too flat, temp from 4 might be useful, seems to go deeper?
+# species <- "Big Skate"
+# species <- "Sandpaper Skate"
+# species <- "Brown Cat Shark"
+# species <- "Spotted Ratfish" # curves not useful because to abundant in shallow
 ```
@@
 # covariates <- "+trawled+mixed"
+# 
+# covariates <- "+as.factor(ssid)"
+# covs <- "-both-tv-depth-ssid"
 
@@
 covs <- "-both-tv-depth"
+
 num_params <- 2
@@
    do[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
+   do[[i]]$ssid <- unique(model_list[[i]]$data$ssid)
    do[[i]]$model <- names(model_list[i])
@@
     td[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
+    td[[i]]$ssid <- unique(model_list[[i]]$data$ssid)
     td[[i]]$model <- names(model_list[i])
@@
   grobs = do_plots,
-  ncol = 2,
+  nrow = 4,
   top = grid::textGrob(paste(species))
@@
   grobs = temp_plots,
-  ncol = 2,
+  nrow = 4,
   top = grid::textGrob(paste(species))
@@
 gridExtra::grid.arrange(grobs = c(raw_do_plots, raw_t_plots),
-  ncol = 2,
+  nrow = 4,
   top = grid::textGrob(paste("Raw densities of", species, "with modeled optimals in vertical lines (if solid includes all surveys)"))
@@
   grobs = depth_plots,
-  ncol = 2,
+  nrow = 4,
   top = grid::textGrob(paste(species))

### 2019-09-10 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
apply changes to vector trimming; save plots of resulting vectors; save dataframes of matched cells to separte mulitspecies folders grouped based on vocc thresholds

@@
 
-Modified to allow optimal temperature to vary by maturity class. 
+Modified to allow thresholds chosen to vary by maturity class. 
 Therefore need to run this file separately for each species-class combination.
@@
 ```{r echo=TRUE}
-# # Species run so far...
- species <- "Arrowtooth Flounder"
-# species <- "Pacific Cod"
-# species <- "Sablefish"
-# species <- "Silvergray Rockfish"
-# species <- "Lingcod"
-# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Quillback Rockfish"
-# species <- "Pacific Ocean Perch"
-# species <- "Yelloweye Rockfish"
+
 ```
@@
 ```{r echo=TRUE}
+biomass_threshold <- "50" # filter for climate values predicted to support 50% of max biomass 
+ biomass_threshold <- "25" # filter for climate values predicted to support 50% of max biomass 
+global_only <- TRUE
+
  climate <- "temperature"
- temp_threshold <- c(0.5)
+ temp_threshold <- c(0.25)
 
+# temp_threshold <- c(0.5)
+# temp_threshold <- c(0.75)
 #  climate <- "do"
-# do_threshold <- c(0.5)
-
+# #  do_threshold <- c(0.5)
+#  do_threshold <- c(0.25)
 # climate <- "do and temperature"
@@
 
-Run all subsequent code...
 ```{r, include=FALSE}
@@
 
-params <- readRDS(paste0("data/optimal-values-by-species-by-age.rds"))
+params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
+
+params <- params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
 
@@
   params <- filter(params, age != "imm")
+  if (any(params$model_type == "split")) { params <- filter(params, model_type != "total") }
 }
@@
 
+if (global_only) {
+ params <- params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
+ params <- filter(params, global == "T")
+} else {
+ #params <- filter(params, ssid %in% model_ssid)
+}
+
 if (climate == "temperature") {
-  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
-  optimal_temps <- na.omit(optimal_temps)
-  optimal_temp <- round(optimal_temps[1], digits = 2)
+  # optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
+  # optimal_temps <- na.omit(optimal_temps)
+  # optimal_temp <- round(optimal_temps[1], digits = 2)
+  # optimal_values <- optimal_temp
+
+  if (biomass_threshold == "50") {
+  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_50 }
+  if (biomass_threshold == "25") {
+  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_25 }
+  .upper <- na.omit(.upper)
+
+  upper_thresholds <- round(.upper[1], digits = 2)
+  lower_thresholds <- NA
+
   variable_names <- c("temp")
-  min_thresholds <- c(Inf)
-  optimal_values <- optimal_temp
-  max_thresholds <- temp_threshold
+
+  lower_change <- c(Inf)
+  upper_change <- temp_threshold
   indices <- c(1, 2)
@@
 if (climate == "do") {
-  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
-  optimal_dos <- na.omit(optimal_dos)
-  optimal_do <- round(optimal_dos[1], digits = 2)
+  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
+  # optimal_dos <- na.omit(optimal_dos)
+  # optimal_do <- round(optimal_dos[1], digits = 2)
+  # optimal_values <- optimal_do
+  params <- filter(params, lower_do_25 > 0.5)
+
+  if (biomass_threshold == "50") {
+  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_50 }
+  if (biomass_threshold == "25") {
+  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_25 }
+
+  .lower <- na.omit(.lower)
+  lower_thresholds <- round(.lower[1], digits = 2)
+  upper_thresholds <- NA
   variable_names <- c("do_est")
-  max_thresholds <- c(Inf)
-  optimal_values <- optimal_do
-  min_thresholds <- do_threshold
+
+  lower_change <- do_threshold
+  upper_change <- c(Inf)
   indices <- c(1, 2)
@@
 if (climate == "do and temperature") {
-  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
-  optimal_temps <- na.omit(optimal_temps)
-  optimal_temp <- round(optimal_temps[1], digits = 2)
-  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
-  optimal_dos <- na.omit(optimal_dos)
-  optimal_do <- round(optimal_dos[1], digits = 2)
-  optimal_values <- c(optimal_do, optimal_temp)
+
+  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
+  # optimal_dos <- na.omit(optimal_dos)
+  # optimal_do <- round(optimal_dos[1], digits = 2)
+  # optimal_values <- optimal_do
+
+  params <- filter(params, lower_do_25 > 0.5)
+  if (biomass_threshold == "50") {
+  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_50 }
+  if (biomass_threshold == "25") {
+  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_25 }
+  .lower <- na.omit(.lower)
+  lower_thresholds <- round(.lower[1], digits = 2)
+
+  # optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
+  # optimal_temps <- na.omit(optimal_temps)
+  # optimal_temp <- round(optimal_temps[1], digits = 2)
+  # optimal_values <- optimal_temp
+
+  if (biomass_threshold == "50") {
+  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_50 }
+  if (biomass_threshold == "25") {
+  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_25 }
+  .upper <- na.omit(.upper)
+  upper_thresholds <- round(.upper[1], digits = 2)
+
   variable_names <- c("do_est", "temp")
-  min_thresholds <- c(do_threshold, Inf)
-  max_thresholds <- c(Inf, temp_threshold)
+  lower_thresholds <- c(lower_thresholds, NA)
+  upper_thresholds <- c(NA, upper_thresholds)
+  lower_change <- c(do_threshold, Inf)
+  upper_change <- c(Inf, temp_threshold)
   indices <- c(1, 2)
@@
 
-min_string <- paste0(min_thresholds, collapse = "-")
-max_string <- paste0(max_thresholds, collapse = "-")
+trim_thresholds <- c(lower_thresholds, upper_thresholds)
+min_string <- paste0(lower_change, collapse = "-")
+max_string <- paste0(upper_change, collapse = "-")
 
-optimal_string <- paste0(optimal_values, collapse = "-")
+optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
 climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
+
+glimpse(params)
 ```
+
+# Run all subsequent code...
 
@@
     indices = indices,
-    min_thresholds = min_thresholds,
-    max_thresholds = max_thresholds,
+    min_thresholds = lower_change,
+    max_thresholds = upper_change,
     round_fact = 10
@@
 
-  vocc$var_1_min <- min_thresholds[1]
-  vocc$var_2_min <- min_thresholds[2]
-  vocc$var_1_max <- max_thresholds[1]
-  vocc$var_2_max <- max_thresholds[2]
+  vocc$var_1_min <- lower_change[1]
+  vocc$var_2_min <- lower_change[2]
+  vocc$var_1_max <- upper_change[1]
+  vocc$var_2_max <- upper_change[2]
 
@@
 
-vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds,
+vectors <- trim_vector_data(vocc, variable_names, 
+  lower_change, upper_change, 
+  lower_thresholds, upper_thresholds,
   cell_size = input_cell_size, dist_intercept = dist_intercept,
@@
 ```
+
+```{r}
+png(
+  file = paste0("figs/", spp,
+    "/vocc-", min_string, "to", max_string, "-", 
+    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
+  ),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 11
+)
+gridExtra::grid.arrange(
+  grobs = trimmed_vector_plots,
+  ncol = 2,
+  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
+)
+dev.off()
+```
+
 # Add species biomass data
@@
 priors <- FALSE
-covariates <- "+muddy+any_rock"
-covs <- gsub("\\+", "-", covariates)
+# covariates <- "+muddy+any_rock"
+# covs <- gsub("\\+", "-", covariates)
+covs <- gsub("\\+", "-", "-trawled+ssid")
 
@@
 # Select matched control cells
+
 ```{r}
@@
 
-match_formula <- paste0("treat ~ 
-  muddy.st + any_rock.st + x + y") #depth.st + trawled + omega_s + 
+#FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?
+
+match_formula <- paste0("treat ~ trawled + x + y") 
+  # depth.st + omega_s + muddy.st + any_rock.st + 
   # for (j in (variable_names)) {
@@
     distance = "probit",
-    discard="both",
+    discard = "both",
     data = alldata)
@@
 if (!exists("vocc_by_sp")){
-match_plots <- apply(matched, function(matched){  
+match_plots <- plot(matched, interactive = F)
 summary(matched)
-  plot(matched)
- })
-match_plots
 }
+
 ```
@@
   for (j in seq_along(variable_names)) {
-    climate <- paste0(variable_names[j])
+    climate_var <- paste0(variable_names[j])
     matched_time0 <- matched_time0 %>% 
-      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
+      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
     matched_time1 <- matched_time1 %>% 
-      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
+      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
    }
@@
 saveRDS(vocc_by_sp, file = paste0(
-    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
+    "data/", spp,
+    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
     min_string, "to", max_string, "-multiyear-all.rds"
@@
 
+## Save spp dataframe
+```{r}
+#biomass_threshold <- "50"
+optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
+#climate <- "temperature"
+try({
+  vocc_by_sp <- readRDS(paste0(
+    "data/", spp,
+    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
+    min_string, "to", max_string, "-multiyear-all.rds"
+  ))
+})
+
+n <- nrow(vocc_by_sp[[1]])
+species <- rep(species, n) 
+data <- vocc_by_sp[[1]]
+df <- cbind(species, data)
+species <- species[[1]]
+
+# folder to hold vocc data
+biomass_threshold_string <- paste0("perc_", biomass_threshold)
+
+dir.create(file.path("data/_all", climate_string, biomass_threshold_string))
+
+if (climate=="temperature") dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))
+
+if (climate=="temperature") {
+  saveRDS(df, file = paste0(
+    "data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", max_string,"/vocc-", climate_string, "-", 
+     max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
+  ))
+}
+
+if (climate=="do") {
+  dir.create(file.path("data/_all", climate_string))  
+  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
+  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string))  
+  }
+if (climate=="do") {
+  saveRDS(df, file = paste0(
+    "data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", min_string,"/vocc-", climate_string, "-",
+    min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
+  ))
+}
+
+if (climate=="do and temperature") {
+  dir.create(file.path("data/_all", climate_string))  
+  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
+}
+
+if (climate=="do and temperature") {
+  saveRDS(df, file = paste0(
+    "data/_all/", climate_string, "/", "perc_", biomass_threshold,"/vocc-", climate_string, "-",
+    min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
+  ))
+}
+
+# View(data)
+```
+
 ```{r}
@@
   vocc_by_sp <- readRDS(paste0(
-    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
+    "data/", spp,
+    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
     min_string, "to", max_string, "-multiyear-all.rds"
@@
 })
-data <- vocc_by_sp [[1]]
-# View(data)
+data <- vocc_by_sp[[1]]
 ```
@@
 ```{r}
+#FIXME: SAVE PLOTS?
+
 dat_summary <- data %>% 
@@
 
+```{r}
+# png(
+#   file = paste0("figs/", spp,
+#     "/matches-", min_string, "to", max_string, "-", 
+#     climate_string, "-", spp, "-", age, "-", ssid_string, "-", 
+#     climate_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
+#   ),
+#   res = 600,
+#   units = "in",
+#   width = 8.5,
+#   height = 11
+# )
+# gridExtra::grid.arrange(
+#   grobs = ,
+#   ncol = 2,
+#   top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
+# )
+# dev.off()
+```
+```
+
 ### Basic BACI
@@
 
-for (j in (variable_names)) {
- gamm_formula <- paste0(gamm_formula, "+ s(std_", j, ")")
- }
+# for (j in (variable_names)) {
+#  gamm_formula <- paste0(gamm_formula, "+ s(std_", j, ")")
+#  }
 

### 2019-09-11 analysis/VOCC/1-biomass-do-models-fixed-both.Rmd
log DO and filter DO outliers

--- 2019-08-29 analysis/VOCC/1-biomass-do-models-fixed-both.Rmd
@@
 ```{r}
-species <- params$species
+# species <- params$species
 
@@
 # species <- "Arrowtooth Flounder"
+# species <- "Petrale Sole"
+# species <- "English Sole"
+# species <- "Dover Sole" # maturity data error for WCHG
+# species <- "Southern Rock Sole"
+# species <- "Flathead Sole"
+
+# species <- "Silvergray Rockfish"
+# species <- "Quillback Rockfish" # summer-birthing
+# species <- "Yelloweye Rockfish" # summer-birthing, overfished, 
+# species <- "Bocaccio" # winter-birthing, overfished
+# species <- "Sharpchin Rockfish"
+# species <- "Splitnose Rockfish"
+
+# species <- "Rougheye/Blackspotted Rockfish Complex"
+# species <- "Redbanded Rockfish"
+# species <- "Greenstriped Rockfish"
+# # species <- "Copper Rockfish" # small sample
+# species <- "Darkblotched Rockfish"
+# # species <- "Shortbelly Rockfish" # small sample
+
+# species <- "Pacific Ocean Perch" # schooling
+# species <- "Widow Rockfish" # schooling
+# species <- "Yellowtail Rockfish" # schooling
+# species <- "Canary Rockfish" # schooling, winter-birthing
+
+# species <- "Longspine Thornyhead"
+# species <- "Shortspine Thornyhead"
+
+# species <- "Walleye Pollock"
 # species <- "Pacific Cod"
 # species <- "Sablefish"
-# species <- "Silvergray Rockfish"
 # species <- "Lingcod"
+# species <- "Pacific Hake"
+
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
- species <- "Quillback Rockfish"
-# species <- "Pacific Ocean Perch"
-# species <- "Yelloweye Rockfish"
+# species <- "Longnose Skate"
+# species <- "Big Skate"
+# species <- "Spotted Ratfish"
+# species <- "Sandpaper Skate"
+# species <- "Brown Cat Shark"
 ```
@@
  covariates <- ""
+# covariates <- "+trawled+as.factor(ssid)"
 # covariates <- "+muddy+mixed+rocky"
@@
 # covariates <- "+trawled+mixed"
-covs <- gsub("\\+", "-", covariates)
-covs <- "-both-tv-depth"
+# covs <- gsub("\\+", "-", covariates)
+# covs <- "-both-tv-depth-ssid"
+# covs <- "-both-tv-depth"
+covs <- "-log-do"
 ```
@@
   # scale predictors before filtering to ensure mean and SD are global
-  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), temp = temperature_c) 
+  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), temp = temperature_c, raw_do = do_mlpl, do_mlpl = log(do_mlpl)) 
   do_data <- scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl))) 
@@
     filter(ssid %in% model_ssid) %>% 
-    filter(!is.na(do_mlpl)) %>% filter(!is.na(depth_scaled)) %>% filter(!is.na(temp))%>% 
-    select(X, Y, X10, Y10, adult_density, imm_density, depth_scaled, depth_scaled2, depth_mean, depth_sd, do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, temp_scaled, temp_scaled2, temp_mean, temp_sd, trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, depth, do_mlpl) %>% filter(!is.na(depth_scaled))
+    filter(!is.na(depth_scaled)) %>% filter(!is.na(do_mlpl)) %>% filter(!is.na(temp))
+
+if (any(names(data) == "adult_density")) { 
+  data <- data %>% select(X, Y, X10, Y10,
+      adult_density, imm_density,
+      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
+      do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, 
+      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
+      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
+      depth, do_mlpl, raw_do) 
+ } else {
+    data <- data %>% select(X, Y, X10, Y10,
+      density,
+      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
+      do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, 
+      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
+      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
+      depth, do_mlpl, raw_do) 
+ }
+
+ data <- filter(data, !(year == 2016 & ssid == 4) ) # 2016 do an outlier for survey 4
+# unique(data$year)
+# table(data$year,data$ssid)
 ```
@@
       file = paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
       file = paste0("data/", spp, 
-        "/mod-imm-biomass-", spp,"-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
     saveRDS(total_biomass_d, file = paste0("data/", spp, 
-      "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
       ))

### 2019-09-12 analysis/VOCC/old-rmds/1-biomass-temp-models.Rmd
move old rmds into separate folder

--- 2019-07-24 analysis/VOCC/1-biomass-do-models.Rmd
@@
 species <- params$species
-
 # # Species run so far...
-# species <- "Arrowtooth Flounder"
+species <- "Arrowtooth Flounder"
 # species <- "Pacific Cod"
-species <- "Sablefish"
+# species <- "Sablefish"
 # species <- "Silvergray Rockfish"
@@
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
+
 # species <- "Quillback Rockfish"
@@
 region <- "Both odd year surveys"
-# region <- "West Coast Vancouver Island"
-#  region <- "West Coast Haida Gwaii"
+ region <- "West Coast Vancouver Island"
+ region <- "West Coast Haida Gwaii"
 ```
@@
 priors <- TRUE
-
 covariates <- "+muddy+any_rock"
  # covariates <- ""
-
 # covariates <- "+muddy+mixed+rocky"
@@
   CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
-  data <- dplyr::left_join(data, CTD) %>% 
-  mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% filter(exclude != 1) %>% # added on july 24 after first round of models run...
-    filter(year > 2007) # 2007 DO data is flawed
+  data <- dplyr::left_join(data, CTD)
 
@@
   data <- do_data %>% mutate(depth = raw_depth) %>% 
-    filter(ssid %in% model_ssid) %>% filter(!is.na(do_mlpl))# %>% filter(year != 2016)
+    filter(ssid %in% model_ssid) %>% filter(!is.na(temp))# %>% filter(year != 2016)
 ```
@@
 
-Run DO sdmTMB model
+Run temperature sdmTMB model
 ```{r}
@@
     starttime1 <- Sys.time()
-    adult_biomass_d <- sdmTMB::sdmTMB(data,
+    adult_biomass_t <- sdmTMB::sdmTMB(data,
       adult_formula,
-      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
+      time_varying = ~ 0 + temp_scaled + temp_scaled2,
       time = "year", spde = spde,
@@
 
-    saveRDS(adult_biomass_d, 
+    saveRDS(adult_biomass_t, 
       file = paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-mat-biomass-", spp,"-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
     starttime2 <- Sys.time()
-    imm_biomass_d <- sdmTMB::sdmTMB(data, 
+    imm_biomass_t <- sdmTMB::sdmTMB(data, 
       imm_formula,
-      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
+      time_varying = ~ 0 + temp_scaled + temp_scaled2,
       time = "year", spde = spde,
@@
 
-    saveRDS(imm_biomass_d, 
+    saveRDS(imm_biomass_t, 
       file = paste0("data/", spp, 
-        "/mod-imm-biomass-", spp,"-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-imm-biomass-", spp,"-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
 
-    total_biomass_d <- sdmTMB::sdmTMB(data,
+    total_biomass_t <- sdmTMB::sdmTMB(data,
       dens_formula,
-      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
+      time_varying = ~ 0 + temp_scaled + temp_scaled2,
       time = "year", spde = spde,
@@
 
-    saveRDS(total_biomass_d, file = paste0("data/", spp, 
-      "/model-total-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    saveRDS(total_biomass_t, file = paste0("data/", spp, 
+      "/model-total-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
       ))

### 2019-09-12 analysis/VOCC/old-rmds/4-vocc-BACI.Rmd
move old rmds into separate folder

--- 2019-08-23 analysis/VOCC/4-vocc-BACI.Rmd
@@
   aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
-  geom_point(alpha=0.4, size = 2) +
+  geom_point(alpha=0.8, size = 1.5) +
   scale_shape_manual(values= c(0, 15)) +

### 2019-09-26 analysis/VOCC/3-fixed-var-response-plots.Rmd
update response plots with log-do, to plot non-sad curves, remove 2016 from models with do...

@@
 #covs <- "-tv-depth"
-covs <- "-both-tv-depth"
-
+# covs <- "-both-tv-depth"
+# covs <- "-log-do"
+covs <- "-log-do-trawled"
 num_params <- 2
@@
 spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-# folder to hold figs for this species
-dir.create(file.path("figs", spp))
-dir.create(file.path("data", spp))
 
@@
 try({
+  # loads global models with 2016 excluded due to strangely high DO 
+  add2016 <- ""
+  x2016 <- "x2016"
+  # for models using 2016 
+  # add2016 <- "-do"
+  #  x2016 <- ""
   survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
@@
   try({adult <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
   imm <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   try({total <- readRDS(paste0("data/", spp,
-    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+    "/model-total-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"))
   })
@@
  if(model_list[[i]]$model$convergence==0){
-  do[[i]] <- fixed_density(model_list[[i]], predictor = "do_mlpl")
+  do[[i]] <- fixed_density(model_list[[i]], predictor = "do_mlpl", quadratics_only = FALSE)
 
@@
    do[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
-   do[[i]]$ssid <- unique(model_list[[i]]$data$ssid)
+   do[[i]]$ssid <- paste0(sort(unique(as.numeric(model_list[[i]]$data$ssid))), collapse = "n")
    do[[i]]$model <- names(model_list[i])
-   do[[i]]$covs <- covs 
+   do[[i]]$covs <- paste0(x2016,covs) 
    roots <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.5)
-   do[[i]]$optimal_do <- roots[["optimal"]]
-   do[[i]]$lower_do_50 <- roots[["lower_threshold"]]
-   do[[i]]$upper_do_50 <- roots[["upper_threshold"]]
+   do[[i]]$optimal_do <- exp(roots[["optimal"]])
+   do[[i]]$lower_do_50 <- exp(roots[["lower_threshold"]])
+   do[[i]]$upper_do_50 <- exp(roots[["upper_threshold"]])
    roots75 <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.75)
-   do[[i]]$lower_do_75 <- roots75[["lower_threshold"]]
-   do[[i]]$upper_do_75 <- roots75[["upper_threshold"]]
+   do[[i]]$lower_do_75 <- exp(roots75[["lower_threshold"]])
+   do[[i]]$upper_do_75 <- exp(roots75[["upper_threshold"]])
    roots25 <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.25)
-   do[[i]]$lower_do_25 <- roots25[["lower_threshold"]]
-   do[[i]]$upper_do_25 <- roots25[["upper_threshold"]]
+   do[[i]]$lower_do_25 <- exp(roots25[["lower_threshold"]])
+   do[[i]]$upper_do_25 <- exp(roots25[["upper_threshold"]])
+   do[[i]]$x <- exp(do[[i]]$x)
    do_plots[[i]] <- 
-     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0, 8)) +
+     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0.2, 8)) +
         ggtitle(paste(species, names(model_list[i]))) +
-        geom_vline(xintercept=do[[i]]$optimal_do, size= 1.2, colour = "darkgrey") + 
-        geom_vline(xintercept=do[[i]]$lower_do_50, linetype="longdash", alpha = 0.75) + 
-        geom_vline(xintercept=do[[i]]$upper_do_50, linetype="longdash", alpha = 0.75) +
-        geom_vline(xintercept=do[[i]]$lower_do_75, linetype="dotted", alpha = 0.5) + 
-        geom_vline(xintercept=do[[i]]$upper_do_75, linetype="dotted", alpha = 0.5) +
-        geom_vline(xintercept=do[[i]]$lower_do_25, linetype="dotted", alpha = 0.5) + 
-        geom_vline(xintercept=do[[i]]$upper_do_25, linetype="dotted", alpha = 0.5) 
+        geom_vline(xintercept=(do[[i]]$optimal_do), size= 1.2, colour = "darkgrey") + 
+        geom_vline(xintercept=(do[[i]]$lower_do_50), linetype="longdash", alpha = 0.75) + 
+        geom_vline(xintercept=(do[[i]]$upper_do_50), linetype="longdash", alpha = 0.75) +
+        geom_vline(xintercept=(do[[i]]$lower_do_75), linetype="dotted", alpha = 0.5) + 
+        geom_vline(xintercept=(do[[i]]$upper_do_75), linetype="dotted", alpha = 0.5) +
+        geom_vline(xintercept=(do[[i]]$lower_do_25), linetype="dotted", alpha = 0.5) + 
+        geom_vline(xintercept=(do[[i]]$upper_do_25), linetype="dotted", alpha = 0.5) 
    } else {
@@
  if(model_list[[i]]$model$convergence==0){
-  td[[i]] <- fixed_density(model_list[[i]], predictor = "temp", fixed_params = num_params)
+  td[[i]] <- fixed_density(model_list[[i]], predictor = "temp", quadratics_only = FALSE, fixed_params = num_params)
   if (length(td[[i]])>0) {
@@
     td[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
-    td[[i]]$ssid <- unique(model_list[[i]]$data$ssid)
+    td[[i]]$ssid <- paste0(sort(unique(as.numeric(model_list[[i]]$data$ssid))), collapse = "n")
     td[[i]]$model <- names(model_list[i])
-    td[[i]]$covs <- covs
+    td[[i]]$covs <- paste0(x2016,covs) 
    roots <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.5)
@@
    temp_plots[[i]] <- 
-     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 10)) +
+     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 12)) +
         ggtitle(paste(species, names(model_list[i]))) +
@@
 png(
-  file = paste0("figs/", spp, "/do-fixed-", spp, covs, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/do-fixed-", spp, x2016, covs, "-priors-", priors, ".png"),
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/temp-fixed-", spp, covs, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/temp-fixed-", spp, x2016, covs, "-priors-", priors, ".png"),
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/check-optimals-", spp, covs, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/check-optimals-", spp, x2016, covs, "-priors-", priors, ".png"),
   res = 600,
@@
     "figs/", spp,
-    "/depth-", spp, covs, "-priors-", priors, ".png"
+    "/depth-", spp, x2016, covs, "-priors-", priors, ".png"
   ),
@@
   if(!is.null(tdat)) {
-    params <- full_join(tdat, odat, by=c("species", "age", "model", "covs"))
+    params <- full_join(tdat, odat, by=c("species", "age", "model", "covs", "ssid"))
     all_params <- full_join(other_params, params)
@@
 # all_params <- subset(other_params, species!="Longnose Skate")
+# # or certain rows need removal
+# all_params <- all_params[-c(#row number#), ] 
 saveRDS(all_params, file = "data/root-values-by-species-by-age.rds")

### 2019-09-26 analysis/VOCC/3-climate-models-fixed-both.Rmd
model do as log-do

--- 2019-09-11 analysis/VOCC/1-biomass-do-models-fixed-both.Rmd
@@
 ```{r}
-# species <- params$species
-
-# # Species run so far...
-# species <- "Arrowtooth Flounder"
-# species <- "Petrale Sole"
-# species <- "English Sole"
-# species <- "Dover Sole" # maturity data error for WCHG
-# species <- "Southern Rock Sole"
-# species <- "Flathead Sole"
-
-# species <- "Silvergray Rockfish"
-# species <- "Quillback Rockfish" # summer-birthing
-# species <- "Yelloweye Rockfish" # summer-birthing, overfished, 
-# species <- "Bocaccio" # winter-birthing, overfished
-# species <- "Sharpchin Rockfish"
-# species <- "Splitnose Rockfish"
-
-# species <- "Rougheye/Blackspotted Rockfish Complex"
-# species <- "Redbanded Rockfish"
-# species <- "Greenstriped Rockfish"
-# # species <- "Copper Rockfish" # small sample
-# species <- "Darkblotched Rockfish"
-# # species <- "Shortbelly Rockfish" # small sample
-
-# species <- "Pacific Ocean Perch" # schooling
-# species <- "Widow Rockfish" # schooling
-# species <- "Yellowtail Rockfish" # schooling
-# species <- "Canary Rockfish" # schooling, winter-birthing
-
-# species <- "Longspine Thornyhead"
-# species <- "Shortspine Thornyhead"
-
-# species <- "Walleye Pollock"
-# species <- "Pacific Cod"
-# species <- "Sablefish"
-# species <- "Lingcod"
-# species <- "Pacific Hake"
-
-# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Longnose Skate"
-# species <- "Big Skate"
-# species <- "Spotted Ratfish"
-# species <- "Sandpaper Skate"
-# species <- "Brown Cat Shark"
+# species <- "Canary Rockfish" 
 ```
@@
 # covariates <- "+muddy+any_rock"
- covariates <- ""
-# covariates <- "+trawled+as.factor(ssid)"
+# covariates <- ""
+# covariates <- "+as.factor(ssid)"
+covariates <- "+trawled"
 # covariates <- "+muddy+mixed+rocky"
@@
 # covs <- "-both-tv-depth"
-covs <- "-log-do"
+# covs <- "-log-do"
+# covs <- "-log-do-ssid"
+covs <- "-log-do-trawled"
+
 ```
@@
 ```{r}
+tictoc::tic()
   if (any(names(data) == "adult_density")) {
@@
   }
+tictoc::toc()
 ```

### 2019-09-26 analysis/VOCC/2-biomass-depth-only-model.Rmd
update model choice and style

@@
 ```{r}
-species <- params$species
 
-# # Species run so far...
-# species <- "Arrowtooth Flounder" # temp good, not DO, imm moving deeper?
-
-# species <- "Petrale Sole"
-# species <- "English Sole"
-# species <- "Dover Sole"
-# species <- "Southern Rock Sole"
-
-# species <- "Walleye Pollock" # survey 4 or 16 might work
-# species <- "Pacific Cod"
-
-## Sebastes
-
-## TRY BOTH - not schooling
- # species <- "Yelloweye Rockfish" # beautiful curves, with depth change ***
- #  species <- "Bocaccio" # good curves, no depth change ... most survey-specific models fail to find minimum ***
-#  species <- "Sharpchin Rockfish" # beautiful curves, imm depth change ***
-#  species <- "Greenstriped Rockfish" # beautiful curves for adults, no depth change ***
-
-# ## TRY BOTH - schooling
-#  species <- "Pacific Ocean Perch" # beautiful curves, no depth change, schooling ***
-#  species <- "Silvergray Rockfish" # beautiful curves, no depth change, schooling ***
-
-## TRY JUST TEMPERATURE 
-# species <- "Rougheye/Blackspotted Rockfish Complex" # good temp, not DO
-# species <- "Redbanded Rockfish" # good temp, not DO, maybe moving shallower
- species <- "Splitnose Rockfish" # good temp, survey 4 DO ok
-
-# species <- "Yellowtail Rockfish" #  good temp for survey 4 only, no DO, schooling
-# species <- "Quillback Rockfish" # like warm = overall a bit flat... territorial! 
-# species <- "Canary Rockfish" # schooling, winter-birthing, and depth change!, but curves at edges... adult1 looks ok?
-#  species <- "Widow Rockfish" # schooling, and depth change, but curves at edges
-
-# species <- "Longspine Thornyhead" # too deep to produce curves
-# species <- "Shortspine Thornyhead"
-# species <- "Sablefish" # temp might work, no DO, no consistant depth change
-# species <- "Lingcod" # curves ok, overall a bit flat
-# # species <- "Pacific Hake" # not useful
-
-# species <- "North Pacific Spiny Dogfish" # too flat, but with depth change! note: pooled maturity 
-# species <- "Longnose Skate" # do too flat, temp from 4 might be useful, seems to go deeper?
-# species <- "Big Skate"
-# species <- "Sandpaper Skate"
-# species <- "Brown Cat Shark"
-# species <- "Spotted Ratfish" # curves not useful because to abundant in shallow
 ```
@@
 
-knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
-                      fig.path=paste0("figs/", spp, "/"),
-                      echo=FALSE, warning=FALSE, message=FALSE)
+knitr::opts_chunk$set(
+  fig.width = 11, fig.height = 8.5,
+  fig.path = paste0("figs/", spp, "/"),
+  echo = FALSE, warning = FALSE, message = FALSE
+)
 ```
@@
 data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth))
-  data <- gfranges::scale_predictors(data, #predictors = c(quo(depth)))
-        predictors = c(quo(depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
-  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)
+data <- gfranges::scale_predictors(data, # predictors = c(quo(depth)))
+  predictors = c(quo(depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock))
+)
+data <- data %>%
+  mutate(depth = raw_depth) %>%
+  filter(ssid %in% model_ssid)
 ```
@@
 ```{r}
-
+tictoc::tic()
 if (any(names(data) == "adult_density")) {
-
   adult_formula <- as.formula(paste(
@@
     adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 , #+ trawled
+    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
     time = "year", spde = spde,
@@
   saveRDS(adult_biomass,
-      file = paste0("data/", spp, 
+    file = paste0(
+      "data/", spp,
       "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-        ))
+    )
+  )
 
@@
   saveRDS(imm_biomass,
-     file = paste0("data/", spp, 
+    file = paste0(
+      "data/", spp,
       "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-       ))
-
+    )
+  )
 } else {
-
   starttime3 <- Sys.time()
@@
     dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,# + trawled
+    time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
     time = "year", spde = spde,
@@
 
-    saveRDS(total_biomass, file = paste0("data/", spp, 
+  saveRDS(total_biomass, file = paste0(
+    "data/", spp,
     "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
 }
+tictoc::toc()
 ```
@@
   bath <- readRDS("data/bathymetry-data")
-  #dat <- bath$data %>% scale_survey_predictors()
+  # dat <- bath$data %>% scale_survey_predictors()
   dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth))
@@
   newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
-  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))
+  nd_all <- left_join(nd_all, newcov, by = c("X", "Y"))
 
@@
 ```{r}
-#covs <- gsub("\\+", "-", covariates)
+# covs <- gsub("\\+", "-", covariates)
 fish <- readRDS(paste0("raw/bio-data-", spp, ""))
@@
 if (length(unique(fish$maturity_code)) > 2) {
-
   nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-    adult_biomass <- readRDS(paste0("data/", spp, 
+  adult_biomass <- readRDS(paste0(
+    "data/", spp,
     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   adult_predictions <- predict(adult_biomass, newdata = nd)
-    saveRDS(adult_predictions, file = paste0("data/", spp, 
+  saveRDS(adult_predictions, file = paste0(
+    "data/", spp,
     "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
@@
 
-    imm_biomass <- readRDS(paste0("data/", spp, 
+  imm_biomass <- readRDS(paste0(
+    "data/", spp,
     "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   imm_predictions <- predict(imm_biomass, newdata = nd)
-    saveRDS(imm_predictions, file = paste0("data/", spp, 
+  saveRDS(imm_predictions, file = paste0(
+    "data/", spp,
     "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
@@
 
-      total_biomass <- readRDS(paste0("data/", spp, 
+    total_biomass <- readRDS(paste0(
+      "data/", spp,
       "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
     predictions <- predict(total_biomass, newdata = nd)
-      saveRDS(predictions, file = paste0("data/", spp, 
+    saveRDS(predictions, file = paste0(
+      "data/", spp,
       "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"

### 2019-09-26 analysis/VOCC/2-biomass-trends.Rmd
update model choice and style

--- 2019-08-29 analysis/VOCC/2-biomass-trends.Rmd
@@
 ```{r}
-# # Species run so far...
- species <- "Arrowtooth Flounder"
-#  species <- "Pacific Cod"
-# species <- "Sablefish"
-#  species <- "Silvergray Rockfish"
-# species <- "Lingcod"
-#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Quillback Rockfish"
-# species <- "Pacific Ocean Perch"
-# species <- "Yelloweye Rockfish"
+
 ```
@@
 # priors <- TRUE
-
+# covariates <- ""
+# covariates <- "+trawled"
+# covs <- gsub("\\+", "-", covariates)
+covariates <- "+trawled+as.factor(ssid)"
+covs <- gsub("\\+", "-", "+trawled+ssid")
 # covariates <- "+muddy+mixed+rocky"
 # covariates <- "+mixed+rocky"
-covariates <- "+muddy+any_rock"
+# covariates <- "+muddy+any_rock"
 # covariates <- "+trawled+muddy+rocky+mixed"
@@
 
-knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
-                      #fig.path=paste0("figs/", spp, "/"),
-                      echo=FALSE, warning=FALSE, message=FALSE)
+knitr::opts_chunk$set(
+  fig.width = 11, fig.height = 8.5,
+  # fig.path=paste0("figs/", spp, "/"),
+  echo = FALSE, warning = FALSE, message = FALSE
+)
 ```
@@
 rm(pred16_all)
-#if (isFALSE(update_spatial_model)) {
+# if (isFALSE(update_spatial_model)) {
 if (region != "All synoptic surveys") {
-
-    try(adult_predictions <- readRDS(paste0("data/", spp, 
-      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
-    try(imm_predictions <- readRDS(paste0("data/", spp, 
-      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
+  try(adult_predictions <- readRDS(paste0(
+    "data/", spp,
+    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
+  )))
+  try(imm_predictions <- readRDS(paste0(
+    "data/", spp,
+    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
+  )))
 
@@
 
-      try(predictions <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
+    try(predictions <- readRDS(paste0(
+      "data/", spp,
+      "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
+    )))
 
@@
   }
+} else {
+  try({
+    adult_predictions <- readRDS(paste0(
+      "data/", spp,
+      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
+    ))
+  })
+  try({
+    imm_predictions <- readRDS(paste0(
+      "data/", spp,
+      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
+    ))
+  })
 
-} else {
-    # Bind together predictions from different models
+  if (!exists("adult_predictions")) {
+    # OR bind together predictions from different models
     try({
-      pred1n3 <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"))
-      pred4 <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"))
-      pred16 <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"))
+      pred1n3 <- readRDS(paste0(
+        "data/", spp,
+        "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"
+      ))
     })
+    try({
+      pred4 <- readRDS(paste0(
+        "data/", spp,
+        "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"
+      ))
+    })
+    try({
+      pred16 <- readRDS(paste0(
+        "data/", spp,
+        "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"
+      ))
+    })
+
+    if (!exists("pred1n3")) pred1n3 <- NULL
     if (!exists("pred16")) pred16 <- NULL
@@
     try({
-      pred1n3_imm <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"))
-      pred4_imm <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))
-      pred16_imm <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"))
+      pred1n3_imm <- readRDS(paste0(
+        "data/", spp,
+        "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"
+      ))
+    })
+    try({
+      pred4_imm <- readRDS(paste0(
+        "data/", spp,
+        "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"
+      ))
+    })
+    try({
+      pred16_imm <- readRDS(paste0(
+        "data/", spp,
+        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"
+      ))
     })
 
+    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
     if (!exists("pred16_imm")) pred16_imm <- NULL
@@
       try({
-      pred1n3_all <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))
-      pred4_all <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))
-      pred16_all <- readRDS(paste0("data/", spp, 
-        "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))
+        predictions <- readRDS(paste0(
+          "data/", spp,
+          "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
+        ))
+      })
+
+      if (!exists("predictions")) {
+        try({
+          pred1n3_all <- readRDS(paste0(
+            "data/", spp,
+            "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"
+          ))
+        })
+        try({
+          pred4_all <- readRDS(paste0(
+            "data/", spp,
+            "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"
+          ))
+        })
+        try({
+          pred16_all <- readRDS(paste0(
+            "data/", spp,
+            "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"
+          ))
+        })
+        try({
           predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
@@
         if (!exists("predictions")) paste("Total biomass predictions are not available.")
+      }
+    }
   }
@@
 
-  saveRDS(adult_predictions, file = paste0("data/", spp, 
-      "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
-  saveRDS(imm_predictions, file = paste0("data/", spp, 
-      "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))
+  saveRDS(adult_predictions, file = paste0(
+    "data/", spp,
+    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
+  ))
+  saveRDS(imm_predictions, file = paste0(
+    "data/", spp,
+    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
+  ))
 
+  model_ssid <- unique(adult_predictions$ssid)
 } else {
-
   predictions$total_bio <- exp(predictions$est) * 10000
@@
 
-  saveRDS(predictions, file = paste0("data/", spp, 
-      "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
+  saveRDS(predictions, file = paste0(
+    "data/", spp,
+    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
+  ))
+
+  model_ssid <- unique(predictions$ssid)
 }
@@
 if (exists("adult_predictions")) {
-
   p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
@@
   print(p_adult_all)
-
 } else {
-
   p_adult_all <- plot_facet_map(predictions, "total_bio",
@@
   if (3 %in% model_ssid) {
-    adult_predictions3 <- adult_predictions %>% filter(ssid != 16) %>% filter(ssid != 4) 
+    adult_predictions3 <- adult_predictions %>%
+      filter(ssid != 16) %>%
+      filter(ssid != 4)
   } else {
@@
   if (16 %in% model_ssid) {
+    if (any(adult_predictions$ssid == 16)) {
       adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
+    } else {
+      adult_predictions16 <- NULL
+    }
   } else {
@@
 ```{r, message=FALSE, warning=FALSE}
-#try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
+# try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
 rm(out_mature)
@@
         variable_names = "est_exp"
-
       )
@@
 ```{r }
-#FIXME: Attempting to plot of current mature biomass with trawl footprint...
-# if (exists("out_mature")) {
-
-trawl_footprint <- sf::st_read(dsn="data/trawl-footprint",layer="Trawl_footprint")
+if (exists("out_mature")) {
+  trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
   proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
@@
   current_biomass <- current_biomass +
-   geom_sf(data= trawl_footprint, aes(geometry=geometry/1000), colour = "gray", fill = "gray", alpha= 0.2, lwd=0.25, inherit.aes = FALSE) +
-   coord_sf(xlim = c(180,820), ylim = c(5350,6050), datum = NULL) +
+    geom_sf(data = trawl_footprint, aes(geometry = geometry / 1000), colour = "gray", fill = "gray", alpha = 0.2, lwd = 0.25, inherit.aes = FALSE) +
+    coord_sf(xlim = c(180, 820), ylim = c(5350, 6050), datum = NULL) +
     xlab("") +
@@
   current_biomass
-# }
+}
 ```
@@
     white_zero = TRUE,
-    high_fill = "deepskyblue3",#"lightseagreen", # limegreen",
+    high_fill = "deepskyblue3", # "lightseagreen", # limegreen",
     low_fill = "red", # "deeppink3",# "steel blue 4",
@@
     high_fill = "deepskyblue3", # "greenyellow", # limegreen",
-    low_fill = "red", #"steel blue 4",
+    low_fill = "red", # "steel blue 4",
     transform_col = fourth_root_power,
@@
   if (3 %in% model_ssid) {
-    imm_predictions3 <- imm_predictions %>% filter(ssid != 16) %>% filter(ssid != 4) 
+    imm_predictions3 <- imm_predictions %>%
+      filter(ssid != 16) %>%
+      filter(ssid != 4)
   } else {
@@
 ```{r, message=FALSE, warning=FALSE}
-#try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
+# try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
 rm(out_imm)
@@
     white_zero = TRUE,
-    high_fill = "deepskyblue3",# "greenyellow", # limegreen",
-    low_fill = "red", #"steel blue 4",
+    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
+    low_fill = "red", # "steel blue 4",
     transform_col = fourth_root_power,

### 2019-09-26 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
BACI matching, if do threshold must be >0.4, save plots of matches

@@
 age <- "mature" # or mostly so
-# age <- "imm"
+age <- "imm" 
 ```
@@
 ```{r echo=TRUE}
-biomass_threshold <- "50" # filter for climate values predicted to support 50% of max biomass 
- biomass_threshold <- "25" # filter for climate values predicted to support 50% of max biomass 
 global_only <- TRUE
+biomass_threshold <- "75"
+
+# biomass_threshold <- "50" # filter for climate values predicted to support 50% of max biomass 
+#  biomass_threshold <- "25" # filter for climate values predicted to support 25% of max biomass... doesn't work for DO
 
@@
  temp_threshold <- c(0.25)
-
 # temp_threshold <- c(0.5)
 # temp_threshold <- c(0.75)
+# temp_threshold <- c(1)
+
  # climate <- "do"
-# #  do_threshold <- c(0.5)
  # do_threshold <- c(0.25)
+# do_threshold <- c(0.5)
+
 # climate <- "do and temperature"
@@
 
+# Run all subsequent code...
+
+```{r, message=FALSE, warning=FALSE, include=FALSE}
+if (region == "both odd year surveys") {
+  survey <- c("SYN QCS", "SYN HS")
+  model_ssid <- c(1, 3)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  start_years <- list("2009"=2009, "2011"=2011, "2013"=2013, "2015"=2015) 
+  # maybe remove 2011 & 2015 once we can add "2017"=2017?
+  years <- NULL
+}
+
+if (region == "West Coast Vancouver Island") {
+  survey <- c("SYN WCVI")
+  model_ssid <- c(4)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) # prob need to remove 2010 & 2014
+  years <- NULL
+}
+
+if (region == "West Coast Haida Gwaii") {
+  survey <- c("SYN WCHG")
+  model_ssid <- c(16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) # prob need to remove 2010 & 2014
+  years <- NULL
+}
+```
+
 #### Get species-specific optimal values from mountain plots for most abundant year
@@
 params <- params[params$species == species, ] 
+glimpse(params)
 
@@
 } else {
- #params <- filter(params, ssid %in% model_ssid)
+ if (model_ssid == 4) {
+   params <- params %>% mutate(survey4 = ifelse(grepl("4", model), "T", "F"))
+   params <- filter(params, survey4 == "T")
+ }
 }
@@
   upper_thresholds <- round(.upper[1], digits = 2)
+  upper_model <- params[rev(order(params$t_maxy)), ]$model
+
   lower_thresholds <- NA
-
+  lower_model <- NULL
   variable_names <- c("temp")
@@
   # optimal_values <- optimal_do
-  params <- filter(params, lower_do_25 > 0.5)
+  params <- filter(params, lower_do_25 > 0.4)
 
@@
   lower_thresholds <- round(.lower[1], digits = 2)
+  lower_model <- params[rev(order(params$do_maxy)), ]$model
   upper_thresholds <- NA
+  upper_model <- NULL
   variable_names <- c("do_est")
@@
   upper_change <- c(Inf, temp_threshold)
+  upper_model <- params[rev(order(params$t_maxy)), ]$model
+  lower_model <- params[rev(order(params$do_maxy)), ]$model
+
   indices <- c(1, 2)
@@
 trim_thresholds <- c(lower_thresholds, upper_thresholds)
+trim_models <- c(lower_model, upper_model)
+
 min_string <- paste0(lower_change, collapse = "-")
@@
 glimpse(params)
-```
-
-# Run all subsequent code...
-
-```{r, message=FALSE, warning=FALSE, include=FALSE}
-if (region == "both odd year surveys") {
-  survey <- c("SYN QCS", "SYN HS")
-  model_ssid <- c(1, 3)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  start_years <- list("2009"=2009, "2011"=2011, "2013"=2013, "2015"=2015) 
-  # maybe remove 2011 & 2015 once we can add "2017"=2017?
-  years <- NULL
-}
-
-if (region == "West Coast Vancouver Island") {
-  survey <- c("SYN WCVI")
-  model_ssid <- c(4)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) # prob need to remove 2010 & 2014
-  years <- NULL
-}
-
-if (region == "West Coast Haida Gwaii") {
-  survey <- c("SYN WCHG")
-  model_ssid <- c(16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) # prob need to remove 2010 & 2014
-  years <- NULL
-}
 ```
@@
 ```{r}
-
 trimmed_vectors <- lapply(vocc_by_year, function(vocc) {    
@@
   ncol = 2,
-  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
+  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", trim_models, ")"), gp=grid::gpar(fontsize=16))
 )
@@
 rm(vocc_by_sp)
-try({
-  vocc_by_sp <- readRDS(paste0(
-    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
-    min_string, "to", max_string, "-multiyear-all.rds"
-  ))
-})
+# try({
+#   vocc_by_sp <- readRDS(paste0(
+#     "data/BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
+#   ))
+# })
 
@@
 
-match_formula <- paste0("treat ~ trawled + x + y") 
+match_formula <- paste0("treat ~ trawled + x + y + depth_st") 
   # depth.st + omega_s + muddy.st + any_rock.st + 
-  # for (j in (variable_names)) {
-  #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
-  #   }
+  for (j in (variable_names)) {
+    match_formula <- paste0("", match_formula, " + ", j, "_s.st")
+    start_value.st <- paste0("", j, "_s.st")
+    start_value <- paste0("", j, "_s")
+
+    alldata <- alldata %>% mutate((!!start_value.st) := 
+        arm::rescale(alldata[[start_value]]) )
+  #(UQ(rlang::sym(paste0(variable_names[j], "_s")))))
+    }
 
@@
   alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
+  alldata$depth_st <- arm::rescale(log(alldata$depth))
   alldata$depth.st <- round(arm::rescale(log(alldata$depth))*5)/5
@@
   alldata$any_rock.st <- arm::rescale(alldata$any_rock)
-  #alldata$omega_s.st <- arm::rescale(alldata$omega_s)
 
@@
   matched <- MatchIt::matchit(as.formula(match_formula), 
-    exact = c("start_density.st", "start_time", "depth.st"),
+    exact = c("start_density.st", "start_time"),
     method = "nearest", m.order = "random",
@@
     "data/", spp,
-    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
-    min_string, "to", max_string, "-multiyear-all.rds"
+    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
   ))
@@
 species <- rep(species, n) 
+mature <- rep(age, n)
 data <- vocc_by_sp[[1]]
-df <- cbind(species, data)
+class <- cbind(species, mature)
+df <- cbind(class, data)
 species <- species[[1]]
@@
 
-# View(data)
+# View(df)
 ```
@@
 
-ggplot(data, 
+p1 <- ggplot(data, 
 #ggplot(filter(data, as.numeric(matchobs) <= 550), 
@@
   theme_bw()
+p1
 
@@
 
-ggplot(data, aes(std_log_depth, fill = cell_type)) +
+p2 <- ggplot(data, aes(std_log_depth, fill = cell_type)) +
   geom_density(alpha=0.5) +
@@
   theme_bw()
-#  
+p2
+
 # ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
@@
 
-ggplot(data, aes(log_density, fill = cell_type)) +
+p3 <- ggplot(data, aes(log_density, fill = cell_type)) +
   geom_density(alpha=0.5) +
@@
   facet_grid(start_time~time, scales = 'fixed') +
-  theme_bw()
+  theme_bw() + theme(legend.position='none')
+p3
 
-ggplot(data, aes(log_density, fill = cell_type)) +
+p4 <- ggplot(data, aes(log_density, fill = cell_type)) +
   geom_histogram(alpha=0.45) +
@@
   facet_grid(time~start_time) +
-  theme_bw()
+  theme_bw() + theme(legend.position='none')
+p4
 
@@
 
-ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
+p5 <- ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
   geom_boxplot(alpha=0.5) +
@@
   geom_text(data = dat_summary,
-            aes(cell_type, max(data$log_density)-1, label = n), hjust = -0.3) +
+            aes(cell_type, min(data$log_density)+1, label = n), hjust = 1.25) +
   facet_grid(time~start_time) +
-  theme_bw()
+  theme_bw() + theme(legend.position='none')
+p5
 ```
@@
 ```{r}
-# png(
-#   file = paste0("figs/", spp,
-#     "/matches-", min_string, "to", max_string, "-", 
-#     climate_string, "-", spp, "-", age, "-", ssid_string, "-", 
-#     climate_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
-#   ),
-#   res = 600,
-#   units = "in",
-#   width = 8.5,
-#   height = 11
-# )
-# gridExtra::grid.arrange(
-#   grobs = ,
-#   ncol = 2,
-#   top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
-# )
-# dev.off()
-```
+png(
+  file = paste0("figs/", spp,
+    "/matches-", min_string, "to", max_string, "-",
+    climate_string, "-", spp, "-", age, "-", ssid_string, "-",
+    climate_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs1.png"
+  ),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 11
+)
+gridExtra::grid.arrange(
+  p1,p2,p3,p4,p5,
+  ncol = 2,
+  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
+)
+dev.off()
 ```
@@
 ### Basic BACI
-```{r}
+```{r eval=F}
 for (j in (variable_names)) {
@@
 
-```{r}
+```{r eval=F}
 sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "adult"))
@@
 ### GAMM
-```{r}
+```{r eval=F}
  gamm_formula <- paste0("log_density ~ ", 
@@
   #"start_time * cell_type + ",
-  "time * cell_type + s(vect_dist) + s(std_log_depth) + s(X,Y)")
+  "time * cell_type  + s(std_log_depth) + s(X,Y) ") #+ s(vect_dist)
 
@@
   random = ~(1|matchobs) + (1|icell) 
-   + (1|start_time)
-  # + (time|start_time) 
+  # + (1|start_time)
+  + (time|start_time) 
   # + (cell_type|start_time)
@@
 ```
+
+```{r}
+spp
+```

### 2019-09-30 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
revise for auto-render with params

@@
 output: html_document
+params: 
+    species: "Yelloweye Rockfish" 
+    region: "West Coast Vancouver Island"
+    age: "mature"
+    biomass_threshold: "25"
+    climate: "do"
+    threshold: c(0.25) 
 ---
-
-Modified to allow thresholds chosen to vary by maturity class. 
-Therefore need to run this file separately for each species-class combination.
-
-Should matches be selected based on exact depth categories in addition to density (all within time periods)?
-
-What does it mean for the analysis that the depth range at which VOCC occur within a species temperature range varies dramatically between time periods?
 
@@
 
-#### Set species
-```{r echo=TRUE}
-
+```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
+library("dplyr")
+library("ggplot2")
+library("gfplot")
+library("sdmTMB")
+library("gfranges")
 ```
 
-#### Set age class
+#### Set parameters
 ```{r echo=TRUE}
-age <- "mature" # or mostly so
-age <- "imm" 
-```
-
-#### Set vector type
-```{r echo=TRUE}
-global_only <- TRUE
-biomass_threshold <- "75"
-
-# biomass_threshold <- "50" # filter for climate values predicted to support 50% of max biomass 
-#  biomass_threshold <- "25" # filter for climate values predicted to support 25% of max biomass... doesn't work for DO
-
-climate <- "temperature"
- temp_threshold <- c(0.25)
-# temp_threshold <- c(0.5)
-# temp_threshold <- c(0.75)
-# temp_threshold <- c(1)
-
- # climate <- "do"
- # do_threshold <- c(0.25)
-# do_threshold <- c(0.5)
-
-# climate <- "do and temperature"
-```
-
-#### Set Region 
-```{r echo=TRUE}
-# region <- "West Coast Vancouver Island"
-region <- "both odd year surveys"
-# region <- "West Coast Haida Gwaii"
+species <- params$species
+age <- params$age
+region <- params$region
+biomass_threshold <- params$biomass_threshold
+climate <- params$climate
 ```
@@
 ```{r, echo=FALSE}
+global_only <- TRUE
 input_cell_size <- 2
@@
 skip_time <- NULL
-```
-
-```{r, include=FALSE}
-library(dplyr)
-library(ggplot2)
-library(gfplot)
-# library(gfdata)
-library(sdmTMB)
-# library(itsadug)
-library(gfranges)
 ```
@@
   start_years <- list("2009"=2009, "2011"=2011, "2013"=2013, "2015"=2015) 
-  # maybe remove 2011 & 2015 once we can add "2017"=2017?
+  # maybe remove 2011 & 2015 once we can add "2017"=2017 to avoid start and end values overlapping
   years <- NULL
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) # prob need to remove 2010 & 2014
+  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) 
+  # prob need to remove 2010 & 2014
   years <- NULL
@@
 
-params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
+root_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
 
-params <- params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
+root_params <- root_params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
 
 if (age == "imm") {
-  params <- filter(params, age == "imm")
+  root_params <- filter(root_params, age == "imm")
 } else {
-  params <- filter(params, age != "imm")
-  if (any(params$model_type == "split")) { params <- filter(params, model_type != "total") }
+  root_params <- filter(root_params, age != "imm")
+  if (any(root_params$model_type == "split")) { root_params <- filter(root_params, model_type != "total") }
 }
 
-params <- params[params$species == species, ] 
-glimpse(params)
+root_params <- root_params[root_params$species == species, ] 
+glimpse(root_params)
 
 if (global_only) {
- params <- params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
- params <- filter(params, global == "T")
+ root_params <- root_params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
+ root_params <- filter(root_params, global == "T")
 } else {
  if (model_ssid == 4) {
-   params <- params %>% mutate(survey4 = ifelse(grepl("4", model), "T", "F"))
-   params <- filter(params, survey4 == "T")
+   root_params <- root_params %>% mutate(survey4 = ifelse(grepl("4", model), "T", "F"))
+   root_params <- filter(root_params, survey4 == "T")
  }
@@
 if (climate == "temperature") {
-  # optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
-  # optimal_temps <- na.omit(optimal_temps)
-  # optimal_temp <- round(optimal_temps[1], digits = 2)
-  # optimal_values <- optimal_temp
 
+temp_threshold <- params$threshold
+
+  if (biomass_threshold == "75") {
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
   if (biomass_threshold == "50") {
-  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_50 }
+  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 }
   if (biomass_threshold == "25") {
-  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_25 }
+  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25 }
   .upper <- na.omit(.upper)
@@
   upper_thresholds <- round(.upper[1], digits = 2)
-  upper_model <- params[rev(order(params$t_maxy)), ]$model
+   if (is.na(upper_thresholds)) {stop(print("Temp maximum not modelled"))}
+
+  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model
 
@@
 if (climate == "do") {
-  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
-  # optimal_dos <- na.omit(optimal_dos)
-  # optimal_do <- round(optimal_dos[1], digits = 2)
-  # optimal_values <- optimal_do
-  params <- filter(params, lower_do_25 > 0.4)
+  do_threshold <- params$threshold
+  root_params <- filter(root_params, lower_do_25 > 0.4)
 
+  if (biomass_threshold == "75") {
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
   if (biomass_threshold == "50") {
-  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_50 }
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
   if (biomass_threshold == "25") {
-  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_25 }
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }
 
@@
   lower_thresholds <- round(.lower[1], digits = 2)
-  lower_model <- params[rev(order(params$do_maxy)), ]$model
+  if (is.na(lower_thresholds)) {stop(print("DO minimum not modelled"))}
+  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model
+
   upper_thresholds <- NA
@@
 if (climate == "do and temperature") {
+  temp_threshold <- params$threshold
+  do_threshold <- params$threshold
 
-  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
-  # optimal_dos <- na.omit(optimal_dos)
-  # optimal_do <- round(optimal_dos[1], digits = 2)
-  # optimal_values <- optimal_do
-
-  params <- filter(params, lower_do_25 > 0.5)
+  root_params <- filter(root_params, lower_do_25 > 0.5)
+  if (biomass_threshold == "75") {
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
   if (biomass_threshold == "50") {
-  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_50 }
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
   if (biomass_threshold == "25") {
-  .lower <- params[rev(order(params$do_maxy)), ]$lower_do_25 }
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }
   .lower <- na.omit(.lower)
@@
 
-  # optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
-  # optimal_temps <- na.omit(optimal_temps)
-  # optimal_temp <- round(optimal_temps[1], digits = 2)
-  # optimal_values <- optimal_temp
-
+  if (biomass_threshold == "75") {
+  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
   if (biomass_threshold == "50") {
-  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_50 }
+  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 }
   if (biomass_threshold == "25") {
-  .upper <- params[rev(order(params$t_maxy)), ]$upper_t_25 }
+  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25 }
   .upper <- na.omit(.upper)
@@
   upper_change <- c(Inf, temp_threshold)
-  upper_model <- params[rev(order(params$t_maxy)), ]$model
-  lower_model <- params[rev(order(params$do_maxy)), ]$model
+  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model
+  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model
 
@@
 
-glimpse(params)
+glimpse(root_params)
 ```
@@
 ```{r}
-rm(vocc)
-
 vocc_by_year <- lapply(start_years, function(start_time) {
@@
 })  
-# glimpse(vocc)
 ```
@@
 trimmed_vectors <- lapply(vocc_by_year, function(vocc) {
-#start_time <- vocc$start_time[1]
 
@@
     "/vocc-", min_string, "to", max_string, "-", 
-    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
+    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs-test.png"
   ),
@@
 Get all biomass predictions 
-```{r, echo=FALSE, warning=FALSE, error=FALSE, message=F}
+```{r echo=FALSE, warning=FALSE}
 priors <- FALSE
-# covariates <- "+muddy+any_rock"
-# covs <- gsub("\\+", "-", covariates)
-covs <- gsub("\\+", "-", "-trawled+ssid")
-
-rm(biomass_predictions)
+covs <- "-trawled-ssid"
 
 if (age == "imm") {
-  try({
   biomass_predictions <- readRDS(paste0(
-    "data/", spp,
-    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
-  ))
-})
-} else {
-try({
+    "data/", spp, "/all-predictions-", spp, covs, 
+    "-imm-biomass-prior-", priors, ".rds"))
+}
+
+if (age == "mature") {
   biomass_predictions <- readRDS(paste0(
-    "data/", spp,
-    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
-  ))
-})
-if (!exists("biomass_predictions")) {
-try({
-  biomass_predictions <- readRDS(paste0(
-    "data/", spp,
-    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
-  ))
-})
-}
+    "data/", spp, "/all-predictions-", spp, covs, 
+    "-mat-biomass-prior-", priors, ".rds"))
 }
@@
 
-  alldata
+ # alldata
 #})
-```
-
-```{r eval=FALSE, include=FALSE, echo=FALSE}
-# lapply(alldata, function(i){ggplot(i, aes(x=cell_type, y=log(start_density))) + geom_boxplot()+facet_wrap(~start_time)})
 ```
@@
 ```{r}
-rm(vocc_by_sp)
+#rm(vocc_by_sp)
 # try({
@@
 # matched <- lapply(alldata, function(alldata) {  
-
   ## trim non-overlapping values to reduce size of dataset 
@@
   # if (order == "decreasing") alldata <- alldata[rev(order(alldata$start_density)), ]
-
+# set.seed(99999)
   matched <- MatchIt::matchit(as.formula(match_formula), 
@@
 mature <- rep(age, n)
+ssid <- rep(ssid_string, n)
 data <- vocc_by_sp[[1]]
-class <- cbind(species, mature)
-df <- cbind(class, data)
+df<- cbind(species, mature, ssid, data)
 species <- species[[1]]
@@
 
-if (climate=="temperature") dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))
-
 if (climate=="temperature") {
-  saveRDS(df, file = paste0(
-    "data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", max_string,"/vocc-", climate_string, "-", 
-     max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
+dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))
+dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string, age))
+}
+if (climate=="temperature") {
+  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", max_string,"/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
   ))
@@
   dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string))
+  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string, age))  
   }
 if (climate=="do") {
-  saveRDS(df, file = paste0(
-    "data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", min_string,"/vocc-", climate_string, "-",
-    min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
+  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", min_string,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
   ))
@@
   dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
+  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, age))  
 }
@@
 if (climate=="do and temperature") {
-  saveRDS(df, file = paste0(
-    "data/_all/", climate_string, "/", "perc_", biomass_threshold,"/vocc-", climate_string, "-",
-    min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
+  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
   ))
 }
-
-# View(df)
 ```
@@
     climate_string, "-", spp, "-", age, "-", ssid_string, "-",
-    climate_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs1.png"
+    climate_string, "-perc_", biomass_threshold, "-", optimal_string, ".png"
   ),
@@
 ```
-
-### Basic BACI
-```{r eval=F}
-for (j in (variable_names)) {
-  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs) + (1|start_time)")
-}
-
-m <- lmerTest::lmer(as.formula(baci_formula), data=data)
-summary(m)
-```
-
-```{r eval=F}
-sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "adult"))
-```
-
-### GAMM
-```{r eval=F}
- gamm_formula <- paste0("log_density ~ ", 
-  # "time * start_time + ",
-  #"start_time * cell_type + ",
-  "time * cell_type  + s(std_log_depth) + s(X,Y) ") #+ s(vect_dist)
-
-# for (j in (variable_names)) {
-#  gamm_formula <- paste0(gamm_formula, "+ s(std_", j, ")")
-#  }
-
-gam4 <- gamm4::gamm4(as.formula(gamm_formula), 
-  random = ~(1|matchobs) + (1|icell) 
-  # + (1|start_time)
-  + (time|start_time) 
-  # + (cell_type|start_time)
-  , data=data)
-
-summary(gam4$gam)
-
-plot(gam4$gam,pages=1)
-```
-
-```{r, eval=FALSE, results="hide", include=FALSE}
-plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source"), start_time=c("2009","2011","2013","2015")), parametricOnly = TRUE, main="", xlab ="log_density"
-  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
-  ) 
-```
-
-```{r}
-spp
-```

### 2019-09-30 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
correct typo in root value selection for temperature

@@
 
+ # if more than one model option in dataframe will pick the one with highest peak density
   if (biomass_threshold == "75") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
+  .upper <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
   if (biomass_threshold == "50") {
@@
   if (biomass_threshold == "75") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
+  .upper <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
   if (biomass_threshold == "50") {

### 2019-09-30 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
remove global_only as all models now global

@@
 ```{r, echo=FALSE}
-global_only <- TRUE
 input_cell_size <- 2
-dist_intercept <- input_cell_size/2
+dist_intercept <- input_cell_size / 2
 scale_fac <- 1
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  start_years <- list("2009"=2009, "2011"=2011, "2013"=2013, "2015"=2015) 
+  start_years <- list("2009" = 2009, "2011" = 2011, "2013" = 2013, "2015" = 2015)
   # maybe remove 2011 & 2015 once we can add "2017"=2017 to avoid start and end values overlapping
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) 
+  start_years <- list("2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
   # prob need to remove 2010 & 2014
@@
   root_params <- filter(root_params, age != "imm")
-  if (any(root_params$model_type == "split")) { root_params <- filter(root_params, model_type != "total") }
+  if (any(root_params$model_type == "split")) {
+    root_params <- filter(root_params, model_type != "total")
+  }
 }
@@
 
-if (global_only) {
- root_params <- root_params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
- root_params <- filter(root_params, global == "T")
-} else {
- if (model_ssid == 4) {
-   root_params <- root_params %>% mutate(survey4 = ifelse(grepl("4", model), "T", "F"))
-   root_params <- filter(root_params, survey4 == "T")
- }
-}
-
 if (climate == "temperature") {
-
   temp_threshold <- params$threshold
@@
   if (biomass_threshold == "75") {
-  .upper <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
+    .upper <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75
+  }
   if (biomass_threshold == "50") {
-  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 }
+    .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50
+  }
   if (biomass_threshold == "25") {
-  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25 }
+    .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25
+  }
   .upper <- na.omit(.upper)
@@
   upper_thresholds <- round(.upper[1], digits = 2)
-   if (is.na(upper_thresholds)) {stop(print("Temp maximum not modelled"))}
+  if (is.na(upper_thresholds)) {
+    stop(print("Temp maximum not modelled"))
+  }
 
@@
   if (biomass_threshold == "75") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
+    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75
+  }
   if (biomass_threshold == "50") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
+    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50
+  }
   if (biomass_threshold == "25") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }
+    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25
+  }
 
@@
   lower_thresholds <- round(.lower[1], digits = 2)
-  if (is.na(lower_thresholds)) {stop(print("DO minimum not modelled"))}
+  if (is.na(lower_thresholds)) {
+    stop(print("DO minimum not modelled"))
+  }
   lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model
@@
   if (biomass_threshold == "75") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
+    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75
+  }
   if (biomass_threshold == "50") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
+    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50
+  }
   if (biomass_threshold == "25") {
-  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }
+    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25
+  }
   .lower <- na.omit(.lower)
@@
 vocc_by_year <- lapply(start_years, function(start_time) {
-
   try({
@@
 trimmed_vectors <- lapply(vocc_by_year, function(vocc) {
-
   vectors <- trim_vector_data(vocc, variable_names,
@@
 trimmed_vector_plots <- lapply(trimmed_vectors, function(vectors) {
-
   try({
-
     start_time <- vectors$start_time[1]
@@
         fill_col = "diff",
-    fill_label = paste("Change from\n", 
-      start_time, "to", start_time + delta_t_total, ""),
+        fill_label = paste(
+          "Change from\n",
+          start_time, "to", start_time + delta_t_total, ""
+        ),
         raster_alpha = 1,
@@
         fill_col = "temp_diff",
-    fill_label = paste("Change from\n", 
-      start_time, "to", start_time + delta_t_total, ""),
+        fill_label = paste(
+          "Change from\n",
+          start_time, "to", start_time + delta_t_total, ""
+        ),
         raster_alpha = 1,
@@
         fill_col = "do_diff",
-    fill_label = paste("Change from\n", 
-      start_time, "to", start_time + delta_t_total, ""),
+        fill_label = paste(
+          "Change from\n",
+          start_time, "to", start_time + delta_t_total, ""
+        ),
         raster_alpha = 1,
@@
           "or temperature of ", optimal_values[2], "plus", max_thresholds[2],
-      " over ", start_time, "-", start_time + delta_t_total, ""))
+          " over ", start_time, "-", start_time + delta_t_total, ""
+        ))
       )
@@
 png(
-  file = paste0("figs/", spp,
+  file = paste0(
+    "figs/", spp,
     "/vocc-", min_string, "to", max_string, "-",
@@
   ncol = 2,
-  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", trim_models, ")"), gp=grid::gpar(fontsize=16))
+  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", trim_models, ")"), gp = grid::gpar(fontsize = 16))
 )
@@
     "data/", spp, "/all-predictions-", spp, covs,
-    "-imm-biomass-prior-", priors, ".rds"))
+    "-imm-biomass-prior-", priors, ".rds"
+  ))
 }
@@
     "data/", spp, "/all-predictions-", spp, covs,
-    "-mat-biomass-prior-", priors, ".rds"))
+    "-mat-biomass-prior-", priors, ".rds"
+  ))
 }
@@
 ```{r}
-#alldata <- lapply(biomass_list, function(biomass_predictions) {
+# alldata <- lapply(biomass_list, function(biomass_predictions) {
 
@@
 
-  biomass <- left_join(biomass_change, biomass_covs, by = c("x" = "X", "y"="Y"))
+biomass <- left_join(biomass_change, biomass_covs, by = c("x" = "X", "y" = "Y"))
 
 # keep just one row for each cell that exceeded the end climate threshold
-  source_cells <- all_vectors %>% filter(distance > (-dist_intercept)) %>% 
+source_cells <- all_vectors %>%
+  filter(distance > (-dist_intercept)) %>%
   distinct(icell, start_time, .keep_all = TRUE) %>%
@@
 source_biomass <- left_join(source_cells, biomass,
-    by = c("icell", "x", "y", "start_time")) %>%
-    mutate(cell_type = "source") %>% filter(!is.na(start_density))
+  by = c("icell", "x", "y", "start_time")
+) %>%
+  mutate(cell_type = "source") %>%
+  filter(!is.na(start_density))
 
 # identify all possible control cells have been given a distance == - dist_intercept
-  control_cells <- all_vectors %>% filter(distance == (-dist_intercept)) %>% 
+control_cells <- all_vectors %>%
+  filter(distance == (-dist_intercept)) %>%
   distinct(icell, start_time, .keep_all = TRUE) %>%
-    mutate(distance = (- dist_intercept)) %>% 
+  mutate(distance = (-dist_intercept)) %>%
   select(-target_X, -target_Y, -target_values, -tid)
@@
 control_biomass <- left_join(control_cells, biomass,
-    by = c("icell", "x", "y", "start_time")) %>%
-    mutate(cell_type = "control") %>% filter(!is.na(start_density))
+  by = c("icell", "x", "y", "start_time")
+) %>%
+  mutate(cell_type = "control") %>%
+  filter(!is.na(start_density))
 
 alldata <- rbind(source_biomass, control_biomass)
-  alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
+alldata <- mutate(alldata, treat = ifelse(cell_type == "source", 1, 0)) %>%
   select(-var_2_min, -var_2_max, -mean_target_X, -mean_target_Y) %>%
@@
 # alldata
-#})
+# })
 ```
@@
 ```{r}
-#rm(vocc_by_sp)
+# rm(vocc_by_sp)
 # try({
@@
 
-#FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?
+# FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?
 
@@
   alldata <- alldata %>% mutate((!!start_value.st) :=
-        arm::rescale(alldata[[start_value]]) )
-  #(UQ(rlang::sym(paste0(variable_names[j], "_s")))))
+    arm::rescale(alldata[[start_value]]))
+  # (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
 }
 
-if (!exists("vocc_by_sp")){
+if (!exists("vocc_by_sp")) {
   # matched <- lapply(alldata, function(alldata) {
   ## trim non-overlapping values to reduce size of dataset
-  control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.01)
-  source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.01)
+  control_min <- quantile(alldata$start_density[alldata$treat == "0"], 0.01)
+  source_min <- quantile(alldata$start_density[alldata$treat == "1"], 0.01)
   # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 1)
@@
   alldata <- filter(alldata, start_density >= group_min)
-  #%>% filter(start_density <= group_max)
+  # %>% filter(start_density <= group_max)
 
@@
   alldata$depth2 <- alldata$depth^2
-  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
+  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density)) * 5) / 5
   alldata$depth_st <- arm::rescale(log(alldata$depth))
-  alldata$depth.st <- round(arm::rescale(log(alldata$depth))*5)/5
+  alldata$depth.st <- round(arm::rescale(log(alldata$depth)) * 5) / 5
   alldata$muddy.st <- arm::rescale(alldata$muddy)
@@
     discard = "both",
-    data = alldata)
+    data = alldata
+  )
   # matched
@@
 ```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
-if (!exists("vocc_by_sp")){
+if (!exists("vocc_by_sp")) {
   match_plots <- plot(matched, interactive = F)
@@
 }
-
 ```
@@
 ```{r}
-if (!exists("vocc_by_sp")){
-#vocc_by_sp <- lapply(matched, function(matched) {    
+if (!exists("vocc_by_sp")) {
+  # vocc_by_sp <- lapply(matched, function(matched) {
   mdata <- MatchIt::match.data(matched, distance = "pscore")
@@
   sources$origobs <- row.names(sources)
-  sources$matchobs <- matched$match.matrix[,1]
+  sources$matchobs <- matched$match.matrix[, 1]
 
@@
   controls$origobs <- row.names(controls)
-  controls$matchobs <- matched$match.matrix[,1]
+  controls$matchobs <- matched$match.matrix[, 1]
 
-  mdata <- rbind(sources, controls) %>% tidyr::drop_na() %>% filter(matchobs!="-1")
+  mdata <- rbind(sources, controls) %>%
+    tidyr::drop_na() %>%
+    filter(matchobs != "-1")
 
   matched_time0 <- mdata %>% mutate(after = 0, density = start_density)
-  #%>% dplyr::select(-start_density, -end_density)
+  # %>% dplyr::select(-start_density, -end_density)
   matched_time1 <- mdata %>% mutate(after = 1, density = end_density)
-  #%>% dplyr::select(-start_density, -end_density)
+  # %>% dplyr::select(-start_density, -end_density)
 
@@
         mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) /
-          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE))*2),
-      (!!std.var2) := (( UQ(rlang::sym(paste0(variable_names[j]))) - 
-          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) ) / 
-          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)*2))^2
+        (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) * 2),
+      (!!std.var2) := ((UQ(rlang::sym(paste0(variable_names[j]))) -
+        mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) /
+        (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) * 2))^2
     )
@@
   data$vect_dist2 <- data$vect_dist^2
-  data <- mutate(data, time = ifelse(after==1, "end", "begin"))
+  data <- mutate(data, time = ifelse(after == 1, "end", "begin"))
   data$matchobs <- as.factor(data$matchobs)
@@
   #  data
-#})
+  # })
   vocc_by_sp <- list()
@@
 ```{r}
-#biomass_threshold <- "50"
+# biomass_threshold <- "50"
 optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
-#climate <- "temperature"
+# climate <- "temperature"
 try({
@@
 data <- vocc_by_sp[[1]]
-df<- cbind(species, mature, ssid, data)
+df <- cbind(species, mature, ssid, data)
 species <- species[[1]]
@@
 
-if (climate=="temperature") {
+if (climate == "temperature") {
   dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))
@@
 }
-if (climate=="temperature") {
-  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", max_string,"/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
-  ))
+if (climate == "temperature") {
+  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold, "/", max_string, "/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
 }
 
-if (climate=="do") {
+if (climate == "do") {
   dir.create(file.path("data/_all", climate_string))
@@
 }
-if (climate=="do") {
-  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", min_string,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
-  ))
+if (climate == "do") {
+  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold, "/", min_string, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
 }
 
-if (climate=="do and temperature") {
+if (climate == "do and temperature") {
   dir.create(file.path("data/_all", climate_string))
@@
 
-if (climate=="do and temperature") {
-  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
-  ))
+if (climate == "do and temperature") {
+  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
 }
@@
 ```{r}
-#FIXME: SAVE PLOTS?
+# FIXME: SAVE PLOTS?
 
@@
   group_by(start_time, cell_type) %>%
-  tally() %>% filter(cell_type=="source")
+  tally() %>%
+  filter(cell_type == "source")
 data_depth <- data %>%
@@
 
-p1 <- ggplot(data, 
-#ggplot(filter(data, as.numeric(matchobs) <= 550), 
-  aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
-  geom_point(alpha=0.8, size = 1) +
-  scale_shape_manual(values= c(0, 15)) +
-  scale_colour_viridis_d(option="C", guide = FALSE) +
+p1 <- ggplot(
+  data,
+  # ggplot(filter(data, as.numeric(matchobs) <= 550),
+  aes(y = Y, x = X, shape = as.factor(treat), colour = as.factor(matchobs))
+) +
+  geom_point(alpha = 0.8, size = 1) +
+  scale_shape_manual(values = c(0, 15)) +
+  scale_colour_viridis_d(option = "C", guide = FALSE) +
   facet_wrap(~start_time) +
@@
 lapply(variable_names, function(j) {
-ggplot(data, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
-  geom_point(alpha=0.35) +
-  scale_colour_viridis_d(option="C") +
-  facet_grid(start_time~cell_type) +
+  ggplot(data, aes(y = log_density, x = UQ(rlang::sym(paste(j))), colour = time)) +
+    geom_point(alpha = 0.35) +
+    scale_colour_viridis_d(option = "C") +
+    facet_grid(start_time ~ cell_type) +
     theme_bw()
@@
 p2 <- ggplot(data, aes(std_log_depth, fill = cell_type)) +
-  geom_density(alpha=0.5) +
-  scale_fill_viridis_d(option="C") +
+  geom_density(alpha = 0.5) +
+  scale_fill_viridis_d(option = "C") +
   facet_wrap(~start_time) +
@@
 p3 <- ggplot(data, aes(log_density, fill = cell_type)) +
-  geom_density(alpha=0.5) +
-  scale_fill_viridis_d(option="C") +
-  facet_grid(start_time~time, scales = 'fixed') +
-  theme_bw() + theme(legend.position='none')
+  geom_density(alpha = 0.5) +
+  scale_fill_viridis_d(option = "C") +
+  facet_grid(start_time ~ time, scales = "fixed") +
+  theme_bw() + theme(legend.position = "none")
 p3
@@
 p4 <- ggplot(data, aes(log_density, fill = cell_type)) +
-  geom_histogram(alpha=0.45) +
-  scale_fill_viridis_d(option="C") +
-  facet_grid(time~start_time) +
-  theme_bw() + theme(legend.position='none')
+  geom_histogram(alpha = 0.45) +
+  scale_fill_viridis_d(option = "C") +
+  facet_grid(time ~ start_time) +
+  theme_bw() + theme(legend.position = "none")
 p4
@@
   group_by(start_time, cell_type) %>%
-  tally() %>% filter(cell_type=="source")
+  tally() %>%
+  filter(cell_type == "source")
 
-p5 <- ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
-  geom_boxplot(alpha=0.5) +
-  scale_fill_viridis_d(option="C") +
-  geom_text(data = dat_summary,
-            aes(cell_type, min(data$log_density)+1, label = n), hjust = 1.25) +
-  facet_grid(time~start_time) +
-  theme_bw() + theme(legend.position='none')
+p5 <- ggplot(data, aes(y = log_density, x = cell_type, fill = cell_type)) +
+  geom_boxplot(alpha = 0.5) +
+  scale_fill_viridis_d(option = "C") +
+  geom_text(
+    data = dat_summary,
+    aes(cell_type, min(data$log_density) + 1, label = n), hjust = 1.25
+  ) +
+  facet_grid(time ~ start_time) +
+  theme_bw() + theme(legend.position = "none")
 p5
@@
 png(
-  file = paste0("figs/", spp,
+  file = paste0(
+    "figs/", spp,
     "/matches-", min_string, "to", max_string, "-",
@@
 gridExtra::grid.arrange(
-  p1,p2,p3,p4,p5,
+  p1, p2, p3, p4, p5,
   ncol = 2,
-  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
+  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp = grid::gpar(fontsize = 16))
 )

### 2019-10-01 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
update thresholds in df as 'var_1_min', 'var_2_min', etc... make sure data doesn't hold over when looping between species; style

@@
 author: "Philina English"
-date: '2019-08-24'
+date: `r Sys.Date()`
 output: html_document
 params: 
-    species: "Yelloweye Rockfish" 
-    region: "West Coast Vancouver Island"
-    age: "mature"
+    species: "Pacific Ocean Perch" 
+    region: "both odd year surveys"
+    age: "imm"
     biomass_threshold: "25"
     climate: "do"
-    threshold: c(0.25) 
+    threshold: 0.25
 ---
@@
 library("gfranges")
+ggplot2::theme_set(gfplot::theme_pbs())
 ```
@@
 input_cell_size <- 2
-dist_intercept <- input_cell_size / 2
+dist_intercept <- 0 # input_cell_size / 2
 scale_fac <- 1
@@
 
-root_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
+all_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
 
-root_params <- root_params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
+all_params <- all_params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
 
 if (age == "imm") {
-  root_params <- filter(root_params, age == "imm")
+  all_params <- filter(all_params, age == "imm")
 } else {
-  root_params <- filter(root_params, age != "imm")
-  if (any(root_params$model_type == "split")) {
-    root_params <- filter(root_params, model_type != "total")
+  all_params <- filter(all_params, age != "imm")
+  if (any(all_params$model_type == "split")) {
+    all_params <- filter(all_params, model_type != "total")
   }
@@
 
-root_params <- root_params[root_params$species == species, ]
+# rm(root_params)
+root_params <- all_params[all_params$species == species, ]
 glimpse(root_params)
@@
 ```{r echo=FALSE}
+if (!exists("vocc")) {
   all_vector_plots <- lapply(vocc_by_year, function(vocc) {
@@
         vocc_plot_temp, vocc_plot_do,
-      ncol = 2,
-      top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
+        ncol = 2 # ,
+        # top = grid::textGrob(paste("do decline of", lower_change[1], "or temperature increase of", upper_change[2], " over ", start_time, "-", start_time + delta_t_total, ""))
       )
@@
   })
+
+  png(
+    file = paste0(
+      "figs/vocc-", climate_string, min_string, "to", max_string, "-", ssid_string, ".png"
+    ),
+    res = 600,
+    units = "in",
+    width = 8.5,
+    height = 11
+  )
+  gridExtra::grid.arrange(
+    grobs = all_vector_plots,
+    ncol = 2,
+    top = grid::textGrob(paste("Vectors for a DO decline of", lower_change[1], "and/or temperature increase of", upper_change[2], ""))
+  )
+  dev.off()
+
   print(all_vector_plots)
+}
 ```
@@
 ```{r}
+rm(trimmed_vectors)
 trimmed_vectors <- lapply(vocc_by_year, function(vocc) {
@@
 all_vectors <- do.call("rbind", trimmed_vectors)
+
+## If VOCC change is a useful column?
+# all_vectors$var_1_changeA <- all_vectors$var_1_min
+# all_vectors$var_2_changeA <- all_vectors$var_2_min
+# all_vectors$var_1_changeB <- all_vectors$var_1_max
+# all_vectors$var_2_changeB <- all_vectors$var_2_max
+
+all_vectors$var_1_min <- lower_thresholds[1]
+all_vectors$var_2_min <- lower_thresholds[2]
+all_vectors$var_1_max <- lower_thresholds[1]
+all_vectors$var_2_max <- lower_thresholds[2]
 ```
@@
 covs <- "-trawled-ssid"
-
+rm(biomass_predictions)
 if (age == "imm") {
@@
 # keep just one row for each cell that exceeded the end climate threshold
+# FIXME: this code could/maybe should be rewritten with an if_else mutate and then distinct...
 source_cells <- all_vectors %>%
@@
 ```{r}
-# rm(vocc_by_sp)
+rm(vocc_by_sp)
 # try({
@@
 
-if (!exists("vocc_by_sp")) {
+# if (!exists("vocc_by_sp")) {
 # matched <- lapply(alldata, function(alldata) {
@@
 alldata$depth2 <- alldata$depth^2
+alldata$depth_st <- arm::rescale(log(alldata$depth))
+# alldata$muddy.st <- arm::rescale(alldata$muddy)
+# alldata$any_rock.st <- arm::rescale(alldata$any_rock)
+
+# bin some variables for exact matching
+alldata$depth.st <- round(arm::rescale(log(alldata$depth)) * 5) / 5
 alldata$start_density.st <- round(arm::rescale(log(alldata$start_density)) * 5) / 5
-  alldata$depth_st <- arm::rescale(log(alldata$depth))
-  alldata$depth.st <- round(arm::rescale(log(alldata$depth)) * 5) / 5
-  alldata$muddy.st <- arm::rescale(alldata$muddy)
-  alldata$any_rock.st <- arm::rescale(alldata$any_rock)
 
@@
 # })
-}
+# }
 ```
@@
 ```{r}
-if (!exists("vocc_by_sp")) {
+# if (!exists("vocc_by_sp")) {
 # vocc_by_sp <- lapply(matched, function(matched) {
@@
 ))
-}
+# }
 ```
@@
 # climate <- "temperature"
-try({
-  vocc_by_sp <- readRDS(paste0(
-    "data/", spp,
-    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
-  ))
-})
+# try({
+#   vocc_by_sp <- readRDS(paste0(
+#     "data/", spp,
+#     "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
+#   ))
+# })
 
@@
   scale_fill_viridis_d(option = "C") +
-  facet_grid(start_time ~ time, scales = "fixed") +
+  facet_grid(time ~ start_time, scales = "fixed") +
   theme_bw() + theme(legend.position = "none")
@@
   ncol = 2,
-  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp = grid::gpar(fontsize = 16))
+  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", age, ")"), gp = grid::gpar(fontsize = 16))
 )

### 2019-10-02 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
filter NA columns to allow match to run; saving temperature-based matches

@@
 author: "Philina English"
-date: `r Sys.Date()`
+date: "`r Sys.Date()`"
 output: html_document
@@
     biomass_threshold: "25"
-    climate: "do"
+    climate: "temperature"
     threshold: 0.25
@@
 all_vectors$var_2_min <- lower_thresholds[2]
-all_vectors$var_1_max <- lower_thresholds[1]
-all_vectors$var_2_max <- lower_thresholds[2]
+all_vectors$var_1_max <- upper_thresholds[1]
+all_vectors$var_2_max <- upper_thresholds[2]
 ```
@@
 alldata <- mutate(alldata, treat = ifelse(cell_type == "source", 1, 0)) %>%
-  select(-var_2_min, -var_2_max, -mean_target_X, -mean_target_Y) %>%
+  select(-mean_target_X, -mean_target_Y) %>%
   rename(vect_dist = distance)
@@
 
+if(is.na(alldata$var_1_max[2])) {
+  alldata <- alldata %>% select(-var_1_max)
+}
+if(is.na(alldata$var_1_min[2])) {
+  alldata <- alldata %>% select(-var_1_min)
+}
+if(is.na(alldata$var_2_max[2])) {
+  alldata <- alldata %>% select(-var_2_max)
+}
+if(is.na(alldata$var_2_min[2])) {
+  alldata <- alldata %>% select(-var_2_min)
+}
 # alldata
@@
 # FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?
+
+# alldata[ , colSums(is.na(alldata)) == 0]
+# alldata <- alldata %>% select(-var_1_min)
 
@@
 if (climate == "temperature") {
+  dir.create(file.path("data/_all", climate_string))
+  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))
   dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))

### 2019-10-02 analysis/VOCC/3-fixed-var-response-plots.Rmd
remove root lines for increasing DO and decreasing temp; change binning for raw density plots

@@
 ```{r}
-
 # # Species run so far...
@@
         geom_vline(xintercept=(do[[i]]$lower_do_50), linetype="longdash", alpha = 0.75) + 
-        geom_vline(xintercept=(do[[i]]$upper_do_50), linetype="longdash", alpha = 0.75) +
+        #geom_vline(xintercept=(do[[i]]$upper_do_50), linetype="longdash", alpha = 0.75) +
         geom_vline(xintercept=(do[[i]]$lower_do_75), linetype="dotted", alpha = 0.5) + 
-        geom_vline(xintercept=(do[[i]]$upper_do_75), linetype="dotted", alpha = 0.5) +
-        geom_vline(xintercept=(do[[i]]$lower_do_25), linetype="dotted", alpha = 0.5) + 
-        geom_vline(xintercept=(do[[i]]$upper_do_25), linetype="dotted", alpha = 0.5) 
+        #geom_vline(xintercept=(do[[i]]$upper_do_75), linetype="dotted", alpha = 0.5) +
+        geom_vline(xintercept=(do[[i]]$lower_do_25), linetype="dotted", alpha = 0.5) #+ 
+        #geom_vline(xintercept=(do[[i]]$upper_do_25), linetype="dotted", alpha = 0.5) 
    } else {
@@
         geom_vline(xintercept=td[[i]]$optimal_t, size= 1.2, colour = "darkgrey") + 
-        geom_vline(xintercept=td[[i]]$lower_t_50, linetype="longdash", alpha = 0.75) + 
+        #geom_vline(xintercept=td[[i]]$lower_t_50, linetype="longdash", alpha = 0.75) + 
         geom_vline(xintercept=td[[i]]$upper_t_50, linetype="longdash", alpha = 0.75) +
-        geom_vline(xintercept=td[[i]]$lower_t_75, linetype="dotted", alpha = 0.5) + 
+        #geom_vline(xintercept=td[[i]]$lower_t_75, linetype="dotted", alpha = 0.5) + 
         geom_vline(xintercept=td[[i]]$upper_t_75, linetype="dotted", alpha = 0.5) +
-        geom_vline(xintercept=td[[i]]$lower_t_25, linetype="dotted", alpha = 0.5) + 
+        #geom_vline(xintercept=td[[i]]$lower_t_25, linetype="dotted", alpha = 0.5) + 
         geom_vline(xintercept=td[[i]]$upper_t_25, linetype="dotted", alpha = 0.5)
@@
 odata <- data %>% filter(!is.na(do_mlpl)) %>% 
-  mutate(ad_density_class = ifelse(adult_density==0, 0, (round(log10((adult_density*10000)+1)*2)+1)),
-    im_density_class = ifelse(imm_density==0, 0, (round(log10((imm_density*10000)+1)*2)+1))) %>% 
+  mutate(ad_density_class = ifelse(adult_density==0, 0, round(log10((adult_density*100000)+1))),
+    im_density_class = ifelse(imm_density==0, 0, round(log10((imm_density*10000000)+1)))) %>% 
   group_by(ad_density_class) %>% mutate(ad_mean_density = mean(adult_density, na.rm = TRUE)) %>% ungroup() %>%
@@
 tdata <- data %>% filter(!is.na(temp)) %>% 
-  mutate(ad_density_class = ifelse(adult_density==0, 0, (round(log10((adult_density*10000)+1)*2)+1)),
-    im_density_class = ifelse(imm_density==0, 0, (round(log10((imm_density*10000)+1))+1)*2)) %>% 
+  mutate(ad_density_class = ifelse(adult_density==0, 0, round(log10((adult_density*100000)+1))),
+    im_density_class = ifelse(imm_density==0, 0, round(log10((imm_density*10000000)+1)))) %>% 
   group_by(ad_density_class) %>% mutate(ad_mean_density = mean(adult_density, na.rm = TRUE)) %>% ungroup() %>%
@@
   geom_density(position = "stack") +
-  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs, begin=0, end=0.5) + 
   geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
@@
   geom_density(position = "fill") +
-  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs, begin=0, end=0.5) + 
   geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +

### 2019-10-04 analysis/VOCC/1-split-biomass-by-maturity.Rmd
rename 1-split-biomass-by-maturity

--- 2019-09-10 analysis/VOCC/1-split-biomass-by maturity.Rmd
@@
 # species <- "Canary Rockfish" # schooling, winter-birthing
+species <- "Shortraker Rockfish"
 
@@
 ```{r}
-# region <- "All synoptic surveys"
- region <- "Both odd year surveys"
+region <- "All synoptic surveys"
+# region <- "Both odd year surveys"
 # region <- "West Coast Vancouver Island"
@@
                       echo=FALSE, warning=FALSE, message=FALSE)
-```
-
-```{r, message=FALSE, warning=FALSE}
 library(dplyr)
@@
 library(gfranges)
+```
 
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+```{r, message=FALSE, warning=FALSE}
+
+#fish <- readRDS(paste0("raw/bio-data-", spp, ""))
 
@@
     width = 8, 
-    height = 10
+    height = 6
   )

### 2019-10-08 analysis/VOCC/1-split-biomass-by-maturity.Rmd
renderable versions of Rmds

@@
 ---
-title: "Spatiotemporal models of species biomass split by maturity class"
+title: "Split `r paste(params$species)` biomass by maturity class"
 author: "Philina English"
-date: '2019-07-18'
+date: "`r Sys.Date()`"
 output: html_document
 params: 
-    species: "Arrowtooth Flounder"
-
+    species: "Curlfin Sole"
+    region: "All synoptic surveys"
+    #region: "Both odd year surveys"
 ---
 
-Choose one value for each parameter:
-
-Species
-```{r}
-species <- params$species
-
-# # Species run so far...
-# species <- "Arrowtooth Flounder"
-# species <- "Petrale Sole"
-# species <- "English Sole"
-# species <- "Dover Sole" # maturity data error for WCHG
-# species <- "Southern Rock Sole"
-# species <- "Flathead Sole"
-
-# species <- "Silvergray Rockfish"
-# species <- "Quillback Rockfish" # summer-birthing
-species <- "Yelloweye Rockfish" # summer-birthing, overfished, 
-# species <- "Bocaccio" # winter-birthing, overfished
-# species <- "Sharpchin Rockfish"
-# species <- "Splitnose Rockfish"
-
-# species <- "Rougheye/Blackspotted Rockfish Complex"
-# species <- "Redbanded Rockfish"
-# species <- "Greenstriped Rockfish"
-# # species <- "Copper Rockfish" # small sample
-# species <- "Darkblotched Rockfish"
-# # species <- "Shortbelly Rockfish" # small sample
-
-# species <- "Pacific Ocean Perch" # schooling
-# species <- "Widow Rockfish" # schooling
-# species <- "Yellowtail Rockfish" # schooling
-# species <- "Canary Rockfish" # schooling, winter-birthing
-species <- "Shortraker Rockfish"
-
-# species <- "Longspine Thornyhead"
-# species <- "Shortspine Thornyhead"
-
-# species <- "Walleye Pollock"
-# species <- "Pacific Cod"
-# species <- "Sablefish"
-# species <- "Lingcod"
-# species <- "Pacific Hake"
-
-# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-# species <- "Longnose Skate"
-# species <- "Big Skate"
-# species <- "Spotted Ratfish"
-# # species <- "Sandpaper Skate" # small sample
-# # species <- "Brown Cat Shark" # small sample
-```
-
-Region 
-```{r}
-region <- "All synoptic surveys"
-# region <- "Both odd year surveys"
-# region <- "West Coast Vancouver Island"
-# region <- "West Coast Haida Gwaii"
-```
-
-Run all subsequent code...
 ```{r global_options, include=FALSE}
-# covs <- gsub("\\+", "-", covariates)
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-# folder to hold figs for this species
-dir.create(file.path("figs", spp))
-dir.create(file.path("data", spp))
-
 knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r params}
+species <- params$species
+region <- params$region
 
-#fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+
+# folder to hold figs for this species
+dir.create(file.path("figs", spp))
+dir.create(file.path("data", spp))
 
@@
 rm(maturity)
-# try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
+
+#try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
 

### 2019-10-08 analysis/VOCC/2-biomass-depth-only-model.Rmd
renderable versions of Rmds

@@
 ---
-title: "Spatiotemporal models of species biomass with time-varying depth and trawl footprint, and optional substrate variables"
+title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
 author: "Philina English"
-date: '2019-07-18'
+date: "`r Sys.Date()`"
 output: html_document
 params: 
-    species: "Arrowtooth Flounder"
-
+    species: "Copper Rockfish"
+    region: "All synoptic surveys"
+    covariates: "+trawled+as.factor(ssid)" # additional fixed-effect covariates
+    covs: "-trawled-ssid" # covariate string used to name model rds 
 ---
 
-Choose one value for each parameter:
-
-Species
-```{r}
-
-```
-
-Region 
-```{r}
-region <- "All synoptic surveys"
-# region <- "Both odd year surveys"
-# region <- "West Coast Vancouver Island"
-# region <- "West Coast Haida Gwaii"
-```
-
-Choose model details
-```{r}
-priors <- FALSE
-# priors <- TRUE
-covariates <- "+trawled+as.factor(ssid)"
-covs <- "-trawled-ssid"
-
-# covariates <- "+muddy+any_rock"
-# covariates <- "+muddy+mixed+rocky"
-# covariates <- "+mixed+rocky"
-# covariates <- "+rocky"
-# covariates <- "+mixed_scaled+rocky_scaled+muddy_scaled"
-
-# covs <- gsub("\\+", "-", covariates)
-# covs <- gsub("\\+", "-", "+muddy+any_rock")
-# covs <- gsub("\\+", "-", "-variable-substrate")
-# covs <- gsub("\\+", "-", "-all-substrates")
-```
-
-Run all subsequent code...
 ```{r global_options, include=FALSE}
-# covs <- gsub("\\+", "-", covariates)
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-# folder to hold figs for this species
-dir.create(file.path("figs", spp))
-dir.create(file.path("data", spp))
-
-knitr::opts_chunk$set(
-  fig.width = 11, fig.height = 8.5,
-  fig.path = paste0("figs/", spp, "/"),
-  echo = FALSE, warning = FALSE, message = FALSE
-)
-```
-
-```{r, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
+                      #fig.path=paste0("figs/", spp, "/"),
+                      echo=FALSE, warning=FALSE, message=FALSE)
 library(dplyr)
@@
 library(gfranges)
+```
 
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+```{r params}
+species <- params$species
+region <- params$region
+covariates <- params$covariates
+covs <- params$covs
+priors <- FALSE
+
+paste("region =", region)
+paste("model covariates =", covariates)
+paste("model label =", covs)
+paste("priors =", priors)
+```
+
+```{r, message=FALSE, warning=FALSE}
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+
+# folder to hold figs for this species
+dir.create(file.path("figs", spp))
+dir.create(file.path("data", spp))
 
@@
 
-  starttime1 <- Sys.time()
   adult_biomass <- sdmTMB::sdmTMB(data,
@@
   )
-  endtime1 <- Sys.time()
-  time1 <- round(starttime1 - endtime1)
 
@@
 
-  starttime2 <- Sys.time()
   imm_biomass <- sdmTMB::sdmTMB(data,
@@
   )
-  endtime2 <- Sys.time()
-  time2 <- round(starttime2 - endtime2)
 
@@
   )
+
 } else {
-  starttime3 <- Sys.time()
+
   dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
@@
   )
-  endtime3 <- Sys.time()
-  time3 <- round(starttime3 - endtime3)
 
@@
 ```{r}
-# covs <- gsub("\\+", "-", covariates)
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+#fish <- readRDS(paste0("raw/bio-data-", spp, ""))
 
@@
 
-if (length(unique(fish$maturity_code)) > 2) {
+if (any(names(data) == "adult_density")) {
+#if (length(unique(fish$maturity_code)) > 2) {
   nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-  adult_biomass <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
+  adult_biomass <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
 
@@
   adult_predictions <- predict(adult_biomass, newdata = nd)
-  saveRDS(adult_predictions, file = paste0(
-    "data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
-  ))
+  saveRDS(adult_predictions, file = paste0("data/", spp,
+    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"))
 
-  imm_biomass <- readRDS(paste0(
-    "data/", spp,
+  imm_biomass <- readRDS(paste0("data/", spp,
     "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
   imm_predictions <- predict(imm_biomass, newdata = nd)
-  saveRDS(imm_predictions, file = paste0(
-    "data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
-  ))
+  saveRDS(imm_predictions, file = paste0("data/", spp,
+    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"))
+
 } else {
-  rm(adult_predictions)
-  rm(imm_predictions)
-
-  if (!exists(predictions)) {
     nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
@@
 
-    total_biomass <- readRDS(paste0(
-      "data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
+    total_biomass <- readRDS(paste0("data/", spp,
+      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
     predictions <- predict(total_biomass, newdata = nd)
-    saveRDS(predictions, file = paste0(
-      "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
-    ))
-  }
+    saveRDS(predictions, file = paste0("data/", spp,
+      "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
 }

### 2019-10-08 analysis/VOCC/2-biomass-trends.Rmd
renderable versions of Rmds

@@
 ---
-title: "Biomass trends and figures"
+title: "`r paste(params$species)` biomass trends and figures"
 author: "Philina English"
-date: '2019-07-18'
+date: "`r Sys.Date()`"
 output: html_document
+params: 
+    species: "Lingcod"
+    region: "All synoptic surveys"
+    bind_predictions: FALSE
+    covs: "-trawled-ssid" # covariate string used to name model rds 
 ---
 
-This file calls models built by "1-biomass-models.Rmd" and stiches the predictions saved therein together.
-
-# Choose params
-
-Species
-```{r}
-
-```
-
-Region 
-```{r}
-region <- "All synoptic surveys"
-
-# region <- "Both odd year surveys"
-# region <- "West Coast Vancouver Island"
-# region <- "West Coast Haida Gwaii"
-```
-
-Choose model details
-```{r}
-update_spatial_model <- FALSE
-priors <- FALSE
-# priors <- TRUE
-# covariates <- ""
-# covariates <- "+trawled"
-# covs <- gsub("\\+", "-", covariates)
-covariates <- "+trawled+as.factor(ssid)"
-covs <- gsub("\\+", "-", "+trawled+ssid")
-# covariates <- "+muddy+mixed+rocky"
-# covariates <- "+mixed+rocky"
-# covariates <- "+muddy+any_rock"
-# covariates <- "+trawled+muddy+rocky+mixed"
-# covariates <- "+trawled+mixed+rocky"
-# covariates <- "+trawled+mixed"
-```
-
-Run all subsequent code...
 ```{r global_options, include=FALSE}
-
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-# folder to hold figs for this species
-dir.create(file.path("figs", spp))
-dir.create(file.path("data", spp))
-
-knitr::opts_chunk$set(
-  fig.width = 11, fig.height = 8.5,
-  # fig.path=paste0("figs/", spp, "/"),
-  echo = FALSE, warning = FALSE, message = FALSE
-)
-```
-
-```{r, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
+                      #fig.path=paste0("figs/", spp, "/"),
+                      echo=FALSE, warning=FALSE, message=FALSE)
 library(dplyr)
@@
 library(gfranges)
+```
 
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+```{r params}
+species <- params$species
+region <- params$region
+bind_predictions <- params$bind_predictions
+covs <- params$covs
+priors <- FALSE
+
+paste("region =", region)
+paste("ssids modelled separately =", bind_predictions)
+paste("model label =", covs)
+paste("priors =", priors)
+
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
 
@@
 
-Bind together predictions from different models
-```{r}
+Load model predictions
+```{r results = "hide"}
 rm(adult_predictions)
@@
 rm(pred16_all)
-# if (isFALSE(update_spatial_model)) {
+
 if (region != "All synoptic surveys") {
-  try(adult_predictions <- readRDS(paste0(
-    "data/", spp,
+  try(adult_predictions <- readRDS(paste0("data/", spp,
     "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
   )))
-  try(imm_predictions <- readRDS(paste0(
-    "data/", spp,
+  try(imm_predictions <- readRDS(paste0("data/", spp,
     "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
@@
 
-    try(predictions <- readRDS(paste0(
-      "data/", spp,
+    try(predictions <- readRDS(paste0("data/", spp,
       "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
@@
 } else {
-  try({
-    adult_predictions <- readRDS(paste0(
-      "data/", spp,
+
+  try({adult_predictions <- readRDS(paste0("data/", spp,
     "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
@@
   })
-  try({
-    imm_predictions <- readRDS(paste0(
-      "data/", spp,
+  try({imm_predictions <- readRDS(paste0("data/", spp,
     "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
@@
 
-  if (!exists("adult_predictions")) {
   # OR bind together predictions from different models
-    try({
-      pred1n3 <- readRDS(paste0(
-        "data/", spp,
+  if (bind_predictions) {
+    try({pred1n3 <- readRDS(paste0("data/", spp,
       "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"
@@
     })
-    try({
-      pred4 <- readRDS(paste0(
-        "data/", spp,
+    try({pred4 <- readRDS(paste0("data/", spp,
       "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"
@@
     })
-    try({
-      pred16 <- readRDS(paste0(
-        "data/", spp,
+    try({pred16 <- readRDS(paste0("data/", spp,
       "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"
@@
 
-    try({
-      pred1n3_imm <- readRDS(paste0(
-        "data/", spp,
+    try({pred1n3_imm <- readRDS(paste0("data/", spp,
       "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"
@@
     })
-    try({
-      pred4_imm <- readRDS(paste0(
-        "data/", spp,
-        "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"
-      ))
+    try({pred4_imm <- readRDS(paste0("data/", spp,
+      "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))
     })
-    try({
-      pred16_imm <- readRDS(paste0(
-        "data/", spp,
-        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"
-      ))
+    try({pred16_imm <- readRDS(paste0("data/", spp,
+        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"))
     })
@@
     imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)
+  }
+  }
 
@@
 
-      try({
-        predictions <- readRDS(paste0(
-          "data/", spp,
-          "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
-        ))
+      try({predictions <- readRDS(paste0("data/", spp,
+        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
       })
@@
       if (!exists("predictions")) {
-        try({
-          pred1n3_all <- readRDS(paste0(
-            "data/", spp,
-            "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"
-          ))
+        try({pred1n3_all <- readRDS(paste0("data/", spp,
+          "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))
         })
-        try({
-          pred4_all <- readRDS(paste0(
-            "data/", spp,
-            "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"
-          ))
+        try({pred4_all <- readRDS(paste0("data/", spp,
+          "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))
         })
-        try({
-          pred16_all <- readRDS(paste0(
-            "data/", spp,
-            "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"
-          ))
+        try({pred16_all <- readRDS(paste0("data/", spp,
+          "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))
         })
@@
         if (!exists("predictions")) paste("Total biomass predictions are not available.")
-      }
-    }
       }
@@
   adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
-  imm_predictions$prop_imm <- imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
+  imm_predictions$prop_imm <- 
+    imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
 
@@
 
-# Plots and trends
-
+## Mature (or total) biomass predictions
 ```{r, error=TRUE}
@@
 
+### Split by survey (SSID)
 ```{r}
@@
 
+### Recent biomass with trawl footprint
 ```{r }
@@
 
+## Trends in biomass density
 ```{r, error=TRUE}
@@
 
+# Immature biomass density
+### if available
 ```{r, error=TRUE}
@@
 
-Trend plots
-```{r}
+Save trend plots
+```{r results = "hide"}
 png(
@@
 
-Spatiotemporal biomass plots
-```{r}
+Save spatiotemporal biomass plots
+```{r results = "hide"}
 biomass_plots <- c(p_adult, p_imm)
@@
 
-Plot medley
-```{r}
+Save plot medley
+```{r results = "hide"}
 png(

### 2019-10-08 analysis/VOCC/3-climate-models-fixed-both.Rmd
renderable versions of Rmds

@@
 ---
-title: "Spatiotemporal models of species biomass with time-varying DO"
+title: "Spatiotemporal models of `r paste(params$species)` biomass with climate covariates"
 author: "Philina English"
-date: '2019-07-18'
+date: "`r Sys.Date()`"
 output: html_document
 params: 
-    species: "Arrowtooth Flounder"
-
+    species: "Pacific Ocean Perch"
+    region: "All synoptic surveys"
+    #region: "Both odd year surveys"
+    log_temp: TRUE # should temperature be log()
+    priors: FALSE
+    covariates: "+trawled" # additional non-climate variables
+    covs: "-log-both-trawled" # string describing model covariates
 ---
 
-Choose one value for each parameter:
-
-Species
-```{r}
-# species <- "Canary Rockfish" 
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
+                     # fig.path=paste0("figs/", spp, "/"),
+                      echo=FALSE, warning=FALSE, message=FALSE)
+library(dplyr)
+library(ggplot2)
+library(gfplot)
+library(gfdata)
+library(sdmTMB)
+library(gfranges)
 ```
 
-Region 
-```{r}
- region <- "All synoptic surveys"
-  # region <- "Both odd year surveys"
-# region <- "West Coast Vancouver Island"
- #  region <- "West Coast Haida Gwaii"
-```
+```{r params}
+species <- params$species
+region <- params$region
+covariates <- params$covariates 
+covs <- params$covs
+log_temp <- params$log_temp
+priors <- params$priors
 
-Choose model details
-```{r}
-# priors <- FALSE
-priors <- TRUE
-# covariates <- "+muddy+any_rock"
-# covariates <- ""
-# covariates <- "+as.factor(ssid)"
-covariates <- "+trawled"
-# covariates <- "+muddy+mixed+rocky"
-# covariates <- "+mixed+rocky"
-# covariates <- "+trawled+muddy+rocky+mixed"
-# covariates <- "+trawled+mixed+rocky"
-# covariates <- "+trawled+mixed"
-# covs <- gsub("\\+", "-", covariates)
-# covs <- "-both-tv-depth-ssid"
-# covs <- "-both-tv-depth"
-# covs <- "-log-do"
-# covs <- "-log-do-ssid"
-covs <- "-log-do-trawled"
-
-```
-
-Run all subsequent code...
-```{r global_options, include=FALSE}
+paste("region =", region)
+paste("model covariates =", covariates)
+paste("model label =", covs)
+paste("log(temperature) =", log_temp)
+paste("priors =", priors)
 
@@
 dir.create(file.path("data", spp))
-
-knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
-                      fig.path=paste0("figs/", spp, "/"),
-                      echo=FALSE, warning=FALSE, message=FALSE)
-```
-
-```{r, message=FALSE, warning=FALSE}
-library(dplyr)
-library(ggplot2)
-library(gfplot)
-library(gfdata)
-library(sdmTMB)
-library(gfranges)
 
@@
   CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
+
+  min(CTD$temperature_c, na.rm = T) #  
+  mintemp <- 2.6
+
   data <- dplyr::left_join(data, CTD) %>% 
@@
   # scale predictors before filtering to ensure mean and SD are global
-  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), temp = temperature_c, raw_do = do_mlpl, do_mlpl = log(do_mlpl)) 
+  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), 
+     raw_do = do_mlpl, do_mlpl = log(do_mlpl))
+
+  if (log_temp) {data <- mutate(data, temp = log(temperature_c))
+  } else {
+    data <- mutate(data, temp = temperature_c)
+  }
+
   do_data <- scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl))) 
        # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
+
   data <- do_data %>% mutate(depth = raw_depth) %>% #filter(year != 2016) %>%
@@
       trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
-      depth, do_mlpl, raw_do) 
+      depth, do_mlpl, raw_do, temp, temperature_c) 
  } else {
@@
       trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
-      depth, do_mlpl, raw_do) 
+      depth, do_mlpl, raw_do, temp, temperature_c) 
  }
@@
  data <- filter(data, !(year == 2016 & ssid == 4) ) # 2016 do an outlier for survey 4
-# unique(data$year)
-# table(data$year,data$ssid)
 ```
@@
 
-    starttime1 <- Sys.time()
     adult_biomass_d <- sdmTMB::sdmTMB(data,
@@
     )
-    endtime1 <- Sys.time()
-    time1 <- round(starttime1 - endtime1)
 
@@
 
-    starttime2 <- Sys.time()
     imm_biomass_d <- sdmTMB::sdmTMB(data, 
@@
     )
-    endtime2 <- Sys.time()
-    time2 <- round(starttime2 - endtime2)
 
@@
 
-    starttime3 <- Sys.time()
     dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""))
@@
     )
-    endtime3 <- Sys.time()
-    time3 <- round(starttime3 - endtime3)
 

### 2019-10-08 analysis/VOCC/3-fixed-var-response-plots.Rmd
renderable versions of Rmds

--- 2019-09-26 analysis/VOCC/3-fixed-var-response-plots.Rmd
@@
 ---
-title: "Plot responses to model covariates to inform threshold choices for VOCC"
+title: "Plot `r paste(params$species)` responses to climate covariates to inform threshold choices for VOCC"
 author: "Philina English"
-date: '2019-07-18'
+date: "`r Sys.Date()`"
 output: html_document
+params: 
+    species: "Pacific Ocean Perch"
+    priors: FALSE # call model that used priors?
+    log_temp: TRUE
+    covs: "-log-both-trawled" # covariate string used to name model rds
+    cov_number: 2
 ---
 
-Species
-```{r}
-
-# # Species run so far...
-# species <- "Arrowtooth Flounder" # temp good, not DO, imm moving deeper?
-
-# species <- "Petrale Sole"
-# species <- "English Sole"
-# species <- "Dover Sole"
-# species <- "Southern Rock Sole"
-
-# species <- "Walleye Pollock" # survey 4 or 16 might work
-# species <- "Pacific Cod"
-
-## Sebastes
-# 
-# ## TRY BOTH - not schooling
-#  species <- "Yelloweye Rockfish" # beautiful curves, with depth change ***
-#  species <- "Bocaccio" # good curves, no depth change ... most survey-specific models fail to find minimum ***
-#  species <- "Sharpchin Rockfish" # beautiful curves, imm depth change ***
-#  species <- "Greenstriped Rockfish" # beautiful curves for adults, no depth change ***
-
-# ## TRY BOTH - schooling
-#  species <- "Pacific Ocean Perch" # beautiful curves, no depth change, schooling ***
-#  species <- "Silvergray Rockfish" # beautiful curves, no depth change, schooling ***
-
-# ## TRY JUST TEMPERATURE 
-# # species <- "Rougheye/Blackspotted Rockfish Complex" # good temp, not DO
-# # species <- "Redbanded Rockfish" # good temp, not DO, maybe moving shallower
-#  species <- "Splitnose Rockfish" # good temp, survey 4 DO ok
-
-# # species <- "Yellowtail Rockfish" #  good temp for survey 4 only, no DO, schooling
-# # species <- "Quillback Rockfish" # like warm = overall a bit flat... territorial! 
-# # species <- "Canary Rockfish" # schooling, winter-birthing, and depth change!, but curves at edges... adult1 looks ok?
-#  species <- "Widow Rockfish" # schooling, and depth change, but curves at edges
-
-# species <- "Longspine Thornyhead" # too deep to produce curves
-# species <- "Shortspine Thornyhead"
-# species <- "Sablefish" # temp might work, no DO, no consistant depth change
-# species <- "Lingcod" # curves ok, overall a bit flat
-# # species <- "Pacific Hake" # not useful
-
-# species <- "North Pacific Spiny Dogfish" # too flat, but with depth change! note: pooled maturity 
-# species <- "Longnose Skate" # do too flat, temp from 4 might be useful, seems to go deeper?
-# species <- "Big Skate"
-# species <- "Sandpaper Skate"
-# species <- "Brown Cat Shark"
-# species <- "Spotted Ratfish" # curves not useful because to abundant in shallow
-```
-
-Choose model details
-```{r}
-covariates <- ""
-# covariates <- "+muddy+mixed+rocky"
-# covariates <- "+mixed+rocky"
-#covariates <- "+muddy+any_rock"
-# covariates <- "+trawled+muddy+rocky+mixed"
-# covariates <- "+trawled+mixed+rocky"
-# covariates <- "+trawled+mixed"
-# 
-# covariates <- "+as.factor(ssid)"
-# covs <- "-both-tv-depth-ssid"
-
-# priors <- FALSE
-# covariates <- "+muddy+any_rock"
-# covs <- gsub("\\+", "-", covariates)
-priors <- TRUE
-#covs <- "-tv-depth"
-# covs <- "-both-tv-depth"
-# covs <- "-log-do"
-covs <- "-log-do-trawled"
-num_params <- 2
-```
-
-Run all subsequent code...
 ```{r global_options, include=FALSE}
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
-
-knitr::opts_chunk$set(
-  fig.width = 11, fig.height = 8.5,
-  fig.path = paste0("figs/", spp, "/"),
-  echo = FALSE, warning = FALSE, message = FALSE
-)
-```
-
-```{r, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
+                      #fig.path=paste0("figs/", spp, "/"),
+                      echo=FALSE, warning=FALSE, message=FALSE)
 library(dplyr)
@@
 
-Load Dissolved O2 models
-```{r}
+```{r params}
+species <- params$species
+priors <- params$priors
+covs <- params$covs
+log_temp <- params$log_temp
+num_params <- params$cov_number
+
+paste("priors =", priors)
+paste("model =", covs)
+paste("log(temperature) =", log_temp)
+
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+```
+
+Load models with environmental covariates
+```{r, results = "hide"}
 rm(adult, imm,
@@
 ```{r}
-# test load small sdm model to avoid crashing R
-# try({imm})
-# or maybe it's crashing when NaNs produced error goes into density function?
+model_list
 ```
@@
 ## Density curves from model coefficients
-
 #### Calculate DO curves
@@
    do[[i]]$model <- names(model_list[i])
-   do[[i]]$covs <- paste0(x2016,covs) 
+   do[[i]]$covs <- paste0(x2016, covs, priors) 
    roots <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.5)
@@
     td[[i]]$model <- names(model_list[i])
-    td[[i]]$covs <- paste0(x2016,covs) 
+    td[[i]]$covs <- paste0(x2016, covs, priors)
    roots <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.5)
-   td[[i]]$optimal_temp <- roots[["optimal"]]
-   td[[i]]$lower_t_50 <- roots[["lower_threshold"]]
-   td[[i]]$upper_t_50 <- roots[["upper_threshold"]]
+
+   if (log_temp) {
+
+   td[[i]]$optimal_temp <- exp(roots[["optimal"]])
+   td[[i]]$lower_t_50 <- exp(roots[["lower_threshold"]])
+   td[[i]]$upper_t_50 <- exp(roots[["upper_threshold"]])
    roots75 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.75)
-   td[[i]]$lower_t_75 <- roots75[["lower_threshold"]]
-   td[[i]]$upper_t_75 <- roots75[["upper_threshold"]]
+   td[[i]]$lower_t_75 <- exp(roots75[["lower_threshold"]])
+   td[[i]]$upper_t_75 <- exp(roots75[["upper_threshold"]])
    roots25 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.25)
-   td[[i]]$lower_t_25 <- roots25[["lower_threshold"]]
-   td[[i]]$upper_t_25 <- roots25[["upper_threshold"]]
+   td[[i]]$lower_t_25 <- exp(roots25[["lower_threshold"]])
+   td[[i]]$upper_t_25 <- exp(roots25[["upper_threshold"]])
+   td[[i]]$x <- exp(td[[i]]$x)
+
+   } else {
+
+   td[[i]]$optimal_temp <-  (roots[["optimal"]])
+   td[[i]]$lower_t_50 <-  (roots[["lower_threshold"]])
+   td[[i]]$upper_t_50 <-  (roots[["upper_threshold"]])
+   roots75 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.75)
+   td[[i]]$lower_t_75 <-  (roots75[["lower_threshold"]])
+   td[[i]]$upper_t_75 <-  (roots75[["upper_threshold"]])
+   roots25 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.25)
+   td[[i]]$lower_t_25 <-  (roots25[["lower_threshold"]])
+   td[[i]]$upper_t_25 <-  (roots25[["upper_threshold"]])
+   td[[i]]$x <-  (td[[i]]$x)
+
+   }
+
    temp_plots[[i]] <- 
@@
 
-Build time-varying depth plots
+### Build time-varying depth plots
+#### Model without environmental covariates
 ```{r }
 max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
+rm(depth_model_list)
+depth_only <- T
+
+if(depth_only) {
+
+try({  adult_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, "-trawled-ssid-1n3n4n16-prior-FALSE.rds"
+  ))
+  imm_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, "-trawled-ssid-1n3n4n16-prior-FALSE.rds"
+  ))
+ # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
+})
+
+    if (!exists("adult_biomass")) {
+      try({
+        total_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-total-biomass-", spp, "-trawled-ssid-1n3n4n16-prior-FALSE.rds"
+  ))
+      })
+    }
+
+if (!exists("adult_biomass")) adult_biomass <- NULL
+if (!exists("imm_biomass")) imm_biomass <- NULL
+if (!exists("total_biomass")) total_biomass <- NULL
+
+  depth_model_list <- list(adult = adult_biomass, imm = imm_biomass, total = total_biomass)
+  depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
+
+}  else {
+  depth_model_list <- model_list
+}
+
 d <- list()
 depth_plots <- list()
-for (i in seq_len(length(model_list))) {
- if(model_list[[i]]$model$convergence==0){
-  d[[i]] <- time_varying_density(model_list[[i]], predictor = "depth")
+for (i in seq_len(length(depth_model_list))) {
+ if(depth_model_list[[i]]$model$convergence==0){
+  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
 
@@
     d[[i]]$x <- exp(d[[i]]$x)
-    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, max_depth_found)) +
-      ggtitle(paste(species, names(model_list[i])))
+    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (in model without environmental variables)", xlimits = c(0, max_depth_found)) +
+      ggtitle(paste(species, names(depth_model_list[i])))
   }
@@
 
-#### Plot raw DO
+#### Model with environmental covariates
+```{r }
+max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
+
+depth_model_list <- model_list
+d <- list()
+depth_plots2 <- list()
+for (i in seq_len(length(depth_model_list))) {
+ if(depth_model_list[[i]]$model$convergence==0){
+  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
+
+  if (length(d[[i]]) == 0) {
+    depth_plots2[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } else {
+    d[[i]]$x <- exp(d[[i]]$x)
+    depth_plots2[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (controling for DO and temperature)", xlimits = c(0, max_depth_found)) +
+      ggtitle(paste(species, names(depth_model_list[i])))
+  }
+ } else {
+    depth_plots2[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } 
+}
+print(depth_plots2)
+```
+
+### Plot raw DO
 ```{r}
@@
 
-#### Plot raw temperature
+### Plot raw temperature
 ```{r}
@@
   geom_density(position = "stack") +
-  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs, begin=0, end=1) + 
   geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
@@
   geom_density(position = "fill") +
-  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
+  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs, begin=0, end=1) + 
   geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
@@
 Save DO plots
-```{r}
+```{r eval=F}
 png(
-  file = paste0("figs/", spp, "/do-fixed-", spp, x2016, covs, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/do-both-log-", spp, x2016, covs, "-priors-", priors, ".png"),
   res = 600,
@@
 
-Save Temperature plots
-```{r}
+Save temperature plots
+```{r eval=F}
 png(
-  file = paste0("figs/", spp, "/temp-fixed-", spp, x2016, covs, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/temp-both-log-", spp, x2016, covs, "-priors-", priors, ".png"),
   res = 600,
@@
 gridExtra::grid.arrange(
-  grobs = depth_plots,
+  grobs = c(depth_plots, depth_plots2),
   nrow = 4,
@@
 ## Save optimal values
-```{r }
+```{r}
 other_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
@@
   if(!is.null(tdat)) {
-    params <- full_join(tdat, odat, by=c("species", "age", "model", "covs", "ssid"))
-    all_params <- full_join(other_params, params)
+    new_params <- full_join(tdat, odat, by=c("species", "age", "model", "covs", "ssid"))
+    all_params <- full_join(other_params, new_params)
   } else {
@@
 
-Load Depth models (if using old depth only models)
-```{r eval=FALSE}
-rm(
-  adult_survey1, imm_survey1,
-  adult_survey4, imm_survey4,
-  adult_survey16, imm_survey16
-)
-priors <- FALSE
-covariates <- "+muddy+any_rock"
-covs <- gsub("\\+", "-", covariates)
-try({
-  survey <- c("SYN QCS", "SYN HS")
-  model_ssid <- c(1, 3)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey1 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  imm_survey1 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult_survey1")) adult_survey1 <- NULL
-if (!exists("imm_survey1")) imm_survey1 <- NULL
-
-try({
-  survey <- c("SYN WCVI")
-  model_ssid <- c(4)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey4 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  imm_survey4 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-if (!exists("adult_survey4")) adult_survey4 <- NULL
-if (!exists("imm_survey4")) imm_survey4 <- NULL
-
-try({
-  survey <- c("SYN WCHG")
-  model_ssid <- c(16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  adult_survey16 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  imm_survey16 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-
-if (!exists("adult_survey16")) adult_survey16 <- NULL
-if (!exists("imm_survey16")) imm_survey16 <- NULL
-
-depth_model_list <- list(
-  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
-  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
-  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
-)
-depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
-```
-
-Build depth plots
-```{r eval=FALSE}
-
-max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
-d <- list()
-depth_plots <- list()
-for (i in seq_len(length(depth_model_list))) {
-  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
-
-  if (length(d[[i]]) == 0) {
-    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
-  } else {
-    d[[i]]$x <- exp(d[[i]]$x)
-    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, max_depth_found)) +
-      ggtitle(paste(species, names(depth_model_list[i])))
-  }
-}
-print(depth_plots)
-```
-
-```{r eval=FALSE}
-png(
-  file = paste0(
-    "figs/", spp,
-    "/depth-only-", spp, covs, "-priors-", priors, ".png"
-  ),
-  res = 600,
-  units = "in",
-  width = 8.5,
-  height = 11
-)
-gridExtra::grid.arrange(
-  grobs = depth_plots,
-  ncol = 2,
-  top = grid::textGrob(paste(species))
-)
-dev.off()
-```
-
 # TRAWLING
@@
 ```
-
-Get all biomass predictions 
-#### Get all biomass predictions 
-```{r, eval=FALSE, echo=FALSE, warning=FALSE}
-priors <- FALSE
-covariates <- "+muddy+any_rock"
-covs <- gsub("\\+", "-", covariates)
-
-rm(adult_predictions)
-rm(imm_predictions)
-rm(predictions)
-pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
-
-try({adult_predictions <- readRDS(paste0("data/", spp,
-    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
-  ))%>% filter(year > 2007)
-adult_predictions <- left_join(adult_predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
-})
-
-try({imm_predictions <- readRDS(paste0("data/", spp,
-    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
-  )) %>% filter(year > 2007)
-imm_predictions <- left_join(imm_predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
-})
-
-try({predictions <- readRDS(paste0("data/", spp,
-    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
-  ))%>% filter(year > 2007)
-predictions <- left_join(predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
-})
-
-if (!exists("adult_predictions")) adult_predictions <- NULL
-if (!exists("imm_predictions")) imm_predictions <- NULL
-if (!exists("predictions")) predictions <- NULL
-
-biomass_list <- list(
-  adult = adult_predictions,
-  imm = imm_predictions,
-  total = predictions
-)
-
-biomass_list <- biomass_list[!sapply(biomass_list, is.null)]
-
-# 
-# ggplot(biomass_list[[1]], aes(x = do_est, y = log(est_exp))) + 
-#   geom_point(alpha = 0.3) #+ #geom_smooth()
-#   geom_smooth(method = "lm", formula = y ~ x + I(x^2) , size = 1) #+ I(x^3)
-```

### 2019-10-08 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
renderable versions of Rmds

@@
 ---
-title: "BACI of biomass change for VOCC source cells for all years"
+title: "BACI of `r paste(params$species)` biomass for `r paste(params$region)` VOCC"
 author: "Philina English"
@@
     species: "Pacific Ocean Perch" 
-    region: "both odd year surveys"
+    region: "West Coast Vancouver Island"
     age: "imm"
-    biomass_threshold: "25"
+    biomass_threshold: "50"
     climate: "temperature"
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(
+knitr::opts_chunk$set(fig.keep = "all", fig.show = "asis",
   fig.width = 11, fig.height = 8.5,
@@
 ```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
-library("dplyr")
-library("ggplot2")
-library("gfplot")
-library("sdmTMB")
-library("gfranges")
+library(dplyr)
+library(ggplot2)
+library(gfplot)
+library(sdmTMB)
+library(gfranges)
 ggplot2::theme_set(gfplot::theme_pbs())
@@
 #### Set parameters
-```{r echo=TRUE}
+```{r }
 species <- params$species
@@
 climate <- params$climate
+
+paste("species =", species)
+paste("age =", age)
+paste("region =", region)
+paste("biomass_threshold =", biomass_threshold)
+paste("climate =", climate)
 ```
@@
 
-# Run all subsequent code...
-
 ```{r, message=FALSE, warning=FALSE, include=FALSE}
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+
+  if(climate=="temperature") {
+    start_years <- list("2005" = 2005, "2007" = 2007, "2009" = 2009, "2011" = 2011, "2013" = 2013, "2015" = 2015)
+  } else {
   start_years <- list("2009" = 2009, "2011" = 2011, "2013" = 2013, "2015" = 2015)
-  # maybe remove 2011 & 2015 once we can add "2017"=2017 to avoid start and end values overlapping
+  }
   years <- NULL
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  start_years <- list("2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
-  # prob need to remove 2010 & 2014
+  if(climate=="temperature") {
+    start_years <- list("2004" = 2004, "2006" = 2006, "2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
+  } else {
+  start_years <- list("2008" = 2008, "2010" = 2010, "2012" = 2012, "2016" = 2016) # "2014" = 2014, # removed due to extreme DO values
+  }
   years <- NULL
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+  if(climate=="temperature") {
+    start_years <- list("2006" = 2006, "2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
+  } else {
   start_years <- list("2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
-  # prob need to remove 2010 & 2014
+  }
   years <- NULL
@@
 all_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
-
-all_params <- all_params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
+#browser()
+all_params <- all_params %>% filter(covs == "x2016-log-both-trawledFALSE") %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
 
@@
 # Calculate climate vectors
-```{r}
+```{r }
 vocc_by_year <- lapply(start_years, function(start_time) {
@@
   if (!exists("vocc")) {
+
+    if (climate=="temperature") {
+      climate_predictions <- readRDS("data/predicted_temp_allyears.rds") %>% mutate(temp = est)
+
+    } else {
       pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
@@
         climate_predictions <- pred_temp_do
+      }
     }
@@
 ### Plot all possible vectors
-```{r echo=FALSE}
+```{r }
 if (!exists("vocc")) {
@@
   dev.off()
-
-  print(all_vector_plots)
 }
+#print(all_vector_plots)
+invisible(lapply(all_vector_plots, print))
 ```
@@
 
-print(trimmed_vector_plots)
+#print(trimmed_vector_plots)
+invisible(lapply(trimmed_vector_plots, print))
 ```
@@
 
-dir.create(file.path("data/_all", climate_string, biomass_threshold_string))
+dir.create(file.path("data/log-both"))
+dir.create(file.path("data/log-both", climate_string))
+dir.create(file.path("data/log-both", climate_string, biomass_threshold_string))
 
 if (climate == "temperature") {
-  dir.create(file.path("data/_all", climate_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string, age))
+  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, max_string))
+  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, max_string, age))
 }
 if (climate == "temperature") {
-  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold, "/", max_string, "/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
+  saveRDS(df, file = paste0("data/log-both/", climate_string, "/", "perc_", biomass_threshold, "/", max_string, "/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
 }
@@
 if (climate == "do") {
-  dir.create(file.path("data/_all", climate_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string, age))
+  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, min_string))
+  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, min_string, age))
 }
 if (climate == "do") {
-  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold, "/", min_string, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
+  saveRDS(df, file = paste0("data/log-both/", climate_string, "/", "perc_", biomass_threshold, "/", min_string, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
 }
@@
 if (climate == "do and temperature") {
-  dir.create(file.path("data/_all", climate_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))
-  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, age))
+  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, age))
 }
@@
 if (climate == "do and temperature") {
-  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
+  saveRDS(df, file = paste0("data/log-both/", climate_string, "/", "perc_", biomass_threshold, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
 }

### 2019-10-09 analysis/VOCC/3-climate-models-fixed-both.Rmd
add option to model just temperature for more years than is available for DO

@@
     log_temp: TRUE # should temperature be log()
+    model_w_do: TRUE
     priors: FALSE
+    knots: 500
     covariates: "+trawled" # additional non-climate variables
@@
 log_temp <- params$log_temp
+model_w_do <- params$model_w_do
 priors <- params$priors
+knots <- params$knots
 
@@
 paste("log(temperature) =", log_temp)
+paste("model_w_do =", model_w_do)
 paste("priors =", priors)
+paste("knots =", knots)
 
@@
 
-  data <- dplyr::left_join(data, CTD) %>% 
-    mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
-    #mutate(exclude = if_else(do_mlpl>6, 1, 0)) %>% #try excluding high values?
-    filter(exclude != 1) %>% # added on july 24 after first round of models run...
-    filter(year > 2007) # 2007 DO data is flawed
+  data <- dplyr::left_join(data, CTD)
 
   # scale predictors before filtering to ensure mean and SD are global
-  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), 
-     raw_do = do_mlpl, do_mlpl = log(do_mlpl))
+  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth))
 
@@
   }
+
+if (model_w_do) {
+
+  data <- filter(data, !(year == 2016 & ssid == 4) ) # 2016 do an outlier for survey 4
+  data <- data %>% mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
+      filter(exclude != 1) %>% # added on july 24 after first round of models run...
+      filter(year > 2007) # 2007 DO data is flawed
+
+    # scale predictors before filtering to ensure mean and SD are global
+  data <- data %>% mutate(raw_do = do_mlpl, do_mlpl = log(do_mlpl))
 
@@
 
- data <- filter(data, !(year == 2016 & ssid == 4) ) # 2016 do an outlier for survey 4
+} else {
+
+  data <- scale_predictors(data, predictors = c(quo(depth), quo(temp))) 
+       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
+
+  data <- data %>% mutate(depth = raw_depth) %>% #filter(year != 2016) %>%
+    filter(ssid %in% model_ssid) %>% 
+    filter(!is.na(depth_scaled)) %>% filter(!is.na(temp))
+
+if (any(names(data) == "adult_density")) { 
+  data <- data %>% select(X, Y, X10, Y10,
+      adult_density, imm_density,
+      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
+      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
+      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
+      depth, temp, temperature_c) 
+ } else {
+    data <- data %>% select(X, Y, X10, Y10,
+      density,
+      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
+      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
+      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
+      depth, temp, temperature_c) 
+ }
+}
 ```
@@
   if (region == "All synoptic surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
   }
@@
 
-Run DO sdmTMB model
+Run sdmTMB model
 ```{r}
@@
 
+    if (model_w_do) {
     adult_formula <- as.formula(paste(
@@
       ))
+    } else {
+      adult_formula <- as.formula(paste(
+      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", covariates, ""
+      ))
+    }
 
@@
 
+    if (model_w_do) {
     imm_formula <- as.formula(paste(
@@
       ))
+    } else {
+         imm_formula <- as.formula(paste(
+      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", covariates, ""
+      ))
+    }
 

### 2019-10-09 analysis/VOCC/3-fixed-var-response-plots.Rmd
add option to model just temperature for more years than is available for DO

@@
 params: 
-    species: "Pacific Ocean Perch"
+    species: "Dover Sole"
     priors: FALSE # call model that used priors?
-    log_temp: TRUE
+    log_temp: TRUE # should temperature be log()
+    model_w_do: TRUE
     covs: "-log-both-trawled" # covariate string used to name model rds
@@
 num_params <- params$cov_number
+model_w_do <- params$model_w_do
 
@@
 paste("model =", covs)
+paste("model_w_do =", model_w_do)
 paste("log(temperature) =", log_temp)
@@
 do_plots <- list()
-
+if (model_w_do) {
 for (i in seq_len(length(model_list))) {
@@
 }
+} else {  odat <- NULL }
 print(do_plots)
@@
 ```{r}
+if (model_w_do) {
 if(!is.null(data$adult_density)){
@@
 raw_do_plots <- list(do_count,do_prop)
+
 print(raw_do_plots)
+}
+} else {
+  raw_do_plots <- NULL
 }
@@
 ```{r eval=F}
+if(model_w_do) {
 png(
@@
 dev.off()
+}
 ```

### 2019-10-21 analysis/VOCC/1-split-biomass-by-maturity.Rmd
tweaks to depth-only biomass models

@@
 params: 
-    species: "Curlfin Sole"
+    species: "Pacific Cod"
     region: "All synoptic surveys"

### 2019-10-21 analysis/VOCC/2-biomass-depth-only-model.Rmd
tweaks to depth-only biomass models

@@
 params: 
-    species: "Copper Rockfish"
+    species: "Petrale Sole"
     region: "All synoptic surveys"
+    knots: 600
     covariates: "+trawled+as.factor(ssid)" # additional fixed-effect covariates
     covs: "-trawled-ssid" # covariate string used to name model rds 
+    update_model: TRUE
 ---
@@
 covs <- params$covs
+knots <- params$knots
 priors <- FALSE
@@
 paste("priors =", priors)
+paste("knots =", knots)
 ```
@@
   mutate(depth = raw_depth) %>%
-  filter(ssid %in% model_ssid)
+  filter(ssid %in% model_ssid) %>% 
+  filter(year < 2019)
 ```
@@
 if (region == "All synoptic surveys") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
+  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
 }
@@
 Run climate independent sdmTMB model
-```{r}
+```{r }
+if (params$update_model) {
 tictoc::tic()
@@
     family = tweedie(link = "log"),
-    ar1_fields = FALSE,
+    ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
     include_spatial = TRUE,
@@
 tictoc::toc()
+}
 ```
@@
 Make prediction grids for all surveys and years, if not done before.
-```{r, message=FALSE, warning=FALSE}
+```{r, eval=FALSE, message=FALSE, warning=FALSE}
 try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
@@
   nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
+
   unique(nd_1$year)
@@
   nd_all <- left_join(nd_all, newcov, by = c("X", "Y"))
-
+  nd_all$trawled[is.na(nd_all$trawled)] <- 0 # correct missing values for small area on edge of QCS grid
   # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
@@
   ))
+
+    nd <- nd_all %>%
+    filter(ssid %in% model_ssid) %>%
+    filter(year %in% unique(imm_biomass$data$year))
+  nd <- na.omit(nd)
+  nd$year <- as.integer(nd$year)
+
   imm_predictions <- predict(imm_biomass, newdata = nd)
@@
 } else {
+
+    total_biomass <- readRDS(paste0("data/", spp,
+      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
     nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-    nd <- nd_all %>% filter(ssid %in% model_ssid)
+    nd <- nd_all %>% 
+      filter(ssid %in% model_ssid) %>%
+      filter(year %in% unique(total_biomass$data$year))
     nd <- na.omit(nd)
@@
 
-    total_biomass <- readRDS(paste0("data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
     predictions <- predict(total_biomass, newdata = nd)

### 2019-10-21 analysis/VOCC/2-biomass-trends.Rmd
change years plotted for biomass trends

@@
 params: 
-    species: "Lingcod"
+    species: "Petrale Sole"
     region: "All synoptic surveys"
+    start_year: 2008
     bind_predictions: FALSE
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
+knitr::opts_chunk$set( fig.width = 9, fig.height = 6.5, #fig.width=11, fig.height=8.5, 
                       #fig.path=paste0("figs/", spp, "/"),
@@
 region <- params$region
+start_year <- params$start_year
+
 bind_predictions <- params$bind_predictions
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  years <- NULL
 }
 
-if (region == "Both odd year surveys") {
+if (region == "both odd year surveys") {
   survey <- c("SYN QCS", "SYN HS")
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  years <- NULL
 }
@@
   try(adult_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
-  )))
+    "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-", priors, ".rds"
+  )) %>% filter(ssid %in% model_ssid))
   try(imm_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
-  )))
+    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-", priors, ".rds"
+  )) %>% filter(ssid %in% model_ssid))
 
@@
     try(predictions <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
-    )))
+      "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-", priors, ".rds"
+    )) %>% filter(ssid %in% model_ssid))
 
@@
 if (exists("adult_predictions")) {
-
   # to convert from kg/m2 to kg/hectare multiply by 10000
@@
   imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
-
-  max_raster <- quantile(adult_predictions$est_exp, 0.999)
-  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)
 
@@
 
+  adult_predictions <- filter(adult_predictions, year >= start_year)
+  imm_predictions <- filter(imm_predictions, year >= start_year)
+
+  max_raster <- quantile(adult_predictions$est_exp, 0.999)
+  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)
+
   model_ssid <- unique(adult_predictions$ssid)
 } else {
+  predictions <- filter(predictions, year >= start_year)
+
+  predictions$est_exp <- exp(predictions$est) * 10000
   predictions$total_bio <- exp(predictions$est) * 10000
   max_raster <- quantile(predictions$total_bio, .999)
-  max_bio <- round(max(predictions$total_bio))
+  #max_bio <- round(max(predictions$total_bio))
+  max_bio <- signif(max(predictions$est_exp), digits = 2)
 
@@
 ```{r}
-if (exists("adult_predictions")) {
+if (exists("adult_predictions")) predictions <- adult_predictions
+
   if (1 %in% model_ssid) {
-    adult_predictions1 <- adult_predictions %>% filter(ssid == 1)
+    adult_predictions1 <- predictions %>% filter(ssid == 1)
   } else {
@@
   if (3 %in% model_ssid) {
-    adult_predictions3 <- adult_predictions %>%
+    adult_predictions3 <- predictions %>%
       filter(ssid != 16) %>%
@@
   if (4 %in% model_ssid) {
-    adult_predictions4 <- adult_predictions %>% filter(ssid == 4)
+    adult_predictions4 <- predictions %>% filter(ssid == 4)
   } else {
@@
   if (16 %in% model_ssid) {
-    if (any(adult_predictions$ssid == 16)) {
-      adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
+    if (any(predictions$ssid == 16)) {
+      adult_predictions16 <- predictions %>% filter(ssid == 16)
     } else {
@@
   p_adult <- list()
+  if(exists("adult_predictions")) {
   for (i in seq_len(length(model_ssid))) {
@@
   }
+  } else {
+    for (i in seq_len(length(model_ssid))) {
+    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
+      raster_limits = c(0, max_raster),
+      # legend_position = "none"
+      transform_col = fourth_root_power
+    ) +
+      labs(fill = "kg/ha") +
+      ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
+    p_adult[[i]]  
+    }
+  }
   print(p_adult)
-}
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
+rm(out_mature)
+rm(out_total)
 # try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
-rm(out_mature)
+
 if (!exists("out_mature")) {
@@
   } else {
-    try(out_total <- readRDS(paste0(
-      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
-    )))
+   try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
   }
@@
     if (1 %in% model_ssid) {
+
+      .predictions <- filter(predictions, ssid == 1)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
       out_mature1 <- make_trend_data(predictions,
         ssid = c(1),
-        start_time = 2003,
+        start_time = start_year,
         skip_time = 2004,
@@
         delta_t_step = 2,
-        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
+        indices = indices,
         variable_names = "est_exp"
@@
     if (3 %in% model_ssid) {
+      .predictions <- filter(predictions, ssid == 3)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
       out_mature3 <- make_trend_data(predictions,
         ssid = c(3),
-        start_time = 2005,
+        start_time = start_year,
         input_cell_size = 2,
@@
         delta_t_step = 2,
-        indices = c(1, 1, 1, 2, 2, 2, 2),
+        indices = indices,
         variable_names = "est_exp"
@@
     if (4 %in% model_ssid) {
+
+      .predictions <- filter(predictions, ssid == 4)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
       out_mature4 <- make_trend_data(predictions,
         ssid = c(4),
-        start_time = 2004,
+        start_time = start_year,
         input_cell_size = 2,
@@
     if (16 %in% model_ssid) {
+
+      .predictions <- filter(predictions, ssid == 16)
+      #unique(.predictions$year)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
       out_mature16 <- make_trend_data(predictions,
         ssid = c(16),
-        start_time = 2004,
+        start_time = start_year,
         skip_time = 2007,
@@
 
-    saveRDS(out_mature, file = paste0(
-      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"
-    ))
+    if (exists("adult_predictions")) {
+            saveRDS(out_mature, file = paste0("data/", spp, "/trends-", 
+        spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"))
 
-    if (!exists("adult_predictions")) {
+         } else {
       out_total <- out_mature
-      saveRDS(out_total, file = paste0(
-        "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
-      ))
+      rm(out_mature)
+      saveRDS(out_total, file = paste0("data/", spp, "/trends-", 
+        spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"))
     }
@@
 ```{r }
-if (exists("out_mature")) {
+if (!exists("out_mature")) {out_recent <- out_total} else {out_recent <- out_mature}
+
   trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
@@
 
-  current_biomass <- plot_vocc(out_mature,
+  current_biomass <- plot_vocc(out_recent,
     fill_col = "est_exp_e",
-    fill_label = "Mean mature \nbiomass (kg/ha)\npost-2010",
+    fill_label = "Mean\nbiomass (kg/ha)\npost-2010",
     raster_alpha = 1,
@@
   current_biomass
-}
+
 ```
@@
     if (1 %in% model_ssid) {
+
+      .predictions <- filter(predictions, ssid == 1)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
       out_im1 <- make_trend_data(predictions,
         ssid = c(1),
-        start_time = 2003,
+        start_time = start_year,
         skip_time = 2004,
@@
     if (3 %in% model_ssid) {
+      .predictions <- filter(predictions, ssid == 3)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
       out_im3 <- make_trend_data(predictions,
         ssid = c(3),
-        start_time = 2005,
+        start_time = start_year,
         input_cell_size = 2,
@@
     if (4 %in% model_ssid) {
+      .predictions <- filter(predictions, ssid == 4)
+      unique(.predictions$year)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
       out_im4 <- make_trend_data(predictions,
         ssid = c(4),
-        start_time = 2004,
+        start_time = start_year,
         input_cell_size = 2,
@@
     if (16 %in% model_ssid) {
+      .predictions <- filter(predictions, ssid == 16)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
       out_im16 <- make_trend_data(predictions,
         ssid = c(16),
-        start_time = 2004,
+        start_time = start_year,
         input_cell_size = 2,
@@
 ```{r results = "hide"}
-biomass_plots <- c(p_adult, p_imm)
+if(exists("adult_predictions")) { biomass_plots <- c(p_adult, p_imm) } else {
+  biomass_plots <- c(p_adult)
+}
 

### 2019-10-21 analysis/VOCC/3-fixed-var-response-plots.Rmd
change custom breaks and fig sizes

@@
 params: 
-    species: "Dover Sole"
+    species: "Redbanded Rockfish"
     priors: FALSE # call model that used priors?
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
-                      #fig.path=paste0("figs/", spp, "/"),
+knitr::opts_chunk$set(dpi = 300,
+  fig.width = 5, fig.height = 2.5,
   echo=FALSE, warning=FALSE, message=FALSE)
@@
    do_plots[[i]] <- 
-     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0.2, 8)) +
+     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0.2, 8), x_breaks = c(1,2,3,4,5,6,7,8)) +
         ggtitle(paste(species, names(model_list[i]))) +
@@
    temp_plots[[i]] <- 
-     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 12)) +
+     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 12), x_breaks = c(3,4,5,6,7,8,9,10,11)) +
         ggtitle(paste(species, names(model_list[i]))) +
@@
 saveRDS(all_params, file = "data/root-values-by-species-by-age.rds")
-all_params
+#spp_params <- all_params[all_params$species==species,]
+glimpse(new_params)
 ```

### 2019-10-21 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
add option for threshold selection based on width

@@
 params: 
-    species: "Pacific Ocean Perch" 
-    region: "West Coast Vancouver Island"
-    age: "imm"
+    species: "Redbanded Rockfish" 
+    region: "both odd year surveys"
+    age: "mature"
     biomass_threshold: "50"
     climate: "temperature"
-    threshold: 0.25
+    threshold: 0.5
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.keep = "all", fig.show = "asis",
-  fig.width = 11, fig.height = 8.5,
+knitr::opts_chunk$set(fig.keep = "all", fig.show = "asis", dpi = 600,
+  fig.width = 6, fig.height = 5,
   echo = FALSE, warning = FALSE, message = FALSE
@@
 climate <- params$climate
+threshold <- params$threshold
 
@@
 paste("climate =", climate)
+paste("climate change threshold =", threshold)
 ```
@@
 ```{r, echo=FALSE}
+max_vector <- 80
 input_cell_size <- 2
@@
 #browser()
-all_params <- all_params %>% filter(covs == "x2016-log-both-trawledFALSE") %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
+all_params <- all_params %>% filter(covs == "x2016-log-both-trawledFALSE") #%>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))
 
@@
   all_params <- filter(all_params, age != "imm")
-  if (any(all_params$model_type == "split")) {
-    all_params <- filter(all_params, model_type != "total")
-  }
+  # if (any(all_params$model_type == "split")) {
+  #   all_params <- filter(all_params, model_type != "total")
+  # }
 }
@@
 if (climate == "temperature") {
-  temp_threshold <- params$threshold
-
+
+  if (threshold) { temp_threshold <- params$threshold } else {
+  temp_threshold <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 - root_params[rev(order(root_params$t_maxy)), ]$upper_t_75
+  temp_threshold <- round(temp_threshold, 2)
+  }
   # if more than one model option in dataframe will pick the one with highest peak density
   if (biomass_threshold == "75") {
-    .upper <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75
+    .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_75
   }
@@
 
+```{r}
+if (climate=="temperature") {
+      climate_predictions <- readRDS("data/predicted_temp_allyears_revised") %>% mutate(temp = est)
+
+    } else {
+      pred_temp_do <- readRDS("data/predicted-DO-new.rds")
+
+      if (length(variable_names) > 1) {
+        climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
+      } else {
+        climate_predictions <- pred_temp_do
+      }
+    }
+```
+
 # Calculate climate vectors
@@
       "data/vocc-", ssid_string, "-", climate_string, "-",
-      min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
+      min_string, "to", max_string, "-", start_time, "-", delta_t_total, "-update.rds"
     ))
@@
 if (!exists("vocc")) {
-
-    if (climate=="temperature") {
-      climate_predictions <- readRDS("data/predicted_temp_allyears.rds") %>% mutate(temp = est)
-
-    } else {
-      pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
-
-      if (length(variable_names) > 1) {
-        climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
-      } else {
-        climate_predictions <- pred_temp_do
-      }
-    }
     end_time <- start_time + 2
@@
 if (!exists("vocc")) {
-  all_vector_plots <- lapply(vocc_by_year, function(vocc) {
+
+  all_vector_plots_w <- lapply(vocc_by_year, function(vocc) {
     start_time <- vocc$start_time[1]
@@
       vocc$diff <- vocc$units_per_decade / 10 * delta_t_total
-
       vocc_plot <- plot_vocc(vocc,
         vec_aes = "distance",
-        min_vec_plotted = input_cell_size,
-        NA_label = ".",
-        fill_col = "diff",
-        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
-        raster_alpha = 1,
-        raster_limits = c(-4, 3),
-        white_zero = TRUE,
-        vec_alpha = 0.25,
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "temp_s",
+        fill_label = paste(" Temperature\n change\n", 
+          start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        # raster_limits = c(raster_min, raster_max),
+        na_colour = "gray",
+        # white_zero = TRUE,
+        vec_alpha = 0.35,
         axis_lables = FALSE,
@@
       vocc$diff <- vocc$units_per_decade / 10 * delta_t_total
-
       vocc_plot <- plot_vocc(vocc,
         vec_aes = "distance",
-        min_vec_plotted = input_cell_size,
-        NA_label = ".",
-        fill_col = "diff",
-        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
-        raster_alpha = 1,
-        white_zero = TRUE,
-        high_fill = "Steel Blue 4",
-        low_fill = "Red 3",
-        vec_alpha = 0.25,
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "do_est_s",
+        fill_label = paste(" DO change\n", start_time, 
+          "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        # raster_limits = c(raster_min, raster_max),
+        na_colour = "gray",
+        # white_zero = TRUE,
+        # high_fill = "Steel Blue 4",
+        # low_fill = "Red 3",
+        vec_alpha = 0.35,
         axis_lables = FALSE,
@@
         vec_aes = "distance",
-        min_vec_plotted = input_cell_size,
-        NA_label = ".",
-        fill_col = "temp_diff",
-        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
-        raster_alpha = 1,
-        white_zero = TRUE,
-        vec_alpha = 0.25,
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "temp_s",
+        fill_label = paste(" Temperature\n change\n", 
+          start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        raster_limits = c(raster_min_temp, raster_max_temp),
+        # white_zero = TRUE,
+        vec_alpha = 0.35,
         axis_lables = FALSE,
@@
         vec_aes = "distance",
-        min_vec_plotted = input_cell_size,
-        NA_label = ".",
-        fill_col = "do_diff",
-        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
-        raster_alpha = 1,
-        white_zero = TRUE,
-        high_fill = "Steel Blue 4",
-        low_fill = "Red 3",
-        vec_alpha = 0.25,
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "do_est_s",
+        fill_label = paste(" DO change\n", start_time, "to", 
+          start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        # raster_limits = c(raster_min_do, raster_max_do),
+        # white_zero = TRUE,
+        # high_fill = "Steel Blue 4",
+        # low_fill = "Red 3",
+        vec_alpha = 0.35,
         axis_lables = FALSE,
@@
     file = paste0(
-      "figs/vocc-", climate_string, min_string, "to", max_string, "-", ssid_string, ".png"
+      "figs/vocc-", climate_string, min_string, "to", max_string, "-", ssid_string, "-absolute.png"
     ),
@@
   gridExtra::grid.arrange(
-    grobs = all_vector_plots,
+    grobs = all_vector_plots_w,
     ncol = 2,
@@
 #print(all_vector_plots)
-invisible(lapply(all_vector_plots, print))
+invisible(lapply(all_vector_plots_w, print))
 ```
@@
 all_vectors$var_2_max <- upper_thresholds[2]
-```
-
-### Plot remaining climate vectors for regions where change exceeds thresholds
-```{r echo=FALSE, warning=FALSE}
+
 if (climate == "temperature") {
   all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
-  raster_min <- min(all_vectors$diff)
-  raster_max <- max(all_vectors$diff)
+  # raster_min <- min(all_vectors$diff)
+  # raster_max <- max(all_vectors$diff)
+  raster_min <- quantile(all_vectors$diff, 0.005)
+  raster_max <- quantile(all_vectors$diff, 0.995)
 }
@@
   all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
-  raster_min <- min(all_vectors$diff)
-  raster_max <- max(all_vectors$diff)
+  # raster_min <- min(all_vectors$diff)
+  # raster_max <- max(all_vectors$diff)
+  # 
+  raster_min <- quantile(all_vectors$diff, 0.0005)
+  raster_max <- quantile(all_vectors$diff, 0.9995)
+
 }
@@
 }
+```
+
+### Plot all possible vectors
+```{r }
+if (!exists("vocc")) {
+
+  all_vector_plots <- lapply(vocc_by_year, function(vocc) {
+    start_time <- vocc$start_time[1]
+
+    if (climate == "temperature") {
+      vocc$diff <- vocc$units_per_decade / 10 * delta_t_total
+      vocc_plot <- plot_vocc(vocc,
+        vec_aes = "distance",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "diff",
+        fill_label = paste(" Temperature\n change\n", start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        #raster_limits = c(raster_min, raster_max),
+        na_colour = "#9E0142",
+        # mid_fill = NULL, #"#FFFEC9", #cream
+        # high_fill = "#D53E4F",#"Steel Blue 4"
+        # low_fill = "#5E4FA2",
+        white_zero = TRUE,
+        vec_col = "white",
+        vec_alpha = 0.45,
+        axis_lables = FALSE,
+        transform_col =  sqrt #fourth_root_power
+      ) + ggtitle(climate) #+ ggpubr::theme_transparent()# +  theme_black()
+    }
+
+    if (climate == "do") {
+      vocc$diff <- vocc$units_per_decade / 10 * delta_t_total
+      vocc_plot <- plot_vocc(vocc,
+        vec_aes = "distance",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "diff",
+        fill_label = paste(" DO change\n", start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        #raster_limits = c(raster_min, raster_max),
+        na_colour = "gray",
+        white_zero = TRUE,
+        high_fill = "Steel Blue 4",
+        low_fill = "Red 3",
+        vec_alpha = 0.65,
+        axis_lables = FALSE,
+        transform_col = sqrt# no_trans
+      ) + ggtitle(climate)
+    }
+
+    if (climate == "do and temperature") {
+      vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total
+
+      vocc_plot_temp <- plot_vocc(vocc,
+        vec_aes = "distance",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "temp_diff",
+        fill_label = paste(" Temperature\n change\n", start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        raster_limits = c(raster_min_temp, raster_max_temp),
+        white_zero = TRUE,
+      vec_alpha = 0.65,
+        axis_lables = FALSE,
+        transform_col = sqrt #no_trans
+      ) + ggtitle("temperature")
+
+      vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total
+
+      vocc_plot_do <- plot_vocc(vocc,
+        vec_aes = "distance",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "do_diff",
+        fill_label = paste(" DO change\n", start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        raster_limits = c(raster_min_do, raster_max_do),
+        white_zero = TRUE,
+        high_fill = "Steel Blue 3",
+        low_fill = "Red 2",
+      vec_alpha = 0.65,
+        axis_lables = FALSE,
+        transform_col = sqrt #no_trans
+      ) + ggtitle("do")
+
+      vocc_plot <- gridExtra::grid.arrange(
+        vocc_plot_temp, vocc_plot_do,
+        ncol = 2 # ,
+        # top = grid::textGrob(paste("do decline of", lower_change[1], "or temperature increase of", upper_change[2], " over ", start_time, "-", start_time + delta_t_total, ""))
+      )
+    }
+    vocc_plot
+  })
+
+  png(
+    file = paste0(
+      "figs/vocc-", climate_string, min_string, "to", max_string, "-", ssid_string, ".png"
+    ),
+    bg = "transparent",
+    res = 600,
+    units = "in",
+    width = 11,
+    height = 17
+  )
+  gridExtra::grid.arrange(
+    grobs = all_vector_plots,
+    ncol = 2,
+    top = grid::textGrob(paste("Vectors for a DO decline of", lower_change[1], "and/or temperature increase of", upper_change[2], ""))
+  )
+  dev.off()
+}
+#print(all_vector_plots)
+invisible(lapply(all_vector_plots, print))
+```
+
+### Plot remaining climate vectors for regions where change exceeds thresholds
+```{r echo=FALSE, warning=FALSE}
 
@@
         vec_aes = "distance",
-        min_vec_plotted = input_cell_size,
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
         NA_label = ".",
         fill_col = "diff",
-        fill_label = paste(
-          "Change from\n",
-          start_time, "to", start_time + delta_t_total, ""
-        ),
+        fill_label = paste("Temperature\n change\n", start_time, "to", start_time + delta_t_total, ""),
         raster_alpha = 1,
         raster_limits = c(raster_min, raster_max),
+        na_colour = "gray",
         white_zero = TRUE,
-        vec_alpha = 0.25,
+      vec_alpha = 0.65,
         axis_lables = FALSE,
         transform_col = no_trans
-      ) + ggtitle(climate)
+      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
+        biomass_threshold, "% of peak predicted ", species, " biomass" ))
     }
@@
         vec_aes = "distance",
-        min_vec_plotted = input_cell_size,
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
         NA_label = ".",
         fill_col = "diff",
-        fill_label = paste(
-          "Change from\n",
-          start_time, "to", start_time + delta_t_total, ""
-        ),
+        fill_label = paste("DO change\n", start_time, "to", start_time + delta_t_total, ""),
         raster_alpha = 1,
         raster_limits = c(raster_min, raster_max),
+        na_colour = "gray",
         white_zero = TRUE,
@@
         low_fill = "Red 3",
-        vec_alpha = 0.25,
+        vec_alpha = 0.65,
         axis_lables = FALSE,
         transform_col = no_trans
-      ) + ggtitle(climate)
+      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
+        biomass_threshold, "% of peak predicted ", species, " biomass" ))
     }
@@
         raster_limits = c(raster_min_temp, raster_max_temp),
+        na_colour = "gray",
         white_zero = TRUE,
-        vec_alpha = 0.25,
+        vec_alpha = 0.65,
         axis_lables = FALSE,
@@
         vec_aes = "distance",
-        min_vec_plotted = input_cell_size,
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
         NA_label = ".",
@@
         raster_limits = c(raster_min_do, raster_max_do),
+        na_colour = "gray",
         white_zero = TRUE,
@@
         low_fill = "Red 3",
-        vec_alpha = 0.25,
+        vec_alpha = 0.65,
         axis_lables = FALSE,
@@
 
-#print(trimmed_vector_plots)
-invisible(lapply(trimmed_vector_plots, print))
+print(trimmed_vector_plots)
+#invisible(lapply(trimmed_vector_plots, print))
 ```
@@
   end_time <- start_time + 2
-
+  indices <- c(1,2)
   biomass_by_year[[j]] <- make_trend_data(biomass_predictions,
@@
 
+### Plot climate vectors for cells with starting with greater than 5% percentile of biomass densities
+```{r echo=FALSE, warning=FALSE}
+# # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass across years
+bio5perc <- sum(biomass_change$start_density)*0.05
+s <- sort(biomass_change$start_density)
+bio_sum <- cumsum(s)
+lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+alldata <- alldata %>% filter(start_density > lower_density_threshold)
+
+# set limits same for all years based on percent change for cells with vectors
+# raster_min <- min((log(alldata$end_density+1)-log(alldata$start_density+1)))
+# raster_max <- max((log(alldata$end_density+1)-log(alldata$start_density+1)))
+raster_min <- quantile(((alldata$end_density)-(alldata$start_density)), 0.01)
+raster_max <- max((alldata$end_density)-(alldata$start_density))#, 0.9999)
+
+# if min and max should be global rather than for cells with vectors
+# raster_min <- min((biomass_change$end_density-biomass_change$start_density)/biomass_change$start_density)
+# raster_max <- max((biomass_change$end_density-biomass_change$start_density)/biomass_change$start_density)
+
+#lower_density_threshold <- quantile(alldata$start_density, 0.25, na.rm = TRUE)
+
+# # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass across years
+# bio5perc <- sum(biomass_change$start_density)*0.05
+# s <- sort(biomass_change$start_density)
+# bio_sum <- cumsum(s)
+# lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+biomass_vector_plots <- lapply(trimmed_vectors, function(vectors) {
+  try({
+    # #offset vectors so that they represent the time step before the one for biomass change
+    # vectors <- vectors %>% mutate(start_time = start_time + 2, end_time = end_time + 2)
+
+    start_time <- vectors$start_time[1]
+
+    # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass within years
+    annual_biomass <- filter(biomass_change, start_time == !!start_time) 
+    bio5perc <- sum(annual_biomass$start_density)*0.05
+    s <- sort(annual_biomass$start_density)
+    bio_sum <- cumsum(s)
+    lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+    #vectors <- vectors %>% filter(start_time == !!start_time)
+
+    biomass <- left_join(biomass_change, vectors, 
+      by = c("icell", "x", "y", "start_time")) %>% 
+      filter(start_time == !!start_time) %>% 
+     # mutate(density_diff = (log(end_density+1)-log(start_density+1)) ) %>%
+      mutate(density_diff = end_density-start_density,
+        distance = if_else(is.na(distance), 0, distance),
+        distance = if_else(start_density < !!lower_density_threshold, 0, distance),
+        target_X = as.numeric(target_X), target_Y = as.numeric(target_Y),
+        target_X = if_else(start_density < !!lower_density_threshold, NA_real_, target_X),
+        target_Y = if_else(start_density < !!lower_density_threshold, NA_real_, target_Y),
+        n_targets = if_else(distance==0, 0, n_targets)
+        )
+    # start_time <- biomass$start_time[1]
+
+    vocc_plot <- plot_vocc(biomass,
+        vec_aes = "distance",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "density_diff",
+        fill_label = paste("Biomass\n change (kg/ha)\n", 
+            start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        raster_limits = c(raster_min, raster_max),
+        # varidis_option = 
+        white_zero = FALSE,
+        # mid_fill = NULL,
+        # high_fill = "#5E4FA2", #"Steel Blue 4"
+        # low_fill = "#D53E4F",
+        # high_fill = "#21908CFF", # "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
+        # low_fill = "#FDE725FF", #"#FEE08B", # "#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B"   
+        na_colour = "red",
+        vec_col = "white",
+        vec_alpha = 0.35,
+        axis_lables = FALSE,
+        transform_col =  sqrt # no_trans ##fourth_root_power #
+      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
+        biomass_threshold, "% of peak predicted ", species, " biomass" ))
+
+    vocc_plot
+  })
+})
+
+#print(biomass_vector_plots)
+invisible(lapply(biomass_vector_plots, print))
+```
+
+### Plot climate vectors for cells with starting with greater than 25% percentile of biomass densities
+```{r echo=FALSE, warning=FALSE}
+# # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass across years
+bio5perc <- sum(biomass_change$start_density)*0.05
+s <- sort(biomass_change$start_density)
+bio_sum <- cumsum(s)
+lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+alldata <- alldata %>% filter(start_density > lower_density_threshold)
+
+biomass_vector_plots_raw <- lapply(vocc_by_year, function(vectors) {
+  try({
+    # #offset vectors so that they represent the time step before the one for biomass change
+    # vectors <- vectors %>% mutate(start_time = start_time + 2, end_time = end_time + 2)
+
+    start_time <- vectors$start_time[1]
+
+    # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass within years
+    annual_biomass <- filter(biomass_change, start_time == !!start_time) 
+    bio5perc <- sum(annual_biomass$start_density)*0.05
+    s <- sort(annual_biomass$start_density)
+    bio_sum <- cumsum(s)
+    lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+    #vectors <- vectors %>% filter(start_time == !!start_time)
+
+    biomass <- left_join(biomass_change, vectors, 
+      by = c("icell", "x", "y", "start_time")) %>% 
+      filter(start_time == !!start_time)
+
+       vocc_plot <- plot_vocc(biomass,
+        vec_aes = "distance",
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "start_density",
+        fill_label = paste("Biomass\n change (kg/ha)\n", 
+            start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        viridis_option = "C",
+        #raster_limits = c(raster_min, raster_max),
+        #white_zero = TRUE,
+        #high_fill = "#21908CFF", # "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
+        #low_fill = "#FDE725FF", #"#FEE08B", # "#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B"
+        na_colour = "red",
+        vec_alpha = 0.35,
+        axis_lables = FALSE,
+        transform_col = fourth_root_power # no_trans #
+      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
+        biomass_threshold, "% of peak predicted ", species, " biomass" ))
+
+    vocc_plot
+  })
+})
+
+#print(biomass_vector_plots)
+invisible(lapply(biomass_vector_plots_raw, print))
+```
+
+```{r}
+
+biomass_vector_plots_trim <- lapply(trimmed_vectors, function(vectors) {
+  try({
+    # #offset vectors so that they represent the time step before the one for biomass change
+    # vectors <- vectors %>% mutate(start_time = start_time + 2, end_time = end_time + 2)
+
+    start_time <- vectors$start_time[1]
+
+    # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass within years
+    annual_biomass <- filter(biomass_change, start_time == !!start_time) 
+    bio5perc <- sum(annual_biomass$start_density)*0.05
+    s <- sort(annual_biomass$start_density)
+    bio_sum <- cumsum(s)
+    lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+    #vectors <- vectors %>% filter(start_time == !!start_time)
+
+    biomass <- left_join(biomass_change, vectors, 
+      by = c("icell", "x", "y", "start_time")) %>% 
+      filter(start_time == !!start_time)
+
+       vocc_plot <- plot_vocc(biomass,
+        vec_aes = "distance",
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "start_density",
+        fill_label = paste("Biomass\n change (kg/ha)\n", 
+            start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        viridis_option = "C",
+        #raster_limits = c(raster_min, raster_max),
+        #white_zero = TRUE,
+        #high_fill = "#21908CFF", # "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
+        #low_fill = "#FDE725FF", #"#FEE08B", # "#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B"
+        na_colour = "red",
+        vec_alpha = 0.35,
+        axis_lables = FALSE,
+        transform_col = fourth_root_power # no_trans #
+      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
+        biomass_threshold, "% of peak predicted ", species, " biomass" ))
+
+    vocc_plot
+  })
+})
+
+#print(biomass_vector_plots)
+invisible(lapply(biomass_vector_plots_trim, print))
+```
+
+```{r}
+
+biomass_vector_plots_present <- lapply(trimmed_vectors, function(vectors) {
+  try({
+    # #offset vectors so that they represent the time step before the one for biomass change
+    # vectors <- vectors %>% mutate(start_time = start_time + 2, end_time = end_time + 2)
+
+    start_time <- vectors$start_time[1]
+
+    # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass within years
+    annual_biomass <- filter(biomass_change, start_time == !!start_time) 
+    bio5perc <- sum(annual_biomass$start_density)*0.05
+    s <- sort(annual_biomass$start_density)
+    bio_sum <- cumsum(s)
+    lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+    #vectors <- vectors %>% filter(start_time == !!start_time)
+
+    biomass <- left_join(biomass_change, vectors, 
+      by = c("icell", "x", "y", "start_time")) %>% 
+      filter(start_time == !!start_time) %>% 
+      mutate(density_diff = end_density-start_density,
+       distance = if_else(is.na(distance), 0, distance),
+       distance = if_else(start_density < !!lower_density_threshold, 0, distance),
+       target_X = as.numeric(target_X), target_Y = as.numeric(target_Y),
+       target_X = if_else(start_density < !!lower_density_threshold, NA_real_, target_X),
+       target_Y = if_else(start_density < !!lower_density_threshold, NA_real_, target_Y),
+       n_targets = if_else(distance==0, 0, n_targets)
+       )
+
+       vocc_plot <- plot_vocc(biomass,
+        vec_aes = "distance",
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "start_density",
+        fill_label = paste("Biomass\n change (kg/ha)\n", 
+            start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        viridis_option = "C",
+        #raster_limits = c(raster_min, raster_max),
+        #white_zero = TRUE,
+        #high_fill = "#21908CFF", # "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
+        #low_fill = "#FDE725FF", #"#FEE08B", # "#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B"
+        na_colour = "red",
+        vec_alpha = 0.35,
+        axis_lables = FALSE,
+        transform_col = fourth_root_power # no_trans #
+      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
+        biomass_threshold, "% of peak predicted ", species, " biomass" ))
+
+    vocc_plot
+  })
+})
+
+#print(biomass_vector_plots)
+invisible(lapply(biomass_vector_plots_present, print))
+```
+
 # Select matched control cells
@@
 rm(vocc_by_sp)
-# try({
-#   vocc_by_sp <- readRDS(paste0(
-#     "data/BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
-#   ))
-# })
+try({
+  vocc_by_sp <- readRDS(paste0("data/", spp,
+  "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
+))
+})
 
@@
 
-# if (!exists("vocc_by_sp")) {
+if (!exists("vocc_by_sp")) {
 # matched <- lapply(alldata, function(alldata) {
@@
 # })
-# }
+ }
 ```
@@
 ```{r}
-# if (!exists("vocc_by_sp")) {
+ if (!exists("vocc_by_sp")) {
 # vocc_by_sp <- lapply(matched, function(matched) {
@@
 ))
-# }
+}
 ```

### 2019-10-21 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
refine plot-map functions: remove utm axis lables from faceted maps and make bathymetry lines finer
update climate models/predictions

@@
 d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
-d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
-d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
+#d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
+#d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
 d_trawl <- mutate(d_trawl,
@@
 
-d_trawl <- d_trawl %>% dplyr::filter(year>2003)
+#d_trawl <- d_trawl %>% dplyr::filter(year>2003)
 ```
@@
 ```{r}
-d_trawl1 <- d_trawl %>% filter(ssid == 3)
-spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 150)
-plot_spde(spde)
-sum(d_trawl1$shallow)
-```
-
-```{r}
-d_trawl1$year_scaled <- arm::rescale(d_trawl1$year)
-d_trawl1$x_scaled <- arm::rescale(d_trawl1$X)
-d_trawl1$y_scaled <- arm::rescale(d_trawl1$Y)
-
-m_temp_SST <- sdmTMB(d_trawl1,
-  temperature_c ~ year_scaled + poly(SST_scaled,3)*depth_scaled , #+ SST_scaled*x_scaled + SST_scaled*y_scaled
-  #time_varying = ~ 0 + depth_scaled2,
-  ar1_fields = FALSE, include_spatial = FALSE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = FALSE
-)
-
-p_SST <- predict(m_temp_SST)
-
-p_SST$residuals <- residuals(m_temp_SST)
-
-ggplot(p_SST, aes(depth_scaled, residuals)) +
-  geom_point(alpha=0.4) +
-  ylim(-8,8) +
-  geom_smooth()
-
-```
-
-```{r}
 library(gbm)
 
-m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl1, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
-
-m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl1, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
+m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
+m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
 
@@
   geom_smooth()
-```
-
-```{r}
-
-p_SST <- predict(m_temp_SST)
-p <- predict(m_temp)
-#predictions3b <- predict(m_temp3_nosst)
-
-p_SST$residuals <- residuals(m_temp_SST)
-p$residuals <- residuals(m_temp)
-
-# nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth)
-# 
-# ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
-#   geom_point() + geom_abline(intercept = 0, slope = 1)
-```
-Residual plots:
-```{r}
-ggplot(p_SST, aes(est, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  ylim(-8,8) +
-  xlim(2,15) +
-  facet_wrap(~year)
-ggplot(p, aes(est, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  ylim(-8,8) +
-   xlim(2,15) +
-  facet_wrap(~year)
-```
-
-```{r}
-ggplot(p_SST, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  ylim(-7,8) +
-  facet_wrap(~year)
-ggplot(p, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  ylim(-7,8) +
-  facet_wrap(~year)
-```
-
-```{r}
-p_SST$depth_mean <- exp(p_SST$depth_scaled*0.66+5)
-glimpse(d_trawl)
-ggplot(d_trawl, aes(depth_scaled, r)) +
-  geom_point(alpha=0.4) +
-  ylim(-8,8) +
-  geom_smooth()
-```
-
-```{r}
-p1 <- ggplot(p_SST, aes(depth_scaled, residuals)) +
-  geom_point() +
-  ylim(-8,8) +
-  geom_smooth()
-p2 <- ggplot(p, aes(depth_scaled, residuals)) +
-  geom_point() +
-  ylim(-8,8) +
-  geom_smooth()
-gridExtra::grid.arrange(p1, p2, nrow = 2)
 ```
@@
 Current top model
-```{r}
-plan(multiprocess)
-temp_rw <- sdmTMB_cv(
-  data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
-  formula = temperature_c ~ 0 + as.factor(year) + SST_scaled,
-  time_varying = ~ 0 + depth_scaled + depth_scaled2,
-  time = "year", family = gaussian(link = "identity"),
-  ar1_fields = FALSE, include_spatial = TRUE, seed = 999
-)
-sum(temp_rw$sum_loglik)
-```
 
 ```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 220)
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 400)
 plot_spde(spde)
@@
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
-  ar1_fields = FALSE, include_spatial = TRUE,
+  ar1_fields = TRUE, include_spatial = TRUE, # changed AR1_field to true
   time = "year", spde = spde, family = gaussian(link = "identity"),
-  silent = TRUE
+  silent = FALSE
 )
 
+saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn.rds")))
+
+```
+
+```{r}
 predictions_rw <- predict(m_temp_rw)
@@
 # aic(m_temp3_rw)
+
 ```
@@
 
-```{r}
+```{r eval=FALSE}
 m_temp3 <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year) + SST_scaled*depth_scaled + SST_scaled*depth_scaled2 + SST_scaled*depth_scaled3,
+  temperature_c ~ 0 + as.factor(year) + SST_scaled*depth_scaled + SST_scaled*depth_scaled2 ,
   ar1_fields = FALSE, include_spatial = TRUE,
@@
 
-```{r}
+```{r eval=FALSE}
 ggplot(predictions3, aes(est, residuals)) +
@@
 
-```{r}
+```{r eval=FALSE}
 ggplot(predictions3, aes(depth_scaled, residuals)) +
@@
 ```{r}
-p_nd <- readRDS("../VOCC/data/predicted_temp_allyears.rds")
+nd_all <- readRDS(paste0("../VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
+unique(nd_all$year)
+glimpse(nd_all)
+predict_temp <- predict(m_temp_rw, newdata = nd_all)
+saveRDS(predict_temp, file = "../VOCC/data/predicted_temp_allyears_revised")
+p_nd <- predict_temp
+#p_nd <- readRDS("../VOCC/data/predicted_temp_allyears.rds")
 # View(p_nd)
@@
     scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3", limits = c(-1.5,2.5))+
-  labs(fill = "") + 
+ high = "Red 3", limits = c(-1.5,1.5),
+    na.value = "red") +  labs(fill = "") + 
 gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))
@@
   ggtitle("Spatiotemporal random effects only") +
-  facet_wrap(~year) +
-  scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3")
+  facet_wrap(~year) +   scale_fill_gradient2( low = "Steel Blue 4",
+ high = "Red 3", limits = c(-1.5,1.5),
+    na.value = "red")
 ```
@@
   facet_wrap(~year) +
-  scale_fill_gradient2(
-    low = "Steel Blue 4",
-    high = "Red 3", 
-    limits = c(-1,1), 
+   scale_fill_gradient2( low = "Steel Blue 4",
+ high = "Red 3", limits = c(-1.5,1.5),
     na.value = "red") +
@@
      scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3", limits = c(-1,1)) +
+ high = "Red 3", limits = c(-1.5,1.5),
+    na.value = "red") +
           labs(fill = " ") + 
@@
    scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3", limits = c(-1,1)) +
+ high = "Red 3", limits = c(-1.5,1.5),
+    na.value = "red") +
   xlim(65,345) +

### 2019-10-21 analysis/tmb-sensor-explore/3-fit-do.Rmd
update climate models/predictions

@@
 
-d_trawl1 <- d_trawl #%>% filter(ssid != 16) %>% filter(ssid != 4)    
+d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4)) #%>% filter(ssid != 16) %>% filter(ssid != 4)    
+#test <- d_trawl1 %>% filter(year == 2016 & ssid ==16) 
 spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 800)
@@
 
-saveRDS(do_model_all_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY2.rds")))
+saveRDS(do_model_all_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds")))
 time1
@@
 
-```{r}
-do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY.rds"))
-do_model_DOY
+```{r eval=FALSE}
+# do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY.rds"))
+# do_model_DOY
 
-get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")
+get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
 
@@
 Predict DO for whole grid using quadratic of predicted temperature values and mean DOY
-```{r}
+```{r eval=FALSE}
 p_nd <- predict(do_model_DOY, newdata = newdata)
@@
 
+Save new predictions without QCS gap and WCVI 2016 DO
+```{r eval=FALSE}
+#new temp and do prediction grid for vectors
+
+model1 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-temp-2008-on-used.rds")
+#model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-all-DOY-used.rds")
+#update model to exclude WCVI 2016 outliers
+model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds")
+
+nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
+nd <- filter(nd_all, year >2007) %>% filter(year<2019)
+pred_temp <- predict(model1, newdata = nd)
+predtemp <- pred_temp %>% mutate(temp_scaled = est, temp_scaled2 = est^2, depth_scaled3 = depth_scaled^3) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
+predtemp$DOY_scaled <- 0
+pred_do <- predict(model2, newdata = predtemp)
+pred_do$temp <- pred_do$temp_scaled*1.296 + 6.49
+
+pred_do$log_do <- pred_do$est
+pred_do$do_est <- exp(pred_do$log_do)
+
+# may want to filter estimates for depths less than 15 m
+# pred_do <- pred_do %>% filter(depth>15)
+
+saveRDS(pred_do, file="~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-new.rds")
+
+```
+
 Plot predictions from saved DO predictions file
@@
 ```{r}
-p_nd <- readRDS((here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
+#p_nd <- readRDS((here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
+p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-new.rds")))
 
@@
 ```{r}
-p_nd4 <- p_nd %>% filter(ssid==4)
+p_nd4 <- p_nd %>% filter(ssid==4) %>% filter(year!=2016)
 pred_4 <- plot_map_raster(p_nd4, "do_est") +
@@
 ```{r}
-p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
+p_nd4 <- p_nd %>% filter(ssid==4) %>% filter(year!=2016)
 st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
@@
 png(
-  file = "figs/do-predictions.png", 
+  file = "figs/do-predictions-without-wcvi2016.png", 
   res = 600,

### 2019-11-05 analysis/VOCC/1-split-biomass-by-maturity.Rmd
remove shallow Redbanded rockfish

@@
 params: 
-    species: "Pacific Cod"
+    species: "Redbanded Rockfish"
     region: "All synoptic surveys"
@@
   bath <- readRDS("data/bathymetry-data")
+
   # for POP: exclude sample with unusually large fish 
@@
   # fish <- fish[fish$fishing_event_id!=1506954,]
+
+  # for Redbanded: 1 fish found at too shallow a depth... 
+  # probably leftover in net?
+  # survey_sets <- survey_sets[survey_sets$fishing_event_id!=2536031,] 
+  # fish <- fish[fish$fishing_event_id!=2536031,]
 

### 2019-11-08 analysis/VOCC/3-bioclimatic-trends.Rmd
plot data from biomass model with climate variables

---
title: "`r paste(params$species)` biomass trends and figures"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    start_year: 2008
    bind_predictions: FALSE
    covs: "-log-both-noAR1" # covariate string used to name model rds 
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set( fig.width = 9, fig.height = 6.5, #fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r params}
species <- params$species
region <- params$region
start_year <- params$start_year

bind_predictions <- params$bind_predictions
covs <- params$covs
priors <- FALSE

paste("region =", region)
paste("ssids modelled separately =", bind_predictions)
paste("model label =", covs)
paste("priors =", priors)

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

Save predictions for full spatial grid
```{r}
#fish <- readRDS(paste0("raw/bio-data-", spp, ""))

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

maturity <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))

if (any(names(maturity) == "adult_density")) {
#if (length(unique(fish$maturity_code)) > 2) {
  nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
  adult_biomass <- readRDS(paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
  #nd$year <- as.integer(nd$year)
  nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(adult_biomass$data$year)) %>% 
    mutate(
      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
      do_mlpl_scaled2 = do_mlpl_scaled^2)
  nd <- na.omit(nd)

  adult_predictions_obj <- predict(adult_biomass, newdata = nd, return_tmb_object = TRUE)
  adult_predictions <- adult_predictions_obj$data

  adult_predictions$est_rw_i <- adult_predictions_obj$obj$report()$proj_rw_i
  # test_eta <- adult_predictions_obj$obj$report()$proj_eta
  # test_fe <- adult_predictions_obj$obj$report()$proj_fe
  # 
  # test_diff <- test_eta - test_fe
  # range(adult_predictions$est_rw_i)
  # report <- adult_predictions_obj$obj$report()

  saveRDS(adult_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"))

  imm_biomass <- readRDS(paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

    nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(imm_biomass$data$year))%>% 
    mutate(do_mlpl_scaled = (log_do - imm_biomass$data$do_mlpl_mean[1]) /(imm_biomass$data$do_mlpl_sd[1]*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
  nd <- na.omit(nd)
  #nd$year <- as.numeric(nd$year)

  imm_predictions_obj <- predict(imm_biomass, newdata = nd, return_tmb_object = TRUE)
  imm_predictions <- imm_predictions_obj$data

  imm_predictions$est_rw_i <- imm_predictions_obj$obj$report()$proj_rw_i

  saveRDS(imm_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"))

} else {

    total_biomass <- readRDS(paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))

    nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
    nd <- nd_all %>% 
      filter(ssid %in% model_ssid) %>%
      filter(year %in% unique(total_biomass$data$year))%>% 
    mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    predictions <- predict(total_biomass, newdata = nd)
    saveRDS(predictions, file = paste0("data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
}
```

Load model predictions
```{r results = "hide"}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

rm(pred1n3)
rm(pred4)
rm(pred16)

rm(pred1n3_imm)
rm(pred4_imm)
rm(pred16_imm)

rm(pred1n3_all)
rm(pred4_all)
rm(pred16_all)

if (region != "All synoptic surveys") {
  try(adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-mature-bioclimatic-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))
  try(imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-imm-bioclimatic-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))

  if (!exists("adult_predictions")) {
    paste("Maturity split biomass predictions are not available for this survey.")

    try(predictions <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3n4n16-total-bioclimatic-prior-", priors, ".rds"
    )) %>% filter(ssid %in% model_ssid))

    if (!exists("predictions")) {
      paste("Total biomass predictions for this survey are also not available.")
    }
  }
} else {

  try({adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"
    ))
  })
  try({imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"
    ))
  })

  # OR bind together predictions from different models
  if (bind_predictions) {
    try({pred1n3 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3-mature-bioclimatic-prior-", priors, ".rds"
      ))
    })
    try({pred4 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-4-mature-bioclimatic-prior-", priors, ".rds"
      ))
    })
    try({pred16 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-16-mature-bioclimatic-prior-", priors, ".rds"
      ))
    })

    if (!exists("pred1n3")) pred1n3 <- NULL
    if (!exists("pred16")) pred16 <- NULL
    if (!exists("pred4")) pred4 <- NULL
    adult_predictions <- rbind(pred1n3, pred4, pred16)

    try({pred1n3_imm <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3-imm-bioclimatic-prior-", priors, ".rds"
      ))
    })
    try({pred4_imm <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-4-imm-bioclimatic-prior-", priors, ".rds"))
    })
    try({pred16_imm <- readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-16-imm-bioclimatic-prior-", priors, ".rds"))
    })

    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
    if (!exists("pred16_imm")) pred16_imm <- NULL
    if (!exists("pred4_imm")) pred4_imm <- NULL
    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)
  }
  }

  if (!exists("adult_predictions")) {
      paste("Maturity split biomass predictions are not available.")

      try({predictions <- readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
      })

      if (!exists("predictions")) {
        try({pred1n3_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-1n3-total-bioclimatic-prior-", priors, ".rds"))
        })
        try({pred4_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-4-total-bioclimatic-prior-", priors, ".rds"))
        })
        try({pred16_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-16-total-bioclimatic-prior-", priors, ".rds"))
        })
        try({
          predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
        })

        if (!exists("predictions")) paste("Total biomass predictions are not available.")
      }
    }
```

Caluclate variables of interest from predictions
```{r}
rm(max_raster)
rm(max_adult)
rm(max_bio)

if (exists("adult_predictions")) {
  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000
  adult_predictions$bioclimatic <- exp(adult_predictions$est - adult_predictions$epsilon_st) * 10000
  imm_predictions$bioclimatic <- exp(imm_predictions$est - imm_predictions$epsilon_st) * 10000

  adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
  imm_predictions$prop_imm <- 
  imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)

  saveRDS(adult_predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-mat-bioclimatic-prior-", priors, ".rds"
  ))
  saveRDS(imm_predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-imm-bioclimatic-prior-", priors, ".rds"
  ))

  adult_predictions <- filter(adult_predictions, year >= start_year)
  imm_predictions <- filter(imm_predictions, year >= start_year)

  max_raster <- quantile(adult_predictions$est_exp, 0.999)
  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)

  model_ssid <- unique(adult_predictions$ssid)
} else {
  predictions <- filter(predictions, year >= start_year)

  predictions$est_exp <- exp(predictions$est) * 10000
  predictions$bioclimatic <- exp(predictions$est - predictions$epsilon_st) * 10000

  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .999)
  #max_bio <- round(max(predictions$total_bio))
  max_bio <- signif(max(predictions$est_exp), digits = 2)

  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-total-bioclimatic-prior-", priors, ".rds"
  ))

  model_ssid <- unique(predictions$ssid)
}
```

## Mature (or total) biomass predictions
```{r, error=TRUE}
if (exists("adult_predictions")) {
  p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
    raster_limits = c(0, max_raster),
    legend_position = "none",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
  print(p_adult_all)
} else {
  p_adult_all <- plot_facet_map(predictions, "total_bio",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))

  print(p_adult_all)
}
```

### Split by survey (SSID)
```{r}
if (exists("adult_predictions")) predictions <- adult_predictions

  if (1 %in% model_ssid) {
    adult_predictions1 <- predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    if (any(predictions$ssid == 16)) {
      adult_predictions16 <- predictions %>% filter(ssid == 16)
    } else {
      adult_predictions16 <- NULL
    }
  } else {
    adult_predictions16 <- NULL
  }

  ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)
  ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]

  p_adult <- list()
  if(exists("adult_predictions")) {
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  } else {
    for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
    p_adult[[i]]  
    }
  }
  print(p_adult)
```

```{r, message=FALSE, warning=FALSE}
rm(out_mature)
rm(out_total)
# try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-bioclimatic-priors-", priors, ".rds")))

if (!exists("out_mature")) {
  if (exists("adult_predictions")) {
    predictions <- adult_predictions
  } else {
   try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-bioclimatic-priors-", priors, ".rds")))
  }

  if (!exists("out_total")) {
    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature3 <- NULL
    }

    if (4 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 4)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature4 <- NULL
    }

    if (16 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 16)
      #unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        skip_time = 2007,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature16 <- NULL
    }

    out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)

    if (exists("adult_predictions")) {
            saveRDS(out_mature, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-mature-bioclimatic-priors-", priors, ".rds"))

         } else {
      out_total <- out_mature
      rm(out_mature)
      saveRDS(out_total, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-total-bioclimatic-priors-", priors, ".rds"))
    }
  }
}
```

### Recent biomass with trawl footprint
```{r }
if (!exists("out_mature")) {out_recent <- out_total} else {out_recent <- out_mature}

  trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
  proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
  trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)

  current_biomass <- plot_vocc(out_recent,
    fill_col = "est_exp_e",
    fill_label = "Mean\nbiomass (kg/ha)\npost-2010",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
  current_biomass <- current_biomass +
    geom_sf(data = trawl_footprint, aes(geometry = geometry / 1000), colour = "gray", fill = "gray", alpha = 0.2, lwd = 0.25, inherit.aes = FALSE) +
    coord_sf(xlim = c(180, 820), ylim = c(5350, 6050), datum = NULL) +
    xlab("") +
    ylab("") +
    theme(panel.grid.major = element_line(colour = "transparent")) +
    ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
  current_biomass

```

## Trends in biomass density
```{r, error=TRUE}
if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "lightseagreen", # limegreen",
    low_fill = "red", # "deeppink3",# "steel blue 4",
    transform_col = fourth_root_power,
    #raster_limits = c(0, quantile(data$units_per_decade, 0.99999)),
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(as.numeric(adult_predictions[["year"]]))), "-", max(unique(as.numeric(adult_predictions[["year"]])))
  ))
  biotrend_ad
} else {
  biotrend_ad <- plot_vocc(out_total,
    fill_col = "units_per_decade",
    fill_label = "Biomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
    low_fill = "red", # "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(as.numeric(predictions[["year"]]))), "-", max(unique(as.numeric(predictions[["year"]])))
  ))
  biotrend_ad
}
```

# Immature biomass density
### if available
```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_imm_all <- plot_facet_map(imm_predictions, "est_exp",
    transform_col = fourth_root_power,
    raster_limits = c(0, max_raster)
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n "))

  p_imm_all
} else {
  p_imm_all <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    imm_predictions1 <- imm_predictions %>% filter(ssid == 1)
  } else {
    imm_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    imm_predictions3 <- imm_predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    imm_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    imm_predictions4 <- imm_predictions %>% filter(ssid == 4)
  } else {
    imm_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    imm_predictions16 <- imm_predictions %>% filter(ssid == 16)
  } else {
    imm_predictions16 <- NULL
  }

  im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)
  im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]

  p_imm <- list()
  for (i in seq_len(length(model_ssid))) {
    p_imm[[i]] <- plot_facet_map(im_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))

    p_imm[[i]]
  }
  print(p_imm)
}
```

```{r, message=FALSE, warning=FALSE}
# try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-bioclimatic-priors-", priors, ".rds")))
rm(out_imm)
if (!exists("out_imm")) {
  if (exists("imm_predictions")) {
    predictions <- imm_predictions

    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
      out_im1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im3 <- NULL
    }

    if (4 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 4)
      unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im4 <- NULL
    }

    if (16 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 16)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im16 <- NULL
    }

    out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
    saveRDS(out_imm, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-bioclimatic-priors-", priors, ".rds"
    ))
  } else {
    rm(out_imm)
  }
}
```

```{r}
if (exists("out_imm")) {
  data <- out_imm
  biotrend_im <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
    low_fill = "red", # "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))
  ))
  biotrend_im
} else {
  biotrend_im <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

Save trend plots
```{r results = "hide"}
png(
  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 11,
  height = 8.5
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

Save spatiotemporal biomass plots
```{r results = "hide"}
if(exists("adult_predictions")) { biomass_plots <- c(p_adult, p_imm) } else {
  biomass_plots <- c(p_adult)
}

png(
  file = paste0("figs/", spp, "/bioclimatic-biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

Save plot medley
```{r results = "hide"}
png(
  file = paste0("figs/", spp, "/bioclimatic-exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 9.5,
  height = 12.5
)

gridExtra::grid.arrange(
  current_biomass,
  p_prop_imm,
  p_adult_all, p_imm_all,
  biotrend_ad, biotrend_im,
  nrow = 3
)
dev.off()
```

```{r}
# plot_map_raster <- function(dat, column = "est") {
#     ggplot(dat, aes_string("X", "Y", fill = column)) +
#     geom_raster() +
#     coord_fixed()+
# gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
# }

 # plot_map_raster(adult_predictions, "epsilon_st") +
 #  ggtitle("Spatiotemporal random effects only") +
 #  facet_wrap(~year) +
 #  scale_fill_gradient2( high = "Steel Blue 4",
 #  low = "Red 3") + coord_fixed()

 if (exists("adult_predictions")) {
  # p_st <- list()
  # for (i in seq_len(length(model_ssid))) {
  #   p_st[[i]] <- plot_map_raster(ad_prediction_list[[i]], "epsilon_st") +
  #     ggtitle("Spatiotemporal random effects only") +
  #     facet_wrap(~year) +
  #     scale_fill_gradient2(high = "#66C2A5", #mid = "black", 
  #       low = "#5E4FA2"#, trans = fourth_root_power
  #     )
  #   p_st[[i]]
  #  }
  # print(p_st)

  p_adult_st <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult_st[[i]] <-plot_facet_map(ad_prediction_list[[i]], "epsilon_st",
    white_zero = TRUE,
    high_fill = "#66C2A5",
    low_fill = "#5E4FA2"
    #raster_limits = c(0, max_raster),
    #legend_position = "none",
    #transform_col = fourth_root_power
  ) +
    #labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature spatiotemporal variation"))
    p_adult_st[[i]]
  }
  print(p_adult_st)
#} else {
  # p_adult_st <- plot_facet_map(predictions, "epsilon_st",
  #   transform_col = fourth_root_power
  # ) +
  #   #labs(fill = "kg/ha") +
  #   ggtitle(paste0("", species, "total spatiotemporal variation"))
  # 
  # print(p_adult_st)
}
```

```{r}
adult_predictions$year.f <- as.factor(adult_predictions$year)

glimpse(adult_predictions)

data <- mutate(adult_predictions, cell_id = paste0(X,Y)) %>% 
  group_by(cell_id) %>% 
  mutate(cell_max = max(bioclimatic)) %>% group_by(year.f) %>% mutate(ann_biomass = sum(est_exp))
unique(data$ann_biomass)
 glimpse(data)

# start_biomass <- filter(adult_predictions, year == min(as.numeric(adult_predictions$year))) 
#     bio5perc <- sum(adult_predictions$start_density)*0.05
#     s <- sort(annual_biomass$start_density)
#     bio_sum <- cumsum(s)
#     lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]

threshold <- quantile(data$cell_max, 0.2)
library(forcats)
violin_data <- data %>% filter (cell_max > threshold) %>% mutate(year = fct_reorder(year.f, desc(ann_biomass)))
 glimpse(violin_data)

ggplot(violin_data) + 
  geom_hline(aes(yintercept=0)) +
  geom_violin(aes(fct_reorder(year.f, (ann_biomass)), epsilon_st, fill = year.f)) +   
  geom_text(aes(fct_reorder(year.f, (ann_biomass)), y=2.8, label = round(ann_biomass)), check_overlap = TRUE) +
  scale_fill_viridis_d(option = "C") +
  ylab("Spatiotemporal variation") +
  xlab("Year") +
  #coord_flip() +
  theme(legend.position="none") +
  ggtitle(paste("Local", species, "abundance deviations not explained by climate and overall abundance"))
```

```{r}
# data <- adult_predictions %>% group_by (year) %>% 
#   mutate(mean_temp = mean(temp)) %>% 
#   filter (est_exp > quantile(est_exp, 0.05)) %>% 
#   filter (do_est > 1.5)
# 
# ggplot(data, aes(temp, epsilon_st)) + 
#   geom_point(alpha = 0.025) + 
#   geom_smooth(method="lm", formula = y~poly(x,2)) +
#   facet_wrap(~year) +
#   scale_color_viridis_d(option = "C") +
#   guides(color = FALSE) +
#   ylab("Spatiotemporal variation") +
#   ggtitle(paste("Local", species, "abundance deviations not explained by climate and overall abundance"))

```

Save spatiotemporal residual plots
```{r results = "hide"}
# if(exists("adult_predictions")) { resid_plots <- c(p_st, p_imm) } else {
#   biomass_plots <- c(p_adult)
# }

png(
  file = paste0("figs/", spp, "/bioclimatic-residuals-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = p_adult_st,
  nrow = 2
)
dev.off()
```

```{r}
 if (exists("adult_predictions")) {
  p_bioclim <- list()
  for (i in seq_len(length(model_ssid))) {
    p_bioclim[[i]] <- plot_facet_map(ad_prediction_list[[i]], "bioclimatic",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " predicted biomass based on climate \nand fixed spatial variation"))
    p_bioclim[[i]]
  }
 }
  print(p_bioclim)

```

Save bioclimatic predictions plots
```{r results = "hide"}
# if(exists("adult_predictions")) { resid_plots <- c(p_st, p_imm) } else {
#   biomass_plots <- c(p_adult)
# }

png(
  file = paste0("figs/", spp, "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = p_bioclim,
  nrow = 2
)
dev.off()
```

```{r}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
# Velocity of climate change ---------------------------------------------

d <- ad_prediction_list[[2]]

d$layer <- d$bioclimatic

# Just to confirm the trends we're seeing are real,
# mod = mgcv::gam(est ~ s(year, k=5) + te(X,Y) + log(depth),
#   data=d)
# plot(mod)

# This parameter controls how the original 2-km projection is aggregated.
# for example, a value of 5 means that the raster would be reprojected to 10km grid
scale_fac <- 1

# create a RasterBrick
# raster for each year
rlist <- list()
for (i in 1:length(unique(d$year))) {
  rlist[[i]] <- raster::rasterFromXYZ(dplyr::filter(d, year == unique(d$year)[i]) %>%
    dplyr::select(X, Y, layer))
  rlist[[i]] <- raster::aggregate(rlist[[i]], fact = scale_fac)
}
# stack rasters into layers -> rasterbrick
rstack <- raster::stack(rlist[[1]], rlist[[2]])
for (i in 3:length(rlist)) {
  rstack <- raster::stack(rstack, rlist[[i]])
}
rbrick <- raster::brick(rstack)

# Then calculate the trend per pixel:
slopedat <- vocc::calcslope(rbrick)
# Then get the mean temperature for a time period and calculate the spatial gradient:
allyears <- rep(1, raster::nlayers(rbrick))
mnsst <- raster::stackApply(rbrick, indices = allyears, fun = mean)
spatx <- vocc::spatialgrad(mnsst)
# Now we can calculate the VoCC:
velodf <- vocc::calcvelocity(spatx, slopedat, y_dist = 1)

# Mapping it again is straightforward:
rtrend <- rgrad_lon <- rgrad_lat <- rvocc <- angle <- magn <- raster::raster(rbrick)
rgrad_lat[spatx$icell] <- spatx$NS # latitude shift, NS
rgrad_lon[spatx$icell] <- spatx$WE # longitude shift, WE
rtrend[slopedat$icell] <- -1 * slopedat$slope
rvocc[velodf$icell] <- velodf$velocity

# convert to data frames for ggplot
rtrend_df <- as.data.frame(raster::rasterToPoints(rtrend)) %>%
  dplyr::rename(trend = layer)
rgradlat_df <- as.data.frame(raster::rasterToPoints(rgrad_lat)) %>%
  dplyr::rename(gradNS = layer)
rgradlon_df <- as.data.frame(raster::rasterToPoints(rgrad_lon)) %>%
  dplyr::rename(gradWE = layer)
rvocc_df <- as.data.frame(raster::rasterToPoints(rvocc)) %>%
  dplyr::rename(velocity = layer)

# create ggquiver plots. need dataframe of lon, lat, delta_lon, delta_lat, trend, velocity
df <- dplyr::left_join(rtrend_df, rgradlat_df, by = c("x", "y")) %>%
  dplyr::left_join(rgradlon_df, by = c("x", "y")) %>%
  dplyr::left_join(rvocc_df, by = c("x", "y"))

gtrend <- ggplot(df, aes(x, y, fill = -trend)) +
  geom_raster() +
  scale_fill_gradient2(low = scales::muted("blue"), high = scales::muted("red")) +
  xlab("Longitude") + ylab("Latitude") + ggtitle("Trend") + coord_fixed()

# spatial gradient plot
quantile_cutoff <- 0.05 # for plotting
df <- dplyr::mutate(df,
  u_velo = trend / gradWE,
  v_velo = trend / gradNS,
  ulow = quantile(u_velo, quantile_cutoff),
  uhi = quantile(u_velo, 1 - quantile_cutoff),
  vlow = quantile(v_velo, quantile_cutoff),
  vhi = quantile(v_velo, 1 - quantile_cutoff),
  u_velo = ifelse(u_velo < ulow, ulow, u_velo),
  u_velo = ifelse(u_velo > uhi, uhi, u_velo),
  v_velo = ifelse(v_velo < vlow, vlow, v_velo),
  v_velo = ifelse(v_velo > vhi, vhi, v_velo)
) %>%
  dplyr::select(-ulow, -uhi, -vlow, -vhi)

# gradient plot with ggquiver
ggrad <- ggplot(df) +
  ggquiver::geom_quiver(aes(x, y, u = gradNS, v = gradWE)) +
  xlab("Longitude") + ylab("Latitude") + ggtitle("Gradient") + coord_fixed()

# velocity plot
gvocc <- ggplot(mutate(df, `Trend\n(kg/decade)` = -trend)) +
  ggquiver::geom_quiver(aes(x, y, u = u_velo, v = v_velo,
    colour = `Trend\n(kg/decade)`),
    vecsize = 2) +
  scale_colour_gradient2(low = scales::muted("blue"), high = scales::muted("red")) +
  xlab("UTM") + ylab("UTM") + #ggtitle("Velocity") +
  gfplot::theme_pbs()

# coast <- gfplot:::load_coastline(range(surv$longitude) + c(-1, 1),
#   range(surv$latitude) + c(-1, 1),
#   utm_zone = 9
# )

# isobath <- gfplot:::load_isobath(range(surv$longitude) + c(-5, 5),
#   range(surv$latitude) + c(-5, 5),
#   bath = c(100, 200, 300, 500), utm_zone = 9
# )

# library(ggnewscale)

gvocc <- gvocc +
  labs(colour = "Local\nclimate trend\n(°C/decade)") +
  ggnewscale::new_scale_color() +
# geom_path(
#   data = isobath, aes_string(
#     x = "X", y = "Y",
#     group = "paste(PID, SID)", colour = "PID"
#   ),
#   inherit.aes = FALSE, lwd = 0.4, alpha = 0.4
# ) +
  scale_colour_continuous(low = "grey70", high = "grey10") +
  # labs(colour = "Depth") +
  guides(colour = FALSE) +
  coord_fixed(xlim = range(df$x) + c(-3, 3), ylim = range(df$y) + c(-3, 3))

# gvocc <- gvocc + geom_polygon(
#   data = coast, aes_string(x = "X", y = "Y", group = "PID"),
#   fill = "grey87", col = "grey70", lwd = 0.2
# )

```

### 2019-11-13 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
plots for PICES talk

@@
     region: "both odd year surveys"
+    #region: "West Coast Vancouver Island"
     age: "mature"
@@
     climate: "temperature"
-    threshold: 0.5
+    threshold: 0.25
 editor_options: 
@@
   do_threshold <- params$threshold
+
+   if (threshold) { do_threshold <- params$threshold } else {
+  do_threshold <- root_params[rev(order(root_params$do_maxy)), ]$upper_do_50 - root_params[rev(order(root_params$do_maxy)), ]$upper_do_75
+  do_threshold <- round(do_threshold, 2)
+  }
+
   root_params <- filter(root_params, lower_do_25 > 0.4)
@@
 vocc_by_year <- lapply(start_years, function(start_time) {
-  try({
-    vocc <- readRDS(paste0(
-      "data/vocc-", ssid_string, "-", climate_string, "-",
-      min_string, "to", max_string, "-", start_time, "-", delta_t_total, "-update.rds"
-    ))
-  })
+  # try({
+  #   vocc <- readRDS(paste0(
+  #     "data/vocc-", ssid_string, "-", climate_string, "-",
+  #     min_string, "to", max_string, "-", start_time, "-", delta_t_total, "-update.rds"
+  #   ))
+  # })
 
@@
 
+  all_vectors <- do.call("rbind", vocc_by_year[3:7])
+
+  raster_min <- min(all_vectors$temp_e)
+  raster_max <- max(all_vectors$temp_e)
+
   all_vector_plots_w <- lapply(vocc_by_year, function(vocc) {
@@
       vocc_plot <- plot_vocc(vocc,  
+        # theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
         NA_label = ".",
-        fill_col = "temp_s",
+        fill_col = "temp_e",
         fill_label = paste(" Temperature\n change\n", 
@@
         raster_alpha = 1,
-        # raster_limits = c(raster_min, raster_max),
+        raster_limits = c(raster_min, raster_max),
         na_colour = "gray",
@@
       vocc_plot <- plot_vocc(vocc, 
+        # theme = "black",
         vec_aes = "distance",
@@
         NA_label = ".",
-        fill_col = "do_est_s",
+        fill_col = "do_est_e",
         fill_label = paste(" DO change\n", start_time, 
@@
         raster_alpha = 1,
-        raster_limits = c(raster_min_temp, raster_max_temp),
+        # raster_limits = c(raster_min_temp, raster_max_temp),
         # white_zero = TRUE,
@@
         NA_label = ".",
-        fill_col = "do_est_s",
+        fill_col = "do_est_e",
         fill_label = paste(" DO change\n", start_time, "to", 
@@
 
-all_vectors <- do.call("rbind", trimmed_vectors)
+all_vectors_trim <- do.call("rbind", trimmed_vectors)
 
 ## If VOCC change is a useful column?
-# all_vectors$var_1_changeA <- all_vectors$var_1_min
-# all_vectors$var_2_changeA <- all_vectors$var_2_min
-# all_vectors$var_1_changeB <- all_vectors$var_1_max
-# all_vectors$var_2_changeB <- all_vectors$var_2_max
-
-all_vectors$var_1_min <- lower_thresholds[1]
-all_vectors$var_2_min <- lower_thresholds[2]
-all_vectors$var_1_max <- upper_thresholds[1]
-all_vectors$var_2_max <- upper_thresholds[2]
+# all_vectors_trim$var_1_changeA <- all_vectors_trim$var_1_min
+# all_vectors_trim$var_2_changeA <- all_vectors_trim$var_2_min
+# all_vectors_trim$var_1_changeB <- all_vectors_trim$var_1_max
+# all_vectors_trim$var_2_changeB <- all_vectors_trim$var_2_max
+
+all_vectors_trim$var_1_min <- lower_thresholds[1]
+all_vectors_trim$var_2_min <- lower_thresholds[2]
+all_vectors_trim$var_1_max <- upper_thresholds[1]
+all_vectors_trim$var_2_max <- upper_thresholds[2]
 
 if (climate == "temperature") {
-  all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
-  # raster_min <- min(all_vectors$diff)
-  # raster_max <- max(all_vectors$diff)
-  raster_min <- quantile(all_vectors$diff, 0.005)
-  raster_max <- quantile(all_vectors$diff, 0.995)
+  all_vectors_trim$diff <- all_vectors_trim$units_per_decade / 10 * delta_t_total
+  # raster_min <- min(all_vectors_trim$diff)
+  # raster_max <- max(all_vectors_trim$diff)
+  raster_min <- quantile(all_vectors_trim$diff, 0.005)
+  raster_max <- quantile(all_vectors_trim$diff, 0.995)
 }
@@
 if (climate == "do") {
-  all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
-  # raster_min <- min(all_vectors$diff)
-  # raster_max <- max(all_vectors$diff)
+  all_vectors_trim$diff <- all_vectors_trim$units_per_decade / 10 * delta_t_total
+  # raster_min <- min(all_vectors_trim$diff)
+  # raster_max <- max(all_vectors_trim$diff)
   # 
-  raster_min <- quantile(all_vectors$diff, 0.0005)
-  raster_max <- quantile(all_vectors$diff, 0.9995)
+  raster_min <- quantile(all_vectors_trim$diff, 0.0005)
+  raster_max <- quantile(all_vectors_trim$diff, 0.9995)
 
@@
 if (climate == "do and temperature") {
-  all_vectors$temp_diff <- all_vectors$temp.units_per_decade / 10 * delta_t_total
-  all_vectors$do_diff <- all_vectors$do_est.units_per_decade / 10 * delta_t_total
-  raster_min_temp <- min(all_vectors$temp_diff)
-  raster_max_temp <- max(all_vectors$temp_diff)
-  raster_min_do <- min(all_vectors$do_diff)
-  raster_max_do <- max(all_vectors$do_diff)
+  all_vectors_trim$temp_diff <- all_vectors_trim$temp.units_per_decade / 10 * delta_t_total
+  all_vectors_trim$do_diff <- all_vectors_trim$do_est.units_per_decade / 10 * delta_t_total
+  raster_min_temp <- min(all_vectors_trim$temp_diff)
+  raster_max_temp <- max(all_vectors_trim$temp_diff)
+  raster_min_do <- min(all_vectors_trim$do_diff)
+  raster_max_do <- max(all_vectors_trim$do_diff)
 }
@@
       vocc_plot <- plot_vocc(vocc, 
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
       vocc_plot <- plot_vocc(vocc,  
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
       vocc_plot_temp <- plot_vocc(vocc,  
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
       vocc_plot_do <- plot_vocc(vocc,  
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
 ### Plot remaining climate vectors for regions where change exceeds thresholds
-```{r echo=FALSE, warning=FALSE}
+```{r , eval=FALSE, echo=FALSE, warning=FALSE}
 
@@
       vocc_plot <- plot_vocc(vectors,
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
       vocc_plot_temp <- plot_vocc(vectors,
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
       vocc_plot_do <- plot_vocc(vectors,
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
 
-```{r}
+```{r, eval=FALSE}
 png(
@@
 # FIXME: this code could/maybe should be rewritten with an if_else mutate and then distinct...
-source_cells <- all_vectors %>%
+source_cells <- all_vectors_trim %>%
   filter(distance > (-dist_intercept)) %>%
@@
 # identify all possible control cells have been given a distance == - dist_intercept
-control_cells <- all_vectors %>%
+control_cells <- all_vectors_trim %>%
   filter(distance == (-dist_intercept)) %>%
@@
 alldata <- rbind(source_biomass, control_biomass)
-alldata <- mutate(alldata, treat = ifelse(cell_type == "source", 1, 0)) %>%
+alldata <- mutate(alldata, treat = ifelse(cell_type == "source", 1, 0), dist_x_biomass_s = start_density*distance, dist_x_biomass_e = end_density*distance) %>%
   select(-mean_target_X, -mean_target_Y) %>%
@@
 
+alldata <- alldata %>% group_by(start_time) %>% mutate(
+  start_total_biomass = sum(start_density), 
+  end_total_biomass = sum(end_density),
+  sum_weighted_dist = sum(dist_x_biomass_s),  
+  ann_climate_pressure = sum_weighted_dist/start_total_biomass)
+
+pressure <- unique(alldata$ann_climate_pressure)
+start_biomass <- unique(alldata$start_total_biomass)
+end_biomass <- unique(alldata$end_total_biomass)
+score_year <- unique(alldata$start_time) + 2
+
+scores <- as.data.frame(cbind(score_year, pressure, start_biomass, end_biomass))
+
 if(is.na(alldata$var_1_max[2])) {
@@
 # })
+
+ggplot(scores, aes(pressure, (end_biomass-start_biomass), colour = as.factor(score_year))) + 
+  geom_point(size =3) + 
+  #geom_point(aes(end_biomass, pressure, colour = as.factor(score_year)), 
+    #shape = 1, size = 3) + 
+  scale_colour_viridis_d()
+
+print(scores)
 ```
@@
 # raster_max <- max((log(alldata$end_density+1)-log(alldata$start_density+1)))
-raster_min <- quantile(((alldata$end_density)-(alldata$start_density)), 0.01)
-raster_max <- max((alldata$end_density)-(alldata$start_density))#, 0.9999)
+# raster_min <- quantile(((alldata$end_density)-(alldata$start_density)), 0.01)
+# raster_max <- max((alldata$end_density)-(alldata$start_density))#, 0.9999)
 
@@
 
+    raster_min <- min(biomass$density_diff) #, 0.01)
+    raster_max <- -(raster_min)#, 0.9999)
+
     vocc_plot <- plot_vocc(biomass,
+        theme = "black",
+        coast = TRUE,
         vec_aes = "distance",
@@
         # varidis_option = 
-        white_zero = FALSE,
-        # mid_fill = NULL,
-        # high_fill = "#5E4FA2", #"Steel Blue 4"
-        # low_fill = "#D53E4F",
+        #white_zero = FALSE,
+        white_zero = TRUE,
+        mid_fill = "black",
+        high_fill = "#66C2A5", #"Steel Blue 4", "#3288BD", #
+        low_fill = "#5E4FA2FF", #"Red",
         # high_fill = "#21908CFF", # "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
         # low_fill = "#FDE725FF", #"#FEE08B", # "#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B"   
-        na_colour = "red",
+        na_colour = "#66C2A5", #"3288BD", #"Steel Blue 4",
         vec_col = "white",
@@
         axis_lables = FALSE,
-        transform_col =  sqrt # no_trans ##fourth_root_power #
+        transform_col =  fourth_root_power #sqrt # no_trans ##
       ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
-        biomass_threshold, "% of peak predicted ", species, " biomass" ))
+        biomass_threshold, "% of peak predicted ", species, " biomass" )) #+ theme_black()
 
@@
 
-### Plot climate vectors for cells with starting with greater than 25% percentile of biomass densities
+### Plot all climate vectors against biomass
 ```{r echo=FALSE, warning=FALSE}
@@
     vocc_plot <- plot_vocc(biomass,
+      theme = "black",
+       coast = TRUE,
+        vec_aes = "distance",
+        vec_col = "white",
+        arrowhead_size = 0.01,
+        vec_lwd_range = c(.7, 1),
+        min_vec_plotted = input_cell_size*2,
+        #max_vec_plotted = max_vector,
+        NA_label = ".",
+        fill_col = "start_density",
+        fill_label = paste("Biomass\n change (kg/ha)\n", 
+            start_time, "to", start_time + delta_t_total, ""),
+        raster_alpha = 1,
+        viridis_option = "C",
+        #raster_limits = c(raster_min, raster_max),
+        #white_zero = TRUE,
+        #high_fill = "#21908CFF", # "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
+        #low_fill = "#FDE725FF", #"#FEE08B", # "#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B"
+        na_colour = "red",
+        vec_alpha = 0.35,
+        axis_lables = FALSE,
+        transform_col = fourth_root_power # no_trans #
+      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
+        biomass_threshold, "% of peak predicted ", species, " biomass" )) 
+       # +
+       #  theme_black()
+       # 
+    vocc_plot
+  })
+})
+
+#print(biomass_vector_plots)
+invisible(lapply(biomass_vector_plots_raw, print))
+```
+
+### Trim for vectors on cells accounting for 95% of biomass
+```{r echo=FALSE, warning=FALSE}
+# # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass across years
+bio5perc <- sum(biomass_change$start_density)*0.05
+s <- sort(biomass_change$start_density)
+bio_sum <- cumsum(s)
+lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+alldata <- alldata %>% filter(start_density > lower_density_threshold)
+
+biomass_vector_plots_raw <- lapply(vocc_by_year, function(vectors) {
+  try({
+    # #offset vectors so that they represent the time step before the one for biomass change
+    # vectors <- vectors %>% mutate(start_time = start_time + 2, end_time = end_time + 2)
+
+    start_time <- vectors$start_time[1]
+
+    # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass within years
+    annual_biomass <- filter(biomass_change, start_time == !!start_time) 
+    bio5perc <- sum(annual_biomass$start_density)*0.05
+    s <- sort(annual_biomass$start_density)
+    bio_sum <- cumsum(s)
+    lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
+
+    #vectors <- vectors %>% filter(start_time == !!start_time)
+
+    biomass <- left_join(biomass_change, vectors, 
+      by = c("icell", "x", "y", "start_time")) %>% 
+      filter(start_time == !!start_time) %>% 
+      mutate(density_diff = end_density-start_density,
+       distance = if_else(is.na(distance), 0, distance),
+       distance = if_else(start_density < !!lower_density_threshold, 0, distance),
+       target_X = as.numeric(target_X), target_Y = as.numeric(target_Y),
+       target_X = if_else(start_density < !!lower_density_threshold, NA_real_, target_X),
+       target_Y = if_else(start_density < !!lower_density_threshold, NA_real_, target_Y))
+
+       vocc_plot <- plot_vocc(biomass,
+         theme = "black",
+         coast = TRUE,
         vec_aes = "distance",
@@
 
+### Further trimmed for vectors crossing species sensitivity threshold
 ```{r}
 
-biomass_vector_plots_trim <- lapply(trimmed_vectors, function(vectors) {
+biomass_vector_plots_present <- lapply(trimmed_vectors, function(vectors) {
   try({
@@
       by = c("icell", "x", "y", "start_time")) %>% 
-      filter(start_time == !!start_time)
+      filter(start_time == !!start_time) %>% 
+      mutate(density_diff = end_density-start_density,
+       distance = if_else(is.na(distance), 0, distance),
+       distance = if_else(start_density < !!lower_density_threshold, 0, distance),
+       target_X = as.numeric(target_X), target_Y = as.numeric(target_Y),
+       target_X = if_else(start_density < !!lower_density_threshold, NA_real_, target_X),
+       target_Y = if_else(start_density < !!lower_density_threshold, NA_real_, target_Y))
 
        vocc_plot <- plot_vocc(biomass,
+         theme = "black",
+         coast = TRUE,
         vec_aes = "distance",
@@
 #print(biomass_vector_plots)
-invisible(lapply(biomass_vector_plots_trim, print))
-```
-
-```{r}
-
-biomass_vector_plots_present <- lapply(trimmed_vectors, function(vectors) {
-  try({
-    # #offset vectors so that they represent the time step before the one for biomass change
-    # vectors <- vectors %>% mutate(start_time = start_time + 2, end_time = end_time + 2)
-
-    start_time <- vectors$start_time[1]
-
-    # set lower_density_threshold of cells with cumulative densities of less than 5% of peak biomass within years
-    annual_biomass <- filter(biomass_change, start_time == !!start_time) 
-    bio5perc <- sum(annual_biomass$start_density)*0.05
-    s <- sort(annual_biomass$start_density)
-    bio_sum <- cumsum(s)
-    lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]
-
-    #vectors <- vectors %>% filter(start_time == !!start_time)
-
-    biomass <- left_join(biomass_change, vectors, 
-      by = c("icell", "x", "y", "start_time")) %>% 
-      filter(start_time == !!start_time) %>% 
-      mutate(density_diff = end_density-start_density,
-       distance = if_else(is.na(distance), 0, distance),
-       distance = if_else(start_density < !!lower_density_threshold, 0, distance),
-       target_X = as.numeric(target_X), target_Y = as.numeric(target_Y),
-       target_X = if_else(start_density < !!lower_density_threshold, NA_real_, target_X),
-       target_Y = if_else(start_density < !!lower_density_threshold, NA_real_, target_Y),
-       n_targets = if_else(distance==0, 0, n_targets)
-       )
-
-       vocc_plot <- plot_vocc(biomass,
-        vec_aes = "distance",
-        vec_col = "white",
-        arrowhead_size = 0.01,
-        vec_lwd_range = c(.7, 1),
-        min_vec_plotted = input_cell_size*2,
-        max_vec_plotted = max_vector,
-        NA_label = ".",
-        fill_col = "start_density",
-        fill_label = paste("Biomass\n change (kg/ha)\n", 
-            start_time, "to", start_time + delta_t_total, ""),
-        raster_alpha = 1,
-        viridis_option = "C",
-        #raster_limits = c(raster_min, raster_max),
-        #white_zero = TRUE,
-        #high_fill = "#21908CFF", # "#E6F598" "#ABDDA4" "#66C2A5" "#3288BD" "#5E4FA2"
-        #low_fill = "#FDE725FF", #"#FEE08B", # "#9E0142" "#D53E4F" "#F46D43" "#FDAE61" "#FEE08B"
-        na_colour = "red",
-        vec_alpha = 0.35,
-        axis_lables = FALSE,
-        transform_col = fourth_root_power # no_trans #
-      ) + ggtitle(paste0(threshold, " units of ", climate, " change below ", 
-        biomass_threshold, "% of peak predicted ", species, " biomass" ))
-
-    vocc_plot
-  })
-})
-
-#print(biomass_vector_plots)
 invisible(lapply(biomass_vector_plots_present, print))
@@
 
-```{r}
+```{r, eval = FALSE}
 rm(vocc_by_sp)
@@
 Print diagnostics for matching
-```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
+```{r eval=FALSE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
 if (!exists("vocc_by_sp")) {
@@
 Prep matched data for analysis
-```{r}
+```{r, eval = FALSE}
  if (!exists("vocc_by_sp")) {
@@
 ## Save spp dataframe
-```{r}
+```{r, eval = FALSE}
 # biomass_threshold <- "50"
@@
 
-```{r}
+```{r eval = FALSE}
 try({
@@
 ### Plot matches
-```{r}
+```{r ,eval = FALSE}
 # FIXME: SAVE PLOTS?
@@
 
-```{r}
+```{r, eval = FALSE}
 png(

### 2019-11-13 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
change BACI rmd to use theme_black for black background argument

@@
 params: 
-    species: "Redbanded Rockfish" 
-    region: "both odd year surveys"
-    #region: "West Coast Vancouver Island"
+    species: "Pacific Cod" 
+    #region: "both odd year surveys"
+    region: "West Coast Vancouver Island"
     age: "mature"
@@
     climate: "temperature"
-    threshold: 0.25
+    threshold: 0.5
 editor_options: 
@@
 vocc_by_year <- lapply(start_years, function(start_time) {
-  # try({
-  #   vocc <- readRDS(paste0(
-  #     "data/vocc-", ssid_string, "-", climate_string, "-",
-  #     min_string, "to", max_string, "-", start_time, "-", delta_t_total, "-update.rds"
-  #   ))
-  # })
+  try({
+    vocc <- readRDS(paste0(
+      "data/vocc-", ssid_string, "-", climate_string, "-",
+      min_string, "to", max_string, "-", start_time, "-", delta_t_total, "-update.rds"
+    ))
+  })
 
@@
       vocc_plot <- plot_vocc(vocc,  
-        # theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
       vocc_plot <- plot_vocc(vocc, 
-        # theme = "black",
+        theme_black = TRUE,
+        coast = TRUE,
         vec_aes = "distance",
@@
         raster_alpha = 1,
-        # raster_limits = c(raster_min_temp, raster_max_temp),
+        #raster_limits = c(raster_min_temp, raster_max_temp),
         # white_zero = TRUE,
@@
       vocc_plot <- plot_vocc(vocc, 
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
       vocc_plot <- plot_vocc(vocc,  
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
       vocc_plot_temp <- plot_vocc(vocc,  
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
       vocc_plot_do <- plot_vocc(vocc,  
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
       vocc_plot <- plot_vocc(vectors,
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
       vocc_plot_temp <- plot_vocc(vectors,
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
       vocc_plot_do <- plot_vocc(vectors,
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
     vocc_plot <- plot_vocc(biomass,
-        theme = "black",
+        theme_black = TRUE,
         coast = TRUE,
@@
     vocc_plot <- plot_vocc(biomass,
-      theme = "black",
+      theme_black = TRUE,
        coast = TRUE,
@@
        vocc_plot <- plot_vocc(biomass,
-         theme = "black",
+         theme_black = TRUE,
          coast = TRUE,

### 2019-11-28 analysis/VOCC/3-climate-models-fixed-depth.Rmd
model for bioclimatic estimates without fixed spatial field, use AR1, and fixed response to depth

--- 2019-10-09 analysis/VOCC/3-climate-models-fixed-both.Rmd
@@
 params: 
-    species: "Pacific Ocean Perch"
+    species: "Redbanded Rockfish"
     region: "All synoptic surveys"
@@
     knots: 500
-    covariates: "+trawled" # additional non-climate variables
-    covs: "-log-both-trawled" # string describing model covariates
+    covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
+    covs: "-AR1-spatial" #
+    #covs: "-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
 ---
@@
 if (model_w_do) {
-
   data <- filter(data, !(year == 2016 & ssid == 4) ) # 2016 do an outlier for survey 4
@@
  }
-
 } else {
-
   data <- scale_predictors(data, predictors = c(quo(depth), quo(temp))) 
@@
 }
+
+# d <- data %>% filter(year==2011) #%>% filter(depth> 400) 
+# = plot((adult_density)~(depth), data=d)
+
 ```
@@
     adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""
+      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
       ))
@@
       adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", covariates, ""
+      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
       ))
@@
 
-    adult_biomass_d <- sdmTMB::sdmTMB(data,
+    adult_biomass_d <- sdmTMB(data = data,
       adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
+      #time_varying = ~ trawled, #0 + depth_scaled + depth_scaled2, 
       time = "year", spde = spde,
@@
       ar1_fields = TRUE, 
-      include_spatial = TRUE,
+      include_spatial = FALSE, # default = TRUE
+      reml = TRUE,
       enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
@@
     imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""
+      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
       ))
@@
          imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2", covariates, ""
+      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
       ))
@@
 
-    imm_biomass_d <- sdmTMB::sdmTMB(data, 
+    imm_biomass_d <- sdmTMB::sdmTMB(data = data, 
       imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
+      #time_varying = ~ trawled, # 0 + depth_scaled + depth_scaled2, 
       time = "year", spde = spde,
       family = tweedie(link = "log"),
-      ar1_fields = FALSE,
-      include_spatial = TRUE,
+      ar1_fields = TRUE,
+      include_spatial = FALSE, # default = TRUE
+      reml = TRUE,
       enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
@@
 
-    total_biomass_d <- sdmTMB::sdmTMB(data,
+    total_biomass_d <- sdmTMB::sdmTMB(data = data,
       dens_formula,
@@
       family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
+      #ar1_fields = TRUE,
+      include_spatial = FALSE, # default = TRUE
       enable_priors = priors,

### 2019-11-28 analysis/VOCC/3-bioclimatic-trends.Rmd
update bioclimatic calculation

@@
     bind_predictions: FALSE
-    covs: "-log-both-noAR1" # covariate string used to name model rds 
+    covs: "-AR1-spatial" 
+    #covs: "-log-both-fixed-depth"
+    #covs: "-log-both-noAR1" # covariate string used to name model rds 
 ---
@@
   predictions$est_exp <- exp(predictions$est) * 10000
-  predictions$bioclimatic <- exp(predictions$est - predictions$epsilon_st) * 10000
+
+  predictions <- predictions %>% group_by(X,Y) %>% arrange(year) %>% mutate(lag_epsilon = lag(epsilon_st), bioclimatic = est - epsilon_st + lag_epsilon)
 

### 2019-12-04 analysis/VOCC/3-bioclimatic-trends.Rmd
calculate bioclimatic estimates from series of separate models and prediction work-arounds

--- 2019-11-08 analysis/VOCC/3-bioclimatic-trends.Rmd
@@
 params: 
+    #species: "Pacific Cod" # 
     species: "Redbanded Rockfish"
@@
     bind_predictions: FALSE
-    covs: "-log-both-noAR1" # covariate string used to name model rds 
+    covs: "-all" 
+    # year_cutoff: 2012
+    #covs: "-log-both-fixed-depth"
+    #covs: "-log-both-noAR1" # covariate string used to name model rds 
 ---
@@
 
+paste("species =", species)
 paste("region =", region)
@@
   nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
+
   adult_biomass <- readRDS(paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+        "/mod-mat-biomass-", spp,"-fixed", covs, "--", ssid_string, "-prior-", priors, ".rds"))
   #nd$year <- as.integer(nd$year)
+
+  sort(unique(nd_all$year))
   nd <- nd_all %>%
@@
       do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
-      do_mlpl_scaled2 = do_mlpl_scaled^2)
+      do_mlpl_scaled2 = do_mlpl_scaled^2) %>% 
+    select(-est, -est_non_rf, - est_rf, -omega_s, -epsilon_st, -zeta_s)
   nd <- na.omit(nd)
+  sort(unique(nd$year))
 
@@
 
-  adult_predictions$est_rw_i <- adult_predictions_obj$obj$report()$proj_rw_i
-  # test_eta <- adult_predictions_obj$obj$report()$proj_eta
-  # test_fe <- adult_predictions_obj$obj$report()$proj_fe
-  # 
-  # test_diff <- test_eta - test_fe
-  # range(adult_predictions$est_rw_i)
-  # report <- adult_predictions_obj$obj$report()
+  nd_pre_2013 <- nd %>% filter(year < 2013)
+  ad_model_2013 <- readRDS(paste0("data/", spp, 
+        "/mod-mat-biomass-", spp,"-fixed-past-2013-", ssid_string, "-prior-", priors, ".rds"))
+  adult_pre_2013 <- predict(ad_model_2013, newdata = nd_pre_2013, return_tmb_object = TRUE)
+
+  fake_2013 <- nd %>% 
+    mutate(year = year - 2) %>% 
+    filter(year > 2007) %>% filter(year < 2013 )
+  sort(unique(fake_2013$year))
+  sort(unique(nd_pre_2013$year))
+  adult_2013 <- predict(ad_model_2013, newdata = fake_2013, return_tmb_object = TRUE)
+  adult_2013$data$year <- adult_2013$data$year + 2
+
+  nd_pre_2015 <- nd %>% filter(year < 2015)
+  ad_model_2015 <- readRDS(paste0("data/", spp, 
+        "/mod-mat-biomass-", spp,"-fixed-past-2015-", ssid_string, "-prior-", priors, ".rds"))
+  adult_pre_2015 <- predict(ad_model_2015, newdata = nd_pre_2015, return_tmb_object = TRUE)
+
+  fake_2015 <- nd %>% 
+    mutate(year = year - 2) %>% 
+    filter(year > 2007) %>% filter(year < 2015)
+  sort(unique(nd_pre_2015$year))
+  sort(unique(fake_2015$year))
+  adult_2015 <- predict(ad_model_2015, newdata = fake_2015, return_tmb_object = TRUE)
+  adult_2015$data$year <- adult_2015$data$year + 2
+
+  nd_pre_2017 <- nd %>% filter(year < 2017)
+  ad_model_2017 <- readRDS(paste0("data/", spp, 
+        "/mod-mat-biomass-", spp,"-fixed-past-2017-", ssid_string, "-prior-", priors, ".rds"))
+  adult_pre_2017 <- predict(ad_model_2017, newdata = nd_pre_2017, return_tmb_object = TRUE)
+
+  fake_2017 <- nd %>% 
+    mutate(year = year - 2) %>% 
+    filter(year > 2007) #%>% filter(year < 2013 )
+  sort(unique(fake_2017$year))
+  sort(unique(nd_pre_2017$year))
+
+  adult_2017 <- predict(ad_model_2017, newdata = fake_2017, return_tmb_object = TRUE)
+  adult_2017$data$year <- adult_2017$data$year + 2
+
+  fake_2017_minus4 <- nd %>% 
+    mutate(year = year - 4) %>% 
+    filter(year > 2007)
+  sort(unique(fake_2017_minus4$year))
+  adult_2017_multi <- predict(ad_model_past, newdata = fake_2017_minus4, return_tmb_object = TRUE)
+  adult_2017_multi$data$year <- adult_2017_multi$data$year + 4
+  sort(unique(adult_2017_multi$data$year))
+
+ past_predictions <- adult_pre_2015$data %>% 
+   mutate(past_est = est, blob_bioclim = est) %>% 
+   select(X, Y, year, past_est, blob_bioclim)
+
+   bioclim_2013_start <- filter(adult_pre_2013$data, year >= 2011) %>% filter(year <= 2012) %>%
+    mutate(bioclim_2013 = est)
+  bioclim_2013_end <- filter(adult_2013$data, year >= 2013) %>% filter(year <= 2014) %>%
+    mutate(bioclim_2013 = est_non_rf + omega_s)
+  bioclim_2013 <- dplyr::bind_rows(bioclim_2013_start, bioclim_2013_end)
+
+  bioclim_2015_start <- filter(adult_past$data, year >= 2013) %>% filter(year <= 2014) %>%
+    mutate(bioclim_2015 = est) %>% select(X,Y, year, bioclim_2015)
+  bioclim_2013 <- left_join(bioclim_2013, bioclim_2015_start) %>% 
+    select(X,Y, year, bioclim_2013, bioclim_2015)
+  past_predictions <- left_join(past_predictions, bioclim_2013)
+
+  bioclim_2015 <- filter(adult_2015$data, year >= 2015) %>% filter(year <= 2016) %>%
+    mutate(blob_bioclim = est_non_rf + omega_s, bioclim_2015 = est_non_rf + omega_s)
+
+  bioclim_2017_start <- filter(adult_2017$data, year >= 2015) %>% filter(year <= 2016) %>%
+    mutate(bioclim_2017 = est) %>% select(X,Y, year, bioclim_2017)
+
+  bioclim_2015 <- left_join(bioclim_2015, bioclim_2017_start)
+
+  bioclim_2017_multi <- filter(adult_2017_multi$data, year >= 2017) %>% filter(year <= 2018) %>%
+    mutate(blob_bioclim = est_non_rf + omega_s) %>% 
+    select(X,Y, year, blob_bioclim)
+
+  bioclim_2017 <- filter(adult_2017$data, year >= 2017) %>% filter(year <= 2018) %>%
+    mutate(bioclim_2017 = est_non_rf + omega_s)
+  bioclim_2017 <- left_join(bioclim_2017, bioclim_2017_multi)
+
+  adult_predictions <- adult_predictions_obj$data
+  adult_bioclim <- dplyr::bind_rows(past_predictions, bioclim_2015, bioclim_2017) %>% 
+    select(X, Y, year, past_est, blob_bioclim, bioclim_2013, bioclim_2015, bioclim_2017)
+
+  adult_predictions <- left_join(adult_predictions, adult_bioclim)
 
   saveRDS(adult_predictions, file = paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"))
+    "/bioclimatic-predictions-", spp, covs,"-", ssid_string, "-mat-2015-cutoff.rds"))
+
+  #################
+  #### Immature 
+  #################
 
   imm_biomass <- readRDS(paste0("data/", spp, 
-        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-imm-biomass-", spp,"-fixed", covs, "--", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   nd <- na.omit(nd)
-  #nd$year <- as.numeric(nd$year)
 
@@
 
-  imm_predictions$est_rw_i <- imm_predictions_obj$obj$report()$proj_rw_i
+  nd_pre_2013 <- nd %>% filter(year < 2013)
+  im_model_2013 <- readRDS(paste0("data/", spp, 
+        "/mod-imm-biomass-", spp,"-fixed-past-2013-", ssid_string, "-prior-", priors, ".rds"))
+  imm_pre_2013 <- predict(im_model_2013, newdata = nd_pre_2013, return_tmb_object = TRUE)
+
+  fake_2013 <- nd %>% 
+    mutate(year = year - 2) %>% 
+    filter(year > 2007) %>% filter(year < 2013 )
+  sort(unique(fake_2013$year))
+  sort(unique(nd_pre_2013$year))
+  imm_2013 <- predict(im_model_2013, newdata = fake_2013, return_tmb_object = TRUE)
+  imm_2013$data$year <- imm_2013$data$year + 2
+
+  nd_pre_2015 <- nd %>% filter(year < 2015)
+  im_model_2015 <- readRDS(paste0("data/", spp, 
+        "/mod-imm-biomass-", spp,"-fixed-past-2015-", ssid_string, "-prior-", priors, ".rds"))
+  imm_pre_2015 <- predict(im_model_2015, newdata = nd_pre_2015, return_tmb_object = TRUE)
+
+  fake_2015 <- nd %>% 
+    mutate(year = year - 2) %>% 
+    filter(year > 2007) %>% filter(year < 2015)
+  sort(unique(nd_pre_2015$year))
+  sort(unique(fake_2015$year))
+  imm_2015 <- predict(im_model_2015, newdata = fake_2015, return_tmb_object = TRUE)
+  imm_2015$data$year <- imm_2015$data$year + 2
+
+  nd_pre_2017 <- nd %>% filter(year < 2017)
+  im_model_2017 <- readRDS(paste0("data/", spp, 
+        "/mod-imm-biomass-", spp,"-fixed-past-2017-", ssid_string, "-prior-", priors, ".rds"))
+  imm_pre_2017 <- predict(im_model_2017, newdata = nd_pre_2017, return_tmb_object = TRUE)
+
+  fake_2017 <- nd %>% 
+    mutate(year = year - 2) %>% 
+    filter(year > 2007) #%>% filter(year < 2013 )
+  sort(unique(fake_2017$year))
+  sort(unique(nd_pre_2017$year))
+
+  imm_2017 <- predict(im_model_2017, newdata = fake_2017, return_tmb_object = TRUE)
+  imm_2017$data$year <- imm_2017$data$year + 2
+
+  fake_2017_minus4 <- nd %>% 
+    mutate(year = year - 4) %>% 
+    filter(year > 2007)
+  sort(unique(fake_2017_minus4$year))
+  imm_2017_multi <- predict(im_model_past, newdata = fake_2017_minus4, return_tmb_object = TRUE)
+  imm_2017_multi$data$year <- imm_2017_multi$data$year + 4
+  sort(unique(imm_2017_multi$data$year))
+
+ past_predictions <- imm_pre_2015$data %>% 
+   mutate(past_est = est, blob_bioclim = est) %>% 
+   select(X, Y, year, past_est, blob_bioclim)
+
+  bioclim_2013_start <- filter(imm_pre_2013$data, year >= 2011) %>% filter(year <= 2012) %>%
+    mutate(bioclim_2013 = est)
+  bioclim_2013_end <- filter(imm_2013$data, year >= 2013) %>% filter(year <= 2014) %>%
+    mutate(bioclim_2013 = est_non_rf + omega_s)
+  bioclim_2013 <- dplyr::bind_rows(bioclim_2013_start, bioclim_2013_end)
+
+  bioclim_2015_start <- filter(imm_past$data, year >= 2013) %>% filter(year <= 2014) %>%
+    mutate(bioclim_2015 = est) %>% select(X,Y, year, bioclim_2015)
+  bioclim_2013 <- left_join(bioclim_2013, bioclim_2015_start) %>% 
+    select(X,Y, year, bioclim_2013, bioclim_2015)
+  past_predictions <- left_join(past_predictions, bioclim_2013)
+
+  bioclim_2015 <- filter(imm_2015$data, year >= 2015) %>% filter(year <= 2016) %>%
+    mutate(blob_bioclim = est_non_rf + omega_s, bioclim_2015 = est_non_rf + omega_s) 
+  bioclim_2017_start <- filter(imm_2017$data, year >= 2015) %>% filter(year <= 2016) %>%
+    mutate(bioclim_2017 = est) %>% select(X,Y, year, bioclim_2017)
+
+  bioclim_2015 <- left_join(bioclim_2015, bioclim_2017_start)
+
+  bioclim_2017_multi <- filter(imm_2017_multi$data, year >= 2017) %>% filter(year <= 2018) %>%
+    mutate(blob_bioclim = est_non_rf + omega_s) %>% 
+    select(X,Y, year, blob_bioclim)
+
+  bioclim_2017 <- filter(imm_2017$data, year >= 2017) %>% filter(year <= 2018) %>%
+    mutate(bioclim_2017 = est_non_rf + omega_s)
+  bioclim_2017 <- left_join(bioclim_2017, bioclim_2017_multi)
+
+   imm_predictions <- imm_predictions_obj$data
+   imm_bioclim <- dplyr::bind_rows(past_predictions, bioclim_2015, bioclim_2017) %>% 
+    select(X, Y, year, past_est, blob_bioclim, bioclim_2013, bioclim_2015, bioclim_2017)
+
+  imm_predictions <- left_join(imm_predictions, imm_bioclim)
 
   saveRDS(imm_predictions, file = paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"))
-
-} else {
-
-    total_biomass <- readRDS(paste0("data/", spp, 
-      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
-
-    nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
-    nd <- nd_all %>% 
-      filter(ssid %in% model_ssid) %>%
-      filter(year %in% unique(total_biomass$data$year))%>% 
-    mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    predictions <- predict(total_biomass, newdata = nd)
-    saveRDS(predictions, file = paste0("data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
+    "/bioclimatic-predictions-", spp, covs,"-", ssid_string, "-imm-2015-cutoff.rds"))
+
+} else { # need to add non-split-by-age version
+    # total_biomass <- readRDS(paste0("data/", spp, 
+    #   "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+    # 
+    # nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
+    # nd <- nd_all %>% 
+    #   filter(ssid %in% model_ssid) %>%
+    #   filter(year %in% unique(total_biomass$data$year))%>% 
+    # mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
+    # nd <- na.omit(nd)
+    # nd$year <- as.integer(nd$year)
+    # 
+    # predictions <- predict(total_biomass, newdata = nd)
+    # saveRDS(predictions, file = paste0("data/", spp,
+    #   "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
 }
@@
 if (region != "All synoptic surveys") {
-  try(adult_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-1n3n4n16-mature-bioclimatic-prior-", priors, ".rds"
+  try(adult_predictions <- readRDS( paste0("data/", spp,
+    "/bioclimatic-predictions-", spp, covs,"-", ssid_string, "-mat-2015-cutoff.rds"
   )) %>% filter(ssid %in% model_ssid))
   try(imm_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-1n3n4n16-imm-bioclimatic-prior-", priors, ".rds"
+    "/bioclimatic-predictions-", spp, covs,"-", ssid_string, "-imm-2015-cutoff.rds"
   )) %>% filter(ssid %in% model_ssid))
@@
 
-  try({adult_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"
-    ))
-  })
+ try({adult_predictions <- readRDS( paste0("data/", spp,
+    "/bioclimatic-predictions-", spp, covs,"-", ssid_string, "-mat-2015-cutoff.rds"
+  )) %>% filter(ssid %in% model_ssid)})
  try({imm_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"
-    ))
-  })
-
-  # OR bind together predictions from different models
-  if (bind_predictions) {
-    try({pred1n3 <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-1n3-mature-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-    try({pred4 <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-4-mature-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-    try({pred16 <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-16-mature-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-
-    if (!exists("pred1n3")) pred1n3 <- NULL
-    if (!exists("pred16")) pred16 <- NULL
-    if (!exists("pred4")) pred4 <- NULL
-    adult_predictions <- rbind(pred1n3, pred4, pred16)
-
-    try({pred1n3_imm <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-1n3-imm-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-    try({pred4_imm <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-4-imm-bioclimatic-prior-", priors, ".rds"))
-    })
-    try({pred16_imm <- readRDS(paste0("data/", spp,
-        "/predictions-", spp, covs, "-16-imm-bioclimatic-prior-", priors, ".rds"))
-    })
-
-    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
-    if (!exists("pred16_imm")) pred16_imm <- NULL
-    if (!exists("pred4_imm")) pred4_imm <- NULL
-    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)
-  }
-  }
+    "/bioclimatic-predictions-", spp, covs,"-", ssid_string, "-imm-2015-cutoff.rds"
+  )) %>% filter(ssid %in% model_ssid)})
 
@@
         "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
-      })
-
-      if (!exists("predictions")) {
-        try({pred1n3_all <- readRDS(paste0("data/", spp,
-          "/predictions-", spp, covs, "-1n3-total-bioclimatic-prior-", priors, ".rds"))
-        })
-        try({pred4_all <- readRDS(paste0("data/", spp,
-          "/predictions-", spp, covs, "-4-total-bioclimatic-prior-", priors, ".rds"))
-        })
-        try({pred16_all <- readRDS(paste0("data/", spp,
-          "/predictions-", spp, covs, "-16-total-bioclimatic-prior-", priors, ".rds"))
-        })
-        try({
-          predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
       })
@@
   imm_predictions$est_exp <- exp(imm_predictions$est) * 10000
-  adult_predictions$bioclimatic <- exp(adult_predictions$est - adult_predictions$epsilon_st) * 10000
-  imm_predictions$bioclimatic <- exp(imm_predictions$est - imm_predictions$epsilon_st) * 10000
-
   adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
+
+  # check that both dataframes are arranged similarly
+  if (adult_predictions[30000,6] == imm_predictions[30000,6]) {
   imm_predictions$prop_imm <- 
   imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
+  }
 
@@
   predictions$est_exp <- exp(predictions$est) * 10000
-  predictions$bioclimatic <- exp(predictions$est - predictions$epsilon_st) * 10000
-
+
+  # predictions <- predictions %>% group_by(X,Y) %>% arrange(year) %>% mutate(lag_epsilon = lag(epsilon_st), bioclimatic = est - epsilon_st + lag_epsilon)
+  # 
   predictions$total_bio <- exp(predictions$est) * 10000
@@
 png(
-  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-fixed-everything-noAR1.png"),
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/bioclimatic-biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/bioclimatic-biomass-", spp, covs, "-", ssid_string, "-fixed-everything-noAR1.png"),
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/bioclimatic-exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/bioclimatic-exploratory-", spp, covs, "-", ssid_string, "-fixed-everything-noAR1.png"),
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/bioclimatic-residuals-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/bioclimatic-residuals-", spp, covs, "-", ssid_string, "-fixed-everything-noAR1.png"),
   res = 600,

### 2019-12-04 analysis/VOCC/4-vocc-bioclimatic.R

@@ -6,12 +6,18 @@ library("dplyr")
 library("ggplot2")
 
 species <- "Redbanded Rockfish"
+# species <- "Pacific Ocean Perch"
+# species <- "Pacific Cod"
 spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
 
 # covs <- "-log-both-fixed-depth"
 # covs <- "-log-both-noAR1"
-covs <- "-AR1-spatial"
-all_years <- TRUE
+covs <- "-all"
+#covs <- "-tv-depth-all"
+average <- TRUE
+
+average <- FALSE
+start_year <- 2011
 
 #########################
 ### CHOOSE SPATIAL EXTENT
@@ -28,16 +34,13 @@ scale_fac <- 5 # means that the raster is reprojected to _ X original grid (2 km
 ### LOAD DATA
 #################
 
-d_all <- readRDS(paste0("data/", spp, "/all-predictions-", 
-  spp, covs, "-mat-bioclimatic-prior-FALSE.rds")) %>% 
+d_all <- readRDS(paste0("data/", spp, "/bioclimatic-predictions-", 
+  spp, covs, "-1n3n4n16-mat-2015-cutoff.rds")) %>% 
   filter(ssid %in% model_ssid)
-
-# d_all <- readRDS(paste0("data/", spp, "/all-predictions-", 
-#   spp, covs, "-imm-bioclimatic-prior-FALSE.rds")) %>% 
-#   filter(ssid %in% model_ssid)
+glimpse(d_all)
 
 # # # for climate independent biotic values...
-# covs <- "-trawled-ssid"
+# covs <- "-trawled-ssid-reml"
 # d_all <- readRDS(paste0("data/", spp, "/all-predictions-",
 #   spp, covs, "-mat-biomass-prior-FALSE.rds")) %>%
 #   filter(ssid %in% model_ssid)
@@ -46,13 +49,12 @@ d_all <- readRDS(paste0("data/", spp, "/all-predictions-",
 ### SET TIME FRAME
 #########################
 
-if (all_years){
+if (average){
  #unique(d_all$year)
  d <- d_all
  indices <- c(1, 1, 1, 2, 2)
  year_range <- 5
 } else{ 
-start_year <- 2011
 d <- d_all %>% filter(year > start_year-1) %>% filter(year < start_year+3)
 unique(d$year)
 indices <- c(1,2)
@@ -223,39 +225,43 @@ grad1
[…]

### 2019-12-05 analysis/VOCC/3-bioclimatic-trends.Rmd
start drafting addition of 2011 cutoff and tv depth option

@@
     bind_predictions: FALSE
-    covs: "-all" 
+    covs: "-fixed"  
     # year_cutoff: 2012
@@
   adult_biomass <- readRDS(paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed", covs, "--", ssid_string, "-prior-", priors, ".rds"))
+        "/mod-mat-biomass-", spp, covs, "-all--", ssid_string, "-prior-", priors, ".rds"))
   #nd$year <- as.integer(nd$year)
@@
 
+  # nd_pre_2011 <- nd %>% filter(year < 2011)
+  # ad_model_2011 <- readRDS(paste0("data/", spp,
+  #       "/mod-mat-biomass-", spp, covs,"-past-2011-", ssid_string, "-prior-", priors, ".rds"))
+  # sort(unique(ad_model_2011$data$year))
+  # adult_pre_2011 <- predict(ad_model_2011, newdata = nd_pre_2011, return_tmb_object = TRUE)
+  # 
+  # fake_2011 <- nd %>%
+  #   mutate(year = year - 2) %>%
+  #   filter(year > 2007) %>% filter(year < 2011 )
+  # sort(unique(fake_2011$year))
+  # sort(unique(nd_pre_2011$year))
+  # adult_2011 <- predict(ad_model_2011, newdata = fake_2011, return_tmb_object = TRUE)
+  # adult_2011$data$year <- adult_2011$data$year + 2
+
   nd_pre_2013 <- nd %>% filter(year < 2013)
   ad_model_2013 <- readRDS(paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed-past-2013-", ssid_string, "-prior-", priors, ".rds"))
+        "/mod-mat-biomass-", spp, covs,"-past-2013-", ssid_string, "-prior-", priors, ".rds"))
   adult_pre_2013 <- predict(ad_model_2013, newdata = nd_pre_2013, return_tmb_object = TRUE)
@@
   ad_model_2015 <- readRDS(paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed-past-2015-", ssid_string, "-prior-", priors, ".rds"))
+        "/mod-mat-biomass-", spp, covs,"-past-2015-", ssid_string, "-prior-", priors, ".rds"))
   adult_pre_2015 <- predict(ad_model_2015, newdata = nd_pre_2015, return_tmb_object = TRUE)
@@
   ad_model_2017 <- readRDS(paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed-past-2017-", ssid_string, "-prior-", priors, ".rds"))
+        "/mod-mat-biomass-", spp, covs,"-past-2017-", ssid_string, "-prior-", priors, ".rds"))
   adult_pre_2017 <- predict(ad_model_2017, newdata = nd_pre_2017, return_tmb_object = TRUE)
@@
   sort(unique(fake_2017_minus4$year))
-  adult_2017_multi <- predict(ad_model_past, newdata = fake_2017_minus4, return_tmb_object = TRUE)
+  sort(unique(nd_pre_2015$year))
+  adult_2017_multi <- predict(ad_model_2015, newdata = fake_2017_minus4, return_tmb_object = TRUE)
   adult_2017_multi$data$year <- adult_2017_multi$data$year + 4
@@
 
+  # bioclim_2011_start <- filter(adult_pre_2011$data, year >= 2009) %>% filter(year <= 2010) %>%
+  #   mutate(bioclim_2011 = est)
+  # bioclim_2011_end <- filter(adult_2013$data, year >= 2011) %>% filter(year <= 2012) %>%
+  #   mutate(bioclim_2011 = est_non_rf + omega_s)
+  # bioclim_2011_all <- dplyr::bind_rows(bioclim_2011_start, bioclim_2011_end)
+
   bioclim_2013_start <- filter(adult_pre_2013$data, year >= 2011) %>% filter(year <= 2012) %>%
-    mutate(bioclim_2013 = est)
+    mutate(bioclim_2013 = est) #%>% select(X,Y, year, bioclim_2013)
+  # bioclim_2013_all <- left_join(bioclim_2011, bioclim_2013_start) %>% 
+  #   select(X,Y, year, bioclim_2011, bioclim_2013)
+  #past_predictions <- left_join(past_predictions, bioclim_2011_all)
+
   bioclim_2013_end <- filter(adult_2013$data, year >= 2013) %>% filter(year <= 2014) %>%
     mutate(bioclim_2013 = est_non_rf + omega_s)
-  bioclim_2013 <- dplyr::bind_rows(bioclim_2013_start, bioclim_2013_end)
-
-  bioclim_2015_start <- filter(adult_past$data, year >= 2013) %>% filter(year <= 2014) %>%
-    mutate(bioclim_2015 = est) %>% select(X,Y, year, bioclim_2015)
-  bioclim_2013 <- left_join(bioclim_2013, bioclim_2015_start) %>% 
+
+  bioclim_2013_all <- dplyr::bind_rows(bioclim_2013_start, bioclim_2013_end)
+
+  bioclim_2015_start <- filter(adult_2015$data, year >= 2013) %>% filter(year <= 2014) %>%
+    mutate(bioclim_2015 = est) %>% 
+    select(X,Y, year, bioclim_2015)
+
+  bioclim_2013_all <- left_join(bioclim_2013_all, bioclim_2015_start) %>% 
     select(X,Y, year, bioclim_2013, bioclim_2015)
-  past_predictions <- left_join(past_predictions, bioclim_2013)
-
-  bioclim_2015 <- filter(adult_2015$data, year >= 2015) %>% filter(year <= 2016) %>%
+  past_predictions <- left_join(past_predictions, bioclim_2013_all)
+
+  bioclim_2015_end <- filter(adult_2015$data, year >= 2015) %>% filter(year <= 2016) %>%
     mutate(blob_bioclim = est_non_rf + omega_s, bioclim_2015 = est_non_rf + omega_s)
@@
 
-  bioclim_2015 <- left_join(bioclim_2015, bioclim_2017_start)
+  bioclim_2015 <- left_join(bioclim_2015_end, bioclim_2017_start)
 

### 2019-12-06 analysis/VOCC/3-bioclimatic-trends.Rmd
functionize bioclimatic calculations

--- 2019-11-28 analysis/VOCC/3-bioclimatic-trends.Rmd
@@
     bind_predictions: FALSE
-    covs: "-AR1-spatial" 
+    covs: "-fixed"  
+    # year_cutoff: 2012
     #covs: "-log-both-fixed-depth"
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set( fig.width = 9, fig.height = 6.5, #fig.width=11, fig.height=8.5, 
-                      #fig.path=paste0("figs/", spp, "/"),
-                      echo=FALSE, warning=FALSE, message=FALSE)
+knitr::opts_chunk$set(
+  fig.width = 9, fig.height = 6.5, # fig.width=11, fig.height=8.5,
+  # fig.path=paste0("figs/", spp, "/"),
+  echo = FALSE, warning = FALSE, message = FALSE
+)
 library(dplyr)
@@
 ```{r}
-#fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+# fish <- readRDS(paste0("raw/bio-data-", spp, ""))
 
@@
 if (any(names(maturity) == "adult_density")) {
-#if (length(unique(fish$maturity_code)) > 2) {
+  # if (length(unique(fish$maturity_code)) > 2) {
   nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
-  adult_biomass <- readRDS(paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
-  #nd$year <- as.integer(nd$year)
+
+  adult_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-all--", ssid_string, "-prior-", priors, ".rds"
+  ))
+  # nd$year <- as.integer(nd$year)
+
+  sort(unique(nd_all$year))
   nd <- nd_all %>%
@@
     mutate(
-      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
-      do_mlpl_scaled2 = do_mlpl_scaled^2)
+      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) / (adult_biomass$data$do_mlpl_sd[1] * 2),
+      do_mlpl_scaled2 = do_mlpl_scaled^2
+    ) %>%
+    select(-est, -est_non_rf, -est_rf, -omega_s, -epsilon_st, -zeta_s)
   nd <- na.omit(nd)
+  sort(unique(nd$year))
 
@@
 
-  adult_predictions$est_rw_i <- adult_predictions_obj$obj$report()$proj_rw_i
-  # test_eta <- adult_predictions_obj$obj$report()$proj_eta
-  # test_fe <- adult_predictions_obj$obj$report()$proj_fe
-  # 
-  # test_diff <- test_eta - test_fe
-  # range(adult_predictions$est_rw_i)
-  # report <- adult_predictions_obj$obj$report()
-
-  saveRDS(adult_predictions, file = paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"))
-
-  imm_biomass <- readRDS(paste0("data/", spp, 
-        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
+  # write as function
+  bioclim_by_year <- function(yr, model_for_prior_year, pred_grid) {
+    nd <- pred_grid
+
+    nd_trim <- nd %>% filter(year < yr)
+    fake_years <- nd %>%
+      mutate(year = year - 2) %>%
+      filter(year > 2007) %>%
+      filter(year < yr)
+
+    fake <- sort(unique(fake_years$year))
+    true <- sort(unique(nd$year))
+
+    if (!identical(fake, true)) "Something is wrong with the pred_grid years."
+
+    prior_years <- predict(model_for_prior_year, newdata = nd_trim, return_tmb_object = FALSE)
+    fake_est <- predict(model_for_prior_year, newdata = fake_years, return_tmb_object = FALSE)
+    fake_est$year <- fake_est$year + 2
+
+    bioclim_start <- filter(prior_years, year >= yr - 2) %>%
+      filter(year <= yr - 1) %>%
+      mutate(bioclim = est)
+    bioclim_end <- filter(fake_est, year >= yr) %>%
+      filter(year <= yr + 1) %>%
+      mutate(bioclim = est_non_rf + omega_s)
+    bioclim_all <- dplyr::bind_rows(bioclim_start, bioclim_end) %>%
+      select(X, Y, year, bioclim)
+
+    bioclim_all
+  }
+
+  model_pre_2013 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2013-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(model_pre_2013$data$year))
+
+  bioclim_2013 <- bioclim_by_year(2013, model_pre_2013, nd)
+  bioclim_2013 <- rename(bioclim_2013, bioclim_2013 = bioclim)
+  glimpse(bioclim_2013)
+
+  model_pre_2015 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(model_pre_2015$data$year))
+
+  bioclim_2015 <- bioclim_by_year(2015, model_pre_2015, nd)
+  bioclim_2015 <- rename(bioclim_2015, bioclim_2015 = bioclim)
+  glimpse(bioclim_2015)
+
+  model_pre_2017 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2017-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(model_pre_2017$data$year))
+
+  bioclim_2017 <- bioclim_by_year(2017, model_pre_2017, nd)
+  bioclim_2017 <- rename(bioclim_2017, bioclim_2017 = bioclim)
+  glimpse(bioclim_2017)
+
+  # for multiyear bioclim
+  bioclim_multi_year <- function(yr, model_for_prior_year, pred_grid) {
+    nd <- pred_grid
+    period2_range <- max(nd$year) - yr + 1
+
+    nd_trim <- nd %>% filter(year < yr)
+    fake_years <- nd %>%
+      mutate(year = year - 2) %>%
+      filter(year > 2007) %>%
+      filter(year < yr)
+
+    fake <- sort(unique(fake_years$year))
+    true <- sort(unique(model_for_prior_year$data$year))
+
+    if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years."
+
+    prior_years <- predict(model_for_prior_year, newdata = nd_trim, return_tmb_object = FALSE)
+
+    past_predictions <- prior_years %>%
+      mutate(past_est = est, bioclim = est) %>%
+      select(X, Y, year, past_est, bioclim)
+
+    fake_est <- predict(model_for_prior_year, newdata = fake_years, return_tmb_object = FALSE)
+    fake_est$year <- fake_est$year + 2
+
+    bioclim <- filter(fake_est, year >= yr + 2 - 2) %>%
+      filter(year <= yr + 2 - 1) %>%
+      mutate(bioclim = est_non_rf + omega_s)
+
+    if (period2_range >= 4) {
+      fake_minus4 <- nd %>%
+        mutate(year = year - 4) %>%
+        filter(year > 2007)
+
+      fake <- sort(unique(fake_minus4$year))
+      true <- sort(unique(nd_trim$year))
+
+      if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years -4."
+
+      fake_est_minus4 <- predict(model_for_prior_year, newdata = fake_minus4, return_tmb_object = FALSE)
+      fake_est_minus4$year <- fake_est_minus4$year + 4
+      # sort(unique(fake_est_minus4$year))
+
+      bioclim_4 <- filter(fake_est_minus4, year >= yr + 4 - 2) %>%
+        filter(year <= yr + 4 - 1) %>%
+        mutate(bioclim = est_non_rf + omega_s)
+
+      bioclim <- dplyr::bind_rows(bioclim, bioclim_4)
+    }
+
+    if (period2_range >= 6) {
+      fake_minus6 <- nd %>%
+        mutate(year = year - 6) %>%
+        filter(year > 2007)
+      fake <- sort(unique(fake_minus6$year))
+      true <- sort(unique(nd_trim$year))
+
+      if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years -6."
+
+      fake_est_minus6 <- predict(model_for_prior_year, newdata = fake_minus6, return_tmb_object = FALSE)
+      fake_est_minus6$year <- fake_est_minus6$year + 6
+
+      bioclim_6 <- filter(fake_est_minus6, year >= yr + 6 - 2) %>%
+        filter(year <= yr + 6 - 1) %>%
+        mutate(bioclim = est_non_rf + omega_s)
+
+      bioclim <- dplyr::bind_rows(bioclim, bioclim_6)
+    }
+
+    bioclim_all <- dplyr::bind_rows(past_predictions, bioclim) %>%
+      select(X, Y, year, past_est, bioclim)
+
+    bioclim_all
+  }
+
+  model_pre_2015 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
+  ))
+
+  bioclim_blob <- bioclim_multi_year(2015, model_pre_2015, nd)
+  bioclim_blob <- rename(bioclim_blob, bioclim_blob = bioclim)
+  glimpse(bioclim_blob)
+  #View(bioclim_blob)
+
+  adult_predictions <- left_join(adult_predictions, bioclim_blob)
+  adult_predictions <- left_join(adult_predictions, bioclim_2013)
+  adult_predictions <- left_join(adult_predictions, bioclim_2015)
+  adult_predictions <- left_join(adult_predictions, bioclim_2017)
+
+  View(adult_predictions)
+
+  saveRDS(adult_predictions, file = paste0(
+    "data/", spp,
+    "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-mat-2015-cutoff.rds"
+  ))
+
+  #################
+  #### Immature
+  #################
+
+  imm_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, covs, "-all--", ssid_string, "-prior-", priors, ".rds"
   ))
@@
     filter(ssid %in% model_ssid) %>%
-    filter(year %in% unique(imm_biomass$data$year))%>% 
-    mutate(do_mlpl_scaled = (log_do - imm_biomass$data$do_mlpl_mean[1]) /(imm_biomass$data$do_mlpl_sd[1]*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
+    filter(year %in% unique(imm_biomass$data$year)) %>%
+    mutate(do_mlpl_scaled = (log_do - imm_biomass$data$do_mlpl_mean[1]) / (imm_biomass$data$do_mlpl_sd[1] * 2), do_mlpl_scaled2 = do_mlpl_scaled^2)
   nd <- na.omit(nd)
-  #nd$year <- as.numeric(nd$year)
 
@@
 
-  imm_predictions$est_rw_i <- imm_predictions_obj$obj$report()$proj_rw_i
-
-  saveRDS(imm_predictions, file = paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"))
-
-} else {
-
-    total_biomass <- readRDS(paste0("data/", spp, 
-      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
-
-    nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
-    nd <- nd_all %>% 
-      filter(ssid %in% model_ssid) %>%
-      filter(year %in% unique(total_biomass$data$year))%>% 
-    mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    predictions <- predict(total_biomass, newdata = nd)
-    saveRDS(predictions, file = paste0("data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
+  imodel_pre_2013 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, covs, "-past-2013-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(imodel_pre_2013$data$year))
+
+  bioclim_2013 <- bioclim_by_year(2013, imodel_pre_2013, nd)
+  bioclim_2013 <- rename(bioclim_2013, bioclim_2013 = bioclim)
+  glimpse(bioclim_2013)
+
+  imodel_pre_2015 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(imodel_pre_2015$data$year))
+
+  bioclim_2015 <- bioclim_by_year(2015, imodel_pre_2015, nd)
+  bioclim_2015 <- rename(bioclim_2015, bioclim_2015 = bioclim)
+  glimpse(bioclim_2015)
+
+  imodel_pre_2017 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, covs, "-past-2017-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(imodel_pre_2017$data$year))
+
+  bioclim_2017 <- bioclim_by_year(2017, imodel_pre_2017, nd)
+  bioclim_2017 <- rename(bioclim_2017, bioclim_2017 = bioclim)
+  glimpse(bioclim_2017)
+
+  imodel_pre_2015 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
+  ))
+
+  bioclim_blob <- bioclim_multi_year(2015, imodel_pre_2015, nd)
+  bioclim_blob <- rename(bioclim_blob, bioclim_blob = bioclim)
+  glimpse(bioclim_blob)
+  #View(bioclim_blob)
+
+  imm_predictions <- left_join(imm_predictions, bioclim_blob)
+  imm_predictions <- left_join(imm_predictions, bioclim_2013)
+  imm_predictions <- left_join(imm_predictions, bioclim_2015)
+  imm_predictions <- left_join(imm_predictions, bioclim_2017)
+
+  saveRDS(imm_predictions, file = paste0(
+    "data/", spp,
+    "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-imm-2015-cutoff.rds"
+  ))
+} else { # need to add non-split-by-age version
+  # total_biomass <- readRDS(paste0("data/", spp,
+  #   "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  #
+  # nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
+  # nd <- nd_all %>%
+  #   filter(ssid %in% model_ssid) %>%
+  #   filter(year %in% unique(total_biomass$data$year))%>%
+  # mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
+  # nd <- na.omit(nd)
+  # nd$year <- as.integer(nd$year)
+  #
+  # predictions <- predict(total_biomass, newdata = nd)
+  # saveRDS(predictions, file = paste0("data/", spp,
+  #   "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
 }
@@
 if (region != "All synoptic surveys") {
-  try(adult_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-1n3n4n16-mature-bioclimatic-prior-", priors, ".rds"
+  try(adult_predictions <- readRDS(paste0(
+    "data/", spp,
+    "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-mat-2015-cutoff.rds"
   )) %>% filter(ssid %in% model_ssid))
-  try(imm_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-1n3n4n16-imm-bioclimatic-prior-", priors, ".rds"
+  try(imm_predictions <- readRDS(paste0(
+    "data/", spp,
+    "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-imm-2015-cutoff.rds"
   )) %>% filter(ssid %in% model_ssid))
@@
 
-    try(predictions <- readRDS(paste0("data/", spp,
+    try(predictions <- readRDS(paste0(
+      "data/", spp,
       "/predictions-", spp, covs, "-1n3n4n16-total-bioclimatic-prior-", priors, ".rds"
@@
 } else {
-
-  try({adult_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"
-    ))
+  try({
+    adult_predictions <- readRDS(paste0(
+      "data/", spp,
+      "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-mat-2015-cutoff.rds"
+    )) %>% filter(ssid %in% model_ssid)
   })
-  try({imm_predictions <- readRDS(paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"
-    ))
+  try({
+    imm_predictions <- readRDS(paste0(
+      "data/", spp,
+      "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-imm-2015-cutoff.rds"
+    )) %>% filter(ssid %in% model_ssid)
   })
-
-  # OR bind together predictions from different models
-  if (bind_predictions) {
-    try({pred1n3 <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-1n3-mature-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-    try({pred4 <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-4-mature-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-    try({pred16 <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-16-mature-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-
-    if (!exists("pred1n3")) pred1n3 <- NULL
-    if (!exists("pred16")) pred16 <- NULL
-    if (!exists("pred4")) pred4 <- NULL
-    adult_predictions <- rbind(pred1n3, pred4, pred16)
-
-    try({pred1n3_imm <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-1n3-imm-bioclimatic-prior-", priors, ".rds"
-      ))
-    })
-    try({pred4_imm <- readRDS(paste0("data/", spp,
-      "/predictions-", spp, covs, "-4-imm-bioclimatic-prior-", priors, ".rds"))
-    })
-    try({pred16_imm <- readRDS(paste0("data/", spp,
-        "/predictions-", spp, covs, "-16-imm-bioclimatic-prior-", priors, ".rds"))
-    })
-
-    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
-    if (!exists("pred16_imm")) pred16_imm <- NULL
-    if (!exists("pred4_imm")) pred4_imm <- NULL
-    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)
-  }
-  }
 
@@
 
-      try({predictions <- readRDS(paste0("data/", spp,
-        "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
-      })
-
-      if (!exists("predictions")) {
-        try({pred1n3_all <- readRDS(paste0("data/", spp,
-          "/predictions-", spp, covs, "-1n3-total-bioclimatic-prior-", priors, ".rds"))
-        })
-        try({pred4_all <- readRDS(paste0("data/", spp,
-          "/predictions-", spp, covs, "-4-total-bioclimatic-prior-", priors, ".rds"))
-        })
-        try({pred16_all <- readRDS(paste0("data/", spp,
-          "/predictions-", spp, covs, "-16-total-bioclimatic-prior-", priors, ".rds"))
-        })
     try({
-          predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
+      predictions <- readRDS(paste0(
+        "data/", spp,
+        "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"
+      ))
     })
@@
   imm_predictions$est_exp <- exp(imm_predictions$est) * 10000
-  adult_predictions$bioclimatic <- exp(adult_predictions$est - adult_predictions$epsilon_st) * 10000
-  imm_predictions$bioclimatic <- exp(imm_predictions$est - imm_predictions$epsilon_st) * 10000
-
   adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
+
+  # check that both dataframes are arranged similarly
+  if (adult_predictions[30000, 6] == imm_predictions[30000, 6]) {
     imm_predictions$prop_imm <-
       imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
+  }
 
@@
 
-  predictions <- predictions %>% group_by(X,Y) %>% arrange(year) %>% mutate(lag_epsilon = lag(epsilon_st), bioclimatic = est - epsilon_st + lag_epsilon)
-
+  # predictions <- predictions %>% group_by(X,Y) %>% arrange(year) %>% mutate(lag_epsilon = lag(epsilon_st), bioclimatic = est - epsilon_st + lag_epsilon)
+  #
   predictions$total_bio <- exp(predictions$est) * 10000
   max_raster <- quantile(predictions$total_bio, .999)
-  #max_bio <- round(max(predictions$total_bio))
+  # max_bio <- round(max(predictions$total_bio))
   max_bio <- signif(max(predictions$est_exp), digits = 2)
@@
 p_adult <- list()
-  if(exists("adult_predictions")) {
+if (exists("adult_predictions")) {
   for (i in seq_len(length(model_ssid))) {
@@
     if (1 %in% model_ssid) {
-
       .predictions <- filter(predictions, ssid == 1)
-      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
-      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
-      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
-      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 5) indices <- c(1, 1, 1, 2, 2)
+      if (length(unique(.predictions$year)) == 6) indices <- c(1, 1, 1, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 
@@
       .predictions <- filter(predictions, ssid == 3)
-      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
-      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
-      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
-      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 5) indices <- c(1, 1, 1, 2, 2)
+      if (length(unique(.predictions$year)) == 6) indices <- c(1, 1, 1, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 
@@
     if (4 %in% model_ssid) {
-
       .predictions <- filter(predictions, ssid == 4)
-      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
-      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
-      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
-      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 5) indices <- c(1, 1, 1, 2, 2)
+      if (length(unique(.predictions$year)) == 6) indices <- c(1, 1, 1, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 
@@
     if (16 %in% model_ssid) {
-
       .predictions <- filter(predictions, ssid == 16)
-      #unique(.predictions$year)
-      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
-      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
-      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
-      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      # unique(.predictions$year)
+      if (length(unique(.predictions$year)) == 5) indices <- c(1, 1, 1, 2, 2)
+      if (length(unique(.predictions$year)) == 6) indices <- c(1, 1, 1, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 
@@
     if (exists("adult_predictions")) {
-            saveRDS(out_mature, file = paste0("data/", spp, "/trends-", 
-        spp, covs, "-", ssid_string, "-mature-bioclimatic-priors-", priors, ".rds"))
-
+      saveRDS(out_mature, file = paste0(
+        "data/", spp, "/trends-",
+        spp, covs, "-", ssid_string, "-mature-bioclimatic-priors-", priors, ".rds"
+      ))
     } else {
@@
       rm(out_mature)
-      saveRDS(out_total, file = paste0("data/", spp, "/trends-", 
-        spp, covs, "-", ssid_string, "-total-bioclimatic-priors-", priors, ".rds"))
+      saveRDS(out_total, file = paste0(
+        "data/", spp, "/trends-",
+        spp, covs, "-", ssid_string, "-total-bioclimatic-priors-", priors, ".rds"
+      ))
     }
@@
 ```{r }
-if (!exists("out_mature")) {out_recent <- out_total} else {out_recent <- out_mature}
+if (!exists("out_mature")) {
+  out_recent <- out_total
+} else {
+  out_recent <- out_mature
+}
 
@@
 current_biomass
-
 ```
@@
     transform_col = fourth_root_power,
-    #raster_limits = c(0, quantile(data$units_per_decade, 0.99999)),
+    # raster_limits = c(0, quantile(data$units_per_decade, 0.99999)),
     legend_position = c(0.2, 0.25)
@@
     if (1 %in% model_ssid) {
-
       .predictions <- filter(predictions, ssid == 1)
-      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
-      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
-      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
-      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 5) indices <- c(1, 1, 1, 2, 2)
+      if (length(unique(.predictions$year)) == 6) indices <- c(1, 1, 1, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
       out_im1 <- make_trend_data(predictions,
@@
       unique(.predictions$year)
-      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
-      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
-      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
-      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 5) indices <- c(1, 1, 1, 2, 2)
+      if (length(unique(.predictions$year)) == 6) indices <- c(1, 1, 1, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 
@@
       .predictions <- filter(predictions, ssid == 16)
-      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
-      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
-      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
-      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 5) indices <- c(1, 1, 1, 2, 2)
+      if (length(unique(.predictions$year)) == 6) indices <- c(1, 1, 1, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if (length(unique(.predictions$year)) == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 
@@
 ```{r results = "hide"}
-if(exists("adult_predictions")) { biomass_plots <- c(p_adult, p_imm) } else {
+if (exists("adult_predictions")) {
+  biomass_plots <- c(p_adult, p_imm)
+} else {
   biomass_plots <- c(p_adult)
@@
   for (i in seq_len(length(model_ssid))) {
-    p_adult_st[[i]] <-plot_facet_map(ad_prediction_list[[i]], "epsilon_st",
+    p_adult_st[[i]] <- plot_facet_map(ad_prediction_list[[i]], "epsilon_st",
       white_zero = TRUE,
@@
       low_fill = "#5E4FA2"
-    #raster_limits = c(0, max_raster),
-    #legend_position = "none",
-    #transform_col = fourth_root_power
+      # raster_limits = c(0, max_raster),
+      # legend_position = "none",
+      # transform_col = fourth_root_power
     ) +
-    #labs(fill = "kg/ha") +
+      # labs(fill = "kg/ha") +
       ggtitle(paste0("", species, " mature spatiotemporal variation"))
@@
   print(p_adult_st)
-#} else {
+  # } else {
   # p_adult_st <- plot_facet_map(predictions, "epsilon_st",
@@
 
-data <- mutate(adult_predictions, cell_id = paste0(X,Y)) %>% 
+data <- mutate(adult_predictions, cell_id = paste0(X, Y)) %>%
   group_by(cell_id) %>%
-  mutate(cell_max = max(bioclimatic)) %>% group_by(year.f) %>% mutate(ann_biomass = sum(est_exp))
+  mutate(cell_max = max(bioclimatic)) %>%
+  group_by(year.f) %>%
+  mutate(ann_biomass = sum(est_exp))
 unique(data$ann_biomass)
@@
 library(forcats)
-violin_data <- data %>% filter (cell_max > threshold) %>% mutate(year = fct_reorder(year.f, desc(ann_biomass)))
+violin_data <- data %>%
+  filter(cell_max > threshold) %>%
+  mutate(year = fct_reorder(year.f, desc(ann_biomass)))
 glimpse(violin_data)
@@
 ggplot(violin_data) +
-  geom_hline(aes(yintercept=0)) +
+  geom_hline(aes(yintercept = 0)) +
   geom_violin(aes(fct_reorder(year.f, (ann_biomass)), epsilon_st, fill = year.f)) +
-  geom_text(aes(fct_reorder(year.f, (ann_biomass)), y=2.8, label = round(ann_biomass)), check_overlap = TRUE) +
+  geom_text(aes(fct_reorder(year.f, (ann_biomass)), y = 2.8, label = round(ann_biomass)), check_overlap = TRUE) +
   scale_fill_viridis_d(option = "C") +
@@
   xlab("Year") +
-  #coord_flip() +
-  theme(legend.position="none") +
+  # coord_flip() +
+  theme(legend.position = "none") +
   ggtitle(paste("Local", species, "abundance deviations not explained by climate and overall abundance"))
@@
 #   ggtitle(paste("Local", species, "abundance deviations not explained by climate and overall abundance"))
-
 ```
@@
 print(p_bioclim)
-
 ```
@@
 gvocc <- ggplot(mutate(df, `Trend\n(kg/decade)` = -trend)) +
-  ggquiver::geom_quiver(aes(x, y, u = u_velo, v = v_velo,
-    colour = `Trend\n(kg/decade)`),
-    vecsize = 2) +
+  ggquiver::geom_quiver(aes(x, y,
+    u = u_velo, v = v_velo,
+    colour = `Trend\n(kg/decade)`
+  ),
+  vecsize = 2
+  ) +
   scale_colour_gradient2(low = scales::muted("blue"), high = scales::muted("red")) +
-  xlab("UTM") + ylab("UTM") + #ggtitle("Velocity") +
+  xlab("UTM") + ylab("UTM") + # ggtitle("Velocity") +
   gfplot::theme_pbs()
@@
 # )
-
-```
+```

### 2019-12-11 analysis/VOCC/2-biomass-depth-only-model.Rmd
update for reml version of sdmTMB; add year_cutoff for bioclimatic calculations

@@
 params: 
-    species: "Petrale Sole"
+    species: "Redbanded Rockfish"
     region: "All synoptic surveys"
     knots: 600
-    covariates: "+trawled+as.factor(ssid)" # additional fixed-effect covariates
-    covs: "-trawled-ssid" # covariate string used to name model rds 
+    covariates: "+trawled+as.factor(ssid) + trawled:as.factor(ssid)" # additional fixed-effect covariates
+    covs: "-trawled-ssid-reml" # covariate string used to name model rds 
     update_model: TRUE
@@
 
-  adult_biomass <- sdmTMB::sdmTMB(data,
+  adult_biomass <- sdmTMB::sdmTMB(data=data,
     adult_formula,
@@
     family = tweedie(link = "log"),
-    ar1_fields = TRUE,
+    #ar1_fields = TRUE,
     include_spatial = TRUE,
+    reml = TRUE,
     enable_priors = priors,
-    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
     silent = FALSE
@@
 
-  imm_biomass <- sdmTMB::sdmTMB(data,
+  imm_biomass <- sdmTMB::sdmTMB(data=data,
     imm_formula,
@@
     family = tweedie(link = "log"),
-    ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
+    #ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
+    reml = TRUE,
     include_spatial = TRUE,
     enable_priors = priors,
-    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
     silent = FALSE
@@
 
-  total_biomass <- sdmTMB::sdmTMB(data,
+  total_biomass <- sdmTMB::sdmTMB(data=data,
     dens_formula,

### 2019-12-11 analysis/VOCC/3-climate-models-fixed-both.Rmd
update for reml version of sdmTMB; add year_cutoff for bioclimatic calculations

@@
     priors: FALSE
-    knots: 500
-    covariates: "+trawled" # additional non-climate variables
-    covs: "-log-both-trawled" # string describing model covariates
+    knots: 800
+    covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
+    # covs: "-tv-depth-past" #
+    # year_cutoff: 2013
+    covs: "-tv-depth-all" #"-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
+    year_cutoff: NULL
+    #covs: "-log-both-noAR1" #"-log-both-noAR1" # string describing model covariates
 ---
@@
 knots <- params$knots
+year_cutoff <- params$year_cutoff
 
@@
 paste("knots =", knots)
+paste("year cutoff <", year_cutoff)
 
@@
 }
+
+#   d <- data %>% filter(year==2011) #%>% filter(depth> 400) 
+# = plot((adult_density)~(depth), data=d)
+
+if (!is.null(year_cutoff)) data <- filter(data, year < year_cutoff)
 ```
@@
 
-    adult_biomass_d <- sdmTMB::sdmTMB(data,
+    adult_biomass_d <- sdmTMB::sdmTMB(data = data,
       adult_formula,
@@
       family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
+      #ar1_fields = TRUE, 
+      #include_spatial = FALSE, # default = TRUE
+      reml = TRUE,
       enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
@@
       file = paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-mat-biomass-", spp, covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
 
-    imm_biomass_d <- sdmTMB::sdmTMB(data, 
+    imm_biomass_d <- sdmTMB::sdmTMB(data = data, 
       imm_formula,
@@
       family = tweedie(link = "log"),
-      ar1_fields = FALSE,
-      include_spatial = TRUE,
+      #ar1_fields = TRUE,
+      #include_spatial = FALSE, # default = TRUE
+      reml = TRUE,
       enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
@@
       file = paste0("data/", spp, 
-        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-imm-biomass-", spp, covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
 
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""))
-
-    total_biomass_d <- sdmTMB::sdmTMB(data,
-      dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-
-    saveRDS(total_biomass_d, file = paste0("data/", spp, 
-      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
-      ))
+    # dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""))
+    # 
+    # total_biomass_d <- sdmTMB::sdmTMB(data,
+    #   dens_formula,
+    #   time_varying = ~ 0 + depth_scaled + depth_scaled2, 
+    #   time = "year", spde = spde,
+    #   family = tweedie(link = "log"),
+    #   #ar1_fields = TRUE,
+    #   #include_spatial = FALSE, # default = TRUE
+    #   enable_priors = priors,
+    #   control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+    #   silent = FALSE
+    # )
+    # 
+    # saveRDS(total_biomass_d, file = paste0("data/", spp, 
+    #   "/model-total-biomass-", spp, covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
+    #   ))
   }
@@
 ```
+
+Save predictions for full spatial grid
+
+```{r eval=FALSE}
+#fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+
+rm(adult_predictions)
+rm(imm_predictions)
+rm(predictions)
+
+if (any(names(data) == "adult_density")) {
+#if (length(unique(fish$maturity_code)) > 2) {
+  nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
+
+  adult_biomass <- readRDS(paste0("data/", spp, 
+        "/mod-mat-biomass-", spp, covs,"-", year_cutoff, "-", ssid_string, 
+        "-prior-", priors, ".rds"))
+
+  nd <- nd_all %>%
+    filter(ssid %in% model_ssid) %>%
+    filter(year %in% unique(adult_biomass$data$year)) %>% 
+    mutate(
+      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
+      do_mlpl_scaled2 = do_mlpl_scaled^2) %>% 
+    select(-est, -est_non_rf, - est_rf, -omega_s, -epsilon_st, -zeta_s)
+  nd <- na.omit(nd)
+  #nd$year <- as.integer(nd$year)
+
+  adult_predictions <- predict(adult_biomass, newdata = nd)
+  saveRDS(adult_predictions, file = paste0("data/", spp,
+    "/predictions-", spp, covs,"-", year_cutoff, "-", ssid_string, 
+        "-prior-", priors, ".rds"))
+
+  imm_biomass <- readRDS(paste0("data/", spp, 
+        "/mod-imm-biomass-", spp, covs,"-", year_cutoff, "-", ssid_string, 
+        "-prior-", priors, ".rds"))
+
+    nd <- nd_all %>%
+    filter(ssid %in% model_ssid) %>%
+    filter(year %in% unique(imm_biomass$data$year)) %>% 
+    mutate(
+      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
+      do_mlpl_scaled2 = do_mlpl_scaled^2) %>% 
+    select(-est, -est_non_rf, - est_rf, -omega_s, -epsilon_st, -zeta_s)
+  nd <- na.omit(nd)
+  #nd$year <- as.integer(nd$year)
+
+  imm_predictions <- predict(imm_biomass, newdata = nd)
+  saveRDS(imm_predictions, file = paste0("data/", spp,
+    "/predictions-", spp, covs,"-", year_cutoff, "-", ssid_string, 
+        "-prior-", priors, ".rds"))
+
+} else {
+
+    # total_biomass <- readRDS(paste0("data/", spp,
+    #   "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+    # 
+    # nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
+    # nd <- nd_all %>% 
+    #   filter(ssid %in% model_ssid) %>%
+    #   filter(year %in% unique(total_biomass$data$year))%>% 
+    # mutate(
+    #   do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
+    #   do_mlpl_scaled2 = do_mlpl_scaled^2) %>% 
+    # select(-est, -est_non_rf, - est_rf, -omega_s, -epsilon_st, -zeta_s)
+    # nd <- na.omit(nd)
+    # #nd$year <- as.integer(nd$year)
+    # 
+    # predictions <- predict(total_biomass, newdata = nd)
+    # saveRDS(predictions, file = paste0("data/", spp,
+    #   "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
+}
+```

### 2019-12-11 analysis/VOCC/3-climate-models-fixed-depth.Rmd
update for reml version of sdmTMB; add year_cutoff for bioclimatic calculations

@@
 params: 
-    species: "Redbanded Rockfish"
+    #species: "Redbanded Rockfish"
+    #species: "Pacific Cod"
+    #species: "Pacific Ocean Perch"
+    #species: "Arrowtooth Flounder" #all DO smiley
+    #species: "Petrale Sole" #adult DO almost quadratic...
+    #species: "English Sole" #all DO smiley
+    #species: "Dover Sole" #all DO smiley
+    species: "Flathead Sole"
     region: "All synoptic surveys"
@@
     priors: FALSE
-    knots: 500
+    knots: 600
+    # knots: 500
     covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
-    covs: "-AR1-spatial" #
-    #covs: "-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
+    covs: "-past" # hardcoded "-fixed" makes covs = "-fixed-past"
+    year_cutoff: 2013
+    # covs: "-all" # hardcoded -fixed makes covs = "-fixed-all"
+    # year_cutoff: NULL
+    #"-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
 ---
@@
 # = plot((adult_density)~(depth), data=d)
-
+if (covs == "-past") data <- filter(data, year < year_cutoff)
 ```
@@
       family = tweedie(link = "log"),
-      ar1_fields = TRUE, 
-      include_spatial = FALSE, # default = TRUE
+      #ar1_fields = TRUE, 
+      #include_spatial = FALSE, # default = TRUE
       reml = TRUE,
@@
       file = paste0("data/", spp, 
-        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-mat-biomass-", spp,"-fixed", covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
       family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = FALSE, # default = TRUE
+      #ar1_fields = TRUE,
+      #include_spatial = FALSE, # default = TRUE
       reml = TRUE,
@@
       file = paste0("data/", spp, 
-        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-imm-biomass-", spp,"-fixed", covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
         ))
@@
     saveRDS(total_biomass_d, file = paste0("data/", spp, 
-      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
+      "/model-total-biomass-", spp, "-fixed", covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
       ))

### 2019-12-11 analysis/VOCC/3-fixed-var-response-plots.Rmd
add fix_depth plots, update_model to model loading so that old models can still be loaded

@@
 params: 
-    species: "Redbanded Rockfish"
+    species: "Flathead Sole" 
     priors: FALSE # call model that used priors?
@@
     model_w_do: TRUE
-    covs: "-log-both-trawled" # covariate string used to name model rds
-    cov_number: 2
+    fix_depth: TRUE
+    #covs: "-all-" #"-log-both-fixed-depth"
+    #covs: "-past-2017"
+    # fix_depth: FALSE #TRUE
+    # #covs: "-log-both-noAR1" #"-log-both-fixed-depth"
+    # covs: "-log-both-AR1"
+    cov_number: 2 #covs: "-log-both-noOMEGA" # covariate string used to name model rds
 ---
@@
 model_w_do <- params$model_w_do
+fix_depth <- params$fix_depth
 
@@
 paste("log(temperature) =", log_temp)
+paste("depth fixed =", fix_depth)
 
@@
   ))
+  adult<-sdmTMB:::update_model(adult) 
   imm <- readRDS(paste0("data/", spp,
@@
   ))
+  imm <-sdmTMB:::update_model(imm) 
   })
@@
         ggtitle(paste(species, names(model_list[i]))) +
-        geom_vline(xintercept=(do[[i]]$optimal_do), size= 1.2, colour = "darkgrey") + 
         geom_vline(xintercept=(do[[i]]$lower_do_50), linetype="longdash", alpha = 0.75) +
@@
         #geom_vline(xintercept=(do[[i]]$upper_do_75), linetype="dotted", alpha = 0.5) +
-        geom_vline(xintercept=(do[[i]]$lower_do_25), linetype="dotted", alpha = 0.5) #+ 
+        geom_vline(xintercept=(do[[i]]$lower_do_25), linetype="dotted", alpha = 0.5) +
+        geom_vline(xintercept=(do[[i]]$optimal_do), size= 1.2, colour = "darkgrey") 
         #geom_vline(xintercept=(do[[i]]$upper_do_25), linetype="dotted", alpha = 0.5) 
@@
         ggtitle(paste(species, names(model_list[i]))) + 
-        geom_vline(xintercept=td[[i]]$optimal_t, size= 1.2, colour = "darkgrey") + 
-        #geom_vline(xintercept=td[[i]]$lower_t_50, linetype="longdash", alpha = 0.75) + 
+        geom_vline(xintercept=td[[i]]$lower_t_50, linetype="longdash", alpha = 0.75) +
         geom_vline(xintercept=td[[i]]$upper_t_50, linetype="longdash", alpha = 0.75) +
-        #geom_vline(xintercept=td[[i]]$lower_t_75, linetype="dotted", alpha = 0.5) + 
+        geom_vline(xintercept=td[[i]]$lower_t_75, linetype="dotted", alpha = 0.5) +
         geom_vline(xintercept=td[[i]]$upper_t_75, linetype="dotted", alpha = 0.5) +
-        #geom_vline(xintercept=td[[i]]$lower_t_25, linetype="dotted", alpha = 0.5) + 
-        geom_vline(xintercept=td[[i]]$upper_t_25, linetype="dotted", alpha = 0.5)
-
+        geom_vline(xintercept=td[[i]]$lower_t_25, linetype="dotted", alpha = 0.5) +
+        geom_vline(xintercept=td[[i]]$upper_t_25, linetype="dotted", alpha = 0.5) +
+        geom_vline(xintercept=td[[i]]$optimal_t, size= 1.2, colour = "darkgrey") 
   } else {
@@
     "data/", spp,
-    "/mod-mat-biomass-", spp, "-trawled-ssid-1n3n4n16-prior-FALSE.rds"
+    "/mod-mat-biomass-", spp, "-trawled-ssid-reml-1n3n4n16-prior-FALSE.rds"
   ))
+adult_biomass<-sdmTMB:::update_model(adult_biomass) 
   imm_biomass <- readRDS(paste0(
     "data/", spp,
-    "/mod-imm-biomass-", spp, "-trawled-ssid-1n3n4n16-prior-FALSE.rds"
+    "/mod-imm-biomass-", spp, "-trawled-ssid-reml-1n3n4n16-prior-FALSE.rds"
   ))
+  imm_biomass<-sdmTMB:::update_model(imm_biomass) 
  # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
@@
     "data/", spp,
-    "/mod-total-biomass-", spp, "-trawled-ssid-1n3n4n16-prior-FALSE.rds"
+    "/mod-total-biomass-", spp, "-trawled-ssid-reml-1n3n4n16-prior-FALSE.rds"
   ))
@@
  if(depth_model_list[[i]]$model$convergence==0){
+
+   if(fix_depth) {
+     d[[i]] <- fixed_density(model_list[[i]], predictor = "depth", 
+       quadratics_only = FALSE, fixed_params = 3)
+   } else {
    d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
+   }
 
@@
 odata <- data %>% filter(!is.na(do_mlpl)) %>% 
-  mutate(ad_density_class = ifelse(adult_density==0, 0, round(log10((adult_density*100000)+1))),
+  mutate(#do_mlpl = raw_do, 
+    ad_density_class = ifelse(adult_density==0, 0, round(log10((adult_density*100000)+1))),
     im_density_class = ifelse(imm_density==0, 0, round(log10((imm_density*10000000)+1)))) %>% 
@@
   scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
-  geom_vline( xintercept = tdat[tdat$model =="imm", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
   geom_vline( xintercept = tdat[tdat$age =="imm", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
@@
 png(
-  file = paste0("figs/", spp, "/do-both-log-", spp, x2016, covs, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/do-", spp, x2016, covs, "-priors-", priors, ".png"),
   res = 600,
@@
 png(
-  file = paste0("figs/", spp, "/temp-both-log-", spp, x2016, covs, "-priors-", priors, ".png"),
+  file = paste0("figs/", spp, "/temp-", spp, x2016, covs, "-priors-", priors, ".png"),
   res = 600,
@@
     "figs/", spp,
-    "/depth-", spp, x2016, covs, "-priors-", priors, ".png"
+    "/depth-", spp, x2016, covs, "-priors-", priors, "-revised.png"
   ),
@@
 other_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
+# # If data far a particular species needs redoing
+# other_params <- subset(other_params, species!="Redbanded Rockfish")
 
@@
 all_params <- all_params %>% distinct()
-# # If data far a particular species needs redoing
-# all_params <- subset(other_params, species!="Longnose Skate")
 # # or certain rows need removal
@@
 saveRDS(all_params, file = "data/root-values-by-species-by-age.rds")
-#spp_params <- all_params[all_params$species==species,]
+
+# spp_params <- all_params[all_params$species==species,]
+# View(spp_params)
+
 glimpse(new_params)

### 2019-12-11 analysis/VOCC/3-bioclimatic-trends.Rmd
move bioclim functions to package; add option for tv-depth model; add 2009-2011 year pair

@@
 params: 
-    #species: "Pacific Cod" # 
-    species: "Redbanded Rockfish"
+    # species: "Redbanded Rockfish"
+    # species: "Pacific Cod" 
+    species: "Pacific Ocean Perch" 
     region: "All synoptic surveys"
@@
     covs: "-fixed"  
+    # covs: "-tv-depth"
     # year_cutoff: 2012
@@
   adult_predictions_obj <- predict(adult_biomass, newdata = nd, return_tmb_object = TRUE)
-  adult_predictions <- adult_predictions_obj$data
-
-  # write as function
-  bioclim_by_year <- function(yr, model_for_prior_year, pred_grid) {
-    nd <- pred_grid
-
-    nd_trim <- nd %>% filter(year < yr)
-    fake_years <- nd %>%
-      mutate(year = year - 2) %>%
-      filter(year > 2007) %>%
-      filter(year < yr)
-
-    fake <- sort(unique(fake_years$year))
-    true <- sort(unique(nd$year))
-
-    if (!identical(fake, true)) "Something is wrong with the pred_grid years."
-
-    prior_years <- predict(model_for_prior_year, newdata = nd_trim, return_tmb_object = FALSE)
-    fake_est <- predict(model_for_prior_year, newdata = fake_years, return_tmb_object = FALSE)
-    fake_est$year <- fake_est$year + 2
-
-    bioclim_start <- filter(prior_years, year >= yr - 2) %>%
-      filter(year <= yr - 1) %>%
-      mutate(bioclim = est)
-    bioclim_end <- filter(fake_est, year >= yr) %>%
-      filter(year <= yr + 1) %>%
-      mutate(bioclim = est_non_rf + omega_s)
-    bioclim_all <- dplyr::bind_rows(bioclim_start, bioclim_end) %>%
-      select(X, Y, year, bioclim)
-
-    bioclim_all
-  }
-
-  model_pre_2013 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-past-2013-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  sort(unique(model_pre_2013$data$year))
-
-  bioclim_2013 <- bioclim_by_year(2013, model_pre_2013, nd)
-  bioclim_2013 <- rename(bioclim_2013, bioclim_2013 = bioclim)
-  glimpse(bioclim_2013)
-
-  model_pre_2015 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  sort(unique(model_pre_2015$data$year))
-
-  bioclim_2015 <- bioclim_by_year(2015, model_pre_2015, nd)
-  bioclim_2015 <- rename(bioclim_2015, bioclim_2015 = bioclim)
-  glimpse(bioclim_2015)
-
-  model_pre_2017 <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-past-2017-", ssid_string, "-prior-", priors, ".rds"
-  ))
-  sort(unique(model_pre_2017$data$year))
-
-  bioclim_2017 <- bioclim_by_year(2017, model_pre_2017, nd)
-  bioclim_2017 <- rename(bioclim_2017, bioclim_2017 = bioclim)
-  glimpse(bioclim_2017)
-
-  # for multiyear bioclim
-  bioclim_multi_year <- function(yr, model_for_prior_year, pred_grid) {
-    nd <- pred_grid
-    period2_range <- max(nd$year) - yr + 1
-
-    nd_trim <- nd %>% filter(year < yr)
-    fake_years <- nd %>%
-      mutate(year = year - 2) %>%
-      filter(year > 2007) %>%
-      filter(year < yr)
-
-    fake <- sort(unique(fake_years$year))
-    true <- sort(unique(model_for_prior_year$data$year))
-
-    if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years."
-
-    prior_years <- predict(model_for_prior_year, newdata = nd_trim, return_tmb_object = FALSE)
-
-    past_predictions <- prior_years %>%
-      mutate(past_est = est, bioclim = est) %>%
-      select(X, Y, year, past_est, bioclim)
-
-    fake_est <- predict(model_for_prior_year, newdata = fake_years, return_tmb_object = FALSE)
-    fake_est$year <- fake_est$year + 2
-
-    bioclim <- filter(fake_est, year >= yr + 2 - 2) %>%
-      filter(year <= yr + 2 - 1) %>%
-      mutate(bioclim = est_non_rf + omega_s)
-
-    if (period2_range >= 4) {
-      fake_minus4 <- nd %>%
-        mutate(year = year - 4) %>%
-        filter(year > 2007)
-
-      fake <- sort(unique(fake_minus4$year))
-      true <- sort(unique(nd_trim$year))
-
-      if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years -4."
-
-      fake_est_minus4 <- predict(model_for_prior_year, newdata = fake_minus4, return_tmb_object = FALSE)
-      fake_est_minus4$year <- fake_est_minus4$year + 4
-      # sort(unique(fake_est_minus4$year))
-
-      bioclim_4 <- filter(fake_est_minus4, year >= yr + 4 - 2) %>%
-        filter(year <= yr + 4 - 1) %>%
-        mutate(bioclim = est_non_rf + omega_s)
-
-      bioclim <- dplyr::bind_rows(bioclim, bioclim_4)
-    }
-
-    if (period2_range >= 6) {
-      fake_minus6 <- nd %>%
-        mutate(year = year - 6) %>%
-        filter(year > 2007)
-      fake <- sort(unique(fake_minus6$year))
-      true <- sort(unique(nd_trim$year))
-
-      if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years -6."
-
-      fake_est_minus6 <- predict(model_for_prior_year, newdata = fake_minus6, return_tmb_object = FALSE)
-      fake_est_minus6$year <- fake_est_minus6$year + 6
-
-      bioclim_6 <- filter(fake_est_minus6, year >= yr + 6 - 2) %>%
-        filter(year <= yr + 6 - 1) %>%
-        mutate(bioclim = est_non_rf + omega_s)
-
-      bioclim <- dplyr::bind_rows(bioclim, bioclim_6)
-    }
-
-    bioclim_all <- dplyr::bind_rows(past_predictions, bioclim) %>%
-      select(X, Y, year, past_est, bioclim)
-
-    bioclim_all
-  }
-
+
+  # multiyear bioclimatic calculation
   model_pre_2015 <- readRDS(paste0(
@@
 
+  adult_predictions <- adult_predictions_obj$data
   adult_predictions <- left_join(adult_predictions, bioclim_blob)
+
+  # saveRDS(adult_predictions, file = paste0(
+  #   "data/", spp,
+  #   "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-mat-2015-cutoff.rds"
+  # ))
+  # 
+  # add biennial bioclimatic calculation
+  model_pre_2011 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2011-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(model_pre_2011$data$year))
+
+  bioclim_2011 <- bioclim_by_year(2011, model_pre_2011, nd)
+  bioclim_2011 <- rename(bioclim_2011, bioclim_2011 = bioclim)
+  glimpse(bioclim_2011)
+
+  model_pre_2013 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2013-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(model_pre_2013$data$year))
+
+  bioclim_2013 <- bioclim_by_year(2013, model_pre_2013, nd)
+  bioclim_2013 <- rename(bioclim_2013, bioclim_2013 = bioclim)
+  glimpse(bioclim_2013)
+
+  model_pre_2015 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(model_pre_2015$data$year))
+
+  bioclim_2015 <- bioclim_by_year(2015, model_pre_2015, nd)
+  bioclim_2015 <- rename(bioclim_2015, bioclim_2015 = bioclim)
+  glimpse(bioclim_2015)
+
+  model_pre_2017 <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-past-2017-", ssid_string, "-prior-", priors, ".rds"
+  ))
+  sort(unique(model_pre_2017$data$year))
+
+  bioclim_2017 <- bioclim_by_year(2017, model_pre_2017, nd)
+  bioclim_2017 <- rename(bioclim_2017, bioclim_2017 = bioclim)
+  glimpse(bioclim_2017)
+
+  adult_predictions <- left_join(adult_predictions, bioclim_2011)
   adult_predictions <- left_join(adult_predictions, bioclim_2013)
@@
 
-  View(adult_predictions)
+#  View(adult_predictions)
 
@@
   ))
+
 } else { # need to add non-split-by-age version
@@
   for (i in seq_len(length(model_ssid))) {
-    p_adult_st[[i]] <- plot_facet_map(ad_prediction_list[[i]], "epsilon_st",
-      white_zero = TRUE,
-      high_fill = "#66C2A5",
-      low_fill = "#5E4FA2"
-      # raster_limits = c(0, max_raster),
+    p_adult_st[[i]] <- plot_facet_map(ad_prediction_list[[i]], "do_est",
+      # white_zero = TRUE,
+      # high_fill = "#66C2A5",
+      # low_fill = "#5E4FA2"
+      # # raster_limits = c(0, max_raster),
       # legend_position = "none",

### 2019-12-11 R/bioclimatic-predictions.R

@@ -1,0 +1,118 @@
+#' Biennial bioclimatic estimates
+#'
+#' @param yr Year for which bioclimatic predictions will be estimated
+#' @param model_for_prior_years sdmTMB model for all years prior to yr above
+#' @param pred_grid Appropriate prediction grid containing unique covariate values 
+#'    for each year in model and the year to be estimated. Can include additional 
+#'    years before and/or after above period.
+#'
+#' @export
+bioclim_by_year <- function(yr, model_for_prior_years, pred_grid) {
+  nd <- pred_grid
+  
+  nd_trim <- nd %>% filter(year < yr)
+  fake_years <- nd %>%
+    mutate(year = year - 2) %>%
+    filter(year > 2007) %>%
+    filter(year < yr)
+  
+  fake <- sort(unique(fake_years$year))
+  true <- sort(unique(nd$year))
+  
+  if (!identical(fake, true)) "Something is wrong with the pred_grid years."
+  
+  prior_years <- predict(model_for_prior_years, newdata = nd_trim, return_tmb_object = FALSE)
+  fake_est <- predict(model_for_prior_years, newdata = fake_years, return_tmb_object = FALSE)
+  fake_est$year <- fake_est$year + 2
+  
+  bioclim_start <- filter(prior_years, year >= yr - 2) %>%
+    filter(year <= yr - 1) %>%
+    mutate(bioclim = est)
+  bioclim_end <- filter(fake_est, year >= yr) %>%
+    filter(year <= yr + 1) %>%
+    mutate(bioclim = est_non_rf + omega_s)
+  bioclim_all <- dplyr::bind_rows(bioclim_start, bioclim_end) %>%
+    select(X, Y, year, bioclim)
+  
+  bioclim_all
+}
+
+#' Bioclimatic estimates for multiyear period 
+#' 
+#' @param yr First year for which bioclimatic predictions will be estimated
+#'
+#' @export
+bioclim_multi_year <- function(yr, model_for_prior_years, pred_grid) {
+  nd <- pred_grid
+  period2_range <- max(nd$year) - yr + 1
+  
+  nd_trim <- nd %>% filter(year < yr)
+  fake_years <- nd %>%
+    mutate(year = year - 2) %>%
+    filter(year > 2007) %>%
+    filter(year < yr)
+  
+  fake <- sort(unique(fake_years$year))
[…]

### 2019-12-17 analysis/VOCC/3-fixed-var-response-plots.Rmd
change plotted ranges to 95% of observed

@@
 params: 
-    species: "Flathead Sole" 
+    #species: "Redbanded Rockfish"
+    #species: "Pacific Ocean Perch"
+    #species: "Pacific Cod"
+    #species: "Petrale Sole" 
+    species: "Arrowtooth Flounder" #"Curlfin Sole"
     priors: FALSE # call model that used priors?
@@
     model_w_do: TRUE
+    # covs is a covariate string used to name model
+    # if
+    #fix_depth: FALSE 
+    #covs: "-tv-depth-all-"
+    #covs: "-tv-depth-past-2015" 
+    # or if 
     fix_depth: TRUE
-    #covs: "-all-" #"-log-both-fixed-depth"
-    #covs: "-past-2017"
-    # fix_depth: FALSE #TRUE
-    # #covs: "-log-both-noAR1" #"-log-both-fixed-depth"
-    # covs: "-log-both-AR1"
-    cov_number: 2 #covs: "-log-both-noOMEGA" # covariate string used to name model rds
+    covs: "-fixed-all-" 
+    #covs: "-fixed-past-2017"
+    cov_number: 2 
 ---
@@
   try({adult <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   imm <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   try({total <- readRDS(paste0("data/", spp,
-    "/model-total-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+    "/model-total-biomass-", spp, add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"))
   })
@@
   try({adult_survey1 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
   imm_survey1 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   try({total1 <- readRDS(paste0("data/", spp,
-    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
   })
@@
   try({adult_survey4 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
   imm_survey4 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
@@
   try({total4 <- readRDS(paste0("data/", spp,
-    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
   })
@@
   try({adult_survey16 <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
    imm_survey16 <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
@@
   try({total16 <- readRDS(paste0("data/", spp,
-    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
   })
@@
    do_plots[[i]] <- 
-     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0.2, 8), x_breaks = c(1,2,3,4,5,6,7,8)) +
+     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0.65, 6.25), x_breaks = c(1,2,3,4,5,6,7,8)) +
         ggtitle(paste(species, names(model_list[i]))) +
@@
    temp_plots[[i]] <- 
-     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 12), x_breaks = c(3,4,5,6,7,8,9,10,11)) +
+     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(4.68, 9.93), x_breaks = c(3,4,5,6,7,8,9,10,11)) +
         ggtitle(paste(species, names(model_list[i]))) + 
@@
 Save DO plots
-```{r eval=F}
+```{r eval=T}
 if(model_w_do) {
@@
 Save temperature plots
-```{r eval=F}
+```{r eval=T}
 png(

### 2019-12-17 analysis/VOCC/2-biomass-depth-only-model.Rmd
add model assessment plots for depth only and dixed depth climate model building

@@
 params: 
-    species: "Redbanded Rockfish"
+    #species: "Redbanded Rockfish"
+    #species: "Pacific Ocean Perch"
+    species: "Pacific Cod"
+    #species: "Petrale Sole"
     region: "All synoptic surveys"
-    knots: 600
-    covariates: "+trawled+as.factor(ssid) + trawled:as.factor(ssid)" # additional fixed-effect covariates
+    knots: 800
+    covariates: "+trawled+as.factor(ssid) + trawled:as.factor(ssid)" 
+    # additional fixed-effect covariates
     covs: "-trawled-ssid-reml" # covariate string used to name model rds 
@@
 ```
+
+```{r}
+adult_biomass <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
+point_predictions <- predict(adult_biomass)
+point_predictions$residuals <- residuals(adult_biomass)
+
+saveRDS(point_predictions, file = paste0("data/", spp,
+    "/check-mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
+```
+
+```{r}
+depth_only_predictions <- readRDS(paste0("data/", spp,
+    "/check-mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
+depth_only_predictions <- filter(depth_only_predictions, ssid %in% c(1,3)) %>% filter (year > 2007)
+
+ggplot(depth_only_predictions, aes(adult_density, residuals, colour = adult_density)) + 
+  geom_point() + scale_x_continuous(trans = 'log10') + facet_wrap(~year)
+```
+
+```{r}
+ggplot(depth_only_predictions, aes((est), residuals, colour = adult_density)) +
+  geom_point() + 
+  #scale_x_continuous(trans = 'log10') + 
+  facet_wrap(~year)
+```
+
+```{r}
+ggplot(depth_only_predictions, aes(X, Y, colour = (residuals))) + 
+  geom_point() + 
+  scale_colour_gradient2() + 
+  scale_x_continuous(trans = 'log10') + 
+  facet_wrap(~year) 
+```

### 2019-12-17 analysis/VOCC/3-climate-models-fixed-depth.Rmd
add model assessment plots for depth only and dixed depth climate model building

@@
     #species: "Arrowtooth Flounder" #all DO smiley
-    #species: "Petrale Sole" #adult DO almost quadratic...
+    species: "Petrale Sole" #adult DO almost quadratic...
     #species: "English Sole" #all DO smiley
     #species: "Dover Sole" #all DO smiley
-    species: "Flathead Sole"
+    #species: "Flathead Sole"
     region: "All synoptic surveys"
-    #region: "Both odd year surveys"
     log_temp: TRUE # should temperature be log()
@@
     knots: 600
-    # knots: 500
+    #knots: 500
     covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
     covs: "-past" # hardcoded "-fixed" makes covs = "-fixed-past"
-    year_cutoff: 2013
-    # covs: "-all" # hardcoded -fixed makes covs = "-fixed-all"
-    # year_cutoff: NULL
+    year_cutoff: 2015
+    #covs: "-all" # hardcoded -fixed makes covs = "-fixed-all"
+    #year_cutoff: NULL
     #"-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
@@
 if (covs == "-past") data <- filter(data, year < year_cutoff)
+
+ # saveRDS(data, file = paste0(spp, "-with-covariates-all-years.rds"))
+ # glimpse(data)
 ```
@@
 ```
+
+```{r}
+adult_biomass <- readRDS(paste0("data/", spp, 
+        "/mod-mat-biomass-", spp,"-fixed", covs, "-", year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"
+        ))
+
+point_predictions <- predict(adult_biomass)
+point_predictions$residuals <- residuals(adult_biomass)
+
+saveRDS(point_predictions, file = paste0("data/", spp,
+    "/check-mod-mat-biomass-", spp, "-fixed", covs, year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"))
+```
+
+```{r}
+all_fixed_predictions <- readRDS(paste0("data/", spp,
+    "/check-mod-mat-biomass-", spp, "-fixed", covs, year_cutoff, "-", ssid_string, "-prior-", priors, ".rds"))
+
+all_fixed_predictions <- filter(all_fixed_predictions, ssid %in% c(1,3))
+
+ggplot(all_fixed_predictions, aes(adult_density, residuals)) + 
+  geom_point() + scale_x_continuous(trans = 'log10') + facet_wrap(~year)
+```
+
+```{r}
+ggplot(all_fixed_predictions, aes(est, residuals, colour = adult_density)) + 
+  geom_point() + 
+  #scale_x_continuous(trans = 'log10') + 
+  facet_wrap(~year)
+```
+
+```{r}
+
+ggplot(all_fixed_predictions, aes(X, Y,  colour = residuals)) + 
+  geom_point() + 
+  scale_colour_gradient2() + 
+  scale_x_continuous(trans = 'log10') + 
+  facet_wrap(~year) 
+```

### 2020-01-01 analysis/VOCC/2-biomass-depth-only-model.Rmd
revise rmd for depth-only model to plot residuals, save predictions, and save time-varying depth plots

--- 2019-12-11 analysis/VOCC/2-biomass-depth-only-model.Rmd
@@
 params: 
-    species: "Redbanded Rockfish"
+    #species: "Redbanded Rockfish"
+    #species: "Pacific Ocean Perch"
+    #species: "Pacific Cod"
+    #species: "Petrale Sole"
+    species: "Pacific Halibut"
     region: "All synoptic surveys"
-    knots: 600
-    covariates: "+trawled+as.factor(ssid) + trawled:as.factor(ssid)" # additional fixed-effect covariates
-    covs: "-trawled-ssid-reml" # covariate string used to name model rds 
-    update_model: TRUE
+    knots: 500
+    covariates: "" # +as.factor(ssid)
+    #additional fixed-effect covariates
+    #covs: "-ssid-only" # covariate string used to name model rds 
+    covs: "-tv-depth-only"
+    #covs: "-trawled-ssid"
+    update_model: FALSE
 ---
@@
     include_spatial = TRUE,
-    reml = TRUE,
+    #reml = TRUE,
     enable_priors = priors,
@@
   nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-  adult_biomass <- readRDS(paste0("data/", spp,
+
+  try({adult_biomass <- readRDS(paste0("data/", spp,
     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
@@
     "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"))
+  })
 
+  try({
     imm_biomass <- readRDS(paste0("data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
+    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
 
@@
     "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"))
+ })
 
@@
 ```
+
+```{r}
+  try({ biomass <- readRDS(paste0("data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  })
+
+  try({ biomass <- readRDS(paste0("data/", spp,
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  })
+
+point_predictions <- predict(biomass)
+point_predictions$residuals <- residuals(biomass)
+
+saveRDS(point_predictions, file = paste0("data/", spp,
+    "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
+```
+
+```{r}
+depth_only_predictions <- readRDS(paste0("data/", spp,
+    "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
+depth_only_predictions <- filter(depth_only_predictions, 
+  #ssid %in% c(1,3)) %>% filter (
+  year > 2007)
+
+ try({ 
+   g <- ggplot(depth_only_predictions, aes(adult_density, residuals, colour = adult_density)) + 
+          geom_point() + scale_x_continuous(trans = 'log10') + facet_wrap(~year)
+ })
+
+ try({ 
+   g <- ggplot(depth_only_predictions, aes(density, residuals, colour = density)) + 
+          geom_point() + scale_x_continuous(trans = 'log10') + facet_wrap(~year)
+ })
+
+g
+```
+
+```{r}
+ try({ 
+   g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = adult_density)) +
+            geom_point() + #scale_x_continuous(trans = 'log10') + 
+            facet_wrap(~year)
+ })
+
+ try({ 
+   g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = density)) +
+            geom_point() + #scale_x_continuous(trans = 'log10') + 
+            facet_wrap(~year)
+ })
+
+g2
+```
+
+```{r}
+ggplot(depth_only_predictions, aes(X, Y, colour = (residuals))) + 
+  geom_point() + 
+  scale_colour_gradient2() + 
+  scale_x_continuous(trans = 'log10') + 
+  facet_wrap(~year) 
+```
+
+### Build time-varying depth plots
+#### Model without environmental covariates
+```{r }
+max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
+rm(depth_model_list)
+rm(adult_biomass)
+rm(imm_biomass)
+rm(total_biomass)
+
+try({  adult_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))})  
+#adult_biomass<-sdmTMB:::update_model(adult_biomass)
+
+  try({  imm_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, covs, "-1n3n4n16-prior-FALSE.rds"
+  ))
+  #imm_biomass<-sdmTMB:::update_model(imm_biomass) 
+ # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
+})
+
+    if (!exists("adult_biomass")) {
+      try({
+        total_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+      })
+    }
+
+if (!exists("adult_biomass")) adult_biomass <- NULL
+if (!exists("imm_biomass")) imm_biomass <- NULL
+if (!exists("total_biomass")) total_biomass <- NULL
+
+  depth_model_list <- list(adult = adult_biomass, imm = imm_biomass, total = total_biomass)
+  depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
+
+d <- list()
+depth_plots <- list()
+for (i in seq_len(length(depth_model_list))) {
+ if(depth_model_list[[i]]$model$convergence==0){
+  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
+
+  if (length(d[[i]]) == 0) {
+    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } else {
+    d[[i]]$x <- exp(d[[i]]$x)
+    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (without environmental variables)", xlimits = c(0, max_depth_found)) +
+      ggtitle(paste(species, names(depth_model_list[i])))
+  }
+ } else {
+    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  } 
+}
+print(depth_plots)
+```
+
+```{r}
+png(
+  file = paste0(
+    "figs/", spp,
+    "/depth-", spp, covs, ".png"
+  ),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 6
+)
+gridExtra::grid.arrange(
+  grobs = c(depth_plots),
+  nrow = 2,
+  top = grid::textGrob(paste(species, "(", covs, ")"))
+)
+dev.off()
+```
+
+```{r}
+depth_model_list
+```

### 2020-01-01 analysis/VOCC/4-vocc-biotic.Rmd
batch calculation of biotic variability and  gradient velocities

---
title: "`r paste(params$species)` climatic and biotic velocities for `r paste(params$region)` (immature: `r paste(params$immature)`"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document 
params: 
    species: "Arrowtooth Flounder" 
    immature: FALSE
    region: "both odd year surveys"
    #region: "West Coast Vancouver Island"
    covs: "-tv-depth-only"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.keep = "all", fig.show = "asis", dpi = 600,
  fig.width = 6, fig.height = 5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("gfranges")
library("dplyr")
library("ggplot2")

ggplot2::theme_set(gfplot::theme_pbs())
```

#### Set parameters
```{r }
species <- params$species
region <- params$region
covs <- params$covs
immature <- params$immature

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

paste("species =", species)
paste("region =", region)
paste("covs =", covs)
```

#### Set cell sizes and scale
```{r, echo=TRUE}
dist_intercept <- 0 # input_cell_size / 2
input_cell_size <- 2
scale_fac <- 2 # means that the raster is reprojected to _ X original grid (2 km)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
}
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, include=FALSE}
if (immature) {
  age <- "immature"
  d_all <-readRDS(paste0("data/", spp,
  "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-FALSE.rds")) %>% filter(year > 2007)

  } else {
    age <- "mature" 
      try({
        d_all <-readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-FALSE.rds")) %>% filter(year > 2007) 
      })
      try({
        d_all <-readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds")) %>% filter(year > 2007) 
      })

   }

nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select (X, Y, year, temp, do_est)

d_all <- left_join(d_all, nd_all, by=c("X","Y", "year"))

glimpse(d_all)
```

#### SET TIME FRAME
```{r}
multiyear <- TRUE

if (multiyear){
 #unique(d_all$year)
 d <- d_all %>% filter(ssid %in% model_ssid)
 length(unique(d$year))
 years_sampled <- length(unique(d$year))
 year_range <- (years_sampled -1)*2
 start_year <- 2007
 if (years_sampled==5) indices <- c(1, 2, 2, 2, 2)
 if (years_sampled==6) indices <- c(1, 2, 2, 2, 2, 2)
} else{ 
d <- d_all %>% filter(year >= start_year) %>% filter(year <= start_year+3)
unique(d$year)
indices <- c(1,2)
year_range <- 2
}

skip_time <- NULL
delta_t_step <- year_range
delta_t_total <- year_range
```

#### TRIM environment to range sampled
```{r}
#d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
d <- d %>% filter(do_est > 0.28) %>% filter(do_est < 7.06) # 0.005 and 0.995
#d <- d %>% filter(temp > 2.61) %>% filter(temp < 14.31) # full range
d <- d %>% filter(temp > 3.07) %>% filter(temp < 11.3) # 0.005 and 0.995
```

# Calculate climate velocity
### Dissolved Oxygen
```{r }
gf0 <- vocc_gradient_calc(d, "do_est",
  scale_fac = scale_fac,
  # all layers with max indice will be averaged for spatial gradient
  # if default (NULL) will use all years
  # indices = indices,
  divisor = 1,
  quantile_cutoff = 0.05
)

gf0 <- gf0 %>%
  mutate(
    trend_per_year = trend,
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>% dplyr::mutate(X = x, Y = y) %>%
   gfplot:::utm2ll(., utm_zone = 9)

grad0 <- plot_vocc(gf0,
  #theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "velocity",
  fill_label = "gradient velocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt # no_trans # 
) + labs(
  title = "do velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad0
```

### Temperature
```{r}
gf2 <- vocc_gradient_calc(d, "temp",
  scale_fac = scale_fac,
  # all layers with max indice will be averaged for spatial gradient
  # if default (NULL) will use all years
  # indices = indices,
  divisor = 1,
  quantile_cutoff = 0.05
)

gf2 <- gf2 %>%
  mutate(
    trend_per_year = trend ,
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>%
  dplyr::mutate(X = x, Y = y)

grad2 <- plot_vocc(gf2,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "velocity",
  fill_label = "gradient velocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt # no_trans # log10 # fourth_root_power # sqrt # 
) + labs(
  title = "temp velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad2
```

### Gradient-based BIOTIC VECTORS
```{r}
#indices <- c(1,2,3,4,5)

gf3 <- vocc_gradient_calc(d, "est",
  scale_fac = scale_fac,
  #indices = indices,
  divisor = 1,
  log_space = TRUE,
  quantile_cutoff = 0.025
)

gf3 <- gf3 %>%
  mutate(
    CV_formula = sqrt(exp(cv^2)-1),
    CV = sd/exp(mean), 
    mean_biomass = exp(mean)*10000,
    trend_per_year = exp(trend),
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>%
  dplyr::mutate(X = x, Y = y)

bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.0001
s <- sort(gf3$mean_biomass)

bio_sum <- cumsum(s)
lower_density_threshold <- s[which(bio_sum >= bio5perc)[1]]

gf3t <- filter(gf3, mean_biomass > lower_density_threshold)

grad3 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "velocity",
  fill_label = "gradient velocity",
  raster_alpha = 1,
 # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col =  sqrt # no_trans #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biotic velocity ", min(d$year), "-", max(d$year))
)
grad3
```

```{r}

grad4 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "mean_biomass",
  fill_label = "mean",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col =  sqrt#
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Mean biomass ", min(d$year), "-", max(d$year))
)
grad4

```

```{r}
grad5 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "trend_per_year",
  fill_label = "biomass trend",
  raster_alpha = 1,
  #raster_limits = c(0, 10),
  na_colour = "black",
  white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biomass trend ", min(d$year), "-", max(d$year))
)
grad5

```

```{r}
grad6 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "cv",
  fill_label = "sd in log space",
  raster_alpha = 1,
  #raster_limits = c(0, 7),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biotic variability (cv from vocc_gradient_calc funcition) ", min(d$year), "-", max(d$year))
)
grad6
```

```{r}
raster_max <- quantile(gf3t$CV, 0.90, na.rm = TRUE)
grad7 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "CV",
  fill_label = "CV",
  raster_alpha = 1,
  raster_limits = c(0, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("CV in estimated biomass ", min(d$year), "-", max(d$year))
)
grad7
```

### Save gradient-based VOCCs
```{r}

biotic_values <- gf3t %>% rename(
    biotic_vel = velocity, 
    biotic_trend = trend_per_year, 
    biotic_CV = CV,
    sd_est = cv,
    biotic_grad = gradient) %>% 
  mutate(sd_biomass = sd*10000,
    mean_biomass = exp(mean)*10000) %>%
  select(x, y, biotic_vel, biotic_trend, biotic_grad, biotic_CV, sd_est, sd_biomass, mean_biomass) %>% 
  mutate(start_year = !!start_year)

DO_values <- gf0 %>% rename( 
    DO_vel = velocity, 
    DO_trend = trend_per_year, 
    DO_grad = gradient) %>% 
  select(x, y, DO_vel, DO_trend, DO_grad)

temp_values <- gf2 %>% rename( 
    temp_vel = velocity, 
    temp_trend = trend_per_year, 
    temp_grad = gradient) %>% 
  select(x, y, temp_vel, temp_trend, temp_grad)

vocc <- left_join(biotic_values, DO_values) # add DO
vocc  <- left_join(vocc , temp_values) # add temp
vocc  <- mutate(vocc, species = !!species, method = "gradient")

write.csv(vocc, file = paste0(
  "data/biotic_test/biotic-vocc-", spp, "-", age, covs, "-", ssid_string, "-", min(d$year), "-", max(d$year), ".csv"
))

```

### 2020-01-01 R/vocc-gradient.R

@@ -20,6 +20,7 @@ vocc_gradient_calc <- function(data,
                                indices = NULL,
                                divisor = 10,
                                latlon = FALSE,
+                               log_space = FALSE,
                                quantile_cutoff = 0.05) {
 
   # # devtools::install_github("seananderson/vocc")
@@ -31,7 +32,28 @@ vocc_gradient_calc <- function(data,
     scale_fac = scale_fac,
     time_step = time_step
   )
-
+  
+  if(log_space) {
+  data$est_exp <- exp(data[[layer]])
+  
+  rbrick_exp <- make_gradient_brick(data, "est_exp",
+    scale_fac = scale_fac,
+    time_step = time_step
+  )
+  sdraster <- raster::calc(rbrick_exp, sd)
+  cvraster <- raster::calc(rbrick, sd)
+  
+  } else {
+    
+  sdraster <- raster::calc(rbrick, sd)
+  cvraster <- raster::calc(rbrick, sd)
+  
+  }
+  
+  
+  
+  
+  
   # Then calculate the trend per pixel:
   slopedat <- vocc::calcslope(rbrick, divisor = divisor)
 
@@ -45,9 +67,9 @@ vocc_gradient_calc <- function(data,
     #   } else {
     mnraster_brick <- raster::stackApply(rbrick, indices = indices, fun = mean)
     # use first time period
-    mnraster <- mnraster_brick[[1]]
+    # mnraster <- mnraster_brick[[1]]
     # to use last time period
-    # mnraster <- mnraster_brick[[raster::nlayers(mnraster_brick)]]
+    mnraster <- mnraster_brick[[raster::nlayers(mnraster_brick)]]
     
   } else {
     # uses average spatial gradient
@@ -81,6 +103,10 @@ vocc_gradient_calc <- function(data,
     dplyr::rename(trend = layer)
   rmnvalues_df <- as.data.frame(raster::rasterToPoints(mnraster))
   names(rmnvalues_df)[3] <- "mean"
+  rsdvalues_df <- as.data.frame(raster::rasterToPoints(sdraster))
+  names(rsdvalues_df)[3] <- "sd"
+  rcvvalues_df <- as.data.frame(raster::rasterToPoints(cvraster))
+  names(rcvvalues_df)[3] <- "cv"
[…]

### 2020-01-01 analysis/VOCC/4-pool-biotic-velocities.R

@@ -1,0 +1,26 @@
+getwd()
+setwd(here::here("/analysis/VOCC/data"))
+library("dplyr")
+
+#covs <- "-fixed"
+# mydir = paste0( "biotic_test1")
+mydir = paste0( "biotic_mature")
+#mydir = paste0( "biotic_immature")
+
+
+myfiles = list.files(path=mydir, pattern="*.csv", full.names=TRUE)
+myfiles
+biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% select(-X, -start_year)
+glimpse(biotic)
+
+# saveRDS(biotic, file = paste0(
+#   "multi-spp-biotic-vocc.rds"
+# ))
+# 
+saveRDS(biotic, file = paste0(
+  "multi-spp-biotic-vocc-mature.rds"
+))
+
+saveRDS(biotic, file = paste0(
+  "multi-spp-biotic-vocc-immature.rds"
+))

### 2020-01-01 analysis/VOCC/render-biotic-variability.R

@@ -1,0 +1,120 @@
+getwd()
+setwd(here::here("/analysis/VOCC"))
+env <- new.env() #parent = baseenv()
+
+list_species <- c(
+  "North Pacific Spiny Dogfish",
+  "Spotted Ratfish",
+  # "Pacific Tomcod",
+   "Walleye Pollock",
+  # "Pacific Cod",
+  "Sablefish",
+  "Lingcod",
+  # "Pacific Hake",
+  # "Rosethorn Rockfish",
+  "Redstripe Rockfish",
+  "Yellowmouth Rockfish",
+  # "Harlequin Rockfish",
+  # "Bocaccio", # winter-birthing, overfished
+   "Canary Rockfish", # schooling, winter-birthing
+  # "Copper Rockfish", # small sample
+   "Darkblotched Rockfish",
+   "Greenstriped Rockfish",
+  # "Pacific Ocean Perch", # schooling
+   "Quillback Rockfish",
+  # "Redbanded Rockfish",
+  # "Rougheye/Blackspotted Rockfish Complex",
+   "Sharpchin Rockfish",
+   "Shortbelly Rockfish", # small sample
+   "Silvergray Rockfish",
+   "Splitnose Rockfish",
+   "Widow Rockfish", # schooling
+   "Yellowtail Rockfish", # schooling
+   "Yelloweye Rockfish", # summer-birthing, overfished,
+   "Longspine Thornyhead",
+   "Shortspine Thornyhead",
+   "Arrowtooth Flounder",
+   "Rex Sole",
+  # "Petrale Sole",
+  # "English Sole",
+   "Dover Sole",
+   "Southern Rock Sole",
+   "Flathead Sole",
+   "Curlfin Sole",
+   "Sand Sole",
+ # "Slender Sole",
+ # "Pacific Sanddab",
+   "Pacific Halibut"
+)
+
+### SUBSETS OF SPECIES
+#  list_species <- c(
+# # #   "Arrowtooth Flounder"
+# # #    
+#   # "Petrale Sole",
+#   # "English Sole",
+#   # "Pacific Ocean Perch", # schooling
+#   # "Redbanded Rockfish",
+#   "Bocaccio"
+#   # "Pacific Cod"
+# # #   
+# # "Pacific Halibut"
+# #   #"Yelloweye Rockfish"
+# )
+
+list_regions <- c("All synoptic surveys")
+
+
+# dir.create(file.path("html/biomass-by-depth"))
+
[…]

### 2020-01-02 analysis/VOCC/2-biomass-depth-only-model.Rmd
style recent files

--- 2019-12-11 analysis/VOCC/2-biomass-depth-only-model.Rmd
@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
-                      #fig.path=paste0("figs/", spp, "/"),
-                      echo=FALSE, warning=FALSE, message=FALSE)
+knitr::opts_chunk$set(
+  fig.width = 11, fig.height = 8.5,
+  # fig.path=paste0("figs/", spp, "/"),
+  echo = FALSE, warning = FALSE, message = FALSE
+)
 library(dplyr)
@@
 
-  adult_biomass <- sdmTMB::sdmTMB(data=data,
+    adult_biomass <- sdmTMB::sdmTMB(
+      data = data,
       adult_formula,
@@
       family = tweedie(link = "log"),
-    #ar1_fields = TRUE,
+      # ar1_fields = TRUE,
       include_spatial = TRUE,
-    reml = TRUE,
+      # reml = TRUE,
       enable_priors = priors,
-    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
@@
 
-  imm_biomass <- sdmTMB::sdmTMB(data=data,
+    imm_biomass <- sdmTMB::sdmTMB(
+      data = data,
       imm_formula,
@@
       family = tweedie(link = "log"),
-    #ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
+      # ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
       reml = TRUE,
@@
       enable_priors = priors,
-    #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE
@@
     )
-
   } else {
-
     dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
 
-  total_biomass <- sdmTMB::sdmTMB(data=data,
+    total_biomass <- sdmTMB::sdmTMB(
+      data = data,
       dens_formula,
@@
 if (any(names(data) == "adult_density")) {
-#if (length(unique(fish$maturity_code)) > 2) {
+  # if (length(unique(fish$maturity_code)) > 2) {
   nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-  adult_biomass <- readRDS(paste0("data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
+  try({
+    adult_biomass <- readRDS(paste0(
+      "data/", spp,
+      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    ))
 
@@
     adult_predictions <- predict(adult_biomass, newdata = nd)
-  saveRDS(adult_predictions, file = paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"))
+    saveRDS(adult_predictions, file = paste0(
+      "data/", spp,
+      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
+    ))
+  })
 
-  imm_biomass <- readRDS(paste0("data/", spp,
+  try({
+    imm_biomass <- readRDS(paste0(
+      "data/", spp,
       "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
@@
     imm_predictions <- predict(imm_biomass, newdata = nd)
-  saveRDS(imm_predictions, file = paste0("data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"))
-
+    saveRDS(imm_predictions, file = paste0(
+      "data/", spp,
+      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
+    ))
+  })
 } else {
-
-    total_biomass <- readRDS(paste0("data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+  total_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
 
@@
   predictions <- predict(total_biomass, newdata = nd)
-    saveRDS(predictions, file = paste0("data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
+  saveRDS(predictions, file = paste0(
+    "data/", spp,
+    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
+  ))
 }
 ```
+
+```{r}
+try({
+  biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+
+try({
+  biomass <- readRDS(paste0(
+    "data/", spp,
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+
+point_predictions <- predict(biomass)
+point_predictions$residuals <- residuals(biomass)
+
+saveRDS(point_predictions, file = paste0(
+  "data/", spp,
+  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+))
+```
+
+```{r}
+depth_only_predictions <- readRDS(paste0(
+  "data/", spp,
+  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+))
+
+depth_only_predictions <- filter(
+  depth_only_predictions,
+  # ssid %in% c(1,3)) %>% filter (
+  year > 2007
+)
+
+try({
+  g <- ggplot(depth_only_predictions, aes(adult_density, residuals, colour = adult_density)) +
+    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
+})
+
+try({
+  g <- ggplot(depth_only_predictions, aes(density, residuals, colour = density)) +
+    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
+})
+
+g
+```
+
+```{r}
+try({
+  g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = adult_density)) +
+    geom_point() + # scale_x_continuous(trans = 'log10') +
+    facet_wrap(~year)
+})
+
+try({
+  g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = density)) +
+    geom_point() + # scale_x_continuous(trans = 'log10') +
+    facet_wrap(~year)
+})
+
+g2
+```
+
+```{r}
+ggplot(depth_only_predictions, aes(X, Y, colour = (residuals))) +
+  geom_point() +
+  scale_colour_gradient2() +
+  scale_x_continuous(trans = "log10") +
+  facet_wrap(~year)
+```
+
+### Build time-varying depth plots
+#### Model without environmental covariates
+```{r }
+max_depth_found <- max(data[data$present == 1, ]$raw_depth, na.rm = TRUE)
+rm(depth_model_list)
+rm(adult_biomass)
+rm(imm_biomass)
+rm(total_biomass)
+
+try({
+  adult_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  ))
+})
+# adult_biomass<-sdmTMB:::update_model(adult_biomass)
+
+try({
+  imm_biomass <- readRDS(paste0(
+    "data/", spp,
+    "/mod-imm-biomass-", spp, covs, "-1n3n4n16-prior-FALSE.rds"
+  ))
+  # imm_biomass<-sdmTMB:::update_model(imm_biomass)
+  # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
+})
+
+if (!exists("adult_biomass")) {
+  try({
+    total_biomass <- readRDS(paste0(
+      "data/", spp,
+      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    ))
+  })
+}
+
+if (!exists("adult_biomass")) adult_biomass <- NULL
+if (!exists("imm_biomass")) imm_biomass <- NULL
+if (!exists("total_biomass")) total_biomass <- NULL
+
+depth_model_list <- list(adult = adult_biomass, imm = imm_biomass, total = total_biomass)
+depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
+
+d <- list()
+depth_plots <- list()
+for (i in seq_len(length(depth_model_list))) {
+  if (depth_model_list[[i]]$model$convergence == 0) {
+    d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
+
+    if (length(d[[i]]) == 0) {
+      depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+    } else {
+      d[[i]]$x <- exp(d[[i]]$x)
+      depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (without environmental variables)", xlimits = c(0, max_depth_found)) +
+        ggtitle(paste(species, names(depth_model_list[i])))
+    }
+  } else {
+    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  }
+}
+print(depth_plots)
+```
+
+```{r}
+png(
+  file = paste0(
+    "figs/", spp,
+    "/depth-", spp, covs, ".png"
+  ),
+  res = 600,
+  units = "in",
+  width = 8.5,
+  height = 6
+)
+gridExtra::grid.arrange(
+  grobs = c(depth_plots),
+  nrow = 2,
+  top = grid::textGrob(paste(species, "(", covs, ")"))
+)
+dev.off()
+```
+
+```{r}
+depth_model_list
+```

### 2020-01-02 analysis/VOCC/3-bioclimatic-trends.Rmd
style recent files

@@
     # species: "Pacific Cod" 
-    species: "Pacific Ocean Perch" 
+    # species: "Pacific Ocean Perch" 
+    species: "Petrale Sole"
     region: "All synoptic surveys"

### 2020-01-02 analysis/VOCC/4-vocc-biotic.Rmd
style recent files

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.keep = "all", fig.show = "asis", dpi = 600,
+knitr::opts_chunk$set(
+  fig.keep = "all", fig.show = "asis", dpi = 600,
   fig.width = 6, fig.height = 5,
@@
   age <- "immature"
-  d_all <-readRDS(paste0("data/", spp,
-  "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-FALSE.rds")) %>% filter(year > 2007)
-
+  d_all <- readRDS(paste0(
+    "data/", spp,
+    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-FALSE.rds"
+  )) %>% filter(year > 2007)
 } else {
@@
   try({
-        d_all <-readRDS(paste0("data/", spp,
-        "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-FALSE.rds")) %>% filter(year > 2007) 
+    d_all <- readRDS(paste0(
+      "data/", spp,
+      "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-FALSE.rds"
+    )) %>% filter(year > 2007)
   })
   try({
-        d_all <-readRDS(paste0("data/", spp,
-        "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds")) %>% filter(year > 2007) 
+    d_all <- readRDS(paste0(
+      "data/", spp,
+      "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds"
+    )) %>% filter(year > 2007)
   })
-
 }
 
-nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select (X, Y, year, temp, do_est)
+nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
 
-d_all <- left_join(d_all, nd_all, by=c("X","Y", "year"))
+d_all <- left_join(d_all, nd_all, by = c("X", "Y", "year"))
 
@@
 
-if (multiyear){
- #unique(d_all$year)
+if (multiyear) {
+  # unique(d_all$year)
   d <- d_all %>% filter(ssid %in% model_ssid)
@@
   years_sampled <- length(unique(d$year))
- year_range <- (years_sampled -1)*2
+  year_range <- (years_sampled - 1) * 2
   start_year <- 2007
- if (years_sampled==5) indices <- c(1, 2, 2, 2, 2)
- if (years_sampled==6) indices <- c(1, 2, 2, 2, 2, 2)
-} else{ 
-d <- d_all %>% filter(year >= start_year) %>% filter(year <= start_year+3)
+  if (years_sampled == 5) indices <- c(1, 2, 2, 2, 2)
+  if (years_sampled == 6) indices <- c(1, 2, 2, 2, 2, 2)
+} else {
+  d <- d_all %>%
+    filter(year >= start_year) %>%
+    filter(year <= start_year + 3)
   unique(d$year)
-indices <- c(1,2)
+  indices <- c(1, 2)
   year_range <- 2
@@
 ```{r}
-#d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
-d <- d %>% filter(do_est > 0.28) %>% filter(do_est < 7.06) # 0.005 and 0.995
-#d <- d %>% filter(temp > 2.61) %>% filter(temp < 14.31) # full range
-d <- d %>% filter(temp > 3.07) %>% filter(temp < 11.3) # 0.005 and 0.995
+# d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
+d <- d %>%
+  filter(do_est > 0.28) %>%
+  filter(do_est < 7.06) # 0.005 and 0.995
+# d <- d %>% filter(temp > 2.61) %>% filter(temp < 14.31) # full range
+d <- d %>%
+  filter(temp > 3.07) %>%
+  filter(temp < 11.3) # 0.005 and 0.995
 ```
@@
     velocity = velocity
-  ) %>% dplyr::mutate(X = x, Y = y) %>%
+  ) %>%
+  dplyr::mutate(X = x, Y = y) %>%
   gfplot:::utm2ll(., utm_zone = 9)
@@
 grad0 <- plot_vocc(gf0,
-  #theme_black = TRUE,
+  # theme_black = TRUE,
   coast = TRUE,
   vec_aes = NULL,
-  #grad_vec_aes = "velocity",
-  #vec_lwd_range = c(0.5, 0.5),
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
   NA_label = "NA",
@@
   mutate(
-    trend_per_year = trend ,
+    trend_per_year = trend,
     trend_per_decade = trend_per_year * 10,
@@
   vec_aes = NULL,
-  #grad_vec_aes = "velocity",
-  #vec_lwd_range = c(0.5, 0.5),
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
   NA_label = "NA",
@@
 ```{r}
-#indices <- c(1,2,3,4,5)
+# indices <- c(1,2,3,4,5)
 
@@
   scale_fac = scale_fac,
-  #indices = indices,
+  # indices = indices,
   divisor = 1,
@@
   mutate(
-    CV_formula = sqrt(exp(cv^2)-1),
-    CV = sd/exp(mean), 
-    mean_biomass = exp(mean)*10000,
+    CV_formula = sqrt(exp(cv^2) - 1),
+    CV = sd / exp(mean),
+    mean_biomass = exp(mean) * 10000,
     trend_per_year = exp(trend),
@@
   axis_lables = FALSE,
-  transform_col =  sqrt#
+  transform_col = sqrt #
 ) + labs(
@@
 grad4
-
 ```
@@
   raster_alpha = 1,
-  #raster_limits = c(0, 10),
+  # raster_limits = c(0, 10),
   na_colour = "black",
@@
 grad5
-
 ```
@@
   raster_alpha = 1,
-  #raster_limits = c(0, 7),
+  # raster_limits = c(0, 7),
   na_colour = "black",
@@
 
-biotic_values <- gf3t %>% rename(
+biotic_values <- gf3t %>%
+  rename(
     biotic_vel = velocity,
@@
     sd_est = cv,
-    biotic_grad = gradient) %>% 
-  mutate(sd_biomass = sd*10000,
-    mean_biomass = exp(mean)*10000) %>%
+    biotic_grad = gradient
+  ) %>%
+  mutate(
+    sd_biomass = sd * 10000,
+    mean_biomass = exp(mean) * 10000
+  ) %>%
   select(x, y, biotic_vel, biotic_trend, biotic_grad, biotic_CV, sd_est, sd_biomass, mean_biomass) %>%
@@
 
-DO_values <- gf0 %>% rename( 
+DO_values <- gf0 %>%
+  rename(
     DO_vel = velocity,
     DO_trend = trend_per_year,
-    DO_grad = gradient) %>% 
+    DO_grad = gradient
+  ) %>%
   select(x, y, DO_vel, DO_trend, DO_grad)
 
-temp_values <- gf2 %>% rename( 
+temp_values <- gf2 %>%
+  rename(
     temp_vel = velocity,
     temp_trend = trend_per_year,
-    temp_grad = gradient) %>% 
+    temp_grad = gradient
+  ) %>%
   select(x, y, temp_vel, temp_trend, temp_grad)
@@
 vocc <- left_join(biotic_values, DO_values) # add DO
-vocc  <- left_join(vocc , temp_values) # add temp
+vocc <- left_join(vocc, temp_values) # add temp
 vocc <- mutate(vocc, species = !!species, method = "gradient")
@@
 ))
-
 ```

### 2020-01-02 analysis/VOCC/4-pool-biotic-velocities.R

@@ -2,13 +2,13 @@ getwd()
 setwd(here::here("/analysis/VOCC/data"))
 library("dplyr")
 
-#covs <- "-fixed"
+# covs <- "-fixed"
 # mydir = paste0( "biotic_test1")
-mydir = paste0( "biotic_mature")
-#mydir = paste0( "biotic_immature")
+mydir <- paste0("biotic_mature")
+# mydir = paste0( "biotic_immature")
 
 
-myfiles = list.files(path=mydir, pattern="*.csv", full.names=TRUE)
+myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
 myfiles
 biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% select(-X, -start_year)
 glimpse(biotic)
@@ -16,7 +16,7 @@ glimpse(biotic)
 # saveRDS(biotic, file = paste0(
 #   "multi-spp-biotic-vocc.rds"
 # ))
-# 
+#
 saveRDS(biotic, file = paste0(
   "multi-spp-biotic-vocc-mature.rds"
 ))

### 2020-01-02 analysis/VOCC/render-biotic-variability.R

@@ -1,12 +1,12 @@
 getwd()
 setwd(here::here("/analysis/VOCC"))
-env <- new.env() #parent = baseenv()
+env <- new.env() # parent = baseenv()
 
 list_species <- c(
   "North Pacific Spiny Dogfish",
   "Spotted Ratfish",
   # "Pacific Tomcod",
-   "Walleye Pollock",
+  "Walleye Pollock",
   # "Pacific Cod",
   "Sablefish",
   "Lingcod",
@@ -16,105 +16,112 @@ list_species <- c(
   "Yellowmouth Rockfish",
   # "Harlequin Rockfish",
   # "Bocaccio", # winter-birthing, overfished
-   "Canary Rockfish", # schooling, winter-birthing
+  "Canary Rockfish", # schooling, winter-birthing
   # "Copper Rockfish", # small sample
-   "Darkblotched Rockfish",
-   "Greenstriped Rockfish",
+  "Darkblotched Rockfish",
+  "Greenstriped Rockfish",
   # "Pacific Ocean Perch", # schooling
-   "Quillback Rockfish",
+  "Quillback Rockfish",
   # "Redbanded Rockfish",
   # "Rougheye/Blackspotted Rockfish Complex",
-   "Sharpchin Rockfish",
-   "Shortbelly Rockfish", # small sample
-   "Silvergray Rockfish",
-   "Splitnose Rockfish",
-   "Widow Rockfish", # schooling
-   "Yellowtail Rockfish", # schooling
-   "Yelloweye Rockfish", # summer-birthing, overfished,
-   "Longspine Thornyhead",
-   "Shortspine Thornyhead",
-   "Arrowtooth Flounder",
-   "Rex Sole",
+  "Sharpchin Rockfish",
+  "Shortbelly Rockfish", # small sample
+  "Silvergray Rockfish",
+  "Splitnose Rockfish",
+  "Widow Rockfish", # schooling
+  "Yellowtail Rockfish", # schooling
+  "Yelloweye Rockfish", # summer-birthing, overfished,
+  "Longspine Thornyhead",
+  "Shortspine Thornyhead",
+  "Arrowtooth Flounder",
+  "Rex Sole",
   # "Petrale Sole",
   # "English Sole",
-   "Dover Sole",
-   "Southern Rock Sole",
-   "Flathead Sole",
-   "Curlfin Sole",
-   "Sand Sole",
[…]

### 2020-01-02 analysis/VOCC/vocc-regression.R

@@ -6,9 +6,9 @@ setwd(here::here("analysis", "VOCC"))
 compile("vocc_regression.cpp")
 dyn.load(dynlib("vocc_regression"))
 
-#model <- "multi-spp-biotic-vocc"
+# model <- "multi-spp-biotic-vocc"
 model <- "multi-spp-biotic-vocc-mature"
-#model <- "multi-spp-biotic-vocc-immature"
+# model <- "multi-spp-biotic-vocc-immature"
 
 d <- readRDS(paste0("data/", model, ".rds"))
 d <- na.omit(d) %>% as_tibble()
@@ -16,213 +16,211 @@ d <- na.omit(d) %>% as_tibble()
 stats <- readRDS(paste0("data/life-history-stats.rds"))
 stats$rockfish <- if_else(stats$group == "ROCKFISH", "ROCKFISH", "OTHER")
 
-d <- left_join(d, stats) %>% filter(species!="Longspine Thornyhead")
+d <- left_join(d, stats) %>% filter(species != "Longspine Thornyhead")
 
-#unique(d$species)
+# unique(d$species)
 
 vocc_regression <- function(y_i, X_ij, knots = 150) {
 
- # y_i <- d$biotic_vel
- # X_ij <- model.matrix(~scale(temp_vel), data = d)
- #  
- # hist(X_ij[,2], breaks = 100)
- # range(X_ij[,2])
- # 
- # hist(y_i, breaks = 100)
- # range(y_i)
- #
- # ggplot(d, aes(x, y, colour = temp_vel)) + geom_point() +
- #   facet_wrap(~species) +
- #   scale_color_viridis_c()
- # 
- # ggplot(d, aes(x, y, colour = biotic_vel)) + geom_point() +
- #   facet_wrap(~species) #+ scale_color_viridis_c()
- #
-# -----------------------------
+  # y_i <- d$biotic_vel
+  # X_ij <- model.matrix(~scale(temp_vel), data = d)
+  #
+  # hist(X_ij[,2], breaks = 100)
+  # range(X_ij[,2])
+  #
+  # hist(y_i, breaks = 100)
+  # range(y_i)
+  #
+  # ggplot(d, aes(x, y, colour = temp_vel)) + geom_point() +
+  #   facet_wrap(~species) +
+  #   scale_color_viridis_c()
+  #
+  # ggplot(d, aes(x, y, colour = biotic_vel)) + geom_point() +
[…]

### 2020-01-02 analysis/VOCC/1-split-biomass-by-maturity.Rmd
tweak split_by_maturity and maturity ogive to output mean length at 50% mature for model with year as a random effect

@@
 params: 
-    species: "Redbanded Rockfish"
+    species: "Longnose Skate"
+    #species:"Redbanded Rockfish"
+    #species: "Pacific Ocean Perch"
     region: "All synoptic surveys"
-    #region: "Both odd year surveys"
+    # region: "Both odd year surveys"
+    split: TRUE
 ---
@@
 ```{r global_options, include=FALSE}
+species <- params$species
+region <- params$region
+split <- params$split
+
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+
 knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
@@
 
-```{r params}
-species <- params$species
-region <- params$region
-
-spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+```{r}
+paste("region =", region)
 
@@
     years = years,
+    year_re = TRUE,
     cutoff_quantile = c(.9995),
@@
   }
+
+  if (split) {
   maturity$data <- filter(maturity$data, ssid == model_ssid) 
   # %>% select(-adult_density, -imm_density) # if maturity model is deemed unuseful
+  if (!is.null(maturity$maturity)) {
+  saveRDS(maturity$model, file = paste0("data/", spp, 
+    "/maturity-ogive-", spp, "-", ssid_string, ".rds"
+    ))
+  }
+  } else {
+
+  maturity$data <- filter(maturity$data, ssid == model_ssid) %>% 
+    select(-adult_density, -imm_density) # if maturity model is deemed unuseful
+
+  }
 

### 2020-01-02 R/5-life-history-stats.R

@@ -43,7 +43,7 @@ species <- c(
 )
 
 # species <- c("Arrowtooth Flounder")
-# species <- c("Widow Rockfish")
+# species <- c("Pacific Ocean Perch")
 # spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
 
 life_history <- purrr::map_dfr(species, function(x) {
@@ -59,24 +59,57 @@ life_history <- purrr::map_dfr(species, function(x) {
   bath <- readRDS("data/bathymetry-data")
   depth <- bath$data %>% select(fishing_event_id, depth)
   fish <- left_join(fish, depth)
+  
+  rm(maturity)
+  try({maturity <- readRDS(paste0("data/", spp, 
+    "/maturity-ogive-", spp, "-1n3n4n16.rds"
+  ))})
+  
+  if(exists("maturity")) {
+  length_50_mat_m <- maturity$mat_perc$mean$m.mean.p0.5
+  length_50_mat_f <- maturity$mat_perc$mean$f.mean.p0.5
+  mat_m <- filter(fish, sex == 1) %>% filter(length > length_50_mat_m)
+  mat_f <- filter(fish, sex == 2) %>% filter(length > length_50_mat_f)
+  
+  imm_m <- filter(fish, sex == 1) %>% filter(length < length_50_mat_m)
+  imm_f <- filter(fish, sex == 2) %>% filter(length < length_50_mat_f)
+  
+  large <- rbind(mat_m, mat_f)
+  small <- rbind(imm_m, imm_f)
+  large_threshold <- NA
+  small_threshold <- NA
+  
+  } else {
+    
+  length_50_mat_m <- NA
+  length_50_mat_f <- NA
   large_threshold <- quantile(fish$length, 0.75, na.rm = TRUE)
   small_threshold <- quantile(fish$length, 0.25, na.rm = TRUE)
   large <- filter(fish, length > large_threshold)
   small <- filter(fish, length < small_threshold)
-
+  }
+  
+  large_depth <- mean(large$depth, na.rm = TRUE)
+  small_depth <- mean(small$depth, na.rm = TRUE)
+  
   list(
     species = x, group = group[1],
-    large_depth = mean(large$depth, na.rm = TRUE),
-    small_depth = mean(small$depth, na.rm = TRUE),
+    length_50_mat_m = length_50_mat_m,
+    length_50_mat_f = length_50_mat_f,
+    mat_depth = round(large_depth),
+    imm_depth = round(small_depth),
[…]

### 2020-01-02 R/maturity.R

@@ -527,6 +527,11 @@ extract_maturity_perc_re <- function(betas, random_intercepts) {
       f.p0.5 = f.p0.5, f.p0.95 = f.p0.95, f.p0.05 = f.p0.05
     )
   }
+  m.mean.p0.5 <- logit_perc(a = b[[1]], b = b[[2]], perc = 0.5)
+  f.mean.p0.5 <- logit_perc(a = b[[1]] + b[[3]],
+    b = b[[2]] + b[[4]], perc = 0.5)
+  #nrow(re) + 1
+  out[["mean"]] <- list(m.mean.p0.5 = m.mean.p0.5, f.mean.p0.5 = f.mean.p0.5)
   out
 }
 

### 2020-01-03 analysis/VOCC/1-split-biomass-by-maturity.Rmd
tweaks to saving of maturity splits

@@
     # region: "Both odd year surveys"
-    split: TRUE
+    split: FALSE
 ---
@@
 ```{r}
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
 paste("region =", region)
@@
   maturity$data <- filter(maturity$data, ssid == model_ssid)
-  # %>% select(-adult_density, -imm_density) # if maturity model is deemed unuseful
+
   if (!is.null(maturity$maturity)) {
@@
   } else {
-
+  if (is.null(maturity$maturity)) {
+    maturity$data <- filter(maturity$data, ssid == model_ssid) 
+  } else { # if maturity model is deemed unuseful
   maturity$data <- filter(maturity$data, ssid == model_ssid) %>% 
-    select(-adult_density, -imm_density) # if maturity model is deemed unuseful
-
+    select(-adult_density, -imm_density) 
+  }
   }
@@
 }
+
+positive_sets <- filter(survey_sets, density_kgpm2 != 0)
+prop_pos <- round(nrow(positive_sets)/nrow(survey_sets), digits = 3)
+
+paste("proportion of positive sets =", prop_pos)
 ```
@@
 ```{r}
-if (exists("maturity")) {
   png(
-    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), 
+    file = paste0("figs/_maturity_ogives/maturity-", prop_pos,"-", spp, "-", ssid_string, ".png"), 
     res = 600,
@@
   )
+  if (exists("maturity")) {
   print(maturity$maturity)
+  } else {
+  ggplot(fish, aes(length)) + geom_histogram(bins=50) + facet_grid(~sex)  
+    }
+
   dev.off()
-}
 

### 2020-01-03 R/5-life-history-stats.R

@@ -39,14 +39,22 @@ species <- c(
   "Widow Rockfish",
   "Yelloweye Rockfish",
   "Yellowmouth Rockfish",
-  "Yellowtail Rockfish"
+  "Yellowtail Rockfish",
+  "Big Skate",
+  "Longnose Skate",
+  "Spotted Ratfish",
+  "Bocaccio",
+  "Redstripe Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Shortraker Rockfish"
 )
 
-# species <- c("Arrowtooth Flounder")
-# species <- c("Pacific Ocean Perch")
+# species <- c("Spotted Ratfish")
+# species <- c("Shortraker Rockfish")
 # spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
 
 life_history <- purrr::map_dfr(species, function(x) {
+  
   spp <- gsub(" ", "-", gsub("\\/", "-", tolower(x)))
   fish <- readRDS(paste0("raw/bio-data-", spp, ""))
   group <- unique(fish$maturity_convention_desc)
@@ -61,13 +69,21 @@ life_history <- purrr::map_dfr(species, function(x) {
   fish <- left_join(fish, depth)
   
   rm(maturity)
-  try({maturity <- readRDS(paste0("data/", spp, 
-    "/maturity-ogive-", spp, "-1n3n4n16.rds"
-  ))})
+  try({
+    maturity <- readRDS(paste0("data/", spp, 
+    "/maturity-ogive-", spp, "-1n3n4n16.rds"))
+  })
   
   if(exists("maturity")) {
-  length_50_mat_m <- maturity$mat_perc$mean$m.mean.p0.5
-  length_50_mat_f <- maturity$mat_perc$mean$f.mean.p0.5
+    
+    if(maturity$year_re) {
+    length_50_mat_m <- maturity$mat_perc$mean$m.mean.p0.5
+    length_50_mat_f <- maturity$mat_perc$mean$f.mean.p0.5
+    } else {
+    length_50_mat_m <- maturity$mat_perc$m.p0.5
+    length_50_mat_f <- maturity$mat_perc$f.p0.5
+    }
+    
   mat_m <- filter(fish, sex == 1) %>% filter(length > length_50_mat_m)
   mat_f <- filter(fish, sex == 2) %>% filter(length > length_50_mat_f)
   
[…]

### 2020-01-03 analysis/VOCC/render-maturity.R

@@ -2,38 +2,40 @@ getwd()
 setwd(here::here("/analysis/VOCC"))
 env <- new.env() #parent = baseenv()
 
+# SPECIES with maturity
 list_species <- c(
-  # "Walleye Pollock",
-  # "Pacific Cod",
-  # "Sablefish",
-  # "Lingcod",
+  "Bocaccio",
+  "Redstripe Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Shortraker Rockfish",
+  "Walleye Pollock",
+  "Pacific Cod",
+  "Sablefish",
+  "Lingcod",
   # "Pacific Hake",
-  # "North Pacific Spiny Dogfish",
-  # "Spotted Ratfish",
   # "Pacific Tomcod",
   # "Rosethorn Rockfish",
-  # "Redstripe Rockfish",
-  # "Yellowmouth Rockfish",
+  "Redstripe Rockfish",
+  "Yellowmouth Rockfish",
   # "Harlequin Rockfish"
-   # "Bocaccio", # winter-birthing, overfished
-  # "Canary Rockfish", # schooling, winter-birthing
-  # "Copper Rockfish", # small sample
-  # "Darkblotched Rockfish",
-  # "Greenstriped Rockfish",
-  # "Pacific Ocean Perch", # schooling
-  # "Quillback Rockfish",
-  # "Redbanded Rockfish",
-  # "Rougheye/Blackspotted Rockfish Complex",
-  # "Sharpchin Rockfish",
-  # "Shortbelly Rockfish", # small sample
-  # "Silvergray Rockfish",
-  # "Splitnose Rockfish",
-  # "Widow Rockfish", # schooling
-  # "Yellowtail Rockfish", # schooling
-  # "Yelloweye Rockfish", # summer-birthing, overfished,
-  # "Longspine Thornyhead",
-  # "Shortspine Thornyhead",
-  #"Arrowtooth Flounder",
+  "Bocaccio", # winter-birthing, overfished
+  "Canary Rockfish", # schooling, winter-birthing
+  "Darkblotched Rockfish",
+  "Greenstriped Rockfish",
+  "Pacific Ocean Perch", # schooling
+  "Quillback Rockfish",
+  "Redbanded Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Sharpchin Rockfish",
+  "Shortbelly Rockfish", # small sample
+  "Silvergray Rockfish",
[…]

### 2020-01-06 analysis/VOCC/2-biomass-depth-only-model.Rmd
update checks on biomass models

@@
     #species: "Petrale Sole"
-    species: "Pacific Halibut"
+    #species: "Widow Rockfish"
+    species: "Bocaccio"
     region: "All synoptic surveys"
@@
     update_model: FALSE
+    update_predictions: FALSE
+    update_model_check: FALSE
 ---
@@
 ))
+
+if(nrow(biomass)<4000) {stop("Need to recalculate split by maturity!")}
+
 covars <- readRDS("data/event-covariates.rds")
@@
       family = tweedie(link = "log"),
-      # ar1_fields = TRUE,
+      #ar1_fields = TRUE,
       include_spatial = TRUE,
@@
 
-    imm_formula <- as.formula(paste(
+   try({ imm_formula <- as.formula(paste(
       "imm_density ~ 0 + as.factor(year)", covariates, ""
@@
       family = tweedie(link = "log"),
-      # ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
-      reml = TRUE,
+      ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
+      # reml = TRUE,
       include_spatial = TRUE,
@@
     saveRDS(imm_biomass,
-      file = paste0(
-        "data/", spp,
-        "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-      )
-    )
+      file = paste0("data/", spp, "/mod-imm-biomass-", 
+        spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+   })
+
   } else {
@@
 Save predictions for full spatial grid
-
 ```{r}
@@
 rm(predictions)
-
+if(params$update_predictions) {
 if (any(names(data) == "adult_density")) {
@@
 }
+}
 ```
@@
 ```{r}
+
+if(params$update_model_check) {
 try({
@@
 ))
+}
 ```
@@
 rm(total_biomass)
+rm(depth_plots)
+#max_depth_found <- 800
 
@@
 for (i in seq_len(length(depth_model_list))) {
-  if (depth_model_list[[i]]$model$convergence == 0) {
+  #if (depth_model_list[[i]]$model$convergence == 0) {
     d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
@@
     }
-  } else {
-    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
-  }
+  # } else {
+  #   depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
+  # }
 }
@@
 ```
+
+```{r}
+sdmTMB:::get_convergence_diagnostics(depth_model_list[[1]])
+```

### 2020-01-06 analysis/VOCC/4-vocc-biotic.Rmd
update velocity and life history data

@@
     mean_biomass = exp(mean) * 10000,
-    trend_per_year = exp(trend),
-    trend_per_decade = trend_per_year * 10,
+    #trend_per_year = exp(trend),
+    #trend_per_decade = trend_per_year * 10,
     velocity = velocity
@@
 
-bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.0001
+bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.05
 s <- sort(gf3$mean_biomass)
@@
   NA_label = "NA",
-  fill_col = "trend_per_year",
+  fill_col = "trend",
   fill_label = "biomass trend",
@@
     biotic_vel = velocity,
-    biotic_trend = trend_per_year,
+    biotic_trend = trend,
     biotic_CV = CV,
@@
     DO_trend = trend_per_year,
-    DO_grad = gradient
+    DO_grad = gradient,
+    mean_DO = mean
   ) %>%
-  select(x, y, DO_vel, DO_trend, DO_grad)
+  select(x, y, DO_vel, DO_trend, DO_grad, mean_DO)
 
@@
     temp_trend = trend_per_year,
-    temp_grad = gradient
+    temp_grad = gradient,
+    mean_temp = mean
   ) %>%
-  select(x, y, temp_vel, temp_trend, temp_grad)
+  select(x, y, temp_vel, temp_trend, temp_grad, mean_temp)
 
@@
 write.csv(vocc, file = paste0(
-  "data/biotic_test/biotic-vocc-", spp, "-", age, covs, "-", ssid_string, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_new_biotic/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))

### 2020-01-06 analysis/VOCC/4-pool-biotic-velocities.R

@@ -4,8 +4,8 @@ library("dplyr")
 
 # covs <- "-fixed"
 # mydir = paste0( "biotic_test1")
-mydir <- paste0("biotic_mature")
-# mydir = paste0( "biotic_immature")
+mydir <- paste0("_biotic/mat")
+mydir = paste0( "_biotic/imm")
 
 
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
@@ -13,10 +13,6 @@ myfiles
 biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% select(-X, -start_year)
 glimpse(biotic)
 
-# saveRDS(biotic, file = paste0(
-#   "multi-spp-biotic-vocc.rds"
-# ))
-#
 saveRDS(biotic, file = paste0(
   "multi-spp-biotic-vocc-mature.rds"
 ))

### 2020-01-06 analysis/VOCC/5-life-history-stats.R

@@ -50,7 +50,8 @@ species <- c(
 )
 
 # species <- c("Spotted Ratfish")
-
+# species <- "Quillback Rockfish"
+species <- "North Pacific Spiny Dogfish"
 # species <- c("Bocaccio")
 # species <- c("Shortraker Rockfish")
 # spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
@@ -69,16 +70,15 @@ life_history <- purrr::map_dfr(species, function(x) {
   bath <- readRDS("data/bathymetry-data")
   depth <- bath$data %>% select(fishing_event_id, depth)
   fish <- left_join(fish, depth)
-   plot(length~depth, data=fish)
   
   rm(maturity)
   try({
     maturity <- readRDS(paste0("data/", spp, 
     "/maturity-ogive-", spp, "-1n3n4n16.rds"))
   })
-  
+  print(spp)
   if(exists("maturity")) {
-    
+ 
     if(maturity$year_re) {
     length_50_mat_m <- maturity$mat_perc$mean$m.mean.p0.5
     length_50_mat_f <- maturity$mat_perc$mean$f.mean.p0.5
@@ -97,7 +97,8 @@ life_history <- purrr::map_dfr(species, function(x) {
   small <- rbind(imm_m, imm_f)
   large_threshold <- NA
   small_threshold <- NA
-  
+#  })
+    
   } else {
     
   length_50_mat_m <- NA
@@ -108,17 +109,19 @@ life_history <- purrr::map_dfr(species, function(x) {
   small <- filter(fish, length < 20)#small_threshold)
   }
   
-  ## FOR EXPLORING SPECIAL CASES
-  fish_f <- filter(fish, sex == 2)
-  fish_m <- filter(fish, sex == 1)
-  hist(fish_m$length, breaks = 50)
-  hist(fish_f$length, breaks = 50)
-  hist(fish$weight, breaks = 50)
-  plot(weight~length, data=fish)
+  # ## FOR EXPLORING SPECIAL CASES
+  # #x <- species
+  # plot(length~depth, data = fish, main = x)
+  # fish_f <- filter(fish, sex == 2)
+  # fish_m <- filter(fish, sex == 1)
+  # hist(fish_m$length, breaks = 50, main = x)
+  # hist(fish_f$length, breaks = 50, main = x)
+  # hist(fish$weight, breaks = 50, main = x)
+  # plot(weight~length, data=fish, main = x)
[…]

### 2020-01-06 analysis/VOCC/render-biotic-variability.R

@@ -3,6 +3,9 @@ setwd(here::here("/analysis/VOCC"))
 env <- new.env() # parent = baseenv()
 
 list_species <- c(
+  "Big Skate",
+  "Longnose Skate",
+  "Spotted Ratfish",
   "North Pacific Spiny Dogfish",
   "Spotted Ratfish",
   # "Pacific Tomcod",
@@ -23,17 +26,15 @@ list_species <- c(
   # "Pacific Ocean Perch", # schooling
   "Quillback Rockfish",
   # "Redbanded Rockfish",
-  # "Rougheye/Blackspotted Rockfish Complex",
   "Sharpchin Rockfish",
   "Shortbelly Rockfish", # small sample
   "Silvergray Rockfish",
   "Splitnose Rockfish",
   "Widow Rockfish", # schooling
   "Yellowtail Rockfish", # schooling
-  "Yelloweye Rockfish", # summer-birthing, overfished,
-  "Longspine Thornyhead",
+   "Longspine Thornyhead",
   "Shortspine Thornyhead",
-  "Arrowtooth Flounder",
+  #"Arrowtooth Flounder",
   "Rex Sole",
   # "Petrale Sole",
   # "English Sole",
@@ -44,29 +45,44 @@ list_species <- c(
   "Sand Sole",
   # "Slender Sole",
   # "Pacific Sanddab",
-  "Pacific Halibut"
+  # "Pacific Halibut"
 )
 
 ### SUBSETS OF SPECIES
-#  list_species <- c(
+ list_species <- c(
+  "Pacific Ocean Perch", # schooling
+  "Redbanded Rockfish",
+  "Pacific Cod",
+  "Pacific Halibut"
+)
+
 #   # "Arrowtooth Flounder"
 #   # "Petrale Sole",
 #   # "English Sole",
-#   # "Pacific Ocean Perch", # schooling
-#   # "Redbanded Rockfish",
-#     "Bocaccio"
-#   # "Pacific Cod"
-#   # "Pacific Halibut"
-#   # "Yelloweye Rockfish"
-# )
 
+ 
+ list_species <- c(
+   # "Redstripe Rockfish",
+   # "Rougheye/Blackspotted Rockfish Complex",
+   # "Widow Rockfish",
+   # "Quillback Rockfish",
+   "Bocaccio",
+   "Shortraker Rockfish",
+   "Yelloweye Rockfish"
+ )
+ 
[…]

### 2020-01-06 analysis/VOCC/render-maturity.R

@@ -4,10 +4,9 @@ env <- new.env() #parent = baseenv()
 
 # SPECIES with maturity
 list_species <- c(
-  "Bocaccio",
   "Redstripe Rockfish",
   "Rougheye/Blackspotted Rockfish Complex",
-  "Shortraker Rockfish",
+  #"Shortraker Rockfish", # too few imm so ogive unreliable
   "Walleye Pollock",
   "Pacific Cod",
   "Sablefish",
@@ -18,21 +17,20 @@ list_species <- c(
   "Redstripe Rockfish",
   "Yellowmouth Rockfish",
   # "Harlequin Rockfish"
-  "Bocaccio", # winter-birthing, overfished
   "Canary Rockfish", # schooling, winter-birthing
   "Darkblotched Rockfish",
   "Greenstriped Rockfish",
   "Pacific Ocean Perch", # schooling
-  "Quillback Rockfish",
+  #"Quillback Rockfish",
   "Redbanded Rockfish",
   "Rougheye/Blackspotted Rockfish Complex",
   "Sharpchin Rockfish",
   "Shortbelly Rockfish", # small sample
   "Silvergray Rockfish",
   "Splitnose Rockfish",
-  "Widow Rockfish", # schooling
+  #"Widow Rockfish", # schooling
   "Yellowtail Rockfish", # schooling
-  "Yelloweye Rockfish", # summer-birthing, overfished,
+  #"Yelloweye Rockfish", # summer-birthing, overfished,
   "Longspine Thornyhead",
   "Shortspine Thornyhead",
   "Arrowtooth Flounder",
@@ -49,16 +47,25 @@ list_species <- c(
 )
 
 
-# # SPECIES WITHOUT maturity
+# # SPECIES to try setting maturity threshold to 0.05
 # list_species <- c(
-#   "North Pacific Spiny Dogfish",
+#   "Widow Rockfish",
+#   "Quillback Rockfish",
+#   "Bocaccio",
+#   "Shortraker Rockfish",   
+#   "Yelloweye Rockfish"
+# )
+
+# # # SPECIES WITHOUT maturity
+# list_species <- c(
 #   "Big Skate",
 #   "Longnose Skate",
 #   "Spotted Ratfish",
 #   "Pacific Halibut"
 # )
 
-
+## NEEDS SPECIAL SETTINGS
+list_species <- c("North Pacific Spiny Dogfish","Longspine Thornyhead","Sand Sole")
 
[…]

### 2020-01-10 analysis/VOCC/4-vocc-biotic.Rmd
merge fishing effort with voccs

@@
 nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
+fishing <- readRDS("data/fishing-effort-grid.rds")
 
 d_all <- left_join(d_all, nd_all, by = c("X", "Y", "year"))
+d_all <- left_join(d_all, fishing, by = c("X", "Y", "year"))
+d_all$log_effort[is.na(d_all$log_effort)] <- -2.3
 
 glimpse(d_all)
+#hist(d_all$log_effort)
 ```
@@
   filter(temp < 11.3) # 0.005 and 0.995
+```
+
+Calculate mean and trend in log fishing pressure
+```{r }
+gf <- vocc_gradient_calc(d, "log_effort",
+  scale_fac = scale_fac,
+  # all layers with max indice will be averaged for spatial gradient
+  # if default (NULL) will use all years
+  # indices = indices,
+  divisor = 1,
+  quantile_cutoff = 0.05
+)
+
+gf <- gf %>%
+  mutate(
+    fishing_trend = trend,
+    mean_effort = exp(mean)
+  ) %>%
+  dplyr::mutate(X = x, Y = y) %>%
+  gfplot:::utm2ll(., utm_zone = 9)
+
+effort <- plot_vocc(gf,
+  # theme_black = TRUE,
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  #NA_label = "NA",
+  fill_col = "mean_effort",
+  fill_label = "mean effort",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+   na_colour = "white",
+  # high_fill = "Steel Blue 4",
+  # low_fill = "Red 3",
+  # white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = fourth_root_power#
+) + labs(
+  title = "Fishing presssure",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+
+effort
+
+fish_t <- plot_vocc(gf,
+  # theme_black = TRUE,
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  #NA_label = "NA",
+  fill_col = "fishing_trend",
+  fill_label = "trend in fishing effort",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+
+  # high_fill = "Steel Blue 4",
+  # low_fill = "Red 3",
+  # white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = sqrt # no_trans #
+) + labs(
+  title = "Fishing presssure",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+
+fish_t
 ```
@@
 
-bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.05
+bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.1
 s <- sort(gf3$mean_biomass)
@@
 
+fished <- gf %>%
+  select(x, y, fishing_trend, mean_effort)
+
 DO_values <- gf0 %>%
@@
 
-vocc <- left_join(biotic_values, DO_values) # add DO
+vocc <- left_join(biotic_values, fished) # add fishing effort
+vocc <- left_join(vocc, DO_values) # add DO
 vocc <- left_join(vocc, temp_values) # add temp
@@
 write.csv(vocc, file = paste0(
-  "data/_new_biotic/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_new_biotic/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
 ))

### 2020-01-10 analysis/VOCC/4-pool-biotic-velocities.R

@@ -4,8 +4,8 @@ library("dplyr")
 
 # covs <- "-fixed"
 # mydir = paste0( "biotic_test1")
-mydir <- paste0("_biotic/mat")
-mydir = paste0( "_biotic/imm")
+mydir <- paste0("_new_biotic/mat")
+mydir = paste0( "_new_biotic/imm")
 
 
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
@@ -14,9 +14,9 @@ biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% select(-X, -start_year)
 glimpse(biotic)
 
 saveRDS(biotic, file = paste0(
-  "multi-spp-biotic-vocc-mature.rds"
+  "multi-spp-biotic-vocc-mature-with-fished.rds"
 ))
 
 saveRDS(biotic, file = paste0(
-  "multi-spp-biotic-vocc-immature.rds"
+  "multi-spp-biotic-vocc-immature-with-fished.rds"
 ))

### 2020-01-10 analysis/VOCC/render-biotic-variability.R

@@ -62,10 +62,10 @@ list_species <- c(
 
  
  list_species <- c(
-   # "Redstripe Rockfish",
-   # "Rougheye/Blackspotted Rockfish Complex",
-   # "Widow Rockfish",
-   # "Quillback Rockfish",
+   "Redstripe Rockfish",
+   "Rougheye/Blackspotted Rockfish Complex",
+   "Widow Rockfish",
+   "Quillback Rockfish",
    "Bocaccio",
    "Shortraker Rockfish",
    "Yelloweye Rockfish"
@@ -145,17 +145,14 @@ list_species <- c(
   "Southern Rock Sole",
   "Flathead Sole",
   "Curlfin Sole",
-  "Sand Sole",
+  # "Sand Sole",
   # "Slender Sole",
   # "Pacific Sanddab",
-  "Pacific Halibut"
-)
-
-list_species <- c(
-  # "Redstripe Rockfish",
-  # "Rougheye/Blackspotted Rockfish Complex",
-  # "Widow Rockfish",
-  # "Quillback Rockfish",
+  "Pacific Halibut",
+  "Redstripe Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Widow Rockfish",
+  "Quillback Rockfish",
   "Bocaccio",
   "Shortraker Rockfish",
   "Yelloweye Rockfish"
@@ -186,7 +183,7 @@ for (r_h in seq_along(list_regions)) {
         output_file = paste0(
           "html/VOCC-plots/biotic-gradients-", spp,
           # "-imm",
-          covs, "-", reg, "-trim5.html"
+          covs, "-", reg, "-trim10.html"
         ),
         envir = env
       )

### 2020-01-13 analysis/VOCC/4-vocc-biotic.Rmd
code for scrambling

@@
     region: "both odd year surveys"
+    #region: "West Coast Haida Gwaii"
     #region: "West Coast Vancouver Island"
     covs: "-tv-depth-only"
+    scrambled_years: FALSE
 ---
@@
 immature <- params$immature
+scrambled_years <- params$scrambled_years
 
@@
 paste("covs =", covs)
+paste("scrambled =", scrambled_years) 
 ```
@@
 }
-
-nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
-fishing <- readRDS("data/fishing-effort-grid.rds")
-
-d_all <- left_join(d_all, nd_all, by = c("X", "Y", "year"))
-d_all <- left_join(d_all, fishing, by = c("X", "Y", "year"))
-d_all$log_effort[is.na(d_all$log_effort)] <- -2.3
-
-glimpse(d_all)
 #hist(d_all$log_effort)
@@
 delta_t_total <- year_range
+
+glimpse (d)
+if (scrambled_years) {
+  d2 <- list()
+
+  set.seed(9999) 
+  # set.seed(12345)
+  # set.seed(98765) 
+  original_years <- unique(d$year)
+  new_years <- sample(original_years)
+  convert <- cbind(original_years,new_years)
+  convert
+  for (i in seq_len(length(original_years))) {
+    old_year <- as.integer(convert[[i,1]])
+    fake_year <- as.integer(convert[[i,2]])
+    d2[[i]] <- filter(d, year == old_year) %>% 
+      mutate(year = fake_year )
+     # browser()
+  }
+  d <- do.call(rbind, d2)
+}
 ```
@@
 ```{r}
+nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
+fishing <- readRDS("data/fishing-effort-grid.rds")
+#fishing <- readRDS("data/fishing-effort-grid-6hr.rds")
+
+d <- left_join(d, nd_all, by = c("X", "Y", "year"))
+d <- left_join(d, fishing, by = c("X", "Y", "year"))
+d$log_effort[is.na(d$log_effort)] <- -2.3
+
+glimpse(d)
+
 # d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
@@
 fish_t
+
+# fished <- gf %>%
+#   select(x, y, fishing_trend, mean_effort) 
+# saveRDS(fished, file = paste0("data/fishing-effort-trend-6hr-", ssid_string, ".rds"))
 ```
@@
 write.csv(vocc, file = paste0(
-  "data/_new_biotic/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
+  "data/_scrambled3/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
 ))

### 2020-01-23 analysis/VOCC/4-vocc-temp.Rmd
calc vocc for all years with temp data

--- 2020-01-13 analysis/VOCC/4-vocc-biotic.Rmd
@@
     immature: FALSE
-    region: "both odd year surveys"
+    #region: "both odd year surveys"
     #region: "West Coast Haida Gwaii"
-    #region: "West Coast Vancouver Island"
+    region: "West Coast Vancouver Island"
     covs: "-tv-depth-only"
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+  start_year <- 2005
 }
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+  start_year <- 2004
 }
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
+  start_year <- 2006
 }
@@
     "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-FALSE.rds"
-  )) %>% filter(year > 2007)
+  )) #%>% filter(year > 2007)
 } else {
@@
       "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-FALSE.rds"
-    )) %>% filter(year > 2007)
+    )) #%>% filter(year > 2007)
   })
@@
       "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds"
-    )) %>% filter(year > 2007)
+    )) #%>% filter(year > 2007)
   })
@@
   year_range <- (years_sampled - 1) * 2
-  start_year <- 2007
-  if (years_sampled == 5) indices <- c(1, 2, 2, 2, 2)
-  if (years_sampled == 6) indices <- c(1, 2, 2, 2, 2, 2)
+
+  if (years_sampled == 7) indices <- c(1, 1, 1, 2, 2, 2)
+  if (years_sampled == 8) indices <- c(1, 1, 1, 1, 2, 2, 2,2)
 } else {
@@
   d2 <- list()
-
-  set.seed(9999) 
+  # set.seed(9999) 
   # set.seed(12345)
@@
   original_years <- unique(d$year)
-  new_years <- sample(original_years)
-  convert <- cbind(original_years,new_years)
+  new_years <- sample(original_years, replace=F)
+  convert <- cbind(original_years, new_years)
   convert
@@
   d <- do.call(rbind, d2)
+  convert
 }
@@
 ```{r}
-nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
+#nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
+nd_all <- readRDS("data/predicted_temp_allyears_revised.rds") %>% select(X, Y, year, est) %>% rename(temp = est)
 fishing <- readRDS("data/fishing-effort-grid.rds")
@@
 d$log_effort[is.na(d$log_effort)] <- -2.3
+d$effort[is.na(d$effort)] <- 0
 
@@
 # d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
-d <- d %>%
-  filter(do_est > 0.28) %>%
-  filter(do_est < 7.06) # 0.005 and 0.995
+# d <- d %>%
+#   filter(do_est > 0.28) %>%
+#   filter(do_est < 7.06) # 0.005 and 0.995
 # d <- d %>% filter(temp > 2.61) %>% filter(temp < 14.31) # full range
@@
 ```{r }
-gf0 <- vocc_gradient_calc(d, "do_est",
-  scale_fac = scale_fac,
-  # all layers with max indice will be averaged for spatial gradient
-  # if default (NULL) will use all years
-  # indices = indices,
-  divisor = 1,
-  quantile_cutoff = 0.05
-)
-
-gf0 <- gf0 %>%
-  mutate(
-    trend_per_year = trend,
-    trend_per_decade = trend_per_year * 10,
-    velocity = velocity
-  ) %>%
-  dplyr::mutate(X = x, Y = y) %>%
-  gfplot:::utm2ll(., utm_zone = 9)
-
-grad0 <- plot_vocc(gf0,
-  # theme_black = TRUE,
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "velocity",
-  fill_label = "gradient velocity",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "black",
-  high_fill = "Steel Blue 4",
-  low_fill = "Red 3",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = sqrt # no_trans #
-) + labs(
-  title = "do velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-grad0
+# gf0 <- vocc_gradient_calc(d, "do_est",
+#   scale_fac = scale_fac,
+#   # all layers with max indice will be averaged for spatial gradient
+#   # if default (NULL) will use all years
+#   # indices = indices,
+#   divisor = 1,
+#   quantile_cutoff = 0.05
+# )
+# 
+# gf0 <- gf0 %>%
+#   mutate(
+#     trend_per_year = trend,
+#     trend_per_decade = trend_per_year * 10,
+#     velocity = velocity
+#   ) %>%
+#   dplyr::mutate(X = x, Y = y) %>%
+#   gfplot:::utm2ll(., utm_zone = 9)
+# 
+# grad0 <- plot_vocc(gf0,
+#   # theme_black = TRUE,
+#   coast = TRUE,
+#   vec_aes = NULL,
+#   # grad_vec_aes = "velocity",
+#   # vec_lwd_range = c(0.5, 0.5),
+#   NA_label = "NA",
+#   fill_col = "trend",
+#   fill_label = "trend",
+#   raster_alpha = 1,
+#   # raster_limits = c(raster_min, raster_max),
+#   na_colour = "black",
+#   high_fill = "Steel Blue 4",
+#   low_fill = "Red 3",
+#   white_zero = TRUE,
+#   vec_alpha = 0.35,
+#   axis_lables = FALSE,
+#   transform_col = no_trans #
+# ) + labs(
+#   title = "do velocity",
+#   subtitle = paste0(min(d$year), "-", max(d$year))
+# )
+# 
+# grad0
 ```
@@
 ```{r}
+# d2 <- filter(d, year>2007)
+# indices <- c(1, 1, 1, 2, 2, 2)
+# indices <- c(1, 1, 1, 1, 2, 2)
+# indices <- c(1, 1, 1, 1, 1, 2)
+# indices <- c(1, 1, 1, 1, 1, 1, 1, 2)
+# indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
 gf2 <- vocc_gradient_calc(d, "temp",
@@
 
+gf2_indices <- select(gf2, X, Y, gradient, velocity) %>% rename(grad_all = gradient, velo_grad_all = velocity)
+
+gf2 <- vocc_gradient_calc(d, "temp",
+  scale_fac = scale_fac,
+  # all layers with max indice will be averaged for spatial gradient
+  # if default (NULL) will use all years
+  indices = indices,
+  divisor = 1,
+  quantile_cutoff = 0.05
+)
+gf2 <- gf2 %>%
+  mutate(
+    trend_per_year = trend,
+    trend_per_decade = trend_per_year * 10,
+    velocity = velocity
+  ) %>%
+  dplyr::mutate(X = x, Y = y)
+
+gf2 <- left_join(gf2, gf2_indices)
+
+gf2$gradient_diff <- gf2$gradient - gf2$grad_all
+range(gf2$gradient_diff)
+
 grad2 <- plot_vocc(gf2,
@@
   NA_label = "NA",
-  fill_col = "velocity",
-  fill_label = "gradient velocity",
+  fill_col = "velocity",# "velo_t2_grad", # "trend", #"gradient_diff", 
+  fill_label = "velocity", #"trend",
   raster_alpha = 1,
+  # raster_limits = c(-6, 6),
   # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
+  na_colour = "black",
   white_zero = TRUE,
@@
   axis_lables = FALSE,
-  transform_col = sqrt # no_trans # log10 # fourth_root_power # sqrt #
+  transform_col = fourth_root_power # sqrt #no_trans # log10 # 
 ) + labs(
@@
 grad2
+
+grad2b <- plot_vocc(gf2,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "trend", #"gradient_diff", 
+  fill_label = "trend", #"trend",
+  raster_alpha = 1,
+  # raster_limits = c(-6, 6),
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "black",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = no_trans # log10 # fourth_root_power # sqrt #
+) + labs(
+  title = "temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+
+grad2b
+
+grad2c <- plot_vocc(gf2,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "gradient", 
+  fill_label = "spatial gradient T2", #"trend",
+  raster_alpha = 1,
+  # raster_limits = c(-6, 6),
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "black",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = no_trans # log10 # fourth_root_power # sqrt #
+) + labs(
+  title = "temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+
+grad2c
 ```
@@
 # indices <- c(1,2,3,4,5)
-
+# indices <- c(1, 1, 1, 1, 1, 1, 1, 2)
+# indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 gf3 <- vocc_gradient_calc(d, "est",
   scale_fac = scale_fac,
-  # indices = indices,
+  indices = indices,
   divisor = 1,
@@
   raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
+  # raster_limits = c(-9, 9),
   na_colour = "black",
@@
 
-DO_values <- gf0 %>%
-  rename(
-    DO_vel = velocity,
-    DO_trend = trend_per_year,
-    DO_grad = gradient,
-    mean_DO = mean
-  ) %>%
-  select(x, y, DO_vel, DO_trend, DO_grad, mean_DO)
+# DO_values <- gf0 %>%
+#   rename(
+#     DO_vel = velocity,
+#     DO_trend = trend_per_year,
+#     DO_grad = gradient,
+#     mean_DO = mean
+#   ) %>%
+#   select(x, y, DO_vel, DO_trend, DO_grad, mean_DO)
 
@@
     temp_vel = velocity,
+    temp_vel_all = velo_grad_all,
     temp_trend = trend_per_year,
     temp_grad = gradient,
+    temp_grad_all = grad_all,
     mean_temp = mean
@@
 vocc <- left_join(biotic_values, fished) # add fishing effort
-vocc <- left_join(vocc, DO_values) # add DO
+# vocc <- left_join(vocc, DO_values) # add DO
 vocc <- left_join(vocc, temp_values) # add temp
@@
 write.csv(vocc, file = paste0(
-  "data/_scrambled3/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
+  "data/_all_temp/", age,  "-all-temp-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
 ))

### 2020-01-24 analysis/VOCC/4-vocc-biotic.Rmd
scramble all spp and regions individually

@@
 }
+
+convert
 ```
@@
   NA_label = "NA",
-  fill_col = "velocity",
-  fill_label = "gradient velocity",
+  fill_col = "trend",
+  fill_label = "trend",
   raster_alpha = 1,
@@
   axis_lables = FALSE,
-  transform_col = sqrt # no_trans #
+  transform_col = no_trans #
 ) + labs(
@@
   axis_lables = FALSE,
-  transform_col = sqrt # no_trans # log10 # fourth_root_power # sqrt #
+  transform_col = no_trans # log10 # fourth_root_power # sqrt #
 ) + labs(
@@
 write.csv(vocc, file = paste0(
-  "data/_scrambled3/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
+  "data/_scrambled5/", age,  "-biotic-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
 ))

### 2020-01-24 analysis/VOCC/render-scrambled-years.R

@@ -1,0 +1,86 @@
+getwd()
+setwd(here::here("/analysis/VOCC"))
+env <- new.env() # parent = baseenv()
+
+# dir.create(file.path("html/biomass-trends"))
+list_species <- c(
+  "Big Skate",
+  "Longnose Skate",
+  "Spotted Ratfish",
+  "North Pacific Spiny Dogfish",
+  # "Pacific Tomcod",
+  "Walleye Pollock",
+  "Pacific Cod",
+  "Sablefish",
+  "Lingcod",
+  # "Pacific Hake",
+  # "Rosethorn Rockfish",
+  "Yellowmouth Rockfish",
+  # "Harlequin Rockfish",
+  "Canary Rockfish", # schooling, winter-birthing
+  # "Copper Rockfish", # small sample
+  "Darkblotched Rockfish",
+  "Greenstriped Rockfish",
+  "Pacific Ocean Perch", # schooling
+  "Redbanded Rockfish",
+  "Sharpchin Rockfish",
+  #"Shortbelly Rockfish", # small sample
+  "Silvergray Rockfish",
+  "Splitnose Rockfish",
+  "Yellowtail Rockfish", # schooling
+  "Longspine Thornyhead",
+  "Shortspine Thornyhead",
+  "Arrowtooth Flounder",
+  "Rex Sole",
+  "Petrale Sole",
+  "English Sole",
+  "Dover Sole",
+  "Southern Rock Sole",
+  "Flathead Sole",
+  "Curlfin Sole",
+  # "Sand Sole",
+  # "Slender Sole",
+  # "Pacific Sanddab",
+  "Pacific Halibut",
+  "Redstripe Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Widow Rockfish",
+  "Quillback Rockfish",
+  "Bocaccio",
+  "Shortraker Rockfish",
+  "Yelloweye Rockfish"
+)
+
+### build biotic gradients
+list_regions <- c(
+  "West Coast Vancouver Island",
+  "West Coast Haida Gwaii",
+  "both odd year surveys"
+  #   "All synoptic surveys"
+)
+
+
+for (r_h in seq_along(list_regions)) {
+  for (spp_i in seq_along(list_species)) {
+    spp <- gsub(" ", "-", gsub("\\/", "-", tolower(list_species[spp_i])))
+    covs <- "-tv-depth-only" # string describing model
[…]

### 2020-02-06 analysis/SOPO/1-trawl-split-by-maturity.Rmd
start sopo project

--- 2020-01-03 analysis/VOCC/1-split-biomass-by-maturity.Rmd
@@
 params: 
-    species: "Longnose Skate"
-    #species:"Redbanded Rockfish"
-    #species: "Pacific Ocean Perch"
+    #species: "Bocaccio"
+    #species: "Widow Rockfish"
+    #species: "Quillback Rockfish"
+    #species: "Yelloweye Rockfish"
+    #species: "Shortraker Rockfish"
+    species: "Pacific Ocean Perch"
     region: "All synoptic surveys"
     # region: "Both odd year surveys"
-    split: FALSE
+    threshold: 0.5
+    split: TRUE
 ---
@@
 split <- params$split
+threshold <- params$threshold
 
@@
 paste("region =", region)
+paste("threshold =", threshold)
+paste("split =", split)
 
@@
     cutoff_quantile = c(.9995),
+    p_threshold = threshold,
     plot = TRUE)
@@
     cutoff_quantile = c(.9995),
+    p_threshold = threshold,
     plot = TRUE
@@
   if (split) {
-  maturity$data <- filter(maturity$data, ssid == model_ssid)
+  maturity$data <- filter(maturity$data, ssid %in% model_ssid)
 
@@
   if (is.null(maturity$maturity)) {
-    maturity$data <- filter(maturity$data, ssid == model_ssid) 
+
+    maturity$data <- filter(maturity$data, ssid %in% model_ssid)
+
   } else { # if maturity model is deemed unuseful
-  maturity$data <- filter(maturity$data, ssid == model_ssid) %>% 
+    maturity$data <- filter(maturity$data, ssid %in% model_ssid) %>% 
       select(-adult_density, -imm_density) 

### 2020-02-06 analysis/VOCC/bioclimatic/4-vocc-bioclimatic.Rmd
organize old code

--- 2019-11-13 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
@@
 ---
-title: "BACI of `r paste(params$species)` biomass for `r paste(params$region)` VOCC"
+title: "`r paste(params$species)` bioclimatic and biotic vectors for `r paste(params$region)` VOCC"
 author: "Philina English"

### 2020-02-06 analysis/VOCC/old-code/1-biomass-do-models-fixed-both.Rmd
organize old code

--- 2019-09-11 analysis/VOCC/1-biomass-do-models-fixed-both.Rmd
@@
 # covariates <- "+muddy+any_rock"
- covariates <- ""
+ covariates <- "+as.factor(ssid)"
 # covariates <- "+trawled+as.factor(ssid)"
@@
 covs <- "-log-do"
+covs <- "-log-do-ssid"
 ```

### 2020-02-06 analysis/VOCC/old-code/1-biomass-do-models.Rmd
organize old code

---
title: "Spatiotemporal models of species biomass with time-varying DO"
author: "Philina English"
date: '2019-07-18'
output: html_document
params: 
    species: "Arrowtooth Flounder"

---

Choose one value for each parameter:

Species
```{r}
species <- params$species

# # Species run so far...
 species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

Region 
```{r}
region <- "All synoptic surveys"
# region <- "Both odd year surveys"
# region <- "West Coast Vancouver Island"
#  region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
#priors <- FALSE
priors <- TRUE

covariates <- "+muddy+any_rock"
# covariates <- ""

# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
#update_spatial_model <- TRUE
```

Calculate maturity weighted biomass if not done before
```{r}
rm(biomass)
try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))

if (!exists("biomass")) {
  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
  bath <- readRDS("data/bathymetry-data")
  # for POP: exclude sample with unusually large fish 
  # sizes are impossible given recorded catch weight
  # fish <- fish[fish$fishing_event_id!=1506954,]
 rm(maturity)
 try( maturity <- split_catch_maturity(survey_sets, fish, bath,
    # survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
    # survey = c("SYN HS", "SYN WCHG"),
    survey = survey,
    years = years,
    cutoff_quantile = c(.9995),
    plot = TRUE
  ))

  if(!exists("maturity")) {
  maturity <- split_catch_maturity(survey_sets, fish, bath,
    survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
    years = years,
    cutoff_quantile = c(.9995),
    plot = TRUE
  )}

  saveRDS(maturity$data, file = paste0("data/", spp, 
    "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
}
```

```{r}
if (exists("maturity")) {
  print(maturity$mass_model)
  ggsave(
    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
    dpi = 600,
    width = 8, 
    height = 10
  )
}
```

```{r}
if (exists("maturity")) {
  png(
    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), 
    res = 600,
    units = "in",
    width = 8, 
    height = 10
  )
  print(maturity$maturity)
  dev.off()
}

try(print(maturity$maturity))
```

Combine biomass and sensor data
```{r}
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars) %>% mutate(depth_akima = depth) %>% select(-depth)

  CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
  data <- dplyr::left_join(data, CTD) %>% 
    mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
    #mutate(exclude = if_else(do_mlpl>6, 1, 0)) %>% #try excluding high values?
    filter(exclude != 1) %>% # added on july 24 after first round of models run...
    filter(year > 2007) # 2007 DO data is flawed

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth), temp = temperature_c) 
  do_data <- gfranges::scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- do_data %>% mutate(depth = raw_depth) %>% #filter(year != 2016) %>%
    filter(ssid %in% model_ssid) %>% filter(!is.na(do_mlpl)) 
```
```{r}
opt_data <- data %>% filter(depth > 150) %>% filter(temp < 7.5) 
ggplot(filter(opt_data, adult_density < 0.02), aes(do_mlpl, (adult_density), colour = as.factor(ssid))) + 
  geom_point() +
  facet_wrap(~year, scales="free_y")
```
```{r}
#filter(data, adult_density < log(0.002))
ggplot(filter(data, ssid ==16), aes(do_mlpl, (adult_density), colour = as.factor(ssid))) + geom_point() + facet_wrap(~ssid)
```

```{r}
  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "All synoptic surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
  }

  sdmTMB::plot_spde(spde)
```

Run DO sdmTMB model
```{r}
  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime1 <- Sys.time()
    adult_biomass_d <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-do3", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime2 <- Sys.time()
    imm_biomass_d <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-do3", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

  } else {

    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

    total_biomass_d <- sdmTMB::sdmTMB(data,
      dens_formula,
      time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2 + do_mlpl_scaled3, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-do3", covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
```

Test gbm 
```{r eval = FALSE}
library(gbm)
#glimpse(data)
data$year.f <- as.factor(data$year)
data$log_density_ad <- log(data$adult_density*100)
data$log_do <- log(data$do_mlpl)

 adult_formula <- as.formula(paste(
       "adult_density ~ 0 + year.f * raw_depth + do_mlpl * temperature_c + muddy + sandy + rocky + mixed + X * Y * year.f + raw_depth:do_mlpl + year.f:do_mlpl"
      ))

m <- gbm( adult_formula ,
  # log_density_ad ~ 0 + year.f + raw_depth + muddy + rocky + mixed + # sandy +
  # do_mlpl + temperature_c + X * Y, 
  data = data, 
  distribution = "gaussian",
  n.trees = 2000, interaction.depth = 3, 
  shrinkage = 0.02)
```

```{r}
hist(data$log_density_ad)
hist(log(data$do_mlpl))
```

```{r eval = FALSE}
m

plot(m,i.var=1)
plot(m,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)
plot(m,i.var=5)
plot(m,i.var=6)
plot(m,i.var=7)
plot(m,i.var=8)
plot(m,i.var=c(1,2))

plot(m,i.var=c(1,3))
plot(m,i.var=c(2,3))
plot(m,i.var=c(3,4))

plot(m,i.var=c(9,10))

```

```{r eval = FALSE}
data$resid <- predict(m, n.trees = 2000) - data$adult_density

ggplot(data, aes(raw_depth, resid)) +
  geom_point(alpha=0.4) +
  ylim(-0.0025,0.0025) +
  geom_smooth()

ggplot(data, aes(do_mlpl, resid)) +
  geom_point(alpha=0.4) +
  ylim(-0.0025,0.0025) +
  geom_smooth()

ggplot(data, aes(temperature_c, resid)) +
  geom_point(alpha=0.4) +
  ylim(-0.0025,0.0025) +
  geom_smooth()

```

Run DO sdmTMB model for all regions together
```{r eval = FALSE}

if (region=="All synoptic surveys")
  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2"#, covariates, ""
      ))

    starttime1 <- Sys.time()
    adult_biomass_d <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-do3", #covs, 
        "-", ssid_string, "-prior-", priors, ".rds"
        ))

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2"#, covariates, ""
      ))

    starttime2 <- Sys.time()
    imm_biomass_d <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-do3", #covs, 
        "-", ssid_string, "-prior-", priors, ".rds"
        ))

  } else {

    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + do_mlpl_scaled3"#, covariates, ""
      ))

    total_biomass_d <- sdmTMB::sdmTMB(data,
      dens_formula,
      #time_varying = ~ 0 + do_mlpl_scaled + do_mlpl_scaled2 + do_mlpl_scaled3, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-do3", #covs, 
      "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
```

```{r eval = FALSE}
#d <- time_varying_density3(adult_biomass_d, predictor = "do_mlpl")

d <- fixed_density(adult_biomass_d, predictor = "do_mlpl")

do_plot <- plot_mountains(d, variable_label = "Dissolved O2", xlimits = c(0,2)) + 
    ggtitle(" ")

do_plot
```

### 2020-02-06 analysis/VOCC/old-code/1-biomass-temp-models-fixed-both.Rmd
organize old code

--- 2019-08-29 analysis/VOCC/1-biomass-temp-models-fixed-both.Rmd
@@
 # species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-species <- "Quillback Rockfish"
+# species <- "Quillback Rockfish"
 # species <- "Pacific Ocean Perch"

### 2020-02-06 analysis/VOCC/old-code/3-fixed-var-response-hist.Rmd
organize old code

---
title: "Plot responses to model covariates to inform threshold choices for VOCC"
author: "Philina English"
date: '2019-07-18'
output: html_document
---

Species
```{r}

# # Species run so far...
# species <- "Arrowtooth Flounder" # temp good, not DO, imm moving deeper?

# species <- "Petrale Sole"
# species <- "English Sole"
# species <- "Dover Sole"
# species <- "Southern Rock Sole"

# species <- "Walleye Pollock" # survey 4 or 16 might work
# species <- "Pacific Cod"

## Sebastes
# 
# ## TRY BOTH - not schooling
#  species <- "Yelloweye Rockfish" # beautiful curves, with depth change ***
#  species <- "Bocaccio" # good curves, no depth change ... most survey-specific models fail to find minimum ***
#  species <- "Sharpchin Rockfish" # beautiful curves, imm depth change ***
#  species <- "Greenstriped Rockfish" # beautiful curves for adults, no depth change ***

# ## TRY BOTH - schooling
#  species <- "Pacific Ocean Perch" # beautiful curves, no depth change, schooling ***
#  species <- "Silvergray Rockfish" # beautiful curves, no depth change, schooling ***

# ## TRY JUST TEMPERATURE 
# # species <- "Rougheye/Blackspotted Rockfish Complex" # good temp, not DO
# # species <- "Redbanded Rockfish" # good temp, not DO, maybe moving shallower
#  species <- "Splitnose Rockfish" # good temp, survey 4 DO ok

# # species <- "Yellowtail Rockfish" #  good temp for survey 4 only, no DO, schooling
# # species <- "Quillback Rockfish" # like warm = overall a bit flat... territorial! 
# # species <- "Canary Rockfish" # schooling, winter-birthing, and depth change!, but curves at edges... adult1 looks ok?
#  species <- "Widow Rockfish" # schooling, and depth change, but curves at edges

# species <- "Longspine Thornyhead" # too deep to produce curves
# species <- "Shortspine Thornyhead"
# species <- "Sablefish" # temp might work, no DO, no consistant depth change
# species <- "Lingcod" # curves ok, overall a bit flat
# # species <- "Pacific Hake" # not useful

# species <- "North Pacific Spiny Dogfish" # too flat, but with depth change! note: pooled maturity 
# species <- "Longnose Skate" # do too flat, temp from 4 might be useful, seems to go deeper?
# species <- "Big Skate"
# species <- "Sandpaper Skate"
# species <- "Brown Cat Shark"
# species <- "Spotted Ratfish" # curves not useful because to abundant in shallow
```

Choose model details
```{r}
covariates <- ""
# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
#covariates <- "+muddy+any_rock"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
# 
# covariates <- "+as.factor(ssid)"
# covs <- "-both-tv-depth-ssid"

# priors <- FALSE
# covariates <- "+muddy+any_rock"
# covs <- gsub("\\+", "-", covariates)
priors <- TRUE
#covs <- "-tv-depth"
# covs <- "-both-tv-depth"
# covs <- "-log-do"
covs <- "-log-do-trawled"
num_params <- 2
```

Run all subsequent code...
```{r global_options, include=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  fig.path = paste0("figs/", spp, "/"),
  echo = FALSE, warning = FALSE, message = FALSE
)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

Load Dissolved O2 models
```{r}
rm(adult, imm,
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16,
  total, total1, total4, total16
  )

try({
  # loads global models with 2016 excluded due to strangely high DO 
  add2016 <- ""
  x2016 <- "x2016"
  # for models using 2016 
  # add2016 <- "-do"
  #  x2016 <- ""
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  try({adult <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  imm <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  })
  try({total <- readRDS(paste0("data/", spp,
    "/model-total-biomass-", spp, "-fixed", add2016, covs, "-", ssid_string, "-prior-", priors, ".rds"))
  })
})
if (!exists("adult")) adult <- NULL
if (!exists("imm")) imm <- NULL
if (!exists("total")) total <- NULL

try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  try({adult_survey1 <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  imm_survey1 <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  })
  try({total1 <- readRDS(paste0("data/", spp,
    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
  })
  })
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL
if (!exists("total1")) total1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  try({adult_survey4 <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  imm_survey4 <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  })
  try({total4 <- readRDS(paste0("data/", spp,
    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
  })
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL
if (!exists("total4")) total4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  try({adult_survey16 <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
   imm_survey16 <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
   ))
  })
  try({total16 <- readRDS(paste0("data/", spp,
    "/model-total-biomass-", spp, "-fixed-do", covs, "-", ssid_string, "-prior-", priors, ".rds"))
  })
})
if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL
if (!exists("total16")) total16 <- NULL

model_list <- list(total1 = total1, total4 = total4, total16 = total16, 
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16,
  total = total, adult = adult, imm = imm
)
model_list <- model_list[!sapply(model_list, is.null)]
```

```{r}
# test load small sdm model to avoid crashing R
# try({imm})
# or maybe it's crashing when NaNs produced error goes into density function?
```

## Density curves from model coefficients

#### Calculate DO curves
```{r}
do <- list()
do_plots <- list()

for (i in seq_len(length(model_list))) {

 # check if models converged
 if(model_list[[i]]$model$convergence==0){
  do[[i]] <- fixed_density(model_list[[i]], predictor = "do_mlpl", quadratics_only = FALSE)

  if (length(do[[i]])>0) {
   do[[i]]$species <- species
   do[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
   do[[i]]$ssid <- paste0(sort(unique(as.numeric(model_list[[i]]$data$ssid))), collapse = "n")
   do[[i]]$model <- names(model_list[i])
   do[[i]]$covs <- paste0(x2016,covs) 
   roots <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.5)
   do[[i]]$optimal_do <- exp(roots[["optimal"]])
   do[[i]]$lower_do_50 <- exp(roots[["lower_threshold"]])
   do[[i]]$upper_do_50 <- exp(roots[["upper_threshold"]])
   roots75 <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.75)
   do[[i]]$lower_do_75 <- exp(roots75[["lower_threshold"]])
   do[[i]]$upper_do_75 <- exp(roots75[["upper_threshold"]])
   roots25 <- get_quadratic_roots(model_list[[i]], predictor = "do_mlpl", fixed_param = 1, threshold = 0.25)
   do[[i]]$lower_do_25 <- exp(roots25[["lower_threshold"]])
   do[[i]]$upper_do_25 <- exp(roots25[["upper_threshold"]])
   do[[i]]$x <- exp(do[[i]]$x)
   do_plots[[i]] <- 
     plot_mountains(do[[i]], variable_label = "Dissolved O2", xlimits = c(0.2, 8)) +
        ggtitle(paste(species, names(model_list[i]))) +
        geom_vline(xintercept=(do[[i]]$optimal_do), size= 1.2, colour = "darkgrey") + 
        geom_vline(xintercept=(do[[i]]$lower_do_50), linetype="longdash", alpha = 0.75) + 
        geom_vline(xintercept=(do[[i]]$upper_do_50), linetype="longdash", alpha = 0.75) +
        geom_vline(xintercept=(do[[i]]$lower_do_75), linetype="dotted", alpha = 0.5) + 
        geom_vline(xintercept=(do[[i]]$upper_do_75), linetype="dotted", alpha = 0.5) +
        geom_vline(xintercept=(do[[i]]$lower_do_25), linetype="dotted", alpha = 0.5) + 
        geom_vline(xintercept=(do[[i]]$upper_do_25), linetype="dotted", alpha = 0.5) 
   } else {
    do_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
   }
 } else {
  do_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
 }
}

odat <- do.call("rbind", do) 
if (length(odat)>0) {
  odat <- odat %>% group_by(model) %>%
  mutate(do_maxy = max(y_hat)) %>% 
  select(-x, -y_hat, -year) %>% 
  distinct() 
} else {
  odat <- NULL
}
print(do_plots)
```

#### Calculate temperature curves
```{r}
td <- list()
temp_plots <- list()
for (i in seq_len(length(model_list))) {
 if(model_list[[i]]$model$convergence==0){
  td[[i]] <- fixed_density(model_list[[i]], predictor = "temp", quadratics_only = FALSE, fixed_params = num_params)
  if (length(td[[i]])>0) {
    td[[i]]$species <- species
    td[[i]]$age <- ifelse(grepl("imm", names(model_list[i])), "imm", "adult")
    td[[i]]$ssid <- paste0(sort(unique(as.numeric(model_list[[i]]$data$ssid))), collapse = "n")
    td[[i]]$model <- names(model_list[i])
    td[[i]]$covs <- paste0(x2016,covs) 
   roots <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.5)
   td[[i]]$optimal_temp <- roots[["optimal"]]
   td[[i]]$lower_t_50 <- roots[["lower_threshold"]]
   td[[i]]$upper_t_50 <- roots[["upper_threshold"]]
   roots75 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.75)
   td[[i]]$lower_t_75 <- roots75[["lower_threshold"]]
   td[[i]]$upper_t_75 <- roots75[["upper_threshold"]]
   roots25 <- get_quadratic_roots(model_list[[i]], predictor = "temp", fixed_param = num_params, threshold = 0.25)
   td[[i]]$lower_t_25 <- roots25[["lower_threshold"]]
   td[[i]]$upper_t_25 <- roots25[["upper_threshold"]]
   temp_plots[[i]] <- 
     plot_mountains(td[[i]], variable_label = "Temperature", xlimits = c(2, 12)) + #
        ggtitle(paste(species, names(model_list[i]))) +
        geom_vline(xintercept=td[[i]]$optimal_t, size= 1.2, colour = "darkgrey") + 
        geom_vline(xintercept=td[[i]]$lower_t_50, linetype="longdash", alpha = 0.75) + 
        geom_vline(xintercept=td[[i]]$upper_t_50, linetype="longdash", alpha = 0.75) +
        geom_vline(xintercept=td[[i]]$lower_t_75, linetype="dotted", alpha = 0.5) + 
        geom_vline(xintercept=td[[i]]$upper_t_75, linetype="dotted", alpha = 0.5) +
        geom_vline(xintercept=td[[i]]$lower_t_25, linetype="dotted", alpha = 0.5) + 
        geom_vline(xintercept=td[[i]]$upper_t_25, linetype="dotted", alpha = 0.5)

  } else {
    temp_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  } 
 } else {
    temp_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  } 
}

tdat <- do.call("rbind", td) 
if (length(tdat)>0) {
  tdat <- tdat %>% group_by(model) %>%
  mutate(t_maxy = max(y_hat)) %>% 
  select(-x, -y_hat, -year) %>% 
  distinct() 
} else {
  tdat <- NULL
}
print(temp_plots)
```

## Check optimals against raw capture denstiies

#### Combine biomass and sensor data
```{r}
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars) %>% mutate(depth_akima = depth) %>% select(-depth)

  CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")
  data <- dplyr::left_join(data, CTD) %>% 
    mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
    #mutate(exclude = if_else(do_mlpl>6, 1, 0)) %>% #try excluding high values?
    filter(exclude != 1) %>% # added on july 24 after first round of models run...
    filter(year > 2007) # 2007 DO data is flawed

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth), temp = temperature_c) 
```

Build time-varying depth plots
```{r }
max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
d <- list()
depth_plots <- list()
for (i in seq_len(length(model_list))) {
 if(model_list[[i]]$model$convergence==0){
  d[[i]] <- time_varying_density(model_list[[i]], predictor = "depth")

  if (length(d[[i]]) == 0) {
    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  } else {
    d[[i]]$x <- exp(d[[i]]$x)
    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, max_depth_found)) +
      ggtitle(paste(species, names(model_list[i])))
  }
 } else {
    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  } 
}
print(depth_plots)
```

#### Plot raw DO
```{r}
if(!is.null(data$adult_density)){
odata <- data %>% filter(!is.na(do_mlpl)) %>% 
  mutate(ad_density_class = ifelse(adult_density==0, 0, round(log10((adult_density*100000)+1))),
    im_density_class = ifelse(imm_density==0, 0, round(log10((imm_density*10000000)+1)))) %>% 
  group_by(ad_density_class) %>% mutate(ad_mean_density = mean(adult_density, na.rm = TRUE)) %>% ungroup() %>%
  group_by(im_density_class) %>% mutate(im_mean_density = mean(imm_density, na.rm = TRUE)) %>% ungroup()

ad_density_class_labs <- unique(signif(odata$ad_mean_density*10000),2) # convert to kg/ha
ad_density_class_labs <- ad_density_class_labs[order(ad_density_class_labs)]
ad_density_class_labs <- formatC(ad_density_class_labs, digits = 1, format = "fg")
ad_density_class_labs

im_density_class_labs <- unique(signif(odata$im_mean_density*10000),2) # convert to kg/ha
im_density_class_labs <- im_density_class_labs[order(im_density_class_labs)]
im_density_class_labs <- formatC(im_density_class_labs, digits = 1, format = "fg")
im_density_class_labs

do_count_ad <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(ad_density_class))) +
geom_histogram(binwidth = 0.5, position = "stack") +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
  geom_vline(xintercept=odat[odat$model =="adult", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Count") +
  ggtitle(paste("Adult", species))

do_prop_ad <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(ad_density_class))) +
geom_histogram(binwidth = 0.5, position = "fill") +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
  geom_vline(xintercept=odat[odat$model =="adult", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Proportion") +
  ggtitle(paste("Adult", species))

do_count_im <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(im_density_class))) +
  geom_histogram(binwidth = 0.5, position = "stack") +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
  geom_vline( xintercept = odat[odat$model =="imm", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
  geom_vline( xintercept = odat[odat$age =="imm", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Count") +
  ggtitle(paste("Immature", species))

do_prop_im <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(im_density_class))) +
geom_histogram(binwidth =0.5, position = "fill") +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
  geom_vline( xintercept = odat[odat$model =="imm", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
  geom_vline( xintercept = odat[odat$age =="imm", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Proportion") +
  ggtitle(paste("Immature", species))

raw_do_plots <- list(do_count_ad, do_prop_ad, do_count_im, do_prop_im )
print(raw_do_plots)

} else {

odata <- data %>% filter(!is.na(do_mlpl)) %>% 
  mutate(density_class = ifelse(density==0, 0, (round(log10((density*10000)+1)*2)+1))) %>% 
  group_by(density_class) %>% mutate(mean_density = mean(density, na.rm = TRUE)) %>% ungroup()

density_class_labs <- unique(signif(odata$mean_density*10000),2) # convert to kg/ha
density_class_labs <- density_class_labs[order(density_class_labs)]
density_class_labs <- formatC(density_class_labs, digits = 1, format = "fg")
density_class_labs

do_count <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(density_class))) +
geom_histogram(binwidth = 0.5, position = "stack" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
  geom_vline(xintercept=odat[odat$model =="total", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Count") +
  ggtitle(paste("Total", species))

do_prop <- ggplot(odata, aes(x = do_mlpl, stat(count), fill = as.factor(density_class))) +
geom_histogram(binwidth = 0.5, position = "fill" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
  geom_vline(xintercept=odat[odat$model =="total", ]$optimal_do, color="black",  size=1, na.rm = TRUE) +
  geom_vline(xintercept=odat[odat$age =="adult", ]$optimal_do, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Proportion") +
  ggtitle(paste("Total", species))

raw_do_plots <- list(do_count,do_prop)
print(raw_do_plots)
}
```

#### Plot raw temperature
```{r}
if(!is.null(data$adult_density)){
tdata <- data %>% filter(!is.na(temp)) %>% 
  mutate(ad_density_class = ifelse(adult_density==0, 0, round(log10((adult_density*100000)+1))),
    im_density_class = ifelse(imm_density==0, 0, round(log10((imm_density*10000000)+1)))) %>% 
  group_by(ad_density_class) %>% mutate(ad_mean_density = mean(adult_density, na.rm = TRUE)) %>% ungroup() %>%
  group_by(im_density_class) %>% mutate(im_mean_density = mean(imm_density, na.rm = TRUE)) %>% ungroup()

ad_density_class_labs <- unique(signif(tdata$ad_mean_density*10000),2) # convert to kg/ha
ad_density_class_labs <- ad_density_class_labs[order(ad_density_class_labs)]
ad_density_class_labs <- formatC(ad_density_class_labs, digits = 1, format = "fg")
ad_density_class_labs

im_density_class_labs <- unique(signif(tdata$im_mean_density*10000),2) # convert to kg/ha
im_density_class_labs <- im_density_class_labs[order(im_density_class_labs)]
im_density_class_labs <- formatC(im_density_class_labs, digits = 1, format = "fg")
im_density_class_labs

t_count_ad <- ggplot(tdata, aes(x = temp, stat(count),fill = as.factor(ad_density_class))) +
  geom_histogram(binwidth = 0.5, position = "stack" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
  geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Temperature from CTD") +
  ylab("Count") +
  ggtitle(paste("Adult", species))

t_prop_ad <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(ad_density_class))) +
  geom_histogram(binwidth = 0.5, position = "fill" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=ad_density_class_labs) + 
  geom_vline(xintercept=tdat[tdat$model =="adult", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Temperature from CTD") +
  ylab("Proportion") +
  ggtitle(paste("Adult", species))

t_count_im <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(im_density_class))) + 
  geom_histogram(binwidth = 0.5, position = "stack" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
  geom_vline( xintercept = tdat[tdat$model =="imm", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
  geom_vline( xintercept = tdat[tdat$age =="imm", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Temperature from CTD") +
  ylab("Count") +
  ggtitle(paste("Immature", species))

t_prop_im <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(im_density_class))) +
  #geom_density(position = "fill") +
geom_histogram(binwidth = 0.5, position = "fill" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=im_density_class_labs) + 
  geom_vline( xintercept = tdat[tdat$model =="imm", ]$optimal_temp, color="black", size=1, na.rm = TRUE) +
  geom_vline( xintercept = tdat[tdat$age =="imm", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Temperature from CTD") +
  ylab("Proportion") +
  ggtitle(paste("Immature", species))

raw_t_plots <- list(t_count_ad, t_prop_ad, t_count_im, t_prop_im )
print(raw_t_plots)

} else {

tdata <- data %>% filter(!is.na(temp)) %>% 
  mutate(density_class = ifelse(density==0, 0, (round(log10((density*10000)+1)*2)+1))) %>% 
  group_by(density_class) %>% mutate(mean_density = mean(density, na.rm = TRUE)) %>% ungroup()

density_class_labs <- unique(signif(tdata$mean_density*10000),2) # convert to kg/ha
density_class_labs <- density_class_labs[order(density_class_labs)]
density_class_labs <- formatC(density_class_labs, digits = 1, format = "fg")
density_class_labs

t_count <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(density_class))) +
geom_histogram(binwidth = 0.5, position = "stack" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
  geom_vline(xintercept=tdat[tdat$model =="total", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Temperature from CTD") +
  ylab("Count") +
  ggtitle(paste("Total", species))

t_prop <- ggplot(tdata, aes(x = temp, stat(count), fill = as.factor(density_class))) +
geom_histogram(binwidth = 0.5, position = "fill" ) +
  scale_fill_viridis_d(name = "Mean Density \n(kg/ha)", label=density_class_labs) + 
  geom_vline(xintercept=tdat[tdat$model =="total", ]$optimal_temp, color="black",  size=1, na.rm = TRUE) +
  geom_vline(xintercept=tdat[tdat$age =="adult", ]$optimal_temp, color="black", linetype="dashed", size=0.5, na.rm = TRUE) +
  theme_pbs() +
  xlab("Temperature from CTD") +
  ylab("Proportion") +
  ggtitle(paste("Total", species))

raw_t_plots <- list(t_count,t_prop)
print(raw_t_plots)
}
```

Save DO plots
```{r}
png(
  file = paste0("figs/", spp, "/do-fixed-", spp, x2016, covs, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = do_plots,
  nrow = 4,
  top = grid::textGrob(paste(species))
)
dev.off()
```

Save Temperature plots
```{r}
png(
  file = paste0("figs/", spp, "/temp-fixed-", spp, x2016, covs, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = temp_plots,
  nrow = 4,
  top = grid::textGrob(paste(species))
)
dev.off()
```

#### Save plots of raw densities vs model optimals 
```{r}
png(
  file = paste0("figs/", spp, "/check-optimals-", spp, x2016, covs, "-priors-", priors, "-hist.png"),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(grobs = c(raw_do_plots, raw_t_plots),
  nrow = 4,
  top = grid::textGrob(paste("Raw densities of", species, "with modeled optimals in vertical lines (if solid includes all surveys)"))
)
dev.off()
```

#### Save plots of time-varying depth from climate model
```{r}
png(
  file = paste0(
    "figs/", spp,
    "/depth-", spp, x2016, covs, "-priors-", priors, ".png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = depth_plots,
  nrow = 4,
  top = grid::textGrob(paste(species))
)
dev.off()
```

## Save optimal values
```{r }
other_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))

if(!is.null(odat)) {
  if(!is.null(tdat)) {
    params <- full_join(tdat, odat, by=c("species", "age", "model", "covs", "ssid"))
    all_params <- full_join(other_params, params)
  } else {
  try({all_params <- full_join(other_params, odat)})
  }
} else {
  try({all_params <- full_join(other_params, tdat)})
}

all_params <- all_params %>% distinct()
# # If data far a particular species needs redoing
# all_params <- subset(other_params, species!="Longnose Skate")
# # or certain rows need removal
# all_params <- all_params[-c(#row number#), ] 
saveRDS(all_params, file = "data/root-values-by-species-by-age.rds")
all_params
```

Alternative presense-absense code
```{r eval=FALSE}
odata <- data %>% filter(!is.na(do_mlpl)) #%>% filter(do_mlpl < 1.5)

odata1 <- filter(odata, adult_density > 0)
# ymax <- which.max(density(odata1$do_mlpl)$y)
# do_optimal_raw <- density(odata1$do_mlpl)$x[ymax]

oplot <- ggplot(odata1) + 
  geom_density(aes(x = do_mlpl), colour = "blue") +
  # geom_vline(aes(xintercept=do_optimal_raw),
  #           color="blue", linetype="dashed", size=1) +
  geom_density(data=filter(odata, adult_density == 0), aes(x = do_mlpl), colour = "red", inherit.aes = FALSE) + theme_bw() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Frequency") +
  ggtitle(paste("Adult"))
oplot

oploti <- ggplot(filter(odata, imm_density > 0)) + geom_density(aes(x = do_mlpl), colour = "blue") + 
  geom_density(data=filter(odata, imm_density == 0), aes(x = do_mlpl), colour = "red", inherit.aes = FALSE) + theme_bw() +
  xlab("Dissolved O2 from CTD (ml/L)") +
  ylab("Frequency") +
  ggtitle(paste("Immature"))
oploti

tdata <- data %>% filter(!is.na(temp)) #%>% filter(do_mlpl < 1.5)
tplot <-ggplot(filter(tdata, adult_density > 0)) + geom_density(aes(x = temp), colour = "blue") + 
  geom_density(data=filter(tdata, adult_density == 0), aes(x = temp), colour = "red", inherit.aes = FALSE) + theme_bw() +
  xlab("Temperature from CTD") +
  ylab("Frequency") +
  ggtitle(paste("Adult"))
tplot
tploti <-ggplot(filter(tdata, imm_density > 0)) + geom_density(aes(x = temp), colour = "blue") + 
  geom_density(data=filter(tdata, imm_density == 0), aes(x = temp), colour = "red", inherit.aes = FALSE) + theme_bw() +
  xlab("Temperature from CTD") +
  ylab("Frequency") +
  ggtitle(paste("Immature"))
tploti
```

Save raw presense-absence density plots
```{r eval=FALSE}
png(
  file = paste0("figs/", spp, "/raw-presense-by-climate-", spp, "-density-plots.png"),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  oplot,  tplot, oploti, tploti,
  ncol = 2,
  top = grid::textGrob(paste("Proportion of tows with", species, "(blue) vs without (red)"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

Load Depth models (if using old depth only models)
```{r eval=FALSE}
rm(
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16
)
priors <- FALSE
covariates <- "+muddy+any_rock"
covs <- gsub("\\+", "-", covariates)
try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  adult_survey1 <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  imm_survey1 <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  adult_survey4 <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  imm_survey4 <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  adult_survey16 <- readRDS(paste0("data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
  imm_survey16 <- readRDS(paste0("data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL

depth_model_list <- list(
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
)
depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
```

Build depth plots
```{r eval=FALSE}

max_depth_found <- max(data[data$present==1,]$raw_depth, na.rm = TRUE)
d <- list()
depth_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")

  if (length(d[[i]]) == 0) {
    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  } else {
    d[[i]]$x <- exp(d[[i]]$x)
    depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0, max_depth_found)) +
      ggtitle(paste(species, names(depth_model_list[i])))
  }
}
print(depth_plots)
```

```{r eval=FALSE}
png(
  file = paste0(
    "figs/", spp,
    "/depth-only-", spp, covs, "-priors-", priors, ".png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = depth_plots,
  ncol = 2,
  top = grid::textGrob(paste(species))
)
dev.off()
```

# TRAWLING
Plot time-varying effect of trawling at mean depth
```{r eval=FALSE}
trawl_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  r <- depth_model_list[[i]]$tmb_obj$report()
  b3 <- r$b_rw_t[, 3]
  yrs <- sort(unique(depth_model_list[[i]]$data$year))
  b_j <- depth_model_list[[i]]$model$par[1:length(yrs)]

  SE <- 0.35 # Place-holder

  trawled <- data.frame(year = yrs, intercept = b_j, trawling = b3)

  # trawl_plots[[i]] <- ggplot(trawled, aes(year, y=exp(trawling) ) ) +
  #   geom_point(colour ="red") +
  #   #geom_pointrange(aes(ymin=exp(intercept-0.35), ymax=exp(intercept+0.35))) +
  #   ylab("Difference in kg/ha within trawled zone") + ggtitle(names(depth_model_list[i]))

  trawl_plots[[i]] <- ggplot(trawled, aes(year, exp(intercept))) + geom_point() +
    geom_point(aes(year, y = exp(intercept + trawling)), colour = "red") +
    geom_pointrange(aes(ymin = exp(intercept - (1.96 * SE)), ymax = exp(intercept + (1.96 * SE)))) + # using mean se on year effects
    ylim(min(exp(trawled$intercept - trawled$trawling - (1.96 * SE))), max(exp(trawled$intercept + trawled$trawling + (1.96 * SE)))) +
    ylab("Predicted kg/ha at mean depth") + ggtitle(names(depth_model_list[i]))
}
print(trawl_plots)
```

### Code for calculation SE on time-varying effects... 
```{r eval=FALSE}
# FIXME: need to add to sdmTMB!
r <- m$tmb_obj$report()
summary(m)
yrs <- sort(unique(m$data$year))
.sd <- summary(m$sd_report)
.sd <- as.data.frame(.sd)
.sd$par <- row.names(.sd)
.sd <- .sd[grepl("^b_rw_t", .sd$par), ]
varnames <- colnames(model.matrix(m$time_varying, m$data))

rep(varnames, each = length(yrs))
.sd$parameter <- rep(varnames, each = length(yrs))
.sd$year <- rep(seq_along(yrs), each = length(varnames))
row.names(.sd) <- NULL
.sd$par <- NULL

names(.sd$Estimate) <- "estimate"
names(.sd[2]) <- "se"
.sd$ci <- .sd$se * 1.96
m$time_varying

nrow(.sd)

b3 <- r$b_rw_t[, 3]
yrs <- sort(unique(m$data$year))
b_j <- m$tmb_params$b_j[1:8]

trawled <- data.frame(year = yrs, intercept = b_j, trawling = b3)
```

Get all biomass predictions 
#### Get all biomass predictions 
```{r, eval=FALSE, echo=FALSE, warning=FALSE}
priors <- FALSE
covariates <- "+muddy+any_rock"
covs <- gsub("\\+", "-", covariates)

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)
pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

try({adult_predictions <- readRDS(paste0("data/", spp,
    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
  ))%>% filter(year > 2007)
adult_predictions <- left_join(adult_predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
})

try({imm_predictions <- readRDS(paste0("data/", spp,
    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
  )) %>% filter(year > 2007)
imm_predictions <- left_join(imm_predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
})

try({predictions <- readRDS(paste0("data/", spp,
    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
  ))%>% filter(year > 2007)
predictions <- left_join(predictions, pred_temp_do, c("X", "Y", "year", "ssid"))
})

if (!exists("adult_predictions")) adult_predictions <- NULL
if (!exists("imm_predictions")) imm_predictions <- NULL
if (!exists("predictions")) predictions <- NULL

biomass_list <- list(
  adult = adult_predictions,
  imm = imm_predictions,
  total = predictions
)

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]

# 
# ggplot(biomass_list[[1]], aes(x = do_est, y = log(est_exp))) + 
#   geom_point(alpha = 0.3) #+ #geom_smooth()
#   geom_smooth(method = "lm", formula = y ~ x + I(x^2) , size = 1) #+ I(x^3)
```

### 2020-02-06 analysis/VOCC/old-code/4-vocc-by-spp.Rmd
organize old code

---
title: "Linking biomass change with VOCC"
author: "Philina English"
date: '2019-08-14'
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

#### Set species
```{r echo=TRUE}
# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, echo=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

params <- readRDS(paste0("data/optimal-values-by-species.rds"))
params <- params[rev(order(as.Date(params$date))), ]
params <- params[!duplicated(params$species), ]

optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)

optimal_values <- c(optimal_do, optimal_temp)
```

#### Set vector type
```{r echo=TRUE}
 climate <- "temperature"
# climate <- "do"
# climate <- "do and temperature"
```

#### Make climate vectors for biannual changes between surveys
```{r echo=TRUE}
trend <- FALSE
start_time <- 2012
end_time <- 2015
temp_threshold <- c(0.1)
do_threshold <- c(0.1)
```

# OR

#### Make climate vectors for average changes and lower threshold values
```{r echo=TRUE}
# trend <- TRUE
# temp_trend_threshold <- c(0.25)
# do_trend_threshold <- c(0.2)
```

#### Set Region 
```{r echo=TRUE}
# region <- "West Coast Vancouver Island"
region <- "both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

Run all subsequent code...
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
# library(gfplot)
# library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2, 2)
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}
```

```{r, include=FALSE}
# Set default cell sizes and scale
input_cell_size <- 2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  delta_threshold <- temp_threshold
  optimal_values <- optimal_temp

  if (trend) {
    max_thresholds <- temp_trend_threshold
    start_time <- 2008 # change if can add back in prior 2008 data
    end_time <- NULL
    delta_t_total <- 6 # change if can add back in prior 2008 data
    indices <- trend_indices_temp
  } else {
    max_thresholds <- temp_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  delta_threshold <- -1 * (do_threshold)
  optimal_values <- optimal_do

  if (trend) {
    min_thresholds <- do_trend_threshold
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- do_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do and temperature") {
  variable_names <- c("do_est", "temp")

  if (trend) {
    min_thresholds <- c(do_trend_threshold, Inf)
    max_thresholds <- c(Inf, temp_trend_threshold)
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- c(do_threshold, Inf)
    max_thresholds <- c(Inf, temp_threshold)
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")

optimal_string <- paste0(optimal_values, collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
```

# Calculate climate vectors

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
})

if (!exists("vocc")) {
  pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }

  starttime <- Sys.time()
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = min_thresholds,
    max_thresholds = max_thresholds,
    round_fact = 10
  )
  endtime <- Sys.time()
  time_vocc <- round(starttime - endtime)

  vocc$var_1_min <- min_thresholds[1]
  vocc$var_2_min <- min_thresholds[2]
  vocc$var_1_max <- max_thresholds[1]
  vocc$var_2_max <- max_thresholds[2]

  saveRDS(vocc, file = paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
}
# glimpse(vocc)
```

Plot all possible vectors
```{r echo=FALSE}
if (climate == "temperature") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```

Filter vectors to match optimal range for chosen species
```{r}
vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds,
  cell_size = input_cell_size, dist_intercept = input_cell_size,
  max_dist = 120
)

if (length(vectors[, 2]) == 0) {
  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
  coefs <- other_coefs[1, ]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$model <- "NA"
  coefs$species <- spp
  coefs$lifestage <- "NA"
  coefs$climate <- climate
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  coefs <- rbind(other_coefs, coefs) %>% distinct()
  saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))

  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
}
# glimpse(vocc)
# glimpse(vectors)
```

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r echo=FALSE, warning=FALSE}

if (climate == "temperature") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vectors$temp_diff <- vectors$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vectors$do_diff <- vectors$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("DO of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```
# Add species biomass data

#### Get all biomass predictions 
```{r, echo=FALSE, warning=FALSE}
priors <- FALSE
# priors <- TRUE

covariates <- "+muddy+any_rock"

covs <- gsub("\\+", "-", covariates)

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

try({
  adult_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
  ))
})

try({
  imm_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
  ))
})

try({
  predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
  ))
})

if (!exists("adult_predictions")) adult_predictions <- NULL
if (!exists("imm_predictions")) imm_predictions <- NULL
if (!exists("predictions")) predictions <- NULL

biomass_list <- list(
  adult = adult_predictions,
  imm = imm_predictions,
  total = predictions
)

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]

# glimpse(biomass_list)
```

#### Extract biomass values to match spatial and temporal range of climate vectors

```{r}
vocc_by_sp <- lapply(biomass_list, function(i) {
  biomass_change <- make_trend_data(i,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )
  # browser()
  vectors <- vectors %>%
    mutate(
      start_cell = icell,
      vector_id = paste0(id, "_to_", tid),
      distance2 = distance * distance,
      source_end_value = e, 
      target_value = target_values
    )

  start <- biomass_change %>%
    rename(start_density = est_exp_s, end_density = est_exp_e) %>%
    dplyr::select(icell, x, y, start_density, end_density)

  vector_source <- left_join(vectors, start, by = c("icell", "x", "y")) %>%
    mutate(
      cell_type = "source", end_climate = source_end_value,
      source_density = start_density, source_end_density = end_density
    )

  end <- biomass_change %>%
    rename(
      start_density = est_exp_s,
      end_density = est_exp_e,
      target_X = x, target_Y = y
    ) %>%
    dplyr::select(start_density, end_density, target_X, target_Y)

  vector_target <- left_join(vectors, end, by = c("target_X", "target_Y")) %>%
    mutate(cell_type = "target", end_climate = target_value)

  source_density <- biomass_change %>%
    rename(
      source_density = est_exp_s,
      source_end_density = est_exp_e
    ) %>%
    dplyr::select(source_density, source_end_density, x, y)

  vector_target <- left_join(vector_target, source_density, by = c("x", "y"))

  # combine source and target cell data
  vocc_by_sp <- rbind(vector_source, vector_target) %>%
    mutate(cell_type = as.factor(cell_type)) %>%
    filter(!is.na(icell)) %>%
    filter(!is.na(source_density)) %>%
    filter(!is.na(end_density)) %>%

    # remove potential source cells with lowest 10% quantile of fish biomass
    filter(source_density > quantile(source_density, 0.1)) %>%

    # calculate percent change in biomass
    mutate(
      log_change = log(end_density / start_density),
      perc_change = end_density / start_density,
      log_source_density = source_density
    )
  # browser()
  end_climate <- purrr::map_df(
    strsplit(vocc_by_sp$end_climate, " "),
    function(x) data.frame(t(as.numeric(x)))
  )

  names(end_climate) <- paste0("end_", variable_names)

  vocc_by_sp <- cbind(vocc_by_sp, end_climate) %>%
    rename(end_climate_chr = end_climate)

  # get mean and sd for densities without cell duplication
  icells <- vocc_by_sp %>% distinct(icell, .keep_all = TRUE)

  mean.log.density <- mean(icells$log_source_density, na.rm = TRUE)
  sd.log.density <- sd(icells$log_source_density, na.rm = TRUE)
  mean.change <- mean(icells$perc_change, na.rm = TRUE)
  sd.change <- sd(icells$perc_change, na.rm = TRUE)

  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])

    icells <- mutate(
      icells,
      (!!mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j], "_e"))), na.rm = TRUE),
      (!!sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j], "_e"))), na.rm = TRUE)
    )
  }

  # calculate standardized density variables
  vocc_by_sp <- vocc_by_sp %>%
    mutate(
      std_source_density = (log_source_density - mean.log.density) / (sd.log.density * 2),
      std_source_density2 = (log_source_density - mean.log.density) / (sd.log.density * 2)^2,
      mean_source_density = mean.log.density, sd_density = sd.log.density,
      mean_change = mean.change, sd_change = sd.change
    )

  # calculate standardized climate variables
  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])
    std.var <- paste0("std.", variable_names[j])
    std.var2 <- paste0("std.", variable_names[j], "2")

    vocc_by_sp <- mutate(
      vocc_by_sp,
      (!!mean.var) := icells[1, mean.var],
      (!!sd.var) := icells[1, sd.var],
      (!!std.var) := (UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var]) / (icells[1, sd.var]*2),
      (!!std.var2) := ((UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var]) / icells[1, sd.var])^2
    )
  }

  # rename spatial coordinates for inclusion in models and plots without naming conflicts
  vocc_by_sp <- vocc_by_sp %>%
    rename(X = x, Y = y)

  vocc_by_sp
})
```

## Adult plots

Difference between target and source cells split by nearest cells and further cells
```{r}
targets_only <- vocc_by_sp[["adult"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)

ggplot(filter(targets_only, distance < 2), aes(change_diff, fill = "green")) +
  geom_histogram(bins = 60) +
  geom_histogram(
    data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"),
    bins = 60, alpha = 0.5, inherit.aes = FALSE
  ) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))

ggplot(filter(targets_only, distance < 2), aes(change_diff)) +
  geom_density(aes(fill = "green")) +
  geom_density(data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"), alpha = 0.5, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))
```

Difference between target and source cells split by starting source biomass
```{r}
targets_only <- vocc_by_sp[["adult"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)

range(targets_only$source_density)

ggplot(targets_only, aes(change_diff)) +
  geom_histogram(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), bins = 60, alpha = 0.9, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))

ggplot(targets_only, aes(change_diff)) +
  geom_density(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), alpha = 0.9, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), alpha = 0.3, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), alpha = 0.6, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))
```

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["adult"]]

mature_plot <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = UQ(rlang::sym(paste0("end_", j))), y = end_density, colour = cell_type, size = start_density,
    group = cell_type
  )) +
    geom_point(alpha = 0.5) +
    geom_smooth(aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
      # method="lm", formula = y ~ x + I(x^2) + I(x^3),
      inherit.aes = FALSE
    ) + # + I(x^3)
    # scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = log10) +
    xlab(paste("Final value of", j, "for cells")) +
    ylab(paste("Final biomass")) +
    scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.7)) +
    ggtitle(paste("Mature", species, "for", region, "and", climate, "based VOCC"))
})
print(mature_plot)
```

Check pattens in distance and end climate
```{r}
data <- vocc_by_sp[["adult"]]
# data <- filter(vocc_by_sp[["adult"]],  distance < 90) # distance > 1 &

mature_plot2 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = (distance + 1), y = log_change,
    group = cell_type, size = start_density,
    colour = UQ(rlang::sym(paste0("end_", j)))
  )) +
    geom_point(alpha = 0.5) +
    # geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method = "lm", se = FALSE) + # formula = y~x+I(x^2),
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    facet_wrap(~cell_type) + # , scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j, "")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) +
    ggtitle(paste("Mature", species, "for", region, "and", climate, "based VOCC"))
})
mature_plot2
```

Check pattens in log source biomass and end climate
```{r warnings=FALSE}
data <- vocc_by_sp[["adult"]]
# data <- filter(vocc_by_sp[["adult"]], distance < 50)
# data <- filter(vocc_by_sp[["adult"]], distance > 4 & distance < 50)

mature_plot3 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = source_density, y = log_change,
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1 / (distance + input_cell_size), # size = 1/distance,
    group = cell_type
  )) +
    geom_jitter(width = 0.01) +
    geom_smooth(method = "lm", se = FALSE) +
    facet_wrap(~cell_type) + # , scales = "free") +
    scale_alpha_continuous(guide = "none") +
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    # scale_y_continuous(trans = sqrt) +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Source density (dark points are closer to source)")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) +
    ggtitle(paste("Mature", species, "for", region, "and", climate, "based VOCC"))
})
mature_plot3
```

## Adult models

### Boosted regression
```{r}
data <- vocc_by_sp[["adult"]]

gbm_formula <- paste0("log_change ~ std_source_density + distance + cell_type + X + Y") # + cell_type * distance2

for (j in (variable_names)) {
  gbm_formula <- paste0(gbm_formula, " + std.", j)
}

mboost <- gbm::gbm(
  as.formula(gbm_formula),
  data = data,
  distribution = "gaussian",
  n.trees = 2000,
  interaction.depth = 3,
  shrinkage = 0.02
)
```

```{r}
summary(mboost)
```

```{r}
plot(mboost, 3)
plot(mboost, 1)
plot(mboost, 2)
plot(mboost, 6)
if (length(variable_names) > 1) plot(mboost, 7)
if (length(variable_names) > 2) plot(mboost, 8)

plot(mboost, c(1, 6, 3))
if (length(variable_names) > 1) plot(mboost, c(1, 7, 3))
if (length(variable_names) > 2) plot(mboost, c(1, 8, 3))

plot(mboost, c(2, 6, 3))
if (length(variable_names) > 1) plot(mboost, c(2, 7, 3))
if (length(variable_names) > 2) plot(mboost, c(2, 8, 3))
```

Other gbm plots
```{r eval=FALSE}
plot(mboost, c(1, 2))
plot(mboost, c(1, 3))
plot(mboost, c(1, 6))

if (length(variable_names) > 1) plot(mboost, c(1, 7))
if (length(variable_names) > 2) plot(mboost, c(1, 8))

plot(mboost, c(2, 3))
plot(mboost, c(2, 6))
if (length(variable_names) > 1) plot(mboost, c(2, 7))
if (length(variable_names) > 2) plot(mboost, c(2, 8))

plot(mboost, c(4, 5))
```

### Regular mixed-model

```{r}
data <- vocc_by_sp[["adult"]] # %>% filter(distance < 50)
glimpse(data)

length(unique(data$icell))
length(unique(data$vector_id))

model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + cell_type * std_source_density2") # + cell_type * distance2

for (j in (variable_names)) {
  model_formula <- paste0("", model_formula, " + std.", j, "+ std.", j, "2")
}

model_formula <- paste0("", model_formula, "+ (1 | vector_id) + (1 | icell)")

test <- lmerTest::lmer(as.formula(model_formula), REML = F, data = data)

summary(test)
# plot(test2)

# library(sjPlot)
# library(sjlabelled)
# library(sjmisc)
sjPlot::plot_model(test, title = paste("Log adult", species, "biomass change"))
```

```{r}
beta <- lme4::fixef(test)
se <- arm::se.fixef(test)
var <- names(beta)

coefs <- as.data.frame(cbind(beta, se))
coefs$ci <- coefs$se * 1.96
var[1] <- "intercept"
row.names(coefs) <- NULL
coefs <- cbind(var, coefs)
coefs$model <- model_formula
coefs$species <- spp
coefs$lifestage <- "adult"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

rm(other_coefs)
try({
  other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}

saveRDS(coefs, file = paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
```

Robust lmm
```{r eval=FALSE}
library(lme4)
data <- vocc_by_sp[["adult"]]

test3 <- robustlmm::rlmer(as.formula(model_formula), data = data)

summary(test3)
plot(test3, ask = FALSE)
```

Maybe try GAM?
```{r}
# library()
#
# model_formula <- paste0("log_change ~ cell_type + s(std_source_density) + s(distance) ") # + cell_type * distance2
#
# for (j in (variable_names) ){
#   model_formula <- paste0("", model_formula, " + s(std.", j, ") + "
#   )
# }
#
# model_formula <- paste0("", model_formula, "+ (1 | vector_id) + (1 | icell)")
#
#
# test2 <- gamm4::....(as.formula(model_formula), REML = F, data = data)
#
```

### sdmTMB

```{r}
data <- vocc_by_sp[["adult"]]
# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$log_change, 0.9995)
#
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

xy <- as.data.frame(cbind(data$X, data$Y))

if (region == "both odd year surveys") {
  n_knots <- 200
}

if (region == "West Coast Vancouver Island") {
  n_knots <- 100
}

# if (region == "West Coast Haida Gwaii") {
#   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
# }

if (n_knots > nrow(unique(xy))) {
  n_knots <- nrow(unique(xy))
}

spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = n_knots)
sdmTMB::plot_spde(spde)
```

```{r}
data <- vocc_by_sp[["adult"]]

tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance") # + cell_type * distance2 cell_type * std_source_density2 +

for (j in (variable_names)) {
  tmb_formula <- paste0("", tmb_formula, " + std.", j, "+ std.", j, "2")
}

mtmb <- sdmTMB(
  data = data,
  as.formula(tmb_formula),
  spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
  silent = TRUE
)
mtmb
```

```{r}
p <- predict(mtmb)
p$residuals <- residuals(mtmb)
# idata$resids <- p$residuals
# glimpse(p)
ggplot(p, aes(est, residuals)) +
  geom_point(alpha = 0.4) +
  ylim(-8, 8) +
  geom_smooth()

plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed() +
    gfplot::theme_pbs() +
    scale_fill_gradient2(trans=fourth_root_power) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(p, "est")
plot_map_raster(p, "est_rf")
plot_map_raster(p, "omega_s")
plot_map_raster(p, "residuals")
```

# Immature

### Immature plots

Difference between target and source cells split by nearest cells and further cells
```{r}
targets_only <- vocc_by_sp[["imm"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)

ggplot(filter(targets_only, distance < 2), aes(change_diff, fill = "green")) +
  geom_histogram(bins = 60) +
  geom_histogram(
    data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"),
    bins = 60, alpha = 0.5, inherit.aes = FALSE
  ) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))

ggplot(filter(targets_only, distance < 2), aes(change_diff)) +
  geom_density(aes(fill = "green")) +
  geom_density(data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"), alpha = 0.5, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))
```

Difference between target and source cells split by starting source biomass
```{r}
targets_only <- vocc_by_sp[["imm"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)
# range(targets_only$source_density)

ggplot(targets_only, aes(change_diff)) +
  geom_histogram(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), bins = 60, alpha = 0.9, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))

ggplot(targets_only, aes(change_diff)) +
  geom_density(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), alpha = 0.9, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), alpha = 0.3, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), alpha = 0.6, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))
```

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["imm"]]
# glimpse(data)

immature_plot <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = UQ(rlang::sym(paste0("end_", j))), y = (end_density),
    colour = cell_type, size = start_density,
    group = cell_type
  )) +
    geom_point(alpha = 0.5) +
    geom_smooth(aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
      # method="lm", formula = y ~ x + I(x^2) + I(x^3),
      inherit.aes = FALSE
    ) + # + I(x^3)
    # scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = log10) +
    xlab(paste("Final value of", j, "for cells")) +
    ylab(paste("Final biomass")) +
    scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) +
    ggtitle(paste("Immature", species, "for", region, "and", climate, "based VOCC"))
})
immature_plot
```

Check pattens in distance and end climate
```{r}
data <- vocc_by_sp[["imm"]]
# data <- filter(vocc_by_sp[["imm"]],  distance < 90) # distance > 1 &
# data <- filter(vocc_by_sp[["imm"]], distance > 1 ) #  &

immature_plot2 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = (distance + 1), y = log_change,
    group = cell_type, size = start_density,
    colour = UQ(rlang::sym(paste0("end_", j)))
  )) +
    geom_point(alpha = 0.5) +
    # geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method = "lm") + # , formula = y~x+I(x^2)
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    facet_wrap(~cell_type) + # , scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j, "")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.7)) +
    ggtitle(paste("Immature", species, "for", region, "and", climate, "based VOCC"))
})
immature_plot2
```

Check pattens in log source biomass and end climate
```{r warnings=FALSE}
data <- vocc_by_sp[["imm"]]
# data <- filter(vocc_by_sp[["imm"]], distance < 50)
# data <- filter(vocc_by_sp[["imm"]], distance > 4 & distance < 50)

immature_plot3 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = source_density, y = log_change,
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1 / (distance + input_cell_size), # size = 1/distance,
    group = cell_type
  )) +
    geom_jitter(width = 0.01) +
    geom_smooth(method = "lm") +
    facet_wrap(~cell_type) + # , scales = "free") +
    scale_alpha_continuous(guide = "none") +
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Source density (dark points are closer to source)")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) +
    ggtitle(paste("Immature", species, "for", region, "and", climate, "based VOCC"))
})
immature_plot3
```

## Immature models

### Boosted regression
```{r}
idata <- vocc_by_sp[["imm"]]

im <- gbm::gbm(
  as.formula(gbm_formula),
  data = idata,
  distribution = "gaussian",
  n.trees = 2000,
  interaction.depth = 3,
  shrinkage = 0.02
)
```

```{r}
summary(im)
```

```{r}
plot(im, 3)
plot(im, 1)
plot(im, 2)
plot(im, 6)
if (length(variable_names) > 1) plot(im, 7)
if (length(variable_names) > 2) plot(im, 8)

plot(im, c(1, 6, 3))
if (length(variable_names) > 1) plot(im, c(1, 7, 3))
if (length(variable_names) > 2) plot(im, c(1, 8, 3))

plot(im, c(2, 6, 3))
if (length(variable_names) > 1) plot(im, c(2, 7, 3))
if (length(variable_names) > 2) plot(im, c(2, 8, 3))
```

Other gbm plots
```{r eval=FALSE}
plot(im, c(1, 2))
plot(im, c(1, 3))
plot(im, c(1, 6))

if (length(variable_names) > 1) plot(im, c(1, 7))
if (length(variable_names) > 2) plot(im, c(1, 8))

plot(im, c(2, 3))
plot(im, c(2, 6))
if (length(variable_names) > 1) plot(im, c(2, 7))
if (length(variable_names) > 2) plot(im, c(2, 8))

plot(im, c(4, 5))
```

### Regular mixed-model

```{r}
idata <- vocc_by_sp[["imm"]]

itest <- lmerTest::lmer(as.formula(model_formula), REML = F, data = idata)

summary(itest)

# library(sjlabelled)
# library(sjmisc)
sjPlot::plot_model(itest, title = paste("Log immature", species, "biomass change"))
```

```{r}
beta <- lme4::fixef(itest)
se <- arm::se.fixef(itest)
var <- names(beta)

coefs <- as.data.frame(cbind(beta, se))
coefs$ci <- coefs$se * 1.96
var[1] <- "intercept"
row.names(coefs) <- NULL
coefs <- cbind(var, coefs)
coefs$model <- model_formula
coefs$species <- spp
coefs$lifestage <- "immature"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

rm(other_coefs)
try({
  other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}

coeds <- coefs %>% distinct()

saveRDS(coefs, file = paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
```

Robust lmm
```{r eval=FALSE}
library(lme4)

test3 <- robustlmm::rlmer(as.formula(model_formula), data = idata)

summary(test3)
plot(test3, ask = FALSE)
```

### sdmTMB

```{r}
idata <- vocc_by_sp[["imm"]]

xy <- as.data.frame(cbind(idata$X, idata$Y))

if (region == "both odd year surveys") {
  n_knots <- 200
}

if (region == "West Coast Vancouver Island") {
  n_knots <- 100
}

# if (region == "West Coast Haida Gwaii") {
#   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
# }

if (n_knots > nrow(unique(xy))) {
  n_knots <- nrow(unique(xy))
}

ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = n_knots)
sdmTMB::plot_spde(ispde)
```

```{r}
idata <- vocc_by_sp[["imm"]]

# tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + distance2") # + cell_type * std_source_density2
#
# for (j in (variable_names) ){
#   tmb_formula <- paste0("", tmb_formula, " + std.", j, "+ std.", j, "2"
#   )
# }

itmb <- sdmTMB(
  data = idata,
  as.formula(tmb_formula),
  spde = ispde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
  silent = TRUE
)
itmb
```

```{r}
ip <- predict(itmb)
ip$residuals <- residuals(itmb)
# idata$resids <- p$residuals
# glimpse(p)
ggplot(ip, aes(est, residuals)) +
  geom_point(alpha = 0.4) +
  ylim(-8, 8) +
  geom_smooth()

plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed() +
    gfplot::theme_pbs() +
    scale_fill_gradient2(trans=fourth_root_power) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(ip, "est")
plot_map_raster(ip, "est_rf")
plot_map_raster(ip, "omega_s")
plot_map_raster(ip, "residuals")
```

```{r}
m <- mtmb

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta", "se")
coefs$ci <- coefs$se * 1.96
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL

coefs <- cbind(var, coefs)
# coefs$var
# coefs$beta
# coefs$se
coefs$model <- model_formula
coefs$species <- spp
coefs$lifestage <- "adult"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

rm(other_coefs)
try({
  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}
saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
```

```{r}
m <- itmb

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta", "se")
# coefs$beta <- signif(coefs$beta, digits = 3)
# coefs$se <- signif(coefs$se, digits = 3)
coefs$ci <- signif(coefs$se * 1.96, digits = 3)
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL
coefs <- cbind(var, coefs)
coefs$species <- spp
coefs$lifestage <- "immature"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string
coefs$model <- model_formula

try({
  other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-sdmTMB-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}

# glimpse(coefs)

saveRDS(coefs, file = paste0("data/vocc-", climate_string, "-by-spp-sdmTMB-coefs.RDS"))
```

```{r eval=FALSE}
lmer_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))

lmer_coefs$beta_upper <- lmer_coefs$beta + lmer_coefs$ci
lmer_coefs$beta_lower <- lmer_coefs$beta - lmer_coefs$ci

# View(lmer_coefs)

beta_target <- lmer_coefs %>%
  filter(ssid_string == "1n3") %>%
  filter(var == "cell_typetarget")

View(beta_target)
```

### 2020-02-06 analysis/VOCC/old-code/dist-vocc-temp-do-4.Rmd
organize old code

---
title: "Distanced-based VOCC analysis for WCVI"
author: "Philina English"
date: '2019-07-11'
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfranges)
```

Load predicted values
```{r}
pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))
```

Define subset of the spatial range 
```{r eval=FALSE}
predicted <- pred_temp_do %>% filter(ssid == 4)

predicted <- list(predicted, predicted)

do <- plot_facet_map(predicted[[1]], "do_est", transform_col = no_trans) +
  labs(fill = "ml/L") +
  ggtitle("Bottom DO")
print(do)
```

Biannual changes starting 2005 and ending 2017
```{r}
out1 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"),
  ssid = c(4),
  start_time = 2008,
  end_time = 2010,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),  
  match_logic = c(">=", "=="),
  thresholds = c(1, 1) # vector of plus/minus threshold(s) to define climate match.
)
out1$do_est.units_per_decade <- out1$do_est.units_per_decade / 5
out1$temp.units_per_decade <- out1$temp.units_per_decade / 5

gvocc1a <- plot_vocc(out1,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # viridis_option = "C",
  # viridis_dir = -1#,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc1b <- plot_vocc(out1,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc1 <- gridExtra::grid.arrange(gvocc1b, gvocc1a, nrow = 1, top = "VOCC vectors to maintain <1°C or 1 ml/L change between 2008 and 2010")
```

```{r}
out2 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"),
  ssid = c(4),
  start_time = 2010,
  end_time = 2012,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
    match_logic = c(">=", "=="),
  thresholds = c(1, 1) # vector of plus/minus threshold(s) to define climate match.
)
out2$do_est.units_per_decade <- out2$do_est.units_per_decade / 5
out2$temp.units_per_decade <- out2$temp.units_per_decade / 5

gvocc2a <- plot_vocc(out2,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc2b <- plot_vocc(out2,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc2 <- gridExtra::grid.arrange(gvocc2b, gvocc2a, nrow = 1, top = "VOCC vectors to maintain <1°C or 1 ml/L change between 2010 and 2012")
```

`
```{r}
out3 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"),
  ssid = c(4),
  start_time = 2012,
  end_time = 2014,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
    match_logic = c(">=", "=="),
  thresholds = c(1, 1) # vector of plus/minus threshold(s) to define climate match.
)
out3$do_est.units_per_decade <- out3$do_est.units_per_decade / 5
out3$temp.units_per_decade <- out3$temp.units_per_decade / 5

gvocc3a <- plot_vocc(out3,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc3b <- plot_vocc(out3,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc3 <- gridExtra::grid.arrange(gvocc3b, gvocc3a, nrow = 1, top = "VOCC vectors to maintain <1°C or 1 ml/L change between 2012 and 2014")
```

```{r}
out4 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"),
  ssid = c(4),
  start_time = 2014,
  end_time = 2016,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
    match_logic = c(">=", "=="),
  thresholds = c(1, 1) # vector of plus/minus threshold(s) to define climate match.
)
out4$do_est.units_per_decade <- out4$do_est.units_per_decade / 5
out4$temp.units_per_decade <- out4$temp.units_per_decade / 5

gvocc4a <- plot_vocc(out4,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc4b <- plot_vocc(out4,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc4 <- gridExtra::grid.arrange(gvocc4b, gvocc4a, nrow = 1, top = "VOCC vectors to maintain <1°C or 1 ml/L change between 2014 and 2016")
```

```{r}
out5 <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"),
  ssid = c(4),
  start_time = 2016,
  end_time = 2018,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 2,
  delta_t_step = 2,
  indices = c(1, 2),
    match_logic = c(">=", "=="),
  thresholds = c(1, 1) # vector of plus/minus threshold(s) to define climate match.
)
out5$do_est.units_per_decade <- out5$do_est.units_per_decade / 5
out5$temp.units_per_decade <- out5$temp.units_per_decade / 5

gvocc5a <- plot_vocc(out5,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L DO",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # viridis_option = "C",
  # viridis_dir = -1,
  # transform_col = fourth_root_power,
  # raster_limits = c(0, 6)
)

gvocc5b <- plot_vocc(out5,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 50,
  NA_label = ".",
  fill_col = "temp.units_per_decade",
  fill_label = "°C",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE # ,
  # transform_col = fourth_root_power,
  # raster_limits = c(5, 13)
)

gvocc5 <- gridExtra::grid.arrange(gvocc5b, gvocc5a, nrow = 1, top = "VOCC vectors to maintain <1°C or 1 ml/L change between 2016 and 2018")
```

```{r}
png(
  file = "figs/biannual_temp-do_change-4.png", # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 8, # The width of the plot in inches
  height = 17
) # The height of the plot in inches

gridExtra::grid.arrange(gvocc1, gvocc2, gvocc3, gvocc4, gvocc5, nrow = 5, top = grid::textGrob("VOCC vectors for <1 ml/L biannual changes in bottom DO")) # ,gp=gpar(fontsize=20,font=3))
dev.off()
```

Cumulative change between last two decades 
VOCC for 1 ml/L change and 1 degree change
```{r}
out0.5d <- make_vector_data(predicted,
  variable_names = c("do_est", "temp"),
  ssid = c(4),
  start_time = 2008,
  skip_time = 2016,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 6,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2),
  match_logic = c(">=", "<="),
  thresholds = c(0.5, 1) # vector of plus/minus threshold(s) to define climate match.
)
```

```{r}
gvocc <- plot_vocc(out0.5d,
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 100,
  fill_col = "1_e",
  fill_label = "mean DO",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  viridis_option = "C",
  viridis_dir = -1,
  transform_col = no_trans # ,
)
gvocc <- gvocc + ggtitle("VOCC <0.5 ml/L DO/>1°C (excludes 2016)")
gvocc
```

```{r}
gtrend1 <- plot_vocc(out0.5d,
  fill_col = "temp.units_per_decade",
  fill_label = "°C ",
  # high_fill = "Steel Blue 4",
  # low_fill = "Red 3",
  raster_alpha = 1,
  vec_aes = NULL,
  viridis_option = "A",
  transform_col = no_trans
)
gtrend1 <- gtrend1 + ggtitle("Temperature trend 2008-2018")
gtrend1
```

```{r}
gtrend <- plot_vocc(out0.5d,
  fill_col = "do_est.units_per_decade",
  fill_label = "ml/L",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = no_trans
)
gtrend <- gtrend + ggtitle("Bottom DO trend 2008-2018")
gtrend
```

```{r}
png(
  file = "figs/temp-do-2008-2018-4.png", # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 10, # The width of the plot in inches
  height = 5
) # The height of the plot in inches

gridExtra::grid.arrange(gtrend1, gtrend, gvocc, nrow = 1)
dev.off()
```

### 2020-02-06 analysis/VOCC/old-code/dist-vocc-temp.Rmd
organize old code

---
title: "dist-vocc-temp"
author: "Philina English"
date: '2019-05-14'
output: html_document
---

```{r setup, include=FALSE}
# install.packages("yaImpute") # install package for k-nearest neighbour (kNN) search
# install.packages("raster") # required for calcslope + all raster functions
# install.package(gfplot) # required for some plotting and data functions

library(dplyr)
library(ggplot2)
library(sdmTMB)
```

Retrieve raw data
```{r}
all_depth <- readRDS("analysis/tmb-sensor-explore/data/dat-sensor-trawl-all-depth.rds")

# add and scale predictors after filtering for survey(s) of interest
data <- all_depth$data %>%
  filter(depth_max>10) %>% # removes data if sensor stayed too near surface
# dplyr::filter(year > 2003) %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

data <- data[data$fishing_event_id != 481861, ]
# add any new predictors
# data <- data %>%
#   dplyr::mutate(
#   DOY = lubridate::yday(date),
#   shallow = ifelse(depth > 35, 0, 1)
# )
```

Create prediction grids for each survey area so that only years with data are included
```{r}
# choose base year(s) to create grid from
dummy_year <- c(2004, 2005)

# create grids for each ssid separately 
ssid <- 1
survey_abbrev <- "SYN QCS"
nd_1 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_1$year)

ssid <- 3
survey_abbrev <- "SYN HS"
nd_3 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_3$year)

ssid <- 4
survey_abbrev <- "SYN WCVI"
nd_4 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_4$year)

ssid <- 16
survey_abbrev <- "SYN WCHG"
nd_16 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year = 2006)
unique(nd_16$year)

nd <- rbind(nd_1, nd_3, nd_4, nd_16)
# nd <- nd_4 # for faster test option
nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))

```

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
# # choose the spatial range to model
dat <- data # %>% filter(ssid == 4)
nd <- nd # %>% filter(ssid == 4)

spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 500)
sdmTMB::plot_spde(spde)

m_temp <- sdmTMB::sdmTMB(dat,
  temperature_c ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = gaussian(link = "identity"),
  ar1_fields = TRUE, # maybe TRUE is better for all areas combined?
  include_spatial = TRUE,
  silent = FALSE
)

# stopifnot(m_temp$model$convergence == 0L)
# m_temp
# # Warning messages:
# #   1: In doTryCatch(return(expr), name, parentenv, handler) :
# #   restarting interrupted promise evaluation
# 
saveRDS(m_temp, file = "data/m_temp_allyears.rds")
```

Calculate predicted values
```{r}
m_temp <- readRDS("analysis/data/m_temp_allpost2003.rds")

predictions <- predict(m_temp, newdata = nd)

plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

p <- plot_map(predictions , "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```

```{r}
# choose the spatial range to build raster on
predicted1 <- predictions %>% filter(ssid == 1) 
predicted3 <- predictions %>% filter(ssid == 3) 
predicted4 <- predictions %>% filter(ssid == 4) 
predicted16 <- predictions %>% filter(ssid == 16)

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted3, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-hs.rds")

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted1, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-qcs.rds")

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted4, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-wcvi.rds")

# scale_fac = 2 means that the raster is reprojected to 2 X original grid (2 km)
rbrick <- make_raster_brick(predicted16, scale_fac = 2)
saveRDS(rbrick, file = "analysis/data/rbrick-temp-wchg.rds")
```

```{r}

predicted <- predictions %>% filter(ssid == 1)

vector_data <- function(climate_data,
  input_cell_size = 2, 
  scale_fac = 2,
  indices = c(1, 1, 1, 2, 2, 3, 3, 3),
  thresholds = c(0.75),
  delta_t = 10,

) {

rbrick <- make_raster_brick(climate_data, scale_fac = scale_fac)

slopedat <-  calcslope(temp_rbrick_qcs) # vocc::calcslope for comparison
mnraster_brick <- raster::stackApply(rbrick, indices = indices, fun = mean)
start_raster <- mnraster_brick[[1]]
end_raster <- mnraster_brick[[3]]

# make sparate named lists containing climate rasters or dataframes
# element names should describe which variable they contain
# the variable_names vector will contain what the column or layer within each element is called

# data with just one climate variable
start_data <- list(var_1 = start_raster)
end_data <- list(var_1 = end_raster)

  out <- dist_based_vocc(
  start_data = start_data,
  end_data = end_data,
  x = "x",
  y = "y",
  variable_names = c("index_1"),
  thresholds = thresholds,
  cell_size =  input_cell_size*scal_fac,
  delta_t = delta_t,
  raster = TRUE
  )

out1_qcs <- left_join(out1_qcs, slopedat_qcs, by = c("x", "y")) %>% select(-icell)
out1_qcs$C_per_decade <- out1_qcs$slope 
out1_qcs$km_per_decade <- out1_qcs$distance * 10 / 10 # dived by delta_t
head(out1_qcs)
#saveRDS(out1, file = "analysis/data/simple-dist-vocc-qcs1.rds")

}
```

### 2020-02-06 analysis/VOCC/old-code/dist-vocc.Rmd
organize old code

--- 2019-06-12 analysis/VOCC/dist-vocc.Rmd
@@
 # # choose the spatial range to model
-dat <- data %>% dplyr::filter(year > 2003) # %>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
+dat <- data #%>% dplyr::filter(year > 2003) # %>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
 #nd <- nd_1 # %>% filter(ssid == 4)

### 2020-02-06 analysis/VOCC/old-code/4-vocc-biotic.R

@@ -1,0 +1,513 @@
+getwd()
+setwd(here::here("/analysis/VOCC"))
+
+library("gfranges")
+library("dplyr")
+library("ggplot2")
+
+# species <- "Redbanded Rockfish"
+# species <- "Pacific Ocean Perch"
+# species <- "Pacific Cod"
+species <- "Arrowtooth Flounder"
+# species <- "Petrale Sole"
+
+
+spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
+covs <- "-tv-depth-only"
+
+multiyear <- TRUE
+
+#########################
+### CHOOSE SPATIAL EXTENT
+#########################
+
+model_ssid <- c(4)
+# model_ssid <- c(16)
+# model_ssid <- c(1,3)
+ssid_string <- paste0(model_ssid, collapse = "n")
+dist_intercept <- 0 # input_cell_size / 2
+input_cell_size <- 2
+scale_fac <- 2 # means that the raster is reprojected to _ X original grid (2 km)
+
+
+#################
+### LOAD DATA
+#################
+
+d_all <-readRDS(paste0("data/", spp,
+  "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds")) %>% filter(year > 2007) 
+
+nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select (X, Y, year, temp, do_est)
+
+d_all <- left_join(d_all, nd_all, by=c("X","Y", "year"))
+
+glimpse(d_all)
+
+#########################
+### SET TIME FRAME
+#########################
+
+if (multiyear){
+ #unique(d_all$year)
+ d <- d_all %>% filter(ssid %in% model_ssid)
+ length(unique(d$year))
+ years_sampled <- length(unique(d$year))
+ year_range <- (years_sampled -1)*2
+ start_year <- 2007
+ if (years_sampled==5) indices <- c(1, 2, 2, 2, 2)
+ if (years_sampled==6) indices <- c(1, 2, 2, 2, 2, 2)
+ 
+} else{ 
+d <- d_all %>% filter(year >= start_year) %>% filter(year <= start_year+3)
+unique(d$year)
+indices <- c(1,2)
+year_range <- 2
+}
+
+# end_time <- start_time + 2
+skip_time <- NULL
+delta_t_step <- year_range
[…]

### 2020-02-06 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
updates to temp and do prediction grids

@@
 m_temp_rw <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year),
+  temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
-  ar1_fields = TRUE, include_spatial = TRUE, # changed AR1_field to true
+  ar1_fields = TRUE, # changed AR1_field to true
+  include_spatial = TRUE, 
   time = "year", spde = spde, family = gaussian(link = "identity"),
@@
 ```{r}
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised")
+
 plot_map_raster <- function(dat, column = "est") {
@@
    scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3", limits = c(-1.5,1.5),
+ high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
     na.value = "red") +
@@
   axis.title.x = element_blank(), axis.title.y = element_blank())
+
+png(
+  file = "figs/temp-predictions-1n3.png", 
+  res = 400,
+  units = "in",
+  width = 8, 
+  height = 8
+) 
 st_1n3
+dev.off()
+
 ```
@@
      scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3", limits = c(-1.5,1.5),
+ high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
     na.value = "red") +

### 2020-02-06 analysis/tmb-sensor-explore/3-fit-do.Rmd
updates to temp and do prediction grids

@@
 ```{r}
-p_nd4 <- p_nd %>% filter(ssid==4) %>% filter(year!=2016)
+p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(year!=2016)
 pred_4 <- plot_map_raster(p_nd4, "do_est") +
@@
 ```{r}
-p_nd4 <- p_nd %>% filter(ssid==4) %>% filter(year!=2016)
+p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(year!=2016)
 st_4 <- plot_map_raster(p_nd4, "epsilon_st") +

### 2020-02-06 analysis/tmb-sensor-explore/climate-predictions.R

@@ -1,0 +1,23 @@
+#new temp and do prediction grid for vectors
+
+model1 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-temp-2008-on-used.rds")
+#model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-all-DOY-used.rds")
+#update model to exclude WCVI 2016 outliers
+model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds")
+
+nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
+nd <- filter(nd_all, year >2007) %>% filter(year<2019)
+pred_temp <- predict(model1, newdata = nd)
+predtemp <- pred_temp %>% mutate(temp_scaled = est, temp_scaled2 = est^2, depth_scaled3 = depth_scaled^3) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
+predtemp$DOY_scaled <- 0
+pred_do <- predict(model2, newdata = predtemp)
+pred_do$temp <- pred_do$temp_scaled*1.296 + 6.49
+
+
+pred_do$log_do <- pred_do$est
+pred_do$do_est <- exp(pred_do$log_do)
+
+# may want to filter estimates for depths less than 15 m
+# pred_do <- pred_do %>% filter(depth>15)
+
+saveRDS(pred_do, file="~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-new.rds")

### 2020-02-07 analysis/VOCC/3-climate-model-temp-only.Rmd
more organizing of vocc folder

--- 2019-10-09 analysis/VOCC/3-climate-models-fixed-both.Rmd
@@
 params: 
-    species: "Pacific Ocean Perch"
+    species: "Flathead Sole"
     region: "All synoptic surveys"
@@
     priors: FALSE
-    knots: 500
-    covariates: "+trawled" # additional non-climate variables
-    covs: "-log-both-trawled" # string describing model covariates
+    covariates: "" #"+trawled" # additional non-climate variables #+as.factor(ssid)
+    # covs: "-tv-depth-past" #
+    # year_cutoff: 2013
+    covs: "-do-tv-depth-all" #"-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
+    year_cutoff: NULL
+    # covariates: "+trawled" # additional non-climate variables
+    # covs: "-log-temp-trawled" # string describing model covariates
 ---
@@
 priors <- params$priors
-knots <- params$knots
 
@@
 paste("priors =", priors)
-paste("knots =", knots)
 
@@
 
+  data <- data %>% filter(raw_do > 0.28) %>% filter(raw_do < 7.06) # 0.005 and 0.995
+  data <- data %>% filter(temperature_c > 3.07) %>% filter(temperature_c < 11.3) # 0.005 and 0.995
+
   if (any(names(data) == "adult_density")) { 
@@
   if (region == "All synoptic surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
+    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
   }
@@
     adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""
+      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""
       ))
@@
       family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
+      #ar1_fields = TRUE,
+      #include_spatial = FALSE,
+      reml = TRUE,
       enable_priors = priors,
@@
     imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""
+      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2", covariates, ""
       ))
@@
       family = tweedie(link = "log"),
-      ar1_fields = FALSE,
-      include_spatial = TRUE,
+      #ar1_fields = TRUE,
+      #include_spatial = FALSE,
+      reml = TRUE,
       enable_priors = priors,

### 2020-02-07 analysis/VOCC/BACI/4-vocc-BACI-multiyear.Rmd
more organizing of vocc folder

--- 2019-11-13 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
@@
     species: "Pacific Cod" 
-    #region: "both odd year surveys"
-    region: "West Coast Vancouver Island"
+    #species: "Pacific Ocean Perch"
+    #species: "Redbanded Rockfish"
+    region: "both odd year surveys"
+    #region: "West Coast Vancouver Island"
     age: "mature"
     biomass_threshold: "50"
-    climate: "temperature"
-    threshold: 0.5
+    climate: "do" # if DO use "do"
+    threshold: 0.2
 editor_options: 
@@
     if (climate == "do") {
+
       vocc$diff <- vocc$units_per_decade / 10 * delta_t_total
@@
 covs <- "-trawled-ssid"
+covs <- "-trawled-ssid-reml"
 rm(biomass_predictions)
@@
 # depth.st + omega_s + muddy.st + any_rock.st +
-for (j in (variable_names)) {
-  match_formula <- paste0("", match_formula, " + ", j, "_s.st")
-  start_value.st <- paste0("", j, "_s.st")
-  start_value <- paste0("", j, "_s")
-
-  alldata <- alldata %>% mutate((!!start_value.st) :=
-    arm::rescale(alldata[[start_value]]))
-  # (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
-}
+# for (j in (variable_names)) {
+#   match_formula <- paste0("", match_formula, " + ", j, "_s.st")
+#   start_value.st <- paste0("", j, "_s.st")
+#   start_value <- paste0("", j, "_s")
+# 
+#   alldata <- alldata %>% mutate((!!start_value.st) :=
+#     arm::rescale(alldata[[start_value]]))
+#   # (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
+# }
 

### 2020-02-07 analysis/VOCC/BACI/4-vocc-BACI-render.Rmd
more organizing of vocc folder

---
title: "BACI of biomass change for VOCC source cells for all years"
author: "Philina English"
date: '2019-08-24'
output: html_document
params: 
    species: "Yelloweye Rockfish" 
    region: "West Coast Vancouver Island"
    age: "mature"
    biomass_threshold: "25"
    climate: "do"
    threshold: c(0.25) 
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("dplyr")
library("ggplot2")
library("gfplot")
library("sdmTMB")
library("gfranges")
```

#### Set parameters
```{r echo=TRUE}
species <- params$species
age <- params$age
region <- params$region
biomass_threshold <- params$biomass_threshold
climate <- params$climate
```

#### Set default cell sizes and scale
```{r, echo=FALSE}
global_only <- TRUE
input_cell_size <- 2
dist_intercept <- input_cell_size/2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL
```

# Run all subsequent code...

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_years <- list("2009"=2009, "2011"=2011, "2013"=2013, "2015"=2015) 
  # maybe remove 2011 & 2015 once we can add "2017"=2017 to avoid start and end values overlapping
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) 
  # prob need to remove 2010 & 2014
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) 
  # prob need to remove 2010 & 2014
  years <- NULL
}
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, include=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

root_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))

root_params <- root_params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))

if (age == "imm") {
  root_params <- filter(root_params, age == "imm")
} else {
  root_params <- filter(root_params, age != "imm")
  if (any(root_params$model_type == "split")) { root_params <- filter(root_params, model_type != "total") }
}

root_params <- root_params[root_params$species == species, ] 
glimpse(root_params)

if (global_only) {
 root_params <- root_params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
 root_params <- filter(root_params, global == "T")
} else {
 if (model_ssid == 4) {
   root_params <- root_params %>% mutate(survey4 = ifelse(grepl("4", model), "T", "F"))
   root_params <- filter(root_params, survey4 == "T")
 }
}

if (climate == "temperature") {

temp_threshold <- params$threshold

  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
  if (biomass_threshold == "50") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 }
  if (biomass_threshold == "25") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25 }
  .upper <- na.omit(.upper)

  upper_thresholds <- round(.upper[1], digits = 2)
   if (is.na(upper_thresholds)) {stop(print("Temp maximum not modelled"))}

  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model

  lower_thresholds <- NA
  lower_model <- NULL
  variable_names <- c("temp")

  lower_change <- c(Inf)
  upper_change <- temp_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do") {
  do_threshold <- params$threshold
  root_params <- filter(root_params, lower_do_25 > 0.4)

  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
  if (biomass_threshold == "50") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
  if (biomass_threshold == "25") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }

  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)
  if (is.na(lower_thresholds)) {stop(print("DO minimum not modelled"))}
  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model

  upper_thresholds <- NA
  upper_model <- NULL
  variable_names <- c("do_est")

  lower_change <- do_threshold
  upper_change <- c(Inf)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do and temperature") {
  temp_threshold <- params$threshold
  do_threshold <- params$threshold

  root_params <- filter(root_params, lower_do_25 > 0.5)
  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
  if (biomass_threshold == "50") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
  if (biomass_threshold == "25") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }
  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)

  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
  if (biomass_threshold == "50") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 }
  if (biomass_threshold == "25") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25 }
  .upper <- na.omit(.upper)
  upper_thresholds <- round(.upper[1], digits = 2)

  variable_names <- c("do_est", "temp")
  lower_thresholds <- c(lower_thresholds, NA)
  upper_thresholds <- c(NA, upper_thresholds)
  lower_change <- c(do_threshold, Inf)
  upper_change <- c(Inf, temp_threshold)
  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model
  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model

  indices <- c(1, 2)
  delta_t_total <- 2
  }

trim_thresholds <- c(lower_thresholds, upper_thresholds)
trim_models <- c(lower_model, upper_model)

min_string <- paste0(lower_change, collapse = "-")
max_string <- paste0(upper_change, collapse = "-")

optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))

glimpse(root_params)
```

# Calculate climate vectors
```{r}
vocc_by_year <- lapply(start_years, function(start_time) {

try({
  vocc <- readRDS(paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
})

if (!exists("vocc")) {
  pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }
  end_time <- start_time + 2
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = lower_change,
    max_thresholds = upper_change,
    round_fact = 10
  )

  vocc$var_1_min <- lower_change[1]
  vocc$var_2_min <- lower_change[2]
  vocc$var_1_max <- upper_change[1]
  vocc$var_2_max <- upper_change[2]

  saveRDS(vocc, file = paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
}
 vocc
})  
```

### Plot all possible vectors
```{r echo=FALSE}
all_vector_plots <- lapply(vocc_by_year, function(vocc) {    
start_time <- vocc$start_time[1]

if (climate == "temperature") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(-4, 3),
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
})
print(all_vector_plots)
```

### Filter vectors to match optimal range for chosen species
```{r}
trimmed_vectors <- lapply(vocc_by_year, function(vocc) {

vectors <- trim_vector_data(vocc, variable_names, 
  lower_change, upper_change, 
  lower_thresholds, upper_thresholds,
  cell_size = input_cell_size, dist_intercept = dist_intercept,
  max_dist = 120
)
})

all_vectors <- do.call("rbind", trimmed_vectors)
```

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r echo=FALSE, warning=FALSE}
if (climate == "temperature") {
all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
raster_min <- min(all_vectors$diff)
raster_max <- max(all_vectors$diff)
}

if (climate == "do") {
all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
raster_min <- min(all_vectors$diff)
raster_max <- max(all_vectors$diff)
}

if (climate == "do and temperature") {
  all_vectors$temp_diff <- all_vectors$temp.units_per_decade / 10 * delta_t_total
  all_vectors$do_diff <- all_vectors$do_est.units_per_decade / 10 * delta_t_total
  raster_min_temp <- min(all_vectors$temp_diff)
  raster_max_temp <- max(all_vectors$temp_diff)
  raster_min_do <- min(all_vectors$do_diff)
  raster_max_do <- max(all_vectors$do_diff)
}

trimmed_vector_plots <- lapply(trimmed_vectors, function(vectors) {

try({

  start_time <- vectors$start_time[1]

  if (climate == "temperature") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min, raster_max),
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min, raster_max),
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vectors$temp_diff <- vectors$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min_temp, raster_max_temp),
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vectors$do_diff <- vectors$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min_do, raster_max_do),
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste(
      "DO of ", optimal_values[1], "less", min_thresholds[1], 
      "or temperature of ", optimal_values[2], "plus", max_thresholds[2], 
      " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
})
})

print(trimmed_vector_plots)
```

```{r}
png(
  file = paste0("figs/", spp,
    "/vocc-", min_string, "to", max_string, "-", 
    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs-test.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = trimmed_vector_plots,
  ncol = 2,
  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", trim_models, ")"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

# Add species biomass data

Get all biomass predictions 
```{r echo=FALSE, warning=FALSE}
priors <- FALSE
covs <- "-trawled-ssid"

if (age == "imm") {
  biomass_predictions <- readRDS(paste0(
    "data/", spp, "/all-predictions-", spp, covs, 
    "-imm-biomass-prior-", priors, ".rds"))
}

if (age == "mature") {
  biomass_predictions <- readRDS(paste0(
    "data/", spp, "/all-predictions-", spp, covs, 
    "-mat-biomass-prior-", priors, ".rds"))
}
```

Extract biomass values to match spatial and temporal range of climate vectors
```{r}
#alldata <- lapply(biomass_list, function(biomass_predictions) {

  biomass_by_year <- list()
  for (j in seq_along(start_years)) {
   start_time <- start_years[[j]]
   end_time <- start_time + 2

   biomass_by_year[[j]] <- make_trend_data(biomass_predictions,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )
  }

  biomass_change <- do.call("rbind", biomass_by_year)

  biomass_covs <- biomass_predictions %>% 
    select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>% 
    distinct()

  biomass_change <- biomass_change %>% 
    rename(start_density = est_exp_s, end_density = est_exp_e) %>%
    dplyr::select(icell, x, y, start_time, start_density, end_density)
    # dplyr::select(-N, -time_step, - units_per_decade, -slope)

  biomass <- left_join(biomass_change, biomass_covs, by = c("x" = "X", "y"="Y"))

  # keep just one row for each cell that exceeded the end climate threshold 
  source_cells <- all_vectors %>% filter(distance > (-dist_intercept)) %>% 
    distinct(icell, start_time, .keep_all = TRUE) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  source_biomass <- left_join(source_cells, biomass, 
    by = c("icell", "x", "y", "start_time")) %>%
    mutate(cell_type = "source") %>% filter(!is.na(start_density))

  # identify all possible control cells have been given a distance == - dist_intercept
  control_cells <- all_vectors %>% filter(distance == (-dist_intercept)) %>% 
    distinct(icell, start_time, .keep_all = TRUE) %>% 
    mutate(distance = (- dist_intercept)) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  control_biomass <- left_join(control_cells, biomass, 
    by = c("icell", "x", "y", "start_time")) %>%
    mutate(cell_type = "control") %>% filter(!is.na(start_density))

  alldata <- rbind(source_biomass, control_biomass) 
  alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
    select(-var_2_min, -var_2_max, -mean_target_X, -mean_target_Y) %>% 
    rename(vect_dist = distance)
  alldata$treat <- as.factor(alldata$treat)

 # alldata
#})
```

# Select matched control cells

```{r}
#rm(vocc_by_sp)
# try({
#   vocc_by_sp <- readRDS(paste0(
#     "data/BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
#   ))
# })

#FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?

match_formula <- paste0("treat ~ trawled + x + y + depth_st") 
  # depth.st + omega_s + muddy.st + any_rock.st + 
  for (j in (variable_names)) {
    match_formula <- paste0("", match_formula, " + ", j, "_s.st")
    start_value.st <- paste0("", j, "_s.st")
    start_value <- paste0("", j, "_s")

    alldata <- alldata %>% mutate((!!start_value.st) := 
        arm::rescale(alldata[[start_value]]) )
  #(UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    }

if (!exists("vocc_by_sp")){
# matched <- lapply(alldata, function(alldata) {  
  ## trim non-overlapping values to reduce size of dataset 
  control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.01)
  source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.01)
  # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 1)
  # source_max <- quantile(alldata$start_density[alldata$treat=="1"], 1)

  group_min <- max(control_min, source_min)
  # group_max <- min(control_max, source_max)
  alldata <- filter(alldata, start_density >= group_min) 
  #%>% filter(start_density <= group_max)

  # need to be standardized? seems not
  alldata$depth2 <- alldata$depth^2
  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
  alldata$depth_st <- arm::rescale(log(alldata$depth))
  alldata$depth.st <- round(arm::rescale(log(alldata$depth))*5)/5
  alldata$muddy.st <- arm::rescale(alldata$muddy) 
  alldata$any_rock.st <- arm::rescale(alldata$any_rock)

 ## Using optimization ... doesn't work   
  # alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
  # fit_formula <- 
  #   paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") 
  # # for (j in (variable_names)) {
  # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  # #   }
  # # options("optmatch_max_problem_size" = Inf)
  # matched <- MatchIt::matchit(as.formula(fit_formula), 
  #     method = "optimal", data = alldata)

  # # Using exact match and rounding
  # # Choose data order if not using random
  # alldata <- alldata[(order(alldata$y)), ]
  # if (order == "increasing") alldata <- alldata[rev(order(alldata$start_density)), ]
  # if (order == "decreasing") alldata <- alldata[rev(order(alldata$start_density)), ]
# set.seed(99999)
  matched <- MatchIt::matchit(as.formula(match_formula), 
    exact = c("start_density.st", "start_time"),
    method = "nearest", m.order = "random",
    distance = "probit",
    discard = "both",
    data = alldata)
  # matched
# })
}
```

Print diagnostics for matching
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if (!exists("vocc_by_sp")){
match_plots <- plot(matched, interactive = F)
summary(matched)
}

```

Prep matched data for analysis
```{r}
if (!exists("vocc_by_sp")){
#vocc_by_sp <- lapply(matched, function(matched) {    
  mdata <- MatchIt::match.data(matched, distance = "pscore")

  sources <- mdata[row.names(matched$match.matrix), ]
  sources$origobs <- row.names(sources)   
  sources$matchobs <- matched$match.matrix[,1]

  controls <- mdata[matched$match.matrix, ]
  controls$origobs <- row.names(controls)   
  controls$matchobs <- matched$match.matrix[,1]

  mdata <- rbind(sources, controls) %>% tidyr::drop_na() %>% filter(matchobs!="-1")

  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) 
  #%>% dplyr::select(-start_density, -end_density)
  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) 
  #%>% dplyr::select(-start_density, -end_density)

  for (j in seq_along(variable_names)) {
    climate_var <- paste0(variable_names[j])
    matched_time0 <- matched_time0 %>% 
      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    matched_time1 <- matched_time1 %>% 
      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
   }

  vocc_by_sp <- rbind(matched_time0, matched_time1)

  # calculate standardized climate variables
  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])
    std.var <- paste0("std_", variable_names[j])
    std.var2 <- paste0("std_", variable_names[j], "2")

    vocc_by_sp <- mutate(
      vocc_by_sp,
      (!!mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!std.var) := ((UQ(rlang::sym(paste0(variable_names[j]))) - 
          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) / 
          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE))*2),
      (!!std.var2) := (( UQ(rlang::sym(paste0(variable_names[j]))) - 
          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) ) / 
          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)*2))^2
    )
  }

  # rename spatial coordinates to avoid naming conflict with y~x notation
  data <- vocc_by_sp %>%
    rename(X = x, Y = y)
    # %>% rename(distance = vect_dist) %>% 
    # select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)

  # other variable prep
  data$log_density <-  log(data$density)
  data$log_depth <-  log(data$depth)
  data$std_log_depth <- arm::rescale(log(data$depth))
  data$vect_dist2 <- data$vect_dist^2
  data <- mutate(data, time = ifelse(after==1, "end", "begin"))
  data$matchobs <- as.factor(data$matchobs)
  data$origobs  <- as.factor(data$origobs)   
  data$start_time <- as.factor(data$start_time)

#  data
#})
vocc_by_sp <- list()
vocc_by_sp[[1]] <- data
vocc_by_sp[[3]] <- match_formula  
vocc_by_sp[[4]] <- match_plots
saveRDS(vocc_by_sp, file = paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
  ))
}
```

## Save spp dataframe
```{r}
#biomass_threshold <- "50"
optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
#climate <- "temperature"
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

n <- nrow(vocc_by_sp[[1]])
species <- rep(species, n) 
mature <- rep(age, n)
ssid <- rep(ssid_string, n)
data <- vocc_by_sp[[1]]
df<- cbind(species, mature, ssid, data)
species <- species[[1]]

# folder to hold vocc data
biomass_threshold_string <- paste0("perc_", biomass_threshold)

dir.create(file.path("data/_all", climate_string, biomass_threshold_string))

if (climate=="temperature") {
dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))
dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string, age))
}
if (climate=="temperature") {
  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", max_string,"/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}

if (climate=="do") {
  dir.create(file.path("data/_all", climate_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string))
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string, age))  
  }
if (climate=="do") {
  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", min_string,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}

if (climate=="do and temperature") {
  dir.create(file.path("data/_all", climate_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, age))  
}

if (climate=="do and temperature") {
  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}
```

```{r}
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})
data <- vocc_by_sp[[1]]
```

### Plot matches
```{r}
#FIXME: SAVE PLOTS?

dat_summary <- data %>% 
  group_by(start_time, cell_type) %>%
  tally() %>% filter(cell_type=="source")
data_depth <- data %>% 
  group_by(start_time) %>% 
  mutate(mean_depth = mean(depth))

p1 <- ggplot(data, 
#ggplot(filter(data, as.numeric(matchobs) <= 550), 
  aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
  geom_point(alpha=0.8, size = 1) +
  scale_shape_manual(values= c(0, 15)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  facet_wrap(~start_time) +
  theme_bw()
p1

lapply(variable_names, function(j) {
ggplot(data, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") +
  facet_grid(start_time~cell_type) +
  theme_bw()
})

p2 <- ggplot(data, aes(std_log_depth, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~start_time) +
  theme_bw()
p2

# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
#   geom_point(alpha=0.35) +
#   scale_colour_viridis_d(option="C") +
#     facet_grid(start_time~time) +
#   theme_bw()

p3 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_grid(start_time~time, scales = 'fixed') +
  theme_bw() + theme(legend.position='none')
p3

p4 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.45) +
  scale_fill_viridis_d(option="C") +
  facet_grid(time~start_time) +
  theme_bw() + theme(legend.position='none')
p4

dat_summary <- data %>% 
  group_by(start_time, cell_type) %>% 
  tally() %>% filter(cell_type=="source")

p5 <- ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  geom_text(data = dat_summary,
            aes(cell_type, min(data$log_density)+1, label = n), hjust = 1.25) +
  facet_grid(time~start_time) +
  theme_bw() + theme(legend.position='none')
p5
```

```{r}
png(
  file = paste0("figs/", spp,
    "/matches-", min_string, "to", max_string, "-",
    climate_string, "-", spp, "-", age, "-", ssid_string, "-",
    climate_string, "-perc_", biomass_threshold, "-", optimal_string, ".png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  p1,p2,p3,p4,p5,
  ncol = 2,
  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

### 2020-02-07 analysis/VOCC/BACI/gamm-draft.R

@@ -1,0 +1,58 @@
+
+mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + s(std_temp) + s(distance) + s(as.factor(idata$matchobs), bs="re")
+  + s(X,Y, bs="re"), data=idata)
+summary(mi)
+
+
+mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(distance) + s(as.factor(idata$matchobs), bs="re"), data=idata) 
+
++ s(X,Y, bs="re")
+
+
+
+### REMOVED DOCUMENTATION
+# @import mgcv
+# @import stats
+# @import grDevices
+# @import graphics
+# @import plotfunctions
+
+#' @description Plots a smooth from a \code{\link[mgcv]{gam}} or
+#' \code{\link[mgcv]{bam}} model based on predictions.
+#' In contrast with the default \code{\link[mgcv]{plot.gam}}, this function
+#' plots the summed effects and optionally removes the random effects.
+#' 
+#' @param x A gam object, produced by \code{\link[mgcv]{gam}} or
+#' \code{\link[mgcv]{bam}}.
+#' @param pred A named list of the values to use for the predictor terms
+#' to plot.
+#' @param cond A named list of the values to use for the other predictor terms
+#' (not in view). Used for choosing between smooths that share the same view
+#' predictors.
+#' @param parametricOnly Logical: whether or not to cancel out all smooth
+#' terms and only use the predictors in the parametric summary.
+#' @param rm.ranef Logical: whether or not to remove random effects.
+#' Default is TRUE.
+#' @param col The colors for the lines and the error bars of the plot.
+#' @param se If less than or equal to zero then only the predicted surface is
+#' plotted, but if greater than zero, then the predicted values plus
+#' confidence intervals are plotted. The value of se will be multiplied with
+#' the standard error (i.e., 1.96 results in 95\%CI and 2.58).
+#' @param print.summary Logical: whether or not to print summary.
+#' Default set to the print info messages option
+#' (see \code{\link{infoMessages}}).
+#' @param main Changing the main title for the plot, see also title.
+#' @param xlab Changing the label for the x axis,
+#' defaults to a description of x.
+#' @param ... other options to pass on to \code{\link{dotplot_error}},
+#' see \code{\link[graphics]{par}}
+#' @section Warning:
+#' Use \code{parametricOnly} with care! When set to TRUE, all smooth
+#' predictors are set to 0. Note that this might result in strange
[…]

### 2020-02-07 analysis/VOCC/old-code/vocc-regression-vel.R

@@ -1,0 +1,472 @@
+library(TMB)
+library(dplyr)
+library(ggplot2)
+library(gfranges)
+
+setwd(here::here("analysis", "VOCC"))
+compile("vocc_regression.cpp")
+dyn.load(dynlib("vocc_regression"))
+source("vocc-regression-functions.R")
+
+stats <- readRDS(paste0("data/life-history-stats.rds"))
+stats$rockfish <- if_else(stats$group == "ROCKFISH", "ROCKFISH", "OTHER")
+stats$genus <- tolower(stats$group)
+
+#### LOAD MATURE VOCC DATA 
+# model_age <- "multi-spp-biotic-vocc-mature"
+# model_age <- "scrambled-vocc-mature"
+# model_age <- "scrambled2-vocc-mature"
+# model_age <- "scrambled3-vocc-mature"
+# model_age <- "scrambled4-mature"
+# model_age <- "scrambled5-mature"
+
+model_age <- "all-temp-mature"
+d <- readRDS(paste0("data/", model_age, "-with-fished.rds"))
+d <- na.omit(d) %>% as_tibble()
+
+d <- suppressWarnings(left_join(d, stats, by = "species")) %>%
+  # filter(species != "Bocaccio") %>%
+  # filter(species != "Sand Sole") %>%
+  # filter(species != "Longspine Thornyhead") %>%
+  filter(species != "Shortbelly Rockfish")
+
+# #### LOAD IMMATURE VOCC DATA 
+# model_age <- "multi-spp-biotic-vocc-immature"
+# d <- readRDS(paste0("data/", model_age, "-with-fished.rds"))
+# d <- na.omit(d) %>% as_tibble()
+# 
+# d <- suppressWarnings(left_join(d, stats, by = "species")) %>%
+#   filter(species != "Curlfin Sole") %>%
+#   # filter(species != "Bocaccio") %>%
+#   filter(species != "Longspine Thornyhead")
+
+select(d, genus, species) %>%
+  distinct() %>%
+  arrange(genus, species) %>%
+  as.data.frame()
+
+
+#### PREP TEMP VARIABLES ####
+d$squashed_temp_vel <- collapse_outliers(d$temp_vel, c(0.005, 0.995))
+# plot(squashed_do_vel ~ squashed_temp_vel, data = d, col = "#00000010")
[…]

### 2020-02-07 analysis/VOCC/vocc-gradient.R

@@ -1,140 +1,0 @@
-#' Velocity of climate change
-#'
-#' @param data Gridded data.
-#' @param layer Variable to retain. Defaults to "est", for raw sdmTMB predictions.
-#' @param scale_fac Controls how the original 2-km projection is aggregated
-#'    (e.g. a value of 5 means that the raster would be reprojected to 10km grid).
-#' @param time_step Time variable.
-#' @param grad_time_steps Define time periods.
-#'    If NULL will use spatial gradient in last time period.
-#' @param latlon Default is TRUE. False for UTMs or other true distance grid.
-#' @param quantile_cutoff Used for plotting to set min and max angles of vectors.
-#'    Defaults to 0.05.
-#'
-#' @export
-#'
-vocc_gradient_calc <- function(data,
-                               layer,
-                               scale_fac = 1,
-                               time_step = "year",
-                               indices = NULL,
-                               divisor = 10,
-                               latlon = FALSE,
-                               quantile_cutoff = 0.05) {
-
-  # # devtools::install_github("seananderson/vocc")
-  # # install.package(ggnewscale)
-  # # install.package(gfplot)
-
-  # make raster brick
-  rbrick <- make_gradient_brick(data, layer,
-    scale_fac = scale_fac,
-    time_step = time_step
-  )
-
-  # Then calculate the trend per pixel:
-  slopedat <- vocc::calcslope(rbrick, divisor = divisor)
-
-  # Then get the mean values for a time period
-  if (!is.null(indices)) {
-    # if (indices = "all") {
-    #   # calculates mean temp across all time slices for each grid cell
-    #   indices <- rep(1, raster::nlayers(rbrick))
-    #  mnraster_brick <- raster::stackApply(rbrick, indices = indices, fun = mean)
-    #  mnraster <-mnraster_brick[[raster::nlayers(mnraster_brick)]]
-    #   } else {
-    mnraster_brick <- raster::stackApply(rbrick, indices = indices, fun = mean)
-    mnraster <- mnraster_brick[[raster::nlayers(mnraster_brick)]]
-  } else {
[…]

### 2020-02-07 analysis/VOCC/3-vocc-temp.Rmd
rename vocc rendering

--- 2020-01-23 analysis/VOCC/4-vocc-temp.Rmd
@@
 params: 
-    species: "Arrowtooth Flounder" 
+    #species: "Arrowtooth Flounder" 
+    species: "Yelloweye Rockfish"
     immature: FALSE
-    #region: "both odd year surveys"
+    region: "both odd year surveys"
     #region: "West Coast Haida Gwaii"
-    region: "West Coast Vancouver Island"
+    #region: "West Coast Vancouver Island"
     covs: "-tv-depth-only"
-    scrambled_years: FALSE
+    #scrambled_years: FALSE
 ---
@@
   # unique(d_all$year)
-  d <- d_all %>% filter(ssid %in% model_ssid)
+  d <- d_all %>% filter(ssid %in% model_ssid) %>% filter(year >= start_year)
   length(unique(d$year))
@@
 glimpse (d)
-if (scrambled_years) {
-  d2 <- list()
-  # set.seed(9999) 
-  # set.seed(12345)
-  # set.seed(98765) 
-  original_years <- unique(d$year)
-  new_years <- sample(original_years, replace=F)
-  convert <- cbind(original_years, new_years)
-  convert
-  for (i in seq_len(length(original_years))) {
-    old_year <- as.integer(convert[[i,1]])
-    fake_year <- as.integer(convert[[i,2]])
-    d2[[i]] <- filter(d, year == old_year) %>% 
-      mutate(year = fake_year )
-     # browser()
-  }
-  d <- do.call(rbind, d2)
-  convert
-}
+# 
+# if (scrambled_years) {
+#   d2 <- list()
+#   # set.seed(9999) 
+#   # set.seed(12345)
+#   # set.seed(98765) 
+#   original_years <- unique(d$year)
+#   new_years <- sample(original_years, replace=F)
+#   convert <- cbind(original_years, new_years)
+#   convert
+#   for (i in seq_len(length(original_years))) {
+#     old_year <- as.integer(convert[[i,1]])
+#     fake_year <- as.integer(convert[[i,2]])
+#     d2[[i]] <- filter(d, year == old_year) %>% 
+#       mutate(year = fake_year )
+#      # browser()
+#   }
+#   d <- do.call(rbind, d2)
+#   convert
+# }
 ```
@@
 d <- left_join(d, fishing, by = c("X", "Y", "year"))
+d$effort[is.na(d$effort)] <- 0
 d$log_effort[is.na(d$log_effort)] <- -2.3
-d$effort[is.na(d$effort)] <- 0
 
@@
     fishing_trend = trend,
-    mean_effort = exp(mean)
+    mean_effort = exp(mean) # gives geometric mean
   ) %>%
@@
   gfplot:::utm2ll(., utm_zone = 9)
+```
 
+```{r eval = FALSE }
 effort <- plot_vocc(gf,
@@
 range(gf2$gradient_diff)
+```
 
+```{r eval = FALSE}
 grad2 <- plot_vocc(gf2,
@@
   dplyr::mutate(X = x, Y = y)
+```
+
+```{r eval = FALSE}
 
@@
 
-```{r}
-
+```{r eval = FALSE}
 grad4 <- gfranges::plot_vocc(gf3t,
@@
 
-```{r}
+```{r eval = FALSE}
 grad5 <- gfranges::plot_vocc(gf3t,
@@
 
-```{r}
+```{r eval = FALSE}
 grad6 <- gfranges::plot_vocc(gf3t,
@@
 
-```{r}
+```{r eval = FALSE}
 raster_max <- quantile(gf3t$CV, 0.90, na.rm = TRUE)
@@
 ```{r}
-
-biotic_values <- gf3t %>%
+biotic_values <- gf3 %>%
   rename(
@@
 vocc <- mutate(vocc, species = !!species, method = "gradient")
+vocc$ssid <- ssid_string
+```
 
+```{r}
+# glimpse(vocc)
+```
+
+```{r}
 write.csv(vocc, file = paste0(
-  "data/_all_temp/", age,  "-all-temp-vocc-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), "-trim10.csv"
+  "data/_new_null/", age,  "-all-temp-nulls-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))

### 2020-02-07 analysis/VOCC/render-biotic-vocc.R

@@ -173,7 +173,7 @@ for (r_h in seq_along(list_regions)) {
     covs <- "-tv-depth-only" # string describing model
     reg <- gsub(" ", "-", gsub("\\/", "-", tolower(list_regions[r_h])))
     try({
-      rmarkdown::render("4-vocc-biotic.Rmd",
+      rmarkdown::render("3-vocc-w-do.Rmd",
         params = list(
           species = list_species[spp_i],
           # immature = TRUE,
@@ -199,7 +199,7 @@ for (r_h in seq_along(list_regions)) {
     covs <- "-tv-depth-only" # string describing model
     reg <- gsub(" ", "-", gsub("\\/", "-", tolower(list_regions[r_h])))
     try({
-      rmarkdown::render("4-vocc-temp.Rmd",
+      rmarkdown::render("3-vocc-temp.Rmd",
         params = list(
           species = list_species[spp_i],
           # immature = TRUE,

### 2020-02-07 analysis/VOCC/biomass-trends.Rmd
rename with numbers ordering steps

--- 2019-06-25 analysis/VOCC/biomass-trends.Rmd
@@
 ---
-title: "Biomass trend maps"
+title: "`r paste(params$species)` biomass trends and figures"
 author: "Philina English"
-date: '2019-06-07'
+date: "`r Sys.Date()`"
 output: html_document
+params: 
+    species: "Pacific Ocean Perch" 
+    #species: "Redbanded Rockfish"
+    region: "All synoptic surveys"
+    start_year: 2008
+    bind_predictions: FALSE
+    covs: "-trawled-ssid-reml"
+    #covs: "-log-both-noAR1" #"-trawled-ssid" # covariate string used to name model rds 
 ---
 
-Choose one value for each parameter:
-
-Species
-```{r}
-species <- "Quillback Rockfish"
-# species <- "Yelloweye Rockfish"
-# priors <- TRUE
-covariates <- "+trawled+mixed+rocky"
-
-species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
-species <- "Silvergray Rockfish"
-species <- "Arrowtooth Flounder"
-species <- "Lingcod"
-```
-
-# Choose covariates  
-```{r}
-covariates <- "+trawled+muddy+rocky+mixed"
-covariates <- "+trawled+mixed+rocky"
-covariates <- "+muddy+mixed+rocky"
-covariates <- "+mixed+rocky"
-covariates <- "+trawled+mixed"
-```
-
-Is a new spatial needed?
-```{r}
-update_spatial_model <- TRUE
-priors <- FALSE
-#priors <- TRUE
-
-# Choose region to update model for:
-region <- "Both odd year surveys"
-region <- "West Coast Vancouver Island"
-region <- "West Coast Haida Gwaii"
-
-```
-
-If not...
-```{r}
-update_spatial_model <- FALSE
-priors <- FALSE
-# priors <- TRUE
-region <- "All synoptic surveys"
-
-if (region == "All synoptic surveys") {
-  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
-  model_ssid <- c(1, 3, 4, 16)
-  ssid_string <- paste0(model_ssid, collapse = "n")
-  years <- NULL
-}
-```
-
-Run all subsequent code...
-
-```{r, message=FALSE, warning=FALSE}
+```{r global_options, include=FALSE}
+knitr::opts_chunk$set( fig.width = 9, fig.height = 6.5, #fig.width=11, fig.height=8.5, 
+                      #fig.path=paste0("figs/", spp, "/"),
+                      echo=FALSE, warning=FALSE, message=FALSE)
 library(dplyr)
@@
 library(gfranges)
+```
+
+```{r params}
+species <- params$species
+region <- params$region
+start_year <- params$start_year
+
+bind_predictions <- params$bind_predictions
+covs <- params$covs
+priors <- FALSE
+
+paste("region =", region)
+paste("ssids modelled separately =", bind_predictions)
+paste("model label =", covs)
+paste("priors =", priors)
 
@@
 
-# folder to hold figs for this species
-dir.create(file.path("figs", spp))
-dir.create(file.path("data", spp))
-fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+if (region == "All synoptic surveys") {
+  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
+  model_ssid <- c(1, 3, 4, 16)
+  ssid_string <- paste0(model_ssid, collapse = "n")
+}
 
-if (region == "Both odd year surveys") {
+if (region == "both odd year surveys") {
   survey <- c("SYN QCS", "SYN HS")
@@
   ssid_string <- paste0(model_ssid, collapse = "n")
-  years <- NULL
-}
-
-```
-
-```{r}
-
-# if(region == "Queen Charlotte Sound") {
-# survey <- c("SYN QCS")
-# model_ssid <- c(1)
-# ssid_string <- paste0(model_ssid, collapse = "n")
-# years <- NULL
-# }
-#
-# if(region == "Hecate Strait") {
-# survey <- c("SYN HS")
-# model_ssid <- c(3)
-# ssid_string <- paste0(model_ssid, collapse = "n")
-# years <- NULL
-# }
-```
-
-Make prediction grids for all surveys and years, if not done before.
-```{r, message=FALSE, warning=FALSE}
-try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
-
-if (!exists("nd_all")) {
-
-  # choose base year(s) to create grid from
-  dummy_year <- c(2005, 2006)
-
-  bath <- readRDS("data/bathymetry-data")
-  dat <- bath$data %>% scale_survey_predictors()
-
-  # create grids for each ssid separately
-  ssid <- 1
-  survey_abbrev <- "SYN QCS"
-  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_1$year)
-
-  ssid <- 3
-  survey_abbrev <- "SYN HS"
-  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_3$year)
-
-  ssid <- 4
-  survey_abbrev <- "SYN WCVI"
-  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_4$year)
-
-  ssid <- 16
-  survey_abbrev <- "SYN WCHG"
-  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_16$year)
-
-  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
-    mutate(year = as.numeric(year))
-
-  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed)
-  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))
-
-  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
-
-  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
 }
@@
 
-Calculate maturity weighted biomass if not done before
-```{r}
-rm(biomass)
-try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
-
-if (!exists("biomass")) {
-  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
-  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
-  bath <- readRDS("data/bathymetry-data")
-  # for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
-  # fish <- fish[fish$fishing_event_id!=1506954,]
-
-  maturity <- split_catch_maturity(survey_sets, fish, bath,
-    # survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
-    # survey = c("SYN HS", "SYN WCHG"),
-    survey = survey,
-    years = years,
-    cutoff_quantile = c(.9995),
-    plot = TRUE
-  )
-
-  saveRDS(maturity$data, file = paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
-}
-```
-
-```{r}
-if (exists("maturity")) {
-  print(maturity$mass_model)
-  ggsave(
-    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
-    dpi = 600,
-    width = 8, # The width of the plot in inches
-    height = 10
-  )
-}
-
-# try(print(maturity$mass_model))
-```
-
-```{r}
-if (exists("maturity")) {
-  png(
-    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), # The directory you want to save the file in
-    res = 600,
-    units = "in",
-    width = 8, # The width of the plot in inches
-    height = 10
-  )
-  print(maturity$maturity)
-  dev.off()
-}
-
-try(print(maturity$maturity))
-```
-
-Run sdmTMB model (if not done previously).
-```{r}
-if (update_spatial_model) {
-  biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
-  covars <- readRDS("data/event-covariates.rds")
-  data <- dplyr::left_join(biomass,covars)
-
-  data <- data %>% gfranges:::scale_survey_predictors()
-
-  data <- data %>% filter(ssid %in% model_ssid)
-
-  if (region == "All synoptic surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 1200)
-  }
-
-  if (region == "Both odd year surveys") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
-  }
-
-  if (region == "West Coast Vancouver Island") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
-  }
-
-  if (region == "West Coast Haida Gwaii") {
-    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 300)
-  }
-
-  sdmTMB::plot_spde(spde)
-
-  if (any(names(data) == "adult_density")) {
-
-    adult_formula <- as.formula(paste("adult_density ~ 0 + as.factor(year)", covariates, ""))
-
-    starttime1 <- Sys.time()
-    adult_biomass <- sdmTMB::sdmTMB(data,
-      adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime1 <- Sys.time()
-    time1 <- round(starttime1 - endtime1)
-
-    saveRDS(adult_biomass, file = paste0("data/", spp, "/mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time1, ".rds"))
-
-    imm_formula <- as.formula(paste("imm_density ~ 0 + as.factor(year)", covariates, ""))
-
-    starttime2 <- Sys.time()
-    imm_biomass <- sdmTMB::sdmTMB(data, 
-      imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = FALSE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime2 <- Sys.time()
-    time2 <- round(starttime2 - endtime2)
-
-    saveRDS(imm_biomass, file = paste0("data/", spp, "/mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, time2, ".rds"))
-  } else {
-    starttime3 <- Sys.time()
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
-
-    total_biomass <- sdmTMB::sdmTMB(dat,
-      dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime3 <- Sys.time()
-    time3 <- round(starttime3 - endtime3)
-
-    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
-  }
-}
-```
-
-Calculate predicted values for available models, if needed
-```{r, error=TRUE}
+Load model predictions
+```{r results = "hide"}
 rm(adult_predictions)
@@
 
-if (isFALSE(update_spatial_model)) {
+rm(pred1n3)
+rm(pred4)
+rm(pred16)
+
+rm(pred1n3_imm)
+rm(pred4_imm)
+rm(pred16_imm)
+
+rm(pred1n3_all)
+rm(pred4_all)
+rm(pred16_all)
+
 if (region != "All synoptic surveys") {
-    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
-    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
-    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
+  try(adult_predictions <- readRDS(paste0("data/", spp,
+    "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-", priors, ".rds"
+  )) %>% filter(ssid %in% model_ssid))
+  try(imm_predictions <- readRDS(paste0("data/", spp,
+    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-", priors, ".rds"
+  )) %>% filter(ssid %in% model_ssid))
+
+  if (!exists("adult_predictions")) {
+    paste("Maturity split biomass predictions are not available for this survey.")
+
+    try(predictions <- readRDS(paste0("data/", spp,
+      "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-", priors, ".rds"
+    )) %>% filter(ssid %in% model_ssid))
+
+    if (!exists("predictions")) {
+      paste("Total biomass predictions for this survey are also not available.")
+    }
+  }
 } else {
-    # Bind together predictions from different models
-    try ({
-      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-prior-", priors, ".rds"))
-      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-prior-", priors, ".rds"))
-      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-prior-", priors, ".rds"))
-      adult_predictions <- rbind(pred1n3, pred4, pred16) 
-      # model_ssid <- unique(adult_predictions$ssid)
-      # ssid_string <- paste0(model_ssid, collapse = "n")
+
+  try({adult_predictions <- readRDS(paste0("data/", spp,
+    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
+    ))
+  })
+  try({imm_predictions <- readRDS(paste0("data/", spp,
+    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
+    ))
   })
 
-    try ({
-      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-imm-biomass-prior-", priors, ".rds"))
-      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-imm-biomass-prior-", priors, ".rds"))
-      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-imm-biomass-prior-", priors, ".rds"))
-      imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm) 
-      # model_ssid <- unique(imm_predictions$ssid)
-      # ssid_string <- paste0(model_ssid, collapse = "n")
+  # OR bind together predictions from different models
+  if (bind_predictions) {
+    try({pred1n3 <- readRDS(paste0("data/", spp,
+      "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"
+      ))
+    })
+    try({pred4 <- readRDS(paste0("data/", spp,
+      "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"
+      ))
+    })
+    try({pred16 <- readRDS(paste0("data/", spp,
+      "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"
+      ))
     })
 
-    try ({
-      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-prior-", priors, ".rds"))
-      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-prior-", priors, ".rds"))
-      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-prior-", priors, ".rds"))
-      predictions <- rbind(pred1n3_all, pred4_all, pred16_all) # ,
-      #  model_ssid <- unique(predictions$ssid),
-      # ssid_string <- paste0(model_ssid, collapse = "n")
+    if (!exists("pred1n3")) pred1n3 <- NULL
+    if (!exists("pred16")) pred16 <- NULL
+    if (!exists("pred4")) pred4 <- NULL
+    adult_predictions <- rbind(pred1n3, pred4, pred16)
+
+    try({pred1n3_imm <- readRDS(paste0("data/", spp,
+      "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"
+      ))
+    })
+    try({pred4_imm <- readRDS(paste0("data/", spp,
+      "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))
+    })
+    try({pred16_imm <- readRDS(paste0("data/", spp,
+        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"))
     })
 
-    if (!exists("adult_predictions")) paste("Maturity split biomass predictions are not available.")
-    if (!exists("predictions")) paste("Total biomass predictions are not available.")
-  }
-} else {
-  if (region == "All synoptic surveys") {
-    paste("To update spatial models, you need to build and save each region's model separately.")
+    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
+    if (!exists("pred16_imm")) pred16_imm <- NULL
+    if (!exists("pred4_imm")) pred4_imm <- NULL
+    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)
   }
@@
   if (!exists("adult_predictions")) {
-    fish <- readRDS(paste0("raw/bio-data-", spp, ""))
+      paste("Maturity split biomass predictions are not available.")
 
-  if (length(unique(fish$maturity_code)) > 2) {
+      try({predictions <- readRDS(paste0("data/", spp,
+        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
+      })
 
-    adult_biomass <- readRDS(list.files(
-      path = paste0("data/", spp, "/"),
-      #pattern = paste0("mod-mat-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
-      pattern = paste0("mod-mat-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
-      all.files = F, full.names = T
-    ))
+      if (!exists("predictions")) {
+        try({pred1n3_all <- readRDS(paste0("data/", spp,
+          "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))
+        })
+        try({pred4_all <- readRDS(paste0("data/", spp,
+          "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))
+        })
+        try({pred16_all <- readRDS(paste0("data/", spp,
+          "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))
+        })
+        try({
+          predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
+        })
 
-    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-    nd <- nd_all %>% filter(ssid %in% model_ssid) %>% filter(year %in% unique(adult_biomass$data$year))
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    adult_predictions <- predict(adult_biomass, newdata = nd)
-    saveRDS(adult_predictions,
-      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-mature-biomass-prior-", priors, ".rds")
-    )
-
-    imm_biomass <- readRDS(list.files(
-      path = paste0("data/", spp, "/"),
-      #pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
-      pattern = paste0("mod-imm-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, "-*"),
-      all.files = F, full.names = T
-    ))
-    imm_predictions <- predict(imm_biomass, newdata = nd)
-    saveRDS(imm_predictions,
-      file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-imm-biomass-prior-", priors, ".rds")
-    )
-  } else {
-    rm(adult_predictions)
-    rm(imm_predictions)
-
-    if (!exists(predictions)) {
-      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-      nd <- nd_all %>% filter(ssid %in% model_ssid)
-      nd <- na.omit(nd)
-      nd$year <- as.integer(nd$year)
-
-      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, covariates, "-", ssid_string, "-prior-", priors, ".rds"))
-      nd$year <- as.integer(nd$year)
-      predictions <- predict(total_biomass, newdata = nd)
-      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, covariates,"-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
+        if (!exists("predictions")) paste("Total biomass predictions are not available.")
       }
     }
-}
-```
-
-Bind together predictions from different models
-```{r, eval=FALSE}
-if (region == "All synoptic surveys") {
-  if (isFALSE(update_spatial_model)) {
-    try(
-      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-mature-biomass-prior-", priors, ".rds")),
-      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-mature-biomass-prior-", priors, ".rds")),
-      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-mature-biomass-prior-", priors, ".rds")),
-
-      adult_predictions <- rbind(pred1n3, pred4, pred16),
-      model_ssid <- unique(adult_predictions$ssid),
-      ssid_string <- paste0(model_ssid, collapse = "n")
-    )
-
-    try(
-      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-1n3-imm-biomass-prior-", priors, ".rds")),
-      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-4-imm-biomass-prior-", priors, ".rds")),
-      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covariates, "-16-imm-biomass-prior-", priors, ".rds")),
-
-      imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm),
-      model_ssid <- unique(imm_predictions$ssid),
-      ssid_string <- paste0(model_ssid, collapse = "n")
-    )
-
-    try(
-      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-prior-", priors, ".rds")),
-      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-prior-", priors, ".rds")),
-      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-prior-", priors, ".rds")),
-
-      predictions <- rbind(pred1n3_all, pred4_all, pred16_all),
-      model_ssid <- unique(predictions$ssid),
-      ssid_string <- paste0(model_ssid, collapse = "n")
-    )
-  } else {
-    paste("To update spatial models, you need to build and save each region's model separately.")
-  }
-}
-
-if (!exists(adult_predictions)) paste("Maturity split biomass predictions were not available.")
-if (!exists(predictions)) paste("Total biomass predictions were not available.")
 ```
@@
 if (exists("adult_predictions")) {
+  # to convert from kg/m2 to kg/hectare multiply by 10000
   adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
@@
   adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
-  imm_predictions$prop_imm <- imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
+  imm_predictions$prop_imm <- 
+  imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)
 
-  max_raster <- quantile(adult_predictions$est_exp, .9999)
+  saveRDS(adult_predictions, file = paste0(
+    "data/", spp,
+    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
+  ))
+  saveRDS(imm_predictions, file = paste0(
+    "data/", spp,
+    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
+  ))
+
+  adult_predictions <- filter(adult_predictions, year >= start_year)
+  imm_predictions <- filter(imm_predictions, year >= start_year)
+
+  max_raster <- quantile(adult_predictions$est_exp, 0.999)
   max_adult <- signif(max(adult_predictions$est_exp), digits = 2)
+
+  model_ssid <- unique(adult_predictions$ssid)
 } else {
+  predictions <- filter(predictions, year >= start_year)
+
+  predictions$est_exp <- exp(predictions$est) * 10000
   predictions$total_bio <- exp(predictions$est) * 10000
-  max_raster <- quantile(predictions$total_bio, .9999)
-  max_bio <- round(max(predictions$total_bio))
+  max_raster <- quantile(predictions$total_bio, .999)
+  #max_bio <- round(max(predictions$total_bio))
+  max_bio <- signif(max(predictions$est_exp), digits = 2)
+
+  saveRDS(predictions, file = paste0(
+    "data/", spp,
+    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
+  ))
+
+  model_ssid <- unique(predictions$ssid)
 }
@@
 
+## Mature (or total) biomass predictions
 ```{r, error=TRUE}
@@
   p_adult_all <- plot_facet_map(predictions, "total_bio",
-    # raster_limits = c(0, max_raster),
     transform_col = fourth_root_power
@@
     ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
+
   print(p_adult_all)
@@
 
+### Split by survey (SSID)
 ```{r}
-if (exists("adult_predictions")) {
-  # if (length(model_ssid)>1){
+if (exists("adult_predictions")) predictions <- adult_predictions
+
   if (1 %in% model_ssid) {
-    adult_predictions1 <- adult_predictions %>% filter(ssid == 1)
+    adult_predictions1 <- predictions %>% filter(ssid == 1)
   } else {
@@
   if (3 %in% model_ssid) {
-    adult_predictions3 <- adult_predictions %>% filter(ssid == 3)
+    adult_predictions3 <- predictions %>%
+      filter(ssid != 16) %>%
+      filter(ssid != 4)
   } else {
@@
   if (16 %in% model_ssid) {
-    adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
+    if (any(predictions$ssid == 16)) {
+      adult_predictions16 <- predictions %>% filter(ssid == 16)
+    } else {
+      adult_predictions16 <- NULL
+    }
   } else {
@@
   ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)
-
   ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]
@@
   }
+  } else {
+    for (i in seq_len(length(model_ssid))) {
+    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
+      raster_limits = c(0, max_raster),
+      # legend_position = "none"
+      transform_col = fourth_root_power
+    ) +
+      labs(fill = "kg/ha") +
+      ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
+    p_adult[[i]]  
+    }
+  }
   print(p_adult)
-}
-# }
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
+rm(out_mature)
+rm(out_total)
+# try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
+
+if (!exists("out_mature")) {
   if (exists("adult_predictions")) {
     predictions <- adult_predictions
-  # model_ssid <- unique(predictions$ssid)
-  # ssid_string <- paste0(model_ssid, collapse = "n")
+  } else {
+   try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
   }
 
+  if (!exists("out_total")) {
     if (1 %in% model_ssid) {
-  out_mature1 <- make_vector_data(predictions,
+
+      .predictions <- filter(predictions, ssid == 1)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
+      out_mature1 <- make_trend_data(predictions,
         ssid = c(1),
-    start_time = 2003,
-    # start_time = 2005,
+        start_time = start_year,
         skip_time = 2004,
@@
         delta_t_step = 2,
-    indices = c(1, 1, 1, 1, 2, 2, 2, 2),
-    # indices = c(1, 1, 1, 2, 2, 2, 2),
-    variable_names = "est_exp",
-    thresholds = c(0.001)
+        indices = indices,
+        variable_names = "est_exp"
       )
@@
     if (3 %in% model_ssid) {
-  out_mature3 <- make_vector_data(predictions,
+      .predictions <- filter(predictions, ssid == 3)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
+      out_mature3 <- make_trend_data(predictions,
         ssid = c(3),
-    start_time = 2005,
+        start_time = start_year,
         input_cell_size = 2,
@@
         delta_t_step = 2,
-    indices = c(1, 1, 1, 2, 2, 2, 2),
-    variable_names = "est_exp",
-    thresholds = c(0.001)
+        indices = indices,
+        variable_names = "est_exp"
       )
@@
     if (4 %in% model_ssid) {
-  out_mature4 <- make_vector_data(predictions,
+
+      .predictions <- filter(predictions, ssid == 4)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
+      out_mature4 <- make_trend_data(predictions,
         ssid = c(4),
-    start_time = 2004,
+        start_time = start_year,
         input_cell_size = 2,
@@
         delta_t_step = 2,
-    indices = c(1, 1, 1, 1, 2, 2, 2, 2),
-    variable_names = "est_exp",
-    thresholds = c(0.001)
+        indices = indices,
+        variable_names = "est_exp"
       )
@@
     if (16 %in% model_ssid) {
-  out_mature16 <- make_vector_data(predictions,
+
+      .predictions <- filter(predictions, ssid == 16)
+      #unique(.predictions$year)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
+      out_mature16 <- make_trend_data(predictions,
         ssid = c(16),
-    start_time = 2004,
+        start_time = start_year,
         skip_time = 2007,
@@
     }
+
     out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)
 
-saveRDS(out_mature,
-  file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
-)
+    if (exists("adult_predictions")) {
+            saveRDS(out_mature, file = paste0("data/", spp, "/trends-", 
+        spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"))
 
-if (!exists("adult_predictions")) {
+         } else {
       out_total <- out_mature
-  saveRDS(out_total,
-    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
-  )
+      rm(out_mature)
+      saveRDS(out_total, file = paste0("data/", spp, "/trends-", 
+        spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"))
+    }
+  }
 }
@@
 
-Optional plot of current total biomass 
-```{r eval=FALSE}
-if (exists("out_total")) {
-  current_biomass <- plot_vocc(out_total,
-    fill_col = "var_1_e",
-    fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
+### Recent biomass with trawl footprint
+```{r }
+if (!exists("out_mature")) {out_recent <- out_total} else {out_recent <- out_mature}
+
+  trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
+  proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
+  trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)
+
+  current_biomass <- plot_vocc(out_recent,
+    fill_col = "est_exp_e",
+    fill_label = "Mean\nbiomass (kg/ha)\npost-2010",
     raster_alpha = 1,
@@
   current_biomass <- current_biomass +
-    ggtitle(paste0("", species, " total biomass"))
+    geom_sf(data = trawl_footprint, aes(geometry = geometry / 1000), colour = "gray", fill = "gray", alpha = 0.2, lwd = 0.25, inherit.aes = FALSE) +
+    coord_sf(xlim = c(180, 820), ylim = c(5350, 6050), datum = NULL) +
+    xlab("") +
+    ylab("") +
+    theme(panel.grid.major = element_line(colour = "transparent")) +
+    ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
+  current_biomass
 
-  current_biomass
-}
 ```
 
+## Trends in biomass density
 ```{r, error=TRUE}
-try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
-
 if (exists("out_mature")) {
@@
     vec_aes = NULL,
-    # axis_lables = TRUE,
     white_zero = TRUE,
+    high_fill = "deepskyblue3", # "lightseagreen", # limegreen",
+    low_fill = "red", # "deeppink3",# "steel blue 4",
     transform_col = fourth_root_power,
     legend_position = c(0.2, 0.25)
-  ) + ggtitle(paste0("", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))))
+  ) + ggtitle(paste0(
+    "", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))
+  ))
   biotrend_ad
 } else {
-  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
-
   biotrend_ad <- plot_vocc(out_total,
@@
     vec_aes = NULL,
-    # axis_lables = TRUE,
     white_zero = TRUE,
+    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
+    low_fill = "red", # "steel blue 4",
     transform_col = fourth_root_power,
     legend_position = c(0.2, 0.25)
-  ) + ggtitle(paste0("", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))))
+  ) + ggtitle(paste0(
+    "", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))
+  ))
   biotrend_ad
@@
     ggtitle(paste0("", species, " immature biomass \n "))
+
   p_imm_all
@@
 if (exists("adult_predictions")) {
-  # if (length(model_ssid)>1){
   if (1 %in% model_ssid) {
@@
   if (3 %in% model_ssid) {
-    imm_predictions3 <- imm_predictions %>% filter(ssid == 3)
+    imm_predictions3 <- imm_predictions %>%
+      filter(ssid != 16) %>%
+      filter(ssid != 4)
   } else {
@@
   im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)
-
   im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]
@@
       ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))
+
     p_imm[[i]]
@@
 }
-# }
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
+# try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
+rm(out_imm)
+if (!exists("out_imm")) {
   if (exists("imm_predictions")) {
@@
     if (1 %in% model_ssid) {
-    out_im1 <- make_vector_data(predictions,
+
+      .predictions <- filter(predictions, ssid == 1)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+      out_im1 <- make_trend_data(predictions,
         ssid = c(1),
-      start_time = 2003,
-      # start_time = 2005,
+        start_time = start_year,
         skip_time = 2004,
@@
         delta_t_step = 2,
-      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
-      # indices = c(1, 1, 1, 2, 2, 2, 2),
-      variable_names = "est_exp",
-      thresholds = c(0.001)
+        indices = indices,
+        variable_names = "est_exp"
       )
@@
     if (3 %in% model_ssid) {
-    out_im3 <- make_vector_data(predictions,
+      .predictions <- filter(predictions, ssid == 3)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
+      out_im3 <- make_trend_data(predictions,
         ssid = c(3),
-      start_time = 2005,
+        start_time = start_year,
         input_cell_size = 2,
@@
         delta_t_step = 2,
-      indices = c(1, 1, 1, 2, 2, 2, 2),
-      variable_names = "est_exp",
-      thresholds = c(0.001)
+        indices = indices,
+        variable_names = "est_exp"
       )
@@
     if (4 %in% model_ssid) {
-    out_im4 <- make_vector_data(predictions,
+      .predictions <- filter(predictions, ssid == 4)
+      unique(.predictions$year)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
+      out_im4 <- make_trend_data(predictions,
         ssid = c(4),
-      start_time = 2004,
+        start_time = start_year,
         input_cell_size = 2,
@@
         delta_t_step = 2,
-      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
-      variable_names = "est_exp",
-      thresholds = c(0.001)
+        indices = indices,
+        variable_names = "est_exp"
       )
@@
     if (16 %in% model_ssid) {
-    out_im16 <- make_vector_data(predictions,
+      .predictions <- filter(predictions, ssid == 16)
+      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
+      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
+      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
+      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
+
+      out_im16 <- make_trend_data(predictions,
         ssid = c(16),
-      start_time = 2004,
+        start_time = start_year,
         input_cell_size = 2,
@@
     out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
-  saveRDS(out_imm,
-    file = paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
-  )
+    saveRDS(out_imm, file = paste0(
+      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds"
+    ))
   } else {
     rm(out_imm)
+  }
 }
@@
 ```{r}
-try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covariates, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
-
 if (exists("out_imm")) {
@@
     white_zero = TRUE,
+    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
+    low_fill = "red", # "steel blue 4",
     transform_col = fourth_root_power,
-    # axis_lables = TRUE,
     legend_position = c(0.2, 0.25)
-  ) + ggtitle(paste0("", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))))
+  ) + ggtitle(paste0(
+    "", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))
+  ))
   biotrend_im
@@
 
-Retrieve (or calculate) predicted values for temperature
-```{r, eval=FALSE}
-# m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
-# temperature <- predict(m_temp, newdata = nd_all)
-# saveRDS(temperature, file = "data/predicted_temp_allyears.rds")
-
-temperature <- readRDS("data/predicted_temp_allyears.rds")
-temperature <- filter(temperature, ssid %in% model_ssid)
-```
-
-```{r, eval=FALSE}
-temp <- plot_facet_map(temperature, "est", transform_col = no_trans, viridis_option = "D") +
-  labs(fill = "°C") +
-  ggtitle("Bottom temperature")
-print(temp)
-```
-
-```{r}
+Save trend plots
+```{r results = "hide"}
 png(
-  file = paste0("figs/", spp, "/trends-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
   res = 600,
   units = "in",
-  width = 11, # The width of the plot in inches
-  height = 8
+  width = 11,
+  height = 8.5
 )
@@
 
-```{r}
-biomass_plots <- c(p_adult, p_imm)
+Save spatiotemporal biomass plots
+```{r results = "hide"}
+if(exists("adult_predictions")) { biomass_plots <- c(p_adult, p_imm) } else {
+  biomass_plots <- c(p_adult)
+}
 
 png(
-  file = paste0("figs/", spp, "/biomass-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
   res = 600,
   units = "in",
-  width = 16, # The width of the plot in inches
+  width = 16,
   height = 8.5
@@
 
-```{r}
+Save plot medley
+```{r results = "hide"}
 png(
-  file = paste0("figs/", spp, "/exploratory-", spp, covariates, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
+  file = paste0("figs/", spp, "/exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
   res = 600,
   units = "in",
-  width = 9.5, # The width of the plot in inches
+  width = 9.5,
   height = 12.5
-) # The height of the plot in inches
+)
 
 gridExtra::grid.arrange(
+  current_biomass,
+  p_prop_imm,
   p_adult_all, p_imm_all,
   biotrend_ad, biotrend_im,
-  # current_biomass,
-  p_prop_imm, temp,
   nrow = 3
@@
 ```
+
+```{r}
+plot_map_raster <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    coord_fixed()+
+gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
+}
+
+ # plot_map_raster(adult_predictions, "epsilon_st") +
+ #  ggtitle("Spatiotemporal random effects only") +
+ #  facet_wrap(~year) +
+ #  scale_fill_gradient2( high = "Steel Blue 4",
+ #  low = "Red 3") + coord_fixed()
+
+ if (exists("adult_predictions")) {
+  p_st <- list()
+  for (i in seq_len(length(model_ssid))) {
+    p_st[[i]] <- plot_map_raster(ad_prediction_list[[i]], "epsilon_st") +
+      ggtitle("Spatiotemporal random effects only") +
+      facet_wrap(~year) +
+      scale_fill_gradient2(high = "#66C2A5", #mid = "black", 
+        low = "#5E4FA2"#, trans = fourth_root_power
+      )
+    p_st[[i]]
+   }
+  print(p_st)
+
+  # p_adult_st <- plot_facet_map(adult_predictions, "epsilon_st",
+  #   white_zero = TRUE
+  #   #raster_limits = c(0, max_raster),
+  #   #legend_position = "none",
+  #   #transform_col = fourth_root_power
+  # ) +
+  #   #labs(fill = "kg/ha") +
+  #   ggtitle(paste0("", species, "mature spatiotemporal variation"))
+  # print(p_adult_st)
+} else {
+  # p_adult_st <- plot_facet_map(predictions, "epsilon_st",
+  #   transform_col = fourth_root_power
+  # ) +
+  #   #labs(fill = "kg/ha") +
+  #   ggtitle(paste0("", species, "total spatiotemporal variation"))
+  # 
+  # print(p_adult_st)
+}
+```
+
+Save spatiotemporal residual plots
+```{r results = "hide"}
+# if(exists("adult_predictions")) { resid_plots <- c(p_st, p_imm) } else {
+#   biomass_plots <- c(p_adult)
+# }
+
+png(
+  file = paste0("figs/", spp, "/biotic-residuals-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
+  res = 600,
+  units = "in",
+  width = 16,
+  height = 8.5
+)
+gridExtra::grid.arrange(
+  grobs = p_st,
+  nrow = 2
+)
+dev.off()
+```

### 2020-02-07 analysis/VOCC/6-vocc-regression.R

@@ -9,7 +9,7 @@ dyn.load(dynlib("vocc_regression"))
 source("vocc-regression-functions.R")
 
 y_type <- "vel"
-# y_type <- "trend"
+y_type <- "trend"
 
 model_type <- "-vel"
 model_type <- "-vel-no-fishing"
@@ -19,7 +19,7 @@ model_type <- "-vel-no-fishing"
 #  model_type <- "-trend-no-covs"
 #  model_type <- "-trend-only"
  # model_type <- "-trend-grad"
- # model_type <- "-trend-with-do"
+model_type <- "-trend-with-do"
 w_genus <- F
 is_null <- T
 
@@ -43,7 +43,7 @@ data_type <- "mature-95-all-do"
 #  null_number <- ""
 null_number <- "-1"
  null_number <- "-2"
-#  null_number <- "-3"
+  null_number <- "-3"
 
 d <- readRDS(paste0("data/", data_type, "-with-null", null_number, ".rds"))
 d <- na.omit(d) %>% as_tibble()

### 2020-02-07 analysis/VOCC/7-vocc-chopsticks.R

@@ -1,0 +1,95 @@
+##############################
+##############################
+#### CHOPSTICK PLOTS
+##############################
+##############################
+model <- new_model
+
+model <- readRDS("data/trend_01-25-mature-95-all-temp-all-main-2.rds")
+data_type <- "mature-95-all-temp"
+
+
+model <- readRDS("data/trend_01-25-mature-80-all-temp-all-main-2.rds")
+data_type <- "mature-80-all-temp"
+
+model <- readRDS("data/trend_01-25-mature-80-all-temp-all-main-sim-2.rds")
+data_type <- "mature-80-all-temp-sim"
+
+
+model <- readRDS("data/trend_01-25-mature-90-all-temp-all-main.rds") # 90
+data_type <- "mature-90-all-temp"
+
+model <- readRDS("data/trend_01-25-mature-90-all-temp-all-main-sim-2.rds")
+data_type <- "mature-90-all-temp-sim"
+
+data_type <- "mature-90-vel-sim"
+
+data_type <- "mature-95-all-temp"
+
+#### INTERACTIONS WIHT TEMP TREND
+title_all <- "Interation plots "
+y_label <- "Predicted biomass trend"
+
+plot_fuzzy_chopsticks(model,
+  x_variable = "temp_trend_scaled", type = "temp",
+  y_label = y_label
+) + #ylim(-1.5, 1) + # 
+  #facet_wrap(vars(species), scales="free_y") +
+  ggtitle(paste(title_all, "(", data_type, ")"))
+
+plot_fuzzy_chopsticks(model,
+  x_variable = "temp_trend_scaled", type = "biomass",
+  y_label = y_label
+) + # facet_wrap(vars(species), scales="free_y") +
+  ggtitle(paste(title_all, "(", data_type, ")"))
+
+plot_fuzzy_chopsticks(model,
+  x_variable = "temp_trend_scaled", type = "temp_grad",
+  y_label = y_label
+) + facet_wrap(vars(species), scales = "free_y") +
+  ggtitle(paste(title_all, "(", data_type, ")"))
+
+
+title_all <- "Interation plots "
+y_label <- "Predicted biomass vel"
+
+plot_fuzzy_chopsticks(model,
+  x_variable = "squashed_temp_vel_scaled", type = "temp",
+  y_label = y_label
+) +# ylim(-1, 1) + # facet_wrap(vars(species), scales="free_y") +
[…]

### 2020-02-12 analysis/SOPO/1-trawl-split-by-maturity.Rmd
tweak and comment to maturity split function

@@
     # region: "Both odd year surveys"
-    threshold: 0.5
+    threshold: 0.05
     split: TRUE
@@
 ```{r}
-rm(biomass)
 rm(maturity)
 
-#try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
-
-if (!exists("biomass")) {
-  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
-  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
-  bath <- readRDS("data/bathymetry-data")
+survey_sets <- readRDS(paste0("../VOCC/raw/event-data-", spp, ""))
+fish <- readRDS(paste0("../VOCC/raw/bio-data-", spp, ""))
+bath <- readRDS("../VOCC/data/bathymetry-data")
 
@@
 saveRDS(maturity$data, file = paste0("data/", spp, 
-    "/data-by-maturity-", spp, "-", ssid_string, ".rds"
-    ))
-}
+    "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
 
@@
 png(
-    file = paste0("figs/_maturity_ogives/maturity-", prop_pos,"-", spp, "-", ssid_string, ".png"), 
+    file = paste0("figs/maturity-", prop_pos,"-", spp, "-", ssid_string, ".png"), 
     res = 600,
@@
 }
-
 dev.off()
-
 try(print(maturity$maturity))

### 2020-02-12 R/mature-biomass-est.R

@@ -55,7 +55,7 @@ split_catch_maturity <- function(survey_sets, fish, bath,
     filter(year %in% years)
 
   # use internal version of tidy_survey_sets to include ssid and month columns
-  tidy_sets <- tidy_survey_sets(survey_sets, survey = survey, years = years)
+  tidy_sets <- gfranges::tidy_survey_sets(survey_sets, survey = survey, years = years)
   tidy_sets <- add_missing_depths(tidy_sets, survey = survey, years = years, bath = bath)
 
   model_ssid <- unique(tidy_sets$ssid)
@@ -179,19 +179,20 @@ split_catch_maturity <- function(survey_sets, fish, bath,
 
     sets_w_ratio <- left_join(tidy_sets, set_ratio, by = "fishing_event_id")
 
-
+# browser()
     # chose value to use when a sample mass ratio is not available
     na_value <- mean(sets_w_ratio$mass_ratio_mature, na.rm = TRUE)
     sets_w_ratio$mass_ratio_mature[is.na(sets_w_ratio$mass_ratio_mature)] <- na_value
 
     # add column to check for discrepencies between total catch weight and biological sample weights
+    # est_sample_mass is in g while catch_weight is in kg?
     sets_w_ratio$errors <- sets_w_ratio$est_sample_mass - sets_w_ratio$catch_weight * 1000
 
     data <- sets_w_ratio %>%
       mutate(adult_density = density * mass_ratio_mature, imm_density = density * (1 - mass_ratio_mature))
 
     if (plot) {
-      try(maturity_plot <- plot_mat_ogive(m) +
+      try(maturity_plot <- gfranges::plot_mat_ogive(m) +
         ggplot2::ggtitle(paste("Length at maturity for", species, "surveys", ssid_string, "")))
 
       try(mass_plot <- ggplot(fish_maturity, aes(length, new_mass, colour = as.factor(sex))) +

### 2020-02-12 analysis/VOCC/1-render-maturity.R

@@ -64,6 +64,36 @@ list_species <- c(
 #   "Pacific Halibut"
 # )
 
+# species with small sample
+
+### NEW ADDITIONS FOR SOPO #1
+list_species <- c(
+  "Pacific Hake",
+  "Pacific Tomcod",
+  "Rosethorn Rockfish",
+  "Slender Sole",
+  "Pacific Sanddab",
+  "Harlequin Rockfish",
+  "Copper Rockfish" ,
+  "Shortbelly Rockfish",
+  "Sandpaper Skate",
+  "Brown Cat Shark",
+  "Sand Sole",
+  "Butter Sole",
+  "Starry Flounder")
+
+### NEW ADDITIONS FOR SOPO #2
+list_species <- list(
+  "Buffalo Sculpin",
+  "Cabezon",
+  "Pacifc Staghorn Sculpin",
+  "Red Irish Lord",
+  "Sturgeon Poacher",
+  "Bigmouth Sculpin",
+  "Kelp Greenling",
+  "Threadfn Sculpin"
+)
+
 ## NEEDS SPECIAL SETTINGS
 list_species <- c("North Pacific Spiny Dogfish","Longspine Thornyhead","Sand Sole")
 

### 2020-02-12 analysis/SOPO/2-trawl-no-covs-model.Rmd
first draft of SOPO models with maturity split and both HS and QCS combined

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    # species: "Redbanded Rockfish"
    # species: "North Pacific Spiny Dogfish"
    # species: "Pacific Ocean Perch"
    # species: "Big Skate"
    # species: "Pacific Cod"
    #species: "Petrale Sole"
    species: "Starry Flounder"
    # species: "Bocaccio"
    # region: "All synoptic surveys"
    region: "Both odd year surveys"
    start_year: 2003
    knots: 250
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    covs: "-no-covs" # covariate string used to name model rds
    AR1: TRUE
    fixed_spatial: FALSE
    update_model: FALSE # TRUE
    update_model_check: FALSE #TRUE
    update_predictions: FALSE # TRUE
    update_index: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  # fig.path=paste0("figs/", spp, "/"),
  echo = FALSE, warning = FALSE, message = FALSE
)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

theme_set(
    gfplot::theme_pbs(base_size = 14) 
)
```

```{r params}
species <- params$species
region <- params$region
covariates <- params$covariates
covs <- params$covs
knots <- params$knots
priors <- FALSE

paste("region =", region)
paste("model covariates =", covariates)
paste("model label =", covs)
paste("priors =", priors)
paste("knots =", knots)
```

```{r, message=FALSE, warning=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

# Build spatiotemporal density model
Load and filter data 
```{r}
biomass <- readRDS(paste0(
  "data/", spp, "/data-by-maturity-", spp, "-1n3n4n16.rds"
))

if(nrow(biomass)<2000) {stop("Need to recalculate split by maturity!")}

# covars <- readRDS("data/event-covariates.rds")
# data <- dplyr::left_join(biomass, covars)
data <- biomass
# scale predictors before filtering to ensure mean and SD are global
data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth))
data <- gfranges::scale_predictors(data, # predictors = c(quo(depth)))
  predictors = c(quo(depth))
)
data <- data %>%
  mutate(depth = raw_depth) %>% 
  filter(ssid %in% model_ssid) %>%
  filter(year >= params$start_year)

rm(adult_biomass)
rm(imm_biomass)
rm(total_biomass)
```

Make mesh
```{r}
if (region == "Both odd year surveys") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 250)
}

if (region == "West Coast Vancouver Island") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
}

if (region == "West Coast Haida Gwaii") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
}

if (region == "All synoptic surveys") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
}

sdmTMB::plot_spde(spde)
```

Run climate independent sdmTMB model
```{r }
if (params$update_model) {
  tictoc::tic()
  if (any(names(data) == "adult_density")) {
    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
    ))

    adult_biomass <- sdmTMB::sdmTMB(
      data = data,
      adult_formula,
      # time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = params$AR1,
      include_spatial = params$fixed_spatial,
      reml = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(adult_biomass,
      file = paste0("data/", spp,
        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"))

   try({ imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
    ))

    imm_biomass <- sdmTMB::sdmTMB(
      data = data,
      imm_formula,
      # time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = params$AR1, # changed to TRUE on Oct 17 2019
      reml = TRUE,
      include_spatial = params$fixed_spatial,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass,
      file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"))
   })

  } else {
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
    # dens_formula <- as.formula(paste("density ~ 0"))

    total_biomass <- sdmTMB::sdmTMB(
      data = data,
      dens_formula,
      # time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = params$AR1,
      reml = TRUE,
      include_spatial = params$fixed_spatial,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(total_biomass, file = paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
    ))
  }
  tictoc::toc()
}
```

Save predictions for spatial grid
```{r}
rm(ad_predictions)
rm(im_predictions)
rm(predicted)

if(params$update_predictions) {
  # nd_all <- readRDS(paste0("data/nd_just_depth.rds"))
  # nd_all <- readRDS(paste0("data/nd_whole_coast_index.rds"))
  nd_all <- readRDS(paste0("data/nd_odd.rds"))

if (any(names(data) == "adult_density")) {

try({

if (!exists("adult_biomass")) {    
  adult_biomass <- readRDS(paste0("data/", spp, 
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, 
    "-ar1-", params$AR1, "-reml.rds"))
}

    nd <- nd_all %>%
      filter(ssid %in% model_ssid) %>%
      filter(year %in% unique(adult_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    ad_predictions <- predict(adult_biomass, newdata = nd, return_tmb_object = TRUE)
    # saveRDS(ad_predictions, file = paste0("data/", spp,
    #   "/predictions-", spp, covs, "-", ssid_string, 
    #   "-mature-biomass-ar1-", params$AR1, "-reml.rds"))
    saveRDS(ad_predictions,
      file = paste0("data/", spp,
        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"))
})

  try({

    if (!exists("imm_biomass")) {
    imm_biomass <- readRDS(paste0("data/", spp,
      "/mod-imm-biomass-", spp, covs, "-", ssid_string, 
      "-ar1-", params$AR1, "-reml.rds"))
    }

    nd <- nd_all %>%
      filter(ssid %in% model_ssid) %>%
      filter(year %in% unique(imm_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)
    im_predictions <- predict(imm_biomass, newdata = nd, return_tmb_object = TRUE)

    # saveRDS(im_predictions, file = paste0("data/", spp,
    #   "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-ar1-", params$AR1, "-reml.rds"
    # ))

     saveRDS(im_predictions, file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"))

  })

} else {

  if (!exists("total_biomass")) {
  total_biomass <- readRDS(paste0("data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, 
    "-ar1-", params$AR1, "-reml.rds"))
  }

  nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(total_biomass$data$year))
  nd <- na.omit(nd)
  nd$year <- as.integer(nd$year)

  predicted <- predict(total_biomass, newdata = nd, return_tmb_object = TRUE)

  # saveRDS(predicted, file = paste0("data/", spp,
  #   "/predictions-", spp, covs, "-", ssid_string, 
  #   "-total-biomass-ar1-", params$AR1,"-reml.rds"))

  saveRDS(predicted, file = paste0("data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, 
    "-ar1-", params$AR1, "-reml.rds"))

}
}
```

### Check residuals
```{r}

if(params$update_model_check) {
try({

if (!exists("ad_predictions")) {    
  ad_predictions <- readRDS(paste0("data/", spp, 
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, 
    "-ar1-", params$AR1, "-reml.rds"))
}
  point_predictions <- predict(ad_predictions$fit_obj)
  point_predictions$residuals <- residuals(ad_predictions$fit_obj)
})

try({

if (!exists("total_biomass")) {    
  predicted <- readRDS(paste0("data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, 
    "-ar1-", params$AR1, "-reml.rds"
  ))
}
  point_predictions <- predict(predicted$fit_obj)
  point_predictions$residuals <- residuals(predicted$fit_obj)
})

saveRDS(point_predictions, file = paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
))
}
```

```{r}
check_predictions <- readRDS(paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
))

check_predictions <- filter(
  check_predictions,
  # ssid %in% c(1,3)) %>% filter (
  year > 2003
)

try({
  g <- ggplot(check_predictions, aes(adult_density, residuals, colour = adult_density)) +
    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
})

try({
  g <- ggplot(check_predictions, aes(density, residuals, colour = density)) +
    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
})

g
```

```{r}
try({
  g2 <- ggplot(check_predictions, aes((est), residuals, colour = adult_density)) +
    geom_point() + # scale_x_continuous(trans = 'log10') +
    facet_wrap(~year)
})

try({
  g2 <- ggplot(check_predictions, aes((est), residuals, colour = density)) +
    geom_point() + # scale_x_continuous(trans = 'log10') +
    facet_wrap(~year)
})

g2
```

```{r}
ggplot(check_predictions, aes(X, Y, colour = (residuals))) +
  geom_point() +
  scale_colour_gradient2() +
  scale_x_continuous(trans = "log10") +
  facet_wrap(~year)
```

Transform estimates to kg/ha
```{r}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

if (!exists("ad_predictions")) {
try({
  ad_predictions <- readRDS(paste0("data/", spp, 
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, 
    "-ar1-", params$AR1, "-reml.rds"))
  })
  try({
    im_predictions <- readRDS(paste0("data/", spp, 
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, 
    "-ar1-", params$AR1, "-reml.rds"))
  })
  if (!exists("ad_predictions")) {
    try({
      predicted <-  readRDS(paste0("data/", spp,
        "/model-total-biomass-", spp, covs, "-", ssid_string, 
        "-ar1-", params$AR1, "-reml.rds"))
    })
  }
}

if (exists("ad_predictions")) {
  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions <- ad_predictions$data
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000

  try({
    imm_predictions <- im_predictions$data
    imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

    adult_predictions$total_bio <- imm_predictions$est_exp +   
        adult_predictions$est_exp
    imm_predictions$prop_imm <- imm_predictions$est_exp /
        (adult_predictions$est_exp + imm_predictions$est_exp)

  plot_imm <- filter(imm_predictions, year>2004)
  max_imm <-  signif(max(plot_imm$est_exp), digits = 2)
  max_raster_imm <- quantile(plot_imm$est_exp, 0.999)

  saveRDS(imm_predictions, file = paste0("data/", spp,
    "/sopo-predictions-", spp, covs, "-imm-biomass-ar1-", 
    params$AR1, "-reml.rds"))
  })

  saveRDS(adult_predictions, file = paste0("data/", spp,
    "/sopo-predictions-", spp, covs, "-mat-biomass-ar1-", 
    params$AR1, "-reml.rds"))

  plot_ad <- filter(adult_predictions, year>2004)

  max_raster_ad <- quantile(plot_ad$est_exp, 0.999)
  max_adult <- signif(max(plot_ad$est_exp), digits = 2)

  # model_ssid <- unique(adult_predictions$ssid)
} else {

  predictions <- predicted$data
  predictions$est_exp <- exp(predictions$est) * 10000
  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .99)
  max_bio <- signif(quantile(predictions$total_bio, .999), digits = 2)

  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/sopo-predictions-", spp, covs, "-total-biomass-ar1-", params$AR1, "-reml.rds"
  ))

}
```

### Model summary
```{r}
 try({ ad_predictions$fit_obj})
 try({ im_predictions$fit_obj})
 try({ predicted$fitobj})
```

### Plot biomass predictions

```{r, error=TRUE}
legend_coords <- c(0.9, 0.17) #"none"

if (exists("adult_predictions")) {

 p_adult_all <- adult_predictions %>% filter(year>2004) %>%
   mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>%  
   plot_facet_map("est_exp", 
      raster_limits = c(0, max_raster_ad),
      legend_position =  legend_coords, 
      transform_col = fourth_root_power
    ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))

 print(p_adult_all)

try({
 p_imm_all <- imm_predictions %>% filter(year>2004) %>%
    mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>%
    plot_facet_map("est_exp",
      raster_limits = c(0, max_raster_imm),
      legend_position =  legend_coords, 
      transform_col = fourth_root_power
    ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n(max = ", max_imm, " kg/ha)"))

print(p_imm_all)
   })

} else {

p_adult_all <- predictions  %>% 
  mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>% 
  plot_facet_map("total_bio",
    raster_limits = c(0, max_raster),
    legend_position =  legend_coords, 
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))

  print(p_adult_all)
}
```

### Calculate index of total biomass for full survey grid

```{r}
rm(ind)
rm(ind_imm)
if(params$update_index) {
if (exists("ad_predictions")) {
ind <- get_index(ad_predictions, bias_correct = FALSE)
ind$species <- species
ind$maturity <- "mature"
ind$ssid <- ssid_string
ind$covs <- covs

write.csv(ind, file = paste0("data/_indices/sopo-index-", 
  spp,"-", ssid_string, "-reml.csv"))

try ({
  ind_imm <- get_index(im_predictions, bias_correct = FALSE)
  ind_imm$species <- species
  ind_imm$maturity <- "immature"
  ind_imm$ssid <- ssid_string
  ind_imm$covs <- covs

  write.csv(ind_imm, file = paste0("data/_indices/sopo-index-imm-", 
    spp,"-", ssid_string, "-reml.csv"))
})
} else {
  ind <- get_index(predicted, bias_correct = FALSE)
  ind$species <- species
  ind$maturity <- "all"
  ind$ssid <- ssid_string
  ind$covs <- covs

  write.csv(ind, file = paste0("data/_indices/sopo-index-", 
    spp,"-", ssid_string, "-reml.csv"))
}
}

try({head(ind)})
try({head(ind_imm)})
```

```{r}
# scale <- 2 * 2 / 1000 # if density was in kg/km2: 2 x 2 km grid and converted from kg to tonnes 
scale <- 1000000 * 4 /1000 # if density was in kg/m2: m2 to km2 * 4 km2 grid size and kg to tonnes
ad_col <- "deepskyblue4"

ind <- read.csv(paste0("data/_indices/sopo-index-", 
  spp,"-", ssid_string, "-reml.csv"))

ind %>% filter(year > 2004) %>%
ggplot(aes(year, est*scale)) + geom_line(col=ad_col) +
  geom_ribbon(aes(ymin = lwr*scale, ymax = upr*scale), fill=ad_col, alpha = 0.4) +
  xlab('Year') +  
  ylab('Mature biomass estimate (metric tonnes)') +
  gfplot::theme_pbs()
```

Add immature trend if available
```{r}
try({
  ind_imm <- read.csv(paste0("data/_indices/sopo-index-imm-", 
    spp,"-", ssid_string, "-reml.csv"))# %>% filter(year > 2004)
})

if (exists("ind_imm")) {
ratio <- mean(ind$est)/max(ind_imm$est)
imm_col <- "orangered"
ind %>% #filter(year > 2004) %>%
ggplot(aes(year, est*scale)) + geom_line(col = ad_col, ) +
  geom_ribbon(aes(ymin = lwr*scale, ymax = upr*scale), fill = ad_col, alpha = 0.4) +
# adding the relative humidity data, transformed to match roughly the range of the temperature
  geom_ribbon(aes(ymin = ind_imm$lwr*scale*ratio, ymax = ind_imm$upr*scale*ratio), 
    fill = "orangered", alpha = 0.4) + 
  geom_line(aes(ind_imm$year, ind_imm$est*scale*ratio), col = "orangered") +
# now adding the secondary axis, following the example in the help file ?scale_y_continuous and, reverting the above transformation
  scale_y_continuous(sec.axis = sec_axis(~./ratio, name = "Immature biomass estimate (metric tonnes)")) +
  xlab('Year') + 
  ylab('Mature biomass estimate (metric tonnes)')+
  gfplot::theme_pbs(base_size = 16)
}
```

### 2020-02-12 analysis/SOPO/1-render-maturity.R

@@ -1,0 +1,137 @@
+getwd()
+setwd(here::here("/analysis/SOPO"))
+env <- new.env() #parent = baseenv()
+
+# SPECIES with maturity
+list_species <- c(
+  "Redstripe Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Walleye Pollock",
+  "Pacific Cod",
+  "Sablefish",
+  "Lingcod",
+  "Redstripe Rockfish",
+  "Yellowmouth Rockfish",
+  "Canary Rockfish", # schooling, winter-birthing
+  "Darkblotched Rockfish",
+  "Greenstriped Rockfish",
+  "Pacific Ocean Perch", # schooling
+  "Redbanded Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Sharpchin Rockfish",
+  "Shortbelly Rockfish", # small sample
+  "Silvergray Rockfish",
+  "Splitnose Rockfish",
+  "Yellowtail Rockfish", # schooling
+  "Longspine Thornyhead",
+  "Shortspine Thornyhead",
+  "Arrowtooth Flounder",
+  "Rex Sole",
+  "Petrale Sole",
+  "English Sole",
+  "Dover Sole",
+  "Southern Rock Sole",
+  "Flathead Sole",
+  "Curlfin Sole",
+  "Sand Sole",
+  "Widow Rockfish",
+  "Quillback Rockfish",
+  "Bocaccio",
+  "Shortraker Rockfish",
+  "Yelloweye Rockfish"
+)
+
+
+# species with small sample
+list_species <- c(
+  "Pacific Hake",
+  "Pacific Tomcod",
+  "Rosethorn Rockfish",
+  "Slender Sole",
+  "Pacific Sanddab",
+  "Harlequin Rockfish"
+  # "Copper Rockfish" ,
+  # "Shortbelly Rockfish",
+  # "Sandpaper Skate",
+  # "Brown Cat Shark" 
+)
+
+list_species <- c( 
+  "Sand Sole",
+  "Butter Sole",
+  "Starry Flounder")
+
+### more species for SOPO
+# yet to be retrieved
+list_species <- list(
+  "Buffalo Sculpin",
+  "Cabezon",
+  "Pacifc Staghorn Sculpin",
+  "Red Irish Lord",
+  "Sturgeon Poacher",
+  "Bigmouth Sculpin",
+  "Kelp Greenling",
[…]

### 2020-02-12 analysis/SOPO/2-render-models.R

@@ -1,0 +1,209 @@
+getwd()
+setwd(here::here("/analysis/SOPO"))
+env <- new.env() # parent = baseenv()
+
+
+### build time-varying depth models
+# list_regions <- c("All synoptic surveys")
+list_regions <- c("Both odd year surveys")
+
+
+list_species <- c(
+  # "Big Skate",
+  # "Longnose Skate",
+  # "Spotted Ratfish",
+  # "North Pacific Spiny Dogfish",
+  # "Pacific Tomcod",
+  # "Walleye Pollock",
+  # "Pacific Cod",
+  "Lingcod",
+  "Sablefish",
+  
+  # "Bocaccio",
+  "Canary Rockfish",
+  "Copper Rockfish", # small sample
+  # "Darkblotched Rockfish", # need predictions still
+  "Greenstriped Rockfish",
+  "Pacific Ocean Perch",
+  "Quillback Rockfish",
+  "Redbanded Rockfish",
+  # "Redstripe Rockfish", # faild to converge?
+  "Rougheye/Blackspotted Rockfish Complex",
+  # "Sharpchin Rockfish",
+  # "Shortbelly Rockfish", # small sample
+  # "Shortraker Rockfish",
+  "Silvergray Rockfish",
+  "Splitnose Rockfish",
+  "Widow Rockfish", # schooling
+  # "Yellowmouth Rockfish",
+  # "Yellowtail Rockfish", # schooling
+  # "Yelloweye Rockfish" # ran mature only with AR1, rerunning without
+  # "Longspine Thornyhead",
+  "Shortspine Thornyhead",
+
+  # "Pacific Halibut",
+  "Arrowtooth Flounder",
+  "Curlfin Sole",
+  "Dover Sole",
+  "English Sole",
+  # "Flathead Sole", # needs predictions
+  "Petrale Sole",
+  # "Rex Sole", # needs predictions
+  "Southern Rock Sole"
+)
+
+
+list_species <- c(
+  # "Pacific Hake",
+  # "Pacific Tomcod",
+  # "Rosethorn Rockfish",
+  # "Slender Sole",
+  # "Pacific Sanddab",
+  # "Harlequin Rockfish",
+  # "Copper Rockfish" ,
+  # # "Shortbelly Rockfish",
+  "Sandpaper Skate"
+  # "Brown Cat Shark"
+)
+
+list_species <- c( 
+#   "Sand Sole",
+# "Butter Sole",
+"Starry Flounder"
[…]

### 2020-02-12 analysis/SOPO/3-pool-indices.R

@@ -1,0 +1,17 @@
+getwd()
+setwd(here::here("/analysis/SOPO/data"))
+library("dplyr")
+
+mydir = paste0( "_indices")
+
+####################
+myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
+myfiles
+indices <- do.call(rbind, lapply(myfiles, read.csv)) 
+glimpse(indices)
+
+#####################
+
+saveRDS(indices, file = paste0(
+  "sopo-2019-indices.rds"
+))

### 2020-02-12 analysis/SOPO/4-plot-index.R

@@ -1,0 +1,116 @@
+getwd()
+setwd(here::here("/analysis/SOPO"))
+library("dplyr")
+library("ggplot2")
+
+
+all_indices <- readRDS(paste0("data/sopo-2019-indices.rds"))
+# View(all_indices)
+
+all_indices <- filter(all_indices, covs == "-no-covs")
+# all_indices <- filter(all_indices, covs != "-no-covs")
+
+scale <- 1000000 * 4 /1000 # if density was in kg/m2: m2 to km2 * 4 km2 grid size and kg to tonnes
+
+all_indices$upr[is.na(all_indices$se) ] <- NA
+all_indices$upr[all_indices$se > 1.25 ] <- NA
+
+all_indices$reliable_est <- all_indices$est
+all_indices$reliable_est[all_indices$se > 1.25 ] <- NA
+
+# filter years after 2004 because incomplete spatial sampling prior causes errors to blow up
+indices <- all_indices %>% filter(year > 2004) %>%
+  # and calculate index to base ratio on (max, mean, or median)
+  group_by(species, maturity) %>% mutate( max_est = max(est), min_est = min(est), mean_est = mean(est)) %>% ungroup()
+
+# add back in est for years prior because potentially still interesting despite uncertainty
+est_only <- all_indices %>% select(year, species, maturity, est) %>% filter(year < 2005)
+indices <- full_join(indices, est_only , by=c("species", "year", "maturity", "est")) 
+
+indices <- indices %>% group_by(species, maturity) %>% mutate (max_upr = max(upr, na.rm = T), max_est2 = max(est), axis_max = pmax(max_upr, max_est2)) %>% ungroup()
+
+# calculate ratio of immature est to mature est for plotting on same figure
+ind <- indices %>% filter(maturity != "immature")
+ind_imm <- indices %>% filter(maturity == "immature")
+
+ind_spp <- list()
+for(i in unique(ind$species)) {
+  rm(ind_i)
+  rm(ind_ii)
+  ind_i <- filter(ind, species == i) %>% select(mean_est)
+  ind_ii <- filter(ind_imm, species == i) %>% select(max_est)
+   # browser()
+  if(nrow(ind_ii)<1) { ind_ii <- 0 }
+  
+  ratio <- ind_i[[1]]/ind_ii[[1]]
+  
+  ind_spp[[i]] <- filter(ind, species == i)
[…]

### 2020-02-12 analysis/SOPO/make-grids.R

@@ -1,0 +1,79 @@
+
+# choose base year(s) to create grid from
+# dummy_year <- c(2005, 2006)
+bath <- readRDS("../VOCC/data/bathymetry-data")
+# dat <- bath$data %>% scale_survey_predictors()
+dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth))
+dat <- gfranges::scale_predictors(dat, predictors = c(quo(depth)))
+
+# create grids for each ssid separately
+grid_locs <- gfplot::synoptic_grid %>%
+  dplyr::filter(survey == "SYN QCS") %>%
+  dplyr::select(X, Y, depth)
+grid_locs$depth_scaled <- (log(grid_locs$depth) - dat$depth_mean[1])/dat$depth_sd[1]
+grid_locs$depth_scaled2 <- grid_locs$depth_scaled^2
+# Expand the prediction grid to create a slice for each time:
+original_time1 <- sort(unique(filter(dat, ssid == 1)$year))
+nd_1 <- do.call("rbind",
+  replicate(length(original_time1), grid_locs, simplify = FALSE))
+nd_1[["year"]] <- rep(original_time1, each = nrow(grid_locs))
+nd_1[["ssid"]] <- 1
+unique(nd_1$year)
+
+grid_locs <- gfplot::synoptic_grid %>%
+  dplyr::filter(survey == "SYN HS") %>%
+  dplyr::select(X, Y, depth)
+grid_locs$depth_scaled <- (log(grid_locs$depth) - dat$depth_mean[1])/dat$depth_sd[1]
+grid_locs$depth_scaled2 <- grid_locs$depth_scaled^2
+# original_time3 <- sort(unique(filter(dat, ssid == 3)$year))
+nd_3 <- do.call("rbind",
+  replicate(length(original_time1), grid_locs, simplify = FALSE))
+nd_3[["year"]] <- rep(original_time1, each = nrow(grid_locs))
+nd_3[["ssid"]] <- 3
+unique(nd_3$year)
+
+# grid_locs <- gfplot::synoptic_grid %>%
+#       dplyr::filter(survey == "SYN WCVI") %>%
+#       dplyr::select(X, Y, depth)
+# grid_locs$depth_scaled <- (log(grid_locs$depth) - dat$depth_mean[1]) / dat$depth_sd[1]
+#   grid_locs$depth_scaled2 <- grid_locs$depth_scaled^2
[…]

### 2020-02-19 analysis/VOCC/1-split-biomass-by-maturity.Rmd
add options for changing ogive threshold and overriding maturity split to maturity rmd

@@
 params: 
-    species: "Longnose Skate"
-    #species:"Redbanded Rockfish"
-    #species: "Pacific Ocean Perch"
+    #species: "Bocaccio"
+    species: "Threadfin Sculpin"
+    #species: "Quillback Rockfish"
+    #species: "Yelloweye Rockfish"
+    #species: "Shortraker Rockfish"
+    # species: "Pacific Ocean Perch"
     region: "All synoptic surveys"
     # region: "Both odd year surveys"
+    threshold: 0.05
     split: FALSE

### 2020-02-19 analysis/VOCC/2-biomass-depth-only-model.Rmd
turn on control helper option in sdmTMB for depth only models

@@
       enable_priors = priors,
-      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
       silent = FALSE

### 2020-02-19 analysis/VOCC/3-vocc-temp.Rmd
correct indices for all temp years vocc

@@
 
-  if (years_sampled == 7) indices <- c(1, 1, 1, 2, 2, 2)
-  if (years_sampled == 8) indices <- c(1, 1, 1, 1, 2, 2, 2,2)
+  if (years_sampled == 7) indices <- c(1, 1, 1, 2, 2, 2, 2) # was 6 for some reason?
+  if (years_sampled == 8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
 } else {

### 2020-02-19 analysis/SOPO/2-trawl-no-covs-model.Rmd
options to change knot number and update models for just rare speciesf

@@
     # species: "Pacific Ocean Perch"
-    # species: "Big Skate"
+    # species: "Pacific Sand Lance"
     # species: "Pacific Cod"
-    #species: "Petrale Sole"
-    species: "Starry Flounder"
-    # species: "Bocaccio"
+    # species: "Buffalo Sculpin"
+    # species: "Cabezon"
+    species: "Harlequin Rockfish"
     # region: "All synoptic surveys"
@@
     start_year: 2003
-    knots: 250
+    knots: 300
     covariates: "" # +as.factor(ssid)
     #additional fixed-effect covariates
-    covs: "-no-covs" # covariate string used to name model rds
+    covs: "-no-covs-300" # covariate string used to name model rds
     AR1: TRUE
     fixed_spatial: FALSE
-    update_model: FALSE # TRUE
+    update_model: FALSE #TRUE
     update_model_check: FALSE #TRUE
-    update_predictions: FALSE # TRUE
+    update_predictions: FALSE #TRUE
     update_index: TRUE
@@
 priors <- FALSE
+update_model <- params$update_model
+update_model_check <- params$update_model_check
+update_predictions <- params$update_predictions
+update_index <- params$update_index
 
@@
 paste("priors =", priors)
-paste("knots =", knots)
 ```
@@
 
-if(nrow(biomass)<2000) {stop("Need to recalculate split by maturity!")}
+  positive_sets <- filter(biomass, density != 0)
+  prop_pos <- round(nrow(positive_sets)/nrow(biomass), digits = 3)
+      if (prop_pos < 0.1) {
+        update_model <- T
+        update_model_check <- T
+        update_predictions <- T
+        update_index <- T
+        knots <- 150
+        }
+    if (prop_pos < 0.05) {
+        update_model <- F
+        update_model_check <- F
+        update_predictions <- F
+        update_index <- T
+        knots <- 150
+        }
+
+# View(biomass)
+if (nrow(biomass)<2000) {
+  # update_model <- T
+  # update_model_check <- T
+  # update_predictions <- T
+  # update_index <- T
+  knots <- 100
+  if(prop_pos < 0.05) { knots <- 50 }
+  }
+if(nrow(biomass)<1000) {stop("Need to recalculate split by maturity!")}
 
@@
 rm(total_biomass)
+
 ```
@@
 if (region == "Both odd year surveys") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 250)
+  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
 }
@@
 
+```{r}
+paste("knots =", knots)
+```
+
 Run climate independent sdmTMB model
 ```{r }
-if (params$update_model) {
+
+if (update_model) {
   tictoc::tic()
@@
 
-if(params$update_predictions) {
+if(update_predictions) {
+
   # nd_all <- readRDS(paste0("data/nd_just_depth.rds"))
@@
 ```{r}
-
-if(params$update_model_check) {
+if(update_model_check) {
 try({
@@
 
-```{r}
+```{r }
+
+if(update_model_check) {
 check_predictions <- readRDS(paste0(
@@
   # ssid %in% c(1,3)) %>% filter (
-  year > 2003
+  year > 2004
 )
@@
 g
+}
 ```
@@
 
-g2
+try ({g2})
 ```
@@
 ```{r}
+try({
   ggplot(check_predictions, aes(X, Y, colour = (residuals))) +
@@
     facet_wrap(~year)
+})
 ```
+
+Load predictions
 
@@
   })
+
+  if (!exists("ad_predictions")) {
+     try({ ad_predictions <- readRDS(paste0("data/", spp,
+      "/mod-mat-biomass-", spp, "-no-covs-", ssid_string,
+      "-ar1-", params$AR1, "-reml.rds")) })
+  }
+
   try({
@@
   })
+    if (!exists("im_predictions")) {
+     try({ im_predictions <- readRDS(paste0("data/", spp,
+      "/mod-imm-biomass-", spp, "-no-covs-", ssid_string,
+      "-ar1-", params$AR1, "-reml.rds")) })
+  }
+
   if (!exists("ad_predictions")) {
@@
     })
+    if (!exists("predicted")) {
+     try({
+       predicted <- readRDS(paste0("data/", spp,
+      "/model-total-biomass-", spp, "-no-covs-", ssid_string,
+      "-ar1-", params$AR1, "-reml.rds"))
+       })
+  }
   }
@@
  try({ im_predictions$fit_obj})
- try({ predicted$fitobj})
+ try({ predicted$fit_obj})
 ```
@@
 rm(ind_imm)
-if(params$update_index) {
+library(data.table)
+if(update_index) {
+
 if (exists("ad_predictions")) {
+
   ind <- get_index(ad_predictions, bias_correct = FALSE)
@@
   ind$covs <- covs
-
+  ind <- as.data.frame(ind)
   write.csv(ind, file = paste0("data/_indices/sopo-index-", 
-  spp,"-", ssid_string, "-reml.csv"))
+    spp,"-", ssid_string, "-", covs, "-reml.csv"), row.names = FALSE)
 
@@
     ind_imm$covs <- covs
+    ind_imm <- as.data.frame(ind_imm)
+    write.csv(ind_imm, file = paste0("data/_indices/sopo-index-imm-", 
+      spp,"-", ssid_string, "-", covs, "-reml.csv"), row.names = FALSE)
+  })
 
-  write.csv(ind_imm, file = paste0("data/_indices/sopo-index-imm-", 
-    spp,"-", ssid_string, "-reml.csv"))
-})
 } else {
@@
 
+  ind <- as.data.frame(ind)
+
   write.csv(ind, file = paste0("data/_indices/sopo-index-", 
-    spp,"-", ssid_string, "-reml.csv"))
+    spp,"-", ssid_string, "-", covs, "-reml.csv"), row.names = FALSE)
 }
+try({ glimpse(ind_imm) })
+try({ glimpse(ind) })
 }  
-
-try({head(ind)})
-try({head(ind_imm)})
 ```
@@
 ind <- read.csv(paste0("data/_indices/sopo-index-", 
-  spp,"-", ssid_string, "-reml.csv"))
+  spp,"-", ssid_string, "-", covs, "-reml.csv"))
 
@@
   ylab('Mature biomass estimate (metric tonnes)') +
-  gfplot::theme_pbs()
+  gfplot::theme_pbs() +
+    ggtitle(paste0("", species, " "))
 ```
@@
   ind_imm <- read.csv(paste0("data/_indices/sopo-index-imm-", 
-    spp,"-", ssid_string, "-reml.csv"))# %>% filter(year > 2004)
+    spp,"-", ssid_string, "-", covs, "-reml.csv")) %>% filter(year > 2004)
 })
@@
 imm_col <- "orangered"
-ind %>% #filter(year > 2004) %>%
+
+ind %>% filter(year > 2004) %>%
 ggplot(aes(year, est*scale)) + geom_line(col = ad_col, ) +

### 2020-02-21 analysis/VOCC/3-vocc-dist.Rmd
experiment with a species-independent version of a distance-based vocc

---
title: "`r paste(params$species)` climatic and biotic velocities for `r paste(params$region)` (immature: `r paste(params$immature)`"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document 
params: 
    # species: "Redbanded Rockfish"
    # species: "Canary Rockfish"
    # species: "English Sole"
    # species: "Dover Sole"
    species: "Flathead Sole"
    # species: "Sablefish"
    # species: "Arrowtooth Flounder"
    immature: FALSE
    # region: "West Coast Vancouver Island"
    region: "both odd year surveys"
    # region: "West Coast Haida Gwaii"
    covs: "-tv-depth-only"
    scrambled_years: FALSE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.keep = "all", fig.show = "asis", dpi = 600,
  fig.width = 6, fig.height = 5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("gfranges")
library("dplyr")
library("ggplot2")

ggplot2::theme_set(gfplot::theme_pbs())
```

#### Set parameters
```{r }
species <- params$species
region <- params$region
covs <- params$covs
immature <- params$immature
scrambled_years <- params$scrambled_years

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

paste("species =", species)
paste("region =", region)
paste("covs =", covs)
paste("scrambled =", scrambled_years)
```

#### Set cell sizes and scale
```{r, echo=TRUE}
dist_intercept <- 0 # input_cell_size / 2
input_cell_size <- 2
scale_fac <- 2 # means that the raster is reprojected to _ X original grid (2 km)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_year <- 2008
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_year <- 2008
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_year <- 2008
}
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, include=FALSE}
if (immature) {
  age <- "immature"
  d_all <- readRDS(paste0(
    "data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-FALSE.rds"
  )) %>% filter(year >= start_year )
} else {
  age <- "mature"
  try({
    d_all <- readRDS(paste0(
      "data/", spp,
      "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-FALSE.rds"
    )) %>% filter(year >= start_year )
  })
  try({
    d_all <- readRDS(paste0(
      "data/", spp,
      "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds"
    )) %>% filter(year >= start_year )
  })
}
# hist(d_all$log_effort)
```

#### SET TIME FRAME
```{r}
multiyear <- TRUE

if (multiyear) {
  # unique(d_all$year)
  d <- d_all %>% filter(ssid %in% model_ssid)
  length(unique(d$year))
  years_sampled <- length(unique(d$year))
  year_range <- (years_sampled - 1) * 2

  # # alternate version of this model
  # if (years_sampled == 5) indices <- c(1, 2, 2, 2, 2)
  # if (years_sampled == 6) indices <- c(1, 2, 2, 2, 2, 2)

  # should try running with gradient based on only later years?
  if (years_sampled == 5) {
    indices <- c(1, 1, 1, 2, 2)
    #delta_t_total <- 5
  }
  if (years_sampled == 6) {
    indices <- c(1, 1, 1, 2, 2, 2)
    #delta_t_total <- 6
  }

} else {
  d <- d_all %>%
    filter(year >= start_year) %>%
    filter(year <= start_year + 3)
  unique(d$year)
  indices <- c(1, 2)
  year_range <- 2
}

skip_time <- NULL
delta_t_step <- year_range
delta_t_total <- year_range

glimpse(d)

if (scrambled_years) {
  d2 <- list()
  # set.seed(9999)
  # set.seed(12345)
  # set.seed(98765)
  original_years <- unique(d$year)
  new_years <- sample(original_years, replace = F)
  convert <- cbind(original_years, new_years)
  convert
  for (i in seq_len(length(original_years))) {
    old_year <- as.integer(convert[[i, 1]])
    fake_year <- as.integer(convert[[i, 2]])
    d2[[i]] <- filter(d, year == old_year) %>%
      mutate(year = fake_year)
    # browser()
  }
  d <- do.call(rbind, d2)
  convert
}
```

#### TRIM environment to range sampled
```{r}
nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
fishing <- readRDS("data/_fishing_effort/fishing-effort-grid.rds")
# fishing <- readRDS("data/fishing-effort-grid-6hr.rds")

d <- left_join(d, nd_all, by = c("X", "Y", "year"))
d <- left_join(d, fishing, by = c("X", "Y", "year"))
d$log_effort[is.na(d$log_effort)] <- -2.3

glimpse(d)

# d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
d <- d %>%
  filter(do_est > 0.28) %>%
  filter(do_est < 7.06) # 0.005 and 0.995
# d <- d %>% filter(temp > 2.61) %>% filter(temp < 14.31) # full range
d <- d %>%
  filter(temp > 3.07) %>%
  filter(temp < 11.3) # 0.005 and 0.995
```

Calculate mean and trend in log fishing pressure
```{r }
gf <- vocc_gradient_calc(d, "log_effort",
  scale_fac = scale_fac,
  # all layers with max indice will be averaged for spatial gradient
  # if default (NULL) will use all years
  # indices = indices,
  divisor = 1,
  quantile_cutoff = 0.05
)

gf <- gf %>%
  mutate(
    fishing_trend = trend,
    mean_effort = exp(mean)
  ) %>%
  dplyr::mutate(X = x, Y = y) %>%
  gfplot:::utm2ll(., utm_zone = 9)

gf$squashed_effort <- collapse_outliers(gf$mean_effort, c(0, 0.95))

# fished <- gf %>%
#   select(x, y, fishing_trend, mean_effort)
# saveRDS(fished, file = paste0("data/fishing-effort-trend-6hr-", ssid_string, ".rds"))
```

```{r eval = F}

effort <- plot_vocc(gf,
  # theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  # NA_label = "NA",
  fill_col = "squashed_effort",
  fill_label = "mean effort",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  # high_fill = "Steel Blue 4",
  # low_fill = "Red 3",
  # white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = fourth_root_power #
) + labs(
  title = "Fishing presssure",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

effort

fish_t <- plot_vocc(gf,
  # theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  # NA_label = "NA",
  fill_col = "fishing_trend",
  fill_label = "trend in fishing effort",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),

  # high_fill = "Steel Blue 4",
  # low_fill = "Red 3",
  # white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt # no_trans #
) + labs(
  title = "Fishing presssure",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

fish_t

```

# Calculate climate velocity
### Temperature
```{r}
gf2 <- vocc_gradient_calc(d, "temp",
  scale_fac = scale_fac,
  # all layers with max indice will be averaged for spatial gradient
  # if default (NULL) will use all years
  indices = indices,
  divisor = 1,
  quantile_cutoff = 0.05
)

gf2 <- gf2 %>%
  mutate(
    trend_per_year = trend,
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>%
  dplyr::mutate(X = x, Y = y)

# FULL DATASET
# mean(gf2$sd, na.rm = T)
# [1] 0.3434
# min(gf2$sd, na.rm = T)
# [1] 0.1266
# max(gf2$sd, na.rm = T)
# [1] 1.12
# median(gf2$sd, na.rm = T)
# [1] 0.287
# 
# 2008 DATASET
mean(gf2$sd, na.rm = T)
# [1] 0.3274
temp_min_sd <- min(gf2$sd, na.rm = T)
# [1] 0.1035
max(gf2$sd, na.rm = T)
# [1] 1.248
temp_median_sd <- median(gf2$sd, na.rm = T)
# [1] 0.2882

temp_sd_choice <- temp_min_sd

lower_change_t <- temp_sd_choice*2
upper_change_t <- temp_sd_choice*2

# variable_names <- c("temp")
# input_cell_size <- 2

## USED for truncating velocities?
max_dist_all_years <- 10 * year_range

dvocc_t <- make_vector_data(d,
  variable_names = c("temp"),
  ssid = model_ssid,
  # start_time = 2013,
  # end_time = 2015,
  # skip_time = skip_time,
  input_cell_size = 2,
  scale_fac = scale_fac,
  min_dist = 0,
  delta_t_total = years_sampled,
  delta_t_step = 2,
  indices = indices,
  min_thresholds = lower_change_t,
  max_thresholds = upper_change_t,
  round_fact = 10
)

dvocc_t <- dvocc_t %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend=slope)
dvocc_t <- dvocc_t %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
  (max_dist_all_years/ year_range)*10, dvocc))
dvocc_t <- mutate(dvocc_t, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
  -(max_dist_all_years / year_range)*10, dvocc))

dv_t <- dvocc_t %>% select(icell, x, y, dvocc) %>% unique()
glimpse(dv_t)

glimpse(gf2)

gf2$dvocc <- dv_t$dvocc

```

```{r eval=F}
gf2$squashed_velocity <- collapse_outliers(gf2$velocity, c(0.005, 0.995))

grad2v <- plot_vocc(gf2,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "squashed_velocity",
  fill_label = "temperature \nvelocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # fourth_root_power # sqrt #
) + labs(
  title = "gradient-based temp velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad2v
```

```{r eval=F}
grad2v <- plot_vocc(gf2,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "dvocc",
  fill_label = "temperature \ndvocc",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # fourth_root_power # sqrt #
) + labs(
  title = "distance-based temp velocity",
  subtitle = paste0(min(d$year), "-", max(d$year), " min temp SD = ", temp_min_sd, " x 2 for 95% min threshold of change")
)

grad2v
```

```{r eval=F}
grad2 <- plot_vocc(gf2,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "trend",
  fill_label = "temperature \ntrend",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # log10 # fourth_root_power # sqrt #
) + labs(
  title = "temp trend",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad2
```

### Dissolved Oxygen
```{r }
gf0 <- vocc_gradient_calc(d, "do_est",
  scale_fac = scale_fac,
  # all layers with max indice will be averaged for spatial gradient
  # if default (NULL) will use all years
  indices = indices,
  divisor = 1,
  quantile_cutoff = 0.05
)

gf0 <- gf0 %>%
  mutate(
    trend_per_year = trend,
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>%
  dplyr::mutate(X = x, Y = y) %>%
  gfplot:::utm2ll(., utm_zone = 9)

# 2008 DATASET
hist(gf0$sd)
mean(gf0$sd, na.rm = T)
# 0.2828
do_min_sd <- min(gf0$sd, na.rm = T)
# [1] 0.04122
max(gf0$sd, na.rm = T)
# [1] 1.531
do_median_sd <- median(gf0$sd, na.rm = T)
# [1] 0.2343

do_sd_choice <- do_median_sd

lower_change_t <- c(Inf, do_sd_choice*2) 
upper_change_t <- c(temp_sd_choice*2, do_sd_choice*2)

# variable_names <- c("temp")
# input_cell_size <- 2

## USED for truncating velocities?
max_dist_all_years <- 10 * year_range

dvocc_d <- make_vector_data(list(d, d),
  variable_names = c("temp", "do_est"),
  ssid = model_ssid,
  # start_time = 2013,
  # end_time = 2015,
  # skip_time = skip_time,
  input_cell_size = 2,
  scale_fac = scale_fac,
  min_dist = 0,
  delta_t_total = years_sampled,
  delta_t_step = 2,
  indices = indices,
  min_thresholds = lower_change_t,
  max_thresholds = upper_change_t,
  round_fact = 10
)

# glimpse(dvocc_d)
dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(do_est.slope < 0, -km_per_decade, km_per_decade), trend= do_est.slope)

dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
  (max_dist_all_years/ year_range)*10, dvocc))
dvocc_d <- mutate(dvocc_d, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
  -(max_dist_all_years / year_range)*10, dvocc))

dv_d <- dvocc_d %>% select(icell, x, y, dvocc) %>% unique()
# glimpse(dv_b)
# glimpse(gf0)

gf0$dvocc <- dv_d$dvocc
```

```{r eval=F}
gf0$squashed_velocity <- collapse_outliers(gf0$velocity, c(0.005, 0.995))

grad0v <- plot_vocc(gf0,
  # theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "squashed_velocity",
  fill_label = "DO velocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # fourth_root_power
) + labs(
  title = "do velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad0v
```

```{r eval=F}
grad0 <- plot_vocc(gf0,
  # theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "trend",
  fill_label = "DO trend",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans #
) + labs(
  title = "do trend",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad0
```

```{r eval=F}
grad0dv <- plot_vocc(gf0,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "dvocc",
  fill_label = "DO and temp \ndvocc",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # fourth_root_power # sqrt #
) + labs(
  title = "distance-based DO and temp velocity",
  subtitle = paste0(min(d$year), "-", max(d$year), " DO SD = ", signif(do_sd_choice, 2), " x 2 for threshold for do change")
)

grad0dv
```

### Gradient-based BIOTIC VECTORS
```{r}
gf3 <- vocc_gradient_calc(d, "est",
  scale_fac = scale_fac,
  indices = indices,
  divisor = 1,
  log_space = TRUE,
  quantile_cutoff = 0.025
)

gf3 <- gf3 %>%
  mutate(
    CV_formula = sqrt(exp(cv^2) - 1),
    CV = sd / exp(mean),
    mean_biomass = exp(mean) * 10000,
    # trend_per_year = exp(trend),
    # trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>%
  dplyr::mutate(X = x, Y = y)

hist((gf3$trend))
hist((gf3$sd))

b_mean_sd <-mean(gf3$sd, na.rm = T)
# 0.2828
b_min_sd <- min(gf3$sd, na.rm = T)
# [1] 0.04122
max(gf3$sd, na.rm = T)
# [1] 1.531
b_median_sd <- median(gf3$sd, na.rm = T)
# [1] 0.2343

b_choice <- 0.1 # in log space roughly equal 10% change

lower_change_t <- c(b_choice) 
upper_change_t <- c(b_choice)

# variable_names <- c("temp")
# input_cell_size <- 2

## USED for truncating velocities?
max_dist_all_years <- 10 * year_range

dvocc_b <- make_vector_data(d,
  variable_names = c("est"),
  ssid = model_ssid,
  # start_time = 2013,
  # end_time = 2015,
  # skip_time = skip_time,
  input_cell_size = 2,
  scale_fac = scale_fac,
  min_dist = 0,
  delta_t_total = years_sampled,
  delta_t_step = 2,
  indices = indices,
  min_thresholds = lower_change_t,
  max_thresholds = upper_change_t,
  round_fact = 10
)
glimpse(dvocc_b)

dvocc_b <- dvocc_b %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend=slope)
dvocc_b <- dvocc_b %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
  (max_dist_all_years/ year_range)*10, dvocc))
dvocc_b <- mutate(dvocc_b, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
  -(max_dist_all_years / year_range)*10, dvocc))

dv_b <- dvocc_b %>% select(icell, x, y, dvocc) %>% unique()
# glimpse(dv_b)
# glimpse(gf3)

gf3$dvocc <- dv_b$dvocc

bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.05
s <- sort(gf3$mean_biomass)

bio_sum <- cumsum(s)
lower_density_threshold <- s[which(bio_sum >= bio5perc)[1]]

gf3t <- filter(gf3, mean_biomass > lower_density_threshold)
```

```{r eval=F}
gradbv <- plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "dvocc",
  fill_label = "biotic \ndvocc",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # fourth_root_power # sqrt #
) + labs(
  title = "distance-based biotic velocity",
  subtitle = paste0(min(d$year), "-", max(d$year), " threshold for change = ", b_choice, "")
)

gradbv
```

```{r eval=T}
gf3t$squashed_velocity <- collapse_outliers(gf3t$velocity, c(0.005, 0.995))

grad3 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "squashed_velocity",
  fill_label = "biotic \nvelocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt # no_trans #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biotic velocity ", min(d$year), "-", max(d$year))
)
grad3
```

```{r eval=T}
# gf3t$squashed_velocity <- collapse_outliers(gf3t$velocity, c(0.005, 0.995))

grad3 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "gradient",
  fill_label = "biotic \ngradient",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt # no_trans #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biotic gradient for final year")
)
grad3
```

```{r eval=T}
#raster_max <- quantile(gf3$mean_biomass, 0.995, na.rm = T)
gf3$mean_biomass2 <- collapse_outliers(gf3$mean_biomass, c(0, 0.995))
grad4 <- gfranges::plot_vocc(gf3,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "mean_biomass2",
  fill_label = "mean",
  raster_alpha = 1,
 # raster_limits = c(0, raster_max),
  na_colour = "white",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Mean biomass ", min(d$year), "-", max(d$year))
)
grad4
```

```{r eval=T}
grad5 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "trend",
  fill_label = "biomass \ntrend",
  raster_alpha = 1,
  # raster_limits = c(0, 10),
  na_colour = "black",
  white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biomass trend ", min(d$year), "-", max(d$year))
)
grad5
```

```{r eval=T}
grad6 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "cv",
  fill_label = "sd in log space",
  raster_alpha = 1,
  # raster_limits = c(0, 7),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biotic variability (cv from vocc_gradient_calc funcition) ", min(d$year), "-", max(d$year))
)
grad6
```

```{r eval=F}
raster_max <- quantile(gf3t$CV, 0.90, na.rm = TRUE)
grad7 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "CV",
  fill_label = "CV",
  raster_alpha = 1,
  raster_limits = c(0, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("CV in estimated biomass ", min(d$year), "-", max(d$year))
)
grad7
```

### Save gradient-based VOCCs
```{r}

biotic_values <- gf3 %>%
  rename(
    biotic_vel = velocity,
    biotic_dvocc = dvocc,
    biotic_trend = trend,
    biotic_CV = CV,
    sd_est = cv,
    biotic_grad = gradient
  ) %>%
  mutate(
    sd_biomass = sd * 10000,
    mean_biomass = exp(mean) * 10000
  ) %>%
  select(x, y, biotic_vel, biotic_dvocc, biotic_trend, biotic_grad, biotic_CV, sd_est, sd_biomass, mean_biomass) %>%
  mutate(start_year = !!start_year)

fished <- gf %>%
  select(x, y, fishing_trend, mean_effort)

DO_values <- gf0 %>%
  rename(
    DO_vel = velocity,
    DO_dvocc = dvocc,
    DO_trend = trend_per_year,
    DO_grad = gradient,
    mean_DO = mean
  ) %>%
  select(x, y, DO_vel, DO_dvocc, DO_trend, DO_grad, mean_DO)

temp_values <- gf2 %>%
  rename(
    temp_vel = velocity,
    temp_dvocc = dvocc,
    temp_trend = trend_per_year,
    temp_grad = gradient,
    mean_temp = mean
  ) %>%
  select(x, y, temp_vel, temp_dvocc, temp_trend, temp_grad, mean_temp)

vocc <- left_join(biotic_values, fished) # add fishing effort
vocc <- left_join(vocc, DO_values) # add DO
vocc <- left_join(vocc, temp_values) # add temp
vocc <- mutate(vocc, species = !!species, method = "both")

write.csv(vocc, file = paste0(
  "data/_dvocc/", age, "-w-do-nulls-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
))
```

```{r eval=T}
vocc_t <- filter(vocc, mean_biomass > lower_density_threshold)
vocc_t$squashed_temp <- collapse_outliers(vocc_t$temp_vel, c(0.005, 0.995))

grad2t <- plot_vocc(vocc_t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "squashed_temp",
  fill_label = "temperature \nvelocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = "temp velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad2t
```

```{r eval=T}
grad2tt <- plot_vocc(vocc_t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "temp_trend",
  fill_label = "temperature \ntrend",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # log10 # fourth_root_power # sqrt #
) + labs(
  title = "temp trend",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad2tt
```

```{r eval=T}
vocc_t$squashed_do <- collapse_outliers(vocc_t$DO_vel, c(0.005, 0.995))

# grad2dv <- 
plot_vocc(vocc_t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "squashed_do",
  fill_label = "DO \nvelocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = "DO velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

# grad2dv
```

```{r eval=T}
# grad2dt <- 
plot_vocc(vocc_t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "DO_trend",
  fill_label = "DO \ntrend",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = no_trans # log10 # fourth_root_power # sqrt #
) + labs(
  title = "DO trend",
  subtitle = paste0(min(d$year), "-", max(d$year))
)
# 
# grad2dt
```

```{r eval=T}
vocc_t$squashed_effort <- collapse_outliers(vocc_t$mean_effort, c(0.005, 0.995))

plot_vocc(vocc_t,
  # theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  # NA_label = "NA",
  fill_col = "squashed_effort",
  fill_label = "mean effort",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  # high_fill = "Steel Blue 4",
  # low_fill = "Red 3",
  # white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = fourth_root_power #
) + labs(
  title = "Fishing presssure",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

plot_vocc(vocc_t,
  # theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  # grad_vec_aes = "velocity",
  # vec_lwd_range = c(0.5, 0.5),
  # NA_label = "NA",
  fill_col = "fishing_trend",
  fill_label = "trend",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  # high_fill = "Steel Blue 4",
  # low_fill = "Red 3",
  # white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt # no_trans #
) + labs(
  title = "Fishing trend",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

```

### 2020-02-21 analysis/SOPO/2-trawl-no-covs-model.Rmd
correct how maps appear in sopo renders

@@
     # species: "Pacific Ocean Perch"
-    # species: "Pacific Sand Lance"
-    # species: "Pacific Cod"
-    # species: "Buffalo Sculpin"
-    # species: "Cabezon"
-    species: "Harlequin Rockfish"
+    species: "Pacific Sand Lance"
+    # species: "Pacific Herring"
     # region: "All synoptic surveys"
@@
     fixed_spatial: FALSE
-    update_model: FALSE #TRUE
+    update_model: FALSE
     update_model_check: FALSE #TRUE
     update_predictions: FALSE #TRUE
-    update_index: TRUE
+    update_index: FALSE
 ---
@@
       if (prop_pos < 0.1) {
-        update_model <- T
-        update_model_check <- T
-        update_predictions <- T
-        update_index <- T
+        # update_model <- T
+        # update_model_check <- T
+        # update_predictions <- T
+        # update_index <- T
         knots <- 150
@@
     if (prop_pos < 0.05) {
-        update_model <- F
-        update_model_check <- F
-        update_predictions <- F
-        update_index <- T
+        # update_model <- F
+        # update_model_check <- F
+        # update_predictions <- F
+        # update_index <- T
         knots <- 150
@@
 
- p_adult_all <- adult_predictions %>% filter(year>2004) %>%
+ p_adult_all <- adult_predictions %>% filter(!((year<2005) & (ssid == 3)))%>%
    mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>%  
@@
 try({
- p_imm_all <- imm_predictions %>% filter(year>2004) %>%
+ p_imm_all <- imm_predictions %>% filter(!((year<2005) & (ssid == 3)))%>%
     mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>%
@@
 
-p_adult_all <- predictions  %>% 
+p_adult_all <- predictions  %>% filter(!((year<2005) & (ssid == 3))) %>% #filter(year>2003) %>%
   mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>% 

### 2020-02-21 analysis/VOCC/3-vocc-dist.Rmd
change mid_fill to pale yellow so 0 dist visible

@@
 
-### Gradient-based BIOTIC VECTORS
+### BIOTIC VECTORS
 ```{r}
@@
   fill_col = "dvocc",
-  fill_label = "biotic \ndvocc",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
+  fill_label = "km/decade",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "black",
+  # white_zero = TRUE,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
   axis_lables = FALSE,
@@
 
-### Save gradient-based VOCCs
+### Save VOCCs
 ```{r}
@@
   na_colour = "white",
+  mid_fill = "lightyellow",
   white_zero = TRUE,
@@
   title = "temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+
+grad2t
+```
+
+```{r eval=T}
+grad2t <- plot_vocc(vocc_t,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "temp_dvocc",
+  fill_label = "temperature \nkm/decade",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "white",
+   # viridis_option = "C",
+   mid_fill = "lightyellow",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = sqrt #
+) + labs(
+  title = "temp distance-based velocity",
   subtitle = paste0(min(d$year), "-", max(d$year))
@@
   # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
+  na_colour = "grey",
+  white_zero = TRUE,
+   mid_fill = "lightyellow",
   vec_alpha = 0.35,

### 2020-02-24 analysis/VOCC/3-vocc-dist.Rmd
add analog-distance velocities to dataset for meta analysis

--- 2020-02-21 analysis/VOCC/3-vocc-dist.Rmd
@@
 params: 
-    # species: "Redbanded Rockfish"
+    species: "Redbanded Rockfish"
     # species: "Canary Rockfish"
@@
     # species: "Dover Sole"
-    species: "Flathead Sole"
+    # species: "Flathead Sole"
     # species: "Sablefish"
@@
     immature: FALSE
-    # region: "West Coast Vancouver Island"
-    region: "both odd year surveys"
+    region: "West Coast Vancouver Island"
+    # region: "both odd year surveys"
     # region: "West Coast Haida Gwaii"
@@
 
+  ## USED for truncating distance-based velocities?
+  max_dist_all_years <- 10 * year_range
+
   # should try running with gradient based on only later years?
-  if (years_sampled == 5) {
-    indices <- c(1, 1, 1, 2, 2)
-    #delta_t_total <- 5
-  }
-  if (years_sampled == 6) {
-    indices <- c(1, 1, 1, 2, 2, 2)
-    #delta_t_total <- 6
-  }
+  if (years_sampled == 5) { indices <- c(1, 1, 1, 2, 2)  }
+  if (years_sampled == 6) { indices <- c(1, 1, 1, 1, 2, 2) }
 
@@
   # indices = indices,
-  divisor = 1,
+  # divisor = 1, #default of 10 gives per decade
   quantile_cutoff = 0.05
@@
 ```{r eval = F}
-
 effort <- plot_vocc(gf,
@@
   # raster_limits = c(raster_min, raster_max),
-
   # high_fill = "Steel Blue 4",
@@
 fish_t
-
 ```
@@
   indices = indices,
-  divisor = 1,
+  # divisor = 1, #default of 10 gives per decade
   quantile_cutoff = 0.05
@@
 # FULL DATASET
-# mean(gf2$sd, na.rm = T)
-# [1] 0.3434
-# min(gf2$sd, na.rm = T)
-# [1] 0.1266
-# max(gf2$sd, na.rm = T)
-# [1] 1.12
-# median(gf2$sd, na.rm = T)
-# [1] 0.287
+# mean(gf2$sd, na.rm = T) # [1] 0.3434
+# min(gf2$sd, na.rm = T) # [1] 0.1266
+# max(gf2$sd, na.rm = T) # [1] 1.12
+# median(gf2$sd, na.rm = T) # [1] 0.287
 # 
 # 2008 DATASET
-mean(gf2$sd, na.rm = T)
-# [1] 0.3274
-temp_min_sd <- min(gf2$sd, na.rm = T)
-# [1] 0.1035
-max(gf2$sd, na.rm = T)
-# [1] 1.248
-temp_median_sd <- median(gf2$sd, na.rm = T)
-# [1] 0.2882
+mean(gf2$sd, na.rm = T) # [1] 0.3274
+temp_min_sd <- min(gf2$sd, na.rm = T) # [1] 0.1035
+max(gf2$sd, na.rm = T) # [1] 1.248
+temp_median_sd <- median(gf2$sd, na.rm = T) # [1] 0.2882
 
 temp_sd_choice <- temp_min_sd
-
 lower_change_t <- temp_sd_choice*2
 upper_change_t <- temp_sd_choice*2
-
-# variable_names <- c("temp")
-# input_cell_size <- 2
-
-## USED for truncating velocities?
-max_dist_all_years <- 10 * year_range
 
@@
   ssid = model_ssid,
-  # start_time = 2013,
-  # end_time = 2015,
-  # skip_time = skip_time,
   input_cell_size = 2,
@@
 dv_t <- dvocc_t %>% select(icell, x, y, dvocc) %>% unique()
-glimpse(dv_t)
-
-glimpse(gf2)
+
+# glimpse(dv_t)
+# glimpse(gf2)
 
 gf2$dvocc <- dv_t$dvocc
-
+```
+
+```{r eval=F}
+plot_vocc(gf2,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "dvocc",
+  fill_label = "temperature \ndvocc",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "white",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = no_trans # fourth_root_power # sqrt #
+) + labs(
+  title = "distance-based temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year), " min temp SD = ", signif(temp_min_sd, 2), " x 2 for 95% min threshold of change")
+)
 ```
@@
 
-grad2v <- plot_vocc(gf2,
+plot_vocc(gf2,
   # theme = "black",
@@
 )
-
-grad2v
-```
-
-```{r eval=F}
-grad2v <- plot_vocc(gf2,
-  # theme = "black",
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "dvocc",
-  fill_label = "temperature \ndvocc",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans # fourth_root_power # sqrt #
-) + labs(
-  title = "distance-based temp velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year), " min temp SD = ", temp_min_sd, " x 2 for 95% min threshold of change")
-)
-
-grad2v
 ```
@@
 # 2008 DATASET
-hist(gf0$sd)
-mean(gf0$sd, na.rm = T)
+# hist(gf0$sd)
+# mean(gf0$sd, na.rm = T)
 # 0.2828
@@
 ```{r eval=F}
+grad0dv <- plot_vocc(gf0,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "dvocc",
+  fill_label = "DO and temp \ndvocc",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "white",
+  white_zero = TRUE,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = no_trans # fourth_root_power # sqrt #
+) + labs(
+  title = "distance-based DO and temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year), " DO SD = ", signif(do_sd_choice, 2), " x 2 for threshold for do change")
+)
+
+grad0dv
+```
+
+```{r eval=F}
 gf0$squashed_velocity <- collapse_outliers(gf0$velocity, c(0.005, 0.995))
@@
 
-```{r eval=F}
-grad0dv <- plot_vocc(gf0,
-  # theme = "black",
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "dvocc",
-  fill_label = "DO and temp \ndvocc",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans # fourth_root_power # sqrt #
-) + labs(
-  title = "distance-based DO and temp velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year), " DO SD = ", signif(do_sd_choice, 2), " x 2 for threshold for do change")
-)
-
-grad0dv
-```
-
-### Gradient-based BIOTIC VECTORS
+### BIOTIC VECTORS
 ```{r}
@@
   indices = indices,
-  divisor = 1,
+  # divisor = 1, #default of 10 gives per decade
   log_space = TRUE,
@@
 
-hist((gf3$trend))
-hist((gf3$sd))
+# hist((gf3$trend))
+# hist((gf3$sd))
 
@@
 )
-glimpse(dvocc_b)
-
-dvocc_b <- dvocc_b %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend=slope)
-dvocc_b <- dvocc_b %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
+# glimpse(dvocc_b)
+
+dvocc_b <- dvocc_b %>% 
+  mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend=slope)
+dvocc_b <- dvocc_b %>% 
+  mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
   (max_dist_all_years/ year_range)*10, dvocc))
-dvocc_b <- mutate(dvocc_b, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
+dvocc_b <-  dvocc_b %>% 
+  mutate(dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
   -(max_dist_all_years / year_range)*10, dvocc))
@@
 
-```{r eval=F}
+```{r eval=T}
 gradbv <- plot_vocc(gf3t,
@@
   fill_col = "dvocc",
-  fill_label = "biotic \ndvocc",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
+  fill_label = "km/decade",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "black",
+  # white_zero = TRUE,
+  mid_fill = "lightyellow",
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
   axis_lables = FALSE,
@@
 ) + labs(
-  title = "distance-based biotic velocity",
+  title = paste0(species, " (", age, ") distance-based biotic velocity"),
   subtitle = paste0(min(d$year), "-", max(d$year), " threshold for change = ", b_choice, "")
@@
   fill_label = "biotic \nvelocity",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "black",
+  # white_zero = TRUE,
+   mid_fill = "lightyellow",
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+  axis_lables = FALSE,
+  transform_col = sqrt # no_trans #
+) + labs(
+  title = paste0(species, " (", age, ")"), #
+  subtitle = paste0("Biotic velocity ", min(d$year), "-", max(d$year))
+)
+grad3
+```
+
+```{r eval=T}
+# gf3t$squashed_velocity <- collapse_outliers(gf3t$velocity, c(0.005, 0.995))
+
+grad3 <- gfranges::plot_vocc(gf3t,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "gradient",
+  fill_label = "biotic \ngradient",
   raster_alpha = 1,
@@
   title = paste0(species, " (", age, ")"), #
-  subtitle = paste0("Biotic velocity ", min(d$year), "-", max(d$year))
+  subtitle = paste0("Biotic gradient for final year")
 )
@@
 ```{r eval=T}
-# gf3t$squashed_velocity <- collapse_outliers(gf3t$velocity, c(0.005, 0.995))
-
-grad3 <- gfranges::plot_vocc(gf3t,
-  # theme = "black",
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "gradient",
-  fill_label = "biotic \ngradient",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "black",
+#raster_max <- quantile(gf3$mean_biomass, 0.995, na.rm = T)
+gf3$mean_biomass2 <- collapse_outliers(gf3$mean_biomass, c(0, 0.995))
+grad4 <- gfranges::plot_vocc(gf3,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "mean_biomass2",
+  fill_label = "mean",
+  raster_alpha = 1,
+ # raster_limits = c(0, raster_max),
+  na_colour = "white",
   # white_zero = TRUE,
@@
   axis_lables = FALSE,
-  transform_col = sqrt # no_trans #
-) + labs(
-  title = paste0(species, " (", age, ")"), #
-  subtitle = paste0("Biotic gradient for final year")
-)
-grad3
-```
-
-```{r eval=T}
-#raster_max <- quantile(gf3$mean_biomass, 0.995, na.rm = T)
-gf3$mean_biomass2 <- collapse_outliers(gf3$mean_biomass, c(0, 0.995))
-grad4 <- gfranges::plot_vocc(gf3,
-  # theme = "black",
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "mean_biomass2",
-  fill_label = "mean",
-  raster_alpha = 1,
- # raster_limits = c(0, raster_max),
-  na_colour = "white",
-  # white_zero = TRUE,
-  high_fill = "Steel Blue 4",
-  low_fill = "Red 3",
-  axis_lables = FALSE,
   transform_col = sqrt #
@@
   white_zero = TRUE,
+   mid_fill = "lightyellow",
   high_fill = "Steel Blue 4",
@@
   title = paste0(species, " (", age, ")"), #
-  subtitle = paste0("Biotic variability (cv from vocc_gradient_calc funcition) ", min(d$year), "-", max(d$year))
+  subtitle = paste0("Biotic variability (cv from vocc_gradient_calc) ", min(d$year), "-", max(d$year))
 )
@@
 ))
-```
-
-```{r eval=T}
+
+```
+
+```{r}
 vocc_t <- filter(vocc, mean_biomass > lower_density_threshold)
+```
+
+```{r eval=T}
+plot_vocc(vocc_t,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "temp_dvocc",
+  fill_label = "temperature \nkm/decade",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "white",
+   # viridis_option = "C",
+   mid_fill = "lightyellow",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = sqrt #
+) + labs(
+    title = "distance-based temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year), " min temp SD = ", temp_min_sd, " x 2 for 95% min threshold of change")
+)
+```
+
+```{r eval=T}
 vocc_t$squashed_temp <- collapse_outliers(vocc_t$temp_vel, c(0.005, 0.995))
 
-grad2t <- plot_vocc(vocc_t,
+plot_vocc(vocc_t,
   # theme = "black",
@@
 ) + labs(
-  title = "temp velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-grad2t
-```
-
-```{r eval=T}
-grad2tt <- plot_vocc(vocc_t,
+  title = "gradient-based temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+```
+
+```{r eval=T}
+plot_vocc(vocc_t,
   # theme = "black",
@@
 )
-
-grad2tt
+```
+
+```{r eval=T}
+plot_vocc(vocc_t,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "DO_dvocc",
+  fill_label = "km/decade",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "white",
+   # viridis_option = "C",
+   mid_fill = "lightyellow",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = sqrt #
+) + labs(
+ title = "distance-based DO and temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year), " DO SD = ", signif(do_sd_choice, 2), " x 2 for threshold for do change")
+)
 ```
@@
 
-# grad2dv <- 
 plot_vocc(vocc_t,
@@
 ) + labs(
-  title = "DO velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-# grad2dv
-```
-
-```{r eval=T}
-# grad2dt <- 
+  title = "DO gradient-based velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+```
+
+```{r eval=T}
 plot_vocc(vocc_t,
@@
   white_zero = TRUE,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+   mid_fill = "lightyellow",
   vec_alpha = 0.35,
@@
 )
-# 
-# grad2dt
 ```
@@
   # white_zero = TRUE,
+   mid_fill = "lightyellow",
   vec_alpha = 0.35,

### 2020-02-24 analysis/VOCC/3-render-biotic-vocc.R

@@ -1,3 +1,4 @@
+.rs.restartR()
 getwd()
 setwd(here::here("/analysis/VOCC"))
 env <- new.env() # parent = baseenv()
@@ -65,7 +66,7 @@ for (r_h in seq_along(list_regions)) {
     covs <- "-tv-depth-only" # string describing model
     reg <- gsub(" ", "-", gsub("\\/", "-", tolower(list_regions[r_h])))
     try({
-      rmarkdown::render("3-vocc-w-do.Rmd",
+      rmarkdown::render("3-vocc-dist.Rmd",
         params = list(
           species = list_species[spp_i],
           # immature = TRUE,
@@ -75,7 +76,7 @@ for (r_h in seq_along(list_regions)) {
         output_file = paste0(
           "html/VOCC-plots/vocc-w-do-", spp,
           # "-imm",
-          covs, "-", reg, "-untrimmed.html"
+          covs, "-", reg, "-dvocc.html"
         ),
         envir = env
       )

### 2020-02-24 analysis/VOCC/4-pool-vocc.R

@@ -5,6 +5,7 @@ library("dplyr")
 # covs <- "-fixed"
 mydir = paste0( "_new_null")
 mydir = paste0( "_do_null")
+mydir = paste0( "_dvocc")
 
 ####################
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
@@ -20,4 +21,8 @@ saveRDS(biotic, file = paste0(
 
 saveRDS(biotic, file = paste0(
   "mature-all-do-untrimmed.rds"
+))
+
+saveRDS(biotic, file = paste0(
+  "mature-all-do-dvocc.rds"
 ))

### 2020-02-24 analysis/VOCC/7-vocc-chopsticks.R

@@ -22,9 +22,14 @@ data_type <- "mature-90-all-temp"
 model <- readRDS("data/trend_01-25-mature-90-all-temp-all-main-sim-2.rds")
 data_type <- "mature-90-all-temp-sim"
 
-data_type <- "mature-90-vel-sim"
 
-data_type <- "mature-95-all-temp"
+model <- readRDS("data/trend-mature-95-all-do-02-20-trend-with-do-sim-1.rds")
+data_type <- "mature-95-with-do-sim"
+
+
+model <- readRDS("data/trend-mature-95-all-do-02-20-trend-with-do-1.rds")
+data_type <- "mature-95-with-do"
+
 
 #### INTERACTIONS WIHT TEMP TREND
 title_all <- "Interation plots "
@@ -37,17 +42,26 @@ plot_fuzzy_chopsticks(model,
   #facet_wrap(vars(species), scales="free_y") +
   ggtitle(paste(title_all, "(", data_type, ")"))
 
-plot_fuzzy_chopsticks(model,
-  x_variable = "temp_trend_scaled", type = "biomass",
-  y_label = y_label
-) + # facet_wrap(vars(species), scales="free_y") +
-  ggtitle(paste(title_all, "(", data_type, ")"))
+# plot_fuzzy_chopsticks(model,
+#   x_variable = "temp_trend_scaled", type = "biomass",
+#   y_label = y_label
+# ) + # facet_wrap(vars(species), scales="free_y") +
+#   ggtitle(paste(title_all, "(", data_type, ")"))
+
+# plot_fuzzy_chopsticks(model,
+#   x_variable = "temp_trend_scaled", type = "temp_grad",
+#   y_label = y_label
+# ) + facet_wrap(vars(species), scales = "free_y") +
+#   ggtitle(paste(title_all, "(", data_type, ")"))
 
 plot_fuzzy_chopsticks(model,
-  x_variable = "temp_trend_scaled", type = "temp_grad",
+  x_variable = "DO_trend_scaled", type = "do",
   y_label = y_label
-) + facet_wrap(vars(species), scales = "free_y") +
+) + #ylim(-1.5, 1) + # 
+  #facet_wrap(vars(species), scales="free_y") +
   ggtitle(paste(title_all, "(", data_type, ")"))
+
+
 
 
 title_all <- "Interation plots "
@@ -59,11 +73,38 @@ plot_fuzzy_chopsticks(model,
[…]

### 2020-02-24 analysis/VOCC/3-vocc-dist.Rmd
add DO by its self to dvocc code, but haven't run it yet (only needed if dvocc stays in manuscript)

@@
 
-lower_change_t <- c(Inf, do_sd_choice*2) 
-upper_change_t <- c(temp_sd_choice*2, do_sd_choice*2)
+lower_change_d <- c(do_sd_choice*2) 
+upper_change_d <- c(do_sd_choice*2)
 
@@
 
-dvocc_d <- make_vector_data(list(d, d),
+dvocc_d <- make_vector_data(d,
+  variable_names = c("do_est"),
+  ssid = model_ssid,
+  # start_time = 2013,
+  # end_time = 2015,
+  # skip_time = skip_time,
+  input_cell_size = 2,
+  scale_fac = scale_fac,
+  min_dist = 0,
+  delta_t_total = years_sampled,
+  delta_t_step = 2,
+  indices = indices,
+  min_thresholds = lower_change_d,
+  max_thresholds = upper_change_d,
+  round_fact = 10
+)
+
+# glimpse(dvocc_d)
+dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(do_est.slope < 0, -km_per_decade, km_per_decade), trend= do_est.slope)
+
+dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
+  (max_dist_all_years/ year_range)*10, dvocc))
+dvocc_d <- mutate(dvocc_d, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
+  -(max_dist_all_years / year_range)*10, dvocc))
+
+dv_d <- dvocc_d %>% select(icell, x, y, dvocc) %>% unique()
+# glimpse(dv_b)
+# glimpse(gf0)
+
+gf0$dvocc <- dv_d$dvocc
+
+lower_change_td <- c(Inf, do_sd_choice*2) 
+upper_change_td <- c(temp_sd_choice*2, do_sd_choice*2)
+
+# variable_names <- c("temp")
+# input_cell_size <- 2
+
+## USED for truncating velocities?
+max_dist_all_years <- 10 * year_range
+
+dvocc_td <- make_vector_data(list(d, d),
   variable_names = c("temp", "do_est"),
@@
   indices = indices,
-  min_thresholds = lower_change_t,
-  max_thresholds = upper_change_t,
+  min_thresholds = lower_change_td,
+  max_thresholds = upper_change_td,
   round_fact = 10
@@
 # glimpse(dvocc_d)
-dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(do_est.slope < 0, -km_per_decade, km_per_decade), trend= do_est.slope)
-
-dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
+dvocc_td <- dvocc_td %>% mutate(dvocc = if_else(do_est.slope < 0, -km_per_decade, km_per_decade), trend= do_est.slope)
+
+dvocc_td <- dvocc_td %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
   (max_dist_all_years/ year_range)*10, dvocc))
-dvocc_d <- mutate(dvocc_d, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
+dvocc_td <- mutate(dvocc_td, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
   -(max_dist_all_years / year_range)*10, dvocc))
 
-dv_d <- dvocc_d %>% select(icell, x, y, dvocc) %>% unique()
+dv_td <- dvocc_td %>% select(icell, x, y, dvocc) %>% unique()
 # glimpse(dv_b)
@@
 
-gf0$dvocc <- dv_d$dvocc
+gf0$dvocc_both <- dv_td$dvocc
 ```
@@
   ) %>%
-  select(x, y, DO_vel, DO_dvocc, DO_trend, DO_grad, mean_DO)
+  select(x, y, DO_vel, DO_dvocc, DO_trend, DO_grad, mean_DO, dvocc_both)
 

### 2020-02-24 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
misc inconsequential

@@
 
-```{r}
+```{r eval = FALSE}
 library(gbm)
@@
 ## Just compare fixed slope models with simplest random effect structure
-```{r}
+```{r eval=F}
 cv_args <- generate_cv_args(
@@
     #temperature_c ~ 0 + SST_scaled*shallow,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
-    #temperature_c ~ 1 + SST_scaled,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled,
+    # temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
+    # #temperature_c ~ 1 + SST_scaled,
+    # temperature_c ~ 0 + as.factor(year) + SST_scaled,
     temperature_c ~ 0 + as.factor(year)
@@
   time_varying = list(
+    # ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
     ~ 0 + depth_scaled + depth_scaled2,
-    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
   ),
-  ar1_fields = c(FALSE),
-  include_spatial = c(FALSE),
-  n_knots = c(100), # keep same n_knots
+  ar1_fields = c(TRUE),
+  include_spatial = c(TRUE),
+  n_knots = c(400), # keep same n_knots
   seed = c(1111, 2222, 3333, 4444) # keep same seeds

### 2020-02-24 analysis/VOCC/7-coef-plots.R

@@ -25,7 +25,6 @@ stats$group[stats$group == "HAKE"] <- "COD"
 #### ONE JUST BUILT
 model <- new_model
 
-
 nrow(model$data)
 
 model2 <- add_colours(model$coefs) %>%
@@ -67,6 +66,7 @@ manipulate::manipulate({
 #    "mean_temp_scaled:mean_DO_scaled", "mean_DO_scaled:mean_temp_scaled"))
 # manipulate::manipulate({plot_coefs(model2b, order_by = order_by)},
 #   order_by = manipulate::picker(as.list(sort(unique(shortener(model2b$coefficient))))))
+
 ### SAVE PLOT WITH SELECTED PARAMS
 # model3 <- plot_coefs(model2, order_by = "squashed_temp_vel_scaled)")
 # model_plot <- model3 +

### 2020-03-09 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
plot new facet map of all temp estimates

@@
 p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised")
+# p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted_temp_allyears_revised.rds")
 
@@
 plot_map_raster(p_nd, "est") +
-  scale_fill_viridis_c(trans = "log") +
+  scale_fill_viridis_c(trans = sqrt) +
   facet_wrap(~year) +
   ggtitle("Prediction (fixed effects + all random effects)")
+
+test <- plot_facet_map(p_nd, "est", viridis_option = "D",
+    raster_limits = c(min(p_nd$est),quantile(p_nd$est, 0.99)),
+    transform_col = sqrt
+  )
 ```

### 2020-03-18 analysis/VOCC/3-vocc-dist.Rmd
tweak for immature analysis

@@
     # species: "Arrowtooth Flounder"
-    immature: FALSE
+    immature: T
     region: "West Coast Vancouver Island"
@@
 # glimpse(dvocc_d)
-dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(do_est.slope < 0, -km_per_decade, km_per_decade), trend= do_est.slope)
+dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend= slope)
 
@@
 write.csv(vocc, file = paste0(
-  "data/_dvocc/", age, "-w-do-nulls-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_dvocc_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))

### 2020-03-18 analysis/VOCC/5-add-null.R

@@ -5,14 +5,19 @@ library(sdmTMB)
 library(gfranges)
 
 setwd(here::here("analysis", "VOCC"))
-d <- readRDS("data/mature-all-do-untrimmed.rds")
-# d <- readRDS("data/mature-all-temp-untrimmed.rds")
-d <- readRDS("data/mature-all-do-dvocc.rds")
+
+age <- "mature"
+age <- "immature"
+
+# d <- readRDS("data/", age, "-all-do-untrimmed.rds")
+# d <- readRDS("data/", age, "-all-temp-untrimmed.rds")
+d <- readRDS("data/", age, "-all-do-dvocc.rds")
+
 d <- na.omit(d) %>% as_tibble()
 
 all_species <- unique(d$species)
 
-null_number <- 3
+null_number <- 1
 trim_threshold <- 0.05
 
 if (trim_threshold == 0.05) { trim_percent <- 95}
@@ -101,7 +106,8 @@ with_nulls[[i]] <- .s
 newdata <- do.call(rbind, with_nulls)
 
 # saveRDS(newdata, file = paste0("data/mature-all-temp-with-null-1-untrimmed.rds"))
-saveRDS(newdata, file = paste0("data/mature-all-do-with-null-", null_number, "-untrimmed.rds"))
+# saveRDS(newdata, file = paste0("data/mature-all-do-with-null-", null_number, "-untrimmed.rds"))
+saveRDS(newdata, file = paste0("data/", age, "-all-do-with-null-", null_number, "-untrimmed.rds"))
 
 
 
@@ -186,14 +192,18 @@ for (i in seq_along(all_species)) {
 data <- do.call(rbind, trimmed.dat)
 
 # saveRDS(data, file = paste0("data/mature-", trim_percent, "-all-temp-with-null-", null_number, ".rds"))
-saveRDS(data, file = paste0("data/mature-", trim_percent, "-all-do-with-null-", null_number, ".rds"))
+
+if (age == "immature") {
+ data <- filter(data, species != "Curlfin Sole")}
+
+saveRDS(data, file = paste0("data/", age, "-", trim_percent, "-all-do-with-null-", null_number, ".rds"))
 
 
 
 
 #####################################
 ### PLOT REAL AND FAKE TREND DATA
-data <- readRDS("data/mature-95-all-do-with-null-3.rds")
[…]

### 2020-03-18 analysis/VOCC/6-vocc-regression.R

@@ -3,8 +3,8 @@ library(dplyr)
 library(tidyr)
 library(ggplot2)
 library(gfranges)
-library(future)
-plan(multiprocess)
+# library(future)
+# plan(multiprocess)
 # plan("multisession")
 setwd(here::here("analysis", "VOCC"))
 compile("vocc_regression.cpp")
@@ -54,11 +54,12 @@ data_type <- "mature-95-all-temp"
 # data_type <- "mature-90-all-do"
 # data_type <- "mature-80-all-do"
 data_type <- "mature-95-all-do"
+data_type <- "immature-95-all-do"
 
 #  null_number <- ""
 null_number <- "-1"
  # null_number <- "-2"
-  null_number <- "-3"
+ # null_number <- "-3"
 
 d <- readRDS(paste0("data/", data_type, "-with-null", null_number, ".rds"))
 d <- na.omit(d) %>% as_tibble()
@@ -653,8 +654,8 @@ date <- format(Sys.time(), "-%m-%d")
 
 saveRDS(new_model, file = paste0("data/", y_type, "-", data_type, date, model_type, null_lab, null_number, genus_lab, "-", knots, ".rds"))
 
-# saveRDS(DO_model, file = paste0("data/", y_type, "-", data_type, date, model_type, null_lab, null_number, genus_lab, "-", knots, "-DO.rds"))
-# saveRDS(temp_model, file = paste0("data/", y_type, "-", data_type, date, model_type, null_lab, null_number, genus_lab, "-", knots, "-temp.rds"))
+saveRDS(DO_model, file = paste0("data/", y_type, "-", data_type, date, model_type, null_lab, null_number, genus_lab, "-", knots, "-DO.rds"))
+saveRDS(temp_model, file = paste0("data/", y_type, "-", data_type, date, model_type, null_lab, null_number, genus_lab, "-", knots, "-temp.rds"))
 
 paste0("data/", y_type, "-", data_type, date, model_type, null_lab,  null_number, genus_lab, "-", knots, ".rds")
 

### 2020-03-18 analysis/SOPO/2-trawl-no-covs-model.Rmd
remove sci notation

@@
     # species: "Redbanded Rockfish"
-    # species: "North Pacific Spiny Dogfish"
+    species: "North Pacific Spiny Dogfish"
     # species: "Pacific Ocean Perch"
-    species: "Pacific Sand Lance"
+    # species: "Bocaccio"
+    # species: "Yelloweye Rockfish"
+    # species: "Pacific Sand Lance"
     # species: "Pacific Herring"
@@
 
+# options(scipen = 999)
 theme_set(

### 2020-03-18 analysis/VOCC/3-vocc-dist.Rmd
add mean temp and mean DO maps

@@
 params: 
-    species: "Redbanded Rockfish"
+    # species: "Redbanded Rockfish"
+    species: "Widow Rockfish"
     # species: "Canary Rockfish"
@@
     immature: T
-    region: "West Coast Vancouver Island"
-    # region: "both odd year surveys"
+    # region: "West Coast Vancouver Island"
+    region: "both odd year surveys"
     # region: "West Coast Haida Gwaii"
@@
 max_dist_all_years <- 10 * year_range
-
+bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.05
+s <- sort(gf3$mean_biomass)
+
+bio_sum <- cumsum(s)
+lower_density_threshold <- s[which(bio_sum >= bio5perc)[1]]
+```
+
+```{r}
 dvocc_b <- make_vector_data(d,
@@
 
-bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.05
-s <- sort(gf3$mean_biomass)
-
-bio_sum <- cumsum(s)
-lower_density_threshold <- s[which(bio_sum >= bio5perc)[1]]
-
 gf3t <- filter(gf3, mean_biomass > lower_density_threshold)
@@
 ```{r}
+vocc  <- read.csv(paste0(
+  "data/_dvocc_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
+))
 vocc_t <- filter(vocc, mean_biomass > lower_density_threshold)
@@
   title = "temp trend",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+```
+
+```{r eval=T}
+plot_vocc(vocc_t,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "mean_temp",
+  fill_label = "mean \ntemperature",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "grey",
+  white_zero = TRUE,
+   mid_fill = "lightyellow",
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = no_trans # log10 # fourth_root_power # sqrt #
+) + labs(
+  title = "mean temp",
   subtitle = paste0(min(d$year), "-", max(d$year))
@@
   title = "DO trend",
+  subtitle = paste0(min(d$year), "-", max(d$year))
+)
+```
+
+```{r eval=T}
+plot_vocc(vocc_t,
+  # theme = "black",
+  coast = TRUE,
+  vec_aes = NULL,
+  # grad_vec_aes = "velocity",
+  # vec_lwd_range = c(0.5, 0.5),
+  NA_label = "NA",
+  fill_col = "mean_DO",
+  fill_label = "mean \nDO",
+  raster_alpha = 1,
+  # raster_limits = c(raster_min, raster_max),
+  na_colour = "white",
+  white_zero = TRUE,
+  high_fill = "Steel Blue 4",
+  low_fill = "Red 3",
+   mid_fill = "lightyellow",
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = no_trans # log10 # fourth_root_power # sqrt #
+) + labs(
+  title = "mean DO",
   subtitle = paste0(min(d$year), "-", max(d$year))

### 2020-03-25 analysis/VOCC/3-vocc-dist.Rmd
update mature data with decadal trends and velocities

@@
     # species: "Arrowtooth Flounder"
-    immature: T
+    immature: F
     # region: "West Coast Vancouver Island"
@@
 
-#### Get species-specific optimal values from mountain plots for most abundant year
-```{r, include=FALSE}
+```{r}
 if (immature) {
@@
 write.csv(vocc, file = paste0(
-  "data/_dvocc_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_dvocc_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))

### 2020-03-25 analysis/VOCC/3-render-biotic-vocc.R

@@ -60,7 +60,7 @@ list_regions <- c(
 )
 
 # age <- "mature"
-if age == "mature") {
+if (age == "mature") {
 for (r_h in seq_along(list_regions)) {
   for (spp_i in seq_along(list_species)) {
     spp <- gsub(" ", "-", gsub("\\/", "-", tolower(list_species[spp_i])))
@@ -71,13 +71,14 @@ for (r_h in seq_along(list_regions)) {
         params = list(
           species = list_species[spp_i],
           # immature = TRUE,
+          immature = FALSE,
           region = list_regions[r_h],
           covs = covs
         ),
         output_file = paste0(
           "html/VOCC-plots/vocc-w-do-", spp,
           # "-imm",
-          covs, "-", reg, "-dvocc.html"
+          covs, "-", reg, "-dvocc-decade.html"
         ),
         envir = env
       )
@@ -87,7 +88,7 @@ for (r_h in seq_along(list_regions)) {
 }
 
 # age <- "immature"
-if age == "immature") {
+if (age == "immature") {
 for (r_h in seq_along(list_regions)) {
   for (spp_i in seq_along(list_species)) {
     spp <- gsub(" ", "-", gsub("\\/", "-", tolower(list_species[spp_i])))

### 2020-03-25 analysis/VOCC/4-pool-vocc.R

@@ -6,6 +6,8 @@ library("dplyr")
 mydir = paste0( "_new_null")
 mydir = paste0( "_do_null")
 mydir = paste0( "_dvocc")
+
+mydir = paste0( "_dvocc_mature")
 mydir = paste0( "_dvocc_immature")
 ####################
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
@@ -23,9 +25,9 @@ glimpse(biotic)
 #   "mature-all-do-untrimmed.rds"
 # ))
 # 
-# saveRDS(biotic, file = paste0(
-#   "mature-all-do-dvocc.rds"
-# ))
+saveRDS(biotic, file = paste0(
+  "mature-all-do-dvocc.rds"
+))
 
 saveRDS(biotic, file = paste0(
   "immature-all-do-dvocc.rds"

### 2020-03-25 analysis/VOCC/5-add-null.R

@@ -7,11 +7,11 @@ library(gfranges)
 setwd(here::here("analysis", "VOCC"))
 
 age <- "mature"
-age <- "immature"
+# age <- "immature"
 
 # d <- readRDS("data/", age, "-all-do-untrimmed.rds")
 # d <- readRDS("data/", age, "-all-temp-untrimmed.rds")
-d <- readRDS("data/", age, "-all-do-dvocc.rds")
+d <- readRDS(paste0("data/", age, "-all-do-dvocc.rds"))
 
 d <- na.omit(d) %>% as_tibble()
 

### 2020-04-01 ms/ms.Rmd
Initial commit of Rmd paper draft

---
title: VOCC outline
author: |
  Philina A. English, Authors, ...\
  ^1^Pacific Biological Station, Fisheries and Oceans Canada,\
     3190 Hammond Bay Road, Nanaimo, BC, V6T 6N7, Canada;\
     philina.english@dfo-mpo.gc.ca\
output:
  bookdown::pdf_document2:
    number_sections: yes
    toc: no
fontfamily: libertine
fontsize: 12pt
header-includes:
    - \usepackage{setspace}
    - \onehalfspacing
geometry: "left=2.54cm, right=2.54cm, top=2.54cm, bottom=2.54cm"
bibliography: "refs.bib"
csl: "ecology.csl"
---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, results='hide', warning=FALSE}
library(knitr)
library(here)
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  autodep = TRUE,
  fig.align = "center",
  fig.pos = "htb"
)
```

# Abstract {-}

# Introduction

**Question framing notes:**

-   *Groundfish*

    -   *occupy a relatively homogenous thermal environment*

        -   *predict less tendency to move in response to climate*

    -   *many migrate and school*

    -   *high possibility of movement in response to climate*

-   *Stock assessments assume stationary distributions*

-   *Choices of marine protected areas*

-   *Approaches used often consider the whole of species ranges and apply forward projections assuming that species responses to climate variables will remain the same through time.*

**Spatial scales of biotic responses to climate change**

-   **Range boundaries** ...

-   **Center of gravity** -- commonly used to assess responses...

-   suggests that fishing pressure is more important than average bottom temperature on annual (maybe seasonal?) basis for predicting COG for entire stocks [@adams2018] however a relatively higher influence of bottom temperature on inertia (variance in biomass relative to COG) suggests that is spatial variability in temperature (or responses to temperature) that aren't captured in COG

-   Because population density often increases faster than area occupied, and organisms likely move towards preferred habitat as populations decline( the basin model of density-dependent habitat selection theory)... populations merit cautious management during population declines, because a compressed range may increase the efficiency of harvest. ... [@thorson2016b]

-   ... Could complicate interpretation of indices such as COG?

-   **Local climate velocities** are one way of quantifying these changes at finer spatial scales.

**Measuring biotic responses in terms of density**

These same principles of change velocity can also be applied to changes in biotic variables.

Known as biotic velocity, these become estimates of the minimum distance and approximate direction an organism might be expected to move to reach an equivalent degree of habitat suitability.
Like with climate velocity, a positive value will be associated with an increase in biotic suitability of the focal location and the distance to the nearest location predicted to have current (or future) suitability equivalent to the lower previous suitability of the focal location, while a negative value represents a decline in suitability of the focal location and the distance to the nearest location that is predicted to meet that prior degreed of biotic suitability.

While COG estimates are intrinsically sensitive to changes in abundance, fine scale estimates of species distributions (SDMs) have traditional been based on presence-absence data both because it's easier to collect and simpler to model than zero inflated abundance or biomass data.

However, with advances in computing power and methods of estimating spatial random effect structures and more flexible distributions like the Tweedie, abundance and biomass based species distribution models can be used to assess habitat suitability in terms more relevant to estimating carrying capacity rather than just occupancy.

These capacity approximating metrics are likely to be much more relevant to fisheries management decision and food security, than simple occupancy estimates.

Like SDMs, most previous applications of biotic velocities have focused on habitat suitability in terms of presence absence rather than using the abundance or density information (REFS).

However, changes in climate may cause shifts carrying capacity long before range shifts are clearly detectable. Therefore, we will extend the application of biotic velocities to estimates of change in fish biomass densities rather than probability of presence.

*DO I need to talk about trends here, if trend analysis is included?*

**Time-scales of change depend on life history**

Just as identifying the appropriate spatial and biotic scales for detecting responses to climate can be fraught, the timescales at which changes occur for each of these other dimensions is likely to differ as well.

At longer timescales it because easier to disentangle variability from overall trends, however standardized long-term datasets are relatively rare, and over shorter timescales, a focus on average trends can obscure meaningful variability.

While overall population status and distribution limits are likely related (e.g. through basin model of density-dependent habitat selection), the time delay between changes in survival and/or reproductive rate and the subsequent redistribution of the population will depend on both spatial scale and species life histories.
For example, this lag should be smaller when generation times are particularly short, or when individuals of all age classes tend to disperse greater distances.
Likewise, changes in local abundances at finer spatial scales should occur more quickly due to shorter dispersal distances involved and *reduced homogenization of responses*.
<!-- TODO: Struggling with how to describe the dilution of responses by greater conflicting processes at larger spatial scales?  -->
Even at a local scale, the rate of change in overall biomass density should still be limited by species' generation times under certain conditions (e.g. when fecundity or early life history stages are most climate sensitive, and when the larger age classes are highly sedentary or territorial).
<!-- TODO: Check 20 & find REF -->
In order to detect the impact of climate change on species like rockfish (many of which are territorial with generation times > 20 yrs), one would require either datasets in excess of 20 years, or to contrast patterns of change between age classes.
Indeed, responses to environmental change can be expected to differ between species, and between age classes within species, depending on the life history of species including physiological tolerances , lifespan and dispersal patterns.
For groundfish it is well established that some species shift their depth distribution as they age (often approximated by fish length bins; REF e.g.
Li et al 2019).
However, choosing the relevant length (or age, if available) thresholds can complicate analyses.
Instead, the reaching of reproductive maturity is likely an important threshold for shifts in dispersal patterns, habitat selectivity, physiological tolerances, that therefore represents the most appropriate break point for understanding the impacts of climate.
Fortunately reproductive maturity information is much easier to collect than fish age, and is therefore more widely available for a wide range of species.

Fishery-independent survey data is often collected to aid in stock assessments and fisheries management.
In Canadian Pacific waters, randomized bottom trawl surveys outfitted with dissolved oxygen and temperature collecting CTD sensors, have been run for about a decade now.
During this time bottom temperatures have tended to rise everywhere, while the direction of change in the availability of oxygen has been spatially variable.
In addition to estimates of biomass density for each species captured, data on size distributions and reproductive maturity are also collected for most species.
***We use these concurrent estimates of both climate and biomass to assess the extent to which groundfish densities are changing in response to climate at the local scales most relevant to management***.
In order to better understand the extent to which the temporal and spatial scales considered are appropriate for different species within this fish community, we also assess the concordance between the life histories and apparent sensitivities across species and maturity classes.
Given the relatively short timescale encompassed in this survey, we expect *stronger patterns in shorter-lived, more pelagic species, and immatures.*
<!-- TODO: Discuss with sean & chris rooper -->

# Methods

Example Figure reference: Fig. \@ref(fig:climate-maps).

## Survey data

Fisheries and Oceans Canada conducts fishery-independent bottom trawl surveys with biennial coverage of the continental shelf and upper slope of most of the BC coast.
These surveys follow a random depth stratified design and use the same bottom trawl fishing gear and fishing protocols (Sinclair et al. 2003.
Surveys are stratified within four separate regions two of which are surveyed each year since 2005 (with surveys also conducted in some regions in 2003 & 2004): Hecate Strait and Queen Charlotte Sound (odd years) and West Coast Vancouver Island and West Coast Haida Gwaii (even years).
Each region is sampled over the same month long period between late May and early August each year.
Only successful tows of 15--20 min duration are included in analyses and total biomass of each species captured was converted to a biomass density estimate based on the speed, distance covered and net opening.
Catches of between 10--50 individuals are weighed and measured individually, while larger catches are subsampled and for commercially important species data on maturity and age are also collected.

## Estimate spatiotemporal variation in adult and immature biomass densities

We chose 36 species of Pacific groundfish most regularly encountered on fisheries-independent bottom trawl surveys and relatively widely distributed within Canadian waters.
Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species (Mindel et al. 2016, Barbeaux and Hollowed 2018, Li et al. 2019), we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low).
To do this, we first estimate length at 50% maturity (defined by gonadal development stages of immature or maturing vs. mature, ripe, or spent) using ogives fit as sex-specific logistic regressions to individual specimens.
When sufficient samples were available in all years, a random effect of year was included to allow for temporal change in size at maturity (REF) and these year specific thresholds were used to split observed catches.
For those species lacking maturity data in some years, a random effect of tow id was included instead, but catches were split based on the global estimate.
For all catches exceeding the species-specific threshold count of individual fish, either a random subset of roughly 50 fish, or all individuals will be measured for both length and weight (occasional missing weights were estimated from $\log(W_i) \sim \mathrm{Student-t} (df = 3, \log(a) + b \log(L_i), \sigma)$, with $W_i$ and $L_i$ representing the weight and length for fish $i$ and $\sigma$ representing the observation error scale).
We then calculate the summed weight of all measured which exceeded the length at 50% maturity threshold, divided by the total weight of all measured fish, and multiply this ratio by the estimated biomass density for each tow.
For each tow which resulted in a catch too small for detailed measurements to have been taken, we apply the mean ratio from all measured tows to estimate mature biomass.
Immature density was estimated as the difference between the total density and mature density.

We then modeled expected biomass density through space and time separately for adults and immatures of each species using spatial GLMMs (e.g., Shelton et al. 2014, Thorson et al. 2015, 2016, Ward et al. 2015).
We model each species and maturity class separately because although these abundance patterns are certainly correlated, explicitly modeling those relationships among groundfish species has not been shown to dramatically improve precision of overall estimates (Thorson and Barnett 2017), and keeping estimates independent should make identification of shared climate responses more conservative.
Because biomass densities contain both zeros and positive continuous values, we modeled the response with a Tweedie distribution and a log link (Anderson et al. in press, Tweedie 1984, Dunn and Smyth 2005).
Each model included fixed-effects for year, year-specific quadratic effects of bottom depth, spatial random effects that represent an amalgamation of spatial processes not explicitly included in the model, spatiotemporal random effects that allow these spatial random effects to differ independently between years, and an observation error component.
The spatial components were modeled as random fields, using a triangulated mesh with vertices at 400 knots to approximate a continuous spatial field (Rue et al. 2009, Lindgren et al. 2011).
The inclusion of year as a factor centered each of the estimated spatiotemporal fields to have mean zero. Maximum likelihood estimation was conducted in R version 3.5.3 (R Core Team 2019) with the package sdmTMB (Anderson et al. In press, Anderson 2019) which interfaces automatic differentiation in Template Model Builder (Kristensen et al. 2016) with INLA (Rue et al. 2009) to .

We modeled the effects of depth and depth squared by first standardizing the log-transformed depth covariate by subtracting its mean and scaling it by its standard deviation.
We then calculated 'depth squared' from this centred and scaled variable.
This ensures the covariate values are not too large or small to avoid computational issues and the centring separates the linear and quadratic predictor components.
We also tried including the proportion of muddy sediment, and proportion of sediment with any rocks, but the values we used were likely not specific enough to the trawl path sampled, and therefore not helpful to our models.
Certainly more spatially precise estimates would be possible with refined substrate variables, but likely irrelevant to this analysis given the resolution and uncertainty in our climate variables.
Furthermore, the spatial random effects can account for spatial factors that are constant across time, like substrate type, while spatiotemporal random effects account for factors that vary from year to year spatially such as bottom temperature, water circulation patterns, species interactions, and species movement.

*MODEL VALIDATION?*

## Estimate climate velocities

A Seabird Electronics 19plus conductivity, temperature, and depth recorder (CTD) with dissolved oxygen (DO) sensor has also been successfully deployed on most tows since 2008.
Therefore, we modeled bottom temperature and DO levels from CTD sensor data collected on each survey tow using a similar geostatistical model as for biomass.
The approach for these models differed only be being Gaussian and by estimating the spatiotemporal random fields using an AR1 process.
Because DO levels cannot be negative and are known to be influenced be both water temperature and seasonal biological processes (REF), we predict log(DO) with added fixed effects for day of year (centered and scaled) and temperature (quadratic of scaled values similar to depth).
Bottom temperature data (but not DO) was also collected on tows prior to 2008, so this data was also included in the model in order to better inform the static spatial field component of the model.

For all climate and biomass models, we then project the model predictions onto a 4 km x 4 km grid (UTM 9 projection) consisting of the particular coastal zone surveyed in each year. We then calculate both gradient-based and distance-based velocities of change, and constituent local trends and spatial gradients, for each cell.

Explain ways of calculating:

- Gradient velocities
- Analog-distance velocities <!-- TODO: Might leave out -->

Both spatial gradients (denominator for gradient-based calculation) and t~2~ conditions (distance-search method) are based on the years 2014-2019 (which correspond to the start of the "Blob", and corresponds with the warm transition in Alaskan waters---Albia et al. 2018).
Most prior application of gradient-based VOCC use a mean spatial gradient across the entire period analyzed, which tends to result in smaller gradients due to reduced variability in means based on larger sample sizes.
Because velocity estimates explode as gradients approach zero, estimated velocities will tend to be smaller (more conservative) when fewer estimates are averaged in the gradient calculation.
Furthermore, the period most relevant to the actual distance travelled is end point, however given that samples were collected only once every two years and that there is variability between species in when dispersal occurs and how long it takes, we use roughly the latter half of the time period analyzed in our estimates of spatial gradients.
<!-- TODO: Cut if analog cut: -->
This gradient is also based on exactly the same cell values are as searched in the analog-distance algorithm.

Because locations varied in the number of years sampled ...

Because the space searched by the analog-distance algorithm was much larger in one dimension than in others, we truncate distance to a max of 100 km.
Likewise, extremes in the gradient-based velocity estimates were collapsed to the 0.005 and 0.995 quantiles.
As a result both estimates produce heavy-tailed distributions...

<!-- TODO: Might leave out? -->
**Control for commercial fishing effort?**

-   include fishing trips up to 9hrs in length

-   calculate mean hrs fished per year per 4x4 km grid cell

-   calculate % change in hrs fished between 2008-2018

**Link biomass change with climate**

To explore the potential for change in climate to predict local changes in biomass density, we used similar spatial GLMMs, but with independent random spatial fields for each species (and genus?)
in order to control for spatial autocorrelation amongst cells.
For each species, only the fewest cells that encompass 95% of total mean biomass were included.
In all models mean biomass density was included as a covariate in order to reduce the influence of changes occurring only at either the highest or lowest densities for a particular species.
We attempt to predict change in biomass of each species as a function of local climate change (just temperature, or both temperature and DO) and an interaction between the mean climate of each cell and its local change estimate.
We model both biomass and climate as either velocities or as raw trends. In addition we attempt to assess the importance of the gradient component in velocity estimates by adding gradient to the model based on trends.

All covariates were scaled to their standard deviations.
Covariates capturing cell means were also centered (except fishing effort---so that estimates can be interpreted as in the absence of fishing), but change variables were kept uncentred because an intercept of 0 change is more intuitive to interpret.
In addition to checking the spatial distribution of model residuals, we simulated fake biomass trends for each species (using variance and spatial correlation parameters unique to each species) and reran each model 3 times and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
<!-- TODO: Maybe up to 10 times? -->

In order to better understand the extent to which the temporal and spatial scales considered are appropriate for different species within this groundfish community, we also assess the concordance between the life histories and apparent sensitivities across species and maturity classes.

# Results

Very little difference whether biotic velocity or trend are the response variable...

Big difference whether temp velocity or temp trend are the predictor...

# Discussion

# Acknowledgements

# References

<div id="refs"></div>

\clearpage

# Figures

<!-- Reference this figure as: Fig. \@ref(fig:climate-maps) -->

```{r climate-maps, fig.cap="Caption goes here", out.width="5in"}
include_graphics(here("ms", "figs", "example.pdf"))
```

\clearpage

```{r another, fig.cap="Caption goes here", out.width="5in"}
include_graphics(here("ms", "figs", "example.pdf"))
```

\clearpage

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\setcounter{section}{0}
\setcounter{subsection}{0}
\setcounter{subsubsection}{0}

# Supporting Information {-}

# TODO Repeat the title here {-}

TODO Repeat the authors here

# Supporting Methods

<!-- TODO: rewrite the following: -->

These models can be represented as Tweedie GLMMs with a log link, spatiotemporal random effects defined for locations in space $s$ and time $t$), and annual predictors for the mean biomass each year:
\begin{align}
  y_{s,t} &\sim \mathrm{Tweedie}(\mu_{s,t}, p, \phi), \quad 1 < p < 2,\\
  \mu_{s,t} &= \exp \left( \mathbf{X}_{s,t} \mathbf{\beta} + \omega_s + \epsilon_{s,t} \right).
\end{align}
where $s$ represents a spatial location, $y_s$ represents observed fish density
for a survey set, $\mu_s$ represents expected fish density, $p$ represents the
Tweedie power parameter, and $\phi$ represents the Tweedie dispersion
parameter. The symbol $\mathbf{X_s}$ represents a vector of predictors (an
intercept, log depth, and log depth squared) and $\mathbf{\beta}$ represents
a corresponding vector of coefficients. The spatial random effects $\omega_s$
<!-- TODO: Need to specify time-varying depth -->
are assumed to be drawn from a multivariate normal distribution with
a covariance matrix $\mathbf{\Sigma}_\omega$ that is centered on zero:

$$
\mathbf{\omega} \sim \mathrm{MVNormal} \left( \mathbf{0}, \mathbf{\Sigma}_\omega \right),
$$

and we assume the same for the spatiotemporal random effects, with each time slice being given its own independent set of random effects ($t$) with covariance matrix $\mathbf{\Sigma}_{\epsilon,t}$::

$$
\mathbf{\epsilon}_t \sim \mathrm{MVNormal} \left( \mathbf{0}, \mathbf{\Sigma}_{\epsilon,t} \right).
$$

All spatial random effects are constrained to follow a Mat\'ern covariance function, which defines the rate with which spatial correlation decays with distance. The Matern function describes the covariance XX (sj , sk ) between spatial locations sj and sk as:

TODO: finish and simplify

\clearpage

# Supporting Figures and Tables

```{r yet-another, fig.cap="Caption goes here", out.width="5in"}
include_graphics(here("ms", "figs", "example.pdf"))
```

### 2020-04-01 ms/ms.Rmd
Add some references

@@
 Indeed, responses to environmental change can be expected to differ between species, and between age classes within species, depending on the life history of species including physiological tolerances , lifespan and dispersal patterns.
-For groundfish it is well established that some species shift their depth distribution as they age (often approximated by fish length bins; REF e.g.
-Li et al 2019).
+For groundfish it is well established that some species shift their depth distribution as they age [often approximated by fish length bins; e.g. @li2019].
 However, choosing the relevant length (or age, if available) thresholds can complicate analyses.
@@
 Fisheries and Oceans Canada conducts fishery-independent bottom trawl surveys with biennial coverage of the continental shelf and upper slope of most of the BC coast.
-These surveys follow a random depth stratified design and use the same bottom trawl fishing gear and fishing protocols (Sinclair et al. 2003.
+These surveys follow a random depth stratified design and use the same bottom trawl fishing gear and fishing protocols (TODO: Sinclair et al. 2003).
 Surveys are stratified within four separate regions two of which are surveyed each year since 2005 (with surveys also conducted in some regions in 2003 & 2004): Hecate Strait and Queen Charlotte Sound (odd years) and West Coast Vancouver Island and West Coast Haida Gwaii (even years).
@@
 We chose 36 species of Pacific groundfish most regularly encountered on fisheries-independent bottom trawl surveys and relatively widely distributed within Canadian waters.
-Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species (Mindel et al.
-2016, Barbeaux and Hollowed 2018, Li et al.
-2019), we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low).
+Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species [@mindel2016; @barbeaux2018; @li2019], we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low)
 To do this, we first estimate length at 50% maturity (defined by gonadal development stages of immature or maturing vs.
@@
 
-We then modeled expected biomass density through space and time separately for adults and immatures of each species using spatial GLMMs (e.g., Shelton et al.
-2014, Thorson et al.
-2015, 2016, Ward et al.
-2015).
+We then modeled expected biomass density through space and time separately for adults and immatures of each species using spatial GLMMs [e.g., @shelton2014; @thorson2015; @thorson2016a; @ward2015].
 We model each species and maturity class separately because although these abundance patterns are certainly correlated, explicitly modeling those relationships among groundfish species has not been shown to dramatically improve precision of overall estimates (Thorson and Barnett 2017), and keeping estimates independent should make identification of shared climate responses more conservative.
@@
 Each model included fixed-effects for year, year-specific quadratic effects of bottom depth, spatial random effects that represent an amalgamation of spatial processes not explicitly included in the model, spatiotemporal random effects that allow these spatial random effects to differ independently between years, and an observation error component.
-The spatial components were modeled as random fields, using a triangulated mesh with vertices at 400 knots to approximate a continuous spatial field (Rue et al. 2009, Lindgren et al. 2011).
+The spatial components were modeled as random fields, using a triangulated mesh with vertices at 400 knots to approximate a continuous spatial field (TODO: Rue et al. 2009, Lindgren et al. 2011).
 The inclusion of year as a factor centered each of the estimated spatiotemporal fields to have mean zero.

### 2020-04-01 ms/ms.Rmd
Add ^1^

@@
 author: |
-  Philina A. English, Authors, ...\
+  Philina A. English^1^, Authors, ...\
   ^1^Pacific Biological Station, Fisheries and Oceans Canada,\

### 2020-04-01 ms/ms.Rmd
Add line numbers

--- 2020-04-01 ms/ms.Rmd
@@
   - \onehalfspacing
-geometry: "left=2.54cm, right=2.54cm, top=2.54cm, bottom=2.54cm"
+  - \usepackage{lineno}
+  - \linenumbers
+  - \modulolinenumbers[2]
+geometry: "left=2.64cm, right=2.64cm, top=2.54cm, bottom=2.54cm"
 bibliography: "refs.bib"

### 2020-04-07 ms/ms.Rmd
add list of proposed figs to ms and link to early draft figs

@@
 
+<!-- TODO: figures to consider
+-   temperature and O2 maps and velocities; 4 panel
+-   coefficient plots temp + O2; add meta-analytical coefficients?
+-   some example temperature maps plus chopsticks
+-   some example O2 maps plus chopsticks
+-   plot coefficients against depth etc.... adults and immature separately?
+-->
+
 # Discussion
@@
 
-```{r climate-maps, fig.cap="Caption goes here", out.width="5in"}
-include_graphics(here("ms", "figs", "example.pdf"))
+```{r climate-maps, fig.cap="Caption goes here", out.width="6in"}
+include_graphics(here("ms", "figs", "climate-maps.pdf"))
 ```
@@
 
-```{r another, fig.cap="Caption goes here", out.width="5in"}
-include_graphics(here("ms", "figs", "example.pdf"))
+```{r worm-plots, fig.cap="Caption goes here", out.width="8in"}
+# include_graphics(here("ms", "figs", "worm-plots.pdf"))
+```
+
+\clearpage
+
+```{r temp-chops, fig.cap="Caption goes here", out.width="8in"}
+# include_graphics(here("ms", "figs", "temp-chops.pdf"))
+```
+
+\clearpage
+
+```{r do-chops, fig.cap="Caption goes here", out.width="8in"}
+include_graphics(here("ms", "figs", "do-chops.pdf"))
 ```
@@
 
+Supplemental figures:
+(In no particular order)
+-   flow diagram
+-   spatial residuals
+-   null model stuff
+-   all the chopstick panels
+-   other coefs plot
+-   velocity versions
+-   pairs plot of predictors
+-   plot vocc regression residuals against the main predictors... (and not in model and fitted vs. residuals plot; but don't necessarily include in paper)
+
 ```{r yet-another, fig.cap="Caption goes here", out.width="5in"}

### 2020-04-11 ms/ms.Rmd
add draft figs to ms

@@
 ```{r worm-plots, fig.cap="Caption goes here", out.width="8in"}
-# include_graphics(here("ms", "figs", "worm-plots.pdf"))
+# include_graphics(here("ms", "figs", "worm-plot-trend.pdf")) # 
+include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
 ```
@@
 ```{r do-chops, fig.cap="Caption goes here", out.width="8in"}
-include_graphics(here("ms", "figs", "do-chops.pdf"))
+# include_graphics(here("ms", "figs", "do-chops.pdf"))
+```
+
+\clearpage
+
+```{r slope-by-depth, fig.cap="Caption goes here", out.width="5in"}
+include_graphics(here("ms", "figs", "slope-by-depth.pdf"))
+```
+
+\clearpage
+
+```{r coef-scatterplots, fig.cap="Caption goes here", out.width="8in"}
+include_graphics(here("ms", "figs", "coef-scatterplots.pdf"))
 ```

### 2020-04-14 ms/ms.Rmd
start drafting fig captions

--- 2020-04-01 ms/ms.Rmd
@@
   - \usepackage{lineno}
+  - \usepackage{fullpage}
+  - \usepackage{pdflscape}
   - \linenumbers
@@
 when fecundity or early life history stages are most climate sensitive, and when the larger age classes are highly sedentary or territorial).
-<!-- TODO: Check 20 & find REF -->
+<!-- TODO: Check generation time & find REF -->
 In order to detect the impact of climate change on species like rockfish (many of which are territorial with generation times > 20 yrs), one would require either datasets in excess of 20 years, or to contrast patterns of change between age classes.
@@
 We attempt to predict change in biomass of each species as a function of local climate change (just temperature, or both temperature and DO) and an interaction between the mean climate of each cell and its local change estimate.
-We model both biomass and climate as either velocities or as raw trends.
+We model both biomass and climate as either velocities or as raw trends (Fig.
+\@ref(fig:climate-maps)).
 In addition we attempt to assess the importance of the gradient component in velocity estimates by adding gradient to the model based on trends.
@@
 
+<!-- TODO: remaining figures to consider
+-   some example temperature maps plus chopsticks
+-   some example O2 maps plus chopsticks
+-->
+
 # Discussion
@@
 <!-- Reference this figure as: Fig. \@ref(fig:climate-maps) -->
+<!-- Reference next figure as: Fig. \@ref(fig:worm-plots) -->
+<!-- TODO: could include velocity worm plots too: include_graphics(here("ms", "figs", "worm-plot-both.pdf")) -->
 
-```{r climate-maps, fig.cap="Caption goes here", out.width="5in"}
-include_graphics(here("ms", "figs", "example.pdf"))
+```{r climate-maps, fig.cap="Maps of predicted trends (top panels) and velocities (bottom panels) of change for bottom temperature and dissolved oxygen during the trawl surveys conducted between 2008 to 2018.
+Values are estimated from a spatiotemporal models of CTD sensor data collected during late spring/early summer groundfish trawl surveys.", out.width="5in"}
+include_graphics(here("ms", "figs", "climate-maps-updated.pdf"))
 ```
@@
 
-```{r another, fig.cap="Caption goes here", out.width="5in"}
-include_graphics(here("ms", "figs", "example.pdf"))
+```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss for groundfish species.
+Open circles indicate patterns for immatures, while closed circles represent individuals large enough to have a 5 percent chance of being reproductively mature or belonging to species for which maturity data was not available.
+Black indicates overall slope where the estimated interaction with mean climate condition overlapped zero.
+Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).", out.width="6in"}
+include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
+```
+
+\clearpage
+
+<!-- TODO: Species of interest...
+
+redbanded - pretty normal response to temp and DO (sig -ve for adults only) 
+        widow rockfish  - strong interactions
+canary - not interaction, -ve temp none for DO
+shortspine thornyhead - both temp (increasing good) and DO (increasing good)
+
+dogfish - increasing DO and warming both bad
+pacific cod - 
+English - warming neutral when cool, bad when warm
+Arrowtooth or Dover - warming good when it's cool, neutral when warm 
+sablefish - warming good, increasing DO not
+
+-->
+
+```{r temp-chops, fig.cap="Temperature ...", out.width="6in"}
+# include_graphics(here("ms", "figs", "temp-chops.pdf"))
+```
+
+\clearpage
+
+```{r do-chops, fig.cap="Dissolved oxygen ...", out.width="6in"}
+# include_graphics(here("ms", "figs", "do-chops.pdf"))
+```
+
+\clearpage
+
+```{r slope-by-depth, fig.cap="Variation in slopes across species with different mean depths of occurrence.", out.width="4in"}
+include_graphics(here("ms", "figs", "slope-by-depth.pdf"))
+```
+
+\clearpage
+
+<!-- TODO: May be we should change this to mean age of mature fish sampled for each species? That would make it easier to relate values to the timespan of the study. 
+-->
+
+```{r coef-by-max-age.pdf, fig.cap="The maximum lifespan of the species may influence the relationships between climate trends and biomass change.", out.width="4in"}
+include_graphics(here("ms", "figs", "coef-by-max-age.pdf"))
 ```
@@
 are assumed to be drawn from a multivariate normal distribution with
-a covariance matrix $\mathbf{\Sigma}_\omega$ that is centered on zero:
+a covariance matrix $\mathbf{\Sigma}_\omega$ that is centred on zero:
 
@@
 
-All spatial random effects are constrained to follow a Mat\'ern covariance function, which defines the rate with which spatial correlation decays with distance.
+All spatial random effects are constrained to follow a Matern covariance function, which defines the rate with which spatial correlation decays with distance.
 The Matern function describes the covariance XX (sj , sk ) between spatial locations sj and sk as:
 
-TODO: finish and simplify
+<!-- TODO: finish and simplify -->
 
@@
 
+Supplemental figures:
+(In no particular order)
+-   flow diagram
+-   spatial residuals
+-   null model stuff
+
+-   all the chopstick panels both trend and velocity
+
+<!-- TODO: change orientation to landscape? -->
+
+\begin{landscape}
+
+```{r trend-chopsticks, fig.cap="Caption goes here", out.width="11in"}
+include_graphics(here("ms", "figs", "supp-trend-chopsticks.pdf"))
+```
+
+\clearpage
+
+```{r vel-chopsticks, fig.cap="Caption goes here", out.width="11in"}
+include_graphics(here("ms", "figs", "supp-vel-chopsticks.pdf"))
+```
+
+\end{landscape}
+
+-   other coefs plot
+<!-- TODO: This fig might go here instead of in results -->
+
+```{r supp-coef-scatterplots, fig.cap="Caption goes here", out.width="6in"}
+include_graphics(here("ms", "figs", "supp-coef-scatterplots.pdf"))
+```
+
+-   pairs plot of predictors
+
+-   plot vocc regression residuals against the main predictors... (and not in model and fitted vs. residuals plot; but don't necessarily include in paper)
+
 ```{r yet-another, fig.cap="Caption goes here", out.width="5in"}

### 2020-04-15 ms/ms.Rmd
more draft fig captions

@@
 
+<!-- TODO: can a % symbol be used in fig caption? -->
+
 ```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss for groundfish species.
@@
 English - warming neutral when cool, bad when warm
-Arrowtooth or Dover - warming good when it's cool, neutral when warm 
+Dover - warming good when it's cool, neutral when warm  
+(arrowtooth shows similar pattern but likely influenced strongly by fishing... could include for that reason though?)
+
 sablefish - warming good, increasing DO not
@@
 
-```{r temp-chops, fig.cap="Temperature ...", out.width="6in"}
-# include_graphics(here("ms", "figs", "temp-chops.pdf"))
+\begin{landscape}
+
+```{r temp-chops, fig.cap="Temperature ...", out.width="10in"}
+include_graphics(here("ms", "figs", "temp-chops.pdf"))
 ```
 
-\clearpage
+\end{landscape}
 
 ```{r do-chops, fig.cap="Dissolved oxygen ...", out.width="6in"}
-# include_graphics(here("ms", "figs", "do-chops.pdf"))
+include_graphics(here("ms", "figs", "do-chops.pdf"))
 ```
 
-\clearpage
+\begin{landscape}
 
-```{r slope-by-depth, fig.cap="Variation in slopes across species with different mean depths of occurrence.", out.width="4in"}
-include_graphics(here("ms", "figs", "slope-by-depth.pdf"))
+```{r coef-scatterplots, fig.cap="Main effect coefficients for relationship between biotic trends and changes in climate for species with different life history strategies.
+Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50 percent maturity threshold) within a species.
+Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of ages recorded for immatures of each species.", out.width="9in"}
+include_graphics(here("ms", "figs", "coef-scatterplots.pdf"))
 ```
 
-\clearpage
-
-<!-- TODO: May be we should change this to mean age of mature fish sampled for each species? That would make it easier to relate values to the timespan of the study. 
--->
-
-```{r coef-by-max-age.pdf, fig.cap="The maximum lifespan of the species may influence the relationships between climate trends and biomass change.", out.width="4in"}
-include_graphics(here("ms", "figs", "coef-by-max-age.pdf"))
-```
-
-\clearpage
+\end{landscape}
 
@@
 Supplemental figures:
-(In no particular order)
+(not yet done)
 -   flow diagram
@@
 -   null model stuff
-
--   all the chopstick panels both trend and velocity
-
-<!-- TODO: change orientation to landscape? -->
+-   pairs plot of predictors
+-   plot vocc regression residuals against the main predictors... (and not in model and fitted vs. residuals plot; but don't necessarily include in paper)
 
@@
 
+\clearpage
+
+```{r supp-coef-scatterplots, fig.cap="Differences slope of effect of climate change on biotic change at how high and low mean conditions (columns) plotted against species life histories.
+Simple regression lines plotted for immature abundance only.
+Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50 percent maturity threshold) within a species.
+Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of ages recorded for immatures of each species.", out.width="8in"}
+include_graphics(here("ms", "figs", "supp-slope-scatterplots.pdf"))
+```
+
 \end{landscape}
 
--   other coefs plot
-<!-- TODO: This fig might go here instead of in results -->
-
-```{r supp-coef-scatterplots, fig.cap="Caption goes here", out.width="6in"}
-include_graphics(here("ms", "figs", "supp-coef-scatterplots.pdf"))
+```{r supp-slope-by-depth, fig.cap="Shape of relationship between estimated mean dissolved oxygen and mean depth of each 4x4 km grid cell (pattern is vary similar for temperature so not illustrated) for comparison with slopes for the environmental extremes experienced at the deeper depths showing expected tendancy towards increasing variation in strength of responses amongst species with deeper mean depth of occurrence.
+Outliers amongst immatures are Widow Rockfish (both DO and temperature) and Bocaccio (temperature).", out.width="4in"}
+include_graphics(here("ms", "figs", "supp-slope-by-depth.pdf"))
 ```
 
--   pairs plot of predictors
-
--   plot vocc regression residuals against the main predictors... (and not in model and fitted vs. residuals plot; but don't necessarily include in paper)
-
-```{r yet-another, fig.cap="Caption goes here", out.width="5in"}
-include_graphics(here("ms", "figs", "example.pdf"))
-```
+\clearpage

### 2020-04-17 ms/ms.Rmd
tidy TODOs; start writing results

@@
 )
+
+# TODO: is this where this should go?
+alldata <- readRDS(paste0("../analysis/VOCC/data/all-do-with-null-1-untrimmed-allvars.rds"))
+model <- readRDS("../analysis/VOCC/data/trend-all-95-all-do-04-11-trend-with-do-family-family-1-500.rds")
 ```
@@
 Likewise, changes in local abundances at finer spatial scales should occur more quickly due to shorter dispersal distances involved and *reduced homogenization of responses*.
+
 <!-- TODO: Struggling with how to describe the dilution of responses by greater conflicting processes at larger spatial scales?  -->
+
 Even at a local scale, the rate of change in overall biomass density should still be limited by species' generation times under certain conditions (e.g.
 when fecundity or early life history stages are most climate sensitive, and when the larger age classes are highly sedentary or territorial).
+
 <!-- TODO: Check generation time & find REF -->
+
 In order to detect the impact of climate change on species like rockfish (many of which are territorial with generation times > 20 yrs), one would require either datasets in excess of 20 years, or to contrast patterns of change between age classes.
@@
 Given the relatively short timescale encompassed in this survey, we expect *stronger patterns in shorter-lived, more pelagic species, and immatures.*
+
 <!-- TODO: Discuss with sean & chris rooper -->
@@
 
-*MODEL VALIDATION?*
+<!-- TODO: MODEL VALIDATION -->
 
@@
 For all climate and biomass models, we then project the model predictions onto a 4 km x 4 km grid (UTM 9 projection) consisting of the particular coastal zone surveyed in each year.
-We then calculate both gradient-based and distance-based velocities of change, and constituent local trends and spatial gradients, for each cell.
+We then calculate gradient-based velocities of change, <!-- If analog included: both gradient-based and distance-based velocities of change, --> and constituent local trends and spatial gradients, for each cell.
 
-Explain ways of calculating:
+<!-- TODO: Explain calculation of gradient velocity-->
+<!-- TODO: Explain collapsing to max of 100 km ... and update code and models to do this more precisely-->
 
-- Gradient velocities
-- Analog-distance velocities <!-- TODO: Might leave out -->
+<!-- If analog included: Explain ways of calculating analog velocity and change next sentence to: Both spatial gradients (denominator for gradient-based calculation) and t~2~ conditions (distance-search method) are based on the years 2014-2019 (which correspond to the start of the "Blob", and corresponds with the warm transition in Alaskan waters---Albia et al.
+2018).
+-->
 
-Both spatial gradients (denominator for gradient-based calculation) and t~2~ conditions (distance-search method) are based on the years 2014-2019 (which correspond to the start of the "Blob", and corresponds with the warm transition in Alaskan waters---Albia et al.
+Spatial gradients (denominator for gradient-based calculation) are based on the years 2014-2019 (which correspond to the start of the "Blob", and corresponds with the warm transition in Alaskan waters---Albia et al.
 2018).
@@
 Furthermore, the period most relevant to the actual distance travelled is end point, however given that samples were collected only once every two years and that there is variability between species in when dispersal occurs and how long it takes, we use roughly the latter half of the time period analyzed in our estimates of spatial gradients.
-<!-- TODO: Cut if analog cut: -->
+
+<!-- If analog included: 
 This gradient is also based on exactly the same cell values are as searched in the analog-distance algorithm. 
+-->
 
@@
 
-<!-- TODO: Might leave out? -->
-**Control for commercial fishing effort?**
-
+<!-- If fishing included: Control for commercial fishing effort? 
   -   include fishing trips up to 9hrs in length
-
   -   calculate mean hrs fished per year per 4x4 km grid cell
-
   -   calculate % change in hrs fished between 2008-2018
+-->
 
@@
 
-To explore the potential for change in climate to predict local changes in biomass density, we used similar spatial GLMMs, but with independent random spatial fields for each species (and genus?)
-in order to control for spatial autocorrelation amongst cells.
+To explore the potential for change in climate to predict local percent changes in biomass density, we used similar spatial GLMMs in order to control for spatial autocorrelation amongst cells.
+For each maturity class of each species the model estimates an independent random spatial field was and with and without either genus or family as possible phylogenetic grouping factors.
+.
 For each species, only the fewest cells that encompass 95% of total mean biomass were included.
@@
 We attempt to predict change in biomass of each species as a function of local climate change (just temperature, or both temperature and DO) and an interaction between the mean climate of each cell and its local change estimate.
-We model both biomass and climate as either velocities or as raw trends (Fig.
-\@ref(fig:climate-maps)).
+We model both biomass and climate as either velocities or as raw trends<!-- TODO: should (Fig.
+\@ref(fig:climate-maps)) be here?-->.
 In addition we attempt to assess the importance of the gradient component in velocity estimates by adding gradient to the model based on trends.
-
 All covariates were scaled to their standard deviations.
-Covariates capturing cell means were also centered (except fishing effort---so that estimates can be interpreted as in the absence of fishing), but change variables were kept uncentred because an intercept of 0 change is more intuitive to interpret.
-In addition to checking the spatial distribution of model residuals, we simulated fake biomass trends for each species (using variance and spatial correlation parameters unique to each species) and reran each model 3 times and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
-<!-- TODO: Maybe up to 10 times? -->
+Covariates capturing cell means were also centered <!-- If fishing included: (except fishing effort---so that estimates can be interpreted as in the absence of fishing)-->, but change variables were kept uncentred because an intercept of 0 change is more intuitive to interpret.
+Models exploring change as simple trends use Gaussian distributions, while velocities are  
+In addition to checking the spatial distribution of model residuals, we simulated fake biomass trends for each species (using variance and spatial correlation parameters unique to each species) and reran each model 3 <!-- TODO: simulated null models up to 10 times?
+--> times and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
 
@@
 
-Very little difference whether biotic velocity or trend are the response variable...
+Our models suggest that over the decade spanning 2008 to 2018, seafloor temperature increased across almost the entirety of Canada's Pacific shelf region (95% interquartile range of `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[2]], digits = 2)`), with warming tending to be most pronounced in shallower areas (Fig.
+\@ref(fig:climate-maps)<!-- TODO: should letters be added and included here?
+eg.
+"C" -->).
+In contrast, change in seafloor DO was much more variable in the direction of change (95% interquartile range of `r signif(quantile(alldata$DO_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$DO_trend, c(0.025, 0.975))[[2]], digits = 2)`), with both the greatest increases also occurring in the shallowest areas and the greatest decreases occuring across a wide range of depths <!-- TODO: should letters be added and included here?
+eg.
+(Fig.
+\@ref(fig:climate-maps)D) -->.
+When these local changes are placed in there geographic context in the form of a gradient-based velocity estimate, the majority of the highest velocities occur in the bottom of troughs in Queen Charlotte Sound and Dixon Entrance <!-- TODO: should letters be added and included here?
+eg.
+(Fig.
+\@ref(fig:climate-maps)E & F) -->.
 
-Big difference whether temp velocity or temp trend are the predictor...
+<!-- TODO:  Species of interest...
 
-<!-- TODO: remaining figures to consider
--   some example temperature maps plus chopsticks
--   some example O2 maps plus chopsticks
+redbanded - pretty normal response to temp and DO (sig -ve for adults only) 
+        widow rockfish  - strong interactions
+canary - not interaction, -ve temp none for DO
+shortspine thornyhead - both temp (increasing good) and DO (increasing good)
+
+dogfish - increasing DO and warming both bad
+pacific cod - 
+English - warming neutral when cool, bad when warm
+Dover - warming good when it's cool, neutral when warm  
+(arrowtooth shows similar pattern but likely influenced strongly by fishing... could include for that reason though?)
+
+sablefish - warming good, increasing DO not
+
 -->
@@
 ```{r climate-maps, fig.cap="Maps of predicted trends (top panels) and velocities (bottom panels) of change for bottom temperature and dissolved oxygen during the trawl surveys conducted between 2008 to 2018.
-Values are estimated from a spatiotemporal models of CTD sensor data collected during late spring/early summer groundfish trawl surveys.", out.width="5in"}
+Values are estimated from a spatiotemporal models of CTD sensor data collected during late spring/early summer groundfish trawl surveys.
+Bathymetry lines at every 100 m are overlaid in shades of grey that increase with depth.", out.width="5in"}
 include_graphics(here("ms", "figs", "climate-maps-updated.pdf"))
@@
 
-\clearpage
-
-<!-- TODO: Species of interest...
-
-redbanded - pretty normal response to temp and DO (sig -ve for adults only) 
-        widow rockfish  - strong interactions
-canary - not interaction, -ve temp none for DO
-shortspine thornyhead - both temp (increasing good) and DO (increasing good)
-
-dogfish - increasing DO and warming both bad
-pacific cod - 
-English - warming neutral when cool, bad when warm
-Dover - warming good when it's cool, neutral when warm  
-(arrowtooth shows similar pattern but likely influenced strongly by fishing... could include for that reason though?)
-
-sablefish - warming good, increasing DO not
-
--->
-
 \begin{landscape}
 
-```{r temp-chops, fig.cap="Temperature ...", out.width="10in"}
+```{r temp-chops, fig.cap="Maps of biotic (top) and temperature (bottom) trends underlying the interactions (middle) where y-axis is predicted biotic trend for a given standardized change in bottom temperature (y-axis) for each .
+", out.width="10in"}
 include_graphics(here("ms", "figs", "temp-chops.pdf"))
@@
 Outliers amongst immatures are Widow Rockfish (both DO and temperature) and Bocaccio (temperature).", out.width="4in"}
-include_graphics(here("ms", "figs", "supp-slope-by-depth.pdf"))
+include_graphics(here("ms", "figs", "supp-slope-by-depth.png"))
 ```

### 2020-04-17 analysis/SOPO/2-trawl-depth-only-model.Rmd
misc sopo code that wasn't used

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    # species: "Redbanded Rockfish"
    # species: "North Pacific Spiny Dogfish"
    # species: "Pacific Ocean Perch"
    # species: "Big Skate"
    species: "Pacific Cod"
    #species: "Petrale Sole"
    #species: "Widow Rockfish"
    # species: "Bocaccio"
    # region: "All synoptic surveys"
    region: "Both odd year surveys"
    knots: 300
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    AR1: FALSE
    fixed_spatial: TRUE
    update_model: FALSE # TRUE
    update_model_check: FALSE # TRUE
    update_predictions: FALSE # TRUE
    update_index: FALSE # TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  # fig.path=paste0("figs/", spp, "/"),
  echo = FALSE, warning = FALSE, message = FALSE
)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

theme_set(
    gfplot::theme_pbs(base_size = 14) 
)
```

```{r params}
species <- params$species
region <- params$region
covariates <- params$covariates
covs <- params$covs
knots <- params$knots
priors <- FALSE

paste("region =", region)
paste("model covariates =", covariates)
paste("model label =", covs)
paste("priors =", priors)
paste("knots =", knots)
```

```{r, message=FALSE, warning=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

# Build spatiotemporal density model
Load and filter data 
```{r}
biomass <- readRDS(paste0(
  "data/", spp, "/data-by-maturity-", spp, "-1n3n4n16.rds"
))

if(nrow(biomass)<4000) {stop("Need to recalculate split by maturity!")}

# covars <- readRDS("data/event-covariates.rds")
# data <- dplyr::left_join(biomass, covars)
data <- biomass
# scale predictors before filtering to ensure mean and SD are global
data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth))
data <- gfranges::scale_predictors(data, # predictors = c(quo(depth)))
  predictors = c(quo(depth))
)
data <- data %>%
  mutate(depth = raw_depth) %>%
  filter(ssid %in% model_ssid)

```

Make mesh
```{r}
if (region == "Both odd year surveys") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 250)
}

if (region == "West Coast Vancouver Island") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
}

if (region == "West Coast Haida Gwaii") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
}

if (region == "All synoptic surveys") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
}

sdmTMB::plot_spde(spde)
```

Run climate independent sdmTMB model
```{r }
if (params$update_model) {
  tictoc::tic()
  if (any(names(data) == "adult_density")) {
    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
    ))

    adult_biomass <- sdmTMB::sdmTMB(
      data = data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = params$AR1,
      include_spatial = params$fixed_spatial,
      reml = TRUE,
      enable_priors = priors,
      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(adult_biomass,
      file = paste0(
        "data/", spp,
        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
      )
    )

   try({ imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
    ))

    imm_biomass <- sdmTMB::sdmTMB(
      data = data,
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = params$AR1, # changed to TRUE on Oct 17 2019
      reml = TRUE,
      include_spatial = params$fixed_spatial,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass,
      file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"))
   })

  } else {
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
    # dens_formula <- as.formula(paste("density ~ 0"))

    total_biomass <- sdmTMB::sdmTMB(
      data = data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = params$AR1,
      reml = TRUE,
      include_spatial = params$fixed_spatial,
      enable_priors = priors,
      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(total_biomass, file = paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
    ))
  }
  tictoc::toc()
}
```

### Check residuals
```{r}
if(params$update_model_check) {
try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
  ))
})

try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
  ))
})

point_predictions <- predict(biomass)
point_predictions$residuals <- residuals(biomass)

saveRDS(point_predictions, file = paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
))
}
```

```{r}
depth_only_predictions <- readRDS(paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
))

depth_only_predictions <- filter(
  depth_only_predictions,
  # ssid %in% c(1,3)) %>% filter (
  year > 2003
)

try({
  g <- ggplot(depth_only_predictions, aes(adult_density, residuals, colour = adult_density)) +
    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
})

try({
  g <- ggplot(depth_only_predictions, aes(density, residuals, colour = density)) +
    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
})

g
```

```{r}
try({
  g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = adult_density)) +
    geom_point() + # scale_x_continuous(trans = 'log10') +
    facet_wrap(~year)
})

try({
  g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = density)) +
    geom_point() + # scale_x_continuous(trans = 'log10') +
    facet_wrap(~year)
})

g2
```

```{r}
ggplot(depth_only_predictions, aes(X, Y, colour = (residuals))) +
  geom_point() +
  scale_colour_gradient2() +
  scale_x_continuous(trans = "log10") +
  facet_wrap(~year)
```

### Time-varying depth plots

```{r }
max_depth_found <- max(data[data$present == 1, ]$raw_depth, na.rm = TRUE)
rm(depth_model_list)
rm(adult_biomass)
rm(imm_biomass)
rm(total_biomass)
rm(depth_plots)
#max_depth_found <- 800

try({
  adult_biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
  ))
})
# adult_biomass<-sdmTMB:::update_model(adult_biomass)

try({
  imm_biomass <- readRDS(paste0("data/", spp, "/mod-imm-biomass-", 
         spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"))
  # imm_biomass<-sdmTMB:::update_model(imm_biomass)
  # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
})

if (!exists("adult_biomass")) {
  try({
    total_biomass <- readRDS(paste0(
      "data/", spp,
      "/model-total-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1,"-reml.rds"
    ))
  })
}

if (!exists("adult_biomass")) adult_biomass <- NULL
if (!exists("imm_biomass")) imm_biomass <- NULL
if (!exists("total_biomass")) total_biomass <- NULL

depth_model_list <- list(adult = adult_biomass, imm = imm_biomass, total = total_biomass)
depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]

d <- list()
depth_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  #if (depth_model_list[[i]]$model$convergence == 0) {
    d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")

    if (length(d[[i]]) == 0) {
      depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
    } else {
      d[[i]]$x <- exp(d[[i]]$x)
      depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (without environmental variables)", xlimits = c(0, max_depth_found)) +
        ggtitle(paste(species, names(depth_model_list[i])))
    }
  # } else {
  #   depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  # }
}
print(depth_plots)
```

```{r}
png(
  file = paste0(
    "figs/", spp,
    "/depth-", spp, covs, "-1n3-ar1-reml.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 6
)
gridExtra::grid.arrange(
  grobs = c(depth_plots),
  nrow = 2,
  top = grid::textGrob(paste(species, "(", covs, ")"))
)
dev.off()
```

### Model summary
```{r}
depth_model_list
```

### Plot biomass predictions

Save predictions for spatial grid
```{r}
rm(ad_predictions)
rm(im_predictions)
rm(predicted)

if(params$update_predictions) {
  # nd_all <- readRDS(paste0("data/nd_just_depth.rds"))
  # nd_all <- readRDS(paste0("data/nd_whole_coast_index.rds"))
  nd_all <- readRDS(paste0("data/nd_odd.rds"))

if (any(names(data) == "adult_density")) {

try({
  adult_biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
  ))

# adult_biomass<-sdmTMB:::update_model(adult_biomass)
    nd <- nd_all %>%
      filter(ssid %in% model_ssid) %>%
      filter(year %in% unique(adult_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    ad_predictions <- predict(adult_biomass, newdata = nd, return_tmb_object = TRUE)
    saveRDS(ad_predictions, file = paste0(
      "data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-ar1-", params$AR1, "-reml.rds"
    ))
})

  try({
    imm_biomass <- readRDS(paste0(
      "data/", spp,
      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
    ))
    nd <- nd_all %>%
      filter(ssid %in% model_ssid) %>%
      filter(year %in% unique(imm_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)
    im_predictions <- predict(imm_biomass, newdata = nd, return_tmb_object = TRUE)

    saveRDS(im_predictions, file = paste0(
      "data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-ar1-", params$AR1, "-reml.rds"
    ))
  })

} else {

  total_biomass <- readRDS(paste0(
    "data/", spp,
    "/model-total-biomass-", spp, covs, "-", ssid_string, "-ar1-", params$AR1, "-reml.rds"
  ))

  nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(total_biomass$data$year))
  nd <- na.omit(nd)
  nd$year <- as.integer(nd$year)

  predicted <- predict(total_biomass, newdata = nd, return_tmb_object = TRUE)

  saveRDS(predicted, file = paste0(
    "data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-ar1-", params$AR1,"-reml.rds"
  ))
}
}
```

Transform estimates to kg/ha
```{r}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

rm(ad_predictions)
rm(im_predictions)
rm(predicted)

try({ad_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-ar1-", params$AR1, "-reml.rds"
    ))
  })
  try({im_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-ar1-", params$AR1, "-reml.rds"
    ))
  })
  if (!exists("ad_predictions")) {
  try({
    predicted <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-ar1-", params$AR1,"-reml.rds"
    ))
  })
  }

if (exists("ad_predictions")) {
  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions <- ad_predictions$data
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000

  try({
    imm_predictions <- im_predictions$data
    imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

    adult_predictions$total_bio <- imm_predictions$est_exp +   
        adult_predictions$est_exp
    imm_predictions$prop_imm <- imm_predictions$est_exp /
        (adult_predictions$est_exp + imm_predictions$est_exp)

  saveRDS(imm_predictions, file = paste0("data/", spp,
    "/sopo-predictions-", spp, covs, "-imm-biomass-ar1-", 
    params$AR1, "-reml.rds"))
  })

  saveRDS(adult_predictions, file = paste0("data/", spp,
    "/sopo-predictions-", spp, covs, "-mat-biomass-ar1-", 
    params$AR1, "-reml.rds"))

  max_raster <- quantile(adult_predictions$est_exp, 0.999)
  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)
  max_imm <- signif(max(imm_predictions$est_exp), digits = 2)

  model_ssid <- unique(adult_predictions$ssid)
} else {

  predictions <- predicted$data
  predictions$est_exp <- exp(predictions$est) * 10000
  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .99)
  max_bio <- signif(quantile(predictions$total_bio, .999), digits = 2)

  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/sopo-predictions-", spp, covs, "-total-biomass-ar1-", params$AR1, "-reml.rds"
  ))

  #model_ssid <- unique(predictions$ssid)
}
```

```{r, error=TRUE}
legend_coords <- c(0.9, 0.17) #"none"

if (exists("adult_predictions")) {

 p_adult_all <- adult_predictions %>% 
   mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>%  
   plot_facet_map("est_exp", 
      raster_limits = c(0, max_raster),
      legend_position =  legend_coords, 
      transform_col = fourth_root_power
    ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))

 print(p_adult_all)

try({
p_imm_all <- imm_predictions %>%
  mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>%
  plot_facet_map("est_exp",
    # raster_limits = c(0, max_raster),
    legend_position =  legend_coords, 
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n(max = ", max_imm, " kg/ha)"))

print(p_imm_all)
   })

} else {

p_adult_all <- predictions  %>% 
  mutate(x = X, y = Y, X = 2 * round(X/2), Y = 2 * round(Y/2)) %>% 
  plot_facet_map("total_bio",
    raster_limits = c(0, max_raster),
    legend_position =  legend_coords, 
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))

  print(p_adult_all)
}
```

### Calculate index of total biomass for full survey grid

```{r}
if(params$update_index) {
if (exists("ad_predictions")) {
ind <- get_index(ad_predictions, bias_correct = FALSE)
ind$species <- species
ind$maturity <- "mature"
ind$ssid <- ssid_string
ind$covs <- covs

write.csv(ind, file = paste0("data/_indices/sopo-index-", 
  spp,"-", ssid_string, "-reml.csv"))

try ({
  ind_imm <- get_index(im_predictions, bias_correct = FALSE)
  ind_imm$species <- species
  ind_imm$maturity <- "immature"
  ind_imm$ssid <- ssid_string
  ind_imm$covs <- covs

  write.csv(ind_imm, file = paste0("data/_indices/sopo-index-imm-", 
    spp,"-", ssid_string, "-reml.csv"))
})
} else {
  ind <- get_index(predicted, bias_correct = FALSE)
  ind$species <- species
  ind$maturity <- "all"
  ind$ssid <- ssid_string
  ind$covs <- covs

  write.csv(ind, file = paste0("data/_indices/sopo-index-", 
    spp,"-", ssid_string, "-reml.csv"))
}
}
```

```{r}
# scale <- 2 * 2 / 1000 # if density was in kg/km2: 2 x 2 km grid and converted from kg to tonnes 
scale <- 1000000 * 4 /1000 # if density was in kg/m2: m2 to km2 * 4 km2 grid size and kg to tonnes

ind <- read.csv(paste0("data/_indices/sopo-index-", 
  spp,"-", ssid_string, "-reml.csv"))

ind %>% filter(year > 2004) %>%
ggplot(aes(year, est*scale)) + geom_line(col="darkred") +
  geom_ribbon(aes(ymin = lwr*scale, ymax = upr*scale), fill="darkred", alpha = 0.4) +
  xlab('Year') +  
  ylab('Mature biomass estimate (metric tonnes)') +
  gfplot::theme_pbs()
```

Add immature trend if available
```{r}
rm(ind_imm)

try({
  ind_imm <- read.csv(paste0("data/_indices/sopo-index-imm-", 
    spp,"-", ssid_string, "-reml.csv"))
})

if (exists("ind_imm")) {
ratio <- max(ind$upr)/max(ind_imm$upr)

ggplot(ind, aes(year, est*scale)) + geom_line(col = "darkred", ) +
  geom_ribbon(aes(ymin = lwr*scale, ymax = upr*scale), fill = "darkred", alpha = 0.4) +
# adding the relative humidity data, transformed to match roughly the range of the temperature
  geom_ribbon(aes(ymin = ind_imm$lwr*scale*ratio, ymax = ind_imm$upr*scale*ratio), 
    fill = "orangered", alpha = 0.4) + 
  geom_line(aes(ind_imm$year, ind_imm$est*scale*ratio), col = "orangered") +
# now adding the secondary axis, following the example in the help file ?scale_y_continuous and, reverting the above transformation
  scale_y_continuous(sec.axis = sec_axis(~./ratio, name = "Immature biomass estimate (metric tonnes)")) +
  xlab('Year') + 
  ylab('Mature biomass estimate (metric tonnes)')+
  gfplot::theme_pbs(base_size = 16)
}
```

### 2020-04-17 analysis/SOPO/hbll.R

@@ -1,0 +1,49 @@
+library(dplyr)
+
+### HBLL DATA
+getwd()
+setwd(here::here("/analysis/SOPO"))
+
+gfdata::get_ssids() %>% View()
+
+# gfdata::get_survey_index(222, ssid = c(22,36))
+
+gfdata::get_survey_index(222, ssid = c(22,36, 39, 40))
+
+dat <- gfdata::get_survey_sets(222, ssid = c(22))
+  
+dat <- gfdata::get_survey_sets(222, ssid = c(39, 40))
+
+saveRDS(dat, "data/hbll-inside.rds")
+
+gfdata::get_survey_ids()
+gfdata::get_other_surveys()
+# survey surveys_conducted_since_2008
+# 1        Queen Charlotte Sound Multispecies Small-mesh Bottom Trawl                            7
+# 2  West Coast Vancouver Island Multispecies Small-mesh Bottom Trawl                           12
+# 3          Strait of Georgia Ecosystem Research Initiative Acoustic                            4
+# 4                                 Sablefish Research and Assessment                            3
+# 5                                Hard Bottom Longline Inside North                             6
+# 6                                Hard Bottom Longline Inside South                             5
+# 7                                 Inlet Standardized Sablefish Trap                           11
+# 8                              Offshore Standardized Sablefish Trap                            3
+# 9                         Offshore Stratified Random Sablefish Trap                           12
+# 10                          Strait of Georgia Synoptic Bottom Trawl                            2
+# 11                                    Joint Canada/US Hake Acoustic                            9
+# 12                               Strait of Georgia Dogfish Longline                            4
+# 13                    Eulachon Migration Study Bottom Trawl (South)                            3
+# 14                    Eulachon Migration Study Bottom Trawl (North)                            2
+
+dat <- gfdata::get_survey_sets("Sablefish", ssid = c(3))
+saveRDS(dat, file = "data/sablefish3.rds")
+
+#dat <- gfdata::get_survey_sets(species_list, ssid = c(39, 40))
+
+# for (i in seq_along(species_list)){
+# dat <- gfdata::get_survey_sets(species_list[i], ssid = c(1,3))
+# }
+
+
+range(dat$year)
+range(dat2$year)
+

### 2020-04-17 ms/ms.Rmd
adding TODOs

@@
 
-Like SDMs, most previous applications of biotic velocities have focused on habitat suitability in terms of presence absence rather than using the abundance or density information (REFS).
+Like SDMs, most previous applications of biotic velocities have focused on habitat suitability in terms of presence absence rather than using the abundance or density information (TODO: REFS).
 
@@
 mature, ripe, or spent) using ogives fit as sex-specific logistic regressions to individual specimens.
-When sufficient samples were available in all years, a random effect of year was included to allow for temporal change in size at maturity (REF) and these year specific thresholds were used to split observed catches.
+When sufficient samples were available in all years, a random effect of year was included to allow for temporal change in size at maturity (TODO: REF) and these year specific thresholds were used to split observed catches.
 For those species lacking maturity data in some years, a random effect of tow id was included instead, but catches were split based on the global estimate.
@@
 We then modeled expected biomass density through space and time separately for adults and immatures of each species using spatial GLMMs [e.g., @shelton2014; @thorson2015; @thorson2016a; @ward2015].
-We model each species and maturity class separately because although these abundance patterns are certainly correlated, explicitly modeling those relationships among groundfish species has not been shown to dramatically improve precision of overall estimates (Thorson and Barnett 2017), and keeping estimates independent should make identification of shared climate responses more conservative.
-Because biomass densities contain both zeros and positive continuous values, we modeled the response with a Tweedie distribution and a log link (Anderson et al.
+We model each species and maturity class separately because although these abundance patterns are certainly correlated, explicitly modeling those relationships among groundfish species has not been shown to dramatically improve precision of overall estimates (TODO: Thorson and Barnett 2017), and keeping estimates independent should make identification of shared climate responses more conservative.
+Because biomass densities contain both zeros and positive continuous values, we modeled the response with a Tweedie distribution and a log link (TODO: Anderson et al.
 in press, Tweedie 1984, Dunn and Smyth 2005).
@@
 The inclusion of year as a factor centered each of the estimated spatiotemporal fields to have mean zero.
-Maximum likelihood estimation was conducted in R version 3.5.3 (R Core Team 2019) with the package sdmTMB (Anderson et al.
-In press, Anderson 2019) which interfaces automatic differentiation in Template Model Builder (Kristensen et al.
-2016) with INLA (Rue et al.
-2009) to .
+Maximum likelihood estimation was conducted in R version 3.5.3 (TODO: R Core Team 2019) with the package sdmTMB (Anderson et al.
+In press, Anderson 2019) which interfaces automatic differentiation in Template Model Builder (TODO: Kristensen et al.
+2016) with INLA (TODO: Rue et al.
+2009).
 
@@
 
-Spatial gradients (denominator for gradient-based calculation) are based on the years 2014-2019 (which correspond to the start of the "Blob", and corresponds with the warm transition in Alaskan waters---Albia et al.
+Spatial gradients (denominator for gradient-based calculation) are based on the years 2014-2019 (which correspond to the start of the "Blob", and corresponds with the warm transition in Alaskan waters--- TODO: Albia et al.
 2018).
@@
 In all models mean biomass density was included as a covariate in order to reduce the influence of changes occurring only at either the highest or lowest densities for a particular species.
-We attempt to predict change in biomass of each species as a function of local climate change (just temperature, or both temperature and DO) and an interaction between the mean climate of each cell and its local change estimate.
+We attempt to predict change in biomass of each species as a function of local climate change (just temperature, just DO, or both temperature and DO) and an interaction between the mean climate of each cell and its local change estimate.
 We model both biomass and climate as either velocities or as raw trends<!-- TODO: should (Fig.
@@
 Covariates capturing cell means were also centered <!-- If fishing included: (except fishing effort---so that estimates can be interpreted as in the absence of fishing)-->, but change variables were kept uncentred because an intercept of 0 change is more intuitive to interpret.
-Models exploring change as simple trends use Gaussian distributions, while velocities are  
-In addition to checking the spatial distribution of model residuals, we simulated fake biomass trends for each species (using variance and spatial correlation parameters unique to each species) and reran each model 3 <!-- TODO: simulated null models up to 10 times?
---> times and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
+Models exploring change as simple trends use Gaussian distributions, while velocities are heavy-tailed and therefore modeled using a student-t (degrees of freedom fixed at 7).
+Trend models converged with 500 knots and both temperature and DO, while velocity models would only converge with one environment variable at a time and 200 knots.
+<!-- TODO: make sure knots are explained somewhere...
+-->
+In addition to checking the spatial distribution of model residuals, we simulated fake biomass trends for each species (using variance and spatial correlation parameters unique to each species) and reran each model 3 times <!-- TODO: simulated null models 10 times?
+--> and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
 
@@
 
-Our models suggest that over the decade spanning 2008 to 2018, seafloor temperature increased across almost the entirety of Canada's Pacific shelf region (95% interquartile range of `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[2]], digits = 2)`), with warming tending to be most pronounced in shallower areas (Fig.
+Our models suggest that from 2008 to 2018, survey season seafloor temperature increased across almost the entirety of Canada's Pacific shelf region (95% interquartile range of `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[2]], digits = 2)`), with warming tending to be most pronounced in shallower areas (Fig.
 \@ref(fig:climate-maps)<!-- TODO: should letters be added and included here?

### 2020-04-19 ms/ms.Rmd
update panel plot sketch

@@
 -->
-In addition to checking the spatial distribution of model residuals, we simulated fake biomass trends for each species (using variance and spatial correlation parameters unique to each species) and reran each model 3 times <!-- TODO: simulated null models 10 times?
+In addition to checking the spatial distribution of model residuals, we simulated fake biomass trends for each species (using variance and spatial correlation parameters unique to each species) and reran each model 3 times <!-- TODO: simulate null model 10 times?
 --> and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
@@
 
-```{r temp-chops, fig.cap="Maps of biotic (top) and temperature (bottom) trends underlying the interactions (middle) where y-axis is predicted biotic trend for a given standardized change in bottom temperature (y-axis) for each .
-", out.width="10in"}
-include_graphics(here("ms", "figs", "temp-chops.pdf"))
+```{r species-map-chop-panel, fig.cap="Maps of biotic (top), temperature (middle) and DO (bottom) trends underlying interaction plots (sandwiched between maps) of predicted biotic trend (y-xis) for each standard deviation in change in seafloor temperature and DO (x-axis) for sample species representing a range of taxonomic groups and response patterns.", out.width="9in"}
+include_graphics(here("ms", "figs", "species-map-chop-panel.png"))
 ```
 
-\end{landscape}
-
-```{r do-chops, fig.cap="Dissolved oxygen ...", out.width="6in"}
-include_graphics(here("ms", "figs", "do-chops.pdf"))
-```
-
-\begin{landscape}
+\clearpage
 
@@
 Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50 percent maturity threshold) within a species.
-Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of ages recorded for immatures of each species.", out.width="9in"}
+Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of ages recorded for immatures of each species.", out.width="8in"}
 include_graphics(here("ms", "figs", "coef-scatterplots.pdf"))

### 2020-04-19 ms/make-figs.R

@@ -309,25 +309,36 @@ species_panels <- function(species, model, x_type, alpha_range = c(0.9, 0.9)) {
 }
 
 
+species_panels ("mature North Pacific Spiny Dogfish", model, "temp")
+species_panels ("mature Sablefish", model, "temp", alpha_range = c(0.25,0.9))
+species_panels ("mature Pacific Cod", model, "temp", alpha_range = c(0.25,0.9))
+
 species_panels ("mature Widow Rockfish", model, "temp")
-species_panels ("mature Widow Rockfish", model, "DO")
+species_panels ("mature Bocaccio", model, "temp") 
+species_panels ("mature Canary Rockfish", model, "temp", alpha_range = c(0.25,0.9))
+species_panels ("mature Redbanded Rockfish", model, "temp")
+species_panels ("mature Shortspine Thornyhead", model, "temp")
 
-A <- species_panels ("mature Canary Rockfish", model, "temp", temp_slopes, alpha_range = c(0.25,0.9))
-B <- species_panels ("mature Redbanded Rockfish", model, "temp")
-C <- species_panels ("mature Shortspine Thornyhead", model, "temp")
+species_panels ("mature English Sole", model, "temp")
+# species_panels ("mature Dover Sole", model, "temp")
+species_panels ("mature Flathead Sole", model, "temp")
+species_panels ("mature Arrowtooth Flounder", model, "temp")
 
-D <- species_panels ("mature English Sole", model, "temp")
-# E <- species_panels ("mature Dover Sole", model, "temp")
-E <- species_panels ("mature Flathead Sole", model, "temp")
-F <- species_panels ("mature Arrowtooth Flounder", model, "temp")
+species_panels ("mature North Pacific Spiny Dogfish", model, "DO")
 species_panels ("mature Sablefish", model, "DO")
+species_panels ("mature Pacific Cod", model, "DO")
+
+species_panels ("mature Canary Rockfish", model, "DO", alpha_range = c(0.25,0.9))
+species_panels ("mature Yelloweye Rockfish", model, "DO", alpha_range = c(0.25,0.9))
+species_panels ("mature Bocaccio", model, "DO", alpha_range = c(0.25,0.9))
+species_panels ("mature Redbanded Rockfish", model, "DO")
+species_panels ("mature Widow Rockfish", model, "DO")
 species_panels ("mature Shortspine Thornyhead", model, "DO")
[…]

### 2020-04-19 ms/ms.Rmd
refine worm plot for both trends and velocities; should both be included in results?

@@
 
-<!-- TODO: MODEL VALIDATION -->
+TODO: DESCRIBE MODEL VALIDATION FOR EACH SPECIES AND CLIMATE MODEL
 
@@
 --> and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
+These simulated biomass trends were divided by the real spatial gradients in biomass in order to simulate biotic velocities.
 
@@
 \@ref(fig:climate-maps)D) -->.
-When these local changes are placed in there geographic context in the form of a gradient-based velocity estimate, the majority of the highest velocities occur in the bottom of troughs in Queen Charlotte Sound and Dixon Entrance <!-- TODO: should letters be added and included here?
+When these local changes are placed in their geographic context in the form of a gradient-based velocity estimate, the majority of the highest velocities occur in the bottom of troughs in Queen Charlotte Sound and Dixon Entrance <!-- TODO: should letters be added and included here?
 eg.
@@
 \@ref(fig:climate-maps)E & F) -->.
+
+<!-- TODO: Figure out what to say about velocity "NULLs", if anything...
+
+This spatial pattern largely remains for simulated biotic velocities, which results in models of simulated biotic velocity that have a similar number of significant effects as the model based on real biomass changes.
+These models often tend to show more significant effects for flatfish species, possibly due to distributions that encompass a greater proportion of flatter spatial gradients for both climate and biomass (TODO see Supplementary...).
+Both climate variables and species distributions were modeled in response to seafloor depth ...
+
+-->
 
@@
 
-<!-- TODO: can a % symbol be used in fig caption? -->
+<!-- TODO: can I use a % symbol in fig captions? -->
 
-```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss for groundfish species.
+```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A, controling for effect of the other climate variable) or biotic and climate velocities (B, without controling for the effect of the other climate variable) for groundfish species.
 Open circles indicate patterns for immatures, while closed circles represent individuals large enough to have a 5 percent chance of being reproductively mature or belonging to species for which maturity data was not available.
@@
 Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).", out.width="6in"}
-include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
+# include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
+# TODO: decide if velocity slopes belong here
+# TODO: should species be in same order on both trend and velocity panels?
+include_graphics(here("ms", "figs", "worm-plot-both.pdf")) 
 ```

### 2020-04-19 ms/make-figs.R

@@ -189,22 +189,47 @@ do_vel_slopes$species[do_vel_slopes$species == "Rougheye/Blackspotted Rockfish C
 # ggsave(here::here("ms", "figs", "worm-plot-vel.pdf"), width = 8, height = 6)
 
 #### IF WE WANT PLOT OF BOTH TREND AND VELOCITY SLOPES TOGETHER
+p_temp_worm <- plot_chopstick_slopes(temp_slopes,
+  type = "temp",
+  legend_position = c(.25, .95)
+) + theme(plot.margin = margin(0, 0, 0.2, 0.2, "cm"),
+  axis.title.x = element_blank()) +
+  ggtitle("temperature") + 
+  xlab("% biomass change for a SD of climate change")
+p_do_worm <- plot_chopstick_slopes(do_slopes,
+  type = "DO",
+  legend_position = c(.25, .95)
+) + coord_flip(ylim = c(-3.1, 1.4)) +
+  ggtitle("DO") +
+  # ylab("slopes")
+  theme(plot.margin = margin(0, 0, 0.2, 0, "cm"),
+    axis.title.x = element_blank())
+
 p_temp_worm3 <- plot_chopstick_slopes(temp_vel_slopes,
   type = "temp",
   legend_position = c(.25, .95)
 ) + coord_flip(ylim = c(-14, 7.95)) +
-  theme(axis.title.x = element_blank())
+  theme(plot.margin = margin(0, 0, 0, 0.2, "cm"),
+    axis.title.x = element_blank(),
+    legend.position = "none") +
+  xlab("biotic velocity change for a SD increase in climate velocity")
 p_do_worm3 <- plot_chopstick_slopes(do_vel_slopes,
   type = "DO",
   legend_position = c(.25, .95)
 ) + coord_flip(ylim = c(-5.75, 3.25)) +
-  theme(axis.title.x = element_blank())
+  theme(plot.margin = margin(0, 0, 0, 0, "cm"),
+    axis.title.x = element_blank(),
+    legend.position = "none")
 
-((p_temp_worm | p_do_worm) / grid::textGrob("slope of biomass change with SD/decade of change in climate", just = 0.31) +
-  plot_layout(height = c(10, 0.25))) /
[…]

### 2020-04-21 ms/ms.Rmd
start describing overall results

@@
 \@ref(fig:climate-maps)E & F) -->.
+Separate geostatistical models exploring either trend- or velocity-based indices of change resolved similar overall interactions for both temperature and DO (TODO: cite supplementary materials <!--\@ref(fig:global-coefs)--> and (Fig.
+\@ref(fig:worm-plots)) if both model types included), but the overall main effects of temperature change and mean DO did differ.
+This difference is likely due to covariation between temperature and DO within our dataset <!-- TODO: run temp only trend model to check if this is true -->, which the trend model could control for, but the the velocity one could not.
+Despite this difference, the qualitative patterns for most species and maturity class tended to agree between model types (TODO: cite supplementary materials <!--\@ref(fig:trend-chopsticks), \@ref(fig:vel-chopsticks)-->);
+therefore, we will focus on interpreting the results of trend-based model (Fig. \@ref(fig:worm-plots) <!-- TODO: add A, if both models are in this plot-->). 
+Across species, the most common patterns were for 1) the impact of warming on biomass change was more likely to be negative (or at least less positive) in locations with higher mean temperatures than in cooler locations, and while controling for the effect of temperature 2) the impact of increasing DO levels  was negative in locations with lower mean DO levels than in locations with higher mean DO levels.
 
@@
 
+```{r global-coefs, fig.cap="Overall coefficient estimates from each model.", out.width="4in"}
+include_graphics(here("ms", "figs", "supp-global-coefs.pdf"))
+```
+
 \begin{landscape}

### 2020-04-21 ms/ms.Rmd
to knit

@@
 Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).", out.width="6in"}
-# include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
+include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
 # TODO: decide if velocity slopes belong here
 # TODO: should species be in same order on both trend and velocity panels?
-include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
+# include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
 ```

### 2020-04-21 ms/ms.Rmd
more results; add captions to some supplementary figs

@@
 We attempt to predict change in biomass of each species as a function of local climate change (just temperature, just DO, or both temperature and DO) and an interaction between the mean climate of each cell and its local change estimate.
-We model both biomass and climate as either velocities or as raw trends<!-- TODO: should (Fig.
-\@ref(fig:climate-maps)) be here?-->.
+We model both biomass and climate as either velocities or as raw trends<!-- TODO: should (Fig. \@ref(fig:climate-maps)) be here?-->. 
 In addition we attempt to assess the importance of the gradient component in velocity estimates by adding gradient to the model based on trends.
@@
 --> and compared the distribution of effect sizes and number of sig effects in our simulated null models with the actual estimates.
-These simulated biomass trends were divided by the real spatial gradients in biomass in order to simulate biotic velocities.
+These simulated biomass trends were divided by the real spatial gradients in biomass in order to simulate biotic velocities. <!-- TODO: where to report on NULL results? -->
 
@@
 This difference is likely due to covariation between temperature and DO within our dataset <!-- TODO: run temp only trend model to check if this is true -->, which the trend model could control for, but the the velocity one could not.
-Despite this difference, the qualitative patterns for most species and maturity class tended to agree between model types (TODO: cite supplementary materials <!--\@ref(fig:trend-chopsticks), \@ref(fig:vel-chopsticks)-->);
-therefore, we will focus on interpreting the results of trend-based model (Fig. \@ref(fig:worm-plots) <!-- TODO: add A, if both models are in this plot-->). 
-Across species, the most common patterns were for 1) the impact of warming on biomass change was more likely to be negative (or at least less positive) in locations with higher mean temperatures than in cooler locations, and while controling for the effect of temperature 2) the impact of increasing DO levels was negative in locations with lower mean DO levels than in locations with higher mean DO levels.
+Despite this difference, the qualitative patterns for most species and maturity class tended to agree between model types (TODO: cite supplementary materials <!--\@ref(fig:trend-chopsticks), \@ref(fig:vel-chopsticks)-->); therefore, we will focus on interpreting the results of trend-based model (Fig.
+\@ref(fig:worm-plots) <!-- TODO: add A, if both models are in this plot-->).
+Across species, the most common pattern was that 1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (eg.
+Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole), 2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (eg.
+Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole), and 3) overall immature abundances were more likely than mature abundances to show negative relationships (eg.
+Sablefish, Widow Rockfish, and Dover Sole (\@ref(fig:trend-chopsticks)).
+<!-- TODO: checking this with model including age as fixed effect (mature biomass as the intercept)...--> 
+There appeared to be no strong patterns of similarity amoungst species belong to the same genus or family (tested as potential random effects, but not included in the final model) as illustrated by negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
+Widow Rockfish and Flathead Sole) impacts on biotic trends being exibited by members of both the largest families represented in our analysis Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.
+\@ref(fig:species-map-chop-panel)).
 
 <!-- TODO: Figure out what to say about velocity "NULLs", if anything...
-
 This spatial pattern largely remains for simulated biotic velocities, which results in models of simulated biotic velocity that have a similar number of significant effects as the model based on real biomass changes.
@@
 Both climate variables and species distributions were modeled in response to seafloor depth ...
-
 -->
@@
 <!-- TODO:  Species of interest...
-
 redbanded - pretty normal response to temp and DO (sig -ve for adults only) 
@@
 shortspine thornyhead - both temp (increasing good) and DO (increasing good)
-
 dogfish - increasing DO and warming both bad
@@
 (arrowtooth shows similar pattern but likely influenced strongly by fishing... could include for that reason though?)
-
 sablefish - warming good, increasing DO not
-
 -->
@@
 
-```{r species-map-chop-panel, fig.cap="Maps of biotic (top), temperature (middle) and DO (bottom) trends underlying interaction plots (sandwiched between maps) of predicted biotic trend (y-xis) for each standard deviation in change in seafloor temperature and DO (x-axis) for sample species representing a range of taxonomic groups and response patterns.", out.width="9in"}
+```{r species-map-chop-panel, fig.cap="Maps of biotic (top), temperature (middle) and DO (bottom) trends underlying interaction plots (sandwiched between maps) of predicted biotic trend (y-xis) for each standard deviation in change in seafloor temperature and DO (x-axis) for mature populations of sample species representing a range of taxonomic groups and response patterns.", out.width="9in"}
 include_graphics(here("ms", "figs", "species-map-chop-panel.png"))
@@
 
-```{r global-coefs, fig.cap="Overall coefficient estimates from each model.", out.width="4in"}
+```{r global-coefs, fig.cap="Overall coefficient estimates from each model.
+The trend model included all terms, while the velocity models only include terms involving either temperature (T) variables or DO (disolved oxygen) variables, but controling for biomass (measured as estimated local mean log(biomass) across all years) in both cases.
+Interactions are betweeen the the mean value and change in the climate varible indicated.", out.width="4in"}
 include_graphics(here("ms", "figs", "supp-global-coefs.pdf"))
@@
 
-```{r trend-chopsticks, fig.cap="Caption goes here", out.width="11in"}
+```{r trend-chopsticks, fig.cap="Predicted percent change in biomass based on interactions betweeen seafloor temperature trend and mean temperature or DO trend and mean DO levels, while controling for the other interaction and mean local log(biomass).
+Mature and immature estmated slopes are represented by closed/solid versus open/dashed dots and lines respectively.
+Colours represent the highest and lowest 95th quantiles for mean climate conditions.", out.width="7.5in"}
 include_graphics(here("ms", "figs", "supp-trend-chopsticks.pdf"))
@@
 
-```{r vel-chopsticks, fig.cap="Caption goes here", out.width="11in"}
+```{r vel-chopsticks, fig.cap="Predicted interactions from separate models predicting biotic velocity based on seafloor temperature velocities and the mean temperature, or seaflour DO velocity and mean DO levels, while only controling for mean local log(biomass) in both cases.
+Colours and symbols are the same as in the previous figure.", out.width="7.5in"}
 include_graphics(here("ms", "figs", "supp-vel-chopsticks.pdf"))

### 2020-04-22 ms/ms.Rmd
current ms

@@
 \@ref(fig:climate-maps)E & F) -->.
-Separate geostatistical models exploring either trend- or velocity-based indices of change resolved similar overall interactions for both temperature and DO (TODO: cite supplementary materials <!--\@ref(fig:global-coefs)--> and (Fig.
-\@ref(fig:worm-plots)) if both model types included), but the overall main effects of temperature change and mean DO did differ.
+
+Separate geostatistical models exploring either trend- or velocity-based indices of change resolved similar overall interactions for both temperature and DO, but the overall main effects of temperature change and mean DO did differ (Fig \@ref(fig:global-coefs) TODO and \@ref(fig:worm-plots)) if both model types included).
 This difference is likely due to covariation between temperature and DO within our dataset <!-- TODO: run temp only trend model to check if this is true -->, which the trend model could control for, but the the velocity one could not.
-Despite this difference, the qualitative patterns for most species and maturity class tended to agree between model types (TODO: cite supplementary materials <!--\@ref(fig:trend-chopsticks), \@ref(fig:vel-chopsticks)-->); therefore, we will focus on interpreting the results of trend-based model (Fig.
+Despite this difference, the qualitative patterns for most species and maturity class tended to agree between model types (Figs.
+\@ref(fig:trend-chopsticks) and \@ref(fig:vel-chopsticks)); therefore, we will focus on interpreting the results of trend-based model (Fig.
 \@ref(fig:worm-plots) <!-- TODO: add A, if both models are in this plot-->).
-Across species, the most common pattern was that 1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (eg.
-Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole), 2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (eg.
+Null models based on simulated biotic trends still showed slightly more significant relationships than would be expected by chance, but far fewer than the model based on modeled changes in abundance and all the effect sizes for parameters of interest were much nearly an order of magnitude smaller (Fig.
+TODO: add supplementary fig <!-- \@ref(fig:null-coefs) -->).
+Therefore, we will explore both the overall and species-level patterns resolved, but keeping in mind that not all species-level patterns are expected to be biologically meaningful.
+
+Across species, the most common patterns were that 1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (e.g.
+Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole), 2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (e.g.
 Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole), and 3) overall immature abundances were more likely than mature abundances to show negative relationships (eg.
-Sablefish, Widow Rockfish, and Dover Sole (\@ref(fig:trend-chopsticks)).
-<!-- TODO: checking this with model including age as fixed effect (mature biomass as the intercept)...--> 
-There appeared to be no strong patterns of similarity amoungst species belong to the same genus or family (tested as potential random effects, but not included in the final model) as illustrated by negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
-Widow Rockfish and Flathead Sole) impacts on biotic trends being exibited by members of both the largest families represented in our analysis Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.
-\@ref(fig:species-map-chop-panel)).
+Sablefish, Widow Rockfish, and Dover Sole (Fig.
+\@ref(fig:trend-chopsticks)).
+<!-- TODO: checking this with model including age as fixed effect (mature biomass as the intercept)... --> 
+There were no strong patterns of similarity amoungst species belong to the same genus or family (
+TODO: should we mention here that they were included as random effects in trial models, but not the final model?
+) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
+Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.
+Fig.
+\@ref(fig:worm-plots and \@ref(fig:species-map-chop-panel)).
+Likewise, most species in all families showed relatively little impact of DO change when DO levels were high, but a somewhat counter-intuitive decline in biomass if DO increased when mean DO was at it's lowest (\@ref(fig:trend-chopsticks).
+However, several species do show increasing biomass with increasing DO across their distributions and most of these are rockfish (mature Bocaccio, Greenstripe, Redstripe, Yelloweye, Yellowmouth and Yellowtail; immature Sharpchin Rockfish; and the one exception, immature Sablefish).
+The few exceptions to these pattern include a strong positive effect of increasing DO at low mean levels for adult Splitnose Rockfish, and both maturity classes of Shortspine Thornyheads, all populations occupying some of the deeper habitats.
+However, three species averaging even deeper in their mean occurance show the more common pattern (mature Sablefish and both maturity classes of the Rougheye/Blackspotted Rockfish complex), or no relationship at all with DO change (Shortraker Rockfish).
 
@@
 -->
-
 <!-- TODO:  Species of interest...
@@
 Bathymetry lines at every 100 m are overlaid in shades of grey that increase with depth.", out.width="5in"}
-include_graphics(here("ms", "figs", "climate-maps-updated.pdf"))
+include_graphics(here("ms", "figs", "climate-maps-updated.png"))
 ```
@@
 Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).", out.width="6in"}
-include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
+# include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
 # TODO: decide if velocity slopes belong here
 # TODO: should species be in same order on both trend and velocity panels?
-# include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
+include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
 ```
@@
 
-```{r species-map-chop-panel, fig.cap="Maps of biotic (top), temperature (middle) and DO (bottom) trends underlying interaction plots (sandwiched between maps) of predicted biotic trend (y-xis) for each standard deviation in change in seafloor temperature and DO (x-axis) for mature populations of sample species representing a range of taxonomic groups and response patterns.", out.width="9in"}
+```{r species-map-chop-panel, fig.cap="Maps of biotic (top), temperature (middle) and DO (bottom) trends underlying interaction plots (sandwiched between maps) of predicted biotic trend (y-xis) for each standard deviation in change in seafloor temperature and DO (x-axis) for mature populations of species representing a range of taxonomic groups and response patterns.
+Colours and lines in the interaction plots match the slopes represented in Fig.
+2, where low (0.025) and high (0.975) quantiles of mean climate conditions are respectively illustrated in blue and red for temperatures and yellow and purple for DO.", out.width="9in"}
 include_graphics(here("ms", "figs", "species-map-chop-panel.png"))
@@
 
+\clearpage
+
 ```{r global-coefs, fig.cap="Overall coefficient estimates from each model.

### 2020-04-23 ms/ms.Rmd
more results and figure edits

@@
 
-# Introduction
+# **Introduction**
 
@@
 ***We use these concurrent estimates of both climate and biomass to assess the extent to which groundfish densities are changing in response to climate at the local scales most relevant to management***.
-In order to better understand the extent to which the temporal and spatial scales considered are appropriate for different species within this fish community, we also assess the concordance between the life histories and apparent sensitivities across species and maturity classes.
-Given the relatively short timescale encompassed in this survey, we expect *stronger patterns in shorter-lived, more pelagic species, and immatures.*
+In order to better understand the extent to which the temporal and spatial scales considered are appropriate for different species within this fish community, we also assess the concordance between the life histories and climate sensitivities across species and maturity classes.
+Given the relatively short timescale encompassed in this survey, we expect *stronger patterns in shorter-lived, more pelagic species (TODO: how do I assess this?), and immatures.*
 
-<!-- TODO: Discuss with sean & chris rooper -->
+<!-- TODO: need an index of pelagic-ness? 
+Need to add mention of growth rates?
+Do I need why?
+Discuss further with sean & chris rooper -->
 
-# Methods
+# **Methods**
 
-Example Figure reference: Fig. \@ref(fig:climate-maps).
-
-## Survey data
+## *Survey data*
 
@@
 
-## Estimate spatiotemporal variation in adult and immature biomass densities
+## *Estimate spatiotemporal variation in adult and immature biomass densities*
 
@@
 
-TODO: DESCRIBE MODEL VALIDATION FOR EACH SPECIES AND CLIMATE MODEL
+<!-- TODO: DESCRIBE MODEL VALIDATION FOR EACH SPECIES AND CLIMATE MODEL -->
 
-## Estimate climate velocities
+## *Estimate climate velocities*
 
@@
 
-<!-- TODO: Explain calculation of gradient velocity-->
+<!-- TODO: Explain calculation of gradient velocity -->
 <!-- TODO: Explain collapsing to max of 100 km ... and update code and models to do this more precisely-->
@@
 
-**Link biomass change with climate**
+## *Link biomass change with climate*
 
@@
 We attempt to predict change in biomass of each species as a function of local climate change (just temperature, just DO, or both temperature and DO) and an interaction between the mean climate of each cell and its local change estimate.
-We model both biomass and climate as either velocities or as raw trends<!-- TODO: should (Fig. \@ref(fig:climate-maps)) be here?-->. 
+We model both biomass and climate as either velocities or as raw trends <!-- TODO: should (Fig. \@ref(fig:climate-maps)) be here?-->. 
 In addition we attempt to assess the importance of the gradient component in velocity estimates by adding gradient to the model based on trends.
@@
 
-In order to better understand the extent to which the temporal and spatial scales considered are appropriate for different species within this groundfish community, we also assess the concordance between the life histories and apparent sensitivities across species and maturity classes.
+In order to better understand the extent to which the temporal and spatial scales considered are appropriate for different species within this groundfish community, we also assess the concordance between the life histories and climate sensitivities across species and maturity classes.
+Life history parameters were estimated from all individual fish measured in these same trawl surveys. 
+For all immatures and adults sampled for each species, we calculated the average depth at which they were captured, the average age estimated from otoliths or fin clips, and an approximation of immature female growth rates as the length at maturity for females divided by the 0.95 quantile of the ages of immature females.
+A similar value was calculated for immature males, but it only differed substantially for a couple of the fastest growing species and did not change any overall patterns.
 
-# Results
+# **Results**
 
 Our models suggest that from 2008 to 2018, survey season seafloor temperature increased across almost the entirety of Canada's Pacific shelf region (95% interquartile range of `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[2]], digits = 2)`), with warming tending to be most pronounced in shallower areas (Fig.
-\@ref(fig:climate-maps)<!-- TODO: should letters be added and included here?
+\@ref(fig:climate-maps) <!-- TODO: should letters be added and included here?
 eg.
@@
 
+<!-- TODO: Figure out what to say about velocity "NULLs", if anything...
+This spatial pattern largely remains for simulated biotic velocities, which results in models of simulated biotic velocity that have a similar number of significant effects as the model based on real biomass changes.
+These models often tend to show more significant effects for flatfish species, possibly due to distributions that encompass a greater proportion of flatter spatial gradients for both climate and biomass (TODO see Supplementary...).
+Both climate variables and species distributions were modeled in response to seafloor depth ...
+-->
+
 Across species, the most common patterns were that 1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (e.g.
@@
 <!-- TODO: checking this with model including age as fixed effect (mature biomass as the intercept)... --> 
-There were no strong patterns of similarity amoungst species belong to the same genus or family (
-TODO: should we mention here that they were included as random effects in trial models, but not the final model?
-) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
+There were no strong patterns of similarity amoungst species belong to the same genus or family (TODO: should we mention here that they were included as random effects in trial models, but not the final model?) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
 Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.
@@
 \@ref(fig:worm-plots and \@ref(fig:species-map-chop-panel)).
-Likewise, most species in all families showed relatively little impact of DO change when DO levels were high, but a somewhat counter-intuitive decline in biomass if DO increased when mean DO was at it's lowest (\@ref(fig:trend-chopsticks).
-However, several species do show increasing biomass with increasing DO across their distributions and most of these are rockfish (mature Bocaccio, Greenstripe, Redstripe, Yelloweye, Yellowmouth and Yellowtail; immature Sharpchin Rockfish; and the one exception, immature Sablefish).
-The few exceptions to these pattern include a strong positive effect of increasing DO at low mean levels for adult Splitnose Rockfish, and both maturity classes of Shortspine Thornyheads, all populations occupying some of the deeper habitats.
-However, three species averaging even deeper in their mean occurance show the more common pattern (mature Sablefish and both maturity classes of the Rougheye/Blackspotted Rockfish complex), or no relationship at all with DO change (Shortraker Rockfish).
+Likewise, most species in all families showed relatively little impact of DO change when DO levels were high, but a potentially counter-intuitive decline in biomass if DO increased when mean DO was at it's lowest (\@ref(fig:trend-chopsticks).
+However, several species, mostly rockfish, do show increasing biomass with increasing DO across their distributions (mature Bocaccio, Greenstripe, Redstripe, Yelloweye, Yellowmouth and Yellowtail Rockfishes, and immature Sharpchin Rockfish and Sablefish).
+The few exceptions to these pattern include a strong positive effect of increasing DO at low mean levels for adult Splitnose Rockfish, and both maturity classes of Shortspine Thornyheads, all groups occupying some of the deeper habitats (Fig.
+\@ref(fig:slope-by-depth)).
+However, three species with deeper mean occurence show the more common pattern (mature Sablefish and all of the Rougheye/Blackspotted Rockfish complex), or no relationship at all with DO change (Shortraker Rockfish).
 
-<!-- TODO: Figure out what to say about velocity "NULLs", if anything...
-This spatial pattern largely remains for simulated biotic velocities, which results in models of simulated biotic velocity that have a similar number of significant effects as the model based on real biomass changes.
-These models often tend to show more significant effects for flatfish species, possibly due to distributions that encompass a greater proportion of flatter spatial gradients for both climate and biomass (TODO see Supplementary...).
-Both climate variables and species distributions were modeled in response to seafloor depth ...
--->
+Overall climate sensitivities did correlate with life history traits and were strongest for the youngest populations (Fig. \@ref(fig:coef-scatterplots)). 
+This included overall more negative responses to both warming and increasing DO in populations with mean age less than ten years (Fig. \@ref(fig:coef-scatterplots) TODO: add panel letters?). 
+Consistent with this, the strongest negative effects of warming (estimated for the warmest parts of a species distribution) were in species occurring at intermediate depths, while most species with mean encounter depths beyond 290 m appeared to increase in abundance with greater warming (Fig.
+\@ref(fig:slope-by-depth) TODO: add panel letter A).
+In contrast, both strong negative and positive responses to changes in DO occur across a range of depth preferences, although increasing DO in the lowest DO locations is mostly negative for species averaging shallower than 150 m (Fig.
+\@ref(fig:slope-by-depth) TODO: add panel letter B).
+Finally, species with faster growth rates in immature females (cm/year) tended to have more negative responses to increases in temperature or DO (Fig.
+\@ref(fig:coef-scatterplots) TODO: add panel letters?).
+
 <!-- TODO:  Species of interest...
@@
 
-# Discussion
+# **Discussion**
 
-# Acknowledgements
+# **Acknowledgements**
 
-# References
+# **References**
 
@@
 
-# Figures
+# **Figures**
 
@@
 
-\clearpage
+\end{landscape}
+
+```{r slope-by-depth, fig.cap="Slopes of the relationship between changes in biomass and climate for the environmental extremes expected to have the greatest physiological impact (highest temperatures [A] and lowest DO [B]) plotted against each species mean depth of occurrence.
+Smoothed (solid lines represent GAM) relationship between estimated temperature (right axis and blue) and DO (left axis and purple) with mean depth in each 4x4 km grid cell (dots).
+An increase in variablity of slopes for DO change occurs for species whose mean depth is at or below depths where estimated mean DO levels approach hypoxic conditions (dashed line at <1.4 ml/L).", out.width="4in"}
+include_graphics(here("ms", "figs", "slope-by-depth.png"))
+```
+
+\begin{landscape}
 
@@
 Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50 percent maturity threshold) within a species.
-Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of ages recorded for immatures of each species.", out.width="8in"}
+Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of all otolith or fin-ray ages from immature females of each species.
+Separate regression slopes are plotted for each maturity class except in relation to mean age given that maturity is largelly a function of age.", out.width="8in"}
 include_graphics(here("ms", "figs", "coef-scatterplots.pdf"))
@@
 
-# Supporting Information {-}
+# **Supporting Information** {-}
 
@@
 
-# Supporting Methods
+# **Supporting Methods**
 
@@
 
-# Supporting Figures and Tables
+# **Supporting Figures and Tables**
 
-Supplemental figures:
-(not yet done)
+<!-- TODO: Possible additional figures:
 -   flow diagram
--   spatial residuals
 -   null model stuff
 -   pairs plot of predictors
--   plot vocc regression residuals against the main predictors... (and not in model and fitted vs. residuals plot; but don't necessarily include in paper)
+-   plot vocc regression residuals against the main predictors... 
+-   fitted vs. residuals plot
+-->
 
@@
 include_graphics(here("ms", "figs", "supp-global-coefs.pdf"))
+```
+
+```{r global-coefs-w-age, fig.cap="Overall coefficient estimates for model with mature as the intercept and two-way interactions with immature.", out.width="3in"}
+include_graphics(here("ms", "figs", "supp-global-coefs-w-age.pdf"))
 ```
@@
 
-```{r supp-slope-by-depth, fig.cap="Shape of relationship between estimated mean dissolved oxygen and mean depth of each 4x4 km grid cell (pattern is vary similar for temperature so not illustrated) for comparison with slopes for the environmental extremes experienced at the deeper depths showing expected tendancy towards increasing variation in strength of responses amongst species with deeper mean depth of occurrence.
-Outliers amongst immatures are Widow Rockfish (both DO and temperature) and Bocaccio (temperature).", out.width="4in"}
-include_graphics(here("ms", "figs", "supp-slope-by-depth.png"))
+```{r mat-residuals, fig.cap="Spatial residuals of mature biomass of each species in the trend-based model of biomass change predicted by environmental change.", out.width="7in"}
+include_graphics(here("ms", "figs", "supp-mat-residuals.pdf"))
 ```
@@
 \clearpage
+
+```{r imm-residuals, fig.cap="Spatial residuals of immature biomass in the trend-based model of biomass change predicted by environmental change.", out.width="7in"}
+include_graphics(here("ms", "figs", "supp-imm-residuals.pdf"))
+```
+
+\clearpage

### 2020-04-23 ms/ms.Rmd
# new figs out for knitting

@@
 ```{r global-coefs-w-age, fig.cap="Overall coefficient estimates for model with mature as the intercept and two-way interactions with immature.", out.width="3in"}
-include_graphics(here("ms", "figs", "supp-global-coefs-w-age.pdf"))
+# include_graphics(here("ms", "figs", "supp-global-coefs-w-age.pdf"))
 ```
@@
 ```{r mat-residuals, fig.cap="Spatial residuals of mature biomass of each species in the trend-based model of biomass change predicted by environmental change.", out.width="7in"}
-include_graphics(here("ms", "figs", "supp-mat-residuals.pdf"))
+# include_graphics(here("ms", "figs", "supp-mat-residuals.pdf"))
 ```
@@
 ```{r imm-residuals, fig.cap="Spatial residuals of immature biomass in the trend-based model of biomass change predicted by environmental change.", out.width="7in"}
-include_graphics(here("ms", "figs", "supp-imm-residuals.pdf"))
+# include_graphics(here("ms", "figs", "supp-imm-residuals.pdf"))
 ```

### 2020-04-23 ms/ms.Rmd
difference between trend and vel NOT due to covariation between temperature and DO within our dataset...

@@
 Separate geostatistical models exploring either trend- or velocity-based indices of change resolved similar overall interactions for both temperature and DO, but the overall main effects of temperature change and mean DO did differ (Fig \@ref(fig:global-coefs) TODO and \@ref(fig:worm-plots)) if both model types included).
-This difference is likely due to covariation between temperature and DO within our dataset <!-- TODO: run temp only trend model to check if this is true -->, which the trend model could control for, but the the velocity one could not.
+<!-- TODO: This difference is NOT due to covariation between temperature and DO within our dataset, which the trend model could control for, but the the velocity one could not.  ... now what? -->
 Despite this difference, the qualitative patterns for most species and maturity class tended to agree between model types (Figs.
@@
 
+```{r trend-chopsticks-temp-only, fig.cap="Trend model without DO for comparison with velocity model.", out.width="7.5in"}
+# include_graphics(here("ms", "figs", "supp-trend-chopsticks-temp-only.pdf"))
+```
+
+\clearpage
+
 ```{r supp-coef-scatterplots, fig.cap="Differences slope of effect of climate change on biotic change at how high and low mean conditions (columns) plotted against species life histories.

### 2020-04-23 ms/ms.Rmd
update supp figs exploring different model configurations (specifically gradients and maturity)

@@
 
-```{r global-coefs, fig.cap="Overall coefficient estimates from each model.
-The trend model included all terms, while the velocity models only include terms involving either temperature (T) variables or DO (disolved oxygen) variables, but controling for biomass (measured as estimated local mean log(biomass) across all years) in both cases.
-Interactions are betweeen the the mean value and change in the climate varible indicated.", out.width="4in"}
+```{r global-coefs, fig.cap="Overall coefficient estimates from a range of models.
+The 'trend' model included trends and interactions between trend and mean conditions for both climate variables, while the velocity models only converge when including terms involving either temperature (T) variables or DO (disolved oxygen) variables.
+All models control for biomass (measured as estimated local mean log(biomass) across all years).
+The 'T w gradient' model includes interactions between temperature variables and local spatial gradient in temperature in an effort to partition variation explained by some of the components of a velocity estimate.
+The model with the maturity effect has mature populations as the intercept and only clearly differs from the simplier trend model in the direction of overall estimates for mean DO.", out.width="6in"}
 include_graphics(here("ms", "figs", "supp-global-coefs.pdf"))
-```
-
-```{r global-coefs-w-age, fig.cap="Overall coefficient estimates for model with mature as the intercept and two-way interactions with immature.", out.width="3in"}
-# include_graphics(here("ms", "figs", "supp-global-coefs-w-age.pdf"))
 ```
@@
 
+```{r trend-chopsticks-temp-grad, fig.cap="Trend model without DO but with gradient for comparison with velocity model.", out.width="7.5in"}
+# include_graphics(here("ms", "figs", "supp-trend-chopsticks-temp-grad.pdf"))
+```
+
+\clearpage
+
 ```{r supp-coef-scatterplots, fig.cap="Differences slope of effect of climate change on biotic change at how high and low mean conditions (columns) plotted against species life histories.

### 2020-04-23 analysis/VOCC/6-vocc-regression.R

@@ -23,10 +23,11 @@ y_type <- "trend"
 
 no_chopsticks <- F
 
-model_type <- "-trend"
+# model_type <- "-trend"
 # model_type <- "-trend-w-age" # an experiment that lacks true chops for imm
 # model_type <- "-trend-with-do"
-# model_type <- "-trend-grad"
+model_type <- "-trend-grad"
+model_type <- "-trend-w-grad"
 
 
 # model_type <- "-vel-temp"
@@ -182,22 +183,25 @@ formula <- ~ temp_trend_scaled +
   x_type <- "trend"
 }
 
-if (model_type == "-trend-with-grad") {
+if (model_type == "-trend-w-grad") {
   
   formula <- ~ temp_trend_scaled +
     mean_temp_scaled + 
     temp_trend_scaled:mean_temp_scaled +
-    log_effort_scaled + fishing_trend_scaled +
+    # log_effort_scaled + fishing_trend_scaled +
     temp_grad_scaled +
-    DO_trend_scaled +
-    mean_DO_scaled +  
-    DO_trend_scaled:mean_DO_scaled +
+    temp_grad_scaled:mean_temp_scaled + 
+    temp_grad_scaled:temp_trend_scaled + 
+    temp_grad_scaled:mean_temp_scaled:mean_temp_scaled + 
+    # DO_trend_scaled +
+    # mean_DO_scaled +  
+    # DO_trend_scaled:mean_DO_scaled +
     log_biomass_scaled 
   
   x <- model.matrix(formula, data = d)
   
   temp_chopstick <- T
-  DO_chopstick <- T
+  DO_chopstick <- F
   x_type <- "trend"
 }
 

### 2020-04-23 ms/make-supp-figs.R

@@ -167,6 +167,8 @@ ggplot(filter(model$data, age == "mature"), aes(temp_grad, biotic_grad)) + geom_
 #########################
 #### GLOBAL COEFS 
 #########################
+
+# trend model
 coef_names <- shortener(unique(model$coefs$coefficient))
 coef_names <- c("intercept", "change in T", "mean T", "change in DO", "mean DO", 
   "biomass", "interaction (T)", "interaction (DO)")
@@ -178,6 +180,7 @@ overall_betas <- cbind.data.frame(coef_names, betas, SE, lowerCI, upperCI)
 overall_betas$model <- "trend"
 ggplot(overall_betas, aes(coef_names, betas)) + geom_pointrange(aes(ymin = lowerCI, ymax = upperCI)) + coord_flip()
 
+# trend model with temp only
 model_temp <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/trend-all-95-all-do-04-22-trend-1-500-temp.rds")
 coef_names <- shortener(unique(model_temp$coefs$coefficient))
 coef_names <- c("intercept", "change in T", "mean T", "biomass", "interaction (T)")
@@ -189,6 +192,27 @@ overall_betas_t <- cbind.data.frame(coef_names, betas, SE, lowerCI, upperCI)
 overall_betas_t$model <- "trend (T only)"
 ggplot(overall_betas_t, aes(coef_names, betas)) + geom_pointrange(aes(ymin = lowerCI, ymax = upperCI)) + coord_flip()
 
+# trend model with temp and gradient
+# no interaction
+model_grad <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/trend-all-95-all-do-04-23-trend-grad-1-500-temp.rds")
+coef_names <- shortener(unique(model_grad$coefs$coefficient))
+coef_names <- c("intercept", "change in T", "mean T", "gradient", "biomass", "interaction (T)")
+
+# version with interactions
+model_grad <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/trend-all-95-all-do-04-23-trend-w-grad-1-500.rds")
+coef_names <- shortener(unique(model_grad$coefs$coefficient))
+coef_names <- c("intercept", "change in T", "mean T", "gradient", "biomass", "interaction (T)", 
+  "mean T:gradient", "change in T:gradient", "change in T:mean:gradient")
+
[…]

### 2020-04-24 ms/ms.Rmd
edits to intro outline, methods and results

@@
 
-**Question framing notes:**
+-   **Why study species responses to climate (Ecological and management perspectives)?**
 
--   *Groundfish*
-
-    -   *occupy a relatively homogenous thermal environment*
-
-        -   *predict less tendency to move in response to climate*
-
-    -   *many migrate and school*
-
-    -   *high possibility of movement in response to climate*
-
--   *Stock assessments assume stationary distributions*
-
--   *Choices of marine protected areas*
-
--   *Approaches used often consider the whole of species ranges and apply forward projections assuming that species responses to climate variables will remain the same through time.*
+    -   Stock assessments assume stationary distributions
+    -   Choices of marine protected areas
 
@@
 
--   **Range boundaries** ...
+Approaches used often consider the whole of species ranges and apply forward projections assuming that species responses to climate variables will remain the same through time.
 
--   **Center of gravity** -- commonly used to assess responses...
+-   ***Range boundaries*** ...
+-   ***Center of gravity*** -- commonly used to assess responses...
 
@@
 
--   ... Could complicate interpretation of indices such as COG?
+    -  Could complicate interpretation of indices such as COG?
 
--   **Local climate velocities** are one way of quantifying these changes at finer spatial scales.
+-   ***Local climate velocities*** are one way of quantifying these changes at finer spatial scales.
 
@@
 
-*DO I need to talk about trends here, if trend analysis is included?*
+<!-- TODO: DO I need to talk about trends here, if trend analysis is included? -->
 
@@
 <!-- TODO: Check generation time & find REF -->
+
+- 	**Why groundfish in BC are a good case study / application**
+
+    -   occupy a relatively homogenous thermal environment
+        -   predict less tendency to move in response to climate
+    -   many migrate and school
+    -   high possibility of movement in response to climate
+
+- 	**Past work on VOCC in BC groundfish, prevailing hypotheses, limitations/unanswered questions**
 
@@
 
+<!-- TODO: Consider starting the methods with one short paragraph giving an overview of the methods -->
+
 ## *Survey data*
 
-Fisheries and Oceans Canada conducts fishery-independent bottom trawl surveys with biennial coverage of the continental shelf and upper slope of most of the BC coast.
-These surveys follow a random depth stratified design and use the same bottom trawl fishing gear and fishing protocols (TODO: Sinclair et al. 2003).
+We analyze abundance, distribution, and morphometric data for 36 species of Pacific groundfish most regularly encountered on fisheries-independent bottom trawl surveys and relatively widely distributed within Canadian waters.
 Surveys are stratified within four separate regions two of which are surveyed each year since 2005 (with surveys also conducted in some regions in 2003 & 2004): Hecate Strait and Queen Charlotte Sound (odd years) and West Coast Vancouver Island and West Coast Haida Gwaii (even years).
 Each region is sampled over the same month long period between late May and early August each year.
+Combined these surveys cover most of the upper continental slope and shelf in Canada and share the same random depth stratified design, fishing gear, and fishing protocols (TODO: Sinclair et al.
+2003).
 Only successful tows of 15--20 min duration are included in analyses and total biomass of each species captured was converted to a biomass density estimate based on the speed, distance covered and net opening.
@@
 
-We chose 36 species of Pacific groundfish most regularly encountered on fisheries-independent bottom trawl surveys and relatively widely distributed within Canadian waters.
-Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species [@mindel2016; @barbeaux2018; @li2019], we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low)
+Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species [@mindel2016; @barbeaux2018; @li2019], we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low).
 To do this, we first estimate length at 50% maturity (defined by gonadal development stages of immature or maturing vs.
@@
 
-A Seabird Electronics 19plus conductivity, temperature, and depth recorder (CTD) with dissolved oxygen (DO) sensor has also been successfully deployed on most tows since 2008.
-Therefore, we modeled bottom temperature and DO levels from CTD sensor data collected on each survey tow using a similar geostatistical model as for biomass.
-The approach for these models differed only be being Gaussian and by estimating the spatiotemporal random fields using an AR1 process.
-Because DO levels cannot be negative and are known to be influenced be both water temperature and seasonal biological processes (REF), we predict log(DO) with added fixed effects for day of year (centered and scaled) and temperature (quadratic of scaled values similar to depth).
+We modeled bottom temperature and DO levels from sensor data collected by Seabird Electronics 19plus conductivity, temperature, and depth recorder (CTD) with dissolved oxygen (DO) sensor was successfully deployed on most survey tows since 2008.
+Our modeling approach was similar to that used for biomass densities except that the distribution was Gaussian and we estimated the spatiotemporal random fields using an AR1 process.
+Because DO levels cannot be negative and are known to be influenced be both water temperature and seasonal biological processes (REF), we predicted log(DO) with fixed effects for day of year (centered and scaled) and temperature (quadratic of scaled values similar to depth).
 Bottom temperature data (but not DO) was also collected on tows prior to 2008, so this data was also included in the model in order to better inform the static spatial field component of the model.
@@
 
-Across species, the most common patterns were that 1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (e.g.
-Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole), 2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (e.g.
-Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole), and 3) overall immature abundances were more likely than mature abundances to show negative relationships (eg.
-Sablefish, Widow Rockfish, and Dover Sole (Fig.
-\@ref(fig:trend-chopsticks)).
-<!-- TODO: checking this with model including age as fixed effect (mature biomass as the intercept)... --> 
-There were no strong patterns of similarity amoungst species belong to the same genus or family (TODO: should we mention here that they were included as random effects in trial models, but not the final model?) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
+TODO: Add comment on significant negative global effects of both temperature and DO change, and negative interact for temperature.
+Maybe also add a simplier version of (Fig.
+\@ref(fig:supp-global-coefs)) to paper?
+
+Across species, the most common patterns were that 
+1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (e.g.
+Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole),
+2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (e.g.
+Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole) <!-- Note that negative effect of DO change is sig overall, but not interaction...
+so may change this?
+-->, and
+3) overall immature abundances showed a tendency to respond more positively to both mean DO and changes in DO (Fig. \@ref(fig:supp-global-coefs)). 
+There were no strong patterns of similarity amoungst species belonging to the same genus or family (TODO: should we mention here that they were included as random effects in trial models, but not the final model?) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
 Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.

### 2020-04-24 ms/ms.Rmd
Start editing

@@
   bookdown::pdf_document2:
-    number_sections: yes
+    number_sections: no
     toc: no
@@
   - \usepackage{setspace}
-  - \onehalfspacing
+#  - \onehalfspacing
   - \usepackage{lineno}
@@
   - \usepackage{pdflscape}
-  - \linenumbers
-  - \modulolinenumbers[2]
+#  - \linenumbers
+#  - \modulolinenumbers[2]
+abstract: TODO
 geometry: "left=2.64cm, right=2.64cm, top=2.54cm, bottom=2.54cm"
@@
 )
+```
 
-# TODO: is this where this should go?
+```{r data, cache=TRUE}
 alldata <- readRDS(paste0("../analysis/VOCC/data/all-do-with-null-1-untrimmed-allvars.rds"))
@@
 
-# Abstract {-}
-
-# **Introduction**
+# Introduction
 
@@
 
-# **Methods**
+# Methods
 
@@
 
-## *Survey data*
+## Survey data
 
-We analyze abundance, distribution, and morphometric data for 36 species of Pacific groundfish most regularly encountered on fisheries-independent bottom trawl surveys and relatively widely distributed within Canadian waters.
-Surveys are stratified within four separate regions two of which are surveyed each year since 2005 (with surveys also conducted in some regions in 2003 & 2004): Hecate Strait and Queen Charlotte Sound (odd years) and West Coast Vancouver Island and West Coast Haida Gwaii (even years).
-Each region is sampled over the same month long period between late May and early August each year.
-Combined these surveys cover most of the upper continental slope and shelf in Canada and share the same random depth stratified design, fishing gear, and fishing protocols (TODO: Sinclair et al.
+We analyze abundance, distribution, and morphometric data for 36 species of Pacific groundfish that are regularly encountered on fisheries-independent bottom trawl surveys and widely distributed within western Canadian waters.
+The surveys are stratified within four regions, two of which are generally surveyed each year since 2003: Hecate Strait and Queen Charlotte Sound (odd years) and West Coast Vancouver Island and West Coast Haida Gwaii (even years).
+Each region is sampled over the same month-long period between late May and early August each year.
+Combined, these surveys cover most of the upper continental slope and shelf in western Canada and share the same random depth-stratified design, fishing gear, and fishing protocols (TODO: Sinclair et al.
 2003).
-Only successful tows of 15--20 min duration are included in analyses and total biomass of each species captured was converted to a biomass density estimate based on the speed, distance covered and net opening.
-Catches of between 10--50 individuals are weighed and measured individually, while larger catches are subsampled and for commercially important species data on maturity and age are also collected.
+We only include successful tows of 15--20 min duration in analyses and convert total biomass of each species to a biomass density based on the speed, distance covered, and net opening.
+Catches of between 10--50 individuals are weighed and measured individually, while larger catches are subsampled and, for commercially important species, data on maturity and age are collected.
 
-## *Estimate spatiotemporal variation in adult and immature biomass densities*
+## Estimate spatiotemporal variation in adult and immature biomass densities
 
-Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species [@mindel2016; @barbeaux2018; @li2019], we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low).
+Because ontogenetic shifts in habitat, particularly depth, are well documented in groundfish species [e.g., @mindel2016; @barbeaux2018; @li2019], we estimate biomass densities separately for mature and immature size classes whenever possible (maturity is not assessed for certain Chondrichthians, or when catches are particularly low).
 To do this, we first estimate length at 50% maturity (defined by gonadal development stages of immature or maturing vs.
@@
 
-## *Estimate climate velocities*
+## Estimate climate velocities
 
@@
 
-## *Link biomass change with climate*
+## Link biomass change with climate
 
@@
 
-# **Results**
+# Results
 

### 2020-04-24 ms/ms.Rmd
Strip whitespace

@@
 
-We modeled bottom temperature and DO levels from sensor data collected by Seabird Electronics 19plus conductivity, temperature, and depth recorder (CTD) with dissolved oxygen (DO) sensor was successfully deployed on most survey tows since 2008.
+We modeled bottom temperature and DO levels from sensor data collected by Seabird Electronics 19, plus conductivity, temperature, and depth recorder (CTD) with dissolved oxygen (DO) sensor was successfully deployed on most survey tows since 2008.
 Our modeling approach was similar to that used for biomass densities except that the distribution was Gaussian and we estimated the spatiotemporal random fields using an AR1 process.
@@
 For each maturity class of each species the model estimates an independent random spatial field was and with and without either genus or family as possible phylogenetic grouping factors.
-.
 For each species, only the fewest cells that encompass 95% of total mean biomass were included.
@@
 Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.
-Fig.
-\@ref(fig:worm-plots and \@ref(fig:species-map-chop-panel)).
+\@ref(fig:worm-plots) and \@ref(fig:species-map-chop-panel)).
 Likewise, most species in all families showed relatively little impact of DO change when DO levels were high, but a potentially counter-intuitive decline in biomass if DO increased when mean DO was at it's lowest (\@ref(fig:trend-chopsticks).

### 2020-04-24 ms/ms.Rmd
Editing and commenting on captions

@@
 
-```{r climate-maps, fig.cap="Maps of predicted trends (top panels) and velocities (bottom panels) of change for bottom temperature and dissolved oxygen during the trawl surveys conducted between 2008 to 2018.
-Values are estimated from a spatiotemporal models of CTD sensor data collected during late spring/early summer groundfish trawl surveys.
+<!-- SA: Can we capitalize the first letter of the panel labels here and throughout? -->
+
+```{r climate-maps, fig.cap="Maps of predicted mean conditions (top row), decadal trends (middle row), and decadal velocities (bottom row) of bottom temperature and dissolved oxygen for 2008--2018.
+Values are estimated using geostatistical spatiotemporal models of CTD sensor data collected during late spring/early summer groundfish trawl surveys.
 Bathymetry lines at every 100 m are overlaid in shades of grey that increase with depth.", out.width="5in"}
@@
 <!-- TODO: can I use a % symbol in fig captions? -->
+<!-- SA: yes, need to escape it with \\ -->
 
-```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A, controling for effect of the other climate variable) or biotic and climate velocities (B, without controling for the effect of the other climate variable) for groundfish species.
-Open circles indicate patterns for immatures, while closed circles represent individuals large enough to have a 5 percent chance of being reproductively mature or belonging to species for which maturity data was not available.
-Black indicates overall slope where the estimated interaction with mean climate condition overlapped zero.
-Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).", out.width="6in"}
+<!-- SA: we may need to choose one here for the main paper (presumably trend) and include this version or just the velocity version in the supplement. -->
+
+<!-- SA: "only included when the confidence intervals on the slopes did not overlap" <- can we always include the CIs? -->
+
+```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A; controling for effect of the other climate variable) or biotic and climate velocities (B; without controling for the effect of the other climate variable) for groundfish species.
+Open circles indicate patterns for immatures, while closed circles represent individuals large enough to have a 5\\% chance of being reproductively mature or belonging to species for which maturity data was not available.
+Black indicates the overall effect when the 95\\% CI of the estimated interaction with mean climate condition overlapped zero.
+Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).", out.width="5.5in"}
 # include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
@@
 
+<!-- SA: probably need to reduce the number of examples for the main figures in the next one -->
+<!-- SA: writing out the caption like this lets us reference other figures or citations -->
+
+(ref:chop-panel) Maps of biotic (top), temperature (middle), and DO (bottom) trends underlying interaction plots (between map rows) of predicted biotic trend (y-xis) for each standard deviation change in seafloor temperature and DO (x-axis) for mature populations of species representing a range of taxonomic groups and response patterns.
+Colours and lines in the interaction plots match the slopes represented in Fig.
+\@ref(fig:worm-plots), where low (0.025) and high (0.975) quantiles of mean climate conditions are illustrated in blue and red for temperatures and yellow and purple for DO.
+
 \begin{landscape}
 
-```{r species-map-chop-panel, fig.cap="Maps of biotic (top), temperature (middle) and DO (bottom) trends underlying interaction plots (sandwiched between maps) of predicted biotic trend (y-xis) for each standard deviation in change in seafloor temperature and DO (x-axis) for mature populations of species representing a range of taxonomic groups and response patterns.
-Colours and lines in the interaction plots match the slopes represented in Fig.
-2, where low (0.025) and high (0.975) quantiles of mean climate conditions are respectively illustrated in blue and red for temperatures and yellow and purple for DO.", out.width="9in"}
+```{r species-map-chop-panel, fig.cap="(ref:chop-panel)", out.width="9in"}
 include_graphics(here("ms", "figs", "species-map-chop-panel.png"))
@@
 
-```{r slope-by-depth, fig.cap="Slopes of the relationship between changes in biomass and climate for the environmental extremes expected to have the greatest physiological impact (highest temperatures [A] and lowest DO [B]) plotted against each species mean depth of occurrence.
-Smoothed (solid lines represent GAM) relationship between estimated temperature (right axis and blue) and DO (left axis and purple) with mean depth in each 4x4 km grid cell (dots).
-An increase in variablity of slopes for DO change occurs for species whose mean depth is at or below depths where estimated mean DO levels approach hypoxic conditions (dashed line at <1.4 ml/L).", out.width="4in"}
+```{r slope-by-depth, fig.cap="Slopes of the relationship between changes in biomass and climate for the environmental extremes expected to have the greatest physiological impact---highest temperatures (A) and lowest DO (B)---plotted against each species' mean depth of occurrence.
+Smoothed (solid lines represent a GAM) relationship between estimated temperature (right axis; blue) and DO (left axis; purple) with mean depth in each 4x4 km survey grid cell (dots).
+An increase in variablity of slopes for changes in DO occurs for species whose mean depth is at or below depths where estimated mean DO levels approach hypoxic conditions (dashed line at <1.4 ml/L).", out.width="4in"}
 include_graphics(here("ms", "figs", "slope-by-depth.png"))
@@
 
+<!--
+SA: can we spell out what a negative (or positive) slope means here again to remind people?
+Negative slope means that...
+easy to lose the thread
+-->
+
 \begin{landscape}
 
-```{r coef-scatterplots, fig.cap="Main effect coefficients for relationship between biotic trends and changes in climate for species with different life history strategies.
-Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50 percent maturity threshold) within a species.
-Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of all otolith or fin-ray ages from immature females of each species.
-Separate regression slopes are plotted for each maturity class except in relation to mean age given that maturity is largelly a function of age.", out.width="8in"}
+```{r coef-scatterplots, fig.cap="Main effect coefficients of the relationship between biotic trends and changes in climate for species with different life history strategies.
+Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50\\% maturity) within a species.
+Growth rate is estimated as the length at maturity for females divided by the 0.95 quantile of all otolith or fin-ray ages from immature females of each species.
+Separate regression slopes are plotted for each maturity class except in relation to mean age given that maturity is largely a function of age.", out.width="8in"}
 include_graphics(here("ms", "figs", "coef-scatterplots.pdf"))
@@
 
+<!-- SA: I went into the code and capitalized the xlab and ylabs for this plot. Can you also do the facet labels? 
+E.g.
+firstup <- function(x){
+  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
+  x
+}
+Same for other figures (at least the main paper ones; supplement doesn't matter as much)
+-->
+
 \end{landscape}
+
+<!-- SA: I'm not sure I follow the intent of the immature growth rate calculation -->
+<!-- SA: I'm not sure I follow the intent of the immature growth rate calculation -->
 

### 2020-04-24 ms/ms.Rmd
Add more comments and edits to captions

@@
 
+<!-- SA: Am I reading it right that Bocaccio really don't like it warming?
+That's interesting considering they've had one of the biggest recruitment events ever seen for rockfish in this warming decade.
+-->
+
+<!-- SA: Point of discussion: if I'm reading it right, Pacific Cod don't like it warming. Tie into collapse in Alaska with warm blob -->
+
 ```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A; controling for effect of the other climate variable) or biotic and climate velocities (B; without controling for the effect of the other climate variable) for groundfish species.
@@
 Black indicates the overall effect when the 95\\% CI of the estimated interaction with mean climate condition overlapped zero.
-Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).", out.width="5.5in"}
+Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).
+Coefficients are ordered by TODO within each panel.", out.width="5.5in"}
 # include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
@@
 \@ref(fig:worm-plots), where low (0.025) and high (0.975) quantiles of mean climate conditions are illustrated in blue and red for temperatures and yellow and purple for DO.
+For example, for Widow Rockfish, TODO: walk through the temperature panels.
 
@@
 
-```{r slope-by-depth, fig.cap="Slopes of the relationship between changes in biomass and climate for the environmental extremes expected to have the greatest physiological impact---highest temperatures (A) and lowest DO (B)---plotted against each species' mean depth of occurrence.
-Smoothed (solid lines represent a GAM) relationship between estimated temperature (right axis; blue) and DO (left axis; purple) with mean depth in each 4x4 km survey grid cell (dots).
-An increase in variablity of slopes for changes in DO occurs for species whose mean depth is at or below depths where estimated mean DO levels approach hypoxic conditions (dashed line at <1.4 ml/L).", out.width="4in"}
+```{r slope-by-depth, fig.cap="Slopes of the relationship between changes in biomass and climate for the environmental extremes expected to have the greatest physiological impact---(A) highest temperatures and (B) lowest DO---plotted against each species' mean depth of occurrence.
+(C) Smoothed (solid lines represent a GAM) relationship between estimated temperature (right axis; blue) and DO (left axis; purple) with mean depth in each 4x4 km survey grid cell (dots).
+An increase in variablity of slopes for changes in DO occurs for species whose mean depth is at or below depths where estimated mean DO levels approach hypoxic conditions (dashed line at <1.4 ml/L in panel C).", out.width="4in"}
 include_graphics(here("ms", "figs", "slope-by-depth.png"))
@@
 easy to lose the thread
+
+and add panel letters obviously
+
+am I reading it correctly that negative slopes mean that as DO goes up, biomass goes down? Can we spell that out again?
+
+if so, does that mean that the slopes in the bottom right of panel B are particularly perplexing? near to hypoxic conditions and seemed to like it more hypoxic?
+
+I guess that points to hypoxia not being the main driver and just that these really deep species may be more O2 driven as a proxy for other things?
 -->
@@
 
+\end{landscape}
+
 <!-- SA: I went into the code and capitalized the xlab and ylabs for this plot. Can you also do the facet labels? 
@@
 
-\end{landscape}
-
-<!-- SA: I'm not sure I follow the intent of the immature growth rate calculation -->
 <!-- SA: I'm not sure I follow the intent of the immature growth rate calculation -->

### 2020-04-24 ms/ms.Rmd
Add comment and example to caption

@@
 \@ref(fig:worm-plots), where low (0.025) and high (0.975) quantiles of mean climate conditions are illustrated in blue and red for temperatures and yellow and purple for DO.
-For example, for Widow Rockfish, TODO: walk through the temperature panels.
+Black lines are shown when TODO.
+As an example, for Widow Rockfish, we estimate that biomass density has increased with increasing temperatures in cooler regions and decreased with increasing temperatures in warmer regions.
+
+<!-- SA: I almost wonder if we should show these chopstick plots before the moderately overwhelming slopes for all the species.
+Not sure, I can see an argument either way.
+Maybe at least allude in the caption for the coefficient plot that this next plot exists and may help with interpretation.
+
+The alternative would be to help the reader along by giving some examples like in this chopstick plot, and then zooming out to show the coefficients for all the species.
+-->
 

### 2020-04-24 ms/ms.Rmd
More caption editing

@@
 
-```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A; controling for effect of the other climate variable) or biotic and climate velocities (B; without controling for the effect of the other climate variable) for groundfish species.
+```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A; controlling for effect of the other climate variable) or biotic and climate velocities (B; without controlling for the effect of the other climate variable) for groundfish species.
 Open circles indicate patterns for immatures, while closed circles represent individuals large enough to have a 5\\% chance of being reproductively mature or belonging to species for which maturity data was not available.
@@
 Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).
-Coefficients are ordered by TODO within each panel.", out.width="5.5in"}
+Coefficients are ordered by TODO within each panel.", out.width="5.2in"}
 # include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
@@
 
-```{r species-map-chop-panel, fig.cap="(ref:chop-panel)", out.width="9in"}
+```{r species-map-chop-panel, fig.cap="(ref:chop-panel)", out.width="8.2in"}
 include_graphics(here("ms", "figs", "species-map-chop-panel.png"))
@@
 
-\begin{landscape}
+<!-- \begin{landscape} -->
 
@@
 Growth rate is estimated as the length at maturity for females divided by the 0.95 quantile of all otolith or fin-ray ages from immature females of each species.
-Separate regression slopes are plotted for each maturity class except in relation to mean age given that maturity is largely a function of age.", out.width="8in"}
+Separate regression slopes are plotted for each maturity class except in relation to mean age given that maturity is largely a function of age.", out.width="\\textwidth"}
 include_graphics(here("ms", "figs", "coef-scatterplots.pdf"))
@@
 
-\end{landscape}
+<!-- \end{landscape} -->
 
@@
 
-```{r global-coefs, fig.cap="Overall coefficient estimates from a range of models.
+<!-- SA: can you reduce the saved size of this figure so that the text is bigger? -->
+
+(ref:fig-global-coefs) Overall coefficient estimates from a range of models.
 The 'trend' model included trends and interactions between trend and mean conditions for both climate variables, while the velocity models only converge when including terms involving either temperature (T) variables or DO (disolved oxygen) variables.
@@
 The 'T w gradient' model includes interactions between temperature variables and local spatial gradient in temperature in an effort to partition variation explained by some of the components of a velocity estimate.
-The model with the maturity effect has mature populations as the intercept and only clearly differs from the simplier trend model in the direction of overall estimates for mean DO.", out.width="6in"}
+The model with the maturity effect has mature populations as the intercept and only clearly differs from the simplier trend model in the direction of overall estimates for mean DO.
+
+```{r global-coefs, fig.cap="(ref:fig-global-coefs)", out.width="\\textwidth"}
 include_graphics(here("ms", "figs", "supp-global-coefs.pdf"))
@@
 
-```{r trend-chopsticks, fig.cap="Predicted percent change in biomass based on interactions betweeen seafloor temperature trend and mean temperature or DO trend and mean DO levels, while controling for the other interaction and mean local log(biomass).
+```{r trend-chopsticks, fig.cap="Predicted percent change in biomass based on interactions betweeen seafloor temperature trend and mean temperature or DO trend and mean DO levels, while controlling for the other interaction and mean local log(biomass).
 Mature and immature estmated slopes are represented by closed/solid versus open/dashed dots and lines respectively.
@@
 
+\end{landscape}
+
 \clearpage
 
-```{r vel-chopsticks, fig.cap="Predicted interactions from separate models predicting biotic velocity based on seafloor temperature velocities and the mean temperature, or seaflour DO velocity and mean DO levels, while only controling for mean local log(biomass) in both cases.
-Colours and symbols are the same as in the previous figure.", out.width="7.5in"}
+(ref:fig-vel-chopsticks) Predicted interactions from separate models predicting biotic \textbf{velocity} based on seafloor temperature \textbf{velocity} and the mean temperature, or seaflour DO \textbf{velocity} and mean DO levels, while only controlling for mean local log(biomass) in both cases.
+Colours and symbols are the same as in Fig.
+\@ref(fig:trend-chopsticks).
+
+\begin{landscape}
+
+```{r vel-chopsticks, fig.cap="(ref:fig-vel-chopsticks)", out.width="7.5in"}
 include_graphics(here("ms", "figs", "supp-vel-chopsticks.pdf"))
@@
 Simple regression lines plotted for immature abundance only.
-Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50 percent maturity threshold) within a species.
+Mean age and depth are calculated across all sampled individuals of each maturity class (as defined by the length at 50\\% maturity threshold) within a species.
 Growth rate is estimated as the length at maturity for females divide by the 0.95 quantile of ages recorded for immatures of each species.", out.width="8in"}

### 2020-04-24 ms/ms.Rmd
caption edits and responses

--- 2020-04-24 ms/ms.Rmd
@@
 
-<!-- TODO: can I use a % symbol in fig captions? -->
-<!-- SA: yes, need to escape it with \\ -->
+<!-- SA: "only included when the confidence intervals on the slopes did not overlap" <- can we always include the CIs? -->
 
-<!-- SA: we may need to choose one here for the main paper (presumably trend) and include this version or just the velocity version in the supplement. -->
-
-<!-- SA: "only included when the confidence intervals on the slopes did not overlap" <- can we always include the CIs? -->
+<!-- PE: The line you quote was meant to convey something different, so I've edited it to make its intent clearer.
+The CI have always been included, so those that don't show up are too small to extend outside the dots...
+-->
 
@@
 
-```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A; controling for effect of the other climate variable) or biotic and climate velocities (B; without controling for the effect of the other climate variable) for groundfish species.
+```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss controling for effect of the other climate variable for groundfish species.
 Open circles indicate patterns for immatures, while closed circles represent individuals large enough to have a 5\\% chance of being reproductively mature or belonging to species for which maturity data was not available.
 Black indicates the overall effect when the 95\\% CI of the estimated interaction with mean climate condition overlapped zero.
-Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).
-Coefficients are ordered by TODO within each panel.", out.width="5.5in"}
-# include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
-# TODO: decide if velocity slopes belong here
+Colours in legends indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions.
+Black indicates global slopes when the confidence intervals of the low and high slopes were found to overlap.
+Coefficients are ordered by mean slope for each species within each panel.", out.width="5.5in"}
+include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
+# TODO: decide if velocity slopes belong here, removed them for now (04-23)
 # TODO: should species be in same order on both trend and velocity panels?
-include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
+# include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
 ```
@@
 -->
+<!-- PE: I like the idea of putting the chops first... I'll see how that works... -->
 
@@
 <!-- SA: I went into the code and capitalized the xlab and ylabs for this plot. Can you also do the facet labels? 
-E.g.
-firstup <- function(x){
-  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
-  x
-}
 Same for other figures (at least the main paper ones; supplement doesn't matter as much)
 -->
+
+<!-- PE: Titles in main figures have capitals now.  Should legend groups like "high temp" and "immature" be capitals too? I think not, but... -->
 

### 2020-04-25 ms/ms.Rmd
Some results copyediting

--- 2020-04-24 ms/ms.Rmd
@@
 )
+options(tinytex.clean = FALSE) # keep TeX auxiliary files for faster compiling
 ```
@@
 
-Our models suggest that from 2008 to 2018, survey season seafloor temperature increased across almost the entirety of Canada's Pacific shelf region (95% interquartile range of `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[2]], digits = 2)`), with warming tending to be most pronounced in shallower areas (Fig.
-\@ref(fig:climate-maps) <!-- TODO: should letters be added and included here?
-eg.
-"C" -->).
-In contrast, change in seafloor DO was much more variable in the direction of change (95% interquartile range of `r signif(quantile(alldata$DO_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$DO_trend, c(0.025, 0.975))[[2]], digits = 2)`), with both the greatest increases also occurring in the shallowest areas and the greatest decreases occuring across a wide range of depths <!-- TODO: should letters be added and included here?
-eg.
-(Fig.
-\@ref(fig:climate-maps)D) -->.
-When these local changes are placed in their geographic context in the form of a gradient-based velocity estimate, the majority of the highest velocities occur in the bottom of troughs in Queen Charlotte Sound and Dixon Entrance <!-- TODO: should letters be added and included here?
-eg.
-(Fig.
-\@ref(fig:climate-maps)E & F) -->.
+Our models suggest that from 2008 to 2018, seafloor temperature from late May to early August increased across almost the entirety of Canada's Pacific shelf region (95% interquantile range of `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$temp_trend, c(0.025, 0.975))[[2]], digits = 2)`), with warming tending to be most pronounced in shallower areas (Fig.
+\@ref(fig:climate-maps)C).
+<!-- TODO: should letters be added and included here? eg. "C" -->
+<!-- SA: yeah, might be helpful for referring to specific panels here-->
+In contrast, change in seafloor DO was much more variable in the direction of change (95% interquantile range of `r signif(quantile(alldata$DO_trend, c(0.025, 0.975))[[1]], digits = 2)` to `r signif(quantile(alldata$DO_trend, c(0.025, 0.975))[[2]], digits = 2)`), with the greatest increases also occurring in the shallowest areas and the greatest decreases occuring across a wide range of depths (Fig.
+\@ref(fig:climate-maps)D).
+When these local changes are placed in their geographic context in the form of a gradient-based velocity estimate, the majority of the highest velocities occur in the bottom of troughs in Queen Charlotte Sound and Dixon Entrance (Fig.
+\@ref(fig:climate-maps)E, F).
 
-Separate geostatistical models exploring either trend- or velocity-based indices of change resolved similar overall interactions for both temperature and DO, but the overall main effects of temperature change and mean DO did differ (Fig \@ref(fig:global-coefs) TODO and \@ref(fig:worm-plots)) if both model types included).
+Geostatistical models exploring either trend- or velocity-based indices of change resolved similar overall interactions for both temperature and DO, but the overall main effects of temperature change and mean DO differed (Fig \@ref(fig:global-coefs) TODO and \@ref(fig:worm-plots)) if both model types included).
 <!-- TODO: This difference is NOT due to covariation between temperature and DO within our dataset, which the trend model could control for, but the the velocity one could not.  ... now what? -->
 Despite this difference, the qualitative patterns for most species and maturity class tended to agree between model types (Figs.
-\@ref(fig:trend-chopsticks) and \@ref(fig:vel-chopsticks)); therefore, we will focus on interpreting the results of trend-based model (Fig.
+\@ref(fig:trend-chopsticks) and \@ref(fig:vel-chopsticks)); therefore, we focus on interpreting the results of the trend-based model (Fig.
 \@ref(fig:worm-plots) <!-- TODO: add A, if both models are in this plot-->).
-Null models based on simulated biotic trends still showed slightly more significant relationships than would be expected by chance, but far fewer than the model based on modeled changes in abundance and all the effect sizes for parameters of interest were much nearly an order of magnitude smaller (Fig.
-TODO: add supplementary fig <!-- \@ref(fig:null-coefs) -->).
-Therefore, we will explore both the overall and species-level patterns resolved, but keeping in mind that not all species-level patterns are expected to be biologically meaningful.
+Null models based on simulated biotic trends showed slightly more significant relationships than would be expected by chance, but far fewer than the model based on modeled changes in abundance
+<!-- SA: what is this "modeled changes in abundance"? you mean density? even still I'm not sure -->
+and all the effect sizes for parameters of interest were much nearly an order of magnitude smaller (Fig. TODO: add supplementary fig).
+<!-- \@ref(fig:null-coefs) -->
+Therefore, we explore both the overall and species-level patterns, but keep in mind that a slightly higher percentage than the stated alpha level of 0.05 (TODO %) may be spurious.
 
@@
 \@ref(fig:supp-global-coefs)) to paper?
+<!-- SA: I think it's fine to leave the gnarly complicated thing in the supplement; maybe we can just add the global effect to the overall coefficient plot at the bottom?
+maybe emphasized with something fancy like a diamond?-->
 
+<!-- SA: I wonder if each of these main findings should be split off into its own paragraph with the point as a topic sentence? This is a super dense results paragraph -->
 Across species, the most common patterns were that
-1) the impact of warming on biomass change was more negative (or at least less positive) in locations with higher mean temperatures than in cooler locations (e.g.
-Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole),
-2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (e.g.
-Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole) <!-- Note that negative effect of DO change is sig overall, but not interaction...
-so may change this?
--->, and
-3) overall immature abundances showed a tendency to respond more positively to both mean DO and changes in DO (Fig. \@ref(fig:supp-global-coefs)).
-There were no strong patterns of similarity amoungst species belonging to the same genus or family (TODO: should we mention here that they were included as random effects in trial models, but not the final model?) as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (eg.
-Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.
+(1) the impact of warming on biomass change was more negative (or less positive) in locations with higher mean temperatures than in cooler locations (e.g., Widow Rockfish, mature Shortspine Thornyhead, and mature English and Flathead Sole),
+(2) the impact of increasing DO levels was more negative in locations with lower mean DO levels than in locations with higher mean DO levels (e.g., Dogfish, mature Sablefish, Widow Rockfish, and English and Flathead Sole),
+<!-- Note that negative effect of DO change is sig overall, but not interaction... so may change this? -->
+and (3) overall immature abundances showed a tendency to respond more positively to both mean DO and changes in DO (Fig. \@ref(fig:supp-global-coefs)).
+There were no strong patterns of similarity amoungst species belonging to the same genus or family
+<!-- (TODO: should we mention here that they were included as random effects in trial models, but not the final model?)  -->
+<!-- SA: yeah, we should probably mention it in the methods -->
+as illustrated by relationships with biotic trends that were negative at high temperatures (Canary Rockfish and English Sole) or positive at low temperatures (e.g., Widow Rockfish and Flathead Sole) occuring in members of both the largest families represented in our analysis, Sebastidae (rockfish) and Pleuronectidae (righteye flounders; Fig.
 \@ref(fig:worm-plots) and \@ref(fig:species-map-chop-panel)).
+<!-- SA: it occurs to me that we need to get the latin names and for all the species and that will make the text get kind of ridiculous.
+I wonder if we should have a table with all the species in latin names and just refer to that?
+-->
 Likewise, most species in all families showed relatively little impact of DO change when DO levels were high, but a potentially counter-intuitive decline in biomass if DO increased when mean DO was at it's lowest (\@ref(fig:trend-chopsticks).
@@
 
-Overall climate sensitivities did correlate with life history traits and were strongest for the youngest populations (Fig. \@ref(fig:coef-scatterplots)).
+Overall climate sensitivities correlated with life-history traits and were strongest for the youngest populations (Fig. \@ref(fig:coef-scatterplots)).
 This included overall more negative responses to both warming and increasing DO in populations with mean age less than ten years (Fig. \@ref(fig:coef-scatterplots) TODO: add panel letters?).
-Consistent with this, the strongest negative effects of warming (estimated for the warmest parts of a species distribution) were in species occurring at intermediate depths, while most species with mean encounter depths beyond 290 m appeared to increase in abundance with greater warming (Fig.
+Consistent with this finding, the strongest negative effects of warming (estimated for the warmest parts of a species distribution) were for species occurring at intermediate depths, whereas most species with mean encounter depths deeper than 290 m appeared to increase in density with greater warming (Fig.
 \@ref(fig:slope-by-depth) TODO: add panel letter A).
-In contrast, both strong negative and positive responses to changes in DO occur across a range of depth preferences, although increasing DO in the lowest DO locations is mostly negative for species averaging shallower than 150 m (Fig.
+In contrast, both strong negative and positive responses to changes in DO occurred across a range of depth preferences, although increasing DO in the lowest DO locations is mostly negative for species averaging shallower than 150 m (Fig.
 \@ref(fig:slope-by-depth) TODO: add panel letter B).

### 2020-04-25 ms/ms.Rmd

--- 2020-04-24 ms/ms.Rmd
@@
 
-```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss (A; controlling for effect of the other climate variable) or biotic and climate velocities (B; without controlling for the effect of the other climate variable) for groundfish species.
+```{r worm-plots, fig.cap="Slopes of relationships between climate trends and percent change in biomasss controling for effect of the other climate variable for groundfish species.
 Open circles indicate patterns for immatures, while closed circles represent individuals large enough to have a 5\\% chance of being reproductively mature or belonging to species for which maturity data was not available.
 Black indicates the overall effect when the 95\\% CI of the estimated interaction with mean climate condition overlapped zero.
-Colours indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions (only included when the confidence intervals on the slopes did not overlap).
-Coefficients are ordered by TODO within each panel.", out.width="5.2in"}
-# include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
-# TODO: decide if velocity slopes belong here
+Colours in legends indicate slopes estimated for the low (0.025) and high (0.975) quantiles of mean climate conditions.
+Coefficients are ordered by mean slope for each species within each panel.", out.width="5.2in"}
+include_graphics(here("ms", "figs", "worm-plot-trend.pdf"))
+# TODO: decide if velocity slopes belong here, removed them for now (04-23)
 # TODO: should species be in same order on both trend and velocity panels?
-include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
+# include_graphics(here("ms", "figs", "worm-plot-both.pdf"))
 ```

### 2020-05-15 analysis/VOCC/3-vocc-dist.Rmd
tidy up units on velocities and trends in vocc-dist output

@@
   mutate(
-    trend_per_year = trend,
+    ## these are only relevant if divisor = 1 
+    # trend_per_year = trend,
+    # trend_per_decade = trend_per_year * 10,
+    trend_per_year = trend/10,
     trend_per_decade = trend_per_year * 10,
@@
 
-dvocc_t <- dvocc_t %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend=slope)
+dvocc_t <- dvocc_t %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend = slope)
 dvocc_t <- dvocc_t %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
@@
   mutate(
-    trend_per_year = trend,
+        ## these are only relevant if divisor = 1 
+    # trend_per_year = trend,
+    # trend_per_decade = trend_per_year * 10,
+    trend_per_year = trend/10,
     trend_per_decade = trend_per_year * 10,
@@
 # glimpse(dvocc_d)
-dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend= slope)
+dvocc_d <- dvocc_d %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend = slope)
 
@@
 # glimpse(dvocc_d)
-dvocc_td <- dvocc_td %>% mutate(dvocc = if_else(do_est.slope < 0, -km_per_decade, km_per_decade), trend= do_est.slope)
+dvocc_td <- dvocc_td %>% mutate(dvocc = if_else(do_est.slope < 0, -km_per_decade, km_per_decade), trend = do_est.slope)
 
@@
     mean_biomass = exp(mean) * 10000,
-    # trend_per_year = exp(trend),
+    ## these are only relevant if divisor = 1 
+      # trend_per_year = trend,
       # trend_per_decade = trend_per_year * 10,
+    ## if default of divisor = 10
+      trend_per_year = exp(trend/10),
+      trend_per_decade = trend_per_year * 10,
     velocity = velocity

### 2020-05-15 analysis/VOCC/1-split-biomass-by-maturity.Rmd
add option for sample_id_re to maturity splits with year effect

@@
     year_re = TRUE,
+    # sample_id_re = TRUE,
     cutoff_quantile = c(.9995),

### 2020-05-17 analysis/VOCC/1-split-biomass-by-maturity.Rmd
add option to use median instead of mean mass ratios

--- 2020-02-19 analysis/VOCC/1-split-biomass-by-maturity.Rmd
@@
     year_re = TRUE,
+    # sample_id_re = TRUE,
     cutoff_quantile = c(.9995),
     p_threshold = threshold,
+    # use_median_ratio = TRUE,
     plot = TRUE)
@@
     year_re = TRUE,
+    # sample_id_re = TRUE,
     cutoff_quantile = c(.9995),
+    # use_median_ratio = TRUE,
     p_threshold = threshold,

### 2020-05-17 R/mature-biomass-est.R

@@ -28,6 +28,7 @@ split_catch_maturity <- function(survey_sets, fish, bath,
                                  sample_id_re = FALSE, 
                                  cutoff_quantile = 0.9995,
                                  p_threshold = 0.5,
+                                 use_median_ratio = FALSE,
                                  plot = FALSE) {
 
   if (is.null(years)) years <- unique(survey_sets[["year"]])
@@ -203,7 +204,12 @@ split_catch_maturity <- function(survey_sets, fish, bath,
 
 # browser()
     # chose value to use when a sample mass ratio is not available
-    na_value <- mean(sets_w_ratio$mass_ratio_mature, na.rm = TRUE)
+    if (use_median_ratio) {
+      na_value <- median(sets_w_ratio$mass_ratio_mature, na.rm = TRUE)
+      } else {
+      na_value <- mean(sets_w_ratio$mass_ratio_mature, na.rm = TRUE)
+      }
+    
     sets_w_ratio$mass_ratio_mature[is.na(sets_w_ratio$mass_ratio_mature)] <- na_value
 
     # add column to check for discrepencies between total catch weight and biological sample weights

### 2020-06-18 analysis/VOCC/1-split-biomass-by-maturity.Rmd
change Bocaccio to a global maturity estimate

@@
 params: 
-    #species: "Bocaccio"
-    species: "Threadfin Sculpin"
+    species: "Bocaccio"
+    # species: "Threadfin Sculpin"
     #species: "Quillback Rockfish"
@@
     # region: "Both odd year surveys"
-    threshold: 0.05
-    split: FALSE
+    threshold: 0.5
+    split: TRUE
 ---

### 2020-06-18 analysis/VOCC/2-biomass-depth-only-model.Rmd
change Bocaccio to a global maturity estimate

@@
 params: 
-    #species: "Redbanded Rockfish"
+    # species: "Redbanded Rockfish"
     #species: "Pacific Ocean Perch"
@@
     #species: "Petrale Sole"
-    #species: "Widow Rockfish"
+    # species: "Widow Rockfish"
     species: "Bocaccio"
+    # species: "Sablefish"
     region: "All synoptic surveys"
@@
 
+```{r}
+fits <- readRDS("climate-fits.rds")
+check_fits <- left_join(depth_only_predictions,fits)
+
+try({
+  g3 <- ggplot(check_fits, aes(t_resids, residuals, colour = density)) +
+    geom_point() + # scale_x_continuous(trans = 'log10') +
+    facet_wrap(~year)
+  g4 <- ggplot(check_fits, aes(d_resids, residuals, colour = density)) +
+    geom_point() + # scale_x_continuous(trans = 'log10') +
+    facet_wrap(~year)
+})
+try({
+  g3 <- ggplot(check_fits, aes(t_resids, residuals, colour = adult_density)) +
+    geom_point() + # scale_x_continuous(trans = 'log10') +
+    facet_wrap(~year)
+  g4 <- ggplot(check_fits, aes(d_resids, residuals, colour = adult_density)) +
+    geom_point() + # scale_x_continuous(trans = 'log10') +
+    facet_wrap(~year)
+})
+g3
+g4
+```
+
 ### Build time-varying depth plots

### 2020-06-18 analysis/VOCC/2-render-biomass-model.R

@@ -78,7 +78,10 @@ list_species <- c(
    "Longnose Skate",
    "Spotted Ratfish"
 )
- 
+
+list_species <- c(
+  "Bocaccio"
+)
 
 ### build time-varying depth models
 list_regions <- c("All synoptic surveys")
@@ -95,12 +98,12 @@ for (r_h in seq_along(list_regions)) {
           region = list_regions[r_h],
           covariates = "", # additional non-climate variables
           covs = covs,
-          knots = 400,
+          knots = 500,
           update_model = TRUE # FALSE #
         ),
         output_file = paste0(
           "html/biomass-by-depth/biomass-by",
-          covs, "-", spp, "-400.html"
+          covs, "-", spp, "-500.html"
         ),
         envir = env
       )

### 2020-06-18 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
rerun climate models with current sdmTMB coding

@@
 
-plan(multiprocess)
+# plan(multiprocess)
 ```
@@
 ```{r}
-ggplot(dplyr::filter(d_trawl1, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
+ggplot(dplyr::filter(d_trawl, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
 ```
@@
 
-```{r eval = FALSE}
+```{r, eval = FALSE}
 library(gbm)
@@
 ## Just compare fixed slope models with simplest random effect structure
-```{r eval=F}
+```{r, eval=F}
 cv_args <- generate_cv_args(
@@
 All are better than fixed slope models
-```{r}
+```{r, eval=FALSE}
 cv_args <- generate_cv_args(
@@
 AR1 doesn't make much difference (differences between years appear independent)
-```{r}
+```{r, eval=FALSE}
 cv_args <- generate_cv_args(
@@
 ## Change knot number
-
-```{r}
+```{r, eval=FALSE}
 cv_args <- generate_cv_args(
@@
 Models without SST_scaled are usually better
-
-```{r}
+```{r, eval=FALSE}
 cv_args <- generate_cv_args(
@@
 Current top model
-
 ```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 400)
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 500)
 plot_spde(spde)
@@
 ```{r}
-m_temp_rw <- sdmTMB(d_trawl,
+m_temp_rw <- sdmTMB(
   temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  data = d_trawl,
   ar1_fields = TRUE, # changed AR1_field to true
@@
 
-saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn.rds")))
+saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
 
@@
 predict_temp <- predict(m_temp_rw, newdata = nd_all)
-saveRDS(predict_temp, file = "../VOCC/data/predicted_temp_allyears_revised")
+saveRDS(predict_temp, file = "../VOCC/data/predicted_temp_allyears_revised2")
 p_nd <- predict_temp
@@
 ```{r}
-p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised")
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised.rds")
 # p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted_temp_allyears_revised.rds")
+p_nd2 <- readRDS("../VOCC/data/predicted_temp_allyears_revised2")
+
+plot(p_nd$est ~ p_nd2$est)
 
@@
     #breaks = c(0,2,4,6,8), 
-    limits = c(5, 11), 
+    limits = c(4, 11),
     na.value = "red") +
@@
   labs(fill = "°C") + 
-  theme(legend.position = "none") +
+  # theme(legend.position = "none") +
   ggtitle("Prediction (fixed effects + all random effects)") 

### 2020-06-18 analysis/tmb-sensor-explore/3-fit-do.Rmd
rerun climate models with current sdmTMB coding

@@
 #test <- d_trawl1 %>% filter(year == 2016 & ssid ==16) 
-spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 800)
+spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 800) ### TODO: rerun with 500
 plot_spde(spde)
@@
 Add temperature predictions to newdata grid for DO predictions
-```{r}
-temperature <- readRDS(here::here("analysis/VOCC/data/predicted_temp_allyears.rds"))
+```{r eval=FALSE}
+temperature <- readRDS(here::here("analysis/VOCC/data/predicted_temp_allyears_revised2"))
 years <- unique(d_trawl1$year) 
@@
 
-    do_model_all_DOY <- sdmTMB::sdmTMB(d_trawl1,
+    do_model_all_DOY <- sdmTMB::sdmTMB(
       log_do ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + DOY_scaled,
+      data = d_trawl1,
       time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
-      time = "year", spde = spde,
+      time = "year", 
+      spde = spde,
       family = gaussian(link = "identity"),
@@
 Other models tried:
+```{r eval=FALSE}
+do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds"))
+do_model_DOY
+#p_nd <- predict(do_model_DOY, newdata = newdata)
+get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")
+```
 
@@
 ```{r eval=FALSE}
-p_nd <- predict(do_model_DOY, newdata = newdata)
+do_model_all <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-poly.rds"))
+p_nd <- predict(do_model_all_DOY, newdata = newdata)
 p_nd$do_est <- exp(p_nd$est)
@@
 
-model1 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-temp-2008-on-used.rds")
+# model1 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-temp-2008-on-used.rds")
 #model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-all-DOY-used.rds")
-#update model to exclude WCVI 2016 outliers
-model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds")
 
-nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-nd <- filter(nd_all, year >2007) %>% filter(year<2019)
-pred_temp <- predict(model1, newdata = nd)
-predtemp <- pred_temp %>% mutate(temp_scaled = est, temp_scaled2 = est^2, depth_scaled3 = depth_scaled^3) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
+# fix RW additon error
+model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds"))
+# update model to exclude WCVI 2016 outliers
+model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds"))
+
+nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
+
+pred_temp <- predict(model1, newdata = nd_all)
+
+# get scaling used in model
+d <- all_sensor %>%
+  # filter missing do data 
+  dplyr::filter(!is.na(do_mlpl)) %>%
+  # remove 2007 data because clearly faulty or scaled differently
+  dplyr::filter(year>2007)
+d$temp_scaled <- scale(d$temperature_c)
+attributes(d$temp_scaled)
+
+predtemp <- pred_temp %>% mutate(temp_scaled = (est - 6.49 )/1.296, temp_scaled2 = ((est - 6.49 )/1.296)^2, depth_scaled3 = depth_scaled^3) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
+predtemp <- predtemp %>% filter(year > 2007) 
 predtemp$DOY_scaled <- 0
+
 pred_do <- predict(model2, newdata = predtemp)
+
 pred_do$temp <- pred_do$temp_scaled*1.296 + 6.49
@@
 
-saveRDS(pred_do, file="~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-new.rds")
+saveRDS(pred_do, file=here::here("analysis/VOCC/data/predicted-DO-june-2020.rds"))
 
@@
 #p_nd <- readRDS((here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
-p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-new.rds")))
+# p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-new.rds")))
+
+p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-june-2020.rds")))
 
@@
 pred_1 <- plot_map_raster(p_nd1, "do_est") +
-  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 9), na.value = "red") +
+  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), 
+    limits = c(1, 8), 
+    na.value = "red") +
   facet_wrap(~year) +
@@
 pred_4 <- plot_map_raster(p_nd4, "do_est") +
-  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 9), na.value = "red")+
+  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(1, 7), 
+     na.value = "red")+
   facet_wrap(~year) +

### 2020-06-18 analysis/VOCC/3-vocc-dist.Rmd
run vocc with updated climate est

@@
 ```{r}
-nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select(X, Y, year, temp, do_est)
+nd_all <- readRDS(paste0("data/predicted-DO-june-2020.rds")) %>% select(X, Y, year, temp, do_est)
 fishing <- readRDS("data/_fishing_effort/fishing-effort-grid.rds")
@@
 effort <- plot_vocc(gf,
+  raster_cell_size = 4,
   # theme_black = TRUE,
@@
 fish_t <- plot_vocc(gf,
+  raster_cell_size = 4,
   # theme_black = TRUE,
@@
   # theme = "black",
+  raster_cell_size = 4,
   coast = TRUE,
@@
 grad2 <- plot_vocc(gf2,
+  raster_cell_size = 4,
   # theme = "black",
@@
 grad0dv <- plot_vocc(gf0,
+  raster_cell_size = 4,
   # theme = "black",
@@
 grad0v <- plot_vocc(gf0,
+  raster_cell_size = 4,
   # theme_black = TRUE,
@@
 grad0 <- plot_vocc(gf0,
+  raster_cell_size = 4,
   # theme_black = TRUE,
@@
 gradbv <- plot_vocc(gf3t,
+  raster_cell_size = 4,
   # theme = "black",
@@
 grad3 <- gfranges::plot_vocc(gf3t,
+  raster_cell_size = 4,
   # theme = "black",
@@
 grad4 <- gfranges::plot_vocc(gf3,
+  raster_cell_size = 4,
   # theme = "black",
@@
 grad5 <- gfranges::plot_vocc(gf3t,
+  raster_cell_size = 4,
   # theme = "black",
@@
 grad6 <- gfranges::plot_vocc(gf3t,
+  raster_cell_size = 4,
   # theme = "black",
@@
 grad7 <- gfranges::plot_vocc(gf3t,
+  raster_cell_size = 4,
   # theme = "black",
@@
 write.csv(vocc, file = paste0(
-  "data/_dvocc_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_newclim_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))
@@
 vocc  <- read.csv(paste0(
-  "data/_dvocc_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, "-", covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_newclim_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))

### 2020-06-18 analysis/VOCC/3-render-biotic-vocc.R

@@ -59,7 +59,7 @@ list_regions <- c(
   #   "All synoptic surveys"
 )
 
-# age <- "mature"
+age <- "mature"
 if (age == "mature") {
 for (r_h in seq_along(list_regions)) {
   for (spp_i in seq_along(list_species)) {
@@ -78,7 +78,7 @@ for (r_h in seq_along(list_regions)) {
         output_file = paste0(
           "html/VOCC-plots/vocc-w-do-", spp,
           # "-imm",
-          covs, "-", reg, "-dvocc-decade.html"
+          covs, "-", reg, "-updated-climate.html"
         ),
         envir = env
       )
@@ -87,7 +87,7 @@ for (r_h in seq_along(list_regions)) {
 }
 }
 
-# age <- "immature"
+age <- "immature"
 if (age == "immature") {
 for (r_h in seq_along(list_regions)) {
   for (spp_i in seq_along(list_species)) {
@@ -105,7 +105,7 @@ for (r_h in seq_along(list_regions)) {
         output_file = paste0(
           "html/VOCC-plots/vocc-w-do-", spp,
           "-imm",
-          covs, "-", reg, "-dvocc.html"
+          covs, "-", reg, "-updated-climate.html"
         ),
         envir = env
       )

### 2020-06-18 analysis/VOCC/4-pool-vocc.R

@@ -3,12 +3,15 @@ setwd(here::here("/analysis/VOCC/data"))
 library("dplyr")
 
 # covs <- "-fixed"
-mydir = paste0( "_new_null")
-mydir = paste0( "_do_null")
-mydir = paste0( "_dvocc")
+# mydir = paste0( "_new_null")
+# mydir = paste0( "_do_null")
+# mydir = paste0( "_dvocc")
 
-mydir = paste0( "_dvocc_mature")
-mydir = paste0( "_dvocc_immature")
+# mydir = paste0( "_dvocc_mature")
+# mydir = paste0( "_dvocc_immature")
+
+# mydir = paste0( "_newclim_mature")
+mydir = paste0( "_newclim_immature")
 ####################
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
 myfiles
@@ -25,10 +28,17 @@ glimpse(biotic)
 #   "mature-all-do-untrimmed.rds"
 # ))
 # 
-saveRDS(biotic, file = paste0(
-  "mature-all-do-dvocc.rds"
-))
+# saveRDS(biotic, file = paste0(
+#   "mature-all-do-dvocc.rds"
+# ))
+# 
+# saveRDS(biotic, file = paste0(
+#   "immature-all-do-dvocc.rds"
+# ))
+# saveRDS(biotic, file = paste0(
+#   "mature-all-do-newclim.rds"
+# ))
 
 saveRDS(biotic, file = paste0(
-  "immature-all-do-dvocc.rds"
+  "immature-all-do-newclim.rds"
 ))

### 2020-06-18 analysis/VOCC/5-add-null.R

@@ -15,16 +15,26 @@ if (trim_threshold == 0.1) { trim_percent <- 90}
 if (trim_threshold == 0.2) { trim_percent <- 80}
 if (trim_threshold == 0.5) { trim_percent <- 50}
 
+# age <- "mature"
+# d1 <- readRDS(paste0("data/", age, "-all-do-dvocc.rds")) %>% mutate(age = "mature")
+# 
+# age <- "immature"
+# d2 <- readRDS(paste0("data/", age, "-all-do-dvocc.rds")) %>% mutate(age = "immature")
+
 age <- "mature"
-d1 <- readRDS(paste0("data/", age, "-all-do-dvocc.rds")) %>% mutate(age = "mature")
+d1 <- readRDS(paste0("data/", age, "-all-do-newclim.rds")) %>% mutate(age = "mature")
 
 age <- "immature"
-d2 <- readRDS(paste0("data/", age, "-all-do-dvocc.rds")) %>% mutate(age = "immature")
+d2 <- readRDS(paste0("data/", age, "-all-do-newclim.rds")) %>% mutate(age = "immature")
 
 
 d <- rbind(d1, d2) 
 
 d <- na.omit(d) %>% as_tibble() %>% mutate(species_age = paste(age, species))
+
+d <- filter(d, species_age != "immature Shortraker Rockfish") 
+d <- filter(d, species_age != "immature Curlfin Sole")
+
 
 all_species <- unique(d$species_age)
 
@@ -111,14 +121,14 @@ newdata <- do.call(rbind, with_nulls)
 
 # saveRDS(newdata, file = paste0("data/mature-all-temp-with-null-1-untrimmed.rds"))
 # saveRDS(newdata, file = paste0("data/", age, "-all-do-with-null-", null_number, "-untrimmed.rds"))
-saveRDS(newdata, file = paste0("data/all-do-with-null-", null_number, "-untrimmed.rds"))
+saveRDS(newdata, file = paste0("data/all-newclim-with-null-", null_number, "-untrimmed.rds"))
 
 
 ### ADD SPECIES TRAITS ANS SCALE VARIABLES
 # age <- "mature"
 # d <- readRDS(paste0("data/", age, "-all-do-with-null-", null_number, "-untrimmed.rds"))
 
-d <- readRDS(paste0("data/all-do-with-null-", null_number, "-untrimmed.rds"))
[…]

### 2020-06-18 analysis/VOCC/3-vocc-dist.Rmd
make dvocc code robust to missing dvocc estimates

@@
     immature: F
-    # region: "West Coast Vancouver Island"
-    region: "both odd year surveys"
+    region: "West Coast Vancouver Island"
+    # region: "both odd year surveys"
     # region: "West Coast Haida Gwaii"
@@
 
-gf2$dvocc <- dv_t$dvocc
+# gf2$dvocc <- dv_t$dvocc
+
+gf2 <- left_join(gf2, dv_t)
 ```
@@
   # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
+  na_colour = "black",
   white_zero = TRUE,
@@
 
-gf0$dvocc <- dv_d$dvocc
+# gf0$dvocc <- dv_d$dvocc
+
+gf0 <- left_join(gf0, dv_d)
 
@@
 
-dv_td <- dvocc_td %>% select(icell, x, y, dvocc) %>% unique()
+dv_td <- dvocc_td %>% select(icell, x, y, dvocc_both = dvocc) %>% unique()
 # glimpse(dv_b)
@@
 
-gf0$dvocc_both <- dv_td$dvocc
+# gf0$dvocc_both <- dv_td$dvocc
+
+gf0 <- left_join(gf0, dv_td)
+
 ```
@@
 
-gf3$dvocc <- dv_b$dvocc
+# gf3$dvocc <- dv_b$dvocc
+
+gf3 <- left_join(gf3, dv_b)
 
@@
 plot_vocc(vocc_t,
+    raster_cell_size = 4,
   # theme = "black",
@@
 
-plot_vocc(vocc_t,
+plot_vocc(vocc_t,  raster_cell_size = 4,
   # theme = "black",
@@
 plot_vocc(vocc_t,
-  # theme = "black",
+    raster_cell_size = 4,# theme = "black",
   coast = TRUE,
@@
 plot_vocc(vocc_t,
-  # theme_black = TRUE,
+    raster_cell_size = 4,# theme_black = TRUE,
   coast = TRUE,

### 2020-06-18 analysis/VOCC/3-vocc-dist.Rmd
adjust climate data for mistaken use of trend_per_year use in vocc code

@@
     ## if default of divisor = 10
-      trend_per_year = exp(trend/10),
+      trend_per_year = (trend/10), # percent change not raw
       trend_per_decade = trend_per_year * 10,
@@
     biotic_dvocc = dvocc,
-    biotic_trend = trend,
+    biotic_trend = trend, # currently per decade
     biotic_CV = CV,
@@
     DO_dvocc = dvocc,
-    DO_trend = trend_per_year,
+    DO_trend = trend_per_year, # should change to trend_per_decade
     DO_grad = gradient,
@@
     temp_dvocc = dvocc,
-    temp_trend = trend_per_year,
+    temp_trend = trend_per_year, # should change to trend_per_decade
     temp_grad = gradient,

### 2020-06-18 analysis/VOCC/5-add-null.R

@@ -99,7 +99,8 @@ omega_s <- omega_s - mean(omega_s)
 # observed <- rnorm(length(omega_s), m$model$par[["b_j"]], exp(m$model$par[["ln_phi"]]))
 # observed <- rnorm(length(omega_s), omega_s + m$model$par[["b_j"]], exp(m$model$par[["ln_phi"]]))
 # observed <- rnorm(length(omega_s), omega_s + mean(.x$biotic_trend), exp(m$model$par[["ln_phi"]]))
-observed <- rnorm(length(omega_s), omega_s + mean(.x$biotic_trend), 0.001)
+# observed <- rnorm(length(omega_s), omega_s + mean(.x$biotic_trend), 0.001)
+observed <- omega_s + mean(.x$biotic_trend)
 
 s <- data.frame(x = .x$x, y = .x$y, fake_trend = observed)
 
@@ -165,7 +166,7 @@ hist((d$mean_temp_scaled2))
 hist(d$squashed_temp_vel_scaled, breaks = 100)
 
 # hist(d$temp_trend)
-d$temp_trend_scaled <- scale(d$temp_trend, center = FALSE)
+d$temp_trend_scaled <- scale(d$temp_trend*10, center = FALSE)
 # hist(d$temp_trend_scaled)
 d$temp_grad_scaled <- scale(d$temp_grad)
 # hist(d$temp_grad_scaled)
@@ -191,12 +192,12 @@ hist((d$mean_DO_scaled))
 hist((d$mean_DO_scaled2))
 
 hist(d$mean_DO_scaled)
-hist(d$DO_trend)
-d$squashed_DO_trend <- collapse_outliers(d$DO_trend, c(0.005, 0.995))
+hist(d$DO_trend*10)
+d$squashed_DO_trend <- collapse_outliers(d$DO_trend*10, c(0.005, 0.995))
 hist(d$squashed_DO_trend)
-d$squashed_DO_trend_scaled <- scale(d$quashed_DO_trend, center = FALSE)
+d$squashed_DO_trend_scaled <- scale(d$squashed_DO_trend, center = FALSE)
 
-d$DO_trend_scaled <- scale(d$DO_trend, center = FALSE)
+d$DO_trend_scaled <- scale(d$DO_trend*10, center = FALSE)
 hist(d$DO_trend_scaled)
 d$DO_grad_scaled <- scale(d$DO_grad)
 # hist(d$DO_grad_scaled)

### 2020-06-19 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
fitting climate models

--- 2020-02-24 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 library(dplyr)
-
-plan(multiprocess)
 ```
 
-Load data
+Load data if not using SST
 ```{r}
-all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
-# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
-# glimpse(all_sensor)
-# 
-# replace obviously faulty depth data with NA
-#all_sensor$depth_m[all_sensor$depth_m < 10] <- NA # 5,139
+all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
 
 all_depth <- all_sensor %>%
-  #filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
+  # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
   # filter missing location data and trial year
@@
 
-unique(all_depth$data$year)
-
 d_trawl <- all_depth$data %>% 
@@
   filter(depth_max>10) %>% 
-  # filter(ssid != 1) %>%
-  # filter(ssid != 3) %>%
-  # filter(ssid != 4) %>%
-  # filter(ssid != 16) %>%
   gfplot:::scale_survey_predictors()
-
-hist(d_trawl$temperature_c)
 
@@
 d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
-#d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
-#d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
+
 d_trawl <- mutate(d_trawl,
@@
   DOY_scaled = arm::rescale(lubridate::yday(date)),
-  DOY_scaled2 = arm::rescale(lubridate::yday(date))^2,
-  #depth_scaled = arm::rescale(log(depth_m)),
-  SST_scaled = arm::rescale(meanSST),
-  #depth_scaled2 = depth_scaled^2,
-  depth_scaled3 = depth_scaled^3,
-  shallow = ifelse(depth>35,0,1),
-  deep = ifelse(depth<35,0,1)
-)
-# d_trawl[is.na(d_trawl$shallow), ]
-# d_trawl[is.na(d_trawl$SST_scaled), ]
-
-glimpse(d_trawl)
-
-#d_trawl <- d_trawl %>% dplyr::filter(year>2003)
-```
-
-Explore SST 
-```{r}
-ggplot(d_trawl, aes_string("X", "Y", colour = "SST")) +
-    geom_point() +
-    facet_wrap(~year) +
-    coord_fixed() +
-    scale_color_viridis_c()
-```
-```{r}
-ggplot(d_trawl, aes(DOY, SST, colour = -depth)) + geom_point() + facet_wrap(~year) 
-```
-
-```{r}
-ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = -depth)) + geom_point()
-ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = DOY)) + geom_point()
-
-```
-
-Explore mean SST 
-```{r}
-ggplot(d_trawl, aes_string("X", "Y", colour = "meanSST")) +
-    geom_point() +
-    facet_wrap(~year) +
-    coord_fixed() +
-    scale_color_viridis_c()
-```
-
-```{r}
-ggplot(dplyr::filter(d_trawl1, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
+  DOY_scaled2 = arm::rescale(lubridate::yday(date))^2)
 ```
@@
 EXLORE TEMPERATURE at DEPTH
-
 ```{r}
@@
 ```{r}
-ggplot(d_trawl, aes(log(depth),temperature_c, colour = depth_max)) + geom_point() + facet_wrap(~ssid)
-```
-
-```{r}
 ggplot(d_trawl, aes(DOY, temperature_c, colour = -depth, alpha=0.5)) + geom_point() + facet_wrap(~year) 
@@
 
-```{r eval = FALSE}
-library(gbm)
-
-m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
-m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
-
-plot(m,i.var=2)
-plot(m,i.var=3)
-plot(m,i.var=4)
-plot(m,i.var=5)
-
-plot(m2,i.var=1)
-plot(m2,i.var=2)
-plot(m,i.var=3)
-plot(m,i.var=4)
-
-plot(m2,i.var=1:2)
-plot(m2,i.var=c(1,3))
-plot(m2,i.var=c(1,4))
-d_trawl$r <- predict(m, n.trees = 2000) - d_trawl1$temperature_c
-
-ggplot(d_trawl, aes(depth_scaled, r)) +
-  geom_point(alpha=0.4) +
-  ylim(-8,8) +
-  geom_smooth()
-```
-
-### Functions for model exploration:
+Current temperature model
 ```{r}
-# aic <- function(m) {
-#   k <- length(m$model$par)
-#   nll <- m$model$objective
-#   2 * k - 2 * (-nll)
-# }
-
-generate_cv_args <- function(
-                             formula = formula,
-                             time_varying = time_varying,
-                             ar1_fields = c(FALSE), # simiple, fast defaults
-                             include_spatial = c(FALSE), # simiple, fast defaults
-                             n_knots = c(100), # simiple, fast defaults
-                             seed = c(999)) {
-  arguments <- expand.grid(
-    formula = formula,
-    time_varying = time_varying,
-    include_spatial = include_spatial,
-    ar1_fields = ar1_fields,
-    n_knots = n_knots,
-    seed = seed
-  )
-  as.data.frame(arguments)
-}
-
-sdmTMB_cv_batch <- function(cv_args, data, ...) {
-  cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl, ...)
-  cv_out <- cv_args
-  cv_out$cv_loglik <- c()
-  cv_out$all_converged <- c()
-  cv_out$max_grad <- c()
-  for (i in seq_len(nrow(cv_args))) {
-    cv_out$cv_loglik[[i]] <- round(cv_batch[[i]]$sum_loglik, 1)
-    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converged)
-    cv_out$max_grad[[i]] <- max(cv_batch[[i]]$max_grad)
-    if (is.null(cv_out$time_varying[[i]])) {
-      cv_out$time_varying[[i]] <- as.character(NA)
-    }
-  }
-  cv_out$formula <- as.character(cv_out$formula)
-  cv_out$time_varying <- as.character(cv_out$time_varying)
-
-  cv_out <- cv_out %>%
-    group_by(seed) %>%
-    mutate(., max_ll_per_seed = max(cv_loglik)) %>%
-    mutate(., delta_ll = cv_loglik - max_ll_per_seed) %>%
-    ungroup()
-  row.names(cv_out) <- NULL
-  list(summary = as.data.frame(cv_out), cv_model_set = cv_batch)
-}
-
-# # this code keeps breaking... working, fails, worked again, then failed again...
-# plot_cv_meshes <- function(sdmTMB_cv_output) {
-#   cv_output <- sdmTMB_cv_output
-#   k <- max(cv_output$data$cv_fold) # should probably rename cv_fold to fold_id
-#   if (k == 2 | k == 5 | k == 10) {
-#     op <- graphics::par(
-#       mfrow = c((sqrt(k)), ceiling(sqrt(k))),
-#       mar = c(1, 1, 1, 1)
-#       )
-#   } else {
-#     op <- graphics::par(
-#       mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
-#       mar = c(1, 1, 1, 1)
-#     )
-#   }
-#   for (i in seq_len(k)){
-#     plot_spde(cv_output$models[[i]]$spde)
-# #FIXME: Error in xy.coords(x, y, xlabel, ylabel, log) :
-# #  'x' is a list, but does not have components 'x' and 'y'
-#   }
-#   graphics::par(op)
-# }
-```
-
-Short example: 
-```{r, eval=FALSE}
-cv_args <- generate_cv_args(
-  formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3
-  ),
-  time_varying = list(NULL),
-  ar1_fields = c(FALSE),
-  include_spatial = c(FALSE),
-  n_knots = c(135, 140, 145, 150),
-  seed = c(1)
-)
-
-cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
-glimpse(cv$summary)
-write.csv(cv$summary, file = "cv-output/cv-test.csv")
-cv_test <- read_csv(file = "cv-output/cv-test.csv")
-View(cv$summary)
-# plot_cv_meshes(cv$cv_model_set[[1]])
-# plot_cv_meshes(cv$cv_model_set[[2]])
-```
-
-# 10-fold cross-validation
-
-## Just compare fixed slope models with simplest random effect structure
-```{r eval=F}
-cv_args <- generate_cv_args(
-  formula = list(
-    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled,
-    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2,
-    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
-    temperature_c ~ 1 + SST_scaled + depth_scaled,
-    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2,
-    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
-    temperature_c ~ 0 + as.factor(year) + depth_scaled,
-    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
-    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2 + SST_scaled * depth_scaled3
-  ),
-  time_varying = list(NULL),
-  ar1_fields = c(FALSE),
-  include_spatial = c(FALSE),
-  n_knots = c(100),
-  seed = c(1111, 2222, 3333, 4444)
-)
-
-non_time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
-View(non_time_varying$summary)
-write.csv(non_time_varying$summary, file = "cv-output/cv-non-time-varying.csv")
-saveRDS(non_time_varying, file = "cv-output/cv-non-time-varying.rds")
-glimpse(non_time_varying$summary)
-```
-
-## Compare time-varying slope models still with simplest random effect structure
-All of top models for any seed contained year and none showed a linear effect of depth
-All are better than fixed slope models
-```{r}
-cv_args <- generate_cv_args(
-  formula = list(
-    #temperature_c ~ 0 + SST_scaled*shallow,
-    # temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
-    # #temperature_c ~ 1 + SST_scaled,
-    # temperature_c ~ 0 + as.factor(year) + SST_scaled,
-    temperature_c ~ 0 + as.factor(year)
-  ),
-  time_varying = list(
-    # ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
-    ~ 0 + depth_scaled + depth_scaled2,
-  ),
-  ar1_fields = c(TRUE),
-  include_spatial = c(TRUE),
-  n_knots = c(400), # keep same n_knots
-  seed = c(1111, 2222, 3333, 4444) # keep same seeds
-)
-
-time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
-
-View(time_varying$summary)
-write.csv(time_varying$summary, file = "cv-output/cv-time-varying.csv")
-saveRDS(time_varying, file = "cv-output/cv-time-varying.rds")
-glimpse(time_varying$summary)
-```
-
-## Now vary random effect configuation for top model(s) 
-a separate spatial field is in all models with highest likelihoods for each seed
-AR1 doesn't make much difference (differences between years appear independent)
-```{r}
-cv_args <- generate_cv_args(
-  formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled,
-    temperature_c ~ 0 + as.factor(year)
-  ),
-  time_varying = list(
-    ~ 0 + depth_scaled + depth_scaled2
-  ),
-  ar1_fields = c(FALSE, TRUE),
-  include_spatial = c(FALSE, TRUE),
-  n_knots = c(100),
-  seed = c(1111, 2222, 3333, 4444)
-)
-
-vary_rf <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
-
-View(vary_rf$summary)
-write.csv(vary_rf$summary, file = "cv-output/cv-varying-rf.csv")
-saveRDS(vary_rf, file = "cv-output/cv-varying-rf.rds")
-glimpse(vary_rf$summary)
-```
-
-## Change knot number
-
-```{r}
-cv_args <- generate_cv_args(
-    formula = list(
-    # temperature_c ~ 0 + as.factor(year) + SST_scaled,
-    temperature_c ~ 0 + as.factor(year)
-  ),
-  time_varying = list(
-    ~ 0 + depth_scaled + depth_scaled2
-  ),
-  ar1_fields = c(FALSE,TRUE),
-  include_spatial = c(TRUE),
-  n_knots = c(140,180,220,250),
-  #seed = c(1111, 2222, 3333, 4444) 
-  seed = c(5555, 6666, 7777, 8888, 9999)
-)
-
-more_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
-
-View(more_knots$summary)
-write.csv(more_knots$summary, file = "cv-output/cv-more-knots2.csv")
-saveRDS(more_knots, file = "cv-output/cv-more-knots2.rds")
-glimpse(more_knots$summary)
-```
-
-## Change seeds 
-Also increased knot number too 
-Models without SST_scaled are usually better
-
-```{r}
-cv_args <- generate_cv_args(
-    formula = list(
-    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
-    temperature_c ~ 0 + as.factor(year) + SST_scaled,
-    temperature_c ~ 0 + as.factor(year)
-  ),
-  time_varying = list(
-    ~ 0 + depth_scaled + depth_scaled2
-  ),
-  ar1_fields = c(FALSE, TRUE),
-  include_spatial = c(TRUE, FALSE),
-  n_knots = c(130),
-  seed = c(5555, 6666, 7777, 8888, 9999)
-)
-
-more_seeds <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
-
-View(more_seeds$summary)
-write.csv(more_seeds$summary, file = "cv-output/cv-more-seeds.csv")
-saveRDS(more_seeds, file = "cv-output/cv-more-seeds.rds")
-glimpse(more_seeds$summary)
-```
-
-Load saved batches:
-```{r, eval=FALSE}
-# non_time_varying <- readRDS(file = "cv-output/cv-non-time-varying.rds")
-# time_varying <- readRDS(file = "cv-output/cv-time-varying.rds")
-# vary_rf <- readRDS(file = "cv-output/cv-varying-rf.rds")
-# knot_refining <- readRDS(file = "cv-output/cv-refine-knots.rds")
-
-non_time_varying <- read.csv(file = "cv-output/cv-non-time-varying.csv")
-time_varying <- read.csv(file = "cv-output/cv-time-varying.csv")
-vary_rf <- read.csv(file = "cv-output/cv-varying-rf.csv")
-non_time_varying1 <- read.csv(file = "cv-output/cv-non-time-varying1.csv")
-time_varying1 <- read.csv(file = "cv-output/cv-time-varying1.csv")
-vary_rf1 <- read.csv(file = "cv-output/cv-varying-rf1.csv")
-refine_knots1 <- read.csv(file = "cv-output/cv-refine-knots1.csv")
-more_seeds <- read.csv(file = "cv-output/cv-more-seeds.csv")
-more_seeds1 <- read.csv(file = "cv-output/cv-more-seeds1.csv")
-
-all_cv_loglik <- rbind(
-  non_time_varying,
-  time_varying,
-  vary_rf,
-  non_time_varying1,
-  time_varying1,
-  vary_rf1,
-  refine_knots1,
-  more_seeds,
-  more_seeds1,
-  deparse.level = 1, make.row.names = TRUE
-)
-
-by_seed <- all_cv_loglik %>%
-  group_by(seed) %>%
-  mutate(., max_ll_per_seed = max(cv_loglik)) %>%
-  mutate(., delta_ll = cv_loglik - max_ll_per_seed)
-
-View(by_seed)
-all_cv_loglik <- by_seed %>% ungroup()
-
-write.csv(all_cv_loglik, file = "all_cv_loglik.csv")
-```
-
-Current top model
-
-```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 400)
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 500)
 plot_spde(spde)
@@
 
-```{r}
-m_temp_rw <- sdmTMB(d_trawl,
+```{r eval = F}
+m_temp_rw <- sdmTMB(
   temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  data = d_trawl,
   ar1_fields = TRUE, # changed AR1_field to true
@@
 
-saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn.rds")))
-
+saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
 ```
 
+Check predictions against raw data
 ```{r}
-predictions_rw <- predict(m_temp_rw)
-predictions_rw$residuals <- residuals(m_temp_rw)
-cor(predictions_rw$est, predictions_rw$temperature_c)^2 #
-# aic(m_temp3_rw)
-
+predictions <- predict(m_temp_rw)
+predictions$residuals <- residuals(m_temp_rw)
+cor(predictions$est, predictions$temperature_c)^2 
 ```
 
-Residual plots:
+Model diagnostics functions
 ```{r}
-ggplot(predictions_rw, aes(est, residuals)) +
+get_diag <- function(m, response = "log_do", variable = "depth_scaled") {
+  predictions <- predict(m)
+  predictions$residuals <- residuals(m)
+
+  plot_map <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", colour = column)) +
+      geom_point() +
+      coord_fixed()
+  }
+
+  g <- plot_map(predictions, "est") +
+    scale_colour_viridis_c() +
+    ggtitle("Prediction (fixed effects + all random effects)")
+  print(g)
+
+  g <- plot_map(predictions, "est_non_rf") +
+    ggtitle("Prediction (fixed effects only)") +
+    scale_colour_viridis_c()
+  print(g)
+
+  g <- plot_map(predictions, "est_rf") +
+    ggtitle("All random effects only") +
+    scale_colour_gradient2()
+  print(g)
+
+  g <- plot_map(predictions, "omega_s") +
+    ggtitle("Spatial random effects only") +
+    scale_colour_gradient2()
+ print(g)
+
+   g <- plot_map(predictions, "epsilon_st") +
+    ggtitle("Spatiotemporal random effects only") +
+    facet_wrap(~year) +
+    scale_colour_gradient2()
+ print(g)
+
+  g <- ggplot(predictions, aes_string("est", response)) +
+    geom_point() +
+    facet_wrap(~year) +
+    coord_fixed() +
+    geom_abline()
+  print(g)
+
+  g <- ggplot(predictions, aes(est, residuals)) +
+    geom_point() +
+    geom_smooth()
+  print(g)
+
+  g <- ggplot(predictions, aes_string(variable, "residuals")) +
     geom_point() +
@@
     facet_wrap(~year)
+  print(g)
+
+  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
+    geom_point() +
+    coord_fixed() +
+    scale_color_gradient2()+
+    facet_wrap(~year)
+  print(g)
+
+  aic <- function(m) {
+    k <- length(m$model$par)
+    nll <- m$model$objective
+    2 * k - 2 * (-nll)
+  }
+
+  print("R^2:")
+  r2 <- cor(predictions$est, predictions[[response]])^2
+  print(r2)
+
+  print("")
+  print("AIC:")
+  print(aic(m))
+}
 ```
 
-```{r}
-ggplot(predictions_rw, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
+```{r eval=FALSE}
+m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
+get_diag(m_temp_rw, response = "temperature_c", variable = "depth_scaled")
 ```
 
-```{r}
-ggplot(predictions_rw, aes(depth_scaled, residuals)) +
+WOULD SST HELP?
+
+Load and prep SST data
+```{r eval=FALSE}
+all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
+# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
+# glimpse(all_sensor)
+
+all_depth <- all_sensor %>%
+  # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
+  # filter missing location data and trial year
+  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%
+
+  # convert lat and lon to UTMs
+  dplyr::mutate(X = longitude, Y = latitude) %>%
+  gfplot:::ll2utm(., utm_zone = 9) %>%
+
+  # interpolate missing depth values
+  dplyr::rename(depth = depth_m) %>%
+  gfplot:::interp_survey_bathymetry()
+
+d_trawl <- all_depth$data %>% 
+  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
+  filter(depth_max>10) %>% 
+  gfplot:::scale_survey_predictors()
+
+d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
+d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
+
+d_trawl2 <- d_trawl[!is.na(d_trawl$SST), ]
+d_trawl2 <- d_trawl2[!is.na(d_trawl2$meanSST), ]
+
+d_trawl2 <- mutate(d_trawl2,
+  DOY = lubridate::yday(date),
+  DOY_scaled = arm::rescale(DOY),
+  DOY_scaled2 = DOY_scaled^2,
+  depth_scaled = arm::rescale(log(depth)),
+  SST_scaled = arm::rescale(meanSST),
+  depth_scaled3 = depth_scaled^3,
+  shallow = ifelse(depth>35,0,1),
+  deep = ifelse(depth<35,0,1)
+)
+
+d_trawl2[is.na(d_trawl2$shallow), ]
+d_trawl2[is.na(d_trawl2$SST_scaled), ]
+
+# glimpse(d_trawl2)
+```
+
+Explore SST on survey day
+```{r eval=FALSE}
+ggplot(d_trawl2, aes_string("X", "Y", colour = "SST")) +
     geom_point() +
-  geom_smooth()
+    facet_wrap(~year) +
+    coord_fixed() +
+    scale_color_viridis_c()
+```
+
+Is bottom temp related to SST in shallower samples?
+```{r eval=FALSE}
+ggplot(dplyr::filter(d_trawl2, depth <35), aes(SST,temperature_c, colour = -depth)) + geom_point()
+ggplot(dplyr::filter(d_trawl2, depth <35), aes(SST,temperature_c, colour = DOY)) + geom_point()
+```
+
+Explore mean SST 
+```{r eval=FALSE}
+ggplot(d_trawl2, aes_string("X", "Y", colour = "meanSST")) +
+    geom_point() +
+    facet_wrap(~year) +
+    coord_fixed() +
+    scale_color_viridis_c()
+```
+
+```{r eval=FALSE}
+ggplot(dplyr::filter(d_trawl2, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
 ```
@@
 Does a fixed interaction with SST_scaled do any better?
+```{r}
+spde2 <- make_spde(d_trawl2$X, d_trawl2$Y, n_knots = 500)
+plot_spde(spde2)
+```
 
 ```{r eval=FALSE}
-m_temp3 <- sdmTMB(d_trawl,
-  temperature_c ~ 0 + as.factor(year) + SST_scaled*depth_scaled + SST_scaled*depth_scaled2 ,
-  ar1_fields = FALSE, include_spatial = TRUE,
-  time = "year", spde = spde, family = gaussian(link = "identity"),
+m_temp3 <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow + DOY_scaled * shallow + DOY_scaled2 * shallow,
+  data = d_trawl2,
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  ar1_fields = TRUE, include_spatial = TRUE,
+  time = "year", spde = spde2, family = gaussian(link = "identity"),
   silent = TRUE
 )
+```
 
+```{r eval=FALSE}
 predictions3 <- predict(m_temp3)
@@
 
-```{r eval=FALSE}
-ggplot(predictions3, aes(est, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r eval=FALSE}
-ggplot(predictions3, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(~year)
-```
-
-```{r eval=FALSE}
-ggplot(predictions3, aes(depth_scaled, residuals)) +
-  geom_point() +
-  geom_smooth()
-```
-
-#FIXME: DON'T HAVE SST TEMP FOR PREDICITON GRID YET...
-
 ## Plotting predictions from saved temperature model
 
-```{r}
-nd_all <- readRDS(paste0("../VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
+```{r }
+m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
+nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
+
 unique(nd_all$year)
 glimpse(nd_all)
-predict_temp <- predict(m_temp_rw, newdata = nd_all)
-saveRDS(predict_temp, file = "../VOCC/data/predicted_temp_allyears_revised")
-p_nd <- predict_temp
-#p_nd <- readRDS("../VOCC/data/predicted_temp_allyears.rds")
-# View(p_nd)
+pred_temp <- predict(m_temp_rw, newdata = nd_all)
+
+# saveRDS(pred_temp, file = "../VOCC/data/predicted_temp_allyears_revised2")
 ```
@@
 ```{r}
-p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised")
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised2")
 
@@
 plot_map_raster(p_nd, "est") +
-  scale_fill_viridis_c(trans = "log") +
+  scale_fill_viridis_c(trans = sqrt) +
   facet_wrap(~year) +
   ggtitle("Prediction (fixed effects + all random effects)")
+```
+
+```{r}
+plot_facet_map(p_nd, "est", viridis_option = "D",
+    raster_limits = c(min(p_nd$est), quantile(p_nd$est, 0.99)),
+    transform_col = sqrt
+  )
 ```
@@
   scale_fill_viridis_c(labels = labels, 
-    #trans = "sqrt",
     option = "C",
-    #breaks = c(0,2,4,6,8), 
-    limits = c(5, 11), 
+    limits = c(4, 11),
     na.value = "red") +
@@
   labs(fill = "°C") + 
-  theme(legend.position = "none") +
   ggtitle("Prediction (fixed effects + all random effects)") 
@@
   scale_fill_viridis_c(labels = labels, 
-     #trans = "sqrt",
     option = "C",
-    #breaks = c(0,2,4,6,8), 
     limits = c(5, 11), 
@@
   labs(fill = "°C")
-  #ggtitle(" ")
 pred_4
@@
 ```{r}
-#glimpse(p_nd)
-
 omega_s_all <- plot_map_raster(p_nd, "omega_s") +
@@
 p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
-st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
+
+(st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
   ggtitle("Spatiotemporal random effects") +
@@
   legend.position = "none", 
-  axis.title.x = element_blank(), axis.title.y = element_blank())
+  axis.title.x = element_blank(), axis.title.y = element_blank()))
 
-png(
-  file = "figs/temp-predictions-1n3.png", 
-  res = 400,
-  units = "in",
-  width = 8, 
-  height = 8
-) 
-st_1n3
-dev.off()
+# png(
+#   file = "figs/temp-predictions-1n3.png", 
+#   res = 400,
+#   units = "in",
+#   width = 8, 
+#   height = 8
+# ) 
+# st_1n3
+# dev.off()
 
@@
 
-```{r}
+```{r, eval= F}
 png(
@@
 
-```{r}
-temp_cog <- get_cog(p_nd)
+BOOSTED REGRESSION MODEL
+```{r, eval = FALSE}
+library(gbm)
+
+m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
+m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
+
+plot(m,i.var=2)
+plot(m,i.var=3)
+plot(m,i.var=4)
+plot(m,i.var=5)
+
+plot(m2,i.var=1)
+plot(m2,i.var=2)
+plot(m,i.var=3)
+plot(m,i.var=4)
+
+plot(m2,i.var=1:2)
+plot(m2,i.var=c(1,3))
+plot(m2,i.var=c(1,4))
+d_trawl$r <- predict(m, n.trees = 2000) - d_trawl1$temperature_c
+
+ggplot(d_trawl, aes(depth_scaled, r)) +
+  geom_point(alpha=0.4) +
+  ylim(-8,8) +
+  geom_smooth()
+```
+
+CENTER OF GRAVITY
+```{r, eval = F}
+p_cog <- predict(m_temp_rw, newdata = nd_all, return_tmb_object = TRUE)
+
+temp_cog <- get_cog(p_cog)
 

### 2020-06-19 analysis/tmb-sensor-explore/3-fit-do.Rmd
fitting climate models

@@
 ---
-# Fit spatial temporal models to bottom temperature
+# Fit spatial temporal models of bottom DO
 
@@
 
-glimpse(d_trawl)
+# glimpse(d_trawl)
+```
+
+Ranges sampled
+```{r}
+paste("depth") 
+range(d_trawl$raw_depth)
+paste("DO") 
+range(d_trawl$do_mlpl)
 ```
@@
 
+Remove WCVI in 2016 due to extreme values and make 500 knot mesh
 ```{r}
-# # filter surveys if desired
-# d_trawl <- d_trawl %>% 
-  # filter(ssid != 1) %>%
-  # filter(ssid != 3) %>%
-  # filter(ssid != 4) %>%
-  # filter(ssid != 16) %>%
-  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
-  # filter(depth_m_max>10) %>% 
-```
+d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4))
 
-```{r}
-
-d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4)) #%>% filter(ssid != 16) %>% filter(ssid != 4)    
-#test <- d_trawl1 %>% filter(year == 2016 & ssid ==16) 
-spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 800) ### TODO: rerun with 500
+spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 500) 
 plot_spde(spde)
@@
 
-Test temp model on this subset of data
-DOY doesn't improve model
-```{r eval=FALSE}
-    starttime1 <- Sys.time()
-
-    temp_model_DOY <- sdmTMB::sdmTMB(d_trawl1,
-      temp_scaled ~ 0 + as.factor(year), # + DOY_scaled,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2,
-      time = "year", spde = spde,
-      family = gaussian(link = "identity"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = FALSE,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-    endtime1 <- Sys.time()
-    time1 <- round(starttime1 - endtime1)
-
-saveRDS(temp_model_all, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-2008-on.rds")))
-
-# saveRDS(temp_model_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-2008-on-DOY.rds")))
-# time1
-# temp_model_DOY
-temp_model_all
-```
-
-```{r eval=FALSE}
-get_diag(temp_model_all, response = "temp_scaled", variable = "depth_scaled")
-```
-
-Add temperature predictions to newdata grid for DO predictions
-```{r eval=FALSE}
-temperature <- readRDS(here::here("analysis/VOCC/data/predicted_temp_allyears_revised2"))
-years <- unique(d_trawl1$year) 
-temperature <- filter(temperature, year %in% years)
-unique(temperature$year)
-
-temperature$temp <- temperature$est
-#temperature$temp_scaled2 <- temperature$est^2
-temperature$raw_depth <- temperature$depth
-temperature$depth <- log(temperature$depth)
-temperature$DOY_scaled <- 0
-
-newdata <- temperature %>% 
-  select(-est, - est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st, -temperature_c) %>%
-  scale_predictors(predictors = c(quo(depth),quo(temp))) %>% 
-  mutate(depth = raw_depth)
-
-glimpse(newdata)
-```
-
 Current best model of DO
-```{r eval=FALSE}
-    starttime1 <- Sys.time()
-
+```{r }
 do_model_all_DOY <- sdmTMB::sdmTMB(
@@
       data = d_trawl1,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
       time = "year", 
@@
     )
-    endtime1 <- Sys.time()
-    time1 <- round(starttime1 - endtime1)
 
 saveRDS(do_model_all_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds")))
-time1
-do_model_all_DOY
 ```
@@
 ```{r eval=FALSE}
-# do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all-DOY.rds"))
-# do_model_DOY
-
 get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
-
-# get_diag(do_model_DOY, response = "log_do", variable = "temp_scaled")
-```
-
-Other models tried:
-```{r eval=FALSE}
-do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds"))
-do_model_DOY
-#p_nd <- predict(do_model_DOY, newdata = newdata)
-get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")
-```
-
-```{r eval=FALSE}
-#with both DOY and depth as time-varying random effects
-do_model_re_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-re-DOY.rds"))
-do_model_re_DOY
-get_diag(do_model_re_DOY, response = "log_do", variable = "depth_scaled")
-```
-
-```{r eval=FALSE}
-do_model_DOY <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-DOY.rds"))
-do_model_DOY
-#p_nd <- predict(do_model_DOY, newdata = newdata)
-get_diag(do_model_DOY, response = "log_do", variable = "depth_scaled")
-
-```
-
-```{r eval=FALSE}
-do_model_depth2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth2.rds"))
-do_model_depth2
-get_diag(do_model_depth2,response = "log_do", variable = "depth_scaled")
-#p_nd <- predict(do_model_depth2, newdata = newdata)
-```
-
-```{r eval = FALSE}
-do_model_depth3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-depth3.rds"))
-get_diag(do_model_depth3,response = "log_do")
-#p_nd <- predict(do_model_depth3, newdata = newdata)
-```
-
-```{r eval=FALSE}
-do_model_all <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-all.rds"))
-#p_nd <- predict(do_model_all, newdata = newdata)
-do_model_all
-get_diag(do_model_all, response = "log_do", variable = "depth_scaled")
 ```
@@
 ```{r eval=FALSE}
-do_model_all <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-poly.rds"))
-p_nd <- predict(do_model_all_DOY, newdata = newdata)
-p_nd$do_est <- exp(p_nd$est)
-
-p_nd <- p_nd %>% filter(depth>15)
-hist(p_nd$do_est)
-
-max(p_nd$do_est)
-max(d_trawl1$do_mlpl)
-glimpse(p_nd)
-min(p_nd$raw_depth)
-min(d_trawl1$raw_depth)
-p_nd[p_nd$do_est>8, ]
-
-p_nd <- p_nd %>% filter(depth>15)
-p_nd_filtered[p_nd_filtered$do_est>8, ]
-
-saveRDS(p_nd, file = (here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
-```
-
-Save new predictions without QCS gap and WCVI 2016 DO
-```{r eval=FALSE}
-#new temp and do prediction grid for vectors
-
-# model1 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-temp-2008-on-used.rds")
-#model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-all-DOY-used.rds")
-
-# fix RW additon error
 model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds"))
-# update model to exclude WCVI 2016 outliers
-model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds"))
+model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth2.rds"))
 
@@
 # get scaling used in model
-d <- all_sensor %>%
-  # filter missing do data 
-  dplyr::filter(!is.na(do_mlpl)) %>%
-  # remove 2007 data because clearly faulty or scaled differently
-  dplyr::filter(year>2007)
-d$temp_scaled <- scale(d$temperature_c)
-attributes(d$temp_scaled)
+sensor_temp_scaled <- scale(d_trawl$temp)
+attributes(sensor_temp_scaled)
 
-predtemp <- pred_temp %>% mutate(temp_scaled = (est - 6.49 )/1.296, temp_scaled2 = ((est - 6.49 )/1.296)^2, depth_scaled3 = depth_scaled^3) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
+predtemp <- pred_temp %>% mutate(
+  temp_scaled = (est - attributes(sensor_temp_scaled)[[2]] )/attributes(sensor_temp_scaled)[[3]], 
+  temp_scaled2 = (temp_scaled)^2) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
 predtemp <- predtemp %>% filter(year > 2007) 
@@
 
+# make DO predictions
 pred_do <- predict(model2, newdata = predtemp)
 
-pred_do$temp <- pred_do$temp_scaled*1.296 + 6.49
-
+# rename and back transform climate variables
+pred_do$temp <- pred_do$temp_scaled * attributes(sensor_temp_scaled)[[3]] + attributes(sensor_temp_scaled)[[2]]
 pred_do$log_do <- pred_do$est
@@
 
-saveRDS(pred_do, file=here::here("analysis/VOCC/data/predicted-DO-june-2020.rds"))
-
+saveRDS(pred_do, file=here::here("analysis/VOCC/data/predicted-DO-depth2.rds"))
 ```
@@
 Plot predictions from saved DO predictions file
-
 ```{r}
-#p_nd <- readRDS((here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds")))
-# p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-new.rds")))
-
-p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-june-2020.rds")))
+p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-depth2.rds")))
 
@@
 
+Plot each survey separately
 ```{r}
@@
 
-breaks <- scales::trans_breaks(no_trans[["transform"]], no_trans[["inverse"]],n = 6)
+breaks <- scales::trans_breaks(no_trans[["transform"]], no_trans[["inverse"]], n = 6)
 labels <- function(x) format(x, digits = 2, scientific = FALSE)
@@
 ```{r, eval=FALSE}
-#glimpse(p_nd)
-
 omega_s_all <- plot_map_raster(p_nd, "omega_s") +
@@
 )
-
-# gridExtra::grid.arrange( nrow = 2, heights = c(2,1.5), 
-#   top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")) 
 dev.off()
 ```
-
-```{r}
-#do_cog <- get_cog(p_nd)
-
-#temp_cog %>%
-#  reshape2::dcast(year ~ coord, value.var = "est") %>%
-#  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
-#  scale_color_viridis_c()
-```
-
-```{r}
-#temp_cog %>%
-#  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
-#  geom_line() +
-#  geom_ribbon(alpha = 0.2) +
-#  facet_wrap(~coord, scales = "free_y")
-```

### 2020-06-19 analysis/tmb-sensor-explore/3-fit-do.Rmd
refining climate models

@@
 
-# get scaling used in model
-sensor_temp_scaled <- scale(d_trawl$temp)
-attributes(sensor_temp_scaled)
-
 predtemp <- pred_temp %>% mutate(
-  temp_scaled = (est - attributes(sensor_temp_scaled)[[2]] )/attributes(sensor_temp_scaled)[[3]], 
+  temp = est,
+  temp_omega = omega_s,
+  temp_epsilon = epsilon_st,
+  temp_scaled = (est - model2$data$temp_mean[1])/model2$data$temp_sd[1], 
   temp_scaled2 = (temp_scaled)^2) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
@@
 # rename and back transform climate variables
-pred_do$temp <- pred_do$temp_scaled * attributes(sensor_temp_scaled)[[3]] + attributes(sensor_temp_scaled)[[2]]
 pred_do$log_do <- pred_do$est
@@
 # pred_do <- pred_do %>% filter(depth>15)
-
-saveRDS(pred_do, file=here::here("analysis/VOCC/data/predicted-DO-depth2.rds"))
+saveRDS(pred_do, file=here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(),".rds")))
 ```

### 2020-06-19 analysis/tmb-sensor-explore/climate-predictions.R

@@ -1,41 +1,46 @@
-### new temp and do prediction grid for vectors
-# model1 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-temp-2008-on-used.rds")
-#model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-all-DOY-used.rds")
-# update model to exclude WCVI 2016 outliers 
-# model2 <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds")
+### REFINING CLIMATE PREDICTIONS
+library(dplyr)
+library(sdmTMB)
 
-# fix RW additon error
+# load temp model
 model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds"))
-# update model to exclude WCVI 2016 outliers
+max(model1$gradients)
+
+model1 <- run_extra_optimization(model1)
+max(model1$gradients)
+
+
+# load DO model 
+model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth2.rds")) 
 model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds"))
 
+max(model2$gradients)
+
+model2 <- run_extra_optimization(model2)
+max(model2$gradients)
+
+# load prediction grid
 nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
 
 pred_temp <- predict(model1, newdata = nd_all)
 
-# get scaling used in model
-d <- all_sensor %>%
-  # filter missing do data 
-  dplyr::filter(!is.na(do_mlpl)) %>%
-  # remove 2007 data because clearly faulty or scaled differently
-  dplyr::filter(year>2007)
-d$temp_scaled <- scale(d$temperature_c)
-attributes(d$temp_scaled)
-
-predtemp <- pred_temp %>% mutate(temp_scaled = (est - 6.49 )/1.296, temp_scaled2 = ((est - 6.49 )/1.296)^2, depth_scaled3 = depth_scaled^3) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
+predtemp <- pred_temp %>% mutate(
+  temp = est,
[…]

### 2020-06-20 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
improve convergence

@@
   include_spatial = TRUE, 
-  time = "year", spde = spde, family = gaussian(link = "identity"),
+  time = "year", spde = spde, 
+  family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
   silent = FALSE
@@
   time = "year", spde = spde2, family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
   silent = TRUE

### 2020-06-20 analysis/tmb-sensor-explore/3-fit-do.Rmd
improve convergence

@@
 library(gfranges)
-```
-
-Load data
-```{r}
-all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/all-sensor-data-processed.rds"))
-# glimpse(all_sensor)
-
-d_trawl <- all_sensor %>%
-  # filter missing do data 
-  dplyr::filter(!is.na(do_mlpl)) %>%
-  # remove 2007 data because clearly faulty or scaled differently
-  dplyr::filter(year>2007) %>%
-  # add day of year variable
-  mutate(DOY = lubridate::yday(fe_event_start), temp = temperature_c, 
-    raw_depth = depth, depth = log(depth), log_do = log(do_mlpl)) %>%
-  # scale potential covariates across all surveys
-  scale_predictors(predictors = c(quo(depth), quo(DOY), quo(temp)))
-
-d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
-d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
-
-d_trawl <- d_trawl %>% mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
-  filter(exclude != 1)
-
-# glimpse(d_trawl)
-```
-
-Ranges sampled
-```{r}
-paste("depth") 
-range(d_trawl$raw_depth)
-paste("DO") 
-range(d_trawl$do_mlpl)
 ```
@@
 }
+```
+
+Load data
+```{r}
+all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/all-sensor-data-processed.rds"))
+# glimpse(all_sensor)
+
+d_trawl <- all_sensor %>%
+  # filter missing do data 
+  dplyr::filter(!is.na(do_mlpl)) %>%
+  # remove 2007 data because clearly faulty or scaled differently
+  dplyr::filter(year>2007) %>%
+  # add day of year variable
+  mutate(DOY = lubridate::yday(fe_event_start), temp = temperature_c, 
+    raw_depth = depth, depth = log(depth), log_do = log(do_mlpl)) %>%
+  # scale potential covariates across all surveys
+  scale_predictors(predictors = c(quo(depth), quo(DOY), quo(temp)))
+
+d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
+d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
+
+d_trawl <- d_trawl %>% mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
+  filter(exclude != 1)
+
+# glimpse(d_trawl)
+```
+
+Ranges sampled
+```{r}
+paste("depth") 
+range(d_trawl$raw_depth)
+paste("DO") 
+range(d_trawl$do_mlpl)
 ```
@@
       enable_priors = FALSE,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      nlminb_loops = 2,
+      newton_steps = 1,
       silent = FALSE

### 2020-06-20 analysis/tmb-sensor-explore/3-fit-do.Rmd
DO model gradient improved by 800kn

@@
 
-spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 500) 
+spde <- make_spde(d_trawl1$X, d_trawl1$Y, n_knots = 800) 
 plot_spde(spde)
@@
 
-saveRDS(do_model_all_DOY, file = (here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016.rds")))
+saveRDS(do_model_all_DOY, file = (here::here(
+  paste0("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-", 
+    "800kn", #Sys.Date(), 
+    ".rds")))
+  )
 ```
@@
 ```{r eval=FALSE}
+do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds")))
 get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
@@
 model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds"))
-model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth2.rds"))
+model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds"))
 
@@
 ```{r}
-p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-depth2.rds")))
+p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-19.rds")))
 
@@
   scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), 
-    limits = c(1, 8), 
-    na.value = "red") +
+    limits = c(0, 7), 
+    na.value = "gold") +
   facet_wrap(~year) +
@@
 pred_4 <- plot_map_raster(p_nd4, "do_est") +
-  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(1, 7), 
+  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 7), 
      na.value = "red")+
@@
 pred_4
+```
+
+```{r}
+p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(year!=2016)
+pred_16 <- plot_map_raster(p_nd16, "do_est") +
+  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 7), 
+     na.value = "red")+
+  facet_wrap(~year) +
+  labs(fill = "ml/L DO") +
+  #ggtitle(" ")+ 
+  theme(legend.position = "none")
+pred_16
 ```
@@
   facet_wrap(~year) +
-  scale_fill_viridis_c()
+  scale_fill_gradient2( high = "Steel Blue 4",
+  low = "Red 3")#, limits = c(-0.65,1)
+  # scale_fill_viridis_c()
 ```

### 2020-06-20 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
other temp models

--- 2020-06-19 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
@@
 ```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 500)
+spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 800)
 plot_spde(spde)
@@
   data = d_trawl,
-  ar1_fields = TRUE, # changed AR1_field to true
+  ar1_fields = TRUE, # changed AR1 to true
   include_spatial = TRUE, 
-  time = "year", spde = spde, family = gaussian(link = "identity"),
+  time = "year", spde = spde, 
+  family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
   silent = FALSE
@@
 
-saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
+saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds")))
 ```
@@
 ```{r}
+m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds"))
 predictions <- predict(m_temp_rw)
@@
 ```{r eval=FALSE}
-m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
+m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds")))
 get_diag(m_temp_rw, response = "temperature_c", variable = "depth_scaled")
@@
 ```{r eval=FALSE}
+m_temp2 <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year) + DOY_scaled * shallow + DOY_scaled2 * shallow,
+  data = d_trawl2,
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  ar1_fields = TRUE, include_spatial = TRUE,
+  time = "year", spde = spde2, family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
+  silent = TRUE
+)
+
+saveRDS(m_temp2, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY.rds")))
+
+m_temp2
+```
+
+```{r eval=FALSE}
 m_temp3 <- sdmTMB(
@@
   time = "year", spde = spde2, family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
   silent = TRUE
 )
+
+saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds")))
+
+m_temp3
 ```
@@
 cor(predictions3$est, predictions3$temperature_c)^2 #
+```
+
+```{r eval=FALSE}
+get_diag(m_temp3, response = "temperature_c", variable = "depth_scaled")
 ```

### 2020-06-24 analysis/VOCC/3-vocc-dist.Rmd
remove duplication in vocc calculations

@@
     # species: "Redbanded Rockfish"
-    species: "Widow Rockfish"
+    # species: "Widow Rockfish"
     # species: "Canary Rockfish"
@@
     # species: "Sablefish"
-    # species: "Arrowtooth Flounder"
-    immature: F
-    region: "West Coast Vancouver Island"
+    species: "Arrowtooth Flounder"
+    immature: T
+    # region: "West Coast Vancouver Island"
     # region: "both odd year surveys"
-    # region: "West Coast Haida Gwaii"
+    region: "West Coast Haida Gwaii"
     covs: "-tv-depth-only"
+    climate_mod: "2020-06-20-more2016"
     scrambled_years: FALSE
@@
 covs <- params$covs
+climate_mod <- params$climate_mod
+
 immature <- params$immature
@@
 
-#### TRIM environment to range sampled
+#### Load climate data
 ```{r}
-nd_all <- readRDS(paste0("data/predicted-DO-june-2020.rds")) %>% select(X, Y, year, temp, do_est)
-fishing <- readRDS("data/_fishing_effort/fishing-effort-grid.rds")
-# fishing <- readRDS("data/fishing-effort-grid-6hr.rds")
+nd_all <- readRDS(paste0("data/predicted-DO-", climate_mod, ".rds")) %>% select(X, Y, year, temp, do_est)
 
 d <- left_join(d, nd_all, by = c("X", "Y", "year"))
-d <- left_join(d, fishing, by = c("X", "Y", "year"))
-d$log_effort[is.na(d$log_effort)] <- -2.3
 
 glimpse(d)
-
-# d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
-d <- d %>%
-  filter(do_est > 0.28) %>%
-  filter(do_est < 7.06) # 0.005 and 0.995
-# d <- d %>% filter(temp > 2.61) %>% filter(temp < 14.31) # full range
-d <- d %>%
-  filter(temp > 3.07) %>%
-  filter(temp < 11.3) # 0.005 and 0.995
-```
-
-Calculate mean and trend in log fishing pressure
-```{r }
-gf <- vocc_gradient_calc(d, "log_effort",
-  scale_fac = scale_fac,
-  # all layers with max indice will be averaged for spatial gradient
-  # if default (NULL) will use all years
-  # indices = indices,
-  # divisor = 1, #default of 10 gives per decade
-  quantile_cutoff = 0.05
-)
-
-gf <- gf %>%
-  mutate(
-    fishing_trend = trend,
-    mean_effort = exp(mean)
-  ) %>%
-  dplyr::mutate(X = x, Y = y) %>%
-  gfplot:::utm2ll(., utm_zone = 9)
-
-gf$squashed_effort <- collapse_outliers(gf$mean_effort, c(0, 0.95))
-
-# fished <- gf %>%
-#   select(x, y, fishing_trend, mean_effort)
-# saveRDS(fished, file = paste0("data/fishing-effort-trend-6hr-", ssid_string, ".rds"))
-```
-
-```{r eval = F}
-effort <- plot_vocc(gf,
-  raster_cell_size = 4,
-  # theme_black = TRUE,
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  # NA_label = "NA",
-  fill_col = "squashed_effort",
-  fill_label = "mean effort",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  # high_fill = "Steel Blue 4",
-  # low_fill = "Red 3",
-  # white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = fourth_root_power #
-) + labs(
-  title = "Fishing presssure",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-effort
-
-fish_t <- plot_vocc(gf,
-  raster_cell_size = 4,
-  # theme_black = TRUE,
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  # NA_label = "NA",
-  fill_col = "fishing_trend",
-  fill_label = "trend in fishing effort",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  # high_fill = "Steel Blue 4",
-  # low_fill = "Red 3",
-  # white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = sqrt # no_trans #
-) + labs(
-  title = "Fishing presssure",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-fish_t
 ```
@@
 ```{r}
+rm(gf2)
+try({gf2 <- readRDS(paste0("data/_climate_vocc/temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+})
+
+if(!exists("gf2")){
+
 gf2 <- vocc_gradient_calc(d, "temp",
@@
 gf2 <- left_join(gf2, dv_t)
-```
-
-```{r eval=F}
-plot_vocc(gf2,
-  # theme = "black",
-  raster_cell_size = 4,
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "dvocc",
-  fill_label = "temperature \ndvocc",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "black",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans # fourth_root_power # sqrt #
-) + labs(
-  title = "distance-based temp velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year), " min temp SD = ", signif(temp_min_sd, 2), " x 2 for 95% min threshold of change")
-)
-```
-
-```{r eval=F}
-gf2$squashed_velocity <- collapse_outliers(gf2$velocity, c(0.005, 0.995))
-
-plot_vocc(gf2,
-  # theme = "black",
-  raster_cell_size = 4,
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "squashed_velocity",
-  fill_label = "temperature \nvelocity",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans # fourth_root_power # sqrt #
-) + labs(
-  title = "gradient-based temp velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-```
-
-```{r eval=F}
-grad2 <- plot_vocc(gf2,
-  raster_cell_size = 4,
-  # theme = "black",
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "trend",
-  fill_label = "temperature \ntrend",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans # log10 # fourth_root_power # sqrt #
-) + labs(
-  title = "temp trend",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-grad2
+
+saveRDS(gf2, file = paste0("data/_climate_vocc/temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+}
+
+# mean(gf2$sd, na.rm = T) # [1] 0.3274
+temp_min_sd <- min(gf2$sd, na.rm = T) # [1] 0.1035
+# max(gf2$sd, na.rm = T) # [1] 1.248
+temp_median_sd <- median(gf2$sd, na.rm = T) # [1] 0.2882
+
+# temp_sd_choice <- temp_min_sd
+# lower_change_t <- temp_sd_choice*2
+# upper_change_t <- temp_sd_choice*2
+
 ```
@@
 ```{r }
+
+rm(gf0)
+try({gf0 <- readRDS(paste0("data/_climate_vocc/DO-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+})
+
+if(!exists("gf0")){
+
 gf0 <- vocc_gradient_calc(d, "do_est",
@@
 
-```
-
-```{r eval=F}
-grad0dv <- plot_vocc(gf0,
-  raster_cell_size = 4,
-  # theme = "black",
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "dvocc",
-  fill_label = "DO and temp \ndvocc",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  white_zero = TRUE,
-  high_fill = "Steel Blue 4",
-  low_fill = "Red 3",
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans # fourth_root_power # sqrt #
-) + labs(
-  title = "distance-based DO and temp velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year), " DO SD = ", signif(do_sd_choice, 2), " x 2 for threshold for do change")
-)
-
-grad0dv
-```
-
-```{r eval=F}
-gf0$squashed_velocity <- collapse_outliers(gf0$velocity, c(0.005, 0.995))
-
-grad0v <- plot_vocc(gf0,
-  raster_cell_size = 4,
-  # theme_black = TRUE,
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "squashed_velocity",
-  fill_label = "DO velocity",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "white",
-  high_fill = "Steel Blue 4",
-  low_fill = "Red 3",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans # fourth_root_power
-) + labs(
-  title = "do velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-grad0v
-```
-
-```{r eval=F}
-grad0 <- plot_vocc(gf0,
-  raster_cell_size = 4,
-  # theme_black = TRUE,
-  coast = TRUE,
-  vec_aes = NULL,
-  # grad_vec_aes = "velocity",
-  # vec_lwd_range = c(0.5, 0.5),
-  NA_label = "NA",
-  fill_col = "trend",
-  fill_label = "DO trend",
-  raster_alpha = 1,
-  # raster_limits = c(raster_min, raster_max),
-  na_colour = "black",
-  high_fill = "Steel Blue 4",
-  low_fill = "Red 3",
-  white_zero = TRUE,
-  vec_alpha = 0.35,
-  axis_lables = FALSE,
-  transform_col = no_trans #
-) + labs(
-  title = "do trend",
-  subtitle = paste0(min(d$year), "-", max(d$year))
-)
-
-grad0
+saveRDS(gf0, file = paste0("data/_climate_vocc/DO-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+}
+
+do_min_sd <- min(gf0$sd, na.rm = T)
+do_median_sd <- median(gf0$sd, na.rm = T)
+do_sd_choice <- do_median_sd
 ```
@@
     ## if default of divisor = 10
-      trend_per_year = (trend/10), # percent change not raw
+      trend_per_year = exp(trend/10), # percent change not raw
       trend_per_decade = trend_per_year * 10,
@@
 
-fished <- gf %>%
-  select(x, y, fishing_trend, mean_effort)
-
 DO_values <- gf0 %>%
@@
     DO_dvocc = dvocc,
-    DO_trend = trend_per_year, # should change to trend_per_decade
+    DO_trend = trend_per_decade, 
     DO_grad = gradient,
@@
     temp_dvocc = dvocc,
-    temp_trend = trend_per_year, # should change to trend_per_decade
+    temp_trend = trend_per_decade, 
     temp_grad = gradient,
@@
 
+fished <- readRDS("data/fishing-effort-change.rds")
+
 vocc <- left_join(biotic_values, fished) # add fishing effort
@@
 write.csv(vocc, file = paste0(
-  "data/_newclim_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_more2016_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))
-
 ```
@@
 vocc  <- read.csv(paste0(
-  "data/_newclim_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_more2016_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))

### 2020-06-24 analysis/VOCC/4-fishing-trend.Rmd
recalculate fishing trend and add depth to 4x4 grids

---
title: "Fishing trends"
output: html_notebook
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.keep = "all", fig.show = "asis", dpi = 600,
  fig.width = 6, fig.height = 5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("gfranges")
library("dplyr")
library("ggplot2")

ggplot2::theme_set(gfplot::theme_pbs())
```

#### Set cell sizes and scale
```{r, echo=TRUE}
scale_fac <- 2 # means that the raster is reprojected to _ X original grid (2 km)
```

```{r}
d_all <- readRDS(paste0(
    "data/sablefish/predictions-sablefish-tv-depth-only-1n3n4n16-mature-biomass-prior-FALSE.rds"
  )) %>% filter(year >= 2008)

all_years <- unique(d_all$year)
length(all_years)
grid <- select(d_all, X,Y, depth, ssid) %>% unique()

grid_by_year <- data.frame(matrix(ncol=0, nrow= nrow(grid) * length(all_years)))
grid_by_year$ssid <- rep(grid$ssid, times = length(all_years))
grid_by_year$X <- rep(grid$X, times = length(all_years))
grid_by_year$Y <- rep(grid$Y, times = length(all_years))
grid_by_year$depth <- rep(grid$depth, times = length(all_years))
grid_by_year$year <- rep(all_years, each = nrow(grid))
```

Calculate mean and trend in log fishing pressure
```{r }
fishing <- readRDS("data/_fishing_effort/fishing-effort-grid.rds")

d <- left_join(grid_by_year, fishing, by = c("X", "Y", "year"))

d$log_effort[is.na(d$log_effort)] <- -2.3
d$effort[is.na(d$effort)] <- 0

### BOTH ODD YEAR ###
d1n3 <- d %>% filter(ssid %in% c(1,3))

ssid1n3 <- vocc_gradient_calc(d1n3, "log_effort",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)

depth1n3 <- vocc_gradient_calc(d1n3, "depth",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
ssid1n3$depth <- depth1n3$mean

d4 <- d %>% filter(ssid %in% c(4))

### WCVI ###
ssid4 <- vocc_gradient_calc(d4, "log_effort",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
depth4 <- vocc_gradient_calc(d4, "depth",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
ssid4$depth <- depth4$mean

### WCHG ###
d16 <- d %>% filter(ssid %in% c(16))

ssid16 <- vocc_gradient_calc(d16, "log_effort",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
depth16 <- vocc_gradient_calc(d16, "depth",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
ssid16$depth <- depth16$mean

fished <- rbind(ssid1n3, ssid4, ssid16)

fished <- fished %>%
  mutate(
    fishing_trend = trend,
    mean_effort = exp(mean),
    fishing_vel = velocity,
    fishing_grad = gradient
  ) %>%
  dplyr::mutate(X = x, Y = y) %>%
  gfplot:::utm2ll(., utm_zone = 9)

# fished$squashed_effort <- collapse_outliers(fished$mean_effort, c(0, 0.99))
# hist(fished$squashed_effort)

# fished <- left_join(fished, grid)
fished <- fished %>%
  select(X, Y, x, y, fishing_trend, mean_effort, fishing_vel, fishing_grad, depth)

saveRDS(fished, file = paste0("data/fishing-effort-change-w-depth.rds"))
```

### 2020-06-24 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
add good DO values from WCVI 2016 into models and predictions

@@
 ```{r}
-m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds"))
+# m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds"))
+m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds"))
 predictions <- predict(m_temp_rw)
@@
 cor(predictions$est, predictions$temperature_c)^2 
+sum(abs(predictions$residuals))
+
+# 2908.879-2789.322
+# 119.557/2908.879 # 4% reduction in abs errors
 ```
@@
   g <- ggplot(predictions, aes_string("est", response)) +
-    geom_point() +
+    geom_point(alpha = 0.2) +
     facet_wrap(~year) +
@@
   g <- ggplot(predictions, aes(est, residuals)) +
-    geom_point() +
+    geom_point(alpha = 0.2) +
     geom_smooth()
@@
   g <- ggplot(predictions, aes_string(variable, "residuals")) +
-    geom_point() +
+    geom_point(alpha = 0.2) +
     geom_smooth() +
@@
   g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
-    geom_point() +
+    geom_point(alpha = 0.2) +
     coord_fixed() +
@@
 ```{r }
-m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
+# m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
 nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
@@
 glimpse(nd_all)
-pred_temp <- predict(m_temp_rw, newdata = nd_all)
+p_nd <- predict(m_temp_rw, newdata = nd_all)
 
-# saveRDS(pred_temp, file = "../VOCC/data/predicted_temp_allyears_revised2")
+saveRDS(p_nd, file = "../VOCC/data/predicted_temp_allyears_800kn.rds")
 ```
@@
 ```{r}
-p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised2")
+# p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")
 
@@
    scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
+ high = "Red 3", limits = c(-1.5,1.5), #trans = sqrt,
     na.value = "red") +
@@
      scale_fill_gradient2( low = "Steel Blue 4",
- high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
+ high = "Red 3", limits = c(-1.5,1.5), #trans = sqrt,
     na.value = "red") +

### 2020-06-24 analysis/tmb-sensor-explore/3-fit-do.Rmd
add good DO values from WCVI 2016 into models and predictions

@@
   print(aic(m))
+
 }
@@
 ```{r}
+ggplot(filter(d_trawl, depth_m_max < 80), aes(DOY, log_do, colour = depth)) +
+  geom_point() +
+  facet_wrap(~year) +
+  scale_color_viridis_c()
+ggplot(filter(d_trawl, depth_m_max > 80 & depth_m_max < 120), aes(DOY, log_do, colour = depth)) +
+  geom_point() +
+  facet_wrap(~year) +
+  scale_color_viridis_c()
+```
+
+```{r}
 ggplot(d_trawl, aes(depth, do_mlpl, colour = DOY, size = sqrt(do_change))) +
@@
 ```{r}
-ggplot(d_trawl, aes(temp, do_mlpl, colour = as.factor(ssid))) +
+ggplot(filter(d_trawl, ssid == 4), aes(DOY,log_do,  colour = temp)) +
   geom_point() +
-  facet_wrap(~year) +
-  scale_color_viridis_d()
+  facet_grid(ssid~year, scales = "free") +
+  scale_color_viridis_c()
 ```
@@
 ```{r}
-ggplot(filter(d_trawl, year==2016), aes(temp, do_mlpl, colour = as.factor(ssid))) +
+ggplot(filter(d_trawl, ssid == 4), aes(DOY,temp,  colour = temp)) +
+  geom_point() +
+  facet_grid(ssid~year, scales = "free") +
+  scale_color_viridis_c()
+```
+
+```{r}
+ggplot(filter(d_trawl, ssid == 4 & !(year ==2016 & DOY > 149)), aes(X, Y, colour = log_do)) +
+  geom_point() +
+  facet_grid(ssid~year, scales = "free") +
+  scale_color_viridis_c()
+```
+
+```{r}
+ggplot(filter(d_trawl, year==2016), aes((temp), log(do_mlpl), colour = as.factor(ssid))) +
   geom_point() +
@@
 ```{r}
-ggplot(filter(d_trawl, year==2017), aes(temp, do_mlpl, colour = as.factor(ssid))) +
+ggplot(filter(d_trawl, year==2017), aes(x = (temp), log(do_mlpl), colour = as.factor(ssid))) +
   geom_point() +
@@
 
-Remove WCVI in 2016 due to extreme values and make 500 knot mesh
+Remove most samples from WCVI in 2016 due to extreme values and make mesh
 ```{r}
-d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4))
+d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4 & DOY > 149))
 
@@
 saveRDS(do_model_all_DOY, file = (here::here(
-  paste0("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-", 
+  paste0("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-", 
     "800kn", #Sys.Date(), 
@@
 ```{r eval=FALSE}
-do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds")))
+# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth3.rds")))
+# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth2.rds")))
+# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds")))
 get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
+```
+
+```{r}
+predictions <- predict(do_model_all_DOY)
+predictions$residuals <- residuals(do_model_all_DOY)
+predictions$abs <- exp(abs(predictions$residuals))
+
+good_rows <- filter (predictions, abs != "Inf")
+sum(good_rows$abs)
+19719.73-17459.49
+2260.24/19719.73  # 11% reduction in abs errors
 ```
@@
 ```{r eval=FALSE}
-model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds"))
-model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds"))
+model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds"))
+model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-800kn.rds"))
 
@@
 # pred_do <- pred_do %>% filter(depth>15)
-saveRDS(pred_do, file=here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(),".rds")))
+saveRDS(pred_do, file=here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(),"-more2016.rds")))
 ```
@@
 ```{r}
-p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-19.rds")))
+p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016.rds")))
 

### 2020-06-25 analysis/tmb-sensor-explore/3-fit-do.Rmd
style fit-do

@@
     coord_fixed() +
-    scale_color_gradient2()+
+    scale_color_gradient2() +
     facet_wrap(~year)
@@
   print(aic(m))
-
 }
@@
   # remove 2007 data because clearly faulty or scaled differently
-  dplyr::filter(year>2007) %>%
+  dplyr::filter(year > 2007) %>%
   # add day of year variable
-  mutate(DOY = lubridate::yday(fe_event_start), temp = temperature_c, 
-    raw_depth = depth, depth = log(depth), log_do = log(do_mlpl)) %>%
+  mutate(
+    DOY = lubridate::yday(fe_event_start), temp = temperature_c,
+    raw_depth = depth, depth = log(depth), log_do = log(do_mlpl)
+  ) %>%
   # scale potential covariates across all surveys
@@
 
-d_trawl <- d_trawl %>% mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
+d_trawl <- d_trawl %>%
+  mutate(exclude = if_else(do_mlpl > 8, 1, 0)) %>%
   filter(exclude != 1)
@@
 ```{r}
-ggplot(d_trawl, aes(X,Y, colour = do_mlpl)) +
+ggplot(d_trawl, aes(X, Y, colour = do_mlpl)) +
   geom_point() +
@@
 ```{r}
-ggplot(d_trawl, aes(X,Y, colour = temp)) +
+ggplot(d_trawl, aes(X, Y, colour = temp)) +
   geom_point() +
@@
   scale_color_viridis_c()
-
 ```
@@
 ```{r}
-ggplot(filter(d_trawl, ssid == 4), aes(DOY,log_do,  colour = temp)) +
+ggplot(filter(d_trawl, ssid == 4), aes(DOY, log_do, colour = temp)) +
   geom_point() +
-  facet_grid(ssid~year, scales = "free") +
+  facet_grid(ssid ~ year, scales = "free") +
   scale_color_viridis_c()
@@
 ```{r}
-ggplot(filter(d_trawl, ssid == 4), aes(DOY,temp,  colour = temp)) +
+ggplot(filter(d_trawl, ssid == 4), aes(DOY, temp, colour = temp)) +
   geom_point() +
-  facet_grid(ssid~year, scales = "free") +
+  facet_grid(ssid ~ year, scales = "free") +
   scale_color_viridis_c()
@@
 ```{r}
-ggplot(filter(d_trawl, ssid == 4 & !(year ==2016 & DOY > 149)), aes(X, Y, colour = log_do)) +
+ggplot(filter(d_trawl, ssid == 4 & !(year == 2016 & DOY > 149)), aes(X, Y, colour = log_do)) +
   geom_point() +
-  facet_grid(ssid~year, scales = "free") +
+  facet_grid(ssid ~ year, scales = "free") +
   scale_color_viridis_c()
@@
 ```{r}
-ggplot(filter(d_trawl, year==2016), aes((temp), log(do_mlpl), colour = as.factor(ssid))) +
+ggplot(filter(d_trawl, year == 2016), aes((temp), log(do_mlpl), colour = as.factor(ssid))) +
   geom_point() +
-  #facet_wrap(~year) +
-  scale_color_viridis_d(begin=0.65)
+  # facet_wrap(~year) +
+  scale_color_viridis_d(begin = 0.65)
 ```
@@
 ```{r}
-ggplot(filter(d_trawl, year==2017), aes(x = (temp), log(do_mlpl), colour = as.factor(ssid))) +
+ggplot(filter(d_trawl, year == 2017), aes(x = (temp), log(do_mlpl), colour = as.factor(ssid))) +
   geom_point() +
-  #facet_wrap(~year) +
+  # facet_wrap(~year) +
   scale_color_viridis_d(end = 0.35)
@@
 ```{r}
-ggplot(d_trawl, aes(DOY, -depth , colour = do_mlpl)) + geom_point() + facet_wrap(~year) +
+ggplot(d_trawl, aes(DOY, -depth, colour = do_mlpl)) + geom_point() + facet_wrap(~year) +
   scale_color_viridis_c() + ylab("-log(depth)")
@@
 ```{r}
-d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4 & DOY > 149))
+d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid == 4 & DOY > 149))
 
@@
 saveRDS(do_model_all_DOY, file = (here::here(
-  paste0("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-", 
-    "800kn", #Sys.Date(), 
-    ".rds")))
+  paste0(
+    "analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-",
+    "800kn", # Sys.Date(),
+    ".rds"
   )
+)))
 ```
@@
 predictions$residuals <- residuals(do_model_all_DOY)
-predictions$abs <- exp(abs(predictions$residuals))
 
-good_rows <- filter (predictions, abs != "Inf")
-sum(good_rows$abs)
-19719.73-17459.49
-2260.24/19719.73  # 11% reduction in abs errors
+# predictions$abs <- exp(abs(predictions$residuals))
+# pred_rows <- filter (predictions, abs != "Inf")
+# sum(pred_rows$abs) # sum absolute error
+# 19719.73-17459.49 # difference in absolute error between 500 and 800 knots
+# 2260.24/19719.73  # 11% reduction in abs errors
 ```
@@
 
-nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
+nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year < 2019)
 
@@
 
-predtemp <- pred_temp %>% mutate(
+predtemp <- pred_temp %>%
+  mutate(
     temp = est,
@@
     temp_epsilon = epsilon_st,
-  temp_scaled = (est - model2$data$temp_mean[1])/model2$data$temp_sd[1], 
-  temp_scaled2 = (temp_scaled)^2) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
+    temp_scaled = (est - model2$data$temp_mean[1]) / model2$data$temp_sd[1],
+    temp_scaled2 = (temp_scaled)^2
+  ) %>%
+  select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
 predtemp <- predtemp %>% filter(year > 2007)
@@
 
-# may want to filter estimates for depths less than 15 m
-# pred_do <- pred_do %>% filter(depth>15)
-saveRDS(pred_do, file=here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(),"-more2016.rds")))
+saveRDS(pred_do, file = here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(), "-more2016.rds")))
 ```
@@
     geom_raster() +
-    coord_fixed()+
+    coord_fixed() +
     gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
@@
 ```{r}
-p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
+p_nd1 <- p_nd %>%
+  filter(ssid != 16) %>%
+  filter(ssid != 4)
 
@@
 pred_1 <- plot_map_raster(p_nd1, "do_est") +
-  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), 
+  scale_fill_viridis_c(
+    labels = labels, breaks = c(0, 2, 4, 6, 8),
     limits = c(0, 7),
-    na.value = "gold") +
+    na.value = "gold"
+  ) +
   facet_wrap(~year) +
@@
   ggtitle("Prediction (fixed effects + all random effects)") +
-  theme(legend.position = c(.85,.25))
+  theme(legend.position = c(.85, .25))
 pred_1
@@
 ```{r}
-p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(year!=2016)
+p_nd4 <- p_nd %>% filter(ssid == 4) # %>% filter(year!=2016)
 pred_4 <- plot_map_raster(p_nd4, "do_est") +
-  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 7), 
-     na.value = "red")+
+  scale_fill_viridis_c(
+    labels = labels, breaks = c(0, 2, 4, 6, 8), limits = c(0, 7),
+    na.value = "red"
+  ) +
   facet_wrap(~year) +
   labs(fill = "ml/L DO") +
-  #ggtitle(" ")+ 
+  # ggtitle(" ")+
   theme(legend.position = "none")
@@
 ```{r}
-p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(year!=2016)
+p_nd16 <- p_nd %>% filter(ssid == 16) # %>% filter(year!=2016)
 pred_16 <- plot_map_raster(p_nd16, "do_est") +
-  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 7), 
-     na.value = "red")+
+  scale_fill_viridis_c(
+    labels = labels, breaks = c(0, 2, 4, 6, 8), limits = c(0, 7),
+    na.value = "red"
+  ) +
   facet_wrap(~year) +
   labs(fill = "ml/L DO") +
-  #ggtitle(" ")+ 
+  # ggtitle(" ")+
   theme(legend.position = "none")
@@
   facet_wrap(~year) +
-  scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3")#, limits = c(-0.65,1)
+  scale_fill_gradient2(
+    high = "Steel Blue 4",
+    low = "Red 3"
+  ) # , limits = c(-0.65,1)
 # scale_fill_viridis_c()
@@
   ggtitle("Spatial random effects only") +
-   scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3", limits = c(-0.65,1))+
+  scale_fill_gradient2(
+    high = "Steel Blue 4",
+    low = "Red 3", limits = c(-0.65, 1)
+  ) +
   labs(fill = " ") +
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))
+  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82, .75))
 
@@
   facet_wrap(~year) +
-  scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3")
+  scale_fill_gradient2(
+    high = "Steel Blue 4",
+    low = "Red 3"
+  )
 ```
@@
 ```{r}
-p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
+p_nd1 <- p_nd %>%
+  filter(ssid != 16) %>%
+  filter(ssid != 4)
 st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
@@
   facet_wrap(~year) +
-  scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3", limits = c(-0.5,0.5)) +
+  scale_fill_gradient2(
+    high = "Steel Blue 4",
+    low = "Red 3", limits = c(-0.5, 0.5)
+  ) +
   labs(fill = " ") +
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.25))
+  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85, .25))
 st_1n3
@@
 ```{r}
-p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(year!=2016)
+p_nd4 <- p_nd %>% filter(ssid == 4) # %>% filter(year!=2016)
 st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
-  #ggtitle(" ") +
+  # ggtitle(" ") +
   facet_wrap(~year) +
-  scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3", limits = c(-0.5,0.5)) +
+  scale_fill_gradient2(
+    high = "Steel Blue 4",
+    low = "Red 3", limits = c(-0.5, 0.5)
+  ) +
   labs(fill = "epsilon") +
@@
 ```{r}
-p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
+p_nd16 <- p_nd %>% filter(ssid == 16) # %>% filter(ssid!=4)
 st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
-  #ggtitle(" ") +
+  # ggtitle(" ") +
   facet_wrap(~year) +
-  scale_fill_gradient2( high = "Steel Blue 4",
-  low = "Red 3", limits = c(-0.5,0.5)) +
-  xlim(65,345) +
-  ylim(5800,6100) +
+  scale_fill_gradient2(
+    high = "Steel Blue 4",
+    low = "Red 3", limits = c(-0.5, 0.5)
+  ) +
+  xlim(65, 345) +
+  ylim(5800, 6100) +
   gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
@@
 gridExtra::grid.arrange(
-  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16 ),
+  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16),
   widths = c(2, 2, 1.75),
   heights = c(2, 1.5),
-  layout_matrix = rbind(c(1, 2, 3),
-                        c(4, 5, 6)),
+  layout_matrix = rbind(
+    c(1, 2, 3),
+    c(4, 5, 6)
+  ),
   top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")

### 2020-06-25 analysis/tmb-sensor-explore/3-fit-do.Rmd
annotate

@@
 ---
+title: "Spatiotemporal models of DO"
+author: "Philina English"
+date: "`r Sys.Date()`"
 output: html_document
@@
 
-Model diagnostics functions
+Model diagnostics function
 ```{r}
@@
 
-Ranges sampled
+# Data exploration
+Check range of depths and DO sampled
 ```{r}
@@
 
+# Define spatial and temporal components to include in model
 Remove most samples from WCVI in 2016 due to extreme values and make mesh
@@
 
+# Define and run model
 Current best model of DO
-```{r }
+```{r eval=FALSE}
 do_model_all_DOY <- sdmTMB::sdmTMB(
@@
   include_spatial = TRUE,
-  enable_priors = FALSE,
   nlminb_loops = 2,
@@
 
-```{r eval=FALSE}
-# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth3.rds")))
-# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth2.rds")))
-# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds")))
+# Check diagnostics
+```{r eval=T}
+# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth3.rds"))) # 500 knots, poly-3 of depth
+# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth2.rds"))) # 500 knots, poly-2 of depth
+# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds"))) # 800 knots
+do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-800kn.rds"))) # add in some WCVI 2016 data
 get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
@@
 
-```{r}
+```{r eval=FALSE}
 predictions <- predict(do_model_all_DOY)
@@
 
-# predictions$abs <- exp(abs(predictions$residuals))
-# pred_rows <- filter (predictions, abs != "Inf")
-# sum(pred_rows$abs) # sum absolute error
+predictions$abs <- exp(abs(predictions$residuals))
+pred_rows <- filter (predictions, abs != "Inf")
+sum(pred_rows$abs) # sum absolute error, divided by nrow for mean absolute error
 # 19719.73-17459.49 # difference in absolute error between 500 and 800 knots
@@
 
-Predict DO for whole grid using quadratic of predicted temperature values and mean DOY
+# DO predictions
+For whole grid using quadratic of predicted temperature values and mean DOY
 ```{r eval=FALSE}
@@
 
-```{r, eval=FALSE}
+```{r}
 omega_s_all <- plot_map_raster(p_nd, "omega_s") +
@@
 
+Save selection of prediction plots
 ```{r, eval= F}

### 2020-07-09 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
merge ROMs data with trawl data and check against CTD and sdmTMB model

@@
 library(sdmTMB)
-library(future)
+# library(future)
 library(dplyr)
@@
 # 119.557/2908.879 # 4% reduction in abs errors
+```
+
+```{r eval=F}
+# test against ROMS
+events_roms <- readRDS("data/roms_temp_by_events2.rds") %>% select(fishing_event_id, roms, year)
+
+predictions <- left_join(predictions, events_roms)
+
+cor(predictions$roms, predictions$est, use = "pairwise.complete.obs")
+cor(predictions$roms, predictions$temperature_c, use = "pairwise.complete.obs")
+
+ggplot(predictions, aes(roms, est, colour = -depth)) + geom_point(alpha=0.3) + 
+  facet_wrap(~year) + coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
+  geom_abline(intercept = 0, slope = 1) +
+  gfplot::theme_pbs()
+
+ggplot(predictions, aes(roms, temperature_c, colour = DOY)) + geom_point(alpha=0.3) +
+  facet_wrap(~year) + coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
+  geom_abline(intercept = 0, slope = 1) +
+  scale_colour_viridis_c() +
+  gfplot::theme_pbs()
+
+ggplot(predictions, aes(DOY, (roms-temperature_c), colour = -depth)) + geom_point(alpha=0.3) +
+  facet_wrap(~year) + #coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
+  geom_abline(intercept = 0, slope = 0) +
+  scale_colour_viridis_c() +
+  gfplot::theme_pbs()
+
+ggplot(predictions, aes(DOY, (est-roms), colour = -depth)) + geom_point(alpha=0.3) +
+  facet_wrap(~year) + #coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
+  geom_abline(intercept = 0, slope = 0) +
+  scale_colour_viridis_c() +
+  gfplot::theme_pbs()
+
+p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")))
+
+cor(p_nd$roms_temp, p_nd$temp, use = "pairwise.complete.obs")
+
+ggplot(p_nd, aes(roms_temp, temp, colour = -depth)) + geom_point(alpha=0.04) + 
+  facet_wrap(~year) + coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
+  geom_abline(intercept = 0, slope = 1) +
+  gfplot::theme_pbs()
+
 ```

### 2020-07-09 analysis/tmb-sensor-explore/extract_ROMS.R

@@ -1,0 +1,156 @@
+library(sp)
+library(raster)
+library(tidyverse)
+
+setwd(here::here("analysis", "tmb-sensor-explore"))
+
+temp_all_years <- stack()
+
+for(i in 2003:2018){
+  temp_one_year <- raster(paste("data/roms_temp/temp_bcroms_y", i, ".tif", sep = ""))
+  names(temp_one_year) <- paste0("T", i)
+  temp_all_years <- stack(temp_all_years, temp_one_year)
+}
+
+# bath <- readRDS(here::here("analysis/VOCC/data/bathymetry-data")) 
+# events <- bath$data %>% mutate(X_old=X*1000, Y_old=Y*1000) %>% select(-X, -Y)
+# xx <- events %>%
+#   sf::st_as_sf(coords = c("lon", "lat"),
+#     crs = 4326 , 
+#     remove = FALSE)
+
+
+events <- m_temp_rw$data %>% mutate(X_old=X*1000, Y_old=Y*1000) %>% select(-X, -Y)
+
+
+xx <- events %>%
+  sf::st_as_sf(coords = c("longitude", "latitude"),
+    crs = 4326 ,
+    remove = FALSE)
+
+trawl <- sf::st_transform(xx, crs = 
+    CRS("+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"))
+
+# this should be your x,y coordinates that you want to extract. They will need to be in the CRS specified on the next line
+trawl_coord <- sf::st_coordinates(trawl)
+ggplot(as.data.frame(trawl_coord), aes(X,Y)) + geom_point()
+
+# ### alternative method gets same result
+# lonlat <- events %>% dplyr::select(lon, lat)
+# coordinates(lonlat) <- events %>% dplyr::select(lon, lat)
+# proj4string(lonlat) <-  CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
[…]

### 2020-07-09 analysis/VOCC/3-vocc-dist.Rmd
vocc based on ROMs

@@
     immature: T
-    # region: "West Coast Vancouver Island"
+    region: "West Coast Vancouver Island"
     # region: "both odd year surveys"
-    region: "West Coast Haida Gwaii"
+    # region: "West Coast Haida Gwaii"
     covs: "-tv-depth-only"
@@
 ```{r}
-nd_all <- readRDS(paste0("data/predicted-DO-", climate_mod, ".rds")) %>% select(X, Y, year, temp, do_est)
+nd_all <- readRDS(paste0("data/predicted-DO-", climate_mod, "-roms.rds")) %>% select(X, Y, year, temp, do_est, roms_temp)
 
@@
 # upper_change_t <- temp_sd_choice*2
-
+```
+
+### Annual temperature (ROMS)
+```{r}
+rm(roms)
+try({roms <- readRDS(paste0("data/_climate_vocc/annual-temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+})
+
+if(!exists("roms")){
+
+roms <- vocc_gradient_calc(d, "roms_temp",
+  scale_fac = scale_fac,
+  # all layers with max indice will be averaged for spatial gradient
+  # if default (NULL) will use all years
+  indices = indices,
+  # divisor = 1, #default of 10 gives per decade
+  quantile_cutoff = 0.05
+)
+
+roms <- roms %>%
+  mutate(
+    ## these are only relevant if divisor = 1 
+    # trend_per_year = trend,
+    # trend_per_decade = trend_per_year * 10,
+    trend_per_year = trend/10,
+    trend_per_decade = trend_per_year * 10,
+    velocity = velocity
+  ) %>%
+  dplyr::mutate(X = x, Y = y)
+
+# # FULL DATASET
+# # mean(roms$sd, na.rm = T) # [1] 0.3434
+# # min(roms$sd, na.rm = T) # [1] 0.1266
+# # max(roms$sd, na.rm = T) # [1] 1.12
+# # median(roms$sd, na.rm = T) # [1] 0.287
+# # 
+# # 2008 DATASET
+# mean(roms$sd, na.rm = T) # [1] 0.3274
+# temp_min_sd <- min(roms$sd, na.rm = T) # [1] 0.1035
+# max(roms$sd, na.rm = T) # [1] 1.248
+# temp_median_sd <- median(roms$sd, na.rm = T) # [1] 0.2882
+# 
+# temp_sd_choice <- temp_min_sd
+# lower_change_t <- temp_sd_choice*2
+# upper_change_t <- temp_sd_choice*2
+# 
+# dvocc_t <- make_vector_data(d,
+#   variable_names = c("temp"),
+#   ssid = model_ssid,
+#   input_cell_size = 2,
+#   scale_fac = scale_fac,
+#   min_dist = 0,
+#   delta_t_total = years_sampled,
+#   delta_t_step = 2,
+#   indices = indices,
+#   min_thresholds = lower_change_t,
+#   max_thresholds = upper_change_t,
+#   round_fact = 10
+# )
+# 
+# dvocc_t <- dvocc_t %>% mutate(dvocc = if_else(slope < 0, -km_per_decade, km_per_decade), trend = slope)
+# dvocc_t <- dvocc_t %>% mutate(dvocc = if_else(dvocc > (max_dist_all_years/ year_range)*10, 
+#   (max_dist_all_years/ year_range)*10, dvocc))
+# dvocc_t <- mutate(dvocc_t, dvocc = if_else(dvocc < -(max_dist_all_years / year_range)*10,
+#   -(max_dist_all_years / year_range)*10, dvocc))
+# 
+# dv_t <- dvocc_t %>% select(icell, x, y, dvocc) %>% unique()
+# 
+# # glimpse(dv_t)
+# # glimpse(roms)
+# 
+# # roms$dvocc <- dv_t$dvocc
+# 
+# roms <- left_join(roms, dv_t)
+
+saveRDS(roms, file = paste0("data/_climate_vocc/annual-temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+
+# mine <- gf2 %>% select(x, y, temp_trend = trend, temp_vel = velocity, temp_grad= gradient)
+# other <- roms %>% select(x, y, roms_temp_trend = trend, roms_temp_vel = velocity, roms_temp_grad= gradient)
+# 
+# check <- left_join(mine, other)
+# 
+# ggplot(check, aes(roms_temp_trend, temp_trend)) + 
+#   geom_point(alpha=0.2) +
+#   coord_cartesian(xlim = c(-1,2.5), ylim = c(-1,2.5)) +
+#   geom_abline(intercept = 0, slope = 1) +
+#   gfplot::theme_pbs()
+
+}
+```
+
+```{r}
+# ggplot(check, aes(roms_temp_vel, temp_vel)) + 
+#   geom_point(alpha=0.2) +
+#   coord_cartesian(xlim = c(-20,80), ylim = c(-20,80)) +
+#   geom_abline(intercept = 0, slope = 1) +
+#   gfplot::theme_pbs()
 ```

### 2020-07-09 analysis/VOCC/4-prep-var-vocc.R

@@ -1,4 +1,7 @@
 # ### JUST CLIMATE AND FISHING COVARIATES
+library(tidyverse)
+library(sdmTMB)
+library(gfranges)
 
 setwd(here::here("analysis", "VOCC"))
 
@@ -42,9 +45,35 @@ d <- left_join(fished, DO_values)
 d <- left_join(d, t_values)
 
 
+# ANNUAL TEMP
+at16 <- readRDS(paste0("data/_climate_vocc/annual-temp-vocc-2020-06-20-more2016-scale-2-ssid16-years-6.rds"))
+at1n3 <- readRDS(paste0("data/_climate_vocc/annual-temp-vocc-2020-06-20-more2016-scale-2-ssid1n3-years-5.rds"))
+at4 <- readRDS(paste0("data/_climate_vocc/annual-temp-vocc-2020-06-20-more2016-scale-2-ssid4-years-6.rds"))
+
+at_vocc <- rbind(at16, at1n3, at4)
+
+at_values <- at_vocc %>%
+  rename(
+    ann_temp_vel = velocity,
+    # ann_temp_dvocc = dvocc,
+    ann_temp_trend = trend_per_decade,
+    ann_temp_grad = gradient,
+    ann_mean_temp = mean
+  ) %>%
+  select(x, y, ann_temp_vel, #ann_temp_dvocc, 
+    ann_temp_trend, ann_temp_grad, ann_mean_temp)
+
+
+d <- left_join(d, at_values)
+
+
 #### PREP DATA ####
 d$squashed_temp_vel <- collapse_outliers(d$temp_vel, c(0.005, 0.995)) #99th quantile and then droped to be less than 100km
 hist(d$squashed_temp_vel, breaks = 100)
+
+d$squashed_ann_temp_vel <- collapse_outliers(d$ann_temp_vel, c(0.005, 0.995)) #99th quantile and then droped to be less than 100km
+hist(d$squashed_ann_temp_vel, breaks = 100)
+
 d$squashed_DO_vel <- collapse_outliers(d$DO_vel, c(0.025, 0.975)) #95th quantile and then droped to be less than 100km
 hist(d$squashed_DO_vel, breaks = 100)
 
@@ -113,4 +142,27 @@ d$log_effort_scaled <- scale(d$log_effort, center = F)
 
 
[…]

### 2020-07-17 analysis/VOCC/3-vocc-dist.Rmd
tweak dvocc calculation for DO and run model on result

@@
     immature: T
-    region: "West Coast Vancouver Island"
+    # region: "West Coast Vancouver Island"
     # region: "both odd year surveys"
-    # region: "West Coast Haida Gwaii"
+    region: "West Coast Haida Gwaii"
     covs: "-tv-depth-only"
@@
 # 2008 DATASET
+# values currently depend on survey region
 mean(gf2$sd, na.rm = T) # [1] 0.3274
@@
 temp_median_sd <- median(gf2$sd, na.rm = T) # [1] 0.2882
+
+print( paste("min SD =", round(temp_min_sd, digits = 3)))
+print( paste("median SD =", round(temp_median_sd, digits = 3)))
+
+print( paste("treshold chosen =", round(temp_min_sd*2, digits = 3), "degrees C"))
 
@@
 ```{r }
-
 rm(gf0)
+
 try({gf0 <- readRDS(paste0("data/_climate_vocc/DO-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
@@
 # [1] 0.2343
-
-do_sd_choice <- do_median_sd
+do_min_sd
+do_median_sd 
+do_sd_choice <- do_min_sd
 
@@
 do_median_sd <- median(gf0$sd, na.rm = T)
-do_sd_choice <- do_median_sd
+do_sd_choice <- do_min_sd
+
+print( paste("min SD =", round(do_min_sd, digits = 3)))
+print( paste("median SD =", round(do_median_sd, digits = 3)))
+
+print( paste("treshold chosen =", round(do_min_sd*2, digits = 3), "ml/L"))
+
 ```

### 2020-07-17 analysis/VOCC/4-prep-var-vocc.R

@@ -141,7 +141,7 @@ hist(log(d$mean_effort + 1))
 d$log_effort_scaled <- scale(d$log_effort, center = F)
 
 
-saveRDS(d, file = paste0("data/all-newclim-untrimmed.rds"))
+saveRDS(d, file = paste0("data/all-newclim-untrimmed-2.rds"))
 
 
 cor(d$ann_temp_trend, d$temp_trend, use = "pairwise.complete.obs")
@@ -165,4 +165,19 @@ ggplot(d, aes(ann_temp_grad, temp_grad)) +
   geom_abline(intercept = 0, slope = 1) +
   gfplot::theme_pbs()
 
+ggplot(d, aes(abs(temp_dvocc), abs(squashed_temp_vel))) + 
+  geom_point(alpha=0.2) +
+  # coord_cartesian(xlim = c(-1,2.5), ylim = c(-1,2.5)) +
+  geom_smooth() +
+  geom_abline(intercept = 0, slope = 1) +
+  gfplot::theme_pbs()
 
+ggplot(d, aes(abs(DO_dvocc), abs(squashed_DO_vel))) + 
+  geom_point(alpha=0.2) +
+  geom_smooth() +
+  # coord_cartesian(xlim = c(-1,2.5), ylim = c(-1,2.5)) +
+  geom_abline(intercept = 0, slope = 1) +
+  gfplot::theme_pbs()
+
+
+hist(d$DO_dvocc , breaks = 100)

### 2020-07-23 analysis/VOCC/3-vocc-dist.Rmd
dvocc calc based on med

--- 2020-07-09 analysis/VOCC/3-vocc-dist.Rmd
@@
 rm(gf2)
-try({gf2 <- readRDS(paste0("data/_climate_vocc/temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
-})
+
+# try({
+#   gf2 <- readRDS(paste0("data/_climate_vocc/temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+# })
 
@@
 # 2008 DATASET
-mean(gf2$sd, na.rm = T) # [1] 0.3274
+# values currently depend on survey region
+temp_mean_sd <- mean(gf2$sd, na.rm = T) # [1] 0.3274
+temp_mean_sd 
+temp_median_sd <- median(gf2$sd, na.rm = T) # [1] 0.2882
+temp_median_sd
 temp_min_sd <- min(gf2$sd, na.rm = T) # [1] 0.1035
+temp_min_sd*2
 max(gf2$sd, na.rm = T) # [1] 1.248
-temp_median_sd <- median(gf2$sd, na.rm = T) # [1] 0.2882
-
-temp_sd_choice <- temp_min_sd
-lower_change_t <- temp_sd_choice*2
-upper_change_t <- temp_sd_choice*2
+
+# temp_sd_choice <- temp_min_sd*2
+# temp_sd_choice <- temp_mean_sd
+temp_sd_choice <- temp_median_sd
+
+# lower_change_t <- temp_sd_choice
+lower_change_t <- Inf
+upper_change_t <- temp_sd_choice
+
+rm(dvocc_t)
 
@@
 saveRDS(gf2, file = paste0("data/_climate_vocc/temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+# saveRDS(gf2, file = paste0("data/_climate_vocc/temp-dvocc-min2SD-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+# saveRDS(gf2, file = paste0("data/_climate_vocc/temp-dvocc-medSDasy-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
 }
@@
 # mean(gf2$sd, na.rm = T) # [1] 0.3274
-temp_min_sd <- min(gf2$sd, na.rm = T) # [1] 0.1035
+temp_min_sd <- round(min(gf2$sd, na.rm = T), digits = 3) #
+
+# temp_mean_sd <- mean(gf2$sd, na.rm = T) # 
 # max(gf2$sd, na.rm = T) # [1] 1.248
-temp_median_sd <- median(gf2$sd, na.rm = T) # [1] 0.2882
-
-# temp_sd_choice <- temp_min_sd
-# lower_change_t <- temp_sd_choice*2
-# upper_change_t <- temp_sd_choice*2
+temp_median_sd <- round(median(gf2$sd, na.rm = T), digits = 3) # [1] 0.2882
+temp_mean_sd <- round(mean(gf2$sd, na.rm = T), digits = 3) #
+
+print( paste("min SD =", round(temp_min_sd, digits = 2)))
+print( paste("median SD =", round(temp_median_sd, digits = 2)))
+print( paste("mean SD =", round(temp_mean_sd, digits = 2)))
+print( paste("treshold chosen =", round(temp_mean_sd, digits = 2), "degrees C"))
+```
+
+```{r eval=F}
+plot_vocc(gf2,
+    raster_cell_size = 4,
+  coast = TRUE,
+  vec_aes = NULL,
+  NA_label = "NA",
+  fill_col = "dvocc",
+  fill_label = "temperature \nkm/decade",
+  raster_alpha = 1,
+  na_colour = "white",
+  mid_fill = "mistyrose1", grey_water = F,
+  # mid_fill = "lavenderblush1", grey_water = F,
+  low_fill = "royalblue4", # low_fill = "#5E4FA2",
+  high_fill = "Red 3",
+  raster_limits = c(0, 100),
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = sqrt #
+) + labs(
+    title = "distance-based temp velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year), " mean temp SD = ", signif(temp_mean_sd, digits = 2), " for min threshold of change")
+)
 ```
@@
 ```{r }
-
 rm(gf0)
-try({gf0 <- readRDS(paste0("data/_climate_vocc/DO-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+
+try({
+  gf0 <- readRDS(paste0("data/_climate_vocc/DO-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
 })
@@
 
+do_mean_sd <- mean(gf0$sd, na.rm = T)
+
+do_min_sd
+do_median_sd 
+do_mean_sd
+
+# do_sd_choice <- do_min_sd*2
 do_sd_choice <- do_median_sd
 
-lower_change_d <- c(do_sd_choice*2) 
-upper_change_d <- c(do_sd_choice*2)
+lower_change_d <- c(do_sd_choice) 
+upper_change_d <- c(do_sd_choice)
 
@@
 
-lower_change_td <- c(Inf, do_sd_choice*2) 
-upper_change_td <- c(temp_sd_choice*2, do_sd_choice*2)
+lower_change_td <- c(Inf, do_sd_choice) 
+upper_change_td <- c(temp_sd_choice, do_sd_choice)
 
@@
 gf0 <- left_join(gf0, dv_td)
-
+# 
 saveRDS(gf0, file = paste0("data/_climate_vocc/DO-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+# saveRDS(gf0, file = paste0("data/_climate_vocc/DO-dvocc-min2SD-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+# saveRDS(gf0, file = paste0("data/_climate_vocc/DO-dvocc-meanSD-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+# saveRDS(gf0, file = paste0("data/_climate_vocc/DO-dvocc-medSD-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
 }
 
-do_min_sd <- min(gf0$sd, na.rm = T)
-do_median_sd <- median(gf0$sd, na.rm = T)
-do_sd_choice <- do_median_sd
-```
-
+do_min_sd <- round(min(gf0$sd, na.rm = T), digits = 3)
+do_mean_sd <- round(mean(gf0$sd, na.rm = T), digits = 3)
+do_median_sd <- round(median(gf0$sd, na.rm = T), digits = 3)
+
+print( paste("min SD =", do_min_sd))
+print( paste("mean SD =", do_mean_sd))
+print( paste("median SD =", do_median_sd))
+
+do_sd_choice <- do_mean_sd
+print( paste("treshold chosen =", signif(do_sd_choice, digits = 2), "ml/L"))
+
+```
+
+```{r eval=F}
+plot_vocc(gf0,
+    raster_cell_size = 4,
+  coast = TRUE,
+  vec_aes = NULL,
+  NA_label = "NA",
+  fill_col = "dvocc",
+  fill_label = "km/decade",
+  raster_alpha = 1,
+  na_colour = "white",
+  mid_fill = "honeydew", grey_water = F,
+  # mid_fill = "lightyellow", grey_water = F,
+  # mid_fill = "azure", grey_water = F,
+  high_fill = "gold",
+  low_fill = "darkcyan", # "lightseagreen",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = sqrt #
+) + labs(
+ title = "distance-based DO velocity",
+  subtitle = paste0(min(d$year), "-", max(d$year), " DO SD = ", signif(do_sd_choice, 2), " for min threshold of DO change")
+)
+```
+
+```{r eval=F}
+plot_vocc(gf0,
+    raster_cell_size = 4,
+  coast = TRUE,
+  vec_aes = NULL,
+  NA_label = "NA",
+  fill_col = "dvocc_both",
+  fill_label = "km/decade",
+  raster_alpha = 1,
+  na_colour = "white",
+  mid_fill = "honeydew", grey_water = F,
+  # mid_fill = "lightyellow", grey_water = F,
+  # mid_fill = "azure", grey_water = F,
+  high_fill = "gold",
+  low_fill = "darkcyan", # "lightseagreen",
+  white_zero = TRUE,
+  vec_alpha = 0.35,
+  axis_lables = FALSE,
+  transform_col = sqrt #
+) + labs(
+ title = "distance-based temp and DO velocity"
+)
+```
 ### BIOTIC VECTORS
@@
     title = "distance-based temp velocity",
-  subtitle = paste0(min(d$year), "-", max(d$year), " min temp SD = ", temp_min_sd, " x 2 for 95% min threshold of change")
+  subtitle = paste0(min(d$year), "-", max(d$year), " min temp SD = ", temp_mean_sd, " for 95% min threshold of change")
 )

### 2020-07-23 analysis/VOCC/4-prep-var-vocc.R

@@ -141,7 +141,7 @@ hist(log(d$mean_effort + 1))
 d$log_effort_scaled <- scale(d$log_effort, center = F)
 
 
-saveRDS(d, file = paste0("data/all-newclim-untrimmed-2.rds"))
+saveRDS(d, file = paste0("data/all-newclim-untrimmed-dvocc-med.rds"))
 
 
 cor(d$ann_temp_trend, d$temp_trend, use = "pairwise.complete.obs")

### 2020-07-23 analysis/VOCC/5-add-null.R

@@ -168,7 +168,7 @@ data <- do.call(rbind, trimmed.dat)
 
 
 #### add in globally scaled and squashed climate data 
-vars <- readRDS(("data/all-newclim-untrimmed-2.rds"))  %>% rename(cell_depth = depth)
+vars <- readRDS(("data/all-newclim-untrimmed-dvocc-med.rds"))  %>% rename(cell_depth = depth)
 
 data <- left_join(data, vars) %>% select(-X.1)
 

### 2020-07-31 analysis/VOCC/2-biomass-depth-only-model.Rmd
add extra optimizations; remove all AR1; save all gradients

@@
 params: 
-    # species: "Redbanded Rockfish"
-    #species: "Pacific Ocean Perch"
-    #species: "Pacific Cod"
-    #species: "Petrale Sole"
-    # species: "Widow Rockfish"
-    species: "Bocaccio"
-    # species: "Sablefish"
+    species: "Redbanded Rockfish"
+    # species: "Redstripe Rockfish"
+    # species: "Shortbelly Rockfish"
     region: "All synoptic surveys"
@@
       # reml = TRUE,
+      nlminb_loops = 2,
+      newton_steps = 2,
       enable_priors = priors,
@@
       family = tweedie(link = "log"),
-      ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
+      # ar1_fields = TRUE, # changed to FALSE July 2020
       # reml = TRUE,
       include_spatial = TRUE,
+      nlminb_loops = 2, 
+      newton_steps = 2,
       enable_priors = priors,
@@
       file = paste0("data/", spp, "/mod-imm-biomass-", 
-        spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+        spp, covs, "-", ssid_string, ".rds"))
    })
@@
       family = tweedie(link = "log"),
-      ar1_fields = TRUE,
+      # ar1_fields = TRUE, # all off in July 2020
       include_spatial = TRUE,
+      nlminb_loops = 2, 
+      newton_steps = 2,
       enable_priors = priors,
@@
       "data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+      "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
     ))
@@
       "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
+      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass.rds"
     ))
@@
       "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
+      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
     ))
@@
     "data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
+    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass.rds"
   ))
@@
 ```{r}
-sdmTMB:::get_convergence_diagnostics(depth_model_list[[1]])
+write_grads <- function(x, macro, ...) {
+  paste0("\\newcommand{\\", macro, "}{", x, "}") %>%
+    readr::write_lines("model-grads.tex", append = TRUE)
+}
 ```
+
+```{r}
+write_grads(signif(max(depth_model_list$adult$gradients), digits = 8), paste0(spp, "-mat"))
+write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0(spp, "-imm"))
+write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0(spp, "-total"))
+```
+
+```{r eval = FALSE}
+try({
+  if(max(depth_model_list$adult$gradients)>0.001) {
+    adult_biomass <- sdmTMB:::run_extra_optimization(depth_model_list$adult, nlminb_loops = 2, newton_steps = 2)
+    write_grads(signif(max(adult_biomass$gradients), digits = 8), paste0(spp, "-mat2"))
+  }
+})
+```
+
+```{r eval = FALSE}
+try({
+  if(max(depth_model_list$imm$gradients)>0.001) {
+    imm_biomass <- sdmTMB::run_extra_optimization(depth_model_list$imm, nlminb_loops = 2, newton_steps = 2)
+    write_grads(signif(max(imm_biomass$gradients), digits = 8), paste0(spp, "-imm2"))
+  }
+})
+```
+
+```{r eval = FALSE}
+try({
+  if(max(depth_model_list$total$gradients)>0.001) {
+    total_biomass <- sdmTMB::run_extra_optimization(depth_model_list$total, nlminb_loops = 2, newton_steps = 2)
+    write_grads(signif(max(total_biomass$gradients), digits = 8), paste0(spp, "-total2"))
+  }
+})
+```
+
+```{r eval = FALSE}
+ if (any(names(data) == "adult_density")) {
+
+    saveRDS(adult_biomass,
+      file = paste0(
+        "data/", spp,
+        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+      )
+    )
+
+    saveRDS(imm_biomass,
+      file = paste0("data/", spp, "/mod-imm-biomass-", 
+        spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+
+  } else {
+    saveRDS(total_biomass, file = paste0(
+      "data/", spp,
+      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    ))
+  }
+```

### 2020-07-31 analysis/VOCC/2-render-biomass-model.R

@@ -10,7 +10,7 @@ list_species <- c(
   "Spotted Ratfish",
   # "Pacific Tomcod",
   "Walleye Pollock",
-  # "Pacific Cod",
+  "Pacific Cod",
   "Sablefish",
   "Lingcod",
   # "Pacific Hake",
@@ -18,34 +18,34 @@ list_species <- c(
   "Redstripe Rockfish",
   "Yellowmouth Rockfish",
   # "Harlequin Rockfish",
-  # "Bocaccio", # winter-birthing, overfished
+# "Bocaccio", # winter-birthing, overfished
   "Canary Rockfish", # schooling, winter-birthing
   # "Copper Rockfish", # small sample
   "Darkblotched Rockfish",
   "Greenstriped Rockfish",
-  # "Pacific Ocean Perch", # schooling
+  "Pacific Ocean Perch", # schooling
   "Quillback Rockfish",
-  # "Redbanded Rockfish",
+  "Redbanded Rockfish",
   "Sharpchin Rockfish",
   "Shortbelly Rockfish", # small sample
   "Silvergray Rockfish",
   "Splitnose Rockfish",
   "Widow Rockfish", # schooling
   "Yellowtail Rockfish", # schooling
-   "Longspine Thornyhead",
+   # "Longspine Thornyhead",
   "Shortspine Thornyhead",
-  #"Arrowtooth Flounder",
+  "Arrowtooth Flounder",
   "Rex Sole",
-  # "Petrale Sole",
-  # "English Sole",
+  "Petrale Sole",
+  "English Sole",
   "Dover Sole",
   "Southern Rock Sole",
   "Flathead Sole",
   "Curlfin Sole",
-  "Sand Sole",
+  # "Sand Sole",
   # "Slender Sole",
   # "Pacific Sanddab",
-  # "Pacific Halibut"
+  "Pacific Halibut"
 )
 
 ### SUBSETS OF SPECIES
@@ -56,17 +56,14 @@ list_species <- c(
   "Pacific Halibut"
 )
 
-#   # "Arrowtooth Flounder"
-#   # "Petrale Sole",
-#   # "English Sole",
 
  
  list_species <- c(
-   "Redstripe Rockfish",
+   # "Redstripe Rockfish",
    "Rougheye/Blackspotted Rockfish Complex",
-   "Widow Rockfish",
-   "Quillback Rockfish",
-   "Bocaccio",
+   # "Widow Rockfish",
[…]

### 2020-07-31 analysis/VOCC/3-vocc-dist.Rmd
update predictions and voccs

@@
     # species: "Widow Rockfish"
-    # species: "Canary Rockfish"
+    species: "Sharpchin Rockfish"
     # species: "English Sole"
@@
     # species: "Sablefish"
-    species: "Arrowtooth Flounder"
-    immature: T
+    # species: "Shortbelly Rockfish"
+    # species: "Arrowtooth Flounder"
+    immature: F
     # region: "West Coast Vancouver Island"
-    region: "both odd year surveys"
-    # region: "West Coast Haida Gwaii"
+    # region: "both odd year surveys"
+    region: "West Coast Haida Gwaii"
     covs: "-tv-depth-only"
@@
     "data/", spp,
-    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-FALSE.rds"
+    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass.rds" #-prior-FALSE.rds"
   )) %>% filter(year >= start_year )
@@
       "data/", spp,
-      "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-FALSE.rds"
+      "/predictions-", spp, covs, "-1n3n4n16-total-biomass.rds"
     )) %>% filter(year >= start_year )
@@
       "data/", spp,
-      "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds"
+      "/predictions-", spp, covs, "-1n3n4n16-mature-biomass.rds"
     )) %>% filter(year >= start_year )
@@
 
-# try({
-#   gf2 <- readRDS(paste0("data/_climate_vocc/temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
-# })
+try({
+  gf2 <- readRDS(paste0("data/_climate_vocc/temp-vocc-", climate_mod, "-scale-", scale_fac, "-ssid", ssid_string, "-years-", years_sampled, ".rds"))
+})
 
@@
 
-```{r eval=T}
+```{r eval=F}
 gradbv <- plot_vocc(gf3t,
@@
 
-```{r eval=T}
+```{r eval=F}
 gf3t$squashed_velocity <- collapse_outliers(gf3t$velocity, c(0.005, 0.995))
@@
 
-```{r eval=T}
+```{r eval=F}
 # gf3t$squashed_velocity <- collapse_outliers(gf3t$velocity, c(0.005, 0.995))
@@
 
-```{r eval=T}
+```{r eval=F}
 #raster_max <- quantile(gf3$mean_biomass, 0.995, na.rm = T)
@@
 
-```{r eval=T}
+```{r eval=F}
 grad5 <- gfranges::plot_vocc(gf3t,
@@
 
-```{r eval=T}
+```{r eval=F}
 grad6 <- gfranges::plot_vocc(gf3t,
@@
 write.csv(vocc, file = paste0(
-  "data/_more2016_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_optimized_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))
@@
 vocc  <- read.csv(paste0(
-  "data/_more2016_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_optimized_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))
@@
 )
-
-```
+```

### 2020-07-31 analysis/VOCC/3-render-biotic-vocc.R

@@ -70,14 +70,12 @@ for (r_h in seq_along(list_regions)) {
       rmarkdown::render("3-vocc-dist.Rmd",
         params = list(
           species = list_species[spp_i],
-          # immature = TRUE,
           immature = FALSE,
           region = list_regions[r_h],
           covs = covs
         ),
         output_file = paste0(
           "html/VOCC-plots/vocc-w-do-", spp,
-          # "-imm",
           covs, "-", reg, "-more2016.html"
         ),
         envir = env
@@ -86,6 +84,27 @@ for (r_h in seq_along(list_regions)) {
   }
 }
 }
+
+######
+# TODO: will need to rerun these imm
+######
+list_species <- c(
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Widow Rockfish",
+  "Quillback Rockfish",
+  "Yelloweye Rockfish"
+)
+
+list_species <- c(
+  "Bocaccio"
+)
+
+list_regions <- c(
+  "West Coast Haida Gwaii",
+  "West Coast Vancouver Island",
+  "both odd year surveys"
+  #   "All synoptic surveys"
+)
 
 age <- "immature"
 if (age == "immature") {

### 2020-07-31 analysis/VOCC/4-pool-biotic-vocc.R

@@ -6,19 +6,21 @@ library("dplyr")
 # for models on the 06-20 that use 500 knots for temp and 800 for DO
 # mydir = paste0( "_newclim2_mature")
 # 800 kn for both
-mydir = paste0( "_more2016_mature")
-
+# mydir = paste0( "_more2016_mature")
+mydir = paste0( "_optimized_mature")
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
 myfiles
 biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% select(-X, -start_year)
 glimpse(biotic)
 
 saveRDS(biotic, file = paste0(
-  "mature-all-do-newclim3.rds"
+  # "mature-all-do-newclim3.rds"
+  "mature-optimized-vocc.rds"
 ))
 
 #### IMMATURE #####
-mydir = paste0( "_more2016_immature")
+# mydir = paste0( "_more2016_immature")
+mydir = paste0( "_optimized_immature")
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
 myfiles
 biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% 
@@ -26,5 +28,6 @@ biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>%
 glimpse(biotic)
 
 saveRDS(biotic, file = paste0(
-  "immature-all-do-newclim3.rds"
+  # "immature-all-do-newclim3.rds"
+  "immature-optimized-vocc.rds"
 ))

### 2020-07-31 analysis/VOCC/2-check-biomass-grads.Rmd
more code for checking gradients and extra optimizations

--- 2020-02-19 analysis/VOCC/2-biomass-depth-only-model.Rmd
@@
 ---
-title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
+title: "Check convergence of biomass models"
 author: "Philina English"
@@
 params: 
-    #species: "Redbanded Rockfish"
-    #species: "Pacific Ocean Perch"
-    #species: "Pacific Cod"
-    #species: "Petrale Sole"
-    #species: "Widow Rockfish"
-    species: "Bocaccio"
+    # species: "Bocaccio"
+    species: "Shortbelly Rockfish"
+    # species: "Sablefish"
     region: "All synoptic surveys"
-    knots: 500
-    covariates: "" # +as.factor(ssid)
-    #additional fixed-effect covariates
-    #covs: "-ssid-only" # covariate string used to name model rds 
+    covariates: "" 
     covs: "-tv-depth-only"
-    #covs: "-trawled-ssid"
-    update_model: FALSE
-    update_predictions: FALSE
-    update_model_check: FALSE
 ---
@@
 covs <- params$covs
-knots <- params$knots
 priors <- FALSE
@@
 
-Make mesh
-```{r}
-if (region == "Both odd year surveys") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
-}
-
-if (region == "West Coast Vancouver Island") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-}
-
-if (region == "West Coast Haida Gwaii") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-}
-
-if (region == "All synoptic surveys") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
-}
-
-sdmTMB::plot_spde(spde)
-```
-
-Run climate independent sdmTMB model
-```{r }
-if (params$update_model) {
-  tictoc::tic()
-  if (any(names(data) == "adult_density")) {
-    adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year)", covariates, ""
-    ))
-
-    adult_biomass <- sdmTMB::sdmTMB(
-      data = data,
-      adult_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      #ar1_fields = TRUE,
-      include_spatial = TRUE,
-      # reml = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-
-    saveRDS(adult_biomass,
-      file = paste0(
-        "data/", spp,
-        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-      )
-    )
-
-   try({ imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year)", covariates, ""
-    ))
-
-    imm_biomass <- sdmTMB::sdmTMB(
-      data = data,
-      imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
-      # reml = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-
-    saveRDS(imm_biomass,
-      file = paste0("data/", spp, "/mod-imm-biomass-", 
-        spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
-   })
-
-  } else {
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
-
-    total_biomass <- sdmTMB::sdmTMB(
-      data = data,
-      dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      ar1_fields = TRUE,
-      include_spatial = TRUE,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-
-    saveRDS(total_biomass, file = paste0(
-      "data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
-  }
-  tictoc::toc()
-}
-```
-
-Make prediction grids for all surveys and years, if not done before.
-```{r, eval=FALSE, message=FALSE, warning=FALSE}
-try(nd_all <- readRDS("data/nd_all_synoptic.rds"))
-
-if (!exists("nd_all")) {
-
-  # choose base year(s) to create grid from
-  dummy_year <- c(2005, 2006)
-
-  bath <- readRDS("data/bathymetry-data")
-  # dat <- bath$data %>% scale_survey_predictors()
-  dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth))
-  dat <- gfranges::scale_predictors(dat, predictors = c(quo(depth)))
-
-  # create grids for each ssid separately
-  ssid <- 1
-  survey_abbrev <- "SYN QCS"
-  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-
-  unique(nd_1$year)
-
-  ssid <- 3
-  survey_abbrev <- "SYN HS"
-  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_3$year)
-
-  ssid <- 4
-  survey_abbrev <- "SYN WCVI"
-  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_4$year)
-
-  ssid <- 16
-  survey_abbrev <- "SYN WCHG"
-  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
-  unique(nd_16$year)
-
-  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>%
-    mutate(year = as.numeric(year))
-
-  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
-  nd_all <- left_join(nd_all, newcov, by = c("X", "Y"))
-  nd_all$trawled[is.na(nd_all$trawled)] <- 0 # correct missing values for small area on edge of QCS grid
-  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
-
-  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
-}
-```
-
-Save predictions for full spatial grid
-```{r}
-# fish <- readRDS(paste0("raw/bio-data-", spp, ""))
-
-rm(adult_predictions)
-rm(imm_predictions)
-rm(predictions)
-if(params$update_predictions) {
-if (any(names(data) == "adult_density")) {
-  # if (length(unique(fish$maturity_code)) > 2) {
-  nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-
-  try({
-    adult_biomass <- readRDS(paste0(
-      "data/", spp,
-      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
-
-    nd <- nd_all %>%
-      filter(ssid %in% model_ssid) %>%
-      filter(year %in% unique(adult_biomass$data$year))
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    adult_predictions <- predict(adult_biomass, newdata = nd)
-    saveRDS(adult_predictions, file = paste0(
-      "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
-    ))
-  })
-
-  try({
-    imm_biomass <- readRDS(paste0(
-      "data/", spp,
-      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
-
-    nd <- nd_all %>%
-      filter(ssid %in% model_ssid) %>%
-      filter(year %in% unique(imm_biomass$data$year))
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    imm_predictions <- predict(imm_biomass, newdata = nd)
-    saveRDS(imm_predictions, file = paste0(
-      "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
-    ))
-  })
-} else {
-  total_biomass <- readRDS(paste0(
-    "data/", spp,
-    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-
-  nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
-  nd <- nd_all %>%
-    filter(ssid %in% model_ssid) %>%
-    filter(year %in% unique(total_biomass$data$year))
-  nd <- na.omit(nd)
-  nd$year <- as.integer(nd$year)
-
-  predictions <- predict(total_biomass, newdata = nd)
-  saveRDS(predictions, file = paste0(
-    "data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
-  ))
-}
-}
-```
-
-```{r}
-
-if(params$update_model_check) {
-try({
-  biomass <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-
-try({
-  biomass <- readRDS(paste0(
-    "data/", spp,
-    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
-
-point_predictions <- predict(biomass)
-point_predictions$residuals <- residuals(biomass)
-
-saveRDS(point_predictions, file = paste0(
-  "data/", spp,
-  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-))
-}
-```
-
-```{r}
-depth_only_predictions <- readRDS(paste0(
-  "data/", spp,
-  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-))
-
-depth_only_predictions <- filter(
-  depth_only_predictions,
-  # ssid %in% c(1,3)) %>% filter (
-  year > 2007
-)
-
-try({
-  g <- ggplot(depth_only_predictions, aes(adult_density, residuals, colour = adult_density)) +
-    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
-})
-
-try({
-  g <- ggplot(depth_only_predictions, aes(density, residuals, colour = density)) +
-    geom_point() + scale_x_continuous(trans = "log10") + facet_wrap(~year)
-})
-
-g
-```
-
-```{r}
-try({
-  g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = adult_density)) +
-    geom_point() + # scale_x_continuous(trans = 'log10') +
-    facet_wrap(~year)
-})
-
-try({
-  g2 <- ggplot(depth_only_predictions, aes((est), residuals, colour = density)) +
-    geom_point() + # scale_x_continuous(trans = 'log10') +
-    facet_wrap(~year)
-})
-
-g2
-```
-
-```{r}
-ggplot(depth_only_predictions, aes(X, Y, colour = (residuals))) +
-  geom_point() +
-  scale_colour_gradient2() +
-  scale_x_continuous(trans = "log10") +
-  facet_wrap(~year)
-```
-
-### Build time-varying depth plots
 #### Model without environmental covariates
@@
     "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, covs, "-1n3n4n16-prior-", priors, ".rds"
   ))
+  # adult_biomass <- sdmTMB:::update_model(adult_biomass)
+
 })
-# adult_biomass<-sdmTMB:::update_model(adult_biomass)
 
@@
     "data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-1n3n4n16-prior-FALSE.rds"
+    "/mod-imm-biomass-", spp, covs, "-1n3n4n16.rds"
   ))
-  # imm_biomass<-sdmTMB:::update_model(imm_biomass)
+
+  # imm_biomass <- sdmTMB:::update_model(imm_biomass)
   # depth_model_list <- list(adult = adult_biomass, imm = imm_biomass)
@@
       "data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+      "/model-total-biomass-", spp, covs, "-1n3n4n16.rds"
     ))
@@
 depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
-
-d <- list()
-depth_plots <- list()
-for (i in seq_len(length(depth_model_list))) {
-  #if (depth_model_list[[i]]$model$convergence == 0) {
-    d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
-
-    if (length(d[[i]]) == 0) {
-      depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
-    } else {
-      d[[i]]$x <- exp(d[[i]]$x)
-      depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (without environmental variables)", xlimits = c(0, max_depth_found)) +
-        ggtitle(paste(species, names(depth_model_list[i])))
-    }
-  # } else {
-  #   depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
-  # }
-}
-print(depth_plots)
-```
-
-```{r}
-png(
-  file = paste0(
-    "figs/", spp,
-    "/depth-", spp, covs, ".png"
-  ),
-  res = 600,
-  units = "in",
-  width = 8.5,
-  height = 6
-)
-gridExtra::grid.arrange(
-  grobs = c(depth_plots),
-  nrow = 2,
-  top = grid::textGrob(paste(species, "(", covs, ")"))
-)
-dev.off()
 ```
@@
 ```{r}
-sdmTMB:::get_convergence_diagnostics(depth_model_list[[1]])
+write_grads <- function(x, macro, ...) {
+  paste0("\\newcommand{\\", macro, "}{", x, "}") %>%
+    readr::write_lines("model-grads.tex", append = TRUE)
+}
 ```
+
+```{r}
+write_grads(signif(max(depth_model_list$adult$gradients), digits = 8), paste0("new-", spp, "-mat"))
+write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0("new-", spp, "-imm"))
+write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0("new-", spp, "-total"))
+```

### 2020-11-12 analysis/VOCC/2-biomass-depth-only-model.Rmd
fix to work with dplyr and sdmTMB updates

@@
 ---
-title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
+title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth"
 author: "Philina English"
@@
 params: 
-    species: "Redbanded Rockfish"
+    species: "English Sole"
     # species: "Redstripe Rockfish"
@@
 
-covars <- readRDS("data/event-covariates.rds")
-data <- dplyr::left_join(biomass, covars)
+covars <- readRDS("data/event-covariates.rds") %>% select(-geometry)
+data <- left_join(biomass, covars)
 
@@
 if (params$update_model) {
+
+  # if updating model, will need new version of mesh to work with changes to sdmTMB
+  spde <- sdmTMB::make_mesh(data, c("X", "Y"), n_knots = knots, type = "kmeans")
+
   tictoc::tic()
@@
 
-   try({ imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year)", covariates, ""
-    ))
-
-    imm_biomass <- sdmTMB::sdmTMB(
-      data = data,
-      imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      # ar1_fields = TRUE, # changed to FALSE July 2020
-      # reml = TRUE,
-      include_spatial = TRUE,
-      nlminb_loops = 2, 
-      newton_steps = 2,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-
-    saveRDS(imm_biomass,
-      file = paste0("data/", spp, "/mod-imm-biomass-", 
-        spp, covs, "-", ssid_string, ".rds"))
-   })
-
+   # try({ imm_formula <- as.formula(paste(
+   #    "imm_density ~ 0 + as.factor(year)", covariates, ""
+   #  ))
+   # 
+   #  imm_biomass <- sdmTMB::sdmTMB(
+   #    data = data,
+   #    imm_formula,
+   #    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
+   #    time = "year", spde = spde,
+   #    family = tweedie(link = "log"),
+   #    # ar1_fields = TRUE, # changed to FALSE July 2020
+   #    # reml = TRUE,
+   #    include_spatial = TRUE,
+   #    nlminb_loops = 2, 
+   #    newton_steps = 2,
+   #    enable_priors = priors,
+   #    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+   #    silent = FALSE
+   #  )
+   # 
+   #  saveRDS(imm_biomass,
+   #    file = paste0("data/", spp, "/mod-imm-biomass-", 
+   #      spp, covs, "-", ssid_string, ".rds"))
+   # })
+   #  
   } else {
@@
 
-  try({
-    imm_biomass <- readRDS(paste0(
-      "data/", spp,
-      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
-
-    nd <- nd_all %>%
-      filter(ssid %in% model_ssid) %>%
-      filter(year %in% unique(imm_biomass$data$year))
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    imm_predictions <- predict(imm_biomass, newdata = nd)
-    saveRDS(imm_predictions, file = paste0(
-      "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
-    ))
-  })
+  # try({
+  #   imm_biomass <- readRDS(paste0(
+  #     "data/", spp,
+  #     "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  #   ))
+  # 
+  #   nd <- nd_all %>%
+  #     filter(ssid %in% model_ssid) %>%
+  #     filter(year %in% unique(imm_biomass$data$year))
+  #   nd <- na.omit(nd)
+  #   nd$year <- as.integer(nd$year)
+  # 
+  #   imm_predictions <- predict(imm_biomass, newdata = nd)
+  #   saveRDS(imm_predictions, file = paste0(
+  #     "data/", spp,
+  #     "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
+  #   ))
+  # })
 } else {

### 2020-11-12 analysis/VOCC/6-vocc-regression.R

@@ -23,10 +23,10 @@ age <- "both"
 no_chopsticks <- F
 w_genus <- F
 w_family <- F
-is_null <- T
+is_null <- F
 
-null_number <- "-3" # 2 failed, 3 NA
-setseed <- 42
+null_number <- "-1" # 2 failed, 3 NA
+setseed <- 10
 
 # ## for trends ###
 # knots <- 500

### 2020-11-14 analysis/VOCC/2-biomass-depth-only-model.Rmd
recheck for SST effect; plot covariate effects, residuals, and spatial random fields

--- 2020-07-31 analysis/VOCC/2-biomass-depth-only-model.Rmd
@@
 params: 
-    species: "Redbanded Rockfish"
-    # species: "Redstripe Rockfish"
+    # species: "Redbanded Rockfish"
+    #species: "Pacific Ocean Perch"
+    #species: "Pacific Cod"
+    #species: "Petrale Sole"
+    species: "Canary Rockfish"
+    # # species: "Redstripe Rockfish"
     # species: "Shortbelly Rockfish"
+    # species: "Lingcod"
     region: "All synoptic surveys"
@@
     #covs: "-trawled-ssid"
-    update_model: FALSE
+    update_model: TRUE
     update_predictions: FALSE
@@
 
-covars <- readRDS("data/event-covariates.rds")
+covars <- readRDS("data/event-covariates.rds") %>% select(-geometry)
 data <- dplyr::left_join(biomass, covars)
@@
 if (region == "All synoptic surveys") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
+  # spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
+  spde <- sdmTMB::make_mesh(data, c("X", "Y"), n_knots = knots, type = "kmeans")
 }
@@
         "data/", spp,
-        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-new.rds"
       )
@@
 
-   try({ imm_formula <- as.formula(paste(
-      "imm_density ~ 0 + as.factor(year)", covariates, ""
-    ))
-
-    imm_biomass <- sdmTMB::sdmTMB(
-      data = data,
-      imm_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      # ar1_fields = TRUE, # changed to FALSE July 2020
-      # reml = TRUE,
-      include_spatial = TRUE,
-      nlminb_loops = 2, 
-      newton_steps = 2,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-
-    saveRDS(imm_biomass,
-      file = paste0("data/", spp, "/mod-imm-biomass-", 
-        spp, covs, "-", ssid_string, ".rds"))
-   })
+   # try({ imm_formula <- as.formula(paste(
+   #    "imm_density ~ 0 + as.factor(year)", covariates, ""
+   #  ))
+   # 
+   #  imm_biomass <- sdmTMB::sdmTMB(
+   #    data = data,
+   #    imm_formula,
+   #    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
+   #    time = "year", spde = spde,
+   #    family = tweedie(link = "log"),
+   #    # ar1_fields = TRUE, # all off for final version
+   #    # reml = TRUE,
+   #    include_spatial = TRUE,
+   #    nlminb_loops = 2, 
+   #    newton_steps = 2,
+   #    enable_priors = priors,
+   #    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+   #    silent = FALSE
+   #  )
+   # 
+   #  saveRDS(imm_biomass,
+   #    file = paste0("data/", spp, "/mod-imm-biomass-", 
+   #      spp, covs, "-", ssid_string, "-new.rds"))
+   # })
 
   } else {
-    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
-
-    total_biomass <- sdmTMB::sdmTMB(
-      data = data,
-      dens_formula,
-      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
-      time = "year", spde = spde,
-      family = tweedie(link = "log"),
-      # ar1_fields = TRUE, # all off in July 2020
-      include_spatial = TRUE,
-      nlminb_loops = 2, 
-      newton_steps = 2,
-      enable_priors = priors,
-      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-      silent = FALSE
-    )
-
-    saveRDS(total_biomass, file = paste0(
-      "data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
-    ))
+    # dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
+    # 
+    # total_biomass <- sdmTMB::sdmTMB(
+    #   data = data,
+    #   dens_formula,
+    #   time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
+    #   time = "year", spde = spde,
+    #   family = tweedie(link = "log"),
+    #   # ar1_fields = TRUE, # all off for final version
+    #   include_spatial = TRUE,
+    #   nlminb_loops = 2, 
+    #   newton_steps = 2,
+    #   enable_priors = priors,
+    #   control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+    #   silent = FALSE
+    # )
+    # 
+    # saveRDS(total_biomass, file = paste0(
+    #   "data/", spp,
+    #   "/model-total-biomass-", spp, covs, "-", ssid_string, "-new.rds"
+    # ))
   }
@@
       "data/", spp,
-      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+      "/mod-imm-biomass-", spp, covs, "-", ssid_string, ".rds"
     ))
@@
     "data/", spp,
-    "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
   ))
@@
     "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-imm-biomass-", spp, covs, "-", ssid_string, ".rds"
   ))
@@
   "data/", spp,
-  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  "/check-mod-predictions-", spp, covs, "-", ssid_string, ".rds"
 ))
@@
     "data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-1n3n4n16-prior-FALSE.rds"
+    "/mod-imm-biomass-", spp, covs, "-1n3n4n16.rds"
   ))
@@
     "figs/", spp,
-    "/depth-", spp, covs, ".png"
+    "/depth-", spp, covs, "-no-AR1.png"
   ),
@@
 write_grads(signif(max(depth_model_list$adult$gradients), digits = 8), paste0(spp, "-mat"))
-write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0(spp, "-imm"))
-write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0(spp, "-total"))
+write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0(spp, "-imm-redo"))
+write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0(spp, "-total-redo"))
 ```
@@
     imm_biomass <- sdmTMB::run_extra_optimization(depth_model_list$imm, nlminb_loops = 2, newton_steps = 2)
-    write_grads(signif(max(imm_biomass$gradients), digits = 8), paste0(spp, "-imm2"))
+    write_grads(signif(max(imm_biomass$gradients), digits = 8), paste0(spp, "-imm3"))
   }
@@
     total_biomass <- sdmTMB::run_extra_optimization(depth_model_list$total, nlminb_loops = 2, newton_steps = 2)
-    write_grads(signif(max(total_biomass$gradients), digits = 8), paste0(spp, "-total2"))
+    write_grads(signif(max(total_biomass$gradients), digits = 8), paste0(spp, "-total3"))
   }
@@
       file = paste0("data/", spp, "/mod-imm-biomass-", 
-        spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
+        spp, covs, "-", ssid_string, ".rds"))
 

### 2020-11-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
recheck for SST effect; plot covariate effects, residuals, and spatial random fields

@@
 library(dplyr)
+library(gfranges)
+
+plot_map_raster <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    coord_fixed()+
+gfplot::theme_pbs() + theme(
+  axis.title = element_blank(), 
+  axis.ticks = element_blank(),
+  axis.text = element_blank()
+  )
+}
+```
+
+Model diagnostics function
+```{r}
+get_diag <- function(m, response = "temperature_c", variable = "depth_scaled", colour_var = "depth", start_year = 2007, ssids = c(1,3,4,16)) {
+  predictions <- predict(m)
+  predictions$residuals <- residuals(m)
+
+  predictions <- predictions %>% filter (year >= start_year)
+
+  plot_map <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", colour = column)) +
+      geom_point() +
+      coord_fixed()+
+      gfplot::theme_pbs() 
+  }
+
+  g <- plot_map(predictions, "est") +
+    scale_colour_viridis_c() +
+    ggtitle("Prediction (fixed effects + all random effects)")
+  print(g)
+
+  g <- plot_map(predictions, "est_non_rf") +
+    ggtitle("Prediction (fixed effects only)") +
+    scale_colour_viridis_c()
+  print(g)
+
+  g <- plot_map(predictions, "est_rf") +
+    ggtitle("All random effects only") +
+    scale_colour_gradient2()
+  print(g)
+
+  g <- plot_map(predictions, "omega_s") +
+    ggtitle("Spatial random effects only") +
+    scale_colour_gradient2()
+  print(g)
+
+  g <- plot_map(predictions, "epsilon_st") +
+    ggtitle("Spatiotemporal random effects only") +
+    facet_wrap(~year) +
+    scale_colour_gradient2()
+  print(g)
+
+  g <- ggplot(predictions, aes_string("est", response)) +
+    geom_point(alpha = 0.2) +
+    facet_wrap(~year) +
+    coord_fixed() +
+    geom_abline()+
+    gfplot::theme_pbs() 
+  print(g)
+
+  g <- ggplot(predictions, aes(est, residuals)) +
+    geom_point(alpha = 0.2) +
+    geom_smooth()+
+    gfplot::theme_pbs() 
+  print(g)
+
+  g <- ggplot(predictions, aes_string(variable, "residuals", colour = colour_var)) +
+    geom_point(alpha = 0.4, size = 1.2) +
+    geom_smooth(colour="grey", size = 1.2) +
+    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
+      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
+    # facet_wrap(~year, scales = "free_x")+
+    gfplot::theme_pbs() + theme(
+      axis.text.y = element_text(size = 14),
+      axis.text.x = element_blank(), axis.ticks.x = element_blank())
+  print(g)
+
+  g <- ggplot(predictions, aes_string(variable, "residuals", colour = colour_var)) +
+    geom_point(alpha = 0.4, size = 0.7) +
+    geom_smooth(colour="grey", size = 0.7) +
+    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
+      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
+    facet_wrap(~year, scales = "free_x")+
+    gfplot::theme_pbs() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
+  print(g)
+
+    g <- ggplot(filter(predictions, ssid %in% ssids), aes_string(variable, "residuals", colour = colour_var)) +
+    geom_point(alpha = 0.4, size = 1) +
+    geom_smooth(colour="grey", size = 1) +
+    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
+      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
+    facet_wrap(~ssid, scales = "free_x")+
+    gfplot::theme_pbs() 
+  print(g)
+
+  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
+    geom_point(alpha = 0.2) +
+    coord_fixed() +
+    scale_color_gradient2() +
+    facet_wrap(~year)+
+    gfplot::theme_pbs() 
+  print(g)
+
+  aic <- function(m) {
+    k <- length(m$model$par)
+    nll <- m$model$objective
+    2 * k - 2 * (-nll)
+  }
+
+  print("R^2:")
+  r2 <- cor(predictions$est, predictions[[response]])^2
+  print(r2)
+
+  print("")
+  print("AIC:")
+  print(aic(m))
+}
 ```
@@
 ```{r}
-spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 800)
-plot_spde(spde)
+spde <- sdmTMB::make_mesh(d_trawl, c("X", "Y"), n_knots = 800, type = "kmeans")
+plot(spde)
+```
+Original model code
+```{r eval = F}
+# spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 800)
+# m_temp_rw <- sdmTMB(
+#   temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
+#   time_varying = ~ 0 + depth_scaled + depth_scaled2,
+#   data = d_trawl,
+#   ar1_fields = TRUE, # changed AR1 to true
+#   include_spatial = TRUE, 
+#   time = "year", spde = spde, 
+#   family = gaussian(link = "identity"),
+#   nlminb_loops = 2,
+#   newton_steps = 1,
+#   silent = FALSE
+# )
+# saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds")))
 ```
 
+New CV model code
 ```{r eval = F}
-m_temp_rw <- sdmTMB(
+m_temp_cv <- sdmTMB_cv(
   temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
@@
   family = gaussian(link = "identity"),
-  nlminb_loops = 2,
-  newton_steps = 1,
+  k_folds = 5,
+  # nlminb_loops = 2,
+  # newton_steps = 1,
   silent = FALSE
 )
+saveRDS(m_temp_cv, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-cv.rds")))
 
-saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds")))
+m_temp_cv <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-cv.rds")))
+m_temp_cv$fold_loglik
+m_temp_cv$sum_loglik
+m_temp_cv$max_gradients
 ```
@@
 
-Model diagnostics functions
-```{r}
-get_diag <- function(m, response = "log_do", variable = "depth_scaled") {
-  predictions <- predict(m)
-  predictions$residuals <- residuals(m)
-
-  plot_map <- function(dat, column = "est") {
-    ggplot(dat, aes_string("X", "Y", colour = column)) +
-      geom_point() +
-      coord_fixed()
-  }
-
-  g <- plot_map(predictions, "est") +
-    scale_colour_viridis_c() +
-    ggtitle("Prediction (fixed effects + all random effects)")
-  print(g)
-
-  g <- plot_map(predictions, "est_non_rf") +
-    ggtitle("Prediction (fixed effects only)") +
-    scale_colour_viridis_c()
-  print(g)
-
-  g <- plot_map(predictions, "est_rf") +
-    ggtitle("All random effects only") +
-    scale_colour_gradient2()
-  print(g)
-
-  g <- plot_map(predictions, "omega_s") +
-    ggtitle("Spatial random effects only") +
-    scale_colour_gradient2()
- print(g)
-
-   g <- plot_map(predictions, "epsilon_st") +
-    ggtitle("Spatiotemporal random effects only") +
-    facet_wrap(~year) +
-    scale_colour_gradient2()
- print(g)
-
-  g <- ggplot(predictions, aes_string("est", response)) +
-    geom_point(alpha = 0.2) +
-    facet_wrap(~year) +
-    coord_fixed() +
-    geom_abline()
-  print(g)
-
-  g <- ggplot(predictions, aes(est, residuals)) +
-    geom_point(alpha = 0.2) +
-    geom_smooth()
-  print(g)
-
-  g <- ggplot(predictions, aes_string(variable, "residuals")) +
-    geom_point(alpha = 0.2) +
-    geom_smooth() +
-    facet_wrap(~year)
-  print(g)
-
-  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
-    geom_point(alpha = 0.2) +
-    coord_fixed() +
-    scale_color_gradient2()+
-    facet_wrap(~year)
-  print(g)
-
-  aic <- function(m) {
-    k <- length(m$model$par)
-    nll <- m$model$objective
-    2 * k - 2 * (-nll)
-  }
-
-  print("R^2:")
-  r2 <- cor(predictions$est, predictions[[response]])^2
-  print(r2)
-
-  print("")
-  print("AIC:")
-  print(aic(m))
-}
-```
-
 ```{r eval=FALSE}
@@
 ```{r eval=FALSE}
-all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
-# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
-# glimpse(all_sensor)
+# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
+# # all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
+# # glimpse(all_sensor)
+# 
+# all_depth <- all_sensor %>%
+#   # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
+#   # filter missing location data and trial year
+#   dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%
+# 
+#   # convert lat and lon to UTMs
+#   dplyr::mutate(X = longitude, Y = latitude) %>%
+#   gfplot:::ll2utm(., utm_zone = 9) %>% 
+# 
+#   # interpolate missing depth values
+#   dplyr::rename(depth = depth_m) %>%
+#   gfplot:::interp_survey_bathymetry()
 
-all_depth <- all_sensor %>%
-  # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
-  # filter missing location data and trial year
-  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%
+# d_trawl <- all_depth$data %>% 
+#   # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
+#   filter(depth_max>10) %>% 
+#   gfplot:::scale_survey_predictors()
 
-  # convert lat and lon to UTMs
-  dplyr::mutate(X = longitude, Y = latitude) %>%
-  gfplot:::ll2utm(., utm_zone = 9) %>%
-
-  # interpolate missing depth values
-  dplyr::rename(depth = depth_m) %>%
-  gfplot:::interp_survey_bathymetry()
-
-d_trawl <- all_depth$data %>% 
-  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
-  filter(depth_max>10) %>% 
-  gfplot:::scale_survey_predictors()
+olddata <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn.rds"))) 
+d_trawl <- olddata$data
 
@@
 
-d_trawl2[is.na(d_trawl2$shallow), ]
-d_trawl2[is.na(d_trawl2$SST_scaled), ]
-
-# glimpse(d_trawl2)
+# d_trawl2[is.na(d_trawl2$shallow), ]
+# d_trawl2[is.na(d_trawl2$SST_scaled), ]
 ```
@@
 Does a fixed interaction with SST_scaled do any better?
+Rerun main model with SST dataset for comparisons
 ```{r}
-spde2 <- make_spde(d_trawl2$X, d_trawl2$Y, n_knots = 500)
-plot_spde(spde2)
+spde <- make_mesh(d_trawl2, c("X", "Y"), n_knots = 800)
+
+m_temp_new <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year), 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  data = d_trawl2, spde = spde,
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
+  silent = FALSE
+)
+saveRDS(m_temp_new, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-sstdata.rds")))
+m_temp_new
+AIC(m_temp_new)
 ```
 
-```{r eval=FALSE}
-m_temp2 <- sdmTMB(
-  temperature_c ~ 0 + as.factor(year) + DOY_scaled * shallow + DOY_scaled2 * shallow,
-  data = d_trawl2,
+```{r}
+spde <- make_mesh(d_trawl2, c("X", "Y"), n_knots = 800)
+
+m_temp_sst <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow, 
+  data = d_trawl2, spde = spde,
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
-  ar1_fields = TRUE, include_spatial = TRUE,
-  time = "year", spde = spde2, family = gaussian(link = "identity"),
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
   nlminb_loops = 2,
@@
 
-saveRDS(m_temp2, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY.rds")))
-
-m_temp2
+saveRDS(m_temp_sst, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-SST.rds")))
+m_temp_sst
+AIC(m_temp_sst)
 ```
 
+```{r}
+m_temp_new <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-sstdata.rds")))
+m_temp_sst <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-SST.rds")))
+get_diag(m_temp_new, response = "temperature_c")
+get_diag(m_temp_sst, response = "temperature_c")
+```
+
+# try with fewer knots
+```{r eval = F}
+spde3 <- sdmTMB::make_mesh(d_trawl2, c("X", "Y"), n_knots = 400, type = "kmeans")
+
+m_temp1aic <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year),
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  data = d_trawl2, spde = spde3,
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
+  silent = FALSE
+)
+saveRDS(m_temp1aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn-new.rds")))
+
+m_temp1aic <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn-new.rds")))
+# m_temp1aic<- sdmTMB:::update_model(m_temp1aic)
+AIC(m_temp1aic)
+```
+
+Run CV on main model
+```{r}
+# spde2 <- make_spde(d_trawl2$X, d_trawl2$Y, n_knots = 800)
+spde2 <- sdmTMB::make_mesh(d_trawl2, c("X", "Y"), n_knots = 200, type = "kmeans")
+plot(spde2)
+```
+
+```{r}
+m_temp1 <- sdmTMB_cv(
+  temperature_c ~ 0 + as.factor(year), 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  data = d_trawl2, spde = spde3, 
+  ar1_fields = TRUE, 
+  include_spatial = TRUE, 
+  time = "year", 
+  family = gaussian(link = "identity"),
+  k_folds = 20,
+  # nlminb_loops = 2,
+  # newton_steps = 1,
+  silent = FALSE
+)
+# saveRDS(m_temp1, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn-cv20.rds")))
+saveRDS(m_temp1, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-150kn-cv20.rds")))
+
+# m_temp1 <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-300kn-cv.rds")))
+m_temp1$fold_loglik
+m_temp1$sum_loglik
+m_temp1$max_gradients
+m_temp1$pdHess
+```
+
+Check if DOY helps model
 ```{r eval=FALSE}
-m_temp3 <- sdmTMB(
-  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow + DOY_scaled * shallow + DOY_scaled2 * shallow,
-  data = d_trawl2,
+spde3 <- sdmTMB::make_mesh(d_trawl2, c("X", "Y"), n_knots = 400, type = "kmeans")
+
+m_temp2aic <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year) + DOY_scaled + DOY_scaled2,
+  data = d_trawl2, spde = spde3, 
   time_varying = ~ 0 + depth_scaled + depth_scaled2,
-  ar1_fields = TRUE, include_spatial = TRUE,
-  time = "year", spde = spde2, family = gaussian(link = "identity"),
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
+  # k_folds = 20,
+  silent = TRUE
+)
+
+saveRDS(m_temp2aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-400kn.rds")))
+m_temp2aic <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-400kn.rds")))
+m_temp2aic
+AIC(m_temp2aic)
+
+# m_temp2cv <- sdmTMB_cv(
+#   temperature_c ~ 0 + as.factor(year) + DOY_scaled + DOY_scaled2,
+#   data = d_trawl2, spde = spde2, 
+#   time_varying = ~ 0 + depth_scaled + depth_scaled2,
+#   ar1_fields = TRUE, 
+#   include_spatial = TRUE,
+#   time = "year", 
+#   family = gaussian(link = "identity"),
+#   # nlminb_loops = 2,
+#   # newton_steps = 1,
+#   k_folds = 20,
+#   silent = TRUE
+# )
+# 
+# saveRDS(m_temp2, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-cv20.rds")))
+# m_temp2cv <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-cv20.rds")))
+# m_temp2cv
+# m_temp2cv$fold_loglik
+# m_temp2cv$sum_loglik
+# m_temp2cv$max_gradients
+```
+
+Check if SST helps model 
+```{r}
+m_temp3aic <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow, 
+  data = d_trawl2, spde = spde3, 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
   nlminb_loops = 2,
@@
 )
+saveRDS(m_temp3aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-400kn.rds")))
+# m_temp3aic <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-400kn.rds"))
+m_temp3aic <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-400kn.rds")))
+AIC(m_temp1aic, m_temp2aic, m_temp3aic)
 
-saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds")))
+m_temp3aic
+```
 
-m_temp3
+Check if SST and DOY help model
+```{r eval=FALSE}
+m_temp4aic <- sdmTMB(
+  temperature_c ~ 0 + as.factor(year) + SST_scaled * DOY_scaled * shallow, 
+  data = d_trawl2, spde = spde3, 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
+  nlminb_loops = 2,
+  newton_steps = 1,
+  silent = TRUE
+)
+saveRDS(m_temp4aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-400kn.rds")))
+
+m_temp4aic <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-400kn.rds"))
+# m_temp4aic
+AIC(m_temp1aic, m_temp2aic,  m_temp3aic, m_temp4aic)
+```
+
+```{r eval=FALSE}
+m_temp3 <- sdmTMB_cv(
+  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow, 
+  data = d_trawl2, spde = spde2, 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
+  # nlminb_loops = 2,
+  # newton_steps = 1,
+  k_folds = 20,
+  silent = TRUE
+)
+
+# 500 knot early version
+# saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds")))
+
+# new 800 knot version for cross validation
+saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-cv20.rds")))
+
+# m_temp3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-cv20.rds"))
+# m_temp3
+m_temp3$fold_loglik
+m_temp3$sum_loglik
+m_temp3$max_gradients
+```
+
+Check if SST and DOY help model using cv
+```{r eval=FALSE}
+m_temp4 <- sdmTMB_cv(
+  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow + SST_scaled * DOY_scaled * shallow, 
+  data = d_trawl2, spde = spde2, 
+  time_varying = ~ 0 + depth_scaled + depth_scaled2,
+  ar1_fields = TRUE, 
+  include_spatial = TRUE,
+  time = "year", 
+  family = gaussian(link = "identity"),
+  # nlminb_loops = 2,
+  # newton_steps = 1,
+  k_folds = 20,
+  silent = TRUE
+)
+
+# 500 knot early version
+# saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds")))
+
+# new 800 knot version for cross validation
+saveRDS(m_temp4, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-cv20.rds")))
+
+# m_temp4 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-cv20.rds"))
+# m_temp4
+m_temp4$fold_loglik
+m_temp4$sum_loglik
+m_temp4$max_gradients
+m_temp4$pdHess
+```
+
+```{r}
+m_temp_sst <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds"))
+
+m_temp_sst <- sdmTMB:::update_model(m_temp_sst)
+
 ```
@@
 ```{r eval=FALSE}
-get_diag(m_temp3, response = "temperature_c", variable = "depth_scaled")
+get_diag(m_temp_new, response = "temperature_c", variable = "depth_scaled", colour_var = "depth", start_year = 2002)
+get_diag(m_temp_new, response = "temperature_c", variable = "DOY", colour_var = "depth", start_year = 2002, ssids = c(1,3,4, 16))
 ```
@@
 
-saveRDS(p_nd, file = "../VOCC/data/predicted_temp_allyears_800kn.rds")
+# saveRDS(p_nd, file = "../VOCC/data/predicted_temp_allyears_800kn.rds")
 ```
@@
 ```{r}
-# p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")
-
-plot_map_raster <- function(dat, column = "est") {
-    ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    coord_fixed()+
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
-}
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")
 
@@
 
+Time-varying depth 
+```{r eval}
+# pd <- expand.grid(
+#   depth_scaled = seq(min(m_temp_rw$data$depth_scaled),
+#     max(m_temp_rw$data$depth_scaled), length.out = 50),
+#   year = unique(m_temp_rw$data$year)
+# )
+# pd$depth_scaled2 <- pd$depth_scaled^2
+# 
+# # p <- predict(m_temp_rw, newdata = pd, se_fit = TRUE, re_form = NA, xy_cols = c("X", "Y"))
+# p <- predict(m_temp_new, newdata = pd, se_fit = TRUE, re_form = NA)
+# 
+# p$Depth <- exp(p$depth_scaled*0.661 + 5.07)
+# p$Temperature <- p$est
+# # 
+# p$lowCI <- (p$est - 1.96 * p$est_se)
+# p$highCI <- (p$est + 1.96 * p$est_se)
+# # 
+# saveRDS(p, here::here("analysis/tmb-sensor-explore/data/depth-tv-m-temp-rw-800-kn.rds"))
+p <- readRDS(here::here("analysis/tmb-sensor-explore/data/depth-tv-m-temp-rw-800-kn.rds"))
+
+ggplot(filter(p, Depth >15 & year >2007), aes(Depth, Temperature,
+  ymin = lowCI, 
+  ymax = highCI, 
+  group = as.factor(year), fill=year, colour = year)) +
+  geom_line(size = 0.5, alpha =0.85) +
+  # geom_smooth(method= "loess", size = 0.4, alpha =0.85, se =F ) +
+  # geom_smooth(span = 0.5, size = 0.4, alpha =0.85, se =F ) +
+  scale_fill_viridis_c(option = "C") +
+  scale_colour_viridis_c(option = "C") +
+  # ylab("DO") +
+  geom_ribbon(alpha = 0.2, colour = NA) +
+  # scale_y_continuous(limits = c(0,4.5)) +
+  scale_x_reverse(limits = c(600,20)) +
+  ylim( 2, 15) +
+  coord_flip() +
+  # scale_x_continuous(limits = c(18,210)) +
+  gfplot::theme_pbs() +
+  theme(legend.position = c(.8,.4))
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/depth-tv-m-temp-rw-800-kn.png"), width = 3, height = 2.5)
+
+# ggplot(filter(p, Depth >15 & year == 2018), aes(Depth, Temperature,
+#   ymin = (Temperature - 1.96 * est_se), 
+#   ymax = (Temperature + 1.96 * est_se))) +
+#   geom_line(size = 1.5, alpha =0.85) + 
+#   # scale_color_viridis_c(option = "C") +
+#   geom_ribbon(alpha=0.2) +
+#   scale_x_reverse(limits = c(210,20)) +
+#   ylim(5, 15) +
+#   coord_flip() +
+#   gfplot::theme_pbs() + 
+#   theme(legend.position = c(.8,.4))
+
+```
+
 ```{r}
@@
 ```{r}
-omega_s_all <- plot_map_raster(p_nd, "omega_s") +
+(omega_s_all <- plot_map_raster(p_nd, "omega_s") +
   ggtitle("Spatial random effects only") +
@@
  high = "Red 3", limits = c(-1.5,1.5),
-    na.value = "red") +  labs(fill = "") + 
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))
+    na.value = "red") +  labs(fill = "") + theme(legend.position = c(0.2, 0.3))
+  )
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/omegas-temp-rw-800-kn.png"), width = 3, height = 3)
 
 omega_s_all
+
 ```
@@
 ```{r}
-p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")
+
+p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4) %>% filter(year>2007)
 
@@
   ggtitle("Spatiotemporal random effects") +
-  facet_wrap(~year) +
+  facet_grid(~year) +
    scale_fill_gradient2( low = "Steel Blue 4",
@@
   labs(fill = "epsilon") + 
-  gfplot::theme_pbs() + 
   theme(
-  legend.position = "none", 
-  axis.title.x = element_blank(), axis.title.y = element_blank()))
+  legend.position = "none"))
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/temp-epsilon-1n3.png"), width =6, height = 2)
 
@@
 ```{r}
-p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
-st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")
+
+p_nd4 <- p_nd %>% filter(ssid==4) %>% filter(year>2007)#%>% filter(ssid!=4)
+(st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
   #ggtitle(" ") +
-  facet_wrap(~year) +
+  facet_grid(~year) +
      scale_fill_gradient2( low = "Steel Blue 4",
@@
     na.value = "red") +
-          labs(fill = " ") + 
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.15))
+          labs(fill = " ") + theme( legend.position = "none") )
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/temp-epsilon-4.png"), width = 6, height = 2)
+
 st_4
@@
 ```{r}
-p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
-st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
+p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")
+
+p_nd16 <- p_nd %>% filter(ssid==16) %>% filter(year>2007)#%>% filter(ssid!=4)
+(st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
   #ggtitle(" ") +
-  facet_wrap(~year) +
+  facet_grid(~year) +
    scale_fill_gradient2( low = "Steel Blue 4",
@@
     na.value = "red") +
-  xlim(65,345) +
-    ylim(5800,6100) +
-gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
+  xlim(20,360) + ylim(5850,6050)+ 
+  # xlim(65,345) + ylim(5800,6100)+ 
+    theme(panel.border = element_blank(), legend.position = "none") )
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/temp-epsilon-16.png"), width = 6, height = 1)
 st_16

### 2020-11-14 analysis/tmb-sensor-explore/3-fit-do.Rmd
for DO model: plot covariate effects, residuals, and spatial random fields

@@
 library(gfranges)
+
+plot_map_raster <- function(dat, column = "est") {
+    ggplot(dat, aes_string("X", "Y", fill = column)) +
+    geom_raster() +
+    coord_fixed()+
+gfplot::theme_pbs() + theme(
+  axis.title = element_blank(), 
+  axis.ticks = element_blank(),
+  axis.text = element_blank()
+  )
+}
 ```
@@
 ```{r}
-get_diag <- function(m, response = "log_do", variable = "depth_scaled") {
+get_diag <- function(m, response = "log_do", variable = "depth_scaled", colour_var =  "raw_depth") {
   predictions <- predict(m)
@@
       geom_point() +
-      coord_fixed()
+      coord_fixed()+
+      gfplot::theme_pbs() 
   }
@@
     coord_fixed() +
-    geom_abline()
+    geom_abline()+
+    gfplot::theme_pbs() 
   print(g)
@@
     geom_point(alpha = 0.2) +
-    geom_smooth()
+    geom_smooth()+
+    gfplot::theme_pbs() 
   print(g)
 
-  g <- ggplot(predictions, aes_string(variable, "residuals")) +
-    geom_point(alpha = 0.2) +
-    geom_smooth() +
-    facet_wrap(~year)
+g <- ggplot(predictions, aes_string(variable, "residuals", colour = colour_var)) +
+    # geom_point(alpha = 0.2) +
+    geom_point(alpha = 0.4, size = 0.7) +    
+    geom_smooth(colour="grey", size = 0.7) +
+    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, trans= sqrt) +
+    facet_wrap(~year, scales = "free_x")+
+    gfplot::theme_pbs() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
+  # g <- ggplot(predictions, aes_string(variable, "residuals")) +
+  #   geom_point(alpha = 0.2) +
+  #   geom_smooth() +
+  #   facet_wrap(~year)+
+  #   gfplot::theme_pbs() 
   print(g)
@@
     scale_color_gradient2() +
-    facet_wrap(~year)
+    facet_wrap(~year)+
+    gfplot::theme_pbs() 
   print(g)
@@
 ```{r eval=FALSE}
+
+# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-800kn.rds")))
+d_trawl1 <- do_model_all_DOY$data
+spde <- make_mesh(d_trawl1, xy_cols=c("X", "Y"), n_knots = 800, type = "kmeans")
+plot(spde)
 do_model_all_DOY <- sdmTMB::sdmTMB(
@@
     "800kn", # Sys.Date(),
-    ".rds"
+    "-update.rds"
   )
@@
 # do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds"))) # 800 knots
-do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-800kn.rds"))) # add in some WCVI 2016 data
-get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
+do_model_all_DOY_old <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-800kn.rds"))) # add in some WCVI 2016 data
+do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-800kn-update.rds"))) # rerun in new format. No other changes
+# do_model_all_DOY <- sdmTMB:::update_model(do_model_all_DOY)
+get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled", colour_var =  "raw_depth")
+
+get_diag(do_model_all_DOY, response = "log_do", variable = "DOY_scaled", colour_var =  "raw_depth")
+get_diag(do_model_all_DOY, response = "log_do", variable = "temp_scaled", colour_var =  "raw_depth")
 ```
@@
 
+Time-varying depth 
+```{r eval = F}
+# pd <- expand.grid(
+#   temp_scaled = 0,
+#   temp_scaled2 = 0,
+#   DOY_scaled = 0, 
+#   depth_scaled = seq(min(do_model_all_DOY$data$depth_scaled),
+#     max(do_model_all_DOY$data$depth_scaled), length.out = 30),
+#   year = unique(do_model_all_DOY$data$year)
+# )
+# pd$depth_scaled2 <- pd$depth_scaled^2
+# 
+# p <- predict(do_model_all_DOY, newdata = pd, se_fit = TRUE, re_form = NA)
+# 
+# p$DO <- exp(p$est)
+# p$lowCI <- exp(p$est - 1.96 * p$est_se)
+# p$highCI <- exp(p$est + 1.96 * p$est_se)
+# 
+# p$Depth <- exp(p$depth_scaled*do_model_all_DOY$data$depth_sd[1]+do_model_all_DOY$data$depth_mean[1])
+# saveRDS(p, here::here("analysis/tmb-sensor-explore/data/depth-tv-m-do-DOY-800kn.rds"))
+
+p <- readRDS(here::here("analysis/tmb-sensor-explore/data/depth-tv-m-do-DOY-800kn.rds"))
+
+ggplot(filter(p, year != 2007), aes(Depth, exp(est), 
+  ymin = lowCI,
+  ymax = highCI,
+  group = as.factor(year), fill = year, colour = year)) +
+  geom_line(size = 0.5, alpha =0.85) +
+  # geom_smooth(method= "loess", size = 0.4, alpha =0.85, se =F ) +
+  # geom_smooth(span = 0.5, size = 0.4, alpha =0.85, se =F ) +
+  scale_fill_viridis_c(option = "C") +
+  scale_colour_viridis_c(option = "C") +
+  ylab("DO") +
+  geom_ribbon(alpha = 0.2, colour = NA) +
+  scale_y_continuous(limits = c(0.5,5)) +
+  scale_x_reverse(limits = c(650,20)) +
+  coord_flip() +
+  # scale_x_continuous(limits = c(18,210)) +
+  gfplot::theme_pbs() +
+  theme(legend.position = c(.8,.4))
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/depth-tv-m-do-DOY-800-kn.png"), width = 3, height = 2.5)
+```
+
+Temperature effect
+```{r}
+### temp
+min(do_model_all_DOY$data$temp_scaled)
+# 
+# pd <- expand.grid(
+#   temp_scaled = seq(min(do_model_all_DOY$data$temp_scaled) + 0.1,
+#     max(do_model_all_DOY$data$temp_scaled) - 0.1, length.out = 20),
+#   DOY_scaled = 0, 
+#   depth_scaled = 0,
+#   depth_scaled2 = 0,
+#   # year = c(2017L, 2018L)
+#   year = unique(do_model_all_DOY$data$year) # seems to need multiple years? why?
+# ) %>% filter(year>2016)
+# # pd$depth_scaled2 <- pd$depth_scaled^2
+# pd$temp_scaled2 <-  pd$temp_scaled^2
+# # p <- predict(do_model_all_DOY, newdata = pd, se_fit = TRUE, re_form = NA, xy_cols = c("X", "Y"))
+# p <- predict(do_model_all_DOY, newdata = pd, se_fit = TRUE, re_form = NA)
+# 
+# p$DO <- exp(p$est)
+# p$lowCI <- exp(p$est - 1.96 * p$est_se)
+# p$highCI <- exp(p$est + 1.96 * p$est_se)
+# 
+# 
+p$Temperature <- do_model_all_DOY$data$temp_sd[1]*p$temp_scaled+do_model_all_DOY$data$temp_mean[1]
+saveRDS(p, here::here("analysis/tmb-sensor-explore/data/temp-m-do-DOY-800-kn.rds"))
+p <- readRDS(here::here("analysis/tmb-sensor-explore/data/temp-m-do-DOY-800-kn.rds"))
+
+ggplot(filter(p, Temperature < 15 & year <2018), aes(Temperature, DO, 
+  ymin = lowCI, 
+  ymax = highCI)) +
+  geom_line(size = 1.5, alpha =0.85) + 
+  # scale_color_viridis_c(option = "C") +
+  geom_ribbon(alpha=0.2) +
+  # ylim(20, 210) +
+  gfplot::theme_pbs() + 
+  theme(legend.position = c(.8,.4))
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/temp-m-do-DOY-800-kn.png"), width = 3, height = 2.5)
+
+```
+
+Seasonal effect
+```{r}
+# pd <- expand.grid(
+#   temp_scaled = 0,
+#   temp_scaled2 = 0,
+#   DOY_scaled = seq(min(do_model_all_DOY$data$DOY_scaled) + 0.1,
+#     max(do_model_all_DOY$data$DOY_scaled) - 0.1, length.out = 5),
+#   depth_scaled = 0,
+#   depth_scaled2 = 0,
+#   year = unique(do_model_all_DOY$data$year) # seems to need multiple years? why?
+# ) 
+# 
+# p <- predict(do_model_all_DOY, newdata = pd, se_fit = TRUE, re_form = NA)
+# 
+# p$DO <- exp(p$est)
+# p$lowCI <- exp(p$est - 1.96 * p$est_se)
+# p$highCI <- exp(p$est + 1.96 * p$est_se)
+# 
+# p$DOY <- p$DOY_scaled*do_model_all_DOY$data$DOY_sd[1]+ do_model_all_DOY$data$DOY_mean[1]
+# 
+# saveRDS(p, here::here("analysis/tmb-sensor-explore/data/DOY-do-800-kn.rds"))
+p <- readRDS(here::here("analysis/tmb-sensor-explore/data/DOY-do-800-kn.rds"))
+
+ggplot(filter(p), aes(DOY, DO, 
+  ymin = lowCI, 
+  ymax = highCI, 
+  group= as.factor(year), fill = year, colour = year)) +
+  geom_line(size = 1.5, alpha =0.85) + 
+  scale_fill_viridis_c(option = "C") +
+  scale_color_viridis_c(option = "C") +
+  geom_ribbon(alpha = 0.05, colour = NA) +
+  # ylim(20, 210) +
+  coord_cartesian(ylim = c(1,5)) +
+  gfplot::theme_pbs()# + 
+  # theme(legend.position = c(.8,.7))
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/do-DOY-800-kn.png"), width = 3, height = 2.5)
+```
+
 Plot predictions from saved DO predictions file
@@
 p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016.rds")))
-
-plot_map_raster <- function(dat, column = "est") {
-  ggplot(dat, aes_string("X", "Y", fill = column)) +
-    geom_raster() +
-    coord_fixed() +
-    gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
-}
-
+p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")
 plot_map_raster(p_nd, "do_est") +
@@
 ```{r}
-omega_s_all <- plot_map_raster(p_nd, "omega_s") +
+p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")
+
+(omega_s_all <- plot_map_raster(p_nd, "omega_s") +
   ggtitle("Spatial random effects only") +
   scale_fill_gradient2(
-    high = "Steel Blue 4",
-    low = "Red 3", limits = c(-0.65, 1)
+    # high = "Steel Blue 4", low = "Red 3",
+    high = "gold", low = "darkcyan",
+    limits = c(-0.65, 1)
   ) +
   labs(fill = " ") +
-  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82, .75))
-
+  theme(legend.position = c(.2, .3)))
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/omegas-do-rw-800-kn-yg.png"), width = 3, height = 3)
 omega_s_all
@@
 ```{r}
+p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")
+
 p_nd1 <- p_nd %>%
@@
   filter(ssid != 4)
-st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
+
+(st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
   ggtitle("Spatiotemporal random effects") +
-  facet_wrap(~year) +
+  facet_grid(~year) +
   scale_fill_gradient2(
-    high = "Steel Blue 4",
-    low = "Red 3", limits = c(-0.5, 0.5)
+    # high = "Steel Blue 4", low = "Red 3", 
+    high = "gold", low = "darkcyan",
+    limits = c(-0.65, 1)
   ) +
-  labs(fill = " ") +
-  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85, .25))
-st_1n3
+  labs(fill = " ") + theme( legend.position = "none") )
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/do-epsilon-1n3-yg.png"), width = 6, height = 2)
+# st_1n3
 ```
@@
 ```{r}
+p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")
+
 p_nd4 <- p_nd %>% filter(ssid == 4) # %>% filter(year!=2016)
-st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
+(st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
   # ggtitle(" ") +
-  facet_wrap(~year) +
+  facet_grid(~year) +
   scale_fill_gradient2(
-    high = "Steel Blue 4",
-    low = "Red 3", limits = c(-0.5, 0.5)
+    # high = "Steel Blue 4", low = "Red 3", 
+    high = "gold", low = "darkcyan",
+    limits = c(-0.65, 1)
   ) +
-  labs(fill = "epsilon") +
-  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
-st_4
+  labs(fill = "epsilon") + theme( legend.position = "none") )
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/do-epsilon-4-yg.png"), width = 6, height = 2)
+# st_4
 ```
@@
 ```{r}
+p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")
+
 p_nd16 <- p_nd %>% filter(ssid == 16) # %>% filter(ssid!=4)
-st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
+(st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
   # ggtitle(" ") +
-  facet_wrap(~year) +
+  facet_grid(~year) +
   scale_fill_gradient2(
-    high = "Steel Blue 4",
-    low = "Red 3", limits = c(-0.5, 0.5)
+    # high = "Steel Blue 4", low = "Red 3", 
+    high = "gold", low = "darkcyan",
+    limits = c(-0.65, 1)
   ) +
-  xlim(65, 345) +
-  ylim(5800, 6100) +
-  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
-st_16
+  xlim(20,360) + ylim(5850,6050)+ 
+  # xlim(65,345) + ylim(5800,6100)+ 
+    theme(panel.border = element_blank(), legend.position = "none") )
+
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/do-epsilon-16-yg.png"), width = 6, height = 1)
+# st_16
+```
+
+```{r}
+p_nd <- readRDS("~/github/dfo/gfranges/analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")
+
+ggplot(p_nd, aes(roms_temp, temp, colour = depth)) +
+  geom_point(alpha = 0.1, size = 0.5) + 
+  scale_color_viridis_c(option = "B", begin = 0, end = 0.7, direction = -1, trans= sqrt) + 
+  xlim(2,15) +
+  ylim(2,15) +
+  geom_abline(intercept = 0, slope = 1, colour="grey") +
+  ylab("Mean May-August temperature from CTD only model ") +
+  xlab("Mean April-September temperature from ROMS model ") +
+  gfplot::theme_pbs() + 
+    theme(legend.position = "none") 
+ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/roms.png"), width = 4.5, height =4.5)
+
 ```

### 2020-11-14 analysis/VOCC/1-split-biomass-by-maturity.Rmd
plot saved maturity ogives

@@
 params: 
-    species: "Bocaccio"
+    # species: "Bocaccio"
+    # species: "Redbanded Rockfish"
+    species: "Lingcod"
     # species: "Threadfin Sculpin"
@@
 
-#try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
+try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))
 
@@
 ```
+
+From previously saved model
+```{r}
+savedmodel <- readRDS( file = paste0("data/", spp, 
+    "/maturity-ogive-", spp, "-", ssid_string, ".rds"))
+(gfranges::plot_mat_ogive(savedmodel) +
+        scale_colour_viridis_d(option = "C") +
+        ggplot2::ggtitle(paste("Length at maturity for", species)))
+ggsave(here::here("ms", "figs", paste0("maturity-", spp, ".png")), width = 4, height = 5)
+```

### 2020-11-16 analysis/VOCC/3-vocc-dist.Rmd
calculate vocc for new predictions; avoid retention of last species predictions; include biomass plots in html

@@
 ```{r}
+rm(d_all)
+
 if (immature) {
@@
       "data/", spp,
-      "/predictions-", spp, covs, "-1n3n4n16-mature-biomass.rds"
+      "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-new.rds"
     )) %>% filter(year >= start_year )
@@
 
-```{r eval=F}
+```{r }
 gf3t$squashed_velocity <- collapse_outliers(gf3t$velocity, c(0.005, 0.995))
@@
 
-```{r eval=F}
+```{r }
 #raster_max <- quantile(gf3$mean_biomass, 0.995, na.rm = T)
@@
 
-```{r eval=F}
+```{r }
 grad5 <- gfranges::plot_vocc(gf3t,
@@
 write.csv(vocc, file = paste0(
-  "data/_optimized_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_optimized_", age, "_new/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))
@@
 vocc  <- read.csv(paste0(
-  "data/_optimized_", age, "/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
+  "data/_optimized_", age, "_new/", age, "-w-do-nulls-", spp, "-", ssid_string, covs, "-", min(d$year), "-", max(d$year), ".csv"
 ))

### 2020-11-16 analysis/VOCC/2-biomass-depth-only-model.Rmd
save new mature models and predictions

@@
     #species: "Petrale Sole"
-    species: "Canary Rockfish"
+    # species: "Canary Rockfish"
     # # species: "Redstripe Rockfish"
-    # species: "Shortbelly Rockfish"
-    # species: "Lingcod"
+    species: "Shortraker Rockfish"
+    # species: "Bocaccio"
     region: "All synoptic surveys"
-    knots: 500
+    knots: 300
     covariates: "" # +as.factor(ssid)
@@
     #covs: "-trawled-ssid"
-    update_model: TRUE
-    update_predictions: FALSE
-    update_model_check: FALSE
+    update_model: FALSE
+    update_predictions: TRUE
+    update_model_check: TRUE
 ---
@@
       "data/", spp,
-      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-new.rds"
     ))
@@
       "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass.rds"
+      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-new.rds"
     ))
@@
       "data/", spp,
-      "/mod-imm-biomass-", spp, covs, "-", ssid_string, ".rds"
+      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-new.rds"
     ))
@@
       "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass.rds"
+      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-new.rds"
     ))
@@
 } else {
+  try({
   total_biomass <- readRDS(paste0(
     "data/", spp,
-    "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-new.rds"
   ))
@@
     "data/", spp,
-    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass.rds"
+    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-new.rds"
   ))
+  })
 }
@@
     "data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-", ssid_string, ".rds"
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-new.rds"
   ))
@@
     "data/", spp,
-    "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
+    "/model-total-biomass-", spp, covs, "-", ssid_string, "-new.rds"
   ))
@@
   "data/", spp,
-  "/check-mod-predictions-", spp, covs, "-", ssid_string, ".rds"
+  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-new.rds"
 ))
@@
 
-```{r}
-fits <- readRDS("climate-fits.rds")
+```{r eval=FALSE}
+fits <- readRDS("data/climate-fits.rds")
 check_fits <- left_join(depth_only_predictions,fits)
@@
 ### Build time-varying depth plots
-#### Model without environmental covariates
 ```{r }
@@
     "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-new.rds"
   ))
@@
     "data/", spp,
-    "/mod-imm-biomass-", spp, covs, "-1n3n4n16.rds"
+    "/mod-imm-biomass-", spp, covs, "-1n3n4n16-new.rds"
   ))
@@
       "data/", spp,
-      "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
+      "/model-total-biomass-", spp, covs, "-", ssid_string, "-new.rds"
     ))
@@
     "figs/", spp,
-    "/depth-", spp, covs, "-no-AR1.png"
+    "/depth-", spp, covs, "-nov2020.png"
   ),
@@
 ```{r}
-write_grads(signif(max(depth_model_list$adult$gradients), digits = 8), paste0(spp, "-mat"))
-write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0(spp, "-imm-redo"))
-write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0(spp, "-total-redo"))
+write_grads(signif(max(depth_model_list$adult$gradients), digits = 8), paste0(spp, "-mat-new"))
+# write_grads(signif(max(depth_model_list$imm$gradients), digits = 8), paste0(spp, "-imm-redo"))
+# write_grads(signif(max(depth_model_list$total$gradients), digits = 8), paste0(spp, "-total-redo"))
 ```
 
+to manually improve optimization run all of next 4 chunks
 ```{r eval = FALSE}
@@
       file = paste0("data/", spp, "/mod-imm-biomass-",
-        spp, covs, "-", ssid_string, ".rds"))
+        spp, covs, "-", ssid_string, "-new.rds"))
 

### 2020-11-16 analysis/VOCC/2-render-biomass-model.R

@@ -144,60 +144,6 @@ for (r_h in seq_along(list_regions)) {
 }
 
 
-
-getwd()
-setwd(here::here("/analysis/VOCC"))
-env <- new.env() # parent = baseenv()
-
-# all species
-list_species <- c(
-  "Big Skate",
-  "Longnose Skate",
-  "Spotted Ratfish",
-  "North Pacific Spiny Dogfish",
-  # "Pacific Tomcod",
-  "Walleye Pollock",
-  "Pacific Cod",
-  "Sablefish",
-  "Lingcod",
-  # "Pacific Hake",
-  # "Rosethorn Rockfish",
-  "Yellowmouth Rockfish",
-  # "Harlequin Rockfish",
-  "Canary Rockfish", # schooling, winter-birthing
-  # "Copper Rockfish", # small sample
-  "Darkblotched Rockfish",
-  "Greenstriped Rockfish",
-  "Pacific Ocean Perch", # schooling
-  "Redbanded Rockfish",
-  "Sharpchin Rockfish",
-  "Shortbelly Rockfish", # small sample
-  "Silvergray Rockfish",
-  "Splitnose Rockfish",
-  "Yellowtail Rockfish", # schooling
-  # "Longspine Thornyhead",
-  "Shortspine Thornyhead",
-  "Arrowtooth Flounder",
-  "Rex Sole",
-  "Petrale Sole",
-  "English Sole",
-  "Dover Sole",
-  "Southern Rock Sole",
-  "Flathead Sole",
-  "Curlfin Sole",
-  # "Sand Sole",
-  # "Slender Sole",
-  # "Pacific Sanddab",
-  "Pacific Halibut",
-  "Redstripe Rockfish",
-  "Rougheye/Blackspotted Rockfish Complex",
-  "Widow Rockfish",
-  "Quillback Rockfish",
-  "Bocaccio",
-  "Shortraker Rockfish",
-  "Yelloweye Rockfish"
-)
-
 ### check biomass model convergence 
 list_regions <- c("All synoptic surveys")
 # dir.create(file.path("html/biomass-by-depth"))
@@ -421,3 +367,94 @@ totdata <- do.call(rbind, checktdata)
 
 m1 <- readRDS(paste0("data/canary-rockfish/mod-mat-biomass-canary-rockfish-tv-depth-only-1n3n4n16-prior-FALSE.rds"))
 View(m1$data)
+
+
+
+getwd()
[…]

### 2020-11-16 analysis/VOCC/3-render-biotic-vocc.R

@@ -51,6 +51,59 @@ list_species <- c(
   "Yelloweye Rockfish"
 )
 
+### Nov 2020 run:
+# with new sdmTMB models for mature
+# a few added species
+# full data set for few spp with missing data
+list_species <- c(
+  "Big Skate",
+  "Longnose Skate",
+  "Spotted Ratfish",
+  "North Pacific Spiny Dogfish",
+  # "Pacific Tomcod",
+  "Walleye Pollock",
+  "Pacific Cod",
+  "Sablefish",
+  "Lingcod", # already rerun
+  "Pacific Hake", # previously excluded
+  "Rosethorn Rockfish", # previously excluded
+  "Yellowmouth Rockfish",
+  # "Harlequin Rockfish", # too small sample
+  "Canary Rockfish", # already rerun
+  # "Copper Rockfish", # too small sample
+  "Darkblotched Rockfish",
+  "Greenstriped Rockfish",
+  "Pacific Ocean Perch", # schooling
+  "Redbanded Rockfish", # already rerun
+  "Sharpchin Rockfish",
+  #"Shortbelly Rockfish", # too small sample
+  "Silvergray Rockfish",
+  "Splitnose Rockfish",
+  "Yellowtail Rockfish", # schooling
+  # "Longspine Thornyhead", # too small sample
+  "Shortspine Thornyhead",
+  "Arrowtooth Flounder",
+  "Rex Sole",
+  "Petrale Sole",
+  "English Sole", # already rerun
+  "Dover Sole",
+  "Southern Rock Sole",
+  "Flathead Sole",
+  "Curlfin Sole",
+  # "Sand Sole",# too small sample
+  "Slender Sole",# previously excluded
+  "Pacific Sanddab",# previously excluded
+  "Pacific Halibut",
+  "Redstripe Rockfish",
+  "Rougheye/Blackspotted Rockfish Complex",
+  "Widow Rockfish",
+  "Quillback Rockfish",
+  "Bocaccio", # rerun with fewer knots
+  "Shortraker Rockfish",# small sample so run with fewer knots
+  "Yelloweye Rockfish"
+)
+
+
 ### build biotic gradients
 list_regions <- c(
   "West Coast Haida Gwaii",
@@ -76,7 +129,7 @@ for (r_h in seq_along(list_regions)) {
         ),
         output_file = paste0(
           "html/VOCC-plots/vocc-w-do-", spp,
[…]

### 2020-11-16 analysis/VOCC/4-pool-biotic-vocc.R

@@ -7,7 +7,7 @@ library("dplyr")
 # mydir = paste0( "_newclim2_mature")
 # 800 kn for both
 # mydir = paste0( "_more2016_mature")
-mydir = paste0( "_optimized_mature")
+mydir = paste0( "_optimized_mature_new")
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
 myfiles
 biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% select(-X, -start_year)

### 2020-11-18 analysis/VOCC/2-biomass-depth-only-model2.Rmd
fix maturity split and models for species with data issues/errors

--- 2020-11-14 analysis/VOCC/2-biomass-depth-only-model.Rmd
@@
     region: "All synoptic surveys"
-    knots: 500
+    knots: 400
     covariates: "" # +as.factor(ssid)
@@
 
+# if(spp=="redbanded-rockfish"){ 
+# 1 redbanded fish found at too shallow a depth... probably leftover in net?
+# slender sole also has an outlier here, so maybe should throw out this event for everyone?
+# could mean there's a location or depth error  
+biomass <- biomass[biomass$fishing_event_id!=2536031,]
+# }
+
+if(spp=="pacific-ocean-perch"){
+# fish sizes are impossible given recorded catch weight
+biomass <- biomass[biomass$fishing_event_id!=1506954,]
+}
+
 if(nrow(biomass)<4000) {stop("Need to recalculate split by maturity!")}
@@
 ```{r}
-if (region == "Both odd year surveys") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
-}
-
-if (region == "West Coast Vancouver Island") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-}
-
-if (region == "West Coast Haida Gwaii") {
-  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
-}
 
@@
   if (any(names(data) == "adult_density")) {
-    adult_formula <- as.formula(paste(
-      "adult_density ~ 0 + as.factor(year)", covariates, ""
+    # adult_formula <- as.formula(paste(
+    #   "adult_density ~ 0 + as.factor(year)", covariates, ""
+    # ))
+    # 
+    # adult_biomass <- sdmTMB::sdmTMB(
+    #   data = data,
+    #   adult_formula,
+    #   time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
+    #   time = "year", spde = spde,
+    #   family = tweedie(link = "log"),
+    #   #ar1_fields = TRUE,
+    #   include_spatial = TRUE,
+    #   # reml = TRUE,
+    #   nlminb_loops = 2,
+    #   newton_steps = 2,
+    #   enable_priors = priors,
+    #   control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+    #   silent = FALSE
+    # )
+    # 
+    # saveRDS(adult_biomass,
+    #   file = paste0(
+    #     "data/", spp,
+    #     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-new.rds"
+    #   )
+    # )
+
+   try({ imm_formula <- as.formula(paste(
+      "imm_density ~ 0 + as.factor(year)", covariates, ""
     ))
 
-    adult_biomass <- sdmTMB::sdmTMB(
+    imm_biomass <- sdmTMB::sdmTMB(
       data = data,
-      adult_formula,
+      imm_formula,
       time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
@@
       family = tweedie(link = "log"),
-      #ar1_fields = TRUE,
+      # ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
+      # reml = TRUE,
       include_spatial = TRUE,
-      # reml = TRUE,
       nlminb_loops = 2,
@@
 
-    saveRDS(adult_biomass,
-      file = paste0(
-        "data/", spp,
-        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-new.rds"
-      )
+    saveRDS(imm_biomass,
+      file = paste0("data/", spp, "/mod-imm-biomass-",
+        spp, covs, "-", ssid_string, ".rds"))
+   })
+
+  } else {
+    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
+
+    total_biomass <- sdmTMB::sdmTMB(
+      data = data,
+      dens_formula,
+      time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
+      time = "year", spde = spde,
+      family = tweedie(link = "log"),
+      # ar1_fields = TRUE, # all off for final version
+      include_spatial = TRUE,
+      nlminb_loops = 2, 
+      newton_steps = 2,
+      enable_priors = priors,
+      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
+      silent = FALSE
     )
 
-   # try({ imm_formula <- as.formula(paste(
-   #    "imm_density ~ 0 + as.factor(year)", covariates, ""
-   #  ))
-   # 
-   #  imm_biomass <- sdmTMB::sdmTMB(
-   #    data = data,
-   #    imm_formula,
-   #    time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
-   #    time = "year", spde = spde,
-   #    family = tweedie(link = "log"),
-   #    # ar1_fields = TRUE, # all off for final version
-   #    # reml = TRUE,
-   #    include_spatial = TRUE,
-   #    nlminb_loops = 2, 
-   #    newton_steps = 2,
-   #    enable_priors = priors,
-   #    control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-   #    silent = FALSE
-   #  )
-   # 
-   #  saveRDS(imm_biomass,
-   #    file = paste0("data/", spp, "/mod-imm-biomass-", 
-   #      spp, covs, "-", ssid_string, "-new.rds"))
-   # })
-
-  } else {
-    # dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))
-    # 
-    # total_biomass <- sdmTMB::sdmTMB(
-    #   data = data,
-    #   dens_formula,
-    #   time_varying = ~ 0 + depth_scaled + depth_scaled2, # + trawled
-    #   time = "year", spde = spde,
-    #   family = tweedie(link = "log"),
-    #   # ar1_fields = TRUE, # all off for final version
-    #   include_spatial = TRUE,
-    #   nlminb_loops = 2, 
-    #   newton_steps = 2,
-    #   enable_priors = priors,
-    #   control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
-    #   silent = FALSE
-    # )
-    # 
-    # saveRDS(total_biomass, file = paste0(
-    #   "data/", spp,
-    #   "/model-total-biomass-", spp, covs, "-", ssid_string, "-new.rds"
-    # ))
+    saveRDS(total_biomass, file = paste0(
+      "data/", spp,
+      "/model-total-biomass-", spp, covs, "-", ssid_string, ".rds"
+    ))
   }
@@
 
-  try({
-    adult_biomass <- readRDS(paste0(
-      "data/", spp,
-      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-    ))
-
-    nd <- nd_all %>%
-      filter(ssid %in% model_ssid) %>%
-      filter(year %in% unique(adult_biomass$data$year))
-    nd <- na.omit(nd)
-    nd$year <- as.integer(nd$year)
-
-    adult_predictions <- predict(adult_biomass, newdata = nd)
-    saveRDS(adult_predictions, file = paste0(
-      "data/", spp,
-      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass.rds"
-    ))
-  })
+  # try({
+  #   adult_biomass <- readRDS(paste0(
+  #     "data/", spp,
+  #     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+  #   ))
+  # 
+  #   nd <- nd_all %>%
+  #     filter(ssid %in% model_ssid) %>%
+  #     filter(year %in% unique(adult_biomass$data$year))
+  #   nd <- na.omit(nd)
+  #   nd$year <- as.integer(nd$year)
+  # 
+  # 
+  #   adult_predictions <- predict(adult_biomass, newdata = nd)
+  #   saveRDS(adult_predictions, file = paste0(
+  #     "data/", spp,
+  #     "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass.rds"
+  #   ))
+  # })
 
@@
 
-```{r}
-fits <- readRDS("climate-fits.rds")
+```{r eval=F}
+fits <- readRDS("data/climate-fits.rds")
 check_fits <- left_join(depth_only_predictions,fits)
@@
 
-try({
-  adult_biomass <- readRDS(paste0(
-    "data/", spp,
-    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-  ))
-})
+# try({
+#   adult_biomass <- readRDS(paste0(
+#     "data/", spp,
+#     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+#   ))
+# })
 # adult_biomass<-sdmTMB:::update_model(adult_biomass)
@@
 
-    saveRDS(adult_biomass,
-      file = paste0(
-        "data/", spp,
-        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
-      )
-    )
+    # saveRDS(adult_biomass,
+    #   file = paste0(
+    #     "data/", spp,
+    #     "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
+    #   )
+    # )
 

### 2020-11-18 analysis/VOCC/1-render-maturity.R

@@ -41,13 +41,16 @@ list_species <- c(
   "Southern Rock Sole",
   "Flathead Sole",
   "Curlfin Sole",
-  "Sand Sole"
-  # "Slender Sole",
-  # "Pacific Sanddab"
+  # "Sand Sole"
+  "Slender Sole",
+  "Pacific Sanddab"
 )
 
 
 # # SPECIES to try setting maturity threshold to 0.05
+# # Bocaccio reverted to 0.5 in June 2020
+# # Shortraker not split due to tiny # of imm
+# # Others seem to be left at 0.5?
 # list_species <- c(
 #   "Widow Rockfish",
 #   "Quillback Rockfish",
@@ -112,6 +115,11 @@ list_species <- c("Pacific Herring")
 
 ## NEEDS SPECIAL SETTINGS
 list_species <- c("North Pacific Spiny Dogfish","Longspine Thornyhead","Sand Sole")
+
+list_species <- c(
+  "Pacific Ocean Perch"
+)
+
 
 list_regions <- c("All synoptic surveys")
 # list_regions <- c(

### 2020-11-18 analysis/VOCC/2-render-biomass-model.R

@@ -364,9 +364,9 @@ for (spp_i in seq_along(list_species)) {
 
 totdata <- do.call(rbind, checktdata)
 
-
-m1 <- readRDS(paste0("data/canary-rockfish/mod-mat-biomass-canary-rockfish-tv-depth-only-1n3n4n16-prior-FALSE.rds"))
-View(m1$data)
+# 
+# m1 <- readRDS(paste0("data/canary-rockfish/mod-mat-biomass-canary-rockfish-tv-depth-only-1n3n4n16-prior-FALSE.rds"))
+# View(m1$data)
 
 
 
@@ -429,6 +429,20 @@ list_species <- c(
 # "Shortraker Rockfish", # small sample so run with fewer knots
 # "Bocaccio" # rerun with fewer knots
 )
+
+# rerun excluding fishing event where depth seems be wrong
+list_species <- c(
+  # "Redbanded Rockfish", # problem with tv depth estimate due to shallow fish left in net
+  # "Dover Sole", # deep fish found shallow... 
+  # "Slender Sole" # deep fish found shallow... 
+  "Longnose Skate" # deep fish found shallow... 
+)
+
+# rerun excluding maturity data for fish that don't total to appropriate weight
+list_species <- c(
+  "Pacific Ocean Perch"
+)
+
 ### rerun all mat models with new sdmTMB version (Nov 14 2020)
 ### also replaces data missing from 9 of the Dec 28 2019 runs 
 list_regions <- c("All synoptic surveys")
@@ -444,7 +458,7 @@ for (r_h in seq_along(list_regions)) {
           region = list_regions[r_h],
           covariates = "", # additional non-climate variables
           covs = covs,
-          knots = 300,
+          knots = 500,
           update_model = TRUE,
           update_predictions = TRUE, 
           update_model_check = TRUE 
@@ -458,3 +472,53 @@ for (r_h in seq_along(list_regions)) {
     })
   }
 }
+
+
+
+# add new imm but label with Jul ending because not rerunning everything
+list_species <- c(
+  # "Rosethorn Rockfish", # previously excluded
+  "Slender Sole",# previously excluded
+  "Pacific Sanddab"# previously excluded
+)
+
+# rerun excluding fishing event where depth seems be wrong
+list_species <- c(
+  "Redbanded Rockfish", # problem with tv depth estimate due to possible fish left in net
+  "Dover Sole", # deep fish found shallow...
+  "Slender Sole" # deep fish found shallow...
+)
+
[…]

### 2020-11-18 analysis/VOCC/3-render-biotic-vocc.R

@@ -3,54 +3,6 @@ getwd()
 setwd(here::here("/analysis/VOCC"))
 env <- new.env() # parent = baseenv()
 
-list_species <- c(
-  "Big Skate",
-  "Longnose Skate",
-  "Spotted Ratfish",
-  "North Pacific Spiny Dogfish",
-  # "Pacific Tomcod",
-  "Walleye Pollock",
-  "Pacific Cod",
-  "Sablefish",
-  "Lingcod",
-  # "Pacific Hake",
-  # "Rosethorn Rockfish",
-  "Yellowmouth Rockfish",
-  # "Harlequin Rockfish",
-  "Canary Rockfish", # schooling, winter-birthing
-  # "Copper Rockfish", # small sample
-  "Darkblotched Rockfish",
-  "Greenstriped Rockfish",
-  "Pacific Ocean Perch", # schooling
-  "Redbanded Rockfish",
-  "Sharpchin Rockfish",
-  "Shortbelly Rockfish", # small sample
-  "Silvergray Rockfish",
-  "Splitnose Rockfish",
-  "Yellowtail Rockfish", # schooling
-  "Longspine Thornyhead",
-  "Shortspine Thornyhead",
-  "Arrowtooth Flounder",
-  "Rex Sole",
-  "Petrale Sole",
-  "English Sole",
-  "Dover Sole",
-  "Southern Rock Sole",
-  "Flathead Sole",
-  "Curlfin Sole",
-  # "Sand Sole",
-  # "Slender Sole",
-  # "Pacific Sanddab",
-  "Pacific Halibut",
-  "Redstripe Rockfish",
-  "Rougheye/Blackspotted Rockfish Complex",
-  "Widow Rockfish",
-  "Quillback Rockfish",
-  "Bocaccio",
-  "Shortraker Rockfish",
-  "Yelloweye Rockfish"
-)
-
 ### Nov 2020 run:
 # with new sdmTMB models for mature
 # a few added species
@@ -103,6 +55,17 @@ list_species <- c(
   "Yelloweye Rockfish"
 )
 
+# 
+# list_species <- c(
+#   # rerun excluding fishing event where depth seems be wrong
+#   "Longnose Skate",
+#   "Redbanded Rockfish", # problem with tv depth estimate due to possible fish left in net
+#   "Dover Sole", # deep fish found shallow...
+#   "Slender Sole", # deep fish found shallow...
[…]

### 2020-11-18 analysis/VOCC/4-pool-biotic-vocc.R

@@ -20,7 +20,7 @@ saveRDS(biotic, file = paste0(
 
 #### IMMATURE #####
 # mydir = paste0( "_more2016_immature")
-mydir = paste0( "_optimized_immature")
+mydir = paste0( "_optimized_immature_new")
 myfiles <- list.files(path = mydir, pattern = "*.csv", full.names = TRUE)
 myfiles
 biotic <- do.call(rbind, lapply(myfiles, read.csv)) %>% 

### 2020-11-18 analysis/VOCC/5-add-null.R

@@ -28,6 +28,8 @@ d <- rbind(d1, d2)
 d <- na.omit(d) %>% as_tibble() %>% mutate(species_age = paste(age, species))
 
 # remove models that did not converge
+
+d <- filter(d, species_age != "immature Bocaccio") 
 d <- filter(d, species_age != "immature Shortraker Rockfish") 
 d <- filter(d, species_age != "immature Curlfin Sole")
 

### 2020-11-25 analysis/SOPO/2-trawl-no-covs-model.Rmd
minor housekeeping

@@
 ---
-title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
+title: "Spatiotemporal models of `r paste(params$species)` biomass with AR1"
 author: "Philina English"

### 2020-11-25 analysis/VOCC/0-packages.R

@@ -1,6 +1,12 @@
+if (!require(tidyverse)) install.packages("tidyverse")
 if (!require(patchwork)) install.packages("patchwork")
 if (!require(ggpubr)) install.packages("ggpubr")
 if (!require(grid)) install.packages("grid")
 
 
 if (!require(interactions)) install.packages("interactions")
+
+
+
+
+

### 2020-11-25 analysis/VOCC/5-add-null.R

@@ -37,7 +37,7 @@ all_species <- unique(d$species_age)
 
 
 ### ADD SPECIES TRAITS 
-stats <- readRDS(paste0("data/life-history-stats.rds"))
+stats <- readRDS(paste0("data/life-history-stats5.rds"))
 stats$rockfish <- if_else(stats$group == "ROCKFISH", "ROCKFISH", "OTHER")
 
 stats <- stats %>% separate(species_science_name, " ", into = c("genus","specific"))
@@ -49,6 +49,7 @@ stats$group[stats$group == "HAKE"] <- "COD"
 
 # d <- na.omit(d) %>% as_tibble()
 d <- suppressWarnings(left_join(d, stats, by = "species")) 
+
 select(d, genus, species) %>%
   distinct() %>%
   arrange(genus, species) %>%

### 2020-11-25 analysis/VOCC/5-life-history-stats.R

@@ -1,5 +1,5 @@
 # library(TMB)
-library(dplyr)
+library(tidyverse)
 library(ggplot2)
 
 setwd(here::here("analysis", "VOCC"))

### 2020-11-25 analysis/tmb-sensor-explore/extract_ROMS.R

@@ -19,9 +19,8 @@ for(i in 2003:2018){
 #     crs = 4326 , 
 #     remove = FALSE)
 
-
-events <- m_temp_rw$data %>% mutate(X_old=X*1000, Y_old=Y*1000) %>% select(-X, -Y)
-
+m_temp_rw <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-sstdata.rds")
+events <- m_temp_rw$data %>% mutate(X_old=X*1000, Y_old=Y*1000) %>% dplyr::select(-X, -Y)
 
 xx <- events %>%
   sf::st_as_sf(coords = c("longitude", "latitude"),
@@ -140,7 +139,7 @@ d$Y <- 2 * round(d$Y/2)
 data <- d %>% group_by(X, Y) %>% summarise_all(mean) 
 
 
-data <- data %>% select(-x,-y) %>% 
+data <- data %>% dplyr::select(-x,-y) %>% 
   pivot_longer(T2003:T2018, names_to = "year", values_to = "roms_temp") %>% mutate(
   year = as.numeric(gsub("T", "", year)) 
 ) 
@@ -148,9 +147,16 @@ data <- data %>% select(-x,-y) %>%
 
 # d <- st_set_geometry(d, NULL) 
 
-pred <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016.rds")))
+# pred <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016.rds")))
+# both <- left_join(pred, data)
+# glimpse(both)
+# saveRDS(both, file = here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds"))
 
-both <- left_join(pred, data)
-glimpse(both)
-
-saveRDS(both, file = here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds"))
+# pred <- readRDS((here::here("analysis/VOCC/data/predicted_temp_allyears_800kn.rds")))
+# both <- left_join(pred, data) %>% mutate(
+#   temp = est,
+#   temp_omega = omega_s,
+#   temp_epsilon = epsilon_st
+# ) %>% dplyr::select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st, -temperature_c)
+# glimpse(both)
[…]

### 2020-11-25 analysis/sand-lance/biomass-density-model.Rmd
draft sand lance models

---
title: "Spatiotemporal model of `r paste(params$species)` biomass"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Pacific Sand Lance"
    region: "Non-HG surveys"
    knots: 400
    # covs: "-sand-depth-roms"
    # covs: "-depth-roms"
    # covs: "-sand-depth-temp"
    covs: "-mud-depth-temp"
    REML: FALSE
    AR1: TRUE
    fixed_spatial: TRUE
    update_model: TRUE
    # update_model: FALSE
    update_model_check: TRUE
    update_predictions: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  # fig.path=paste0("figs/", spp, "/"),
  echo = FALSE, warning = FALSE, message = FALSE
)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

theme_set(
    gfplot::theme_pbs(base_size = 14) 
)
```

```{r formula}
if(params$covs == "-sand-depth-roms"){
mod.formula <- as.formula(paste0(
  "density ~ 0 + as.factor(year_pair) + 
  s(sandy_scaled, k = 3) + 
  s(depth_scaled, k = 3) + 
  s(roms_scaled, k = 3)"
  ))
clim_cov <- "roms"
}

if(params$covs == "-depth-roms"){
mod.formula <- as.formula(paste0(
  "density ~ 0 + as.factor(year_pair) + 
  s(depth_scaled, k = 3) + 
  s(roms_scaled, k = 3)"
  ))
clim_cov <- "roms"
}

if(params$covs == "-depth-roms"){
mod.formula <- as.formula(paste0(
  "density ~ 0 + as.factor(year_pair) + 
  s(depth_scaled, k = 3) + 
  s(roms_scaled, k = 3)"
  ))
clim_cov <- "roms"
}

if(params$covs == "-sand-depth-temp"){
mod.formula <- as.formula(paste0(
  "density ~ 0 + as.factor(year_pair) + 
  s(sandy_scaled, k = 3) + 
  s(depth_scaled, k = 3) + 
  s(temp_scaled, k = 3)"
  ))
clim_cov <- "temp"
}

if(params$covs == "-mud-depth-temp"){
mod.formula <- as.formula(paste0(
  "density ~ 0 + as.factor(year_pair) + 
  s(muddy_scaled, k = 3) + 
  s(depth_scaled, k = 3) + 
  s(temp_scaled, k = 3)"
  ))
clim_cov <- "temp"
}

```

```{r params}
species <- params$species
region <- params$region
covs <- params$covs
knots <- params$knots
REML <- params$REML
AR1 <- params$AR1

paste("region =", region)
paste("climate covariate =", clim_cov)
paste("model label =", covs)
paste("knots =", knots)
paste("REML =", REML)
paste("AR1 =", AR1)
```

```{r, message=FALSE, warning=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "Non-HG surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI")
  model_ssid <- c(1, 3, 4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

# Build spatiotemporal density model
Load and filter data 
```{r}
# # if on dfo network trawl data can be retrieved
# events <- gfdata::get_survey_sets(species, ssid = c(1, 3, 4, 16))
# saveRDS(events, file = paste0("data/event-data-", spp, "")) 
# biomass <- readRDS(paste0("data/event-data-", spp, ""))

# on PE's system
biomass <- readRDS(paste0("../VOCC/raw/event-data-", spp, ""))

# trace samples aren't always recorded in 'density_kgpm2' variable
# but are recorded in 'catch_count'
# likewise larger catches are recorded as 0 in 'catch_count'
biomass <- biomass %>% mutate(
  present = if_else(density_kgpm2 > 0, 1, if_else(catch_count > 0, 1, 0)),
  density = (if_else(density_kgpm2 > 0, density_kgpm2, 
    # use roughly the min recorded density for these trace samples
    if_else(catch_count > 0, 1e-07, 0))) * 10000  # convert kgpm2 to kg per ha
  ) %>% select(fishing_event_id, year, present, density) %>% 
  # don't currently have covariate layers for 2019
  filter(year < 2019)

# get substrate data
covars <- readRDS("data/event-covariates.rds") %>% select(-geometry)
data <- dplyr::left_join(biomass, covars)

# get ctd sensor data
sensor <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/data/all-sensor-data-processed.rds") %>% select(-X, -Y, -lat, -lon)
data <- left_join(data, sensor) %>% 
  # remove depth data to bring in complete set later
  select(-exclude, -akima_depth, -depth_edit, -log_depth, -depth_bath)

# get ROMS mean temperature for April through September each year
roms <- readRDS("~/github/dfo/gfranges/analysis/tmb-sensor-explore/data/roms_temp_by_events2.rds") %>% select(fishing_event_id, roms)
data <- left_join(data, roms)

# get interpolated bathymetry data for comparison with sensor recorded depths
bath <- readRDS("data/bathymetry-data")
bath <- bath$data %>% select(X,Y, ssid, fishing_event_id, depth_bath = depth)

d <- left_join(data, bath) %>% mutate(
  depth_sam = depth,# rename sensor depth
  depth = coalesce(depth, depth_bath),# use bath depth if sensor depth missing
  raw_depth = depth, # create 'depth' var that's actually log depth for scaling
  log_depth = log(depth), depth = log(raw_depth)) %>% 
  rename(temp = temperature_c, do = do_mlpl)

# scale predictors before filtering to ensure mean and SD are global
d <- gfranges::scale_predictors(d, 
  predictors = c(quo(depth), quo(sandy), quo(muddy), quo(temp), quo(do), quo(roms))
)

# # rename raw_depth for plots 
# d <- d %>% mutate(depth = raw_depth) %>%
#   filter(ssid %in% model_ssid)

### trim for the climate variable of interest
if (clim_cov %in%  "temp"){ d <- d[!is.na(d$temp), ]}
if (clim_cov %in% "roms"){ d <- d[!is.na(d$roms), ]}
if (clim_cov %in% "do"){ d <- d[!is.na(d$do), ]}

data <- d

# merge survey pairs such that even year surveys are modelled as though they occured in the previous odd year
data <- data %>% 
  filter(year > 2003) %>% 
  # trick to round to odd years
  # 0.8 is required to avoid 5's always rounding to even in R
  mutate(year_pair = (2 * round((year+0.8)/2))-1) %>% 
  filter(year < 2019)
```

Plots illustrating year pairs and location of samples where species was caught
```{r}
ggplot() + geom_point(data = filter(data, present == 0), 
    aes(X, Y, colour = sandy), # color absent points with proportion sandy
    shape= 1, alpha=0.5, size=2) +
  geom_point(data = filter(data, present == 1), aes(X, Y, size = density), 
    shape = 4) +
  scale_color_viridis_c(trans=fourth_root_power, option = "A", begin = 0.4) +
  facet_wrap(~year)

ggplot() + geom_point(data = filter(data, present == 0), 
    aes(X, Y, colour = sandy), 
    shape= 1, alpha=0.5, size=2) +
    geom_point(data = filter(data, present == 1), aes(X, Y, size = density), 
    shape = 4) +
  scale_color_viridis_c(trans=fourth_root_power, option = "A", begin = 0.4) +
  facet_wrap(~year_pair)
```

All years plotted at once
```{r}
ggplot() + geom_point(data = filter(data, present == 0), 
    aes(X, Y, colour = sandy), 
    shape= 1, alpha=0.65, size=2) +
  geom_point(data = filter(data, present == 1), aes(X, Y, size = density), 
    shape = 4) +
  scale_color_viridis_c(option = "A", trans=fourth_root_power, begin = 0.4) +
  geom_abline(intercept = 0, slope = 1)
```

```{r eval= FALSE}
# depth diff = difference between bath and sample depths
# depth range is difference between start and end sample depths
ggplot(data, aes(depth_sam, depth_bath, colour = depth_range)) + 
  geom_point(alpha=0.75) +
  scale_color_viridis_c(trans=fourth_root_power, direction = -1) +
  geom_abline(intercept = 0, slope = 1)

# # current data set includes only sampled depths with no bathymetry based depths
# ggplot(data, aes(raw_depth, depth_bath, colour = depth_range)) + 
#   geom_point(alpha=0.75) +
#   scale_color_viridis_c(trans=fourth_root_power, direction = -1) +
#   geom_abline(intercept = 0, slope = 1)

ggplot(data, aes(depth_sam, depth_range, colour = as.factor(ssid))) + 
  geom_point(alpha=0.75) +
  scale_color_viridis_d() 
```

Check ROMS vs. CTD temperature measurements
```{r eval= FALSE}
# differences between bath and sample depths doesn't seem to explain differences between sample values and roms
# these roms values are an average of April through September vs. single CTD measurements 
ggplot(data, aes(roms, temp, colour = depth_diff)) + 
  geom_point(alpha=0.75, size=0.5) +
  scale_color_viridis_c(trans=fourth_root_power, direction = -1) +
  geom_abline(intercept = 0, slope = 1) +
  xlab("ROMS mean (April - September)") +
  ylab("ctd temperature")

# differences between start and end depths also don't seem to explain differences between sample values and roms
ggplot(data, aes(roms, temp, colour = depth_range)) + 
  geom_point(alpha=0.75, size=0.5)  +
  scale_color_viridis_c(trans=fourth_root_power, direction = -1) +
  geom_abline(intercept = 0, slope = 1) +
  xlab("ROMS mean (April - September)") +
  ylab("ctd temperature")
```

Make mesh
```{r}
data <- data %>% filter(ssid %in% model_ssid)

if (region == "Both odd year surveys") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 250)
}

if (region == "West Coast Vancouver Island") {
  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
}

if (region == "All synoptic surveys") {
  spde <- sdmTMB::make_mesh(data, c("X", "Y"), n_knots = knots)
}

if (region == "Non-HG surveys") {
  spde <- sdmTMB::make_mesh(data, c("X", "Y"), n_knots = knots)
}
plot(spde)
```

Run sdmTMB model
```{r }
if (params$update_model) {

    total_biomass <- sdmTMB::sdmTMB(
      data = data, mod.formula,
      # time_varying = ~ 0 + depth_scaled + depth_scaled2, 
      time = "year_pair", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = params$AR1,
      reml = params$REML,
      include_spatial = params$fixed_spatial
    )

    saveRDS(total_biomass, file = paste0(
      "models/", spp, "/mod-tot-biomass-", spp, 
      covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
    ))
}
```

### Check residuals
```{r}
if(params$update_model_check) {

m <- readRDS(paste0(
  "models/", spp, "/mod-tot-biomass-", spp, 
  covs, "-", ssid_string, "-ar1-", params$AR1, 
  "-reml-", params$REML, params$knots, ".rds"
  ))

point_predictions <- predict(m)
point_predictions$residuals <- residuals(m)

saveRDS(point_predictions, file = paste0(
  "data/", spp, "/check-mod-predictions-", spp, 
  covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", 
  params$REML, params$knots, ".rds"
))
}
```

```{r}
pred <- readRDS(paste0(
  "data/", spp, "/check-mod-predictions-", spp, 
  covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
))

qqnorm(pred$residuals);abline(a = 0, b = 1)
```

```{r}
g <- ggplot(pred, 
    aes(density, residuals, colour = as.factor(ssid))) +
  geom_point() + scale_x_continuous(trans = "log10") +  
  facet_wrap(~year_pair)
g
```

```{r}
g1 <- ggplot(pred,
      aes(density, exp(est), colour = as.factor(ssid))) +
    geom_abline(slope=1, intercept = 0) +
    scale_x_continuous(trans = 'log10') +
    scale_y_continuous(trans = 'log10') +
    geom_point(alpha = 0.2) 
g1
```

```{r}
g2 <- ggplot(pred, 
      aes((est), residuals, colour = density)) +
    geom_point(alpha = 0.2) + # scale_x_continuous(trans = 'log10') +
    facet_wrap(~year)
g2
```

```{r}
# ggplot(pred, aes(X, Y, colour = (residuals))) +
#   geom_point() +
#   scale_colour_gradient2() +
#   scale_x_continuous(trans = "log10") +
#   facet_wrap(~year)

ggplot(pred, aes(X, Y, colour = residuals)) +
  geom_point(size=0.5) +
  scale_colour_gradient2() +
  facet_wrap(~year_pair)
```

### Model summary
```{r }
if(!exists("m")){
m <- readRDS(paste0(
    "models/", spp, "/mod-tot-biomass-", spp, 
      covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
  ))
}
m
max(m$gradients)
```

### Effect plots
```{r}
if(params$update_model_check){

nd_depth <- data.frame(
  depth_scaled = seq(min(data$depth_scaled), 
    max(data$depth_scaled), length.out = 30), 
  sandy_scaled = 0,
  muddy_scaled = 0,
  temp_scaled = 0,
  roms_scaled = 0,
  year_pair = 2011 # a chosen year
)

p_depth <- predict(m, newdata = nd_depth, se_fit = TRUE, re_form = NA)
p_depth$depth <- exp((p_depth$depth_scaled*data$depth_sd[1])+ data$depth_mean[1])

(g_depth <- ggplot(p_depth, aes(depth, exp(est), 
  ymin = exp(est - 1.96 * est_se), 
  ymax = exp(est + 1.96 * est_se))) +
  coord_cartesian(ylim = c(0, max(exp(p_depth$est))*20), xlim = c(0, 150)) +
  geom_line() + geom_ribbon(alpha = 0.4))

ggsave(paste0("figs/depth-", spp, covs, ".png"), width = 5, height = 5)
}
try({g_depth})
```

```{r}
if(clim_cov == "roms"){
nd_roms <- data.frame(
  depth_scaled = min(data$depth_scaled)-0.2, 
  sandy_scaled = 0,
  muddy_scaled = 0,
  roms_scaled = seq(min(data$roms_scaled), 
    max(data$roms_scaled), length.out = 50),
  year_pair = 2011 # a chosen year
)
p_roms <- predict(m, newdata = nd_roms, se_fit = TRUE, re_form = NA)
p_roms$roms <- (p_roms$roms_scaled*data$roms_sd[1])+ data$roms_mean[1]

(g_roms <- ggplot(p_roms, aes(roms, exp(est), 
  ymin = exp(est - 1.96 * est_se), 
  ymax = exp(est + 1.96 * est_se))) +
  coord_cartesian(ylim = c(0, max(exp(p_roms$est))*20)) +
  geom_line() + geom_ribbon(alpha = 0.4))

ggsave(paste0("figs/roms-", spp, covs, ".png"), width = 5, height = 5)
}

try({g_roms})
```

```{r}
if(clim_cov == "temp"){
nd_temp <- data.frame(
  depth_scaled = min(data$depth_scaled)-0.2, 
  sandy_scaled = 0,
  muddy_scaled = 0,
  temp_scaled = seq(min(data$temp_scaled), 
    max(data$temp_scaled), length.out = 50),
  year_pair = 2011 # a chosen year
)
p_temp <- predict(m, newdata = nd_temp, se_fit = TRUE, re_form = NA)
p_temp$temp <- (p_temp$temp_scaled*data$temp_sd[1])+ data$temp_mean[1]

(g_temp <- ggplot(p_temp, aes(temp, exp(est), 
  ymin = exp(est - 1.96 * est_se), 
  ymax = exp(est + 1.96 * est_se))) +
  # coord_cartesian(ylim = c(0, max(exp(p_temp$est))*20)) +
  geom_line() + geom_ribbon(alpha = 0.4))

ggsave(paste0("figs/temp-", spp, covs, ".png"), width = 5, height = 5)
}
try({g_temp})
```

```{r}
if(covs %in% c("-sand-depth-roms", "-sand-depth-temp")){
nd_sand <- data.frame(
  depth_scaled = min(data$depth_scaled)-0.2, 
  sandy_scaled = seq(min(data$sandy_scaled), 
    max(data$sandy_scaled), length.out = 50),
  muddy_scaled = 0,
  roms_scaled = 0,
  temp_scaled = 0,
  year_pair = 2011 # a chosen year
)

p_sand <- predict(m, newdata = nd_sand, se_fit = TRUE, re_form = NA)
p_sand$sandy <- (p_sand$sandy_scaled*data$sandy_sd[1])+ data$sandy_mean[1]

(g_sand <- ggplot(p_sand, aes(sandy, exp(est),
  ymin = exp(est - 1.96 * est_se), 
  ymax = exp(est + 1.96 * est_se))) +
  # coord_cartesian(ylim = c(0,max(exp(p_sand$est))*20)) +
  geom_line() + geom_ribbon(alpha = 0.4))

# ggplot(p_sand, aes(sandy, exp(est),
#   ymin = exp(est - 1.96 * est_se), 
#   ymax = exp(est + 1.96 * est_se))) +
#   coord_cartesian(ylim = c(0, max(exp(p_sand$est))*2), xlim = c(0,1)) +
#   geom_line() + geom_ribbon(alpha = 0.4)

ggsave(paste0("figs/sand-", spp, covs, ".png"), width = 5, height = 5)
}
try({g_sand})
```

```{r}
if(covs %in% c("-mud-depth-roms", "-mud-depth-temp")){
nd_mud <- data.frame(
  depth_scaled = min(data$depth_scaled)-0.2, 
  muddy_scaled = seq(min(data$muddy_scaled), 
    max(data$muddy_scaled), length.out = 50),
  sandy_scaled = 0,
  roms_scaled = 0,
  temp_scaled = 0,
  year_pair = 2011 # a chosen year
)

p_mud <- predict(m, newdata = nd_mud, se_fit = TRUE, re_form = NA)
p_mud$muddy <- (p_mud$muddy_scaled*data$muddy_sd[1])+ data$muddy_mean[1]

(g_mud <- ggplot(p_mud, aes(muddy, exp(est),
  ymin = exp(est - 1.96 * est_se), 
  ymax = exp(est + 1.96 * est_se))) +
  # coord_cartesian(ylim = c(0,max(exp(p_mud$est))*20)) +
  geom_line() + geom_ribbon(alpha = 0.4))

# ggplot(p_mud, aes(muddy, exp(est),
#   ymin = exp(est - 1.96 * est_se), 
#   ymax = exp(est + 1.96 * est_se))) +
#   coord_cartesian(ylim = c(0, max(exp(p_mud$est))*2), xlim = c(0,1)) +
#   geom_line() + geom_ribbon(alpha = 0.4)

ggsave(paste0("figs/mud-", spp, covs, ".png"), width = 5, height = 5)
}
try({g_mud})
```

### Map biomass predictions

Save predictions for spatial grid
```{r}
rm(ad_predictions)
rm(im_predictions)
rm(predicted)

if(params$update_predictions) {
  newcov <- readRDS("data/new_covariates.rds") %>% 
    dplyr::select(X, Y, sandy, muddy, rocky, mixed)

  nd_all <- readRDS(
    here::here("analysis/VOCC/data/predicted-all-temp-with-roms.rds")
    # paste0("data/predicted-DO-2020-06-20-more2016-roms.rds")
    ) %>% dplyr::select(X, Y, year, ssid, depth, roms = roms_temp, temp)

  scaledata <- data %>% 
    dplyr::select(depth_mean, depth_sd, 
      sandy_mean, sandy_sd,  muddy_mean, muddy_sd, 
      roms_mean, roms_sd, temp_mean, temp_sd)

  nd_all <- left_join(nd_all, newcov)

  nd_all$depth_mean <- scaledata$depth_mean[1]
  nd_all$depth_sd <- scaledata$depth_sd[1]
  nd_all$sandy_mean <- scaledata$sandy_mean[1]
  nd_all$sandy_sd <- scaledata$sandy_sd[1]
  nd_all$muddy_mean <- scaledata$muddy_mean[1]
  nd_all$muddy_sd <- scaledata$muddy_sd[1]
  nd_all$roms_mean <- scaledata$roms_mean[1]
  nd_all$roms_sd <- scaledata$roms_sd[1]
  nd_all$temp_mean <- scaledata$temp_mean[1]
  nd_all$temp_sd <- scaledata$temp_sd[1]

  nd_all <- nd_all %>% mutate(
    depth_scaled = (log(depth) - depth_mean)/depth_sd,
    sandy_scaled = (sandy - sandy_mean)/sandy_sd,
    muddy_scaled = (muddy - muddy_mean)/muddy_sd,
    roms_scaled = (roms - roms_mean)/roms_sd,
    temp_scaled = (temp - temp_mean)/temp_sd
  ) 
  nd_all <- nd_all %>% filter(year > 2003) %>% 
  # merge survey pairs such that even year surveys are pooled w the previous odd year
    mutate(year_pair = (2 * round((year+0.8)/2))-1) %>% 
    # 0.8 is required to avoid 5s rounding to even in R
    filter(year < 2019)

  m <- readRDS(paste0("models/", spp, "/mod-tot-biomass-", spp, 
    covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
  ))

  nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(m$data$year))
  nd <- na.omit(nd)
  # nd$year <- as.integer(nd$year)

  predicted <- predict(m, newdata = nd, se_fit = F, return_tmb_object = T)

  saveRDS(predicted, file = paste0(
    "data/", spp, "/predictions-", spp, 
    covs, "-", ssid_string, "-tot-biomass-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
  ))
}
```

Transform estimates to biomass density in kg/ha
```{r}
rm(predictions)
rm(predicted)

  try({
    predicted <- readRDS(paste0("data/", spp, "/predictions-", spp, 
      covs, "-", ssid_string, "-tot-biomass-ar1-", 
      params$AR1, "-reml-", params$REML, params$knots, ".rds"
    ))
  })

  predictions <- predicted$data
  predictions$total_bio <- exp(predictions$est) 
  max_raster <- quantile(predictions$total_bio, 0.999)
  max_bio <- signif(quantile(predictions$total_bio, 1), digits = 2)

  predictions$prop_max <- predictions$total_bio/max_raster

  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/just-predictions-", spp, covs, "-total-biomass-ar1-", 
    params$AR1, "-reml-", params$REML, params$knots, ".rds"
  ))
```

```{r, error=TRUE}
legend_coords <- c(0.12, 0.2) #"none"

p_mean <- predictions  %>% group_by(X, Y) %>%
  mutate(mean_est = mean(exp(est))) %>% 
  filter(year_pair == 2011) %>% 
  mutate(year = "all years") %>%
  plot_facet_map("mean_est",
    raster_limits = c(0, max_raster),
    transform_col = fourth_root_power,
    legend_position = legend_coords
  ) + labs(fill = "kg/ha") +
  ggtitle(paste0("Estimated biomass density of ", species, 
    " \n(max = ", max_bio, " kg/ha)"))
print(p_mean)
ggsave(paste0("figs/map-", spp, covs, "-mean-biomass.png"), width = 5, height = 5)
```

Spatial random field - captures variance not explained by depth and temperature
```{r, error=TRUE}
legend_coords <- c(0.12, 0.2) #"none"

p_omega <- predictions  %>% filter(year_pair == 2011) %>%
  mutate(x = X, y = Y, year = year_pair) %>%
  filter(year_pair == 2011) %>% 
  mutate(year = "all years") %>%
  plot_facet_map("omega_s",
    # raster_limits = c(0, 1),
    # transform_col = fourth_root_power,
    legend_position =  legend_coords
  ) + ggtitle(paste0("Spatial random field"))

print(p_omega)

ggsave(paste0("figs/map-", spp, covs, "-spatial-rf.png"), 
  width = 5, height = 5)
```

Spatiotemporal random fields
```{r, error=TRUE}
legend_coords <- c(0.9, 0.12) #"none"

p_st <- predictions  %>% 
  mutate(x = X, y = Y) %>%
  plot_facet_map("epsilon_st",
    legend_position = legend_coords,
    transform_col = fourth_root_power
  ) +
    ggtitle(paste0("Spatiotemporal random fields"))

print(p_st)
ggsave(paste0("figs/map-", spp, covs, "-spatiotemporal-rf.png"), 
  width = 5, height = 5)
```

### 2020-11-25 analysis/sand-lance/occupancy-model.Rmd
draft sand lance models

--- 2020-11-25 analysis/sand-lance/biomass-density-model.Rmd
@@
 ---
-title: "Spatiotemporal model of `r paste(params$species)` biomass"
+title: "Spatiotemporal model of `r paste(params$species)` occupancy"
 author: "Philina English"
@@
     # covs: "-sand-depth-temp"
-    covs: "-mud-depth-temp"
+    # covs: "-mud-depth-temp"
+    # covs: "-sand-mud-depth-temp"
+    covs: "-sand-by-depth-mud-temp"
     REML: FALSE
@@
 
+logit2prob <- function(logit){
+  odds <- exp(logit)
+  prob <- odds / (1 + odds)
+  return(prob)
+}
+
 theme_set(
@@
 mod.formula <- as.formula(paste0(
-  "density ~ 0 + as.factor(year_pair) + 
+  "present ~ 0 + as.factor(year_pair) + 
   s(sandy_scaled, k = 3) + 
@@
 mod.formula <- as.formula(paste0(
-  "density ~ 0 + as.factor(year_pair) + 
+  "present ~ 0 + as.factor(year_pair) + 
   s(depth_scaled, k = 3) + 
@@
 mod.formula <- as.formula(paste0(
-  "density ~ 0 + as.factor(year_pair) + 
+  "present ~ 0 + as.factor(year_pair) + 
   s(muddy_scaled, k = 3) + 
   s(depth_scaled, k = 3) + 
+  s(temp_scaled, k = 3)"
+  ))
+clim_cov <- "temp"
+}
+
+if(params$covs == "-sand-mud-depth-temp"){
+mod.formula <- as.formula(paste0(
+  "present ~ 0 + as.factor(year_pair) + 
+  s(sandy_scaled, k = 3) + 
+  s(muddy_scaled, k = 3) + 
+  s(depth_scaled, k = 3) + 
+  s(temp_scaled, k = 3)"
+  ))
+clim_cov <- "temp"
+}
+
+if(params$covs == "-sand-by-depth-mud-temp"){
+mod.formula <- as.formula(paste0(
+  "present ~ 0 + as.factor(year_pair) + 
+  te(sandy_scaled, depth_scaled, k = 5) + 
+  s(muddy_scaled, k = 3) + 
   s(temp_scaled, k = 3)"
@@
 paste("AR1 =", AR1)
+paste("formula =", mod.formula)
 ```
@@
 )
-
-# # rename raw_depth for plots 
-# d <- d %>% mutate(depth = raw_depth) %>%
-#   filter(ssid %in% model_ssid)
 
@@
 
-    total_biomass <- sdmTMB::sdmTMB(
+    presence_mod <- sdmTMB::sdmTMB(
       data = data, mod.formula,
@@
       time = "year_pair", spde = spde,
-      family = tweedie(link = "log"),
+      family = binomial(link = "logit"),
       ar1_fields = params$AR1,
@@
 
-    saveRDS(total_biomass, file = paste0(
-      "models/", spp, "/mod-tot-biomass-", spp, 
+    saveRDS(presence_mod, file = paste0(
+      "models/", spp, "/mod-tot-binomial-", spp, 
       covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
@@
 m <- readRDS(paste0(
-  "models/", spp, "/mod-tot-biomass-", spp, 
+  "models/", spp, "/mod-tot-binomial-", spp, 
   covs, "-", ssid_string, "-ar1-", params$AR1, 
@@
 saveRDS(point_predictions, file = paste0(
-  "data/", spp, "/check-mod-predictions-", spp, 
+  "data/", spp, "/check-binomial-mod-predictions-", spp, 
   covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", 
@@
 pred <- readRDS(paste0(
-  "data/", spp, "/check-mod-predictions-", spp, 
+  "data/", spp, "/check-binomial-mod-predictions-", spp, 
   covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
@@
 ```{r}
-g <- ggplot(pred, 
-    aes(density, residuals, colour = as.factor(ssid))) +
-  geom_point() + scale_x_continuous(trans = "log10") +  
-  facet_wrap(~year_pair)
-g
-```
-
-```{r}
 g1 <- ggplot(pred,
-      aes(density, exp(est), colour = as.factor(ssid))) +
-    geom_abline(slope=1, intercept = 0) +
-    scale_x_continuous(trans = 'log10') +
-    scale_y_continuous(trans = 'log10') +
+      aes(logit2prob(est), present)) +
+      # aes(as.factor(present), logit2prob(est))) +
+    geom_smooth(method = "glm", 
+    method.args = list(family = "binomial"), se=F)+
     geom_point(alpha = 0.2) 
@@
 g2 <- ggplot(pred, 
-      aes((est), residuals, colour = density)) +
+      aes(est, residuals, colour = as.factor(present))) +
     geom_point(alpha = 0.2) + # scale_x_continuous(trans = 'log10') +
@@
 g2
+```
+
+```{r}
+g3 <- ggplot(pred, 
+      aes(log_depth, residuals, colour = as.factor(present))) +
+    geom_point(alpha = 0.2) + # scale_x_continuous(trans = 'log10') +
+    facet_wrap(~year)
+g3
 ```
@@
 m <- readRDS(paste0(
-    "models/", spp, "/mod-tot-biomass-", spp, 
+    "models/", spp, "/mod-tot-binomial-", spp, 
       covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
@@
 
-(g_depth <- ggplot(p_depth, aes(depth, exp(est), 
-  ymin = exp(est - 1.96 * est_se), 
-  ymax = exp(est + 1.96 * est_se))) +
-  coord_cartesian(ylim = c(0, max(exp(p_depth$est))*20), xlim = c(0, 150)) +
+(g_depth <- ggplot(p_depth, aes(depth, logit2prob(est), 
+  ymin = logit2prob(est - 1.96 * est_se), 
+  ymax = logit2prob(est + 1.96 * est_se))) +
+  # coord_cartesian(ylim = c(0, max(logit2prob(p_depth$est))*20), xlim = c(0, 150)) +
   geom_line() + geom_ribbon(alpha = 0.4))
 
-ggsave(paste0("figs/depth-", spp, covs, ".png"), width = 5, height = 5)
+ggsave(paste0("figs/depth-binomial-", spp, covs, ".png"), width = 5, height = 5)
 }
@@
 
-(g_roms <- ggplot(p_roms, aes(roms, exp(est), 
-  ymin = exp(est - 1.96 * est_se), 
-  ymax = exp(est + 1.96 * est_se))) +
-  coord_cartesian(ylim = c(0, max(exp(p_roms$est))*20)) +
+(g_roms <- ggplot(p_roms, aes(roms, logit2prob(est), 
+  ymin = logit2prob(est - 1.96 * est_se), 
+  ymax = logit2prob(est + 1.96 * est_se))) +
+  # coord_cartesian(ylim = c(0, max(logit2prob(p_roms$est))*20)) +
   geom_line() + geom_ribbon(alpha = 0.4))
 
-ggsave(paste0("figs/roms-", spp, covs, ".png"), width = 5, height = 5)
+ggsave(paste0("figs/roms-binomial-", spp, covs, ".png"), width = 5, height = 5)
 }
-
 try({g_roms})
@@
 nd_temp <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled)-0.2, # use min depth for estimated error at relevant depths
   sandy_scaled = 0,
@@
 
-(g_temp <- ggplot(p_temp, aes(temp, exp(est), 
-  ymin = exp(est - 1.96 * est_se), 
-  ymax = exp(est + 1.96 * est_se))) +
-  # coord_cartesian(ylim = c(0, max(exp(p_temp$est))*20)) +
+(g_temp <- ggplot(p_temp, aes(temp, logit2prob(est), 
+  ymin = logit2prob(est - 1.96 * est_se), 
+  ymax = logit2prob(est + 1.96 * est_se))) +
+  # coord_cartesian(ylim = c(0, max(logit2prob(p_temp$est))*20)) +
   geom_line() + geom_ribbon(alpha = 0.4))
 
-ggsave(paste0("figs/temp-", spp, covs, ".png"), width = 5, height = 5)
+ggsave(paste0("figs/temp-binomial-", spp, covs, ".png"), width = 5, height = 5)
 }
@@
 ```{r}
-if(covs %in% c("-sand-depth-roms", "-sand-depth-temp")){
+if(covs %in% c("-sand-depth-roms", "-sand-depth-temp", "-sand-mud-depth-temp", "-sand-by-depth-mud-temp")){
 nd_sand <- data.frame(
@@
 
-(g_sand <- ggplot(p_sand, aes(sandy, exp(est),
-  ymin = exp(est - 1.96 * est_se), 
-  ymax = exp(est + 1.96 * est_se))) +
-  # coord_cartesian(ylim = c(0,max(exp(p_sand$est))*20)) +
+(g_sand <- ggplot(p_sand, aes(sandy, logit2prob(est),
+  ymin = logit2prob(est - 1.96 * est_se), 
+  ymax = logit2prob(est + 1.96 * est_se))) +
+  # coord_cartesian(ylim = c(0,max(logit2prob(p_sand$est))*20)) +
   geom_line() + geom_ribbon(alpha = 0.4))
 
-# ggplot(p_sand, aes(sandy, exp(est),
-#   ymin = exp(est - 1.96 * est_se), 
-#   ymax = exp(est + 1.96 * est_se))) +
-#   coord_cartesian(ylim = c(0, max(exp(p_sand$est))*2), xlim = c(0,1)) +
+# ggplot(p_sand, aes(sandy, logit2prob(est),
+#   ymin = logit2prob(est - 1.96 * est_se), 
+#   ymax = logit2prob(est + 1.96 * est_se))) +
+#   coord_cartesian(ylim = c(0, max(logit2prob(p_sand$est))*2), xlim = c(0,1)) +
 #   geom_line() + geom_ribbon(alpha = 0.4)
 
-ggsave(paste0("figs/sand-", spp, covs, ".png"), width = 5, height = 5)
+ggsave(paste0("figs/sand-binomial-", spp, covs, ".png"), width = 5, height = 5)
 }
@@
 ```{r}
-if(covs %in% c("-mud-depth-roms", "-mud-depth-temp")){
+if(covs %in% c("-mud-depth-roms", "-mud-depth-temp", "-sand-mud-depth-temp", "-sand-by-depth-mud-temp")){
 nd_mud <- data.frame(
@@
 
-(g_mud <- ggplot(p_mud, aes(muddy, exp(est),
-  ymin = exp(est - 1.96 * est_se), 
-  ymax = exp(est + 1.96 * est_se))) +
-  # coord_cartesian(ylim = c(0,max(exp(p_mud$est))*20)) +
+(g_mud <- ggplot(p_mud, aes(muddy, logit2prob(est),
+  ymin = logit2prob(est - 1.96 * est_se), 
+  ymax = logit2prob(est + 1.96 * est_se))) +
+  # coord_cartesian(ylim = c(0,max(logit2prob(p_mud$est))*20)) +
   geom_line() + geom_ribbon(alpha = 0.4))
 
-# ggplot(p_mud, aes(muddy, exp(est),
-#   ymin = exp(est - 1.96 * est_se), 
-#   ymax = exp(est + 1.96 * est_se))) +
-#   coord_cartesian(ylim = c(0, max(exp(p_mud$est))*2), xlim = c(0,1)) +
+# ggplot(p_mud, aes(muddy, logit2prob(est),
+#   ymin = logit2prob(est - 1.96 * est_se), 
+#   ymax = logit2prob(est + 1.96 * est_se))) +
+#   coord_cartesian(ylim = c(0, max(logit2prob(p_mud$est))*2), xlim = c(0,1)) +
 #   geom_line() + geom_ribbon(alpha = 0.4)
 
-ggsave(paste0("figs/mud-", spp, covs, ".png"), width = 5, height = 5)
+ggsave(paste0("figs/mud-binomial-", spp, covs, ".png"), width = 5, height = 5)
 }
@@
 
-  m <- readRDS(paste0("models/", spp, "/mod-tot-biomass-", spp, 
+  m <- readRDS(paste0("models/", spp, "/mod-tot-binomial-", spp, 
     covs, "-", ssid_string, "-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
@@
     "data/", spp, "/predictions-", spp, 
-    covs, "-", ssid_string, "-tot-biomass-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
+    covs, "-", ssid_string, "-tot-binomial-ar1-", params$AR1, "-reml-", params$REML, params$knots, ".rds"
   ))
@@
 
-Transform estimates to biomass density in kg/ha
+Transform estimates to probability of occupancy
 ```{r}
@@
     predicted <- readRDS(paste0("data/", spp, "/predictions-", spp, 
-      covs, "-", ssid_string, "-tot-biomass-ar1-", 
+      covs, "-", ssid_string, "-tot-binomial-ar1-", 
       params$AR1, "-reml-", params$REML, params$knots, ".rds"
@@
   predictions <- predicted$data
-  predictions$total_bio <- exp(predictions$est) 
-  max_raster <- quantile(predictions$total_bio, 0.999)
-  max_bio <- signif(quantile(predictions$total_bio, 1), digits = 2)
+  predictions$total_prob <- logit2prob(predictions$est) 
+  max_raster <- quantile(predictions$total_prob, 0.999)
+  max_prob <- signif(quantile(predictions$total_prob, 1), digits = 2)
 
-  predictions$prop_max <- predictions$total_bio/max_raster
+  predictions$prop_max <- predictions$total_prob/max_raster
 
@@
     "data/", spp,
-    "/just-predictions-", spp, covs, "-total-biomass-ar1-", 
+    "/just-predictions-", spp, covs, "-total-binomial-ar1-", 
     params$AR1, "-reml-", params$REML, params$knots, ".rds"
@@
 p_mean <- predictions  %>% group_by(X, Y) %>%
-  mutate(mean_est = mean(exp(est))) %>% 
+  mutate(mean_est = mean(logit2prob(est))) %>% 
   filter(year_pair == 2011) %>% 
@@
     legend_position = legend_coords
-  ) + labs(fill = "kg/ha") +
+  ) + labs(fill = "prob") +
   ggtitle(paste0("Estimated biomass density of ", species, 
-    " \n(max = ", max_bio, " kg/ha)"))
+    " \n(max = ", max_prob, ")"))
 print(p_mean)
-ggsave(paste0("figs/map-", spp, covs, "-mean-biomass.png"), width = 5, height = 5)
+ggsave(paste0("figs/map-", spp, covs, "-mean-distribution.png"), width = 5, height = 5)
 ```
@@
 
-ggsave(paste0("figs/map-", spp, covs, "-spatial-rf.png"), 
+ggsave(paste0("figs/map-", spp, covs, "-binomial-spatial-rf.png"), 
   width = 5, height = 5)
@@
 print(p_st)
-ggsave(paste0("figs/map-", spp, covs, "-spatiotemporal-rf.png"), 
+ggsave(paste0("figs/map-", spp, covs, "-binomial-spatiotemporal-rf.png"), 
   width = 5, height = 5)

### 2020-11-25 analysis/sand-lance/biomass-density-model.Rmd
refine models and effect plots

@@
     # covs: "-depth-roms"
-    # covs: "-sand-depth-temp"
-    covs: "-mud-depth-temp"
+    covs: "-sand-depth-temp"
+    # covs: "-mud-depth-temp"
+    # covs: "-mud-sand-depth-temp"
     REML: FALSE
@@
 
-if(params$covs == "-depth-roms"){
+if(params$covs == "-depth-temp"){
 mod.formula <- as.formula(paste0(
@@
   s(depth_scaled, k = 3) + 
-  s(roms_scaled, k = 3)"
+  s(temp_scaled, k = 3)"
   ))
-clim_cov <- "roms"
+clim_cov <- "temp"
 }
@@
   "density ~ 0 + as.factor(year_pair) + 
-  s(sandy_scaled, k = 3) + 
+  s(sandy_scaled, by = shallow, k = 3) + 
   s(depth_scaled, k = 3) + 
@@
 }
+if(params$covs == "-mud-sand-depth-temp"){
+mod.formula <- as.formula(paste0(
+  "density ~ 0 + as.factor(year_pair) + 
+  s(muddy_scaled, k = 3) + 
+  s(sandy_scaled, k = 3) + 
+  s(depth_scaled, k = 3) + 
+  s(temp_scaled, k = 3)"
+  ))
+clim_cov <- "temp"
+}
 
@@
 knots <- params$knots
-REML <- params$REML
-AR1 <- params$AR1
+# REML <- params$REML
+# AR1 <- params$AR1
 
-paste("region =", region)
+paste("region =", params$region)
 paste("climate covariate =", clim_cov)
-paste("model label =", covs)
-paste("knots =", knots)
-paste("REML =", REML)
-paste("AR1 =", AR1)
+paste("model label =", params$covs)
+paste("knots =", params$knots)
+paste("REML =", params$REML)
+paste("AR1 =", params$AR1)
+paste("fixed spatial =", params$fixed_spatial)
+paste("formula =", mod.formula[3])
 ```
@@
   raw_depth = depth, # create 'depth' var that's actually log depth for scaling
-  log_depth = log(depth), depth = log(raw_depth)) %>% 
+  log_depth = log(depth), depth = log(raw_depth),
+  shallow = if_else(raw_depth < 100, 1, 0)
+  ) %>% 
   rename(temp = temperature_c, do = do_mlpl)
@@
 
-```{r eval= FALSE}
+```{r eval = FALSE}
 # depth diff = difference between bath and sample depths
@@
 
+Check for substrate correlations -- sandy somewhat correlated with depth
+```{r eval= FALSE}
+ggplot(data, aes(raw_depth, sandy, colour = raw_depth)) + 
+  geom_point(alpha=0.75, size=0.5) +
+  scale_color_viridis_c(trans=fourth_root_power, direction = -1)
+
+ggplot(data, aes(raw_depth, sandy-muddy, colour = raw_depth)) + 
+  geom_point(alpha=0.75, size=0.5) +
+  scale_color_viridis_c(trans=fourth_root_power, direction = -1)
+
+ggplot(data, aes(raw_depth, muddy, colour = raw_depth)) + 
+  geom_point(alpha=0.75, size=0.5) +
+  scale_color_viridis_c(trans=fourth_root_power, direction = -1)
+
+ggplot(filter(data, raw_depth < 100), aes(depth, sandy, colour = depth)) + 
+  geom_point(alpha=0.75, size=0.5) +
+  scale_color_viridis_c(trans=fourth_root_power, direction = -1) 
+```
+
 Make mesh
@@
       data = data, mod.formula,
-      # time_varying = ~ 0 + depth_scaled + depth_scaled2, 
       time = "year_pair", spde = spde,
@@
 g2 <- ggplot(pred, 
-      aes((est), residuals, colour = density)) +
-    geom_point(alpha = 0.2) + # scale_x_continuous(trans = 'log10') +
+      aes(exp(est), residuals, colour = density)) +
+    geom_point(alpha = 0.2) + 
+    scale_x_continuous(trans = 'log10') +
     facet_wrap(~year)
@@
 ```{r}
+threshold <- quantile(data$density, 0.999)
 if(params$update_model_check){
@@
     max(data$depth_scaled), length.out = 30), 
+  shallow = 1,
   sandy_scaled = 0,
@@
 
+max_CI <- max(exp(p_depth$est + 1.96 * p_depth$est_se))
+max_Y <- if_else(threshold > max_CI, max_CI, threshold)
+
 (g_depth <- ggplot(p_depth, aes(depth, exp(est), 
@@
   ymax = exp(est + 1.96 * est_se))) +
-  coord_cartesian(ylim = c(0, max(exp(p_depth$est))*20), xlim = c(0, 150)) +
+  coord_cartesian(ylim = c(0, max_Y), xlim = c(0, 150)) +
   geom_line() + geom_ribbon(alpha = 0.4))
@@
 nd_roms <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled), # use min depth for estimated prob at most relevant depth
+  shallow = 1,
   sandy_scaled = 0,
@@
 p_roms$roms <- (p_roms$roms_scaled*data$roms_sd[1])+ data$roms_mean[1]
+
+max_CI <- max(exp(p_roms$est + 1.96 * p_roms$est_se))
+max_Y <- if_else(threshold > max_CI, max_CI, threshold)
 
@@
 nd_temp <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled), 
+  shallow = 1,
   sandy_scaled = 0,
@@
 
+max_CI <- max(exp(p_temp$est + 1.96 * p_temp$est_se))
+max_Y <- if_else(threshold > max_CI, max_CI, threshold)
+
 (g_temp <- ggplot(p_temp, aes(temp, exp(est), 
@@
   ymax = exp(est + 1.96 * est_se))) +
-  # coord_cartesian(ylim = c(0, max(exp(p_temp$est))*20)) +
+  coord_cartesian(ylim = c(0, max_Y)) +
   geom_line() + geom_ribbon(alpha = 0.4))
@@
 ```{r}
-if(covs %in% c("-sand-depth-roms", "-sand-depth-temp")){
+if(covs %in% c("-sand-depth-roms", "-sand-depth-temp", "-mud-sand-depth-temp")){
 nd_sand <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled), 
+  shallow = 1,
   sandy_scaled = seq(min(data$sandy_scaled), 
     max(data$sandy_scaled), length.out = 50),
-  muddy_scaled = 0,
+  muddy_scaled = min(data$muddy_scaled),
   roms_scaled = 0,
@@
   ymax = exp(est + 1.96 * est_se))) +
-  # coord_cartesian(ylim = c(0,max(exp(p_sand$est))*20)) +
+  coord_cartesian(ylim = c(0, max_Y)) +
   geom_line() + geom_ribbon(alpha = 0.4))
-
-# ggplot(p_sand, aes(sandy, exp(est),
-#   ymin = exp(est - 1.96 * est_se), 
-#   ymax = exp(est + 1.96 * est_se))) +
-#   coord_cartesian(ylim = c(0, max(exp(p_sand$est))*2), xlim = c(0,1)) +
-#   geom_line() + geom_ribbon(alpha = 0.4)
 
@@
 ```{r}
-if(covs %in% c("-mud-depth-roms", "-mud-depth-temp")){
+if(covs %in% c("-mud-depth-roms", "-mud-depth-temp","-mud-sand-depth-temp")){
 nd_mud <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled), 
+  shallow = 1,
   muddy_scaled = seq(min(data$muddy_scaled), 
     max(data$muddy_scaled), length.out = 50),
-  sandy_scaled = 0,
+  sandy_scaled = max(data$sandy_scaled),
   roms_scaled = 0,
@@
 
+max_CI <- max(exp(p_mud$est + 1.96 * p_mud$est_se))
+max_Y <- if_else(threshold > max_CI, max_CI, threshold)
+
 (g_mud <- ggplot(p_mud, aes(muddy, exp(est),
@@
   ymax = exp(est + 1.96 * est_se))) +
-  # coord_cartesian(ylim = c(0,max(exp(p_mud$est))*20)) +
+  coord_cartesian(ylim = c(0, max_Y)) +
   geom_line() + geom_ribbon(alpha = 0.4))
-
-# ggplot(p_mud, aes(muddy, exp(est),
-#   ymin = exp(est - 1.96 * est_se), 
-#   ymax = exp(est + 1.96 * est_se))) +
-#   coord_cartesian(ylim = c(0, max(exp(p_mud$est))*2), xlim = c(0,1)) +
-#   geom_line() + geom_ribbon(alpha = 0.4)
 

### 2020-11-25 analysis/sand-lance/occupancy-model.Rmd
refine models and effect plots

@@
     # covs: "-sand-depth-temp"
-    # covs: "-mud-depth-temp"
+    covs: "-mud-depth-temp"
     # covs: "-sand-mud-depth-temp"
-    covs: "-sand-by-depth-mud-temp"
+    # covs: "-sand-by-depth-mud-temp"
     REML: FALSE
-    AR1: TRUE
-    fixed_spatial: TRUE
+    AR1: FALSE # true doesn't seem to work with occupancy models
+    fixed_spatial: FALSE
     update_model: TRUE
@@
   "present ~ 0 + as.factor(year_pair) + 
-  te(sandy_scaled, depth_scaled, k = 5) + 
+  te(sandy_scaled, depth_scaled, k = 3) + 
   s(muddy_scaled, k = 3) + 
@@
 knots <- params$knots
-REML <- params$REML
-AR1 <- params$AR1
+# REML <- params$REML
+# AR1 <- params$AR1
 
-paste("region =", region)
+paste("region =", params$region)
 paste("climate covariate =", clim_cov)
-paste("model label =", covs)
-paste("knots =", knots)
-paste("REML =", REML)
-paste("AR1 =", AR1)
-paste("formula =", mod.formula)
+paste("model label =", params$covs)
+paste("knots =", params$knots)
+paste("REML =", params$REML)
+paste("AR1 =", params$AR1)
+paste("fixed spatial =", params$fixed_spatial)
+paste("formula =", mod.formula[3])
 ```
@@
 nd_roms <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled), # use min depth for estimated prob at most relevant depth
   sandy_scaled = 0,
@@
 nd_temp <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, # use min depth for estimated error at relevant depths
+  depth_scaled = min(data$depth_scaled), 
   sandy_scaled = 0,
@@
 nd_sand <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled), 
   sandy_scaled = seq(min(data$sandy_scaled), 
@@
   ) + labs(fill = "prob") +
-  ggtitle(paste0("Estimated biomass density of ", species, 
+  ggtitle(paste0("Estimated occupancy of ", species, 
     " \n(max = ", max_prob, ")"))
 print(p_mean)
-ggsave(paste0("figs/map-", spp, covs, "-mean-distribution.png"), width = 5, height = 5)
+ggsave(paste0("figs/map-", spp, covs, "-mean-occupancy.png"), width = 5, height = 5)
 ```

### 2020-11-26 analysis/sand-lance/biomass-density-model.Rmd
working occupancy model

@@
     # covs: "-depth-roms"
-    covs: "-sand-depth-temp"
+    # covs: "-sand-depth-temp"
     # covs: "-mud-depth-temp"
     # covs: "-mud-sand-depth-temp"
+    covs: "-sand-x-depth-temp" # doesn't converge
     REML: FALSE
@@
   "density ~ 0 + as.factor(year_pair) + 
+  s(sandy_scaled, k = 3) + 
   s(sandy_scaled, by = shallow, k = 3) + 
@@
   "density ~ 0 + as.factor(year_pair) + 
-  s(muddy_scaled, k = 3) + 
-  s(depth_scaled, k = 3) + 
-  s(temp_scaled, k = 3)"
-  ))
-clim_cov <- "temp"
-}
-if(params$covs == "-mud-sand-depth-temp"){
-mod.formula <- as.formula(paste0(
-  "density ~ 0 + as.factor(year_pair) + 
-  s(muddy_scaled, k = 3) + 
-  s(sandy_scaled, k = 3) + 
+  s(muddy_scaled, k = 2) + 
   s(depth_scaled, k = 3) + 
@@
 
+if(params$covs == "-mud-sand-depth-temp"){
+mod.formula <- as.formula(paste0(
+  "density ~ 0 + as.factor(year_pair) + 
+  muddy_scaled + 
+  sandy_scaled + 
+  s(depth_scaled, k = 3) + 
+  s(temp_scaled, k = 3)"
+  ))
+clim_cov <- "temp"
+}
+
+if(params$covs == "-sand-x-depth-temp"){
+mod.formula <- as.formula(paste0(
+  "density ~ 0 + as.factor(year_pair) + 
+  sandy_scaled + 
+  s(depth_scaled, k = 3) + 
+  te(depth_scaled, sandy_scaled, k = 3, m=1) + 
+  s(temp_scaled, k = 3)"
+  ))
+clim_cov <- "temp"
+}
 ```
@@
 Check for substrate correlations -- sandy somewhat correlated with depth
-```{r eval= FALSE}
+```{r}
 ggplot(data, aes(raw_depth, sandy, colour = raw_depth)) + 
@@
 
-ggplot(data, aes(raw_depth, sandy-muddy, colour = raw_depth)) + 
+ggplot(filter(data, raw_depth < 100), aes(raw_depth, sandy, colour = raw_depth)) + 
   geom_point(alpha=0.75, size=0.5) +
@@
 
-ggplot(filter(data, raw_depth < 100), aes(depth, sandy, colour = depth)) + 
-  geom_point(alpha=0.75, size=0.5) +
-  scale_color_viridis_c(trans=fourth_root_power, direction = -1) 
+# ggplot(data, aes(raw_depth, sandy-muddy, colour = raw_depth)) + 
+#   geom_point(alpha=0.75, size=0.5) +
+#   scale_color_viridis_c(trans=fourth_root_power, direction = -1) 
 ```
@@
 }
-try({g_depth})
+try({g_depth}, silent = TRUE)
 ```
@@
 
-try({g_roms})
+try({g_roms}, silent = TRUE)
 ```
@@
 }
-try({g_temp})
+try({g_temp}, silent = TRUE)
 ```
@@
 }
-try({g_sand})
+try({g_sand}, silent = TRUE)
 ```
@@
 }
-try({g_mud})
+try({g_mud}, silent = TRUE)
 ```
@@
   # merge survey pairs such that even year surveys are pooled w the previous odd year
-    mutate(year_pair = (2 * round((year+0.8)/2))-1) %>% 
     # 0.8 is required to avoid 5s rounding to even in R
+    mutate(year_pair = (2 * round((year+0.8)/2))-1,
+      shallow = if_else(depth < 100, 1, 0)) %>% 
     filter(year < 2019)
@@
 
-  try({
   predicted <- readRDS(paste0("data/", spp, "/predictions-", spp, 
@@
     ))
-  })
 

### 2020-11-26 analysis/sand-lance/occupancy-model.Rmd
working occupancy model

@@
     # covs: "-depth-roms"
-    # covs: "-sand-depth-temp"
-    covs: "-mud-depth-temp"
+    covs: "-sand-depth-temp"
+    # covs: "-mud-depth-temp"
     # covs: "-sand-mud-depth-temp"
@@
     AR1: FALSE # true doesn't seem to work with occupancy models
-    fixed_spatial: FALSE
+    fixed_spatial: TRUE
     update_model: TRUE
@@
   "present ~ 0 + as.factor(year_pair) + 
-  s(sandy_scaled, k = 3) + 
+  sandy_scaled + 
   s(depth_scaled, k = 3) + 
@@
   "present ~ 0 + as.factor(year_pair) + 
-  s(muddy_scaled, k = 3) + 
+  muddy_scaled + 
   s(depth_scaled, k = 3) + 
@@
   "present ~ 0 + as.factor(year_pair) + 
-  s(sandy_scaled, k = 3) + 
-  s(muddy_scaled, k = 3) + 
+  sandy_scaled +
+  muddy_scaled + 
   s(depth_scaled, k = 3) + 
@@
 
+# probably too complex to fit
 if(params$covs == "-sand-by-depth-mud-temp"){
@@
   "present ~ 0 + as.factor(year_pair) + 
+  s(sandy_scaled, k = 3) + 
+  s(depth_scaled, k = 3) + 
   te(sandy_scaled, depth_scaled, k = 3) + 
@@
   raw_depth = depth, # create 'depth' var that's actually log depth for scaling
-  log_depth = log(depth), depth = log(raw_depth)) %>% 
+  log_depth = log(depth), depth = log(raw_depth),
+  shallow = if_else(raw_depth < 100, 1, 0)) %>% 
   rename(temp = temperature_c, do = do_mlpl)
@@
 g2 <- ggplot(pred, 
-      aes(est, residuals, colour = as.factor(present))) +
+      aes(logit2prob(est), residuals, colour = as.factor(present))) +
     geom_point(alpha = 0.2) + # scale_x_continuous(trans = 'log10') +
@@
     max(data$depth_scaled), length.out = 30), 
+  shallow = 1, 
   sandy_scaled = 0,
@@
   depth_scaled = min(data$depth_scaled), # use min depth for estimated prob at most relevant depth
+  shallow = 1,
   sandy_scaled = 0, 
@@
 }
-try({g_roms})
+try({g_roms}, silent = TRUE)
 ```
@@
   depth_scaled = min(data$depth_scaled), 
+  shallow = 1, 
   sandy_scaled = 0,
@@
   depth_scaled = min(data$depth_scaled), 
+  shallow = 1, 
   sandy_scaled = seq(min(data$sandy_scaled), 
@@
 nd_mud <- data.frame(
-  depth_scaled = min(data$depth_scaled)-0.2, 
+  depth_scaled = min(data$depth_scaled), 
+  shallow = 1, 
   muddy_scaled = seq(min(data$muddy_scaled), 
@@
 
-  try({
 predicted <- readRDS(paste0("data/", spp, "/predictions-", spp, 
@@
   ) + labs(fill = "prob") +
-  ggtitle(paste0("Estimated occupancy of ", species, 
-    " \n(max = ", max_prob, ")"))
+  ggtitle(paste0("Probability of occupancy by ", species))
 print(p_mean)

### 2020-11-28 analysis/sand-lance/biomass-density-model.Rmd
make sandy linear only in sand lance density model

@@
     # covs: "-sand-depth-temp"
-    # covs: "-mud-depth-temp"
+    covs: "-mud-depth-temp"
     # covs: "-mud-sand-depth-temp"
-    covs: "-sand-x-depth-temp" # doesn't converge
+    # covs: "-sand-shallow-depth-temp"
+    # covs: "-sand-x-depth-temp" # doesn't converge
     REML: FALSE
@@
     fixed_spatial: TRUE
+    # update_model: FALSE
     update_model: TRUE
-    # update_model: FALSE
     update_model_check: TRUE
@@
   "density ~ 0 + as.factor(year_pair) + 
-  s(sandy_scaled, k = 3) + 
-  s(sandy_scaled, by = shallow, k = 3) + 
+  sandy_scaled + 
   s(depth_scaled, k = 3) + 
@@
 
+if(params$covs == "-sand-shallow-depth-temp"){
+mod.formula <- as.formula(paste0(
+  "density ~ 0 + as.factor(year_pair) + 
+  sandy_scaled * shallow + 
+  s(depth_scaled, k = 3) + 
+  s(temp_scaled, k = 3)"
+  ))
+clim_cov <- "temp"
+}
+  # s(sandy_scaled, k = 3) + 
+  # s(sandy_scaled, by = shallow, k = 3) +
+
 if(params$covs == "-mud-depth-temp"){
@@
   "density ~ 0 + as.factor(year_pair) + 
-  s(muddy_scaled, k = 2) + 
+  muddy_scaled + 
   s(depth_scaled, k = 3) + 
@@
 g1 <- ggplot(pred,
-      aes(density, exp(est), colour = as.factor(ssid))) +
+      aes(density+0.01, exp(est), colour = as.factor(ssid))) +
     geom_abline(slope=1, intercept = 0) +
@@
 ```{r}
-if(covs %in% c("-sand-depth-roms", "-sand-depth-temp", "-mud-sand-depth-temp")){
+if(covs %in% c("-sand-depth-roms", "-sand-depth-temp", "-mud-sand-depth-temp", "-sand-shallow-depth-temp")){
 nd_sand <- data.frame(

### 2021-05-20 analysis/misc/halibut-south-outside.Rmd
exploring model structures for outside S HBLL

---
title: "outside halibut"
author: "Philina English"
date: "18/05/2021"
output: html_document
---

```{r}
# setwd(here::here()) # for RStudio Jobs
library(dplyr)
library(ggplot2)
library(sdmTMB)
# devtools::install_github("seananderson/ggsidekick")
library(ggsidekick) # for fourth_root_power_trans
theme_set(ggsidekick::theme_sleek())

# data <- gfdata::get_ll_hook_data("Yelloweye Rockfish", ssid = c(36))
# 
# d1 <- gfdata::get_survey_sets("Yelloweye Rockfish", ssid = c(36))

```

```{r}
convert2utm <- function(df, coords = c("X", "Y"), out_crs = 3156) {
  x <- sf::st_as_sf(df,
    coords = coords, crs = 4326
  ) %>%
    sf::st_transform(crs = out_crs) %>%
    sf::st_coordinates() %>%
    as.data.frame()
  x$X <- x$X / 100000
  x$Y <- x$Y / 100000
  dplyr::bind_cols(x,
    df[, which(!names(df) %in% coords), drop = FALSE])
}

expand_prediction_grid <- function(grid, years) {
  nd <- do.call("rbind",
    replicate(length(years), grid, simplify = FALSE))
  nd[["year"]] <- rep(years, each = nrow(grid))
  nd
}
```

Model diagnostics function
```{r}
get_diag <- function(m, response = "catch_count", variable = "depth_scaled", colour_var = "depth_m", start_year = 2007) {
  predictions <- predict(m)
  predictions$residuals <- residuals(m)

  predictions <- predictions %>% filter (year >= start_year)
  predictions$est_exp <- exp(predictions$est)

  print("R^2:")
  r2 <- cor(predictions$est, predictions[[response]])^2
  print(r2)

  print("")
  print("AIC:")
  print(AIC(m))

  print("")
  print("MSE:")
  print(mean(predictions$residuals^2))

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point(alpha = 0.5) +
      coord_fixed()+
      gfplot::theme_pbs() 
  }

  g <- plot_map(predictions, "est") +
    scale_colour_viridis_c() +
    ggtitle("Prediction (fixed effects + all random effects)")
  print(g)

  # g <- plot_map(predictions, "est_non_rf") +
  #   ggtitle("Prediction (fixed effects only)") +
  #   scale_colour_viridis_c()
  # print(g)
# 
#   g <- plot_map(predictions, "est_rf") +
#     ggtitle("All random effects only") +
#     scale_colour_gradient2()
#   print(g)

  g <- plot_map(predictions, "omega_s") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- plot_map(predictions, "epsilon_st") +
    ggtitle("Spatiotemporal random effects only") +
    facet_wrap(~year) +
    scale_colour_gradient2()
  print(g)

  g <- plot_map(predictions, "residuals") +
    ggtitle("Residuals") +
    facet_wrap(~year) +
    scale_colour_gradient2()
  print(g)

  g <- ggplot(predictions, aes_string("est_exp", response)) +
    geom_point(alpha = 0.2) +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()+
    gfplot::theme_pbs() 
  print(g)

  g <- ggplot(predictions, aes(est, residuals)) +
    geom_point(alpha = 0.2) +
    geom_smooth()+
    gfplot::theme_pbs() 
  print(g)

  g <- ggplot(predictions, aes_string(variable, "residuals", colour = colour_var)) +
    geom_point(alpha = 0.4, size = 1.2) +
    geom_smooth(colour="grey", size = 1.2) +
    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
    # facet_wrap(~year, scales = "free_x")+
    gfplot::theme_pbs() + theme(
      axis.text.y = element_text(size = 14),
      axis.text.x = element_blank(), axis.ticks.x = element_blank())
  print(g)

  g <- ggplot(predictions, aes_string(variable, "residuals", colour = colour_var)) +
    geom_point(alpha = 0.4, size = 0.7) +
    geom_smooth(colour="grey", size = 0.7) +
    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
    facet_wrap(~year, scales = "free_x")+
    gfplot::theme_pbs() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  print(g)

}
```

# combine event and hook data if not done already
```{r}
hook_adjusted_data <- "data-generated/halibut_hook_adjusted_36.rds"
if (!file.exists(hook_adjusted_data)) {
data <- readRDS("data/halibut-hook-dat36.rds")
d1 <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")

# the following code should deal with the issue of 0 baited hooks being observed.
adjust <- data %>%
  group_by(year, fishing_event_id) %>%
  mutate(total_hooks = count_target_species + count_non_target_species +
    count_bait_only + count_empty_hooks - count_bent_broken) %>%
  mutate(count_bait_only = replace(count_bait_only, which(count_bait_only == 0), 1)) %>%
  mutate(prop_bait_hooks = count_bait_only / total_hooks) %>%
  mutate(hook_adjust_factor = -log(prop_bait_hooks) / (1 - prop_bait_hooks)) %>%
  mutate(expected_catch = round(count_target_species * hook_adjust_factor))

# find out which events are in hook data but not set data:
setdiff(data$fishing_event_id, d1$fishing_event_id)

hook_adjusted_data <- left_join(d1, adjust, by = c("fishing_event_id", "year"))

saveRDS(hook_adjusted_data, file = "data-generated/halibut_hook_adjusted_36.rds")
}
```

# add block ids and offsets
```{r}
f <- "data-generated/halibut-surv-36-blocks.rds"
if (file.exists(f)) {
  d <- readRDS(f)
} else {
d <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
  block_ids <- readRDS("data/survey-blocks-36.rds") 
  d <- dplyr::left_join(d, block_ids)
  d$block_designation <- as.numeric(d$block_designation)
  saveRDS(d, file = "data-generated/halibut-surv-36-blocks.rds")
}

hook <- readRDS("data-generated/halibut_hook_adjusted_36.rds") %>%
  select(
    survey_series_id, year, fishing_event_id, total_hooks, prop_bait_hooks,
    hook_adjust_factor, expected_catch
  )

d <- left_join(d, hook, by = c("survey_series_id", "year", "fishing_event_id"))

d <- d %>%
  select(
    survey_abbrev, year, longitude, latitude, catch_count, hook_count,
    grouping_code, depth_m, block_designation, hook_adjust_factor, prop_bait_hooks
  ) %>% mutate(lon = longitude, lat = latitude) %>%
  rename(survey = survey_abbrev, block = block_designation)

d_utm <- convert2utm(d, coords = c("lon", "lat"))
d_utm <- filter(d_utm, grouping_code > 270 & grouping_code < 330) # not all part of survey

load("data/hbll_s_grid.rda")

d_utm$depth_log <- log(d_utm$depth_m)
d_utm$depth_centred <- d_utm$depth_log - mean(d_utm$depth_log)
d_utm$depth_mean <- mean(d_utm$depth_log)
d_utm$depth_sd <- sd(d_utm$depth_centred)
d_utm$depth_scaled <- d_utm$depth_centred / sd(d_utm$depth_centred)
d_utm$Y_cent <- d_utm$Y - mean(d_utm$Y)
d_utm$X_cent <- d_utm$X - mean(d_utm$X)
d_utm$area_swept <- d_utm$hook_count * 0.0024384 * 0.009144 * 1000
d_utm$offset_area_swept <- log(d_utm$area_swept)
d_utm$offset_area_hook <- log(d_utm$area_swept / d_utm$hook_adjust_factor)

s_grid_utm <- convert2utm(hbll_s_grid$grid, coords = c("X", "Y"))

s_grid_utm$longitude <- hbll_s_grid$grid$X 
s_grid_utm$latitude <- hbll_s_grid$grid$Y

s_grid_utm$offset_area_hook <- mean(d_utm$offset_area_hook)
s_grid_utm$offset_area_swept <- mean(d_utm$offset_area_swept)
s_grid_utm$hook_adjust_factor <- mean(d_utm$hook_adjust_factor)
s_grid_utm$hook_count <- 100 #mean(d_utm$hook_count)
s_grid_utm$offset <- log(100)

years <- sort(unique(d_utm$year))
s_grid_utm <- expand_prediction_grid(s_grid_utm, years = years) %>%
  mutate(depth_centred = log(depth) - mean(d_utm$depth_log)) %>%
  mutate(depth_scaled = depth_centred / sd(d_utm$depth_centred))
s_grid_utm <- mutate(s_grid_utm, Y_cent = Y - mean(d_utm$Y))
s_grid_utm <- mutate(s_grid_utm, X_cent = X - mean(d_utm$X))
```

```{r }

sp <- make_mesh(d_utm, c("X", "Y"), n_knots = 200)
plot(sp)
# png("figs/hbll-outside-S-spde.png", width = 7, height = 7, units = "in", res = 180)
```

```{r}
glimpse(d_utm)
```

# just depth
```{r}
# m0 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) +
#           depth_scaled + I(depth_scaled^2),
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m0, file = "models/halibut-m0-depth-200kn.rds")
m0  <-  readRDS(file = "models/halibut-m0-depth-200kn.rds")
m0
get_diag(m0)
```

```{r}
d_utm$offset <- d_utm$offset_area_hook

m1 <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m1, file = "models/halibut-m1-offset-area-hook-200kn.rds")

# m1  <- readRDS(file = "models/halibut-m1-offset-area-hook-200kn.rds")
m1
get_diag(m1)
```

# estimate effect of hook area 
```{r}
m1b <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset_area_hook,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m1b, file = "models/halibut-m1b-area-hook-200kn.rds")
# m1b  <- readRDS(file = "models/halibut-m1b-area-hook-200kn.rds")
m1b
get_diag(m1b)
```

# area swept worse than area hook
```{r}
# d_utm$offset <- d_utm$offset_area_swept
# 
# m2 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) +
#           depth_scaled + I(depth_scaled^2) + offset,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m2, file = "models/halibut-m2-offset-area-swept-200kn.rds")
m2  <-  readRDS(file = "models/halibut-m2-offset-area-swept-200kn.rds")
m2
get_diag(m2)
```

# estimate effect of area swept better AIC but worse MSE than area hook 
```{r}
m2b <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) +
          depth_scaled + I(depth_scaled^2) + offset_area_swept,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m2b, file = "models/halibut-m2b-area-swept-200kn.rds")

# m2b  <-  readRDS(file = "models/halibut-m2b-area-swept-200kn.rds")
m2b
get_diag(m2b)
```
# just an estimate for the hook adjustment factor
## second best AIC to model predicting effect of area swept
```{r}
# m3 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) +
#           depth_scaled + I(depth_scaled^2) +
#           hook_adjust_factor,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m3, file = "models/halibut-m3-hook-factor-200kn.rds")

m3  <- readRDS(file = "models/halibut-m3-hook-factor-200kn.rds")
m3
get_diag(m3)
```

# use swept area offset and estimate hook adjustment
```{r}
d_utm$offset <- d_utm$offset_area_swept

m4 <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year)+
          depth_scaled + I(depth_scaled^2) +
          hook_adjust_factor +
          offset,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m4, file = "models/halibut-m4-hookfact-offset-area-swept-200kn.rds")
# m4  <-  readRDS(file = "models/halibut-m4-hookfact-offset-area-swept-200kn.rds")
m4
get_diag(m4)
```

# estimate swept area and hook adjustment
```{r}
# m4b <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year)+
#           depth_scaled + I(depth_scaled^2) +
#           hook_adjust_factor +
#           offset_area_swept,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m4b, file = "models/halibut-m4b-hookfact-area-swept-200kn.rds")
m4b  <-  readRDS(file = "models/halibut-m4b-hookfact-area-swept-200kn.rds")
m4b
get_diag(m4b)
```

# Try with more knots
```{r }
sp4 <- make_mesh(d_utm, c("X", "Y"), n_knots = 400)
plot(sp4)
# png("figs/hbll-outside-S-spde.png", width = 7, height = 7, units = "in", res = 180)
```

# estimate swept area and hook adjustment
```{r}
m4c <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year)+
          depth_scaled + I(depth_scaled^2) +
          hook_adjust_factor +
          offset_area_swept,
      data = d_utm,
      spde = sp4,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m4c, file = "models/halibut-m4c-hookfact-area-swept-400kn.rds")
# m4c  <-  readRDS(file = "models/halibut-m4c-hookfact-area-swept-400kn.rds")
m4c
get_diag(m4c)
```

# offset of hook_count
```{r}
d_utm$offset <- log(d_utm$hook_count)

m5 <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
          depth_scaled + I(depth_scaled^2) +
          offset,
      data = d_utm,
      spde = sp4,
      ar1_fields = T,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m5, file = "models/halibut-m5-hook-count-400kn-AR1.rds")
# m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
m5
tidy(m5, "ran_pars",conf.int = TRUE)
get_diag(m5)
```

# spatial only model
```{r}
# d_utm$offset <- log(d_utm$hook_count)
# 
# m5s <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
#           depth_scaled + I(depth_scaled^2) +
#           offset,
#       data = d_utm,
#       spde = sp4,
#       # ar1_fields = T,
#       # time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m5s, file = "models/halibut-m5-hook-count-400kn-spatial.rds")
m5s  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-spatial.rds")
m5s
tidy(m5s, "ran_pars",conf.int = TRUE)
get_diag(m5s)
```

# Predictions from best model
```{r}
# p <- predict(m5, newdata = s_grid_utm, se_fit = F, re_form = NULL, return_tmb_object = T)
# saveRDS(p, "models/halibut-m5-AR1-predictions.rds")
```

```{r}
plot_map <- function(dat, column, wrap = TRUE) {
  gg <- ggplot(data = dat) +
    geom_tile(mapping = aes(X, Y, fill = {{ column }}), width = 0.025, height = 0.025) +
    coord_fixed() +
    scale_fill_viridis_c(option = "D")
  if (wrap) gg + facet_wrap(~year) else gg
}

# p <- readRDS("models/halibut-m4c-AR1-predictions.rds")
p <- readRDS("models/halibut-m5-AR1-predictions.rds")
g <- plot_map(p$data, exp(est)) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), option = "D") +
  labs(
    fill = "Estimated\nadjusted\ncount",
    size = "Observed\nadjusted\ncount"
  ) +
  geom_point(
    data = d_utm, pch = 21, mapping = aes(
      x = X, y = Y,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey20", alpha = 0.5
  ) +
  scale_size_area(max_size = 7)
g
```

```{r}

  gg <- ggplot(filter(p$data, year == 2020)) +
    geom_tile(mapping = aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
    coord_fixed() +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), option = "D") +
  labs(
    fill = "Estimated\nadjusted\ncount",
    size = "Observed\nadjusted\ncount"
  ) +
  geom_point(
    data = filter(d_utm, year == 2020), pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey20", alpha = 0.5
  ) +
  scale_size_area(max_size = 7)
gg
```

# Add court defined area shape file

```{r}
library(sf)
# library(raster)
focal_area <-sf::st_read(dsn="shape-files/taaqwiihak_areaVer2.shp",layer="taaqwiihak_areaVer2")
focal_area <- sf::st_transform(focal_area, crs = 4326)

# make prediction grid st

s_grid_3cd <- filter(s_grid_utm, latitude<50.5)

s_grid_3cd_sf <- st_as_sf(s_grid_3cd, coords = c("longitude", "latitude"))
st_crs(s_grid_3cd_sf) <- 4326 # set the coordinate reference system
s_grid_3cd_sf

intersected <- sf::st_intersects(s_grid_3cd_sf, focal_area)

# cda <- which(lengths(intersected) > 0)
cda_grid_sf <- s_grid_3cd_sf[which(lengths(intersected) > 0), ]
noncda_grid_sf <- s_grid_3cd_sf[which(lengths(intersected) == 0), ]

sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

cda_grid <- sfc_as_cols(cda_grid_sf, c("longitude", "latitude")) 
noncda_grid <- sfc_as_cols(noncda_grid_sf, c("longitude", "latitude")) 
st_geometry(cda_grid) <- NULL
st_geometry(noncda_grid) <- NULL
```

# Predictions for CDA only
```{r}
m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
p_cda <- predict(m5, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(p_cda, "models/halibut-m5-AR1-predictions-cda.rds")

i_cda <- get_index(p_cda)
saveRDS(i_cda, "models/halibut-m5-AR1-index-cda.rds")

ggplot(i_cda, aes(year, est)) + geom_line() +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4) +
  xlab('Year') + ylab('Biomass estimate (metric tonnes)')

```

# Predictions outside CDA but within 3cd
```{r}
# m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
p_3cd <- predict(m5, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(p_3cd, "models/halibut-m5-AR1-predictions-3cd.rds")

i_3cd <- get_index(p_3cd)
saveRDS(i_3cd, "models/halibut-m5-AR1-index-3cd.rds")

ggplot(i_3cd, aes(year, est)) + geom_line() +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.4) +
  xlab('Year') + ylab('Biomass estimate (metric tonnes)')

```

# or without time
```{r}
m5s  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-spatial.rds")
sp_cda <- predict(m5s, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(ps, "models/halibut-m5-spatial-predictions-cda.rds")

i_cda_sp <- get_index(sp_cda)
# saveRDS(i_sp, "models/ye-m5-cda-spatial-index.rds")

sp_noncda <- predict(m5s, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(ps, "models/halibut-m5-spatial-predictions-cda.rds")

i_noncda_sp <- get_index(sp_noncda)
# saveRDS(i_sp, "models/halibut-m5-cda-spatial-index.rds")

i_cda_sp$est / (i_cda_sp$est + i_noncda_sp$est)
```

# map all of S HBLL
```{r}
# creates coast lines for area defined in lat lon
load_coastll <- function(xlim_ll, ylim_ll, utm_zone, buffer = 0) {
  data("nepacLLhigh", package = "PBSmapping", envir = environment())
  np <- PBSmapping::clipPolys(nepacLLhigh,
    xlim = xlim_ll + c(-buffer, buffer),
    ylim = ylim_ll + c(-buffer, buffer)
  )
  # ll2utm(np, utm_zone = utm_zone)
}

coast <- load_coastll(
    range(d_utm$longitude) ,
    range(d_utm$latitude) ,
    utm_zone = 9
  )

load_boundaries <- function(utm_zone) {
  data("major", package = "PBSdata", envir = environment())
  gfplot:::ll2utm(major, utm_zone = utm_zone)
}

majorbound  <- load_boundaries(9)
# to see what's what
library(PBSmapping) # needs this for some reason
attributes(majorbound)$PolyData
bound3CDnorth <- fortify(majorbound) %>% 
  filter(PID %in% c(3,4)) %>%  filter(X > 400 & X < 600 & Y > 5550)%>% gfplot:::utm2ll(utm_zone = 9)

# predictions <- p$data %>% 
#   st_as_sf(crs = 4326, coords = c("longitude", "latitude"))

# p_hal <- readRDS("models/halibut-m4c-AR1-predictions.rds")
p_hal <- readRDS("models/halibut-m5-AR1-predictions.rds")
gh <- ggplot(data = focal_area) +
  # plot predictions from best model with mean swept area and hook adjustment factor
  geom_tile(data = filter(p_hal$data, year == 2020), 
    aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), 
    option = "D", limits=c(0, 40), na.value = "yellow") +
  labs(
    fill = "Estimated\nadjusted\ncount",
    size = "Observed\nadjusted\ncount"
  ) +
  geom_point(
    data = filter(d_utm, year == 2020), #just most recent year
    # data = d_utm, # plot all years together
    pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_line(# add major management region boundaries
    data = bound3CDnorth,
    aes(X, Y), colour = "black", lty = 2, 
    inherit.aes = F
  ) +
  annotate("text", x = -129.5, y = 50.25, label = "3CD") +
  annotate("text", x = -129.5, y = 51.25, label = "5AB") +
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.5, 52.1))  +# trim points beyond grid boundaries
  ggtitle("HBLL observed and predicted Pacific Halibut catch per 100 hooks")
gh
ggsave("halibut-m5-AR1-2020-predictions.png", width = 6, height = 6)
```

```{r}
gh3cd <- ggplot(data = focal_area) +
   geom_tile(data = filter(p_cda$data, year == 2020),
   aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
   geom_tile(data = filter(p_3cd$data, year == 2020),
    aes(longitude, latitude, fill = exp(est)), width = 0.05, height = 0.025) +
  scale_fill_viridis_c(
    trans = ggsidekick::fourth_root_power_trans(),
    limits=c(0, 7), na.value = "yellow",
    option = "D") +
  labs(
    fill = "Estimated\ncount\nin 3CD",
    size = "Observed\ncount"
  ) +
  geom_point(# plot all years together
    data = d_utm, pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.4, 50.6), xlim=c(-128.5, -125.15))  +# trim points beyond grid boundaries
  ggtitle("HBLL observed and predicted Pacific Halibut catch per 100 hooks")
gh3cd
ggsave("halibut-m5-AR1-2020-predictions-3cd.png", width = 6, height = 6)
```

```{r}
py <- readRDS("models/ye-m5-AR1-predictions.rds")

pyd <- py$data %>% rename(ye_est = est) 
ph <- readRDS("models/halibut-m5-AR1-predictions.rds")
phd <- ph$data %>% rename(hal_est = est) %>% select(hal_est, X, Y, year)

ratio_df <- left_join(pyd, phd) %>% mutate(halibut = exp(hal_est), yelloweye = exp(ye_est), ye_per_hal = yelloweye/halibut)

gratio <- ggplot(data = focal_area) +
  # plot predictions from best model with mean swept area and hook adjustment factor
  geom_tile(data = filter(ratio_df, year == 2020), 
    aes(longitude, latitude, fill = ye_per_hal), width = 0.06, height = 0.025) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), 
    # limits=c(0, 40), na.value = "yellow", 
    option = "D"
    ) +
  labs(
    fill = "Yelloweye\ncaught\nper halibut",
    size = "Observed\nhalibut\ncatch"
  ) +
  geom_point(# plot all years together
    data = d_utm, pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_line(# add major management region boundaries
    data = bound3CDnorth,
    aes(X, Y), colour = "black", lty = 2, 
    inherit.aes = F
  ) +
  annotate("text", x = -129.5, y = 50.25, label = "3CD") +
  annotate("text", x = -129.5, y = 51.25, label = "5AB") +
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.5, 52.1))  +# trim points beyond grid boundaries
  ggtitle("Estimated number of Yelloweye Rockfish caught per halibut")
gratio
ggsave("ye-per-halibut-w-hal-catch.png", width = 6, height = 6)

```
# try tv depth
### not useful
```{r}
# m5 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year)+
#           hook_adjust_factor +
#           offset_area_swept,
#       time_varying = ~ 0 + depth_scaled + I(depth_scaled^2),
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m5, file = "models/halibut-m5-tv-depth-200kn.rds")
m5  <-  readRDS(file = "models/halibut-m5-tv-depth-200kn.rds")
m5
get_diag(m5)
```

```{r}
tvd <- gfranges::time_varying_density(m5, predictor = "depth")
tvd$x <- exp(tvd$x)
p <- gfranges::plot_mountains(tvd,
    variable_label = "Depth (m)",
    peaklines = F,
    mountains_only = F,
    xlimits = c(15, 270))
p
```

```{r}
# make_tv_depth_grid <- function(m){
#   expand.grid(
#     depth_scaled = seq(min(m$data$depth_scaled), max(m$data$depth_scaled), length.out = 100),
#     hook_adjust_factor = mean(m$data$hook_adjust_factor), 
#     offset_area_swept = mean(m$data$offset_area_swept), 
#     year = as.factor(unique(m$data$year))
#   )}
# 
# depth_grid <- make_tv_depth_grid(m5)
# 
# species_depth_profile <- function(m, depth_grid,
#   min_depth_m = 15, 
#   max_depth_m = 240,
#   min_year = min(m$data$year), 
#   response = "count"
#   ){
# 
#   p <- predict(m, newdata = depth_grid, se_fit = TRUE, re_form = NA)
#   p <- p %>% mutate(
#     Depth = exp(depth_scaled * m$data$depth_sd[1] + m$data$depth_mean[1]),
#     response = exp(est),
#     lowCI = exp(est - 1.96 * est_se),
#     highCI = exp(est + 1.96 * est_se))
#   
# g <- ggplot(filter(p, Depth > min_depth_m & year > min_year), 
#   aes(Depth, response,
#   ymin = lowCI, 
#   ymax = highCI, 
#   group = as.factor(year), fill = year, colour = year)) +
#   geom_ribbon(alpha = 0.1, colour = NA) +
#   geom_line(size = 0.5, alpha =0.85) +
#   scale_fill_viridis_c(option = "C") +
#   scale_colour_viridis_c(option = "C") +
#   coord_cartesian(xlim = c(min_depth_m, max_depth_m), ylim=c(0, quantile(p$Biomass, 1))) +
#   xlab("Depth (m)") +
#   gfplot::theme_pbs() 
# 
# if(response=="biomass") {
# g <- g + ylab("Biomass")
# }
# 
# if(response=="count") {
# g <- g + ylab("Individuals")
# }
# 
# g
# }
# 
# (tv_plot <- species_depth_profile(m5, depth_grid))
```

### 2021-05-20 analysis/misc/yelloweye-south-outside.Rmd
exploring model structures for outside S HBLL

--- 2021-05-20 analysis/misc/halibut-south-outside.Rmd
@@
 ---
-title: "outside halibut"
+title: "outside yelloweye"
 author: "Philina English"
@@
     ggplot(dat, aes_string("X", "Y", colour = column)) +
-      geom_point(alpha = 0.5) +
+      geom_point(alpha = 0.25) +
       coord_fixed()+
@@
 ```{r}
-hook_adjusted_data <- "data-generated/halibut_hook_adjusted_36.rds"
+hook_adjusted_data <- "data-generated/yelloweye_hook_adjusted_36.rds"
 if (!file.exists(hook_adjusted_data)) {
-data <- readRDS("data/halibut-hook-dat36.rds")
-d1 <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
+data <- readRDS("data/yelloweye-rockfish-hook-dat36.rds")
+d1 <- readRDS("data/yelloweye-rockfish-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
 
@@
 
-saveRDS(hook_adjusted_data, file = "data-generated/halibut_hook_adjusted_36.rds")
+saveRDS(hook_adjusted_data, file = "data-generated/yelloweye_hook_adjusted_36.rds")
 }
@@
 ```{r}
-f <- "data-generated/halibut-surv-36-blocks.rds"
+f <- "data-generated/yelloweye-rockfish-surv-36-blocks.rds"
 if (file.exists(f)) {
@@
 } else {
-d <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
+d <- readRDS("data/yelloweye-rockfish-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
   block_ids <- readRDS("data/survey-blocks-36.rds") 
@@
   d$block_designation <- as.numeric(d$block_designation)
-  saveRDS(d, file = "data-generated/halibut-surv-36-blocks.rds")
+  saveRDS(d, file = "data-generated/yelloweye-rockfish-surv-36-blocks.rds")
 }
 
-hook <- readRDS("data-generated/halibut_hook_adjusted_36.rds") %>%
+hook <- readRDS("data-generated/yelloweye_hook_adjusted_36.rds") %>%
   select(
@@
   rename(survey = survey_abbrev, block = block_designation)
+
+# hb <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
 
@@
 #     )
-# saveRDS(m0, file = "models/halibut-m0-depth-200kn.rds")
-m0  <-  readRDS(file = "models/halibut-m0-depth-200kn.rds")
+# saveRDS(m0, file = "models/ye-m0-depth-200kn.rds")
+m0  <-  readRDS(file = "models/ye-m0-depth-200kn.rds")
 m0
+
 get_diag(m0)
@@
 ```{r}
-d_utm$offset <- d_utm$offset_area_hook
+# d_utm$offset <- d_utm$offset_area_hook
+# 
+# m1 <- sdmTMB(
+#       formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset,
+#       data = d_utm,
+#       spde = sp,
+#       time = "year",
+#       silent = FALSE,
+#       anisotropy = TRUE,
+#       #cores = cores,
+#       reml = TRUE,
+#       family = nbinom2(link = "log")
+#     )
+# saveRDS(m1, file = "models/ye-m1-offset-area-hook-200kn.rds")
 
-m1 <- sdmTMB(
-      formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset,
-      data = d_utm,
-      spde = sp,
-      time = "year",
-      silent = FALSE,
-      anisotropy = TRUE,
-      #cores = cores,
-      reml = TRUE,
-      family = nbinom2(link = "log")
-    )
-saveRDS(m1, file = "models/halibut-m1-offset-area-hook-200kn.rds")
-
-# m1  <- readRDS(file = "models/halibut-m1-offset-area-hook-200kn.rds")
+m1  <- readRDS(file = "models/ye-m1-offset-area-hook-200kn.rds")
 m1
@@
 ```{r}
-m1b <- sdmTMB(
-      formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset_area_hook,
-      data = d_utm,
-      spde = sp,
-      time = "year",
-      silent = FALSE,
-      anisotropy = TRUE,
-      #cores = cores,
-      reml = TRUE,
-      family = nbinom2(link = "log")
-    )
-saveRDS(m1b, file = "models/halibut-m1b-area-hook-200kn.rds")
-# m1b  <- readRDS(file = "models/halibut-m1b-area-hook-200kn.rds")
+# m1b <- sdmTMB(
+#       formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset_area_hook,
+#       data = d_utm,
+#       spde = sp,
+#       time = "year",
+#       silent = FALSE,
+#       anisotropy = TRUE,
+#       #cores = cores,
+#       reml = TRUE,
+#       family = nbinom2(link = "log")
+#     )
+# saveRDS(m1b, file = "models/ye-m1b-area-hook-200kn.rds")
+m1b  <- readRDS(file = "models/ye-m1b-area-hook-200kn.rds")
 m1b
@@
 
-# area swept worse than area hook
 ```{r}
 # d_utm$offset <- d_utm$offset_area_swept
-# 
+
 # m2 <- sdmTMB(
@@
 #     )
-# saveRDS(m2, file = "models/halibut-m2-offset-area-swept-200kn.rds")
-m2  <-  readRDS(file = "models/halibut-m2-offset-area-swept-200kn.rds")
+# saveRDS(m2, file = "models/ye-m2-offset-area-swept-200kn.rds")
+m2  <-  readRDS(file = "models/ye-m2-offset-area-swept-200kn.rds")
 m2
@@
 ```
+# estimate effect of area swept 
+## best so far
+```{r}
+# m2b <- sdmTMB(
+#       formula = catch_count ~ 0 + as.factor(year) + 
+#           depth_scaled + I(depth_scaled^2) + offset_area_swept,
+#       data = d_utm,
+#       spde = sp,
+#       time = "year",
+#       silent = FALSE,
+#       anisotropy = TRUE,
+#       #cores = cores,
+#       reml = TRUE,
+#       family = nbinom2(link = "log")
+#     )
+# saveRDS(m2b, file = "models/ye-m2b-area-swept-200kn.rds")
 
-# estimate effect of area swept better AIC but worse MSE than area hook 
-```{r}
-m2b <- sdmTMB(
-      formula = catch_count ~ 0 + as.factor(year) +
-          depth_scaled + I(depth_scaled^2) + offset_area_swept,
-      data = d_utm,
-      spde = sp,
-      time = "year",
-      silent = FALSE,
-      anisotropy = TRUE,
-      #cores = cores,
-      reml = TRUE,
-      family = nbinom2(link = "log")
-    )
-saveRDS(m2b, file = "models/halibut-m2b-area-swept-200kn.rds")
-
-# m2b  <-  readRDS(file = "models/halibut-m2b-area-swept-200kn.rds")
+m2b  <-  readRDS(file = "models/ye-m2b-area-swept-200kn.rds")
 m2b
@@
 #     )
-# saveRDS(m3, file = "models/halibut-m3-hook-factor-200kn.rds")
+# saveRDS(m3, file = "models/ye-m3-hook-factor-200kn.rds")
 
-m3  <- readRDS(file = "models/halibut-m3-hook-factor-200kn.rds")
+m3  <- readRDS(file = "models/ye-m3-hook-factor-200kn.rds")
 m3
@@
 ```{r}
-d_utm$offset <- d_utm$offset_area_swept
-
-m4 <- sdmTMB(
-      formula = catch_count ~ 0 + as.factor(year)+
-          depth_scaled + I(depth_scaled^2) +
-          hook_adjust_factor +
-          offset,
-      data = d_utm,
-      spde = sp,
-      time = "year",
-      silent = FALSE,
-      anisotropy = TRUE,
-      #cores = cores,
-      reml = TRUE,
-      family = nbinom2(link = "log")
-    )
-saveRDS(m4, file = "models/halibut-m4-hookfact-offset-area-swept-200kn.rds")
-# m4  <-  readRDS(file = "models/halibut-m4-hookfact-offset-area-swept-200kn.rds")
+# d_utm$offset <- d_utm$offset_area_swept
+# 
+# m4 <- sdmTMB(
+#       formula = catch_count ~ 0 + as.factor(year)+
+#           depth_scaled + I(depth_scaled^2) + 
+#           hook_adjust_factor +
+#           offset,
+#       data = d_utm,
+#       spde = sp,
+#       time = "year",
+#       silent = FALSE,
+#       anisotropy = TRUE,
+#       #cores = cores,
+#       reml = TRUE,
+#       family = nbinom2(link = "log")
+#     )
+# saveRDS(m4, file = "models/ye-m4-hookfact-offset-area-swept-200kn.rds")
+m4  <-  readRDS(file = "models/ye-m4-hookfact-offset-area-swept-200kn.rds")
 m4
@@
 #     )
-# saveRDS(m4b, file = "models/halibut-m4b-hookfact-area-swept-200kn.rds")
-m4b  <-  readRDS(file = "models/halibut-m4b-hookfact-area-swept-200kn.rds")
+# saveRDS(m4b, file = "models/ye-m4b-hookfact-area-swept-200kn.rds")
+m4b  <-  readRDS(file = "models/ye-m4b-hookfact-area-swept-200kn.rds")
 m4b
@@
       spde = sp4,
-      time = "year",
-      silent = FALSE,
-      anisotropy = TRUE,
-      #cores = cores,
-      reml = TRUE,
-      family = nbinom2(link = "log")
-    )
-saveRDS(m4c, file = "models/halibut-m4c-hookfact-area-swept-400kn.rds")
-# m4c  <-  readRDS(file = "models/halibut-m4c-hookfact-area-swept-400kn.rds")
-m4c
-get_diag(m4c)
-```
-
-# offset of hook_count
-```{r}
-d_utm$offset <- log(d_utm$hook_count)
-
-m5 <- sdmTMB(
-      formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
-          depth_scaled + I(depth_scaled^2) +
-          offset,
-      data = d_utm,
-      spde = sp4,
       ar1_fields = T,
@@
     )
-saveRDS(m5, file = "models/halibut-m5-hook-count-400kn-AR1.rds")
-# m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
+saveRDS(m4c, file = "models/ye-m4c-hookfact-area-swept-400kn-AR1.rds")
+m4c  <-  readRDS(file = "models/ye-m4c-hookfact-area-swept-400kn-AR1.rds")
+m4c
+tidy(m4c, "ran_pars",conf.int = TRUE)
+get_diag(m4c)
+```
+
+# offset of hook_count
+```{r}
+# d_utm$offset <- log(d_utm$hook_count)
+# 
+# m5 <- sdmTMB(
+#       formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
+#           depth_scaled + I(depth_scaled^2) + 
+#           offset,
+#       data = d_utm,
+#       spde = sp4,
+#       ar1_fields = T,
+#       time = "year",
+#       silent = FALSE,
+#       anisotropy = TRUE,
+#       #cores = cores,
+#       reml = TRUE,
+#       family = nbinom2(link = "log")
+#     )
+saveRDS(m5, file = "models/ye-m5-hook-count-400kn-AR1.rds")
+# m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
 m5
@@
 ```{r}
-# d_utm$offset <- log(d_utm$hook_count)
-# 
-# m5s <- sdmTMB(
-#       formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
-#           depth_scaled + I(depth_scaled^2) +
-#           offset,
-#       data = d_utm,
-#       spde = sp4,
-#       # ar1_fields = T,
-#       # time = "year",
-#       silent = FALSE,
-#       anisotropy = TRUE,
-#       #cores = cores,
-#       reml = TRUE,
-#       family = nbinom2(link = "log")
-#     )
-# saveRDS(m5s, file = "models/halibut-m5-hook-count-400kn-spatial.rds")
-m5s  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-spatial.rds")
+d_utm$offset <- log(d_utm$hook_count)
+
+m5s <- sdmTMB(
+      formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
+          depth_scaled + I(depth_scaled^2) +
+          offset,
+      data = d_utm,
+      spde = sp4,
+      # ar1_fields = T,
+      # time = "year",
+      silent = FALSE,
+      anisotropy = TRUE,
+      #cores = cores,
+      reml = TRUE,
+      family = nbinom2(link = "log")
+    )
+saveRDS(m5s, file = "models/ye-m5-hook-count-400kn-spatial.rds")
+# m5s  <-  readRDS(file = "models/ye-m5-hook-count-400kn-spatial.rds")
 m5s
@@
 ```{r}
+
+# m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
 # p <- predict(m5, newdata = s_grid_utm, se_fit = F, re_form = NULL, return_tmb_object = T)
-# saveRDS(p, "models/halibut-m5-AR1-predictions.rds")
+# saveRDS(p, "models/ye-m5-AR1-predictions.rds")
 ```
@@
 
-# p <- readRDS("models/halibut-m4c-AR1-predictions.rds")
-p <- readRDS("models/halibut-m5-AR1-predictions.rds")
+# p <- readRDS("models/ye-m4c-AR1-predictions.rds")
+p <- readRDS("models/ye-m5-AR1-predictions.rds")
 g <- plot_map(p$data, exp(est)) +
@@
 
-# Add court defined area shape file
+# try adding shape files
 
@@
 ```{r}
-m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
+m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
 p_cda <- predict(m5, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
-saveRDS(p_cda, "models/halibut-m5-AR1-predictions-cda.rds")
+# saveRDS(p_cda, "models/ye-m5-AR1-predictions-cda.rds")
 
 i_cda <- get_index(p_cda)
-saveRDS(i_cda, "models/halibut-m5-AR1-index-cda.rds")
+# saveRDS(i_cda, "models/ye-m5-AR1-index-cda.rds")
 
@@
 ```{r}
-# m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
+# m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
 p_3cd <- predict(m5, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
-saveRDS(p_3cd, "models/halibut-m5-AR1-predictions-3cd.rds")
+saveRDS(p_3cd, "models/ye-m5-AR1-predictions-3cd.rds")
 
 i_3cd <- get_index(p_3cd)
-saveRDS(i_3cd, "models/halibut-m5-AR1-index-3cd.rds")
+saveRDS(i_3cd, "models/ye-m5-AR1-index-3cd.rds")
 
@@
 ```{r}
-m5s  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-spatial.rds")
+m5s  <-  readRDS(file = "models/ye-m5-hook-count-400kn-spatial.rds")
 sp_cda <- predict(m5s, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
-saveRDS(ps, "models/halibut-m5-spatial-predictions-cda.rds")
+saveRDS(ps, "models/ye-m5-spatial-predictions-cda.rds")
 
@@
 sp_noncda <- predict(m5s, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
-saveRDS(ps, "models/halibut-m5-spatial-predictions-cda.rds")
+saveRDS(ps, "models/ye-m5-spatial-predictions-cda.rds")
 
 i_noncda_sp <- get_index(sp_noncda)
-# saveRDS(i_sp, "models/halibut-m5-cda-spatial-index.rds")
+# saveRDS(i_sp, "models/ye-m5-cda-spatial-index.rds")
 
@@
 
-# map all of S HBLL
 ```{r}
+
 # creates coast lines for area defined in lat lon
@@
 
-# p_hal <- readRDS("models/halibut-m4c-AR1-predictions.rds")
-p_hal <- readRDS("models/halibut-m5-AR1-predictions.rds")
-gh <- ggplot(data = focal_area) +
+# p <- readRDS("models/ye-m4c-AR1-predictions.rds")
+p <- readRDS("models/ye-m5-AR1-predictions.rds")
+gye <- ggplot(data = focal_area) +
   # plot predictions from best model with mean swept area and hook adjustment factor
-  geom_tile(data = filter(p_hal$data, year == 2020), 
+  geom_tile(data = filter(p$data, year == 2020),
     aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
@@
   ) +
-  geom_point(
-    data = filter(d_utm, year == 2020), #just most recent year
-    # data = d_utm, # plot all years together
-    pch = 21, mapping = aes(
+  geom_point(# plot all years together
+    data = d_utm, pch = 21, mapping = aes(
       longitude, latitude,
@@
   coord_sf(expand = F, ylim=c(48.5, 52.1))  +# trim points beyond grid boundaries
-  ggtitle("HBLL observed and predicted Pacific Halibut catch per 100 hooks")
-gh
-ggsave("halibut-m5-AR1-2020-predictions.png", width = 6, height = 6)
+  ggtitle("HBLL observed and predicted Yelloweye Rockfish catch per 100 hooks")
+gye
+# ggsave("ye-m5-AR1-2020-predictions.png", width = 6, height = 6)
 ```
@@
 ```{r}
-gh3cd <- ggplot(data = focal_area) +
+g3cd <- ggplot(data = focal_area) +
    geom_tile(data = filter(p_cda$data, year == 2020),
@@
     trans = ggsidekick::fourth_root_power_trans(),
-    limits=c(0, 7), na.value = "yellow",
+    # limits=c(0, 40), na.value = "yellow",
     option = "D") +
@@
   ) +  scale_size_area(max_size = 4) + 
+  # geom_line(# add major management region boundaries
+  #   data = bound3CDnorth,
+  #   aes(X, Y), colour = "black", lty = 2, 
+  #   inherit.aes = F
+  # ) +
+  # annotate("text", x = -128, y = 49.5, label = "3CD") +
   geom_sf(colour = "red", fill = NA) + # add focal area
@@
   coord_sf(expand = F, ylim=c(48.4, 50.6), xlim=c(-128.5, -125.15))  +# trim points beyond grid boundaries
-  ggtitle("HBLL observed and predicted Pacific Halibut catch per 100 hooks")
-gh3cd
-ggsave("halibut-m5-AR1-2020-predictions-3cd.png", width = 6, height = 6)
+  ggtitle("HBLL observed and predicted Yelloweye Rockfish catch per 100 hooks")
+g3cd
+ggsave("ye-m5-AR1-2020-predictions-3cd.png", width = 6, height = 6)
 ```
@@
     fill = "Yelloweye\ncaught\nper halibut",
-    size = "Observed\nhalibut\ncatch"
+    size = "Observed\nyelloweye\ncatch"
   ) +
@@
 gratio
-ggsave("ye-per-halibut-w-hal-catch.png", width = 6, height = 6)
+ggsave("ye-per-halibut.png", width = 6, height = 6)
 
 ```
+
 # try tv depth
@@
 #     )
-# saveRDS(m5, file = "models/halibut-m5-tv-depth-200kn.rds")
-m5  <-  readRDS(file = "models/halibut-m5-tv-depth-200kn.rds")
+# saveRDS(m5, file = "models/ye-m5-tv-depth-200kn.rds")
+m5  <-  readRDS(file = "models/ye-m5-tv-depth-200kn.rds")
 m5
